### 2013-09-15 doc/chapters/output.tex
add manual tex file with the rough structure layout

\section{Output}
\subsection{Files}

Explain the created C++ files and executables, and how to use them.

\subsection{Two-scale algorithm}

Explain the structure of the two-scale algorithm classes and how they
work and are plugged together.

\subsection{Lattice algorithm}

Explain the structure of the lattice algorithm classes and how they
work.

### 2013-09-15 doc/chapters/overview.tex
add manual tex file with the rough structure layout

\section{Overview}

Explain:
\begin{itemize}
\item What is FlexibleSUSY?
\item Features, design goals
\end{itemize}

\subsection{Requirements}

\begin{itemize}
\item Mathematica 7-9
\item SARAH version 4
\item a C++11 compatible compiler (g++ 4.7.2 or higher, clang++ 3.1 or
  higher)
\item a fortran compiler (gfortran etc.)
\item Eigen library version 3
\item Boost library
\item GNU scientific library
\item lapack (needed for the Lattice algorithm only)
\end{itemize}

### 2013-09-15 doc/chapters/quick_start.tex
add manual tex file with the rough structure layout

\section{Quick start}
Create a MSSM spectrum generator
  \begin{lstlisting}[language=bash]
$ ./createmodel --models=MSSM
$ ./configure --with-models=MSSM
$ make
  \end{lstlisting}
Run the spectrum generator using the default parameter point
  \begin{lstlisting}[language=bash]
$ ./models/MSSM/run_MSSM.x
  \end{lstlisting}
Run the spectrum generator using a SLHA input file
  \begin{lstlisting}[language=bash]
$ ./models/MSSM/run_MSSM.x --slha-input-file=templates/MSSM/LesHouches.in.MSSM --slha-output-file=LesHouches.out.MSSM
  \end{lstlisting}

### 2013-09-15 doc/chapters/usage.tex
add manual tex file with the rough structure layout

\section{Usage}
\subsection{Files and directories}

Explain:
\begin{itemize}
\item Soucre code directories and files
\item Model directories and files
\end{itemize}

\subsection{Basic commands}

Explain:
\begin{itemize}
\item ./createmodel
\item ./configure
\item make
\item Mathematica interface (see models/<model>/start.m)
\end{itemize}

\subsection{Model files}

Explain how to write a FlexibleSUSY model file (starting from a SARAH
model file).

\subsection{Makefile modules}

Explain how to write a custom makefile module (module.mk) for a
FlexibleSUSY model.

### 2013-09-16 doc/chapters/usage.tex
add overview for the FlexibleSUSY model file variables

@@
 
+A FlexibleSUSY model file is a Mathematica file, where the following
+pieces are defined:
+
+\paragraph{General model information}
+
+\subparagraph{FSModelName}
+Name of the FlexibleSUSY model.
+Example (MSSM):
+\begin{lstlisting}[language=Mathematica]
+FSModelName = "MSSM";
+\end{lstlisting}
+
+\subparagraph{OnlyLowEnergyFlexibleSUSY} If set to True, creates a
+spectrum generator without high-scale constraint, i.e.\ only a
+low-scale and susy-scale constraint will be created (default: False).
+In this case all model parameters, except for gauge and Yukawa
+couplings are input at the susy scale.  This option is similar to
+\texttt{OnlyLowEnergySPheno} in SARAH/SPheno.
+\\
+Example:
+\begin{lstlisting}[language=Mathematica]
+OnlyLowEnergyFlexibleSUSY = False;  (* default *)
+\end{lstlisting}
+
+\paragraph{Input and output parameters}
+
+\subparagraph{MINPAR} This list of two-component lists defines model
+input parameters and their SLHA keys.  These parameters will be read
+from the MINPAR block in the SLHA file.
+\\
+Example (MSSM):
+\begin{lstlisting}[language=Mathematica]
+MINPAR = { {1, m0},
+           {2, m12},
+           {3, TanBeta},
+           {4, Sign[\[Mu]]},
+           {5, Azero}
+         };
+\end{lstlisting}
+
+\subparagraph{EXPAR} This list of two-component lists defines further
+model input parameters and their SLHA keys.  These parameters will be
+read from the EXTPAR block in the SLHA file.
+\\
+Example (E6SSM):
+\begin{lstlisting}[language=Mathematica]
+EXTPAR = { {61, LambdaInput},
+           {62, KappaInput},
+           {63, muPrimeInput},
+           {64, BmuPrimeInput},
+           {65, vSInput}
+         };
+\end{lstlisting}
+
+\subparagraph{DefaultParameterPoint} This list of two-component lists
+defines default values for the model input parameters.
+\\
+Example (MSSM):
+\begin{lstlisting}[language=Mathematica]
+DefaultParameterPoint = {
+    {m0, 125},
+    {m12, 500},
+    {TanBeta, 10},
+    {Sign[\[Mu]], 1},
+    {Azero, 0}
+};
+\end{lstlisting}
+
+\paragraph{Constraints}
+Currently three constraints are supported: low-scale, susy-scale and
+high-scale constraints.  In FlexibleSUSY they are named as
+\texttt{LowScale}, \texttt{SUSYScale} and \texttt{HighScale}.  For
+each constraint there is a scale definition (named after the
+constraint), an initial guess for the scale (concatenation of the
+constraint name and \texttt{FirstGuess}) and a list settings to be
+applied at the constraint (concatenation of the constraint name and
+\texttt{Input}).
+\\
+Example (MSSM):
+\begin{lstlisting}[language=Mathematica]
+(* susy-scale constraint *)
+SUSYScale = Sqrt[M[Su[1]]*M[Su[6]]];        (* scale definition *)
+SUSYScaleFirstGuess = Sqrt[m0^2 + 4 m12^2]; (* first scale guess *)
+SUSYScaleInput = {};                        (* nothing is set here *)
+
+(* high-scale constraint *)
+HighScale = g1 == g2;                       (* scale definition *)
+HighScaleFirstGuess = 1.0 10^14;            (* first scale guess *)
+HighScaleInput = {
+   {T[Ye], Azero*Ye},
+   {T[Yd], Azero*Yd},
+   {T[Yu], Azero*Yu},
+   {mHd2, m0^2},
+   {mHu2, m0^2},
+   {mq2, UNITMATRIX[3] m0^2},
+   {ml2, UNITMATRIX[3] m0^2},
+   {md2, UNITMATRIX[3] m0^2},
+   {mu2, UNITMATRIX[3] m0^2},
+   {me2, UNITMATRIX[3] m0^2},
+   {MassB, m12},
+   {MassWB,m12},
+   {MassG, m12}
+};
+
+LowScale = SM[MZ];                          (* scale definition *)
+LowScaleFirstGuess = SM[MZ];                (* first scale guess *)
+LowScaleInput = {
+   {vd, 2 MZDRbar / Sqrt[GUTNormalization[g1]^2 g1^2 + g2^2] Cos[ArcTan[TanBeta]]},
+   {vu, 2 MZDRbar / Sqrt[GUTNormalization[g1]^2 g1^2 + g2^2] Sin[ArcTan[TanBeta]]}
+};
+\end{lstlisting}
+
+\paragraph{Initial parameter guesses} At the \texttt{LowScale} and
+\texttt{HighScale} it is recommended to make an initial guess for the
+model parameters.  This can be done via
+\begin{lstlisting}[language=Mathematica]
+InitialGuessAtLowScale = {
+   {vd, SM[vev] Cos[ArcTan[TanBeta]]},
+   {vu, SM[vev] Sin[ArcTan[TanBeta]]}
+};
+
+InitialGuessAtHighScale = {
+   {\[Mu]   , 1.0},
+   {B[\[Mu]], 0.0}
+};
+\end{lstlisting}
+
 \subsection{Makefile modules}

### 2013-09-17 doc/chapters/usage.tex
add table with all model file variables

@@
 
+\begin{sidewaystable}[tb]
+  \centering
+  \begin{tabularx}{\textwidth}{>{\ttfamily}l>{\ttfamily}lX}
+    \toprule
+    variable    & default value & description \\
+    \midrule
+    FSModelName & Model`Name & Name of the FlexibleSUSY model \\
+    OnlyLowEnergyFlexibleSUSY & False & low-energy model \\
+    MINPAR & \{\} & list of input parameters in SLHA MINPAR block \\
+    EXTPAR & \{\} & list of input parameters in SLHA EXTPAR block \\
+    DefaultParameterPoint & \{\} & list of default values for the input parameters \\
+    LowScale & SM[MZ] & Standard Model matching scale (in GeV) \\
+    LowScaleFirstGuess & SM[MZ] & first guess for the Standard Model matching scale (in GeV) \\
+    LowScaleInput & \{\} & settings applied at the low scale \\
+    SUSYScale & 1000 & scale of supersymmetric particle masses (in GeV) \\
+    SUSYScaleFirstGuess & 1000 & first guess for the susy scale (in GeV) \\
+    SUSYScaleInput & \{\} & settings applied at the susy scale \\
+    HighScale & SARAH`hyperchargeCoupling == SARAH`leftCoupling
+      & unification scale (in GeV) \\
+    HighScaleFirstGuess & $10^{14}$ & first guess for the unification scale (in GeV) \\
+    HighScaleInput & \{\} & settings applied at the unification scale \\
+    \bottomrule
+  \end{tabularx}
+  \caption{FlexibleSUSY model file variables}
+  \label{tab:model-file-variables}
+\end{sidewaystable}
+
 \subsection{Makefile modules}

### 2013-09-25 doc/chapters/output.tex
Add class diagram for the two-scale models

@@
 
+\begin{figure}[tb]
+  \centering
+  \tikzumlset{fill class=white}
+  \begin{tikzpicture}
+    \umlclass[x=0, y=0, type=abstract]{Beta\_function}{}{
+      + \umlvirt{display() : const Eigen::ArrayXd} \\
+      + \umlvirt{set(v : const Eigen::ArrayXd\&) : void}\\
+      + \umlvirt{beta() : Eigen::ArrayXd}
+    }
+    \umlclass[x=8, y=0, type=abstract]{Two\_scale\_model}{}{
+      + \umlvirt{calculate\_spectrum() : void}\\
+      + \umlvirt{run\_to(scale : double, eps : double) : int}\\
+      + \umlvirt{set\_precision(precision : double) : void}
+    }
+    \umlclass[x=0, y=-4]{MSSM\_susy\_parameters}{}{
+      + display() : const Eigen::ArrayXd \\
+      + set(v : const Eigen::ArrayXd\&) : void\\
+      + beta() : Eigen::ArrayXd
+    }
+    \umlclass[x=0, y=-8]{MSSM\_soft\_parameters}{}{
+      + display() : const Eigen::ArrayXd \\
+      + set(v : const Eigen::ArrayXd\&) : void\\
+      + beta() : Eigen::ArrayXd
+    }
+    \umlclass[x=0, y=-12]{MSSM$<$Two\_scale$>$}{}{
+      + calculate\_spectrum() : void\\
+      + run\_to(scale : double, eps : double) : int\\
+      + set\_precision(precision : double) : void
+    }
+   \umlinherit{MSSM\_susy\_parameters}{Beta\_function}
+   \umlinherit{MSSM\_soft\_parameters}{MSSM\_susy\_parameters}
+   \umlinherit{MSSM$<$Two\_scale$>$}{MSSM\_soft\_parameters}
+   \umlinherit[geometry=-|]{MSSM$<$Two\_scale$>$}{Two\_scale\_model}
+  \end{tikzpicture}
+  \caption{Two-scale model class hierarchy}
+  \label{fig:two-scale-class-hierarchy}
+\end{figure}
+
 \subsection{Lattice algorithm}

### 2013-09-25 doc/chapters/output.tex
add private members to Beta_funcion and show public run_to()

@@
   \begin{tikzpicture}
-    \umlclass[x=0, y=0, type=abstract]{Beta\_function}{}{
+    \umlclass[x=0, y=0, type=abstract]{Beta\_function}{
+      -- scale : double\\
+      -- loops : unsigned\\
+      -- numPars : unsigned
+    }{
       + \umlvirt{display() : const Eigen::ArrayXd} \\
       + \umlvirt{set(v : const Eigen::ArrayXd\&) : void}\\
-      + \umlvirt{beta() : Eigen::ArrayXd}
+      + \umlvirt{beta() : Eigen::ArrayXd}\\
+      + run\_to(scale : double, eps : double) : int\\
     }

### 2013-09-25 doc/chapters/output.tex
add template specialization for the tow-scale model

@@
     }
-    \umlclass[x=8, y=0, type=abstract]{Two\_scale\_model}{}{
+    \umlclass[x=8, y=-12, template={T}]{MSSM}{}{}
+    \umlclass[x=8, y=-8, type=abstract]{Two\_scale\_model}{}{
       + \umlvirt{calculate\_spectrum() : void}\\
@@
    \umlinherit{MSSM$<$Two\_scale$>$}{MSSM\_soft\_parameters}
-   \umlinherit[geometry=-|]{MSSM$<$Two\_scale$>$}{Two\_scale\_model}
+   \umlinherit{MSSM$<$Two\_scale$>$}{Two\_scale\_model}
+   \umldep[arg1={$<<$bind$>>$}, mult1={T $\rightarrow$ Two\_scale}, pos1=0.5]{MSSM$<$Two\_scale$>$}{MSSM}
   \end{tikzpicture}

### 2013-09-25 doc/chapters/output.tex
start adding RGFlow class hierarchy

@@
   \caption{Two-scale model class hierarchy}
-  \label{fig:two-scale-class-hierarchy}
+  \label{fig:two-scale-model-class-hierarchy}
+\end{figure}
+
+\begin{figure}[tb]
+  \centering
+  \tikzumlset{fill class=white}
+  \begin{tikzpicture}
+    \umlclass[x=0, y=0, type=abstract]{Two\_scale\_model}{}{
+      + \umlvirt{calculate\_spectrum() : void}\\
+      + \umlvirt{run\_to(scale : double, eps : double) : int}\\
+      + \umlvirt{set\_precision(precision : double) : void}
+    }
+    \umlclass[x=0, y=-4]{RGFlow$<$Two\_scale$>$}{}{
+      + solve() : void
+    }
+    \umlclass[x=0, y=-8]{Constraint$<$Two\_scale$>$}{}{
+      + \umlvirt{apply() : void}\\
+      + \umlvirt{get\_scale() : double}\\
+      + \umlvirt{set\_model(model : Two\_scale\_model*) : void}
+    }
+    \umlclass[x=8, y=-4, template={T}]{RGFlow}{}{}
+    \umldep[arg1={$<<$bind$>>$}, mult1={T $\rightarrow$ Two\_scale}, pos1=0.5]{RGFlow$<$Two\_scale$>$}{RGFlow}
+    \umldep[mult1=1..*]{RGFlow$<$Two\_scale$>$}{Two\_scale\_model}
+    \umldep[mult1=1..*]{RGFlow$<$Two\_scale$>$}{Constraint$<$Two\_scale$>$}
+  \end{tikzpicture}
+  \caption{Two-scale renormalization group solver class hierarchy}
+  \label{fig:two-scale-rgflow-class-hierarchy}
 \end{figure}

### 2013-09-25 doc/chapters/output.tex
add initial guesser and convergence tester to the class diagram

@@
     }
-    \umlclass[x=0, y=-8]{Constraint$<$Two\_scale$>$}{}{
+    \umlclass[x=8, y=0]{Constraint$<$Two\_scale$>$}{}{
       + \umlvirt{apply() : void}\\
@@
     }
+    \umlclass[x=0, y=-7]{Initial\_guesser$<$Two\_scale$>$}{}{
+      + \umlvirt{guess() : void}
+    }
+    \umlclass[x=0, y=-11, template={T}]{Initial\_guesser}{}{}
+    \umlclass[x=8, y=-7]{Convergence\_tester$<$Two\_scale$>$}{}{
+      + \umlvirt{accuracy\_goal\_reached() : bool}\\
+      + \umlvirt{max\_iterations() : unsigned}
+    }
+    \umlclass[x=8, y=-11, template={T}]{Convergence\_tester}{}{}
     \umlclass[x=8, y=-4, template={T}]{RGFlow}{}{}
@@
     \umldep[mult1=1..*]{RGFlow$<$Two\_scale$>$}{Constraint$<$Two\_scale$>$}
+    \umldep{RGFlow$<$Two\_scale$>$}{Initial\_guesser$<$Two\_scale$>$}
+    \umldep{RGFlow$<$Two\_scale$>$}{Convergence\_tester$<$Two\_scale$>$}
+    \umldep[mult1={$<<$bind$>>$}, arg1={T $\rightarrow$ Two\_scale}, pos1=0.5]{Initial\_guesser$<$Two\_scale$>$}{Initial\_guesser}
+    \umldep[mult1={$<<$bind$>>$}, arg1={T $\rightarrow$ Two\_scale}, pos1=0.5]{Convergence\_tester$<$Two\_scale$>$}{Convergence\_tester}
   \end{tikzpicture}

### 2013-09-25 doc/chapters/output.tex
add constraint template class
briefly describe the two-scale algorithm

@@
 work and are plugged together.
+
+The two-scale algorithm works like this:
+%
+Initial guess:
+\begin{enumerate}
+\item Guess gauge couplings $g_{1,2,3}$ at $m_Z$
+\item Apply user-defined low-scale constraint (\code{LowScaleInput})
+\item Guess Yukawa couplings $y^{u,d,e}_{ij}$ at $m_Z$
+\item Run model to the high-scale (\code{HighScaleFirstGuess})
+\item Apply high-scale constraint (\code{HighScaleInput})
+\item Run model to the low-scale (\code{LowScaleFirstGuess})
+\item Solve EWSB eqs.\ at the tree-level
+\item Calculate \DRbar\ masses
+\end{enumerate}
+%
+Solving the renormalization group equations:
+\begin{enumerate}
+\item \label{rge-step-one} Run model to the low-scale (\code{LowScale})
+  \begin{enumerate}
+  \item Calculate \DRbar\ masses
+  \item Recalculate low-scale
+  \item Calculate \DRbar\ gauge and Yukawa couplings $g_{1,2,3}$, $y^{u,d,e}_{ij}$
+  \item Apply user-defined low-scale constraint (\code{LowScaleInput})
+  \end{enumerate}
+\item Run model to the high-scale (\code{HighScale})
+  \begin{enumerate}
+  \item Recalculate high-scale
+  \item Apply high-scale constraint (\code{HighScaleInput})
+  \end{enumerate}
+\item Run model to the susy-scale (\code{SUSYScale})
+  \begin{enumerate}
+  \item Calculate \DRbar\ masses
+  \item Recalculate low-scale
+  \item Apply susy-scale constraint (\code{SUSYScaleInput})
+  \item Solve EWSB eqs.\ at the one-loop level
+  \end{enumerate}
+\item If not converged yet, goto \ref{rge-step-one}
+\end{enumerate}
+%
+Calculating the particle spectrum:
+\begin{enumerate}
+\item Run model to the susy-scale
+\item Calculate the pole masses
+\item Run model to the parameter output scale (SLHA block MODSEL 12)
+\end{enumerate}
 
@@
     }
+    \umlclass[x=8, y=4, template={T}]{Constraint}{}{}
     \umlclass[x=0, y=-7]{Initial\_guesser$<$Two\_scale$>$}{}{
@@
     \umldep[mult1={$<<$bind$>>$}, arg1={T $\rightarrow$ Two\_scale}, pos1=0.5]{Convergence\_tester$<$Two\_scale$>$}{Convergence\_tester}
+    \umldep[arg1={$<<$bind$>>$}, mult1={T $\rightarrow$ Two\_scale}, pos1=0.5]{Constraint$<$Two\_scale$>$}{Constraint}
   \end{tikzpicture}

### 2013-09-25 meta/EWSB.m
typo

@@ -271,7 +271,7 @@ ReduceSolution[solution_List] :=
                  ];
                If[Length[flattenedSolution] > 2,
                   Print["Warning: cannot reduce solution for ", flattenedSolution];
-                  Prtin["   because there are more than two solutions"];
+                  Print["   because there are more than two solutions"];
                   Return[{{},{}}];
                  ];
                If[Length[flattenedSolution] == 1,

### 2013-09-25 meta/EWSB.m
Add user help message about manual EWSB solution setting in the model file

@@ -286,7 +286,7 @@ ReduceSolution[solution_List] :=
                      ];
               ];
            If[Length[reducedSolution] != Length[solution],
-              Print["Warning: analytic reduction of EWSB solutions failed"];
+              Print["Warning: analytic reduction of EWSB solutions failed."];
               Return[{{},{}}];
              ];
            Return[{reducedSolution, freePhases}];

### 2013-09-25 meta/FlexibleSUSY.m

@@ -824,6 +824,18 @@ MakeFlexibleSUSY[OptionsPattern[]] :=
               (* trying to find an analytic solution for the EWSB eqs. *)
               Print["Solving EWSB equations ..."];
               {ewsbSolution, freePhases} = EWSB`FindSolutionAndFreePhases[ewsbEquations, ParametersToSolveTadpoles];
+              If[ewsbSolution === {},
+                 Print["Warning: could not find an analytic solution to the EWSB eqs."];
+                 Print["   An iterative algorithm will be used.  You can try to set"];
+                 Print["   the solution by hand in the model file like this:"];
+                 Print[""];
+                 Print["TreeLevelEWSBSolution = {"];
+                 For[i = 1, i <= Length[ParametersToSolveTadpoles], i++,
+                     Print["   { ", ParametersToSolveTadpoles[[i]], ", ... }" <>
+                           If[i != Length[ParametersToSolveTadpoles], ",", ""]];
+                    ];
+                 Print["};\n"];
+                ];
               ,
               If[Length[FlexibleSUSY`TreeLevelEWSBSolution] != Length[ewsbEquations],
                  Print["Error: not enought EWSB solutions given!"];

### 2013-09-25 meta/EWSB.m
print out the full solution to help the user with TreeLevelEWSBSolution

@@ -296,6 +296,10 @@ FindSolutionAndFreePhases[equations_List, parametersFixedByEWSB_List] :=
     Module[{solution, reducedSolution, freePhases},
            solution = FindSolution[equations, parametersFixedByEWSB];
            {reducedSolution, freePhases} = ReduceSolution[solution];
+           If[reducedSolution === {},
+              Print["The full solution is:"];
+              Print[InputForm[solution]];
+             ];
            Return[{Flatten[reducedSolution], freePhases}];
           ];
 

### 2013-09-25 doc/chapters/usage.tex
explain the TreeLevelEWSBSolution variable
use code macro instead of explicite \textt

@@
 couplings are input at the susy scale.  This option is similar to
-\texttt{OnlyLowEnergySPheno} in SARAH/SPheno.
+\code{OnlyLowEnergySPheno} in SARAH/SPheno.
 \\
@@
 high-scale constraints.  In FlexibleSUSY they are named as
-\texttt{LowScale}, \texttt{SUSYScale} and \texttt{HighScale}.  For
+\code{LowScale}, \code{SUSYScale} and \code{HighScale}.  For
 each constraint there is a scale definition (named after the
 constraint), an initial guess for the scale (concatenation of the
-constraint name and \texttt{FirstGuess}) and a list settings to be
+constraint name and \code{FirstGuess}) and a list settings to be
 applied at the constraint (concatenation of the constraint name and
-\texttt{Input}).
+\code{Input}).
 \\
@@
 
-\paragraph{Initial parameter guesses} At the \texttt{LowScale} and
-\texttt{HighScale} it is recommended to make an initial guess for the
+\paragraph{Initial parameter guesses} At the \code{LowScale} and
+\code{HighScale} it is recommended to make an initial guess for the
 model parameters.  This can be done via
@@
    {B[\[Mu]], 0.0}
+};
+\end{lstlisting}
+
+\paragraph{Setting the tree-level EWSB eqs.\ solution by hand} In case
+the meta code cannot find an analytic solution to the tree-level EWSB
+eqs., one can set solutions by hand in the model file using the
+variable \code{TreeLevelEWSBSolution}.
+\\
+Example (MSSM):
+\begin{lstlisting}[language=Mathematica]
+TreeLevelEWSBSolution = {
+  {\[Mu]   , ... },
+  {B[\[Mu]], ... }
 };

### 2013-09-26 models/SoftsusyNMSSM/snmssm_two_scale.cpp
correct member return type of the SoftsusyNMSSM classes

@@ -42,11 +42,11 @@ void SNmssm<Two_scale>::calculate_spectrum()
    runto(displayMz());
 }
 
-int SNmssm<Two_scale>::run_to(double scale, double eps)
+void SNmssm<Two_scale>::run_to(double scale, double eps)
 {
    if (eps < 0.0)
       eps = precision;
-   return runto(scale, eps);
+   runto(scale, eps);
 }
 
 }

### 2013-09-26 models/SoftsusyNMSSM/snmssm_two_scale.hpp

@@ -35,7 +35,7 @@ public:
 
    virtual void calculate_spectrum();
    virtual std::string name() const { return "SNmssm"; }
-   virtual int run_to(double, double eps = -1.0);
+   virtual void run_to(double, double eps = -1.0);
    virtual void print(std::ostream& s) const { s << static_cast<NmssmSoftsusy>(*this); }
    virtual void set_precision(double p) { precision = p; }
 

### 2013-09-26 doc/chapters/output.tex
the run_to functions now return void

@@
       + \umlvirt{beta() : Eigen::ArrayXd}\\
-      + run\_to(scale : double, eps : double) : int\\
+      + run\_to(scale : double, eps : double) : void\\
     }
@@
       + \umlvirt{calculate\_spectrum() : void}\\
-      + \umlvirt{run\_to(scale : double, eps : double) : int}\\
+      + \umlvirt{run\_to(scale : double, eps : double) : void}\\
       + \umlvirt{set\_precision(precision : double) : void}
@@
       + calculate\_spectrum() : void\\
-      + run\_to(scale : double, eps : double) : int\\
+      + run\_to(scale : double, eps : double) : void\\
       + set\_precision(precision : double) : void

### 2013-09-28 doc/chapters/usage.tex
briefly explain the ./createmodel script

@@
 
+\subsubsection{\code{createmodel}}
+
+The \code{createmodel} script sets up a \flexisusy\ model.  This involves
+%
+\begin{itemize}
+\item a model directory \code{models/<flexiblesusy-model>/}
+\item a makefile module \code{models/<flexiblesusy-model>/module.mk}
+\item a \flexisusy\ model file \code{models/<flexiblesusy-model>/FlexibleSUSY.m}
+\item a Mathematica start script \code{models/<flexiblesusy-model>/start.m}
+\end{itemize}
+%
+Usage:
+\begin{lstlisting}[language=bash]
+./createmodel --models=<flexiblesusy-model>:<sarah-model>
+\end{lstlisting}
+Here \code{<flexiblesusy-model>} is the name of \flexisusy\ model
+to be created and \code{<sarah-model>} is the name of the \sarah\
+model file which defines the Lagrangian and the particles.\\
+Example:
+\begin{lstlisting}[language=bash]
+./createmodel --models=MyCMSSM:MSSM
+\end{lstlisting}
+%
+Multiple models can be created at once by separating the models by a
+comma.\\
+Example:
+\begin{lstlisting}[language=bash]
+./createmodel --models=MyCMSSM:MSSM,MyCNMSSM:NMSSM
+\end{lstlisting}
+%
+If a certain \sarah\ sub-model should be used, the name of the sub-model
+has to be appended with at preceeding slash.\\
+Example: use the \code{CKM} sub-model of the \code{MSSM}
+\begin{lstlisting}[language=bash]
+./createmodel --models=MyCMSSM:MSSM/CKM
+\end{lstlisting}
+%
+For further information and options see
+\begin{lstlisting}[language=bash]
+./createmodel --help
+\end{lstlisting}
+
 \subsection{Model files}
 
-Explain how to write a FlexibleSUSY model file (starting from a SARAH
+Explain how to write a \flexisusy\ model file (starting from a \sarah\
 model file).
 
-A FlexibleSUSY model file is a Mathematica file, where the following
+A \flexisusy\ model file is a Mathematica file, where the following
 pieces are defined:
@@
 \subparagraph{FSModelName}
-Name of the FlexibleSUSY model.
+Name of the \flexisusy\ model.
 Example (MSSM):
@@
 Currently three constraints are supported: low-scale, susy-scale and
-high-scale constraints.  In FlexibleSUSY they are named as
+high-scale constraints.  In \flexisusy\ they are named as
 \code{LowScale}, \code{SUSYScale} and \code{HighScale}.  For
@@
     \midrule
-    FSModelName & Model`Name & Name of the FlexibleSUSY model \\
+    FSModelName & Model`Name & Name of the \flexisusy\ model \\
     OnlyLowEnergyFlexibleSUSY & False & low-energy model \\
@@
   \end{tabularx}
-  \caption{FlexibleSUSY model file variables}
+  \caption{\flexisusy\ model file variables}
   \label{tab:model-file-variables}
@@
 Explain how to write a custom makefile module (module.mk) for a
-FlexibleSUSY model.
+\flexisusy\ model.

### 2013-10-21 doc/chapters/overview.tex
describe design goals and make requirements more precise

@@
 
-Explain:
-\begin{itemize}
-\item What is FlexibleSUSY?
-\item Features, design goals
-\end{itemize}
+FlexibleSUSY provides Mathematica and C++ code to create spectrum
+generators for non-minimal supersymmetric models.
+
+\subsection{Design goals}
+
+FlexibleSUSY is designed with the following points in mind:
+
+\paragraph{Speed}
+Exploring the parameter space of supersymmetric models with a high
+number of free parameters is quite time consuming.  Therefore
+FlexibleSUSY aims to produce spectrum generators with a short runtime.
+
+The two most time consuming parts of a spectrum generator are the
+calculation of the beta functions and the pole masses for mixed
+particles.
+
+The reason is the following: The RG solving algorithms usually need
+$O(10)$ iterations to calculate a particle spectrum with $0.1\%$
+precision.  During each iteration the Runge--Kutta algorithm needs to
+calculate all beta functions $O(50)$ times.  Most two-loop beta
+functions involve $O(50)$ matrix multiplications and additions.  All
+together one arrives at $O(2.5\cdot 10^4)$ matrix multiplications.  To
+optimize these matrix multiplications FlexibleSUSY uses the fast
+linear algebra package \href{Eigen}{http://eigen.tuxfamily.org}, which
+exploits C++ expression templates to make the matrix multiplications
+easy to write and optimize.
+
+The second most time consuming part is the precise calculation of the
+pole masses for mixed particles.  For each particle $\psi_i$ the full
+self-energy matrix $\Sigma^\psi_{ij}(p=m^\text{tree}_{\psi_i})$ has to
+be evaluated.  Each self-energy matrix entry again involves the
+evaluation of $O(50)$ Feynman diagrams, which involves the calculation
+of vertices and loop functions.  All in all, one arrives at $O(500)$
+Feynman diagam or $O(3\cdot 10^4)$ loop function evaluations.  To
+speed up the calculation of the pole masses FlexibleSUSY makes use of
+multi-threading, where each pole mass is calculated in a separate
+thread.  By this technique one can gain a speed-up of $20$--$30\%$.
+
+\paragraph{Modularity}
+The large variety of supersymmetric models makes it likely that the
+user needs to modify the generated spectrum generator source code.
+FlexibleSUSY uses object orientation with C++ to modularize the source
+code to make it easy for the user to modify and extend the spectrum
+generator.  An important application here are the constraints: The RG
+solvers provide a constraint interface,
+\figurename~\ref{fig:schematic-two-scale-constraint-interface},
+%
+\begin{figure}
+  \centering
+  \tikzumlset{fill class=white}
+  \begin{tikzpicture}
+    \umlclass[x=0, y=0]{RGFlow$<$Two\_scale$>$}{}{
+      + solve()
+    }
+    \umlclass[x=6, y=0, type=abstract]{Constraint$<$Two\_scale$>$}{}{
+      + \umlvirt{apply()}\\
+      + \umlvirt{get\_scale() : double}
+    }
+    \umlclass[x=6, y=-3]{MyConstraint}{}{
+      + apply()\\
+      + get\_scale() : double
+    }
+    \umldep{RGFlow$<$Two\_scale$>$}{Constraint$<$Two\_scale$>$}
+    \umlinherit{MyConstraint}{Constraint$<$Two\_scale$>$}
+  \end{tikzpicture}
+  \caption{Schematic two-scale constraint interface}
+  \label{fig:schematic-two-scale-constraint-interface}
+\end{figure}
+%
+which can be implemented to create the custom constraint
+\code{MyConstraint}.  In the \code{apply()} function the constraint
+application takes place and the \code{get_scale()} function specifies
+the scale at which the constraint is to be applied.
 
@@
 \begin{itemize}
-\item Mathematica 7-9
-\item SARAH version 4
-\item a C++11 compatible compiler (g++ 4.7.2 or higher, clang++ 3.1 or
+\item Mathematica, version 7 or higher
+\item SARAH, version 4.0.0 or higher \url{http://sarah.hepforge.org}
+\item C++11 compatible compiler (g++ 4.6.0 or higher, clang++ 3.1 or
   higher)
-\item a fortran compiler (gfortran etc.)
-\item Eigen library version 3
-\item Boost library
-\item GNU scientific library
+\item Fortran compiler (gfortran, ifort etc.)
+\item Eigen library, version 3.1 \url{http://eigen.tuxfamily.org}
+\item Boost library, version 1.36.0 or higher
+  \url{http://www.boost.org}
+\item GNU scientific library \url{http://www.gnu.org/software/gsl/}
 \item lapack (needed for the Lattice algorithm only)
+  \url{http://www.netlib.org/lapack/}
 \end{itemize}
+%
+Optional:
+%
+\begin{itemize}
+\item Looptools, version 2.8 or higher
+  \url{http://www.feynarts.de/looptools/}
+\end{itemize}

### 2013-10-23 doc/chapters/usage.tex
add section about the configure scritp

@@
 ./createmodel --help
+\end{lstlisting}
+
+\subsubsection{\code{configure}}
+
+The \code{configure} script checks for the installed compilers,
+libraries and Mathematica.  If all of these exists in a sufficent
+version, the \code{Makefile} is created, which contains the information
+on how to compile the code.  The user has to specify the models which
+should be included in the build via the \code{--with-models=} option,
+e.g.
+%
+\begin{lstlisting}[language=bash]
+./configure --with-models=<flexibsusy-model>
+\end{lstlisting}
+%
+Here \code{<flexibsusy-model>} is either \code{all} or a comma
+separated list of \flexisusy models.
+
+Furthermore, the user can select which RG solver algorithms to use via
+the \code{--with-algorithms=} option.\\
+Example:
+%
+\begin{lstlisting}[language=bash]
+./configure --with-algorithms=two_scale
+\end{lstlisting}
+
+The \code{configure} script further allows to select the C++ and
+Fortran compilers, the Mathematica kernel command as well as paths to
+libraries and header files.  See
+\code{configure --help} for all available options.\\
+Example:
+%
+\begin{lstlisting}[language=bash]
+./configure --with-models=MSSM --with-cxx=clang++
+   --with-boost-incdir=/usr/include/
+   --with-boost-libdir=/usr/lib/
 \end{lstlisting}

### 2013-10-28 doc/paper.tex
Added skeleton of paper.  very rough stuff. probably all of it will be changed.

\documentclass[final,3p,times,pdflatex]{elsarticle}
\usepackage[utf8x]{inputenc}      % input font encoding
\usepackage{amsmath,amssymb}
\usepackage[T1]{fontenc}          % output font encoding
\usepackage{booktabs,tabularx}
\usepackage{rotating}             % for sidewaystable
\usepackage{xspace}
\usepackage[usenames]{xcolor}
\usepackage{tikz,tikz-uml}
\usepackage{listings}   
\bibstyle{elsarticle-num}
% source code highlighting
\lstset{breaklines=true,
  breakatwhitespace=true,
  stepnumber=1,
  basicstyle=\ttfamily\footnotesize,
  commentstyle=\ttfamily\color{gray},
  prebreak={\textbackslash},
  breakindent=10pt,
  breakautoindent=false,
  showspaces=false,
  showstringspaces=false,
  frame=single}
\usepackage[pdftitle={FlexibleSUSY --- A spectrum generator generator for supersymmetric models},
pdfauthor={Peter Athron,Jae-hyeon Park,Dominik Stockinger,Alexander Voigt},
pdfkeywords={FlexibleSUSY,supersymmetry,spectrum,generator,MSSM,NMSSM,E6SSM},
bookmarks=true, linktocpage]{hyperref}

%macros
\newcommand{\sarah}{SARAH\xspace}
\newcommand{\fs}{FlexibleSUSY\xspace}

\newcommand{\code}[1]{\lstinline|#1|}  % inline source code
\newcommand{\textoverline}[1]{$\overline{\mbox{#1}}$}
\newcommand{\DRbar}{\textoverline{DR}}  % inline source code

\journal{Computer Physics Communications}
\begin{document}
\begin{frontmatter}

 \title{\Large\bf FlexibleSUSY --- A spectrum generator generator for supersymmetric models}

\author{Peter Athron}
\address{ARC Centre of Excellence for Particle Physics at 
the Tera-scale, School of Chemistry and Physics, University of Adelaide, 
Adelaide SA 5005 Australia}
\author{Jae-hyeon Park}
\author{Dominik St\"ockinger}
\author{Alexander Voigt}
\address{Institut f\"ur Kern- und Teilchenphysik,
TU Dresden, Zellescher Weg 19, 01069 Dresden, Germany}

  \begin{abstract}
   Here we introduce \fs, a MATHEMATICA package which generates a C++
   spectrum generator for any SUSY model, created with modularity and
   speed in mind.  \fs makes use of the existing \sarah package to
   obtain self energies, tadpoles corrections and renormalisation
   group equations in any model.  \fs then generates C++ code for
   calculating spectra with a structure that is designed to be
   flexible. The user specifies the model supplying the superpotential,
   gauge structure and particle content in a \sarah model file. The
   user should also specify the boundary conditions in a separate \fs
   model file. {\color{red}( Do we want to keep this separate and in
     different places, add a script to put a SARAH model file in the
     right place or combine them?)}  Flexible SUSY then translates
   these into C++ code along with numerical routines for solving the
   renormalisation group equations, one loop electroweak symmetry
   breaking conditions and a solver routine for simultaneously solving
   for the spectrum.  The modular structure of the generated code
   allows for any individual component to be replaced with an
   alternative if available.  Indeed \fs has been intentionally
   designed to grow as alternative solvers and calculators are added.
  \end{abstract}
\begin{keyword}
sparticle, 
supersymmetry, 
Higgs
\PACS 12.60.Jv
\PACS 14.80.Ly
\end{keyword}
\end{frontmatter}

\section{Program Summary}
\noindent{\em Program title:} \fs\\ {\em Program obtainable from:}
         {\tt http://flexiblesusy.hepforge.org/}\\ {\em Distribution
           format:}\/ tar.gz\\ {\em Programming language:} {\tt
           C++}\\ {\em Computer:}\/ Personal computer\\ {\em Operating
           system:}\/ Tested on Linux 4.x\\ {\em Word size:}\/ 32
         bits\\ {\em External routines:}\/ SARAH 4.0.0, Boost library,
         lapack {\color{red} Not sure what counts here?}\\ {\em
           Typical running time:}\/ 0.1-0.3 seconds per parameter
         point.\\ {\em Nature of problem:}\/ Determining the mass
         spectrum and mixings for any supersymmetric model. The
         generated code must find simultaneous solutions to
         constraints which are specified at two or more different
         renormalisation scales, which are connected by
         renormalisation group equations forming a large set of
         coupled differential equations. \\ {\em Solution method:}\/
         Nested iterative algorithm and numerical minimisation of the
         Higgs potential.  \\ {\em Restrictions:} The couplings must
         remain perturbative at all scales between the highest and
         lowest boundary condition.  \fs~ assumes that all couplings
         of the model are real (i.e.\ $CP-$conserving)?  Due to the
         modular nature of the generated code adaption and extension
         to overcome restrictions in scope is quite straight forward.

\newpage
\section{Introduction}
Supersymmetry provides the only way to extend the space-time symmetries of the poincar\'e group\cite{}, leading many to suspect that it may be realised in nature in some form. In particular supersymmetric extensions of the standard model where supersymmetry is broken at the TeV scale have been proposed to solve the hierarchy problem\cite{}, allow gauge coupling unification\cite{} and predict a dark matter candidate\cite{} and can fit the observed relic density\cite{}.  Such models have also been used for baryogenesis\cite{} or leptogensis\cite{} to solve the matter-anti-matter asymmetry of the universe and have been considered as the low energy effective models originating from string theory\cite{}.

Detailed phenonenological studies have been carried out for scenarios with the minimal supersymmetric standard model (MSSM). Many computational tools have been created to aid these studies and to explore the model. In particular there are a number of public spectrum generators for MSSM\cite{} and one for the NMSSM\cite{} which allow one to find the sparticle spectrum and mixings for a particular choices of breaking mechanism inspired boundary conditions and specified parameters.

However none of the fundamental motivations of supersymmetry require minimality. Yet without minimality as a constraint, constructing specialised tools to study all relevant models would require an enormous amount of work.  So general tools which can automate this process and produce fats and reliable programs can greatly enhance our ability to understand and test non-minimal realisations of supersymmetry.

From the recent $7$ TeV and $8$ TeV runs at the Large Hadron Collider (LHC) there have been two important developments.  First low energy signatures expected from such models, such as the classic jets plus missing transverse energy signature, have not been observed, substantially raising the lower limit on sparticle masses\cite{}. No other signature of beyond the standard model(BSM) physics has been observed, leaving the fundamental questions which motivated BSM physics unanswered.

 Secondly ATLAS and CMS discovered\cite{} a light Higgs of $125$ GeV, within the mass range that could be accommodated in the MSSM but requiring stops which are significantly heavier than both the direct collider limits and indirect limits that appears in constrained models from the significantly higher limits on first and second generation squarks.  Alternatively it is possible that a $125$ GeV Higgs in non-minimal models may not require such heavy stops, indeed a number of such models have been constructed to do exactly this by having new tree level contributions to the Higgs mass.

These results lead to a number of different perspectives, for example some may wish to abandon naturalness as a motivation and focus on BSM models which can fit the observed relic density and allow unification of gauge couplings, while others may focus on special models which preserve naturalness.  In both cases exploration of the models can be aided if it is possible to quickly create a spectrum generator to find if there are phenomenologically viable scenarios in the models they are interested and what the general features of such models are.

\subsection{Design goals}

FlexibleSUSY is designed with the following points in mind:

\paragraph{Speed}
Exploring the parameter space of supersymmetric models with a high
number of free parameters is quite time consuming.  Therefore
FlexibleSUSY aims to produce spectrum generators with a short run-time.

The two most time consuming parts of a spectrum generator are the
calculation of the beta functions and the pole masses for mixed
particles.

The reason is the following: The RG solving algorithms usually need
$O(10)$ iterations to calculate a particle spectrum with $0.1\%$
precision.  During each iteration the Runge--Kutta algorithm needs to
calculate all beta functions $O(50)$ times.  Most two-loop beta
functions involve $O(50)$ matrix multiplications and additions.  All
together one arrives at $O(2.5\cdot 10^4)$ matrix multiplications.  To
optimize these matrix multiplications FlexibleSUSY uses the fast
linear algebra package \href{Eigen}{http://eigen.tuxfamily.org}, which
exploits C++ expression templates to make the matrix multiplications
easy to write and optimize.

The second most time consuming part is the precise calculation of the
pole masses for mixed particles.  For each particle $\psi_i$ the full
self-energy matrix $\Sigma^\psi_{ij}(p=m^\text{tree}_{\psi_i})$ has to
be evaluated.  Each self-energy matrix entry again involves the
evaluation of $O(50)$ Feynman diagrams, which involves the calculation
of vertices and loop functions.  All in all, one arrives at $O(500)$
Feynman diagram or $O(3\cdot 10^4)$ loop function evaluations.  To
speed up the calculation of the pole masses FlexibleSUSY makes use of
multi-threading, where each pole mass is calculated in a separate
thread.  By this technique one can gain a speed-up of $20$--$30\%$.

\paragraph{Modularity}
The large variety of supersymmetric models makes it likely that the
user needs to modify the generated spectrum generator source code.
FlexibleSUSY uses object orientation with C++ to modularize the source
code to make it easy for the user to modify and extend the spectrum
generator.  An important application here are the constraints: The RG
solvers provide a constraint interface,
\figurename~\ref{fig:schematic-two-scale-constraint-interface},
%
\begin{figure}
  \centering
  \tikzumlset{fill class=white}
  \begin{tikzpicture}
    \umlclass[x=0, y=0]{RGFlow$<$Two\_scale$>$}{}{
      + solve()
    }
    \umlclass[x=6, y=0, type=abstract]{Constraint$<$Two\_scale$>$}{}{
      + \umlvirt{apply()}\\
      + \umlvirt{get\_scale() : double}
    }
    \umlclass[x=6, y=-3]{MyConstraint}{}{
      + apply()\\
      + get\_scale() : double
    }
    \umldep{RGFlow$<$Two\_scale$>$}{Constraint$<$Two\_scale$>$}
    \umlinherit{MyConstraint}{Constraint$<$Two\_scale$>$}
  \end{tikzpicture}
  \caption{Schematic two-scale constraint interface}
  \label{fig:schematic-two-scale-constraint-interface}
\end{figure}
%
which can be implemented to create the custom constraint
\code{MyConstraint}.  In the \code{apply()} function the constraint
application takes place and the \code{get_scale()} function specifies
the scale at which the constraint is to be applied.

\section{Download and compilation}
\subsection{Quickstart}

\section{Stucture of the spectrum generator}
\subsection{boundary conditions}
Different types, how we match to low energy data.  reference EWSB or merge that section here.
\subsection{RGEs}
\subsection{tree level spectrum}
\subsection{Electroweak symmetry breaking}
\subsection{pole masses}
\subsection{two scale fixed point iteration}
\subsection{Lattice method}
If ready.
\section{Implementing new models}
\section{Adapting the code at the C++ level}
\appendix
\section{Examples}
\section{Tests}

\begin{thebibliography}{100}
%\cite{Allanach:2001kg}
\bibitem{Allanach:2001kg} 
  B.~C.~Allanach,
  %``SOFTSUSY: a program for calculating supersymmetric spectra,''
  Comput.\ Phys.\ Commun.\  {\bf 143}, 305 (2002)
  [hep-ph/0104145].
  %%CITATION = HEP-PH/0104145;%%
  %716 citations counted in INSPIRE as of 20 Sep 2013

%\cite{Degrassi:2009yq}
\bibitem{Degrassi:2009yq} 
  G.~Degrassi and P.~Slavich,
  %``On the radiative corrections to the neutral Higgs boson masses in the NMSSM,''
  Nucl.\ Phys.\ B {\bf 825}, 119 (2010)
  [arXiv:0907.4682 [hep-ph]].
  %%CITATION = ARXIV:0907.4682;%%
  %35 citations counted in INSPIRE as of 21 Sep 2013
%\cite{Allanach:2008qq}
\bibitem{Allanach:2008qq} 
  B.~C.~Allanach, C.~Balazs, G.~Belanger, M.~Bernhardt, F.~Boudjema, D.~Choudhury, K.~Desch and U.~Ellwanger {\it et al.},
  %``SUSY Les Houches Accord 2,''
  Comput.\ Phys.\ Commun.\  {\bf 180}, 8 (2009)
  [arXiv:0801.0045 [hep-ph]].
  %%CITATION = ARXIV:0801.0045;%%
  %177 citations counted in INSPIRE as of 21 Sep 2013

%\cite{Ellwanger:2009dp}
\bibitem{Ellwanger:2009dp} 
  U.~Ellwanger, C.~Hugonie and A.~M.~Teixeira,
  %``The Next-to-Minimal Supersymmetric Standard Model,''
  Phys.\ Rept.\  {\bf 496}, 1 (2010)
  [arXiv:0910.1785 [hep-ph]].
  %%CITATION = ARXIV:0910.1785;%%
  %327 citations counted in INSPIRE as of 21 Sep 2013

%\cite{MV94}
\bibitem{MV94} 
  S.~P.~Martin and M.~T.~Vaughn,
  %``Two loop renormalization group equations for soft supersymmetry breaking couplings,''
  Phys.\ Rev.\ D {\bf 50}, 2282 (1994)
  [Erratum-ibid.\ D {\bf 78}, 039903 (2008)]
  [hep-ph/9311340].
  %%CITATION = HEP-PH/9311340;%%
  %568 citations counted in INSPIRE as of 24 Sep 2013

%\cite{Yam94}
\bibitem{Yam94} 
  Y.~Yamada,
  %``Two loop renormalization group equations for soft SUSY breaking scalar interactions: Supergraph method,''
  Phys.\ Rev.\ D {\bf 50}, 3537 (1994)
  [hep-ph/9401241].
  %%CITATION = HEP-PH/9401241;%%
  %225 citations counted in INSPIRE as of 24 Sep 2013

\cite{Sper13}
\bibitem{Sper13} 
  M.~Sperling, D.~Stöckinger and A.~Voigt,
  %``Renormalization of vacuum expectation values in spontaneously broken gauge theories,''
  JHEP {\bf 1307}, 132 (2013)
  [arXiv:1305.1548 [hep-ph]].
  %%CITATION = ARXIV:1305.1548;%%
  %4 citations counted in INSPIRE as of 14 Oct 2013

%\cite{Ell08}
\bibitem{Ell08} 
  U.~Ellwanger, C.~-C.~Jean-Louis and A.~M.~Teixeira,
  %``Phenomenology of the General NMSSM with Gauge Mediated Supersymmetry Breaking,''
  JHEP {\bf 0805}, 044 (2008)
  [arXiv:0803.2962 [hep-ph]].
  %%CITATION = ARXIV:0803.2962;%%
  %28 citations counted in INSPIRE as of 24 Sep 2013

  %\cite{Staub:2010ty}
\bibitem{Staub:2010ty} 
  F.~Staub, W.~Porod and B.~Herrmann,
  %``The Electroweak sector of the NMSSM at the one-loop level,''
  JHEP {\bf 1010}, 040 (2010)
  [arXiv:1007.4049 [hep-ph]].
  %%CITATION = ARXIV:1007.4049;%%
  %28 citations counted in INSPIRE as of 12 Oct 2013

%\cite{Staub:2009bi}
\bibitem{Staub:2009bi} 
  F.~Staub,
  %``From Superpotential to Model Files for FeynArts and CalcHep/CompHep,''
  Comput.\ Phys.\ Commun.\  {\bf 181}, 1077 (2010)
  [arXiv:0909.2863 [hep-ph]].
  %%CITATION = ARXIV:0909.2863;%%
  %64 citations counted in INSPIRE as of 12 Oct 2013

%\cite{Staub:2010jh}
\bibitem{Staub:2010jh} 
  F.~Staub,
  %``Automatic Calculation of supersymmetric Renormalization Group Equations and Self Energies,''
  Comput.\ Phys.\ Commun.\  {\bf 182}, 808 (2011)
  [arXiv:1002.0840 [hep-ph]].
  %%CITATION = ARXIV:1002.0840;%%
  %60 citations counted in INSPIRE as of 12 Oct 2013

%\cite{Staub:2012pb}
\bibitem{Staub:2012pb} 
  F.~Staub,
  %``SARAH 3.2: Dirac Gauginos, UFO output, and more,''
  Computer Physics Communications {\bf 184}, pp. 1792 (2013)
  [Comput.\ Phys.\ Commun.\  {\bf 184}, 1792 (2013)]
  [arXiv:1207.0906 [hep-ph]].
  %%CITATION = ARXIV:1207.0906;%%
  %19 citations counted in INSPIRE as of 12 Oct 2013

%\cite{Staub:2013tta}
\bibitem{Staub:2013tta} 
  F.~Staub,
  %``SARAH 4: A tool for (not only SUSY) model builders,''
  arXiv:1309.7223 [hep-ph].
  %%CITATION = ARXIV:1309.7223;%%
  %2 citations counted in INSPIRE as of 12 Oct 2013

%

\end{thebibliography}
\end{document}

### 2013-10-28 meta/FlexibleSUSY.m
remove the softSusyCompatibleRGEs flag
Softsusy and SARAH both have now the g^4 terms of the anomalous dim.
of the Higgs fields implemented.  However, SARAH misses the g^4
terms from the anomalous dim. from the background field.  Therefore
I've disabled them in Softsusy for now.

@@ -14,7 +14,6 @@ MakeFlexibleSUSY::usage="";
 LowPrecision::usage="";
 MediumPrecision::usage="";
 HighPrecision::usage="";
-softSusyCompatibleRGEs::usage="";
 GUTNormalization::usage="Returns GUT normalization of a given coupling";
 
 FSModelName;
@@ -727,7 +726,6 @@ Options[MakeFlexibleSUSY] :=
     {
         Eigenstates -> SARAH`EWSB,
         InputFile -> "FlexibleSUSY.m",
-        softSusyCompatibleRGEs -> True,
         DefaultDiagonalizationPrecision -> MediumPrecision,
         HighDiagonalizationPrecision -> {},
         MediumDiagonalizationPrecision -> {},
@@ -771,16 +769,6 @@ MakeFlexibleSUSY[OptionsPattern[]] :=
            SARAH`Xi = 1;
            SARAH`Xip = 1;
            SARAH`rMS = 0;
-
-           If[OptionValue[softSusyCompatibleRGEs] === True && Model`Name == "MSSM",
-              Print["Generating SoftSusy compatible beta function ..."];
-              Print["Note: SoftSusy is missing g^4 two-loop contributions to the VEVs, I'm disabling them."];
-              SARAH`BetaVEV = SARAH`BetaVEV /.
-              { SARAH`g1^4 -> 0,
-                Susyno`LieGroups`g2^4 -> 0,
-                SARAH`g3^4 -> 0,
-                SARAH`g1^2 Susyno`LieGroups`g2^2 -> 0 };
-             ];
 
            FlexibleSUSY`InputParameters = Join[(#[[2]])& /@ SARAH`MINPAR, (#[[2]])& /@ SARAH`EXTPAR];
            Parameters`SetInputParameters[FlexibleSUSY`InputParameters];

### 2013-10-28 models/SoftsusyMSSM/susy.cpp

@@ -282,7 +282,7 @@ MssmSusy MssmSusy::beta(sBrevity & a) const {
 
   const static double twolp = 4.010149318236068e-5; // 1/(16 pi^2)^2
   if (displayLoops() > 1) {
-    const double g4terms = 0.; // 1.035 * g4(1) + 0.45 * gsq(1) * gsq(2) + 5.875 * g4(2);
+    const double g4terms = 1.035 * g4(1) + 0.45 * gsq(1) * gsq(2) + 5.875 * g4(2);
     sH1H1 = sH1H1 + twolp * 
       (-(3.0 * (e2 * e2).trace() + 9.0 * (d2t * d2t).trace() + 3.0 * t) + 
        (16 * gsq(3) - 0.4 * gsq(1)) * ddT + 1.2 * gsq(1) * eeT  + g4terms);
@@ -303,6 +303,7 @@ MssmSusy MssmSusy::beta(sBrevity & a) const {
     dt = dt + displayTanb() * twolp * (3.0 * ddT + eeT - 3.0 * uuT) * feynman;
     dHvev = dHvev - hVev * twolp * (cosb2 * (3.0 * ddT + eeT) +
 				    sinb2 * 3.0 * uuT) * feynman;
+       // + hVev * twolp * 4.5 * g4(2);
   }
   // Contains all susy derivatives:
   MssmSusy ds(du, dd, de, dg, dmu, dt, displayMu(), displayLoops(),

### 2013-10-28 models/SoftsusyNMSSM/nmssmsusy.cpp

@@ -229,7 +229,8 @@ NmssmSusy NmssmSusy::beta(nmsBrevity & a) const {
     /// Two-loop pieces
     dt = dt + displayTanb() * twolp * (3.0 * ddT + eeT - 3.0 * uuT) * feynman;
     dHvev = dHvev - displayHvev() * twolp * (cosb2 * (3.0 * ddT + eeT) +
-				    sinb2 * 3.0 * uuT + lsq) * feynman;
+                                             sinb2 * 3.0 * uuT + lsq) * feynman;
+    // + displayHvev() * twolp * 4.5 * g4(2);
     dSvev = dSvev + sVev * twolp * (8 * k4 + 8 * ksq * lsq
             + 2 * lsq * (-2 * feynman + 2 * lsq + 3 * ddT + eeT + 3 * uuT));
   }

### 2013-10-28 doc/chapters/overview.tex

@@
 number of free parameters is quite time consuming.  Therefore
-FlexibleSUSY aims to produce spectrum generators with a short runtime.
+FlexibleSUSY aims to produce spectrum generators with a short run-time.
 
@@
 of vertices and loop functions.  All in all, one arrives at $O(500)$
-Feynman diagam or $O(3\cdot 10^4)$ loop function evaluations.  To
+Feynman diagram or $O(3\cdot 10^4)$ loop function evaluations.  To
 speed up the calculation of the pole masses FlexibleSUSY makes use of

### 2013-11-03 doc/paper.tex
Added additional subsection.
added download location, Quick start and requirements (just c+p from manual).

@@
 \section{Download and compilation}
-\subsection{Quickstart}
+Flexible SUSY can be downloaded from \url{http://flexiblesusy.hepforge.org}.  
+To use FlexibleSUSY please make sure you have the following installed.  
+\begin{itemize}
+\item Mathematica, version 7 or higher
+\item SARAH, version 4.0.0 or higher \url{http://sarah.hepforge.org}
+\item C++11 compatible compiler (g++ 4.6.0 or higher, clang++ 3.1 or
+  higher)
+\item Fortran compiler (gfortran, ifort etc.)
+\item Eigen library, version 3.1 \url{http://eigen.tuxfamily.org}
+\item Boost library, version 1.36.0 or higher
+  \url{http://www.boost.org}
+\item GNU scientific library \url{http://www.gnu.org/software/gsl/}
+\item lapack (needed for the Lattice algorithm only)
+  \url{http://www.netlib.org/lapack/}
+\end{itemize}
+%
+Optional:
+%
+\begin{itemize}
+\item Looptools, version 2.8 or higher
+  \url{http://www.feynarts.de/looptools/}
+\end{itemize}
+
+\subsection{Quick Start}
+Create a MSSM spectrum generator
+  \begin{lstlisting}[language=bash]
+$ ./createmodel --models=MSSM
+$ ./configure --with-models=MSSM
+$ make
+  \end{lstlisting}
+Run the spectrum generator using the default parameter point
+  \begin{lstlisting}[language=bash]
+$ ./models/MSSM/run_MSSM.x
+  \end{lstlisting}
+Run the spectrum generator using a SLHA input file
+  \begin{lstlisting}[language=bash]
+$ ./models/MSSM/run_MSSM.x --slha-input-file=templates/MSSM/LesHouches.in.MSSM --slha-output-file=LesHouches.out.MSSM
+  \end{lstlisting}
 
@@
 If ready.
+\subsection{multiple effective models at different scales}
 \section{Implementing new models}

### 2013-11-03 doc/paper.tex
added more to introduction.

@@
 \newcommand{\fs}{FlexibleSUSY\xspace}
-
+\newcommand{\ESSM}{E$_6$SSM\xspace}
 \newcommand{\code}[1]{\lstinline|#1|}  % inline source code
@@
 \section{Introduction}
-Supersymmetry provides the only way to extend the space-time symmetries of the poincar\'e group\cite{}, leading many to suspect that it may be realised in nature in some form.
-In particular supersymmetric extensions of the standard model where supersymmetry is broken at the TeV scale have been proposed to solve the hierarchy problem\cite{}, allow gauge coupling unification\cite{} and predict a dark matter candidate\cite{} and can fit the observed relic density\cite{}.
-Such models have also been used for baryogenesis\cite{} or leptogensis\cite{} to solve the matter-anti-matter asymmetry of the universe and have been considered as the low energy effective models originating from string theory\cite{}.
+Supersymmetry provides the only way to extend the space-time
+symmetries of the poincar\'e group\cite{}, leading many to suspect
+that it may be realised in nature in some form. In particular
+supersymmetric extensions of the standard model where supersymmetry is
+broken at the TeV scale have been proposed to solve the hierarchy
+problem\cite{}, allow gauge coupling unification\cite{} and predict a
+dark matter candidate\cite{} and can fit the observed relic
+density\cite{}.  Such models have also been used for
+baryogenesis\cite{} or leptogensis\cite{} to solve the
+matter-anti-matter asymmetry of the universe and have been considered
+as the low energy effective models originating from string
+theory\cite{}.
 
-Detailed phenonenological studies have been carried out for scenarios with the minimal supersymmetric standard model (MSSM).
-Many computational tools have been created to aid these studies and to explore the model.
-In particular there are a number of public spectrum generators for MSSM\cite{} and one for the NMSSM\cite{} which allow one to find the sparticle spectrum and mixings for a particular choices of breaking mechanism inspired boundary conditions and specified parameters.
+Detailed phenonenological studies have been carried out for scenarios
+with the minimal supersymmetric standard model (MSSM). Many
+computational tools have been created to aid these studies and to
+explore the model. In particular there are a number of public spectrum
+generators for MSSM\cite{} and one for the NMSSM\cite{} which allow
+one to find the sparticle spectrum and mixings for a particular
+choices of breaking mechanism inspired boundary conditions and
+specified parameters.
 
-However none of the fundamental motivations of supersymmetry require minimality.
-Yet without minimality as a constraint, constructing specialised tools to study all relevant models would require an enormous amount of work.
-So general tools which can automate this process and produce fats and reliable programs can greatly enhance our ability to understand and test non-minimal realisations of supersymmetry.
+However none of the fundamental motivations of supersymmetry require
+minimality. Yet without minimality as a constraint, constructing
+specialised tools to study all relevant models would require an
+enormous amount of work.  So general tools which can automate this
+process and produce fats and reliable programs can greatly enhance our
+ability to understand and test non-minimal realisations of
+supersymmetry.
 
-From the recent $7$ TeV and $8$ TeV runs at the Large Hadron Collider (LHC) there have been two important developments.
-First low energy signatures expected from such models, such as the classic jets plus missing transverse energy signature, have not been observed, substantially raising the lower limit on sparticle masses\cite{}.
-No other signature of beyond the standard model(BSM) physics has been observed, leaving the fundamental questions which motivated BSM physics unanswered.
+From the recent $7$ TeV and $8$ TeV runs at the Large Hadron Collider
+(LHC) there have been two important developments.  First low energy
+signatures expected from such models, such as the classic jets plus
+missing transverse energy signature, have not been observed,
+substantially raising the lower limit on sparticle masses\cite{}. No
+other signature of beyond the standard model(BSM) physics has been
+observed, leaving the fundamental questions which motivated BSM
+physics unanswered.
 
- Secondly ATLAS and CMS discovered\cite{} a light Higgs of $125$ GeV, within the mass range that could be accommodated in the MSSM but requiring stops which are significantly heavier than both the direct collider limits and indirect limits that appears in constrained models from the significantly higher limits on first and second generation squarks.
-Alternatively it is possible that a $125$ GeV Higgs in non-minimal models may not require such heavy stops, indeed a number of such models have been constructed to do exactly this by having new tree level contributions to the Higgs mass.
+ Secondly ATLAS and CMS discovered\cite{} a light Higgs of $125$ GeV,
+ within the mass range that could be accommodated in the MSSM but
+ requiring stops which are significantly heavier than both the direct
+ collider limits and indirect limits that appears in constrained
+ models from the significantly higher limits on first and second
+ generation squarks.  Alternatively it is possible that a $125$ GeV
+ Higgs in non-minimal models may not require such heavy stops, indeed
+ a number of such models have been constructed to do exactly this by
+ having new tree level contributions to the Higgs mass.
 
-These results lead to a number of different perspectives, for example some may wish to abandon naturalness as a motivation and focus on BSM models which can fit the observed relic density and allow unification of gauge couplings, while others may focus on special models which preserve naturalness.
-In both cases exploration of the models can be aided if it is possible to quickly create a spectrum generator to find if there are phenomenologically viable scenarios in the models they are interested and what the general features of such models are.
+These results lead to a number of different perspectives, for example
+some may wish to abandon naturalness as a motivation and focus on BSM
+models which can fit the observed relic density and allow unification
+of gauge couplings, while others may focus on special models which
+preserve naturalness.  In both cases exploration of the models can be
+aided if it is possible to quickly create a spectrum generator to find
+if there are phenomenologically viable scenarios in the models they
+are interested and what the general features of such models are.
+
+\subsection{The Program}
+\fs is a MATHEMATICA package designed to create a fast and easily
+adaptable spectrum generator in C++ for any SUSY model and some
+non-supersymmetric models {\color{red} can we?}.  \fs makes use of the
+existing \sarah package to obtain one-loop self energies, one-loop
+tadpoles corrections and two-loop renormalisation group
+equations(RGEs) in any model.  \fs then converts these algerbraic
+expressions from \sarah into C++ code using the C++11 standard. \fs
+uses some parts of SOFTSUSY\cite{Allanach:2001kg}, the very fast Eigen
+library \cite{eigen}, as well as the gnu scientific library and the
+boost library to create numerical routines which then use these
+routines to solve the RGEs and EWSB conditions which will be used in
+the spectrum generator. The main solver routine of \fs then calls
+these routines to find simultaneous solutions for the EWSB, low energy
+data, and other, user supplied, boundary conditions (e.g. the usual
+mSUGRA universality conditions) and calculates the mass spectrum.
+
+ The modular structure of the generated code allows for any individual
+ components to be replaced with an alternative if available.  Indeed
+ \fs has been intentionally designed to grow as alternative solvers
+ and calculators are added and a subsequent release with an
+ alternative lattice solver is already planned.{\color{red} Or Indeed
+   flexible SUSY is distributed with two diffrent main solvers, the
+   two-scale solver based omn the approach of \cite{} and designed for
+   speed and the a lattice solver designed to solve tricky cases where
+   the fixed point iteration of the two-scale approach fails to
+   converge.}
 
@@
 \section{Examples}
+\subsection{NMSSM}
+Detailed steps to create NMSSM model file.
+\subsection{\ESSM}
+Detailed steps to create \ESSM model file.
+
 \section{Tests}
@@
 
+\bibitem{eigen}
+Eigen library, version 3.1 \url{http://eigen.tuxfamily.org}
+
 %

### 2013-11-03 doc/paper.tex
fleshed out more of the structure.

@@
 \newcommand{\DRbar}{\textoverline{DR}}  % inline source code
+\def\at{\alpha_t}
+\def\ab{\alpha_b}
+\def\as{\alpha_s}
+\def\atau{\alpha_{\tau}}
+\def\oat{O(\at)}
+\def\oab{O(\ab)}
+\def\oatau{O(\atau)}
+\def\oatab{O(\at\ab)}
+\def\oatas{O(\at\as)}
+\def\oabas{O(\ab\as)}
+\def\oatababq{O(\at\ab + \ab^2)}
+\def\oatqatababq{O(\at^2 + \at\ab + \ab^2)}
+\def\oatasatq{O(\at\as + \at^2)}
+\def\oatasabas{O(\at\as +\ab\as)}
+\def\oatasabasatq{O(\at\as + \at^2 +\ab\as)}
+\def\oatq{O(\at^2)}
+\def\oabq{O(\ab^2)}
+\def\oatauq{O(\atau^2)}
+\def\oabatau{O(\ab \atau)}
+\def\oas{O(\as)}
+\def\oatauqatab{O(\atau^2 +\ab \atau )}
 
@@
  alternative lattice solver is already planned.{\color{red} Or Indeed
-   flexible SUSY is distributed with two diffrent main solvers, the
-   two-scale solver based omn the approach of \cite{} and designed for
+   flexible SUSY is distributed with two different main solvers, the
+   two-scale solver based on the approach of \cite{} and designed for
    speed and the a lattice solver designed to solve tricky cases where
@@
 $ ./models/MSSM/run_MSSM.x --slha-input-file=templates/MSSM/LesHouches.in.MSSM --slha-output-file=LesHouches.out.MSSM
-  \end{lstlisting}
+  \end{lstlisting}%% $
 
@@
 Different types, how we match to low energy data.  reference EWSB or merge that section here.
+mention two loop qcd parts from softsusy.
 \subsection{RGEs}
+The two loop renormalisation group equations are generated using command ?
+Even in the simplest supersymmetric model the RGEs form a very large set of coupled non-linear differential equations.
+To solve these and thereby evolve the parameters between two different scales \fs has...
 \subsection{tree level spectrum}
+The diagonalised tree level $\overline{DR}$ masses are obtained from expressions...
+The numerical diagonalisation is performed by... with an accuracy set by...   
 \subsection{Electroweak symmetry breaking}
+At tree level the EWSB conditions are written...
+The one loop electroweak symmetry breaking are solved using the gsl routines finding multiple roots, \code{gsl_multiroot_fsolver}.  Iteration. For example in the MSSM \fs generates.. show code.
+ If higher accuracy is required additional routines with higher order corrections can be added in a very simple way.
+For example in the MSSM by default \fs adds two loop Higgs FORTRAN routines supplied by P.~Slavich from \cite{slavich-degrassi} to add two loop corrections of $\oatas$, $\oabas$, $\oatq$, $\oabatau$,
+$\oabq$, $\oatauq$ and $\oatab$  as follows...  
+Something similar is done for the NMSSM, but in this case the NMSSM $\oatas$, $\oabas$ pieces come from \cite{Degrassi:2009yq}.
+ Corrections from other sources can be added to the generated code by...
+\subsection{two scale fixed point iteration}
 \subsection{pole masses}
-\subsection{two scale fixed point iteration}
+With the solver routine finished and convergence achieved, the $\overline{DR}$ parameters consistent with one loop EWSB, low energy data and all user supplied boundary conditions are then known at any scale between...
+
+The physical mass spectrum can now be calculated.
+\fs uses the full one-loop self energies and tree level routines to calculate the pole masses at the one-level.
+If higher accuracy is required additional routines with higher order corrections can be added in a very simple way.
+For example in the MSSM by default \fs adds two loop Higgs FORTRAN routines supplied by P.~Slavich from \cite{slavich-degrassi} to add two loop corrections of $\oatas$, $\oabas$, $\oatq$, $\oabatau$,
+$\oabq$, $\oatauq$ and $\oatab$  as follows...  
+Something similar is done for the NMSSM, but in this case the NMSSM $\oatas$, $\oabas$ pieces come from \cite{Degrassi:2009yq}, while for $\oatq$, $\oabatau$, 
+$\oabq$, $\oatauq$ and $\oatab$ the MSSM pieces are used though it should be understood that these are not complete in the NMSSM.
+For other models since the Higgs mass is a very important measurement and the two loop corrections can be larger than the current experimental error \cite{Degrassi:2009yq} we recommend that the leading log two loop corrections are estimated, by generalising those of the MSSM or NMSSM, as has been done, for example, in the E$_6$SSM\cite{King:2005jy}.
+
+Corrections to other masses or further corrections to the Higgs states can be added very easily by...
+
 \subsection{Lattice method}
@@
 \section{Tests}
+Comparisons against existing spectrum generators to test for deviations, e.g. the tests against MSSM and NMSSM (and E6SSM?).
+
+Comparisons against existing spectrum generators and SARAH generated spheno code for speed.
 
@@
 
+%\cite{King:2005jy}
+\bibitem{King:2005jy} 
+  S.~F.~King, S.~Moretti and R.~Nevzorov,
+  %``Theory and phenomenology of an exceptional supersymmetric standard model,''
+  Phys.\ Rev.\ D {\bf 73}, 035009 (2006)
+  [hep-ph/0510419].
+  %%CITATION = HEP-PH/0510419;%%
+  %130 citations counted in INSPIRE as of 03 Nov 2013
+
 %

### 2013-11-03 doc/paper.tex
added more on modularity utility.

@@
    the fixed point iteration of the two-scale approach fails to
-   converge.}
+   converge.}  In addition if one already has an existing code for
+ some other purpose and simply wishes to improve upon this by adding,
+ e.g, new RGEs or self energies the modular fraework makes this fairly
+ straightward{\color{red} Is this true?  because it should be...}
 
@@
 \section{Adapting the code at the C++ level}
+\section{Using individual modules in your own code}
 \appendix

### 2013-11-03 doc/paper.tex
added a couple of references.

@@
 explore the model. In particular there are a number of public spectrum
-generators for MSSM\cite{} and one for the NMSSM\cite{} which allow
+generators for MSSM\cite{Allanach:2001kg,Porod:2003um,Djouadi:2002ze,Baer:1993ae} and two for the NMSSM\cite{Ellwanger:2006rn,Allanach:nmssm} which allow
 one to find the sparticle spectrum and mixings for a particular
@@
   %716 citations counted in INSPIRE as of 20 Sep 2013
+%\cite{Porod:2003um}
+\bibitem{Porod:2003um} 
+  W.~Porod,
+  %``SPheno, a program for calculating supersymmetric spectra, SUSY particle decays and SUSY particle production at e+ e- colliders,''
+  Comput.\ Phys.\ Commun.\  {\bf 153}, 275 (2003)
+  [hep-ph/0301101].
+  %%CITATION = HEP-PH/0301101;%%
+  %413 citations counted in INSPIRE as of 03 Nov 2013
+
+%\cite{Djouadi:2002ze}
+\bibitem{Djouadi:2002ze} 
+  A.~Djouadi, J.~-L.~Kneur and G.~Moultaka,
+  %``SuSpect: A Fortran code for the supersymmetric and Higgs particle spectrum in the MSSM,''
+  Comput.\ Phys.\ Commun.\  {\bf 176}, 426 (2007)
+  [hep-ph/0211331].
+  %%CITATION = HEP-PH/0211331;%%
+  %654 citations counted in INSPIRE as of 03 Nov 2013
+
+%\cite{Baer:1993ae}
+\bibitem{Baer:1993ae} 
+  H.~Baer, F.~E.~Paige, S.~D.~Protopopescu and X.~Tata,
+  %``Simulating Supersymmetry with ISAJET 7.0 / ISASUSY 1.0,''
+  hep-ph/9305342.
+  %%CITATION = HEP-PH/9305342;%%
+  %82 citations counted in INSPIRE as of 03 Nov 2013
+%\cite{Ellwanger:2006rn}
+\bibitem{Ellwanger:2006rn} 
+  U.~Ellwanger and C.~Hugonie,
+  %``NMSPEC: A Fortran code for the sparticle and Higgs masses in the NMSSM with GUT scale boundary conditions,''
+  Comput.\ Phys.\ Commun.\  {\bf 177}, 399 (2007)
+  [hep-ph/0612134].
+  %%CITATION = HEP-PH/0612134;%%
+  %81 citations counted in INSPIRE as of 03 Nov 2013
 

### 2013-11-03 doc/paper.tex
more references and copied over details on two-scale solver.

@@
    spectrum generator for any SUSY model, created with modularity and
-   speed in mind.  \fs makes use of the existing \sarah package to
-   obtain self energies, tadpoles corrections and renormalisation
+   speed in mind.  \fs makes use of the existing \sarah package
+   to obtain self energies, tadpoles corrections and renormalisation
    group equations in any model.  \fs then generates C++ code for
    calculating spectra with a structure that is designed to be
-   flexible. The user specifies the model supplying the superpotential,
-   gauge structure and particle content in a \sarah model file. The
-   user should also specify the boundary conditions in a separate \fs
-   model file. {\color{red}( Do we want to keep this separate and in
-     different places, add a script to put a SARAH model file in the
-     right place or combine them?)}  Flexible SUSY then translates
-   these into C++ code along with numerical routines for solving the
-   renormalisation group equations, one loop electroweak symmetry
-   breaking conditions and a solver routine for simultaneously solving
-   for the spectrum.  The modular structure of the generated code
-   allows for any individual component to be replaced with an
-   alternative if available.  Indeed \fs has been intentionally
-   designed to grow as alternative solvers and calculators are added.
+   flexible. The user specifies the model supplying the
+   superpotential, gauge structure and particle content in a \sarah
+   model file. The user should also specify the boundary conditions in
+   a separate \fs model file. {\color{red}( Do we want to keep this
+     separate and in different places, add a script to put a \sarah
+     model file in the right place or combine them?)}  Flexible SUSY
+   then translates these into C++ code along with numerical routines
+   for solving the renormalisation group equations, one loop
+   electroweak symmetry breaking conditions and a solver routine for
+   simultaneously solving for the spectrum.  The modular structure of
+   the generated code allows for any individual component to be
+   replaced with an alternative if available.  Indeed \fs has been
+   intentionally designed to grow as alternative solvers and
+   calculators are added.
   \end{abstract}
@@
 non-supersymmetric models {\color{red} can we?}.  \fs makes use of the
-existing \sarah package to obtain one-loop self energies, one-loop
+existing \sarah package \cite{Staub:2010ty,Staub:2009bi,Staub:2010jh,Staub:2012pb,Staub:2013tta} to obtain one-loop self energies, one-loop
 tadpoles corrections and two-loop renormalisation group
@@
 \subsection{two scale fixed point iteration}
+The default solver which calls the above routines above and finds the full set of $\overline{DR}$ masses in the model consistent with all constraints is the two-scale solver.
+This can be selected by...
+but if no selection is made this will be used anyway as default.
+
+This approach is decribed in \cite{Barger:1993gh} for the MSSM and is widely implemented in SUSY spectrum generators.
+
+The stucture of the two-scale algorithm used in \fs is as follows,
+%
+Initial guess:
+\begin{enumerate}
+\item Guess gauge couplings $g_{1,2,3}$ at $m_Z$
+\item Apply user-defined low-scale constraint (\code{LowScaleInput})
+\item Guess Yukawa couplings $y^{u,d,e}_{ij}$ at $m_Z$
+\item Run model to the high-scale (\code{HighScaleFirstGuess})
+\item Apply high-scale constraint (\code{HighScaleInput})
+\item Run model to the low-scale (\code{LowScaleFirstGuess})
+\item Solve EWSB eqs.\ at the tree-level
+\item Calculate \DRbar\ masses
+\end{enumerate}
+%
+Solving the renormalization group equations:
+\begin{enumerate}
+\item \label{rge-step-one} Run model to the low-scale (\code{LowScale})
+  \begin{enumerate}
+  \item Calculate \DRbar\ masses
+  \item Recalculate low-scale
+  \item Calculate \DRbar\ gauge and Yukawa couplings $g_{1,2,3}$, $y^{u,d,e}_{ij}$
+  \item Apply user-defined low-scale constraint (\code{LowScaleInput})
+  \end{enumerate}
+\item Run model to the high-scale (\code{HighScale})
+  \begin{enumerate}
+  \item Recalculate high-scale
+  \item Apply high-scale constraint (\code{HighScaleInput})
+  \end{enumerate}
+\item Run model to the susy-scale (\code{SUSYScale})
+  \begin{enumerate}
+  \item Calculate \DRbar\ masses
+  \item Recalculate low-scale
+  \item Apply susy-scale constraint (\code{SUSYScaleInput})
+  \item Solve EWSB eqs.\ at the one-loop level
+  \end{enumerate}
+\item If not converged yet, goto \ref{rge-step-one}
+\end{enumerate}
+%
+Calculating the particle spectrum:
+\begin{enumerate}
+\item Run model to the susy-scale
+\item Calculate the pole masses
+\item Run model to the parameter output scale (SLHA block MODSEL 12)
+\end{enumerate}
+
+\begin{figure}[tb]
+  \centering
+  \tikzumlset{fill class=white}
+  \begin{tikzpicture}
+    \umlclass[x=0, y=0, type=abstract]{Beta\_function}{
+      -- scale : double\\
+      -- loops : unsigned\\
+      -- numPars : unsigned
+    }{
+      + \umlvirt{display() : const Eigen::ArrayXd} \\
+      + \umlvirt{set(v : const Eigen::ArrayXd\&) : void}\\
+      + \umlvirt{beta() : Eigen::ArrayXd}\\
+      + run\_to(scale : double, eps : double) : void\\
+    }
+    \umlclass[x=8, y=-12, template={T}]{MSSM}{}{}
+    \umlclass[x=8, y=-8, type=abstract]{Two\_scale\_model}{}{
+      + \umlvirt{calculate\_spectrum() : void}\\
+      + \umlvirt{run\_to(scale : double, eps : double) : void}\\
+      + \umlvirt{set\_precision(precision : double) : void}
+    }
+    \umlclass[x=0, y=-4]{MSSM\_susy\_parameters}{}{
+      + display() : const Eigen::ArrayXd \\
+      + set(v : const Eigen::ArrayXd\&) : void\\
+      + beta() : Eigen::ArrayXd
+    }
+    \umlclass[x=0, y=-8]{MSSM\_soft\_parameters}{}{
+      + display() : const Eigen::ArrayXd \\
+      + set(v : const Eigen::ArrayXd\&) : void\\
+      + beta() : Eigen::ArrayXd
+    }
+    \umlclass[x=0, y=-12]{MSSM$<$Two\_scale$>$}{}{
+      + calculate\_spectrum() : void\\
+      + run\_to(scale : double, eps : double) : void\\
+      + set\_precision(precision : double) : void
+    }
+   \umlinherit{MSSM\_susy\_parameters}{Beta\_function}
+   \umlinherit{MSSM\_soft\_parameters}{MSSM\_susy\_parameters}
+   \umlinherit{MSSM$<$Two\_scale$>$}{MSSM\_soft\_parameters}
+   \umlinherit{MSSM$<$Two\_scale$>$}{Two\_scale\_model}
+   \umldep[arg1={$<<$bind$>>$}, mult1={T $\rightarrow$ Two\_scale}, pos1=0.5]{MSSM$<$Two\_scale$>$}{MSSM}
+  \end{tikzpicture}
+  \caption{Two-scale model class hierarchy}
+  \label{fig:two-scale-model-class-hierarchy}
+\end{figure}
+
+\begin{figure}[tb]
+  \centering
+  \tikzumlset{fill class=white}
+  \begin{tikzpicture}
+    \umlclass[x=0, y=0, type=abstract]{Two\_scale\_model}{}{
+      + \umlvirt{calculate\_spectrum() : void}\\
+      + \umlvirt{run\_to(scale : double, eps : double) : void}\\
+      + \umlvirt{set\_precision(precision : double) : void}
+    }
+    \umlclass[x=0, y=-4]{RGFlow$<$Two\_scale$>$}{}{
+      + solve() : void
+    }
+    \umlclass[x=8, y=0]{Constraint$<$Two\_scale$>$}{}{
+      + \umlvirt{apply() : void}\\
+      + \umlvirt{get\_scale() : double}\\
+      + \umlvirt{set\_model(model : Two\_scale\_model*) : void}
+    }
+    \umlclass[x=8, y=4, template={T}]{Constraint}{}{}
+    \umlclass[x=0, y=-7]{Initial\_guesser$<$Two\_scale$>$}{}{
+      + \umlvirt{guess() : void}
+    }
+    \umlclass[x=0, y=-11, template={T}]{Initial\_guesser}{}{}
+    \umlclass[x=8, y=-7]{Convergence\_tester$<$Two\_scale$>$}{}{
+      + \umlvirt{accuracy\_goal\_reached() : bool}\\
+      + \umlvirt{max\_iterations() : unsigned}
+    }
+    \umlclass[x=8, y=-11, template={T}]{Convergence\_tester}{}{}
+    \umlclass[x=8, y=-4, template={T}]{RGFlow}{}{}
+    \umldep[arg1={$<<$bind$>>$}, mult1={T $\rightarrow$ Two\_scale}, pos1=0.5]{RGFlow$<$Two\_scale$>$}{RGFlow}
+    \umldep[mult1=1..*]{RGFlow$<$Two\_scale$>$}{Two\_scale\_model}
+    \umldep[mult1=1..*]{RGFlow$<$Two\_scale$>$}{Constraint$<$Two\_scale$>$}
+    \umldep{RGFlow$<$Two\_scale$>$}{Initial\_guesser$<$Two\_scale$>$}
+    \umldep{RGFlow$<$Two\_scale$>$}{Convergence\_tester$<$Two\_scale$>$}
+    \umldep[mult1={$<<$bind$>>$}, arg1={T $\rightarrow$ Two\_scale}, pos1=0.5]{Initial\_guesser$<$Two\_scale$>$}{Initial\_guesser}
+    \umldep[mult1={$<<$bind$>>$}, arg1={T $\rightarrow$ Two\_scale}, pos1=0.5]{Convergence\_tester$<$Two\_scale$>$}{Convergence\_tester}
+    \umldep[arg1={$<<$bind$>>$}, mult1={T $\rightarrow$ Two\_scale}, pos1=0.5]{Constraint$<$Two\_scale$>$}{Constraint}
+  \end{tikzpicture}
+  \caption{Two-scale renormalization group solver class hierarchy}
+  \label{fig:two-scale-rgflow-class-hierarchy}
+\end{figure}
+{\color{red} copied and pasted from latest manual -- not sure how best to maintain this, could split maual up into tiny bits, but we may also want some differences?}
 \subsection{pole masses}
@@
 
+  %\cite{Staub:2010ty}
+\bibitem{Staub:2010ty} 
+  F.~Staub, W.~Porod and B.~Herrmann,
+  %``The Electroweak sector of the NMSSM at the one-loop level,''
+  JHEP {\bf 1010}, 040 (2010)
+  [arXiv:1007.4049 [hep-ph]].
+  %%CITATION = ARXIV:1007.4049;%%
+  %28 citations counted in INSPIRE as of 12 Oct 2013
+
+%\cite{Staub:2009bi}
+\bibitem{Staub:2009bi} 
+  F.~Staub,
+  %``From Superpotential to Model Files for FeynArts and CalcHep/CompHep,''
+  Comput.\ Phys.\ Commun.\  {\bf 181}, 1077 (2010)
+  [arXiv:0909.2863 [hep-ph]].
+  %%CITATION = ARXIV:0909.2863;%%
+  %64 citations counted in INSPIRE as of 12 Oct 2013
+
+%\cite{Staub:2010jh}
+\bibitem{Staub:2010jh} 
+  F.~Staub,
+  %``Automatic Calculation of supersymmetric Renormalization Group Equations and Self Energies,''
+  Comput.\ Phys.\ Commun.\  {\bf 182}, 808 (2011)
+  [arXiv:1002.0840 [hep-ph]].
+  %%CITATION = ARXIV:1002.0840;%%
+  %60 citations counted in INSPIRE as of 12 Oct 2013
+
+%\cite{Staub:2012pb}
+\bibitem{Staub:2012pb} 
+  F.~Staub,
+  %``SARAH 3.2: Dirac Gauginos, UFO output, and more,''
+  Computer Physics Communications {\bf 184}, pp. 1792 (2013)
+  [Comput.\ Phys.\ Commun.\  {\bf 184}, 1792 (2013)]
+  [arXiv:1207.0906 [hep-ph]].
+  %%CITATION = ARXIV:1207.0906;%%
+  %19 citations counted in INSPIRE as of 12 Oct 2013
+
+%\cite{Staub:2013tta}
+\bibitem{Staub:2013tta} 
+  F.~Staub,
+  %``SARAH 4: A tool for (not only SUSY) model builders,''
+  arXiv:1309.7223 [hep-ph].
+  %%CITATION = ARXIV:1309.7223;%%
+  %2 citations counted in INSPIRE as of 12 Oct 2013
+
 %\cite{Degrassi:2009yq}
@@
 
-  %\cite{Staub:2010ty}
-\bibitem{Staub:2010ty} 
-  F.~Staub, W.~Porod and B.~Herrmann,
-  %``The Electroweak sector of the NMSSM at the one-loop level,''
-  JHEP {\bf 1010}, 040 (2010)
-  [arXiv:1007.4049 [hep-ph]].
-  %%CITATION = ARXIV:1007.4049;%%
-  %28 citations counted in INSPIRE as of 12 Oct 2013
-
-%\cite{Staub:2009bi}
-\bibitem{Staub:2009bi} 
-  F.~Staub,
-  %``From Superpotential to Model Files for FeynArts and CalcHep/CompHep,''
-  Comput.\ Phys.\ Commun.\  {\bf 181}, 1077 (2010)
-  [arXiv:0909.2863 [hep-ph]].
-  %%CITATION = ARXIV:0909.2863;%%
-  %64 citations counted in INSPIRE as of 12 Oct 2013
-
-%\cite{Staub:2010jh}
-\bibitem{Staub:2010jh} 
-  F.~Staub,
-  %``Automatic Calculation of supersymmetric Renormalization Group Equations and Self Energies,''
-  Comput.\ Phys.\ Commun.\  {\bf 182}, 808 (2011)
-  [arXiv:1002.0840 [hep-ph]].
-  %%CITATION = ARXIV:1002.0840;%%
-  %60 citations counted in INSPIRE as of 12 Oct 2013
-
-%\cite{Staub:2012pb}
-\bibitem{Staub:2012pb} 
-  F.~Staub,
-  %``SARAH 3.2: Dirac Gauginos, UFO output, and more,''
-  Computer Physics Communications {\bf 184}, pp. 1792 (2013)
-  [Comput.\ Phys.\ Commun.\  {\bf 184}, 1792 (2013)]
-  [arXiv:1207.0906 [hep-ph]].
-  %%CITATION = ARXIV:1207.0906;%%
-  %19 citations counted in INSPIRE as of 12 Oct 2013
-
-%\cite{Staub:2013tta}
-\bibitem{Staub:2013tta} 
-  F.~Staub,
-  %``SARAH 4: A tool for (not only SUSY) model builders,''
-  arXiv:1309.7223 [hep-ph].
-  %%CITATION = ARXIV:1309.7223;%%
-  %2 citations counted in INSPIRE as of 12 Oct 2013
+\bibitem{Barger:1993gh} 
+  V.~D.~Barger, M.~S.~Berger and P.~Ohmann,
+  %``The Supersymmetric particle spectrum,''
+  Phys.\ Rev.\ D {\bf 49}, 4908 (1994)
+  [hep-ph/9311269].
+  %%CITATION = HEP-PH/9311269;%%
+  %365 citations counted in INSPIRE as of 03 Nov 2013
 

### 2013-11-10 test/test_MSSM_model.cpp
correct index access in MSSM model test

@@ -137,19 +137,19 @@ void compare_tree_level_masses(MssmSoftsusy s, MSSM<Two_scale> m)
    // neutral CP even Higgs
    DoubleVector hh(ToDoubleVector(m.get_Mhh()));
    TEST_EQUALITY(hh, s.displayDrBarPars().mh0);
+   TEST_CLOSE(m.get_ZH()(0,0), -sin(alpha), 1.0e-12);
+   TEST_CLOSE(m.get_ZH()(0,1), cos(alpha), 1.0e-12);
+   TEST_CLOSE(m.get_ZH()(1,0), -cos(alpha), 1.0e-12);
    TEST_CLOSE(m.get_ZH()(1,1), -sin(alpha), 1.0e-12);
-   TEST_CLOSE(m.get_ZH()(1,2), cos(alpha), 1.0e-12);
-   TEST_CLOSE(m.get_ZH()(2,1), -cos(alpha), 1.0e-12);
-   TEST_CLOSE(m.get_ZH()(2,2), -sin(alpha), 1.0e-12);
 
    // neutral CP odd Higgs
    DoubleVector Ah(ToDoubleVector(m.get_MAh()));
    TEST_CLOSE(Ah(1), s.displayMzRun(), 1.0e-11);
    TEST_EQUALITY(Ah(2), s.displayDrBarPars().mA0(1));
+   TEST_CLOSE(m.get_ZA()(0,0), -cos(beta), 1.0e-12);
+   TEST_CLOSE(m.get_ZA()(0,1), sin(beta), 1.0e-12);
+   TEST_CLOSE(m.get_ZA()(1,0), -sin(beta), 1.0e-12);
    TEST_CLOSE(m.get_ZA()(1,1), -cos(beta), 1.0e-12);
-   TEST_CLOSE(m.get_ZA()(1,2), sin(beta), 1.0e-12);
-   TEST_CLOSE(m.get_ZA()(2,1), -sin(beta), 1.0e-12);
-   TEST_CLOSE(m.get_ZA()(2,2), -cos(beta), 1.0e-12);
 
    // charged Higgs
    DoubleVector Hpm(ToDoubleVector(m.get_MHpm()));
@@ -293,14 +293,14 @@ void compare_tree_level_masses(MssmSoftsusy s, MSSM<Two_scale> m)
    TEST_EQUALITY(m.get_MGlu(), s.displayDrBarPars().mGluino);
 
    // neutrinos
[…]

### 2013-11-10 test/test_MSSM_low_scale_constraint.cpp
correct index access in MSSM low-scale constraint

@@ -119,9 +119,9 @@ BOOST_AUTO_TEST_CASE( test_low_energy_constraint )
    const double ss_MZ = Sqrt(Sqr(MZ) + pizzt);
    const double ss_new_vev = s.getVev();
 
-   const double fs_mt = m.calculate_MFu_DRbar_1loop(oneset.displayPoleMt(), 3);
-   const double fs_mb = m.calculate_MFd_DRbar_1loop(oneset.displayMass(mBottom), 3);
-   const double fs_me = m.calculate_MFe_DRbar_1loop(oneset.displayMass(mTau), 3);
+   const double fs_mt = m.calculate_MFu_DRbar_1loop(oneset.displayPoleMt(), 2);
+   const double fs_mb = m.calculate_MFd_DRbar_1loop(oneset.displayMass(mBottom), 2);
+   const double fs_me = m.calculate_MFe_DRbar_1loop(oneset.displayMass(mTau), 2);
    const double fs_MZ = m.calculate_MVZ_DRbar_1loop(Electroweak_constants::MZ);
    const double fs_old_vd = m.get_vd();
    const double fs_old_vu = m.get_vu();

### 2013-11-10 doc/paper.tex
added some instructions for nmssm model files in sarah. Needs work and flexibleSUSY parts to be added.

@@
 Detailed steps to create NMSSM model file.
+\begin{enumerate}
+\item Create new directory \code{mkdir NMSSM}
+\item Move into this new directory \code{cd NMSSM}
+\item copy the MSSM model file in \code{cp /path/to/SARAH/Models/MSSM/MSSM.m  NMSSM.m}
+\item Edit NMSSM.m as follows
+\begin{enumerate}
+\item  Change Model name, author and date
+\item Add singlet superfield to SuperFields
+\begin{lstlisting}
+SuperFields[[8]] = {s, 1, sR,    0, 1,  1, RpM};
+\end{lstlisting}
+\item Modify superpotential
+\begin{lstlisting}
+SuperPotential = Yu u.q.Hu - Yd d.q.Hd - Ye e.l.Hd + \[Lambda] S.Hu.Hd +  \[Kappa]/3 s.s.s ;
+\end{lstlisting}
+\item Add new VEV
+\begin{lstlisting} DEFINITION[EWSB][VEVs]= 
+  { {SHd0, {vd, 1/Sqrt[2]}, {sigmad, \[ImaginaryI]/Sqrt[2]},{phid,1/Sqrt[2]}},
+    {SHu0, {vu, 1/Sqrt[2]}, {sigmau, \[ImaginaryI]/Sqrt[2]},{phiu,1/Sqrt[2]}},
+    {SsR,  {vs, 1/Sqrt[2]}, {sigmaS, \[ImaginaryI]/Sqrt[2]},{phiS,1/Sqrt[2]}}
+};
+\end{lstlisting}
+\item Add new VEV
+\begin{lstlisting} 
+  DEFINITION[EWSB][VEVs]= 
+  { {SHd0, {vd, 1/Sqrt[2]}, {sigmad, \[ImaginaryI]/Sqrt[2]},{phid,1/Sqrt[2]}},
+    {SHu0, {vu, 1/Sqrt[2]}, {sigmau, \[ImaginaryI]/Sqrt[2]},{phiu,1/Sqrt[2]}},
+    {SsR,  {vs, 1/Sqrt[2]}, {sigmaS, \[ImaginaryI]/Sqrt[2]},{phiS,1/Sqrt[2]}}
+};
+\end{lstlisting}
+\item  Add the singlino to the list of gauge eigenstate dirac fermions
+\begin{lstlisting}
+  DEFINITION[GaugeES][DiracSpinors]={
+  Bino ->{fB, conj[fB]},
+  Wino -> {fWB, conj[fWB]},
+  Glu -> {fG, conj[fG]},
+  H0 -> {FHd0, conj[FHu0]},
+  HC -> {FHdm, conj[FHup]},
+  Fd1 -> {FdL, 0},
+  Fd2 -> {0, FdR},
+  Fu1 -> {FuL, 0},
+  Fu2 -> {0, FuR},
+  Fe1 -> {FeL, 0},
+  Fe2 -> {0, FeR},
+  Fv -> {FvL,0}
+  FS -> {FsR,conj[FsR]}   
+};
+\end{lstlisting}
+\item Edit mass eigenstates to give how the gauge eigenstates mix into mass eigenstates.
+Here we add to the cp even higgs a real singlet, to the CP odd the imaginary part of the singlet and finally to the neutralinos we add the singlino.
+
+\begin{lstlisting}
+ DEFINITION[EWSB][MatterSector]= 
+{    {{SdL, SdR}, {Sd, ZD}},
+     {{SvL}, {Sv, ZV}},
+     {{SuL, SuR}, {Su, ZU}},
+     {{SeL, SeR}, {Se, ZE}},
+     {{phid, phiu, phiS}, {hh, ZH}},       <----------------
+     {{sigmad, sigmau, sigmaS}, {Ah, ZA}}, <----------------
+     {{SHdm,conj[SHup]},{Hpm,ZP}},
+     {{fB, fW0, FHd0, FHu0, FsR}, {L0, ZN}}, <---------------
+     {{{fWm, FHdm}, {fWp, FHup}}, {{Lm,UM}, {Lp,UP}}},
+     {{{FeL},{conj[FeR]}},{{FEL,ZEL},{FER,ZER}}},
+     {{{FdL},{conj[FdR]}},{{FDL,ZDL},{FDR,ZDR}}},
+     {{{FuL},{conj[FuR]}},{{FUL,ZUL},{FUR,ZUR}}}       };
+
+\end{lstlisting}
+\end{enumerate}
+\item Now we copy over particles.m \code{cp /path/to/SARAH/Models/MSSM/particles.m  ./}
+\item Edit particles.m as follows
+\begin{enumerate}  
+\item First we add to particleDefinitions[GaugeES]:
+\begin{lstlisting}
+ ParticleDefinitions[GaugeES] = {
+      {SdL,  { Description -> "Left Down-Squarks"}},
+      {SdR,  { Description -> "Right Down-Squarks"}},
+      ...........
+      ...........
+      ...........
+      {H0,   { Description -> "Neutral Higgsinos"}},
+      {HC,   { Description -> "Charged Higgsinos"}},
+      {SsR, { Description -> "Singlet"}},         <---------------
+      {FS,   { Description -> "Singlino" }},      <---------------
+      {Fd1,  { Description -> "Dirac Left Down-Quark"}},
+      {Fd2,  { Description -> "Dirac Right Down-Quark"}},
+      ............
+      ............
+      };
+\end{lstlisting}
+Note that we are brief in information here becasue there is already some information contained in the default particles.m file in the model directory:
+\begin{lstlisting} 
+{{ Description -> "Singlet", 
+                 PDG -> {0},
+                 PDG.IX -> 0,
+                 Width -> 0, 
+                 Mass -> Automatic,
+                 FeynArtsNr -> 98,
+                 LaTeX -> "S",
+                 OutputName -> "s"
+
+    }},
+
+(* -------- Weyl Spinor  ------ *)
+
+  {{ Description -> "Weyl Spinor of Singlino", 
+                 PDG -> 99,
+                 Width -> 0, 
+                 Mass -> Automatic,
+                 FeynArtsNr -> 9,
+                 LaTeX -> "\\tilde{S}",
+                 OutputName -> "s"
+
+    }},
+
+\end{lstlisting}
+ \item Now we edit ParticleDefinitions[EWSB].
+Again most states are defined in brief form by a name because they use the default information in the particles.m in the models directory.
+But here we need to over write some of the default options for the cp even and cp odd higgs states:
+\begin{lstlisting}
+ {hh   ,  {  Description -> "Higgs", 
+                 PDG -> {25, 35,45},
+                 PDG.IX ->{101000001,101000002,101000003}
+ }}, 
+       {Ah   ,  {    Description -> "Pseudo-Scalar Higgs",
+                 PDG -> {0, 36, 46},
+                 PDG.IX ->{0,102000001,102000002  } }},
+
+\end{lstlisting}
+\item To the WeylFermionAndIndermediate we add the Weyl singlino and the real scalar singlet field and imaginary scalar singlet field:  
+ \begin{lstlisting}
+  {sigmaS,      {Description -> "Scalar Singlet" }}  ,
+  {phiS,      { Description -> "Pseudo Scalar Singlet"}},
+  {FsR,   { Description -> "Weyl Spinor of Singlino"}},
+ \end{lstlisting}
+
+Although the WeylFermionAndIndermediate also contains, 
+\begin{lstlisting}
+{SHd,  { Description -> "Down-Higgs"}},
+       {SHu,  { Description -> "Up-Higgs"}},
+\end{lstlisting}
+\noindent we do not add a singlet version becasue for the singlet there is no distinction between these and the individual SU(2) components, as there are for Hu and Hd.
+\end{enumerate} 
+\item Now we copy over parameters.m \code{cp /path/to/SARAH/Models/MSSM/parameters.m  ./}
+\item Add the new parameters lambda, kappa, Alambda and Akappa to the list.
+\begin{lstlisting}
+{\[Kappa],   {Description -> "Singlet Self-Interaction"}},              
+{T[\[Kappa]],  { Description -> "Softbreaking Singlet Self-Interaction" }}, 
+{\[Lambda],   { Description -> "Singlet-Higgs-Interaction"   }},
+{T[\[Lambda]],  {Description -> "Softbreaking Singlet-Higgs-Interaction"}}, 
+{ms2,       { Description -> "Softbreaking Singlet Mass" }},
+{vS,        { Description -> "Singlet-VEV"}}       
+\end{lstlisting}
+Once again this is simplified from the general case becasue there are default versions containedin parameters.m in the Models directory:
+\begin{lstlisting}
+{{  Description -> "Singlet Self-Interaction",
+              LaTeX -> "\\kappa",
+             Real ->False,
+             Dependence -> None, 
+             Value -> None, 
+             LesHouches -> {NMSSMRUN,2},
+             OutputName-> kap         }},
+
+{{ Description -> "Softbreaking Singlet Self-Interaction",
+               LaTeX -> "T_{\\kappa}",
+              Real -> False,
+             Dependence -> None, 
+             Value -> None, 
+             LesHouches ->{NMSSMRUN,4},
+             OutputName-> Tk         }},
+
+ {{ Description -> "Singlet-Higgs-Interaction",
+             LaTeX -> "\\lambda",
+             Real -> False,
+	     Dependence -> None, 
+             Value -> None, 
+             LesHouches -> {NMSSMRUN,1},
+             OutputName-> lam          }},
+
+{{Description -> "Softbreaking Singlet-Higgs-Interaction",
+                LaTeX -> "T_{\\lambda}",
+              Real -> False,
+             Dependence -> None, 
+             Value -> None, 
+             LesHouches ->{NMSSMRUN,3},
+             OutputName-> Tlam         }},    
+{{ Description -> "Softbreaking Singlet Mass", 
+              LaTeX -> "m_S^2",
+             DependenceNum ->  None, 
+             LesHouches -> {NMSSMRUN,10},
+             OutputName-> ms2 }},
+
+{{ Description -> "Singlet-VEV", 
+			 LaTeX -> "v_s",
+             Value -> None, 
+             LesHouches -> {NMSSMRUN,5},
+             OutputName-> vS         }},
+\end{lstlisting}
+
+\end{enumerate}   
 \subsection{\ESSM}

### 2013-11-12 meta/LoopMasses.m
adapt diagonalization function calls to the new naming scheme

@@ -114,13 +114,13 @@ DoFastDiagonalization[particle_Symbol /; IsScalar[particle], tadpoles_List] :=
                  U = ToValidCSymbolString[mixingMatrix[[1]]];
                  V = ToValidCSymbolString[mixingMatrix[[2]]];
                  result = result <>
-                          "sarah_svd(M_1loop, PHYSICAL(" <> massName <> "), " <>
+                          "fs_svd(M_1loop, PHYSICAL(" <> massName <> "), " <>
                           "PHYSICAL(" <> U <> "), " <>
                           "PHYSICAL(" <> V <> "));\n";
                  ,
                  U = ToValidCSymbolString[mixingMatrix];
                  result = result <>
-                       "sarah_diagonalize_hermitian(M_1loop, PHYSICAL(" <> massName <> "), " <>
+                       "fs_diagonalize_hermitian(M_1loop, PHYSICAL(" <> massName <> "), " <>
                        "PHYSICAL(" <> U <> "));\n";
                 ];
               result = result <>
@@ -190,14 +190,14 @@ DoFastDiagonalization[particle_Symbol /; IsFermion[particle], _] :=
                  U = ToValidCSymbolString[mixingMatrix[[1]]];
                  V = ToValidCSymbolString[mixingMatrix[[2]]];
                  result = result <>
-                          "sarah_svd(M_1loop, " <>
+                          "fs_svd(M_1loop, " <>
                           "PHYSICAL(" <> massName <> "), " <>
                           "PHYSICAL(" <> U <> "), " <>
                           "PHYSICAL(" <> V <> "));\n";
                  ,
                  U = ToValidCSymbolString[mixingMatrix];
                  result = result <>
-                          "sarah_diagonalize_symmetric(M_1loop, " <>
+                          "fs_diagonalize_symmetric(M_1loop, " <>
                           "PHYSICAL(" <> massName <> "), " <>
                           "PHYSICAL(" <> U <> "));\n";
                 ];
@@ -259,7 +259,7 @@ DoMediumDiagonalization[particle_Symbol /; IsScalar[particle], inputMomentum_, t
               Utemp = "mix_" <> U;
               Vtemp = "mix_" <> V;
               diagSnippet = mixingMatrixType <> " " <> Utemp <> ", " <> Vtemp <> ";\n" <>
-                            "sarah_svd(M_1loop, eigen_values, " <> Utemp <> ", " <> Vtemp <> ");\n\n" <>
[…]

### 2013-11-12 meta/TreeMasses.m

@@ -583,16 +583,16 @@ CreateDiagonalizationFunction[matrix_List, eigenVector_, mixingMatrixSymbol_] :=
               (* use SVD *)
               U = ToValidCSymbolString[mixingMatrixSymbol[[1]]];
               V = ToValidCSymbolString[mixingMatrixSymbol[[2]]];
-              body = body <> "sarah_svd(" <>
+              body = body <> "fs_svd(" <>
                      matrixSymbol <> ", " <> ev <> ", " <> U <> ", " <> V <> ");\n";
               ,
               (* use conventional diagonalization *)
               U = ToValidCSymbolString[mixingMatrixSymbol];
               If[IsSymmetric[matrix] && IsFermion[GetHead[eigenVector]],
-                 body = body <> "sarah_diagonalize_symmetric(" <> matrixSymbol <> ", " <>
+                 body = body <> "fs_diagonalize_symmetric(" <> matrixSymbol <> ", " <>
                         ev <> ", " <> U <> ");\n";
                  ,
-                 body = body <> "sarah_diagonalize_hermitian(" <> matrixSymbol <> ", " <>
+                 body = body <> "fs_diagonalize_hermitian(" <> matrixSymbol <> ", " <>
                         ev <> ", " <> U <> ");\n";
                 ];
              ];

### 2013-11-12 test/test_NMSSM_low_scale_constraint.cpp
relax test precision to make the {N,S}MSSM tests pass
There seems to be a numerical difference in the output of the dia-
gonalization routines between Softsusy and Eigen of the order
1.0e-11 ... 1.0e-12

@@ -85,7 +85,7 @@ BOOST_AUTO_TEST_CASE( test_low_energy_constraint )
    const double fs_new_vev = Sqrt(Sqr(fs_new_vu) + Sqr(fs_new_vd));
 
    BOOST_CHECK_CLOSE_FRACTION(fs_mt, ss_mt, 9.5e-05);
-   BOOST_CHECK_CLOSE_FRACTION(fs_mb, ss_mb, 2.0e-14);
+   BOOST_CHECK_CLOSE_FRACTION(fs_mb, ss_mb, 2.6e-14);
    BOOST_CHECK_CLOSE_FRACTION(fs_me, ss_me, 4.3e-07);
    BOOST_CHECK_CLOSE_FRACTION(fs_MZ, ss_MZ, 5.0e-10);
    BOOST_CHECK_CLOSE_FRACTION(fs_new_vev, ss_new_vev, 5.0e-10);

### 2013-11-12 test/test_SMSSM_one_loop_spectrum.cpp

@@ -267,7 +267,7 @@ BOOST_AUTO_TEST_CASE( test_self_energies )
    const double fsT2 = m.tadpole_hh(1).real() / m.get_vu();
    const double fsTS = m.tadpole_hh(2).real() / m.get_vS();
 
-   BOOST_CHECK_CLOSE(nmT1, fsT1, 1.0e-12);
+   BOOST_CHECK_CLOSE(nmT1, fsT1, 1.1e-12);
    BOOST_CHECK_CLOSE(nmT2, fsT2, 1.0e-12);
    BOOST_CHECK_CLOSE(nmTS, fsTS, 1.0e-12);
 
@@ -288,9 +288,9 @@ BOOST_AUTO_TEST_CASE( test_self_energies )
 
    BOOST_CHECK_CLOSE(nmHSE11, fsHSE11, 1.0e-12);
    BOOST_CHECK_CLOSE(nmHSE12, fsHSE12, 1.0e-12);
-   BOOST_CHECK_CLOSE(nmHSE13, fsHSE13, 1.0e-12);
+   BOOST_CHECK_CLOSE(nmHSE13, fsHSE13, 2.0e-12);
    BOOST_CHECK_CLOSE(nmHSE22, fsHSE22, 1.0e-12);
-   BOOST_CHECK_CLOSE(nmHSE23, fsHSE23, 1.0e-12);
+   BOOST_CHECK_CLOSE(nmHSE23, fsHSE23, 2.0e-12);
    BOOST_CHECK_CLOSE(nmHSE33, fsHSE33, 1.5e-11);
 
    /// CP-odd self-energies
@@ -309,8 +309,8 @@ BOOST_AUTO_TEST_CASE( test_self_energies )
 
    BOOST_CHECK_CLOSE(nmPHSE11, fsPHSE11, 1.0e-12);
    BOOST_CHECK_CLOSE(nmPHSE12, fsPHSE12, 1.0e-12);
-   BOOST_CHECK_CLOSE(nmPHSE13, fsPHSE13, 1.0e-12);
+   BOOST_CHECK_CLOSE(nmPHSE13, fsPHSE13, 7.0e-12);
    BOOST_CHECK_CLOSE(nmPHSE22, fsPHSE22, 1.0e-12);
-   BOOST_CHECK_CLOSE(nmPHSE23, fsPHSE23, 1.0e-12);
[…]

### 2013-11-15 src/wrappers.hpp
fix compile-time error
One cannot create a matrix of type Derived if the arguments to Diag()
is a product of matrices.

@@ -102,9 +102,9 @@ inline int KroneckerDelta(int i, int j)
 }
 
 template <class Derived>
-Eigen::MatrixBase<Derived> Diag(const Eigen::MatrixBase<Derived>& m)
+typename Eigen::MatrixBase<Derived>::PlainObject Diag(const Eigen::MatrixBase<Derived>& m)
 {
-   Derived diag(m);
+   typename Eigen::MatrixBase<Derived>::PlainObject diag(m);
    for (int i = 0; i < m.rows(); ++i) {
       for (int k = 0; k < m.cols(); ++k) {
          if (i != k)

### 2013-11-15 doc/paper.tex

@@
 \begin{enumerate}
-\item Create new directory \code{mkdir NMSSM}
-\item Move into this new directory \code{cd NMSSM}
-\item copy the MSSM model file in \code{cp /path/to/SARAH/Models/MSSM/MSSM.m  NMSSM.m}
-\item Edit NMSSM.m as follows
+\item Create new directory \code{mkdir myNMSSM}
+\item Move into this new directory \code{cd myNMSSM}
+\item copy the MSSM model file in \code{cp /path/to/SARAH/Models/MSSM/MSSM.m  myNMSSM.m}
+\item Edit myNMSSM.m as follows
 \begin{enumerate}
@@
 \begin{lstlisting}
-SuperPotential = Yu u.q.Hu - Yd d.q.Hd - Ye e.l.Hd + \[Lambda] S.Hu.Hd +  \[Kappa]/3 s.s.s ;
+SuperPotential = Yu u.q.Hu - Yd d.q.Hd - Ye e.l.Hd + \[Lambda] s.Hu.Hd +  \[Kappa]/3 s.s.s ;
 \end{lstlisting}
@@
 \end{lstlisting}
+\item Navigate to Flexible SUSY directory  \code{cd /path/to/FlexibleSUSY/}
+\item Create Z3NMSSM directory 
+\begin{lstlisting}
+./createmodel --models=Z3NMSSM:myNMSSM
+\end{lstlisting}
 
@@
 Detailed steps to create \ESSM model file.
+\begin{enumerate}
+\item Create new directory \code{mkdir myE6SSM}
+\item Move into this new directory \code{cd myE6SSM}
+\item copy the MSSM model file in \code{cp /path/to/SARAH/Models/MSSM/MSSM.m  myE6SSM.m}
+\item Edit myE6SSM.m as follows
+\begin{enumerate}
+\item  Change Model name, author and date
+\item Modify MSSM superfield definitions to include new charges.
+\begin{lstlisting}
+SuperFields[[1]] = {q, 3, {uL,  dL},     1/6, 2, 3, Qq, RpM};  
+SuperFields[[2]] = {l, 3, {vL,  eL},    -1/2, 2, 1, Ql, RpM};
+SuperFields[[3]] = {Hd,1, {Hd0, Hdm},   -1/2, 2, 1, QHd, RpP};
+SuperFields[[4]] = {Hu,1, {Hup, Hu0},    1/2, 2, 1, QHu, RpP};
+SuperFields[[5]] = {d, 3, conj[dR],    1/3, 1, -3, Qd, RpM};
+SuperFields[[6]] = {u, 3, conj[uR],   -2/3, 1, -3, Qu, RpM};
+SuperFields[[7]] = {e, 3, conj[eR],      1, 1,  1, Qe, RpM};
+\end{lstlisting}
+Note that in the \ESSM model the charges are fixed to be those of the $U(1)_N$ as shown in e.g.
+\cite{King:2005jy}.
+However here we leave the charges unassigned at the Mathematica level, and will set them as parameters in the C++ code.
+The charges could also simply be fixed here to simplify the C++ code.
 
+\item Add Singlet Higgs superfield to SuperFields
+\begin{lstlisting}
+SuperFields[[8]] = {s, 1, sR,     0, 1,  1, Qs, RpP};
+\end{lstlisting}\end{lstlisting}
+\item Add inert Higgs-like superfields 
+\begin{lstlisting}
+SuperFields[[9]] = {H1I, 2, {H1I0, H1Im},  -1/2, 2, 1, QHu, RpP};
+SuperFields[[10]] = {H2I, 2, {H2Ip, H2I0},   1/2, 2, 1, QHd, RpP};
+SuperFields[[11]] = {sI, 2, sR,    0, 1,  1, Qs, RpM};
+\end{lstlisting}
+\item Add exotics to SuperFields
+\begin{lstlisting}
+SuperFields[[12]] = {Dx, 3, conj[DxR],  1/3, 1, -3, QDx, RpP};
+SuperFields[[13]] = {Dxbar, 3, conj[DxbarR],  -1/3, 1, 3, QDxbar, RpP};
+\end{lstlisting}
+\item Add additional survival $SU(2)$ multiplets to SuperFields
+\begin{lstlisting}
+SuperFields[[14]] = {Hp, 1, conj[HpR],    0.5, 2,  1, QHp, RpP};
+SuperFields[[15]] = {Hpbar, 1, conj[HpbarR],    -0.5, 2,  1, QHpbar, RpP};
+\end{lstlisting}
+\item Modify superpotential
+\begin{lstlisting}
+SuperPotential = Yu u.q.Hu - Yd d.q.Hd - Ye e.l.Hd + \[Lambda] S.Hu.Hd +  La S.H2I.H1I + \[Kappa] S.Dx.Dxbar;
+\end{lstlisting}
+\item Add new VEV
+\begin{lstlisting} DEFINITION[EWSB][VEVs]= 
+  { {SHd0, {vd, 1/Sqrt[2]}, {sigmad, \[ImaginaryI]/Sqrt[2]},{phid,1/Sqrt[2]}},
+    {SHu0, {vu, 1/Sqrt[2]}, {sigmau, \[ImaginaryI]/Sqrt[2]},{phiu,1/Sqrt[2]}},
+    {SsR,  {vs, 1/Sqrt[2]}, {sigmaS, \[ImaginaryI]/Sqrt[2]},{phiS,1/Sqrt[2]}}
+};
+\end{lstlisting}
+
+\end{enumerate}
+
+\end{enumerate}   
 \section{Tests}

### 2013-11-17 doc/paper.tex
fix typo.
added instrcutions for editing FlexibleSUSY.m to make it useful for the NMSSM.

@@
 \item Navigate to Flexible SUSY directory  \code{cd /path/to/FlexibleSUSY/}
-\item Create Z3NMSSM directory 
+\item Create Z3NMSSM directory \code{./createmodel --models=Z3NMSSM:myNMSSM}
+\item Configure with the new model \code{./configure --with-models=Z3NMSSM}
+\item The default FlexibleSUSY model file for the MSSM will now be in the models/NMSSM2 directory.  Edit this for the NMSSM as follows:
+  \begin{enumerate} 
+    \item Add new input parameter $\lambda$.
       \begin{lstlisting}
-./createmodel --models=Z3NMSSM:myNMSSM
+        EXTPAR = { {61, LambdaInput} };
       \end{lstlisting}
+     \item Also add this parameter to the DefaultParameterPoint.
+      \begin{lstlisting}
+       DefaultParameterPoint = {
+    {m0, 125},
+    {m12, 500},
+    {TanBeta, 10},
+    {Sign[\[Mu]], 1},
+    {Azero, 0}
+    {LambdaInput, 0.1}
+    };
+  \end{lstlisting}
+    \item Name the parameters you wish to be output by the EWSB conditions.  Note you should use the names given for them in parameters.m.    A common choice in the NMSSM is:
+\begin{lstlisting}
+EWSBOutputParameters = {vS, \[Kappa], ms2};
+\end{lstlisting}
+which will work for the semi-constrained NMSSM.
+\item choose the Highscale and Highscale first guess, if the default is different from what you wish to use.  
+\begin{lstlisting} 
+ HighScale = g1 == g2;
+ HighScaleFirstGuess = 1.0 10^16;
+\end{lstlisting}
+\item Add how the parameters found in parameters.m that are to be set at the highscale:
+\begin{lstlisting} 
+HighScaleInput={
+   {T[Ye], Azero*Ye},
+   {T[Yd], Azero*Yd},
+   {T[Yu], Azero*Yu},
+   {mHd2, m0^2},
+   {mHu2, m0^2},
+   {mq2, UNITMATRIX[3] m0^2},
+   {ml2, UNITMATRIX[3] m0^2},
+   {md2, UNITMATRIX[3] m0^2},
+   {mu2, UNITMATRIX[3] m0^2},
+   {me2, UNITMATRIX[3] m0^2},
+   {MassB, m12},
+   {MassWB,m12},
+   {MassG,m12},
+{\[Lambda], LambdaInput},
+   {T[\[Kappa]], Azero \[Kappa]},
+   {T[\[Lambda]], Azero LambdaInput}
+};
+\end{lstlisting}
+\item set lowscale boundary condititions if you want something different from the default.  The defaults are:
+\begin{lstlisting} 
+ LowScale = SM[MZ];
+ LowScaleFirstGuess = SM[MZ];
+\end{lstlisting}
+\item Change the LowscaleInput if necessary.  In this example we keep the same choice as in the MSSM
+  \begin{lstlisting} 
+ LowScaleInput={
+   {vd, 2 MZDRbar / Sqrt[GUTNormalization[g1]^2 g1^2 + g2^2] Cos[ArcTan[TanBeta]]},
+   {vu, 2 MZDRbar / Sqrt[GUTNormalization[g1]^2 g1^2 + g2^2] Sin[ArcTan[TanBeta]]}
+};
+  \end{lstlisting}
+
+Finally we set the initial guesses for the paremeters.  In this example we overwtite the MSSM version replacing it with:
+ \begin{lstlisting} 
+InitialGuessAtLowScale = {
+   {vd, SM[vev] Cos[ArcTan[TanBeta]]},
+   {vu, SM[vev] Sin[ArcTan[TanBeta]]},
+   {\[Lambda], LambdaInput},
+   {\[Kappa], 0.1},
+   {vS, 1000},
+   {ms2, SM[MZ]^2}
+};
+
+InitialGuessAtHighScale = {};
+ \end{lstlisting}
+
+The code can now be created and also compiled (if you didn't select --disable-compile option when you ran the configure script) by typing \code{make -j2}.
+
+ \end{enumerate}
 
@@
 SuperFields[[8]] = {s, 1, sR,     0, 1,  1, Qs, RpP};
-\end{lstlisting}\end{lstlisting}
+\end{lstlisting}
 \item Add inert Higgs-like superfields 
@@
 \end{thebibliography}
-\end{document}
+\end{document
+}

### 2013-12-12 doc/paper.tex
minor changes to instructions for nmssm in paper.tex.
typo fix on last line in paper.tex.
minor edit to remove duplication in paper.tex.
edits to paper.tex.

@@
 \item Add new VEV
-\begin{lstlisting} DEFINITION[EWSB][VEVs]= 
-  { {SHd0, {vd, 1/Sqrt[2]}, {sigmad, \[ImaginaryI]/Sqrt[2]},{phid,1/Sqrt[2]}},
-    {SHu0, {vu, 1/Sqrt[2]}, {sigmau, \[ImaginaryI]/Sqrt[2]},{phiu,1/Sqrt[2]}},
-    {SsR,  {vs, 1/Sqrt[2]}, {sigmaS, \[ImaginaryI]/Sqrt[2]},{phiS,1/Sqrt[2]}}
-};
-\end{lstlisting}
-\item Add new VEV
 \begin{lstlisting} 
@@
   DEFINITION[GaugeES][DiracSpinors]={
-  Bino ->{fB, conj[fB]},
-  Wino -> {fWB, conj[fWB]},
-  Glu -> {fG, conj[fG]},
-  H0 -> {FHd0, conj[FHu0]},
-  HC -> {FHdm, conj[FHup]},
-  Fd1 -> {FdL, 0},
-  Fd2 -> {0, FdR},
-  Fu1 -> {FuL, 0},
-  Fu2 -> {0, FuR},
-  Fe1 -> {FeL, 0},
-  Fe2 -> {0, FeR},
-  Fv -> {FvL,0}
+ .............
   FS -> {FsR,conj[FsR]}   
@@
  DEFINITION[EWSB][MatterSector]= 
-{    {{SdL, SdR}, {Sd, ZD}},
-     {{SvL}, {Sv, ZV}},
-     {{SuL, SuR}, {Su, ZU}},
-     {{SeL, SeR}, {Se, ZE}},
-     {{phid, phiu, phiS}, {hh, ZH}},       <----------------
-     {{sigmad, sigmau, sigmaS}, {Ah, ZA}}, <----------------
-     {{SHdm,conj[SHup]},{Hpm,ZP}},
-     {{fB, fW0, FHd0, FHu0, FsR}, {L0, ZN}}, <---------------
-     {{{fWm, FHdm}, {fWp, FHup}}, {{Lm,UM}, {Lp,UP}}},
-     {{{FeL},{conj[FeR]}},{{FEL,ZEL},{FER,ZER}}},
-     {{{FdL},{conj[FdR]}},{{FDL,ZDL},{FDR,ZDR}}},
-     {{{FuL},{conj[FuR]}},{{FUL,ZUL},{FUR,ZUR}}}       };
+{   ................
+    ................
+     {{phid, phiu, phiS}, {hh, ZH}},      
+     {{sigmad, sigmau, sigmaS}, {Ah, ZA}},
+    ................
+     {{fB, fW0, FHd0, FHu0, FsR}, {L0, ZN}},
+     ...............
+};
 
@@
  ParticleDefinitions[GaugeES] = {
-      {SdL,  { Description -> "Left Down-Squarks"}},
-      {SdR,  { Description -> "Right Down-Squarks"}},
-      ...........
-      ...........
-      ...........
-      {H0,   { Description -> "Neutral Higgsinos"}},
-      {HC,   { Description -> "Charged Higgsinos"}},
-      {SsR, { Description -> "Singlet"}},         <---------------
-      {FS,   { Description -> "Singlino" }},      <---------------
-      {Fd1,  { Description -> "Dirac Left Down-Quark"}},
-      {Fd2,  { Description -> "Dirac Right Down-Quark"}},
-      ............
+      {SsR, { Description -> "Singlet"}},        
+      {FS,   { Description -> "Singlino" }},    
       ............
@@
   \begin{enumerate} 
-    \item Add new input parameter $\lambda$.
-      \begin{lstlisting}
+    \item Add new input parameter $\lambda$ to EXTPAR and remove the sign of $\mu$ from MINPAR.
+      \begin{lstlisting}    
+        MINPAR = { {1, m0},
+           {2, m12},
+           {3, TanBeta},
+           {5, Azero} };
+
   EXTPAR = { {61, LambdaInput} };
@@
        DefaultParameterPoint = {
-    {m0, 125},
-    {m12, 500},
+    {m0, 200},
+    {m12, 200},
     {TanBeta, 10},
-    {Sign[\[Mu]], 1},
-    {Azero, 0}
+    {Azero, -500}
     {LambdaInput, 0.1}
@@
 \end{thebibliography}
-\end{document
-}
+\end{document}

### 2013-12-12 doc/paper.tex
added some more E6SSM parts.

@@
 \item  Change Model name, author and date
+\item Add new Vector superfield,
+ \begin{lstlisting} 
+   Gauge[[4]]={Bp,  U[1], Ncharge, g1p, False, RpM};
+ \end{lstlisting}
 \item Modify MSSM superfield definitions to include new charges.
 \begin{lstlisting}
-SuperFields[[1]] = {q, 3, {uL,  dL},     1/6, 2, 3, Qq, RpM};  
-SuperFields[[2]] = {l, 3, {vL,  eL},    -1/2, 2, 1, Ql, RpM};
-SuperFields[[3]] = {Hd,1, {Hd0, Hdm},   -1/2, 2, 1, QHd, RpP};
-SuperFields[[4]] = {Hu,1, {Hup, Hu0},    1/2, 2, 1, QHu, RpP};
-SuperFields[[5]] = {d, 3, conj[dR],    1/3, 1, -3, Qd, RpM};
-SuperFields[[6]] = {u, 3, conj[uR],   -2/3, 1, -3, Qu, RpM};
-SuperFields[[7]] = {e, 3, conj[eR],      1, 1,  1, Qe, RpM};
+SuperFields[[1]] = {q, 3, {uL,  dL},     1/6, 2, 3, 1, RpM};  
+SuperFields[[2]] = {l, 3, {vL,  eL},    -1/2, 2, 1, 2, RpM};
+SuperFields[[3]] = {Hd,1, {Hd0, Hdm},   -1/2, 2, 1, -3, RpP};
+SuperFields[[4]] = {Hu,1, {Hup, Hu0},    1/2, 2, 1, -2, RpP};
+SuperFields[[5]] = {d, 3, conj[dR],    1/3, 1, -3, 2, RpM};
+SuperFields[[6]] = {u, 3, conj[uR],   -2/3, 1, -3, 1, RpM};
+SuperFields[[7]] = {e, 3, conj[eR],      1, 1,  1, 1, RpM};
 \end{lstlisting}
@@
 \cite{King:2005jy}.
-However here we leave the charges unassigned at the Mathematica level, and will set them as parameters in the C++ code.
-The charges could also simply be fixed here to simplify the C++ code.
-
-\item Add Singlet Higgs superfield to SuperFields
-\begin{lstlisting}
-SuperFields[[8]] = {s, 1, sR,     0, 1,  1, Qs, RpP};
+However here one can leave the charges unassigned vby using parameters in place of the fixed charges.
+Here we fix the charges to simplify the C++ code.
+
+\item Add Singlet Higgs superfield to SuperField
+\begin{lstlisting}
+SuperFields[[8]] = {s, 1, sR,     0, 1,  1, 5, RpP};
 \end{lstlisting}
@@
 \begin{lstlisting}
-SuperFields[[9]] = {H1I, 2, {H1I0, H1Im},  -1/2, 2, 1, QHu, RpP};
-SuperFields[[10]] = {H2I, 2, {H2Ip, H2I0},   1/2, 2, 1, QHd, RpP};
-SuperFields[[11]] = {sI, 2, sR,    0, 1,  1, Qs, RpM};
+SuperFields[[9]] = {H1I, 2, {H1I0, H1Im},  -1/2, 2, 1, -3, RpP};
+SuperFields[[10]] = {H2I, 2, {H2Ip, H2I0},   1/2, 2, 1, -2, RpP};
+SuperFields[[11]] = {sI, 2, sIR,    0, 1,  1, 5, RpM};
 \end{lstlisting}
@@
 \begin{lstlisting}
-SuperFields[[12]] = {Dx, 3, conj[DxR],  1/3, 1, -3, QDx, RpP};
-SuperFields[[13]] = {Dxbar, 3, conj[DxbarR],  -1/3, 1, 3, QDxbar, RpP};
+SuperFields[[12]] = {Dx, 3, DxL,  -1/3, 1, 3, -2, RpP};
+SuperFields[[13]] = {Dxbar, 3, conj[DxbarR],  1/3, 1, -3, -3, RpP};
 \end{lstlisting}
@@
 \begin{lstlisting}
-SuperFields[[14]] = {Hp, 1, conj[HpR],    0.5, 2,  1, QHp, RpP};
-SuperFields[[15]] = {Hpbar, 1, conj[HpbarR],    -0.5, 2,  1, QHpbar, RpP};
+SuperFields[[14]] = {Hp, 1, {Hpd0, Hpdm},  1/2, 2,  1, 2, RpP};
+SuperFields[[15]] = {Hpbar, 1, {Hpup, Hpu0}, -1/2, 2,  1, -2, RpP};
 \end{lstlisting}
@@
 \begin{lstlisting}
-SuperPotential = Yu u.q.Hu - Yd d.q.Hd - Ye e.l.Hd + \[Lambda] S.Hu.Hd +  La S.H2I.H1I + \[Kappa] S.Dx.Dxbar;
+SuperPotential = Yu u.q.Hu - Yd d.q.Hd - Ye e.l.Hd + \[Lambda] s.Hu.Hd +  \[Lambda]12 s.H2I.H1I + \[Kappa] s.Dx.Dxbar + \[Mu]Pr Hp.Hpbar;
+\end{lstlisting}
+\item Define new gauge eigenstate Dirac spinors
+\begin{lstlisting}
+  DEFINITION[GaugeES][DiracSpinors]={
+    .............
+    .............
+    FS -> {FsR,conj[FsR]},
+    FBp -> {fBp,conj[fBp]},
+    HI0 -> {FH1I0, conj[FH2I0]},
+    HIC -> {FH1Im, conj[FH2Ip]},
+    FSI -> {FsIR, conj[FsIR]},
+    FDx1 -> {FDxL, 0},
+    FDx2 -> {0, FDxbarR}, 
+    Hp0 -> {FHpd0, conj[FHpu0]},
+    HpC -> {FHpdm, conj[FHpup]}
+  };
 \end{lstlisting}
 \item Add new VEV
-\begin{lstlisting} DEFINITION[EWSB][VEVs]= 
+\begin{lstlisting} 
+  DEFINITION[EWSB][VEVs]= 
   { {SHd0, {vd, 1/Sqrt[2]}, {sigmad, \[ImaginaryI]/Sqrt[2]},{phid,1/Sqrt[2]}},
@@
 \end{lstlisting}
-
+\item Add new gauge boson  mass eigenstate
+\begin{lstlisting} 
+  DEFINITION[EWSB][GaugeSector] =
+  { 
+    {{VB,VWB[3],VBp},{VP,VZ,VZp},ZZ},
+    {{VWB[1],VWB[2]},{VWm,conj[VWm]},ZW},
+    {{fWB[1],fWB[2],fWB[3]},{fWm,fWp,fW0},ZfW}
+  }; 
+\end{lstlisting}
+\item Edit mixings for Higgs and neutralino states and add additional states that mix.
+\begin{lstlisting}
+DEFINITION[EWSB][MatterSector]= 
+{ ...............
+  ...............
+{{phid, phiu, phiS}, {hh, ZH}},
+{{sigmad, sigmau, sigmaS}, {Ah, ZA}},
+  ...............
+{{fB, fW0, FHd0, FHu0, FsR, fBp}, {L0, ZN}},
+ ................ 
+ {{SH1I0,conj[SH2I0]},{SHI0,UHI0}},
+ {{SH1Im,conj[SH2Ip]},{SHIp,UHIp}},
+ {{SHpd0,conj[SHpu0]},{SHp0,UHp0}},
+ {{SHpdm,conj[SHpup]},{SHpp,UHpp}},
+ {{FHpd0,FHpu0},{L0p,ZNp}},
+ {{FH1I0,FH2I0},{L0I,ZNp}}
+\end{lstlisting}
 \end{enumerate}

### 2013-12-16 doc/paper.tex

@@
  {{FHpd0,FHpu0},{L0p,ZNp}},
- {{FH1I0,FH2I0},{L0I,ZNp}}
+ {{FH1I0,FH2I0},{L0I,ZNI}}
 \end{lstlisting}

### 2013-12-28 doc/paper.tex
briefly describe model parameter and beta function classes

@@
 mention two loop qcd parts from softsusy.
-\subsection{RGEs}
-The two loop renormalisation group equations are generated using command ?
-Even in the simplest supersymmetric model the RGEs form a very large set of coupled non-linear differential equations.
-To solve these and thereby evolve the parameters between two different scales \fs has...
+
+\subsection{Model parameters and RGEs}
+Even in the simplest supersymmetric model the RGEs form a very large
+set of coupled non-linear differential equations.  In a general
+(softly broken) supersymmetric theory the soft-breaking parameters
+depend on themselves and the susy parameters.  However, the susy
+parameters are independent of the soft-breaking parameters [cite: MV,
+Staub, Sperling].  This makes it possible to split the model
+parameters and their into the following two subsets:
+%
+\begin{enumerate}
+\item \emph{susy parameters:} gauge couplings, superpotential
+  parameters and VEVs and
+\item soft-breaking parameters.
+\end{enumerate}
+%
+The susy parameters together with their two-loop renormalisation group
+equations for the are stored in the class
+\code{<model>_susy_parameters}.  The soft-breaking model parameters
+together with their two-loop renormalisation group equations for the
+are stored in the class \code{<model>_soft_parameters}.  The
+dependence of the soft-breaking parameters on the susy parameters is
+implemented via inheritance, see
+\figurename~\ref{fig:parameter-classes}.
+%
+\begin{figure}
+  \centering
+  \tikzumlset{fill class=white}
+  \begin{tikzpicture}
+    \umlclass[x=0, y=3]{model\_susy\_parameters}{
+      - susy parameters
+    }{
+      + beta()
+    }
+    \umlclass[x=0, y=0]{model\_soft\_parameters}{
+      - soft-breaking parameters
+    }{
+      + beta()
+    }
+    \umlinherit{model\_soft\_parameters}{model\_susy\_parameters}
+  \end{tikzpicture}
+  \caption{Model parameter class structure.}
+  \label{fig:parameter-classes}
+\end{figure}
+
 \subsection{tree level spectrum}

### 2013-12-28 doc/paper.tex
extend EWSB solving section

@@
 \section{Stucture of the spectrum generator}
-\subsection{boundary conditions}
-Different types, how we match to low energy data.  reference EWSB or merge that section here.
-mention two loop qcd parts from softsusy.
 
@@
 
+\subsection{Boundary conditions}
+In \fs the RGE boundary conditions are implemented by inheriting from
+a common \code{Constraint<Two\_scale>} interface class and
+implementing the pure virtual functions \code{apply()},
+\code{get\_scale()} and \code{set\_model()}, see
+\figurename~\ref{fig:schematic-two-scale-constraint-interface}.  Per
+default, \fs creates three default boundary conditions:
+%
+\begin{itemize}
+\item \emph{GUT constraint:} Here the \code{apply()} function
+  calculates the value of the unification scale $M_X$ (from the
+  condition given in \code{HighScale}) and sets the model parameters
+  given in \code{HighScaleInput}.
+\item \emph{SUSY constraint:} Here the \code{apply()} function
+  calculates the value of the susy scale $M_S$ and sets the model
+  parameters given in \code{SUSYScaleInput}.  Afterwards the EWSB
+  conditions are solved for the parameters given in
+  \code{EWSBOutputParameters} at the one-loop level, i.e., the the
+  parameters in \code{EWSBOutputParameters} are determined such that
+  the one-loop effective potential is minimized.
+\item \emph{Low-scale constraint:} This constraint matches the gauge
+  and Yukawa couplings of the susy model to the Standard Model values
+  using threshold corrections.  Furthermore the parameters given in
+  \code{LowScaleInput} are set.
+\end{itemize}
+
+{\color{red}Different types, how we match to low energy data.
+  reference EWSB or merge that section here.  mention two loop qcd
+  parts from softsusy.}
+
+\subsubsection{Electroweak symmetry breaking}
+In \fs the one-loop electroweak symmetry breaking conditions are
+formulated as
+%
+\begin{align}
+  0 = \frac{\partial V^\text{tree}}{\partial v_i} - t_i,
+  \label{eq:one-loop-ewsb-eq}
+\end{align}
+%
+where $V^\text{tree}$ is the tree-level Higgs potential, $v_i$ is the
+VEVs corresponding to the Higgs field $H_i$ and $t_i$ is the one-loop
+tadpole diagram of $H_i$.
+
+The electroweak symmetry breaking conditions
+\eqref{eq:one-loop-ewsb-eq} are then solved simultaneously using the
+iterative multi-dimensional root finder algorithm
+\code{gsl_multiroot_fsolver_hybrid} from the Gnu Scientific Libraray
+(GSL).  If no root was found, the \code{gsl_multiroot_fsolver_hybrids}
+algorithm is tried as alternative, which uses a variable step size but
+might be a little slower.
+
+For example in the MSSM \fs expresses \eqref{eq:one-loop-ewsb-eq} as
+the function
+%
+\begin{lstlisting}[language=C++]
+int MSSM<Two_scale>::tadpole_equations(const gsl_vector* x, void* params, gsl_vector* f)
+{
+   ...
+
+   const CLASSNAME::Ewsb_parameters* ewsb_parameters
+      = static_cast<CLASSNAME::Ewsb_parameters*>(params);
+   MSSM* model = ewsb_parameters->model;
+   const unsigned ewsb_loop_order = ewsb_parameters->ewsb_loop_order;
+
+   double tadpole[numberOfEWSBEquations];
+
+   model->set_BMu(gsl_vector_get(x, 0));
+   model->set_Mu(INPUT(SignMu) * Abs(gsl_vector_get(x, 1)));
+
+   // calculate tree-level tadpole eqs.
+   tadpole[0] = model->get_ewsb_eq_vd();
+   tadpole[1] = model->get_ewsb_eq_vu();
+
+   // subtract one-loop tadpoles
+   if (ewsb_loop_order > 0) {
+      model->calculate_DRbar_parameters();
+      tadpole[0] -= Re(model->tadpole_hh(0));
+      tadpole[1] -= Re(model->tadpole_hh(1));
+
+   }
+
+   for (std::size_t i = 0; i < numberOfEWSBEquations; ++i)
+      gsl_vector_set(f, i, tadpole[i]);
+
+   return GSL_SUCCESS;
+}
+\end{lstlisting}
+%
+Here \code{x} is the vector of EWSB output parameters
+(\code{EWSBOutputParameters}) and \code{f} is a vector which contains
+the one-loop EWSB eqs.\ \eqref{eq:one-loop-ewsb-eq}.  This
+\code{tadpole_equations()} function is called inside of
+\code{MSSM<Two_scale>::solve_ewsb_iteratively_with} then via
+%
+\begin{lstlisting}[language=C++]
+// initial guess
+double x_init[numberOfEWSBEquations];
+ewsb_initial_guess(x_init);
+
+// setup root finder
+int ewsb_loop_order = 1;
+Ewsb_parameters params = {this, ewsb_loop_order};
+Root_finder<numberOfEWSBEquations> root_finder(MSSM<Two_scale>::tadpole_equations,
+                           &params,
+                           number_of_ewsb_iterations,
+                           ewsb_iteration_precision);
+root_finder.set_solver_type(gsl_multiroot_fsolver_hybrid);
+
+root_finder.find_root(x_init);
+\end{lstlisting}
+
+If higher accuracy is required additional routines with higher order
+corrections can be added in a very simple way.  For example in the
+MSSM by default \fs adds two loop Higgs FORTRAN routines supplied by
+P.~Slavich from \cite{slavich-degrassi} to add two loop corrections of
+$\oatas$, $\oabas$, $\oatq$, $\oabatau$, $\oabq$, $\oatauq$ and
+$\oatab$ as follows...  Something similar is done for the NMSSM, but
+in this case the NMSSM $\oatas$, $\oabas$ pieces come from
+\cite{Degrassi:2009yq}.  Corrections from other sources can be added
+to the generated code by...
+
 \subsection{tree level spectrum}
@@
 The numerical diagonalisation is performed by... with an accuracy set by...   
-\subsection{Electroweak symmetry breaking}
-At tree level the EWSB conditions are written...
-The one loop electroweak symmetry breaking are solved using the gsl routines finding multiple roots, \code{gsl_multiroot_fsolver}.  Iteration. For example in the MSSM \fs generates.. show code.
- If higher accuracy is required additional routines with higher order corrections can be added in a very simple way.
-For example in the MSSM by default \fs adds two loop Higgs FORTRAN routines supplied by P.~Slavich from \cite{slavich-degrassi} to add two loop corrections of $\oatas$, $\oabas$, $\oatq$, $\oabatau$,
-$\oabq$, $\oatauq$ and $\oatab$  as follows...  
-Something similar is done for the NMSSM, but in this case the NMSSM $\oatas$, $\oabas$ pieces come from \cite{Degrassi:2009yq}.
- Corrections from other sources can be added to the generated code by...
 \subsection{two scale fixed point iteration}

### 2013-12-28 doc/paper.tex
extend tree-level spectrum calculation and fixed-point it. section

@@
 
-\subsection{tree level spectrum}
-The diagonalised tree level $\overline{DR}$ masses are obtained from expressions...
-The numerical diagonalisation is performed by... with an accuracy set by...   
-\subsection{two scale fixed point iteration}
-The default solver which calls the above routines above and finds the full set of $\overline{DR}$ masses in the model consistent with all constraints is the two-scale solver.
-This can be selected by...
-but if no selection is made this will be used anyway as default.
-
-This approach is decribed in \cite{Barger:1993gh} for the MSSM and is widely implemented in SUSY spectrum generators.
-
-The stucture of the two-scale algorithm used in \fs is as follows,
-%
-Initial guess:
+\subsection{Tree-level spectrum}
+The tree-level $\overline{DR}$ masses are calculated by diagonalizing
+the mass matrices returned from \code{SARAH`MassMatrix[]}.  The
+numerical diagonalization is performed by the Eigen library routines
+\code{Eigen::JacobiSVD} and \code{Eigen::SelfAdjointEigenSolver} for
+small matrices, and the Lapack routines \code{zgesvd}, \code{dgesvd},
+\code{zheev}, \code{dsyev} for matrices with more than four rows and
+columns.
+
+\subsection{Two-scale fixed point iteration}
+The default two-scale RG solver uses above beta functions and boundary
+conditions to find the full set of $\overline{DR}$ masses in the model
+consistent with all constraints.  It does so by iterating between the
+scales of all boundary conditions, applying the constraints and
+checking for convergence.  This approach is decribed in
+\cite{Barger:1993gh} for the MSSM and is widely implemented in SUSY
+spectrum generators.
+
+The in more detail the two-scale algorithm used in \fs works as
+follows:
+%
+\\Initial guess:
 \begin{enumerate}

### 2013-12-28 doc/paper.tex
correcting required SARAH version and naming convention

@@
            C++}\\ {\em Computer:}\/ Personal computer\\ {\em Operating
-           system:}\/ Tested on Linux 4.x\\ {\em Word size:}\/ 32
-         bits\\ {\em External routines:}\/ SARAH 4.0.0, Boost library,
+           system:}\/ Tested on Linux 3.x\\ {\em Word size:}\/ 64
+         bits\\ {\em External routines:}\/ SARAH 4.0.3, Boost library,
          lapack {\color{red} Not sure what counts here?}\\ {\em
@@
 \item Mathematica, version 7 or higher
-\item SARAH, version 4.0.0 or higher \url{http://sarah.hepforge.org}
+\item SARAH, version 4.0.3 or higher \url{http://sarah.hepforge.org}
 \item C++11 compatible compiler (g++ 4.6.0 or higher, clang++ 3.1 or
@@
 
-   double tadpole[numberOfEWSBEquations];
+   double tadpole[number_of_ewsb_equations];
 
@@
 
-   for (std::size_t i = 0; i < numberOfEWSBEquations; ++i)
+   for (std::size_t i = 0; i < number_of_ewsb_equations; ++i)
       gsl_vector_set(f, i, tadpole[i]);
@@
 // initial guess
-double x_init[numberOfEWSBEquations];
+double x_init[number_of_ewsb_equations];
 ewsb_initial_guess(x_init);
@@
 Ewsb_parameters params = {this, ewsb_loop_order};
-Root_finder<numberOfEWSBEquations> root_finder(MSSM<Two_scale>::tadpole_equations,
+Root_finder<number_of_ewsb_equations> root_finder(
+                           MSSM<Two_scale>::tadpole_equations,
                            &params,

### 2013-12-28 doc/paper.tex
reformulate the pole-mass section a bit

@@
 {\color{red} copied and pasted from latest manual -- not sure how best to maintain this, could split maual up into tiny bits, but we may also want some differences?}
-\subsection{pole masses}
-With the solver routine finished and convergence achieved, the $\overline{DR}$ parameters consistent with one loop EWSB, low energy data and all user supplied boundary conditions are then known at any scale between...
-
-The physical mass spectrum can now be calculated.
-\fs uses the full one-loop self energies and tree level routines to calculate the pole masses at the one-level.
-If higher accuracy is required additional routines with higher order corrections can be added in a very simple way.
-For example in the MSSM by default \fs adds two loop Higgs FORTRAN routines supplied by P.~Slavich from \cite{slavich-degrassi} to add two loop corrections of $\oatas$, $\oabas$, $\oatq$, $\oabatau$,
-$\oabq$, $\oatauq$ and $\oatab$  as follows...  
-Something similar is done for the NMSSM, but in this case the NMSSM $\oatas$, $\oabas$ pieces come from \cite{Degrassi:2009yq}, while for $\oatq$, $\oabatau$, 
-$\oabq$, $\oatauq$ and $\oatab$ the MSSM pieces are used though it should be understood that these are not complete in the NMSSM.
-For other models since the Higgs mass is a very important measurement and the two loop corrections can be larger than the current experimental error \cite{Degrassi:2009yq} we recommend that the leading log two loop corrections are estimated, by generalising those of the MSSM or NMSSM, as has been done, for example, in the E$_6$SSM\cite{King:2005jy}.
-
-Corrections to other masses or further corrections to the Higgs states can be added very easily by...
+
+\subsection{Pole masses}
+After the solver routine has finished and convergence has been
+achieved, all $\overline{DR}$ parameters are consistent with the
+one-loop EWSB conditions, low energy data and all user supplied
+boundary conditions are known at any scale between \code{LowScale} and
+\code{HighScale}.
+
+The physical (pole) mass spectrum can now be calculated.  \fs uses the
+full one-loop self-energies and tree-level mass matrices to calculate
+the pole masses at the one-loop level.  If higher accuracy is required
+additional routines with higher order corrections can be added in a
+very simple way. For example in the MSSM by default \fs adds two-loop
+Higgs FORTRAN routines supplied by P.~Slavich from
+\cite{slavich-degrassi} to add two loop corrections of $\oatas$,
+$\oabas$, $\oatq$, $\oabatau$, $\oabq$, $\oatauq$ and $\oatab$ as
+follows... {\color{red} This is not yet implemented!}
+
+Something similar is done for the NMSSM, but in this case the NMSSM
+$\oatas$, $\oabas$ pieces come from \cite{Degrassi:2009yq}, while for
+$\oatq$, $\oabatau$, $\oabq$, $\oatauq$ and $\oatab$ the MSSM pieces
+are used though it should be understood that these are not complete in
+the NMSSM. For other models since the Higgs mass is a very important
+measurement and the two loop corrections can be larger than the
+current experimental error \cite{Degrassi:2009yq} we recommend that
+the leading log two loop corrections are estimated, by generalising
+those of the MSSM or NMSSM, as has been done, for example, in the
+E$_6$SSM\cite{King:2005jy}.  {\color{red} This is not yet implemented!}
+
+Corrections to other masses or further corrections to the Higgs states
+can be added very easily by...
 

### 2014-01-22 doc/chapters/output.tex
small mlanguage improvement from paper.tex to output.tex.

@@
 
-The two-scale algorithm works like this:
-%
+The stucture of the two-scale algorithm used in \flexisusy\ is as follows,%
+
 Initial guess:
@@
   \item Recalculate low-scale
-  \item Calculate \DRbar\ gauge and Yukawa couplings $g_{1,2,3}$, $y^{u,d,e}_{ij}$
+  \item Calculate \DRbar\ gauge and Yukawa couplings $g_{1,2,3}$,
+    $y^{u,d,e}_{ij}$, see
+    Section~\ref{sec:matching-to-the-standard-model}.
   \item Apply user-defined low-scale constraint (\code{LowScaleInput})
@@
 
+\subsubsection{Matching to the Standard Model}
+\label{sec:matching-to-the-standard-model}
+
+At the low-scale $M_Z$ we calculate
+%
+\begin{align}
+  \alpha_{\text{e.m.},\text{susy}}^{\text{\DRbar}}(M_Z) &=
+  \frac{\alpha_{\text{e.m.},\text{SM}}^{(5),\text{\MSbar}}(M_Z)}{1 -
+    \Delta\alpha_{\text{e.m.},\text{SM}}(M_Z) -
+    \Delta\alpha_{\text{e.m.},\text{susy}}(M_Z)} ,\\
+  \Delta\alpha_{\text{e.m.},\text{SM}}(\mu) &=
+  \frac{\alpha_\text{e.m.}}{2\pi} \left[\frac{1}{3}
+    - \frac{16}{9} \log{\frac{m_t}{\mu}} \right],\\
+  \Delta\alpha_{\text{e.m.},\text{susy}}(\mu) &=
+  \frac{\alpha_\text{e.m.}}{2\pi} \left[ -\sum_{\text{susy particle }
+      i}
+    C_i \log{\frac{m_i}{\mu}} \right],\\
+    e_{\text{susy}}^{\text{\DRbar}}(M_Z) &=
+    \sqrt{4\pi\alpha_{\text{e.m.},\text{susy}}^{\text{\DRbar}}(M_Z)}
+\end{align}
+%
+\begin{align}
+  \alpha_{\text{s},\text{susy}}^{\text{\DRbar}}(M_Z) &=
+  \frac{\alpha_{\text{s},\text{SM}}^{(5),\text{\MSbar}}(M_Z)}{1 -
+    \Delta\alpha_{\text{s},\text{SM}}(M_Z)
+    - \Delta\alpha_{\text{s},\text{susy}}(M_Z)} ,\\
+  \Delta\alpha_{\text{s},\text{SM}}(\mu) &=
+  \frac{\alpha_\text{s}}{2\pi} \left[
+    -\frac{2}{3} \log{\frac{m_t}{\mu}} \right],\\
+  \Delta\alpha_{\text{s},\text{susy}}(\mu) &=
+  \frac{\alpha_\text{s}}{2\pi}\left[ \frac{1}{2}-\sum_{\text{susy
+        particle } i} C_i \log{\frac{m_i}{\mu}} \right]
+\end{align}
+%
+
 \subsection{Lattice algorithm}

### 2014-01-22 doc/chapters/output.tex
adding numerical values for alpha_em and alpha_s
dumping low-energy mass formulas into the manual
Still missing:
 * formatting
 * what does "heavy" mean?

@@
 
+\paragraph{Gauge couplings}
+
 At the low-scale $M_Z$ we calculate
@@
   \frac{\alpha_\text{s}}{2\pi}\left[ \frac{1}{2}-\sum_{\text{susy
-        particle } i} C_i \log{\frac{m_i}{\mu}} \right]
+        particle } i} C_i \log{\frac{m_i}{\mu}} \right] ,\\
+  g_{3,\text{susy}}^{\text{\DRbar}}(M_Z) &=
+  \sqrt{4\pi\alpha_{\text{s},\text{susy}}^{\text{\DRbar}}(M_Z)} ,
 \end{align}
 %
+where
+%
+\begin{align}
+  \alpha_{\text{e.m.},\text{SM}}^{(5),\text{\MSbar}}(M_Z) &= 1/127.944,\\
+  \alpha_{\text{s},\text{SM}}^{(5),\text{\MSbar}}(M_Z) &= 0.1185,
+\end{align}
+%
+are the \MSbar electromagnetic and strong cougling constants in the
+Standard Model including only $5$ quark flavours
+\cite{Beringer:1900zz}.  Afterwards, SARAH's tree-level expressions
+for the Weinberg angle and the Hypercharge and left gauge couplings
+$g_Y$ and $g_2$ in terms of $M_{W,\text{susy}}^{\text{\DRbar}}(M_Z)$,
+$M_{Z,\text{susy}}^{\text{\DRbar}}(M_Z)$ and
+$e_{\text{susy}}^{\text{\DRbar}}(M_Z)$ are used to calculate
+$g_{1,\text{susy}}^{\text{\DRbar}}(M_Z)$ and
+$g_{2,\text{susy}}^{\text{\DRbar}}(M_Z)$.  In the MSSM, for example,
+we'll have
+%
+\begin{align}
+  \theta_{W,\text{susy}}^{\text{\DRbar}}(M_Z) &=
+  \arcsin\sqrt{1 - \left(\frac{M_{W,\text{susy}}^{\text{\DRbar}}(M_Z)}{M_{Z,\text{susy}}^{\text{\DRbar}}(M_Z)}\right)^2} ,\\
+  g_{1,\text{susy}}^{\text{\DRbar}}(M_Z) &=
+  \sqrt{\frac{5}{3}} \frac{e_{\text{susy}}^{\text{\DRbar}}(M_Z)}{\cos\theta_{W,\text{susy}}^{\text{\DRbar}}(M_Z)} ,\\
+  g_{2,\text{susy}}^{\text{\DRbar}}(M_Z) &=
+  \frac{e_{\text{susy}}^{\text{\DRbar}}(M_Z)}{\sin\theta_{W,\text{susy}}^{\text{\DRbar}}(M_Z)} ,
+\end{align}
+%
+where
+%
+\begin{align}
+  \left(M_{W,\text{susy}}^{\text{\DRbar}}(M_Z)\right)^2 &=
+  \left(M_W^{\text{pole}}\right)^2 + \re \Pi_{WW}^T(p^2 = (M_W^{\text{pole}})^2) ,\\
+  \left(M_{Z,\text{susy}}^{\text{\DRbar}}(M_Z)\right)^2 &=
+  \left(M_Z^{\text{pole}}\right)^2 + \re \Pi_{ZZ}^T(p^2 = (M_Z^{\text{pole}})^2)
+\end{align}
+%
+and $M_W^{\text{pole}} = 80.404\unit{GeV}$, $M_Z^{\text{pole}} =
+91.1876\unit{GeV}$ \cite{Beringer:1900zz}.
+
+\paragraph{Fermion masses}
+
+\begin{align}
+  \left(y_u^{\text{\DRbar}}(M_Z)\right)_{ij} &= \frac{\sqrt{2} m_{u,ji}}{v_u} ,\\
+  \left(y_d^{\text{\DRbar}}(M_Z)\right)_{ij} &= \frac{\sqrt{2} m_{d,ji}}{v_d} ,\\
+  \left(y_e^{\text{\DRbar}}(M_Z)\right)_{ij} &= \frac{\sqrt{2} m_{e,ji}}{v_e} ,
+\end{align}
+%
+where the fermion mass matrices are given by
+%
+\begin{align}
+  m_{u,11} &= m_{u}^{\userinput} ,\\
+  m_{u,22} &= m_{c}^{\userinput} ,\\
+  m_{u,33} &= m_{t,\text{susy}}^{\text{\DRbar}}(M_Z),\\
+  m_{d,11} &= m_{d}^{\userinput} ,\\
+  m_{d,22} &= m_{s}^{\userinput} ,\\
+  m_{d,33} &= m_{b,\text{susy}}^{\text{\DRbar}}(M_Z),\\
+  m_{e,11} &= m_{e}^{\userinput} ,\\
+  m_{e,22} &= m_{\mu}^{\userinput} ,\\
+  m_{e,33} &= m_{\tau,\text{susy}}^{\text{\DRbar}}(M_Z),\\
+  m_{u,ij} &= m_{d,ij} = m_{u,ij} = 0 \quad \forall i \neq j.
+\end{align}
+%
+Only the 3rd generation quark masses are calculated in the \DRbar
+scheme from the user input quantities $m_t^\text{pole}$,
+$m_{b,\text{SM}}^{\text{\MSbar}}(M_Z)$,
+$m_{\tau,\text{SM}}^{\text{\MSbar}}(M_Z)$.  The top quark \DRbar mass
+is calculated via
+%
+\begin{align}
+  m_{t,\text{susy}}^{\text{\DRbar}}(\mu) &= m_t^\text{pole}
+  + \re\Sigma_{tt}^{S,\text{heavy}}(m_t^\text{pole})
+  + m_t^\text{pole} \left[
+    \re\Sigma_{tt}^{L,\text{heavy}}(m_t^\text{pole}) +
+    \re\Sigma_{tt}^{R,\text{heavy}}(m_t^\text{pole}) +
+    \Delta m_t^{(1),\text{qcd}} +
+    \Delta m_t^{(2),\text{qcd}}
+  \right].
+\end{align}
+%
+The QCD contributions $\Delta m_t^{(1),\text{qcd}}$ and $\Delta
+m_t^{(2),\text{qcd}}$ are taken from \cite{Bednyakov:2002sf} and read
+%
+\begin{align}
+  \Delta m_t^{(1),\text{qcd}} &= -\frac{g_3^2 \left(5-3 \log\left(\frac{m_t^2}{\mu^2}\right)\right)}{12 \pi^2},\\
+  \Delta m_t^{(2),\text{qcd}} &= \left(\Delta
+    m_t^{(1),\text{qcd}}\right)^2 - \frac{g_3^4 \left[396
+      \log^2\left(\frac{m_t^2}{\mu^2}\right)-1476
+      \log\left(\frac{m_t^2}{\mu^2}\right)-48 \zeta(3)+2011+16 \pi ^2
+      (1+\log(4))\right]}{4608 \pi^4}.
+\end{align}
+%
+The \DRbar mass of the bottom quark is calculated via
+\cite{Baer:2002ek}
+%
+\begin{align}
+  m_{b,\text{susy}}^{\text{\DRbar}}(\mu) &=
+  \frac{m_{b,\text{SM}}^{\text{\DRbar}}(\mu)}{1 -
+    \re\Sigma_{bb}^{S,\text{heavy}}(m_{b,\text{SM}}^\text{\MSbar})/m_b
+    - \re\Sigma_{bb}^{L,\text{heavy}}(m_{b,\text{SM}}^\text{\MSbar}) -
+    \re\Sigma_{bb}^{R,\text{heavy}}(m_{b,\text{SM}}^\text{\MSbar})} ,\\
+  m_{b,\text{SM}}^{\text{\DRbar}}(\mu) &=
+  m_{b,\text{SM}}^{\text{\MSbar}}(\mu) \left(1 - \frac{\alpha_s}{3
+      \pi} - \frac{23}{72} \frac{\alpha_s^2}{\pi^2} + \frac{3
+      g_2^2}{128 \pi^2} + \frac{13 g_1^2}{1152 \pi^2}\right) .
+\end{align}
+%
+The \DRbar mass of the $\tau$ is calculated via
+%
+\begin{align}
+  m_{\tau,\text{susy}}^{\text{\DRbar}}(\mu) &=
+  m_{\tau,\text{SM}}^{\text{\DRbar}}(\mu) +
+  \re\Sigma_{\tau\tau}^{S,\text{heavy}}(m_{\tau,\text{SM}}^\text{\MSbar}) +
+  m_{\tau,\text{SM}}^{\text{\DRbar}}(\mu) \left[
+    \re\Sigma_{\tau\tau}^{L,\text{heavy}}(m_{\tau,\text{SM}}^\text{\MSbar}) +
+    \re\Sigma_{\tau\tau}^{R,\text{heavy}}(m_{\tau,\text{SM}}^\text{\MSbar})
+  \right] ,\\
+  m_{\tau,\text{SM}}^{\text{\DRbar}}(\mu) &= m_{\tau,\text{SM}}^{\text{\MSbar}}(\mu)
+  \left(1 - 3 \frac{g_1^2 - g_2^2}{128 \pi^2}\right).
+\end{align}
 

### 2014-01-22 doc/chapters/output.tex
reformat a bit and explain user input and heavy self-energies

@@
 
+The expressios for the SM-Yukawa coupling matrices extracted from the
+tree-level mass matrices of the SM-fermions.  In the MSSM, for
+example, they read
+%
 \begin{align}
-  \left(y_u^{\text{\DRbar}}(M_Z)\right)_{ij} &= \frac{\sqrt{2} m_{u,ji}}{v_u} ,\\
-  \left(y_d^{\text{\DRbar}}(M_Z)\right)_{ij} &= \frac{\sqrt{2} m_{d,ji}}{v_d} ,\\
-  \left(y_e^{\text{\DRbar}}(M_Z)\right)_{ij} &= \frac{\sqrt{2} m_{e,ji}}{v_e} ,
+  y_u^{\text{\DRbar}}(M_Z) &= \frac{\sqrt{2} m_{u}^T}{v_u} ,\\
+  y_d^{\text{\DRbar}}(M_Z) &= \frac{\sqrt{2} m_{d}^T}{v_d} ,\\
+  y_e^{\text{\DRbar}}(M_Z) &= \frac{\sqrt{2} m_{e}^T}{v_e}.
 \end{align}
 %
-where the fermion mass matrices are given by
+At the low-scale the fermion mass matrices are composed as
 %
 \begin{align}
-  m_{u,11} &= m_{u}^{\userinput} ,\\
-  m_{u,22} &= m_{c}^{\userinput} ,\\
-  m_{u,33} &= m_{t,\text{susy}}^{\text{\DRbar}}(M_Z),\\
-  m_{d,11} &= m_{d}^{\userinput} ,\\
-  m_{d,22} &= m_{s}^{\userinput} ,\\
-  m_{d,33} &= m_{b,\text{susy}}^{\text{\DRbar}}(M_Z),\\
-  m_{e,11} &= m_{e}^{\userinput} ,\\
-  m_{e,22} &= m_{\mu}^{\userinput} ,\\
-  m_{e,33} &= m_{\tau,\text{susy}}^{\text{\DRbar}}(M_Z),\\
-  m_{u,ij} &= m_{d,ij} = m_{u,ij} = 0 \quad \forall i \neq j.
+  m_u =
+  \begin{pmatrix}
+    m_{u}^{\userinput} & 0 & 0 \\
+    0 & m_{c}^{\userinput} & 0 \\
+    0 & 0 & m_{t,\text{susy}}^{\text{\DRbar}}(M_Z)
+  \end{pmatrix} ,\\
+  m_d =
+  \begin{pmatrix}
+    m_{d}^{\userinput} & 0 & 0 \\
+    0 & m_{s}^{\userinput} & 0 \\
+    0 & 0 & m_{b,\text{susy}}^{\text{\DRbar}}(M_Z)
+  \end{pmatrix} ,\\
+  m_e =
+  \begin{pmatrix}
+    m_{e}^{\userinput} & 0 & \\
+    0 & m_{\mu}^{\userinput} & 0 \\
+    0 & 0 & m_{\tau,\text{susy}}^{\text{\DRbar}}(M_Z)
+  \end{pmatrix}.
 \end{align}
@@
 $m_{b,\text{SM}}^{\text{\MSbar}}(M_Z)$,
-$m_{\tau,\text{SM}}^{\text{\MSbar}}(M_Z)$.  The top quark \DRbar mass
-is calculated via
+$m_{\tau,\text{SM}}^{\text{\MSbar}}(M_Z)$.  In detail, the top quark
+\DRbar mass is calculated via
 %
 \begin{align}
-  m_{t,\text{susy}}^{\text{\DRbar}}(\mu) &= m_t^\text{pole}
-  + \re\Sigma_{tt}^{S,\text{heavy}}(m_t^\text{pole})
-  + m_t^\text{pole} \left[
-    \re\Sigma_{tt}^{L,\text{heavy}}(m_t^\text{pole}) +
-    \re\Sigma_{tt}^{R,\text{heavy}}(m_t^\text{pole}) +
-    \Delta m_t^{(1),\text{qcd}} +
-    \Delta m_t^{(2),\text{qcd}}
-  \right].
+  \begin{split}
+    m_{t,\text{susy}}^{\text{\DRbar}}(\mu) &= m_t^\text{pole} +
+    \re\Sigma_{tt}^{S,\text{heavy}}(m_t^\text{pole}) \\
+    &\phantom{=\;} + m_t^\text{pole}
+    \left[ \re\Sigma_{tt}^{L,\text{heavy}}(m_t^\text{pole}) +
+      \re\Sigma_{tt}^{R,\text{heavy}}(m_t^\text{pole}) + \Delta
+      m_t^{(1),\text{qcd}} + \Delta m_t^{(2),\text{qcd}} \right] ,
+  \end{split}
 \end{align}
 %
-The QCD contributions $\Delta m_t^{(1),\text{qcd}}$ and $\Delta
-m_t^{(2),\text{qcd}}$ are taken from \cite{Bednyakov:2002sf} and read
+where the $\Sigma_{ff}^\text{heavy}$ denotes the self-energy of
+fermion $f$ without the Standard Model contributions.  The appearing
+QCD corrections $\Delta m_t^{(1),\text{qcd}}$ and $\Delta
+m_t^{(2),\text{qcd}}$ are taken from
+\cite[Eq. (58),(61)]{Bednyakov:2002sf} and read
 %
@@
   \Delta m_t^{(1),\text{qcd}} &= -\frac{g_3^2 \left(5-3 \log\left(\frac{m_t^2}{\mu^2}\right)\right)}{12 \pi^2},\\
+  \begin{split}
     \Delta m_t^{(2),\text{qcd}} &= \left(\Delta
-    m_t^{(1),\text{qcd}}\right)^2 - \frac{g_3^4 \left[396
+      m_t^{(1),\text{qcd}}\right)^2 \\
+    &\phantom{=\;} - \frac{g_3^4 \left[396
         \log^2\left(\frac{m_t^2}{\mu^2}\right)-1476
-      \log\left(\frac{m_t^2}{\mu^2}\right)-48 \zeta(3)+2011+16 \pi ^2
-      (1+\log(4))\right]}{4608 \pi^4}.
+        \log\left(\frac{m_t^2}{\mu^2}\right)-48 \zeta(3)+2011+16 \pi
+        ^2 (1+\log 4)\right]}{4608 \pi^4}.
+  \end{split}
 \end{align}
@@
       \pi} - \frac{23}{72} \frac{\alpha_s^2}{\pi^2} + \frac{3
-      g_2^2}{128 \pi^2} + \frac{13 g_1^2}{1152 \pi^2}\right) .
+      g_2^2}{128 \pi^2} + \frac{13 g_1^2}{1152 \pi^2}\right) ,
 \end{align}
 %
-The \DRbar mass of the $\tau$ is calculated via
+where $m_{b,\text{SM}}^{\text{\MSbar}}(M_Z)$ is user input.  The
+\DRbar mass of the $\tau$ is calculated via
 %
 \begin{align}
+  \begin{split}
     m_{\tau,\text{susy}}^{\text{\DRbar}}(\mu) &=
     m_{\tau,\text{SM}}^{\text{\DRbar}}(\mu) +
-  \re\Sigma_{\tau\tau}^{S,\text{heavy}}(m_{\tau,\text{SM}}^\text{\MSbar}) +
-  m_{\tau,\text{SM}}^{\text{\DRbar}}(\mu) \left[
-    \re\Sigma_{\tau\tau}^{L,\text{heavy}}(m_{\tau,\text{SM}}^\text{\MSbar}) +
+    \re\Sigma_{\tau\tau}^{S,\text{heavy}}(m_{\tau,\text{SM}}^\text{\MSbar}) \\
+    &\phantom{=\;} + m_{\tau,\text{SM}}^{\text{\DRbar}}(\mu) \left[
+      \re\Sigma_{\tau\tau}^{L,\text{heavy}}(m_{\tau,\text{SM}}^\text{\MSbar})
+      +
       \re\Sigma_{\tau\tau}^{R,\text{heavy}}(m_{\tau,\text{SM}}^\text{\MSbar})
-  \right] ,\\
+    \right] ,
+  \end{split}\\
   m_{\tau,\text{SM}}^{\text{\DRbar}}(\mu) &= m_{\tau,\text{SM}}^{\text{\MSbar}}(\mu)
-  \left(1 - 3 \frac{g_1^2 - g_2^2}{128 \pi^2}\right).
+  \left(1 - 3 \frac{g_1^2 - g_2^2}{128 \pi^2}\right),
 \end{align}
+%
+where $m_{\tau,\text{SM}}^{\text{\MSbar}}(M_Z)$ is user input.
 

### 2014-01-26 doc/chapters/usage.tex
write a workflow section and correcting ./createmodel usage

@@
 \section{Usage}
+
+\subsection{Workflow}
+
+Creating a spectrum generator with \flexisusy consists of the
+following steps:
+%
+\begin{enumerate}
+\item Creating a \emph{SARAH model file}.  Note, that SARAH already
+  ships a lot of pre-defined model files in \code{SARAH/Models/} that
+  can be used.
+\item Creating a \emph{\flexisusy model file} in
+  \code{models/<model-name>/}.  The \code{createmodel} script will
+  help you creating one, see Section~\ref{sec:createmodel}.  For more
+  details on how to write a \flexisusy model file see
+  Section~\ref{sec:model-files}.
+\item \emph{Configure} \flexisusy with your desired model, see
+  Section~\ref{sec:configure}.
+\item Run \code{make}.  This will create the C++ code for your
+  spectrum generator and compile it, see Section~\ref{sec:make}.
+\end{enumerate}
+
 \subsection{Files and directories}
@@
 \subsubsection{\code{createmodel}}
+\label{sec:createmodel}
 
@@
 \begin{lstlisting}[language=bash]
-./createmodel --models=<flexiblesusy-model>:<sarah-model>
+./createmodel --name=<flexiblesusy-model> --sarah-model=<sarah-model>
 \end{lstlisting}
@@
 \begin{lstlisting}[language=bash]
-./createmodel --models=MyCMSSM:MSSM
+./createmodel --name=MyCMSSM --sarah-model=MSSM
 \end{lstlisting}
 %
-Multiple models can be created at once by separating the models by a
-comma.\\
-Example:
-\begin{lstlisting}[language=bash]
-./createmodel --models=MyCMSSM:MSSM,MyCNMSSM:NMSSM
-\end{lstlisting}
 %
@@
 \begin{lstlisting}[language=bash]
-./createmodel --models=MyCMSSM:MSSM/CKM
+./createmodel --name=MyCMSSM --sarah-model=MSSM/CKM
 \end{lstlisting}
@@
 \subsubsection{\code{configure}}
+\label{sec:configure}
 
@@
 
+\subsubsection{\code{make}}
+\label{sec:make}
+
+Running \code{make} will create the C++ source code for your spectrum
+generator and compile it.  These two processes are controled in the
+makefile module \code{models/<model-name>/module.mk}.  See,
+Section~\ref{sec:makefile-modules} for details how to create your own
+makefile module.
+
 \subsection{Model files}
+\label{sec:model-files}
 
@@
 \subsection{Makefile modules}
+\label{sec:makefile-modules}
 

### 2014-01-26 doc/chapters/usage.tex
add illustration of the workflow

@@
 \begin{enumerate}
-\item Creating a \emph{SARAH model file}.  Note, that SARAH already
-  ships a lot of pre-defined model files in \code{SARAH/Models/} that
-  can be used.
+  \setcounter{enumi}{-1}
+\item Creating a \emph{SARAH model file}.  The SARAH model file must
+  be loadable via \code{Start["<model-name>"]}.  Note, that SARAH
+  already ships a lot of pre-defined model files in
+  \code{SARAH/Models/} that can be used.
 \item Creating a \emph{\flexisusy model file} in
@@
 \end{enumerate}
+%
+The steps 1--3 are illustrated in \figref{fig:workflow}.
+
+\begin{figure}[tbh]
+  \centering
+  \begin{tikzpicture}[node distance = 3.5cm, auto]
+    \node[block, text width=14em] (createmodel) {%
+      \code{./createmodel --name=NMSSM}};
+    \node[block, text width=14em, below of=createmodel] (configure) {%
+      \code{./configure --with-models=NMSSM}};
+    \node[block, text width=14em, below of=configure] (make) {%
+      \code{make}};
+    \node[cloud, text width=22em, right=1.6cm of createmodel] (createmodel-output) {%
+      Creates the files:\\%
+      \code{./models/NMSSM/module.mk}\\%
+      \code{./models/NMSSM/start.m}\\%
+      \code{./models/NMSSM/FlexibleSUSY.m}};
+    \node[cloud, text width=22em, right=1.6cm of configure] (configure-output) {%
+      Creates \code{./Makefile} which includes\\%
+      \code{./models/NMSSM/module.mk}};
+    \node[cloud, text width=22em, right=1.6cm of make] (make-output) {%
+      \begin{enumerate}\itemsep0em
+      \item Runs the meta code \code{math -run
+          "Get[\"./models/NMSSM/start.m\"];"} which creates the C++
+        source files.
+      \item The source files are then compiled to
+        \code{./models/NMSSM/libNMSSM.a} and linked against the
+        executable \code{./models/NMSSM/run\_NMSSM.x}.
+      \end{enumerate}};
+    \path[arrow] (createmodel) -- (configure);
+    \path[arrow] (configure) -- (make);
+    \path[draw, dashed] (createmodel) -- (createmodel-output);
+    \path[draw, dashed] (configure) -- (configure-output);
+    \path[draw, dashed] (make) -- (make-output);
+  \end{tikzpicture}
+  \caption{\flexisusy workflow}
+  \label{fig:workflow}
+\end{figure}
 

### 2014-02-05 doc/paper.tex
increase required SARAH version to 4.0.4

@@
            system:}\/ Tested on Linux 3.x\\ {\em Word size:}\/ 64
-         bits\\ {\em External routines:}\/ SARAH 4.0.3, Boost library,
+         bits\\ {\em External routines:}\/ SARAH 4.0.4, Boost library,
          lapack {\color{red} Not sure what counts here?}\\ {\em
@@
 \item Mathematica, version 7 or higher
-\item SARAH, version 4.0.3 or higher \url{http://sarah.hepforge.org}
+\item SARAH, version 4.0.4 or higher \url{http://sarah.hepforge.org}
 \item C++11 compatible compiler (g++ 4.6.0 or higher, clang++ 3.1 or

### 2014-02-20 doc/paper.tex
adding run-time comparison section

@@
 \newcommand{\DRbar}{\textoverline{DR}}  % inline source code
+\newcommand{\unit}[1]{\,\text{#1}}      % units
+\DeclareMathOperator{\sign}{sign}
 \def\at{\alpha_t}
@@
 \section{Using individual modules in your own code}
+
+\section{Run-time comparison with other spectrum generators}
+
+We compare the run-time of the four different spectrum generators
+FlexibleSUSY (version 0.5.3), Softsusy (version 3.4.0), SPheno
+(3.2.4), SPhenoMSSM (generated with SARAH 4.1.0 and linked against
+SPheno 3.2.4).  FlexibleSUSY and Softsusy were compiled with g++ 4.8.0
+and gfortran 4.8.0.  SPheno and SPhenoMSSM were compiled with Intel
+ifort 13.1.3 20130607\footnote{We are using Intel's ifort compiler,
+  instead of gfortran 4.8.0, because it decreases the run-time of
+  SPheno and SPhenoMSSM by approximately a factor $1.5$.}.  We're
+generating random CMSSM parameter points with $m_0\in [50,1000]$
+$m_{1/2}\in [50,1000]$, $\tan\beta\in [1,100]$, $\sign\mu\in
+\{-1,+1\}$ and $A_0\in [-1000,1000]$.  The following SLHA template
+file was used:
+%
+\begin{verbatim}
+Block MODSEL                 # Select model
+    6    0                   # flavour violation
+    1    1                   # mSUGRA
+Block SMINPUTS               # Standard Model inputs
+    1   1.279180000e+02      # alpha^(-1) SM MSbar(MZ)
+    2   1.166390000e-05      # G_Fermi
+    3   1.189000000e-01      # alpha_s(MZ) SM MSbar
+    4   9.118760000e+01      # MZ(pole)
+    5   4.200000000e+00      # mb(mb) SM MSbar
+    6   1.709000000e+02      # mtop(pole)
+    7   1.777000000e+00      # mtau(pole)
+Block SOFTSUSY               # SOFTSUSY specific inputs
+    1   1.000000000e-04      # tolerance
+    2   2                    # up-quark mixing (=1) or down (=2)
+    3   0                    # printout
+    5   1                    # 2-loop running
+    7   2                    # EWSB and Higgs mass loop order
+Block FlexibleSUSY
+    0   1.000000000e-04      # precision goal
+    1   0                    # max. iterations (0 = automatic)
+    2   0                    # algorithm (0 = two_scale, 1 = lattice)
+    3   0                    # calculate SM pole masses
+    4   1                    # calculate loop masses
+    5   1                    # EWSB loop order
+Block SPhenoInput            # SPheno specific input
+    1  -1                    # error level
+    2   1                    # SPA conventions
+    11  0                    # calculate branching ratios
+    13  0                    # include 3-Body decays
+    12  1.000E-04            # write only branching ratios larger than this value
+    31  -1                   # fixed GUT scale (-1: dynamical GUT scale)
+    32  0                    # Strict unification
+    34  1.000E-04            # Precision of mass calculation
+    35  40                   # Maximal number of iterations
+    37  1                    # Set Yukawa scheme
+    38  2                    # 1- or 2-Loop RGEs
+    50  1                    # Majorana phases: use only positive masses
+    51  0                    # Write Output in CKM basis
+    52  0                    # Write spectrum in case of tachyonic states
+    55  1                    # Calculate one loop masses
+    57  0                    # Calculate low energy constraints
+    60  0                    # Include possible, kinetic mixing
+    65  1                    # Solution tadpole equation
+    75  0                    # Write WHIZARD files
+    76  0                    # Write HiggsBounds file
+    86  0.                   # Maximal width to be counted as invisible in Higgs decays
+    510 0.                   # Write tree level values for tadpole solutions
+    515 0                    # Write parameter values at GUT scale
+    520 0.                   # Write effective Higgs couplings (HiggsBounds blocks)
+    525 0.                   # Write loop contributions to diphoton decay of Higgs
+Block MINPAR
+    1   [50..1000]           # m0(MX)
+    2   [50..1000]           # m12(MX)
+    3   [1..100]             # tan(beta)(MZ) DRbar
+    4   {-1,+1}              # sign(mu)
+    5   [-1000..1000]        # A0(MX)
+\end{verbatim}
+%
+The SLHA file was passed to each spectrum generator and the
+(wall-clock) time was measured until the program has finished.  The
+average run-times for three different CPU types can be found in
+\tablename~\ref{tab:run-time-comparison}.  The first row shows the
+run-time on a Intel Core2 Duo (P8600) where only one core was enabled.
+The second row shows the run-time on a Intel Core2 Duo (P8600) where
+both cores were enabled.  In the third row a Intel Xeon (L5640) was
+used, which has $6$ CPU cores.
+%
+\begin{table}[tbh]
+  \centering
+  \begin{tabular}{llll}
+    \toprule
+                          & Intel Core2 Duo            & Intel Core2 Duo             & Intel Xeon\\
+                          & (1 core, $2.40\unit{GHz}$) & (2 cores, $2.40\unit{GHz}$) & (6 cores, $2.27\unit{GHz}$)\\
+    \midrule
+    FlexibleSUSY 0.5.3    & $0.151\unit{s}$    & $0.114\unit{s}$   & $0.081\unit{s}$\\
+    Softsusy 3.4.0        & $0.178\unit{s}$    & $0.174\unit{s}$   & $0.157\unit{s}$\\
+    SPheno 3.2.4          & $0.120\unit{s}$    & $0.118\unit{s}$   & $0.108\unit{s}$\\
+    SPhenoMSSM 4.1.0${}^*$ & $0.463\unit{s}$    & $0.450\unit{s}$   & $0.403\unit{s}$\\
+    \bottomrule
+  \end{tabular}
+  \caption{Average run-time for random CMSSM parameter points.
+    FlexibleSUSY and Softsusy were compiled
+    with g++ 4.8.0 and gfortran 4.8.0.  SPheno and SPhenoMSSM were
+    compiled with ifort 13.1.3 20130607.
+    ${}^*$ SPhenoMSSM was generated with SARAH 4.1.0 and linked
+    against SPheno 3.2.4.}
+  \label{tab:run-time-comparison}
+\end{table}
+%
+We find that on the single-core machine SPheno is fastest with
+$0.120\unit{s}$, and FlexibleSUSY is second with $0.151\unit{s}$.  In
+the case of 2 or 6 CPU cores, FlexibleSUSY is fastest with
+$0.114\unit{s}$ and $0.081\unit{s}$, respectively.  The reason is that
+FlexibleSUSY calculates each pole mass in a separate thread, and
+therefore benefits from multi-core CPUs.
+
 \appendix

### 2014-02-21 doc/paper.tex
adding the CPU type to the table head
remove lattice and EFT sections
extend RGE section

@@
 depend on themselves and the susy parameters.  However, the susy
-parameters are independent of the soft-breaking parameters [cite: MV,
-Staub, Sperling].  This makes it possible to split the model
-parameters and their into the following two subsets:
+parameters are independent of the soft-breaking parameters
+\cite{Jones:1974pg,Jones:1983vk,West:1984dg,Martin:1993yx,Yamada:1993ga,MV94,Fonseca:2011vn,Sperling:2013eva,Sperling:2013xqa}.
+This makes it possible to split the model parameters into the
+following two subsets:
 %
@@
   parameters and VEVs and
-\item soft-breaking parameters.
+\item \emph{soft-breaking parameters} \cite{Girardello:1981wz}: soft
+  linear scalar terms, soft bilinear scalar interactions, soft
+  trilinear scalar interactions, soft gaugino mass terms and soft
+  scalar squared masses.
 \end{enumerate}
@@
 The susy parameters together with their two-loop renormalisation group
-equations for the are stored in the class
-\code{<model>_susy_parameters}.  The soft-breaking model parameters
-together with their two-loop renormalisation group equations for the
-are stored in the class \code{<model>_soft_parameters}.  The
-dependence of the soft-breaking parameters on the susy parameters is
-implemented via inheritance, see
-\figurename~\ref{fig:parameter-classes}.
+equations are stored in the class \code{<model>_susy_parameters}.  The
+soft-breaking model parameters together with their two-loop
+renormalisation group equations are stored in the class
+\code{<model>_soft_parameters}.  The dependence of the soft-breaking
+parameters on the susy parameters is implemented via inheritance, see
+\figurename~\ref{fig:parameter-classes}, where
+\code{<model>_soft_parameters} inherits from
+\code{<model>_susy_parameters}.  In this way, the class
+\code{<model>_soft_parameters} can implement the soft-breaking beta
+functions in terms of the susy and soft-breaking parameters, while the
+class \code{<model>_susy_parameters} is independent of the
+soft-breaking parameters.
 %
@@
 
-\subsection{Lattice method}
-If ready.
-\subsection{multiple effective models at different scales}
 \section{Implementing new models}
@@
     \toprule
-                          & Intel Core2 Duo            & Intel Core2 Duo             & Intel Xeon\\
+                          & Intel Core2 Duo (P8600)    & Intel Core2 Duo (P8600)     & Intel Xeon (L5640)\\
                           & (1 core, $2.40\unit{GHz}$) & (2 cores, $2.40\unit{GHz}$) & (6 cores, $2.27\unit{GHz}$)\\
@@
 \begin{thebibliography}{100}
+%\cite{Girardello:1981wz}
+\bibitem{Girardello:1981wz}
+  L.~Girardello and M.~T.~Grisaru,
+  %``Soft Breaking of Supersymmetry,''
+  Nucl.\ Phys.\ B {\bf 194} (1982) 65.
+  %%CITATION = NUPHA,B194,65;%%
+  %717 citations counted in INSPIRE as of 21 Feb 2014
+
 %\cite{Allanach:2001kg}
@@
 
+%\cite{Jones:1974pg}
+\bibitem{Jones:1974pg}
+  D.~R.~T.~Jones,
+  %``Asymptotic Behavior of Supersymmetric Yang-Mills Theories in the Two Loop Approximation,''
+  Nucl.\ Phys.\ B {\bf 87} (1975) 127.
+  %%CITATION = NUPHA,B87,127;%%
+  %123 citations counted in INSPIRE as of 21 Feb 2014
+
+%\cite{Jones:1983vk}
+\bibitem{Jones:1983vk}
+  D.~R.~T.~Jones and L.~Mezincescu,
+  %``The Beta Function in Supersymmetric {Yang-Mills} Theory,''
+  Phys.\ Lett.\ B {\bf 136} (1984) 242.
+  %%CITATION = PHLTA,B136,242;%%
+  %109 citations counted in INSPIRE as of 21 Feb 2014
+
+%\cite{West:1984dg}
+\bibitem{West:1984dg}
+  P.~C.~West,
+  %``The Yukawa beta Function in N=1 Rigid Supersymmetric Theories,''
+  Phys.\ Lett.\ B {\bf 137} (1984) 371.
+  %%CITATION = PHLTA,B137,371;%%
+  %134 citations counted in INSPIRE as of 21 Feb 2014
+
+%\cite{Martin:1993yx}
+\bibitem{Martin:1993yx}
+  S.~P.~Martin and M.~T.~Vaughn,
+  %``Regularization dependence of running couplings in softly broken supersymmetry,''
+  Phys.\ Lett.\ B {\bf 318} (1993) 331
+  [hep-ph/9308222].
+  %%CITATION = HEP-PH/9308222;%%
+  %220 citations counted in INSPIRE as of 21 Feb 2014
+
+%\cite{Yamada:1993ga}
+\bibitem{Yamada:1993ga}
+  Y.~Yamada,
+  %``Two loop renormalization of gaugino masses in general supersymmetric gauge models,''
+  Phys.\ Rev.\ Lett.\  {\bf 72} (1994) 25
+  [hep-ph/9308304].
+  %%CITATION = HEP-PH/9308304;%%
+  %48 citations counted in INSPIRE as of 21 Feb 2014
+
 %\cite{MV94}
@@
 
+%\cite{Fonseca:2011vn}
+\bibitem{Fonseca:2011vn}
+  R.~M.~Fonseca, M.~Malinsky, W.~Porod and F.~Staub,
+  %``Running soft parameters in SUSY models with multiple U(1) gauge factors,''
+  Nucl.\ Phys.\ B {\bf 854} (2012) 28
+  [arXiv:1107.2670 [hep-ph]].
+  %%CITATION = ARXIV:1107.2670;%%
+  %29 citations counted in INSPIRE as of 21 Feb 2014
+
 %\cite{Yam94}
@@
 
-\cite{Sper13}
-\bibitem{Sper13} 
+%\cite{Sperling:2013eva}
+\bibitem{Sperling:2013eva}
   M.~Sperling, D.~Stöckinger and A.~Voigt,
   %``Renormalization of vacuum expectation values in spontaneously broken gauge theories,''
-  JHEP {\bf 1307}, 132 (2013)
+  JHEP {\bf 1307} (2013) 132
   [arXiv:1305.1548 [hep-ph]].
   %%CITATION = ARXIV:1305.1548;%%
-  %4 citations counted in INSPIRE as of 14 Oct 2013
+  %10 citations counted in INSPIRE as of 21 Feb 2014
+
+%\cite{Sperling:2013xqa}
+\bibitem{Sperling:2013xqa}
+  M.~Sperling, D.~Stöckinger and A.~Voigt,
+  %``Renormalization of vacuum expectation values in spontaneously broken gauge theories: Two-loop results,''
+  JHEP {\bf 1401} (2014) 068
+  [arXiv:1310.7629 [hep-ph]].
+  %%CITATION = ARXIV:1310.7629;%%
+  %4 citations counted in INSPIRE as of 21 Feb 2014
 

### 2014-02-21 doc/paper.tex
extend the boundary conditions section

@@
 \subsection{Boundary conditions}
-In \fs the RGE boundary conditions are implemented by inheriting from
-a common \code{Constraint<Two\_scale>} interface class and
-implementing the pure virtual functions \code{apply()},
-\code{get\_scale()} and \code{set\_model()}, see
-\figurename~\ref{fig:schematic-two-scale-constraint-interface}.  Per
-default, \fs creates three default boundary conditions:
+In \fs the RGE boundary conditions are classes which implement
+\code{Constraint<Two\_scale>} interface, see
+\figurename~\ref{fig:schematic-two-scale-constraint-interface}.
+%
+\begin{lstlisting}[language=C++]
+template<>
+class Constraint<Two_scale> {
+public:
+   virtual ~Constraint() {}
+   virtual void apply() = 0;
+   virtual double get_scale() const = 0;
+   virtual void set_model(Two_scale_model*) = 0;
+};
+\end{lstlisting}
+%
+The three pure virtual functions \code{apply()}, \code{get\_scale()}
+and \code{set\_model()} need to be implemented
+%
+\begin{itemize}
+\item \code{apply()}: implements the constraint by fixing model
+  parameters
+\item \code{get\_scale()}: returns the scale at which the constraint
+  is to be applied.
+\item \code{set\_model()}: connects the constraint to a model class,
+  in order to access the model parameters
+\end{itemize}
+%
+Per default, \fs creates three default boundary conditions for each
+model:
 %
@@
   condition given in \code{HighScale}) and sets the model parameters
-  given in \code{HighScaleInput}.
+  given in \code{HighScaleInput}.  The \code{get\_scale()} returns the
+  value of $M_X$.
 \item \emph{SUSY constraint:} Here the \code{apply()} function
@@
   parameters in \code{EWSBOutputParameters} are determined such that
-  the one-loop effective potential is minimized.
-\item \emph{Low-scale constraint:} This constraint matches the gauge
-  and Yukawa couplings of the susy model to the Standard Model values
-  using threshold corrections.  Furthermore the parameters given in
-  \code{LowScaleInput} are set.
+  the one-loop effective potential is minimized.  The
+  \code{get\_scale()} returns the value of $M_S$.
+\item \emph{Low-scale constraint:} This constraint calculates the
+  gauge and Yukawa couplings of the susy model from the known Standard
+  Model gauge couplings and fermion masses (at $M_Z$), using threshold
+  corrections.  Furthermore the parameters given in
+  \code{LowScaleInput} are set.  The \code{get\_scale()} returns the
+  value of $M_Z$.
 \end{itemize}

### 2014-02-21 models/SoftsusyMSSM/softsusy.cpp
Correcting MS-bar DR-bar conversion of fermion masses

@@ -2682,12 +2682,12 @@ double Softsusy<SoftPars>::calcRunningMt() {
 template<class SoftPars>
 double Softsusy<SoftPars>::calcRunMbDrBarConv() const {
    double    g       = displayGaugeCoupling(2);
-   double    g1      = displayGaugeCoupling(1);
+   double    gp      = displayGaugeCoupling(1) * sqrt(0.6);
    double alphasMZ = sqr(displayGaugeCoupling(3)) / (4.0 * PI);
    double conversion = (1.0 - alphasMZ / (3.0 * PI) 
                         - 23.0 / 72.0 * sqr(alphasMZ) / sqr(PI) +
                         3.0 * sqr(g) / (128.0 * sqr(PI)) +
-                        13.0 * sqr(g1) / (1152. * sqr(PI)));
+                        13.0 * sqr(gp) / (1152. * sqr(PI)));
    return conversion;
 }
 
@@ -2909,7 +2909,8 @@ double Softsusy<SoftPars>::calcRunningMb() const {
 }
 template<class SoftPars>
 double Softsusy<SoftPars>::calcRunMtauDrBarConv() const {
- double conv = (1.0 - 3.0 * (sqr(displayGaugeCoupling(1)) - sqr(displayGaugeCoupling(2))) / (128.0 * sqr(PI)));
+ double gp = displayGaugeCoupling(1) * sqrt(0.6);
+ double conv = (1.0 - 3.0 * (sqr(gp) - sqr(displayGaugeCoupling(2))) / (128.0 * sqr(PI)));
  return conv;
 }
 

### 2014-02-21 test/test_MSSM_low_scale_constraint.cpp

@@ -132,7 +132,7 @@ BOOST_AUTO_TEST_CASE( test_low_energy_constraint )
 
    BOOST_CHECK_CLOSE_FRACTION(fs_mt, ss_mt, 9.5e-5);
    BOOST_CHECK_CLOSE_FRACTION(fs_mb, ss_mb, 3.0e-15);
-   BOOST_CHECK_CLOSE_FRACTION(fs_me, ss_me, 4.3e-7);
+   BOOST_CHECK_CLOSE_FRACTION(fs_me, ss_me, 6.0e-7);
    BOOST_CHECK_CLOSE_FRACTION(fs_MZ, ss_MZ, 4.5e-10);
    BOOST_CHECK_CLOSE_FRACTION(fs_new_vev, ss_new_vev, 4.5e-10);
    BOOST_CHECK_CLOSE_FRACTION(fs_old_vu / fs_old_vd, s.displayTanb(), 1.0e-10);

### 2014-02-21 test/test_NMSSM_low_scale_constraint.cpp

@@ -86,7 +86,7 @@ BOOST_AUTO_TEST_CASE( test_low_energy_constraint )
 
    BOOST_CHECK_CLOSE_FRACTION(fs_mt, ss_mt, 9.5e-05);
    BOOST_CHECK_CLOSE_FRACTION(fs_mb, ss_mb, 2.6e-14);
-   BOOST_CHECK_CLOSE_FRACTION(fs_me, ss_me, 4.3e-07);
+   BOOST_CHECK_CLOSE_FRACTION(fs_me, ss_me, 6.0e-07);
    BOOST_CHECK_CLOSE_FRACTION(fs_MZ, ss_MZ, 5.0e-10);
    BOOST_CHECK_CLOSE_FRACTION(fs_new_vev, ss_new_vev, 5.0e-10);
    BOOST_CHECK_CLOSE_FRACTION(fs_old_vu / fs_old_vd, s.displayTanb(), 1.0e-10);

### 2014-02-21 doc/chapters/output.tex
correcting MS-bar DR-bar conversion for the fermion masses
There is a bug in Softsusy and FlexibleSUSY!
See SLHA paper Eq. (19).
adding reference to the used MS-bar DR-bar conversion of mb and mtau

@@
       \pi} - \frac{23}{72} \frac{\alpha_s^2}{\pi^2} + \frac{3
-      g_2^2}{128 \pi^2} + \frac{13 g_1^2}{1152 \pi^2}\right) ,
+      g_2^2}{128 \pi^2} + \frac{13 g_Y^2}{1152 \pi^2}\right) ,
 \end{align}
@@
   m_{\tau,\text{SM}}^{\text{\DRbar}}(\mu) &= m_{\tau,\text{SM}}^{\text{\MSbar}}(\mu)
-  \left(1 - 3 \frac{g_1^2 - g_2^2}{128 \pi^2}\right),
+  \left(1 - 3 \frac{g_Y^2 - g_2^2}{128 \pi^2}\right),
 \end{align}
 %
-where $m_{\tau,\text{SM}}^{\text{\MSbar}}(M_Z)$ is user input.
+where $m_{\tau,\text{SM}}^{\text{\MSbar}}(M_Z)$ is user input.  To
+convert the fermion masses from the \MSbar to the \DRbar scheme the
+Yukawa coupling conversion from \cite[Eq.~(19)]{Skands:2003cj} was
+used and it was assumed that the VEV is already given in the \DRbar
+scheme.
 

### 2014-02-22 doc/paper.tex
add dummy sections for the details of the low-scale constraint
adding eqs. to calculate the Yukawa couplings y_f(MZ)

@@
 \newcommand{\textoverline}[1]{$\overline{\mbox{#1}}$}
-\newcommand{\DRbar}{\textoverline{DR}}  % inline source code
+\newcommand{\DRbar}{\textoverline{DR}\xspace}
+\newcommand{\MSbar}{\textoverline{MS}\xspace}
 \newcommand{\unit}[1]{\,\text{#1}}      % units
+\newcommand{\userinput}{\text{<input>}}
 \DeclareMathOperator{\sign}{sign}
+\DeclareMathOperator{\re}{Re}
+\DeclareMathOperator{\im}{Im}
 \def\at{\alpha_t}
@@
   parts from softsusy.}
+
+\subsubsection{Calculation of the gauge couplings $g_i(M_Z)$}
+
+The low-scale constraint calculates the gauge couplings $g_i(M_Z)$.
+It does so by first calculating $e_{\text{susy}}^{\text{\DRbar}}(M_Z)$
+and $g_{3,\text{susy}}^{\text{\DRbar}}(M_Z)$ via
+%
+\begin{align}
+  \alpha_{\text{e.m.},\text{susy}}^{\text{\DRbar}}(M_Z) &=
+  \frac{\alpha_{\text{e.m.},\text{SM}}^{(5),\text{\MSbar}}(M_Z)}{1 -
+    \Delta\alpha_{\text{e.m.},\text{SM}}(M_Z) -
+    \Delta\alpha_{\text{e.m.},\text{susy}}(M_Z)} ,\\
+  \Delta\alpha_{\text{e.m.},\text{SM}}(\mu) &=
+  \frac{\alpha_\text{e.m.}}{2\pi} \left[\frac{1}{3}
+    - \frac{16}{9} \log{\frac{m_t}{\mu}} \right],\\
+  \Delta\alpha_{\text{e.m.},\text{susy}}(\mu) &=
+  \frac{\alpha_\text{e.m.}}{2\pi} \left[ -\sum_{\text{susy particle }
+      i}
+    C_i \log{\frac{m_i}{\mu}} \right],\\
+    e_{\text{susy}}^{\text{\DRbar}}(M_Z) &=
+    \sqrt{4\pi\alpha_{\text{e.m.},\text{susy}}^{\text{\DRbar}}(M_Z)}
+\end{align}
+%
+\begin{align}
+  \alpha_{\text{s},\text{susy}}^{\text{\DRbar}}(M_Z) &=
+  \frac{\alpha_{\text{s},\text{SM}}^{(5),\text{\MSbar}}(M_Z)}{1 -
+    \Delta\alpha_{\text{s},\text{SM}}(M_Z)
+    - \Delta\alpha_{\text{s},\text{susy}}(M_Z)} ,\\
+  \Delta\alpha_{\text{s},\text{SM}}(\mu) &=
+  \frac{\alpha_\text{s}}{2\pi} \left[
+    -\frac{2}{3} \log{\frac{m_t}{\mu}} \right],\\
+  \Delta\alpha_{\text{s},\text{susy}}(\mu) &=
+  \frac{\alpha_\text{s}}{2\pi}\left[ \frac{1}{2}-\sum_{\text{susy
+        particle } i} C_i \log{\frac{m_i}{\mu}} \right] ,\\
+  g_{3,\text{susy}}^{\text{\DRbar}}(M_Z) &=
+  \sqrt{4\pi\alpha_{\text{s},\text{susy}}^{\text{\DRbar}}(M_Z)} ,
+\end{align}
+%
+where
+%
+\begin{align}
+  \alpha_{\text{e.m.},\text{SM}}^{(5),\text{\MSbar}}(M_Z) &= 1/127.944,\\
+  \alpha_{\text{s},\text{SM}}^{(5),\text{\MSbar}}(M_Z) &= 0.1185,
+\end{align}
+%
+are the \MSbar electromagnetic and strong cougling constants in the
+Standard Model including only $5$ quark flavours
+\cite{Beringer:1900zz}.  Afterwards, SARAH's tree-level expressions
+for the Weinberg angle and the $U(1)_Y$ and $SU(2)_L$ gauge couplings
+$g_Y$ and $g_2$ in terms of $M_{W,\text{susy}}^{\text{\DRbar}}(M_Z)$,
+$M_{Z,\text{susy}}^{\text{\DRbar}}(M_Z)$ and
+$e_{\text{susy}}^{\text{\DRbar}}(M_Z)$ are used to calculate
+$g_{1,\text{susy}}^{\text{\DRbar}}(M_Z)$ and
+$g_{2,\text{susy}}^{\text{\DRbar}}(M_Z)$.  In the MSSM, for example,
+one has
+%
+\begin{align}
+  \theta_{W,\text{susy}}^{\text{\DRbar}}(M_Z) &=
+  \arcsin\sqrt{1 - \left(\frac{M_{W,\text{susy}}^{\text{\DRbar}}(M_Z)}{M_{Z,\text{susy}}^{\text{\DRbar}}(M_Z)}\right)^2} ,\\
+  g_{1,\text{susy}}^{\text{\DRbar}}(M_Z) &=
+  \sqrt{\frac{5}{3}} \frac{e_{\text{susy}}^{\text{\DRbar}}(M_Z)}{\cos\theta_{W,\text{susy}}^{\text{\DRbar}}(M_Z)} ,\\
+  g_{2,\text{susy}}^{\text{\DRbar}}(M_Z) &=
+  \frac{e_{\text{susy}}^{\text{\DRbar}}(M_Z)}{\sin\theta_{W,\text{susy}}^{\text{\DRbar}}(M_Z)} ,
+\end{align}
+%
+where
+%
+\begin{align}
+  \left(M_{W,\text{susy}}^{\text{\DRbar}}(M_Z)\right)^2 &=
+  \left(M_W^{\text{pole}}\right)^2 + \re \Pi_{WW}^T(p^2 = (M_W^{\text{pole}})^2) ,\\
+  \left(M_{Z,\text{susy}}^{\text{\DRbar}}(M_Z)\right)^2 &=
+  \left(M_Z^{\text{pole}}\right)^2 + \re \Pi_{ZZ}^T(p^2 = (M_Z^{\text{pole}})^2)
+\end{align}
+%
+and $M_W^{\text{pole}} = 80.404\unit{GeV}$, $M_Z^{\text{pole}} =
+91.1876\unit{GeV}$ \cite{Beringer:1900zz}.
+
+\subsubsection{Calculation of the Yukawa couplings $y_f(M_Z)$}
+
+At the low-scale constraint the expressions for the Yukawa coupling
+matrices $y_f(M_Z)$ are calculated from the tree-level mass matrices
+of the SM-fermions $m_u$, $m_d$, $m_e$.  In the MSSM, for example,
+these relations read
+%
+\begin{align}
+  y_u^{\text{\DRbar}}(M_Z) &= \frac{\sqrt{2} m_{u}^T}{v_u} , &
+  y_d^{\text{\DRbar}}(M_Z) &= \frac{\sqrt{2} m_{d}^T}{v_d} , &
+  y_e^{\text{\DRbar}}(M_Z) &= \frac{\sqrt{2} m_{e}^T}{v_d}.
+\end{align}
+%
+At the low-scale the fermion mass matrices are composed as
+%
+\begin{align}
+  m_u =
+  \begin{pmatrix}
+    m_{u}^{\userinput} & 0 & 0 \\
+    0 & m_{c}^{\userinput} & 0 \\
+    0 & 0 & m_{t,\text{susy}}^{\text{\DRbar}}(M_Z)
+  \end{pmatrix} ,\\
+  m_d =
+  \begin{pmatrix}
+    m_{d}^{\userinput} & 0 & 0 \\
+    0 & m_{s}^{\userinput} & 0 \\
+    0 & 0 & m_{b,\text{susy}}^{\text{\DRbar}}(M_Z)
+  \end{pmatrix} ,\\
+  m_e =
+  \begin{pmatrix}
+    m_{e}^{\userinput} & 0 & \\
+    0 & m_{\mu}^{\userinput} & 0 \\
+    0 & 0 & m_{\tau,\text{susy}}^{\text{\DRbar}}(M_Z)
+  \end{pmatrix},
+\end{align}
+%
+where $m_{u,c,d,s,e,\mu}^{\userinput}$ are read from the \code{Block
+  SMINPUTS} of the SLHA input file \cite{Skands:2003cj}.  The 3rd
+generation quark masses are calculated in the \DRbar scheme from the
+SLHA user input quantities $m_t^\text{pole}$,
+$m_{b,\text{SM}}^{\text{\MSbar}}(M_Z)$,
+$m_{\tau,\text{SM}}^{\text{\MSbar}}(M_Z)$ \cite{Skands:2003cj}.  In
+detail, the top quark \DRbar mass is calculated via
+%
+\begin{align}
+  \begin{split}
+    m_{t,\text{susy}}^{\text{\DRbar}}(\mu) &= m_t^\text{pole} +
+    \re\Sigma_{tt}^{S,\text{heavy}}(m_t^\text{pole}) \\
+    &\phantom{=\;} + m_t^\text{pole}
+    \left[ \re\Sigma_{tt}^{L,\text{heavy}}(m_t^\text{pole}) +
+      \re\Sigma_{tt}^{R,\text{heavy}}(m_t^\text{pole}) + \Delta
+      m_t^{(1),\text{qcd}} + \Delta m_t^{(2),\text{qcd}} \right] ,
+  \end{split}
+\end{align}
+%
+where the $\Sigma_{ff}^\text{heavy}$ denotes the self-energy of
+fermion $f$ without the Standard Model contributions.  The appearing
+QCD corrections $\Delta m_t^{(1),\text{qcd}}$ and $\Delta
+m_t^{(2),\text{qcd}}$ are taken from
+\cite[Eq.\ (58),(61)]{Bednyakov:2002sf} and read
+%
+\begin{align}
+  \Delta m_t^{(1),\text{qcd}} &= -\frac{g_3^2 \left(5-3 \log\left(\frac{m_t^2}{\mu^2}\right)\right)}{12 \pi^2},\\
+  \begin{split}
+    \Delta m_t^{(2),\text{qcd}} &= \left(\Delta
+      m_t^{(1),\text{qcd}}\right)^2 \\
+    &\phantom{=\;} - \frac{g_3^4 \left[396
+        \log^2\left(\frac{m_t^2}{\mu^2}\right)-1476
+        \log\left(\frac{m_t^2}{\mu^2}\right)-48 \zeta(3)+2011+16 \pi
+        ^2 (1+\log 4)\right]}{4608 \pi^4}.
+  \end{split}
+\end{align}
+%
+The \DRbar mass of the bottom quark is calculated via
+\cite{Baer:2002ek,Skands:2003cj}
+%
+\begin{align}
+  m_{b,\text{susy}}^{\text{\DRbar}}(\mu) &=
+  \frac{m_{b,\text{SM}}^{\text{\DRbar}}(\mu)}{1 -
+    \re\Sigma_{bb}^{S,\text{heavy}}(m_{b,\text{SM}}^\text{\MSbar})/m_b
+    - \re\Sigma_{bb}^{L,\text{heavy}}(m_{b,\text{SM}}^\text{\MSbar}) -
+    \re\Sigma_{bb}^{R,\text{heavy}}(m_{b,\text{SM}}^\text{\MSbar})} ,\\
+  m_{b,\text{SM}}^{\text{\DRbar}}(\mu) &=
+  m_{b,\text{SM}}^{\text{\MSbar}}(\mu) \left(1 - \frac{\alpha_s}{3
+      \pi} - \frac{23}{72} \frac{\alpha_s^2}{\pi^2} + \frac{3
+      g_2^2}{128 \pi^2} + \frac{13 g_Y^2}{1152 \pi^2}\right) ,
+\end{align}
+%
+where $m_{b,\text{SM}}^{\text{\MSbar}}(M_Z)$ is user input.  The
+\DRbar mass of the $\tau$ is calculated via
+%
+\begin{align}
+  \begin{split}
+    m_{\tau,\text{susy}}^{\text{\DRbar}}(\mu) &=
+    m_{\tau,\text{SM}}^{\text{\DRbar}}(\mu) +
+    \re\Sigma_{\tau\tau}^{S,\text{heavy}}(m_{\tau,\text{SM}}^\text{\MSbar}) \\
+    &\phantom{=\;} + m_{\tau,\text{SM}}^{\text{\DRbar}}(\mu) \left[
+      \re\Sigma_{\tau\tau}^{L,\text{heavy}}(m_{\tau,\text{SM}}^\text{\MSbar})
+      +
+      \re\Sigma_{\tau\tau}^{R,\text{heavy}}(m_{\tau,\text{SM}}^\text{\MSbar})
+    \right] ,
+  \end{split}\\
+  m_{\tau,\text{SM}}^{\text{\DRbar}}(\mu) &= m_{\tau,\text{SM}}^{\text{\MSbar}}(\mu)
+  \left(1 - 3 \frac{g_Y^2 - g_2^2}{128 \pi^2}\right),
+\end{align}
+%
+where $m_{\tau,\text{SM}}^{\text{\MSbar}}(M_Z)$ is user input.  To
+convert the fermion masses from the \MSbar to the \DRbar scheme the
+Yukawa coupling conversion from \cite[Eq.~(19)]{Skands:2003cj} was
+used and it was assumed that the VEV is already given in the \DRbar
+scheme.
 
@@
   %35 citations counted in INSPIRE as of 21 Sep 2013
+
+%\cite{Skands:2003cj}
+\bibitem{Skands:2003cj}
+  P.~Z.~Skands, B.~C.~Allanach, H.~Baer, C.~Balazs, G.~Belanger, F.~Boudjema, A.~Djouadi and R.~Godbole {\it et al.},
+  %``SUSY Les Houches accord: Interfacing SUSY spectrum calculators, decay packages, and event generators,''
+  JHEP {\bf 0407} (2004) 036
+  [hep-ph/0311123].
+  %%CITATION = HEP-PH/0311123;%%
+  %394 citations counted in INSPIRE as of 22 Feb 2014
+
 %\cite{Allanach:2008qq}
@@
 
-%
+%\cite{Beringer:1900zz}
+\bibitem{Beringer:1900zz}
+  J.~Beringer {\it et al.}  [Particle Data Group Collaboration],
+  %``Review of Particle Physics (RPP),''
+  Phys.\ Rev.\ D {\bf 86} (2012) 010001.
+  %%CITATION = PHRVA,D86,010001;%%
+  %3229 citations counted in INSPIRE as of 22 Feb 2014
+
+%\cite{Bednyakov:2002sf}
+\bibitem{Bednyakov:2002sf}
+  A.~Bednyakov, A.~Onishchenko, V.~Velizhanin and O.~Veretin,
+  %``Two loop O(alpha-s**2) MSSM corrections to the pole masses of heavy quarks,''
+  Eur.\ Phys.\ J.\ C {\bf 29} (2003) 87
+  [hep-ph/0210258].
+  %%CITATION = HEP-PH/0210258;%%
+  %35 citations counted in INSPIRE as of 22 Feb 2014
+
+%\cite{Baer:2002ek}
+\bibitem{Baer:2002ek}
+  H.~Baer, J.~Ferrandis, K.~Melnikov and X.~Tata,
+  %``Relating bottom quark mass in DR-BAR and MS-BAR regularization schemes,''
+  Phys.\ Rev.\ D {\bf 66} (2002) 074007
+  [hep-ph/0207126].
+  %%CITATION = HEP-PH/0207126;%%
+  %55 citations counted in INSPIRE as of 22 Feb 2014
 

### 2014-02-27 doc/paper.tex
sketching the meta code section in the paper

@@
 \newcommand{\userinput}{\text{<input>}}
+\newcommand{\pole}{\text{pole}}
 \DeclareMathOperator{\sign}{sign}
@@
 
+\section{Setting up a FlexibleSUSY model}
+
+A \fs model is configured via a (Mathematica) model file, which in
+case of the MSSM looks like this
+%
+\begin{lstlisting}[language=Mathematica]
+FSModelName = "MSSM";
+
+MINPAR = {
+   {1, m0},
+   {2, m12},
+   {3, TanBeta},
+   {4, Sign[\[Mu]]},
+   {5, Azero}
+};
+
+DefaultParameterPoint = {
+   {m0, 125},
+   {m12, 500},
+   {TanBeta, 10},
+   {Sign[\[Mu]], 1},
+   {Azero, 0}
+};
+
+EWSBOutputParameters = { B[\[Mu]], \[Mu] };
+
+HighScale = g1 == g2;
+
+HighScaleFirstGuess = 1.0 10^14;
+
+HighScaleInput = {
+   {T[Ye], Azero*Ye},
+   {T[Yd], Azero*Yd},
+   {T[Yu], Azero*Yu},
+   {mHd2, m0^2},
+   {mHu2, m0^2},
+   {mq2, UNITMATRIX[3] m0^2},
+   {ml2, UNITMATRIX[3] m0^2},
+   {md2, UNITMATRIX[3] m0^2},
+   {mu2, UNITMATRIX[3] m0^2},
+   {me2, UNITMATRIX[3] m0^2},
+   {MassB, m12},
+   {MassWB, m12},
+   {MassG, m12}
+};
+
+SUSYScale = Sqrt[M[Su[1]]*M[Su[6]]];
+
+SUSYScaleFirstGuess = Sqrt[m0^2 + 4 m12^2];
+
+SUSYScaleInput = {};
+
+LowScale = SM[MZ];
+
+LowScaleFirstGuess = SM[MZ];
+
+LowScaleInput = {
+   {vd, 2 MZDRbar / Sqrt[GUTNormalization[g1]^2 g1^2 + g2^2] Cos[ArcTan[TanBeta]]},
+   {vu, 2 MZDRbar / Sqrt[GUTNormalization[g1]^2 g1^2 + g2^2] Sin[ArcTan[TanBeta]]}
+};
+
+InitialGuessAtLowScale = {
+   {vd, SM[vev] Cos[ArcTan[TanBeta]]},
+   {vu, SM[vev] Sin[ArcTan[TanBeta]]}
+};
+
+InitialGuessAtHighScale = {
+   {\[Mu]   , 1.0},
+   {B[\[Mu]], 0.0}
+};
+
+OnlyLowEnergyFlexibleSUSY = False; (* default *)
+\end{lstlisting}
+%
+The variable \code{FSModelName} contains the name of the \fs model.
+
+All non-Standard Model input variables must be specified in the lists
+\code{MINPAR} and \code{EXTPAR}.  Theses two variables refer to the
+MINPAR and EXTPAR blocks in a SLHA input file \cite{Skands:2003cj}.
+The list elements are two-component lists where the first entry is the
+SLHA index in the MINPAR or EXTPAR blocks.  The second entry is the
+name of the parameter.  The input parameters can be set to default
+values in the \code{DefaultParameterPoint} list, see above.
+
+Using the the variable \code{EWSBOutputParameters} the user can
+specify the model parameters that are output of the electroweak
+symmetry breaking conditions.  When doing EWSB, \fs will adjust these
+parameters until the EWSB conditons are fulfilled.  In the CMSSM these
+are \code{Mu} and \code{B[\[Mu]]}.
+
+Furthermore, the user has to specify the form of three model
+constraints: low-scale, susy-scale and high-scale constraints.  In \fs
+they are named as \code{LowScale}, \code{SUSYScale} and
+\code{HighScale}.  For each constraint there is a scale definition
+(named after the constraint), an initial guess for the scale
+(concatenation of the constraint name and \code{FirstGuess}) and a
+list of settings to be applied at the constraint (concatenation of the
+constraint name and \code{Input}).
+%
+\begin{itemize}
+\item \emph{High-scale constraint:} The high-scale constraint is
+  usuall the GUT-scale constraint, applied at the scale where the
+  gauge couplings $g_1$ and $g_2$ unify.  The high-scale can be
+  defined by an equation of the form \code{g1 == g2} or by fixed scale
+  value.
+\item \emph{Susy-scale constraint:} The susy-scale constraint is the
+  constraint where \fs will impose the EWSB conditions and calculate
+  the spectrum.  It is usually defined by
+  $\sqrt{m_{\tilde{t}_1}m_{\tilde{t}_2}}$.
+\item \emph{Low-scale constraint:} The low-scale constraint is the
+  constraint where the gauge couplings $g_1$, $g_2$, $g_3$ and the
+  Yukawa couplings $y_f$ will be calculated from given Standard Model
+  input $\alpha_{\text{e.m.}}(M_Z)$, $\alpha_{s}(M_Z)$, $m_f^\pole$,
+  $M_Z^\pole$, $M_W^\pole$, see
+  Sec.~\ref{sec:calculation-of-gauge-couplings}--\ref{sec:calculation-of-yukawa-couplings}.
+  This should be done at the $Z$ mass scale $M_Z$.  Note that the
+  gauge and Yukawa couplings are calculated automatically and
+  therefore do not need to be specified in the \code{LowScaleInput}
+  variable.
+\end{itemize}
+
+Finally the user can set an initial guess for the model parameters at
+the low- and high-scale using the variables
+\code{InitialGuessAtLowScale} and \code{InitialGuessAtHighScale},
+respectively.  Note, that the gauge and Yukawa couplings will be
+guessed automatically at the low-scale from the known Standard Model
+parameters.
+
+To create a low-energy model one has to set
+\code{OnlyLowEnergyFlexibleSUSY = True}.  In this case the high-scale
+constraint will be ignored and only the low-scale and susy-scale
+constraints are kept.  All model parameters that are not specified in
+\code{MINPAR} or \code{EXTPAR} will be read from the corresponding
+input blocks in the SLHA input file and will be set at the susy-scale.
+
 \section{Stucture of the spectrum generator}
@@
 \subsubsection{Calculation of the gauge couplings $g_i(M_Z)$}
+\label{sec:calculation-of-gauge-couplings}
 
@@
 \subsubsection{Calculation of the Yukawa couplings $y_f(M_Z)$}
+\label{sec:calculation-of-yukawa-couplings}
 

### 2014-02-28 doc/paper.tex
adding diagonalization conventions

@@
 \newcommand{\pole}{\text{pole}}
+\newcommand{\Lagr}{\mathcal{L}}
 \DeclareMathOperator{\sign}{sign}
@@
 columns.
+
+The mass matrix $M^2$ for real scalar fiels $\phi_i$ is diagonalized
+with an orthogonal matrix $O$ via
+%
+\begin{align}
+  \Lagr_{m,\text{real scalar}}
+  &= - \frac{1}{2} \phi^T M^2 \phi
+  = - \frac{1}{2} (\phi^m)^T M^2_D \phi^m, \\
+  \qquad M^2 &= (M^2)^T ,
+  \qquad \phi^m = O \phi ,
+  \qquad M^2_D = O M^2 O^T ,
+  \qquad O^T O = 1 ,
+\end{align}
+%
+where $M^2_D$ is diagonal and $\phi^m_i$ are the mass eigenstates.  In
+case of complex scalar fields $\phi_i$ we use
+%
+\begin{align}
+  \Lagr_{m,\text{complex scalar}}
+  &= - \phi^\dagger M^2 \phi
+  = - (\phi^m)^\dagger M^2_D \phi^m, \\
+  \qquad M^2 &= (M^2)^\dagger ,
+  \qquad \phi^m = U \phi ,
+  \qquad M^2_D = U M^2 U^\dagger ,
+  \qquad U^\dagger U = 1 .
+\end{align}
+%
+A symmetric mass matrix $Y$ for Weyl spinors $\psi_i$ is diagonalized
+as
+%
+\begin{align}
+  \Lagr_{m,\text{fermion}}^\text{symm.}
+  &= - \frac{1}{2} \psi^T Y \psi + \text{h.c.}
+  = - \frac{1}{2} \chi^T Y_D \chi + \text{h.c.}, \\
+  \qquad Y &= Y^T ,
+  \qquad Y_D = Z Y Z^\dagger ,
+  \qquad \chi = Z \psi ,
+  \qquad Z^\dagger Z = 1 ,
+\end{align}
+%
+where $Y_D$ is diagonal and $\chi_i$ are the mass eigenstates.  The
+phases of $Z$ are chosen such that all mass eigenvalues are positive.
+In case of a non-symmetric mass matrix $X$ for Weyl spinors $\psi_i$
+we use
+%
+\begin{align}
+  \Lagr_{m,\text{fermion}}^\text{svd}
+  &= - (\psi^-)^T X \psi^+ + \text{h.c.}
+  = - (\chi^-)^T X_D \chi^+ + \text{h.c.}, \\
+  \qquad \chi^+ &= V \psi^+ ,
+  \qquad \chi^- = U \psi^- ,
+  \qquad X_D = U^* X V^{-1} ,
+  \qquad U^\dagger U = 1 = V^\dagger V ,
+\end{align}
+%
+where we're again chosing the phases of $U$ and $V$ such that all mass
+eigenvalues are positive.
 

### 2014-02-28 doc/paper.tex
Fixing a typo in convention of symmetric diagonalization
Clarifying uses of diagonalization routines

@@
 the mass matrices returned from \code{SARAH`MassMatrix[]}.  The
-numerical diagonalization is performed by the Eigen library routines
-\code{Eigen::JacobiSVD} and \code{Eigen::SelfAdjointEigenSolver} for
-small matrices, and the Lapack routines \code{zgesvd}, \code{dgesvd},
-\code{zheev}, \code{dsyev} for matrices with more than four rows and
-columns.
+numerical singular value decomposition is performed by the Eigen
+library routine \code{Eigen::JacobiSVD} for matrices with less than
+four rows and columns, and the Lapack routines \code{zgesvd},
+\code{dgesvd} for larger matrices.  For the other types of
+diagonalization, \code{Eigen::SelfAdjointEigenSolver} from Eigen is
+used regardless of the matrix size.
 
@@
 %
-A symmetric mass matrix $Y$ for Weyl spinors $\psi_i$ is diagonalized
-as
+A (possibly complex) symmetric mass matrix $Y$ for Weyl spinors
+$\psi_i$ is diagonalized as
 %
@@
   \qquad Y &= Y^T ,
-  \qquad Y_D = Z Y Z^\dagger ,
+  \qquad Y_D = Z^* Y Z^\dagger ,
   \qquad \chi = Z \psi ,

### 2014-03-01 doc/paper.tex
describe the three differend diagonalization methods (pole masses)

@@
 \newcommand{\Lagr}{\mathcal{L}}
+\newcommand{\unity}{\mathbf{1}}
 \DeclareMathOperator{\sign}{sign}
@@
       tadpole[1] -= Re(model->tadpole_hh(1));
-
    }
@@
   \qquad M^2_D = O M^2 O^T ,
-  \qquad O^T O = 1 ,
+  \qquad O^T O = \unity ,
 \end{align}
@@
   \qquad M^2_D = U M^2 U^\dagger ,
-  \qquad U^\dagger U = 1 .
+  \qquad U^\dagger U = \unity .
 \end{align}
@@
   \qquad \chi = Z \psi ,
-  \qquad Z^\dagger Z = 1 ,
+  \qquad Z^\dagger Z = \unity ,
 \end{align}
@@
   \qquad X_D = U^* X V^{-1} ,
-  \qquad U^\dagger U = 1 = V^\dagger V ,
+  \qquad U^\dagger U = \unity = V^\dagger V ,
 \end{align}
@@
 full one-loop self-energies and tree-level mass matrices to calculate
-the pole masses at the one-loop level.  If higher accuracy is required
-additional routines with higher order corrections can be added in a
-very simple way. For example in the MSSM by default \fs adds two-loop
-Higgs FORTRAN routines supplied by P.~Slavich from
-\cite{slavich-degrassi} to add two loop corrections of $\oatas$,
-$\oabas$, $\oatq$, $\oabatau$, $\oabq$, $\oatauq$ and $\oatab$ as
-follows... {\color{red} This is not yet implemented!}
+the pole masses at the one-loop level, which means finding the
+eigenvalues $p$ that solve the following equations:
+%
+\begin{align}
+  &\text{scalars } \phi:  & 0 &= p^2\unity - m^2_{\phi,1L}(p^2), \\
+  &\text{fermions } \xi: & 0 &= p\unity - m_{\xi,1L}(p^2),
+\end{align}
+%
+where the one-loop mass matrices are given by
+%
+\begin{align}
+  &\text{scalars } \phi: &
+  m^2_{\phi,1L}(p^2) &= m^2_{\phi} - \Sigma_\phi(p^2), \\
+  &\text{majorana fermions } \chi: &
+  m_{\chi,1L}(p^2) &= m_{\chi} - \frac{1}{2}\Big[
+    \Sigma_\chi^S(p^2) + \Sigma_\chi^{S,T}(p^2)
+    + \Big( \Sigma_\chi^{L,T}(p^2) + \Sigma_\chi^R(p^2) \Big) m_{\chi} \notag \\
+    &&&\phantom{= m_{\chi} - \frac{1}{2}\Big[}
+    + m_{\chi} \Big( \Sigma_\chi^L(p^2) + \Sigma_\chi^{R,T}(p^2) \Big)
+  \Big], \\
+  &\text{dirac fermions } \psi: &
+  m_{\psi,1L}(p^2) &= m_{\psi}
+  - \Sigma_\psi^S(p^2)
+  - \Sigma_\psi^R(p^2) m_{\psi}
+  - m_{\psi} \Sigma_\psi^L(p^2) .
+\end{align}
+%
+Since the one-loop mass matrices depend on $p$, an iterative procedure
+must be used.  \fs provides three different methods with different
+precision and execution speed to determine the mass eigenvalues:
+
+\code{LowDiagonalizationPrecision}: This option provides the lowest
+precision but is also the fastest one.  Here the one-loop mass matrix
+$m_{\phi,1L,\text{low}}$ is calculated exactly once as
+%
+\begin{align}
+  (m_{\phi,1L,\text{low}})_{ij} = (m_{\phi,1L}(p^2 = m_{\phi_i} m_{\phi_j}))_{ij} ,
+\end{align}
+%
+where $m_{\phi_i}$ is the $i$th mass eigenvalue of the tree-level mass
+matrix $m_\phi$.  Afterwards $m_{\phi,1L,\text{low}}$ is diagonalized
+and the eigenvalues are interpreted as pole masses $m_{\phi_i}^\pole$.
+
+\code{MediumDiagonalizationPrecision} (default): This option provides
+calculation with medium precision with a medium execution time.  Here
+the one-loop mass matrix $m_{\phi,1L,\text{medium}}$ is calculated $n$
+times as
+%
+\begin{align}
+  (m_{\phi,1L,\text{medium}})_{ij}^{(k)} = (m_{\phi,1L}(p^2 = m_{\phi_k}^2))_{ij} ,
+  \qquad k = 1,\ldots,n ,
+\end{align}
+%
+where $m_{\phi_k}$ is the $k$th mass eigenvalue of the tree-level mass
+matrix $m_\phi$.  Afterwards, each mass matrix
+$(m_{\phi,1L,\text{medium}})^{(k)}$ is diagonalized and the $k$th
+eigenvalue is interpreted as pole mass $m_{\phi_k}^\pole$.
+
+\code{HighDiagonalizationPrecision}: This option provides
+diagonalization with highest precision but has also the highest
+execution time.  Here the one-loop mass matrix
+$m_{\phi,1L,\text{high}}$ is diagonalized $n$ times as in the case of
+\code{MediumDiagonalizationPrecision}, resulting in $n$ pole masses
+$m_{\phi_k}^\pole$ ($k = 1,\ldots,n$).  Afterwards, the
+diagonalization is repeated, this time using the so calculated pole
+masses $m_{\phi_k}^\pole$ for the momentum calculation $p^2 =
+(m_{\phi_k}^\pole)^2$, until convergence is reached.
+
+If higher accuracy is required additional routines with higher order
+corrections can be added in a very simple way. For example in the MSSM
+by default \fs adds two-loop Higgs FORTRAN routines supplied by
+P.~Slavich from \cite{slavich-degrassi} to add two loop corrections of
+$\oatas$, $\oabas$, $\oatq$, $\oabatau$, $\oabq$, $\oatauq$ and
+$\oatab$ as follows... {\color{red} This is not yet implemented!}
 

### 2014-03-05 test/test_MSSM.hpp
generalize n-loop EWSB functions

@@ -61,7 +61,18 @@ void ensure_one_loop_ewsb(MSSM<Two_scale>& m)
    TEST_CLOSE(m.get_ewsb_eq_vu() - m.tadpole_hh(1).real(), 0.0, precision);
 }
 
-void ensure_one_loop_ewsb(MssmSoftsusy& s)
+void ensure_n_loop_ewsb(MSSM<Two_scale>& m, int loop_level)
+{
+   ensure_tree_level_ewsb(m);
+
+   const double precision = m.get_ewsb_iteration_precision();
+   m.set_pole_mass_loop_order(loop_level);
+   m.solve_ewsb();
+   TEST_CLOSE(m.get_ewsb_eq_vd() - m.tadpole_hh(0).real(), 0.0, precision);
+   TEST_CLOSE(m.get_ewsb_eq_vu() - m.tadpole_hh(1).real(), 0.0, precision);
+}
+
+void ensure_n_loop_ewsb(MssmSoftsusy& s, int loop_level)
 {
    ensure_tree_level_ewsb(s);
 
@@ -69,7 +80,7 @@ void ensure_one_loop_ewsb(MssmSoftsusy& s)
    const int signMu = s.displaySusyMu() >= 0.0 ? 1 : -1;
    const double mtrun = s.displayDrBarPars().mt;
    const DoubleVector pars(3);
-   softsusy::numRewsbLoops = 1;
+   softsusy::numRewsbLoops = loop_level;
    s.rewsb(signMu, mtrun, pars);
 }
 

### 2014-03-05 test/test_MSSM_model.cpp

@@ -1060,8 +1060,8 @@ void compare_loop_masses(MssmSoftsusy s, MSSM<Two_scale> m)
    TEST_CLOSE(m.get_Mu(), s.displaySusyMu(), 1.0e-10);
    TEST_CLOSE(m.get_BMu(), s.displayM3Squared(), 1.0e-10);
 
-   ensure_one_loop_ewsb(m);
-   ensure_one_loop_ewsb(s);
+   ensure_n_loop_ewsb(m, 1);
+   ensure_n_loop_ewsb(s, 1);
    // check that the important scalar potential parameters are equal
    TEST_CLOSE(m.get_mHd2(), s.displayMh1Squared(), 1.0e-10);
    TEST_CLOSE(m.get_mHu2(), s.displayMh2Squared(), 1.0e-10);

### 2014-03-05 test/test_MSSM.hpp
adding two-loop Higgs mass tests

@@ -51,16 +51,6 @@ void ensure_tree_level_ewsb(MssmSoftsusy& softSusy)
    TEST_CLOSE(-2 * BMu, (mHd2 - mHu2) * tan(2*beta) + sqr(MZrun) * sin(2*beta), 1.0e-10);
 }
 
-void ensure_one_loop_ewsb(MSSM<Two_scale>& m)
-{
-   ensure_tree_level_ewsb(m);
-
-   const double precision = m.get_ewsb_iteration_precision();
-   m.solve_ewsb_one_loop();
-   TEST_CLOSE(m.get_ewsb_eq_vd() - m.tadpole_hh(0).real(), 0.0, precision);
-   TEST_CLOSE(m.get_ewsb_eq_vu() - m.tadpole_hh(1).real(), 0.0, precision);
-}
-
 void ensure_n_loop_ewsb(MSSM<Two_scale>& m, int loop_level)
 {
    ensure_tree_level_ewsb(m);
@@ -68,8 +58,16 @@ void ensure_n_loop_ewsb(MSSM<Two_scale>& m, int loop_level)
    const double precision = m.get_ewsb_iteration_precision();
    m.set_pole_mass_loop_order(loop_level);
    m.solve_ewsb();
-   TEST_CLOSE(m.get_ewsb_eq_vd() - m.tadpole_hh(0).real(), 0.0, precision);
-   TEST_CLOSE(m.get_ewsb_eq_vu() - m.tadpole_hh(1).real(), 0.0, precision);
+
+   if (loop_level == 1) {
+      TEST_CLOSE(m.get_ewsb_eq_vd() - m.tadpole_hh(0).real(), 0.0, precision);
+      TEST_CLOSE(m.get_ewsb_eq_vu() - m.tadpole_hh(1).real(), 0.0, precision);
+   }
+}
+
+void ensure_one_loop_ewsb(MSSM<Two_scale>& m)
+{
+   ensure_n_loop_ewsb(m, 1);
 }
 
 void ensure_n_loop_ewsb(MssmSoftsusy& s, int loop_level)

### 2014-03-05 test/test_MSSM_model.cpp

@@ -1070,7 +1070,7 @@ void compare_loop_masses(MssmSoftsusy s, MSSM<Two_scale> m)
    m.calculate_pole_masses();
    s.physical(1);
 
-   const DoubleVector hh(ToDoubleVector(m.get_physical().Mhh));
+   DoubleVector hh(ToDoubleVector(m.get_physical().Mhh));
    TEST_CLOSE(s.displayPhys().mh0(1), hh(1), 0.114);
    TEST_CLOSE(s.displayPhys().mh0(2), hh(2), 0.04);
    TEST_CLOSE_REL(s.displayPhys().mh0(1), hh(1), 0.00115);
@@ -1118,6 +1118,27 @@ void compare_loop_masses(MssmSoftsusy s, MSSM<Two_scale> m)
    TEST_CLOSE_REL(s.displayPhys().md.flatten().sort()(4), m.get_physical().MSd(3), 0.000026);
    TEST_CLOSE_REL(s.displayPhys().md.flatten().sort()(5), m.get_physical().MSd(4), 0.000026);
    TEST_CLOSE_REL(s.displayPhys().md.flatten().sort()(6), m.get_physical().MSd(5), 0.00005);
+
+   // test two-loop corrections
+   softsusy::numRewsbLoops = 2;
+   softsusy::numHiggsMassLoops = 2;
+   m.set_pole_mass_loop_order(2);
+   m.set_ewsb_loop_order(2);
+   ensure_n_loop_ewsb(m, 2);
+   ensure_n_loop_ewsb(s, 2);
+   // check that the important scalar potential parameters are equal
+   TEST_CLOSE(m.get_mHd2(), s.displayMh1Squared(), 1.0e-7);
+   TEST_CLOSE(m.get_mHu2(), s.displayMh2Squared(), 1.0e-8);
+   TEST_CLOSE_REL(m.get_Mu(), s.displaySusyMu(), 0.0000001);
+   TEST_CLOSE_REL(m.get_BMu(), s.displayM3Squared(), 0.04);
+   m.calculate_pole_masses();
+   s.physical(2);
+
+   hh = ToDoubleVector(m.get_physical().Mhh);
+   TEST_CLOSE(s.displayPhys().mh0(1), hh(1), 0.013);
+   TEST_CLOSE(s.displayPhys().mh0(2), hh(2), 0.014);
[…]

### 2014-03-05 doc/paper.tex
improve SLHA input file option documentation

@@
 The in more detail the two-scale algorithm used in \fs works as
-follows:
+follows (see \figurename~\ref{fig:two-scale-algorithm}):
 %
@@
 \item Apply user-defined low-scale constraint (\code{LowScaleInput})
-\item Guess Yukawa couplings $y^{u,d,e}_{ij}$ at $m_Z$
+\item Guess Yukawa couplings $y_{u,d,e}$ at $m_Z$
 \item Run model to the high-scale (\code{HighScaleFirstGuess})
@@
   \item Recalculate low-scale
-  \item Calculate \DRbar\ gauge and Yukawa couplings $g_{1,2,3}$, $y^{u,d,e}_{ij}$
+  \item Calculate \DRbar\ gauge and Yukawa couplings $g_{1,2,3}$, $y_{u,d,e}$
   \item Apply user-defined low-scale constraint (\code{LowScaleInput})
@@
 \end{enumerate}
+%
+\begin{figure}[tbh]
+  \centering
+  \begin{tikzpicture}[node distance = 2cm, auto]
+    \tikzstyle{block} = [rectangle, draw, text width=16em, text centered, minimum height=3em]
+    \tikzstyle{arrow} = [draw, -latex, thick]
+    \node[block] (guess) {Guess $g_i(M_Z)$, $y_f(M_Z)$ and soft
+      parameters at \code{LowScale}};
+    \node[block,below of=guess] (MZ) {Calculate $g_i(M_Z)$, $y_f(M_Z)$ and apply
+      low-scale boundary conditions (\code{LowScaleInput})};
+    \path[arrow] (guess) -- node {run to \code{LowScale}} (MZ);
+    \node[block,below of=MZ] (MX) {Apply high-scale boundary conditions
+      (\code{HighScaleInput})};
+    \path[arrow] (MZ) -- node {run to \code{HighScale}} (MX);
+    \node[block,below of=MX] (MS) {Apply susy-scale boundary conditions
+      (\code{SUSYScaleInput}) and solve EWSB};
+    \path[arrow] (MX) -- node {run to \code{SUSYScale}} (MS);
+    %\path[arrow] (MS.east) -| node {run to \code{LowScale}} (MZ.east);
+    \path[-latex, thick] (MS.east) edge[bend right=90] node[right] {run to \code{LowScale}} (MZ.east);
+    \node[block,below of=MS] (spec) {Calculate pole masses};
+    \path[arrow,dashed] (MS) -- node[text width=16em] {on convergence\\ run to \code{SUSYScale}} (spec);
+  \end{tikzpicture}
+  \caption{Iterative two-scale algorithm to calculate the spectrum.}
+  \label{fig:two-scale-algorithm}
+\end{figure}
 
@@
     3   0                    # calculate SM pole masses
-    4   1                    # calculate loop masses
+    4   1                    # pole mass loop order
     5   1                    # EWSB loop order

### 2014-03-17 doc/paper.tex
update benchmark numbers with Higgs two-loop corrections

@@
 We compare the run-time of the four different spectrum generators
-FlexibleSUSY (version 0.5.3), Softsusy (version 3.4.0), SPheno
+FlexibleSUSY (version 0.5.4), Softsusy (version 3.4.0), SPheno
 (3.2.4), SPhenoMSSM (generated with SARAH 4.1.0 and linked against
 SPheno 3.2.4).  FlexibleSUSY and Softsusy were compiled with g++ 4.8.0
-and gfortran 4.8.0.  SPheno and SPhenoMSSM were compiled with Intel
-ifort 13.1.3 20130607\footnote{We are using Intel's ifort compiler,
-  instead of gfortran 4.8.0, because it decreases the run-time of
+and Intel ifort 13.1.3 20130607.  SPheno and SPhenoMSSM were compiled
+with Intel ifort 13.1.3 20130607\footnote{We are using Intel's ifort
+  compiler, instead of gfortran, because it decreases the run-time of
   SPheno and SPhenoMSSM by approximately a factor $1.5$.}.  We're
@@
     \midrule
-    FlexibleSUSY 0.5.3    & $0.151\unit{s}$    & $0.114\unit{s}$   & $0.081\unit{s}$\\
-    Softsusy 3.4.0        & $0.178\unit{s}$    & $0.174\unit{s}$   & $0.157\unit{s}$\\
-    SPheno 3.2.4          & $0.120\unit{s}$    & $0.118\unit{s}$   & $0.108\unit{s}$\\
-    SPhenoMSSM 4.1.0${}^*$ & $0.463\unit{s}$    & $0.450\unit{s}$   & $0.403\unit{s}$\\
+    FlexibleSUSY 0.5.4    & $0.151\unit{s}$    & $0.114\unit{s}$   & $0.091\unit{s}$\\
+    Softsusy 3.4.0        & $0.174\unit{s}$    & $0.169\unit{s}$   & $0.162\unit{s}$\\
+    SPheno 3.2.4          & $0.119\unit{s}$    & $0.115\unit{s}$   & $0.108\unit{s}$\\
+    SPhenoMSSM 4.1.0${}^*$ & $0.416\unit{s}$    & $0.401\unit{s}$   & $0.366\unit{s}$\\
     \bottomrule
@@
     FlexibleSUSY and Softsusy were compiled
-    with g++ 4.8.0 and gfortran 4.8.0.  SPheno and SPhenoMSSM were
+    with g++ 4.8.0 and ifort 13.1.3 20130607.
+    SPheno, SPhenoMSSM were
     compiled with ifort 13.1.3 20130607.
-    ${}^*$ SPhenoMSSM was generated with SARAH 4.1.0 and linked
+    ${}^*$SPhenoMSSM was generated with SARAH 4.1.0 and linked
     against SPheno 3.2.4.}
@@
 We find that on the single-core machine SPheno is fastest with
-$0.120\unit{s}$, and FlexibleSUSY is second with $0.151\unit{s}$.  In
+$0.119\unit{s}$, and FlexibleSUSY is second with $0.151\unit{s}$.  In
 the case of 2 or 6 CPU cores, FlexibleSUSY is fastest with
-$0.114\unit{s}$ and $0.081\unit{s}$, respectively.  The reason is that
+$0.114\unit{s}$ and $0.091\unit{s}$, respectively.  The reason is that
 FlexibleSUSY calculates each pole mass in a separate thread, and

### 2014-03-30 meta/FlexibleSUSY.m
setting first guess for MX to 2 10^16
This improves convergence, because the new guessed MX is closer to the
final value.

@@ -145,7 +145,7 @@ CheckModelFileSettings[] :=
            If[!ValueQ[FlexibleSUSY`HighScaleFirstGuess],
               Print["Warning: FlexibleSUSY`HighScaleFirstGuess should be",
                     " set in the model file!"];
-              FlexibleSUSY`HighScaleFirstGuess = 1.0 10^14;
+              FlexibleSUSY`HighScaleFirstGuess = 2.0 10^16;
              ];
            If[Head[FlexibleSUSY`HighScaleInput] =!= List,
               FlexibleSUSY`HighScaleInput = {};

### 2014-03-30 models/SoftsusyMSSM/SoftsusyMSSM_parameter_point.hpp

@@ -28,7 +28,7 @@ struct SoftsusyMSSM_parameter_point {
       : m0(125.0)
       , m12(500.0)
       , a0(0.0)
-      , mxGuess(1.9e16)
+      , mxGuess(2.0e16)
       , msGuess(1000.)
       , signMu(1)
       , tanBeta(10.0)

### 2014-03-30 test/test_MSSM_initial_guesser.cpp

@@ -94,7 +94,7 @@ BOOST_AUTO_TEST_CASE( test_initial_guess )
    else
       tanBeta = vu/vd;
    const double vev = Sqrt(Sqr(m.get_vu()) + Sqr(m.get_vd()));
-   BOOST_CHECK_CLOSE_FRACTION(smssm.displayTanb(), tanBeta, 6.5e-7);
+   BOOST_CHECK_CLOSE_FRACTION(smssm.displayTanb(), tanBeta, 1.5e-6);
    BOOST_CHECK_CLOSE_FRACTION(smssm.displayHvev(), vev    , 1.3e-6);
 
    BOOST_CHECK_CLOSE_FRACTION(smssm.displayGaugino(1), m.get_MassB() , 3.0e-7);

### 2014-03-30 test/test_MSSM_spectrum.cpp

@@ -600,7 +600,7 @@ BOOST_AUTO_TEST_CASE( test_MSSM_spectrum_with_Softsusy_gauge_couplings )
    BOOST_CHECK_CLOSE_FRACTION(fs.get_g3(), ss.displayGaugeCoupling(3), 0.00010);
 
    BOOST_CHECK_CLOSE_FRACTION(fs.get_Mu() , ss.displaySusyMu(), 0.0012);
-   BOOST_CHECK_CLOSE_FRACTION(fs.get_BMu(), ss.displayM3Squared(), 0.0023);
+   BOOST_CHECK_CLOSE_FRACTION(fs.get_BMu(), ss.displayM3Squared(), 0.0024);
    BOOST_CHECK_CLOSE_FRACTION(fs.get_mHd2(), ss.displayMh1Squared(), 0.0005);
    BOOST_CHECK_CLOSE_FRACTION(fs.get_mHu2(), ss.displayMh2Squared(), 0.0022);
 

### 2014-04-05 doc/paper.tex
setting first guess for MX to 2 10^16
This improves convergence, because the new guessed MX is closer to the
final value.
document how to enable the two-loop corrections in the (N)MSSM

@@
 
-HighScaleFirstGuess = 1.0 10^14;
+HighScaleFirstGuess = 2.0 10^16;
 
@@
 };
+
+UseHiggs2LoopMSSM = True;
+EffectiveMu = \[Mu];
 
@@
 parameters.
+
+\fs allows to add leading two-loop contributions to the CP-even Higgs
+tadpoles and self-energies.  For MSSM-like models (with two CP-even
+Higgs bosons, one CP-odd Higgs boson, one neutral Goldstone boson)
+these corrections can be enabled by setting \code{UseHiggs2LoopMSSM =
+  True;} in the model file and by defining the effectiv $\mu$-term
+\code{EffectiveMu = \[Mu]}.  This will add the zero-momentum
+corrections of the order $O(y_t^4 + y_b^2 y_t^2 + y_b^4)$, $O(y_t^2
+g_3^2)$, $O(y_b^2 g_3^2)$, $O(y_\tau^4)$, $O(y_\tau^2 y_b^2)$ from
+\cite{Degrassi:2001yf,Brignole:2001jy,Dedes:2002dy,Brignole:2002bz,Dedes:2003km}.
+For NMSSM-like models (with three CP-even Higgs bosons, two CP-odd
+Higgs bosons, one neutral Goldstone boson) the two-loop contributions
+are enabled by setting \code{UseHiggs2LoopNMSSM = True;} and by
+defining the effectiv $\mu$-term like \code{EffectiveMu = \[Lambda] vS
+  / Sqrt[2]}, for example.  This will add the the zero-momentum
+corrections of the order $O(y_t^2 g_3^2)$, $O(y_b^2 g_3^2)$ from
+\cite{Degrassi:2009yq}, plus leading MSSM-like contributions of the
+order $O(y_\tau^4)$, $O(y_t^4 + y_t^2 y_b^2 + y_b^4)$
+\cite{Brignole:2001jy,Dedes:2003km}.
 
@@
  HighScale = g1 == g2;
- HighScaleFirstGuess = 1.0 10^16;
+ HighScaleFirstGuess = 2.0 10^16;
 \end{lstlisting}
@@
 
+%\cite{Degrassi:2001yf}
+\bibitem{Degrassi:2001yf}
+  G.~Degrassi, P.~Slavich and F.~Zwirner,
+  %``On the neutral Higgs boson masses in the MSSM for arbitrary stop mixing,''
+  Nucl.\ Phys.\ B {\bf 611} (2001) 403
+  [hep-ph/0105096].
+  %%CITATION = HEP-PH/0105096;%%
+  %166 citations counted in INSPIRE as of 04 Apr 2014
+
+%\cite{Brignole:2001jy}
+\bibitem{Brignole:2001jy}
+  A.~Brignole, G.~Degrassi, P.~Slavich and F.~Zwirner,
+  %``On the O(alpha(t)**2) two loop corrections to the neutral Higgs boson masses in the MSSM,''
+  Nucl.\ Phys.\ B {\bf 631} (2002) 195
+  [hep-ph/0112177].
+  %%CITATION = HEP-PH/0112177;%%
+  %192 citations counted in INSPIRE as of 04 Apr 2014
+
+%\cite{Dedes:2002dy}
+\bibitem{Dedes:2002dy}
+  A.~Dedes and P.~Slavich,
+  %``Two loop corrections to radiative electroweak symmetry breaking in the MSSM,''
+  Nucl.\ Phys.\ B {\bf 657} (2003) 333
+  [hep-ph/0212132].
+  %%CITATION = HEP-PH/0212132;%%
+  %53 citations counted in INSPIRE as of 04 Apr 2014
+
+%\cite{Brignole:2002bz}
+\bibitem{Brignole:2002bz}
+  A.~Brignole, G.~Degrassi, P.~Slavich and F.~Zwirner,
+  %``On the two loop sbottom corrections to the neutral Higgs boson masses in the MSSM,''
+  Nucl.\ Phys.\ B {\bf 643} (2002) 79
+  [hep-ph/0206101].
+  %%CITATION = HEP-PH/0206101;%%
+  %152 citations counted in INSPIRE as of 04 Apr 2014
+
+%\cite{Dedes:2003km}
+\bibitem{Dedes:2003km}
+  A.~Dedes, G.~Degrassi and P.~Slavich,
+  %``On the two loop Yukawa corrections to the MSSM Higgs boson masses at large tan beta,''
+  Nucl.\ Phys.\ B {\bf 672} (2003) 144
+  [hep-ph/0305127].
+  %%CITATION = HEP-PH/0305127;%%
+  %99 citations counted in INSPIRE as of 04 Apr 2014
+
+%\cite{Degrassi:2009yq}
+\bibitem{Degrassi:2009yq}
+  G.~Degrassi and P.~Slavich,
+  %``On the radiative corrections to the neutral Higgs boson masses in the NMSSM,''
+  Nucl.\ Phys.\ B {\bf 825} (2010) 119
+  [arXiv:0907.4682 [hep-ph]].
+  %%CITATION = ARXIV:0907.4682;%%
+  %44 citations counted in INSPIRE as of 04 Apr 2014
+
 \end{thebibliography}

### 2014-04-05 doc/paper.tex
remove "Not yet implemented" remarks from the paper
reformulate calculation of pole masses

@@
 full one-loop self-energies and tree-level mass matrices to calculate
-the pole masses at the one-loop level, which means finding the
-eigenvalues $p$ that solve the following equations:
+the pole masses at the one-loop level, which means finding the values
+$p$ that solve the equation
 %
 \begin{align}
-  &\text{scalars } \phi:  & 0 &= p^2\unity - m^2_{\phi,1L}(p^2), \\
-  &\text{fermions } \xi: & 0 &= p\unity - m_{\xi,1L}(p^2),
+  0 = \det\left[p^2\unity - m^2_{f,1L}(p^2)\right],
 \end{align}
 %
-where the one-loop mass matrices are given by
+where the one-loop mass matrices $m_{f,1L}(p^2)$ are given in terms of
+the tree-level mass matrices $m_f$ and the self-energies
+$\Sigma_f(p^2)$ as
 %
@@
 precision but is also the fastest one.  Here the one-loop mass matrix
-$m_{\phi,1L,\text{low}}$ is calculated exactly once as
+$m_{f,1L}^\text{low}$ is calculated exactly once as
 %
 \begin{align}
-  (m_{\phi,1L,\text{low}})_{ij} = (m_{\phi,1L}(p^2 = m_{\phi_i} m_{\phi_j}))_{ij} ,
+  \forall i,j: (m_{f,1L}^\text{low})_{ij} = (m_{f,1L}(p^2 = m_{f_i} m_{f_j}))_{ij} ,
 \end{align}
 %
-where $m_{\phi_i}$ is the $i$th mass eigenvalue of the tree-level mass
-matrix $m_\phi$.  Afterwards $m_{\phi,1L,\text{low}}$ is diagonalized
-and the eigenvalues are interpreted as pole masses $m_{\phi_i}^\pole$.
+where $m_{f_i}$ is the $i$th mass eigenvalue of the tree-level mass
+matrix $m_f$.  Afterwards $m_{f,1L}^\text{low}$ is diagonalized and
+the eigenvalues are interpreted as pole masses $m_{f_i}^\pole$.
 
@@
 calculation with medium precision with a medium execution time.  Here
-the one-loop mass matrix $m_{\phi,1L,\text{medium}}$ is calculated $n$
+the one-loop mass matrix $m_{f,1L}^\text{medium}$ is calculated $n$
 times as
@@
 \begin{align}
-  (m_{\phi,1L,\text{medium}})_{ij}^{(k)} = (m_{\phi,1L}(p^2 = m_{\phi_k}^2))_{ij} ,
+  (m_{f,1L}^\text{medium})_{ij}^{(k)} = (m_{f,1L}(p^2 = m_{f_k}^2))_{ij} ,
   \qquad k = 1,\ldots,n ,
@@
 %
-where $m_{\phi_k}$ is the $k$th mass eigenvalue of the tree-level mass
-matrix $m_\phi$.  Afterwards, each mass matrix
-$(m_{\phi,1L,\text{medium}})^{(k)}$ is diagonalized and the $k$th
-eigenvalue is interpreted as pole mass $m_{\phi_k}^\pole$.
+where $m_{f_k}$ is the $k$th mass eigenvalue of the tree-level mass
+matrix $m_f$.  Afterwards, each mass matrix
+$(m_{f,1L}^\text{medium})^{(k)}$ is diagonalized and the $k$th
+eigenvalue is interpreted as pole mass $m_{f_k}^\pole$.
 
 \code{HighDiagonalizationPrecision}: This option provides
-diagonalization with highest precision but has also the highest
-execution time.  Here the one-loop mass matrix
-$m_{\phi,1L,\text{high}}$ is diagonalized $n$ times as in the case of
+diagonalization with highest precision, but has also the highest
+execution time.  Here the one-loop mass matrix $m_{f,1L}^\text{high}$
+is diagonalized $n$ times, as in the case of
 \code{MediumDiagonalizationPrecision}, resulting in $n$ pole masses
-$m_{\phi_k}^\pole$ ($k = 1,\ldots,n$).  Afterwards, the
-diagonalization is repeated, this time using the so calculated pole
-masses $m_{\phi_k}^\pole$ for the momentum calculation $p^2 =
-(m_{\phi_k}^\pole)^2$, until convergence is reached.
+$m_{f_k}^\pole$ ($k = 1,\ldots,n$).  Afterwards, the diagonalization
+is repeated, this time using the calculated pole masses
+$m_{f_k}^\pole$ for the momentum calculation $p^2 =
+(m_{f_k}^\pole)^2$.  The iteration stops until convergence is reached.
 
@@
 $\oatas$, $\oabas$, $\oatq$, $\oabatau$, $\oabq$, $\oatauq$ and
-$\oatab$ as follows... {\color{red} This is not yet implemented!}
+$\oatab$ as follows...
 
@@
 those of the MSSM or NMSSM, as has been done, for example, in the
-E$_6$SSM\cite{King:2005jy}.  {\color{red} This is not yet implemented!}
+E$_6$SSM\cite{King:2005jy}.
 

### 2014-04-05 doc/paper.tex
add note about GUT normalization of the gauge couplings in FS

@@
 \end{itemize}
+%
+Note, that \fs GUT-normalizes all gauge couplings.  Thus, the
+high-scale definition takes the simple form \code{g1 == g2}.  Another
+consequence is that in the calculation of the VEVs $v_u$ and $v_d$
+from $M_Z$ and $\tan\beta$ at the low-scale the GUT-normalization has
+to be taken into account, see above.
 

### 2014-04-05 doc/paper.tex
correcting display of the Mu parameter in the code snippets
correcting display of the Lambda parameter in the code snippets

@@
 parameters until the EWSB conditons are fulfilled.  In the CMSSM these
-are \code{Mu} and \code{B[\[Mu]]}.
+are \code{\\[Mu]} and \code{B[\\[Mu]]}.
 
@@
   True;} in the model file and by defining the effectiv $\mu$-term
-\code{EffectiveMu = \[Mu]}.  This will add the zero-momentum
+\code{EffectiveMu = \\[Mu]}.  This will add the zero-momentum
 corrections of the order $O(y_t^4 + y_b^2 y_t^2 + y_b^4)$, $O(y_t^2
@@
 are enabled by setting \code{UseHiggs2LoopNMSSM = True;} and by
-defining the effectiv $\mu$-term like \code{EffectiveMu = \[Lambda] vS
+defining the effectiv $\mu$-term like \code{EffectiveMu = \\[Lambda] vS
   / Sqrt[2]}, for example.  This will add the the zero-momentum

### 2014-04-07 src/slha_io.hpp
replace to_string by ToString to avoid another C++11 feature

@@ -226,8 +226,8 @@ void SLHA_io::set_block(const std::string& name,
       for (int k = 1; k <= NCols; ++k) {
          ss << boost::format(mixing_matrix_formatter) % i % k
             % Re(matrix(i-1,k-1))
-            % ("Re(" + symbol + "(" + std::to_string(i) + ","
-               + std::to_string(k) + "))");
+            % ("Re(" + symbol + "(" + ToString(i) + ","
+               + ToString(k) + "))");
       }
 
    set_block(ss);
@@ -249,7 +249,7 @@ void SLHA_io::set_block(const std::string& name,
    for (int i = 1; i <= rows; ++i)
       for (int k = 1; k <= cols; ++k) {
          ss << boost::format(mixing_matrix_formatter) % i % k % matrix(i-1,k-1)
-            % (symbol + "(" + std::to_string(i) + "," + std::to_string(k) + ")");
+            % (symbol + "(" + ToString(i) + "," + ToString(k) + ")");
       }
 
    set_block(ss);

### 2014-04-17 config/remove_export_markers.sh
Add a missing * in config/remove_export_markers.sh

@@ -4,7 +4,7 @@
 # The markers are removed as well.
 
 begin_marker="# *BEGIN: *NOT EXPORTED *#*"
-end_marker="# END: *NOT EXPORTED *#*"
+end_marker="# *END: *NOT EXPORTED *#*"
 
 if test $# -lt 1; then
     # read from stdin

### 2014-04-17 doc/paper.tex
document minimum required g++ version (4.4.7)
explain allowed constraint forms

@@
 \item SARAH, version 4.0.4 or higher \url{http://sarah.hepforge.org}
-\item C++11 compatible compiler (g++ 4.6.0 or higher, clang++ 3.1 or
+\item C++11 compatible compiler (g++ 4.4.7 or higher, clang++ 3.1 or
   higher)
@@
 they are named as \code{LowScale}, \code{SUSYScale} and
-\code{HighScale}.  For each constraint there is a scale definition
-(named after the constraint), an initial guess for the scale
-(concatenation of the constraint name and \code{FirstGuess}) and a
-list of settings to be applied at the constraint (concatenation of the
-constraint name and \code{Input}).
+\code{HighScale}.  For each constraint there is (i) a scale definition
+(named after the constraint), (ii) an initial guess for the scale
+(concatenation of the constraint name and \code{FirstGuess}) and (iii)
+a list of parameter settings to be applied at the scale (concatenation
+of the constraint name and \code{Input}).
 %
@@
 to be taken into account, see above.
+
+The list of parameter settings can contain as elements any of the
+following:
+%
+\begin{itemize}
+\item Two-component lists of the form \code{\{parameter, value\}},
+  which indicates that the \code{parameter} is set to \code{value} at
+  the defined scale.  If the \code{value} should be read from the SLHA
+  input file, it must be written as \code{LHInput[value]}.  Example:
+  %
+  \begin{lstlisting}
+SUSYScaleInput = {
+   {mHd2, m0^2},
+   {mHu2, LHInput[mHu2]}
+};
+  \end{lstlisting}
+  %
+  In this example the parameter \code{mHd2} is set to the value of
+  \code{m0^2}, and \code{mHu2} is set to the value given in the SLHA
+  input file in block \code{MSOFTIN}, entry 22.  The SLHA block names
+  and keys for the MSSM and NMSSM are defined in SARAH's
+  \code{parameters.m} file, see the SARAH manual or
+  \cite{Staub:2010jh}.
+
+\item The function \code{FSMinimize[parameters, function]} can be
+  given, where \code{parameters} is a list of model parameters and
+  \code{function} is a function of these parameters.
+  \code{FSMinimize[parameters, function]} will numericall vary the
+  \code{parameters} until the \code{function} is minimized.  Example:
+  %
+  \begin{lstlisting}
+FSMinimize[{vd,vu},
+           (SM[MZ] - Pole[M[VZ]])^2 / STANDARDDEVIATION[MZ]^2 +
+           (SM[MH] - Pole[M[hh[1]]])^2 / STANDARDDEVIATION[MH]^2]
+  \end{lstlisting}
+  %
+  Here, the parameters \code{vu} and \code{vd} are varied until the
+  function
+  %
+  \begin{align}
+    \chi^2(v_d,v_u) =
+    \frac{(\texttt{SM[MZ]}-m_Z^\pole)^2}{\sigma_{m_Z}^2} +
+    \frac{(\texttt{SM[MH]}-m_{h_1}^\pole)^2}{\sigma_{m_h}^2}
+  \end{align}
+  %
+  is minimal.  The constants \code{SM[MZ]}, \code{SM[MH]},
+  $\sigma_{m_Z}$ and $\sigma_{m_h}$ are defined in
+  \code{src/ew_input.hpp} to be
+  %
+  \begin{align}
+    \texttt{SM[MZ]} &= 91.1876, &
+    \texttt{SM[MH]} &= 125.9, \\
+    \sigma_{m_Z} &= 0.0021, &
+    \sigma_{m_h} &= 0.4 .
+  \end{align}
+
+\item The function \code{FSFindRoot[parameters, functions]} can be
+  given, where \code{parameters} is a list of model parameters and
+  \code{functions} is a list of functions of these parameters.
+  \code{FSFindRoot[parameters, functions]} will numericall vary the
+  \code{parameters} until the \code{functions} are zero.  Example:
+  %
+  \begin{lstlisting}
+FSFindRoot[{vd,vu},
+           {SM[MZ] - Pole[M[VZ]], SM[MH] - Pole[M[hh[1]]]}]
+  \end{lstlisting}
+  %
+  Here, the parameters \code{vu} and \code{vd} are varied until the
+  vector-like function
+  %
+  \begin{align}
+    f(v_d,v_u) =
+    \begin{pmatrix}
+      \texttt{SM[MZ]} - m_Z^\pole \\
+      \texttt{SM[MH]} - m_{h_1}^\pole
+    \end{pmatrix}
+  \end{align}
+  %
+  is zero.
+\end{itemize}
 

### 2014-04-17 doc/paper.tex
update the example model file
The calculation of the Yukawa couplings must defined explicitely.

@@
 LowScaleInput = {
+   {Yu, Automatic},
+   {Yd, Automatic},
+   {Ye, Automatic},
    {vd, 2 MZDRbar / Sqrt[GUTNormalization[g1]^2 g1^2 + g2^2] Cos[ArcTan[TanBeta]]},
@@
    {vd, SM[vev] Cos[ArcTan[TanBeta]]},
-   {vu, SM[vev] Sin[ArcTan[TanBeta]]}
+   {vu, SM[vev] Sin[ArcTan[TanBeta]]},
+   {Yu, Automatic},
+   {Yd, Automatic},
+   {Ye, Automatic}
 };
@@
   \code{parameters.m} file, see the SARAH manual or
-  \cite{Staub:2010jh}.
+  \cite{Staub:2010jh}.  The Standard Model Yukawa couplings \code{Yu},
+  \code{Yd}, \code{Ye} can be calculated automatically from the known
+  Standard Model quark and lepton masses, see
+  Section~\ref{sec:calculation-of-yukawa-couplings}.  To enable the
+  automatic calculation, \code{value} must be set to \code{Automatic},
+  see the example above.
 
@@
 \code{InitialGuessAtLowScale} and \code{InitialGuessAtHighScale},
-respectively.  Note, that the gauge and Yukawa couplings will be
-guessed automatically at the low-scale from the known Standard Model
-parameters.
+respectively.  Note, that the gauge couplings will be guessed
+automatically at the low-scale from the known Standard Model
+parameters, see Section~\ref{sec:calculation-of-gauge-couplings}.
 

### 2014-04-17 config/remove_export_markers.sh
Simplify config/remove_export_markers.sh

@@ -1,18 +1,10 @@
 #!/bin/sh
 
-# This script removes lines beween two markers from a file (or stdin).
+# This script removes lines between two markers from files (or stdin).
 # The markers are removed as well.
+# All arguments understood by sed are accepted such as `-i'.
 
 begin_marker="# *BEGIN: *NOT EXPORTED *#*"
 end_marker="# *END: *NOT EXPORTED *#*"
 
-if test $# -lt 1; then
-    # read from stdin
-    if [ ! -t 0 ]; then
-        sed -e "/${begin_marker}/,/${end_marker}/d" < /dev/stdin
-    fi
-else
-    # read from file
-    input_file="$1"
-    sed -e "/${begin_marker}/,/${end_marker}/d" ${input_file}
-fi
+exec sed -e "/${begin_marker}/,/${end_marker}/d" "$@"

### 2014-04-24 doc/paper.tex
tiny fix to abstract of paper.

@@
   \begin{abstract}
-   Here we introduce \fs, a MATHEMATICA package which generates a C++
+   We introduce \fs, a MATHEMATICA package which generates a C++
    spectrum generator for any SUSY model, created with modularity and
    speed in mind.  \fs makes use of the existing \sarah package
-   to obtain self energies, tadpoles corrections and renormalisation
+   to obtain self energies, tadpole corrections and renormalisation
    group equations in any model.  \fs then generates C++ code for

### 2014-04-24 doc/paper.tex
removed red comment in abstract.

@@
    model file. The user should also specify the boundary conditions in
-   a separate \fs model file. {\color{red}( Do we want to keep this
-     separate and in different places, add a script to put a \sarah
-     model file in the right place or combine them?)}  Flexible SUSY
+   a separate \fs model file.  Flexible SUSY
    then translates these into C++ code along with numerical routines

### 2014-04-24 doc/paper.tex
Added some basic references to introduction.

@@
 \section{Introduction}
-Supersymmetry provides the only way to extend the space-time
-symmetries of the poincar\'e group\cite{}, leading many to suspect
-that it may be realised in nature in some form. In particular
-supersymmetric extensions of the standard model where supersymmetry is
-broken at the TeV scale have been proposed to solve the hierarchy
-problem\cite{}, allow gauge coupling unification\cite{} and predict a
-dark matter candidate\cite{} and can fit the observed relic
-density\cite{}.  Such models have also been used for
-baryogenesis\cite{} or leptogensis\cite{} to solve the
-matter-anti-matter asymmetry of the universe and have been considered
-as the low energy effective models originating from string
-theory\cite{}.
+Supersymmetry provides the only non-trivial way to extend the
+space-time symmetries of the poincar\'e
+group\cite{Coleman:1967ad,Haag:1974qh}, leading many to suspect that it
+may be realised in nature in some form. In particular supersymmetric
+extensions of the standard model where supersymmetry is broken at the
+TeV scale have been proposed to solve the hierarchy
+problem\cite{Weinberg:1975gm, Weinberg:1979bn, Gildener:1976ai,
+  Susskind:1978ms, 'tHooft:1980xb}, allow gauge coupling
+unification\cite{Langacker:1990jh, Ellis:1990wk, Amaldi:1991cn,
+  Langacker:1991an, Giunti:1991ta} and predict a dark matter
+candidate\cite{} and can fit the observed relic density\cite{}.  Such
+models have also been used for baryogenesis\cite{} or
+leptogensis\cite{} to solve the matter-anti-matter asymmetry of the
+universe and have been considered as the low energy effective models
+originating from string theory\cite{}.
 
@@
 \begin{thebibliography}{100}
+%% %\cite{Nilles:1983ge}
+%% \bibitem{Nilles:1983ge} 
+%%   H.~P.~Nilles,
+%%   %``Supersymmetry, Supergravity and Particle Physics,''
+%%   Phys.\ Rept.\  {\bf 110}, 1 (1984).
+%%   %%CITATION = PRPLC,110,1;%%
+%%   %4282 citations counted in INSPIRE as of 24 Apr 2014
+%% %\cite{Lahanas:1986uc}
+%% \bibitem{Lahanas:1986uc} 
+%%   A.~B.~Lahanas and D.~V.~Nanopoulos,
+%%   %``The Road to No Scale Supergravity,''
+%%   Phys.\ Rept.\  {\bf 145}, 1 (1987).
+%%   %%CITATION = PRPLC,145,1;%%
+%%   %694 citations counted in INSPIRE as of 24 Apr 2014
+%% %\cite{Ellis:1990wk}
+%\cite{Coleman:1967ad}
+
+%\cite{Weinberg:1975gm}
+\bibitem{Weinberg:1975gm} 
+  S.~Weinberg,
+  %``Implications of Dynamical Symmetry Breaking,''
+  Phys.\ Rev.\ D {\bf 13}, 974 (1976).
+  %%CITATION = PHRVA,D13,974;%%
+  %1251 citations counted in INSPIRE as of 24 Apr 2014
+%\cite{Weinberg:1979bn}
+\bibitem{Weinberg:1979bn}
+  S.~Weinberg,
+  %``Implications of Dynamical Symmetry Breaking: An Addendum,''
+  Phys.\ Rev.\ D {\bf 19} (1979) 1277.
+  %%CITATION = PHRVA,D19,1277;%%
+  %1586 citations counted in INSPIRE as of 24 Apr 2014
+%\cite{Gildener:1976ai}
+\bibitem{Gildener:1976ai} 
+  E.~Gildener,
+  %``Gauge Symmetry Hierarchies,''
+  Phys.\ Rev.\ D {\bf 14}, 1667 (1976).
+  %%CITATION = PHRVA,D14,1667;%%
+  %480 citations counted in INSPIRE as of 24 Apr 2014
+%\cite{Susskind:1978ms}
+\bibitem{Susskind:1978ms}
+  L.~Susskind,
+  %``Dynamics of Spontaneous Symmetry Breaking in the Weinberg-Salam Theory,''
+  Phys.\ Rev.\ D {\bf 20} (1979) 2619.
+  %%CITATION = PHRVA,D20,2619;%%
+  %2011 citations counted in INSPIRE as of 24 Apr 2014
+%\cite{'tHooft:1980xb}
+\bibitem{'tHooft:1980xb} 
+  G.~'t Hooft, C.~Itzykson, A.~Jaffe, H.~Lehmann, P.~K.~Mitter, I.~M.~Singer and R.~Stora,
+  %``Recent Developments in Gauge Theories. Proceedings, Nato Advanced Study Institute, Cargese, France, August 26 - September 8, 1979,''
+  NATO Adv.\ Study Inst.\ Ser.\ B Phys.\  {\bf 59}, pp.1 (1980).
+  %%CITATION = NASBD,59,pp.1;%%
+  %10 citations counted in INSPIRE as of 24 Apr 2014
+\bibitem{Coleman:1967ad} 
+  S.~R.~Coleman and J.~Mandula,
+  %``All Possible Symmetries of the S Matrix,''
+  Phys.\ Rev.\  {\bf 159}, 1251 (1967).
+  %%CITATION = PHRVA,159,1251;%%
+  %722 citations counted in INSPIRE as of 24 Apr 2014
+%\cite{Haag:1974qh}
+\bibitem{Haag:1974qh} 
+  R.~Haag, J.~T.~Lopuszanski and M.~Sohnius,
+  %``All Possible Generators of Supersymmetries of the s Matrix,''
+  Nucl.\ Phys.\ B {\bf 88}, 257 (1975).
+  %%CITATION = NUPHA,B88,257;%%
+  %828 citations counted in INSPIRE as of 24 Apr 2014
+
+\bibitem{Ellis:1990wk} 
+  J.~R.~Ellis, S.~Kelley and D.~V.~Nanopoulos,
+  %``Probing the desert using gauge coupling unification,''
+  Phys.\ Lett.\ B {\bf 260}, 131 (1991).
+  %%CITATION = PHLTA,B260,131;%%
+  %862 citations counted in INSPIRE as of 24 Apr 2014
+%\cite{Amaldi:1991cn}
+\bibitem{Amaldi:1991cn} 
+  U.~Amaldi, W.~de Boer and H.~Furstenau,
+  %``Comparison of grand unified theories with electroweak and strong coupling constants measured at LEP,''
+  Phys.\ Lett.\ B {\bf 260}, 447 (1991).
+  %%CITATION = PHLTA,B260,447;%%
+  %1525 citations counted in INSPIRE as of 24 Apr 2014
+%\cite{Langacker:1991an}
+\bibitem{Langacker:1991an} 
+  P.~Langacker and M.~-x.~Luo,
+  %``Implications of precision electroweak experiments for $M_t$, $\rho_{0}$, $\sin^2\theta_W$ and grand unification,''
+  Phys.\ Rev.\ D {\bf 44}, 817 (1991).
+  %%CITATION = PHRVA,D44,817;%%
+  %1219 citations counted in INSPIRE as of 24 Apr 2014
+%\cite{Giunti:1991ta}
+\bibitem{Giunti:1991ta} 
+  C.~Giunti, C.~W.~Kim and U.~W.~Lee,
+  %``Running coupling constants and grand unification models,''
+  Mod.\ Phys.\ Lett.\ A {\bf 6}, 1745 (1991).
+  %%CITATION = MPLAE,A6,1745;%%
+  %273 citations counted in INSPIRE as of 24 Apr 2014
+%\cite{Langacker:1990jh}
+\bibitem{Langacker:1990jh} 
+  P.~Langacker,
+  %``Precision tests of the standard model,''
+  In *Boston 1990, Proceedings, Particles, strings and cosmology* 237-269 and Pennsylvania Univ.
+Philadelphia - UPR-0435T (90,rec.Oct.) 33 p.
+(015721) (see HIGH ENERGY PHYSICS INDEX 29 (1991) No.
+9950)
+  %4 citations counted in INSPIRE as of 24 Apr 2014
+
 %\cite{Girardello:1981wz}

### 2014-04-24 doc/paper.tex
Added more refernces and fixed typos in introduction. many more to do.

@@
 \section{Introduction}
-Supersymmetry provides the only non-trivial way to extend the
+Supersymmetry(SUSY) provides the only non-trivial way to extend the
 space-time symmetries of the poincar\'e
-group\cite{Coleman:1967ad,Haag:1974qh}, leading many to suspect that it
-may be realised in nature in some form. In particular supersymmetric
-extensions of the standard model where supersymmetry is broken at the
-TeV scale have been proposed to solve the hierarchy
+group\cite{Coleman:1967ad,Haag:1974qh}, leading many to suspect that
+SUSY may be realised in nature in some form. In particular
+supersymmetric extensions of the standard model where SUSY is broken
+at the TeV scale have been proposed to solve the hierarchy
 problem\cite{Weinberg:1975gm, Weinberg:1979bn, Gildener:1976ai,
@@
 unification\cite{Langacker:1990jh, Ellis:1990wk, Amaldi:1991cn,
-  Langacker:1991an, Giunti:1991ta} and predict a dark matter
-candidate\cite{} and can fit the observed relic density\cite{}.  Such
-models have also been used for baryogenesis\cite{} or
-leptogensis\cite{} to solve the matter-anti-matter asymmetry of the
-universe and have been considered as the low energy effective models
-originating from string theory\cite{}.
+  Langacker:1991an, Giunti:1991ta} and predict a dark matter candidate
+which can fit the observed relic
+density\cite{Goldberg:1983nd,Ellis:1983ew}.  Such models have also
+been used for baryogenesis\cite{} or leptogensis\cite{} to solve the
+matter-anti-matter asymmetry of the universe and have been considered
+as the low energy effective models originating from string
+theory\cite{}.
 
@@
 explore the model. In particular there are a number of public spectrum
-generators for MSSM\cite{Allanach:2001kg,Porod:2003um,Djouadi:2002ze,Baer:1993ae} and two for the NMSSM\cite{Ellwanger:2006rn,Allanach:nmssm} which allow
+generators for MSSM\cite{Allanach:2001kg,Porod:2003um,Djouadi:2002ze,Baer:1993ae} and two for the NMSSM\cite{Ellwanger:2006rn,Allanach:2013kza} which allow
 one to find the sparticle spectrum and mixings for a particular
@@
 %\cite{Coleman:1967ad}
-
+\bibitem{Coleman:1967ad} 
+  S.~R.~Coleman and J.~Mandula,
+  %``All Possible Symmetries of the S Matrix,''
+  Phys.\ Rev.\  {\bf 159}, 1251 (1967).
+  %%CITATION = PHRVA,159,1251;%%
+  %722 citations counted in INSPIRE as of 24 Apr 2014
+%\cite{Haag:1974qh}
+\bibitem{Haag:1974qh} 
+  R.~Haag, J.~T.~Lopuszanski and M.~Sohnius,
+  %``All Possible Generators of Supersymmetries of the s Matrix,''
+  Nucl.\ Phys.\ B {\bf 88}, 257 (1975).
+  %%CITATION = NUPHA,B88,257;%%
+  %828 citations counted in INSPIRE as of 24 Apr 2014
 %\cite{Weinberg:1975gm}
@@
   %10 citations counted in INSPIRE as of 24 Apr 2014
-\bibitem{Coleman:1967ad} 
-  S.~R.~Coleman and J.~Mandula,
-  %``All Possible Symmetries of the S Matrix,''
-  Phys.\ Rev.\  {\bf 159}, 1251 (1967).
-  %%CITATION = PHRVA,159,1251;%%
-  %722 citations counted in INSPIRE as of 24 Apr 2014
-%\cite{Haag:1974qh}
-\bibitem{Haag:1974qh} 
-  R.~Haag, J.~T.~Lopuszanski and M.~Sohnius,
-  %``All Possible Generators of Supersymmetries of the s Matrix,''
-  Nucl.\ Phys.\ B {\bf 88}, 257 (1975).
-  %%CITATION = NUPHA,B88,257;%%
-  %828 citations counted in INSPIRE as of 24 Apr 2014
+
+\bibitem{Langacker:1990jh} 
+  P.~Langacker,
+  %``Precision tests of the standard model,''
+  In *Boston 1990, Proceedings, Particles, strings and cosmology* 237-269 and Pennsylvania Univ.
+Philadelphia - UPR-0435T (90,rec.Oct.) 33 p.
+(015721) (see HIGH ENERGY PHYSICS INDEX 29 (1991) No.
+9950)
+  %4 citations counted in INSPIRE as of 24 Apr 2014
 
@@
 %\cite{Langacker:1990jh}
-\bibitem{Langacker:1990jh} 
-  P.~Langacker,
-  %``Precision tests of the standard model,''
-  In *Boston 1990, Proceedings, Particles, strings and cosmology* 237-269 and Pennsylvania Univ.
-Philadelphia - UPR-0435T (90,rec.Oct.) 33 p.
-(015721) (see HIGH ENERGY PHYSICS INDEX 29 (1991) No.
-9950)
-  %4 citations counted in INSPIRE as of 24 Apr 2014
+
+%\cite{Goldberg:1983nd}
+\bibitem{Goldberg:1983nd} 
+  H.~Goldberg,
+  %``Constraint on the Photino Mass from Cosmology,''
+  Phys.\ Rev.\ Lett.\  {\bf 50}, 1419 (1983)
+  [Erratum-ibid.\  {\bf 103}, 099905 (2009)].
+  %%CITATION = PRLTA,50,1419;%%
+  %972 citations counted in INSPIRE as of 24 Apr 2014
+%\cite{Ellis:1983ew}
+\bibitem{Ellis:1983ew} 
+  J.~R.~Ellis, J.~S.~Hagelin, D.~V.~Nanopoulos, K.~A.~Olive and M.~Srednicki,
+  %``Supersymmetric Relics from the Big Bang,''
+  Nucl.\ Phys.\ B {\bf 238}, 453 (1984).
+  %%CITATION = NUPHA,B238,453;%%
+  %1282 citations counted in INSPIRE as of 24 Apr 2014
 
@@
   %81 citations counted in INSPIRE as of 03 Nov 2013
-
+%\cite{Allanach:2013kza}
+\bibitem{Allanach:2013kza} 
+  B.~C.~Allanach, P.~Athron, L.~C.~Tunstall, A.~Voigt and A.~G.~Williams,
+  %``Next-to-Minimal SOFTSUSY,''
+  arXiv:1311.7659 [hep-ph].
+  %%CITATION = ARXIV:1311.7659;%%
+  %1 citations counted in INSPIRE as of 24 Apr 2014
   %\cite{Staub:2010ty}

### 2014-04-24 doc/paper.tex
More work on introduction, some restructuring and new refs.

@@
 density\cite{Goldberg:1983nd,Ellis:1983ew}.  Such models have also
-been used for baryogenesis\cite{} or leptogensis\cite{} to solve the
+been used for baryogenesis or leptogensis to solve the
 matter-anti-matter asymmetry of the universe and have been considered
 as the low energy effective models originating from string
-theory\cite{}.
+theory.
 
 Detailed phenonenological studies have been carried out for scenarios
-with the minimal supersymmetric standard model (MSSM). Many
-computational tools have been created to aid these studies and to
-explore the model. In particular there are a number of public spectrum
-generators for MSSM\cite{Allanach:2001kg,Porod:2003um,Djouadi:2002ze,Baer:1993ae} and two for the NMSSM\cite{Ellwanger:2006rn,Allanach:2013kza} which allow
-one to find the sparticle spectrum and mixings for a particular
-choices of breaking mechanism inspired boundary conditions and
-specified parameters.
-
-However none of the fundamental motivations of supersymmetry require
-minimality. Yet without minimality as a constraint, constructing
-specialised tools to study all relevant models would require an
-enormous amount of work.  So general tools which can automate this
-process and produce fats and reliable programs can greatly enhance our
-ability to understand and test non-minimal realisations of
-supersymmetry.
-
-From the recent $7$ TeV and $8$ TeV runs at the Large Hadron Collider
-(LHC) there have been two important developments.  First low energy
-signatures expected from such models, such as the classic jets plus
-missing transverse energy signature, have not been observed,
-substantially raising the lower limit on sparticle masses\cite{}. No
-other signature of beyond the standard model(BSM) physics has been
-observed, leaving the fundamental questions which motivated BSM
-physics unanswered.
+within the minimal supersymmetric standard model (MSSM).  Such work
+has been greatly aided by public spectrum generators for
+MSSM\cite{Allanach:2001kg,Porod:2003um,Djouadi:2002ze,Baer:1993ae},
+allowing fast and reliable exploration of the sparticle spectrum and
+mixings that can be obtained from particular choices of breaking
+mechanism inspired boundary conditions and specified parameters. There
+are also two public spectrum generators
+\cite{Ellwanger:2006rn,Allanach:2013kza} for the next to minimal
+supersymmetric standard model (NMSSM) \cite{NMSSM} .
+
+None of the fundamental motivations of supersymmetry require
+minimality, and specific altternatives to (or extensions of) the MSSM
+are motiovated by, for example,the $\mu-problem$ of the
+MSSM\cite{Kim:1983dt}, explaining the family structure (see e.g,
+\cite{King:2014nza}) or for successful baryogensis or leptogenesis
+(see e.g,\cite{King:2008qb}). However constructing specialised tools
+to study all relevant models would require an enormous amount of work.
+So general tools which can automate this process and produce fast and
+reliable programs can greatly enhance our ability to understand and
+test non-minimal realisations of supersymmetry.
+
+  For instance from the recent $7$ TeV and $8$ TeV runs
+at the Large Hadron Collider (LHC) there have been two important
+developments.  First low energy signatures expected from such models,
+such as the classic jets plus missing transverse energy signature,
+have not been observed, substantially raising the lower limit on
+sparticle masses\cite{}. No other signature of beyond the standard
+model(BSM) physics has been observed, leaving the fundamental
+questions which motivated BSM physics unanswered.
 
@@
 of gauge couplings, while others may focus on special models which
-preserve naturalness.  In both cases exploration of the models can be
-aided if it is possible to quickly create a spectrum generator to find
-if there are phenomenologically viable scenarios in the models they
-are interested and what the general features of such models are.
+preserve naturalness.  In both cases exploration of such models can be
+aided if it is possible to quickly create a fast spectrum generator.
 
@@
   %\cite{Staub:2010ty}
+\bibitem{NMSSM} P. Fayet, Nucl. Phys. B \textbf{90} (1975) 104; Phys. Lett.
+B \textbf{64} (1976) 159; Phys. Lett. B \textbf{69} (1977) 489 and Phys. Lett. B
+\textbf{84} (1979) 416; H.P. Nilles, M. Srednicki and D. Wyler, Phys. Lett. B
+\textbf{120} (1983) 346; J.M. Frere, D.R. Jones and S. Raby, Nucl. Phys. B
+\textbf{222} (1983) 11; J.P. Derendinger and C.A. Savoy, Nucl. Phys. B
+\textbf{237} (1984) 307;  A.I. Veselov, M.I. Vysotsky and K.A. Ter-Martirosian,
+Sov. Phys. JETP \textbf{63} (1986) 489; J.R. Ellis, J.F. Gunion, H.E. Haber, L.
+Roszkowski and F. Zwirner, Phys. Rev. D \textbf{39}  (1989) 844; M. Drees, Int.
+J. Mod. Phys. A \textbf{4}  (1989) 3635; U. Ellwanger, M. Rausch de
+Traubenberg and C.A. Savoy, Phys. 
+Lett. B \textbf{315} (1993) 331, Z. Phys. C {\bf 67} (1995) 665 and Nucl. Phys.
+B \textbf{492} (1997) 307; U.~Ellwanger, Phys.\ Lett.\  B {\bf 303} (1993) 271; P.
+Pandita, Z. Phys. C \textbf{59} (1993) 575; T. Elliott, S.F. King and P.L.
+White, Phys. Rev. D {\bf 49} (1994) 2435; S.F. King and P.L. White, Phys. Rev. D
+\textbf{52} (1995) 4183;  F.~Franke and H.~Fraas, Int.\ J.\ Mod.\ Phys.\  A {\bf
+12} (1997) 479.   D.~J.~Miller, R.~Nevzorov and P.~M.~Zerwas,  Nucl.\ Phys.\ B {\bf 681}, 3 (2004) [hep-ph/0304049].
+%\cite{Kim:1983dt}
+\bibitem{Kim:1983dt} 
+  J.~E.~Kim and H.~P.~Nilles,
+  %``The mu Problem and the Strong CP Problem,''
+  Phys.\ Lett.\ B {\bf 138}, 150 (1984).
+  %%CITATION = PHLTA,B138,150;%%
+  %604 citations counted in INSPIRE as of 24 Apr 2014
+%\cite{King:2014nza}
+\bibitem{King:2014nza} 
+  S.~F.~King, A.~Merle, S.~Morisi, Y.~Shimizu and M.~Tanimoto,
+  %``Neutrino Mass and Mixing: from Theory to Experiment,''
+  arXiv:1402.4271 [hep-ph].
+  %%CITATION = ARXIV:1402.4271;%%
+  %9 citations counted in INSPIRE as of 24 Apr 2014
+%\cite{King:2008qb}
+\bibitem{King:2008qb} 
+  S.~F.~King, R.~Luo, D.~J.~Miller and R.~Nevzorov,
+  %``Leptogenesis in the Exceptional Supersymmetric Standard Model: Flavour dependent lepton asymmetries,''
+  JHEP {\bf 0812}, 042 (2008)
+  [arXiv:0806.0330 [hep-ph]].
+  %%CITATION = ARXIV:0806.0330;%%
+  %26 citations counted in INSPIRE as of 24 Apr 2014
 \bibitem{Staub:2010ty} 

### 2014-04-24 doc/paper.tex
added refs for experimental results and praragraph selling FS.

@@
 
-  For instance from the recent $7$ TeV and $8$ TeV runs
-at the Large Hadron Collider (LHC) there have been two important
-developments.  First low energy signatures expected from such models,
-such as the classic jets plus missing transverse energy signature,
-have not been observed, substantially raising the lower limit on
-sparticle masses\cite{}. No other signature of beyond the standard
-model(BSM) physics has been observed, leaving the fundamental
-questions which motivated BSM physics unanswered.
-
- Secondly ATLAS and CMS discovered\cite{} a light Higgs of $125$ GeV,
- within the mass range that could be accommodated in the MSSM but
- requiring stops which are significantly heavier than both the direct
- collider limits and indirect limits that appears in constrained
- models from the significantly higher limits on first and second
- generation squarks.  Alternatively it is possible that a $125$ GeV
- Higgs in non-minimal models may not require such heavy stops, indeed
- a number of such models have been constructed to do exactly this by
- having new tree level contributions to the Higgs mass.
-
-These results lead to a number of different perspectives, for example
-some may wish to abandon naturalness as a motivation and focus on BSM
-models which can fit the observed relic density and allow unification
-of gauge couplings, while others may focus on special models which
-preserve naturalness.  In both cases exploration of such models can be
-aided if it is possible to quickly create a fast spectrum generator.
+Recent experimental developments have also increased the relevancy of
+such a tool. From the recent $7$ TeV and $8$ TeV runs at the Large
+Hadron Collider (LHC) there have been two important developments.
+First low energy signatures expected from such models, such as the
+classic jets plus missing transverse energy signature, have not been
+observed, substantially raising the lower limit on sparticle masses
+(see e.g.~\cite{Aad:2013wta,Chatrchyan:2014lfa}). No other signature
+of beyond the standard model(BSM) physics has been observed, leaving
+the fundamental questions which motivated BSM physics unanswered.
+
+ Secondly ATLAS and CMS discovered\cite{ATLAS:2012ae,
+   Chatrchyan:2012tx} a light Higgs of $125$ GeV, within the mass
+ range that could be accommodated in the MSSM but requiring stops
+ which are significantly heavier than both the direct collider limits
+ and indirect limits that appears in constrained models from the
+ significantly higher limits on first and second generation squarks.
+ Alternatively it is possible that a $125$ GeV Higgs in non-minimal
+ models may not require such heavy stops, indeed a number of such
+ models have been constructed to do exactly this by having new tree
+ level contributions to the Higgs mass.
+
+These motivate both exploring non-minimal SUSY models which ameliorate
+the naturaless problems introduced, by, for example, raising the tree
+level Higgs mass, and models developed with a fresh persp[ective,
+  based on other considerations.  In both cases exploration of such
+  models can be aided if it is possible to quickly create a fast
+  spectrum generator.
+
+\fs is designed for precisely this task. \fs is a MATHEMATICA package
+which with the help of \sarah
+\cite{Staub:2010ty,Staub:2009bi,Staub:2010jh,Staub:2012pb,Staub:2013tta}
+creates a fast, modular C++ spectrum generator for a user specified
+SUSY model.  The generated code structure is designed to be as
+flexible as possible to accomodate different types of extensions and
+due to it's modular nature it is easy to modify, add new featuires to
+or combine with other programs.
 
 \subsection{The Program}
+\label{Sec:Program}
 \fs is a MATHEMATICA package designed to create a fast and easily
@@
   %26 citations counted in INSPIRE as of 24 Apr 2014
+%\cite{Aad:2013wta}
+\bibitem{Aad:2013wta} 
+  G.~Aad {\it et al.}  [ATLAS Collaboration],
+  %``Search for new phenomena in final states with large jet multiplicities and missing transverse momentum at sqrt(s)=8 TeV proton-proton collisions using the ATLAS experiment,''
+  JHEP {\bf 1310}, 130 (2013)
+  [arXiv:1308.1841 [hep-ex]].
+  %%CITATION = ARXIV:1308.1841;%%
+  %36 citations counted in INSPIRE as of 24 Apr 2014
+\bibitem{Chatrchyan:2014lfa}
+  S.~Chatrchyan {\it et al.}  [CMS Collaboration],
+  %``Search for new physics in the multijet and missing transverse momentum final state in proton-proton collisions at $\sqrt{s}$ = 8 TeV,''
+  arXiv:1402.4770 [hep-ex].
+  %%CITATION = ARXIV:1402.4770;%%
+  %10 citations counted in INSPIRE as of 24 Apr 2014
+%\cite{ATLAS:2012ae}
+\bibitem{ATLAS:2012ae} 
+  G.~Aad {\it et al.}  [ATLAS Collaboration],
+  %``Combined search for the Standard Model Higgs boson using up to 4.9 fb$^{-1}$ of $pp$ collision data at $\sqrt{s}=7$ TeV with the ATLAS detector at the LHC,''
+  Phys.\ Lett.\ B {\bf 710}, 49 (2012)
+  [arXiv:1202.1408 [hep-ex]].
+  %%CITATION = ARXIV:1202.1408;%%
+  %474 citations counted in INSPIRE as of 24 Apr 2014
+%\cite{Chatrchyan:2012tx}
+\bibitem{Chatrchyan:2012tx} 
+  S.~Chatrchyan {\it et al.}  [CMS Collaboration],
+  %``Combined results of searches for the standard model Higgs boson in $pp$ collisions at $\sqrt{s}=7$ TeV,''
+  Phys.\ Lett.\ B {\bf 710}, 26 (2012)
+  [arXiv:1202.1488 [hep-ex]].
+  %%CITATION = ARXIV:1202.1488;%%
+  %592 citations counted in INSPIRE as of 24 Apr 2014
+%\cite{Chatrchyan:2014lfa}
 \bibitem{Staub:2010ty} 

### 2014-04-24 doc/paper.tex
added paragraph at end of introduction describing paper structure. Changed this structure a little.

@@
 
-Detailed phenonenological studies have been carried out for scenarios
+Detailed phenomenological studies have been carried out for scenarios
 within the minimal supersymmetric standard model (MSSM).  Such work
@@
 None of the fundamental motivations of supersymmetry require
-minimality, and specific altternatives to (or extensions of) the MSSM
-are motiovated by, for example,the $\mu-problem$ of the
+minimality, and specific alternatives to (or extensions of) the MSSM
+are motivated by, for example,the $\mu-problem$ of the
 MSSM\cite{Kim:1983dt}, explaining the family structure (see e.g,
@@
 These motivate both exploring non-minimal SUSY models which ameliorate
-the naturaless problems introduced, by, for example, raising the tree
-level Higgs mass, and models developed with a fresh persp[ective,
+the naturalness problems introduced, by, for example, raising the tree
+level Higgs mass, and models developed with a fresh perspective,
   based on other considerations.  In both cases exploration of such
@@
 SUSY model.  The generated code structure is designed to be as
-flexible as possible to accomodate different types of extensions and
-due to it's modular nature it is easy to modify, add new featuires to
-or combine with other programs.
-
-\subsection{The Program}
+flexible as possible to accommodate different types of extensions and
+due to it's modular nature it is easy to modify, add new features to
+or combine with other programs.  The generated code has been
+extensively tested against well known spectrum generators and as well
+providing a solution for new SUSY models the generated MSSM and NMSSM
+codes offer a modern and fast alternative to the existing public
+spectrum generators.
+
+In section \ref{Sec:Program} we describe the program in more detail
+and explain our design goals.  In section \ref{Sec:download}
+information on how to download and compile the code may be found along
+with details on how to get started quickly.  In section
+\ref{Sec:modfile} we describe how the user can create a new
+FlexibleSUSY model file. A detailed description of the structure and
+features of the generated code is then given in section
+\ref{Sec:SpecGenStruct}.  In section \ref{Sec:Flexible} we describe
+the various ways the code can be modified both at the meta code level
+by writing model files and at the C++ code level by modifying the code
+or adding new modules. Finally in section \ref{Sec:comparison} we
+describe detailed comparisons between our generated code and existing
+public spectrum generators as well as against the SPheno-like fortran
+code which can be created using SARAH.
+
+\section{The Program}
 \label{Sec:Program}
@@
 \section{Download and compilation}
+\label{Sec:download}
 Flexible SUSY can be downloaded from \url{http://flexiblesusy.hepforge.org}.  
@@
 \section{Setting up a FlexibleSUSY model}
-
+\label{Sec:modfile}
 A \fs model is configured via a (Mathematica) model file, which in
@@
 \section{Stucture of the spectrum generator}
-
+\label{Sec:SpecGenStruct}
 \subsection{Model parameters and RGEs}
@@
 
-\section{Implementing new models}
-\section{Adapting the code at the C++ level}
-\section{Using individual modules in your own code}
+%%\section{Implementing new models}
+\section{Flexible Spectra}
+\label{Sec:Flexible}
+\subsection{Meta code}
+\subsection{Adapting C++ code}
+\subsection{Combing with  modules of your own code}
 
 \section{Run-time comparison with other spectrum generators}
-
+\label{Sec:comaprison}
 We compare the run-time of the four different spectrum generators

### 2014-04-24 doc/paper.tex
Small corrections, and merged the two paragraphs on recent experimental motivations, cutting the last sentence which was essentially repeated in the paragraph immediately after.

@@
 Supersymmetry(SUSY) provides the only non-trivial way to extend the
-space-time symmetries of the poincar\'e
+space-time symmetries of the Poincar\'e
 group\cite{Coleman:1967ad,Haag:1974qh}, leading many to suspect that
@@
 MSSM\cite{Allanach:2001kg,Porod:2003um,Djouadi:2002ze,Baer:1993ae},
-allowing fast and reliable exploration of the sparticle spectrum and
-mixings that can be obtained from particular choices of breaking
-mechanism inspired boundary conditions and specified parameters. There
-are also two public spectrum generators
-\cite{Ellwanger:2006rn,Allanach:2013kza} for the next to minimal
-supersymmetric standard model (NMSSM) \cite{NMSSM} .
+allowing fast and reliable exploration of the sparticle spectrum,
+mixings and couplings, which can be obtained from particular choices
+of breaking mechanism inspired boundary conditions and specified
+parameters. Beyond the MSSM there are also two public spectrum
+generators \cite{Ellwanger:2006rn,Allanach:2013kza} for the next to
+minimal supersymmetric standard model (NMSSM) \cite{NMSSM}.
 
@@
 minimality, and specific alternatives to (or extensions of) the MSSM
-are motivated by, for example,the $\mu-problem$ of the
-MSSM\cite{Kim:1983dt}, explaining the family structure (see e.g,
+are, for example, motivated by the $\mu$-problem of the
+MSSM \cite{Kim:1983dt}; explaining the family structure (see e.g,
 \cite{King:2014nza}) or for successful baryogensis or leptogenesis
@@
 of beyond the standard model(BSM) physics has been observed, leaving
-the fundamental questions which motivated BSM physics unanswered.
-
- Secondly ATLAS and CMS discovered\cite{ATLAS:2012ae,
-   Chatrchyan:2012tx} a light Higgs of $125$ GeV, within the mass
- range that could be accommodated in the MSSM but requiring stops
- which are significantly heavier than both the direct collider limits
- and indirect limits that appears in constrained models from the
+the fundamental questions which motivated BSM physics
+unanswered. Secondly ATLAS and CMS discovered\cite{ATLAS:2012ae,
+  Chatrchyan:2012tx} a light Higgs of $125$ GeV, within the mass range
+that could be accommodated in the MSSM but requiring stops which are
+significantly heavier than both the direct collider limits and
+indirect limits that appears in constrained models from the
 significantly higher limits on first and second generation squarks.
- Alternatively it is possible that a $125$ GeV Higgs in non-minimal
- models may not require such heavy stops, indeed a number of such
- models have been constructed to do exactly this by having new tree
- level contributions to the Higgs mass.
 
@@
 \section{Run-time comparison with other spectrum generators}
-\label{Sec:comaprison}
+\label{Sec:comparison}
 We compare the run-time of the four different spectrum generators

### 2014-04-24 doc/paper.tex
More tinkering mostly for readability.  Mentioned sarah as the competitor.

@@
 These motivate both exploring non-minimal SUSY models which ameliorate
-the naturalness problems introduced, by, for example, raising the tree
-level Higgs mass, and models developed with a fresh perspective,
-  based on other considerations.  In both cases exploration of such
-  models can be aided if it is possible to quickly create a fast
-  spectrum generator.
-
-\fs is designed for precisely this task. \fs is a MATHEMATICA package
-which with the help of \sarah
-\cite{Staub:2010ty,Staub:2009bi,Staub:2010jh,Staub:2012pb,Staub:2013tta}
-creates a fast, modular C++ spectrum generator for a user specified
-SUSY model.  The generated code structure is designed to be as
-flexible as possible to accommodate different types of extensions and
-due to it's modular nature it is easy to modify, add new features to
-or combine with other programs.  The generated code has been
-extensively tested against well known spectrum generators and as well
-providing a solution for new SUSY models the generated MSSM and NMSSM
-codes offer a modern and fast alternative to the existing public
-spectrum generators.
+the naturalness problems by, for example, raising the tree level Higgs
+mass, and models developed with a fresh perspective, based on other
+considerations.  In both cases exploration of such models can be aided
+if it is possible to quickly create a fast spectrum generator.
+Currently there is only option for this, a SPheno-like FORTRAN code
+which can be generated from
+\sarah\cite{Staub:2010ty,Staub:2009bi,Staub:2010jh,Staub:2012pb,Staub:2013tta}.
+
+\fs provides a much needed alternative to this with a structure which
+has been freshly designed to accommodate as general range of models as
+possible and to be easily adaptable to changing goals and new
+ideas. \fs is a MATHEMATICA package which uses \sarah to create a
+fast, modular C++ spectrum generator for a user specified SUSY model.
+The generated code structure is designed to be as flexible as possible
+to accommodate different types of extensions and due to it's modular
+nature it is easy to modify, add new features and combine with other
+programs.  The generated code has been extensively tested against well
+known spectrum generators and as well providing a solution for new
+SUSY models the generated MSSM and NMSSM codes offer a modern and fast
+alternative to the existing public spectrum generators.
 
@@
   %1 citations counted in INSPIRE as of 24 Apr 2014
-  %\cite{Staub:2010ty}
 \bibitem{NMSSM} P. Fayet, Nucl. Phys. B \textbf{90} (1975) 104; Phys. Lett.
@@
 %\cite{Chatrchyan:2014lfa}
+ %\cite{Staub:2010ty}
 \bibitem{Staub:2010ty} 

### 2014-04-24 doc/paper.tex
Edited abstract for readability.

@@
   \begin{abstract}
-   We introduce \fs, a MATHEMATICA package which generates a C++
-   spectrum generator for any SUSY model, created with modularity and
-   speed in mind.  \fs makes use of the existing \sarah package
-   to obtain self energies, tadpole corrections and renormalisation
-   group equations in any model.  \fs then generates C++ code for
-   calculating spectra with a structure that is designed to be
-   flexible. The user specifies the model supplying the
+   We introduce \fs, a MATHEMATICA package which generates a fast,
+   precise C++ spectrum generator for any SUSY model specified by the
+   user.  The generated code is designed with both speed and
+   modularity in mind, making it easy to adapt and extend with new
+   features. The user specifies the model supplying the
    superpotential, gauge structure and particle content in a \sarah
-   model file. The user should also specify the boundary conditions in
-   a separate \fs model file.  Flexible SUSY
-   then translates these into C++ code along with numerical routines
-   for solving the renormalisation group equations, one loop
-   electroweak symmetry breaking conditions and a solver routine for
-   simultaneously solving for the spectrum.  The modular structure of
-   the generated code allows for any individual component to be
-   replaced with an alternative if available.  Indeed \fs has been
-   intentionally designed to grow as alternative solvers and
-   calculators are added.
+   model file and provides specific the boundary conditions in a
+   separate \fs model file. \fs makes use of the existing \sarah
+   package to obtain self energies, tadpole corrections,
+   renormalisation group equations(RGEs) and tree level mass and
+   electroweak symmetry breaking conditions for the specified model. These are
+   translated into C++ code and combined  with numerical
+   routines for solving the RGEs, EWSB conditions and simultaneously
+   solving for the spectrum conssistent with user specified boundary
+   conditions.  The modular structure of the generated code allows for
+   individual components to be replaced with an alternative if
+   available. \fs has been carefully designed to grow as
+   alternative solvers and calculators are added.
   \end{abstract}
+
 \begin{keyword}

### 2014-04-24 doc/paper.tex
Fixed names to be in one line, as opposed to split style copied from old softsusy.

@@
 
-\author{Peter Athron}
-\address{ARC Centre of Excellence for Particle Physics at 
+\author[adelaide]{Peter Athron}
+\author[dresden]{Jae-hyeon Park}
+\author[dresden]{Dominik St\"ockinger}
+\author[dresden]{Alexander Voigt}
+\address[adelaide]{ARC Centre of Excellence for Particle Physics at 
 the Tera-scale, School of Chemistry and Physics, University of Adelaide, 
 Adelaide SA 5005 Australia}
-\author{Jae-hyeon Park}
-\author{Dominik St\"ockinger}
-\author{Alexander Voigt}
-\address{Institut f\"ur Kern- und Teilchenphysik,
+\address[dresden]{Institut f\"ur Kern- und Teilchenphysik,
 TU Dresden, Zellescher Weg 19, 01069 Dresden, Germany}

### 2014-04-24 doc/paper.tex
removed comment in program summary. these should be checked though.

@@
          bits\\ {\em External routines:}\/ SARAH 4.0.4, Boost library,
-         lapack {\color{red} Not sure what counts here?}\\ {\em
+         lapack\\ {\em
            Typical running time:}\/ 0.1-0.3 seconds per parameter

### 2014-05-04 higgs-study/plot2d.C
make plots nicer and convert them to pdf directly

@@ -12,11 +12,26 @@
 #include "TLegend.h"
 #include "TString.h"
 #include "TStyle.h"
+#include "TSystem.h"
+#include "TColor.h"
 
 #include "plot.h"
 
+void set_plot_style()
+{
+    static const Int_t NRGBs = 5;
+    static const Int_t NCont = 30;
+
+    Double_t stops[NRGBs] = { 0.00, 0.34, 0.61, 0.84, 1.00 };
+    Double_t red[NRGBs]   = { 0.00, 0.00, 0.87, 1.00, 0.51 };
+    Double_t green[NRGBs] = { 0.00, 0.81, 1.00, 0.20, 0.00 };
+    Double_t blue[NRGBs]  = { 0.51, 1.00, 0.12, 0.00, 0.00 };
+    TColor::CreateGradientColorTable(NRGBs, stops, red, green, blue, NCont);
+    gStyle->SetNumberContours(NCont);
+}
+
 void plot2d(const TString& file_name = "higgs-study/data/scan_MSSM.dat",
-            const TString& title = "CMSSM $m_{h}^{\text{pole}}$ / GeV")
+            const TString& title = "m_{h}^{\\text{pole}}\\text{ / GeV}")
 {
    std::ifstream ifs(file_name.Data());
    std::string line;
@@ -61,17 +76,25 @@ void plot2d(const TString& file_name = "higgs-study/data/scan_MSSM.dat",
    }
 
    h->SetTitle(title);
-   h->GetXaxis()->SetTitle("$\tan\beta$");
+   h->GetXaxis()->SetTitle("\\tan\\beta");
    h->GetYaxis()->SetTitle("$m_0$ / TeV");
    h->SetMinimum(95.);
    h->SetMaximum(135.);
    // SetZminZmax(h);
 
-   TCanvas* canvas = new TCanvas("canvas", title, 800, 600);
+   TCanvas* canvas = new TCanvas("canvas", title, 400, 300);
    canvas->cd(1);
+   // gStyle->SetPaperSize(10.,10.); // in cm
+   set_plot_style();
    h->Draw("colz");
 
    TString tex_file(file_name);
-   tex_file.ReplaceAll(".dat",".pdf");
+   tex_file.ReplaceAll(".dat",".tex");
+   TString pdf_file(file_name);
+   pdf_file.ReplaceAll(".dat",".pdf");
    canvas->Print(tex_file);
+
[…]

### 2014-05-05 meta/LoopMasses.m
compare pole masses for the LSP determination

@@ -784,9 +784,9 @@ particle_type = " <> info <> "::NUMBER_OF_PARTICLES;
            For[m = 1, m <= Length[masses], m++,
                mass = masses[[m]];
                body = body <> "\
-tmp_mass = " <>
-CConversion`RValueToCFormString[Abs[GetLightestMassEigenstate[mass]]] <>
-";
+tmp_mass = Abs(PHYSICAL(" <>
+CConversion`RValueToCFormString[GetLightestMassEigenstate[mass]] <>
+"));
 if (tmp_mass < lsp_mass) {
 " <> IndentText["\
 lsp_mass = tmp_mass;

### 2014-05-05 meta/FlexibleSUSY.m
create get_lsp() function only if PotentialLSPParticles is non-empty

@@ -527,7 +527,10 @@ WriteModelClass[massMatrices_List, ewsbEquations_List,
            enablePoleMassThreads = True;
            callAllLoopMassFunctionsInThreads = LoopMasses`CallAllPoleMassFunctions[FlexibleSUSY`FSEigenstates, enablePoleMassThreads];
            masses                       = FlexibleSUSY`M[TreeMasses`GetMassEigenstate[#]]& /@ massMatrices;
-           {lspGetters, lspFunctions}   = LoopMasses`CreateLSPFunctions[FlexibleSUSY`PotentialLSPParticles];
+           If[Head[PotentialLSPParticles] === List &&
+              PotentialLSPParticles =!= {},
+              {lspGetters, lspFunctions} = LoopMasses`CreateLSPFunctions[FlexibleSUSY`PotentialLSPParticles];
+             ];
            printMasses                  = WriteOut`PrintParameters[masses, "ostr"];
            mixingMatrices               = Flatten[TreeMasses`GetMixingMatrixSymbol[#]& /@ massMatrices];
            printMixingMatrices          = WriteOut`PrintParameters[mixingMatrices, "ostr"];

### 2014-05-05 doc/paper.tex
document `PotentialLSPParticles' variable

@@
          lowest boundary condition.  \fs~ assumes that all couplings
-         of the model are real (i.e.\ $CP-$conserving)?  Due to the
+         of the model are real (i.e.\ $CP-$conserving). Due to the
          modular nature of the generated code adaption and extension
@@
 
-\section{The Program}
+\section{Overview of the Program}
 \label{Sec:Program}
@@
 adaptable spectrum generator in C++ for any SUSY model and some
-non-supersymmetric models {\color{red} can we?}.  \fs makes use of the
-existing \sarah package \cite{Staub:2010ty,Staub:2009bi,Staub:2010jh,Staub:2012pb,Staub:2013tta} to obtain one-loop self energies, one-loop
-tadpoles corrections and two-loop renormalisation group
-equations(RGEs) in any model.  \fs then converts these algerbraic
-expressions from \sarah into C++ code using the C++11 standard. \fs
-uses some parts of SOFTSUSY\cite{Allanach:2001kg}, the very fast Eigen
-library \cite{eigen}, as well as the gnu scientific library and the
-boost library to create numerical routines which then use these
-routines to solve the RGEs and EWSB conditions which will be used in
-the spectrum generator. The main solver routine of \fs then calls
-these routines to find simultaneous solutions for the EWSB, low energy
-data, and other, user supplied, boundary conditions (e.g. the usual
-mSUGRA universality conditions) and calculates the mass spectrum.
+non-supersymmetric models.  The user specifies the model by giving the superfield content, superpotential, gauge symmetries and mass mixings in the  
+\sarah model files, examples of which can be found in the \code{Models}
+directory of \sarah and the \code{sarah} directory of FlexibleSUSY.  The user
+then also specifies the boundary conditions on the model parameters in a separate \code{FlexibleSUSY.m} file, examples of which can be found in the \code{model_files} directory.
+A detailed guide on how to write the FlexibleSUSY.m model is given in section \ref{Sec:modfile}.
+
+  \fs makes use of the existing \sarah package
+  \cite{Staub:2010ty,Staub:2009bi,Staub:2010jh,Staub:2012pb,Staub:2013tta}
+  to obtain tree level expressions for the mass matrices and
+  electroweak symmetry breaking conditions, one-loop self energies,
+  one-loop tadpoles corrections and two-loop renormalisation group
+  equations(RGEs) for the model specified in the model
+  files. Additional corrections which have been calculated elsewhere,
+  such as two-loop corrections to the Higgs masses\footnote{By default FlexibleSUSY has two loop corrections to the Higgs masses for the MSSM\cite{Degrassi:2001yf,Brignole:2001jy,Dedes:2002dy,Brignole:2002bz,Dedes:2003km}  and NMSSM \cite{Degrassi:2009yq} in FORTRAN files supplied by Pietro Slavich.
+These are the same corrections which are implemented in many of the public spectrum generators.} may be added in
+  file specified by the user.
+
+ \fs then
+converts these algerbraic expressions from \sarah into C++ code using
+the C++11 standard. \fs uses some parts of
+SOFTSUSY\cite{Allanach:2001kg}, the very fast Eigen library
+\cite{eigen}, as well as the gnu scientific library and the boost
+library to create numerical routines which solve the RGEs and EWSB
+conditions.  The main solver routine of \fs then calls these routines
+to find simultaneous solutions for the EWSB, low energy data, and
+other, user supplied, boundary conditions (e.g. the usual mSUGRA
+universality conditions) and calculates the mass spectrum.
 
@@
  and calculators are added and a subsequent release with an
- alternative lattice solver is already planned.{\color{red} Or Indeed
-   flexible SUSY is distributed with two different main solvers, the
-   two-scale solver based on the approach of \cite{} and designed for
-   speed and the a lattice solver designed to solve tricky cases where
-   the fixed point iteration of the two-scale approach fails to
-   converge.}  In addition if one already has an existing code for
- some other purpose and simply wishes to improve upon this by adding,
- e.g, new RGEs or self energies the modular fraework makes this fairly
- straightward{\color{red} Is this true?  because it should be...}
+ alternative lattice solver is already planned.  In addition if one
+ already has an existing code for some other purpose and simply wishes
+ to improve upon this by adding, e.g, new RGEs or self energies the
+ modular framework makes this straightforward by adding or
+ interfacing with the appropriate routine.
 
@@
 OnlyLowEnergyFlexibleSUSY = False; (* default *)
+
+PotentialLSPParticles = { Chi, Cha, Glu, Sv, Su, Sd, Se };
 \end{lstlisting}
@@
 input blocks in the SLHA input file and will be set at the susy-scale.
+
+\fs can create the helper function \code{get_lsp()}, which finds the
+lightest supersymmetric particle (LSP).  To let this function be
+created by \fs, the model file variable \code{PotentialLSPParticles}
+must be set to a list of SUSY particles which are potential LSPs.  In
+the model file example above, the particles \code{Chi}, \code{Cha},
+\code{Glu}, \code{Sv}, \code{Su}, \code{Sd}, \code{Se} (neutralino,
+chargino, gluino, sneutrino, up-type squark, down-type squark,
+selectron) are considered to be LSPs.
 

### 2014-05-06 doc/paper.tex
document minimum and maximum scale restrictions

@@
 
+HighScaleMinimum = 1.0 10^10; (* optional *)
+
+HighScaleMaximum = 1.0 10^18; (* optional *)
+
 HighScaleInput = {
@@
 OnlyLowEnergyFlexibleSUSY = False; (* default *)
-
-PotentialLSPParticles = { Chi, Cha, Glu, Sv, Su, Sd, Se };
 \end{lstlisting}
@@
 a list of parameter settings to be applied at the scale (concatenation
-of the constraint name and \code{Input}).
+of the constraint name and \code{Input}).  Optional a minimum and a
+maximum value for the scale can be given (concatenation of the
+constraint name and \code{Minimum} or \code{Maximum}, respectively).
+This avoids underflows or overflows of the scale value during the
+iteration.  This is especially useful in models where the iteration is
+very unstable and the value of the scale is very sensitive to the
+model parameters.
 %
@@
 input blocks in the SLHA input file and will be set at the susy-scale.
-
-\fs can create the helper function \code{get_lsp()}, which finds the
-lightest supersymmetric particle (LSP).  To let this function be
-created by \fs, the model file variable \code{PotentialLSPParticles}
-must be set to a list of SUSY particles which are potential LSPs.  In
-the model file example above, the particles \code{Chi}, \code{Cha},
-\code{Glu}, \code{Sv}, \code{Su}, \code{Sd}, \code{Se} (neutralino,
-chargino, gluino, sneutrino, up-type squark, down-type squark,
-selectron) are considered to be LSPs.
 

### 2014-05-08 meta/FlexibleSUSY.m
simplify function call

@@ -527,10 +527,7 @@ WriteModelClass[massMatrices_List, ewsbEquations_List,
            enablePoleMassThreads = True;
            callAllLoopMassFunctionsInThreads = LoopMasses`CallAllPoleMassFunctions[FlexibleSUSY`FSEigenstates, enablePoleMassThreads];
            masses                       = FlexibleSUSY`M[TreeMasses`GetMassEigenstate[#]]& /@ massMatrices;
-           If[Head[PotentialLSPParticles] === List &&
-              PotentialLSPParticles =!= {},
-              {lspGetters, lspFunctions} = LoopMasses`CreateLSPFunctions[FlexibleSUSY`PotentialLSPParticles];
-             ];
+           {lspGetters, lspFunctions}   = LoopMasses`CreateLSPFunctions[FlexibleSUSY`PotentialLSPParticles];
            printMasses                  = WriteOut`PrintParameters[masses, "ostr"];
            mixingMatrices               = Flatten[TreeMasses`GetMixingMatrixSymbol[#]& /@ massMatrices];
            printMixingMatrices          = WriteOut`PrintParameters[mixingMatrices, "ostr"];

### 2014-05-08 meta/LoopMasses.m

@@ -772,6 +772,8 @@ GetLightestMassEigenstate[mass_] :=
        FlexibleSUSY`M[mass],
        FlexibleSUSY`M[mass][GetDimensionStartSkippingGoldstones[mass] - 1]];
 
+CreateLSPFunctions[{}] := {"", ""};
+
 CreateLSPFunctions[masses_List] :=
     Module[{prototype, function, mass, info, particleType, m},
            info = FlexibleSUSY`FSModelName <> "_info";
@@ -804,6 +806,8 @@ CConversion`ToValidCSymbolString[mass /. FlexibleSUSY`M -> Identity] <>
                       "& particle_type)\n{\n" <> IndentText[body] <> "}\n";
            Return[{prototype, function}];
           ];
+
+CreateLSPFunctions[_] := {"", ""};
 
 End[];
 

### 2014-05-08 doc/paper.tex
change interface function display() --> get()

@@
     }{
-      + \umlvirt{display() : const Eigen::ArrayXd} \\
+      + \umlvirt{get() : const Eigen::ArrayXd} \\
       + \umlvirt{set(v : const Eigen::ArrayXd\&) : void}\\
@@
     \umlclass[x=0, y=-4]{MSSM\_susy\_parameters}{}{
-      + display() : const Eigen::ArrayXd \\
+      + get() : const Eigen::ArrayXd \\
       + set(v : const Eigen::ArrayXd\&) : void\\
@@
     \umlclass[x=0, y=-8]{MSSM\_soft\_parameters}{}{
-      + display() : const Eigen::ArrayXd \\
+      + get() : const Eigen::ArrayXd \\
       + set(v : const Eigen::ArrayXd\&) : void\\

### 2014-05-08 src/betafunction.cpp

@@ -49,7 +49,7 @@ void Beta_function::run(double x1, double x2, double eps)
    if (std::fabs(x2) < tol)
       throw NonPerturbativeRunningError(x2);
 
-   Eigen::ArrayXd y(display());
+   Eigen::ArrayXd y(get());
    runge_kutta::Derivs derivs = boost::bind(&Beta_function::derivatives,
                                             this, _1, _2);
 

### 2014-05-08 src/betafunction.hpp

@@ -38,7 +38,7 @@ public:
    unsigned get_loops() const { return loops; }
    unsigned get_thresholds() const { return thresholds; }
 
-   virtual const Eigen::ArrayXd display() const = 0;
+   virtual const Eigen::ArrayXd get() const = 0;
    virtual void set(const Eigen::ArrayXd&) = 0;
    virtual Eigen::ArrayXd beta() const = 0;
 

### 2014-05-09 doc/paper.tex
updating speed test results (including non-fv FS variant)

@@
 \label{Sec:comparison}
-We compare the run-time of the four different spectrum generators
-FlexibleSUSY (version 0.5.4), Softsusy (version 3.4.0), SPheno
-(3.2.4), SPhenoMSSM (generated with SARAH 4.1.0 and linked against
-SPheno 3.2.4).  FlexibleSUSY and Softsusy were compiled with g++ 4.8.0
-and Intel ifort 13.1.3 20130607.  SPheno and SPhenoMSSM were compiled
-with Intel ifort 13.1.3 20130607\footnote{We are using Intel's ifort
-  compiler, instead of gfortran, because it decreases the run-time of
-  SPheno and SPhenoMSSM by approximately a factor $1.5$.}.  We're
-generating random CMSSM parameter points with $m_0\in [50,1000]$
+We compare the run-time of five different CMSSM spectrum generators:
+%
+\begin{itemize}
+\item FlexibleSUSY-FV (version 1.0.0), which is based on SARAH's MSSM
+  model file that allows flavour violation,
+\item FlexibleSUSY-NoFV (version 1.0.0), without flavour violation
+\item Softsusy (version 3.4.0), without flavour violation
+\item SPheno (3.2.4), without flavour violation
+\item SPhenoMSSM (generated with SARAH 4.1.0 and linked against SPheno
+  3.2.4), which uses SARAH's MSSM model file that includes flavour
+  violation.
+\end{itemize}
+%
+FlexibleSUSY and Softsusy were compiled with g++ 4.8.0 and Intel ifort
+13.1.3 20130607.  SPheno and SPhenoMSSM were compiled with Intel ifort
+13.1.3 20130607\footnote{We are using Intel's ifort compiler, instead
+  of gfortran, because it decreases the run-time of SPheno and
+  SPhenoMSSM by approximately a factor $1.5$.}.  We're generating
+$2\cot 10^{4}$ random CMSSM parameter points with $m_0\in [50,1000]$
 $m_{1/2}\in [50,1000]$, $\tan\beta\in [1,100]$, $\sign\mu\in
@@
 average run-times for three different CPU types can be found in
-\tablename~\ref{tab:run-time-comparison}.  The first row shows the
+\tablename~\ref{tab:run-time-comparison}.  The first column shows the
 run-time on a Intel Core2 Duo (P8600) where only one core was enabled.
-The second row shows the run-time on a Intel Core2 Duo (P8600) where
-both cores were enabled.  In the third row a Intel Xeon (L5640) was
-used, which has $6$ CPU cores.
+The second column shows the run-time on a Intel Core2 Duo (P8600)
+where both cores were enabled.  In the third column a Intel Xeon
+(L5640) was used, which has $6$ CPU cores.
 %
@@
     \midrule
-    FlexibleSUSY 0.5.4    & $0.151\unit{s}$    & $0.114\unit{s}$   & $0.091\unit{s}$\\
-    Softsusy 3.4.0        & $0.174\unit{s}$    & $0.169\unit{s}$   & $0.162\unit{s}$\\
-    SPheno 3.2.4          & $0.119\unit{s}$    & $0.115\unit{s}$   & $0.108\unit{s}$\\
-    SPhenoMSSM 4.1.0${}^*$ & $0.416\unit{s}$    & $0.401\unit{s}$   & $0.366\unit{s}$\\
+    FlexibleSUSY-NoFV 0.5.4 & $0.086\unit{s}$    & $0.079\unit{s}$   & $0.060\unit{s}$\\
+    SPheno 3.2.4            & $0.119\unit{s}$    & $0.114\unit{s}$   & $0.101\unit{s}$\\
+    Softsusy 3.4.0          & $0.175\unit{s}$    & $0.171\unit{s}$   & $0.147\unit{s}$\\
+    \midrule
+    FlexibleSUSY-FV 0.5.4   & $0.150\unit{s}$    & $0.113\unit{s}$   & $0.074\unit{s}$\\
+    SPhenoMSSM 4.1.0${}^*$  & $0.415\unit{s}$    & $0.401\unit{s}$   & $0.370\unit{s}$\\
     \bottomrule
   \end{tabular}
-  \caption{Average run-time for random CMSSM parameter points.
-    FlexibleSUSY and Softsusy were compiled
-    with g++ 4.8.0 and ifort 13.1.3 20130607.
-    SPheno, SPhenoMSSM were
-    compiled with ifort 13.1.3 20130607.
-    ${}^*$SPhenoMSSM was generated with SARAH 4.1.0 and linked
-    against SPheno 3.2.4.}
+  \caption{Average run-time of CMSSM spectrum generators for
+    for random parameter points.  The upper }
   \label{tab:run-time-comparison}
 \end{table}
-%
-We find that on the single-core machine SPheno is fastest with
-$0.119\unit{s}$, and FlexibleSUSY is second with $0.151\unit{s}$.  In
-the case of 2 or 6 CPU cores, FlexibleSUSY is fastest with
-$0.114\unit{s}$ and $0.091\unit{s}$, respectively.  The reason is that
-FlexibleSUSY calculates each pole mass in a separate thread, and
-therefore benefits from multi-core CPUs.
+
+Under both the non-flavor violating spectrum generators
+FlexibleSUSY-NoFV, Softsusy and SPheno (first three rows) as well as
+the flavor violating ones FlexibleSUSY-NoFV and SPhenoMSSM (4th and
+5th row) we find that \fs is significantly fastest.  We also find that
+increasing the number of CPU cores reduces the run-time of \fs.  The
+reason is that \fs calculates each pole mass in a separate thread, and
+therefore benefits from multi-core CPUs.  Under the hand-written CMSSM
+spectrum generators SPheno and Softsusy we find that SPheno is fastest
+on average.
 

### 2014-05-10 meta/FlexibleSUSY.m
default initialization of model file variables

@@ -23,27 +23,27 @@ FSModelName;
 FSLesHouchesList;
 FSUnfixedParameters;
 InputParameters;
-DefaultParameterPoint;
-EWSBOutputParameters;
+DefaultParameterPoint = {};
+EWSBOutputParameters = {};
 SUSYScale;
 SUSYScaleFirstGuess;
-SUSYScaleInput;
+SUSYScaleInput = {};
 SUSYScaleMinimum;
 SUSYScaleMaximum;
 HighScale;
 HighScaleFirstGuess;
-HighScaleInput;
+HighScaleInput = {};
 HighScaleMinimum;
 HighScaleMaximum;
 LowScale;
 LowScaleFirstGuess;
-LowScaleInput;
+LowScaleInput = {};
 LowScaleMinimum;
 LowScaleMaximum;
-InitialGuessAtLowScale;
-InitialGuessAtHighScale;
+InitialGuessAtLowScale = {};
+InitialGuessAtHighScale = {};
 OnlyLowEnergyFlexibleSUSY = False;
-TreeLevelEWSBSolution;
+TreeLevelEWSBSolution = {};
 Pole;
 FSMinimize;
 FSFindRoot;

### 2014-05-10 meta/FlexibleSUSY.m
update year

@@ -3,7 +3,7 @@ BeginPackage["FlexibleSUSY`", {"SARAH`", "AnomalousDimension`", "BetaFunction`",
 
 FS`Version = StringTrim[Import[FileNameJoin[{Global`$flexiblesusyConfigDir,"version"}], "String"]];
 FS`Authors = {"P. Athron", "J. Park", "D. Stöckinger", "A. Voigt"};
-FS`Years   = {2013};
+FS`Years   = {2013, 2014};
 
 Print["*****************************************************"];
 Print["FlexibleSUSY ", FS`Version];

### 2014-05-10 doc/paper.tex
update version number of FS
remove duplicate reference
improve description of model class hierarchy and boundary conditions

@@
 \newcommand{\unity}{\mathbf{1}}
+\newcommand{\figref}[1]{\figurename~\ref{#1}}
+\newcommand{\secref}[1]{Sec.~\ref{#1}}
 \DeclareMathOperator{\sign}{sign}
@@
 solvers provide a constraint interface,
-\figurename~\ref{fig:schematic-two-scale-constraint-interface},
+\figref{fig:schematic-two-scale-constraint-interface},
 %
@@
 \fs can create the helper function \code{get_lsp()}, which finds the
-lightest supersymmetric particle (LSP).  To let this function be
+lightest supersymmetric particle (LSP).  To have this function be
 created by \fs, the model file variable \code{PotentialLSPParticles}
@@
 \subsection{Model parameters and RGEs}
-Even in the simplest supersymmetric model the RGEs form a very large
-set of coupled non-linear differential equations.  In a general
-(softly broken) supersymmetric theory the soft-breaking parameters
-depend on themselves and the susy parameters.  However, the susy
-parameters are independent of the soft-breaking parameters
-\cite{Jones:1974pg,Jones:1983vk,West:1984dg,Martin:1993yx,Yamada:1993ga,MV94,Fonseca:2011vn,Sperling:2013eva,Sperling:2013xqa}.
-This makes it possible to split the model parameters into the
-following two subsets:
+
+The parameters of the susy model together with their RGEs are stored
+in the model class hierarchy, see the UML diagram in
+\figref{fig:parameter-classes}.
+%
+\begin{figure}
+  \centering
+  \tikzumlset{fill class=white}
+  \begin{tikzpicture}
+    \umlclass[x=0, y=8, type=abstract]{Beta\_function}{
+    }{
+      + \umlvirt{get()}\\
+      + \umlvirt{set()}\\
+      + \umlvirt{beta()}\\
+      + run\_to()
+    }
+    \umlclass[x=0, y=4]{<model>\_susy\_parameters}{
+      -- susy parameters
+    }{
+      + get()\\
+      + set()\\
+      + beta()\\
+      + run\_to()
+    }
+    \umlclass[x=0, y=0]{<model>\_soft\_parameters}{
+      -- soft-breaking parameters
+    }{
+      + get()\\
+      + set()\\
+      + beta()\\
+      + run\_to()
+    }
+    \umlinherit{<model>\_susy\_parameters}{Beta\_function}
+    \umlinherit{<model>\_soft\_parameters}{<model>\_susy\_parameters}
+  \end{tikzpicture}
+  \caption{Model parameter class hierarchy.}
+  \label{fig:parameter-classes}
+\end{figure}
+
+The top of the hierarchy is formed by the \code{Beta_function}
+interface class, which defines the basic RGE running interface.  It
+provides the user function \code{run_to()}, which integrates the RGEs
+up to a given scale using an adaptive Runge-Kutta algorithm.  Thereby
+it uses the pure virtual functions \code{get()}, \code{set()} and
+\code{beta()}, that need to be implemented by the derived class.  The
+\code{get()} and \code{set()} functions return and set the model
+parameters in form of a long vector, respectively.  The \code{beta()}
+method returns the $\beta$-function for each model parameter in form
+of a long vector as well.
+
+All susy model parameters and their $\beta$-functions are contained in
+the derived classes.  The structure of the $\beta$-functions of a
+general susy model
+\cite{Jones:1974pg,Jones:1983vk,West:1984dg,Martin:1993yx,Yamada:1993ga,MV94,Fonseca:2011vn,Sperling:2013eva,Sperling:2013xqa}
+allows to split the parameters into two classes:
 %
@@
 %
-The susy parameters together with their two-loop renormalisation group
-equations are stored in the class \code{<model>_susy_parameters}.  The
-soft-breaking model parameters together with their two-loop
-renormalisation group equations are stored in the class
-\code{<model>_soft_parameters}.  The dependence of the soft-breaking
-parameters on the susy parameters is implemented via inheritance, see
-\figurename~\ref{fig:parameter-classes}, where
-\code{<model>_soft_parameters} inherits from
-\code{<model>_susy_parameters}.  In this way, the class
-\code{<model>_soft_parameters} can implement the soft-breaking beta
-functions in terms of the susy and soft-breaking parameters, while the
-class \code{<model>_susy_parameters} is independent of the
-soft-breaking parameters.
-%
-\begin{figure}
-  \centering
-  \tikzumlset{fill class=white}
-  \begin{tikzpicture}
-    \umlclass[x=0, y=3]{model\_susy\_parameters}{
-      - susy parameters
-    }{
-      + beta()
-    }
-    \umlclass[x=0, y=0]{model\_soft\_parameters}{
-      - soft-breaking parameters
-    }{
-      + beta()
-    }
-    \umlinherit{model\_soft\_parameters}{model\_susy\_parameters}
-  \end{tikzpicture}
-  \caption{Model parameter class structure.}
-  \label{fig:parameter-classes}
-\end{figure}
+The $\beta$-functions of the susy parameters do only depend on the
+susy parameters themselfs.  They do not depend on soft-breaking
+parameters.  However, the $\beta$-functions of the soft-breaking
+parameters depend on all model parameters in general.
+
+This property is reflected in the C++ code as well: The class
+\code{<model>_susy_parameters} direcly inherits from
+\code{Beta_functions} and implements the $\beta$-functions of the susy
+parameters.  The class of soft-breaking parameters
+\code{<model>_soft_parameters} in turn inherits from
+\code{<model>_susy_parameters} and implements the $\beta$-functions of
+the soft-breaking parameters in terms of all model parameters.
+This hierarchy allows to
+%
+\begin{enumerate}
+\item use the RGE running of all model parameters via the common
+  \code{Beta_function} interface and to
+\item run the susy parameters independently of the soft-breaking
+  parameters.
+\end{enumerate}
 
 \subsection{Boundary conditions}
-In \fs the RGE boundary conditions are classes which implement
+
+The parameters of a susy model have to meet certain boundary
+conditions.  For example, the running gauge and Yukawa couplings
+should match the ones of the Standard Model at the scale $M_Z$.
+Usually also high-scale boundary conditions on the soft-breaking
+parameters are applied, as for example in mSUGRA scenarios.
+
+In \fs the boundary conditions, imposed on the model parameters, are
+(at the C++ level) classes which implement the common
 \code{Constraint<Two\_scale>} interface, see
-\figurename~\ref{fig:schematic-two-scale-constraint-interface}.
+\figref{fig:schematic-two-scale-constraint-interface}.
 %
@@
    virtual double get_scale() const = 0;
-   virtual void set_model(Two_scale_model*) = 0;
 };
@@
 %
-The three pure virtual functions \code{apply()}, \code{get\_scale()}
-and \code{set\_model()} need to be implemented
+The \code{get_scale()} function is supposed to return the
+renormalisation scale at which the constraint is to be imposed.  The
+\code{apply()} method imposes the constraint by setting model
+parameters to values as chosen by the user.
+
+By default, \fs creates three default boundary conditions for each
+susy model:
 %
 \begin{itemize}
-\item \code{apply()}: implements the constraint by fixing model
-  parameters
-\item \code{get\_scale()}: returns the scale at which the constraint
-  is to be applied.
-\item \code{set\_model()}: connects the constraint to a model class,
-  in order to access the model parameters
+\item The \emph{high-scale constraint} is indended to set boundary
+  conditions on the model parameters at some very high scale, e.g.\
+  the GUT scale $M_X$.  The high-scale is defined by the value given
+  in the variable \code{HighScale}.  In the model file in
+  \secref{Sec:modfile} it is defined to be the unification scale $M_X$
+  where $g_1(M_X) = g_2(M_X)$.  The model parameters which, are set in
+  the $apply()$ function at the high-scale, are extracted from the
+  \code{HighScaleInput} variable.
+
+\item The \emph{SUSY-scale constraint} is indended to set boundary
+  conditions at the mass scale $M_S$ of the new susy particles.  The
+  value of $M_S$ is defined by the model file variable
+  \code{SUSYScale}.  In the example model file in \secref{Sec:modfile}
+  it is defined to be $\sqrt{m_{\tilde{u}_1} m_{\tilde{u}_6}}$, where
+  $m_{\tilde{u}_1}$ is the mass of the lightest, and $m_{\tilde{u}_6}$
+  is the mass of the heaviest up-type squark.  The generated
+  \code{apply()} function sets the model parameters, defined in
+  \code{SUSYScaleInput}.  Finally, the \code{apply()} solves the EWSB
+  equations at the loop level, by adjusting the parameters given in
+  \code{EWSBOutputParameters} such that the effective Higgs potential
+  is minimized.  See \secref{sec:ewsb} for a more detailed description
+  of the EWSB in \fs.
+
+\item The \emph{low-scale constraint} is indended to match the susy
+  model to the Standard Model at the scale $M_Z$.  It does so by
+  calculating the gauge and Yukawa couplings of the susy model from
+  the known Standard Model fermion and gauge boson masses, as well as
+  $\alpha_{\text{e.m.},\text{SM}}^{(5),\text{\MSbar}}(M_Z)$ and
+  $\alpha_{\text{s},\text{SM}}^{(5),\text{\MSbar}}(M_Z)$.  See
+  \secref{sec:calculation-of-gauge-couplings} and
+  \ref{sec:calculation-of-yukawa-couplings} for more details.  In
+  addition to the gauge and Yukawa couplings, the model parameter
+  constraints given in \code{LowScaleInput} are imposed here.
 \end{itemize}
-%
-Per default, \fs creates three default boundary conditions for each
-model:
-%
-\begin{itemize}
-\item \emph{GUT constraint:} Here the \code{apply()} function
-  calculates the value of the unification scale $M_X$ (from the
-  condition given in \code{HighScale}) and sets the model parameters
-  given in \code{HighScaleInput}.  The \code{get\_scale()} returns the
-  value of $M_X$.
-\item \emph{SUSY constraint:} Here the \code{apply()} function
-  calculates the value of the susy scale $M_S$ and sets the model
-  parameters given in \code{SUSYScaleInput}.  Afterwards the EWSB
-  conditions are solved for the parameters given in
-  \code{EWSBOutputParameters} at the one-loop level, i.e., the the
-  parameters in \code{EWSBOutputParameters} are determined such that
-  the one-loop effective potential is minimized.  The
-  \code{get\_scale()} returns the value of $M_S$.
-\item \emph{Low-scale constraint:} This constraint calculates the
-  gauge and Yukawa couplings of the susy model from the known Standard
-  Model gauge couplings and fermion masses (at $M_Z$), using threshold
-  corrections.  Furthermore the parameters given in
-  \code{LowScaleInput} are set.  The \code{get\_scale()} returns the
-  value of $M_Z$.
-\end{itemize}
-
-{\color{red}Different types, how we match to low energy data.
-  reference EWSB or merge that section here.  mention two loop qcd
-  parts from softsusy.}
 
@@
 \subsubsection{Electroweak symmetry breaking}
+\label{sec:ewsb}
+
 In \fs the one-loop electroweak symmetry breaking conditions are
@@
     \midrule
-    FlexibleSUSY-NoFV 0.5.4 & $0.086\unit{s}$    & $0.079\unit{s}$   & $0.060\unit{s}$\\
+    FlexibleSUSY-NoFV 1.0.0 & $0.086\unit{s}$    & $0.079\unit{s}$   & $0.060\unit{s}$\\
     SPheno 3.2.4            & $0.119\unit{s}$    & $0.114\unit{s}$   & $0.101\unit{s}$\\
@@
     \midrule
-    FlexibleSUSY-FV 0.5.4   & $0.150\unit{s}$    & $0.113\unit{s}$   & $0.074\unit{s}$\\
+    FlexibleSUSY-FV 1.0.0   & $0.150\unit{s}$    & $0.113\unit{s}$   & $0.074\unit{s}$\\
     SPhenoMSSM 4.1.0${}^*$  & $0.415\unit{s}$    & $0.401\unit{s}$   & $0.370\unit{s}$\\
@@
 
-%\cite{Degrassi:2009yq}
-\bibitem{Degrassi:2009yq}
-  G.~Degrassi and P.~Slavich,
-  %``On the radiative corrections to the neutral Higgs boson masses in the NMSSM,''
-  Nucl.\ Phys.\ B {\bf 825} (2010) 119
-  [arXiv:0907.4682 [hep-ph]].
-  %%CITATION = ARXIV:0907.4682;%%
-  %44 citations counted in INSPIRE as of 04 Apr 2014
-
 \end{thebibliography}

### 2014-05-10 doc/paper.tex
my eyes suffer less with 11pt and CM font
try to make Section 6.2.1 a bit clearer

@@
-\documentclass[final,3p,times,pdflatex]{elsarticle}
+\documentclass[final,3p,11pt,pdflatex]{elsarticle}
 \usepackage[utf8x]{inputenc}      % input font encoding
@@
 
-The low-scale constraint calculates the gauge couplings $g_i(M_Z)$.
-It does so by first calculating $e_{\text{susy}}^{\text{\DRbar}}(M_Z)$
-and $g_{3,\text{susy}}^{\text{\DRbar}}(M_Z)$ via
+The low-scale constraint automatically calculates the \DRbar\ gauge
+couplings $g_{i,\text{susy}}^{\text{\DRbar}}(M_Z)$ in the susy model
+at the scale $M_Z$.  It starts from the known electromagnetic and
+strong \MSbar\ couplings in the Standard Model including only $5$
+quark flavours
+$\alpha_{\text{e.m.},\text{SM}}^{(5),\text{\MSbar}}(M_Z) = 1/127.944$
+and $\alpha_{\text{s},\text{SM}}^{(5),\text{\MSbar}}(M_Z) = 0.1185$
+\cite{Beringer:1900zz}.  These are converted to the electromagnetic
+and strong \DRbar\ couplings in the susy model
+$e_{\text{susy}}^{\text{\DRbar}}(M_Z)$ and
+$g_{3,\text{susy}}^{\text{\DRbar}}(M_Z)$ as
 %
@@
     \Delta\alpha_{\text{e.m.},\text{susy}}(M_Z)} ,\\
+    e_{\text{susy}}^{\text{\DRbar}}(M_Z) &=
+    \sqrt{4\pi\alpha_{\text{e.m.},\text{susy}}^{\text{\DRbar}}(M_Z)}, \\
+  \alpha_{\text{s},\text{susy}}^{\text{\DRbar}}(M_Z) &=
+  \frac{\alpha_{\text{s},\text{SM}}^{(5),\text{\MSbar}}(M_Z)}{1 -
+    \Delta\alpha_{\text{s},\text{SM}}(M_Z)
+    - \Delta\alpha_{\text{s},\text{susy}}(M_Z)} ,\\
+  g_{3,\text{susy}}^{\text{\DRbar}}(M_Z) &=
+  \sqrt{4\pi\alpha_{\text{s},\text{susy}}^{\text{\DRbar}}(M_Z)} .
+\end{align}
+%
+The $\Delta\alpha_i(\mu)$ are threshold corrections and read
+%
+\begin{align}
   \Delta\alpha_{\text{e.m.},\text{SM}}(\mu) &=
@@
       i}
-    C_i \log{\frac{m_i}{\mu}} \right],\\
-    e_{\text{susy}}^{\text{\DRbar}}(M_Z) &=
-    \sqrt{4\pi\alpha_{\text{e.m.},\text{susy}}^{\text{\DRbar}}(M_Z)}
-\end{align}
-%
-\begin{align}
-  \alpha_{\text{s},\text{susy}}^{\text{\DRbar}}(M_Z) &=
-  \frac{\alpha_{\text{s},\text{SM}}^{(5),\text{\MSbar}}(M_Z)}{1 -
-    \Delta\alpha_{\text{s},\text{SM}}(M_Z)
-    - \Delta\alpha_{\text{s},\text{susy}}(M_Z)} ,\\
+    F_i T_i \log{\frac{m_i}{\mu}} \right],\\
   \Delta\alpha_{\text{s},\text{SM}}(\mu) &=
@@
   \frac{\alpha_\text{s}}{2\pi}\left[ \frac{1}{2}-\sum_{\text{susy
-        particle } i} C_i \log{\frac{m_i}{\mu}} \right] ,\\
-  g_{3,\text{susy}}^{\text{\DRbar}}(M_Z) &=
-  \sqrt{4\pi\alpha_{\text{s},\text{susy}}^{\text{\DRbar}}(M_Z)} ,
+        particle } i} F_i T_i \log{\frac{m_i}{\mu}} \right] .
 \end{align}
 %
-where
+The constants $T_i$ are the dynkin indices of the representation of
+particle $i$ with respect to the gauge group, and $F_i$ are particle
+type-specific constants \cite{Hall:1980kf}
 %
 \begin{align}
-  \alpha_{\text{e.m.},\text{SM}}^{(5),\text{\MSbar}}(M_Z) &= 1/127.944,\\
-  \alpha_{\text{s},\text{SM}}^{(5),\text{\MSbar}}(M_Z) &= 0.1185,
+  F_i =
+  \begin{cases}
+    2/3 & \text{if particle $i$ is a Majorana fermion},\\
+    4/3 & \text{if particle $i$ is a Dirac fermion},\\
+    1/6 & \text{if particle $i$ is a real scalar},\\
+    1/3 & \text{if particle $i$ is a complex scalar}.
+  \end{cases}
 \end{align}
 %
-are the \MSbar electromagnetic and strong cougling constants in the
-Standard Model including only $5$ quark flavours
-\cite{Beringer:1900zz}.  Afterwards, SARAH's tree-level expressions
-for the Weinberg angle and the $U(1)_Y$ and $SU(2)_L$ gauge couplings
-$g_Y$ and $g_2$ in terms of $M_{W,\text{susy}}^{\text{\DRbar}}(M_Z)$,
-$M_{Z,\text{susy}}^{\text{\DRbar}}(M_Z)$ and
-$e_{\text{susy}}^{\text{\DRbar}}(M_Z)$ are used to calculate
-$g_{1,\text{susy}}^{\text{\DRbar}}(M_Z)$ and
-$g_{2,\text{susy}}^{\text{\DRbar}}(M_Z)$.  In the MSSM, for example,
-one has
+Afterwards, SARAH's tree-level expression for the Weinberg angle
+$\theta_W$ in terms of $M_{W,\text{susy}}^{\text{\DRbar}}(M_Z)$,
+$M_{Z,\text{susy}}^{\text{\DRbar}}(M_Z)$ is used to calculate
+$\theta_W$ in the susy model in the \DRbar\ scheme.  In the MSSM, for
+example, one has
 %
 \begin{align}
-  \theta_{W,\text{susy}}^{\text{\DRbar}}(M_Z) &=
-  \arcsin\sqrt{1 - \left(\frac{M_{W,\text{susy}}^{\text{\DRbar}}(M_Z)}{M_{Z,\text{susy}}^{\text{\DRbar}}(M_Z)}\right)^2} ,\\
+  \theta_{W,\text{susy}}^{\text{\DRbar}}(M_Z) &= \arcsin\sqrt{1 -
+    \left(\frac{M_{W,\text{susy}}^{\text{\DRbar}}(M_Z)}{M_{Z,\text{susy}}^{\text{\DRbar}}(M_Z)}\right)^2}
+  .
+\end{align}
+%
+The running \DRbar\ $W$ and $Z$ boson masses are calculated in each
+iteration from the corrsponding pole masses as
+%
+\begin{align}
+  \left(M_{W,\text{susy}}^{\text{\DRbar}}(M_Z)\right)^2 &=
+  \left(M_W^{\text{pole}}\right)^2 + \re \Pi_{WW}^T(p^2 = (M_W^{\text{pole}})^2, \mu=M_Z) ,\\
+  \left(M_{Z,\text{susy}}^{\text{\DRbar}}(M_Z)\right)^2 &=
+  \left(M_Z^{\text{pole}}\right)^2 + \re \Pi_{ZZ}^T(p^2 = (M_Z^{\text{pole}})^2, \mu=M_Z) ,
+\end{align}
+%
+where $M_W^{\text{pole}} = 80.404\unit{GeV}$ and $M_Z^{\text{pole}} =
+91.1876\unit{GeV}$ \cite{Beringer:1900zz}.  Having
+$e_{\text{susy}}^{\text{\DRbar}}(M_Z)$ and
+$\theta_{W,\text{susy}}^{\text{\DRbar}}(M_Z)$ allows to calculate the
+(GUT-normalized) $U(1)_Y$ and $SU(2)_L$ gauge couplings in the susy
+model.  In the MSSM one has for instance
+%
+\begin{align}
   g_{1,\text{susy}}^{\text{\DRbar}}(M_Z) &=
@@
   g_{2,\text{susy}}^{\text{\DRbar}}(M_Z) &=
-  \frac{e_{\text{susy}}^{\text{\DRbar}}(M_Z)}{\sin\theta_{W,\text{susy}}^{\text{\DRbar}}(M_Z)} ,
+  \frac{e_{\text{susy}}^{\text{\DRbar}}(M_Z)}{\sin\theta_{W,\text{susy}}^{\text{\DRbar}}(M_Z)} .
 \end{align}
-%
-where
-%
-\begin{align}
-  \left(M_{W,\text{susy}}^{\text{\DRbar}}(M_Z)\right)^2 &=
-  \left(M_W^{\text{pole}}\right)^2 + \re \Pi_{WW}^T(p^2 = (M_W^{\text{pole}})^2) ,\\
-  \left(M_{Z,\text{susy}}^{\text{\DRbar}}(M_Z)\right)^2 &=
-  \left(M_Z^{\text{pole}}\right)^2 + \re \Pi_{ZZ}^T(p^2 = (M_Z^{\text{pole}})^2)
-\end{align}
-%
-and $M_W^{\text{pole}} = 80.404\unit{GeV}$, $M_Z^{\text{pole}} =
-91.1876\unit{GeV}$ \cite{Beringer:1900zz}.
 
@@
 
+%\cite{Hall:1980kf}
+\bibitem{Hall:1980kf}
+  L.~J.~Hall,
+  %``Grand Unification of Effective Gauge Theories,''
+  Nucl.\ Phys.\ B {\bf 178} (1981) 75.
+  %%CITATION = NUPHA,B178,75;%%
+  %265 citations counted in INSPIRE as of 10 May 2014
+
 \end{thebibliography}

### 2014-05-10 doc/paper.tex
improve rest of Section 6 a bit

@@
 \newcommand{\secref}[1]{Sec.~\ref{#1}}
+\DeclareMathOperator{\diag}{diag}
 \DeclareMathOperator{\sign}{sign}
@@
 
-At the low-scale constraint the expressions for the Yukawa coupling
-matrices $y_f(M_Z)$ are calculated from the tree-level mass matrices
-of the SM-fermions $m_u$, $m_d$, $m_e$.  In the MSSM, for example,
-these relations read
+At the low-scale the \DRbar\ Yukawa coupling matrices
+$y_f^{\text{\DRbar}}(M_Z)$ ($f=u,d,e$) in the susy model are
+calculated automatically if the user has set $y_f$ to the value
+\code{Automatic} in the \fs model file.  This is done for example in
+the MSSM model file in \secref{Sec:modfile}.  In this case \fs
+expresses the Yukawa couplings in terms of the \DRbar\ fermion mass
+matrices $m_u$, $m_d$, $m_e$.  In the MSSM, for example, these
+relations read
 %
@@
 %
-At the low-scale the fermion mass matrices are composed as
+The fermion mass matrices are composed as
 %
 \begin{align}
-  m_u =
-  \begin{pmatrix}
-    m_{u}^{\userinput} & 0 & 0 \\
-    0 & m_{c}^{\userinput} & 0 \\
-    0 & 0 & m_{t,\text{susy}}^{\text{\DRbar}}(M_Z)
-  \end{pmatrix} ,\\
-  m_d =
-  \begin{pmatrix}
-    m_{d}^{\userinput} & 0 & 0 \\
-    0 & m_{s}^{\userinput} & 0 \\
-    0 & 0 & m_{b,\text{susy}}^{\text{\DRbar}}(M_Z)
-  \end{pmatrix} ,\\
-  m_e =
-  \begin{pmatrix}
-    m_{e}^{\userinput} & 0 & \\
-    0 & m_{\mu}^{\userinput} & 0 \\
-    0 & 0 & m_{\tau,\text{susy}}^{\text{\DRbar}}(M_Z)
-  \end{pmatrix},
+  m_u = \diag(m_{u}^{\userinput}, m_{c}^{\userinput}, m_{t,\text{susy}}^{\text{\DRbar}}(M_Z)) ,\\
+  m_d = \diag(m_{d}^{\userinput}, m_{s}^{\userinput}, m_{b,\text{susy}}^{\text{\DRbar}}(M_Z)) ,\\
+  m_e = \diag(m_{e}^{\userinput}, m_{\mu}^{\userinput}, m_{\tau,\text{susy}}^{\text{\DRbar}}(M_Z)),
 \end{align}
 %
-where $m_{u,c,d,s,e,\mu}^{\userinput}$ are read from the \code{Block
-  SMINPUTS} of the SLHA input file \cite{Skands:2003cj}.  The 3rd
+where $m_{u,c,d,s,e,\mu}^{\userinput}$ are read from the
+\code{SMINPUTS} block of the SLHA input file \cite{Skands:2003cj}.
+The CKM mixing matrix is currently assumed to be diagonal.  The 3rd
 generation quark masses are calculated in the \DRbar scheme from the
 SLHA user input quantities $m_t^\text{pole}$,
-$m_{b,\text{SM}}^{\text{\MSbar}}(M_Z)$,
+$m_{b,\text{SM}}^{\text{\MSbar}}(M_Z)$ and
 $m_{\tau,\text{SM}}^{\text{\MSbar}}(M_Z)$ \cite{Skands:2003cj}.  In
-detail, the top quark \DRbar mass is calculated via
+detail, the top quark \DRbar mass is calculated as
 %
@@
     m_{t,\text{susy}}^{\text{\DRbar}}(\mu) &= m_t^\text{pole} +
-    \re\Sigma_{tt}^{S,\text{heavy}}(m_t^\text{pole}) \\
+    \re\Sigma_{t}^{S,\text{heavy}}(m_t^\text{pole}) \\
     &\phantom{=\;} + m_t^\text{pole}
-    \left[ \re\Sigma_{tt}^{L,\text{heavy}}(m_t^\text{pole}) +
-      \re\Sigma_{tt}^{R,\text{heavy}}(m_t^\text{pole}) + \Delta
+    \left[ \re\Sigma_{t}^{L,\text{heavy}}(m_t^\text{pole}) +
+      \re\Sigma_{t}^{R,\text{heavy}}(m_t^\text{pole}) + \Delta
       m_t^{(1),\text{qcd}} + \Delta m_t^{(2),\text{qcd}} \right] ,
@@
 %
-where the $\Sigma_{ff}^\text{heavy}$ denotes the self-energy of
-fermion $f$ without the Standard Model contributions.  The appearing
-QCD corrections $\Delta m_t^{(1),\text{qcd}}$ and $\Delta
-m_t^{(2),\text{qcd}}$ are taken from
-\cite[Eq.\ (58),(61)]{Bednyakov:2002sf} and read
+where the $\Sigma_{f}^\text{heavy}$ is the self-energy of fermion $f$
+without the Standard Model contributions.  The labels $L,R,S$ denote
+the left-, right- and non-polarized part of the self-energy.  The
+appearing QCD corrections $\Delta m_t^{(1),\text{qcd}}$ and $\Delta
+m_t^{(2),\text{qcd}}$ are taken from \cite{Bednyakov:2002sf} and read
 %
@@
 %
-The \DRbar mass of the bottom quark is calculated via
+The \DRbar mass of the bottom quark is calculated as
 \cite{Baer:2002ek,Skands:2003cj}
@@
   \frac{m_{b,\text{SM}}^{\text{\DRbar}}(\mu)}{1 -
-    \re\Sigma_{bb}^{S,\text{heavy}}(m_{b,\text{SM}}^\text{\MSbar})/m_b
-    - \re\Sigma_{bb}^{L,\text{heavy}}(m_{b,\text{SM}}^\text{\MSbar}) -
-    \re\Sigma_{bb}^{R,\text{heavy}}(m_{b,\text{SM}}^\text{\MSbar})} ,\\
+    \re\Sigma_{b}^{S,\text{heavy}}(m_{b,\text{SM}}^\text{\MSbar})/m_b
+    - \re\Sigma_{b}^{L,\text{heavy}}(m_{b,\text{SM}}^\text{\MSbar}) -
+    \re\Sigma_{b}^{R,\text{heavy}}(m_{b,\text{SM}}^\text{\MSbar})} ,\\
   m_{b,\text{SM}}^{\text{\DRbar}}(\mu) &=
@@
       \pi} - \frac{23}{72} \frac{\alpha_s^2}{\pi^2} + \frac{3
-      g_2^2}{128 \pi^2} + \frac{13 g_Y^2}{1152 \pi^2}\right) ,
+      g_2^2}{128 \pi^2} + \frac{13 g_Y^2}{1152 \pi^2}\right) .
 \end{align}
 %
-where $m_{b,\text{SM}}^{\text{\MSbar}}(M_Z)$ is user input.  The
-\DRbar mass of the $\tau$ is calculated via
+The \DRbar mass of the $\tau$ is calculated as
 %
@@
     m_{\tau,\text{SM}}^{\text{\DRbar}}(\mu) +
-    \re\Sigma_{\tau\tau}^{S,\text{heavy}}(m_{\tau,\text{SM}}^\text{\MSbar}) \\
+    \re\Sigma_{\tau}^{S,\text{heavy}}(m_{\tau,\text{SM}}^\text{\MSbar}) \\
     &\phantom{=\;} + m_{\tau,\text{SM}}^{\text{\DRbar}}(\mu) \left[
-      \re\Sigma_{\tau\tau}^{L,\text{heavy}}(m_{\tau,\text{SM}}^\text{\MSbar})
+      \re\Sigma_{\tau}^{L,\text{heavy}}(m_{\tau,\text{SM}}^\text{\MSbar})
       +
-      \re\Sigma_{\tau\tau}^{R,\text{heavy}}(m_{\tau,\text{SM}}^\text{\MSbar})
+      \re\Sigma_{\tau}^{R,\text{heavy}}(m_{\tau,\text{SM}}^\text{\MSbar})
     \right] ,
@@
   m_{\tau,\text{SM}}^{\text{\DRbar}}(\mu) &= m_{\tau,\text{SM}}^{\text{\MSbar}}(\mu)
-  \left(1 - 3 \frac{g_Y^2 - g_2^2}{128 \pi^2}\right),
+  \left(1 - 3 \frac{g_Y^2 - g_2^2}{128 \pi^2}\right).
 \end{align}
 %
-where $m_{\tau,\text{SM}}^{\text{\MSbar}}(M_Z)$ is user input.  To
-convert the fermion masses from the \MSbar to the \DRbar scheme the
-Yukawa coupling conversion from \cite[Eq.~(19)]{Skands:2003cj} was
-used and it was assumed that the VEV is already given in the \DRbar
-scheme.
+To convert the fermion masses from the \MSbar to the \DRbar scheme the
+Yukawa coupling conversion from \cite[Eq.~(19)]{Skands:2003cj} is used
+and it is assumed that the VEV is defined in the \DRbar scheme.
 
@@
 where $V^\text{tree}$ is the tree-level Higgs potential, $v_i$ is the
-VEVs corresponding to the Higgs field $H_i$ and $t_i$ is the one-loop
+VEVs corresponding to the Higgs fields $H_i$ and $t_i$ is the one-loop
 tadpole diagram of $H_i$.
@@
 \begin{lstlisting}[language=C++]
-int MSSM<Two_scale>::tadpole_equations(const gsl_vector* x, void* params, gsl_vector* f)
+int MSSM<Two_scale>::tadpole_equations(const gsl_vector* x, void* params,
+                                       gsl_vector* f)
 {
@@
 %
-Here \code{x} is the vector of EWSB output parameters
-(\code{EWSBOutputParameters}) and \code{f} is a vector which contains
+Here \code{x} is the vector of EWSB output parameters (defined via
+\code{EWSBOutputParameters}) and \code{f} is a vector which contains
 the one-loop EWSB eqs.\ \eqref{eq:one-loop-ewsb-eq}.  This
 \code{tadpole_equations()} function is called inside of
-\code{MSSM<Two_scale>::solve_ewsb_iteratively_with} then via
+\code{MSSM<Two_scale>::solve_ewsb_iteratively_with} then as
 %
@@
 If higher accuracy is required additional routines with higher order
-corrections can be added in a very simple way.  For example in the
-MSSM by default \fs adds two loop Higgs FORTRAN routines supplied by
-P.~Slavich from \cite{slavich-degrassi} to add two loop corrections of
-$\oatas$, $\oabas$, $\oatq$, $\oabatau$, $\oabq$, $\oatauq$ and
-$\oatab$ as follows...  Something similar is done for the NMSSM, but
-in this case the NMSSM $\oatas$, $\oabas$ pieces come from
-\cite{Degrassi:2009yq}.  Corrections from other sources can be added
-to the generated code by...
+corrections can be added by setting \code{UseHiggs2LoopMSSM = True;}
+in the model file.  For example in the MSSM by default \fs adds two
+two-loop Higgs FORTRAN routines supplied by P.~Slavich from
+\cite{Degrassi:2001yf,Brignole:2001jy,Dedes:2002dy,Brignole:2002bz,Dedes:2003km}
+to add two loop corrections of $\oatas$, $\oabas$, $\oatq$,
+$\oabatau$, $\oabq$, $\oatauq$ and $\oatab$.  In the NMSSM
+contributions of the order $\oatas$, $\oabas$ \cite{Degrassi:2009yq}
+can be added by setting \code{UseHiggs2LoopNMSSM = True;} in the model
+file.
 
 \subsection{Tree-level spectrum}
-The tree-level $\overline{DR}$ masses are calculated by diagonalizing
-the mass matrices returned from \code{SARAH`MassMatrix[]}.  The
-numerical singular value decomposition is performed by the Eigen
-library routine \code{Eigen::JacobiSVD} for matrices with less than
-four rows and columns, and the Lapack routines \code{zgesvd},
-\code{dgesvd} for larger matrices.  For the other types of
-diagonalization, \code{Eigen::SelfAdjointEigenSolver} from Eigen is
-used regardless of the matrix size.
-
-The mass matrix $M^2$ for real scalar fiels $\phi_i$ is diagonalized
-with an orthogonal matrix $O$ via
+The tree-level \DRbar masses are calculated by diagonalizing the mass
+matrices returned from \code{SARAH`MassMatrix[]}.  The numerical
+singular value decomposition is performed by the Eigen library routine
+\code{Eigen::JacobiSVD} for matrices with less than four rows and
+columns, and the Lapack routines \code{zgesvd}, \code{dgesvd} for
+larger matrices.  For the other types of diagonalization,
+\code{Eigen::SelfAdjointEigenSolver} from Eigen is used regardless of
+the matrix size.
+
+\fs uses the following conventions for the diagonalization: A mass
+matrix $M^2$ for real scalar fiels $\phi_i$ is diagonalized with an
+orthogonal matrix $O$ as
 %
@@
 \subsection{Two-scale fixed point iteration}
-The default two-scale RG solver uses above beta functions and boundary
-conditions to find the full set of $\overline{DR}$ masses in the model
-consistent with all constraints.  It does so by iterating between the
-scales of all boundary conditions, applying the constraints and
-checking for convergence.  This approach is decribed in
-\cite{Barger:1993gh} for the MSSM and is widely implemented in SUSY
-spectrum generators.
+
+The $\beta$-functions for the model parameters together with the
+boundary conditions for a boundary value problem.  The default
+two-scale RG solver tries to find the full set of model parameters
+consistent with all constraints at all scales.  It does so by running
+iteratively between the scales of all boundary conditions, imposing
+the constraints and checking for convergence after each iteration.
+This approach is decribed in \cite{Barger:1993gh} for the MSSM and is
+widely implemented in SUSY spectrum generators.
 
 The in more detail the two-scale algorithm used in \fs works as
-follows (see \figurename~\ref{fig:two-scale-algorithm}):
-%
-\\Initial guess:
+follows, see \figref{fig:two-scale-algorithm}:
+%
+\subparagraph{Initial guess:} The RG solver starts to guess all model
+parameters at the low-scale.
+%
 \begin{enumerate}
-\item Guess gauge couplings $g_{1,2,3}$ at $m_Z$
-\item Apply user-defined low-scale constraint (\code{LowScaleInput})
-\item Guess Yukawa couplings $y_{u,d,e}$ at $m_Z$
-\item Run model to the high-scale (\code{HighScaleFirstGuess})
-\item Apply high-scale constraint (\code{HighScaleInput})
-\item Run model to the low-scale (\code{LowScaleFirstGuess})
-\item Solve EWSB eqs.\ at the tree-level
-\item Calculate \DRbar\ masses
+\item At the $M_Z$ scale the susy model gauge couplings $g_{1,2,3}$
+  are set to the known Standard Model values (ignoring threshold
+  corrections).
+\item Then, the user-defined low-scale boundary conditions (defined in
+  \code{LowScaleInput}) are imposed.
+\item The Yukawa couplings $y_{u,d,e}$ of the susy model are set to
+  the known Standard Model Yukawa couplings.
+\item The susy parameters are run to the user-supplied first guess of
+  the high-scale (\code{HighScaleFirstGuess}).
+\item The high-scale boundary condition is imposed (defined in
+  \code{HighScaleInput}).
+\item All model parameters are run to the first guess of the low-scale
+  (\code{LowScaleFirstGuess})
+\item The EWSB eqs.\ are solved at the tree-level.
+\item The \DRbar\ mass spectrum is calculated.
 \end{enumerate}
 %
-Solving the renormalization group equations:
+At this point all model parameters are set to some initial values and
+a first estimation of the \DRbar\ mass spectrum is known.  Now the
+actual iteration starts
+%
+\subparagraph{Fixed-point iteration:}
+%
 \begin{enumerate}
-\item \label{rge-step-one} Run model to the low-scale (\code{LowScale})
+\item \label{rge-step-one} All model parameters are run to the
+  low-scale (\code{LowScale})
   \begin{enumerate}
-  \item Calculate \DRbar\ masses
-  \item Recalculate low-scale
-  \item Calculate \DRbar\ gauge and Yukawa couplings $g_{1,2,3}$, $y_{u,d,e}$
-  \item Apply user-defined low-scale constraint (\code{LowScaleInput})
+  \item The \DRbar\ mass spectrum is calculated.
+  \item The low-scale is recalculated.
+  \item The \DRbar\ gauge and Yukawa couplings $g_{1,2,3}(M_Z)$,
+    $y_{u,d,e}(M_Z)$ of the susy model are calculated using threshold
+    corrections.
+  \item The user-defined low-scale constraint is imposed
+    (\code{LowScaleInput})
   \end{enumerate}
-\item Run model to the high-scale (\code{HighScale})
+\item Run all model parameters to the high-scale (\code{HighScale})
   \begin{enumerate}
-  \item Recalculate high-scale
-  \item Apply high-scale constraint (\code{HighScaleInput})
+  \item Recalculate the high-scale
+  \item Apply the high-scale constraint (\code{HighScaleInput})
   \end{enumerate}
-\item Run model to the susy-scale (\code{SUSYScale})
+\item Run model parameters to the susy-scale (\code{SUSYScale})
   \begin{enumerate}
-  \item Calculate \DRbar\ masses
-  \item Recalculate low-scale
-  \item Apply susy-scale constraint (\code{SUSYScaleInput})
-  \item Solve EWSB eqs.\ at the one-loop level
+  \item Calculate the \DRbar\ mass spectrum
+  \item Recalculate the susy-scale
+  \item Apply the susy-scale constraint (\code{SUSYScaleInput})
+  \item Solve the EWSB eqs.\ at the loop level
   \end{enumerate}
@@
 %
-Calculating the particle spectrum:
-\begin{enumerate}
-\item Run model to the susy-scale
-\item Calculate the pole masses
-\item Run model to the parameter output scale (SLHA block MODSEL 12)
-\end{enumerate}
+If the fixed-point iteration has converged, all \DRbar\ model
+parameters are known at all scales between \code{LowScale} and
+\code{HighScale}.  In this case all model parameters are run to the
+susy-scale and the pole-mass spectrum is calculated.  If the user has
+chosen a specific output scale for the running \DRbar\ model
+parameters by setting entry $12$ in block \code{MODSEL} in the SLHA
+input file, all model parameters are run to the defined output scale.
 %
@@
 
-\begin{figure}[tb]
-  \centering
-  \tikzumlset{fill class=white}
-  \begin{tikzpicture}
-    \umlclass[x=0, y=0, type=abstract]{Beta\_function}{
-      -- scale : double\\
-      -- loops : unsigned\\
-      -- numPars : unsigned
-    }{
-      + \umlvirt{get() : const Eigen::ArrayXd} \\
-      + \umlvirt{set(v : const Eigen::ArrayXd\&) : void}\\
-      + \umlvirt{beta() : Eigen::ArrayXd}\\
-      + run\_to(scale : double, eps : double) : void\\
-    }
-    \umlclass[x=8, y=-12, template={T}]{MSSM}{}{}
-    \umlclass[x=8, y=-8, type=abstract]{Two\_scale\_model}{}{
-      + \umlvirt{calculate\_spectrum() : void}\\
-      + \umlvirt{run\_to(scale : double, eps : double) : void}\\
-      + \umlvirt{set\_precision(precision : double) : void}
-    }
-    \umlclass[x=0, y=-4]{MSSM\_susy\_parameters}{}{
-      + get() : const Eigen::ArrayXd \\
-      + set(v : const Eigen::ArrayXd\&) : void\\
-      + beta() : Eigen::ArrayXd
-    }
-    \umlclass[x=0, y=-8]{MSSM\_soft\_parameters}{}{
-      + get() : const Eigen::ArrayXd \\
-      + set(v : const Eigen::ArrayXd\&) : void\\
-      + beta() : Eigen::ArrayXd
-    }
-    \umlclass[x=0, y=-12]{MSSM$<$Two\_scale$>$}{}{
-      + calculate\_spectrum() : void\\
-      + run\_to(scale : double, eps : double) : void\\
-      + set\_precision(precision : double) : void
-    }
-   \umlinherit{MSSM\_susy\_parameters}{Beta\_function}
-   \umlinherit{MSSM\_soft\_parameters}{MSSM\_susy\_parameters}
-   \umlinherit{MSSM$<$Two\_scale$>$}{MSSM\_soft\_parameters}
-   \umlinherit{MSSM$<$Two\_scale$>$}{Two\_scale\_model}
-   \umldep[arg1={$<<$bind$>>$}, mult1={T $\rightarrow$ Two\_scale}, pos1=0.5]{MSSM$<$Two\_scale$>$}{MSSM}
-  \end{tikzpicture}
-  \caption{Two-scale model class hierarchy}
-  \label{fig:two-scale-model-class-hierarchy}
-\end{figure}
-
-\begin{figure}[tb]
-  \centering
-  \tikzumlset{fill class=white}
-  \begin{tikzpicture}
-    \umlclass[x=0, y=0, type=abstract]{Two\_scale\_model}{}{
-      + \umlvirt{calculate\_spectrum() : void}\\
-      + \umlvirt{run\_to(scale : double, eps : double) : void}\\
-      + \umlvirt{set\_precision(precision : double) : void}
-    }
-    \umlclass[x=0, y=-4]{RGFlow$<$Two\_scale$>$}{}{
-      + solve() : void
-    }
-    \umlclass[x=8, y=0]{Constraint$<$Two\_scale$>$}{}{
-      + \umlvirt{apply() : void}\\
-      + \umlvirt{get\_scale() : double}\\
-      + \umlvirt{set\_model(model : Two\_scale\_model*) : void}
-    }
-    \umlclass[x=8, y=4, template={T}]{Constraint}{}{}
-    \umlclass[x=0, y=-7]{Initial\_guesser$<$Two\_scale$>$}{}{
-      + \umlvirt{guess() : void}
-    }
-    \umlclass[x=0, y=-11, template={T}]{Initial\_guesser}{}{}
-    \umlclass[x=8, y=-7]{Convergence\_tester$<$Two\_scale$>$}{}{
-      + \umlvirt{accuracy\_goal\_reached() : bool}\\
-      + \umlvirt{max\_iterations() : unsigned}
-    }
-    \umlclass[x=8, y=-11, template={T}]{Convergence\_tester}{}{}
-    \umlclass[x=8, y=-4, template={T}]{RGFlow}{}{}
-    \umldep[arg1={$<<$bind$>>$}, mult1={T $\rightarrow$ Two\_scale}, pos1=0.5]{RGFlow$<$Two\_scale$>$}{RGFlow}
-    \umldep[mult1=1..*]{RGFlow$<$Two\_scale$>$}{Two\_scale\_model}
-    \umldep[mult1=1..*]{RGFlow$<$Two\_scale$>$}{Constraint$<$Two\_scale$>$}
-    \umldep{RGFlow$<$Two\_scale$>$}{Initial\_guesser$<$Two\_scale$>$}
-    \umldep{RGFlow$<$Two\_scale$>$}{Convergence\_tester$<$Two\_scale$>$}
-    \umldep[mult1={$<<$bind$>>$}, arg1={T $\rightarrow$ Two\_scale}, pos1=0.5]{Initial\_guesser$<$Two\_scale$>$}{Initial\_guesser}
-    \umldep[mult1={$<<$bind$>>$}, arg1={T $\rightarrow$ Two\_scale}, pos1=0.5]{Convergence\_tester$<$Two\_scale$>$}{Convergence\_tester}
-    \umldep[arg1={$<<$bind$>>$}, mult1={T $\rightarrow$ Two\_scale}, pos1=0.5]{Constraint$<$Two\_scale$>$}{Constraint}
-  \end{tikzpicture}
-  \caption{Two-scale renormalization group solver class hierarchy}
-  \label{fig:two-scale-rgflow-class-hierarchy}
-\end{figure}
-{\color{red} copied and pasted from latest manual -- not sure how best to maintain this, could split maual up into tiny bits, but we may also want some differences?}
-
 \subsection{Pole masses}
@@
 
-\code{LowDiagonalizationPrecision}: This option provides the lowest
-precision but is also the fastest one.  Here the one-loop mass matrix
-$m_{f,1L}^\text{low}$ is calculated exactly once as
+\begin{itemize}
+\item \code{LowDiagonalizationPrecision}: This option provides the
+  lowest precision but is also the fastest one.  Here the one-loop
+  mass matrix $m_{f,1L}^\text{low}$ is calculated exactly once as
 %
   \begin{align}
-  \forall i,j: (m_{f,1L}^\text{low})_{ij} = (m_{f,1L}(p^2 = m_{f_i} m_{f_j}))_{ij} ,
+    \forall i,j: (m_{f,1L}^\text{low})_{ij} = (m_{f,1L}(p^2 = m_{f_i}
+    m_{f_j}))_{ij} ,
   \end{align}
@@
 
-\code{MediumDiagonalizationPrecision} (default): This option provides
-calculation with medium precision with a medium execution time.  Here
-the one-loop mass matrix $m_{f,1L}^\text{medium}$ is calculated $n$
-times as
+\item \code{MediumDiagonalizationPrecision} (default): This option
+  provides calculation with medium precision with a medium execution
+  time.  Here the one-loop mass matrix $m_{f,1L}^\text{medium}$ is
+  calculated $n$ times as
 %
   \begin{align}
-  (m_{f,1L}^\text{medium})_{ij}^{(k)} = (m_{f,1L}(p^2 = m_{f_k}^2))_{ij} ,
-  \qquad k = 1,\ldots,n ,
+    (m_{f,1L}^\text{medium})_{ij}^{(k)} = (m_{f,1L}(p^2 =
+    m_{f_k}^2))_{ij} , \qquad k = 1,\ldots,n ,
   \end{align}
@@
 
-\code{HighDiagonalizationPrecision}: This option provides
+\item \code{HighDiagonalizationPrecision}: This option provides
   diagonalization with highest precision, but has also the highest
-execution time.  Here the one-loop mass matrix $m_{f,1L}^\text{high}$
-is diagonalized $n$ times, as in the case of
+  execution time.  Here the one-loop mass matrix
+  $m_{f,1L}^\text{high}$ is diagonalized $n$ times, as in the case of
   \code{MediumDiagonalizationPrecision}, resulting in $n$ pole masses
@@
   $m_{f_k}^\pole$ for the momentum calculation $p^2 =
-(m_{f_k}^\pole)^2$.  The iteration stops until convergence is reached.
+  (m_{f_k}^\pole)^2$.  The iteration stops until convergence is
+  reached.
+\end{itemize}
 
 If higher accuracy is required additional routines with higher order
-corrections can be added in a very simple way. For example in the MSSM
-by default \fs adds two-loop Higgs FORTRAN routines supplied by
-P.~Slavich from \cite{slavich-degrassi} to add two loop corrections of
-$\oatas$, $\oabas$, $\oatq$, $\oabatau$, $\oabq$, $\oatauq$ and
-$\oatab$ as follows...
+corrections can be added by setting \code{UseHiggs2LoopMSSM = True;}
+in the model file. For example in the MSSM by default \fs adds
+two-loop Higgs FORTRAN routines supplied by P.~Slavich from
+\cite{Degrassi:2001yf,Brignole:2001jy,Dedes:2002dy,Brignole:2002bz,Dedes:2003km}
+to add two-loop corrections of $\oatas$, $\oabas$, $\oatq$,
+$\oabatau$, $\oabq$, $\oatauq$ and $\oatab$.
 
@@
 
-Corrections to other masses or further corrections to the Higgs states
-can be added very easily by...
-
-%%\section{Implementing new models}
 \section{Flexible Spectra}

### 2014-05-10 doc/paper.tex
moving the SLHA input file to the appendix

@@
 %
-FlexibleSUSY and Softsusy were compiled with g++ 4.8.0 and Intel ifort
-13.1.3 20130607.  SPheno and SPhenoMSSM were compiled with Intel ifort
+FlexibleSUSY and Softsusy are compiled with g++ 4.8.0 and Intel ifort
+13.1.3 20130607.  SPheno and SPhenoMSSM are compiled with Intel ifort
 13.1.3 20130607\footnote{We are using Intel's ifort compiler, instead
@@
   SPhenoMSSM by approximately a factor $1.5$.}.  We're generating
-$2\cot 10^{4}$ random CMSSM parameter points with $m_0\in [50,1000]$
+$2\cdot 10^{4}$ random CMSSM parameter points with $m_0\in [50,1000]$
 $m_{1/2}\in [50,1000]$, $\tan\beta\in [1,100]$, $\sign\mu\in
-\{-1,+1\}$ and $A_0\in [-1000,1000]$.  The following SLHA template
-file was used:
+\{-1,+1\}$ and $A_0\in [-1000,1000]$.  For each point an SLHA input
+file is created by appending the values of $m_0$, $m_{1/2}$,
+$\tan\beta$, $\sign\mu$, $A_0$ in form of a \code{MINPAR} block to the
+SLHA template file given in \ref{sec:speed-test-slha-template-file}.
+The resulting SLHA input file is passed to each spectrum generator and
+the (wall-clock) time is measured until the program has finished.  The
+average run-times for three different CPU types can be found in
+\tablename~\ref{tab:run-time-comparison}.  The first column shows the
+run-time on a Intel Core2 Duo (P8600) where only one core was enabled.
+The second column shows the run-time on a Intel Core2 Duo (P8600)
+where both cores were enabled.  In the third column a Intel Xeon
+(L5640) was used, which has $6$ CPU cores.
+%
+\begin{table}[tbh]
+  \centering
+  \begin{tabular}{llll}
+    \toprule
+                          & Intel Core2 Duo (P8600)    & Intel Core2 Duo (P8600)     & Intel Xeon (L5640)\\
+                          & (1 core, $2.40\unit{GHz}$) & (2 cores, $2.40\unit{GHz}$) & (6 cores, $2.27\unit{GHz}$)\\
+    \midrule
+    FlexibleSUSY-NoFV 1.0.0 & $0.086\unit{s}$    & $0.079\unit{s}$   & $0.060\unit{s}$\\
+    SPheno 3.2.4            & $0.119\unit{s}$    & $0.114\unit{s}$   & $0.101\unit{s}$\\
+    Softsusy 3.4.0          & $0.175\unit{s}$    & $0.171\unit{s}$   & $0.147\unit{s}$\\
+    \midrule
+    FlexibleSUSY-FV 1.0.0   & $0.150\unit{s}$    & $0.113\unit{s}$   & $0.074\unit{s}$\\
+    SPhenoMSSM 4.1.0${}^*$  & $0.415\unit{s}$    & $0.401\unit{s}$   & $0.370\unit{s}$\\
+    \bottomrule
+  \end{tabular}
+  \caption{Average run-time of CMSSM spectrum generators for
+    for random parameter points.  The upper }
+  \label{tab:run-time-comparison}
+\end{table}
+
+Under both the non-flavor violating spectrum generators
+FlexibleSUSY-NoFV, Softsusy and SPheno (first three rows) as well as
+the flavor violating ones FlexibleSUSY-NoFV and SPhenoMSSM (4th and
+5th row) we find that \fs is significantly fastest.  We also find that
+increasing the number of CPU cores reduces the run-time of \fs.  The
+reason is that \fs calculates each pole mass in a separate thread, and
+therefore benefits from multi-core CPUs.  Under the hand-written CMSSM
+spectrum generators SPheno and Softsusy we find that SPheno is fastest
+on average.
+
+\appendix
+\section{Examples}
+\subsection{NMSSM}
+Detailed steps to create NMSSM model file.
+\begin{enumerate}
+\item Create new directory \code{mkdir myNMSSM}
+\item Move into this new directory \code{cd myNMSSM}
+\item copy the MSSM model file in \code{cp /path/to/SARAH/Models/MSSM/MSSM.m  myNMSSM.m}
+\item Edit myNMSSM.m as follows
+\begin{enumerate}
+\item  Change Model name, author and date
+\item Add singlet superfield to SuperFields
+\begin{lstlisting}
+SuperFields[[8]] = {s, 1, sR,    0, 1,  1, RpM};
+\end{lstlisting}
+\item Modify superpotential
+\begin{lstlisting}
+SuperPotential = Yu u.q.Hu - Yd d.q.Hd - Ye e.l.Hd + \[Lambda] s.Hu.Hd +  \[Kappa]/3 s.s.s ;
+\end{lstlisting}
+\item Add new VEV
+\begin{lstlisting} 
+  DEFINITION[EWSB][VEVs]= 
+  { {SHd0, {vd, 1/Sqrt[2]}, {sigmad, \[ImaginaryI]/Sqrt[2]},{phid,1/Sqrt[2]}},
+    {SHu0, {vu, 1/Sqrt[2]}, {sigmau, \[ImaginaryI]/Sqrt[2]},{phiu,1/Sqrt[2]}},
+    {SsR,  {vs, 1/Sqrt[2]}, {sigmaS, \[ImaginaryI]/Sqrt[2]},{phiS,1/Sqrt[2]}}
+};
+\end{lstlisting}
+\item  Add the singlino to the list of gauge eigenstate dirac fermions
+\begin{lstlisting}
+  DEFINITION[GaugeES][DiracSpinors]={
+ .............
+  FS -> {FsR,conj[FsR]}   
+};
+\end{lstlisting}
+\item Edit mass eigenstates to give how the gauge eigenstates mix into mass eigenstates.
+Here we add to the cp even higgs a real singlet, to the CP odd the imaginary part of the singlet and finally to the neutralinos we add the singlino.
+
+\begin{lstlisting}
+ DEFINITION[EWSB][MatterSector]= 
+{   ................
+    ................
+     {{phid, phiu, phiS}, {hh, ZH}},      
+     {{sigmad, sigmau, sigmaS}, {Ah, ZA}},
+    ................
+     {{fB, fW0, FHd0, FHu0, FsR}, {L0, ZN}},
+     ...............
+};
+
+\end{lstlisting}
+\end{enumerate}
+\item Now we copy over particles.m \code{cp /path/to/SARAH/Models/MSSM/particles.m  ./}
+\item Edit particles.m as follows
+\begin{enumerate}  
+\item First we add to particleDefinitions[GaugeES]:
+\begin{lstlisting}
+ ParticleDefinitions[GaugeES] = {
+      {SsR, { Description -> "Singlet"}},        
+      {FS,   { Description -> "Singlino" }},    
+      ............
+      };
+\end{lstlisting}
+Note that we are brief in information here becasue there is already some information contained in the default particles.m file in the model directory:
+\begin{lstlisting} 
+{{ Description -> "Singlet", 
+                 PDG -> {0},
+                 PDG.IX -> 0,
+                 Width -> 0, 
+                 Mass -> Automatic,
+                 FeynArtsNr -> 98,
+                 LaTeX -> "S",
+                 OutputName -> "s"
+
+    }},
+
+(* -------- Weyl Spinor  ------ *)
+
+  {{ Description -> "Weyl Spinor of Singlino", 
+                 PDG -> 99,
+                 Width -> 0, 
+                 Mass -> Automatic,
+                 FeynArtsNr -> 9,
+                 LaTeX -> "\\tilde{S}",
+                 OutputName -> "s"
+
+    }},
+
+\end{lstlisting}
+ \item Now we edit ParticleDefinitions[EWSB].
+Again most states are defined in brief form by a name because they use the default information in the particles.m in the models directory.
+But here we need to over write some of the default options for the cp even and cp odd higgs states:
+\begin{lstlisting}
+ {hh   ,  {  Description -> "Higgs", 
+                 PDG -> {25, 35,45},
+                 PDG.IX ->{101000001,101000002,101000003}
+ }}, 
+       {Ah   ,  {    Description -> "Pseudo-Scalar Higgs",
+                 PDG -> {0, 36, 46},
+                 PDG.IX ->{0,102000001,102000002  } }},
+
+\end{lstlisting}
+\item To the WeylFermionAndIndermediate we add the Weyl singlino and the real scalar singlet field and imaginary scalar singlet field:  
+ \begin{lstlisting}
+  {sigmaS,      {Description -> "Scalar Singlet" }}  ,
+  {phiS,      { Description -> "Pseudo Scalar Singlet"}},
+  {FsR,   { Description -> "Weyl Spinor of Singlino"}},
+ \end{lstlisting}
+
+Although the WeylFermionAndIndermediate also contains, 
+\begin{lstlisting}
+{SHd,  { Description -> "Down-Higgs"}},
+       {SHu,  { Description -> "Up-Higgs"}},
+\end{lstlisting}
+\noindent we do not add a singlet version becasue for the singlet there is no distinction between these and the individual SU(2) components, as there are for Hu and Hd.
+\end{enumerate} 
+\item Now we copy over parameters.m \code{cp /path/to/SARAH/Models/MSSM/parameters.m  ./}
+\item Add the new parameters lambda, kappa, Alambda and Akappa to the list.
+\begin{lstlisting}
+{\[Kappa],   {Description -> "Singlet Self-Interaction"}},              
+{T[\[Kappa]],  { Description -> "Softbreaking Singlet Self-Interaction" }}, 
+{\[Lambda],   { Description -> "Singlet-Higgs-Interaction"   }},
+{T[\[Lambda]],  {Description -> "Softbreaking Singlet-Higgs-Interaction"}}, 
+{ms2,       { Description -> "Softbreaking Singlet Mass" }},
+{vS,        { Description -> "Singlet-VEV"}}       
+\end{lstlisting}
+Once again this is simplified from the general case becasue there are default versions containedin parameters.m in the Models directory:
+\begin{lstlisting}
+{{  Description -> "Singlet Self-Interaction",
+              LaTeX -> "\\kappa",
+             Real ->False,
+             Dependence -> None, 
+             Value -> None, 
+             LesHouches -> {NMSSMRUN,2},
+             OutputName-> kap         }},
+
+{{ Description -> "Softbreaking Singlet Self-Interaction",
+               LaTeX -> "T_{\\kappa}",
+              Real -> False,
+             Dependence -> None, 
+             Value -> None, 
+             LesHouches ->{NMSSMRUN,4},
+             OutputName-> Tk         }},
+
+ {{ Description -> "Singlet-Higgs-Interaction",
+             LaTeX -> "\\lambda",
+             Real -> False,
+	     Dependence -> None, 
+             Value -> None, 
+             LesHouches -> {NMSSMRUN,1},
+             OutputName-> lam          }},
+
+{{Description -> "Softbreaking Singlet-Higgs-Interaction",
+                LaTeX -> "T_{\\lambda}",
+              Real -> False,
+             Dependence -> None, 
+             Value -> None, 
+             LesHouches ->{NMSSMRUN,3},
+             OutputName-> Tlam         }},    
+{{ Description -> "Softbreaking Singlet Mass", 
+              LaTeX -> "m_S^2",
+             DependenceNum ->  None, 
+             LesHouches -> {NMSSMRUN,10},
+             OutputName-> ms2 }},
+
+{{ Description -> "Singlet-VEV", 
+			 LaTeX -> "v_s",
+             Value -> None, 
+             LesHouches -> {NMSSMRUN,5},
+             OutputName-> vS         }},
+\end{lstlisting}
+\item Navigate to Flexible SUSY directory  \code{cd /path/to/FlexibleSUSY/}
+\item Create Z3NMSSM directory \code{./createmodel --models=Z3NMSSM:myNMSSM}
+\item Configure with the new model \code{./configure --with-models=Z3NMSSM}
+\item The default FlexibleSUSY model file for the MSSM will now be in the models/NMSSM2 directory.  Edit this for the NMSSM as follows:
+  \begin{enumerate} 
+    \item Add new input parameter $\lambda$ to EXTPAR and remove the sign of $\mu$ from MINPAR.
+      \begin{lstlisting}    
+        MINPAR = { {1, m0},
+           {2, m12},
+           {3, TanBeta},
+           {5, Azero} };
+
+  EXTPAR = { {61, LambdaInput} };
+      \end{lstlisting}
+     \item Also add this parameter to the DefaultParameterPoint.
+      \begin{lstlisting}
+       DefaultParameterPoint = {
+    {m0, 200},
+    {m12, 200},
+    {TanBeta, 10},
+    {Azero, -500}
+    {LambdaInput, 0.1}
+    };
+  \end{lstlisting}
+    \item Name the parameters you wish to be output by the EWSB conditions.  Note you should use the names given for them in parameters.m.    A common choice in the NMSSM is:
+\begin{lstlisting}
+EWSBOutputParameters = {vS, \[Kappa], ms2};
+\end{lstlisting}
+which will work for the semi-constrained NMSSM.
+\item choose the Highscale and Highscale first guess, if the default is different from what you wish to use.  
+\begin{lstlisting} 
+ HighScale = g1 == g2;
+ HighScaleFirstGuess = 2.0 10^16;
+\end{lstlisting}
+\item Add how the parameters found in parameters.m that are to be set at the highscale:
+\begin{lstlisting} 
+HighScaleInput={
+   {T[Ye], Azero*Ye},
+   {T[Yd], Azero*Yd},
+   {T[Yu], Azero*Yu},
+   {mHd2, m0^2},
+   {mHu2, m0^2},
+   {mq2, UNITMATRIX[3] m0^2},
+   {ml2, UNITMATRIX[3] m0^2},
+   {md2, UNITMATRIX[3] m0^2},
+   {mu2, UNITMATRIX[3] m0^2},
+   {me2, UNITMATRIX[3] m0^2},
+   {MassB, m12},
+   {MassWB,m12},
+   {MassG,m12},
+{\[Lambda], LambdaInput},
+   {T[\[Kappa]], Azero \[Kappa]},
+   {T[\[Lambda]], Azero LambdaInput}
+};
+\end{lstlisting}
+\item set lowscale boundary condititions if you want something different from the default.  The defaults are:
+\begin{lstlisting} 
+ LowScale = SM[MZ];
+ LowScaleFirstGuess = SM[MZ];
+\end{lstlisting}
+\item Change the LowscaleInput if necessary.  In this example we keep the same choice as in the MSSM
+  \begin{lstlisting} 
+ LowScaleInput={
+   {vd, 2 MZDRbar / Sqrt[GUTNormalization[g1]^2 g1^2 + g2^2] Cos[ArcTan[TanBeta]]},
+   {vu, 2 MZDRbar / Sqrt[GUTNormalization[g1]^2 g1^2 + g2^2] Sin[ArcTan[TanBeta]]}
+};
+  \end{lstlisting}
+
+Finally we set the initial guesses for the paremeters.  In this example we overwtite the MSSM version replacing it with:
+ \begin{lstlisting} 
+InitialGuessAtLowScale = {
+   {vd, SM[vev] Cos[ArcTan[TanBeta]]},
+   {vu, SM[vev] Sin[ArcTan[TanBeta]]},
+   {\[Lambda], LambdaInput},
+   {\[Kappa], 0.1},
+   {vS, 1000},
+   {ms2, SM[MZ]^2}
+};
+
+InitialGuessAtHighScale = {};
+ \end{lstlisting}
+
+The code can now be created and also compiled (if you didn't select --disable-compile option when you ran the configure script) by typing \code{make -j2}.
+
+ \end{enumerate}
+
+\end{enumerate}   
+\subsection{\ESSM}
+Detailed steps to create \ESSM model file.
+\begin{enumerate}
+\item Create new directory \code{mkdir myE6SSM}
+\item Move into this new directory \code{cd myE6SSM}
+\item copy the MSSM model file in \code{cp /path/to/SARAH/Models/MSSM/MSSM.m  myE6SSM.m}
+\item Edit myE6SSM.m as follows
+\begin{enumerate}
+\item  Change Model name, author and date
+\item Add new Vector superfield,
+ \begin{lstlisting} 
+   Gauge[[4]]={Bp,  U[1], Ncharge, g1p, False, RpM};
+ \end{lstlisting}
+\item Modify MSSM superfield definitions to include new charges.
+\begin{lstlisting}
+SuperFields[[1]] = {q, 3, {uL,  dL},     1/6, 2, 3, 1, RpM};  
+SuperFields[[2]] = {l, 3, {vL,  eL},    -1/2, 2, 1, 2, RpM};
+SuperFields[[3]] = {Hd,1, {Hd0, Hdm},   -1/2, 2, 1, -3, RpP};
+SuperFields[[4]] = {Hu,1, {Hup, Hu0},    1/2, 2, 1, -2, RpP};
+SuperFields[[5]] = {d, 3, conj[dR],    1/3, 1, -3, 2, RpM};
+SuperFields[[6]] = {u, 3, conj[uR],   -2/3, 1, -3, 1, RpM};
+SuperFields[[7]] = {e, 3, conj[eR],      1, 1,  1, 1, RpM};
+\end{lstlisting}
+Note that in the \ESSM model the charges are fixed to be those of the $U(1)_N$ as shown in e.g.
+\cite{King:2005jy}.
+However here one can leave the charges unassigned vby using parameters in place of the fixed charges.
+Here we fix the charges to simplify the C++ code.
+
+\item Add Singlet Higgs superfield to SuperField
+\begin{lstlisting}
+SuperFields[[8]] = {s, 1, sR,     0, 1,  1, 5, RpP};
+\end{lstlisting}
+\item Add inert Higgs-like superfields 
+\begin{lstlisting}
+SuperFields[[9]] = {H1I, 2, {H1I0, H1Im},  -1/2, 2, 1, -3, RpP};
+SuperFields[[10]] = {H2I, 2, {H2Ip, H2I0},   1/2, 2, 1, -2, RpP};
+SuperFields[[11]] = {sI, 2, sIR,    0, 1,  1, 5, RpM};
+\end{lstlisting}
+\item Add exotics to SuperFields
+\begin{lstlisting}
+SuperFields[[12]] = {Dx, 3, DxL,  -1/3, 1, 3, -2, RpP};
+SuperFields[[13]] = {Dxbar, 3, conj[DxbarR],  1/3, 1, -3, -3, RpP};
+\end{lstlisting}
+\item Add additional survival $SU(2)$ multiplets to SuperFields
+\begin{lstlisting}
+SuperFields[[14]] = {Hp, 1, {Hpd0, Hpdm},  1/2, 2,  1, 2, RpP};
+SuperFields[[15]] = {Hpbar, 1, {Hpup, Hpu0}, -1/2, 2,  1, -2, RpP};
+\end{lstlisting}
+\item Modify superpotential
+\begin{lstlisting}
+SuperPotential = Yu u.q.Hu - Yd d.q.Hd - Ye e.l.Hd + \[Lambda] s.Hu.Hd +  \[Lambda]12 s.H2I.H1I + \[Kappa] s.Dx.Dxbar + \[Mu]Pr Hp.Hpbar;
+\end{lstlisting}
+\item Define new gauge eigenstate Dirac spinors
+\begin{lstlisting}
+  DEFINITION[GaugeES][DiracSpinors]={
+    .............
+    .............
+    FS -> {FsR,conj[FsR]},
+    FBp -> {fBp,conj[fBp]},
+    HI0 -> {FH1I0, conj[FH2I0]},
+    HIC -> {FH1Im, conj[FH2Ip]},
+    FSI -> {FsIR, conj[FsIR]},
+    FDx1 -> {FDxL, 0},
+    FDx2 -> {0, FDxbarR}, 
+    Hp0 -> {FHpd0, conj[FHpu0]},
+    HpC -> {FHpdm, conj[FHpup]}
+  };
+\end{lstlisting}
+\item Add new VEV
+\begin{lstlisting} 
+  DEFINITION[EWSB][VEVs]= 
+  { {SHd0, {vd, 1/Sqrt[2]}, {sigmad, \[ImaginaryI]/Sqrt[2]},{phid,1/Sqrt[2]}},
+    {SHu0, {vu, 1/Sqrt[2]}, {sigmau, \[ImaginaryI]/Sqrt[2]},{phiu,1/Sqrt[2]}},
+    {SsR,  {vs, 1/Sqrt[2]}, {sigmaS, \[ImaginaryI]/Sqrt[2]},{phiS,1/Sqrt[2]}}
+};
+\end{lstlisting}
+\item Add new gauge boson  mass eigenstate
+\begin{lstlisting} 
+  DEFINITION[EWSB][GaugeSector] =
+  { 
+    {{VB,VWB[3],VBp},{VP,VZ,VZp},ZZ},
+    {{VWB[1],VWB[2]},{VWm,conj[VWm]},ZW},
+    {{fWB[1],fWB[2],fWB[3]},{fWm,fWp,fW0},ZfW}
+  }; 
+\end{lstlisting}
+\item Edit mixings for Higgs and neutralino states and add additional states that mix.
+\begin{lstlisting}
+DEFINITION[EWSB][MatterSector]= 
+{ ...............
+  ...............
+{{phid, phiu, phiS}, {hh, ZH}},
+{{sigmad, sigmau, sigmaS}, {Ah, ZA}},
+  ...............
+{{fB, fW0, FHd0, FHu0, FsR, fBp}, {L0, ZN}},
+ ................ 
+ {{SH1I0,conj[SH2I0]},{SHI0,UHI0}},
+ {{SH1Im,conj[SH2Ip]},{SHIp,UHIp}},
+ {{SHpd0,conj[SHpu0]},{SHp0,UHp0}},
+ {{SHpdm,conj[SHpup]},{SHpp,UHpp}},
+ {{FHpd0,FHpu0},{L0p,ZNp}},
+ {{FH1I0,FH2I0},{L0I,ZNI}}
+\end{lstlisting}
+\end{enumerate}
+
+\end{enumerate}   
+\section{Tests}
+Comparisons against existing spectrum generators to test for deviations, e.g. the tests against MSSM and NMSSM (and E6SSM?).
+
+Comparisons against existing spectrum generators and SARAH generated spheno code for speed.
+
+\section{Speed test SLHA input file}
+\label{sec:speed-test-slha-template-file}
 %
@@
 \end{verbatim}
-%
-The SLHA file was passed to each spectrum generator and the
-(wall-clock) time was measured until the program has finished.  The
-average run-times for three different CPU types can be found in
-\tablename~\ref{tab:run-time-comparison}.  The first column shows the
-run-time on a Intel Core2 Duo (P8600) where only one core was enabled.
-The second column shows the run-time on a Intel Core2 Duo (P8600)
-where both cores were enabled.  In the third column a Intel Xeon
-(L5640) was used, which has $6$ CPU cores.
-%
-\begin{table}[tbh]
-  \centering
-  \begin{tabular}{llll}
-    \toprule
-                          & Intel Core2 Duo (P8600)    & Intel Core2 Duo (P8600)     & Intel Xeon (L5640)\\
-                          & (1 core, $2.40\unit{GHz}$) & (2 cores, $2.40\unit{GHz}$) & (6 cores, $2.27\unit{GHz}$)\\
-    \midrule
-    FlexibleSUSY-NoFV 1.0.0 & $0.086\unit{s}$    & $0.079\unit{s}$   & $0.060\unit{s}$\\
-    SPheno 3.2.4            & $0.119\unit{s}$    & $0.114\unit{s}$   & $0.101\unit{s}$\\
-    Softsusy 3.4.0          & $0.175\unit{s}$    & $0.171\unit{s}$   & $0.147\unit{s}$\\
-    \midrule
-    FlexibleSUSY-FV 1.0.0   & $0.150\unit{s}$    & $0.113\unit{s}$   & $0.074\unit{s}$\\
-    SPhenoMSSM 4.1.0${}^*$  & $0.415\unit{s}$    & $0.401\unit{s}$   & $0.370\unit{s}$\\
-    \bottomrule
-  \end{tabular}
-  \caption{Average run-time of CMSSM spectrum generators for
-    for random parameter points.  The upper }
-  \label{tab:run-time-comparison}
-\end{table}
-
-Under both the non-flavor violating spectrum generators
-FlexibleSUSY-NoFV, Softsusy and SPheno (first three rows) as well as
-the flavor violating ones FlexibleSUSY-NoFV and SPhenoMSSM (4th and
-5th row) we find that \fs is significantly fastest.  We also find that
-increasing the number of CPU cores reduces the run-time of \fs.  The
-reason is that \fs calculates each pole mass in a separate thread, and
-therefore benefits from multi-core CPUs.  Under the hand-written CMSSM
-spectrum generators SPheno and Softsusy we find that SPheno is fastest
-on average.
-
-\appendix
-\section{Examples}
-\subsection{NMSSM}
-Detailed steps to create NMSSM model file.
-\begin{enumerate}
-\item Create new directory \code{mkdir myNMSSM}
-\item Move into this new directory \code{cd myNMSSM}
-\item copy the MSSM model file in \code{cp /path/to/SARAH/Models/MSSM/MSSM.m  myNMSSM.m}
-\item Edit myNMSSM.m as follows
-\begin{enumerate}
-\item  Change Model name, author and date
-\item Add singlet superfield to SuperFields
-\begin{lstlisting}
-SuperFields[[8]] = {s, 1, sR,    0, 1,  1, RpM};
-\end{lstlisting}
-\item Modify superpotential
-\begin{lstlisting}
-SuperPotential = Yu u.q.Hu - Yd d.q.Hd - Ye e.l.Hd + \[Lambda] s.Hu.Hd +  \[Kappa]/3 s.s.s ;
-\end{lstlisting}
-\item Add new VEV
-\begin{lstlisting} 
-  DEFINITION[EWSB][VEVs]= 
-  { {SHd0, {vd, 1/Sqrt[2]}, {sigmad, \[ImaginaryI]/Sqrt[2]},{phid,1/Sqrt[2]}},
-    {SHu0, {vu, 1/Sqrt[2]}, {sigmau, \[ImaginaryI]/Sqrt[2]},{phiu,1/Sqrt[2]}},
-    {SsR,  {vs, 1/Sqrt[2]}, {sigmaS, \[ImaginaryI]/Sqrt[2]},{phiS,1/Sqrt[2]}}
-};
-\end{lstlisting}
-\item  Add the singlino to the list of gauge eigenstate dirac fermions
-\begin{lstlisting}
-  DEFINITION[GaugeES][DiracSpinors]={
- .............
-  FS -> {FsR,conj[FsR]}   
-};
-\end{lstlisting}
-\item Edit mass eigenstates to give how the gauge eigenstates mix into mass eigenstates.
-Here we add to the cp even higgs a real singlet, to the CP odd the imaginary part of the singlet and finally to the neutralinos we add the singlino.
-
-\begin{lstlisting}
- DEFINITION[EWSB][MatterSector]= 
-{   ................
-    ................
-     {{phid, phiu, phiS}, {hh, ZH}},      
-     {{sigmad, sigmau, sigmaS}, {Ah, ZA}},
-    ................
-     {{fB, fW0, FHd0, FHu0, FsR}, {L0, ZN}},
-     ...............
-};
-
-\end{lstlisting}
-\end{enumerate}
-\item Now we copy over particles.m \code{cp /path/to/SARAH/Models/MSSM/particles.m  ./}
-\item Edit particles.m as follows
-\begin{enumerate}  
-\item First we add to particleDefinitions[GaugeES]:
-\begin{lstlisting}
- ParticleDefinitions[GaugeES] = {
-      {SsR, { Description -> "Singlet"}},        
-      {FS,   { Description -> "Singlino" }},    
-      ............
-      };
-\end{lstlisting}
-Note that we are brief in information here becasue there is already some information contained in the default particles.m file in the model directory:
-\begin{lstlisting} 
-{{ Description -> "Singlet", 
-                 PDG -> {0},
-                 PDG.IX -> 0,
-                 Width -> 0, 
-                 Mass -> Automatic,
-                 FeynArtsNr -> 98,
-                 LaTeX -> "S",
-                 OutputName -> "s"
-
-    }},
-
-(* -------- Weyl Spinor  ------ *)
-
-  {{ Description -> "Weyl Spinor of Singlino", 
-                 PDG -> 99,
-                 Width -> 0, 
-                 Mass -> Automatic,
-                 FeynArtsNr -> 9,
-                 LaTeX -> "\\tilde{S}",
-                 OutputName -> "s"
-
-    }},
-
-\end{lstlisting}
- \item Now we edit ParticleDefinitions[EWSB].
-Again most states are defined in brief form by a name because they use the default information in the particles.m in the models directory.
-But here we need to over write some of the default options for the cp even and cp odd higgs states:
-\begin{lstlisting}
- {hh   ,  {  Description -> "Higgs", 
-                 PDG -> {25, 35,45},
-                 PDG.IX ->{101000001,101000002,101000003}
- }}, 
-       {Ah   ,  {    Description -> "Pseudo-Scalar Higgs",
-                 PDG -> {0, 36, 46},
-                 PDG.IX ->{0,102000001,102000002  } }},
-
-\end{lstlisting}
-\item To the WeylFermionAndIndermediate we add the Weyl singlino and the real scalar singlet field and imaginary scalar singlet field:  
- \begin{lstlisting}
-  {sigmaS,      {Description -> "Scalar Singlet" }}  ,
-  {phiS,      { Description -> "Pseudo Scalar Singlet"}},
-  {FsR,   { Description -> "Weyl Spinor of Singlino"}},
- \end{lstlisting}
-
-Although the WeylFermionAndIndermediate also contains, 
-\begin{lstlisting}
-{SHd,  { Description -> "Down-Higgs"}},
-       {SHu,  { Description -> "Up-Higgs"}},
-\end{lstlisting}
-\noindent we do not add a singlet version becasue for the singlet there is no distinction between these and the individual SU(2) components, as there are for Hu and Hd.
-\end{enumerate} 
-\item Now we copy over parameters.m \code{cp /path/to/SARAH/Models/MSSM/parameters.m  ./}
-\item Add the new parameters lambda, kappa, Alambda and Akappa to the list.
-\begin{lstlisting}
-{\[Kappa],   {Description -> "Singlet Self-Interaction"}},              
-{T[\[Kappa]],  { Description -> "Softbreaking Singlet Self-Interaction" }}, 
-{\[Lambda],   { Description -> "Singlet-Higgs-Interaction"   }},
-{T[\[Lambda]],  {Description -> "Softbreaking Singlet-Higgs-Interaction"}}, 
-{ms2,       { Description -> "Softbreaking Singlet Mass" }},
-{vS,        { Description -> "Singlet-VEV"}}       
-\end{lstlisting}
-Once again this is simplified from the general case becasue there are default versions containedin parameters.m in the Models directory:
-\begin{lstlisting}
-{{  Description -> "Singlet Self-Interaction",
-              LaTeX -> "\\kappa",
-             Real ->False,
-             Dependence -> None, 
-             Value -> None, 
-             LesHouches -> {NMSSMRUN,2},
-             OutputName-> kap         }},
-
-{{ Description -> "Softbreaking Singlet Self-Interaction",
-               LaTeX -> "T_{\\kappa}",
-              Real -> False,
-             Dependence -> None, 
-             Value -> None, 
-             LesHouches ->{NMSSMRUN,4},
-             OutputName-> Tk         }},
-
- {{ Description -> "Singlet-Higgs-Interaction",
-             LaTeX -> "\\lambda",
-             Real -> False,
-	     Dependence -> None, 
-             Value -> None, 
-             LesHouches -> {NMSSMRUN,1},
-             OutputName-> lam          }},
-
-{{Description -> "Softbreaking Singlet-Higgs-Interaction",
-                LaTeX -> "T_{\\lambda}",
-              Real -> False,
-             Dependence -> None, 
-             Value -> None, 
-             LesHouches ->{NMSSMRUN,3},
-             OutputName-> Tlam         }},    
-{{ Description -> "Softbreaking Singlet Mass", 
-              LaTeX -> "m_S^2",
-             DependenceNum ->  None, 
-             LesHouches -> {NMSSMRUN,10},
-             OutputName-> ms2 }},
-
-{{ Description -> "Singlet-VEV", 
-			 LaTeX -> "v_s",
-             Value -> None, 
-             LesHouches -> {NMSSMRUN,5},
-             OutputName-> vS         }},
-\end{lstlisting}
-\item Navigate to Flexible SUSY directory  \code{cd /path/to/FlexibleSUSY/}
-\item Create Z3NMSSM directory \code{./createmodel --models=Z3NMSSM:myNMSSM}
-\item Configure with the new model \code{./configure --with-models=Z3NMSSM}
-\item The default FlexibleSUSY model file for the MSSM will now be in the models/NMSSM2 directory.  Edit this for the NMSSM as follows:
-  \begin{enumerate} 
-    \item Add new input parameter $\lambda$ to EXTPAR and remove the sign of $\mu$ from MINPAR.
-      \begin{lstlisting}    
-        MINPAR = { {1, m0},
-           {2, m12},
-           {3, TanBeta},
-           {5, Azero} };
-
-  EXTPAR = { {61, LambdaInput} };
-      \end{lstlisting}
-     \item Also add this parameter to the DefaultParameterPoint.
-      \begin{lstlisting}
-       DefaultParameterPoint = {
-    {m0, 200},
-    {m12, 200},
-    {TanBeta, 10},
-    {Azero, -500}
-    {LambdaInput, 0.1}
-    };
-  \end{lstlisting}
-    \item Name the parameters you wish to be output by the EWSB conditions.  Note you should use the names given for them in parameters.m.    A common choice in the NMSSM is:
-\begin{lstlisting}
-EWSBOutputParameters = {vS, \[Kappa], ms2};
-\end{lstlisting}
-which will work for the semi-constrained NMSSM.
-\item choose the Highscale and Highscale first guess, if the default is different from what you wish to use.  
-\begin{lstlisting} 
- HighScale = g1 == g2;
- HighScaleFirstGuess = 2.0 10^16;
-\end{lstlisting}
-\item Add how the parameters found in parameters.m that are to be set at the highscale:
-\begin{lstlisting} 
-HighScaleInput={
-   {T[Ye], Azero*Ye},
-   {T[Yd], Azero*Yd},
-   {T[Yu], Azero*Yu},
-   {mHd2, m0^2},
-   {mHu2, m0^2},
-   {mq2, UNITMATRIX[3] m0^2},
-   {ml2, UNITMATRIX[3] m0^2},
-   {md2, UNITMATRIX[3] m0^2},
-   {mu2, UNITMATRIX[3] m0^2},
-   {me2, UNITMATRIX[3] m0^2},
-   {MassB, m12},
-   {MassWB,m12},
-   {MassG,m12},
-{\[Lambda], LambdaInput},
-   {T[\[Kappa]], Azero \[Kappa]},
-   {T[\[Lambda]], Azero LambdaInput}
-};
-\end{lstlisting}
-\item set lowscale boundary condititions if you want something different from the default.  The defaults are:
-\begin{lstlisting} 
- LowScale = SM[MZ];
- LowScaleFirstGuess = SM[MZ];
-\end{lstlisting}
-\item Change the LowscaleInput if necessary.  In this example we keep the same choice as in the MSSM
-  \begin{lstlisting} 
- LowScaleInput={
-   {vd, 2 MZDRbar / Sqrt[GUTNormalization[g1]^2 g1^2 + g2^2] Cos[ArcTan[TanBeta]]},
-   {vu, 2 MZDRbar / Sqrt[GUTNormalization[g1]^2 g1^2 + g2^2] Sin[ArcTan[TanBeta]]}
-};
-  \end{lstlisting}
-
-Finally we set the initial guesses for the paremeters.  In this example we overwtite the MSSM version replacing it with:
- \begin{lstlisting} 
-InitialGuessAtLowScale = {
-   {vd, SM[vev] Cos[ArcTan[TanBeta]]},
-   {vu, SM[vev] Sin[ArcTan[TanBeta]]},
-   {\[Lambda], LambdaInput},
-   {\[Kappa], 0.1},
-   {vS, 1000},
-   {ms2, SM[MZ]^2}
-};
-
-InitialGuessAtHighScale = {};
- \end{lstlisting}
-
-The code can now be created and also compiled (if you didn't select --disable-compile option when you ran the configure script) by typing \code{make -j2}.
-
- \end{enumerate}
-
-\end{enumerate}   
-\subsection{\ESSM}
-Detailed steps to create \ESSM model file.
-\begin{enumerate}
-\item Create new directory \code{mkdir myE6SSM}
-\item Move into this new directory \code{cd myE6SSM}
-\item copy the MSSM model file in \code{cp /path/to/SARAH/Models/MSSM/MSSM.m  myE6SSM.m}
-\item Edit myE6SSM.m as follows
-\begin{enumerate}
-\item  Change Model name, author and date
-\item Add new Vector superfield,
- \begin{lstlisting} 
-   Gauge[[4]]={Bp,  U[1], Ncharge, g1p, False, RpM};
- \end{lstlisting}
-\item Modify MSSM superfield definitions to include new charges.
-\begin{lstlisting}
-SuperFields[[1]] = {q, 3, {uL,  dL},     1/6, 2, 3, 1, RpM};  
-SuperFields[[2]] = {l, 3, {vL,  eL},    -1/2, 2, 1, 2, RpM};
-SuperFields[[3]] = {Hd,1, {Hd0, Hdm},   -1/2, 2, 1, -3, RpP};
-SuperFields[[4]] = {Hu,1, {Hup, Hu0},    1/2, 2, 1, -2, RpP};
-SuperFields[[5]] = {d, 3, conj[dR],    1/3, 1, -3, 2, RpM};
-SuperFields[[6]] = {u, 3, conj[uR],   -2/3, 1, -3, 1, RpM};
-SuperFields[[7]] = {e, 3, conj[eR],      1, 1,  1, 1, RpM};
-\end{lstlisting}
-Note that in the \ESSM model the charges are fixed to be those of the $U(1)_N$ as shown in e.g.
-\cite{King:2005jy}.
-However here one can leave the charges unassigned vby using parameters in place of the fixed charges.
-Here we fix the charges to simplify the C++ code.
-
-\item Add Singlet Higgs superfield to SuperField
-\begin{lstlisting}
-SuperFields[[8]] = {s, 1, sR,     0, 1,  1, 5, RpP};
-\end{lstlisting}
-\item Add inert Higgs-like superfields 
-\begin{lstlisting}
-SuperFields[[9]] = {H1I, 2, {H1I0, H1Im},  -1/2, 2, 1, -3, RpP};
-SuperFields[[10]] = {H2I, 2, {H2Ip, H2I0},   1/2, 2, 1, -2, RpP};
-SuperFields[[11]] = {sI, 2, sIR,    0, 1,  1, 5, RpM};
-\end{lstlisting}
-\item Add exotics to SuperFields
-\begin{lstlisting}
-SuperFields[[12]] = {Dx, 3, DxL,  -1/3, 1, 3, -2, RpP};
-SuperFields[[13]] = {Dxbar, 3, conj[DxbarR],  1/3, 1, -3, -3, RpP};
-\end{lstlisting}
-\item Add additional survival $SU(2)$ multiplets to SuperFields
-\begin{lstlisting}
-SuperFields[[14]] = {Hp, 1, {Hpd0, Hpdm},  1/2, 2,  1, 2, RpP};
-SuperFields[[15]] = {Hpbar, 1, {Hpup, Hpu0}, -1/2, 2,  1, -2, RpP};
-\end{lstlisting}
-\item Modify superpotential
-\begin{lstlisting}
-SuperPotential = Yu u.q.Hu - Yd d.q.Hd - Ye e.l.Hd + \[Lambda] s.Hu.Hd +  \[Lambda]12 s.H2I.H1I + \[Kappa] s.Dx.Dxbar + \[Mu]Pr Hp.Hpbar;
-\end{lstlisting}
-\item Define new gauge eigenstate Dirac spinors
-\begin{lstlisting}
-  DEFINITION[GaugeES][DiracSpinors]={
-    .............
-    .............
-    FS -> {FsR,conj[FsR]},
-    FBp -> {fBp,conj[fBp]},
-    HI0 -> {FH1I0, conj[FH2I0]},
-    HIC -> {FH1Im, conj[FH2Ip]},
-    FSI -> {FsIR, conj[FsIR]},
-    FDx1 -> {FDxL, 0},
-    FDx2 -> {0, FDxbarR}, 
-    Hp0 -> {FHpd0, conj[FHpu0]},
-    HpC -> {FHpdm, conj[FHpup]}
-  };
-\end{lstlisting}
-\item Add new VEV
-\begin{lstlisting} 
-  DEFINITION[EWSB][VEVs]= 
-  { {SHd0, {vd, 1/Sqrt[2]}, {sigmad, \[ImaginaryI]/Sqrt[2]},{phid,1/Sqrt[2]}},
-    {SHu0, {vu, 1/Sqrt[2]}, {sigmau, \[ImaginaryI]/Sqrt[2]},{phiu,1/Sqrt[2]}},
-    {SsR,  {vs, 1/Sqrt[2]}, {sigmaS, \[ImaginaryI]/Sqrt[2]},{phiS,1/Sqrt[2]}}
-};
-\end{lstlisting}
-\item Add new gauge boson  mass eigenstate
-\begin{lstlisting} 
-  DEFINITION[EWSB][GaugeSector] =
-  { 
-    {{VB,VWB[3],VBp},{VP,VZ,VZp},ZZ},
-    {{VWB[1],VWB[2]},{VWm,conj[VWm]},ZW},
-    {{fWB[1],fWB[2],fWB[3]},{fWm,fWp,fW0},ZfW}
-  }; 
-\end{lstlisting}
-\item Edit mixings for Higgs and neutralino states and add additional states that mix.
-\begin{lstlisting}
-DEFINITION[EWSB][MatterSector]= 
-{ ...............
-  ...............
-{{phid, phiu, phiS}, {hh, ZH}},
-{{sigmad, sigmau, sigmaS}, {Ah, ZA}},
-  ...............
-{{fB, fW0, FHd0, FHu0, FsR, fBp}, {L0, ZN}},
- ................ 
- {{SH1I0,conj[SH2I0]},{SHI0,UHI0}},
- {{SH1Im,conj[SH2Ip]},{SHIp,UHIp}},
- {{SHpd0,conj[SHpu0]},{SHp0,UHp0}},
- {{SHpdm,conj[SHpup]},{SHpp,UHpp}},
- {{FHpd0,FHpu0},{L0p,ZNp}},
- {{FH1I0,FH2I0},{L0I,ZNI}}
-\end{lstlisting}
-\end{enumerate}
-
-\end{enumerate}   
-\section{Tests}
-Comparisons against existing spectrum generators to test for deviations, e.g. the tests against MSSM and NMSSM (and E6SSM?).
-
-Comparisons against existing spectrum generators and SARAH generated spheno code for speed.
 

### 2014-05-11 doc/paper.tex
better spacing of listings

@@
   showstringspaces=false,
-  frame=single}
+  frame=single,
+  abovecaptionskip=0em,
+  aboveskip=1.5em,
+  belowcaptionskip=0.5em,
+  belowskip=1em,
+}
 \usepackage[pdftitle={FlexibleSUSY --- A spectrum generator generator for supersymmetric models},

### 2014-05-12 doc/paper.tex
improve the design goal and compilation sections a bit

@@
 describe detailed comparisons between our generated code and existing
-public spectrum generators as well as against the SPheno-like fortran
+public spectrum generators as well as against the SPheno-like FORTRAN
 code which can be created using SARAH.
@@
 adaptable spectrum generator in C++ for any SUSY model and some
-non-supersymmetric models.  The user specifies the model by giving the superfield content, superpotential, gauge symmetries and mass mixings in the  
-\sarah model files, examples of which can be found in the \code{Models}
-directory of \sarah and the \code{sarah} directory of FlexibleSUSY.  The user
-then also specifies the boundary conditions on the model parameters in a separate \code{FlexibleSUSY.m} file, examples of which can be found in the \code{model_files} directory.
-A detailed guide on how to write the FlexibleSUSY.m model is given in section \ref{Sec:modfile}.
+non-supersymmetric models.  The user specifies the model by giving the
+superfield content, superpotential, gauge symmetries and mass mixings
+in the \sarah model files, examples of which can be found in the
+\code{Models} directory of \sarah and the \code{sarah} directory of
+FlexibleSUSY.  The user then also specifies the boundary conditions on
+the model parameters in a separate \code{FlexibleSUSY.m} file,
+examples of which can be found in the \code{model_files} directory. A
+detailed guide on how to write the FlexibleSUSY.m model is given in
+section \ref{Sec:modfile}.
 
@@
 \cite{Staub:2010ty,Staub:2009bi,Staub:2010jh,Staub:2012pb,Staub:2013tta}
-  to obtain tree level expressions for the mass matrices and
-  electroweak symmetry breaking conditions, one-loop self energies,
-  one-loop tadpoles corrections and two-loop renormalisation group
-  equations(RGEs) for the model specified in the model
-  files. Additional corrections which have been calculated elsewhere,
-  such as two-loop corrections to the Higgs masses\footnote{By default FlexibleSUSY has two loop corrections to the Higgs masses for the MSSM\cite{Degrassi:2001yf,Brignole:2001jy,Dedes:2002dy,Brignole:2002bz,Dedes:2003km}  and NMSSM \cite{Degrassi:2009yq} in FORTRAN files supplied by Pietro Slavich.
-These are the same corrections which are implemented in many of the public spectrum generators.} may be added in
-  file specified by the user.
-
- \fs then
-converts these algerbraic expressions from \sarah into C++ code using
-the C++11 standard. \fs uses some parts of
+to obtain tree level expressions for the mass matrices and electroweak
+symmetry breaking conditions, one-loop self energies, one-loop
+tadpoles corrections and two-loop renormalisation group
+equations(RGEs) for the model specified in the model files. Additional
+corrections which have been calculated elsewhere, such as two-loop
+corrections to the Higgs masses\footnote{By default FlexibleSUSY has
+  two-loop corrections to the Higgs masses for the
+  MSSM\cite{Degrassi:2001yf,Brignole:2001jy,Dedes:2002dy,Brignole:2002bz,Dedes:2003km}
+  and NMSSM \cite{Degrassi:2009yq} in FORTRAN files supplied by Pietro
+  Slavich. These are the same corrections which are implemented in
+  many of the public spectrum generators.} may be added in file
+specified by the user.
+
+\fs then converts these algerbraic expressions from \sarah into C++
+code using the C++11 standard. \fs uses some parts of
 SOFTSUSY\cite{Allanach:2001kg}, the very fast Eigen library
@@
 to find simultaneous solutions for the EWSB, low energy data, and
-other, user supplied, boundary conditions (e.g. the usual mSUGRA
+other, user supplied, boundary conditions (e.g.\ the usual mSUGRA
 universality conditions) and calculates the mass spectrum.
@@
 components to be replaced with an alternative if available.  Indeed
- \fs has been intentionally designed to grow as alternative solvers
- and calculators are added and a subsequent release with an
- alternative lattice solver is already planned.  In addition if one
- already has an existing code for some other purpose and simply wishes
- to improve upon this by adding, e.g, new RGEs or self energies the
- modular framework makes this straightforward by adding or
- interfacing with the appropriate routine.
+\fs has been intentionally designed to grow as alternative solvers and
+calculators are added and a subsequent release with an alternative
+lattice solver is already planned.  In addition if one already has an
+existing code for some other purpose and simply wishes to improve upon
+this by adding, e.g., new RGEs or self energies the modular framework
+makes this straightforward by adding or interfacing with the
+appropriate routine.
 
@@
 \paragraph{Speed}
+
 Exploring the parameter space of supersymmetric models with a high
-number of free parameters is quite time consuming.  Therefore
-FlexibleSUSY aims to produce spectrum generators with a short run-time.
-
-The two most time consuming parts of a spectrum generator are the
-calculation of the beta functions and the pole masses for mixed
-particles.
-
-The reason is the following: The RG solving algorithms usually need
-$O(10)$ iterations to calculate a particle spectrum with $0.1\%$
-precision.  During each iteration the Runge--Kutta algorithm needs to
-calculate all beta functions $O(50)$ times.  Most two-loop beta
-functions involve $O(50)$ matrix multiplications and additions.  All
-together one arrives at $O(2.5\cdot 10^4)$ matrix multiplications.  To
-optimize these matrix multiplications FlexibleSUSY uses the fast
-linear algebra package \href{Eigen}{http://eigen.tuxfamily.org}, which
-exploits C++ expression templates to make the matrix multiplications
-easy to write and optimize.
-
-The second most time consuming part is the precise calculation of the
-pole masses for mixed particles.  For each particle $\psi_i$ the full
-self-energy matrix $\Sigma^\psi_{ij}(p=m^\text{tree}_{\psi_i})$ has to
-be evaluated.  Each self-energy matrix entry again involves the
-evaluation of $O(50)$ Feynman diagrams, which involves the calculation
-of vertices and loop functions.  All in all, one arrives at $O(500)$
-Feynman diagram or $O(3\cdot 10^4)$ loop function evaluations.  To
-speed up the calculation of the pole masses FlexibleSUSY makes use of
-multi-threading, where each pole mass is calculated in a separate
-thread.  By this technique one can gain a speed-up of $20$--$30\%$.
+number of free parameters is quite time consuming.  Therefore \fs aims
+to produce spectrum generators with a short run-time.  The two most
+time consuming parts of a SUSY spectrum generator are usually the
+calculation of the two-loop $\beta$-functions and the pole masses of
+mixed particles:
+%
+\begin{itemize}
+\item \emph{Calculation of the $\beta$-functions:} The RG solving
+  algorithms usually need $O(10)$ iterations between the high and the
+  low scale to find a set of parameters consistent will all boundary
+  conditions with a $0.01\%$ precision goal.  During each iteration
+  the Runge-Kutta algorithm needs to calculate all $\beta$-functions
+  $O(50)$ times.  Most two-loop $\beta$-functions involve $O(50)$
+  matrix multiplications and additions.  All together one arrives at
+  $O(10^4)$ matrix operations.  To optimize these, \fs uses the fast
+  linear algebra package \href{Eigen}{http://eigen.tuxfamily.org}.
+  Eigen uses C++ expression templates to remove temporary objects and
+  enable lazy evaluation of the expressions.  It supports explicite
+  vectorization, and provides fixed-size matrices to avoid dynamic
+  memory allocation.  All of these features in combination result in
+  very fast code for the calculation of the $\beta$-functions in \fs.
+%
+\item \emph{Calculation of the pole masses:} The second most time
+  consuming part is the precise calculation of the pole masses for
+  mixed particles.  For each particle $\psi_k$ in a multiplet the full
+  self-energy matrix $\Sigma^\psi_{ij}(p=m^\text{tree}_{\psi_k})$ has
+  to be evaluated.  Each self-energy matrix entry again involves the
+  calculation of $O(50)$ Feynman diagrams, each involving the
+  calculation of vertices and a loop-function.  All in all, one
+  arrives at $O(500)$ Feynman diagrams and $O(10^4)$ loop function
+  evaluations.  To speed up the calculation of the pole masses \fs
+  makes use of multi-threading, where each pole mass is calculated in
+  a separate thread.  This allows the operating system to distribute
+  these calculations among different CPU cores.  With this technique
+  one can gain a speed-up of $20$--$30\%$.
+\end{itemize}
 
@@
 The large variety of supersymmetric models makes it likely that the
-user needs to modify the generated spectrum generator source code.
-FlexibleSUSY uses object orientation with C++ to modularize the source
-code to make it easy for the user to modify and extend the spectrum
-generator.  An important application here are the constraints: The RG
-solvers provide a constraint interface,
-\figref{fig:schematic-two-scale-constraint-interface},
+user wants to modify the generated spectrum generator source code.
+\fs uses C++ object orientation features to modularize the source code
+in order to make it easy for the user to modify, reuse and extend the
+spectrum generator.  An important application of this concept are the
+boundary conditions: The RG solver provides a common \code{Constraint}
+interface, see \figref{fig:schematic-two-scale-constraint-interface}.
 %
@@
   \begin{tikzpicture}
-    \umlclass[x=0, y=0]{RGFlow$<$Two\_scale$>$}{}{
+    \umlclass[x=0, y=0]{RGFlow<Two\_scale>}{}{
       + solve()
     }
-    \umlclass[x=6, y=0, type=abstract]{Constraint$<$Two\_scale$>$}{}{
+    \umlclass[x=6, y=0, type=abstract]{Constraint<Two\_scale>}{}{
       + \umlvirt{apply()}\\
@@
     }
-    \umldep{RGFlow$<$Two\_scale$>$}{Constraint$<$Two\_scale$>$}
-    \umlinherit{MyConstraint}{Constraint$<$Two\_scale$>$}
+    \umldep{RGFlow<Two\_scale>}{Constraint<Two\_scale>}
+    \umlinherit{MyConstraint}{Constraint<Two\_scale>}
   \end{tikzpicture}
@@
 %
-which can be implemented to create the custom constraint
-\code{MyConstraint}.  In the \code{apply()} function the constraint
-application takes place and the \code{get_scale()} function specifies
-the scale at which the constraint is to be applied.
+To create a custom constraint \code{MyConstraint}, the user implements
+the interface by inheriting from \code{Constraint} and writing the
+\code{apply()} and \code{get_scale()} methods (see
+\secref{sec:boundary-conditions} for more details).  The so created
+constraint class can then be passed to the RG solver to impose the
+user-defined boundary condition at the chosen scale.
 
@@
 \label{Sec:download}
-Flexible SUSY can be downloaded from \url{http://flexiblesusy.hepforge.org}.  
-To use FlexibleSUSY please make sure you have the following installed.  
+
+\subsection{Requirements}
+
+\fs can be downloaded from \url{http://flexiblesusy.hepforge.org}.  To
+create a custom spectrum generator the following requirements are
+necessary:
+%
 \begin{itemize}
-\item Mathematica, version 7 or higher
+\item MATHEMATICA, version 7 or higher
 \item SARAH, version 4.0.4 or higher \url{http://sarah.hepforge.org}
 \item C++11 compatible compiler (g++ 4.4.7 or higher, clang++ 3.1 or
-  higher)
-\item Fortran compiler (gfortran, ifort etc.)
-\item Eigen library, version 3.1 \url{http://eigen.tuxfamily.org}
+  higher, icpc 12.1 or higher)
+\item FORTRAN compiler (gfortran, ifort etc.)
+\item Eigen library, version 3.1 or higher
+  \url{http://eigen.tuxfamily.org}
 \item Boost library, version 1.36.0 or higher
@@
 \item GNU scientific library \url{http://www.gnu.org/software/gsl/}
-\item lapack (needed for the Lattice algorithm only)
-  \url{http://www.netlib.org/lapack/}
+\item lapack \url{http://www.netlib.org/lapack/}
 \end{itemize}
@@
 \subsection{Quick Start}
-Create a MSSM spectrum generator
+
+To create a MSSM spectrum generator the following three commands:
+%
 \begin{lstlisting}[language=bash]
@@
 $ make
-  \end{lstlisting}
-Run the spectrum generator using the default parameter point
+\end{lstlisting}%% $
+%
+The first command creates the model directory \code{models/MSSM/}
+together with a MSSM model file.  The \code{configure} scrip checks
+the system requirements and creates the \code{Makefile}.  See
+\code{./configure --help} for more options.  Executing \code{make}
+will start MATHEMATICA to generate the spectrum generator and compile
+it.  The resulting executable can be run like this:
 \begin{lstlisting}[language=bash]
 $ ./models/MSSM/run_MSSM.x
-  \end{lstlisting}
-Run the spectrum generator using a SLHA input file
+\end{lstlisting}%% $
+%
+An SLHA input file can be given via the \code{--slha-input-file=}
+option, for example:
+%
 \begin{lstlisting}[language=bash]
-$ ./models/MSSM/run_MSSM.x --slha-input-file=templates/MSSM/LesHouches.in.MSSM --slha-output-file=LesHouches.out.MSSM
+$ ./models/MSSM/run_MSSM.x \
+    --slha-input-file=templates/MSSM/LesHouches.in.MSSM
 \end{lstlisting}%% $
@@
 \label{Sec:modfile}
-A \fs model is configured via a (Mathematica) model file, which in
-case of the MSSM looks like this
+A \fs model is configured via a (MATHEMATICA) model file
+\code{FlexibleSUSY.m} in the model directory \code{models/<model>/}.
+Many example model files can be found in \code{model_files/}.  The
+default model file for the CMSSM looks like this:
 %
@@
 name of the parameter.  The input parameters can be set to default
-values in the \code{DefaultParameterPoint} list, see above.
+values in the \code{DefaultParameterPoint} list.
 
@@
 parameters until the EWSB conditons are fulfilled.  In the CMSSM these
-are \code{\\[Mu]} and \code{B[\\[Mu]]}.
+are the superpotential parameter \code{\\[Mu]} and its corresponding
+soft-breaking parameter \code{B[\\[Mu]]}.
 
 Furthermore, the user has to specify the form of three model
-constraints: low-scale, susy-scale and high-scale constraints.  In \fs
+constraints: low-scale, SUSY-scale and high-scale constraints.  In \fs
 they are named as \code{LowScale}, \code{SUSYScale} and
@@
 constraint name and \code{Minimum} or \code{Maximum}, respectively).
-This avoids underflows or overflows of the scale value during the
-iteration.  This is especially useful in models where the iteration is
-very unstable and the value of the scale is very sensitive to the
-model parameters.
+The latter can avoid underflows or overflows of the scale value during
+the iteration.  This is especially useful in models where the
+iteration is very unstable and the value of the scale is very
+sensitive to the model parameters.  The meaning of the three
+constraints is the following:
 %
@@
 \item \emph{High-scale constraint:} The high-scale constraint is
-  usuall the GUT-scale constraint, applied at the scale where the
+  usually the GUT-scale constraint, applied at the scale where the
   gauge couplings $g_1$ and $g_2$ unify.  The high-scale can be
@@
   value.
-\item \emph{Susy-scale constraint:} The susy-scale constraint is the
-  constraint where \fs will impose the EWSB conditions and calculate
-  the spectrum.  It is usually defined by
-  $\sqrt{m_{\tilde{t}_1}m_{\tilde{t}_2}}$.
+%
+\item \emph{SUSY-scale constraint:} The SUSY-scale constraint is the
+  constraint where \fs imposes the EWSB conditions and calculate the
+  pole mass spectrum.  Its scale $M_S$ is usually defined as
+  $\sqrt{m_{\tilde{t}_1}m_{\tilde{t}_2}}$.  In the example above it
+  has the value $M_S = \sqrt{m_{\tilde{u}_1}m_{\tilde{u}_6}}$, where
+  $m_{\tilde{u}_1}$ and $m_{\tilde{u}_6}$ are the \DRbar\ masses of
+  the lightest and heaviest up-type squark, respectively.
+%
 \item \emph{Low-scale constraint:} The low-scale constraint is the
   constraint where the gauge couplings $g_1$, $g_2$, $g_3$ and the
-  Yukawa couplings $y_f$ will be calculated from given Standard Model
+  Yukawa couplings $y_f$ are calculated from given Standard Model
   input $\alpha_{\text{e.m.}}(M_Z)$, $\alpha_{s}(M_Z)$, $m_f^\pole$,
@@
   Sec.~\ref{sec:calculation-of-gauge-couplings}--\ref{sec:calculation-of-yukawa-couplings}.
-  This should be done at the $Z$ mass scale $M_Z$.  Note that the
-  gauge and Yukawa couplings are calculated automatically and
-  therefore do not need to be specified in the \code{LowScaleInput}
-  variable.
+  This should be done at the $Z$ mass scale $M_Z$.  The gauge
+  couplings are calculated automatically and therefore do not need to
+  be specified in the \code{LowScaleInput} variable.
 \end{itemize}
 %
-Note, that \fs GUT-normalizes all gauge couplings.  Thus, the
-high-scale definition takes the simple form \code{g1 == g2}.  Another
-consequence is that in the calculation of the VEVs $v_u$ and $v_d$
-from $M_Z$ and $\tan\beta$ at the low-scale the GUT-normalization has
-to be taken into account, see above.
+\fs GUT-normalizes all gauge couplings.  Thus, the high-scale
+definition takes the simple form \code{g1 == g2}.  Another consequence
+is that in the calculation of the VEVs $v_u$ and $v_d$ from $M_Z$ and
+$\tan\beta$ at the low-scale the GUT-normalization has to be taken
+into account, see the example above.
 
@@
 \code{OnlyLowEnergyFlexibleSUSY = True}.  In this case the high-scale
-constraint will be ignored and only the low-scale and susy-scale
+constraint will be ignored and only the low-scale and SUSY-scale
 constraints are kept.  All model parameters that are not specified in
 \code{MINPAR} or \code{EXTPAR} will be read from the corresponding
-input blocks in the SLHA input file and will be set at the susy-scale.
+input blocks in the SLHA input file and will be set at the SUSY-scale.
 
@@
 lightest supersymmetric particle (LSP).  To have this function be
-created by \fs, the model file variable \code{PotentialLSPParticles}
-must be set to a list of SUSY particles which are potential LSPs.  In
-the model file example above, the particles \code{Chi}, \code{Cha},
+created the model file variable \code{PotentialLSPParticles} must be
+set to a list of SUSY particles which are potential LSPs.  In the
+model file example above, the particles \code{Chi}, \code{Cha},
 \code{Glu}, \code{Sv}, \code{Su}, \code{Sd}, \code{Se} (neutralino,
@@
 
-The parameters of the susy model together with their RGEs are stored
+The parameters of the SUSY model together with their RGEs are stored
 in the model class hierarchy, see the UML diagram in
@@
 
-All susy model parameters and their $\beta$-functions are contained in
+All SUSY model parameters and their $\beta$-functions are contained in
 the derived classes.  The structure of the $\beta$-functions of a
-general susy model
+general SUSY model
 \cite{Jones:1974pg,Jones:1983vk,West:1984dg,Martin:1993yx,Yamada:1993ga,MV94,Fonseca:2011vn,Sperling:2013eva,Sperling:2013xqa}
@@
 \begin{enumerate}
-\item \emph{susy parameters:} gauge couplings, superpotential
+\item \emph{SUSY parameters:} gauge couplings, superpotential
   parameters and VEVs and
@@
 %
-The $\beta$-functions of the susy parameters do only depend on the
-susy parameters themselfs.  They do not depend on soft-breaking
+The $\beta$-functions of the SUSY parameters do only depend on the
+SUSY parameters themselfs.  They do not depend on soft-breaking
 parameters.  However, the $\beta$-functions of the soft-breaking
@@
 \code{<model>_susy_parameters} direcly inherits from
-\code{Beta_functions} and implements the $\beta$-functions of the susy
+\code{Beta_functions} and implements the $\beta$-functions of the SUSY
 parameters.  The class of soft-breaking parameters
@@
 \code{<model>_susy_parameters} and implements the $\beta$-functions of
-the soft-breaking parameters in terms of all model parameters.
-This hierarchy allows to
-%
-\begin{enumerate}
-\item use the RGE running of all model parameters via the common
-  \code{Beta_function} interface and to
-\item run the susy parameters independently of the soft-breaking
+the soft-breaking parameters in terms of all model parameters.  The so
+constructed class hierarchy allows to (i) use the RGE running of all
+model parameters via the common \code{Beta_function} interface and to
+(ii) run the SUSY parameters independently of the soft-breaking
 parameters.
-\end{enumerate}
 
 \subsection{Boundary conditions}
-
-The parameters of a susy model have to meet certain boundary
+\label{sec:boundary-conditions}
+
+The parameters of a SUSY model have to meet certain boundary
 conditions.  For example, the running gauge and Yukawa couplings
@@
 By default, \fs creates three default boundary conditions for each
-susy model:
+SUSY model:
 %
@@
 \item The \emph{SUSY-scale constraint} is indended to set boundary
-  conditions at the mass scale $M_S$ of the new susy particles.  The
+  conditions at the mass scale $M_S$ of the new SUSY particles.  The
   value of $M_S$ is defined by the model file variable
@@
 
-\item The \emph{low-scale constraint} is indended to match the susy
+\item The \emph{low-scale constraint} is indended to match the SUSY
   model to the Standard Model at the scale $M_Z$.  It does so by
-  calculating the gauge and Yukawa couplings of the susy model from
+  calculating the gauge and Yukawa couplings of the SUSY model from
   the known Standard Model fermion and gauge boson masses, as well as
@@
 The low-scale constraint automatically calculates the \DRbar\ gauge
-couplings $g_{i,\text{susy}}^{\text{\DRbar}}(M_Z)$ in the susy model
+couplings $g_{i,\text{susy}}^{\text{\DRbar}}(M_Z)$ in the SUSY model
 at the scale $M_Z$.  It starts from the known electromagnetic and
@@
 \cite{Beringer:1900zz}.  These are converted to the electromagnetic
-and strong \DRbar\ couplings in the susy model
+and strong \DRbar\ couplings in the SUSY model
 $e_{\text{susy}}^{\text{\DRbar}}(M_Z)$ and
@@
 $M_{Z,\text{susy}}^{\text{\DRbar}}(M_Z)$ is used to calculate
-$\theta_W$ in the susy model in the \DRbar\ scheme.  In the MSSM, for
+$\theta_W$ in the SUSY model in the \DRbar\ scheme.  In the MSSM, for
 example, one has
@@
 $\theta_{W,\text{susy}}^{\text{\DRbar}}(M_Z)$ allows to calculate the
-(GUT-normalized) $U(1)_Y$ and $SU(2)_L$ gauge couplings in the susy
+(GUT-normalized) $U(1)_Y$ and $SU(2)_L$ gauge couplings in the SUSY
 model.  In the MSSM one has for instance
@@
 At the low-scale the \DRbar\ Yukawa coupling matrices
-$y_f^{\text{\DRbar}}(M_Z)$ ($f=u,d,e$) in the susy model are
+$y_f^{\text{\DRbar}}(M_Z)$ ($f=u,d,e$) in the SUSY model are
 calculated automatically if the user has set $y_f$ to the value
@@
 \begin{enumerate}
-\item At the $M_Z$ scale the susy model gauge couplings $g_{1,2,3}$
+\item At the $M_Z$ scale the SUSY model gauge couplings $g_{1,2,3}$
   are set to the known Standard Model values (ignoring threshold
@@
   \code{LowScaleInput}) are imposed.
-\item The Yukawa couplings $y_{u,d,e}$ of the susy model are set to
+\item The Yukawa couplings $y_{u,d,e}$ of the SUSY model are set to
   the known Standard Model Yukawa couplings.
-\item The susy parameters are run to the user-supplied first guess of
+\item The SUSY parameters are run to the user-supplied first guess of
   the high-scale (\code{HighScaleFirstGuess}).
@@
   \item The \DRbar\ gauge and Yukawa couplings $g_{1,2,3}(M_Z)$,
-    $y_{u,d,e}(M_Z)$ of the susy model are calculated using threshold
+    $y_{u,d,e}(M_Z)$ of the SUSY model are calculated using threshold
     corrections.
@@
   \end{enumerate}
-\item Run model parameters to the susy-scale (\code{SUSYScale})
+\item Run model parameters to the SUSY-scale (\code{SUSYScale})
   \begin{enumerate}
   \item Calculate the \DRbar\ mass spectrum
-  \item Recalculate the susy-scale
-  \item Apply the susy-scale constraint (\code{SUSYScaleInput})
+  \item Recalculate the SUSY-scale
+  \item Apply the SUSY-scale constraint (\code{SUSYScaleInput})
   \item Solve the EWSB eqs.\ at the loop level
@@
 \code{HighScale}.  In this case all model parameters are run to the
-susy-scale and the pole-mass spectrum is calculated.  If the user has
+SUSY-scale and the pole-mass spectrum is calculated.  If the user has
 chosen a specific output scale for the running \DRbar\ model
@@
     \path[arrow] (MZ) -- node {run to \code{HighScale}} (MX);
-    \node[block,below of=MX] (MS) {Apply susy-scale boundary conditions
+    \node[block,below of=MX] (MS) {Apply SUSY-scale boundary conditions
       (\code{SUSYScaleInput}) and solve EWSB};

### 2014-05-13 doc/paper.tex
improve section 8

@@
 \newcommand{\secref}[1]{Sec.~\ref{#1}}
+\newcommand{\tabref}[1]{\tablename~\ref{#1}}
 \DeclareMathOperator{\diag}{diag}
@@
 \label{Sec:comparison}
-We compare the run-time of five different CMSSM spectrum generators:
+
+One of \fs's design goals is a short run-time.  In this section we
+demonstrate that this goal was achieved by comparing the run-time of
+two different sets of CMSSM spectrum generators:
 %
 \begin{itemize}
-\item FlexibleSUSY-FV (version 1.0.0), which is based on SARAH's MSSM
-  model file that allows flavour violation,
-\item FlexibleSUSY-NoFV (version 1.0.0), without flavour violation
-\item Softsusy (version 3.4.0), without flavour violation
-\item SPheno (3.2.4), without flavour violation
-\item SPhenoMSSM (generated with SARAH 4.1.0 and linked against SPheno
-  3.2.4), which uses SARAH's MSSM model file that includes flavour
-  violation.
+\item \emph{Without flavour violation:} Disallowing flavour violation
+  simplifies the calculation of the pole masses, because
+  flavour-off-diagonal sfermion self-energy matrix elements don't need
+  to be calculated.  Here we compare \fs's non-flavour violating CMSSM
+  spectrum generator FlexibleSUSY-NoFV (version 1.0.0) against SPheno
+  (version 3.2.4) and Softsusy (version 3.4.0).
+%
+\item \emph{With flavour violation:} Allowing for flavour violation in
+  general increases the run-time of spectrum generators, because the
+  full $6\times 6$ sfermion self-energy matrices have to be
+  calculated.  Here we compare FlexibleSUSY-FV (version 1.0.0) and
+  SPhenoMSSM (generated with SARAH 4.1.0 and linked against SPheno
+  3.2.4).  Both spectrum generators are based on SARAH's MSSM model
+  file, which allows for flavour violation.
 \end{itemize}
@@
 13.1.3 20130607.  SPheno and SPhenoMSSM are compiled with Intel ifort
-13.1.3 20130607\footnote{We are using Intel's ifort compiler, instead
-  of gfortran, because it decreases the run-time of SPheno and
-  SPhenoMSSM by approximately a factor $1.5$.}.  We're generating
-$2\cdot 10^{4}$ random CMSSM parameter points with $m_0\in [50,1000]$
-$m_{1/2}\in [50,1000]$, $\tan\beta\in [1,100]$, $\sign\mu\in
-\{-1,+1\}$ and $A_0\in [-1000,1000]$.  For each point an SLHA input
+13.1.3 20130607.\footnote{Intel's ifort compiler decreases the
+  run-time of SPheno and SPhenoMSSM by approximately a factor $1.5$,
+  compared to gfortran.}
+
+For the run-time comparison we're generating $2\cdot 10^{4}$ random
+CMSSM parameter points with $m_0\in [50,1000]\unit{GeV}$, $m_{1/2}\in
+[50,1000]\unit{GeV}$, $\tan\beta\in [1,100]$, $\sign\mu\in \{-1,+1\}$
+and $A_0\in [-1000,1000]\unit{GeV}$.  For each point an SLHA input
 file is created by appending the values of $m_0$, $m_{1/2}$,
@@
 average run-times for three different CPU types can be found in
-\tablename~\ref{tab:run-time-comparison}.  The first column shows the
-run-time on a Intel Core2 Duo (P8600) where only one core was enabled.
-The second column shows the run-time on a Intel Core2 Duo (P8600)
+\tabref{tab:run-time-comparison}.  The first column shows the run-time
+on a Intel Core2 Duo (P8600, $2.40\unit{GHz}$) where only one core was
+enabled.  The second column shows the run-time on the same processor
 where both cores were enabled.  In the third column a Intel Xeon
-(L5640) was used, which has $6$ CPU cores.
+(L5640, $2.27\unit{GHz}$) was used, which has $6$ CPU cores.
 %
@@
     \toprule
-                          & Intel Core2 Duo (P8600)    & Intel Core2 Duo (P8600)     & Intel Xeon (L5640)\\
-                          & (1 core, $2.40\unit{GHz}$) & (2 cores, $2.40\unit{GHz}$) & (6 cores, $2.27\unit{GHz}$)\\
+                            & Intel Core2 Duo    & Intel Core2 Duo   & Intel Xeon\\
+                            & (P8600, 1 core)    & (P8600, 2 cores)  & (L5640, 6 cores)\\
     \midrule
@@
     FlexibleSUSY-FV 1.0.0   & $0.150\unit{s}$    & $0.113\unit{s}$   & $0.074\unit{s}$\\
-    SPhenoMSSM 4.1.0${}^*$  & $0.415\unit{s}$    & $0.401\unit{s}$   & $0.370\unit{s}$\\
+    SPhenoMSSM 4.1.0        & $0.415\unit{s}$    & $0.401\unit{s}$   & $0.370\unit{s}$\\
     \bottomrule
@@
   \caption{Average run-time of CMSSM spectrum generators for
-    for random parameter points.  The upper }
+    for random parameter points.  The first three rows show
+    spectrum generators which disallow flavour violation.  Rows
+    $4$--$5$ contain flavour violatig spectrum generators, based
+    on SARAH's MSSM model file.}
   \label{tab:run-time-comparison}
@@
 
-Under both the non-flavor violating spectrum generators
-FlexibleSUSY-NoFV, Softsusy and SPheno (first three rows) as well as
-the flavor violating ones FlexibleSUSY-NoFV and SPhenoMSSM (4th and
-5th row) we find that \fs is significantly fastest.  We also find that
-increasing the number of CPU cores reduces the run-time of \fs.  The
-reason is that \fs calculates each pole mass in a separate thread, and
-therefore benefits from multi-core CPUs.  Under the hand-written CMSSM
-spectrum generators SPheno and Softsusy we find that SPheno is fastest
-on average.
+Under both the non-flavour violating spectrum generators (first three
+rows) as well as the flavour violating ones (4th and 5th row) we find
+that \fs is significantly fastest.  Compared to SPheno,
+FlexibleSUSY-NoFV is faster by a factor $1.4$--$1.7$, and compared to
+Softsusy around a factor $2$--$2.5$.  Under the flavour violating
+spectrum generators FlexibleSUSY-FV is faster than SPhenoMSSM by a
+factor $2.8$--$5$.  Reason for the long run-time of SPhenoMSSM is the
+long calculation duration of the two-loop $\beta$-functions.  Here \fs
+benefits a lot from Eigen's well-optimizable matrix expressions.  We
+also find that increasing the number of CPU cores reduces the run-time
+of \fs.  The reason is that \fs calculates each pole mass in a
+separate thread, and therefore benefits from multi-core CPUs.
 

### 2014-05-13 doc/paper.tex
extending Appendix B

@@
 \end{enumerate}
-\section{Tests}
-Comparisons against existing spectrum generators to test for deviations, e.g. the tests against MSSM and NMSSM (and E6SSM?).
-
-Comparisons against existing spectrum generators and SARAH generated spheno code for speed.
+
+\section{Numeric tests}
+
+For checking the correctness of \fs's generated spectrum generators
+extensive unit tests against Softsusy's MSSM, and NMSSM
+implementations (both $Z_3$-invariant and $Z_3$-violating variants)
+were done: We checked mass matrices, EWSB equations, one- and two-loop
+$\beta$-functions, one- and two-loop self-energies and one- and
+two-loop tadpoles and found all to agree within double machine
+precision.  We also compared the overall pole mass spectrum after the
+full fixed-point iteration has finished, and found the spectra to
+agree at the sub-permille level.  Analytic tests of the $R$-symmetric
+low-energy model, MRSSM, were done by Philip Diessner and several bugs
+in \fs and SARAH were identified and fixed.
+
+TODO: Adding tests against the CE$_6$SSM?
+
+We also compared the run-time of \fs against SPheno, Softsusy and the
+SARAH generated MSSM spectrum generator SPhenoMSSM.  The test results
+can be found in \secref{Sec:comparison}.
 

### 2014-05-13 doc/paper.tex
adding References title
put SLHA template file into lstlisting environmetn

@@
 %
-\begin{verbatim}
+\begin{lstlisting}
 Block MODSEL                 # Select model
@@
     5   [-1000..1000]        # A0(MX)
-\end{verbatim}
+\end{lstlisting}
+
+\clearpage
+\section*{References}
 

### 2014-05-13 doc/paper.tex
give more details on the fixed-point iteration

@@
 
-The $\beta$-functions for the model parameters together with the
-boundary conditions for a boundary value problem.  The default
-two-scale RG solver tries to find the full set of model parameters
-consistent with all constraints at all scales.  It does so by running
-iteratively between the scales of all boundary conditions, imposing
-the constraints and checking for convergence after each iteration.
-This approach is decribed in \cite{Barger:1993gh} for the MSSM and is
-widely implemented in SUSY spectrum generators.
+The RGEs for the model parameters together with the boundary
+conditions form a boundary value problem.  The default two-scale RG
+solver tries to find a set of model parameters consistent with all
+constraints at all scales.  It does so by running iteratively between
+the scales of all boundary conditions, imposing the constraints and
+checking for convergence after each iteration.  This approach is
+decribed in \cite{Barger:1993gh} for the MSSM and is widely
+implemented in SUSY spectrum generators.
 
 The in more detail the two-scale algorithm used in \fs works as
-follows, see \figref{fig:two-scale-algorithm}:
+follows, see also \figref{fig:two-scale-algorithm}:
 %
@@
 \begin{enumerate}
-\item At the $M_Z$ scale the SUSY model gauge couplings $g_{1,2,3}$
-  are set to the known Standard Model values (ignoring threshold
-  corrections).
-\item Then, the user-defined low-scale boundary conditions (defined in
-  \code{LowScaleInput}) are imposed.
-\item The Yukawa couplings $y_{u,d,e}$ of the SUSY model are set to
-  the known Standard Model Yukawa couplings.
+\item At the $M_Z$ scale the gauge couplings $g_{1,2,3}$ are set to
+  the known Standard Model values (ignoring threshold corrections).
+\item The user-defined initial guess at the low-scale (defined in
+  \code{InitialGuessAtLowScale}) is imposed.  In the example given in
+  \secref{Sec:modfile} the Higgs VEVs are set to
+  \begin{align}
+    v_d &= v \cos\beta, & v_u &= v \sin\beta ,
+  \end{align}
+  where $v=246.22\unit{GeV}$.  Afterwards, the Yukawa couplings
+  $y_{u,d,e}$ of the SUSY model are set to the known Standard Model
+  Yukawa couplings (ignoring SUSY particle corrections).
 \item The SUSY parameters are run to the user-supplied first guess of
@@
 \item The high-scale boundary condition is imposed (defined in
-  \code{HighScaleInput}).
+  \code{HighScaleInput}).  Afterwards, the user-defined initial guess
+  for the remaining model parameters (defined in
+  \code{InitialGuessAtHighScale}) is imposed.  In the example given in
+  \secref{Sec:modfile} the superpotential parameter $\mu$ is set to
+  the value $1.0$ and its corresponding soft-breaking parameter $B\mu$
+  is set to zero.
 \item All model parameters are run to the first guess of the low-scale
@@
 \item \label{rge-step-one} All model parameters are run to the
-  low-scale (\code{LowScale})
+  low-scale (\code{LowScale}).
   \begin{enumerate}
   \item The \DRbar\ mass spectrum is calculated.
-  \item The low-scale is recalculated.
-  \item The \DRbar\ gauge and Yukawa couplings $g_{1,2,3}(M_Z)$,
-    $y_{u,d,e}(M_Z)$ of the SUSY model are calculated using threshold
-    corrections.
+  \item The low-scale is recalculated.  In the above example this step
+    is trivial, because the low-scale is fixed to be $M_Z$.
+  \item The \DRbar\ gauge couplings $g_{1,2,3}(M_Z)$ of the SUSY model
+    are calculated using threshold corrections as described in
+    \secref{sec:calculation-of-gauge-couplings}.
   \item The user-defined low-scale constraint is imposed
-    (\code{LowScaleInput})
+    (\code{LowScaleInput}).  In the example above, the Yukawa
+    couplings are calculated automatically as described in
+    \secref{sec:calculation-of-yukawa-couplings} and the Higgs VEVs
+    are set to
+    \begin{align}
+      v_d(M_Z) &= \frac{2 M_Z(M_Z)}{\sqrt{0.6 g_1^2(M_Z) + g_2^2(M_Z)} \cos\beta(M_Z)}, \\
+      v_u(M_Z) &= \frac{2 M_Z(M_Z)}{\sqrt{0.6 g_1^2(M_Z) + g_2^2(M_Z)} \sin\beta(M_Z)}.
+    \end{align}
+    Since the Hypercharge gauge coupling $g_1$ is GUT normalized, the
+    normalization factor $\sqrt{3/5}$ has to be included in the above
+    relations.
   \end{enumerate}
-\item Run all model parameters to the high-scale (\code{HighScale})
+\item Run all model parameters to the high-scale (\code{HighScale}).
   \begin{enumerate}
-  \item Recalculate the high-scale
-  \item Apply the high-scale constraint (\code{HighScaleInput})
+  \item Recalculate the high-scale as
+    \begin{align}
+      M_X' = M_X \exp\left(\frac{g_2(M_X)-g_1(M_X)}{\beta_{g_1} -
+          \beta_{g_2}}\right),
+    \end{align}
+    where $\beta_{g_i}$ is the two-loop $\beta$-function of the gauge
+    coupling $g_i$.  The value $M_X'$ is used as new high-scale in the
+    next iteration.
+  \item Impose the high-scale constraint (\code{HighScaleInput}).  In
+    the MSSM example the following soft-breaking parameters are fixed
+    to the universal values $m_0$, $M_{1/2}$ and $A_0$:
+    \begin{align}
+      A^f(M_X) &= A_0 y_f(M_X) & &(f=u,d,e),\\
+      m_{H_i}^2(M_X) &= m_0^2 & &(i=1,2),\\
+      m_{f}^2(M_X) &= m_0^2\mathbf{1} & &(f=q,l,d,u,e),\\
+      M_{i}(M_X) &= M_{1/2} & &(i=1,2,3).
+    \end{align}\label{eq:fs-cmssm-high-scale-bc}%
   \end{enumerate}
-\item Run model parameters to the SUSY-scale (\code{SUSYScale})
+\item Run model parameters to the SUSY-scale (\code{SUSYScale}).
   \begin{enumerate}
-  \item Calculate the \DRbar\ mass spectrum
-  \item Recalculate the SUSY-scale
-  \item Apply the SUSY-scale constraint (\code{SUSYScaleInput})
-  \item Solve the EWSB eqs.\ at the loop level
+  \item Calculate the \DRbar\ mass spectrum.
+  \item Recalculate the SUSY-scale $M_S$ as
+    \begin{align}
+      M_S = \sqrt{m_{\tilde{u}_1}m_{\tilde{u}_6}},
+    \end{align}
+    where $m_{\tilde{u}_1}$ and $m_{\tilde{u}_6}$ are the lightest and
+    heaviest up-type squarks, respectively.
+  \item Impose the SUSY-scale constraint (\code{SUSYScaleInput}).  In
+    the example above, this step is trivial since
+    \code{SUSYScaleInput} is set to be empty.
+  \item Solve the EWSB equations iteratively at the loop level.  In
+    the MSSM example from above leading two-loop corrections have been
+    enabled by setting \code{UseHiggs2LoopMSSM = True;}.  This will
+    add two-loop tadpole contributions to the effective Higgs
+    potential during the EWSB iteration.
   \end{enumerate}
-\item If not converged yet, goto \ref{rge-step-one}
+\item If not converged yet, goto \ref{rge-step-one}.  Otherwise,
+  finish the iteration.
 \end{enumerate}
@@
   \centering
-  \begin{tikzpicture}[node distance = 2cm, auto]
+  \begin{tikzpicture}[node distance = 2.2cm, auto]
     \tikzstyle{block} = [rectangle, draw, text width=16em, text centered, minimum height=3em]
@@
     \node[block,below of=MS] (spec) {Calculate pole masses};
-    \path[arrow,dashed] (MS) -- node[text width=16em] {on convergence\\ run to \code{SUSYScale}} (spec);
+    \path[arrow,dashed] (MS) -- node[text width=16em] {if converged run to \code{SUSYScale}} (spec);
   \end{tikzpicture}

### 2014-05-15 doc/paper.tex
give example output for the default CMSSM parameter point

@@
 %
-An SLHA input file can be given via the \code{--slha-input-file=}
-option, for example:
+When executed, the spectrum generator tries to find a set of \DRbar\
+model parameters consistent with all CMSSM boundary conditions for the
+default parameter point.  Afterwards, the pole mass spectrum and
+mixing matrices are calculated and written to the command line in SLHA
+format \cite{Skands:2003cj,Allanach:2008qq}.  For the default
+parameter point the calculated pole mass spectrum reads
+%
+\begin{lstlisting}
+Block MASS
+   1000021     1.15236980E+03   # Glu
+   1000024     3.85774164E+02   # Cha_1
+   1000037     6.50457634E+02   # Cha_2
+        25     1.14751169E+02   # hh_1
+        35     7.06790308E+02   # hh_2
+        37     7.11386171E+02   # Hpm_2
+        36     7.06520770E+02   # Ah_2
+   1000012     3.51856408E+02   # Sv_1
+   1000014     3.53042587E+02   # Sv_2
+   1000016     3.53046535E+02   # Sv_3
+   1000022     2.03889708E+02   # Chi_1
+   1000023     3.85760548E+02   # Chi_2
+   1000025     6.36542302E+02   # Chi_3
+   1000035     6.50131337E+02   # Chi_4
+   1000001     9.66656167E+02   # Sd_1
+   1000003     1.00983192E+03   # Sd_2
+   1000005     1.01651888E+03   # Sd_3
+   2000001     1.01653020E+03   # Sd_4
+   2000003     1.06089549E+03   # Sd_5
+   2000005     1.06090253E+03   # Sd_6
+   1000011     2.22570345E+02   # Se_1
+   1000013     2.29864547E+02   # Se_2
+   1000015     2.29888857E+02   # Se_3
+   2000011     3.61946702E+02   # Se_4
+   2000013     3.61950896E+02   # Se_5
+   2000015     3.63136044E+02   # Se_6
+   1000002     8.09787908E+02   # Su_1
+   1000004     1.01454208E+03   # Su_2
+   1000006     1.01981123E+03   # Su_3
+   2000002     1.02015283E+03   # Su_4
+   2000004     1.05807774E+03   # Su_5
+   2000006     1.05808183E+03   # Su_6
+\end{lstlisting}
+%
+To calculate the mass spectrum for a user-defined parameter point, the
+CMSSM input parameters can be given in form of a SLHA input file via
+the \code{--slha-input-file=} option, for example:
 %

### 2014-05-16 higgs-study/scan_MSSM_tc.cpp
do TC scan in the MSSMNoFV

@@ -18,8 +18,8 @@
 
 // File generated at Sun 27 Apr 2014 19:19:50
 
-#include "MSSM_input_parameters.hpp"
-#include "MSSM_spectrum_generator.hpp"
+#include "MSSMNoFV_input_parameters.hpp"
+#include "MSSMNoFV_spectrum_generator.hpp"
 
 #include "error.hpp"
 #include "scan.hpp"
@@ -36,15 +36,19 @@ int main(int argc, const char* argv[])
 
    Options options(argc, argv);
 
-   MSSM_input_parameters input;
+   MSSMNoFV_input_parameters input;
    QedQcd oneset;
    oneset.toMz();
 
-   MSSM_spectrum_generator<algorithm_type> spectrum_generator;
+   MSSMNoFV_spectrum_generator<algorithm_type> spectrum_generator;
    spectrum_generator.set_precision_goal(1.0e-3);
    spectrum_generator.set_max_iterations(0);         // 0 == automatic
    spectrum_generator.set_calculate_sm_masses(0);    // 0 == no
    spectrum_generator.set_parameter_output_scale(0); // 0 == susy scale
+
+   const unsigned loop_order = 2;
+   spectrum_generator.set_pole_mass_loop_order(loop_order);
+   spectrum_generator.set_ewsb_loop_order(loop_order);
 
    const std::vector<double> range_TanBeta(
       float_range(options.tanb_start, options.tanb_stop, options.tanb_npoints));
@@ -57,7 +61,7 @@ int main(int argc, const char* argv[])
    input.Azero = 5000.;
    input.SignMu = 1;
 
-   cout << "# MSSM with "
+   cout << "# MSSMNoFV with "
         << "Azero = " << input.Azero
         << ", m0 = " << input.m0
         << ", m12 = " << input.m12
@@ -84,9 +88,9 @@ int main(int argc, const char* argv[])
 
          spectrum_generator.run(oneset, input);
 
-         const MSSM<algorithm_type>& model = spectrum_generator.get_model();
-         const MSSM_physical& pole_masses = model.get_physical();
-         const Problems<MSSM_info::NUMBER_OF_PARTICLES>& problems
+         const MSSMNoFV<algorithm_type>& model = spectrum_generator.get_model();
+         const MSSMNoFV_physical& pole_masses = model.get_physical();
+         const Problems<MSSMNoFV_info::NUMBER_OF_PARTICLES>& problems
             = spectrum_generator.get_problems();
[…]

### 2014-05-16 doc/paper.tex

@@
     3   0                    # calculate SM pole masses
-    4   1                    # pole mass loop order
-    5   1                    # EWSB loop order
+    4   2                    # pole mass loop order
+    5   2                    # EWSB loop order
+    6   2                    # beta-functions loop order
 Block SPhenoInput            # SPheno specific input

### 2014-05-22 doc/paper.tex
doc/paper.tex: add Eigen to list of external routines

@@
          bits\\ {\em External routines:}\/ SARAH 4.0.4, Boost library,
-         lapack\\ {\em
+         Eigen, lapack\\ {\em
            Typical running time:}\/ 0.1-0.3 seconds per parameter

### 2014-05-27 doc/paper.tex
very minor change to english and added comment about B\mu = 0 claim. Wanted to make change when spotted or I'd forget. hope doesn't cause merge problems with Dominik.

@@
 %
-The $\beta$-functions of the SUSY parameters do only depend on the
-SUSY parameters themselfs.  They do not depend on soft-breaking
+The $\beta$-functions of the SUSY parameters only depend on the
+SUSY parameters and are independent of the soft-breaking
 parameters.  However, the $\beta$-functions of the soft-breaking
@@
   the value $1.0$ and its corresponding soft-breaking parameter $B\mu$
-  is set to zero.
+  is set to zero. {\color{red} Is $B\mu$ set to zero, this would be weird and from looking I think it is $A_0$.}
 \item All model parameters are run to the first guess of the low-scale

### 2014-05-28 meta/FlexibleSUSY.m
remove support for a default parameter point

@@ -23,7 +23,6 @@ FSModelName;
 FSLesHouchesList;
 FSUnfixedParameters;
 InputParameters;
-DefaultParameterPoint = {};
 EWSBOutputParameters = {};
 SUSYScale;
 SUSYScaleFirstGuess;
@@ -192,9 +191,6 @@ CheckModelFileSettings[] :=
               FlexibleSUSY`SUSYScaleInput = {};
              ];
 
-           If[Head[FlexibleSUSY`DefaultParameterPoint] =!= List,
-              FlexibleSUSY`DefaultParameterPoint = {};
-             ];
            If[Head[SARAH`MINPAR] =!= List,
               SARAH`MINPAR = {};
              ];
@@ -317,8 +313,9 @@ WriteRGEClass[betaFun_List, anomDim_List, files_List,
 
 WriteInputParameterClass[inputParameters_List, freePhases_List,
                          lesHouchesInputParameters_List,
-                         defaultValues_List, files_List] :=
+                         files_List] :=
    Module[{defineInputParameters, defaultInputParametersInit},
+          defaultValues = {#, 0}& /@ inputParameters;
           defineInputParameters = Constraint`DefineInputParameters[Join[inputParameters,freePhases,lesHouchesInputParameters]];
           defaultInputParametersInit = Constraint`InitializeInputParameters[Join[defaultValues,freePhases,lesHouchesInputParameters]];
           WriteOut`ReplaceInFiles[files,
@@ -1191,7 +1188,6 @@ MakeFlexibleSUSY[OptionsPattern[]] :=
            Print["Creating class for input parameters ..."];
            WriteInputParameterClass[FlexibleSUSY`InputParameters, Complement[freePhases, FlexibleSUSY`InputParameters],
                                     {#[[2]], #[[3]]}& /@ lesHouchesInputParameters,
-                                    FlexibleSUSY`DefaultParameterPoint,
                                     {{FileNameJoin[{Global`$flexiblesusyTemplateDir, "input_parameters.hpp.in"}],
                                       FileNameJoin[{Global`$flexiblesusyOutputDir, FlexibleSUSY`FSModelName <> "_input_parameters.hpp"}]}}
                                    ];

### 2014-05-29 test/test_NMSSM.hpp
fix NMSSM tests

@@ -10,7 +10,7 @@
 using namespace flexiblesusy;
 using namespace softsusy;
 
-void setup_NMSSM(NMSSM<Two_scale>& m, NmssmSoftsusy& s, const NMSSM_input_parameters& input)
+void setup_NMSSM_const(NMSSM<Two_scale>& m, NmssmSoftsusy& s, const NMSSM_input_parameters& input)
 {
    const double ALPHASMZ = 0.1176;
    const double ALPHAMZ = 1.0 / 127.918;
@@ -118,6 +118,18 @@ void setup_NMSSM(NMSSM<Two_scale>& m, NmssmSoftsusy& s, const NMSSM_input_parame
    s.setTanb(tanBeta);
    s.setSvev(vS);
    s.setMw(s.displayMwRun());
+}
+
+void setup_NMSSM(NMSSM<Two_scale>& m, NmssmSoftsusy& s, NMSSM_input_parameters& input)
+{
+   input.m0 = 200.;
+   input.m12 = 200.;
+   input.TanBeta = 10.;
+   input.Azero = -500.;
+   input.LambdaInput = 0.1;
+   input.SignvS = 1;
+
+   setup_NMSSM_const(m, s, input);
 }
 
 void ensure_n_loop_ewsb(NMSSM<Two_scale>& m, int loop_level)

### 2014-05-29 test/test_NMSSM_ewsb.cpp

@@ -71,9 +71,14 @@ BOOST_AUTO_TEST_CASE( test_NMSSM_one_loop_tadpoles )
 {
    NMSSM_input_parameters input;
    input.m0 = 250.; // avoids tree-level tachyons
-   NMSSM<Two_scale> m;
+   input.m12 = 200.;
+   input.TanBeta = 10.;
+   input.Azero = -500.;
+   input.LambdaInput = 0.1;
+   input.SignvS = 1;
+   NMSSM<Two_scale> m(input);
    NmssmSoftsusy s;
-   setup_NMSSM(m, s, input);
+   setup_NMSSM_const(m, s, input);
 
    s.calcDrBarPars();
    m.calculate_DRbar_parameters();
@@ -105,10 +110,15 @@ BOOST_AUTO_TEST_CASE( test_NMSSM_one_loop_ewsb )
 {
    NMSSM_input_parameters input;
    input.m0 = 250.; // avoids tree-level tachyons
-   NMSSM<Two_scale> m;
+   input.m12 = 200.;
+   input.TanBeta = 10.;
+   input.Azero = -500.;
+   input.LambdaInput = 0.1;
+   input.SignvS = 1;
+   NMSSM<Two_scale> m(input);
    NmssmSoftsusy s;
    const double precision = 1.0e-5;
-   setup_NMSSM(m, s, input);
+   setup_NMSSM_const(m, s, input);
 
    // initial guess
    m.set_Kappa(0.1);
@@ -164,9 +174,14 @@ BOOST_AUTO_TEST_CASE( test_NMSSM_two_loop_tadpoles )
 {
    NMSSM_input_parameters input;
    input.m0 = 250.; // avoids tree-level tachyons
-   NMSSM<Two_scale> m;
+   input.m12 = 200.;
+   input.TanBeta = 10.;
+   input.Azero = -500.;
+   input.LambdaInput = 0.1;
+   input.SignvS = 1;
+   NMSSM<Two_scale> m(input);
    NmssmSoftsusy s;
-   setup_NMSSM(m, s, input);
+   setup_NMSSM_const(m, s, input);
 
    s.calcDrBarPars();
    m.calculate_DRbar_parameters();
@@ -210,10 +225,15 @@ BOOST_AUTO_TEST_CASE( test_NMSSM_two_loop_ewsb )
 {
    NMSSM_input_parameters input;
    input.m0 = 300.; // avoids tree-level tachyons
[…]

### 2014-05-29 test/test_NMSSM_one_loop_spectrum.cpp

@@ -84,9 +84,14 @@ BOOST_AUTO_TEST_CASE( test_NMSSM_pole_masses )
 {
    NMSSM_input_parameters input;
    input.m0 = 250.;
+   input.m12 = 200.;
+   input.TanBeta = 10.;
+   input.Azero = -500.;
+   input.LambdaInput = 0.1;
+   input.SignvS = 1;
    NMSSM<Two_scale> m;
    NmssmSoftsusy s;
-   setup_NMSSM(m, s, input);
+   setup_NMSSM_const(m, s, input);
 
    ensure_tree_level_ewsb(m, s, input);
 

### 2014-05-29 test/test_NMSSM_spectrum.cpp

@@ -121,16 +121,16 @@ public:
    void set_low_scale_constraint(NMSSM_low_scale_constraint<Two_scale>* c) { low_constraint = c; }
    void set_susy_scale_constraint(NMSSM_susy_scale_constraint<Two_scale>* c) { susy_constraint = c; }
    void set_high_scale_constraint(NMSSM_high_scale_constraint<Two_scale>* c) { high_constraint = c; }
-   void setup_default_constaints() {
+   void setup_default_constaints(const NMSSM_input_parameters& pp, const QedQcd& oneset) {
       if (!high_constraint)
-         high_constraint = new NMSSM_high_scale_constraint<Two_scale>();
+         high_constraint = new NMSSM_high_scale_constraint<Two_scale>(pp);
       if (!susy_constraint)
-         susy_constraint = new NMSSM_susy_scale_constraint<Two_scale>();
+         susy_constraint = new NMSSM_susy_scale_constraint<Two_scale>(pp);
       if (!low_constraint)
-         low_constraint = new NMSSM_low_scale_constraint<Two_scale>();
+         low_constraint = new NMSSM_low_scale_constraint<Two_scale>(pp, oneset);
    }
    void test(const NMSSM_input_parameters& pp, const QedQcd& oneset = QedQcd()) {
-      setup_default_constaints();
+      setup_default_constaints(pp, oneset);
       high_constraint->set_input_parameters(pp);
       low_constraint->set_input_parameters(pp);
       low_constraint->set_sm_parameters(oneset);
@@ -184,6 +184,13 @@ private:
 BOOST_AUTO_TEST_CASE( test_NMSSM_spectrum )
 {
    NMSSM_input_parameters pp;
+   pp.m0 = 200.;
+   pp.m12 = 200.;
+   pp.TanBeta = 10.;
+   pp.Azero = -500.;
+   pp.LambdaInput = 0.1;
+   pp.SignvS = 1;
+
    const NMSSM_high_scale_constraint<Two_scale> high_constraint(pp);
    const double mxGuess = high_constraint.get_initial_scale_guess();
 

### 2014-05-29 test/test_MSSM.hpp
consistent naming of input parameter setters

@@ -129,7 +129,7 @@ void setup_MSSM(MSSM<Two_scale>& m, MssmSoftsusy& s, MSSM_input_parameters& inpu
    Ye(2,2) = 1.77699 * root2 / (vev * cosBeta);
    mm0 = Sqr(m0) * Eigen::Matrix<double,3,3>::Identity();
 
-   m.set_input(input);
+   m.set_input_parameters(input);
    m.set_scale(scale);
    m.set_loops(1);
    m.set_thresholds(3);

### 2014-05-29 test/test_MSSM_spectrum.cpp

@@ -298,7 +298,7 @@ public:
       mssm.set_thresholds(1);
       mssm.set_ewsb_loop_order(1);
       mssm.set_pole_mass_loop_order(1);
-      mssm.set_input(pp);
+      mssm.set_input_parameters(pp);
       mssm.set_precision(1.0e-4); // == softsusy::TOLERANCE
 
       std::vector<Constraint<Two_scale>*> upward_constraints;

### 2014-05-29 test/test_NMSSM_spectrum.cpp

@@ -148,7 +148,7 @@ public:
       mssm.set_thresholds(1);
       mssm.set_ewsb_loop_order(1);
       mssm.set_pole_mass_loop_order(1);
-      mssm.set_input(pp);
+      mssm.set_input_parameters(pp);
       mssm.set_precision(1.0e-4); // == softsusy::TOLERANCE
 
       std::vector<Constraint<Two_scale>*> upward_constraints;

### 2014-05-29 doc/paper.tex
remove documentation of default parameter point
the SG cannot be run without a SLHA input file

@@
 \begin{lstlisting}[language=bash]
-$ ./models/MSSM/run_MSSM.x
+$ ./models/MSSM/run_MSSM.x \
+    --slha-input-file=model_fies/MSSM/LesHouches.in.MSSM
 \end{lstlisting}%% $
@@
 model parameters consistent with all CMSSM boundary conditions for the
-default parameter point.  Afterwards, the pole mass spectrum and
-mixing matrices are calculated and written to the command line in SLHA
-format \cite{Skands:2003cj,Allanach:2008qq}.  For the default
-parameter point the calculated pole mass spectrum reads
+parameter point given in the SLHA input file
+\code{templates/MSSM/LesHouches.in.MSSM}.  Afterwards, the pole mass
+spectrum and mixing matrices are calculated and written to the command
+line in SLHA format \cite{Skands:2003cj,Allanach:2008qq}.  For the
+parameter point given in the above example the calculated pole mass
+spectrum reads
 %
@@
 Block MASS
-   1000021     1.15236980E+03   # Glu
-   1000024     3.85774164E+02   # Cha_1
-   1000037     6.50457634E+02   # Cha_2
-        25     1.14751169E+02   # hh_1
-        35     7.06790308E+02   # hh_2
-        37     7.11386171E+02   # Hpm_2
-        36     7.06520770E+02   # Ah_2
-   1000012     3.51856408E+02   # Sv_1
-   1000014     3.53042587E+02   # Sv_2
-   1000016     3.53046535E+02   # Sv_3
-   1000022     2.03889708E+02   # Chi_1
-   1000023     3.85760548E+02   # Chi_2
-   1000025     6.36542302E+02   # Chi_3
-   1000035     6.50131337E+02   # Chi_4
-   1000001     9.66656167E+02   # Sd_1
-   1000003     1.00983192E+03   # Sd_2
-   1000005     1.01651888E+03   # Sd_3
-   2000001     1.01653020E+03   # Sd_4
-   2000003     1.06089549E+03   # Sd_5
-   2000005     1.06090253E+03   # Sd_6
-   1000011     2.22570345E+02   # Se_1
-   1000013     2.29864547E+02   # Se_2
-   1000015     2.29888857E+02   # Se_3
-   2000011     3.61946702E+02   # Se_4
-   2000013     3.61950896E+02   # Se_5
-   2000015     3.63136044E+02   # Se_6
-   1000002     8.09787908E+02   # Su_1
-   1000004     1.01454208E+03   # Su_2
-   1000006     1.01981123E+03   # Su_3
-   2000002     1.02015283E+03   # Su_4
-   2000004     1.05807774E+03   # Su_5
-   2000006     1.05808183E+03   # Su_6
-\end{lstlisting}
-%
-To calculate the mass spectrum for a user-defined parameter point, the
-CMSSM input parameters can be given in form of a SLHA input file via
-the \code{--slha-input-file=} option, for example:
-%
-\begin{lstlisting}[language=bash]
-$ ./models/MSSM/run_MSSM.x \
-    --slha-input-file=templates/MSSM/LesHouches.in.MSSM
-\end{lstlisting}%% $
+   1000021     1.15236966E+03   # Glu
+   1000024     3.85774334E+02   # Cha_1
+   1000037     6.50460073E+02   # Cha_2
+        25     1.14766149E+02   # hh_1
+        35     7.06792640E+02   # hh_2
+        37     7.11388516E+02   # Hpm_2
+        36     7.06523105E+02   # Ah_2
+   1000012     3.51856376E+02   # Sv_1
+   1000014     3.53042556E+02   # Sv_2
+   1000016     3.53046504E+02   # Sv_3
+   1000022     2.03889780E+02   # Chi_1
+   1000023     3.85760714E+02   # Chi_2
+   1000025     6.36544884E+02   # Chi_3
+   1000035     6.50133768E+02   # Chi_4
+   1000001     9.66656018E+02   # Sd_1
+   1000003     1.00983181E+03   # Sd_2
+   1000005     1.01651873E+03   # Sd_3
+   2000001     1.01653005E+03   # Sd_4
+   2000003     1.06089534E+03   # Sd_5
+   2000005     1.06090238E+03   # Sd_6
+   1000011     2.22570305E+02   # Se_1
+   1000013     2.29864536E+02   # Se_2
+   1000015     2.29888846E+02   # Se_3
+   2000011     3.61946671E+02   # Se_4
+   2000013     3.61950866E+02   # Se_5
+   2000015     3.63136031E+02   # Se_6
+   1000002     8.09787818E+02   # Su_1
+   1000004     1.01454197E+03   # Su_2
+   1000006     1.01981109E+03   # Su_3
+   2000002     1.02015269E+03   # Su_4
+   2000004     1.05807759E+03   # Su_5
+   2000006     1.05808168E+03   # Su_6
+\end{lstlisting}
 
@@
    {5, Azero}
-};
-
-DefaultParameterPoint = {
-   {m0, 125},
-   {m12, 500},
-   {TanBeta, 10},
-   {Sign[\[Mu]], 1},
-   {Azero, 0}
 };
@@
 SLHA index in the MINPAR or EXTPAR blocks.  The second entry is the
-name of the parameter.  The input parameters can be set to default
-values in the \code{DefaultParameterPoint} list.
+name of the parameter.
 
@@
   EXTPAR = { {61, LambdaInput} };
-      \end{lstlisting}
-     \item Also add this parameter to the DefaultParameterPoint.
-      \begin{lstlisting}
-       DefaultParameterPoint = {
-    {m0, 200},
-    {m12, 200},
-    {TanBeta, 10},
-    {Azero, -500}
-    {LambdaInput, 0.1}
-    };
       \end{lstlisting}

### 2014-05-29 plot_selected.sh
plot E6SSM scans from scan 6

@@ -5,7 +5,7 @@ root -l -b -q higgs-study/plot2d.C+(\"higgs-study/data/scan_5/scan_MSSM.dat\")
 
 # NMSSM
 
-# root -l -b -q higgs-study/plot2d.C+(\"higgs-study/data/scan_5/scan_NMSSM_lambda-0.1-a0-5000.dat\")
+root -l -b -q higgs-study/plot2d.C+(\"higgs-study/data/scan_5/scan_NMSSM_lambda-0.1-a0-5000.dat\")
 root -l -b -q higgs-study/plot2d.C+(\"higgs-study/data/scan_5/scan_NMSSM_lambda-0.2-a0-5000.dat\")
 
 root -l -b -q higgs-study/plot2d.C+(\"higgs-study/data/scan_5/scan_NMSSM_lambda-0.1-a0--5000.dat\")
@@ -19,9 +19,9 @@ root -l -b -q higgs-study/plot2d.C+(\"higgs-study/data/scan_5/scan_phdUMSSM_lamb
 
 # phdE6SSM
 
-root -l -b -q higgs-study/plot2d.C+(\"higgs-study/data/scan_5/scan_phdE6SSM_lambda-0.1_vs-10000_kappa-0.1.dat\")
-root -l -b -q higgs-study/plot2d.C+(\"higgs-study/data/scan_5/scan_phdE6SSM_lambda-0.1_vs-10000_kappa-0.2.dat\")
-root -l -b -q higgs-study/plot2d.C+(\"higgs-study/data/scan_5/scan_phdE6SSM_lambda-0.2_vs-10000_kappa-0.1.dat\")
-root -l -b -q higgs-study/plot2d.C+(\"higgs-study/data/scan_5/scan_phdE6SSM_lambda-0.2_vs-10000_kappa-0.2.dat\")
-root -l -b -q higgs-study/plot2d.C+(\"higgs-study/data/scan_5/scan_phdE6SSM_lambda-0.4_vs-10000_kappa-0.1.dat\")
-root -l -b -q higgs-study/plot2d.C+(\"higgs-study/data/scan_5/scan_phdE6SSM_lambda-0.4_vs-10000_kappa-0.2.dat\")
+root -l -b -q higgs-study/plot2d.C+(\"higgs-study/data/scan_6/scan_phdE6SSM_lambda-0.1_vs-10000_kappa-0.1.dat\")
[…]

### 2014-05-29 doc/paper.tex
remove example models E6SSM and NMSSM from the appendix

@@
 \appendix
-\section{Examples}
-\subsection{NMSSM}
-Detailed steps to create NMSSM model file.
-\begin{enumerate}
-\item Create new directory \code{mkdir myNMSSM}
-\item Move into this new directory \code{cd myNMSSM}
-\item copy the MSSM model file in \code{cp /path/to/SARAH/Models/MSSM/MSSM.m  myNMSSM.m}
-\item Edit myNMSSM.m as follows
-\begin{enumerate}
-\item  Change Model name, author and date
-\item Add singlet superfield to SuperFields
-\begin{lstlisting}
-SuperFields[[8]] = {s, 1, sR,    0, 1,  1, RpM};
-\end{lstlisting}
-\item Modify superpotential
-\begin{lstlisting}
-SuperPotential = Yu u.q.Hu - Yd d.q.Hd - Ye e.l.Hd + \[Lambda] s.Hu.Hd +  \[Kappa]/3 s.s.s ;
-\end{lstlisting}
-\item Add new VEV
-\begin{lstlisting} 
-  DEFINITION[EWSB][VEVs]= 
-  { {SHd0, {vd, 1/Sqrt[2]}, {sigmad, \[ImaginaryI]/Sqrt[2]},{phid,1/Sqrt[2]}},
-    {SHu0, {vu, 1/Sqrt[2]}, {sigmau, \[ImaginaryI]/Sqrt[2]},{phiu,1/Sqrt[2]}},
-    {SsR,  {vs, 1/Sqrt[2]}, {sigmaS, \[ImaginaryI]/Sqrt[2]},{phiS,1/Sqrt[2]}}
-};
-\end{lstlisting}
-\item  Add the singlino to the list of gauge eigenstate dirac fermions
-\begin{lstlisting}
-  DEFINITION[GaugeES][DiracSpinors]={
- .............
-  FS -> {FsR,conj[FsR]}   
-};
-\end{lstlisting}
-\item Edit mass eigenstates to give how the gauge eigenstates mix into mass eigenstates.
-Here we add to the cp even higgs a real singlet, to the CP odd the imaginary part of the singlet and finally to the neutralinos we add the singlino.
-
-\begin{lstlisting}
- DEFINITION[EWSB][MatterSector]= 
-{   ................
-    ................
-     {{phid, phiu, phiS}, {hh, ZH}},      
-     {{sigmad, sigmau, sigmaS}, {Ah, ZA}},
-    ................
-     {{fB, fW0, FHd0, FHu0, FsR}, {L0, ZN}},
-     ...............
-};
-
-\end{lstlisting}
-\end{enumerate}
-\item Now we copy over particles.m \code{cp /path/to/SARAH/Models/MSSM/particles.m  ./}
-\item Edit particles.m as follows
-\begin{enumerate}  
-\item First we add to particleDefinitions[GaugeES]:
-\begin{lstlisting}
- ParticleDefinitions[GaugeES] = {
-      {SsR, { Description -> "Singlet"}},        
-      {FS,   { Description -> "Singlino" }},    
-      ............
-      };
-\end{lstlisting}
-Note that we are brief in information here becasue there is already some information contained in the default particles.m file in the model directory:
-\begin{lstlisting} 
-{{ Description -> "Singlet", 
-                 PDG -> {0},
-                 PDG.IX -> 0,
-                 Width -> 0, 
-                 Mass -> Automatic,
-                 FeynArtsNr -> 98,
-                 LaTeX -> "S",
-                 OutputName -> "s"
-
-    }},
-
-(* -------- Weyl Spinor  ------ *)
-
-  {{ Description -> "Weyl Spinor of Singlino", 
-                 PDG -> 99,
-                 Width -> 0, 
-                 Mass -> Automatic,
-                 FeynArtsNr -> 9,
-                 LaTeX -> "\\tilde{S}",
-                 OutputName -> "s"
-
-    }},
-
-\end{lstlisting}
- \item Now we edit ParticleDefinitions[EWSB].
-Again most states are defined in brief form by a name because they use the default information in the particles.m in the models directory.
-But here we need to over write some of the default options for the cp even and cp odd higgs states:
-\begin{lstlisting}
- {hh   ,  {  Description -> "Higgs", 
-                 PDG -> {25, 35,45},
-                 PDG.IX ->{101000001,101000002,101000003}
- }}, 
-       {Ah   ,  {    Description -> "Pseudo-Scalar Higgs",
-                 PDG -> {0, 36, 46},
-                 PDG.IX ->{0,102000001,102000002  } }},
-
-\end{lstlisting}
-\item To the WeylFermionAndIndermediate we add the Weyl singlino and the real scalar singlet field and imaginary scalar singlet field:  
- \begin{lstlisting}
-  {sigmaS,      {Description -> "Scalar Singlet" }}  ,
-  {phiS,      { Description -> "Pseudo Scalar Singlet"}},
-  {FsR,   { Description -> "Weyl Spinor of Singlino"}},
- \end{lstlisting}
-
-Although the WeylFermionAndIndermediate also contains, 
-\begin{lstlisting}
-{SHd,  { Description -> "Down-Higgs"}},
-       {SHu,  { Description -> "Up-Higgs"}},
-\end{lstlisting}
-\noindent we do not add a singlet version becasue for the singlet there is no distinction between these and the individual SU(2) components, as there are for Hu and Hd.
-\end{enumerate} 
-\item Now we copy over parameters.m \code{cp /path/to/SARAH/Models/MSSM/parameters.m  ./}
-\item Add the new parameters lambda, kappa, Alambda and Akappa to the list.
-\begin{lstlisting}
-{\[Kappa],   {Description -> "Singlet Self-Interaction"}},              
-{T[\[Kappa]],  { Description -> "Softbreaking Singlet Self-Interaction" }}, 
-{\[Lambda],   { Description -> "Singlet-Higgs-Interaction"   }},
-{T[\[Lambda]],  {Description -> "Softbreaking Singlet-Higgs-Interaction"}}, 
-{ms2,       { Description -> "Softbreaking Singlet Mass" }},
-{vS,        { Description -> "Singlet-VEV"}}       
-\end{lstlisting}
-Once again this is simplified from the general case becasue there are default versions containedin parameters.m in the Models directory:
-\begin{lstlisting}
-{{  Description -> "Singlet Self-Interaction",
-              LaTeX -> "\\kappa",
-             Real ->False,
-             Dependence -> None, 
-             Value -> None, 
-             LesHouches -> {NMSSMRUN,2},
-             OutputName-> kap         }},
-
-{{ Description -> "Softbreaking Singlet Self-Interaction",
-               LaTeX -> "T_{\\kappa}",
-              Real -> False,
-             Dependence -> None, 
-             Value -> None, 
-             LesHouches ->{NMSSMRUN,4},
-             OutputName-> Tk         }},
-
- {{ Description -> "Singlet-Higgs-Interaction",
-             LaTeX -> "\\lambda",
-             Real -> False,
-	     Dependence -> None, 
-             Value -> None, 
-             LesHouches -> {NMSSMRUN,1},
-             OutputName-> lam          }},
-
-{{Description -> "Softbreaking Singlet-Higgs-Interaction",
-                LaTeX -> "T_{\\lambda}",
-              Real -> False,
-             Dependence -> None, 
-             Value -> None, 
-             LesHouches ->{NMSSMRUN,3},
-             OutputName-> Tlam         }},    
-{{ Description -> "Softbreaking Singlet Mass", 
-              LaTeX -> "m_S^2",
-             DependenceNum ->  None, 
-             LesHouches -> {NMSSMRUN,10},
-             OutputName-> ms2 }},
-
-{{ Description -> "Singlet-VEV", 
-			 LaTeX -> "v_s",
-             Value -> None, 
-             LesHouches -> {NMSSMRUN,5},
-             OutputName-> vS         }},
-\end{lstlisting}
-\item Navigate to Flexible SUSY directory  \code{cd /path/to/FlexibleSUSY/}
-\item Create Z3NMSSM directory \code{./createmodel --models=Z3NMSSM:myNMSSM}
-\item Configure with the new model \code{./configure --with-models=Z3NMSSM}
-\item The default FlexibleSUSY model file for the MSSM will now be in the models/NMSSM2 directory.  Edit this for the NMSSM as follows:
-  \begin{enumerate} 
-    \item Add new input parameter $\lambda$ to EXTPAR and remove the sign of $\mu$ from MINPAR.
-      \begin{lstlisting}    
-        MINPAR = { {1, m0},
-           {2, m12},
-           {3, TanBeta},
-           {5, Azero} };
-
-  EXTPAR = { {61, LambdaInput} };
-      \end{lstlisting}
-    \item Name the parameters you wish to be output by the EWSB conditions.  Note you should use the names given for them in parameters.m.    A common choice in the NMSSM is:
-\begin{lstlisting}
-EWSBOutputParameters = {vS, \[Kappa], ms2};
-\end{lstlisting}
-which will work for the semi-constrained NMSSM.
-\item choose the Highscale and Highscale first guess, if the default is different from what you wish to use.  
-\begin{lstlisting} 
- HighScale = g1 == g2;
- HighScaleFirstGuess = 2.0 10^16;
-\end{lstlisting}
-\item Add how the parameters found in parameters.m that are to be set at the highscale:
-\begin{lstlisting} 
-HighScaleInput={
-   {T[Ye], Azero*Ye},
-   {T[Yd], Azero*Yd},
-   {T[Yu], Azero*Yu},
-   {mHd2, m0^2},
-   {mHu2, m0^2},
-   {mq2, UNITMATRIX[3] m0^2},
-   {ml2, UNITMATRIX[3] m0^2},
-   {md2, UNITMATRIX[3] m0^2},
-   {mu2, UNITMATRIX[3] m0^2},
-   {me2, UNITMATRIX[3] m0^2},
-   {MassB, m12},
-   {MassWB,m12},
-   {MassG,m12},
-{\[Lambda], LambdaInput},
-   {T[\[Kappa]], Azero \[Kappa]},
-   {T[\[Lambda]], Azero LambdaInput}
-};
-\end{lstlisting}
-\item set lowscale boundary condititions if you want something different from the default.  The defaults are:
-\begin{lstlisting} 
- LowScale = SM[MZ];
- LowScaleFirstGuess = SM[MZ];
-\end{lstlisting}
-\item Change the LowscaleInput if necessary.  In this example we keep the same choice as in the MSSM
-  \begin{lstlisting} 
- LowScaleInput={
-   {vd, 2 MZDRbar / Sqrt[GUTNormalization[g1]^2 g1^2 + g2^2] Cos[ArcTan[TanBeta]]},
-   {vu, 2 MZDRbar / Sqrt[GUTNormalization[g1]^2 g1^2 + g2^2] Sin[ArcTan[TanBeta]]}
-};
-  \end{lstlisting}
-
-Finally we set the initial guesses for the paremeters.  In this example we overwtite the MSSM version replacing it with:
- \begin{lstlisting} 
-InitialGuessAtLowScale = {
-   {vd, SM[vev] Cos[ArcTan[TanBeta]]},
-   {vu, SM[vev] Sin[ArcTan[TanBeta]]},
-   {\[Lambda], LambdaInput},
-   {\[Kappa], 0.1},
-   {vS, 1000},
-   {ms2, SM[MZ]^2}
-};
-
-InitialGuessAtHighScale = {};
- \end{lstlisting}
-
-The code can now be created and also compiled (if you didn't select --disable-compile option when you ran the configure script) by typing \code{make -j2}.
-
- \end{enumerate}
-
-\end{enumerate}   
-\subsection{\ESSM}
-Detailed steps to create \ESSM model file.
-\begin{enumerate}
-\item Create new directory \code{mkdir myE6SSM}
-\item Move into this new directory \code{cd myE6SSM}
-\item copy the MSSM model file in \code{cp /path/to/SARAH/Models/MSSM/MSSM.m  myE6SSM.m}
-\item Edit myE6SSM.m as follows
-\begin{enumerate}
-\item  Change Model name, author and date
-\item Add new Vector superfield,
- \begin{lstlisting} 
-   Gauge[[4]]={Bp,  U[1], Ncharge, g1p, False, RpM};
- \end{lstlisting}
-\item Modify MSSM superfield definitions to include new charges.
-\begin{lstlisting}
-SuperFields[[1]] = {q, 3, {uL,  dL},     1/6, 2, 3, 1, RpM};  
-SuperFields[[2]] = {l, 3, {vL,  eL},    -1/2, 2, 1, 2, RpM};
-SuperFields[[3]] = {Hd,1, {Hd0, Hdm},   -1/2, 2, 1, -3, RpP};
-SuperFields[[4]] = {Hu,1, {Hup, Hu0},    1/2, 2, 1, -2, RpP};
-SuperFields[[5]] = {d, 3, conj[dR],    1/3, 1, -3, 2, RpM};
-SuperFields[[6]] = {u, 3, conj[uR],   -2/3, 1, -3, 1, RpM};
-SuperFields[[7]] = {e, 3, conj[eR],      1, 1,  1, 1, RpM};
-\end{lstlisting}
-Note that in the \ESSM model the charges are fixed to be those of the $U(1)_N$ as shown in e.g.
-\cite{King:2005jy}.
-However here one can leave the charges unassigned vby using parameters in place of the fixed charges.
-Here we fix the charges to simplify the C++ code.
-
-\item Add Singlet Higgs superfield to SuperField
-\begin{lstlisting}
-SuperFields[[8]] = {s, 1, sR,     0, 1,  1, 5, RpP};
-\end{lstlisting}
-\item Add inert Higgs-like superfields 
-\begin{lstlisting}
-SuperFields[[9]] = {H1I, 2, {H1I0, H1Im},  -1/2, 2, 1, -3, RpP};
-SuperFields[[10]] = {H2I, 2, {H2Ip, H2I0},   1/2, 2, 1, -2, RpP};
-SuperFields[[11]] = {sI, 2, sIR,    0, 1,  1, 5, RpM};
-\end{lstlisting}
-\item Add exotics to SuperFields
-\begin{lstlisting}
-SuperFields[[12]] = {Dx, 3, DxL,  -1/3, 1, 3, -2, RpP};
-SuperFields[[13]] = {Dxbar, 3, conj[DxbarR],  1/3, 1, -3, -3, RpP};
-\end{lstlisting}
-\item Add additional survival $SU(2)$ multiplets to SuperFields
-\begin{lstlisting}
-SuperFields[[14]] = {Hp, 1, {Hpd0, Hpdm},  1/2, 2,  1, 2, RpP};
-SuperFields[[15]] = {Hpbar, 1, {Hpup, Hpu0}, -1/2, 2,  1, -2, RpP};
-\end{lstlisting}
-\item Modify superpotential
-\begin{lstlisting}
-SuperPotential = Yu u.q.Hu - Yd d.q.Hd - Ye e.l.Hd + \[Lambda] s.Hu.Hd +  \[Lambda]12 s.H2I.H1I + \[Kappa] s.Dx.Dxbar + \[Mu]Pr Hp.Hpbar;
-\end{lstlisting}
-\item Define new gauge eigenstate Dirac spinors
-\begin{lstlisting}
-  DEFINITION[GaugeES][DiracSpinors]={
-    .............
-    .............
-    FS -> {FsR,conj[FsR]},
-    FBp -> {fBp,conj[fBp]},
-    HI0 -> {FH1I0, conj[FH2I0]},
-    HIC -> {FH1Im, conj[FH2Ip]},
-    FSI -> {FsIR, conj[FsIR]},
-    FDx1 -> {FDxL, 0},
-    FDx2 -> {0, FDxbarR}, 
-    Hp0 -> {FHpd0, conj[FHpu0]},
-    HpC -> {FHpdm, conj[FHpup]}
-  };
-\end{lstlisting}
-\item Add new VEV
-\begin{lstlisting} 
-  DEFINITION[EWSB][VEVs]= 
-  { {SHd0, {vd, 1/Sqrt[2]}, {sigmad, \[ImaginaryI]/Sqrt[2]},{phid,1/Sqrt[2]}},
-    {SHu0, {vu, 1/Sqrt[2]}, {sigmau, \[ImaginaryI]/Sqrt[2]},{phiu,1/Sqrt[2]}},
-    {SsR,  {vs, 1/Sqrt[2]}, {sigmaS, \[ImaginaryI]/Sqrt[2]},{phiS,1/Sqrt[2]}}
-};
-\end{lstlisting}
-\item Add new gauge boson  mass eigenstate
-\begin{lstlisting} 
-  DEFINITION[EWSB][GaugeSector] =
-  { 
-    {{VB,VWB[3],VBp},{VP,VZ,VZp},ZZ},
-    {{VWB[1],VWB[2]},{VWm,conj[VWm]},ZW},
-    {{fWB[1],fWB[2],fWB[3]},{fWm,fWp,fW0},ZfW}
-  }; 
-\end{lstlisting}
-\item Edit mixings for Higgs and neutralino states and add additional states that mix.
-\begin{lstlisting}
-DEFINITION[EWSB][MatterSector]= 
-{ ...............
-  ...............
-{{phid, phiu, phiS}, {hh, ZH}},
-{{sigmad, sigmau, sigmaS}, {Ah, ZA}},
-  ...............
-{{fB, fW0, FHd0, FHu0, FsR, fBp}, {L0, ZN}},
- ................ 
- {{SH1I0,conj[SH2I0]},{SHI0,UHI0}},
- {{SH1Im,conj[SH2Ip]},{SHIp,UHIp}},
- {{SHpd0,conj[SHpu0]},{SHp0,UHp0}},
- {{SHpdm,conj[SHpup]},{SHpp,UHpp}},
- {{FHpd0,FHpu0},{L0p,ZNp}},
- {{FH1I0,FH2I0},{L0I,ZNI}}
-\end{lstlisting}
-\end{enumerate}
-
-\end{enumerate}
-
 \section{Numeric tests}

### 2014-05-29 doc/paper.tex
typo

@@
 For checking the correctness of \fs's generated spectrum generators
-extensive unit tests against Softsusy's MSSM, and NMSSM
+extensive unit tests against Softsusy's MSSM and NMSSM
 implementations (both $Z_3$-invariant and $Z_3$-violating variants)

### 2014-05-30 doc/paper.tex
correcting description of SM fermion mass calculation.
Note: mt is calculated from mt^pole.  The top self-energies miss the
pure QCD corrections, which are added separately.  mb and mtau are
calculated from MS-bar/DR-bar masses, which already contain photon and
QCD contributions.  Therefore in Sigma_{b,tau} these contributions
must be omitted.

@@
     m_{t,\text{susy}}^{\text{\DRbar}}(\mu) &= m_t^\text{pole} +
-    \re\Sigma_{t}^{S,\text{heavy}}(m_t^\text{pole}) \\
+    \re\Sigma_{t}^{S}(m_t^\text{pole}) \\
     &\phantom{=\;} + m_t^\text{pole}
-    \left[ \re\Sigma_{t}^{L,\text{heavy}}(m_t^\text{pole}) +
-      \re\Sigma_{t}^{R,\text{heavy}}(m_t^\text{pole}) + \Delta
+    \left[ \re\Sigma_{t}^{L}(m_t^\text{pole}) +
+      \re\Sigma_{t}^{R}(m_t^\text{pole}) + \Delta
       m_t^{(1),\text{qcd}} + \Delta m_t^{(2),\text{qcd}} \right] ,
@@
 %
-where the $\Sigma_{f}^\text{heavy}$ is the self-energy of fermion $f$
-without the Standard Model contributions.  The labels $L,R,S$ denote
-the left-, right- and non-polarized part of the self-energy.  The
-appearing QCD corrections $\Delta m_t^{(1),\text{qcd}}$ and $\Delta
-m_t^{(2),\text{qcd}}$ are taken from \cite{Bednyakov:2002sf} and read
+where the $\Sigma_{t}$ is the top one-loop self-energy without QCD
+contributions.  The labels $L,R,S$ denote the left-, right- and
+non-polarized part of the self-energy.  The separated QCD corrections
+$\Delta m_t^{(1),\text{qcd}}$ and $\Delta m_t^{(2),\text{qcd}}$ are
+taken from \cite{Bednyakov:2002sf} and read
 %
@@
       \pi} - \frac{23}{72} \frac{\alpha_s^2}{\pi^2} + \frac{3
-      g_2^2}{128 \pi^2} + \frac{13 g_Y^2}{1152 \pi^2}\right) .
-\end{align}
-%
-The \DRbar mass of the $\tau$ is calculated as
+      g_2^2}{128 \pi^2} + \frac{13 g_Y^2}{1152 \pi^2}\right) ,
+\end{align}
+%
+and the \DRbar mass of the $\tau$ is calculated as
 %
@@
 %
-To convert the fermion masses from the \MSbar to the \DRbar scheme the
-Yukawa coupling conversion from \cite[Eq.~(19)]{Skands:2003cj} is used
-and it is assumed that the VEV is defined in the \DRbar scheme.
+In the above equations $\Sigma_{b,\tau}^{\text{heavy}}$ are the
+one-loop self-energies of the bottom and $\tau$, where contributions
+from the gluon and photon are omitted.  To convert the fermion masses
+from the \MSbar to the \DRbar scheme the Yukawa coupling conversion
+from \cite{Martin:1993yx} is used and it is assumed that the VEV is
+defined in the \DRbar scheme.
 

### 2014-05-30 doc/paper.tex
doc/paper.tex: replace --models= by --name= in createmodel command
doc/paper.tex: shorten an execution command

@@
 \begin{lstlisting}[language=bash]
-$ ./createmodel --models=MSSM
+$ ./createmodel --name=MSSM
 $ ./configure --with-models=MSSM
@@
 \begin{lstlisting}[language=bash]
-$ ./models/MSSM/run_MSSM.x \
-    --slha-input-file=model_fies/MSSM/LesHouches.in.MSSM
+$ models/MSSM/run_MSSM.x --slha-input-file=model_fies/MSSM/LesHouches.in.MSSM
 \end{lstlisting}%% $

### 2014-06-01 doc/paper.tex
doc/paper.tex: update address
doc/paper.tex: introduce command \mathematica
Its definition is Mathematica, as is capitalized on
http://www.wolfram.com/mathematica/ .

@@
 \newcommand{\fs}{FlexibleSUSY\xspace}
+\newcommand{\mathematica}{Mathematica\xspace}
 \newcommand{\ESSM}{E$_6$SSM\xspace}
@@
 \author[adelaide]{Peter Athron}
-\author[dresden]{Jae-hyeon Park}
+\author[valencia]{Jae-hyeon Park}
 \author[dresden]{Dominik St\"ockinger}
@@
 Adelaide SA 5005 Australia}
+\address[valencia]{Departament de F\'{i}sica Te\`{o}rica and IFIC,
+Universitat de Val\`{e}ncia-CSIC,
+46100, Burjassot, Spain}
 \address[dresden]{Institut f\"ur Kern- und Teilchenphysik,
@@
   \begin{abstract}
-   We introduce \fs, a MATHEMATICA package which generates a fast,
+   We introduce \fs, a \mathematica package which generates a fast,
    precise C++ spectrum generator for any SUSY model specified by the
@@
 possible and to be easily adaptable to changing goals and new
-ideas. \fs is a MATHEMATICA package which uses \sarah to create a
+ideas. \fs is a \mathematica package which uses \sarah to create a
 fast, modular C++ spectrum generator for a user specified SUSY model.
@@
 \label{Sec:Program}
-\fs is a MATHEMATICA package designed to create a fast and easily
+\fs is a \mathematica package designed to create a fast and easily
 adaptable spectrum generator in C++ for any SUSY model and some
@@
 \begin{itemize}
-\item MATHEMATICA, version 7 or higher
+\item \mathematica, version 7 or higher
 \item SARAH, version 4.0.4 or higher \url{http://sarah.hepforge.org}
@@
 \code{./configure --help} for more options.  Executing \code{make}
-will start MATHEMATICA to generate the spectrum generator and compile
+will start \mathematica to generate the spectrum generator and compile
 it.  The resulting executable can be run like this:
@@
 \label{Sec:modfile}
-A \fs model is configured via a (MATHEMATICA) model file
+A \fs model is configured via a (\mathematica) model file
 \code{FlexibleSUSY.m} in the model directory \code{models/<model>/}.

### 2014-06-01 doc/paper.tex
doc/paper.tex: insert \@ between a capital letter and \xspace
For a correct intersentence space.
doc/paper.tex: add macros
doc/paper.tex: follow up Peter's comment
doc/paper.tex: add section "Flexible Applications"

@@
 %macros
-\newcommand{\sarah}{SARAH\xspace}
-\newcommand{\fs}{FlexibleSUSY\xspace}
+\newcommand{\sarah}{SARAH\@\xspace}
+\newcommand{\fs}{FlexibleSUSY\@\xspace}
 \newcommand{\mathematica}{Mathematica\xspace}
-\newcommand{\ESSM}{E$_6$SSM\xspace}
+\newcommand{\ESSM}{E$_6$SSM\@\xspace}
 \newcommand{\code}[1]{\lstinline|#1|}  % inline source code
@@
 \newcommand{\tabref}[1]{\tablename~\ref{#1}}
+\newcommand{\ptitle}[1]{\emph{#1}}
+\renewcommand{\ptitle}[1]{}
+\newcommand{\JHtxt}[1]{\textcolor[rgb]{1,.3,0}{\texttt{#1}}}
+\newcommand{\JHrem}[1]{\textcolor[rgb]{1,.3,0}{\textbf{{\textsl{\boldmath #1}}}}}
+\makeatletter
+\lstnewenvironment{numlstlisting}[1][]{%
+  \lstset{%
+    #1,
+    numbers=left,
+    firstnumber=auto,
+    numberstyle=\tiny\sffamily}%
+  \csname\@lst @SetFirstNumber\endcsname
+}{%
+  \csname \@lst @SaveFirstNumber\endcsname
+}
+\makeatother
 \DeclareMathOperator{\diag}{diag}
@@
 \subsection{Design goals}
+\label{sec:design goals}
 
@@
   is set to zero. {\color{red} Is $B\mu$ set to zero, this would be weird and from looking I think it is $A_0$.}
+  \JHrem{The C++ code does set $B\mu$ to zero.  Moreover, it sets $\mu$ to +1 GeV irrespective of SignMu.  This should be no problem as long as the solution is found though.}
 \item All model parameters are run to the first guess of the low-scale
@@
 
-\section{Flexible Spectra}
+\section{Flexible Applications}
 \label{Sec:Flexible}
-\subsection{Meta code}
+
+By definition, research is an endeavour to find something new.
+Therefore, it can often be the case that
+an auto-generated spectrum generator executable is not enough.
+\fs attempts to offer a clean interface through which
+one can exploit its facilities
+while undergoing a minimal amount of frustration,
+when he or she programs for a wide variety of studies.
+In what follows,
+this property shall be demonstrated by presenting a few use cases
+at differing levels of complexity.
+
+To avoid confusion,
+it should be mentioned that
+the code snippets presented below are not
+verbatim listings of the files included in the package.
+They have been tailored retaining the semantics
+for conciseness.
+
+\subsection{Adapting model files}
+
+There are simple but interesting goals
+that one can achieve only by working on
+\mathematica files.
+The outcome thus obtained from \fs
+might already include a fully-fledged program
+that is useful in physics analysis.
+In a more advanced project,
+one might utilize the produced libraries as building blocks
+that constitute the target application.
+For a general account of the \fs model files,
+refer to \secref{Sec:modfile}.
+
+\subsubsection{Changing boundary conditions}
+
+As already emphasized in \secref{sec:design goals}, the modular design of
+\fs makes it straightforward % to change a boundary condition.
+% Within the present C++ code structure,
+% this task reduces
+to replace a boundary condition object.
+The question then becomes how
+one could obtain an alternative boundary condition class,
+apart from writing one by hand.
+The meta code feature of \fs offers great assistance
+in this respect.
+An example shall be presented to illustrate how this works.
+
+In the literature,
+there is a popular alternative to the CMSSM boundary condition
+under which the Higgs soft masses are allowed to be different from
+the universal mass of the other scalars \cite{NUHM}.
+One might implement
+this non-universal Higgs-mass MSSM (NUHMSSM) scenario
+simply by modifying the model description given to \fs.
+A section of the \code{FlexibleSUSY.m.in} file is listed below:
+\begin{numlstlisting}
+EXTPAR = {{1, mHd2In}, {2, mHu2In}};
+
+HighScaleInput={
+  {mHd2, mHd2In}, {mHu2, mHu2In},
+  {T[Ye], Azero*Ye}, {T[Yd], Azero*Yd}, {T[Yu], Azero*Yu},
+  {mq2, UNITMATRIX[3] m0^2}, {ml2, UNITMATRIX[3] m0^2}, {md2, UNITMATRIX[3] m0^2},
+  {mu2, UNITMATRIX[3] m0^2}, {me2, UNITMATRIX[3] m0^2},
+  {MassB, m12}, {MassWB, m12}, {MassG, m12}
+};
+\end{numlstlisting}
+Since \code{mHd2} and \code{mHu2} are to be fixed at constants
+different from \code{m0^2},
+two additional input parameters,
+\code{mHd2In} and \code{mHu2In}, holding those constants,
+are introduced in the list \code{EXTPAR}.
+These input parameters are then declared to be the high-scale values of
+\code{mHd2} and \code{mHu2} in line 4.
+The rest of the boundary conditions is the same as in the CMSSM\@.
+In the SLHA input file,
+the parameter indices \code{1} and \code{2} of
+\code{mHd2In} and \code{mHu2In}, declared in \code{EXTPAR} above,
+must appear
+as the first field in each line in the \code{EXTPAR} block:
+\begin{numlstlisting}
+Block EXTPAR
+    1   10000                # mHd2In
+    2   -2500                # mHu2In
+\end{numlstlisting}
+Note that the two additional input parameters are chosen to have
+mass dimension 2, unlike \code{m0}.
+This makes it easy to
+try both signs of the high-scale value of either soft Higgs mass squared,
+as exemplified in line 3.
+If one were not interested in a negative boundary value of \code{mHu2}
+for instance,
+then a dimension-1 parameter %, \code{mHuIn},
+might instead be introduced whose square is equated with \code{mHu2}.
+
+The full implementation is available
+in \code{model_files/NUHMSSM/}.
+To try it out, do the following:
+\begin{numlstlisting}[language=bash]
+$ ./createmodel --name=NUHMSSM --sarah-model=MSSM
+$ ./configure --with-models=NUHMSSM
+$ make
+$ models/NUHMSSM/run_NUHMSSM.x --slha-input-file=model_files/NUHMSSM/LesHouches.in.NUHMSSM
+\end{numlstlisting}
+Notice the \code{--sarah-model=MSSM} flag in line 1.
+It tells the \code{createmodel} script to reuse
+the MSSM specification in \sarah
+to generate the C++ program.
+
+\subsubsection{Extending existing models}
+
+The preceding example was a modest alteration of a physics scenario
+in that an existing model has been reused.
+A more non-trivial modification might involve
+an extension of the particle content as well as the interactions.
+One of the simplest classes of models beyond the MSSM is
+those with additional gauge-single fields.
+In what follows, a supersymmetric type-I see-saw model
+\cite{see-saw}
+shall be considered in which
+three neutral (heavy) chiral superfields are introduced.
+
+% To create a new model, one should name it.
+In the package, this model is named \code{MSSMRHN}, standing for
+the MSSM plus right-handed neutrinos.
+% Since \fs needs the outcome of \sarah,
+One needs to prepare an input file to \sarah which might be placed in
+\code{<FlexibleSUSY-root>/sarah/MSSMRHN/} or
+\code{<SARAH-root>/Models/MSSMRHN/}.
+The input file \code{MSSMRHN.m} contains
+the declaration of the three-generation singlets \code{v}:
+\begin{numlstlisting}[name=MSSMRHN.m]
+SuperFields[[8]] = {v, 3, conj[vR], 0, 1, 1, RpM};
+\end{numlstlisting}
+as well as the neutrino Yukawa couplings and the Majorana mass terms
+of the singlets:
+\begin{numlstlisting}[name=MSSMRHN.m]
+SuperPotential = Yu u.q.Hu - Yd d.q.Hd - Ye e.l.Hd + \[Mu] Hu.Hd +
+                 Yv v.l.Hu + Mv/2 v.v;
+\end{numlstlisting}
+Further declarations inform \sarah
+of how to form Dirac spinors out of the new Weyl spinors
+and how the scalars and the fermions mix to comprise
+the mass eigenstates:
+\begin{numlstlisting}[name=MSSMRHN.m]
+DEFINITION[GaugeES][DiracSpinors] = {
+  Fu1 -> {FuL, 0}, Fu2 -> {0, FuR},
+  Fv1 -> {FvL, 0}, Fv2 -> {0, FvR},
+  ...
+};
+
+DEFINITION[EWSB][MatterSector] = {
+  {{SuL, SuR}, {Su, ZU}},
+  {{SvL, SvR}, {Sv, ZV}},
+  ...
+  {{fB, fW0, FHd0, FHu0}, {L0, ZN}},
+  {{FvL, conj[FvR]}, {FV, UV}},
+  {{{fWm, FHdm}, {fWp, FHup}}, {{Lm, UM}, {Lp, UP}}},
+  {{{FuL}, {conj[FuR]}}, {{FUL, ZUL}, {FUR, ZUR}}}
+};
+
+DEFINITION[EWSB][DiracSpinors] = {
+  Fu  -> {FUL, conj[FUR]},
+  Fv  -> {FV , conj[FV] },
+  Chi -> {L0 , conj[L0] },
+  Cha -> {Lm , conj[Lp] },
+  ...
+};
+\end{numlstlisting}
+With respect to the MSSM file, newly added lines are
+6, 12, 15, and 22.
+Notice that the (left- and right-handed) neutrino mixing
+in line 15 resembles the neutralino mixing in line 14.
+Due to the Majorana mass term in the superpotential,
+the six neutrino mass eigenstates are described in terms of
+Majorana spinors like the neutralinos.
+
+One should then add descriptions of the new states in the file
+\code{particles.m}:
+\begin{numlstlisting}
+ParticleDefinitions[GaugeES] = {
+  {Fv1, { Description -> "Dirac Left Neutrino" }},
+  {Fv2, { Description -> "Dirac Right Neutrino" }},
+  {SvR, { Description -> "Right Sneutrino", LaTeX ->"\\tilde{\\nu}_R"}},
+  ...
+};
+
+ParticleDefinitions[EWSB] = {
+  {Sv, { Description -> "Sneutrinos",
+         PDG -> {1000012, 1000014, 1000016, 2000012, 2000014, 2000016}}},
+  {Fv, { Description -> "Neutrinos",
+         PDG -> {12, 14, 16, 9900012, 9900014, 9900016}}},
+  ...
+};
+
+WeylFermionAndIndermediate = {
+  {v,   { Description -> "Right Neutrino Superfield" }},
+  {FV,  { Description -> "Neutrino-Masseigenstate"}},
+  {FvL, { Description -> "Left Neutrino"}},
+  {FvR, { Description -> "Right Neutrino"}},
+  ...
+};
+\end{numlstlisting}
+In line 12, one finds PDG codes beginning with \code{99}.
+Such numbers are available for a program author's private use
+\cite{Beringer:1900zz}.
+The new parameters in the superpotential
+and the soft supersymmetry breaking sector
+are to be described in \code{parameters.m}:
+\begin{numlstlisting}
+ParameterDefinitions = {
+  {UV,    { Description -> "Neutrino-Mixing-Matrix"}},
+  {Yv,    { Description -> "Neutrino-Yukawa-Coupling" }},
+  {T[Yv], { Description -> "Trilinear-Neutrino-Coupling"}},
+  {Mv,    { LaTeX -> "M_v", OutputName -> Mv, LesHouches -> Mv}},
+  {B[Mv], { LaTeX -> "B_v", OutputName -> BMv, LesHouches -> BMv}},
+  {mv2,   { Description -> "Softbreaking right Sneutrino Mass"}},
+  ...
+};
+\end{numlstlisting}
+For details on how to write model files for \sarah, consult its manual.
+
+Finally, it remains to put \code{FlexibleSUSY.m.in}
+in \code{model_files/MSSMRHN/}.
+The high-scale boundary conditions therein might read:
+\begin{numlstlisting}
+HighScaleInput = {
+  {Mv, LHInput[Mv]}, {B[Mv], LHInput[B[Mv]]},
+  {Yv, LHInput[Yv]}, {T[Yv], Azero*Yv},
+  {mv2, UNITMATRIX[3] m0^2},
+  ...
+};
+\end{numlstlisting}
+In this particular example,
+the new parameters shown in lines 2--3
+are constrained at the $M_X$ scale
+to the values given in the SLHA input file.
+In a data-driven approach,
+they might alternatively be fixed from a matching against
+a low-energy theory containing the dimension-5
+neutrino mass operator in conjunction with supplementary constraints.
+
+With the above set of input files,
+\fs can generate the C++ code and compile it
+to produce \code{libMSSMRHN}.
+These products shall be employed as the implementation of the theory that
+underlies the MSSM in the next subsection.
+
 \subsection{Adapting C++ code}
-\subsection{Combing with  modules of your own code}
+
+Consider a physics scenario which is best described
+by a tower of effective theories.
+Within the framework of \fs, the C++ object structure is
+a faithful reflection of
+this physicist's view on the given problem.
+This point might be illustrated
+using a well-known configuration in which
+the higher-energy theory is the MSSMRHN discussed above which
+gives rise to the MSSM as the lower-energy effective theory.
+The relevant objects are sketched
+in \figref{fig:class structure for tower}.
+\begin{figure}
+  \centering
+\begin{tikzpicture}
+  \draw (0,-.5) coordinate(bottom)
+        (0,-.4)
+     -- (0,0)  coordinate(MZ) node[left] {$M_Z$}
+     -- (0,.7) coordinate(MS) node[left] {$M_S$}
+     -- (0,2)  coordinate(Mv) node[left] {$M_\nu$}
+     -- (0,3)  coordinate(MX) node[left] {$M_X$}
+     -- (0,3.4)
+        (0,3.5) coordinate(top);
+
+  \path (MZ) ++(.3,0) coordinate(MZR) +(3,0) coordinate(MZRR)
+        (MS) ++(.3,0) coordinate(MSR) +(3,0) coordinate(MSRR)
+        (Mv) ++(.3,0) coordinate(MvR) +(3,0) coordinate(MvRR)
+        (MX) ++(.3,0) coordinate(MXR) +(3,0) coordinate(MXRR);
+
+  \draw (MZ) +(-.1,0) -- +(+.1,0); \draw[<-] (MZ) ++(-.8,0) -- ++(-.5,0)
+     node[left,draw] {\code{MSSM_low_scale_constraint}};
+  \draw (MS) +(-.1,0) -- +(+.1,0); \draw[<-] (MS) ++(-.8,0) -- ++(-.5,0)
+     node[left,draw] {\code{MSSM_susy_scale_constraint}};
+  \draw (Mv) +(-.1,0) -- +(+.1,0); \draw[<-] (Mv) ++(-.8,0) -- ++(-.5,0)
+     node[left,draw] {\code{MSSM_MSSMRHN_matching}};
+  \draw (MX) +(-.1,0) -- +(+.1,0); \draw[<-] (MX) ++(-.8,0) -- ++(-.5,0)
+     node[left,draw] {\code{MSSMRHN_high_scale_constraint}};
+
+  \draw (MvRR) +(0,-.03) rectangle (MZR) node[pos=.5] {\code{MSSM}};
+  \draw (MvRR) +(0, .03) rectangle (MXR) node[pos=.5] {\code{MSSMRHN}};
+
+  \path (bottom) +(3.8,0) coordinate(SE) (top) +(-6.8,0) coordinate(NW);
+  \draw  (NW) +(0,.5) rectangle (SE)
+         (NW) -- +(10.6,0);
+  \path (top) node[above] {\code{RGFlow}};
+\end{tikzpicture}
+  \caption{Schematic object structure in the C++ code for
+    the tower scenario.}
+  \label{fig:class structure for tower}
+\end{figure}
+The \code{MSSMRHN} object is in effect from the $M_X$ scale down to
+the $M_\nu$ scale at which the right-handed neutrinos are decoupled.
+Below this scale, the \code{MSSM} object takes over.
+On the left of the vertical axis,
+the boundary condition objects acting on either model are displayed,
+together with the matching object connecting the two theories.
+Note that each of the boundary condition and matching objects
+maintains and updates its own scale over iterations.
+An arrow in the figure depicts
+the association of a constraint with its scale.
+All these components are plugged into the \code{RGFlow} object
+which then solves the problem.
+
+Since the current implementation of \fs does not (yet)
+produce a spectrum generator spanning multiple models,
+one needs to write the matching object
+% in \figref{fig:class structure for tower}
+as well as gluing codes by hand to build such a program.
+Nevertheless, the clean object structure eases this job.
+
+One could best see the overall code structure in the \code{Makefile}:
+\begin{numlstlisting}
+CPPFLAGS  := -I. $(INCCONFIG) $(INCFLEXI) $(INCLEGACY) $(INCSLHAEA) \
+             $(INCMSSM) $(INCMSSMRHN)
+
+TOWER_SRC := run_tower.cpp \
+	     MSSM_MSSMRHN_two_scale_matching.cpp \
+	     MSSM_MSSMRHN_two_scale_initial_guesser.cpp
+
+TOWER_OBJ := $(patsubst %.cpp, %.o, $(filter %.cpp, $(TOWER_SRC)))
+
+run_tower.x: $(TOWER_OBJ) $(LIBMSSM) $(LIBMSSMRHN) $(LIBFLEXI) $(LIBLEGACY)
+  $(CXX) -o $@ $^ $(LOOPFUNCLIBS) $(GSLLIBS) $(BOOSTTHREADLIBS) $(THREADLIBS) $(LAPACKLIBS) $(FLIBS)
+\end{numlstlisting}
+This file is created in \code{examples/tower/}
+on the execution of the \code{configure} script.
+The include directives in line 2 tell the compiler
+where to find the headers for either \code{MSSM} or \code{MSSMRHN}.
+The \code{.cpp} files in lines 4--6 and the \code{.hpp} files
+that they include are to be written by hand.
+Obviously, the executable \code{run_tower.x}, in line 10, depends on both
+\code{$(LIBMSSM)} and \code{$(LIBMSSMRHN)}
+that implement the auto-generated
+components in \figref{fig:class structure for tower}.
+
+To prepare the main source file \code{run_tower.cpp},
+one can make a straightforward extension to
+\code{run_MSSM.cpp} or \code{run_MSSMRHN.cpp}
+produced in either model directory.
+The shipped example reads:
+\begin{numlstlisting}
+#include "MSSM_MSSMRHN_spectrum_generator.hpp"
+
+int main(int argc, char* argv[])
+{
+  // define objects;
+  QedQcd oneset;
+  MSSM_input_parameters    input_1;
+  MSSMRHN_input_parameters input_2;
+  // fill in input_1 and input_2;
+  oneset.toMz(); // run SM fermion masses to MZ
+  typedef Two_scale algorithm_type;
+  MSSM_MSSMRHN_spectrum_generator<algorithm_type> spectrum_generator;
+  // set up spectrum_generator;
+  spectrum_generator.run(oneset, input_1, input_2);
+  // extract outcome from models;
+}
+\end{numlstlisting}
+where a line in the form \code{// ...;} shall be understood
+to be a pseudo-code.
+Given two models,
+one declares two sets of input parameters,
+\code{input_1} and \code{input_2}, in lines 7--8.
+The suffixes \code{_1} and \code{_2} are not necessary but intended to remind
+the programmer of to which model the object is relevant.
+
+The \code{main} function then delegates
+the \code{MSSM_MSSMRHN_spectrum_generator} object
+to set up and drive the \code{RGFlow} object in
+\figref{fig:class structure for tower}.
+This task is started by the call in line 14 to
+the following member function defined in
+\code{MSSM_MSSMRHN_spectrum_generator.hpp}:
+\begin{numlstlisting}[name=SGrun]
+template<class T> void MSSM_MSSMRHN_spectrum_generator<T>::run
+(const QedQcd& oneset,
+ const MSSM_input_parameters& input_1, const MSSMRHN_input_parameters& input_2)
+{
+  high_scale_constraint_2.clear(); // of type MSSMRHN_high_scale_constraint<T>
+  susy_scale_constraint_1.clear(); // of type MSSM_susy_scale_constraint<T>
+  low_scale_constraint_1 .clear(); // of type MSSM_low_scale_constraint<T>
+  matching.reset();                // of type MSSM_MSSMRHN_matching<T>
+  high_scale_constraint_2.set_input_parameters(input_2);
+  susy_scale_constraint_1.set_input_parameters(input_1);
+  low_scale_constraint_1 .set_input_parameters(input_1);
+  matching.set_upper_input_parameters(input_2);
+  high_scale_constraint_2.initialize();
+  susy_scale_constraint_1.initialize();
+  low_scale_constraint_1 .initialize();
+  if (!is_zero(input_scale_2)) high_scale_constraint_2.set_scale(input_scale_2);
+\end{numlstlisting}
+This piece of code is nearly a verbatim copy of
+the corresponding part of \code{MSSM_spectrum_generator.hpp}.
+The only differences are that the type of
+\code{high_scale_constraint_2} is
+\code{MSSMRHN_high_scale_constraint<T>} and that
+the \code{matching} object has been added.
+Recall that the template parameter \code{T} has been bound to
+\code{Two_scale} in the \code{main} function.
+One then constructs a list of
+the constraints on \code{MSSM}:
+\begin{numlstlisting}[name=SGrun]
+  std::vector<Constraint<T>*> upward_constraints_1;
+  upward_constraints_1.push_back(&low_scale_constraint_1);
+  std::vector<Constraint<T>*> downward_constraints_1;
+  downward_constraints_1.push_back(&susy_scale_constraint_1);
+  downward_constraints_1.push_back(&low_scale_constraint_1);
+\end{numlstlisting}
+and initializes the \code{MSSM} object:
+\begin{numlstlisting}[name=SGrun]
+  model_1.clear();                 // of type MSSM<T>
+  model_1.set_input_parameters(input_1);
+  model_1.do_calculate_sm_pole_masses(calculate_sm_masses);
+\end{numlstlisting}
+Likewise for \code{MSSMRHN}:
+\begin{numlstlisting}[name=SGrun]
+  std::vector<Constraint<T>*> upward_constraints_2;
+  upward_constraints_2.push_back(&high_scale_constraint_2);
+  std::vector<Constraint<T>*> downward_constraints_2;
+  downward_constraints_2.push_back(&high_scale_constraint_2);
+  model_2.clear();                 // of type MSSMRHN<T>
+  model_2.set_input_parameters(input_2);
+\end{numlstlisting}
+Note that \code{model_2} does not have to calculate the pole masses of
+the SM particles since it is active only above $M_\nu$
+which is assumed to be much higher than the weak scale.
+To test the convergence of both models,
+one may construct a composite convergence tester
+out of auto-generated
+\code{MSSM_convergence_tester} and \code{MSSMRHN_convergence_tester}:
+\begin{numlstlisting}[name=SGrun]
+  MSSM_convergence_tester<T> convergence_tester_1(&model_1, precision_goal);
+  MSSMRHN_convergence_tester<T> convergence_tester_2(&model_2, precision_goal);
+  if (max_iterations > 0) {
+    convergence_tester_1.set_max_iterations(max_iterations);
+    convergence_tester_2.set_max_iterations(max_iterations);
+  }
+  Composite_convergence_tester<T> convergence_tester;
+  convergence_tester.add_convergence_tester(&convergence_tester_1);
+  convergence_tester.add_convergence_tester(&convergence_tester_2);
+\end{numlstlisting}
+On construction,
+the initial guesser accepts the following parameters
+including the two model objects:
+\begin{numlstlisting}[name=SGrun]
+  MSSM_MSSMRHN_initial_guesser<T> initial_guesser
+    (&model_1, &model_2, input_2, oneset,
+     low_scale_constraint_1, susy_scale_constraint_1, high_scale_constraint_2,
+     matching);
+\end{numlstlisting}
+The code of the above class shall be presented later on.
+One then passes
+\code{convergence_tester} and \code{initial_guesser}
+to \code{solver}, the \code{RGFlow} object,
+along with the precision specification:
+\begin{numlstlisting}[name=SGrun]
+  Two_scale_increasing_precision precision(10.0, precision_goal);
+  solver.reset();                  // of type RGFlow<T>
+  solver.set_convergence_tester(&convergence_tester);
+  solver.set_running_precision(&precision);
+  solver.set_initial_guesser(&initial_guesser);
+\end{numlstlisting}
+Finally,
+one is ready to construct the tower of effective theories
+by adding to \code{solver}
+each model plus the associated list of constraints
+optionally accompanied by a matching object:
+\begin{numlstlisting}[name=SGrun]
+  solver.add_model(&model_1, &matching, upward_constraints_1, downward_constraints_1);
+  solver.add_model(&model_2, upward_constraints_2, downward_constraints_2);
+\end{numlstlisting}
+The order of addition is from the lowest scale to the highest.
+Notice in line 49
+that the matching object between \code{model_1} and
+\code{model_2} is given when one adds the former, i.e.\ the lower-energy model.
+It then remains to solve the boundary value problem:
+\begin{numlstlisting}[name=SGrun]
+  high_scale_2 = susy_scale_1 = low_scale_1 = 0; matching_scale = 0;
+  solver.solve();
+\end{numlstlisting}
+After the solution is found,
+one can obtain the resulting low-energy spectrum.
+Since \code{model_1} is in contact with the lowest energy,
+let it calculate the spectrum:
+\begin{numlstlisting}[name=SGrun]
+  susy_scale_1 = susy_scale_constraint_1.get_scale();
+  model_1.run_to(susy_scale_1);    // of type MSSM<T>
+  model_1.calculate_spectrum();
+  if (!is_zero(parameter_output_scale_1))
+    model_1.run_to(parameter_output_scale_1);
+}
+\end{numlstlisting}
+In lines 56--57,
+the scale is optionally brought to the value at which
+one wishes to get the \DRbar\ parameters.
+
+One needs to write the matching class for a particular pair of
+models from scratch.
+It shall be based on the abstract class \code{Matching<Two_scale>}
+that comes with \fs.
+In the present example, the class is declared in the header
+\code{MSSM_MSSMRHN_two_scale_matching.hpp}:
+\begin{numlstlisting}
+template<> class MSSM_MSSMRHN_matching<Two_scale> : public Matching<Two_scale> {
+public:
+  MSSM_MSSMRHN_matching();
+  MSSM_MSSMRHN_matching(const MSSMRHN_input_parameters&);
+  void match_low_to_high_scale_model();
+  void match_high_to_low_scale_model();
+  double get_scale() const;
+  void set_models(Two_scale_model *lower, Two_scale_model *upper);
+  double get_initial_scale_guess() const;
+  void set_upper_input_parameters(const MSSMRHN_input_parameters&);
+  void set_scale(double);
+  void reset();
+private:
+  MSSM   <Two_scale> *lower;
+  MSSMRHN<Two_scale> *upper;
+  void make_initial_scale_guess();
+  void update_scale();
+  ...
+};
+\end{numlstlisting}
+As lines 4 and 10 indicate,
+this class takes an \code{MSSMRHN_input_parameters} object as input.
+The \code{MvInput} field thereof is referenced
+for the initial guess of the matching scale
+in line 25 of \code{MSSM_MSSMRHN_two_scale_matching.cpp}:
+\begin{numlstlisting}
+void MSSM_MSSMRHN_matching<Two_scale>::match_low_to_high_scale_model()
+{
+  upper->set_Yd(lower->get_Yd());
+  // copy rest of couplings from lower to upper;
+  upper->set_scale(lower->get_scale());
+}
+
+void MSSM_MSSMRHN_matching<Two_scale>::match_high_to_low_scale_model()
+{
+  update_scale();
+  lower->set_Yd(upper->get_Yd());
+  // copy rest of couplings from upper to lower;
+  lower->set_scale(upper->get_scale());
+}
+
+void MSSM_MSSMRHN_matching<Two_scale>::set_upper_input_parameters
+(const MSSMRHN_input_parameters& inputPars_)
+{
+  inputPars = inputPars_;
+  make_initial_scale_guess();
+}
+
+void MSSM_MSSMRHN_matching<Two_scale>::make_initial_scale_guess()
+{
+  double RHN_scale = pow(abs(inputPars.MvInput.determinant()), 1.0/3);
+  scale = initial_scale_guess = RHN_scale;
+}
+
+void MSSM_MSSMRHN_matching<Two_scale>::update_scale()
+{
+  double RHN_scale = pow(abs(upper->get_Mv().determinant()), 1.0/3);
+  scale = RHN_scale;
+}
+\end{numlstlisting}
+Recall that \code{MvInput} is the high-scale boundary value of \code{Mv}.
+If one had opted for an alternative strategy to fix \code{Mv},
+then \code{MSSM_MSSMRHN_matching} might have required a different input.
+Subsequently, the matching scale is updated at each iteration
+to be the geometric mean of the running \code{Mv} eigenvalues
+in lines 31--32.
+The actual matching process takes place in the two functions
+\code{match_low_to_high_scale_model} and \code{match_high_to_low_scale_model}.
+For brevity,
+their examples shown above simply copy each coupling
+upwards or downwards, neglecting threshold corrections.
+The way to incorporate these corrections should be self-evident
+from the code structure.
+
+% \begin{numlstlisting}
+% template<>
+% class MSSM_MSSMRHN_initial_guesser<Two_scale> : public Initial_guesser<Two_scale> {
+% public:
+%   MSSM_MSSMRHN_initial_guesser(MSSM<Two_scale>*, MSSMRHN<Two_scale>*,
+%                                const MSSMRHN_input_parameters&,
+%                                const QedQcd&,
+%                                const MSSM_low_scale_constraint<Two_scale>&,
+%                                const MSSM_susy_scale_constraint<Two_scale>&,
+%                                const MSSMRHN_high_scale_constraint<Two_scale>&,
+%                                const MSSM_MSSMRHN_matching<Two_scale>&);
+%   virtual ~MSSM_MSSMRHN_initial_guesser();
+%   virtual void guess();
+%
+% private:
+%   MSSM   <Two_scale>* model_1;
+%   MSSMRHN<Two_scale>* model_2;
+%   MSSMRHN_input_parameters input_pars;
+%   QedQcd oneset;
+%   MSSM_low_scale_constraint<Two_scale> low_constraint;
+%   MSSM_susy_scale_constraint<Two_scale> susy_constraint;
+%   MSSMRHN_high_scale_constraint<Two_scale> high_constraint;
+%   MSSM_MSSMRHN_matching<Two_scale> matching;
+% };
+% \end{numlstlisting}
+The last missing piece is the initial guesser.
+One can extend the already available
+\code{MSSM_initial_guesser} class in a straightforward manner.
+The essential task is done by the following member function:
+\begin{numlstlisting}[name=MSSM_MSSMRHN_two_scale_initial_guesser.cpp]
+void MSSM_MSSMRHN_initial_guesser<Two_scale>::guess()
+{
+  // guess SUSY couplings in model-1 at low energy;
+
+  const double low_scale_guess_1 = low_constraint_1.get_initial_scale_guess();
+  const double high_scale_guess_2 = high_constraint_2.get_initial_scale_guess();
+  const double matching_scale_guess = matching.get_initial_scale_guess();
+\end{numlstlisting}
+Compared to the MSSM case,
+the differences are that
+the type of \code{high_constraint_2}
+is \code{MSSMRHN_high_scale_constraint<Two_scale>} and that
+\code{matching_scale_guess} has been inserted.
+Due to this intermediate scale,
+the initial run-up is divided into two steps,
+with a matching procedure in-between:
+\begin{numlstlisting}[name=MSSM_MSSMRHN_two_scale_initial_guesser.cpp]
+  model_1->run_to(matching_scale_guess); // of type MSSM<Two_scale>
+  matching.set_models(model_1, model_2);
+  matching.match_low_to_high_scale_model();
+  model_2->run_to(high_scale_guess_2);   // of type MSSMRHN<Two_scale>
+\end{numlstlisting}
+The high-scale constraints are applied to \code{model_2},
+the higher-energy model,
+and the remaining undetermined parameters are guessed:
+\begin{numlstlisting}[name=MSSM_MSSMRHN_two_scale_initial_guesser.cpp]
+  high_constraint_2.set_model(model_2);
+  high_constraint_2.apply();
+  model_2->set_Mu(1.0); model_2->set_BMu(0.0);
+\end{numlstlisting}
+The initial two-step run-down again involves a matching process:
+\begin{numlstlisting}[name=MSSM_MSSMRHN_two_scale_initial_guesser.cpp]
+  model_2->run_to(matching_scale_guess);
+  matching.match_high_to_low_scale_model();
+  model_1->run_to(low_scale_guess_1);
+\end{numlstlisting}
+At the low scale where \code{MSSM} is valid,
+the code is the same as in \code{MSSM_initial_guesser}:
+\begin{numlstlisting}[name=MSSM_MSSMRHN_two_scale_initial_guesser.cpp]
+  model_1->solve_ewsb_tree_level();
+  model_1->calculate_DRbar_parameters();
+  model_1->set_thresholds(3); model_1->set_loops(2);
+}
+\end{numlstlisting}
+
+Finally,
+one prescribes the additional input parameters in the SLHA input file:
+\begin{numlstlisting}
+Block YvIN                   # neutrino Yukawas at MX
+1 1  0.1568611               # Yv(1,1)
+1 2  0.6400513               # Yv(1,2)
+1 3  0.7521494               # Yv(1,3)
+2 1  0.4663838               # Yv(2,1)
+...                          # remaining 5 entries
+Block MvIN                   # heavy neutrino masses at MX
+1 1  4.150000E+14            # Mv(1,1)
+...                          # remaining 8 entries
+Block BMvIN                  # right-handed sneutrino bilinear terms
+1 1  1.000000E+02            # BMv(1,1)
+...                          # remaining 8 entries
+\end{numlstlisting}
+The file in the package contains the values that approximately reproduce
+the observed neutrino masses and mixing angles
+\cite{Beringer:1900zz}
+through the see-saw mechanism.
+
+For further details, browse the directory \code{examples/tower/}.
+One can build and run the example therein by:
+\begin{numlstlisting}
+$ ./createmodel --name=MSSM
+$ ./createmodel --name=MSSMRHN
+$ ./configure --with-models=MSSM,MSSMRHN
+$ make
+$ cd examples/tower
+$ make
+$ ./run_tower.x --slha-input-file=LesHouches.in.tower
+\end{numlstlisting}
+In the output,
+a part of significant physical interest might be:
+\begin{numlstlisting}
+Block MSL2 Q=   8.97114431E+02
+  1  1     1.25652698E+05   # ml2(1,1)
+  1  2    -7.64196328E+01   # ml2(1,2)
+  1  3    -7.08429254E+01   # ml2(1,3)
+  2  1    -7.64196328E+01   # ml2(2,1)
+  2  2     1.25388732E+05   # ml2(2,2)
+  2  3    -3.15865242E+02   # ml2(2,3)
+  3  1    -7.08429254E+01   # ml2(3,1)
+  3  2    -3.15865242E+02   # ml2(3,2)
+  3  3     1.24556532E+05   # ml2(3,3)
+\end{numlstlisting}
+This result demonstrates the well-known effect on
+the off-diagonal slepton mass matrix elements
+from the non-trivial flavour structure of \code{Yv}
+\cite{Borzumati:1986qx}.
+This leads in turn to the slepton mixing
+matrices, \code{ZE} and \code{ZV},
+which contain intergenerational mixings apart from the
+generic left-right mixings.
 
@@
 
+\bibitem{NUHM}
+%\cite{Berezinsky:1995cj}
+%\bibitem{Berezinsky:1995cj}
+  V.~Berezinsky, A.~Bottino, J.~R.~Ellis, N.~Fornengo, G.~Mignola and S.~Scopel,
+  %``Neutralino dark matter in supersymmetric models with nonuniversal scalar mass terms,''
+  Astropart.\ Phys.\  {\bf 5} (1996) 1
+  [hep-ph/9508249];
+  %%CITATION = HEP-PH/9508249;%%
+  %176 citations counted in INSPIRE as of 27 May 2014
+%\cite{Nath:1997qm}
+%\bibitem{Nath:1997qm}
+  P.~Nath and R.~L.~Arnowitt,
+  %``Nonuniversal soft SUSY breaking and dark matter,''
+  Phys.\ Rev.\ D {\bf 56} (1997) 2820
+  [hep-ph/9701301];
+  %%CITATION = HEP-PH/9701301;%%
+  %196 citations counted in INSPIRE as of 27 May 2014
+%\cite{Bottino:2000jx}
+%\bibitem{Bottino:2000jx}
+  A.~Bottino, F.~Donato, N.~Fornengo and S.~Scopel,
+  %``Probing the supersymmetric parameter space by WIMP direct detection,''
+  Phys.\ Rev.\ D {\bf 63} (2001) 125003
+  [hep-ph/0010203];
+  %%CITATION = HEP-PH/0010203;%%
+  %167 citations counted in INSPIRE as of 27 May 2014
+%\cite{Bertin:2002sq}
+%\bibitem{Bertin:2002sq}
+  V.~Bertin, E.~Nezri and J.~Orloff,
+  %``Neutralino dark matter beyond CMSSM universality,''
+  JHEP {\bf 0302} (2003) 046
+  [hep-ph/0210034];
+  %%CITATION = HEP-PH/0210034;%%
+  %84 citations counted in INSPIRE as of 27 May 2014
+%
+%\cite{Drees:1996pk}
+%\bibitem{Drees:1996pk}
+  M.~Drees, M.~M.~Nojiri, D.~P.~Roy and Y.~Yamada,
+  %``Light Higgsino dark matter,''
+  Phys.\ Rev.\ D {\bf 56} (1997) 276
+   [Erratum-ibid.\ D {\bf 64} (2001) 039901]
+  [hep-ph/9701219];
+  %%CITATION = HEP-PH/9701219;%%
+  %142 citations counted in INSPIRE as of 27 May 2014
+%\cite{Drees:2000he}
+%\bibitem{Drees:2000he}
+  M.~Drees, Y.~G.~Kim, M.~M.~Nojiri, D.~Toya, K.~Hasuko and T.~Kobayashi,
+  %``Scrutinizing LSP dark matter at the CERN LHC,''
+  Phys.\ Rev.\ D {\bf 63} (2001) 035008
+  [hep-ph/0007202];
+  %%CITATION = HEP-PH/0007202;%%
+  %87 citations counted in INSPIRE as of 27 May 2014
+%
+%\cite{Ellis:1998jk}
+%\bibitem{Ellis:1998jk}
+  J.~R.~Ellis, T.~Falk, G.~Ganis, K.~A.~Olive and M.~Schmitt,
+  %``Charginos and neutralinos in the light of radiative corrections: Sealing the fate of Higgsino dark matter,''
+  Phys.\ Rev.\ D {\bf 58} (1998) 095002
+  [hep-ph/9801445];
+  %%CITATION = HEP-PH/9801445;%%
+  %121 citations counted in INSPIRE as of 27 May 2014
+%\cite{Ellis:2000we}
+%\bibitem{Ellis:2000we}
+  J.~R.~Ellis, T.~Falk, G.~Ganis and K.~A.~Olive,
+  %``Supersymmetric dark matter in the light of LEP and the Tevatron collider,''
+  Phys.\ Rev.\ D {\bf 62} (2000) 075010
+  [hep-ph/0004169];
+  %%CITATION = HEP-PH/0004169;%%
+  %138 citations counted in INSPIRE as of 27 May 2014
+%
+%\cite{Ellis:2002wv}
+%\bibitem{Ellis:2002wv}
+  J.~R.~Ellis, K.~A.~Olive and Y.~Santoso,
+  %``The MSSM parameter space with nonuniversal Higgs masses,''
+  Phys.\ Lett.\ B {\bf 539} (2002) 107
+  [hep-ph/0204192];
+  %%CITATION = HEP-PH/0204192;%%
+  %164 citations counted in INSPIRE as of 27 May 2014
+%\cite{Ellis:2002iu}
+%\bibitem{Ellis:2002iu}
+  J.~R.~Ellis, T.~Falk, K.~A.~Olive and Y.~Santoso,
+  %``Exploration of the MSSM with nonuniversal Higgs masses,''
+  Nucl.\ Phys.\ B {\bf 652} (2003) 259
+  [hep-ph/0210205].
+  %%CITATION = HEP-PH/0210205;%%
+  %204 citations counted in INSPIRE as of 27 May 2014
+
+\bibitem{see-saw}
+%\bibitem{Minkowski:1977sc}
+  P.~Minkowski, \ptitle{
+  $\mu \rightarrow e \gamma$ at a Rate of One Out of 1-Billion Muon Decays?,}
+  Phys.\ Lett.\ B {\bf 67} (1977) 421;
+  %%CITATION = PHLTA,B67,421;%%
+  %1696 citations counted in INSPIRE as of 12 Nov 2013
+%\bibitem{Yanagida:1979as}
+  T.~Yanagida, \ptitle{
+  Horizontal Symmetry And Masses Of Neutrinos,}
+  Proc.\ of the
+  Workshop on Unified Theories and the Baryon Number of the Universe,
+  edited by O.~Sawada and A.~Sugamoto, KEK, Japan (1979) 95;
+  %%CITATION = CONFP,C7902131,95;%%
+  %459 citations counted in INSPIRE as of 12 Nov 2013
+%\bibitem{GellMann:1980vs}
+  M.~Gell-Mann, P.~Ramond and R.~Slansky, \ptitle{
+  Complex Spinors and Unified Theories,}
+  Supergravity, edited by F.~Nieuwenhuizen and
+  D.~Friedman, North Holland, Amsterdam (1979) 315
+  [arXiv:1306.4669 [hep-th]];
+  %%CITATION = ARXIV:1306.4669;%%
+  %1456 citations counted in INSPIRE as of 12 Nov 2013
+%\cite{Mohapatra:1979ia}
+%\bibitem{Mohapatra:1979ia}
+  R.~N.~Mohapatra and G.~Senjanovic, \ptitle{
+  Neutrino Mass and Spontaneous Parity Nonconservation,}
+  Phys.\ Rev.\ Lett.\  {\bf 44} (1980) 912.
+  %%CITATION = PRLTA,44,912;%%
+  %3254 citations counted in INSPIRE as of 29 Nov 2013
+
+\bibitem{Borzumati:1986qx}
+  F.~Borzumati and A.~Masiero, \ptitle{
+Large Muon- and electron-Number Nonconservation in Supergravity Theories,}
+  Phys.\ Rev.\ Lett.\  {\bf 57} (1986) 961.
+  %%CITATION = PRLTA,57,961;%%
+  %582 citations counted in INSPIRE as of 28 Oct 2013
+
 %\cite{Hall:1980kf}

### 2014-06-02 doc/paper.tex
doc/paper.tex: refer to Sec:SpecGenStruct in section "Adapting C++ code"

@@
 Nevertheless, the clean object structure eases this job.
+One can author nearly all the components of a multi-model spectrum generator
+by making a straightforward extension
+to each corresponding single-model counterpart
+for one of the models forming the tower.
+For this, it should help to have working knowledge about the basic structure
+of a spectrum generator, set out in \secref{Sec:SpecGenStruct}.
+Obviously, this tip does not work for the matching class,
+which has no analogue in a single-model case.
 

### 2014-06-02 doc/paper.tex
doc/paper.tex: reduce repetition of "straightforward"
doc/paper.tex: change "auto-generated" -> "out of the box"
Since the NUHMSSM executable is also auto-generated to some extent.

@@
 Therefore, it can often be the case that
-an auto-generated spectrum generator executable is not enough.
+a spectrum generator right out of the box is not enough.
 \fs attempts to offer a clean interface through which
@@
 To prepare the main source file \code{run_tower.cpp},
-one can make a straightforward extension to
+one can extend % make a straightforward extension to
 \code{run_MSSM.cpp} or \code{run_MSSMRHN.cpp}
@@
 One can extend the already available
-\code{MSSM_initial_guesser} class in a straightforward manner.
+\code{MSSM_initial_guesser} class. % in a straightforward manner.
 The essential task is done by the following member function:

### 2014-06-02 doc/paper.tex
set pole mass calculation precision in the model file

@@
 PotentialLSPParticles = { Chi, Cha, Glu, Sv, Su, Sd, Se };
+
+DefaultPoleMassPrecision = MediumPrecision;
+HighPoleMassPrecision    = {hh, Ah, Hpm};
+MediumPoleMassPrecision  = {};
+LowPoleMassPrecision     = {};
 \end{lstlisting}
@@
 must be used.  \fs provides three different methods with different
-precision and execution speed to determine the mass eigenvalues:
-
+precision and execution speed to determine the mass eigenvalues.
+These methods are set in the model file and work as follows:
+%
 \begin{itemize}
-\item \code{LowDiagonalizationPrecision}: This option provides the
+\item \code{LowPoleMassPrecision}: This option provides the
   lowest precision but is also the fastest one.  Here the one-loop
@@
 
-\item \code{MediumDiagonalizationPrecision} (default): This option
+\item \code{MediumPoleMassPrecision} (default): This option
   provides calculation with medium precision with a medium execution
@@
 
-\item \code{HighDiagonalizationPrecision}: This option provides
+\item \code{HighPoleMassPrecision}: This option provides
   diagonalization with highest precision, but has also the highest
@@
   $m_{f,1L}^\text{high}$ is diagonalized $n$ times, as in the case of
-  \code{MediumDiagonalizationPrecision}, resulting in $n$ pole masses
+  \code{MediumPoleMassPrecision}, resulting in $n$ pole masses
   $m_{f_k}^\pole$ ($k = 1,\ldots,n$).  Afterwards, the diagonalization

### 2014-06-02 meta/FlexibleSUSY.m

@@ -54,6 +54,12 @@ UseHiggs2LoopNMSSM;
 EffectiveMu;
 PotentialLSPParticles = {};
 
+(* precision of pole mass calculation *)
+DefaultPoleMassPrecision = MediumPrecision;
+HighPoleMassPrecision    = {SARAH`HiggsBoson, SARAH`PseudoScalar, SARAH`ChargedHiggs};
+MediumPoleMassPrecision  = {};
+LowPoleMassPrecision     = {};
+
 FSEigenstates;
 FSSolveEWSBTimeConstraint = 120;
 FSSimplifyBetaFunctionsTimeConstraint = 120;
@@ -901,8 +907,8 @@ PrepareUnrotatedParticles[eigenstates_] :=
            TreeMasses`SetUnrotatedParticles[nonMixedParticles];
           ];
 
-ReadDiagonalizationPrecisions[defaultPrecision_Symbol, highPrecisionList_List,
-                              mediumPrecisionList_List, lowPrecisionList_List, eigenstates_] :=
+ReadPoleMassPrecisions[defaultPrecision_Symbol, highPrecisionList_List,
+                       mediumPrecisionList_List, lowPrecisionList_List, eigenstates_] :=
     Module[{particles, particle, i, precisionList = {}},
            If[!MemberQ[{LowPrecision, MediumPrecision, HighPrecision}, defaultPrecision],
               Print["Error: ", defaultPrecision, " is not a valid",
@@ -946,10 +952,6 @@ Options[MakeFlexibleSUSY] :=
     {
         Eigenstates -> SARAH`EWSB,
         InputFile -> "FlexibleSUSY.m",
-        DefaultDiagonalizationPrecision -> MediumPrecision,
-        HighDiagonalizationPrecision -> {},
-        MediumDiagonalizationPrecision -> {},
-        LowDiagonalizationPrecision -> {},
         SolveEWSBTimeConstraint -> 120 (* in seconds *)
     };
 
@@ -1307,11 +1309,11 @@ MakeFlexibleSUSY[OptionsPattern[]] :=
            phases = ConvertSarahPhases[SARAH`ParticlePhases];
 
            (* determin diagonalization precision for each particle *)
-           diagonalizationPrecision = ReadDiagonalizationPrecisions[
-               OptionValue[DefaultDiagonalizationPrecision],
-               Flatten[{OptionValue[HighDiagonalizationPrecision]}],
-               Flatten[{OptionValue[MediumDiagonalizationPrecision]}],
[…]

### 2014-06-02 meta/FlexibleSUSY.m
set eigenstates and time constraints in the model file

@@ -60,7 +60,7 @@ HighPoleMassPrecision    = {SARAH`HiggsBoson, SARAH`PseudoScalar, SARAH`ChargedH
 MediumPoleMassPrecision  = {};
 LowPoleMassPrecision     = {};
 
-FSEigenstates;
+FSEigenstates = SARAH`EWSB;
 FSSolveEWSBTimeConstraint = 120;
 FSSimplifyBetaFunctionsTimeConstraint = 120;
 FSSolveWeinbergAngleTimeConstraint = 120;
@@ -950,9 +950,7 @@ FindUnfixedParameters[fixed_List] :=
 
 Options[MakeFlexibleSUSY] :=
     {
-        Eigenstates -> SARAH`EWSB,
-        InputFile -> "FlexibleSUSY.m",
-        SolveEWSBTimeConstraint -> 120 (* in seconds *)
+        InputFile -> "FlexibleSUSY.m"
     };
 
 MakeFlexibleSUSY[OptionsPattern[]] :=
@@ -971,8 +969,6 @@ MakeFlexibleSUSY[OptionsPattern[]] :=
               Quit[1];
              ];
            CheckSARAHVersion[];
-           FSEigenstates = OptionValue[Eigenstates];
-           FSSolveEWSBTimeConstraint = OptionValue[SolveEWSBTimeConstraint];
            (* load model file *)
            LoadModelFile[OptionValue[InputFile]];
            Print["FlexibleSUSY model file loaded"];

### 2014-06-03 doc/paper.tex
fix typos in the introduction and add ref. for SuseFlav

@@
 space-time symmetries of the Poincar\'e
-group\cite{Coleman:1967ad,Haag:1974qh}, leading many to suspect that
+group \cite{Coleman:1967ad,Haag:1974qh}, leading many to suspect that
 SUSY may be realised in nature in some form. In particular
@@
 at the TeV scale have been proposed to solve the hierarchy
-problem\cite{Weinberg:1975gm, Weinberg:1979bn, Gildener:1976ai,
+problem \cite{Weinberg:1975gm, Weinberg:1979bn, Gildener:1976ai,
   Susskind:1978ms, 'tHooft:1980xb}, allow gauge coupling
-unification\cite{Langacker:1990jh, Ellis:1990wk, Amaldi:1991cn,
+unification \cite{Langacker:1990jh, Ellis:1990wk, Amaldi:1991cn,
   Langacker:1991an, Giunti:1991ta} and predict a dark matter candidate
 which can fit the observed relic
-density\cite{Goldberg:1983nd,Ellis:1983ew}.  Such models have also
+density \cite{Goldberg:1983nd,Ellis:1983ew}.  Such models have also
 been used for baryogenesis or leptogensis to solve the
@@
 has been greatly aided by public spectrum generators for
-MSSM\cite{Allanach:2001kg,Porod:2003um,Djouadi:2002ze,Baer:1993ae},
+MSSM \cite{Allanach:2001kg,Porod:2003um,Djouadi:2002ze,Baer:1993ae,Chowdhury:2011zr},
 allowing fast and reliable exploration of the sparticle spectrum,
@@
 The generated code structure is designed to be as flexible as possible
-to accommodate different types of extensions and due to it's modular
+to accommodate different types of extensions and due to its modular
 nature it is easy to modify, add new features and combine with other
 programs.  The generated code has been extensively tested against well
-known spectrum generators and as well providing a solution for new
-SUSY models the generated MSSM and NMSSM codes offer a modern and fast
+known spectrum generators. As well as providing a solution for new
+SUSY models, the generated MSSM and NMSSM codes offer a modern and fast
 alternative to the existing public spectrum generators.
@@
 
+%\cite{Chowdhury:2011zr}
+\bibitem{Chowdhury:2011zr}
+  D.~Chowdhury, R.~Garani and S.~KVempati,
+  %``SUSEFLAV: Program for supersymmetric mass spectra with seesaw mechanism and rare lepton flavor violating decays,''
+  Comput.\ Phys.\ Commun.\  {\bf 184} (2013) 899
+  [arXiv:1109.3551 [hep-ph]].
+  %%CITATION = ARXIV:1109.3551;%%
+  %16 citations counted in INSPIRE as of 03 Jun 2014
+
 \end{thebibliography}

### 2014-06-03 doc/paper.tex
doc/paper.tex: add to text on lapack

@@
          bits\\ {\em External routines:}\/ SARAH 4.0.4, Boost library,
-         Eigen, lapack\\ {\em
+         Eigen, LAPACK\\ {\em
            Typical running time:}\/ 0.1-0.3 seconds per parameter
@@
 SOFTSUSY\cite{Allanach:2001kg}, the very fast Eigen library
-\cite{eigen}, as well as the gnu scientific library and the boost
+\cite{eigen} augmented by LAPACK,
+as well as the gnu scientific library and the boost
 library to create numerical routines which solve the RGEs and EWSB
@@
 \item GNU scientific library \url{http://www.gnu.org/software/gsl/}
-\item lapack \url{http://www.netlib.org/lapack/}
+\item an implementation of LAPACK \url{http://www.netlib.org/lapack/}
+  such as ATLAS \url{http://math-atlas.sourceforge.net/} or
+  Intel Math Kernel Library \url{http://software.intel.com/intel-mkl}
 \end{itemize}
@@
 \code{Eigen::JacobiSVD} for matrices with less than four rows and
-columns, and the Lapack routines \code{zgesvd}, \code{dgesvd} for
+columns, and the LAPACK routines \code{zgesvd}, \code{dgesvd} for
 larger matrices.  For the other types of diagonalization,

### 2014-06-03 doc/paper.tex
Improvements suggested by Dominik

@@
 \label{Sec:Program}
+
+To study the properties of SUSY models programs are needed which
+numerically calculate the pole masses and couplings of the SUSY
+particles given a set of theory input parameters.  The output of these
+so-called spectrum generators can be transferred to programs which
+calculate further observables such as branching rations or the dark
+matter relic density.
+
+In order to create a spectrum generator the SUSY model must be defined
+by specifying the gauge group, the field content and mixings as well
+as the superpotential and the soft-breaking terms.  From this
+information the renormalization group equations, mass matrices,
+self-energies, tadpole diagrams and electroweak symmetry breaking
+conditions have to be derived.  These expressions must then be
+combined in a computer program to allow for a numeric calculation of
+the mass spectrum.  In addition most SUSY models require boundary
+conditions for the model parameters at a low and a high scale.  For
+example in the CMSSM mSUGRA boundary conditions for the soft-breaking
+parameters are imposed at the gauge coupling unification scale.
+Furthemore, at the $Z$ mass scale the CMSSM is matched to the Standard
+Model, which implies conditions for the gauge and Yukawa couplings.
+The so defined boundary value problem must be solved numerically until
+a set of model parameters has been found consistent with all
+user-supplied boundary conditions.
+
 \fs is a \mathematica package designed to create a fast and easily
@@
 superfield content, superpotential, gauge symmetries and mass mixings
-in the \sarah model files, examples of which can be found in the
-\code{Models} directory of \sarah and the \code{sarah} directory of
-FlexibleSUSY.  The user then also specifies the boundary conditions on
-the model parameters in a separate \code{FlexibleSUSY.m} file,
-examples of which can be found in the \code{model_files} directory. A
-detailed guide on how to write the FlexibleSUSY.m model is given in
-section \ref{Sec:modfile}.
-
-\fs makes use of the existing \sarah package
+in the \sarah model files.  Furthermore, the boundary conditions on
+the model parameters must be specified in a separate
+\code{FlexibleSUSY.m} file.
+%
+Based on this information \fs uses the existing \sarah package
 \cite{Staub:2010ty,Staub:2009bi,Staub:2010jh,Staub:2012pb,Staub:2013tta}
@@
 symmetry breaking conditions, one-loop self energies, one-loop
-tadpoles corrections and two-loop renormalisation group
-equations(RGEs) for the model specified in the model files. Additional
-corrections which have been calculated elsewhere, such as two-loop
-corrections to the Higgs masses\footnote{By default FlexibleSUSY has
-  two-loop corrections to the Higgs masses for the
-  MSSM\cite{Degrassi:2001yf,Brignole:2001jy,Dedes:2002dy,Brignole:2002bz,Dedes:2003km}
+tadpoles corrections and two-loop renormalisation group equations
+(RGEs) for the model.  Additional corrections which have been
+calculated elsewhere, such as two-loop corrections to the Higgs
+masses\footnote{By default FlexibleSUSY has two-loop corrections to
+  the Higgs masses for the MSSM
+  \cite{Degrassi:2001yf,Brignole:2001jy,Dedes:2002dy,Brignole:2002bz,Dedes:2003km}
   and NMSSM \cite{Degrassi:2009yq} in FORTRAN files supplied by Pietro
@@
 specified by the user.
-
-\fs then converts these algerbraic expressions from \sarah into C++
-code using the C++11 standard. \fs uses some parts of
-SOFTSUSY\cite{Allanach:2001kg}, the very fast Eigen library
-\cite{eigen} augmented by LAPACK,
-as well as the gnu scientific library and the boost
-library to create numerical routines which solve the RGEs and EWSB
-conditions.  The main solver routine of \fs then calls these routines
-to find simultaneous solutions for the EWSB, low energy data, and
-other, user supplied, boundary conditions (e.g.\ the usual mSUGRA
-universality conditions) and calculates the mass spectrum.
-
-The modular structure of the generated code allows for any individual
-components to be replaced with an alternative if available.  Indeed
-\fs has been intentionally designed to grow as alternative solvers and
-calculators are added and a subsequent release with an alternative
-lattice solver is already planned.  In addition if one already has an
-existing code for some other purpose and simply wishes to improve upon
-this by adding, e.g., new RGEs or self energies the modular framework
-makes this straightforward by adding or interfacing with the
-appropriate routine.
-
-\subsection{Design goals}
-\label{sec:design goals}
-
-FlexibleSUSY is designed with the following points in mind:
+%
+These algerbraic expressions are converted into C++ code and are put
+into classes with well-defined interfaces to allow for easy exchange,
+extension and reuse of the modules.  All of these classes are finally
+combined to a complete spectrum generator, which solves the
+user-defined boundary value problem.  For this task \fs uses some
+parts of SOFTSUSY \cite{Allanach:2001kg}, the very fast Eigen library
+\cite{eigen}, augmented by LAPACK, as well as the GNU scientific
+library and the BOOST library to create numerical routines which solve
+the RGEs and boundary conditions simultaneously.  If a solution has
+been found the pole mass spectrum is eventually calculated using full
+one-loop self-energies (and leading two-loop Higgs self-energy
+contributions for the MSSM and NMSSM).
+
+\subsection*{Design goals}
+
+Since the calculation of the pole mass spectrum in a SUSY model is a
+non-trivial task, \fs is designed with the following points in mind:
 
@@
 \paragraph{Modularity}
+
 The large variety of supersymmetric models makes it likely that the
-user wants to modify the generated spectrum generator source code.
-\fs uses C++ object orientation features to modularize the source code
-in order to make it easy for the user to modify, reuse and extend the
-spectrum generator.  An important application of this concept are the
-boundary conditions: The RG solver provides a common \code{Constraint}
-interface, see \figref{fig:schematic-two-scale-constraint-interface}.
-%
-\begin{figure}
-  \centering
-  \tikzumlset{fill class=white}
-  \begin{tikzpicture}
-    \umlclass[x=0, y=0]{RGFlow<Two\_scale>}{}{
-      + solve()
-    }
-    \umlclass[x=6, y=0, type=abstract]{Constraint<Two\_scale>}{}{
-      + \umlvirt{apply()}\\
-      + \umlvirt{get\_scale() : double}
-    }
-    \umlclass[x=6, y=-3]{MyConstraint}{}{
-      + apply()\\
-      + get\_scale() : double
-    }
-    \umldep{RGFlow<Two\_scale>}{Constraint<Two\_scale>}
-    \umlinherit{MyConstraint}{Constraint<Two\_scale>}
-  \end{tikzpicture}
-  \caption{Schematic two-scale constraint interface}
-  \label{fig:schematic-two-scale-constraint-interface}
-\end{figure}
-%
-To create a custom constraint \code{MyConstraint}, the user implements
-the interface by inheriting from \code{Constraint} and writing the
-\code{apply()} and \code{get_scale()} methods (see
-\secref{sec:boundary-conditions} for more details).  The so created
-constraint class can then be passed to the RG solver to impose the
-user-defined boundary condition at the chosen scale.
+user wants to modify the generated spectrum generator source code or
+reuses componets in his own program. \fs uses C++ object orientation
+features to modularize the source code in order to make it easy for
+the user to modify, reuse, replace or extend the individual
+components.  An important application of this concept are the boundary
+conditions: The boundary conditions solver class provides a plugin
+mechanism via a common \code{Constraint} interface, which allows a
+user to exchange or add boundary conditions at any scale.
+Alternatively, if one already has an existing code for some other
+purpose and simply wishes to improve upon this by adding, e.g., new
+RGEs or self-energies the modular framework makes this straightforward
+by adding or interfacing with the appropriate routine.  A possible
+application of reusing the \fs' generated classes would be to further
+improve the precision of the CE$_6$SSM spectrum generator presended in
+\cite{Athron:2009bs,Athron:2012pw} by adding further one-loop
+self-energies and two-loop $\beta$-functions.
+
+\paragraph{Alternative boundary condition problem solvers}
+
+Furthermore, the standard algorithm which solves the user-defined
+boundary value problem via a fixed-point iteration is not guaranteed
+to converge in all regions of the model parameter space.  Therefore,
+\fs has been intentionally designed to allow for alternative solvers
+to search for solutions in such critical parameter regions.  A
+subsequent release with an alternative lattice solver is already
+planned.
+
+\paragraph{Towers of effective theories}
+
+In \fs the standard fixed-point iteration solver has been generalized
+to handle towers of models (effective theories), which are matched at
+intermediate scales.  An example of such a tower construction will be
+given in section \ref{sec:adapting-cpp-code}, where a right-handed
+neutrino is integrated out at the see-saw scale, between the SUSY and
+the GUT scale.
 
@@
 \subsection{Adapting C++ code}
+\label{sec:adapting-cpp-code}
 
@@
 
+%\cite{Athron:2009bs}
+\bibitem{Athron:2009bs}
+  P.~Athron, S.~F.~King, D.~J.~Miller, S.~Moretti and R.~Nevzorov,
+  %``The Constrained Exceptional Supersymmetric Standard Model,''
+  Phys.\ Rev.\ D {\bf 80} (2009) 035009
+  [arXiv:0904.2169 [hep-ph]].
+  %%CITATION = ARXIV:0904.2169;%%
+  %45 citations counted in INSPIRE as of 03 Jun 2014
+
+%\cite{Athron:2012pw}
+\bibitem{Athron:2012pw}
+  P.~Athron, D.~Stockinger and A.~Voigt,
+  %``Threshold Corrections in the Exceptional Supersymmetric Standard Model,''
+  Phys.\ Rev.\ D {\bf 86} (2012) 095012
+  [arXiv:1209.1470 [hep-ph]].
+  %%CITATION = ARXIV:1209.1470;%%
+  %5 citations counted in INSPIRE as of 03 Jun 2014
+
 \end{thebibliography}

### 2014-06-03 doc/paper.tex
extend section 4 a bit and give more quick start examples

@@
 
-To create a MSSM spectrum generator the following three commands:
+\fs can be downloaded as a gzipped tar file from the
+\url{http://flexiblesusy.hepforge.org/}.  To download and install
+version 1.0.0 run:
+%
+\begin{lstlisting}[language=bash]
+$ wget https://www.hepforge.org/archive/flexiblesusy/FlexibleSUSY-1.0.0.tar.gz
+$ tar -xf FlexibleSUSY-1.0.0.tar.gz
+$ cd FlexibleSUSY-1.0.0/
+\end{lstlisting}%% $
+%
+A CMSSM spectrum generator can be created with the following three
+commands:
 %
@@
 The first command creates the model directory \code{models/MSSM/}
-together with a MSSM model file.  The \code{configure} scrip checks
+together with a CMSSM model file.  The \code{configure} script checks
 the system requirements and creates the \code{Makefile}.  See
@@
 parameter point given in the SLHA input file
-\code{templates/MSSM/LesHouches.in.MSSM}.  Afterwards, the pole mass
+\code{model_files/MSSM/LesHouches.in.MSSM}.  Afterwards, the pole mass
 spectrum and mixing matrices are calculated and written to the command
@@
    2000006     1.05808168E+03   # Su_6
+\end{lstlisting}
+
+\subsection{Alternative models}
+
+\fs already comes with plenty of predefined models: the CMSSM (simply
+called \code{MSSM}), the $Z_3$-symmetric NMSSM (called \code{NMSSM}),
+$Z_3$-violating NMSSM (\code{SMSSM}), the USSM (\code{UMSSM}), the
+\ESSM (\code{E6SSM}), the right-handed neutrino extended MSSM
+(\code{MSSMRHN}) and the NUHM-MSSM (\code{NUHMSSM}).  See the content
+of \code{model_files/} for all predefined model files.  A spectrum
+generator for the $Z_3$-symmetric NMSSM for example can be generated
+like this:
+%
+\begin{lstlisting}[language=bash]
+$ ./createmodel --name=NMSSM
+$ ./configure --with-models=NMSSM
+$ make
+\end{lstlisting}%% $
+%
+This NMSSM variant unifies all soft-breaking trilinear scalar
+couplings at the GUT scale.  In order to relax this constraint and use
+a separate value for $A_\lambda$ at the GUT scale one can edit the
+model file \code{models/NMSSM/FlexibleSUSY.m} and change the lines
+%
+\begin{lstlisting}[language=Mathematica]
+EXTPAR = { {61, LambdaInput} };
+
+HighScaleInput = {
+   ...
+   {T[\[Lambda]], Azero LambdaInput},
+   ...
+};
+\end{lstlisting}
+%
+into
+%
+\begin{lstlisting}[language=Mathematica]
+EXTPAR = { {61, LambdaInput},
+           {63, ALambdaInput}
+         };
+
+HighScaleInput = {
+   ...
+   {T[\[Lambda]], ALambdaInput},
+   ...
+};
+\end{lstlisting}
+%
+The value of $A_\lambda$ at the GUT scale can be set in the SLHA input
+file in \code{EXTPAR} block entry $63$ via
+%
+\begin{lstlisting}
+Block EXTPAR
+   61   0.1                  # LambdaInput
+   63   -100                 # ALambdaInput
 \end{lstlisting}

### 2014-06-03 doc/paper.tex
use as GUT condition TLambda = Alambda * lambda
use proper shell command
typo in the path to the SLHA file
Addressing Dominik's suggestions on section 5 (not finished yet)

@@
 self-energies, tadpole diagrams and electroweak symmetry breaking
-conditions have to be derived.  These expressions must then be
+(EWSB) conditions have to be derived.  These expressions must then be
 combined in a computer program to allow for a numeric calculation of
@@
 \begin{lstlisting}[language=bash]
-$ models/MSSM/run_MSSM.x --slha-input-file=model_fies/MSSM/LesHouches.in.MSSM
+$ ./models/MSSM/run_MSSM.x \
+    --slha-input-file=model_files/MSSM/LesHouches.in.MSSM
 \end{lstlisting}%% $
@@
    ...
-   {T[\[Lambda]], ALambdaInput},
+   {T[\[Lambda]], ALambdaInput LambdaInput},
    ...
@@
 \label{Sec:modfile}
-A \fs model is configured via a (\mathematica) model file
-\code{FlexibleSUSY.m} in the model directory \code{models/<model>/}.
-Many example model files can be found in \code{model_files/}.  The
-default model file for the CMSSM looks like this:
+
+In general a (non-constrained) softly broken SUSY model is defined by
+the gauge group, the field content and mixings as well as the
+superpotential and the soft-breaking Lagrangian.  In order to create a
+spectrum generator for such a SUSY model with \fs, the aforementioned
+model properties have to be defined in a SARAH model file.  The SARAH
+model file can be put into the \code{sarah/<model>/} directory.  See
+the SARAH manual \cite{Staub:2008uz,Staub:2013tta} for a detailed
+explanation of how to write such a model file.
+
+The model boundary conditions are defined in the \fs model file
+\code{FlexibleSUSY.m}, which has to be located in the model directory
+\code{models/<model>/}.  Note, that many example model files can
+already be found in \code{model_files/}.  In the following it is
+explained how the boundary conditions are defined on the basis of the
+CMSSM.  The application to other models is straightforward.  The CMSSM
+model file reads:
 %
@@
    {Ye, Automatic},
-   {vd, 2 MZDRbar / Sqrt[GUTNormalization[g1]^2 g1^2 + g2^2] Cos[ArcTan[TanBeta]]},
-   {vu, 2 MZDRbar / Sqrt[GUTNormalization[g1]^2 g1^2 + g2^2] Sin[ArcTan[TanBeta]]}
+   {vd, 2 MZDRbar / Sqrt[GUTNormalization[g1]^2 g1^2 + g2^2]
+           Cos[ArcTan[TanBeta]]},
+   {vu, 2 MZDRbar / Sqrt[GUTNormalization[g1]^2 g1^2 + g2^2]
+           Sin[ArcTan[TanBeta]]}
 };
@@
 All non-Standard Model input variables must be specified in the lists
-\code{MINPAR} and \code{EXTPAR}.  Theses two variables refer to the
+\code{MINPAR} and \code{EXTPAR}.  These two variables refer to the
 MINPAR and EXTPAR blocks in a SLHA input file \cite{Skands:2003cj}.
 The list elements are two-component lists where the first entry is the
-SLHA index in the MINPAR or EXTPAR blocks.  The second entry is the
-name of the parameter.
+SLHA index in the MINPAR or EXTPAR block, respectively.  The second
+entry is the name of the input parameter.  In the above example the
+input parameters are the universal soft-breaking parameters $m_0$,
+$M_{1/2}$, $A_0$ as well as $\tan\beta$ and $\sign\mu$.
 
@@
 specify the model parameters that are output of the electroweak
-symmetry breaking conditions.  When doing EWSB, \fs will adjust these
-parameters until the EWSB conditons are fulfilled.  In the CMSSM these
-are the superpotential parameter \code{\\[Mu]} and its corresponding
-soft-breaking parameter \code{B[\\[Mu]]}.
-
-Furthermore, the user has to specify the form of three model
-constraints: low-scale, SUSY-scale and high-scale constraints.  In \fs
-they are named as \code{LowScale}, \code{SUSYScale} and
-\code{HighScale}.  For each constraint there is (i) a scale definition
-(named after the constraint), (ii) an initial guess for the scale
-(concatenation of the constraint name and \code{FirstGuess}) and (iii)
-a list of parameter settings to be applied at the scale (concatenation
-of the constraint name and \code{Input}).  Optional a minimum and a
-maximum value for the scale can be given (concatenation of the
-constraint name and \code{Minimum} or \code{Maximum}, respectively).
-The latter can avoid underflows or overflows of the scale value during
-the iteration.  This is especially useful in models where the
-iteration is very unstable and the value of the scale is very
-sensitive to the model parameters.  The meaning of the three
-constraints is the following:
+symmetry breaking consistency conditions.  When imposing the EWSB, \fs
+will adjust these parameters until the EWSB conditons are fulfilled.
+In the CMSSM example above these are the superpotential parameter
+\code{\\[Mu]} and its corresponding soft-breaking parameter
+\code{B[\\[Mu]]}.  In the NMSSM the parameters $\kappa$, $|v_s|$ and
+$m_s^2$ are usually chosen for this purpose.
+
+Furthermore, the user has to specify three model constraints:
+low-scale, SUSY-scale and high-scale.  In \fs they are named as
+\code{LowScale}, \code{SUSYScale} and \code{HighScale}.  For each
+constraint there is (i) a scale definition (named after the
+constraint), (ii) an initial guess for the scale (concatenation of the
+constraint name and \code{FirstGuess}) and (iii) a list of parameter
+settings to be applied at the scale (concatenation of the constraint
+name and \code{Input}).  Optionally a minimum and a maximum value for
+the scale can be given (concatenation of the constraint name and
+\code{Minimum} or \code{Maximum}, respectively).  The latter can avoid
+underflows or overflows of the scale value during the iteration.  This
+is especially useful in models where the iteration is very unstable
+and the value of the scale is very sensitive to the model parameters.
+The meaning of the three constraints is the following:
 %
@@
 \item \emph{High-scale constraint:} The high-scale constraint is
-  usually the GUT-scale constraint, applied at the scale where the
+  usually the GUT-scale constraint, imposed at the scale where the
   gauge couplings $g_1$ and $g_2$ unify.  The high-scale can be
-  defined by an equation of the form \code{g1 == g2} or by fixed scale
-  value.
-%
-\item \emph{SUSY-scale constraint:} The SUSY-scale constraint is the
-  constraint where \fs imposes the EWSB conditions and calculate the
-  pole mass spectrum.  Its scale $M_S$ is usually defined as
-  $\sqrt{m_{\tilde{t}_1}m_{\tilde{t}_2}}$.  In the example above it
-  has the value $M_S = \sqrt{m_{\tilde{u}_1}m_{\tilde{u}_6}}$, where
+  defined by an equation of the form \code{g1 == g2} or by a fixed
+  numerical value.
+%
+\item \emph{SUSY-scale constraint:} The SUSY-scale is the typical mass
+  scale of the SUSY particle spectrum.  At this scale \fs imposes the
+  EWSB conditions and calculates the pole mass spectrum.  The
+  SUSY-scale, $M_S$, is usually defined as $M_S =
+  \sqrt{m_{\tilde{t}_1}m_{\tilde{t}_2}}$.  However, in the example
+  above, where sfermion flavour violation is enabled, it has the value
+  $M_S = \sqrt{m_{\tilde{u}_1}m_{\tilde{u}_6}}$, where
   $m_{\tilde{u}_1}$ and $m_{\tilde{u}_6}$ are the \DRbar\ masses of
@@
 \item \emph{Low-scale constraint:} The low-scale constraint is the
-  constraint where the gauge couplings $g_1$, $g_2$, $g_3$ and the
-  Yukawa couplings $y_f$ are calculated from given Standard Model
-  input $\alpha_{\text{e.m.}}(M_Z)$, $\alpha_{s}(M_Z)$, $m_f^\pole$,
-  $M_Z^\pole$, $M_W^\pole$, see
-  Sec.~\ref{sec:calculation-of-gauge-couplings}--\ref{sec:calculation-of-yukawa-couplings}.
-  This should be done at the $Z$ mass scale $M_Z$.  The gauge
-  couplings are calculated automatically and therefore do not need to
-  be specified in the \code{LowScaleInput} variable.
+  constraint where the SUSY model is matched to the Standard Model.
+  This is done by calculating the gauge couplings $g_i$ ($i=1,2,3$)
+  and the Yukawa couplings $y_f$ ($f=u,d,e$) from the known Standard
+  Model quantities $\alpha_{\text{e.m.}}(M_Z^\pole)$,
+  $\alpha_{s}(M_Z^\pole)$, $m_f^\pole$, $M_Z^\pole$, $M_W^\pole$.  The
+  details of the calculation are explained in
+  sections~\ref{sec:calculation-of-gauge-couplings}--\ref{sec:calculation-of-yukawa-couplings}.
+  Currently this scale is fixed to be the $Z$ pole mass scale
+  $M_Z^\pole$.  The calculation of the gauge couplings is done
+  automatically and is not needed to be specified in the
+  \code{LowScaleInput} variable.  The calculation of the Yukawa
+  couplings is optional.  If the Yukawa couplings are set to the value
+  \code{Automatic}, they are determined automatically from the known
+  Standard Model fermion masses $m_f$ as explained in section
+  \ref{sec:calculation-of-yukawa-couplings}.
 \end{itemize}
 %
-\fs GUT-normalizes all gauge couplings.  Thus, the high-scale
-definition takes the simple form \code{g1 == g2}.  Another consequence
-is that in the calculation of the VEVs $v_u$ and $v_d$ from $M_Z$ and
-$\tan\beta$ at the low-scale the GUT-normalization has to be taken
-into account, see the example above.
-
-The list of parameter settings can contain as elements any of the
-following:
+Note, that \fs GUT-normalizes all gauge couplings.  Thus, the
+high-scale definition takes the simple form \code{g1 == g2}.  Another
+consequence is that in the calculation of the VEVs $v_u$ and $v_d$
+from $M_Z$ and $\tan\beta$ at the low-scale the GUT-normalization has
+to be taken into account, see the example above.
+
+The list of parameter settings for imposing a constraint can contain
+as elements any of the following:
 %
@@
 
+%\cite{Staub:2008uz}
+\bibitem{Staub:2008uz}
+  F.~Staub,
+  %``Sarah,''
+  arXiv:0806.0538 [hep-ph].
+  %%CITATION = ARXIV:0806.0538;%%
+  %82 citations counted in INSPIRE as of 03 Jun 2014
+
 \end{thebibliography}

### 2014-06-03 doc/paper.tex
make clear that the Yukawa couplings are not calculated automatically
unless their values are set to Automatic.

@@
   defined by an equation of the form \code{g1 == g2} or by a fixed
-  numerical value.
+  numerical value.  Note, that \fs GUT-normalizes all gauge couplings.
+  Thus, the high-scale definition takes the simple form \code{g1 ==
+    g2}.  As a consequence in the calculation of the VEVs $v_u$ and
+  $v_d$ from $M_Z$ and $\tan\beta$ at the low-scale the
+  GUT-normalization has to be taken into account, see the example
+  above.
 %
@@
   constraint where the SUSY model is matched to the Standard Model.
-  This is done by calculating the gauge couplings $g_i$ ($i=1,2,3$)
-  and the Yukawa couplings $y_f$ ($f=u,d,e$) from the known Standard
-  Model quantities $\alpha_{\text{e.m.}}(M_Z^\pole)$,
-  $\alpha_{s}(M_Z^\pole)$, $m_f^\pole$, $M_Z^\pole$, $M_W^\pole$.  The
-  details of the calculation are explained in
-  sections~\ref{sec:calculation-of-gauge-couplings}--\ref{sec:calculation-of-yukawa-couplings}.
-  Currently this scale is fixed to be the $Z$ pole mass scale
-  $M_Z^\pole$.  The calculation of the gauge couplings is done
-  automatically and is not needed to be specified in the
-  \code{LowScaleInput} variable.  The calculation of the Yukawa
-  couplings is optional.  If the Yukawa couplings are set to the value
-  \code{Automatic}, they are determined automatically from the known
-  Standard Model fermion masses $m_f$ as explained in section
+  This is done by automatically calculating the gauge couplings $g_i$
+  ($i=1,2,3$) of the SUSY model from the known Standard Model
+  quantities $\alpha_{\text{e.m.}}(M_Z^\pole)$,
+  $\alpha_{s}(M_Z^\pole)$, $M_Z^\pole$, $M_W^\pole$.  The details of
+  the calculation are explained in
+  section~\ref{sec:calculation-of-gauge-couplings}.  Currently this
+  scale is fixed to be the $Z$ pole mass scale $M_Z^\pole$.
+  Optionally the Yukawa couplings $y_f$ ($f=u,d,e$) can be calculated
+  automatically from the known Standard Model fermion masses $m_f$ by
+  setting their values to \code{Automatic}.  This automatic
+  calculation is explained in section
   \ref{sec:calculation-of-yukawa-couplings}.
@@
 %
-Note, that \fs GUT-normalizes all gauge couplings.  Thus, the
-high-scale definition takes the simple form \code{g1 == g2}.  Another
-consequence is that in the calculation of the VEVs $v_u$ and $v_d$
-from $M_Z$ and $\tan\beta$ at the low-scale the GUT-normalization has
-to be taken into account, see the example above.
-
 The list of parameter settings for imposing a constraint can contain

### 2014-06-03 doc/paper.tex
Revert "use proper shell command"
This reverts commit b1babd2c67c2e10aabd79ed65529321b6be7cd9b.

Conflicts:

	doc/paper.tex
doc/paper.tex: minor corrections

@@
    superpotential, gauge structure and particle content in a \sarah
-   model file and provides specific the boundary conditions in a
+   model file and provides the specific boundary conditions in a
    separate \fs model file. \fs makes use of the existing \sarah
    package to obtain self energies, tadpole corrections,
-   renormalisation group equations(RGEs) and tree level mass and
+   renormalisation group equations (RGEs) and tree level mass and
    electroweak symmetry breaking conditions for the specified model. These are
@@
    routines for solving the RGEs, EWSB conditions and simultaneously
-   solving for the spectrum conssistent with user specified boundary
+   solving for the spectrum consistent with user specified boundary
    conditions.  The modular structure of the generated code allows for
@@
          renormalisation group equations forming a large set of
-         coupled differential equations. \\ {\em Solution method:}\/
+         coupled first-order differential equations. \\ {\em Solution method:}\/
          Nested iterative algorithm and numerical minimisation of the
@@
          modular nature of the generated code adaption and extension
-         to overcome restrictions in scope is quite straight forward.
+         to overcome restrictions in scope is quite straightforward.
 
@@
 \section{Introduction}
-Supersymmetry(SUSY) provides the only non-trivial way to extend the
+Supersymmetry (SUSY) provides the only non-trivial way to extend the
 space-time symmetries of the Poincar\'e
@@
 SUSY may be realised in nature in some form. In particular
-supersymmetric extensions of the standard model where SUSY is broken
+supersymmetric extensions of the standard model (SM) where SUSY is broken
 at the TeV scale have been proposed to solve the hierarchy
@@
 (see e.g.~\cite{Aad:2013wta,Chatrchyan:2014lfa}). No other signature
-of beyond the standard model(BSM) physics has been observed, leaving
+of beyond the standard model (BSM) physics has been observed, leaving
 the fundamental questions which motivated BSM physics
-unanswered. Secondly ATLAS and CMS discovered\cite{ATLAS:2012ae,
+unanswered. Secondly ATLAS and CMS discovered \cite{ATLAS:2012ae,
   Chatrchyan:2012tx} a light Higgs of $125$ GeV, within the mass range
@@
 which can be generated from
-\sarah\cite{Staub:2010ty,Staub:2009bi,Staub:2010jh,Staub:2012pb,Staub:2013tta}.
+\sarah \cite{Staub:2010ty,Staub:2009bi,Staub:2010jh,Staub:2012pb,Staub:2013tta}.
 
@@
 
-\paragraph{Alternative boundary condition problem solvers}
+\paragraph{Alternative boundary value problem solvers}
 
@@
 it.  The resulting executable can be run like this:
+%
 \begin{lstlisting}[language=bash]
-$ ./models/MSSM/run_MSSM.x \
+$ models/MSSM/run_MSSM.x \
     --slha-input-file=model_files/MSSM/LesHouches.in.MSSM
@@
   where $g_1(M_X) = g_2(M_X)$.  The model parameters which, are set in
-  the $apply()$ function at the high-scale, are extracted from the
+  the \code{apply()} function at the high-scale, are extracted from the
   \code{HighScaleInput} variable.
@@
 where $V^\text{tree}$ is the tree-level Higgs potential, $v_i$ is the
-VEVs corresponding to the Higgs fields $H_i$ and $t_i$ is the one-loop
+VEV corresponding to the Higgs fields $H_i$ and $t_i$ is the one-loop
 tadpole diagram of $H_i$.
@@
 
-The in more detail the two-scale algorithm used in \fs works as
+In more detail the two-scale algorithm used in \fs works as
 follows, see also \figref{fig:two-scale-algorithm}:
@@
 After the solver routine has finished and convergence has been
-achieved, all $\overline{DR}$ parameters are consistent with the
+achieved, all $\overline{DR}$ parameters consistent with the
 one-loop EWSB conditions, low energy data and all user supplied
@@
 the off-diagonal slepton mass matrix elements
-from the non-trivial flavour structure of \code{Yv}
+from a non-trivial flavour structure of \code{Yv}
 \cite{Borzumati:1986qx}.

### 2014-06-03 doc/paper.tex
doc/paper.tex: fix a dangling ref
add Higgs triplet expression for theta_W, as suggested by Dominik

@@
 particle $i$ with respect to the gauge group, and $F_i$ are particle
-type-specific constants \cite{Hall:1980kf}
+type specific constants \cite{Hall:1980kf}
 %
@@
 \begin{align}
+  \theta_{W,\text{susy}}^{\text{\DRbar}}(M_Z) &= \arcsin\sqrt{1
+    -
+    \left(\frac{M_{W,\text{susy}}^{\text{\DRbar}}(M_Z)}{M_{Z,\text{susy}}^{\text{\DRbar}}(M_Z)}\right)^2}
+  ,
+\end{align}
+%
+while in models with Higgs triplets the relation looks like
+%
+\begin{align}
   \theta_{W,\text{susy}}^{\text{\DRbar}}(M_Z) &= \arcsin\sqrt{1 -
-    \left(\frac{M_{W,\text{susy}}^{\text{\DRbar}}(M_Z)}{M_{Z,\text{susy}}^{\text{\DRbar}}(M_Z)}\right)^2}
-  .
+    \frac{\left(M_{W,\text{susy}}^{\text{\DRbar}}(M_Z)\right)^2 -
+      g_2^2v_T^2}{\left(M_{Z,\text{susy}}^{\text{\DRbar}}(M_Z)\right)^2}}.
 \end{align}
@@
   \left(M_{W,\text{susy}}^{\text{\DRbar}}(M_Z)\right)^2 &=
-  \left(M_W^{\text{pole}}\right)^2 + \re \Pi_{WW}^T(p^2 = (M_W^{\text{pole}})^2, \mu=M_Z) ,\\
+  M_W^2 + \re \Pi_{WW}^T(p^2 = M_W^2, \mu=M_Z) ,\\
   \left(M_{Z,\text{susy}}^{\text{\DRbar}}(M_Z)\right)^2 &=
-  \left(M_Z^{\text{pole}}\right)^2 + \re \Pi_{ZZ}^T(p^2 = (M_Z^{\text{pole}})^2, \mu=M_Z) ,
+  M_Z^2 + \re \Pi_{ZZ}^T(p^2 = M_Z^2, \mu=M_Z) ,
 \end{align}
 %
-where $M_W^{\text{pole}} = 80.404\unit{GeV}$ and $M_Z^{\text{pole}} =
-91.1876\unit{GeV}$ \cite{Beringer:1900zz}.  Having
-$e_{\text{susy}}^{\text{\DRbar}}(M_Z)$ and
-$\theta_{W,\text{susy}}^{\text{\DRbar}}(M_Z)$ allows to calculate the
-(GUT-normalized) $U(1)_Y$ and $SU(2)_L$ gauge couplings in the SUSY
-model.  In the MSSM one has for instance
+where $M_W = 80.404\unit{GeV}$ and $M_Z = 91.1876\unit{GeV}$
+\cite{Beringer:1900zz}.  Having $e_{\text{susy}}^{\text{\DRbar}}(M_Z)$
+and $\theta_{W,\text{susy}}^{\text{\DRbar}}(M_Z)$ allows to calculate
+the (GUT-normalized) $U(1)_Y$ and $SU(2)_L$ gauge couplings in the
+SUSY model.  In the MSSM one has for instance
 %
@@
 
-As already emphasized in \secref{sec:design goals}, the modular design of
+As already emphasized in \secref{Sec:Program}, the modular design of
 \fs makes it straightforward % to change a boundary condition.

### 2014-06-04 doc/paper.tex
note about tan(beta) resummation in Delta mb
add brief introduction to Section 5

@@
 
-\section{Stucture of the spectrum generator}
+\section{Structure of the spectrum generator}
 \label{Sec:SpecGenStruct}
+
+In this section we explain the internals of \fs's automatically
+generated spectrum generator.
+%
+As mentioned in \secref{Sec:Program} one of \fs's design goals is to
+create modular C++ code to allow for an easy exchange, extension and
+reuse of the generated modules.  For this reason
+\secref{Sec:ModelParametersAndRGEs} briefly describes the so-called
+``model class'' hierarchy, which contains the general model
+information, such as parameters, $\beta$-functions, \DRbar\ mass
+spectrum, EWSB, self-energies, and the pole mass spectrum.
+%
+\secref{sec:boundary-conditions} describes how boundary conditions on
+the model parameters are implemented in general at the C++ level.  The
+subsections \ref{sec:calculation-of-gauge-couplings}--\ref{sec:ewsb}
+explain how the parameters of the SUSY model are matched to the
+Standard Model at the low scale and how the EWSB is imposed.
+%
+In \secref{sec:TreeLevelSpectrum} we describe the conventions used to
+calculate the \DRbar\ mass spectrum.
+%
+In \secref{sec:TwoScaleFixedPointIteration} the algorithm which solves
+the user-defined boundary value problem is described on the basis of
+the CMSSM example given in \secref{Sec:modfile}.
+%
+Finally, \secref{Sec:PoleMasses} briefly summarizes how the pole mass
+spectrum is obtained from the \DRbar\ model parameters after a
+solution to the boundary value problem has been found.
+
 \subsection{Model parameters and RGEs}
+\label{Sec:ModelParametersAndRGEs}
 
@@
 %
-and the \DRbar mass of the $\tau$ is calculated as
+where $\tan\beta$ enhanced loop self-energy corrections are resummed.
+Finally, the \DRbar mass of the $\tau$ is calculated as
 %
@@
 \subsection{Tree-level spectrum}
+\label{sec:TreeLevelSpectrum}
 The tree-level \DRbar masses are calculated by diagonalizing the mass
@@
 \subsection{Two-scale fixed point iteration}
+\label{sec:TwoScaleFixedPointIteration}
 
@@
 \subsection{Pole masses}
+\label{Sec:PoleMasses}
 After the solver routine has finished and convergence has been

### 2014-06-04 doc/paper.tex
unify section references

--- 2014-06-03 doc/paper.tex
@@
 \newcommand{\figref}[1]{\figurename~\ref{#1}}
-\newcommand{\secref}[1]{Sec.~\ref{#1}}
+\newcommand{\secref}[1]{Section~\ref{#1}}
 \newcommand{\tabref}[1]{\tablename~\ref{#1}}
@@
 
-In section \ref{Sec:Program} we describe the program in more detail
-and explain our design goals.  In section \ref{Sec:download}
-information on how to download and compile the code may be found along
-with details on how to get started quickly.  In section
-\ref{Sec:modfile} we describe how the user can create a new
-FlexibleSUSY model file. A detailed description of the structure and
-features of the generated code is then given in section
-\ref{Sec:SpecGenStruct}.  In section \ref{Sec:Flexible} we describe
-the various ways the code can be modified both at the meta code level
-by writing model files and at the C++ code level by modifying the code
-or adding new modules. Finally in section \ref{Sec:comparison} we
-describe detailed comparisons between our generated code and existing
-public spectrum generators as well as against the SPheno-like FORTRAN
-code which can be created using SARAH.
+In \secref{sec:Program} we describe the program in more detail and
+explain our design goals.  In \secref{sec:download} information on how
+to download and compile the code may be found along with details on
+how to get started quickly.  In \secref{sec:modfile} we describe how
+the user can create a new FlexibleSUSY model file. A detailed
+description of the structure and features of the generated code is
+then given in \secref{sec:SpecGenStruct}.  In \secref{sec:Flexible} we
+describe the various ways the code can be modified both at the meta
+code level by writing model files and at the C++ code level by
+modifying the code or adding new modules. Finally in
+\secref{sec:comparison} we describe detailed comparisons between our
+generated code and existing public spectrum generators as well as
+against the SPheno-like FORTRAN code which can be created using SARAH.
 
 \section{Overview of the Program}
-\label{Sec:Program}
+\label{sec:Program}
 
@@
 intermediate scales.  An example of such a tower construction will be
-given in section \ref{sec:adapting-cpp-code}, where a right-handed
-neutrino is integrated out at the see-saw scale, between the SUSY and
-the GUT scale.
+given in \secref{sec:adapting-cpp-code}, where a right-handed neutrino
+is integrated out at the see-saw scale, between the SUSY and the GUT
+scale.
 
 \section{Download and compilation}
-\label{Sec:download}
+\label{sec:download}
 
@@
 \section{Setting up a FlexibleSUSY model}
-\label{Sec:modfile}
+\label{sec:modfile}
 
@@
   the calculation are explained in
-  section~\ref{sec:calculation-of-gauge-couplings}.  Currently this
-  scale is fixed to be the $Z$ pole mass scale $M_Z^\pole$.
-  Optionally the Yukawa couplings $y_f$ ($f=u,d,e$) can be calculated
-  automatically from the known Standard Model fermion masses $m_f$ by
-  setting their values to \code{Automatic}.  This automatic
-  calculation is explained in section
-  \ref{sec:calculation-of-yukawa-couplings}.
+  \secref{sec:calculation-of-gauge-couplings}.  Currently this scale
+  is fixed to be the $Z$ pole mass scale $M_Z^\pole$.  Optionally the
+  Yukawa couplings $y_f$ ($f=u,d,e$) can be calculated automatically
+  from the known Standard Model fermion masses $m_f$ by setting their
+  values to \code{Automatic}.  This automatic calculation is explained
+  in \secref{sec:calculation-of-yukawa-couplings}.
 \end{itemize}
@@
   Standard Model quark and lepton masses, see
-  Section~\ref{sec:calculation-of-yukawa-couplings}.  To enable the
+  \secref{sec:calculation-of-yukawa-couplings}.  To enable the
   automatic calculation, \code{value} must be set to \code{Automatic},
@@
 automatically at the low-scale from the known Standard Model
-parameters, see Section~\ref{sec:calculation-of-gauge-couplings}.
+parameters, see \secref{sec:calculation-of-gauge-couplings}.
 
@@
 
-\section{Stucture of the spectrum generator}
-\label{Sec:SpecGenStruct}
+\section{Structure of the spectrum generator}
+\label{sec:SpecGenStruct}
+
+In this section we explain the internals of \fs's automatically
+generated spectrum generator.
+%
+As mentioned in \secref{sec:Program} one of \fs's design goals is to
+create modular C++ code to allow for an easy exchange, extension and
+reuse of the generated modules.  For this reason
+\secref{sec:ModelParametersAndRGEs} briefly describes the so-called
+``model class'' hierarchy, which contains the general model
+information, such as parameters, $\beta$-functions, \DRbar\ mass
+spectrum, EWSB, self-energies, and the pole mass spectrum.
+%
+\secref{sec:boundary-conditions} describes how boundary conditions on
+the model parameters are implemented in general at the C++ level.  The
+Subsections \ref{sec:calculation-of-gauge-couplings}--\ref{sec:ewsb}
+explain how the parameters of the SUSY model are matched to the
+Standard Model at the low scale and how the EWSB is imposed.
+%
+In \secref{sec:TreeLevelSpectrum} we describe the conventions used to
+calculate the \DRbar\ mass spectrum.
+%
+In \secref{sec:TwoScaleFixedPointIteration} the algorithm which solves
+the user-defined boundary value problem is described on the basis of
+the CMSSM example given in \secref{sec:modfile}.
+%
+Finally, \secref{sec:PoleMasses} briefly summarizes how the pole mass
+spectrum is obtained from the \DRbar\ model parameters after a
+solution to the boundary value problem has been found.
+
 \subsection{Model parameters and RGEs}
+\label{sec:ModelParametersAndRGEs}
 
@@
   in the variable \code{HighScale}.  In the model file in
-  \secref{Sec:modfile} it is defined to be the unification scale $M_X$
+  \secref{sec:modfile} it is defined to be the unification scale $M_X$
   where $g_1(M_X) = g_2(M_X)$.  The model parameters which, are set in
@@
   value of $M_S$ is defined by the model file variable
-  \code{SUSYScale}.  In the example model file in \secref{Sec:modfile}
+  \code{SUSYScale}.  In the example model file in \secref{sec:modfile}
   it is defined to be $\sqrt{m_{\tilde{u}_1} m_{\tilde{u}_6}}$, where
@@
 \code{Automatic} in the \fs model file.  This is done for example in
-the MSSM model file in \secref{Sec:modfile}.  In this case \fs
+the MSSM model file in \secref{sec:modfile}.  In this case \fs
 expresses the Yukawa couplings in terms of the \DRbar\ fermion mass
@@
   \code{InitialGuessAtLowScale}) is imposed.  In the example given in
-  \secref{Sec:modfile} the Higgs VEVs are set to
+  \secref{sec:modfile} the Higgs VEVs are set to
   \begin{align}
@@
   \code{InitialGuessAtHighScale}) is imposed.  In the example given in
-  \secref{Sec:modfile} the superpotential parameter $\mu$ is set to
+  \secref{sec:modfile} the superpotential parameter $\mu$ is set to
   the value $1.0$ and its corresponding soft-breaking parameter $B\mu$
@@
 \subsection{Pole masses}
+\label{sec:PoleMasses}
 After the solver routine has finished and convergence has been
@@
 \section{Flexible Applications}
-\label{Sec:Flexible}
+\label{sec:Flexible}
 
@@
 For a general account of the \fs model files,
-refer to \secref{Sec:modfile}.
+refer to \secref{sec:modfile}.
 
@@
 
-As already emphasized in \secref{Sec:Program}, the modular design of
+As already emphasized in \secref{sec:Program}, the modular design of
 \fs makes it straightforward % to change a boundary condition.
@@
 For this, it should help to have working knowledge about the basic structure
-of a spectrum generator, set out in \secref{Sec:SpecGenStruct}.
+of a spectrum generator, set out in \secref{sec:SpecGenStruct}.
 Obviously, this tip does not work for the matching class,
@@
 $ ./run_tower.x --slha-input-file=LesHouches.in.tower
-\end{numlstlisting}
+\end{numlstlisting}%% $
 In the output,
@@
 \section{Run-time comparison with other spectrum generators}
-\label{Sec:comparison}
+\label{sec:comparison}
 
@@
 $\tan\beta$, $\sign\mu$, $A_0$ in form of a \code{MINPAR} block to the
-SLHA template file given in \ref{sec:speed-test-slha-template-file}.
-The resulting SLHA input file is passed to each spectrum generator and
-the (wall-clock) time is measured until the program has finished.  The
-average run-times for three different CPU types can be found in
+SLHA template file given in
+\secref{sec:speed-test-slha-template-file}.  The resulting SLHA input
+file is passed to each spectrum generator and the (wall-clock) time is
+measured until the program has finished.  The average run-times for
+three different CPU types can be found in
 \tabref{tab:run-time-comparison}.  The first column shows the run-time
@@
 SARAH generated MSSM spectrum generator SPhenoMSSM.  The test results
-can be found in \secref{Sec:comparison}.
+can be found in \secref{sec:comparison}.
 

### 2014-06-04 doc/paper.tex
remover ref. to missing figure
give a small introduction on what the SG does

@@
 generated spectrum generator.
-%
-As mentioned in \secref{sec:Program} one of \fs's design goals is to
-create modular C++ code to allow for an easy exchange, extension and
-reuse of the generated modules.  For this reason
-\secref{sec:ModelParametersAndRGEs} briefly describes the so-called
-``model class'' hierarchy, which contains the general model
-information, such as parameters, $\beta$-functions, \DRbar\ mass
+
+As mentioned in \secref{sec:Program}, \fs uses SARAH-generated
+expressions for the $\beta$-functions, mass matrices, self-energies
+and EWSB conditions plus the user-defined parameter boundary
+conditions to create a spectrum generator in C++.  This program takes
+the Standard Model and user-defined input paramters and numerically
+solves the boundary value problem, which is defined by the RG
+equations and the boundary conditions.  If a solution is found the
+pole mass spectrum is calculated.
+
+In the following it is explained how this procedure is realized in
+\fs.  As mentioned in \secref{sec:Program} one of \fs's design goals
+is to create modular C++ code to allow for an easy exchange, extension
+and reuse of the generated modules.  For this reason
+\secref{sec:ModelParametersAndRGEs} first of all briefly describes the
+so-called C++ ``model class'' hierarchy, which contains the general
+model information, such as parameters, $\beta$-functions, \DRbar\ mass
 spectrum, EWSB, self-energies, and the pole mass spectrum.
@@
 Subsections \ref{sec:calculation-of-gauge-couplings}--\ref{sec:ewsb}
-explain how the parameters of the SUSY model are matched to the
-Standard Model at the low scale and how the EWSB is imposed.
+then show two concrete boundary conditions, which are always imposed:
+The matching of the model parameters to the Standard Model and the
+electroweak symmetry breaking.
 %
 In \secref{sec:TreeLevelSpectrum} we describe the conventions used to
-calculate the \DRbar\ mass spectrum.
-%
-In \secref{sec:TwoScaleFixedPointIteration} the algorithm which solves
-the user-defined boundary value problem is described on the basis of
-the CMSSM example given in \secref{sec:modfile}.
-%
-Finally, \secref{sec:PoleMasses} briefly summarizes how the pole mass
-spectrum is obtained from the \DRbar\ model parameters after a
-solution to the boundary value problem has been found.
+calculate the \DRbar\ mass spectrum given a set of \DRbar\ model
+parameters.
+%
+Afterwards, in \secref{sec:TwoScaleFixedPointIteration} the algorithm,
+which solves the user-defined boundary value problem is described on
+the basis of the CMSSM example given in \secref{sec:modfile}.
+%
+Finally, \secref{sec:PoleMasses} explains how the pole mass spectrum
+is obtained from the \DRbar\ model parameters after a solution to the
+boundary value problem has been found.
 
@@
 (at the C++ level) classes which implement the common
-\code{Constraint<Two\_scale>} interface, see
-\figref{fig:schematic-two-scale-constraint-interface}.
+\code{Constraint<Two\_scale>} interface:
 %

### 2014-06-04 doc/paper.tex
adding Dominik's suggestions on the boundary condition sections

@@
 
-The parameters of the SUSY model together with their RGEs are stored
-in the model class hierarchy, see the UML diagram in
+The parameters of the SUSY model together with their RGEs, mass
+matrices, self-energies and EWSB equations are stored at the C++ level
+in the model class hierarchy, which is shown in the UML diagram in
 \figref{fig:parameter-classes}.
@@
     }
+    \umlclass[x=0, y=-3.5]{<model>}{}{
+      + calculate\_spectrum()\\
+      + solve\_ewsb()
+    }
     \umlinherit{<model>\_susy\_parameters}{Beta\_function}
     \umlinherit{<model>\_soft\_parameters}{<model>\_susy\_parameters}
+    \umlinherit{<model>}{<model>\_soft\_parameters}
   \end{tikzpicture}
@@
 interface class, which defines the basic RGE running interface.  It
-provides the user function \code{run_to()}, which integrates the RGEs
-up to a given scale using an adaptive Runge-Kutta algorithm.  Thereby
-it uses the pure virtual functions \code{get()}, \code{set()} and
-\code{beta()}, that need to be implemented by the derived class.  The
-\code{get()} and \code{set()} functions return and set the model
-parameters in form of a long vector, respectively.  The \code{beta()}
-method returns the $\beta$-function for each model parameter in form
-of a long vector as well.
+provides the interface function \code{run_to()}, which integrates the
+RGEs up to a given scale using an adaptive Runge-Kutta algorithm.
+This algorithm uses the pure virtual functions \code{get()},
+\code{set()} and \code{beta()}, which need to be implemented by the
+derived class.  The \code{get()} and \code{set()} functions return and
+set the model parameters in form of a vector, respectively.  The
+\code{beta()} method returns the $\beta$-function for each parameter
+in form of a vector as well.
 
 All SUSY model parameters and their $\beta$-functions are contained in
-the derived classes.  The structure of the $\beta$-functions of a
-general SUSY model
+the first and second derived classes.  The structure of the
+$\beta$-functions of a general SUSY model
 \cite{Jones:1974pg,Jones:1983vk,West:1984dg,Martin:1993yx,Yamada:1993ga,MV94,Fonseca:2011vn,Sperling:2013eva,Sperling:2013xqa}
@@
 %
-The $\beta$-functions of the SUSY parameters only depend on the
-SUSY parameters and are independent of the soft-breaking
+The $\beta$-functions of the SUSY parameters in general only depend on
+the SUSY parameters and are independent of the soft-breaking
 parameters.  However, the $\beta$-functions of the soft-breaking
 parameters depend on all model parameters in general.
-
-This property is reflected in the C++ code as well: The class
+%
+This property is reflected in the C++ code: The class
 \code{<model>_susy_parameters} direcly inherits from
@@
 
+\fs creates these two classes from the model parameters defined in the
+SARAH model file.  The corresponding one- and two-loop
+$\beta$-functions are calculated algebraically using SARAH's
+\code{SARAH`CalcRGEs[]} routine and are converted to C++ form and
+written into the corresponding \code{beta()} functions.  These two
+classes now allow to use renormalisation group running of all SUSY
+model parameters.
+
+At the bottom of the hierarchy stands the actual model class, which
+defines \code{calculate_spectrum()} and \code{solve_ewsb()} functions.
+The former uses the mass matrices and self-energies obtained from
+SARAH to calculate the pole mass spectrum.  It will be explained in
+detail in \secref{sec:PoleMasses}.  The \code{solve_ewsb()} function
+numerically solves the electroweak symmetry breaking equations and
+will be explained in the next section.
+
 \subsection{Boundary conditions}
@@
 
-The parameters of a SUSY model have to meet certain boundary
-conditions.  For example, the running gauge and Yukawa couplings
-should match the ones of the Standard Model at the scale $M_Z$.
-Usually also high-scale boundary conditions on the soft-breaking
-parameters are applied, as for example in mSUGRA scenarios.
-
-In \fs the boundary conditions, imposed on the model parameters, are
-(at the C++ level) classes which implement the common
-\code{Constraint<Two\_scale>} interface:
+% The parameters of a SUSY model have to meet certain boundary
+% conditions.  For example, the running gauge and Yukawa couplings
+% should match the ones of the Standard Model at the scale $M_Z$.
+% Usually also high-scale boundary conditions on the soft-breaking
+% parameters are applied, as for example in mSUGRA scenarios.
+
+As described in \secref{sec:modfile}, the user defines three boundary
+conditions in the \fs model file at the \mathematica level.  These
+boundary conditions are converted to C++ form and are put into classes
+with a common \code{Constraint<Two\_scale>} interface.  This interface
+has the form:
 %
@@
 \code{apply()} method imposes the constraint by setting model
-parameters to values as chosen by the user.
-
-By default, \fs creates three default boundary conditions for each
-SUSY model:
+parameters to values as chosen by the user.  The three boundary
+condition classes are generated as follows:
 %
@@
   the GUT scale $M_X$.  The high-scale is defined by the value given
-  in the variable \code{HighScale}.  In the model file in
-  \secref{sec:modfile} it is defined to be the unification scale $M_X$
-  where $g_1(M_X) = g_2(M_X)$.  The model parameters which, are set in
-  the \code{apply()} function at the high-scale, are extracted from the
+  in the variable \code{HighScale}.  In the CMSSM example model file
+  in \secref{sec:modfile} it is defined to be the unification scale
+  $M_X$ where $g_1(M_X) = g_2(M_X)$.  The \code{apply()} function is
+  implemented by setting model parameters to the values defined in the
   \code{HighScaleInput} variable.
@@
 \item The \emph{SUSY-scale constraint} is indended to set boundary
-  conditions at the mass scale $M_S$ of the new SUSY particles.  The
-  value of $M_S$ is defined by the model file variable
-  \code{SUSYScale}.  In the example model file in \secref{sec:modfile}
-  it is defined to be $\sqrt{m_{\tilde{u}_1} m_{\tilde{u}_6}}$, where
-  $m_{\tilde{u}_1}$ is the mass of the lightest, and $m_{\tilde{u}_6}$
-  is the mass of the heaviest up-type squark.  The generated
-  \code{apply()} function sets the model parameters, defined in
-  \code{SUSYScaleInput}.  Finally, the \code{apply()} solves the EWSB
-  equations at the loop level, by adjusting the parameters given in
+  conditions at the mass scale $M_S$ of the SUSY particles.  The value
+  of $M_S$ is defined by the model file variable \code{SUSYScale}.  In
+  the example model file in \secref{sec:modfile} it is defined to be
+  $\sqrt{m_{\tilde{u}_1} m_{\tilde{u}_6}}$, where $m_{\tilde{u}_1}$ is
+  the mass of the lightest, and $m_{\tilde{u}_6}$ is the mass of the
+  heaviest up-type squark.  The \code{apply()} function for this
+  constraint sets the model parameters to the values defined in
+  \code{SUSYScaleInput}.  Afterwards, \code{apply()} solves the EWSB
+  equations at the loop level by adjusting the parameters given in
   \code{EWSBOutputParameters} such that the effective Higgs potential
@@
   model to the Standard Model at the scale $M_Z$.  It does so by
-  calculating the gauge and Yukawa couplings of the SUSY model from
-  the known Standard Model fermion and gauge boson masses, as well as
-  $\alpha_{\text{e.m.},\text{SM}}^{(5),\text{\MSbar}}(M_Z)$ and
-  $\alpha_{\text{s},\text{SM}}^{(5),\text{\MSbar}}(M_Z)$.  See
-  \secref{sec:calculation-of-gauge-couplings} and
-  \ref{sec:calculation-of-yukawa-couplings} for more details.  In
+  calculating the gauge couplings of the SUSY model from the known
+  Standard Model quantities
+  $\alpha_{\text{e.m.},\text{SM}}^{(5),\text{\MSbar}}(M_Z)$
+  $\alpha_{\text{s},\text{SM}}^{(5),\text{\MSbar}}(M_Z)$, $M_Z$ and
+  $M_W$.  This calculation is explained in
+  \secref{sec:calculation-of-gauge-couplings}.  Optionally, the Yukawa
+  couplings of the SUSY model can be calculated automatically from the
+  Standard Model fermion masses.  See
+  \secref{sec:calculation-of-yukawa-couplings} for more details.  In
   addition to the gauge and Yukawa couplings, the model parameter
@@
   \frac{\alpha_\text{s}}{2\pi}\left[ \frac{1}{2}-\sum_{\text{susy
-        particle } i} F_i T_i \log{\frac{m_i}{\mu}} \right] .
+        particle } i} F_i T_i \log{\frac{m_i}{\mu}} \right] ,
 \end{align}
 %
-The constants $T_i$ are the dynkin indices of the representation of
-particle $i$ with respect to the gauge group, and $F_i$ are particle
-type specific constants \cite{Hall:1980kf}
+where the sums on the right-hand sides run over all electrically and
+color charged fields absent from the Standard Model.  The constants
+$T_i$ are the dynkin indices of the representation of particle~$i$
+with respect to the gauge group, and $F_i$ are particle-type specific
+constants \cite{Hall:1980kf}
 %
@@
 %
-Afterwards, SARAH's tree-level expression for the Weinberg angle
-$\theta_W$ in terms of $M_{W,\text{susy}}^{\text{\DRbar}}(M_Z)$,
+Afterwards, SARAH's expression for the Weinberg angle $\theta_W$ in
+terms of $M_{W,\text{susy}}^{\text{\DRbar}}(M_Z)$,
 $M_{Z,\text{susy}}^{\text{\DRbar}}(M_Z)$ is used to calculate
@@
 the (GUT-normalized) $U(1)_Y$ and $SU(2)_L$ gauge couplings in the
-SUSY model.  In the MSSM one has for instance
+SUSY model.  In the MSSM they read for instance
 %
@@
 \code{Automatic} in the \fs model file.  This is done for example in
-the MSSM model file in \secref{sec:modfile}.  In this case \fs
-expresses the Yukawa couplings in terms of the \DRbar\ fermion mass
-matrices $m_u$, $m_d$, $m_e$.  In the MSSM, for example, these
-relations read
+the CMSSM model file in \secref{sec:modfile}.  In this case \fs
+expresses the Yukawa couplings in terms of the fermion mass matrices
+$m_u$, $m_d$, $m_e$.  In the MSSM, for example, these relations read
 %
@@
 %
-where $m_{u,c,d,s,e,\mu}^{\userinput}$ are read from the
-\code{SMINPUTS} block of the SLHA input file \cite{Skands:2003cj}.
-The CKM mixing matrix is currently assumed to be diagonal.  The 3rd
+where the values for $m_{u,c,d,s,e,\mu}^{\userinput}$ are read from
+the \code{SMINPUTS} block of the SLHA input file \cite{Skands:2003cj}.
+The CKM mixing matrix is currently assumed to be diagonal.  The third
 generation quark masses are calculated in the \DRbar scheme from the
@@
 \begin{align}
-  \Delta m_t^{(1),\text{qcd}} &= -\frac{g_3^2 \left(5-3 \log\left(\frac{m_t^2}{\mu^2}\right)\right)}{12 \pi^2},\\
+  \Delta m_t^{(1),\text{qcd}} &= -\frac{g_3^2}{12 \pi^2} \left[5-3 \log\left(\frac{m_t^2}{\mu^2}\right)\right],\\
   \begin{split}
@@
       m_t^{(1),\text{qcd}}\right)^2 \\
-    &\phantom{=\;} - \frac{g_3^4 \left[396
+    &\phantom{=\;} - \frac{g_3^4}{4608 \pi^4} \Bigg[396
     \log^2\left(\frac{m_t^2}{\mu^2}\right)-1476
-        \log\left(\frac{m_t^2}{\mu^2}\right)-48 \zeta(3)+2011+16 \pi
-        ^2 (1+\log 4)\right]}{4608 \pi^4}.
+    \log\left(\frac{m_t^2}{\mu^2}\right)
+    -48 \zeta(3)+2011+16 \pi ^2 (1+\log 4)\Bigg].
   \end{split}
@@
 
-In \fs the one-loop electroweak symmetry breaking conditions are
-formulated as
+\fs assumes that each SUSY model contains Higgs bosons, which allow
+for a spontaneous breaking of the electroweak symmetry.  The
+corresponding EWSB consistency conditions are formulated in \fs at the
+one-loop level as
 %
@@
 VEV corresponding to the Higgs fields $H_i$ and $t_i$ is the one-loop
-tadpole diagram of $H_i$.
-
-The electroweak symmetry breaking conditions
-\eqref{eq:one-loop-ewsb-eq} are then solved simultaneously using the
-iterative multi-dimensional root finder algorithm
-\code{gsl_multiroot_fsolver_hybrid} from the Gnu Scientific Libraray
-(GSL).  If no root was found, the \code{gsl_multiroot_fsolver_hybrids}
-algorithm is tried as alternative, which uses a variable step size but
-might be a little slower.
-
-For example in the MSSM \fs expresses \eqref{eq:one-loop-ewsb-eq} as
-the function
+tadpole diagram of $H_i$.  The electroweak symmetry breaking
+conditions \eqref{eq:one-loop-ewsb-eq} are solved simultaneously using
+the iterative multi-dimensional root finder algorithm
+\code{gsl_multiroot_fsolver_hybrid} from the GNU Scientific Libraray
+(GSL).  If no root can be found, the
+\code{gsl_multiroot_fsolver_hybrids} algorithm is tried as
+alternative, which uses a variable step size but might be a little
+slower.
+
+In the CMSSM example from \secref{sec:modfile} the
+Eqs.~\eqref{eq:one-loop-ewsb-eq} are expressed in the form of the
+following function:
 %
@@
    ...
-
-   const CLASSNAME::Ewsb_parameters* ewsb_parameters
-      = static_cast<CLASSNAME::Ewsb_parameters*>(params);
-   MSSM* model = ewsb_parameters->model;
-   const unsigned ewsb_loop_order = ewsb_parameters->ewsb_loop_order;
 
@@
 %
-Here \code{x} is the vector of EWSB output parameters (defined via
-\code{EWSBOutputParameters}) and \code{f} is a vector which contains
-the one-loop EWSB eqs.\ \eqref{eq:one-loop-ewsb-eq}.  This
-\code{tadpole_equations()} function is called inside of
-\code{MSSM<Two_scale>::solve_ewsb_iteratively_with} then as
-%
-\begin{lstlisting}[language=C++]
-// initial guess
-double x_init[number_of_ewsb_equations];
-ewsb_initial_guess(x_init);
-
-// setup root finder
-int ewsb_loop_order = 1;
-Ewsb_parameters params = {this, ewsb_loop_order};
-Root_finder<number_of_ewsb_equations> root_finder(
-                           MSSM<Two_scale>::tadpole_equations,
-                           &params,
-                           number_of_ewsb_iterations,
-                           ewsb_iteration_precision);
-root_finder.set_solver_type(gsl_multiroot_fsolver_hybrid);
-
-root_finder.find_root(x_init);
-\end{lstlisting}
+The function parameter \code{x} is the vector of EWSB output
+parameters (defined in \code{EWSBOutputParameters}) and \code{f} is a
+vector which contains the one-loop EWSB eqs.\
+\eqref{eq:one-loop-ewsb-eq}.  This \code{tadpole_equations()} function
+is passed to the root finder, which searches for values of the model
+parameters $\mu$ and $B\mu$ until \eqref{eq:one-loop-ewsb-eq} is
+fulfilled.
 

### 2014-06-04 doc/paper.tex
better wording

@@
 \label{sec:TreeLevelSpectrum}
-The tree-level \DRbar masses are calculated by diagonalizing the mass
-matrices returned from \code{SARAH`MassMatrix[]}.  The numerical
-singular value decomposition is performed by the Eigen library routine
-\code{Eigen::JacobiSVD} for matrices with less than four rows and
-columns, and the LAPACK routines \code{zgesvd}, \code{dgesvd} for
-larger matrices.  For the other types of diagonalization,
-\code{Eigen::SelfAdjointEigenSolver} from Eigen is used regardless of
-the matrix size.
+The tree-level \DRbar masses are calculated from the \DRbar\ model
+parameters by diagonalizing the mass matrices returned from
+\code{SARAH`MassMatrix[]}.  The numerical singular value decomposition
+is performed by the Eigen library routine \code{Eigen::JacobiSVD} for
+matrices with less than four rows and columns, and the LAPACK routines
+\code{zgesvd}, \code{dgesvd} for larger matrices.  For the other types
+of diagonalization, \code{Eigen::SelfAdjointEigenSolver} from Eigen is
+used regardless of the matrix size.
 

### 2014-06-04 doc/paper.tex
final suggestions from Dominik on section 6

@@
 
-The RGEs for the model parameters together with the boundary
-conditions form a boundary value problem.  The default two-scale RG
-solver tries to find a set of model parameters consistent with all
-constraints at all scales.  It does so by running iteratively between
-the scales of all boundary conditions, imposing the constraints and
-checking for convergence after each iteration.  This approach is
-decribed in \cite{Barger:1993gh} for the MSSM and is widely
-implemented in SUSY spectrum generators.
+As explained at the beginning of \secref{sec:SpecGenStruct}, the RGEs
+plus the user-defined boundary conditions on the model parameters form
+a boundary value problem.  \fs provides a default two-scale boundary
+value problem solver, which tries to find a set of model parameters
+consistent with all constraints at all scales.  It does so by running
+iteratively between the scales of all boundary conditions, imposing
+the constraints (by calling the \code{apply()} function) and checking
+for convergence after each iteration.  This approach is decribed in
+\cite{Barger:1993gh} for the MSSM and is widely implemented in SUSY
+spectrum generators.
 
@@
     \begin{align}
-      v_d(M_Z) &= \frac{2 M_Z(M_Z)}{\sqrt{0.6 g_1^2(M_Z) + g_2^2(M_Z)} \cos\beta(M_Z)}, \\
-      v_u(M_Z) &= \frac{2 M_Z(M_Z)}{\sqrt{0.6 g_1^2(M_Z) + g_2^2(M_Z)} \sin\beta(M_Z)}.
+      v_d(M_Z) &= \frac{2 M_Z^\text{\DRbar}(M_Z)}{\sqrt{0.6 g_1^2(M_Z) + g_2^2(M_Z)} \cos\beta(M_Z)}, \\
+      v_u(M_Z) &= \frac{2 M_Z^\text{\DRbar}(M_Z)}{\sqrt{0.6 g_1^2(M_Z) + g_2^2(M_Z)} \sin\beta(M_Z)}.
     \end{align}
@@
     \begin{align}
-      A^f(M_X) &= A_0 y_f(M_X) & &(f=u,d,e),\\
+      A^f(M_X) &= A_0 & &(f=u,d,e),\\
       m_{H_i}^2(M_X) &= m_0^2 & &(i=1,2),\\
@@
 \code{HighScale}.  In this case all model parameters are run to the
-SUSY-scale and the pole-mass spectrum is calculated.  If the user has
-chosen a specific output scale for the running \DRbar\ model
-parameters by setting entry $12$ in block \code{MODSEL} in the SLHA
-input file, all model parameters are run to the defined output scale.
+SUSY-scale and the pole-mass spectrum is calculated, see
+\secref{sec:PoleMasses}.  If the user has chosen a specific output
+scale for the running \DRbar\ model parameters by setting entry $12$
+in block \code{MODSEL} in the SLHA input file, all model parameters
+are finally run to the defined output scale.
 %
@@
 After the solver routine has finished and convergence has been
-achieved, all $\overline{DR}$ parameters consistent with the
-one-loop EWSB conditions, low energy data and all user supplied
-boundary conditions are known at any scale between \code{LowScale} and
-\code{HighScale}.
-
-The physical (pole) mass spectrum can now be calculated.  \fs uses the
-full one-loop self-energies and tree-level mass matrices to calculate
-the pole masses at the one-loop level, which means finding the values
-$p$ that solve the equation
+achieved, all \DRbar\ parameters consistent with the EWSB conditions,
+low energy data and all user-supplied boundary conditions and are
+known at any scale between \code{LowScale} and \code{HighScale}.
+
+The (physical) pole mass spectrum can now be calculated.  \fs uses the
+full one-loop self-energies and tree-level mass matrices obtained from
+SARAH to calculate the pole masses, which means finding the values $p$
+that solve the equation
 %
 \begin{align}
-  0 = \det\left[p^2\unity - m^2_{f,1L}(p^2)\right],
+  0 = \det\left[p^2\unity - m_{f,1L}(p^2)\right].
+  \label{eq:pole-mass-def}
 \end{align}
 %
-where the one-loop mass matrices $m_{f,1L}(p^2)$ are given in terms of
-the tree-level mass matrices $m_f$ and the self-energies
+Here the one-loop mass matrix $m_{f,1L}(p^2)$ for field $f$ is given
+in terms of the tree-level mass matrix $m_f$ and the self-energy
 $\Sigma_f(p^2)$ as
@@
   &\text{scalars } \phi: &
-  m^2_{\phi,1L}(p^2) &= m^2_{\phi} - \Sigma_\phi(p^2), \\
+  m_{\phi,1L}(p^2) &= m_{\phi} - \Sigma_\phi(p^2), \\
   &\text{majorana fermions } \chi: &
@@
 %
-Since the one-loop mass matrices depend on $p$, an iterative procedure
-must be used.  \fs provides three different methods with different
-precision and execution speed to determine the mass eigenvalues.
-These methods are set in the model file and work as follows:
+Eq.~\eqref{eq:pole-mass-def} can be solved by diagonalizing the
+one-loop mass matrix $m_{f,1L}(p^2)$.  However, since $m_{f,1L}(p^2)$
+depends on the momentum $p$, an iterative procedure must be used.
+Since this iteration can be very time consuming for large field
+multiplets, \fs provides two approximative procedures with a faster
+run-time.  The used procedure can be set in the model file for each
+field.  They work as follows:
 %
@@
   where $m_{f_i}$ is the $i$th mass eigenvalue of the tree-level mass
-  matrix $m_f$.  Afterwards $m_{f,1L}^\text{low}$ is diagonalized and
+  matrix $m_f$.  Afterwards, $m_{f,1L}^\text{low}$ is diagonalized and
   the eigenvalues are interpreted as pole masses $m_{f_i}^\pole$.
+  This method neglects which are formally of two-loop order.  This
+  method is imprecise if the self-energy corrections to the mass
+  matrix are large or the tree-level mass spectrum of the multiplet is
+  very split.
 
@@
   $(m_{f,1L}^\text{medium})^{(k)}$ is diagonalized and the $k$th
-  eigenvalue is interpreted as pole mass $m_{f_k}^\pole$.
-
-\item \code{HighPoleMassPrecision}: This option provides
-  diagonalization with highest precision, but has also the highest
-  execution time.  Here the one-loop mass matrix
-  $m_{f,1L}^\text{high}$ is diagonalized $n$ times, as in the case of
-  \code{MediumPoleMassPrecision}, resulting in $n$ pole masses
-  $m_{f_k}^\pole$ ($k = 1,\ldots,n$).  Afterwards, the diagonalization
-  is repeated, this time using the calculated pole masses
-  $m_{f_k}^\pole$ for the momentum calculation $p^2 =
-  (m_{f_k}^\pole)^2$.  The iteration stops until convergence is
+  eigenvalue is interpreted as pole mass $m_{f_k}^\pole$.  This method
+  is imprecise if the self-energy corrections to the mass matrix are
+  large.
+
+\item \code{HighPoleMassPrecision}: This option solves
+  Eq.~\eqref{eq:pole-mass-def} exactly by iterating over the momentum
+  $p$.  It therefore provides the determination of the pole masses
+  with highest precision, but has also the highest execution time.
+  Here the one-loop mass matrix $m_{f,1L}^\text{high}$ is diagonalized
+  $n$ times, as in the case of \code{MediumPoleMassPrecision},
+  resulting in $n$ pole masses $m_{f_k}^\pole$ ($k = 1,\ldots,n$).
+  Afterwards, the diagonalization is repeated, this time using the
+  calculated pole masses $m_{f_k}^\pole$ for the momentum calculation
+  $p^2 = (m_{f_k}^\pole)^2$.  The iteration stops if convergence is
   reached.
 \end{itemize}
-
-If higher accuracy is required additional routines with higher order
-corrections can be added by setting \code{UseHiggs2LoopMSSM = True;}
-in the model file. For example in the MSSM by default \fs adds
-two-loop Higgs FORTRAN routines supplied by P.~Slavich from
+%
+If higher accuracy is required, two-loop corrections to the
+self-energies can be added by setting \code{UseHiggs2LoopMSSM = True;}
+in the MSSM or \code{UseHiggs2LoopNMSSM = True;} in the NMSSM in the
+model file.  This enables two-loop Higgs FORTRAN routines supplied by
+P.~Slavich from
 \cite{Degrassi:2001yf,Brignole:2001jy,Dedes:2002dy,Brignole:2002bz,Dedes:2003km}
-to add two-loop corrections of $\oatas$, $\oabas$, $\oatq$,
-$\oabatau$, $\oabq$, $\oatauq$ and $\oatab$.
-
-Something similar is done for the NMSSM, but in this case the NMSSM
-$\oatas$, $\oabas$ pieces come from \cite{Degrassi:2009yq}, while for
-$\oatq$, $\oabatau$, $\oabq$, $\oatauq$ and $\oatab$ the MSSM pieces
-are used though it should be understood that these are not complete in
-the NMSSM. For other models since the Higgs mass is a very important
-measurement and the two loop corrections can be larger than the
+which add two-loop corrections of $\oatas$, $\oabas$, $\oatq$,
+$\oabatau$, $\oabq$, $\oatauq$ and $\oatab$.  Something similar is
+done for the NMSSM, but in this case the NMSSM $\oatas$, $\oabas$
+pieces come from \cite{Degrassi:2009yq}, while for $\oatq$,
+$\oabatau$, $\oabq$, $\oatauq$ and $\oatab$ the MSSM pieces are used
+though it should be understood that these are not complete in the
+NMSSM. For other models since the Higgs mass is a very important
+measurement and the two-loop corrections can be larger than the
 current experimental error \cite{Degrassi:2009yq} we recommend that
-the leading log two loop corrections are estimated, by generalising
+the leading log two-loop corrections are estimated, by generalising
 those of the MSSM or NMSSM, as has been done, for example, in the
-E$_6$SSM\cite{King:2005jy}.
+E$_6$SSM \cite{King:2005jy}.
 

### 2014-06-04 doc/paper.tex
correcting typos

@@
    package to obtain self energies, tadpole corrections,
-   renormalisation group equations (RGEs) and tree level mass and
+   renormalization group equations (RGEs) and tree level mass and
    electroweak symmetry breaking conditions for the specified model. These are
@@
          constraints which are specified at two or more different
-         renormalisation scales, which are connected by
-         renormalisation group equations forming a large set of
+         renormalization scales, which are connected by
+         renormalization group equations forming a large set of
          coupled first-order differential equations. \\ {\em Solution method:}\/
@@
 group \cite{Coleman:1967ad,Haag:1974qh}, leading many to suspect that
-SUSY may be realised in nature in some form. In particular
+SUSY may be realized in nature in some form. In particular
 supersymmetric extensions of the standard model (SM) where SUSY is broken
@@
 MSSM \cite{Kim:1983dt}; explaining the family structure (see e.g,
-\cite{King:2014nza}) or for successful baryogensis or leptogenesis
-(see e.g,\cite{King:2008qb}). However constructing specialised tools
+\cite{King:2014nza}) or for successful baryogenesis or leptogenesis
+(see e.g,\cite{King:2008qb}). However constructing specialized tools
 to study all relevant models would require an enormous amount of work.
@@
 reliable programs can greatly enhance our ability to understand and
-test non-minimal realisations of supersymmetry.
+test non-minimal realizations of supersymmetry.
 
@@
 parameters are imposed at the gauge coupling unification scale.
-Furthemore, at the $Z$ mass scale the CMSSM is matched to the Standard
+Furthermore, at the $Z$ mass scale the CMSSM is matched to the Standard
 Model, which implies conditions for the gauge and Yukawa couplings.
@@
 symmetry breaking conditions, one-loop self energies, one-loop
-tadpoles corrections and two-loop renormalisation group equations
+tadpoles corrections and two-loop renormalization group equations
 (RGEs) for the model.  Additional corrections which have been
@@
 %
-These algerbraic expressions are converted into C++ code and are put
+These algebraic expressions are converted into C++ code and are put
 into classes with well-defined interfaces to allow for easy exchange,
@@
 user-defined boundary value problem.  For this task \fs uses some
-parts of SOFTSUSY \cite{Allanach:2001kg}, the very fast Eigen library
+parts of Softsusy \cite{Allanach:2001kg}, the very fast Eigen library
 \cite{eigen}, augmented by LAPACK, as well as the GNU scientific
@@
   Eigen uses C++ expression templates to remove temporary objects and
-  enable lazy evaluation of the expressions.  It supports explicite
+  enable lazy evaluation of the expressions.  It supports explicit
   vectorization, and provides fixed-size matrices to avoid dynamic
@@
 user wants to modify the generated spectrum generator source code or
-reuses componets in his own program. \fs uses C++ object orientation
+reuses components in his own program. \fs uses C++ object orientation
 features to modularize the source code in order to make it easy for
@@
 application of reusing the \fs' generated classes would be to further
-improve the precision of the CE$_6$SSM spectrum generator presended in
+improve the precision of the CE$_6$SSM spectrum generator presented in
 \cite{Athron:2009bs,Athron:2012pw} by adding further one-loop
@@
 symmetry breaking consistency conditions.  When imposing the EWSB, \fs
-will adjust these parameters until the EWSB conditons are fulfilled.
+will adjust these parameters until the EWSB conditions are fulfilled.
 In the CMSSM example above these are the superpotential parameter
@@
   \code{function} is a function of these parameters.
-  \code{FSMinimize[parameters, function]} will numericall vary the
+  \code{FSMinimize[parameters, function]} will numerical vary the
   \code{parameters} until the \code{function} is minimized.  Example:
@@
   \code{functions} is a list of functions of these parameters.
-  \code{FSFindRoot[parameters, functions]} will numericall vary the
+  \code{FSFindRoot[parameters, functions]} will numerical vary the
   \code{parameters} until the \code{functions} are zero.  Example:
@@
 these corrections can be enabled by setting \code{UseHiggs2LoopMSSM =
-  True;} in the model file and by defining the effectiv $\mu$-term
+  True;} in the model file and by defining the effective $\mu$-term
 \code{EffectiveMu = \\[Mu]}.  This will add the zero-momentum
@@
 are enabled by setting \code{UseHiggs2LoopNMSSM = True;} and by
-defining the effectiv $\mu$-term like \code{EffectiveMu = \\[Lambda] vS
+defining the effective $\mu$-term like \code{EffectiveMu = \\[Lambda] vS
   / Sqrt[2]}, for example.  This will add the the zero-momentum
@@
 conditions to create a spectrum generator in C++.  This program takes
-the Standard Model and user-defined input paramters and numerically
+the Standard Model and user-defined input parameters and numerically
 solves the boundary value problem, which is defined by the RG
@@
 This property is reflected in the C++ code: The class
-\code{<model>_susy_parameters} direcly inherits from
+\code{<model>_susy_parameters} directly inherits from
 \code{Beta_functions} and implements the $\beta$-functions of the SUSY
@@
 written into the corresponding \code{beta()} functions.  These two
-classes now allow to use renormalisation group running of all SUSY
+classes now allow to use renormalization group running of all SUSY
 model parameters.
@@
 The \code{get_scale()} function is supposed to return the
-renormalisation scale at which the constraint is to be imposed.  The
+renormalization scale at which the constraint is to be imposed.  The
 \code{apply()} method imposes the constraint by setting model
@@
 \begin{itemize}
-\item The \emph{high-scale constraint} is indended to set boundary
+\item The \emph{high-scale constraint} is intended to set boundary
   conditions on the model parameters at some very high scale, e.g.\
@@
 
-\item The \emph{SUSY-scale constraint} is indended to set boundary
+\item The \emph{SUSY-scale constraint} is intended to set boundary
   conditions at the mass scale $M_S$ of the SUSY particles.  The value
@@
 
-\item The \emph{low-scale constraint} is indended to match the SUSY
+\item The \emph{low-scale constraint} is intended to match the SUSY
   model to the Standard Model at the scale $M_Z$.  It does so by
@@
 The running \DRbar\ $W$ and $Z$ boson masses are calculated in each
-iteration from the corrsponding pole masses as
+iteration from the corresponding pole masses as
 %
@@
 the iterative multi-dimensional root finder algorithm
-\code{gsl_multiroot_fsolver_hybrid} from the GNU Scientific Libraray
+\code{gsl_multiroot_fsolver_hybrid} from the GNU Scientific Library
 (GSL).  If no root can be found, the
@@
 \fs uses the following conventions for the diagonalization: A mass
-matrix $M^2$ for real scalar fiels $\phi_i$ is diagonalized with an
+matrix $M^2$ for real scalar fields $\phi_i$ is diagonalized with an
 orthogonal matrix $O$ as
@@
 %
-where we're again chosing the phases of $U$ and $V$ such that all mass
+where we're again choosing the phases of $U$ and $V$ such that all mass
 eigenvalues are positive.
@@
 the constraints (by calling the \code{apply()} function) and checking
-for convergence after each iteration.  This approach is decribed in
+for convergence after each iteration.  This approach is described in
 \cite{Barger:1993gh} for the MSSM and is widely implemented in SUSY
@@
 current experimental error \cite{Degrassi:2009yq} we recommend that
-the leading log two-loop corrections are estimated, by generalising
+the leading log two-loop corrections are estimated, by generalizing
 those of the MSSM or NMSSM, as has been done, for example, in the
@@
 
-By definition, research is an endeavour to find something new.
+By definition, research is an endeavor to find something new.
 Therefore, it can often be the case that
@@
     spectrum generators which disallow flavour violation.  Rows
-    $4$--$5$ contain flavour violatig spectrum generators, based
+    $4$--$5$ contain flavour violating spectrum generators, based
     on SARAH's MSSM model file.}

### 2014-06-04 plot_selected.sh
update plots

@@ -1,27 +1,27 @@
 
 # MSSM
 
-root -l -b -q higgs-study/plot2d.C+(\"higgs-study/data/scan_5/scan_MSSM.dat\")
+root -l -b -q higgs-study/plot2d.C+(\"higgs-study/data/scan_7/scan_MSSM.dat\")
 
 # NMSSM
 
-root -l -b -q higgs-study/plot2d.C+(\"higgs-study/data/scan_5/scan_NMSSM_lambda-0.1-a0-5000.dat\")
-root -l -b -q higgs-study/plot2d.C+(\"higgs-study/data/scan_5/scan_NMSSM_lambda-0.2-a0-5000.dat\")
+root -l -b -q higgs-study/plot2d.C+(\"higgs-study/data/scan_7/scan_NMSSM_lambda-0.1-a0-5000.dat\")
+root -l -b -q higgs-study/plot2d.C+(\"higgs-study/data/scan_7/scan_NMSSM_lambda-0.2-a0-5000.dat\")
 
-root -l -b -q higgs-study/plot2d.C+(\"higgs-study/data/scan_5/scan_NMSSM_lambda-0.1-a0--5000.dat\")
-root -l -b -q higgs-study/plot2d.C+(\"higgs-study/data/scan_5/scan_NMSSM_lambda-0.2-a0--5000.dat\")
+root -l -b -q higgs-study/plot2d.C+(\"higgs-study/data/scan_7/scan_NMSSM_lambda-0.1-a0--5000.dat\")
+root -l -b -q higgs-study/plot2d.C+(\"higgs-study/data/scan_7/scan_NMSSM_lambda-0.2-a0--5000.dat\")
 
 # phdUMSSM
 
-root -l -b -q higgs-study/plot2d.C+(\"higgs-study/data/scan_5/scan_phdUMSSM_lambda-0.1_vs-10000.dat\")
-root -l -b -q higgs-study/plot2d.C+(\"higgs-study/data/scan_5/scan_phdUMSSM_lambda-0.2_vs-10000.dat\")
-root -l -b -q higgs-study/plot2d.C+(\"higgs-study/data/scan_5/scan_phdUMSSM_lambda-0.4_vs-10000.dat\")
+root -l -b -q higgs-study/plot2d.C+(\"higgs-study/data/scan_7/scan_phdUMSSM_lambda-0.1_vs-10000.dat\")
[…]

### 2014-06-04 doc/paper.tex
small typos

@@
 if it is possible to quickly create a fast spectrum generator.
-Currently there is only option for this, a SPheno-like FORTRAN code
+Currently there is only one option for this, a SPheno-like FORTRAN code
 which can be generated from
@@
 superfield content, superpotential, gauge symmetries and mass mixings
-in the \sarah model files.  Furthermore, the boundary conditions on
-the model parameters must be specified in a separate
-\code{FlexibleSUSY.m} file.
-%
-Based on this information \fs uses the existing \sarah package
-\cite{Staub:2010ty,Staub:2009bi,Staub:2010jh,Staub:2012pb,Staub:2013tta}
-to obtain tree level expressions for the mass matrices and electroweak
-symmetry breaking conditions, one-loop self energies, one-loop
-tadpoles corrections and two-loop renormalization group equations
-(RGEs) for the model.  Additional corrections which have been
-calculated elsewhere, such as two-loop corrections to the Higgs
-masses\footnote{By default FlexibleSUSY has two-loop corrections to
-  the Higgs masses for the MSSM
+in form of \sarah model files.  The boundary conditions on the model
+parameters must be specified in a separate \code{FlexibleSUSY.m} file.
+Based on this information \fs uses \sarah to obtain tree-level
+expressions for the mass matrices and electroweak symmetry breaking
+conditions, one-loop self energies, one-loop tadpoles corrections and
+two-loop renormalisation group equations (RGEs) for the model.
+Additional corrections which have been calculated elsewhere, such as
+two-loop corrections to the Higgs masses\footnote{By default
+  FlexibleSUSY has two-loop corrections to the Higgs masses for the
+  MSSM
   \cite{Degrassi:2001yf,Brignole:2001jy,Dedes:2002dy,Brignole:2002bz,Dedes:2003km}
@@
   Slavich. These are the same corrections which are implemented in
-  many of the public spectrum generators.} may be added in file
-specified by the user.
+  many of the public spectrum generators.} may be added by the user.
 %
@@
 components.  An important application of this concept are the boundary
-conditions: The boundary conditions solver class provides a plugin
-mechanism via a common \code{Constraint} interface, which allows a
-user to exchange or add boundary conditions at any scale.
-Alternatively, if one already has an existing code for some other
-purpose and simply wishes to improve upon this by adding, e.g., new
-RGEs or self-energies the modular framework makes this straightforward
-by adding or interfacing with the appropriate routine.  A possible
-application of reusing the \fs' generated classes would be to further
-improve the precision of the CE$_6$SSM spectrum generator presented in
+conditions: The boundary conditions solver provides a plugin mechanism
+via a common \code{Constraint} interface, which allows a user to
+exchange or add boundary conditions at any scale.  Alternatively, if
+one already has an existing code for some other purpose and simply
+wishes to improve upon this by adding, e.g., new RGEs or self-energies
+the modular framework makes this straightforward by adding or
+interfacing with the appropriate routine.  A possible application of
+reusing the \fs's generated classes would be to further improve the
+precision of the CE$_6$SSM spectrum generator presented in
 \cite{Athron:2009bs,Athron:2012pw} by adding further one-loop
@@
 
-\fs can be downloaded as a gzipped tar file from the
+\fs can be downloaded as a gzipped tar file from
 \url{http://flexiblesusy.hepforge.org/}.  To download and install
@@
 
-\subsection{Alternative models}
+\subsection{Spectrum generators for alternative models}
 
 \fs already comes with plenty of predefined models: the CMSSM (simply
-called \code{MSSM}), the $Z_3$-symmetric NMSSM (called \code{NMSSM}),
-$Z_3$-violating NMSSM (\code{SMSSM}), the USSM (\code{UMSSM}), the
-\ESSM (\code{E6SSM}), the right-handed neutrino extended MSSM
-(\code{MSSMRHN}) and the NUHM-MSSM (\code{NUHMSSM}).  See the content
-of \code{model_files/} for all predefined model files.  A spectrum
-generator for the $Z_3$-symmetric NMSSM for example can be generated
-like this:
+called \code{MSSM}), the $Z_3$-symmetric NMSSM (called \code{NMSSM})
+\cite{Ellwanger:2009dp}, $Z_3$-violating NMSSM (\code{SMSSM})
+\cite{Ellwanger:2009dp}, the USSM (\code{UMSSM})
+\cite{Fayet:1977yc,Cvetic:1997ky,Erler:2002pr}, the NUHM-\ESSM
+(\code{E6SSM}) \cite{King:2005jy,Athron:2007en}, the right-handed
+neutrino extended MSSM (\code{MSSMRHN}) and the NUHM-MSSM
+(\code{NUHMSSM}).  See the content of \code{model_files/} for all
+predefined model files.  A spectrum generator for the $Z_3$-symmetric
+NMSSM for example can be generated like this:
 %
@@
 EXTPAR = { {61, LambdaInput},
-           {63, ALambdaInput}
-         };
+           {63, ALambdaInput} };
 
@@
 %
-The value of $A_\lambda$ at the GUT scale can be set in the SLHA input
-file in \code{EXTPAR} block entry $63$ via
+The value of $A_\lambda$ at the GUT scale can then be set in the SLHA
+input file in the \code{EXTPAR} block entry $63$ via
 %
@@
 
-In general a (non-constrained) softly broken SUSY model is defined by
-the gauge group, the field content and mixings as well as the
+A general (non-constrained) softly broken SUSY model is defined by the
+gauge group, the field content and mixings as well as the
 superpotential and the soft-breaking Lagrangian.  In order to create a
@@
 the SARAH manual \cite{Staub:2008uz,Staub:2013tta} for a detailed
-explanation of how to write such a model file.
+explanation of how to write such a model file.  Note, that SARAH
+already comes with a lot of predefined models, which can be used with
+\fs immediately.
 
@@
 already be found in \code{model_files/}.  In the following it is
-explained how the boundary conditions are defined on the basis of the
-CMSSM.  The application to other models is straightforward.  The CMSSM
-model file reads:
+explained how the boundary conditions can be defined on the basis of
+the CMSSM.  The application to other models is straightforward.  The
+CMSSM model file reads:
 %
@@
 The list elements are two-component lists where the first entry is the
-SLHA index in the MINPAR or EXTPAR block, respectively.  The second
+SLHA index in the MINPAR or EXTPAR block, respectively, and the second
 entry is the name of the input parameter.  In the above example the
@@
 In the CMSSM example above these are the superpotential parameter
-\code{\\[Mu]} and its corresponding soft-breaking parameter
-\code{B[\\[Mu]]}.  In the NMSSM the parameters $\kappa$, $|v_s|$ and
-$m_s^2$ are usually chosen for this purpose.
+$\mu$ and its corresponding soft-breaking parameter $B\mu$.  In the
+NMSSM the parameters $\kappa$, $|v_s|$ and $m_s^2$ are usually chosen
+for this purpose.
 
@@
 the scale can be given (concatenation of the constraint name and
-\code{Minimum} or \code{Maximum}, respectively).  The latter can avoid
+\code{Minimum} or \code{Maximum}, respectively).  The latter avoids
 underflows or overflows of the scale value during the iteration.  This
@@
   ($i=1,2,3$) of the SUSY model from the known Standard Model
-  quantities $\alpha_{\text{e.m.}}(M_Z^\pole)$,
-  $\alpha_{s}(M_Z^\pole)$, $M_Z^\pole$, $M_W^\pole$.  The details of
-  the calculation are explained in
+  quantities $\alpha_{\text{e.m.}}(M_Z)$, $\alpha_{s}(M_Z)$, $M_Z$,
+  $M_W$.  The details of this calculation are explained in
   \secref{sec:calculation-of-gauge-couplings}.  Currently this scale
-  is fixed to be the $Z$ pole mass scale $M_Z^\pole$.  Optionally the
-  Yukawa couplings $y_f$ ($f=u,d,e$) can be calculated automatically
-  from the known Standard Model fermion masses $m_f$ by setting their
-  values to \code{Automatic}.  This automatic calculation is explained
-  in \secref{sec:calculation-of-yukawa-couplings}.
+  is fixed to be the $Z$ pole mass scale $M_Z$.  Optionally the Yukawa
+  couplings $y_f$ ($f=u,d,e$) can be calculated automatically from the
+  known Standard Model fermion masses $m_f$ by setting their values to
+  \code{Automatic}.  This automatic calculation is explained in
+  \secref{sec:calculation-of-yukawa-couplings}.
 \end{itemize}
 %
-The list of parameter settings for imposing a constraint can contain
-as elements any of the following:
+The variables \code{LowScaleInput}, \code{SUSYScaleInput}
+\code{HighScaleInput}, which list the parameter settings for imposing
+the constraints can contain as elements any of the following:
 %
@@
   \code{m0^2}, and \code{mHu2} is set to the value given in the SLHA
-  input file in block \code{MSOFTIN}, entry 22.  The SLHA block names
-  and keys for the MSSM and NMSSM are defined in SARAH's
-  \code{parameters.m} file, see the SARAH manual or
-  \cite{Staub:2010jh}.  The Standard Model Yukawa couplings \code{Yu},
-  \code{Yd}, \code{Ye} can be calculated automatically from the known
+  input file in block \code{MSOFTIN}, entry 22 at the SUSY scale.  The
+  SLHA block names and keys for the MSSM and NMSSM are defined in
+  SARAH's \code{parameters.m} file, see the SARAH manual or
+  \cite{Staub:2010jh}.  For the Standard Model Yukawa couplings
+  \code{Yu}, \code{Yd}, \code{Ye} the value \code{Automatic} is
+  allowed, which triggers their automatic determination from the known
   Standard Model quark and lepton masses, see
-  \secref{sec:calculation-of-yukawa-couplings}.  To enable the
-  automatic calculation, \code{value} must be set to \code{Automatic},
-  see the example above.
+  \secref{sec:calculation-of-yukawa-couplings}.
 
@@
 \end{itemize}
-
-Finally the user can set an initial guess for the model parameters at
+%
+Finally, the user can set an initial guess for the model parameters at
 the low- and high-scale using the variables
 \code{InitialGuessAtLowScale} and \code{InitialGuessAtHighScale},
-respectively.  Note, that the gauge couplings will be guessed
-automatically at the low-scale from the known Standard Model
-parameters, see \secref{sec:calculation-of-gauge-couplings}.
+respectively.  The gauge couplings will be guessed automatically at
+the low-scale from the known Standard Model parameters.
 
@@
 are enabled by setting \code{UseHiggs2LoopNMSSM = True;} and by
-defining the effective $\mu$-term like \code{EffectiveMu = \\[Lambda] vS
-  / Sqrt[2]}, for example.  This will add the the zero-momentum
+defining the effective $\mu$-term like \code{EffectiveMu = \\[Lambda]
+  vS / Sqrt[2]}, for example.  This will add the the zero-momentum
 corrections of the order $O(y_t^2 g_3^2)$, $O(y_b^2 g_3^2)$ from
-\cite{Degrassi:2009yq}, plus leading MSSM-like contributions of the
-order $O(y_\tau^4)$, $O(y_t^4 + y_t^2 y_b^2 + y_b^4)$
+\cite{Degrassi:2009yq}, plus MSSM-like contributions of the order
+$O(y_\tau^4)$, $O(y_t^4 + y_t^2 y_b^2 + y_b^4)$
 \cite{Brignole:2001jy,Dedes:2003km}.
 
-To create a low-energy model one has to set
+One can create a pure low-energy model by setting
 \code{OnlyLowEnergyFlexibleSUSY = True}.  In this case the high-scale
-constraint will be ignored and only the low-scale and SUSY-scale
-constraints are kept.  All model parameters that are not specified in
-\code{MINPAR} or \code{EXTPAR} will be read from the corresponding
-input blocks in the SLHA input file and will be set at the SUSY-scale.
+constraint is ignored and only the low-scale and SUSY-scale
+constraints are used.  All model parameters which are not specified in
+\code{MINPAR} or \code{EXTPAR} will then be read from the
+corresponding input blocks in the SLHA input file and will be set at
+the SUSY-scale.
 
@@
 Subsections \ref{sec:calculation-of-gauge-couplings}--\ref{sec:ewsb}
-then show two concrete boundary conditions, which are always imposed:
-The matching of the model parameters to the Standard Model and the
-electroweak symmetry breaking.
+then show the two concrete boundary conditions, which are always
+imposed: The matching of the model parameters to the Standard Model
+and the electroweak symmetry breaking.
 %
@@
 
+%\cite{Fayet:1977yc}
+\bibitem{Fayet:1977yc}
+  P.~Fayet,
+  %``Spontaneously Broken Supersymmetric Theories of Weak, Electromagnetic and Strong Interactions,''
+  Phys.\ Lett.\ B {\bf 69} (1977) 489.
+  %%CITATION = PHLTA,B69,489;%%
+  %817 citations counted in INSPIRE as of 04 Jun 2014
+
+%\cite{Cvetic:1997ky}
+\bibitem{Cvetic:1997ky}
+  M.~Cvetic, D.~A.~Demir, J.~R.~Espinosa, L.~L.~Everett and P.~Langacker,
+  %``Electroweak breaking and the mu problem in supergravity models with an additional U(1),''
+  Phys.\ Rev.\ D {\bf 56} (1997) 2861
+   [Erratum-ibid.\ D {\bf 58} (1998) 119905]
+  [hep-ph/9703317].
+  %%CITATION = HEP-PH/9703317;%%
+  %237 citations counted in INSPIRE as of 04 Jun 2014
+
+%\cite{Erler:2002pr}
+\bibitem{Erler:2002pr}
+  J.~Erler, P.~Langacker and T.~-j.~Li,
+  %``The $Z$ - $Z^\prime$ mass hierarchy in a supersymmetric model with a secluded U(1) -prime breaking sector,''
+  Phys.\ Rev.\ D {\bf 66} (2002) 015002
+  [hep-ph/0205001].
+  %%CITATION = HEP-PH/0205001;%%
+  %84 citations counted in INSPIRE as of 04 Jun 2014
+
+%\cite{Athron:2007en}
+\bibitem{Athron:2007en}
+  P.~Athron, S.~F.~King, D.~J.~Miller, S.~Moretti and R.~Nevzorov,
+  %``Electroweak symmetry breaking in the E(6)SSM,''
+  J.\ Phys.\ Conf.\ Ser.\  {\bf 110} (2008) 072001
+  [arXiv:0708.3248 [hep-ph]].
+  %%CITATION = ARXIV:0708.3248;%%
+  %2 citations counted in INSPIRE as of 04 Jun 2014
+
 \end{thebibliography}

### 2014-06-04 doc/paper.tex
a little more style and typos

@@
 \newcommand{\unit}[1]{\,\text{#1}}      % units
-\newcommand{\userinput}{\text{<input>}}
+\newcommand{\userinput}{\text{input}}
 \newcommand{\pole}{\text{pole}}
@@
 these corrections can be enabled by setting \code{UseHiggs2LoopMSSM =
-  True;} in the model file and by defining the effective $\mu$-term
+  True} in the model file and by defining the effective $\mu$-term
 \code{EffectiveMu = \\[Mu]}.  This will add the zero-momentum
@@
 Higgs bosons, one neutral Goldstone boson) the two-loop contributions
-are enabled by setting \code{UseHiggs2LoopNMSSM = True;} and by
+are enabled by setting \code{UseHiggs2LoopNMSSM = True} and by
 defining the effective $\mu$-term like \code{EffectiveMu = \\[Lambda]
@@
   \end{tikzpicture}
-  \caption{Model parameter class hierarchy.}
+  \caption{Model class hierarchy.}
   \label{fig:parameter-classes}
@@
 This algorithm uses the pure virtual functions \code{get()},
-\code{set()} and \code{beta()}, which need to be implemented by the
+\code{set()} and \code{beta()}, which need to be implemented by a
 derived class.  The \code{get()} and \code{set()} functions return and
@@
 \cite{Jones:1974pg,Jones:1983vk,West:1984dg,Martin:1993yx,Yamada:1993ga,MV94,Fonseca:2011vn,Sperling:2013eva,Sperling:2013xqa}
-allows to split the parameters into two classes:
+allows to split these parameters into two classes:
 %
@@
 %
-The $\beta$-functions of the SUSY parameters in general only depend on
+The $\beta$-functions of the SUSY parameters in general depend only on
 the SUSY parameters and are independent of the soft-breaking
@@
 $\beta$-functions are calculated algebraically using SARAH's
-\code{SARAH`CalcRGEs[]} routine and are converted to C++ form and
-written into the corresponding \code{beta()} functions.  These two
-classes now allow to use renormalization group running of all SUSY
-model parameters.
+\code{CalcRGEs[]} routine, converted to C++ form and written into the
+corresponding \code{beta()} functions.  These two classes then allow
+to use renormalization group running of all SUSY model parameters.
 
@@
 conditions in the \fs model file at the \mathematica level.  These
-boundary conditions are converted to C++ form and are put into classes
-with a common \code{Constraint<Two\_scale>} interface.  This interface
-has the form:
+boundary conditions are converted to C++ form and are put into
+classes, which implement the common \code{Constraint<Two\_scale>}
+interface.  This interface has the form:
 %
@@
 $\theta_W$ in the SUSY model in the \DRbar\ scheme.  In the MSSM, for
-example, one has
+example, it yields
 %
@@
 
-\fs assumes that each SUSY model contains Higgs bosons, which allow
-for a spontaneous breaking of the electroweak symmetry.  The
-corresponding EWSB consistency conditions are formulated in \fs at the
-one-loop level as
+\fs assumes that each SUSY model contains Higgs bosons, which trigger
+a spontaneous breaking of the electroweak symmetry.  The corresponding
+EWSB consistency conditions are formulated in \fs at the one-loop
+level as
 %
@@
 Eqs.~\eqref{eq:one-loop-ewsb-eq} are expressed in the form of the
-following function:
+following C++ function:
 %
@@
 parameters (defined in \code{EWSBOutputParameters}) and \code{f} is a
-vector which contains the one-loop EWSB eqs.\
+vector which contains the one-loop EWSB Eqs.\
 \eqref{eq:one-loop-ewsb-eq}.  This \code{tadpole_equations()} function
 is passed to the root finder, which searches for values of the model
-parameters $\mu$ and $B\mu$ until \eqref{eq:one-loop-ewsb-eq} is
-fulfilled.
+parameters $\mu$ and $B\mu$ until the Eqs.\
+\eqref{eq:one-loop-ewsb-eq} are fulfilled.
 
 If higher accuracy is required additional routines with higher order
-corrections can be added by setting \code{UseHiggs2LoopMSSM = True;}
+corrections can be added by setting \code{UseHiggs2LoopMSSM = True}
 in the model file.  For example in the MSSM by default \fs adds two
@@
 contributions of the order $\oatas$, $\oabas$ \cite{Degrassi:2009yq}
-can be added by setting \code{UseHiggs2LoopNMSSM = True;} in the model
+can be added by setting \code{UseHiggs2LoopNMSSM = True} in the model
 file.
@@
 iteratively between the scales of all boundary conditions, imposing
-the constraints (by calling the \code{apply()} function) and checking
-for convergence after each iteration.  This approach is described in
-\cite{Barger:1993gh} for the MSSM and is widely implemented in SUSY
-spectrum generators.
+the constraints (by calling the corresponding \code{apply()} function)
+and checking for convergence after each iteration.  This approach is
+described in \cite{Barger:1993gh} for the MSSM and is widely
+implemented in SUSY spectrum generators.
 
@@
   $y_{u,d,e}$ of the SUSY model are set to the known Standard Model
-  Yukawa couplings (ignoring SUSY particle corrections).
+  Yukawa couplings (ignoring SUSY radiative corrections).
 \item The SUSY parameters are run to the user-supplied first guess of
@@
 \item All model parameters are run to the first guess of the low-scale
-  (\code{LowScaleFirstGuess})
+  (\code{LowScaleFirstGuess}).
 \item The EWSB eqs.\ are solved at the tree-level.
@@
   \item Impose the high-scale constraint (\code{HighScaleInput}).  In
-    the MSSM example the following soft-breaking parameters are fixed
+    the CMSSM example the following soft-breaking parameters are fixed
     to the universal values $m_0$, $M_{1/2}$ and $A_0$:
@@
     the MSSM example from above leading two-loop corrections have been
-    enabled by setting \code{UseHiggs2LoopMSSM = True;}.  This will
+    enabled by setting \code{UseHiggs2LoopMSSM = True}.  This will
     add two-loop tadpole contributions to the effective Higgs
@@
 \code{HighScale}.  In this case all model parameters are run to the
-SUSY-scale and the pole-mass spectrum is calculated, see
+SUSY-scale and the pole-mass spectrum is calculated as described in
 \secref{sec:PoleMasses}.  If the user has chosen a specific output
@@
   m_{\phi,1L}(p^2) &= m_{\phi} - \Sigma_\phi(p^2), \\
-  &\text{majorana fermions } \chi: &
+  &\text{Majorana fermions } \chi: &
   m_{\chi,1L}(p^2) &= m_{\chi} - \frac{1}{2}\Big[
@@
   \Big], \\
-  &\text{dirac fermions } \psi: &
+  &\text{Dirac fermions } \psi: &
   m_{\psi,1L}(p^2) &= m_{\psi}
@@
 one-loop mass matrix $m_{f,1L}(p^2)$.  However, since $m_{f,1L}(p^2)$
-depends on the momentum $p$, an iterative procedure must be used.
+depends on the momentum $p$, an iteration over $p$ must be performed.
 Since this iteration can be very time consuming for large field
@@
 run-time.  The used procedure can be set in the model file for each
-field.  They work as follows:
+field.  These approximations work as follows:
 %
@@
 If higher accuracy is required, two-loop corrections to the
-self-energies can be added by setting \code{UseHiggs2LoopMSSM = True;}
-in the MSSM or \code{UseHiggs2LoopNMSSM = True;} in the NMSSM in the
+self-energies can be added by setting \code{UseHiggs2LoopMSSM = True}
+in the MSSM or \code{UseHiggs2LoopNMSSM = True} in the NMSSM in the
 model file.  This enables two-loop Higgs FORTRAN routines supplied by

### 2014-06-04 doc/paper.tex
merge "Numeric tests" section into last "Tests" section
add conclusions and acknowledgments

@@
 
-\section{Run-time comparison with other spectrum generators}
+\section{Tests and comparisons with other spectrum generators}
 \label{sec:comparison}
+
+\subsection{Run-time comparison}
 
@@
 
-\appendix
-\section{Numeric tests}
+\subsection{Numeric tests}
 
@@
 can be found in \secref{sec:comparison}.
+
+\section{Conclusions}
+
+We presented \fs, a Mathematica package, which generates fast and
+modular spectrum generators for non-minimal SUSY models.  [TODO:
+That's basically all, but we could add more.]
+
+\section*{Acknowledgments}
+
+A.V.\ would like to thank Florian Staub for countless explanations of
+SARAH's internals, discussions, and exceptionally fast bug fixing.
+Finally the authors thank Philip Diessner for testing and identifying
+bugs in the MRSSM model file, as well as Ulrik Günther for compilation
+tests on MacOS.
+
+\appendix
 

### 2014-06-05 sarah/E6SSM/parameters.m
adding definitions for intermediate singlinos to E6SSM model file

@@ -235,6 +235,12 @@ ParameterDefinitions = {
 	         LesHouches -> ESIXZNI,
                  Dependence -> None,
 		 OutputName -> ZNI}},
+{ZFSInert,     { Description -> "Inert singlino Mixing-Matrix",
+		 LaTeX -> "Z^{\\tilde{s}^0}",
+		 Value -> None,
+	         LesHouches -> ESIXZSI,
+                 Dependence -> None,
+		 OutputName -> ZSI}},
 {UMInert,      { Description -> "Inert Chargino-minus Mixing-Matrix",
 		 LaTeX -> "U^{\\tilde{h}^-}",
 		 Value -> None,

### 2014-06-05 sarah/E6SSM/particles.m

@@ -448,6 +448,8 @@ ParticleDefinitions[EWSB] = {
   		     RParity -> -1}},
       {L0Prime,    { Description -> "Neutralino Prime-Weyl-Spinor",
   		     RParity -> -1}},
+      {LS0Inert,   { Description -> "Inert singlino Weyl-Spinor",
+                     RParity -> -1}},
 
       {phid,       { Description -> "Scalar Down",
 		     LaTeX -> "\\phi_{1}"}},

### 2014-06-05 doc/paper.tex
remove comment about mu, Bmu in the initial guess
correcting descr. of initial guess for Yukawa couplinges

@@
   where $v=246.22\unit{GeV}$.  Afterwards, the Yukawa couplings
-  $y_{u,d,e}$ of the SUSY model are set to the known Standard Model
-  Yukawa couplings (ignoring SUSY radiative corrections).
+  $y_{u,d,e}$ of the SUSY model are set from the known Standard Model
+  Yukawa couplings using the tree-level relations (ignoring SUSY
+  radiative corrections).
 \item The SUSY parameters are run to the user-supplied first guess of
@@
   the value $1.0$ and its corresponding soft-breaking parameter $B\mu$
-  is set to zero. {\color{red} Is $B\mu$ set to zero, this would be weird and from looking I think it is $A_0$.}
-  \JHrem{The C++ code does set $B\mu$ to zero.  Moreover, it sets $\mu$ to +1 GeV irrespective of SignMu.  This should be no problem as long as the solution is found though.}
+  is set to zero.
 \item All model parameters are run to the first guess of the low-scale

### 2014-06-05 doc/paper.tex
there are no alpha_b * alpha_tau corrections

@@
 to add two loop corrections of $\oatas$, $\oabas$, $\oatq$,
-$\oabatau$, $\oabq$, $\oatauq$ and $\oatab$.  In the NMSSM
+$\oabq$, $\oatauq$ and $\oatab$.  In the NMSSM
 contributions of the order $\oatas$, $\oabas$ \cite{Degrassi:2009yq}
@@
 which add two-loop corrections of $\oatas$, $\oabas$, $\oatq$,
-$\oabatau$, $\oabq$, $\oatauq$ and $\oatab$.  Something similar is
+$\oabq$, $\oatauq$ and $\oatab$.  Something similar is
 done for the NMSSM, but in this case the NMSSM $\oatas$, $\oabas$
 pieces come from \cite{Degrassi:2009yq}, while for $\oatq$,
-$\oabatau$, $\oabq$, $\oatauq$ and $\oatab$ the MSSM pieces are used
+$\oabq$, $\oatauq$ and $\oatab$ the MSSM pieces are used
 though it should be understood that these are not complete in the

### 2014-06-05 doc/paper.tex
doc/paper.tex: add section "Integrating custom C++ components"

@@
 intermediate scales.  An example of such a tower construction will be
-given in \secref{sec:adapting-cpp-code}, where a right-handed neutrino
-is integrated out at the see-saw scale, between the SUSY and the GUT
+given in \secref{sec:tower construction}, where right-handed neutrinos
+are integrated out at the see-saw scale, between the SUSY and the GUT
 scale.
@@
 \subsubsection{Changing boundary conditions}
+\label{sec:changing boundary conditions}
 
@@
 \label{sec:adapting-cpp-code}
+
+\subsubsection{Stacking models in a tower of effective theories}
+\label{sec:tower construction}
 
@@
 
+\subsubsection{Integrating custom-built C++ components}
+
+In \secref{sec:changing boundary conditions},
+it was explained how one can let \fs generate
+an alternative C++ boundary condition class
+by authoring a model file.
+Nonetheless, the way to employ this class at the C++ level
+might still remain obscure to the reader
+since \fs automatically took care of it.
+Here, an example shall be exhibited
+with the emphasis on the modular C++ code structure
+that helps programming.
+Concretely,
+the auto-generated low-energy boundary condition on the MSSM
+shall be modified so that
+$\alpha_{\text{s},\text{susy}}^{\text{\DRbar}}$
+is determined from
+$\alpha_{\text{s},\text{SM}}^{(5),\text{\MSbar}}$
+by means of a two-loop matching.
+This shall be accompanied by an improvement
+of the $g_3$ $\beta$-function to the three-loop accuracy.
+
+The first step is to alter
+the model class which evaluates the $\beta$-functions.
+Thanks to the \code{beta()} method being virtual,
+one can override it conveniently by deriving a class from \code{MSSM}.
+The declaration might look like:
+% MSSMcbs_two_scale_model.hpp
+\begin{numlstlisting}
+#include "MSSM_two_scale_model.hpp"
+
+template<>
+class MSSMcbs<Two_scale> : public MSSM<Two_scale> {
+public:
+  explicit MSSMcbs(const MSSM_input_parameters& input_ = MSSM_input_parameters());
+  virtual ~MSSMcbs();
+  virtual Eigen::ArrayXd beta() const;
+  MSSM_soft_parameters calc_beta() const;
+};
+\end{numlstlisting}
+where the name \code{MSSMcbs} is an abbreviation of
+the MSSM with custom-built $\beta$'s.
+Note that the objects for \code{MSSM} are reused where possible:
+\code{MSSM_input_parameters} in line 6 as well as
+\code{MSSM_soft_parameters} in line 9.
+This saves the programmer from excessive duplication of codes.
+The member function definitions read:
+\begin{numlstlisting}
+Eigen::ArrayXd MSSMcbs<Two_scale>::beta() const
+{
+  return calc_beta().get();
+}
+
+MSSM_soft_parameters MSSMcbs<Two_scale>::calc_beta() const
+{
+  MSSM_soft_parameters betas(MSSM<Two_scale>::calc_beta());
+  if (get_loops() <= 2) return betas;
+  double bg33 = /* formula in terms of g1, g2, g3, Yu, Yd, Ye */;
+  betas.set_g3(betas.get_g3() + Power(oneOver16PiSqr,3) * g3 * bg33);
+  return betas;
+}
+\end{numlstlisting}
+The full C++ expression of \code{bg33},
+used in \code{MSSMcbs_two_scale_model.cpp},
+has been adapted from the code by Jack and Jones \cite{3-loop MSSM betas}.
+This already completes the amendment of the $g_3$ $\beta$-function.
+
+The next step is to write a substitute for
+the low-energy boundary condition class.
+It must be declared as a descendent of \code{Constraint<Two_scale>}:
+% MSSMcbs_two_scale_low_scale_constraint.hpp
+\begin{numlstlisting}
+#include "two_scale_constraint.hpp"
+
+template<>
+class MSSMcbs_low_scale_constraint<Two_scale> : public Constraint<Two_scale> {
+public:
+  MSSMcbs_low_scale_constraint(const MSSM_input_parameters&, const QedQcd&);
+  virtual ~MSSMcbs_low_scale_constraint();
+  void set_threshold_corrections(unsigned);
+  ...
+private:
+  MSSMcbs<Two_scale>* model;
+  QedQcd oneset;
+  double new_g3;
+  unsigned threshold_corrections;
+  void calculate_DRbar_gauge_couplings();
+  double calculate_alS5DRbar_over_alS5MSbar(double) const;
+  double calculate_zeta_g_QCD_2(double) const;
+  double calculate_zeta_g_SUSY_2(double) const;
+  ...
+};
+\end{numlstlisting}
+In line 11,
+the type of \code{model} has been adapted to the new model.
+In fact, this class should work even if \code{model} remained
+a pointer to \code{MSSM<Two_scale>} because of inheritance.
+The main additions to \code{MSSM_low_scale_constraint}
+in \code{models/MSSM/} are the member functions in lines 16--18,
+which evaluate the two-loop matching coefficients
+from Ref.~\cite{Harlander:2005wm}.
+The following member function then performs
+the two-step decoupling as reported in this reference:
+% MSSMcbs_two_scale_low_scale_constraint.cpp
+\begin{numlstlisting}
+void MSSMcbs_low_scale_constraint<Two_scale>::calculate_DRbar_gauge_couplings()
+{
+  ...
+  double alpha_s = oneset.displayAlpha(ALPHAS);
+  double alS5DRbar_over_alS5MSbar = 1;
+  double zeta_g_QCD_2 = 1;
+  double zeta_g_SUSY_2 = 1;
+  if (model->get_thresholds()) {
+    alS5DRbar_over_als5MSbar = calculate_alS5DRbar_over_alS5MSbar(alpha_s);
+    alpha_s *= alS5DRbar_over_alS5MSbar;  // alS5MSbar -> alS5DRbar
+    zeta_g_QCD_2 = calculate_zeta_g_QCD_2(alpha_s);
+    alpha_s /= zeta_g_QCD_2;		  // alS5DRbar -> alS6DRbar
+    zeta_g_SUSY_2 = calculate_zeta_g_SUSY_2(alpha_s);
+    alpha_s /= zeta_g_SUSY_2;		  // alS6DRbar -> alS6DRbarMSSM
+    ...
+  }
+  new_g3 = Sqrt(4*Pi * alpha_s);
+  ...
+}
+\end{numlstlisting}
+Finally,
+one can integrate the new boundary condition class
+\code{MSSMcbs_low_scale_constraint} together with
+the new model \code{MSSMcbs}
+into the spectrum generator in a straightforward manner.
+They should supersede
+\code{MSSM_low_scale_constraint} and \code{MSSM}, respectively.
+The replacement should be carried out in
+those objects that depend on these classes,
+i.e.\ the initial guesser:
+\begin{numlstlisting}
+template<>
+class MSSMcbs_initial_guesser<Two_scale> : public Initial_guesser<Two_scale> {
+public:
+  MSSMcbs_initial_guesser(MSSMcbs<Two_scale>*,
+                          const MSSM_input_parameters&,
+                          const QedQcd&,
+                          const MSSMcbs_low_scale_constraint<Two_scale>&,
+                          const MSSM_susy_scale_constraint<Two_scale>&,
+                          const MSSM_high_scale_constraint<Two_scale>&);
+  ...
+private:
+  MSSMcbs<Two_scale>* model;
+  MSSM_input_parameters input_pars;
+  QedQcd oneset;
+  MSSMcbs_low_scale_constraint<Two_scale> low_constraint;
+  MSSM_susy_scale_constraint<Two_scale> susy_constraint;
+  MSSM_high_scale_constraint<Two_scale> high_constraint;
+  ...
+};
+\end{numlstlisting}
+as well as the spectrum generator object:
+% MSSMcbs_spectrum_generator.hpp
+\begin{numlstlisting}
+template <class T>
+void MSSMcbs_spectrum_generator<T>::run(const QedQcd& oneset,
+					const MSSM_input_parameters& input)
+{
+  ...
+  MSSMcbs_initial_guesser<T> initial_guesser
+    (&model, input, oneset,
+     low_scale_constraint, susy_scale_constraint, high_scale_constraint);
+  ...
+}
+\end{numlstlisting}
+The procedure described above is essentially all that is needed
+to employ new spectrum generator components,
+which may be composed from scratch or through
+a linkage to external routines.
+In doing so,
+notice that there was an evident limit on the scope of
+modules which one had to deal with.
+For instance, it was clear from the outset
+that one does not have to go through the code of the central
+fixed-point iteration engine, \code{RGFlow}.
+This manifests the power of the clean separation among objects
+each with its well-defined distinct role.
+One can find a working realization of this example
+in \code{examples/customized-betas/}.
+
 \section{Tests and comparisons with other spectrum generators}
@@
 
+\bibitem{3-loop MSSM betas}
+  I.~Jack and D.~R.~T.~Jones, \ptitle{
+    3-loop MSSM beta functions,}
+  \url{http://www.liv.ac.uk/~dij/betas}.
+
+\bibitem{Harlander:2005wm}
+  R.~Harlander, L.~Mihaila and M.~Steinhauser,
+  %``Two-loop matching coefficients for the strong coupling in the MSSM,''
+  Phys.\ Rev.\ D {\bf 72} (2005) 095009
+  [hep-ph/0509048].
+  %%CITATION = HEP-PH/0509048;%%
+  %27 citations counted in INSPIRE as of 05 Jun 2014
+
 %\cite{Hall:1980kf}

### 2014-06-06 src/linalg2.hpp
move sorting function to wrappers module

@@ -420,12 +420,6 @@ void fs_diagonalize_hermitian
     z = (z * p).adjoint().eval();
 }
 
-template<int N>
-void sort(Eigen::Array<double, N, 1>& v)
-{
-   std::sort(v.data(), v.data() + v.size(), CompareAbs<N>(v));
-}
-
 } // namespace flexiblesusy
 
 #endif // linalg2_hpp

### 2014-06-06 src/wrappers.hpp

@@ -252,6 +252,18 @@ inline int Sign(double x)
    return (x >= 0.0 ? 1 : -1);
 }
 
+namespace {
+   struct CompareAbs_d {
+      bool operator() (double a, double b) { return std::abs(a) < std::abs(b); }
+   };
+}
+
+template<int N>
+void Sort(Eigen::Array<double, N, 1>& v)
+{
+   std::sort(v.data(), v.data() + v.size(), CompareAbs_d());
+}
+
 inline double Sqrt(double a)
 {
    return std::sqrt(a);

### 2014-06-06 doc/paper.tex
doc/paper.tex: add to section "Adapting C++ code"

@@
 
+There are problems which one cannot solve only by
+editing \mathematica model files.
+To unlock the full potential of \fs,
+it is an advantage not to avoid programming at the C++ level.
+For this, it should help to have working knowledge about the basic structure
+of a spectrum generator, set out in \secref{sec:SpecGenStruct}.
+In what follows,
+it shall be demonstrated that
+the clean object structure serves as firm guidance on the job.
+
 \subsubsection{Stacking models in a tower of effective theories}
@@
 as well as gluing codes by hand to build such a program.
-Nevertheless, the clean object structure eases this job.
 One can author nearly all the components of a multi-model spectrum generator
@@
 for one of the models forming the tower.
-For this, it should help to have working knowledge about the basic structure
-of a spectrum generator, set out in \secref{sec:SpecGenStruct}.
 Obviously, this tip does not work for the matching class,
@@
 it was explained how one can let \fs generate
-an alternative C++ boundary condition class
+an alternative boundary condition class
 by authoring a model file.
@@
   if (model->get_thresholds()) {
-    alS5DRbar_over_als5MSbar = calculate_alS5DRbar_over_alS5MSbar(alpha_s);
+    alS5DRbar_over_alS5MSbar = calculate_alS5DRbar_over_alS5MSbar(alpha_s);
     alpha_s *= alS5DRbar_over_alS5MSbar;  // alS5MSbar -> alS5DRbar
@@
 \end{numlstlisting}
+One can find a working realization of this example
+in \code{examples/customized-betas/}.
+
 The procedure described above is essentially all that is needed
@@
 each with its well-defined distinct role.
-One can find a working realization of this example
-in \code{examples/customized-betas/}.
+This is just like the fact that one does not need to access the internals
+of the \code{std::sort} function in the C++ Standard Library.
+It might be entertaining to complete the analogy by
+mapping the model objects in \code{RGFlow}
+to the elements that \code{std::sort} sorts
+and the boundary condition objects
+to the comparator function.
 

### 2014-06-06 doc/paper.tex
doc/paper.tex: reminder of a complaint

@@
 adaptable spectrum generator in C++ for any SUSY model and some
-non-supersymmetric models.  The user specifies the model by giving the
+non-supersymmetric models.
+\JHrem{Which non-supersymmetric models do we support?  I failed to make \fs work for the SM from \sarah.}
+The user specifies the model by giving the
 superfield content, superpotential, gauge symmetries and mass mixings
@@
 \bibitem{eigen}
-Eigen library, version 3.1 \url{http://eigen.tuxfamily.org}
+Eigen library, version 3.1 \url{http://eigen.tuxfamily.org}.
 
@@
 \bibitem{Chowdhury:2011zr}
-  D.~Chowdhury, R.~Garani and S.~KVempati,
+  D.~Chowdhury, R.~Garani and S.~K.~Vempati,
   %``SUSEFLAV: Program for supersymmetric mass spectra with seesaw mechanism and rare lepton flavor violating decays,''

### 2014-06-06 meta/LoopMasses.m
Bugfix: sort momenta used for the self-energy calculation

@@ -84,7 +84,10 @@ Do1DimVector[particleName_String, massName_String, selfEnergyFunction_String,
 
 DoFastDiagonalization[particle_Symbol /; IsScalar[particle], tadpoles_List] :=
     Module[{result, dim, dimStr, massName, particleName, mixingMatrix, selfEnergyFunction,
-            tadpoleMatrix, U, V, massMatrixStr, selfEnergyIsSymmetric, selfEnergyMatrixType},
+            tadpoleMatrix, U, V, massMatrixStr, selfEnergyIsSymmetric, selfEnergyMatrixType,
+            momentum = "", massNameSorted = "", eigenArrayType = "",
+            declareSortedDRbarMassEigenvalues = ""
+           },
            dim = GetDimension[particle];
            dimStr = ToString[dim];
            particleName = ToValidCSymbolString[particle];
@@ -94,15 +97,31 @@ DoFastDiagonalization[particle_Symbol /; IsScalar[particle], tadpoles_List] :=
            selfEnergyFunction = SelfEnergies`CreateSelfEnergyFunctionName[particle];
            selfEnergyMatrixType = CreateCType[CConversion`MatrixType[CConversion`realScalarCType, dim, dim]];
            tadpoleMatrix = FillTadpoleMatrix[tadpoles, "tadpoles"];
+           If[ContainsGoldstone[particle],
+              (* DR-bar mass multiplets with Goldstones were re-ordered.
+                 So, we need to ensure that we use sorted masses as momenta. *)
+                 massNameSorted = massName <> "_sorted";
+                 momentum = massNameSorted;
+                 eigenArrayType = CreateCType[CConversion`ArrayType[CConversion`realScalarCType, dim]];
+                 declareSortedDRbarMassEigenvalues = eigenArrayType <> " " <>
+                 massNameSorted <> "(" <> massName <> ");\n" <>
+                 "Sort(" <> massNameSorted <> ");\n";
+                 ,
+                 (* DR-bar mass multiplets without Goldstones are assumed to
+                    be mass ordered. *)
+              momentum = massName;
+              declareSortedDRbarMassEigenvalues = "";
+             ];
            If[dim > 1,
               selfEnergyIsSymmetric = Length[Flatten[{mixingMatrix}]] === 1;
-              result = tadpoleMatrix <>
+              result = declareSortedDRbarMassEigenvalues <>
+                       tadpoleMatrix <>
                        selfEnergyMatrixType <> " self_energy;\n" <>
[…]

### 2014-06-06 doc/paper.tex
doc/paper.tex: minor corrections
correcting complaint

--- 2014-06-06 doc/paper.tex
@@
 \fs is a \mathematica package designed to create a fast and easily
-adaptable spectrum generator in C++ for any SUSY model and some
-non-supersymmetric models.  The user specifies the model by giving the
+adaptable spectrum generator in C++ for any SUSY model.
+The user specifies the model by giving the
 superfield content, superpotential, gauge symmetries and mass mixings

### 2014-06-06 test/test_SMSSM_tree_level_spectrum.cpp
correcting test: Goldstone bosons are now at the front

@@ -35,8 +35,8 @@ BOOST_AUTO_TEST_CASE( test_SMSSM_tree_level_masses )
    const DoubleVector Ah(ToDoubleVector(m.get_MAh()));
    const double mz = s.displayMzRun();
    DoubleVector mA0(3);
-   mA0(1) = std::min(s.displayDrBarPars().mA0(1), mz);
-   mA0(2) = std::max(s.displayDrBarPars().mA0(1), mz);
+   mA0(1) = mz;
+   mA0(2) = s.displayDrBarPars().mA0(1);
    mA0(3) = s.displayDrBarPars().mA0(2);
    BOOST_CHECK_CLOSE(Ah(1), mA0(1), 1.0e-12);
    BOOST_CHECK_CLOSE(Ah(2), mA0(2), 1.0e-12);

### 2014-06-06 test/test_benchmark.cpp
pipe cerr output to /dev/null

@@ -33,7 +33,7 @@ int run_point(const std::string& slha_file, double& fs_time, double& ss_time)
    stopwatch.start();
    status = run_cmd("./models/MSSM/run_MSSM.x --slha-input-file=" +
                     slha_file + " --slha-output-file="
-                    "test/test_benchmark.out.spc > /dev/null");
+                    "test/test_benchmark.out.spc > /dev/null 2>&1");
    stopwatch.stop();
    fs_time = stopwatch.get_time_in_seconds();
 

### 2014-06-06 doc/paper.tex
e.g, --> e.g.~ in introduction.

@@
 are, for example, motivated by the $\mu$-problem of the
-MSSM \cite{Kim:1983dt}; explaining the family structure (see e.g,
-\cite{King:2014nza}) or for successful baryogenesis or leptogenesis
-(see e.g,\cite{King:2008qb}). However constructing specialized tools
+MSSM \cite{Kim:1983dt}; explaining the family structure (see e.g.~\cite{King:2014nza}) or for successful baryogenesis or leptogenesis
+(see e.g.~\cite{King:2008qb}). However constructing specialized tools
 to study all relevant models would require an enormous amount of work.

### 2014-06-06 doc/paper.tex
slight grammar fix in introduction.  some sentences are still too long.

@@
 the naturalness problems by, for example, raising the tree level Higgs
-mass, and models developed with a fresh perspective, based on other
+mass and models developed with a fresh perspective based on other
 considerations.  In both cases exploration of such models can be aided

### 2014-06-06 doc/paper.tex
Dominik's suggestions on sections 3-4

@@
 
-\section{Overview of the Program}
+\section{Overview of the program and design goals}
 \label{sec:Program}
@@
 
+\paragraph{Modularity}
+
+The large variety of supersymmetric models and potential
+investigations makes it likely that the user wants to modify the
+generated spectrum generator source code or reuse components in
+further programs.  \fs offers two levels to influence the code: (i) On
+the \mathematica model file level the model itself or GUT/weak scale
+boundary conditions as well as input and output parameters can be
+controlled (see \secref{sec:quick-start-alternative-models} and
+\secref{sec:adapting-model-files} for examples).  (ii) In particular
+\fs uses C++ object orientation features to modularize the source code
+in order to make it easy for the user to modify, reuse, replace or
+extend the individual components (see \secref{sec:adapting-cpp-code}
+for examples).  An important application of this concept are the
+boundary conditions, for which the C++ level offers a wider range of
+possibilities.  The boundary conditions solver provides a plugin
+mechanism via a common \code{Constraint} interface, which allows a
+user to exchange or add boundary conditions at any scale.  An
+elaborate example of a tower of effective field theories and multiple
+matching scales is presented in \secref{sec:tower construction}.
+Alternatively, the modular structure makes it straightforward to take
+\fs generated code for e.g.\ RGEs or self-energies and reuse it in an
+existing code for some other purpose.  Conversely, it is also easy to
+include code from elsewhere into the spectrum generator.  For an
+example see \secref{sec:integrating-custom-built}.
+
 \paragraph{Speed}
@@
 
-\paragraph{Modularity}
-
-The large variety of supersymmetric models makes it likely that the
-user wants to modify the generated spectrum generator source code or
-reuses components in his own program. \fs uses C++ object orientation
-features to modularize the source code in order to make it easy for
-the user to modify, reuse, replace or extend the individual
-components.  An important application of this concept are the boundary
-conditions: The boundary conditions solver provides a plugin mechanism
-via a common \code{Constraint} interface, which allows a user to
-exchange or add boundary conditions at any scale.  Alternatively, if
-one already has an existing code for some other purpose and simply
-wishes to improve upon this by adding, e.g., new RGEs or self-energies
-the modular framework makes this straightforward by adding or
-interfacing with the appropriate routine.  A possible application of
-reusing the \fs's generated classes would be to further improve the
-precision of the CE$_6$SSM spectrum generator presented in
-\cite{Athron:2009bs,Athron:2012pw} by adding further one-loop
-self-energies and two-loop $\beta$-functions.
-
 \paragraph{Alternative boundary value problem solvers}
@@
 to search for solutions in such critical parameter regions.  A
-subsequent release with an alternative lattice solver is already
-planned.
+subsequent release with an alternative solver is already planned.
 
@@
 
-\section{Download and compilation}
+\section{Quick start}
 \label{sec:download}
@@
 
-\subsection{Quick Start}
+\subsection{Download and generating a first spectrum generator}
+\label{sec:quick-start-cmssm}
 
@@
 \subsection{Spectrum generators for alternative models}
+\label{sec:quick-start-alternative-models}
 
@@
 (\code{E6SSM}) \cite{King:2005jy,Athron:2007en}, the right-handed
-neutrino extended MSSM (\code{MSSMRHN}) and the NUHM-MSSM
-(\code{NUHMSSM}).  See the content of \code{model_files/} for all
-predefined model files.  A spectrum generator for the $Z_3$-symmetric
-NMSSM for example can be generated like this:
+neutrino extended MSSM (\code{MSSMRHN}), the NUHM-MSSM
+(\code{NUHMSSM}) and the $R$-symmetric MSSM (\code{MRSSM})
+\cite{Kribs:2007ac}.  See the content of \code{model_files/} for all
+predefined model files.  For all these models spectrum generators can
+be generated easily like for the CMSSM in
+\secref{sec:quick-start-cmssm}.  The spectrum generator for the
+$Z_3$-symmetric NMSSM for example can be generated like this:
 %
@@
 %
-This NMSSM variant unifies all soft-breaking trilinear scalar
-couplings at the GUT scale.  In order to relax this constraint and use
-a separate value for $A_\lambda$ at the GUT scale one can edit the
-model file \code{models/NMSSM/FlexibleSUSY.m} and change the lines
+One of the design goals is modularity and the possibility to easily
+construct custom spectrum generators.  The details of the
+customization can be found in Sections
+\ref{sec:modfile}--\ref{sec:Flexible}.  As a simple example consider
+the NMSSM.  The NMSSM variant above unifies all soft-breaking
+trilinear scalar couplings at the GUT scale.  In order to relax this
+constraint and use a separate value for $A_\lambda$ at the GUT scale
+one can edit the model file \code{models/NMSSM/FlexibleSUSY.m} and
+change the lines
 %
@@
 \subsection{Adapting model files}
+\label{sec:adapting-model-files}
 
@@
 \subsubsection{Integrating custom-built C++ components}
+\label{sec:integrating-custom-built}
 
@@
 
+%\cite{Kribs:2007ac}
+\bibitem{Kribs:2007ac}
+  G.~D.~Kribs, E.~Poppitz and N.~Weiner,
+  %``Flavor in supersymmetry with an extended R-symmetry,''
+  Phys.\ Rev.\ D {\bf 78} (2008) 055010
+  [arXiv:0712.2039 [hep-ph]].
+  %%CITATION = ARXIV:0712.2039;%%
+  %130 citations counted in INSPIRE as of 06 Jun 2014
+
 \end{thebibliography}

### 2014-06-06 doc/paper.tex
grammar: Note,  -->  Note

@@
 the SARAH manual \cite{Staub:2008uz,Staub:2013tta} for a detailed
-explanation of how to write such a model file.  Note, that SARAH
+explanation of how to write such a model file.  Note that SARAH
 already comes with a lot of predefined models, which can be used with
@@
 \code{FlexibleSUSY.m}, which has to be located in the model directory
-\code{models/<model>/}.  Note, that many example model files can
+\code{models/<model>/}.  Note that many example model files can
 already be found in \code{model_files/}.  In the following it is
@@
   defined by an equation of the form \code{g1 == g2} or by a fixed
-  numerical value.  Note, that \fs GUT-normalizes all gauge couplings.
+  numerical value.  Note that \fs GUT-normalizes all gauge couplings.
   Thus, the high-scale definition takes the simple form \code{g1 ==

### 2014-06-06 doc/paper.tex
note about example for a pure low-energy model
a little language improvement

@@
 corresponding input blocks in the SLHA input file and will be set at
-the SUSY-scale.
+the SUSY-scale.  An example of such a pure low-energy model is the
+MRSSM, where the three gauge couplings do not unify at a common scale.
 
@@
 \secref{sec:boundary-conditions} describes how boundary conditions on
-the model parameters are implemented in general at the C++ level.  The
+the model parameters are implemented in general at the C++ level.
 Subsections \ref{sec:calculation-of-gauge-couplings}--\ref{sec:ewsb}

### 2014-06-06 doc/paper.tex
update Program Summary

@@
            format:}\/ tar.gz\\ {\em Programming language:} {\tt
-           C++}\\ {\em Computer:}\/ Personal computer\\ {\em Operating
-           system:}\/ Tested on Linux 3.x\\ {\em Word size:}\/ 64
+           C++, \mathematica, FORTRAN}\\ {\em Computer:}\/ Personal computer\\ {\em Operating
+           system:}\/ Tested on Linux 3.x, MacOS\\ {\em Word size:}\/ 64
          bits\\ {\em External routines:}\/ SARAH 4.0.4, Boost library,

### 2014-06-06 doc/paper.tex
doc/paper.tex: split command box for tower example
as Dominik suggested.

@@
 
-One could best see the overall code structure in the \code{Makefile}:
+As the target spectrum generator depends on two models,
+one should first build these prerequisites by:
+\begin{numlstlisting}
+$ ./createmodel --name=MSSM
+$ ./createmodel --name=MSSMRHN
+$ ./configure --with-models=MSSM,MSSMRHN
+$ make
+\end{numlstlisting}
+As a by-product, line 3
+also creates a \code{Makefile} in \code{examples/tower/}.
+One could best see the overall code structure of the application in this file:
 \begin{numlstlisting}
@@
 \begin{numlstlisting}
-$ ./createmodel --name=MSSM
-$ ./createmodel --name=MSSMRHN
-$ ./configure --with-models=MSSM,MSSMRHN
-$ make
 $ cd examples/tower

### 2014-06-06 doc/paper.tex
removed intance of y_b^2 y_\taui^2 since we don't inlcude these.

@@
 corrections of the order $O(y_t^4 + y_b^2 y_t^2 + y_b^4)$, $O(y_t^2
-g_3^2)$, $O(y_b^2 g_3^2)$, $O(y_\tau^4)$, $O(y_\tau^2 y_b^2)$ from
+g_3^2)$, $O(y_b^2 g_3^2)$, $O(y_\tau^4)$$ from
 \cite{Degrassi:2001yf,Brignole:2001jy,Dedes:2002dy,Brignole:2002bz,Dedes:2003km}.

### 2014-06-06 doc/paper.tex
the programming language used in the Mathmatica program is called Wolfram
> Hi all,
>
> apparently, the programming language used in the Mathmatica program is
> called "Wolfram", although I am not sure how many people would recognize
> the name.  :)
>
>      http://en.wikipedia.org/wiki/Wolfram_Language
>      http://reference.wolfram.com/language
>
> Best,
> Jae-hyeon

@@
            format:}\/ tar.gz\\ {\em Programming language:} {\tt
-           C++, \mathematica, FORTRAN}\\ {\em Computer:}\/ Personal computer\\ {\em Operating
+           C++, Wolfram/\mathematica, FORTRAN}\\ {\em Computer:}\/ Personal computer\\ {\em Operating
            system:}\/ Tested on Linux 3.x, MacOS\\ {\em Word size:}\/ 64

### 2014-06-06 doc/paper.tex
doc/paper.tex: corrections to section "Adapting C++ code"

@@
 corrections of the order $O(y_t^4 + y_b^2 y_t^2 + y_b^4)$, $O(y_t^2
-g_3^2)$, $O(y_b^2 g_3^2)$, $O(y_\tau^4)$$ from
+g_3^2)$, $O(y_b^2 g_3^2)$, $O(y_\tau^4)$ from
 \cite{Degrassi:2001yf,Brignole:2001jy,Dedes:2002dy,Brignole:2002bz,Dedes:2003km}.
@@
 \end{numlstlisting}
-This file is created in \code{examples/tower/}
-on the execution of the \code{configure} script.
 The include directives in line 2 tell the compiler
@@
 fixed-point iteration engine, \code{RGFlow}.
-This manifests the power of the clean separation among objects
+This manifests the power of the clear separation among objects
 each with its well-defined distinct role.

### 2014-06-06 doc/paper.tex
typo
doc/paper.tex: highlight C++ comments except pseudo-codes

@@
 \code{MSSM_MSSMRHN_spectrum_generator.hpp}:
-\begin{numlstlisting}[name=SGrun]
+\begin{numlstlisting}[name=SGrun,language=C++]
 template<class T> void MSSM_MSSMRHN_spectrum_generator<T>::run
@@
 and initializes the \code{MSSM} object:
-\begin{numlstlisting}[name=SGrun]
+\begin{numlstlisting}[name=SGrun,language=C++]
   model_1.clear();                 // of type MSSM<T>
@@
 Likewise for \code{MSSMRHN}:
-\begin{numlstlisting}[name=SGrun]
+\begin{numlstlisting}[name=SGrun,language=C++]
   std::vector<Constraint<T>*> upward_constraints_2;
@@
 along with the precision specification:
-\begin{numlstlisting}[name=SGrun]
+\begin{numlstlisting}[name=SGrun,language=C++]
   Two_scale_increasing_precision precision(10.0, precision_goal);
@@
 let it calculate the spectrum:
-\begin{numlstlisting}[name=SGrun]
+\begin{numlstlisting}[name=SGrun,language=C++]
   susy_scale_1 = susy_scale_constraint_1.get_scale();
@@
 with a matching procedure in-between:
-\begin{numlstlisting}[name=MSSM_MSSMRHN_two_scale_initial_guesser.cpp]
+\begin{numlstlisting}[name=MSSM_MSSMRHN_two_scale_initial_guesser.cpp,language=C++]
   model_1->run_to(matching_scale_guess); // of type MSSM<Two_scale>
@@
 % MSSMcbs_two_scale_low_scale_constraint.cpp
-\begin{numlstlisting}
+\begin{numlstlisting}[language=C++]
 void MSSMcbs_low_scale_constraint<Two_scale>::calculate_DRbar_gauge_couplings()

### 2014-06-06 doc/paper.tex
Abstract with Dominik's suggestions

@@
   \begin{abstract}
-   We introduce \fs, a \mathematica package which generates a fast,
+    We introduce \fs, a \mathematica package, which generates a fast,
     precise C++ spectrum generator for any SUSY model specified by the
@@
     modularity in mind, making it easy to adapt and extend with new
-   features. The user specifies the model supplying the
-   superpotential, gauge structure and particle content in a \sarah
-   model file and provides the specific boundary conditions in a
-   separate \fs model file. \fs makes use of the existing \sarah
-   package to obtain self energies, tadpole corrections,
-   renormalization group equations (RGEs) and tree level mass and
-   electroweak symmetry breaking conditions for the specified model. These are
-   translated into C++ code and combined  with numerical
-   routines for solving the RGEs, EWSB conditions and simultaneously
-   solving for the spectrum consistent with user specified boundary
-   conditions.  The modular structure of the generated code allows for
-   individual components to be replaced with an alternative if
-   available. \fs has been carefully designed to grow as
-   alternative solvers and calculators are added.
+    features. The model is specified by supplying the superpotential,
+    gauge structure and particle content in a \sarah model file;
+    specific boundary conditions e.g.\ at the GUT, weak or
+    intermediate scales are defined in a separate \fs model file.
+    From these model files, \fs generates C++ code for self-energies,
+    tadpole corrections, renormalization group equations (RGEs) and
+    electroweak symmetry breaking (EWSB) conditions and combines them
+    with numerical routines for solving the RGEs and EWSB conditions
+    simultaneously.  The resulting spectrum generator is then able to
+    solve for the spectrum of the model, including loop-corrected pole
+    masses, consistent with user specified boundary conditions.  The
+    modular structure of the generated code allows for individual
+    components to be replaced with an alternative if available. \fs
+    has been carefully designed to grow as alternative solvers and
+    calculators are added.  Predefined models include the MSSM, NMSSM,
+    \ESSM, USSM, $R$-symmetric models and models with right-handed
+    neutrinos.
   \end{abstract}

### 2014-06-06 doc/paper.tex
doc/paper.tex: differentiate between inheritance and component separation

@@
 \fs uses C++ object orientation features to modularize the source code
-in order to make it easy for the user to modify, reuse, replace or
+so that it is sharply divided into manifestly separate building blocks.
+This modular architecture makes it easy
+for the user to modify, reuse, replace or
 extend the individual components (see \secref{sec:adapting-cpp-code}
@@
 mechanism via a common \code{Constraint} interface, which allows a
-user to exchange or add boundary conditions at any scale.  An
+user to exchange or add boundary conditions at any scale.
+To realized this, all (derived) constraint objects are intentionally
+kept outside the solver.  Despite being independent of one another,
+they can fit together with the aid of class inheritance.  An
 elaborate example of a tower of effective field theories and multiple

### 2014-06-06 doc/paper.tex
doc/paper.tex: avoid redundancy
sharply ~ manifestly
divided ~ separate
addressing Dominik's suggestions on section 7

@@
 \fs uses C++ object orientation features to modularize the source code
-so that it is sharply divided into manifestly separate building blocks.
+so that it is sharply divided into building blocks performing
+distinct duties.
 This modular architecture makes it easy
@@
 when he or she programs for a wide variety of studies.
-In what follows,
-this property shall be demonstrated by presenting a few use cases
-at differing levels of complexity.
+We provide two basic levels for the user to create a custom spectrum
+generator: (i) The \mathematica level, where he or she writes or
+adapts a model file and (ii) the C++ level, where the generated
+classes can be extended, recombined or replaced by self-made modules.
+In what follows, adaptions on these two levels shall be demonstrated
+by presenting a few use cases at differing degrees of complexity.
 
@@
 \end{numlstlisting}
-With respect to the MSSM file, newly added lines are
+With respect to the MSSM file, the newly added lines are
 6, 12, 15, and 22.
@@
 \end{numlstlisting}
-For details on how to write model files for \sarah, consult its manual.
+For further details on how to write model files for \sarah, we refer
+to its manual \cite{Staub:2008uz,Staub:2013tta}.
 
@@
 it shall be demonstrated that
-the clean object structure serves as firm guidance on the job.
+the clean class structure serves as firm guidance on the job.
 
@@
 by a tower of effective theories.
-Within the framework of \fs, the C++ object structure is
+Within the framework of \fs, the C++ class structure is
 a faithful reflection of
 this physicist's view on the given problem.
-This point might be illustrated
+Here we illustrate this point
 using a well-known configuration in which
@@
 gives rise to the MSSM as the lower-energy effective theory.
-The relevant objects are sketched
+The relevant classes are sketched
 in \figref{fig:class structure for tower}.
@@
 \end{tikzpicture}
-  \caption{Schematic object structure in the C++ code for
+  \caption{Schematic class structure in the C++ code for
     the tower scenario.}
@@
 
-Since the current implementation of \fs does not (yet)
+Since the current implementation of \fs does not (yet) automatically
 produce a spectrum generator spanning multiple models,
-one needs to write the matching object
+one needs to write the matching class
 % in \figref{fig:class structure for tower}
@@
 for one of the models forming the tower.
-Obviously, this tip does not work for the matching class,
-which has no analogue in a single-model case.
+The matching class, however, needs to be set-up by hand, because it
+has no analogue in a single-model case.
 
@@
 also creates a \code{Makefile} in \code{examples/tower/}.
-One could best see the overall code structure of the application in this file:
+One can best see the overall code structure of the application in this file:
 \begin{numlstlisting}
@@
 \code{input_1} and \code{input_2}, in lines 7--8.
-The suffixes \code{_1} and \code{_2} are not necessary but intended to remind
-the programmer of to which model the object is relevant.
-
-The \code{main} function then delegates
-the \code{MSSM_MSSMRHN_spectrum_generator} object
-to set up and drive the \code{RGFlow} object in
-\figref{fig:class structure for tower}.
-This task is started by the call in line 14 to
-the following member function defined in
-\code{MSSM_MSSMRHN_spectrum_generator.hpp}:
+
+The crucial point is the definition of the
+\code{MSSM_MSSMRHN_spectrum_generator} object in line 12, which
+creates and drives the \code{RGFlow} object in \figref{fig:class
+  structure for tower}.  This task is started by calling the
+\code{run()} member function in line 14.  It is defined in
+\code{MSSM_MSSMRHN_spectrum_generator.hpp} and reads:
+%
 \begin{numlstlisting}[name=SGrun,language=C++]

### 2014-06-06 doc/paper.tex
extend conclusions

@@
 We presented \fs, a Mathematica package, which generates fast and
-modular spectrum generators for non-minimal SUSY models.  [TODO:
-That's basically all, but we could add more.]
+modular spectrum generators for non-minimal SUSY models.  It comes
+with plenty of predefined models for the CMSSM, NMSSM, USSM, \ESSM,
+MRSSM etc., which can be used out of the box.  In particular the CMSSM
+and NMSSM spectrum generators constitute alternatives for to the
+existing ones Softsusy, SPheno and NMSPEC.  We have shown that the
+generated source code can be influenced at two different levels: The
+\mathematica level where the user provides a model file, and the C++
+level where the generated objects can be easily exchanged, extended,
+modified and reused.  This provides great flexibility for creating
+custom spectrum generators for both the most common and most
+extraordinary models.
 

### 2014-06-07 meta/FlexibleSUSY.m
reorder the goldstone bosons separately

@@ -479,6 +479,7 @@ WriteModelClass[massMatrices_List, ewsbEquations_List,
             saveSoftHiggsMasses, restoreSoftHiggsMasses,
             solveTreeLevelEWSBviaSoftHiggsMasses,
             copyDRbarMassesToPoleMasses = "",
+            reorderDRbarMasses = "", reorderPoleMasses = "",
             higgsToEWSBEqAssociation,
             twoLoopHiggsHeaders = "",
             lspGetters = "", lspFunctions = "",
@@ -562,6 +563,8 @@ WriteModelClass[massMatrices_List, ewsbEquations_List,
            restoreSoftHiggsMasses       = Parameters`RestoreParameter[softHiggsMasses, "old_", ""];
            solveTreeLevelEWSBviaSoftHiggsMasses = EWSB`SolveTreeLevelEwsbVia[ewsbEquations, softHiggsMasses];
            gslEWSBRootFinders           = EWSB`CreateEWSBRootFinders[FlexibleSUSY`FSEWSBSolvers];
+           reorderDRbarMasses           = TreeMasses`ReorderGoldstoneBosons[""];
+           reorderPoleMasses            = TreeMasses`ReorderGoldstoneBosons["PHYSICAL"];
            WriteOut`ReplaceInFiles[files,
                           { "@lspGetters@"           -> IndentText[lspGetters],
                             "@lspFunctions@"         -> lspFunctions,
@@ -576,6 +579,8 @@ WriteModelClass[massMatrices_List, ewsbEquations_List,
                             "@clearOutputParameters@"  -> IndentText[clearOutputParameters],
                             "@clearPhases@"            -> IndentText[clearPhases],
                             "@copyDRbarMassesToPoleMasses@" -> IndentText[copyDRbarMassesToPoleMasses],
+                            "@reorderDRbarMasses@"     -> IndentText[reorderDRbarMasses],
+                            "@reorderPoleMasses@"      -> IndentText[reorderPoleMasses],
                             "@ewsbInitialGuess@"       -> IndentText[ewsbInitialGuess],
                             "@physicalMassesDef@"      -> IndentText[physicalMassesDef],
                             "@mixingMatricesDef@"      -> IndentText[mixingMatricesDef],

### 2014-06-07 meta/LoopMasses.m

@@ -528,8 +528,6 @@ CreateLoopMassFunction[particle_Symbol, precision_Symbol, tadpole_] :=
                      IndentText["return;"] <> "\n";
              ];
            body = body <> DoDiagonalization[particle, precision, tadpole];
-           body = body <> TreeMasses`ReorderGoldstoneBosons[particle,
-                              TreeMasses`FindMixingMatrixSymbolFor[particle], "PHYSICAL"];
            result = "void CLASSNAME::" <> CreateLoopMassFunctionName[particle] <>
                     "()\n{\n" <> IndentText[body] <> "}\n\n";
            Return[result];

### 2014-06-07 meta/TreeMasses.m

@@ -609,13 +609,28 @@ CreateReorderingFunctionCalls[{idx_Integer, vector_, higgs_, mixingMatrix_}, mac
 
 CreateReorderingFunctionCalls[___] := "";
 
-ReorderGoldstoneBosons[particle_, mixingMatrix_, macro_String:""] :=
+ReorderGoldstoneBosons[particle_, mixingMatrix_, macro_String] :=
     Module[{goldstoneList},
            goldstoneList = Cases[SARAH`GoldstoneGhost,
                                  {vector_, particle[{idx_}]} :> {idx, vector, particle, mixingMatrix}];
-           If[goldstoneList =!= {}, "\n", ""] <>
            StringJoin[CreateReorderingFunctionCalls[#,macro]& /@ goldstoneList]
           ];
+
+ReorderGoldstoneBosons[particles_List, macro_String] :=
+    Module[{result = ""},
+           (result = result <> ReorderGoldstoneBosons[#,macro])& /@ particles;
+           Print["reordering for particles ", particles, " = ", result];
+           Return[result];
+          ];
+
+ReorderGoldstoneBosons[particle_Symbol, macro_String] :=
+    ReorderGoldstoneBosons[particle, FindMixingMatrixSymbolFor[particle], macro];
+
+ReorderGoldstoneBosons[particle_[___], macro_String] :=
+    ReorderGoldstoneBosons[particle, macro];
+
+ReorderGoldstoneBosons[macro_String] :=
+    ReorderGoldstoneBosons[GetParticles[], macro];
 
 CreateDiagonalizationFunction[matrix_List, eigenVector_, mixingMatrixSymbol_] :=
     Module[{dim, body = "", result, U = "", V = "", dimStr = "", ev, particle, k},
@@ -652,9 +667,6 @@ CreateDiagonalizationFunction[matrix_List, eigenVector_, mixingMatrixSymbol_] :=
                      IndentText["problems.unflag_tachyon(" <> particle <> ");"] <> "\n\n";
               body = body <> ev <> " = AbsSqrt(" <> ev <> ");\n";
              ];
-           (* reorder goldstone boson masses *)
-           body = body <> ReorderGoldstoneBosons[GetHead[eigenVector],
-                                                 mixingMatrixSymbol];
            Return[result <> IndentText[body] <> "}\n"];
           ];
 

### 2014-06-07 test/test_SMSSM_tree_level_spectrum.cpp

@@ -22,6 +22,7 @@ BOOST_AUTO_TEST_CASE( test_SMSSM_tree_level_masses )
    setup_SMSSM(m, s, input);
 
    m.calculate_DRbar_parameters();
+   m.reorder_DRbar_masses();
    s.calcDrBarPars();
 
    // neutral CP even Higgs

### 2014-06-07 meta/TreeMasses.m
remove debug output

@@ -619,7 +619,6 @@ ReorderGoldstoneBosons[particle_, mixingMatrix_, macro_String] :=
 ReorderGoldstoneBosons[particles_List, macro_String] :=
     Module[{result = ""},
            (result = result <> ReorderGoldstoneBosons[#,macro])& /@ particles;
-           Print["reordering for particles ", particles, " = ", result];
            Return[result];
           ];
 

### 2014-06-07 doc/paper.tex
doc/paper.tex: add Bourne shell as language, delete word size

@@
            format:}\/ tar.gz\\ {\em Programming language:} {\tt
-           C++, Wolfram/\mathematica, FORTRAN}\\ {\em Computer:}\/ Personal computer\\ {\em Operating
-           system:}\/ Tested on Linux 3.x, MacOS\\ {\em Word size:}\/ 64
-         bits\\ {\em External routines:}\/ SARAH 4.0.4, Boost library,
+           C++, Wolfram/\mathematica, FORTRAN, Bourne shell}\\ {\em Computer:}\/ Personal computer\\ {\em Operating
+           system:}\/ Tested on Linux 3.x, MacOS\\
+         {\em External routines:}\/ SARAH 4.0.4, Boost library,
          Eigen, LAPACK\\ {\em

### 2014-06-07 doc/paper.tex
doc/paper.tex: change title of sec:quick-start-cmssm
add and discuss model class methods for DR-bar and pole masses

@@
 
-\subsection{Download and generating a first spectrum generator}
+\subsection{Downloading \fs and generating a first spectrum generator}
 \label{sec:quick-start-cmssm}
@@
     }
-    \umlclass[x=0, y=-3.5]{<model>}{}{
+    \umlclass[x=0, y=-4.2]{<model>}{
+      -- \DRbar\ masses and mixing matrices
+    }{
+      + calculate\_DRbar\_parameters()\\
+      + calculate\_pole\_masses()\\
       + calculate\_spectrum()\\
+      + get\_physical()\\
       + solve\_ewsb()
@@
 At the bottom of the hierarchy stands the actual model class, which
-defines \code{calculate_spectrum()} and \code{solve_ewsb()} functions.
-The former uses the mass matrices and self-energies obtained from
-SARAH to calculate the pole mass spectrum.  It will be explained in
-detail in \secref{sec:PoleMasses}.  The \code{solve_ewsb()} function
-numerically solves the electroweak symmetry breaking equations and
-will be explained in the next section.
+uses the \DRbar\ parameters from the parent classes to calculate
+\DRbar\ and pole mass spectra.  These two calculations are performed
+in the \code{calculate_DRbar_parameters()} and
+\code{calculate_pole_masses()} functions, which make use of the mass
+matrices and self-energies obtained from SARAH.  The calculation of
+the pole mass spectrum will be explained in detail in
+\secref{sec:PoleMasses}.  The resulting masses can be obtained by
+calling \code{get_physical()}.  The \code{calculate_spectrum()}
+function combines these two spectrum computations into one call.  In
+addition, the model class provides a \code{solve_ewsb()} method, which
+solves the electroweak symmetry breaking equations numerically at the
+loop level.  This function is explained in the next section.
 
@@
 \label{sec:boundary-conditions}
-
-% The parameters of a SUSY model have to meet certain boundary
-% conditions.  For example, the running gauge and Yukawa couplings
-% should match the ones of the Standard Model at the scale $M_Z$.
-% Usually also high-scale boundary conditions on the soft-breaking
-% parameters are applied, as for example in mSUGRA scenarios.
 

### 2014-06-07 doc/paper.tex
avoid term "out of the box"
sort references

@@
 $\beta$-functions of a general SUSY model
-\cite{Jones:1974pg,Jones:1983vk,West:1984dg,Martin:1993yx,Yamada:1993ga,MV94,Fonseca:2011vn,Sperling:2013eva,Sperling:2013xqa}
+\cite{Jones:1974pg,Jones:1983vk,West:1984dg,Martin:1993yx,Yamada:1993ga,MV94,Fonseca:2011vn,Yam94,Sperling:2013eva,Sperling:2013xqa}
 allows to split these parameters into two classes:
@@
 
-TODO: Adding tests against the CE$_6$SSM?
+TODO: Adding tests against the CE$_6$SSM?  \cite{Athron:2009bs,Athron:2012pw}
 
@@
 with plenty of predefined models for the CMSSM, NMSSM, USSM, \ESSM,
-MRSSM etc., which can be used out of the box.  In particular the CMSSM
+MRSSM etc., which can be generated immediately without any editing.  In particular the CMSSM
 and NMSSM spectrum generators constitute alternatives for to the
@@
   %82 citations counted in INSPIRE as of 03 Nov 2013
+
+%\cite{Chowdhury:2011zr}
+\bibitem{Chowdhury:2011zr}
+  D.~Chowdhury, R.~Garani and S.~K.~Vempati,
+  %``SUSEFLAV: Program for supersymmetric mass spectra with seesaw mechanism and rare lepton flavor violating decays,''
+  Comput.\ Phys.\ Commun.\  {\bf 184} (2013) 899
+  [arXiv:1109.3551 [hep-ph]].
+  %%CITATION = ARXIV:1109.3551;%%
+  %16 citations counted in INSPIRE as of 03 Jun 2014
+
 %\cite{Ellwanger:2006rn}
@@
 
+%\cite{Degrassi:2001yf}
+\bibitem{Degrassi:2001yf}
+  G.~Degrassi, P.~Slavich and F.~Zwirner,
+  %``On the neutral Higgs boson masses in the MSSM for arbitrary stop mixing,''
+  Nucl.\ Phys.\ B {\bf 611} (2001) 403
+  [hep-ph/0105096].
+  %%CITATION = HEP-PH/0105096;%%
+  %166 citations counted in INSPIRE as of 04 Apr 2014
+
+%\cite{Brignole:2001jy}
+\bibitem{Brignole:2001jy}
+  A.~Brignole, G.~Degrassi, P.~Slavich and F.~Zwirner,
+  %``On the O(alpha(t)**2) two loop corrections to the neutral Higgs boson masses in the MSSM,''
+  Nucl.\ Phys.\ B {\bf 631} (2002) 195
+  [hep-ph/0112177].
+  %%CITATION = HEP-PH/0112177;%%
+  %192 citations counted in INSPIRE as of 04 Apr 2014
+
+%\cite{Dedes:2002dy}
+\bibitem{Dedes:2002dy}
+  A.~Dedes and P.~Slavich,
+  %``Two loop corrections to radiative electroweak symmetry breaking in the MSSM,''
+  Nucl.\ Phys.\ B {\bf 657} (2003) 333
+  [hep-ph/0212132].
+  %%CITATION = HEP-PH/0212132;%%
+  %53 citations counted in INSPIRE as of 04 Apr 2014
+
+%\cite{Brignole:2002bz}
+\bibitem{Brignole:2002bz}
+  A.~Brignole, G.~Degrassi, P.~Slavich and F.~Zwirner,
+  %``On the two loop sbottom corrections to the neutral Higgs boson masses in the MSSM,''
+  Nucl.\ Phys.\ B {\bf 643} (2002) 79
+  [hep-ph/0206101].
+  %%CITATION = HEP-PH/0206101;%%
+  %152 citations counted in INSPIRE as of 04 Apr 2014
+
+%\cite{Dedes:2003km}
+\bibitem{Dedes:2003km}
+  A.~Dedes, G.~Degrassi and P.~Slavich,
+  %``On the two loop Yukawa corrections to the MSSM Higgs boson masses at large tan beta,''
+  Nucl.\ Phys.\ B {\bf 672} (2003) 144
+  [hep-ph/0305127].
+  %%CITATION = HEP-PH/0305127;%%
+  %99 citations counted in INSPIRE as of 04 Apr 2014
+
 %\cite{Degrassi:2009yq}
@@
 
+\bibitem{eigen}
+  Eigen library, version 3.1 \url{http://eigen.tuxfamily.org}.
+
 %\cite{Skands:2003cj}
@@
   %327 citations counted in INSPIRE as of 21 Sep 2013
+
+%\cite{Fayet:1977yc}
+\bibitem{Fayet:1977yc}
+  P.~Fayet,
+  %``Spontaneously Broken Supersymmetric Theories of Weak, Electromagnetic and Strong Interactions,''
+  Phys.\ Lett.\ B {\bf 69} (1977) 489.
+  %%CITATION = PHLTA,B69,489;%%
+  %817 citations counted in INSPIRE as of 04 Jun 2014
+
+%\cite{Cvetic:1997ky}
+\bibitem{Cvetic:1997ky}
+  M.~Cvetic, D.~A.~Demir, J.~R.~Espinosa, L.~L.~Everett and P.~Langacker,
+  %``Electroweak breaking and the mu problem in supergravity models with an additional U(1),''
+  Phys.\ Rev.\ D {\bf 56} (1997) 2861
+   [Erratum-ibid.\ D {\bf 58} (1998) 119905]
+  [hep-ph/9703317].
+  %%CITATION = HEP-PH/9703317;%%
+  %237 citations counted in INSPIRE as of 04 Jun 2014
+
+%\cite{Erler:2002pr}
+\bibitem{Erler:2002pr}
+  J.~Erler, P.~Langacker and T.~-j.~Li,
+  %``The $Z$ - $Z^\prime$ mass hierarchy in a supersymmetric model with a secluded U(1) -prime breaking sector,''
+  Phys.\ Rev.\ D {\bf 66} (2002) 015002
+  [hep-ph/0205001].
+  %%CITATION = HEP-PH/0205001;%%
+  %84 citations counted in INSPIRE as of 04 Jun 2014
+
+%\cite{King:2005jy}
+\bibitem{King:2005jy} 
+  S.~F.~King, S.~Moretti and R.~Nevzorov,
+  %``Theory and phenomenology of an exceptional supersymmetric standard model,''
+  Phys.\ Rev.\ D {\bf 73}, 035009 (2006)
+  [hep-ph/0510419].
+  %%CITATION = HEP-PH/0510419;%%
+  %130 citations counted in INSPIRE as of 03 Nov 2013
+
+%\cite{Athron:2007en}
+\bibitem{Athron:2007en}
+  P.~Athron, S.~F.~King, D.~J.~Miller, S.~Moretti and R.~Nevzorov,
+  %``Electroweak symmetry breaking in the E(6)SSM,''
+  J.\ Phys.\ Conf.\ Ser.\  {\bf 110} (2008) 072001
+  [arXiv:0708.3248 [hep-ph]].
+  %%CITATION = ARXIV:0708.3248;%%
+  %2 citations counted in INSPIRE as of 04 Jun 2014
+
+%\cite{Kribs:2007ac}
+\bibitem{Kribs:2007ac}
+  G.~D.~Kribs, E.~Poppitz and N.~Weiner,
+  %``Flavor in supersymmetry with an extended R-symmetry,''
+  Phys.\ Rev.\ D {\bf 78} (2008) 055010
+  [arXiv:0712.2039 [hep-ph]].
+  %%CITATION = ARXIV:0712.2039;%%
+  %130 citations counted in INSPIRE as of 06 Jun 2014
+
+%\cite{Staub:2008uz}
+\bibitem{Staub:2008uz}
+  F.~Staub,
+  %``Sarah,''
+  arXiv:0806.0538 [hep-ph].
+  %%CITATION = ARXIV:0806.0538;%%
+  %82 citations counted in INSPIRE as of 03 Jun 2014
 
@@
 
-%\cite{Ell08}
-\bibitem{Ell08} 
-  U.~Ellwanger, C.~-C.~Jean-Louis and A.~M.~Teixeira,
-  %``Phenomenology of the General NMSSM with Gauge Mediated Supersymmetry Breaking,''
-  JHEP {\bf 0805}, 044 (2008)
-  [arXiv:0803.2962 [hep-ph]].
-  %%CITATION = ARXIV:0803.2962;%%
-  %28 citations counted in INSPIRE as of 24 Sep 2013
-
-\bibitem{Barger:1993gh} 
-  V.~D.~Barger, M.~S.~Berger and P.~Ohmann,
-  %``The Supersymmetric particle spectrum,''
-  Phys.\ Rev.\ D {\bf 49}, 4908 (1994)
-  [hep-ph/9311269].
-  %%CITATION = HEP-PH/9311269;%%
-  %365 citations counted in INSPIRE as of 03 Nov 2013
-
-\bibitem{eigen}
-Eigen library, version 3.1 \url{http://eigen.tuxfamily.org}.
-
-%\cite{King:2005jy}
-\bibitem{King:2005jy} 
-  S.~F.~King, S.~Moretti and R.~Nevzorov,
-  %``Theory and phenomenology of an exceptional supersymmetric standard model,''
-  Phys.\ Rev.\ D {\bf 73}, 035009 (2006)
-  [hep-ph/0510419].
-  %%CITATION = HEP-PH/0510419;%%
-  %130 citations counted in INSPIRE as of 03 Nov 2013
-
 %\cite{Beringer:1900zz}
@@
   %3229 citations counted in INSPIRE as of 22 Feb 2014
+
+%\cite{Hall:1980kf}
+\bibitem{Hall:1980kf}
+  L.~J.~Hall,
+  %``Grand Unification of Effective Gauge Theories,''
+  Nucl.\ Phys.\ B {\bf 178} (1981) 75.
+  %%CITATION = NUPHA,B178,75;%%
+  %265 citations counted in INSPIRE as of 10 May 2014
 
@@
 
-%\cite{Degrassi:2001yf}
-\bibitem{Degrassi:2001yf}
-  G.~Degrassi, P.~Slavich and F.~Zwirner,
-  %``On the neutral Higgs boson masses in the MSSM for arbitrary stop mixing,''
-  Nucl.\ Phys.\ B {\bf 611} (2001) 403
-  [hep-ph/0105096].
-  %%CITATION = HEP-PH/0105096;%%
-  %166 citations counted in INSPIRE as of 04 Apr 2014
-
-%\cite{Brignole:2001jy}
-\bibitem{Brignole:2001jy}
-  A.~Brignole, G.~Degrassi, P.~Slavich and F.~Zwirner,
-  %``On the O(alpha(t)**2) two loop corrections to the neutral Higgs boson masses in the MSSM,''
-  Nucl.\ Phys.\ B {\bf 631} (2002) 195
-  [hep-ph/0112177].
-  %%CITATION = HEP-PH/0112177;%%
-  %192 citations counted in INSPIRE as of 04 Apr 2014
-
-%\cite{Dedes:2002dy}
-\bibitem{Dedes:2002dy}
-  A.~Dedes and P.~Slavich,
-  %``Two loop corrections to radiative electroweak symmetry breaking in the MSSM,''
-  Nucl.\ Phys.\ B {\bf 657} (2003) 333
-  [hep-ph/0212132].
-  %%CITATION = HEP-PH/0212132;%%
-  %53 citations counted in INSPIRE as of 04 Apr 2014
-
-%\cite{Brignole:2002bz}
-\bibitem{Brignole:2002bz}
-  A.~Brignole, G.~Degrassi, P.~Slavich and F.~Zwirner,
-  %``On the two loop sbottom corrections to the neutral Higgs boson masses in the MSSM,''
-  Nucl.\ Phys.\ B {\bf 643} (2002) 79
-  [hep-ph/0206101].
-  %%CITATION = HEP-PH/0206101;%%
-  %152 citations counted in INSPIRE as of 04 Apr 2014
-
-%\cite{Dedes:2003km}
-\bibitem{Dedes:2003km}
-  A.~Dedes, G.~Degrassi and P.~Slavich,
-  %``On the two loop Yukawa corrections to the MSSM Higgs boson masses at large tan beta,''
-  Nucl.\ Phys.\ B {\bf 672} (2003) 144
-  [hep-ph/0305127].
-  %%CITATION = HEP-PH/0305127;%%
-  %99 citations counted in INSPIRE as of 04 Apr 2014
+\bibitem{Barger:1993gh} 
+  V.~D.~Barger, M.~S.~Berger and P.~Ohmann,
+  %``The Supersymmetric particle spectrum,''
+  Phys.\ Rev.\ D {\bf 49}, 4908 (1994)
+  [hep-ph/9311269].
+  %%CITATION = HEP-PH/9311269;%%
+  %365 citations counted in INSPIRE as of 03 Nov 2013
 
@@
 
-%\cite{Hall:1980kf}
-\bibitem{Hall:1980kf}
-  L.~J.~Hall,
-  %``Grand Unification of Effective Gauge Theories,''
-  Nucl.\ Phys.\ B {\bf 178} (1981) 75.
-  %%CITATION = NUPHA,B178,75;%%
-  %265 citations counted in INSPIRE as of 10 May 2014
-
-%\cite{Chowdhury:2011zr}
-\bibitem{Chowdhury:2011zr}
-  D.~Chowdhury, R.~Garani and S.~K.~Vempati,
-  %``SUSEFLAV: Program for supersymmetric mass spectra with seesaw mechanism and rare lepton flavor violating decays,''
-  Comput.\ Phys.\ Commun.\  {\bf 184} (2013) 899
-  [arXiv:1109.3551 [hep-ph]].
-  %%CITATION = ARXIV:1109.3551;%%
-  %16 citations counted in INSPIRE as of 03 Jun 2014
-
 %\cite{Athron:2009bs}
@@
 
-%\cite{Staub:2008uz}
-\bibitem{Staub:2008uz}
-  F.~Staub,
-  %``Sarah,''
-  arXiv:0806.0538 [hep-ph].
-  %%CITATION = ARXIV:0806.0538;%%
-  %82 citations counted in INSPIRE as of 03 Jun 2014
-
-%\cite{Fayet:1977yc}
-\bibitem{Fayet:1977yc}
-  P.~Fayet,
-  %``Spontaneously Broken Supersymmetric Theories of Weak, Electromagnetic and Strong Interactions,''
-  Phys.\ Lett.\ B {\bf 69} (1977) 489.
-  %%CITATION = PHLTA,B69,489;%%
-  %817 citations counted in INSPIRE as of 04 Jun 2014
-
-%\cite{Cvetic:1997ky}
-\bibitem{Cvetic:1997ky}
-  M.~Cvetic, D.~A.~Demir, J.~R.~Espinosa, L.~L.~Everett and P.~Langacker,
-  %``Electroweak breaking and the mu problem in supergravity models with an additional U(1),''
-  Phys.\ Rev.\ D {\bf 56} (1997) 2861
-   [Erratum-ibid.\ D {\bf 58} (1998) 119905]
-  [hep-ph/9703317].
-  %%CITATION = HEP-PH/9703317;%%
-  %237 citations counted in INSPIRE as of 04 Jun 2014
-
-%\cite{Erler:2002pr}
-\bibitem{Erler:2002pr}
-  J.~Erler, P.~Langacker and T.~-j.~Li,
-  %``The $Z$ - $Z^\prime$ mass hierarchy in a supersymmetric model with a secluded U(1) -prime breaking sector,''
-  Phys.\ Rev.\ D {\bf 66} (2002) 015002
-  [hep-ph/0205001].
-  %%CITATION = HEP-PH/0205001;%%
-  %84 citations counted in INSPIRE as of 04 Jun 2014
-
-%\cite{Athron:2007en}
-\bibitem{Athron:2007en}
-  P.~Athron, S.~F.~King, D.~J.~Miller, S.~Moretti and R.~Nevzorov,
-  %``Electroweak symmetry breaking in the E(6)SSM,''
-  J.\ Phys.\ Conf.\ Ser.\  {\bf 110} (2008) 072001
-  [arXiv:0708.3248 [hep-ph]].
-  %%CITATION = ARXIV:0708.3248;%%
-  %2 citations counted in INSPIRE as of 04 Jun 2014
-
-%\cite{Kribs:2007ac}
-\bibitem{Kribs:2007ac}
-  G.~D.~Kribs, E.~Poppitz and N.~Weiner,
-  %``Flavor in supersymmetry with an extended R-symmetry,''
-  Phys.\ Rev.\ D {\bf 78} (2008) 055010
-  [arXiv:0712.2039 [hep-ph]].
-  %%CITATION = ARXIV:0712.2039;%%
-  %130 citations counted in INSPIRE as of 06 Jun 2014
-
 \end{thebibliography}

### 2014-06-07 doc/paper.tex
typo fix gauge-single -> gauge singlet
inserted \linbreak to stop line overflow.

@@
 One of the simplest classes of models beyond the MSSM is
-those with additional gauge-single fields.
+those with additional gauge-singlet fields.
 In what follows, a supersymmetric type-I see-saw model
@@
 \end{numlstlisting}
-Compared to the MSSM case,
-the differences are that
-the type of \code{high_constraint_2}
-is \code{MSSMRHN_high_scale_constraint<Two_scale>} and that
-\code{matching_scale_guess} has been inserted.
-Due to this intermediate scale,
-the initial run-up is divided into two steps,
-with a matching procedure in-between:
+Compared to the MSSM case, the differences are that the type of
+\code{high_constraint_2} is \linebreak {\code{MSSMRHN_high_scale_constraint<Two_scale>} and that \code{matching_scale_guess} has been inserted.
+Due to this intermediate scale, the initial run-up is divided into two steps, with a matching procedure in-between:
 \begin{numlstlisting}[name=MSSM_MSSMRHN_two_scale_initial_guesser.cpp,language=C++]

### 2014-06-07 doc/paper.tex
for to -> to

@@
 with plenty of predefined models for the CMSSM, NMSSM, USSM, \ESSM,
-MRSSM etc., which can be generated immediately without any editing.  In particular the CMSSM
-and NMSSM spectrum generators constitute alternatives for to the
-existing ones Softsusy, SPheno and NMSPEC.  We have shown that the
-generated source code can be influenced at two different levels: The
-\mathematica level where the user provides a model file, and the C++
-level where the generated objects can be easily exchanged, extended,
-modified and reused.  This provides great flexibility for creating
-custom spectrum generators for both the most common and most
-extraordinary models.
+MRSSM etc., which can be generated immediately without any editing.
+In particular the CMSSM and NMSSM spectrum generators constitute
+alternatives to the existing ones Softsusy, SPheno and NMSPEC.  We
+have shown that the generated source code can be influenced at two
+different levels: The \mathematica level where the user provides a
+model file, and the C++ level where the generated objects can be
+easily exchanged, extended, modified and reused.  This provides great
+flexibility for creating custom spectrum generators for both the most
+common and most extraordinary models.
 

### 2014-06-07 doc/paper.tex
extended conclusions.

@@
 
-We presented \fs, a Mathematica package, which generates fast and
-modular spectrum generators for non-minimal SUSY models.  It comes
-with plenty of predefined models for the CMSSM, NMSSM, USSM, \ESSM,
-MRSSM etc., which can be generated immediately without any editing.
-In particular the CMSSM and NMSSM spectrum generators constitute
-alternatives to the existing ones Softsusy, SPheno and NMSPEC.  We
-have shown that the generated source code can be influenced at two
-different levels: The \mathematica level where the user provides a
+We have presented \fs, a Mathematica package, which generates fast and
+modular spectrum generators for any user specified SUSY model.  \fs is
+distributed with a large number of predefined models for the CMSSM,
+NMSSM, USSM, \ESSM, MRSSM etc., which can be generated immediately
+without any editing.  In particular the CMSSM and NMSSM spectrum
+generators constitute a fast and reliable alternative to the existing
+publicly available spectrum generators, Softsusy, SPheno and
+NMSPEC.
+
+We have described how the generated source code can be influenced at
+two different levels: The \mathematica level where the user provides a
 model file, and the C++ level where the generated objects can be
@@
 flexibility for creating custom spectrum generators for both the most
-common and most extraordinary models.
+common and most extraordinary models.  We have demonstrated these
+features in detailed examples for the NUHMSSM, right handed neutrinos
+and on adding three loop RGEs and two loop matching for the strong
+gauge coupling.
+
+The generated code has extensively tested against Softsusy, and
+additional tests have been carried out for non-minimal models, the
+\ESSM and MRSSM.  Speed tests have also been performed against
+Softsusy, SPheno and SPheno-like MSSM code generated by SARAH,
+demonstrating that \fs runs faster than all three.
+
+As a result \fs enables fast exploitation of new SUSY models with high
+precision and reliability.
 

### 2014-06-07 doc/paper.tex
doc/paper.tex: join a split command line
doc/paper.tex: correct a likely typo

@@
 \begin{lstlisting}[language=bash]
-$ models/MSSM/run_MSSM.x \
-    --slha-input-file=model_files/MSSM/LesHouches.in.MSSM
+$ models/MSSM/run_MSSM.x --slha-input-file=model_files/MSSM/LesHouches.in.MSSM
 \end{lstlisting}%% $
@@
 
-TODO: Adding tests against the CE$_6$SSM?  \cite{Athron:2009bs,Athron:2012pw}
+\fs generated NUH \ESSM has also been compared against a handwritten
+spectrum generator for a constrained version of the \ESSM
+\cite{Athron:2009bs,Athron:2012pw}.  The RGEs could be compared
+directly in unit tests and were found to match.  The handwritten code
+doesn't include full one loop self energies or tadpoles so tests on
+these were not carried.  Although the generators assume different
+constraints they could be compared by using the output of the
+CE$_6$SSM as input to \fs and the spectra were found to be in
+reasonable agreement, given the different levels of precision.
 
@@
 
-The generated code has extensively tested against Softsusy, and
+The generated code has been extensively tested against Softsusy, and
 additional tests have been carried out for non-minimal models, the

### 2014-06-07 sarah/myE6SSM/myE6SSM.m
removed sarah/myE6SSM from repository.

@@ -1,192 +1,0 @@
-Off[General::spell]
-
-Model`Name = "myE6MSSM";
-Model`NameLaTeX ="myE6MSSM";
-Model`Authors = "G.Hellwig, P.Diessner, P.Athron, A.Voigt";
-Model`Date = "2013-11-10";
-
-(*-------------------------------------------*)
-(*   Particle Content*)
-(*-------------------------------------------*)
-
-(* Global symmetries *)
-
-Global[[1]] = {Z[2],RParity};
-RpM = {-1,-1,1};
-RpP = {1,1,-1};
-
-(* Vector Superfields *)
-
-Gauge[[1]]={B,   U[1], hypercharge, g1,False,RpM};
-Gauge[[2]]={WB, SU[2], left,        g2,True, RpM};
-Gauge[[3]]={G,  SU[3], color,       g3,False,RpM};
-Gauge[[4]]={Bp,  U[1], Ncharge,    g1p,False,RpM};
-
-(* Chiral Superfields *)
-
-SuperFields[[1]] = {q, 3, {uL,  dL},     1/6, 2, 3, 1, RpM};  
-SuperFields[[2]] = {l, 3, {vL,  eL},    -1/2, 2, 1, 2, RpM};
-SuperFields[[3]] = {Hd,1, {Hd0, Hdm},   -1/2, 2, 1, -3, RpP};
-SuperFields[[4]] = {Hu,1, {Hup, Hu0},    1/2, 2, 1, -2, RpP};
-
-SuperFields[[5]] = {d, 3, conj[dR],    1/3, 1, -3, 2, RpM};
-SuperFields[[6]] = {u, 3, conj[uR],   -2/3, 1, -3, 1, RpM};
-SuperFields[[7]] = {e, 3, conj[eR],      1, 1,  1, 1, RpM};
-SuperFields[[8]] = {s, 1, sR,     0, 1,  1, 5, RpP};
-
-SuperFields[[9]] = {H1I, 2, {H1I0, H1Im},  -1/2, 2, 1, -3, RpP};
-SuperFields[[10]] = {H2I, 2, {H2Ip, H2I0},   1/2, 2, 1, -2, RpP};
-SuperFields[[11]] = {sI, 2, sIR,    0, 1,  1, 5, RpP};
[…]

### 2014-06-07 doc/paper.tex
added paragraph on testing against CE6SSM.
doc/paper.tex: relate a sentence better to the preceding one
update average run-time in the Program Summary

@@
          Eigen, LAPACK\\ {\em
-           Typical running time:}\/ 0.1-0.3 seconds per parameter
+           Typical running time:}\/ 0.06-0.2 seconds per parameter
          point.\\ {\em Nature of problem:}\/ Determining the mass
@@
 with the emphasis on the modular C++ code structure
-that helps programming.
+that helps such programming tasks.
 Concretely,

### 2014-06-07 doc/paper.tex
doc/paper.tex: replace MacOS by Mac OS X
Mac OS was not Unix-like until version 9.

@@
            C++, Wolfram/\mathematica, FORTRAN, Bourne shell}\\ {\em Computer:}\/ Personal computer\\ {\em Operating
-           system:}\/ Tested on Linux 3.x, MacOS\\
+           system:}\/ Tested on Linux 3.x, Mac OS X\\
          {\em External routines:}\/ SARAH 4.0.4, Boost library,
@@
 bugs in the MRSSM model file, as well as Ulrik Günther for compilation
-tests on MacOS.
+tests on Mac OS X\@.
 

### 2014-06-08 doc/paper.tex
discussing problems during the iteration

@@
 \end{figure}
+
+During the fixed-point iteration several problems can appear.  First
+of all, the iteration is not guaranteed to converge.  If the desired
+accuracy goal is not achieved with the given maximum number of
+iterations, \fs will set the \code{no_convergence} flag in the
+\code{Problems} class.  This class monitors the problem status of the
+spectrum generator during the iteration and can be obtained from the
+model class via the \code{get_problems()} function.  Besides
+non-convergence, solving the EWSB conditions
+\eqref{eq:one-loop-ewsb-eq} numerically with the desired accuracy
+might fail.  In this case the \code{no_ewsb} flag is set.
+Furthermore, in intermediate iteration steps tachyonic states might
+appear, which are ignored but nevertheless monitored in the
+\code{Problems} class.  If tachyons still exist after the iteration
+has converged the mass spectrum is marked as invalid by setting entry
+$4$ in the \code{SPINFO} block in the SLHA output file.  Finally,
+during the RG running some couplings might become non-perturbative.
+In this case the iteration stops setting the \code{no_perturbative}
+flag.
 

### 2014-06-08 doc/paper.tex
added more to acknowledgements.

@@
 A.V.\ would like to thank Florian Staub for countless explanations of
-SARAH's internals, discussions, and exceptionally fast bug fixing.
-Finally the authors thank Philip Diessner for testing and identifying
-bugs in the MRSSM model file, as well as Ulrik Günther for compilation
-tests on Mac OS X\@.
+SARAH's internals, discussions, and exceptionally fast bug
+fixing. P.A.~would like to thank Roman Nevzorov for useful discussions
+about challenges in non-minimal SUSY models. The authors would also
+like to thank Lewis Tunstall for helping with early tests against
+Next-to-Minimal Softsusy; Sophie Underwood for discovering problems
+introducing couplings wih a single family indice; Philip Diessner for
+testing and identifying bugs in the MRSSM model file; Ulrik Günther
+for compilation tests on Mac OS X\@ and Dylan Harries for spotting
+a bug in the configuration script.
 

### 2014-06-08 doc/paper.tex
fixed typos in last edit.

--- 2014-06-08 doc/paper.tex
@@
 A.V.\ would like to thank Florian Staub for countless explanations of
-SARAH's internals, discussions, and exceptionally fast bug fixing.
-Finally the authors thank Philip Diessner for testing and identifying
-bugs in the MRSSM model file, as well as Ulrik Günther for compilation
-tests on Mac OS X\@.
+SARAH's internals, discussions, and exceptionally fast bug
+fixing. P.A.~would like to thank Roman Nevzorov for useful discussions
+about challenges in non-minimal SUSY models. The authors would also
+like to thank Lewis Tunstall for helping with early tests against
+Next-to-Minimal Softsusy; Sophie Underwood for discovering problems
+introducing couplings with a single family index; Philip Diessner for
+testing and identifying bugs in the MRSSM model file; Ulrik Günther
+for compilation tests on Mac OS X\@ and Dylan Harries for spotting
+a bug in the configuration script.
 

### 2014-06-08 doc/paper.tex
added some dsicussion of problem issues. more to follow.

@@
 flag.
+
+It is important to note that in the case of such problem points it is
+non-trivial to judge whether this is because there is no physical
+solution for the given parameter space point or whether a solution
+exists but the fixed point iteration is unable to find the solution.
+While \fs makes it as easy as possible to fin spectra, when studying
+new models a physical understanding of the model is still essential
+and this can help the user determine why the such problems arise.
+
+Nonetheless \fs provides help for such cases in several ways. One may
+adjust initial guesses specified in the FlexibleSUSY modle file, such
+as changing the choice of HighScaleFirstGuess or altering
+HighScaleMinimum and HighScaleMaximum which can be used to push the
+iteration back towards where the solution should be if it gets off
+track.
+
+The modular design of FlexibleSUSY allows solver for the boudary
 

### 2014-06-08 doc/paper.tex
added stuff about E6SSM problems.  We cannot keep it in this form though. But wnated to put it in temporarily to help explain to Alex and Jae-hyeon.

@@
 
-The modular design of FlexibleSUSY allows solver for the boudary
+For experienced users the clear code structure also allows the
+possibility of direct adaption of the code.  For example in the case
+of the E6SSM we observed that for some points the presence of new
+Yuwawas had a big impact on the RGEs for the VEVS.  When the user
+chooses to set these Yukawas at the GUT scale, then on the initial run
+up to the GUT scale they will be set to zero.  As a result the VEV
+RGEs on the run up are quite different to the VEV RGEs on the run down
+so that the VEVs used at the SUSY scale are far from the correct
+values and this can create an instability leading eventually to
+non-perturbative in the couplings.  This can be fixed by adapting the
+algorithm so that thge VEVs are stored at the SUSY scale.  However a
+simpler, if less clean, fix is to directly modify the generated code
+by adding three lines to the file
+\code{<model>_two_scale_susy_parameters.cpp} just before the beta
+functions are returned in method
+\code{<model>_susy_parameters::calc_beta()}.
+
+\code{ 
+if(get_scale()  > 3.0e3) beta_vd =0;
+if(get_scale()  > 3.0e3) beta_vu =0;
+if(get_scale()  > 3.0e3) beta_vs =0;
+return E6SSMnew_susy_parameters(get_scale(), get_loops(), get_thresholds(), input,
+beta_Yd, beta_Ye, beta_Kappa, beta_Lambda12, beta_Lambdax, beta_Yu, beta_MuPr, beta_g1, beta_g2, beta_g3, beta_g1p, beta_vd, beta_vu, beta_vs);
+}
+
+In addition modular design of FlexibleSUSY allows solver for the boundary
 

### 2014-06-08 doc/paper.tex
added reference to jae-hyeon's lattice method as a possible solution to convergence problems.

@@
 
-In addition modular design of FlexibleSUSY allows solver for the boundary
+In addition the modular design of FlexibleSUSY allows for the solver
+for the boundary value problem to be replaced. An alternative solver
+with better convergence properties (at the expense of slower speed) is
+already planned for a later release.
 

### 2014-06-08 doc/paper.tex
added another sentcence regarding additional solver.

@@
 
-In addition the modular design of FlexibleSUSY allows for the solver
+Finally instead of tinkering with the two-scale solver one may wish to replace it entirely. the modular design of FlexibleSUSY allows for the solver
 for the boundary value problem to be replaced. An alternative solver

### 2014-06-08 doc/paper.tex
spell check.

@@
 Nonetheless \fs provides help for such cases in several ways. One may
-adjust initial guesses specified in the FlexibleSUSY modle file, such
-as changing the choice of HighScaleFirstGuess or altering
-HighScaleMinimum and HighScaleMaximum which can be used to push the
-iteration back towards where the solution should be if it gets off
-track.
+adjust initial guesses specified in the FlexibleSUSY model file, such
+as changing the choice of \code{HighScaleFirstGuess} or altering
+\code{HighScaleMinimum} and \code{HighScaleMaximum} which can be used
+to push the iteration back towards where the solution should be if it
+gets off track.
 
@@
 of the E6SSM we observed that for some points the presence of new
-Yuwawas had a big impact on the RGEs for the VEVS.  When the user
+Yukawas had a big impact on the RGEs for the VEVs.  When the user
 chooses to set these Yukawas at the GUT scale, then on the initial run
@@
 non-perturbative in the couplings.  This can be fixed by adapting the
-algorithm so that thge VEVs are stored at the SUSY scale.  However a
+algorithm so that the VEVs are stored at the SUSY scale.  However a
 simpler, if less clean, fix is to directly modify the generated code
@@
 
-Finally instead of tinkering with the two-scale solver one may wish to replace it entirely. the modular design of FlexibleSUSY allows for the solver
-for the boundary value problem to be replaced. An alternative solver
-with better convergence properties (at the expense of slower speed) is
-already planned for a later release.
+Finally instead of tinkering with the two-scale solver one may wish to
+replace it entirely. The modular design of FlexibleSUSY allows for the
+solver for the boundary value problem to be replaced. An alternative
+solver with better convergence properties (at the expense of slower
+speed) is already planned for a later release.
 

### 2014-06-08 doc/paper.tex
removed detailed explanation of VEV problem. I think there is no way to include this, but perhaps Jae-hyeon or Alex can suggest a cleanerf fix that is more suitable.

@@
 to push the iteration back towards where the solution should be if it
-gets off track.
-
-For experienced users the clear code structure also allows the
-possibility of direct adaption of the code.  For example in the case
-of the E6SSM we observed that for some points the presence of new
-Yukawas had a big impact on the RGEs for the VEVs.  When the user
-chooses to set these Yukawas at the GUT scale, then on the initial run
-up to the GUT scale they will be set to zero.  As a result the VEV
-RGEs on the run up are quite different to the VEV RGEs on the run down
-so that the VEVs used at the SUSY scale are far from the correct
-values and this can create an instability leading eventually to
-non-perturbative in the couplings.  This can be fixed by adapting the
-algorithm so that the VEVs are stored at the SUSY scale.  However a
-simpler, if less clean, fix is to directly modify the generated code
-by adding three lines to the file
-\code{<model>_two_scale_susy_parameters.cpp} just before the beta
-functions are returned in method
-\code{<model>_susy_parameters::calc_beta()}.
-
-\code{ 
-if(get_scale()  > 3.0e3) beta_vd =0;
-if(get_scale()  > 3.0e3) beta_vu =0;
-if(get_scale()  > 3.0e3) beta_vs =0;
-return E6SSMnew_susy_parameters(get_scale(), get_loops(), get_thresholds(), input,
-beta_Yd, beta_Ye, beta_Kappa, beta_Lambda12, beta_Lambdax, beta_Yu, beta_MuPr, beta_g1, beta_g2, beta_g3, beta_g1p, beta_vd, beta_vu, beta_vs);
-}
+gets off track.For experienced users the clear code structure also allows the
+possibility of direct adaption of the code.
 

### 2014-06-08 doc/paper.tex
doc/paper.tex: add a little Zurückhaltung
fix two typos

@@
 exists but the fixed point iteration is unable to find the solution.
-While \fs makes it as easy as possible to fin spectra, when studying
+While \fs makes it as easy as possible to find spectra, when studying
 new models a physical understanding of the model is still essential
@@
 to push the iteration back towards where the solution should be if it
-gets off track.For experienced users the clear code structure also allows the
+gets off track. For experienced users the clear code structure also allows the
 possibility of direct adaption of the code.
@@
 solver for the boundary value problem to be replaced. An alternative
-solver with better convergence properties (at the expense of slower
+solver with potentially better convergence properties (at the expense of slower
 speed) is already planned for a later release.

### 2014-06-08 doc/paper.tex
make modifications suggested by Dominik

@@
 user to exchange or add boundary conditions at any scale.
-To realized this, all (derived) constraint objects are intentionally
+To realize this, all (derived) constraint objects are intentionally
 kept outside the solver.  Despite being independent of one another,
@@
 while undergoing a minimal amount of frustration,
-when he or she programs for a wide variety of studies.
+when one programs for a wide variety of studies.
 We provide two basic levels for the user to create a custom spectrum
-generator: (i) The \mathematica level, where he or she writes or
+generator: (i) The \mathematica level, where one writes or
 adapts a model file and (ii) the C++ level, where the generated
@@
 
-Since the current implementation of \fs does not (yet) automatically
-produce a spectrum generator spanning multiple models,
-one needs to write the matching class
-% in \figref{fig:class structure for tower}
-as well as gluing codes by hand to build such a program.
-One can author nearly all the components of a multi-model spectrum generator
+The matching class as well as gluing codes have to be written by hand
+to build such a program.\footnote{It is planned that a future release
+  of \fs will be capable of creating this code automatically.}  All
+remaining components of a multi-model spectrum generator can be authored
 by making a straightforward extension
@@
 for one of the models forming the tower.
-The matching class, however, needs to be set-up by hand, because it
-has no analogue in a single-model case.
 
@@
 testing and identifying bugs in the MRSSM model file; Ulrik Günther
-for compilation tests on Mac OS X\@ and Dylan Harries for spotting
+for compilation tests on Mac OS X and Dylan Harries for spotting
 a bug in the configuration script.
+%
+This work has been supported by the German Research Foundation DFG
+through Grant No.~STO876/2-1.
 

### 2014-06-09 doc/paper.tex
doc/paper.tex: fix typos

@@
 so-called spectrum generators can be transferred to programs which
-calculate further observables such as branching rations or the dark
+calculate further observables such as branching ratios or the dark
 matter relic density.
@@
 
-Using the the variable \code{EWSBOutputParameters} the user can
+Using the variable \code{EWSBOutputParameters} the user can
 specify the model parameters that are output of the electroweak
@@
   \code{function} is a function of these parameters.
-  \code{FSMinimize[parameters, function]} will numerical vary the
+  \code{FSMinimize[parameters, function]} will numerically vary the
   \code{parameters} until the \code{function} is minimized.  Example:
@@
 defining the effective $\mu$-term like \code{EffectiveMu = \\[Lambda]
-  vS / Sqrt[2]}, for example.  This will add the the zero-momentum
+  vS / Sqrt[2]}, for example.  This will add the zero-momentum
 corrections of the order $O(y_t^2 g_3^2)$, $O(y_b^2 g_3^2)$ from
@@
   Standard Model quantities
-  $\alpha_{\text{e.m.},\text{SM}}^{(5),\text{\MSbar}}(M_Z)$
+  $\alpha_{\text{e.m.},\text{SM}}^{(5),\text{\MSbar}}(M_Z)$,
   $\alpha_{\text{s},\text{SM}}^{(5),\text{\MSbar}}(M_Z)$, $M_Z$ and
@@
 achieved, all \DRbar\ parameters consistent with the EWSB conditions,
-low energy data and all user-supplied boundary conditions and are
+low energy data and all user-supplied boundary conditions are
 known at any scale between \code{LowScale} and \code{HighScale}.
@@
   \end{tabular}
-  \caption{Average run-time of CMSSM spectrum generators for
+  \caption{Average run-time of CMSSM spectrum generators
     for random parameter points.  The first three rows show

### 2014-06-09 doc/paper.tex
doc/paper.tex: conform to -ize convention
doc/paper.tex: match firstly ... secondly ...

@@
          coupled first-order differential equations. \\ {\em Solution method:}\/
-         Nested iterative algorithm and numerical minimisation of the
+         Nested iterative algorithm and numerical minimization of the
          Higgs potential.  \\ {\em Restrictions:} The couplings must
@@
 Hadron Collider (LHC) there have been two important developments.
-First low energy signatures expected from such models, such as the
+Firstly low energy signatures expected from such models, such as the
 classic jets plus missing transverse energy signature, have not been

### 2014-06-09 doc/paper.tex
doc/paper.tex: change capitalizations
See http://www.boost.org .

@@
 \cite{eigen}, augmented by LAPACK, as well as the GNU scientific
-library and the BOOST library to create numerical routines which solve
+library and the Boost library to create numerical routines which solve
 the RGEs and boundary conditions simultaneously.  If a solution has
@@
 color charged fields absent from the Standard Model.  The constants
-$T_i$ are the dynkin indices of the representation of particle~$i$
+$T_i$ are the Dynkin indices of the representation of particle~$i$
 with respect to the gauge group, and $F_i$ are particle-type specific

### 2014-06-09 doc/paper.tex
doc/paper.tex: trim optional trailing / from links

@@
   \url{http://www.boost.org}
-\item GNU scientific library \url{http://www.gnu.org/software/gsl/}
-\item an implementation of LAPACK \url{http://www.netlib.org/lapack/}
-  such as ATLAS \url{http://math-atlas.sourceforge.net/} or
+\item GNU scientific library \url{http://www.gnu.org/software/gsl}
+\item an implementation of LAPACK \url{http://www.netlib.org/lapack}
+  such as ATLAS \url{http://math-atlas.sourceforge.net} or
   Intel Math Kernel Library \url{http://software.intel.com/intel-mkl}
@@
 \item Looptools, version 2.8 or higher
-  \url{http://www.feynarts.de/looptools/}
+  \url{http://www.feynarts.de/looptools}
 \end{itemize}
@@
 \fs can be downloaded as a gzipped tar file from
-\url{http://flexiblesusy.hepforge.org/}.  To download and install
+\url{http://flexiblesusy.hepforge.org}.  To download and install
 version 1.0.0 run:

### 2014-06-09 doc/paper.tex
doc/paper.tex: trim optional trailing / from argument to cd

@@
 $ tar -xf FlexibleSUSY-1.0.0.tar.gz
-$ cd FlexibleSUSY-1.0.0/
+$ cd FlexibleSUSY-1.0.0
 \end{lstlisting}%% $

### 2014-06-09 doc/paper.tex
doc/paper.tex: minor changes
Command line is not an output channel but an input.

@@
 \code{model_files/MSSM/LesHouches.in.MSSM}.  Afterwards, the pole mass
-spectrum and mixing matrices are calculated and written to the command
-line in SLHA format \cite{Skands:2003cj,Allanach:2008qq}.  For the
+spectrum and mixing matrices are calculated and written to the standard
+output in SLHA format \cite{Skands:2003cj,Allanach:2008qq}.  For the
 parameter point given in the above example the calculated pole mass
@@
   Here, the parameters \code{vu} and \code{vd} are varied until the
-  vector-like function
+  vector-valued function
   %
@@
 chargino, gluino, sneutrino, up-type squark, down-type squark,
-selectron) are considered to be LSPs.
+selectron) are considered to be LSP candidates.
 
@@
 Since this iteration can be very time consuming for large field
-multiplets, \fs provides two approximative procedures with a faster
+multiplets, \fs provides two approximative procedures with a shorter
 run-time.  The used procedure can be set in the model file for each
@@
   the eigenvalues are interpreted as pole masses $m_{f_i}^\pole$.
-  This method neglects which are formally of two-loop order.  This
+  This method neglects what are formally of two-loop order.  This
   method is imprecise if the self-energy corrections to the mass

### 2014-06-09 doc/paper.tex
doc/paper.tex: add/correct cross-references

@@
 the low-energy boundary condition class.
-It must be declared as a descendent of \code{Constraint<Two_scale>}:
+It must be declared as a descendent of \code{Constraint<Two_scale>}
+whose function is described in \secref{sec:boundary-conditions}:
 % MSSMcbs_two_scale_low_scale_constraint.hpp
@@
 \subsection{Run-time comparison}
+\label{sec:run-time comparison}
 
@@
 SLHA template file given in
-\secref{sec:speed-test-slha-template-file}.  The resulting SLHA input
+\ref{sec:speed-test-slha-template-file}.  The resulting SLHA input
 file is passed to each spectrum generator and the (wall-clock) time is
@@
 SARAH generated MSSM spectrum generator SPhenoMSSM.  The test results
-can be found in \secref{sec:comparison}.
+can be found in \secref{sec:run-time comparison}.
 

### 2014-06-09 doc/paper.tex
added key words 'renormalization group equations' as per Jae-hyeon suggestion.
removed reference to FlexibleSUSY bugs in section 8 as per Jae-hyeon suggestion.

@@
 supersymmetry, 
-Higgs
+Higgs,
+renormalization group equations
 \PACS 12.60.Jv
@@
 low-energy model, MRSSM, were done by Philip Diessner and several bugs
-in \fs and SARAH were identified and fixed.
+in SARAH were identified and fixed.
 
@@
 constraints they could be compared by using the output of the
-CE$_6$SSM as input to \fs and the spectra were found to be in
+CE$_6$SSM generator as an input to \fs and the spectra were found to be in
 reasonable agreement, given the different levels of precision.

### 2014-06-09 doc/paper.tex
Changed obvious instance of FlexibleSUSY.m to FlexibleSUSY.m.in. Thinking about how to calrify distinction elsewhere.

@@
 constraint and use a separate value for $A_\lambda$ at the GUT scale
-one can edit the model file \code{models/NMSSM/FlexibleSUSY.m} and
+one can edit the model file \code{model_files/NMSSM/FlexibleSUSY.m.in} and
 change the lines

### 2014-06-09 doc/paper.tex
added decription of how FlexibleSUSY takes FlexibleSUSY.m.in fle nad creates FlexibleSUSY.m file to section 5. Alos changed the CMSSM file quoted in paper to be FlexiblleSUSY.m.in file, replacing first line.

@@
 in form of \sarah model files.  The boundary conditions on the model
-parameters must be specified in a separate \code{FlexibleSUSY.m} file.
+parameters must be specified in a separate \code{FlexibleSUSY.m.in} file.
 Based on this information \fs uses \sarah to obtain tree-level
@@
 two-loop corrections to the Higgs masses\footnote{By default
-  FlexibleSUSY has two-loop corrections to the Higgs masses for the
+  \fs has two-loop corrections to the Higgs masses for the
   MSSM
@@
 explanation of how to write such a model file.  Note that SARAH
-already comes with a lot of predefined models, which can be used with
+already is distributed with a lot of predefined models, which can be used with
 \fs immediately.
@@
 \code{FlexibleSUSY.m}, which has to be located in the model directory
-\code{models/<model>/}.  Note that many example model files can
-already be found in \code{model_files/}.  In the following it is
-explained how the boundary conditions can be defined on the basis of
-the CMSSM.  The application to other models is straightforward.  The
-CMSSM model file reads:
+\code{models/<model>/}. To add this the user should create a
+\code{FlexibleSUSY.m} in the directory \code{model_files/<model>/}.
+When the \code{./createmodel} command is used Flexible SUSY will create the
+FlexibleSUSY.m file from the FlexibleSUSY.m.in file found in the
+directory in \code{model_files} with the name specified by the
+\code{--model-file=<model-file-name>} option. If no such option is
+given then it will use the directory matching the option
+\code{--name=<model>}.  In either case the the FlexibleSUSY.m file
+which is created is then automatically placed in the directory
+\code{models/<model>/} by \fs. Note that many example model
+files can already be found in \code{model_files/<model>/}.
+
+In the following it is explained how the
+boundary conditions can be defined on the basis of the CMSSM.  The
+application to other models is straightforward.  The CMSSM model file
+reads:
 %
 \begin{lstlisting}[language=Mathematica]
-FSModelName = "MSSM";
+FSModelName = "@CLASSNAME@";
 
@@
 %
-The variable \code{FSModelName} contains the name of the \fs model.
+The first line \code{FSModelName = "@CLASSNAME@";} will be replaced
+with \code{FSModelName = "<model>";} in the generated FlexibleSUSY.m
+file where \code{<model>} is given by the option \code{--name=<model>} for
+command \code{./createmodel}.  So the variable \code{FSModelName} then
+contains the name of the \fs model.
 

### 2014-06-09 doc/paper.tex
FlexibleSUSY --> \fs in some places to make changing notation easier, if we ever wish to do so.

@@
 how to get started quickly.  In \secref{sec:modfile} we describe how
-the user can create a new FlexibleSUSY model file. A detailed
+the user can create a new \fs model file. A detailed
 description of the structure and features of the generated code is
@@
 Nonetheless \fs provides help for such cases in several ways. One may
-adjust initial guesses specified in the FlexibleSUSY model file, such
+adjust initial guesses specified in the \fs model file, such
 as changing the choice of \code{HighScaleFirstGuess} or altering
@@
 Finally instead of tinkering with the two-scale solver one may wish to
-replace it entirely. The modular design of FlexibleSUSY allows for the
+replace it entirely. The modular design of \fs allows for the
 solver for the boundary value problem to be replaced. An alternative
@@
 %
-FlexibleSUSY and Softsusy are compiled with g++ 4.8.0 and Intel ifort
+\fs and Softsusy are compiled with g++ 4.8.0 and Intel ifort
 13.1.3 20130607.  SPheno and SPhenoMSSM are compiled with Intel ifort

### 2014-06-09 sarah/E6SSM/E6SSM.m
deleted E6SSM files then replaced with new E6SSM FlexibleSUSY file for GUT normalised charges.

@@ -1,191 +1,0 @@
-Off[General::spell]
-
-Model`Name = "E6SSM";
-Model`NameLaTeX ="E6SSM";
-Model`Authors = "G.Hellwig, P.Diessner, P.Athron, A.Voigt";
-Model`Date = "2014-03-10";
-
-(*-------------------------------------------*)
-(*   Particle Content*)
-(*-------------------------------------------*)
-
-(* Global symmetries *)
-
-Global[[1]] = {Z[2],RParity};
-RpM = {-1,-1,1};
-RpP = {1,1,-1};
-
-(* Vector Superfields *)
-
-Gauge[[1]]={B,   U[1], hypercharge, g1,False,RpM};
-Gauge[[2]]={WB, SU[2], left,        g2,True, RpM};
-Gauge[[3]]={G,  SU[3], color,       g3,False,RpM};
-Gauge[[4]]={Bp,  U[1], Ncharge,    g1p,False,RpM};
-
-(* Chiral Superfields *)
-
-SuperFields[[1]] = {q, 3, {uL,  dL},     1/6, 2, 3, 1, RpM};  
-SuperFields[[2]] = {l, 3, {vL,  eL},    -1/2, 2, 1, 2, RpM};
-SuperFields[[3]] = {Hd,1, {Hd0, Hdm},   -1/2, 2, 1, -3, RpP};
-SuperFields[[4]] = {Hu,1, {Hup, Hu0},    1/2, 2, 1, -2, RpP};
-
-SuperFields[[5]] = {d, 3, conj[dR],    1/3, 1, -3, 2, RpM};
-SuperFields[[6]] = {u, 3, conj[uR],   -2/3, 1, -3, 1, RpM};
-SuperFields[[7]] = {e, 3, conj[eR],      1, 1,  1, 1, RpM};
-SuperFields[[8]] = {s, 1, sR,     0, 1,  1, 5, RpP};
-
-SuperFields[[9]] = {H1I, 2, {H1I0, H1Im},  -1/2, 2, 1, -3, RpP};
-SuperFields[[10]] = {H2I, 2, {H2Ip, H2I0},   1/2, 2, 1, -2, RpP};
-SuperFields[[11]] = {sI, 2, sIR,    0, 1,  1, 5, RpP};
[…]

### 2014-06-09 sarah/E6SSM/parameters.m

@@ -1,347 +1,0 @@
-ParameterDefinitions = {
-{g1,        { Description -> "Hypercharge-Coupling"}},
-{g2,        { Description -> "Left-Coupling"}},
-{g3,        { Description -> "Strong-Coupling"}},
-{g1p,        { Description -> "Ncharge-Coupling",
-                 Dependence -> None,
-	         LaTeX -> "g_1'",
-		 GUTnormalization -> 1/Sqrt[40],
-		 LesHouches -> {gauge,4},
-	         OutputName -> g1p}},
-{g1gN,        {Description -> "Mixed Gauge Coupling 1"}},
-{gNg1,        {Description -> "Mixed Gauge Coupling 2"}},
-{AlphaS,       { Description -> "Alpha Strong"}},
-{aEWinv,       { Description -> "inverse weak coupling constant at mZ"}},
-{Gf,           { Description -> "Fermi's constant"}},
-{e,            { Description -> "electric charge"}},
-	       
-{Yu,           { Description -> "Up-Yukawa-Coupling",
-	         LaTeX -> "Y_u",
-		 Form -> Diagonal,
-	         Dependence -> None,
-	         Value -> None,
-		 OutputName -> Yu}},
-{Yd,           { Description -> "Down-Yukawa-Coupling",
-	         LaTeX -> "Y_d",
-		 Form -> Diagonal,
-	         Dependence -> None,
-	         Value -> None,
-		 OutputName -> Yd}},
-{Ye,           { Description -> "Lepton-Yukawa-Coupling",
-	         LaTeX -> "Y_e",
-		 Form -> Diagonal,
-	         Dependence -> None,
-	         Value -> None,
-		 OutputName -> Ye}},
-{T[Ye],        { Description -> "Trilinear-Lepton-Coupling",
-		 Form -> Diagonal,
-	         Dependence -> None,
-	         Value -> None}},
-{T[Yd],        { Description -> "Trilinear-Down-Coupling",
-		 Form -> Diagonal,
-	         Dependence -> None,
-	         Value -> None}},
-{T[Yu],        { Description -> "Trilinear-Up-Coupling",
-		 Form -> Diagonal,
-	         Dependence -> None,
-	         Value -> None}},	       
-{\[Mu],        { Description -> "Mu-parameter",
-      Dependence -> \[Lambda]*vs/Sqrt[2]}},
-{B[\[Mu]],     { Description -> "Bmu-parameter",
[…]

### 2014-06-09 sarah/E6SSM/E6SSM_altcharges.m
added sarah/E6SSM files from GUT normalised version.

@@ -1,0 +1,191 @@
+Off[General::spell]
+
+Model`Name = "E6SSM_altchatges";
+Model`NameLaTeX ="E6SSM_altcharges";
+Model`Authors = "G.Hellwig, P.Diessner, P.Athron, A.Voigt";
+Model`Date = "2014-03-10";
+
+(*-------------------------------------------*)
+(*   Particle Content*)
+(*-------------------------------------------*)
+
+(* Global symmetries *)
+
+Global[[1]] = {Z[2],RParity};
+RpM = {-1,-1,1};
+RpP = {1,1,-1};
+
+(* Vector Superfields *)
+
+Gauge[[1]]={B,   U[1], hypercharge, g1,False,RpM};
+Gauge[[2]]={WB, SU[2], left,        g2,True, RpM};
+Gauge[[3]]={G,  SU[3], color,       g3,False,RpM};
+Gauge[[4]]={Bp,  U[1], Ncharge,    gN,False,RpM};
+
+(* Chiral Superfields *)
+
+SuperFields[[1]] = {q, 3, {uL,  dL},     1/6, 2, 3, 1/Sqrt[40], RpM};  
+SuperFields[[2]] = {l, 3, {vL,  eL},    -1/2, 2, 1, 2/Sqrt[40], RpM};
+SuperFields[[3]] = {Hd,1, {Hd0, Hdm},   -1/2, 2, 1, -3/Sqrt[40], RpP};
+SuperFields[[4]] = {Hu,1, {Hup, Hu0},    1/2, 2, 1, -2/Sqrt[40], RpP};
+
+SuperFields[[5]] = {d, 3, conj[dR],    1/3, 1, -3, 2/Sqrt[40], RpM};
+SuperFields[[6]] = {u, 3, conj[uR],   -2/3, 1, -3, 1/Sqrt[40], RpM};
+SuperFields[[7]] = {e, 3, conj[eR],      1, 1,  1, 1/Sqrt[40], RpM};
+SuperFields[[8]] = {s, 1, sR,     0, 1,  1, 5/Sqrt[40], RpP};
+
+SuperFields[[9]] = {H1I, 2, {H1I0, H1Im},  -1/2, 2, 1, -3/Sqrt[40], RpP};
[…]

### 2014-06-09 sarah/E6SSM/parameters.m

@@ -1,0 +1,346 @@
+ParameterDefinitions = {
+{g1,        { Description -> "Hypercharge-Coupling"}},
+{g2,        { Description -> "Left-Coupling"}},
+{g3,        { Description -> "Strong-Coupling"}},
+{gN,        { Description -> "Ncharge-Coupling",
+                 Dependence -> None,
+                 LaTeX -> "g_N",
+		 LesHouches -> {gauge,4},
+	         OutputName -> gN}},
+{g1gN,        {Description -> "Mixed Gauge Coupling 1"}},
+{gNg1,        {Description -> "Mixed Gauge Coupling 2"}},
+{AlphaS,       { Description -> "Alpha Strong"}},
+{aEWinv,       { Description -> "inverse weak coupling constant at mZ"}},
+{Gf,           { Description -> "Fermi's constant"}},
+{e,            { Description -> "electric charge"}},
+	       
+{Yu,           { Description -> "Up-Yukawa-Coupling",
+	         LaTeX -> "Y_u",
+		 Form -> Diagonal,
+	         Dependence -> None,
+	         Value -> None,
+		 OutputName -> Yu}},
+{Yd,           { Description -> "Down-Yukawa-Coupling",
+	         LaTeX -> "Y_d",
+		 Form -> Diagonal,
+	         Dependence -> None,
+	         Value -> None,
+		 OutputName -> Yd}},
+{Ye,           { Description -> "Lepton-Yukawa-Coupling",
+	         LaTeX -> "Y_e",
+		 Form -> Diagonal,
+	         Dependence -> None,
+	         Value -> None,
+		 OutputName -> Ye}},
+{T[Ye],        { Description -> "Trilinear-Lepton-Coupling",
+		 Form -> Diagonal,
+	         Dependence -> None,
+	         Value -> None}},
+{T[Yd],        { Description -> "Trilinear-Down-Coupling",
+		 Form -> Diagonal,
+	         Dependence -> None,
+	         Value -> None}},
+{T[Yu],        { Description -> "Trilinear-Up-Coupling",
+		 Form -> Diagonal,
+	         Dependence -> None,
+	         Value -> None}},	       
+{\[Mu],        { Description -> "Mu-parameter",
+      Dependence -> \[Lambda]*vs/Sqrt[2]}},
+{B[\[Mu]],     { Description -> "Bmu-parameter",
+      Dependence -> T[\[Lambda]]*vs/Sqrt[2]}},	       
[…]

### 2014-06-09 sarah/E6SSM/particles.m

@@ -1,0 +1,455 @@
+
+ParticleDefinitions[GaugeES] = {
+      {SdL,  {  Description -> "Left Down-Squarks"}},
+      {SdR,  { Description -> "Right Down-Squarks"}},
+      {SuL,  { Description -> "Left Up-Squarks"}}, 
+      {SuR,  { Description -> "Right Up-Squarks" }},                
+      {SeL,  { Description -> "Left Selectron"}}, 
+      {SeR,  { Description -> "Right Selectron"}},                         
+      {SvL,  { Description -> "Left Sneutrino"}},                         
+      {SHd0, { Description -> "Neutral Down-Higgs"}},                         
+      {SHdm, { Description -> "Charged Down-Higgs"}}, 
+      {SHu0, { Description -> "Neutral Up-Higgs"}},                         
+      {SHup, { Description -> "Charged Up-Higgs"}},
+      {SsR,  { Description -> "Singlet"}},            
+      {SDxL,        {Description -> "Left SExotics",
+                     FeynArtsNr -> 666,
+	             LaTeX -> "\\tilde{Dx}_L",
+                     OutputName -> "SDxL"}},
+      {SDxbarR,     {Description -> "Right SExotics",
+	             FeynArtsNr -> 667,
+	             LaTeX -> "\\tilde{Dx}_R",
+                     OutputName -> "SDxbR"}},
+      {SH1I0,       {Description -> "Neutral Inert-Down-Higgs",
+  		     FeynArtsNr -> 101,
+  		     LaTeX -> "h^{0Inert}_{1}",
+                     OutputName -> "SH1I0" }},
+      {SH1Im,  {     Description -> "Charged Inert-Down-Higgs",
+                     FeynArtsNr -> 102,
+                     LaTeX -> "h^{-Inert}_{1}",
+                     OutputName -> "SH1Im"}},
+      {SH2I0,  {     Description -> "Neutral Inert-Up-Higgs",
+  		     FeynArtsNr -> 103,
+  		     LaTeX -> "h^{0Inert}_{2}",
+                     OutputName -> "SH2I0"}},
+      {SH2Ip,  { Description -> "Charged Inert-Up-Higgs",
+  		     FeynArtsNr -> 104,
+  		     LaTeX -> "h^{+Inert}_{2}",
+                     OutputName -> "SH2Ip"}},
+      {SsIR,   { Description -> "Inert-Singlet",
+  		     FeynArtsNr -> 105,
[…]

### 2014-06-09 sarah/E6SSM/E6SSM.m
renamed sarah/E6SSM/E6SSM.m and changed model name inside accordingly.

@@ -1,7 +1,7 @@
 Off[General::spell]
 
-Model`Name = "E6SSM_altchatges";
-Model`NameLaTeX ="E6SSM_altcharges";
+Model`Name = "E6SSM";
+Model`NameLaTeX ="E6SSM";
 Model`Authors = "G.Hellwig, P.Diessner, P.Athron, A.Voigt";
 Model`Date = "2014-03-10";
 

### 2014-06-09 doc/paper.tex
intergenerational -> inter-generational

@@
 matrices, \code{ZE} and \code{ZV},
-which contain intergenerational mixings apart from the
+which contain inter-generational mixings apart from the
 generic left-right mixings.

### 2014-06-09 doc/paper.tex
finished full spell check.
explain meaning of vT in Eq. (15)

@@
     \left(\frac{M_{W,\text{susy}}^{\text{\DRbar}}(M_Z)}{M_{Z,\text{susy}}^{\text{\DRbar}}(M_Z)}\right)^2}
-  ,
+  .
 \end{align}
 %
-while in models with Higgs triplets the relation looks like
+In a model with a Higgs triplet the relation looks like
 %
@@
     \frac{\left(M_{W,\text{susy}}^{\text{\DRbar}}(M_Z)\right)^2 -
-      g_2^2v_T^2}{\left(M_{Z,\text{susy}}^{\text{\DRbar}}(M_Z)\right)^2}}.
+      g_2^2v_T^2}{\left(M_{Z,\text{susy}}^{\text{\DRbar}}(M_Z)\right)^2}},
 \end{align}
 %
+where $v_T$ is the vacuum expectation value of the scalar Higgs
+triplet field.
 The running \DRbar\ $W$ and $Z$ boson masses are calculated in each
@@
 the low-energy boundary condition class.
-It must be declared as a descendent of \code{Constraint<Two_scale>}
+It must be declared as a descendant of \code{Constraint<Two_scale>}
 whose function is described in \secref{sec:boundary-conditions}:

### 2014-06-09 doc/paper.tex
clarify the three categories of model parameters
We have now the following three terms:

  "SUSY parameters" := gauge couplings, VEVs,
                       superpotential parameters,

  "soft-breaking parameters := Girardello, Grisaru terms

  "all model parameters" := SUSY parameters + soft-breaking parameters

@@
 
-The parameters of the SUSY model together with their RGEs, mass
+The parameters of the model together with their RGEs, mass
 matrices, self-energies and EWSB equations are stored at the C++ level
@@
 
-All SUSY model parameters and their $\beta$-functions are contained in
+All model parameters and their $\beta$-functions are contained in
 the first and second derived classes.  The structure of the
-$\beta$-functions of a general SUSY model
+$\beta$-functions of a general supersymmetric model
 \cite{Jones:1974pg,Jones:1983vk,West:1984dg,Martin:1993yx,Yamada:1993ga,MV94,Fonseca:2011vn,Yam94,Sperling:2013eva,Sperling:2013xqa}
@@
 corresponding \code{beta()} functions.  These two classes then allow
-to use renormalization group running of all SUSY model parameters.
+to use renormalization group running of all model parameters.
 

### 2014-06-09 doc/paper.tex
add missing \code{} wrapper commands

@@
 \code{models/<model>/}. To add this the user should create a
-\code{FlexibleSUSY.m} in the directory \code{model_files/<model>/}.
-When the \code{./createmodel} command is used Flexible SUSY will create the
-FlexibleSUSY.m file from the FlexibleSUSY.m.in file found in the
+\code{FlexibleSUSY.m.in} in the directory \code{model_files/<model>/}.
+When the \code{./createmodel} command is used \fs will create the
+\code{FlexibleSUSY.m} file from the \code{FlexibleSUSY.m.in} file found in the
 directory in \code{model_files} with the name specified by the
@@
 given then it will use the directory matching the option
-\code{--name=<model>}.  In either case the the FlexibleSUSY.m file
+\code{--name=<model>}.  In either case the the \code{FlexibleSUSY.m} file
 which is created is then automatically placed in the directory
 \code{models/<model>/} by \fs. Note that many example model
-files can already be found in \code{model_files/<model>/}.
+files can already be found in \code{model_files/}.
 
@@
 The first line \code{FSModelName = "@CLASSNAME@";} will be replaced
-with \code{FSModelName = "<model>";} in the generated FlexibleSUSY.m
+with \code{FSModelName = "<model>";} in the generated \code{FlexibleSUSY.m}
 file where \code{<model>} is given by the option \code{--name=<model>} for

### 2014-06-09 doc/paper.tex
remove \JH* macros
FS is a "Mathematica and C++ package"

@@
 \renewcommand{\ptitle}[1]{}
-\newcommand{\JHtxt}[1]{\textcolor[rgb]{1,.3,0}{\texttt{#1}}}
-\newcommand{\JHrem}[1]{\textcolor[rgb]{1,.3,0}{\textbf{{\textsl{\boldmath #1}}}}}
 \makeatletter
@@
   \begin{abstract}
-    We introduce \fs, a \mathematica package, which generates a fast,
+    We introduce \fs, a \mathematica and C++ package, which generates a fast,
     precise C++ spectrum generator for any SUSY model specified by the
@@
 possible and to be easily adaptable to changing goals and new
-ideas. \fs is a \mathematica package which uses \sarah to create a
+ideas. \fs is a \mathematica and C++ package which uses \sarah to create a
 fast, modular C++ spectrum generator for a user specified SUSY model.
@@
 
-\fs is a \mathematica package designed to create a fast and easily
+\fs is a \mathematica and C++ package designed to create a fast and easily
 adaptable spectrum generator in C++ for any SUSY model.
@@
 
-We have presented \fs, a Mathematica package, which generates fast and
+We have presented \fs, a Mathematica and C++ package, which generates fast and
 modular spectrum generators for any user specified SUSY model.  \fs is

### 2014-06-09 doc/paper.tex
correcting citations for the two-loop tadpoles in the MSSM and NMSSM

@@
 corrections can be added by setting \code{UseHiggs2LoopMSSM = True}
-in the model file.  For example in the MSSM by default \fs adds two
+in the model file.  For example in the MSSM by default \fs adds
 two-loop Higgs FORTRAN routines supplied by P.~Slavich from
-\cite{Degrassi:2001yf,Brignole:2001jy,Dedes:2002dy,Brignole:2002bz,Dedes:2003km}
-to add two loop corrections of $\oatas$, $\oabas$, $\oatq$,
-$\oabq$, $\oatauq$ and $\oatab$.  In the NMSSM
-contributions of the order $\oatas$, $\oabas$ \cite{Degrassi:2009yq}
+\cite{Dedes:2002dy,Dedes:2003km}
+to add two-loop corrections of $\oatas$, $\oabas$, $\oatq$,
+$\oabq$, $\oatauq$ and $\oatab$.  In the NMSSM the same contributions
 can be added by setting \code{UseHiggs2LoopNMSSM = True} in the model

### 2014-06-09 meta/FlexibleSUSY.m
adding references to publication on arXiv with temporary arXiv number

@@ -4,11 +4,15 @@ BeginPackage["FlexibleSUSY`", {"SARAH`", "AnomalousDimension`", "BetaFunction`",
 FS`Version = StringTrim[Import[FileNameJoin[{Global`$flexiblesusyConfigDir,"version"}], "String"]];
 FS`Authors = {"P. Athron", "Jae-hyeon Park", "D. Stöckinger", "A. Voigt"};
 FS`Years   = {2013, 2014};
+FS`References = Get[FileNameJoin[{Global`$flexiblesusyConfigDir,"references"}]];
 
 Print["*****************************************************************"];
 Print["FlexibleSUSY ", FS`Version];
 Print["by " <> WriteOut`StringJoinWithSeparator[FS`Authors, ", "] <> ", " <>
       WriteOut`StringJoinWithSeparator[FS`Years, ", "]];
+Print[""];
+Print["References:"];
+Print["  " <> #]& /@ FS`References;
 Print["*****************************************************************"];
 Print[""];
 

### 2014-06-09 doc/paper.tex
explain superscript T and mention SLHA-2 convention
explain creation of model files and directories in a (hopefully) clearer way

@@
 \code{models/<model>/}.  To add this the user should create a
-\code{FlexibleSUSY.m.in} in the directory \code{model_files/<model>/}.
-When the \code{./createmodel} command is used \fs will create the
-\code{FlexibleSUSY.m} file from the \code{FlexibleSUSY.m.in} file found in the
-directory in \code{model_files} with the name specified by the
+\code{FlexibleSUSY.m.in} file in the directory
+\code{model_files/<model>/}.  When the \code{./createmodel} script is
+executed, the \code{FlexibleSUSY.m} file is created from the
+\code{model_files/<model-file-name>/FlexibleSUSY.m.in} file, where the
+directory \code{<model-file-name>} is the specified by the
 \code{--model-file=<model-file-name>} option.  If no such option is
-given then it will use the directory matching the option
-\code{--name=<model>}.  In either case the the \code{FlexibleSUSY.m} file
-which is created is then automatically placed in the directory
-\code{models/<model>/} by \fs. Note that many example model
-files can already be found in \code{model_files/}.
+given the directory matching the \code{--name=<model>} option is used.
+In either case the \code{FlexibleSUSY.m} file which is created is then
+automatically placed in the directory \code{models/<model>/}.  Note
+that many predefined example model files can already be found in
+\code{model_files/}.
 
@@
 The first line \code{FSModelName = "@CLASSNAME@";} will be replaced
-with \code{FSModelName = "<model>";} in the generated \code{FlexibleSUSY.m}
-file where \code{<model>} is given by the option \code{--name=<model>} for
-command \code{./createmodel}.  So the variable \code{FSModelName} then
-contains the name of the \fs model.
+with \code{FSModelName = "<model>";} in the generated
+\code{FlexibleSUSY.m} file, where \code{<model>} is specified by the
+\code{--name=<model>} option for the \code{./createmodel} script.  So
+the variable \code{FSModelName} then contains the name of the \fs
+model.
 
@@
 $m_u$, $m_d$, $m_e$.  In the MSSM, for example, these relations read
+in the SLHA convention \cite{Allanach:2008qq}
 %
@@
   y_d^{\text{\DRbar}}(M_Z) &= \frac{\sqrt{2} m_{d}^T}{v_d} , &
-  y_e^{\text{\DRbar}}(M_Z) &= \frac{\sqrt{2} m_{e}^T}{v_d}.
+  y_e^{\text{\DRbar}}(M_Z) &= \frac{\sqrt{2} m_{e}^T}{v_d} ,
 \end{align}
 %
-The fermion mass matrices are composed as
+where the superscript $T$ means transposition of a matrix.  The
+fermion mass matrices are composed as
 %

### 2014-06-09 doc/paper.tex
added extra references, so we refer to standard papers for the models.

@@
 within the minimal supersymmetric standard model (MSSM).  Such work
-has been greatly aided by public spectrum generators for
-MSSM \cite{Allanach:2001kg,Porod:2003um,Djouadi:2002ze,Baer:1993ae,Chowdhury:2011zr},
+has been greatly aided by public spectrum generators for MSSM
+\cite{Allanach:2001kg,Porod:2003um,Djouadi:2002ze,Baer:1993ae,Chowdhury:2011zr},
 allowing fast and reliable exploration of the sparticle spectrum,
@@
 generators \cite{Ellwanger:2006rn,Allanach:2013kza} for the next to
-minimal supersymmetric standard model (NMSSM) \cite{NMSSM}.
+minimal supersymmetric standard model (NMSSM) \cite{NMSSM} (or for recent
+reviews see \cite{Ellwanger:2009dp,Maniatis:2009re}).
 
@@
 minimality, and specific alternatives to (or extensions of) the MSSM
-are, for example, motivated by the $\mu$-problem of the
-MSSM \cite{Kim:1983dt}; explaining the family structure (see e.g.~\cite{King:2014nza}) or for successful baryogenesis or leptogenesis
-(see e.g.~\cite{King:2008qb}). However constructing specialized tools
-to study all relevant models would require an enormous amount of work.
-So general tools which can automate this process and produce fast and
-reliable programs can greatly enhance our ability to understand and
-test non-minimal realizations of supersymmetry.
+are, for example, motivated by the $\mu$-problem of the MSSM
+\cite{Kim:1983dt}; explaining the family structure (see
+e.g.~\cite{King:2014nza}) or for successful baryogenesis or
+leptogenesis (see e.g.~\cite{King:2008qb}). However constructing
+specialized tools to study all relevant models would require an
+enormous amount of work.  So general tools which can automate this
+process and produce fast and reliable programs can greatly enhance our
+ability to understand and test non-minimal realizations of
+supersymmetry.
 
@@
 
-These motivate both exploring non-minimal SUSY models which ameliorate
-the naturalness problems by, for example, raising the tree level Higgs
-mass and models developed with a fresh perspective based on other
-considerations.  In both cases exploration of such models can be aided
-if it is possible to quickly create a fast spectrum generator.
-Currently there is only one option for this, a SPheno-like FORTRAN code
-which can be generated from
-\sarah \cite{Staub:2010ty,Staub:2009bi,Staub:2010jh,Staub:2012pb,Staub:2013tta}.
+These developments motivate the exploration of non-minimal SUSY models
+which ameliorate the naturalness problems, as can happen in the \ESSM
+\cite{King:2005jy,King:2005my,King:2007uj,Athron:2010zz} and USSM
+\cite{Fayet:1977yc,Suematsu:1994qm,Cvetic:1995rj,deCarlos:1997yv,Cvetic:1997ky,Demir:1998dk,Langacker:1998tc,Erler:2002pr,Choi:2006fz,Ham:2007wc,Langacker:2008yv,Ham:2008xf,Kalinowski:2008iq}
+by raising the tree level Higgs mass. At the same time they can also
+motivates models that are developed with a fresh perspective, based on
+other considerations.  In both cases exploration of such models can be
+aided if it is possible to quickly create a fast spectrum generator.
+Currently there is only one option for this, a SPheno-like FORTRAN
+code which can be generated from \sarah
+\cite{Staub:2010ty,Staub:2009bi,Staub:2010jh,Staub:2012pb,Staub:2013tta}.
 
@@
 \fs already comes with plenty of predefined models: the CMSSM (simply
-called \code{MSSM}), the $Z_3$-symmetric NMSSM (called \code{NMSSM})
-\cite{Ellwanger:2009dp}, $Z_3$-violating NMSSM (\code{SMSSM})
-\cite{Ellwanger:2009dp}, the USSM (\code{UMSSM})
-\cite{Fayet:1977yc,Cvetic:1997ky,Erler:2002pr}, the NUHM-\ESSM
-(\code{E6SSM}) \cite{King:2005jy,Athron:2007en}, the right-handed
-neutrino extended MSSM (\code{MSSMRHN}), the NUHM-MSSM
+called \code{MSSM}), the NMSSM \cite{NMSSM} in it's $Z_3$-symmetric form (called \code{NMSSM}), $Z_3$-violating NMSSM (\code{SMSSM}), the USSM (\code{UMSSM})
+\cite{Fayet:1977yc,Suematsu:1994qm,Cvetic:1995rj,deCarlos:1997yv,Cvetic:1997ky,Demir:1998dk,Langacker:1998tc,Erler:2002pr,Choi:2006fz,Ham:2007wc,Langacker:2008yv,Ham:2008xf,Kalinowski:2008iq},
+the NUHM-\ESSM (\code{E6SSM}) \cite{Athron:2007en}, the
+right-handed neutrino extended MSSM (\code{MSSMRHN}), the NUHM-MSSM
 (\code{NUHMSSM}) and the $R$-symmetric MSSM (\code{MRSSM})
@@
 spectrum generator for a constrained version of the \ESSM
-\cite{Athron:2009bs,Athron:2012pw}.  The RGEs could be compared
-directly in unit tests and were found to match.  The handwritten code
-doesn't include full one loop self energies or tadpoles so tests on
-these were not carried.  Although the generators assume different
-constraints they could be compared by using the output of the
-CE$_6$SSM generator as an input to \fs and the spectra were found to be in
-reasonable agreement, given the different levels of precision.
+\cite{Athron:2009ue, Athron:2009bs, Athron:2011wu, Athron:2012sq,
+  Athron:2012pw, Athron:2013ipa}.  The RGEs could be compared directly
+in unit tests and were found to match.  The handwritten code doesn't
+include full one loop self energies or tadpoles so tests on these were
+not carried.  Although the generators assume different constraints
+they could be compared by using the output of the CE$_6$SSM generator
+as an input to \fs and the spectra were found to be in reasonable
+agreement, given the different levels of precision.
 
@@
 12} (1997) 479.   D.~J.~Miller, R.~Nevzorov and P.~M.~Zerwas,  Nucl.\ Phys.\ B {\bf 681}, 3 (2004) [hep-ph/0304049].
+%\cite{Ellwanger:2009dp}
+\bibitem{Ellwanger:2009dp} 
+  U.~Ellwanger, C.~Hugonie and A.~M.~Teixeira,
+  %``The Next-to-Minimal Supersymmetric Standard Model,''
+  Phys.\ Rept.\  {\bf 496}, 1 (2010)
+  [arXiv:0910.1785 [hep-ph]].
+  %%CITATION = ARXIV:0910.1785;%%
+  %327 citations counted in INSPIRE as of 21 Sep 2013
+%\cite{Maniatis:2009re}
+\bibitem{Maniatis:2009re} 
+  M.~Maniatis,
+  %``The Next-to-Minimal Supersymmetric extension of the Standard Model reviewed,''
+  Int.\ J.\ Mod.\ Phys.\ A {\bf 25}, 3505 (2010)
+  [arXiv:0906.0777 [hep-ph]].
+  %%CITATION = ARXIV:0906.0777;%%
+  %160 citations counted in INSPIRE as of 09 Jun 2014
+
 %\cite{Kim:1983dt}
@@
 %\cite{Chatrchyan:2014lfa}
+
+%\cite{King:2005jy}
+\bibitem{King:2005jy} 
+  S.~F.~King, S.~Moretti and R.~Nevzorov,
+  %``Theory and phenomenology of an exceptional supersymmetric standard model,''
+  Phys.\ Rev.\ D {\bf 73}, 035009 (2006)
+  [hep-ph/0510419].
+  %%CITATION = HEP-PH/0510419;%%
+  %130 citations counted in INSPIRE as of 03 Nov 2013
+
+%\cite{King:2005my}
+\bibitem{King:2005my} 
+  S.~F.~King, S.~Moretti and R.~Nevzorov,
+  %``Exceptional supersymmetric standard model,''
+  Phys.\ Lett.\ B {\bf 634}, 278 (2006)
+  [hep-ph/0511256].
+  %%CITATION = HEP-PH/0511256;%%
+  %90 citations counted in INSPIRE as of 09 Jun 2014
+
+%\cite{King:2007uj}
+\bibitem{King:2007uj} 
+  S.~F.~King, S.~Moretti and R.~Nevzorov,
+  %``Gauge coupling unification in the exceptional supersymmetric standard model,''
+  Phys.\ Lett.\ B {\bf 650}, 57 (2007)
+  [hep-ph/0701064].
+  %%CITATION = HEP-PH/0701064;%%
+  %46 citations counted in INSPIRE as of 09 Jun 2014
+
+%\cite{Athron:2010zz}
+\bibitem{Athron:2010zz} 
+  P.~Athron, J.~P.~Hall, R.~Howl, S.~F.~King, D.~J.~Miller, S.~Moretti and R.~Nevzorov,
+  %``Aspects of the exceptional supersymmetric standard model,''
+  Nucl.\ Phys.\ Proc.\ Suppl.\  {\bf 200-202}, 120 (2010).
+  %%CITATION = NUPHZ,200-202,120;%%
+  %5 citations counted in INSPIRE as of 09 Jun 2014
+
+%\cite{Fayet:1977yc}
+\bibitem{Fayet:1977yc}
+  P.~Fayet,
+  %``Spontaneously Broken Supersymmetric Theories of Weak, Electromagnetic and Strong Interactions,''
+  Phys.\ Lett.\ B {\bf 69} (1977) 489.
+  %%CITATION = PHLTA,B69,489;%%
+  %817 citations counted in INSPIRE as of 04 Jun 2014
+%\cite{deCarlos:1997yv}
+%\cite{Suematsu:1994qm}
+\bibitem{Suematsu:1994qm} 
+  D.~Suematsu and Y.~Yamagishi,
+  %``Radiative symmetry breaking in a supersymmetric model with an extra U(1),''
+  Int.\ J.\ Mod.\ Phys.\ A {\bf 10}, 4521 (1995)
+  [hep-ph/9411239].
+  %%CITATION = HEP-PH/9411239;%%
+  %116 citations counted in INSPIRE as of 09 Jun 2014
+
+  %\cite{Cvetic:1995rj}
+\bibitem{Cvetic:1995rj} 
+  M.~Cvetic and P.~Langacker,
+  %``Implications of Abelian extended gauge structures from string models,''
+  Phys.\ Rev.\ D {\bf 54}, 3570 (1996)
+  [hep-ph/9511378].
+  %%CITATION = HEP-PH/9511378;%%
+  %207 citations counted in INSPIRE as of 09 Jun 2014
+
+\bibitem{deCarlos:1997yv} 
+  B.~de Carlos and J.~R.~Espinosa,
+  %``Cold dark matter candidate in a class of supersymmetric models with an extra U(1),''
+  Phys.\ Lett.\ B {\bf 407}, 12 (1997)
+  [hep-ph/9705315].
+  %%CITATION = HEP-PH/9705315;%%
+  %51 citations counted in INSPIRE as of 09 Jun 2014
+%\cite{Cvetic:1997ky}
+\bibitem{Cvetic:1997ky}
+  M.~Cvetic, D.~A.~Demir, J.~R.~Espinosa, L.~L.~Everett and P.~Langacker,
+  %``Electroweak breaking and the mu problem in supergravity models with an additional U(1),''
+  Phys.\ Rev.\ D {\bf 56} (1997) 2861
+   [Erratum-ibid.\ D {\bf 58} (1998) 119905]
+  [hep-ph/9703317].
+  %%CITATION = HEP-PH/9703317;%%
+  %237 citations counted in INSPIRE as of 04 Jun 2014
+
+%\cite{Demir:1998dk}
+\bibitem{Demir:1998dk} 
+  D.~A.~Demir and N.~K.~Pak,
+  %``One loop effects in supergravity models with an additional U(1),''
+  Phys.\ Rev.\ D {\bf 57}, 6609 (1998)
+  [hep-ph/9809357].
+  %%CITATION = HEP-PH/9809357;%%
+  %19 citations counted in INSPIRE as of 09 Jun 2014
+
+%\cite{Langacker:1998tc}
+\bibitem{Langacker:1998tc} 
+  P.~Langacker and J.~Wang,
+  %``U(1)-prime symmetry breaking in supersymmetric E(6) models,''
+  Phys.\ Rev.\ D {\bf 58}, 115010 (1998)
+  [hep-ph/9804428].
+  %%CITATION = HEP-PH/9804428;%%
+  %108 citations counted in INSPIRE as of 09 Jun 2014
+
+%\cite{Erler:2002pr}
+\bibitem{Erler:2002pr}
+  J.~Erler, P.~Langacker and T.~-j.~Li,
+  %``The $Z$ - $Z^\prime$ mass hierarchy in a supersymmetric model with a secluded U(1) -prime breaking sector,''
+  Phys.\ Rev.\ D {\bf 66} (2002) 015002
+  [hep-ph/0205001].
+  %%CITATION = HEP-PH/0205001;%%
+  %84 citations counted in INSPIRE as of 04 Jun 2014
+
+%\cite{Choi:2006fz}
+\bibitem{Choi:2006fz} 
+  S.~Y.~Choi, H.~E.~Haber, J.~Kalinowski and P.~M.~Zerwas,
+  %``The Neutralino sector in the U(1)-extended supersymmetric standard model,''
+  Nucl.\ Phys.\ B {\bf 778}, 85 (2007)
+  [hep-ph/0612218].
+  %%CITATION = HEP-PH/0612218;%%
+  %43 citations counted in INSPIRE as of 09 Jun 2014
+
+%\cite{Ham:2007wc}
+\bibitem{Ham:2007wc} 
+  S.~W.~Ham and S.~K.~OH,
+  %``Electroweak phase transition in MSSM with U(1)-prime in explicit CP violation scenario,''
+  Phys.\ Rev.\ D {\bf 76}, 095018 (2007)
+  [arXiv:0708.1785 [hep-ph]].
+  %%CITATION = ARXIV:0708.1785;%%
+  %20 citations counted in INSPIRE as of 09 Jun 2014
+
+%\cite{Langacker:2008yv}
+\bibitem{Langacker:2008yv} 
+  P.~Langacker,
+  %``The Physics of Heavy $Z^\prime$ Gauge Bosons,''
+  Rev.\ Mod.\ Phys.\  {\bf 81}, 1199 (2009)
+  [arXiv:0801.1345 [hep-ph]].
+  %%CITATION = ARXIV:0801.1345;%%
+  %519 citations counted in INSPIRE as of 09 Jun 2014
+
+%\cite{Ham:2008xf}
+\bibitem{Ham:2008xf} 
+  S.~W.~Ham, T.~Hur, P.~Ko and S.~K.~Oh,
+  %``Neutral scalar Higgs bosons in the USSM at the LHC,''
+  J.\ Phys.\ G {\bf 35}, 095007 (2008)
+  [arXiv:0801.2361 [hep-ph]].
+  %%CITATION = ARXIV:0801.2361;%%
+  %10 citations counted in INSPIRE as of 09 Jun 2014
+
+%\cite{Kalinowski:2008iq}
+\bibitem{Kalinowski:2008iq} 
+  J.~Kalinowski, S.~F.~King and J.~P.~Roberts,
+  %``Neutralino Dark Matter in the USSM,''
+  JHEP {\bf 0901}, 066 (2009)
+  [arXiv:0811.2204 [hep-ph]].
+  %%CITATION = ARXIV:0811.2204;%%
+  %30 citations counted in INSPIRE as of 09 Jun 2014
+
  %\cite{Staub:2010ty}
@@
   %177 citations counted in INSPIRE as of 21 Sep 2013
-
-%\cite{Ellwanger:2009dp}
-\bibitem{Ellwanger:2009dp} 
-  U.~Ellwanger, C.~Hugonie and A.~M.~Teixeira,
-  %``The Next-to-Minimal Supersymmetric Standard Model,''
-  Phys.\ Rept.\  {\bf 496}, 1 (2010)
-  [arXiv:0910.1785 [hep-ph]].
-  %%CITATION = ARXIV:0910.1785;%%
-  %327 citations counted in INSPIRE as of 21 Sep 2013
-
-%\cite{Fayet:1977yc}
-\bibitem{Fayet:1977yc}
-  P.~Fayet,
-  %``Spontaneously Broken Supersymmetric Theories of Weak, Electromagnetic and Strong Interactions,''
-  Phys.\ Lett.\ B {\bf 69} (1977) 489.
-  %%CITATION = PHLTA,B69,489;%%
-  %817 citations counted in INSPIRE as of 04 Jun 2014
-
-%\cite{Cvetic:1997ky}
-\bibitem{Cvetic:1997ky}
-  M.~Cvetic, D.~A.~Demir, J.~R.~Espinosa, L.~L.~Everett and P.~Langacker,
-  %``Electroweak breaking and the mu problem in supergravity models with an additional U(1),''
-  Phys.\ Rev.\ D {\bf 56} (1997) 2861
-   [Erratum-ibid.\ D {\bf 58} (1998) 119905]
-  [hep-ph/9703317].
-  %%CITATION = HEP-PH/9703317;%%
-  %237 citations counted in INSPIRE as of 04 Jun 2014
-
-%\cite{Erler:2002pr}
-\bibitem{Erler:2002pr}
-  J.~Erler, P.~Langacker and T.~-j.~Li,
-  %``The $Z$ - $Z^\prime$ mass hierarchy in a supersymmetric model with a secluded U(1) -prime breaking sector,''
-  Phys.\ Rev.\ D {\bf 66} (2002) 015002
-  [hep-ph/0205001].
-  %%CITATION = HEP-PH/0205001;%%
-  %84 citations counted in INSPIRE as of 04 Jun 2014
-
-%\cite{King:2005jy}
-\bibitem{King:2005jy} 
-  S.~F.~King, S.~Moretti and R.~Nevzorov,
-  %``Theory and phenomenology of an exceptional supersymmetric standard model,''
-  Phys.\ Rev.\ D {\bf 73}, 035009 (2006)
-  [hep-ph/0510419].
-  %%CITATION = HEP-PH/0510419;%%
-  %130 citations counted in INSPIRE as of 03 Nov 2013
 
@@
 
+%\cite{Athron:2009ue}
+\bibitem{Athron:2009ue} 
+  P.~Athron, S.~F.~King, D.~J.~Miller, S.~Moretti and R.~Nevzorov,
+  %``Predictions of the Constrained Exceptional Supersymmetric Standard Model,''
+  Phys.\ Lett.\ B {\bf 681}, 448 (2009)
+  [arXiv:0901.1192 [hep-ph]].
+  %%CITATION = ARXIV:0901.1192;%%
+  %37 citations counted in INSPIRE as of 09 Jun 2014
+
 %\cite{Athron:2009bs}
@@
 
+%\cite{Athron:2011wu}
+\bibitem{Athron:2011wu} 
+  P.~Athron, S.~F.~King, D.~J.~Miller, S.~Moretti and R.~Nevzorov,
+  %``LHC Signatures of the Constrained Exceptional Supersymmetric Standard Model,''
+  Phys.\ Rev.\ D {\bf 84}, 055006 (2011)
+  [arXiv:1102.4363 [hep-ph]].
+  %%CITATION = ARXIV:1102.4363;%%
+  %25 citations counted in INSPIRE as of 09 Jun 2014
+
+%\cite{Athron:2012sq}
+\bibitem{Athron:2012sq} 
+  P.~Athron, S.~F.~King, D.~J.~Miller, S.~Moretti and R.~Nevzorov,
+  %``Constrained Exceptional Supersymmetric Standard Model with a Higgs Near 125 GeV,''
+  Phys.\ Rev.\ D {\bf 86}, 095003 (2012)
+  [arXiv:1206.5028 [hep-ph]].
+  %%CITATION = ARXIV:1206.5028;%%
+  %21 citations counted in INSPIRE as of 09 Jun 2014
+
 %\cite{Athron:2012pw}
@@
 
+  %\cite{Athron:2013ipa}
+\bibitem{Athron:2013ipa} 
+  P.~Athron, M.~Binjonaid and S.~F.~King,
+  %``Fine Tuning in the Constrained Exceptional Supersymmetric Standard Model,''
+  Phys.\ Rev.\ D {\bf 87}, no. 11, 115023 (2013)
+  [arXiv:1302.5291 [hep-ph]].
+  %%CITATION = ARXIV:1302.5291;%%
+  %10 citations counted in INSPIRE as of 09 Jun 2014
+
 \end{thebibliography}

### 2014-06-09 doc/paper.tex
Removed names from testin g sevction and mention instead user testing modified acknowledgements slightly to reflect content taken out of testing section.

@@
 
-For checking the correctness of \fs's generated spectrum generators
-extensive unit tests against Softsusy's MSSM and NMSSM
+To check the correctness of \fs's generated spectrum generators
+extensive unit testing against Softsusy's MSSM and NMSSM
 implementations (both $Z_3$-invariant and $Z_3$-violating variants)
-were done: We checked mass matrices, EWSB equations, one- and two-loop
+was carried out. We checked mass matrices, EWSB equations, one- and two-loop
 $\beta$-functions, one- and two-loop self-energies and one- and
@@
 full fixed-point iteration has finished, and found the spectra to
-agree at the sub-permille level.  Analytic tests of the $R$-symmetric
-low-energy model, MRSSM, were done by Philip Diessner and several bugs
-in SARAH were identified and fixed.
+agree at the sub-permille level.
 
@@
 
+In addition \fs has already undergone some user testing. This includes
+analytic tests of the $R$-symmetric low-energy model (MRSSM) and
+alternative E6 inspired SUSY scenarios.  The users who have helped us
+with this are thanked in the acknowledgements.
+
 We also compared the run-time of \fs against SPheno, Softsusy and the
@@
 Next-to-Minimal Softsusy; Sophie Underwood for discovering problems
-introducing couplings with a single family index; Philip Diessner for
-testing and identifying bugs in the MRSSM model file; Ulrik Günther
-for compilation tests on Mac OS X and Dylan Harries for spotting
-a bug in the configuration script.
+when introducing couplings with a single family index; Philip Diessner
+for testing and identifying several bugs in FlexibleSUSY and in SARAH,
+in work on the MRSSM; Ulrik Günther for compilation tests on Mac OS X
+and Dylan Harries for spotting a bug in the configuration script.
 %

### 2014-06-09 doc/paper.tex
fix typos

@@
 within the minimal supersymmetric standard model (MSSM).  Such work
-has been greatly aided by public spectrum generators for MSSM
+has been greatly aided by public spectrum generators for the MSSM
 \cite{Allanach:2001kg,Porod:2003um,Djouadi:2002ze,Baer:1993ae,Chowdhury:2011zr},
@@
 by raising the tree level Higgs mass. At the same time they can also
-motivates models that are developed with a fresh perspective, based on
+motivate models that are developed with a fresh perspective, based on
 other considerations.  In both cases exploration of such models can be
@@
 \cite{Fayet:1977yc,Suematsu:1994qm,Cvetic:1995rj,deCarlos:1997yv,Cvetic:1997ky,Demir:1998dk,Langacker:1998tc,Erler:2002pr,Choi:2006fz,Ham:2007wc,Langacker:2008yv,Ham:2008xf,Kalinowski:2008iq},
-the NUHM-\ESSM (\code{E6SSM}) \cite{Athron:2007en}, the
+the NUHM \ESSM (\code{E6SSM}) \cite{Athron:2007en}, the
 right-handed neutrino extended MSSM (\code{MSSMRHN}), the NUHM-MSSM
@@
 
-\fs generated NUH \ESSM has also been compared against a handwritten
+\fs generated NUHM \ESSM has also been compared against a handwritten
 spectrum generator for a constrained version of the \ESSM
@@
 in unit tests and were found to match.  The handwritten code doesn't
-include full one loop self energies or tadpoles so tests on these were
-not carried.  Although the generators assume different constraints
+include full one-loop self-energies or tadpoles so tests on these were
+not carried out.  Although the generators assume different constraints
 they could be compared by using the output of the CE$_6$SSM generator

### 2014-06-09 doc/paper.tex
fix another typo

@@
 analytic tests of the $R$-symmetric low-energy model (MRSSM) and
-alternative E6 inspired SUSY scenarios.  The users who have helped us
+alternative $E_6$-inspired SUSY scenarios.  The users who have helped us
 with this are thanked in the acknowledgements.
@@
 when introducing couplings with a single family index; Philip Diessner
-for testing and identifying several bugs in FlexibleSUSY and in SARAH,
+for testing and identifying several bugs in \fs and in SARAH,
 in work on the MRSSM; Ulrik Günther for compilation tests on Mac OS X

### 2014-06-09 doc/paper.tex
typo

@@
 features in detailed examples for the NUHMSSM, right handed neutrinos
-and on adding three loop RGEs and two loop matching for the strong
+and on adding three-loop RGEs and two-loop matching for the strong
 gauge coupling.

### 2014-06-09 doc/paper.tex
stick to -ize convention

@@
 conditions, one-loop self energies, one-loop tadpoles corrections and
-two-loop renormalisation group equations (RGEs) for the model.
+two-loop renormalization group equations (RGEs) for the model.
 Additional corrections which have been calculated elsewhere, such as

### 2014-06-09 doc/flexiblesusy-paper.tex
rename paper TeX file

--- 2014-06-09 doc/paper.tex
@@
 and checking for convergence after each iteration.  This approach is
-described in \cite{Barger:1993gh} for the MSSM and is widely
+described in \cite{Barger:1993gh} originally for the MSSM and is widely
 implemented in SUSY spectrum generators.
-
-In more detail the two-scale algorithm used in \fs works as
-follows, see also \figref{fig:two-scale-algorithm}:
+Despite sharing the same algorithm with others,
+the boundary value problem solver class from \fs, named \code{RGFlow},
+has two notable properties.
+First, it extends the aforementioned procedure to towers of models.
+If the problem involves more than one model,
+\code{RGFlow} matches one model to the next after
+running the model parameters to the matching scale.
+Second, \code{RGFlow} is an abstract implementation of
+the algorithm, unaware of physics, in that
+it is free of hard-wired model-dependent code related to
+RGEs, boundary or matching conditions, or initial guesses.
+All these pieces of physics information are
+carried by separate objects which one then
+links to \code{RGFlow} to set up a boundary value problem.
+This modular design makes it easy
+to replace any of the above components,
+as shall be demonstrated in \secref{sec:integrating-custom-built}.
+
+In more detail the two-scale algorithm used in \fs,
+as applied to a problem with a single MSSM-like model,
+works as follows, see also \figref{fig:two-scale-algorithm}:
 %

### 2014-06-09 doc/flexiblesusy-paper.tex
Add report numbers and acknowledgement
Intel Xeon has 2 CPUs with 6 cores each

@@
 \usepackage{listings}
+\usepackage[absolute]{textpos}
 \bibstyle{elsarticle-num}
@@
 \end{frontmatter}
+
+% report numbers
+\begin{textblock*}{7em}(\textwidth,1cm)
+\noindent\footnotesize
+FTUV--14--3904 \\
+IFIC--14--40
+\end{textblock*}
 
@@
 enabled.  The second column shows the run-time on the same processor
-where both cores were enabled.  In the third column a Intel Xeon
-(L5640, $2.27\unit{GHz}$) was used, which has $6$ CPU cores.
+where both cores were enabled.  In the third column an Intel Xeon
+(L5640, $2.27\unit{GHz}$) was used, which has two CPUs with $6$ cores
+each.
 %
@@
                             & Intel Core2 Duo    & Intel Core2 Duo   & Intel Xeon\\
-                            & (P8600, 1 core)    & (P8600, 2 cores)  & (L5640, 6 cores)\\
+                            & (P8600, $1$ core)  & (P8600, $2$ cores)& (L5640, $2\times 6$ cores)\\
     \midrule
@@
 %
+J.P. acknowledges support from the MEC and FEDER (EC) Grants
+FPA2011--23596 and the Generalitat Valenciana under grant PROMETEOII/2013/017.
 This work has been supported by the German Research Foundation DFG

### 2014-06-09 doc/flexiblesusy-paper.tex
fix long line

@@
 [50,1000]\unit{GeV}$, $\tan\beta\in [1,100]$, $\sign\mu\in \{-1,+1\}$
-and $A_0\in [-1000,1000]\unit{GeV}$.  For each point an SLHA input
+and $A_0\in [-1000,$ $1000]\unit{GeV}$.  For each point an SLHA input
 file is created by appending the values of $m_0$, $m_{1/2}$,

### 2014-06-09 doc/flexiblesusy-paper.tex
correcting the intel Xeon machine type again
The machine has 2 CPUs of type Intel Xeon.
Each Xeon CPU has 6 physical cores.

@@
 enabled.  The second column shows the run-time on the same processor
-where both cores were enabled.  In the third column an Intel Xeon
-(L5640, $2.27\unit{GHz}$) was used, which has two CPUs with $6$ cores
-each.
+where both cores were enabled.  In the third column a machine with two
+Intel Xeon CPUs (L5640, $2.27\unit{GHz}$, $6$ cores) was used.
 %
@@
     \toprule
-                            & Intel Core2 Duo    & Intel Core2 Duo   & Intel Xeon\\
-                            & (P8600, $1$ core)  & (P8600, $2$ cores)& (L5640, $2\times 6$ cores)\\
+                            & Intel Core2 Duo    & Intel Core2 Duo   & $2$ $\times$ Intel Xeon\\
+                            & (P8600, $1$ core)  & (P8600, $2$ cores)& (L5640, $6$ cores)\\
     \midrule

### 2014-06-12 doc/flexiblesusy-paper.tex
added adelade pre-print number to paper.
changed some of the commands using LesHouches files to refelct Jae-hyeon's update. I didn't change all references to this though as the leshoucesh files inside model_files/<model> directory are still there. Not sure what is optimal way to describe in paper.

@@
 \begin{frontmatter}
-
  \title{\Large\bf FlexibleSUSY --- A spectrum generator generator for supersymmetric models}
@@
 FTUV--14--3904 \\
-IFIC--14--40
+IFIC--14--40 \\
+ADP-14-17/T875
 \end{textblock*}
@@
 \begin{lstlisting}[language=bash]
-$ models/MSSM/run_MSSM.x --slha-input-file=model_files/MSSM/LesHouches.in.MSSM
-\end{lstlisting}%% $
+$ cd models/MSSM 
+$ ./run_MSSM.x --slha-input-file=LesHouches.in.MSSM
+\end{lstlisting}
 %
@@
 $ make
-$ models/NUHMSSM/run_NUHMSSM.x --slha-input-file=model_files/NUHMSSM/LesHouches.in.NUHMSSM
+$ models/NUHMSSM/run_NUHMSSM.x --slha-input-file=models/NUHMSSM/LesHouches.in.NUHMSSM
 \end{numlstlisting}
@@
 \clearpage
-\section*{References}
 

### 2014-06-12 doc/flexiblesusy-paper.tex
added reference from Moritz McGarrie to USSM list.  may not be entirtely appropriate.

@@
 \cite{King:2005jy,King:2005my,King:2007uj,Athron:2010zz} and USSM
-\cite{Fayet:1977yc,Suematsu:1994qm,Cvetic:1995rj,deCarlos:1997yv,Cvetic:1997ky,Demir:1998dk,Langacker:1998tc,Erler:2002pr,Choi:2006fz,Ham:2007wc,Langacker:2008yv,Ham:2008xf,Kalinowski:2008iq}
+\cite{Fayet:1977yc,Suematsu:1994qm,Cvetic:1995rj,deCarlos:1997yv,Cvetic:1997ky,Demir:1998dk,Langacker:1998tc,Erler:2002pr,Choi:2006fz,Ham:2007wc,Langacker:2008yv,Ham:2008xf,Kalinowski:2008iq,Bharucha:2013ela}
 by raising the tree level Higgs mass. At the same time they can also
@@
   %30 citations counted in INSPIRE as of 09 Jun 2014
+%\cite{Bharucha:2013ela}
+\bibitem{Bharucha:2013ela} 
+  A.~Bharucha, A.~Goudelis and M.~McGarrie,
+  %``En-gauging Naturalness,''
+  arXiv:1310.4500 [hep-ph].
+  %%CITATION = ARXIV:1310.4500;%%
+  %9 citations counted in INSPIRE as of 12 Jun 2014
 

### 2014-06-12 doc/flexiblesusy-paper.tex
added more references to non-decoupled D-term approach, citing the kaplan paper. and Anibal Medina's paper.  there are probably many more.

@@
 which ameliorate the naturalness problems, as can happen in the \ESSM
-\cite{King:2005jy,King:2005my,King:2007uj,Athron:2010zz} and USSM
-\cite{Fayet:1977yc,Suematsu:1994qm,Cvetic:1995rj,deCarlos:1997yv,Cvetic:1997ky,Demir:1998dk,Langacker:1998tc,Erler:2002pr,Choi:2006fz,Ham:2007wc,Langacker:2008yv,Ham:2008xf,Kalinowski:2008iq,Bharucha:2013ela}
+\cite{King:2005jy,King:2005my,King:2007uj,Athron:2010zz}, USSM
+\cite{Fayet:1977yc,Suematsu:1994qm,Cvetic:1995rj,deCarlos:1997yv,Cvetic:1997ky,Demir:1998dk,Langacker:1998tc,Erler:2002pr,Choi:2006fz,Ham:2007wc,Langacker:2008yv,Ham:2008xf,Kalinowski:2008iq} or from other gauge extensions \cite{Batra:2003nj, Medina:2009ey, Bharucha:2013ela},
 by raising the tree level Higgs mass. At the same time they can also
@@
 %\cite{Bharucha:2013ela}
+
 \bibitem{Bharucha:2013ela} 
@@
   %%CITATION = ARXIV:1310.4500;%%
+  %9 citations counted in INSPIRE as of 12 Jun 2014
+
+%\cite{Batra:2003nj}
+\bibitem{Batra:2003nj} 
+  P.~Batra, A.~Delgado, D.~E.~Kaplan and T.~M.~P.~Tait,
+  %``The Higgs mass bound in gauge extensions of the minimal supersymmetric standard model,''
+  JHEP {\bf 0402}, 043 (2004)
+  [hep-ph/0309149].
+  %%CITATION = HEP-PH/0309149;%%
+  %174 citations counted in INSPIRE as of 12 Jun 2014
+%\cite{Medina:2009ey}
+\bibitem{Medina:2009ey} 
+  A.~D.~Medina, N.~R.~Shah and C.~E.~M.~Wagner,
+  %``A Heavy Higgs and a Light Sneutrino NLSP in the MSSM with Enhanced SU(2) D-terms,''
+  Phys.\ Rev.\ D {\bf 80}, 015001 (2009)
+  [arXiv:0904.1625 [hep-ph]].
+  %%CITATION = ARXIV:0904.1625;%%
   %9 citations counted in INSPIRE as of 12 Jun 2014

### 2014-06-12 doc/flexiblesusy-paper.tex
fixed order of refs I just added.
added Gregor Hellwig and Gambit to acknowledgements.

@@
 fixing. P.A.~would like to thank Roman Nevzorov for useful discussions
-about challenges in non-minimal SUSY models. The authors would also
-like to thank Lewis Tunstall for helping with early tests against
+about challenges in non-minimal SUSY models and the GAMBIT
+collaboration for helpful suggestions, particularly regarding
+lowering our compiler prerequisites. The authors would also like to
+thank Lewis Tunstall for helping with early tests against
 Next-to-Minimal Softsusy; Sophie Underwood for discovering problems
-when introducing couplings with a single family index; Philip Diessner
-for testing and identifying several bugs in \fs and in SARAH,
-in work on the MRSSM; Ulrik Günther for compilation tests on Mac OS X
-and Dylan Harries for spotting a bug in the configuration script.
+when introducing couplings with a single family index; Gregor Hellwig
+for writing the first version of the E6SSM model file; Philip Diessner
+for further adaptions to the E6SSM model file and for supplying it to
+us and for for testing and identifying several bugs in \fs and in
+SARAH, in work on the MRSSM; Ulrik Günther for compilation tests on
+Mac OS X and Dylan Harries for spotting a bug in the configuration
+script.
 %
@@
 
-\bibitem{Bharucha:2013ela} 
-  A.~Bharucha, A.~Goudelis and M.~McGarrie,
-  %``En-gauging Naturalness,''
-  arXiv:1310.4500 [hep-ph].
-  %%CITATION = ARXIV:1310.4500;%%
-  %9 citations counted in INSPIRE as of 12 Jun 2014
-
 %\cite{Batra:2003nj}
@@
   %%CITATION = ARXIV:0904.1625;%%
+  %9 citations counted in INSPIRE as of 12 Jun 2014
+
+\bibitem{Bharucha:2013ela} 
+  A.~Bharucha, A.~Goudelis and M.~McGarrie,
+  %``En-gauging Naturalness,''
+  arXiv:1310.4500 [hep-ph].
+  %%CITATION = ARXIV:1310.4500;%%
   %9 citations counted in INSPIRE as of 12 Jun 2014

### 2014-06-12 doc/flexiblesusy-paper.tex
Adapt paper further to installation of LesHouches files

@@
 The first command creates the model directory \code{models/MSSM/}
-together with a CMSSM model file.  The \code{configure} script checks
+together with a CMSSM model file accompanied by
+a specimen SLHA input file.  The \code{configure} script checks
 the system requirements and creates the \code{Makefile}.  See
@@
 parameter point given in the SLHA input file
-\code{model_files/MSSM/LesHouches.in.MSSM}.  Afterwards, the pole mass
+\code{LesHouches.in.MSSM}.  Afterwards, the pole mass
 spectrum and mixing matrices are calculated and written to the standard
@@
 to generate the C++ program.
+Another remark is in order regarding the naming convention
+of specimen SLHA input files.  The \code{createmodel} script assumes
+that their names are in the form \code{LesHouches.in*} (case-insensitive).
+If the script finds such files in \code{model_files/<model>/},
+it installs them into the model directory.  The argument to
+\code{--slha-input-file=} in line 4 has been thus created.
 

### 2014-06-16 doc/flexiblesusy-paper.tex
added s few of Roman's E6 custodial symmetry papers to manual, since that is a really nice alternative formulation of the E6SSM and can ~fit with our model files since we neglect many couplings.
updated pulication info for nmssm-softsusy.
update FS version number in download instructions to 1.0.1
Version 1.0.1 fixes a `make install-src' bug.

@@
 which ameliorate the naturalness problems, as can happen in the \ESSM
-\cite{King:2005jy,King:2005my,King:2007uj,Athron:2010zz}, USSM
+\cite{King:2005jy,King:2005my,King:2007uj,Athron:2010zz,Nevzorov:2012hs,Nevzorov:2013tta,Nevzorov:2013ixa}, USSM
 \cite{Fayet:1977yc,Suematsu:1994qm,Cvetic:1995rj,deCarlos:1997yv,Cvetic:1997ky,Demir:1998dk,Langacker:1998tc,Erler:2002pr,Choi:2006fz,Ham:2007wc,Langacker:2008yv,Ham:2008xf,Kalinowski:2008iq} or from other gauge extensions \cite{Batra:2003nj, Medina:2009ey, Bharucha:2013ela},
@@
 \url{http://flexiblesusy.hepforge.org}.  To download and install
-version 1.0.0 run:
+version 1.0.1 run:
 %
 \begin{lstlisting}[language=bash]
-$ wget https://www.hepforge.org/archive/flexiblesusy/FlexibleSUSY-1.0.0.tar.gz
-$ tar -xf FlexibleSUSY-1.0.0.tar.gz
-$ cd FlexibleSUSY-1.0.0
+$ wget https://www.hepforge.org/archive/flexiblesusy/FlexibleSUSY-1.0.1.tar.gz
+$ tar -xf FlexibleSUSY-1.0.1.tar.gz
+$ cd FlexibleSUSY-1.0.1
 \end{lstlisting}%% $
@@
   %``Next-to-Minimal SOFTSUSY,''
-  arXiv:1311.7659 [hep-ph].
+  Comput.\ Phys.\ Commun.\  {\bf 185}, 2322 (2014)
+  [arXiv:1311.7659 [hep-ph]].
   %%CITATION = ARXIV:1311.7659;%%
-  %1 citations counted in INSPIRE as of 24 Apr 2014
+  %2 citations counted in INSPIRE as of 15 Jun 2014
 \bibitem{NMSSM} P. Fayet, Nucl. Phys. B \textbf{90} (1975) 104; Phys. Lett.
@@
 
+%\cite{Nevzorov:2012hs}
+\bibitem{Nevzorov:2012hs} 
+  R.~Nevzorov,
+  %``E6 Inspired SUSY Models with Exact Custodial Symmetry,''
+  Phys.\ Rev.\ D {\bf 87}, 015029 (2013)
+  [arXiv:1205.5967 [hep-ph]].
+  %%CITATION = ARXIV:1205.5967;%%
+  %4 citations counted in INSPIRE as of 13 Jun 2014
+
+%\cite{Nevzorov:2013tta}
+\bibitem{Nevzorov:2013tta} 
+  R.~Nevzorov and S.~Pakvasa,
+  %``Exotic Higgs decays in the $E_6$ inspired SUSY models,''
+  Phys.\ Lett.\ B {\bf 728}, 210 (2014)
+  [arXiv:1308.1021 [hep-ph]].
+  %%CITATION = ARXIV:1308.1021;%%
+  %2 citations counted in INSPIRE as of 13 Jun 2014
+
+%\cite{Nevzorov:2013ixa}
+\bibitem{Nevzorov:2013ixa} 
+  R.~Nevzorov,
+  %``Quasi-fixed point scenarios and the Higgs mass in the E6 inspired SUSY models,''
+  Phys.\ Rev.\ D {\bf 89}, 055010 (2014)
+  [arXiv:1309.4738 [hep-ph]].
+  %%CITATION = ARXIV:1309.4738;%%
+  %1 citations counted in INSPIRE as of 13 Jun 2014
+
 %\cite{Fayet:1977yc}

### 2014-06-23 doc/flexiblesusy-paper.tex
recommend to set 2-loop switches to True in any MSSM- and NMSSM- like model

@@
 which add two-loop corrections of $\oatas$, $\oabas$, $\oatq$,
-$\oabq$, $\oatauq$ and $\oatab$.  Something similar is
-done for the NMSSM, but in this case the NMSSM $\oatas$, $\oabas$
-pieces come from \cite{Degrassi:2009yq}, while for $\oatq$,
-$\oabq$, $\oatauq$ and $\oatab$ the MSSM pieces are used
-though it should be understood that these are not complete in the
-NMSSM. For other models since the Higgs mass is a very important
-measurement and the two-loop corrections can be larger than the
-current experimental error \cite{Degrassi:2009yq} we recommend that
-the leading log two-loop corrections are estimated, by generalizing
-those of the MSSM or NMSSM, as has been done, for example, in the
-E$_6$SSM \cite{King:2005jy}.
+$\oabq$, $\oatauq$ and $\oatab$.  Something similar is done for the
+NMSSM, but in this case the NMSSM $\oatas$, $\oabas$ pieces come from
+\cite{Degrassi:2009yq}, while for $\oatq$, $\oabq$, $\oatauq$ and
+$\oatab$ the MSSM pieces are used though it should be understood that
+these are not complete in the NMSSM.  Since the Higgs mass is a very
+important measurement and the two-loop corrections can be larger than
+the current experimental error \cite{Degrassi:2009yq} we recommend to
+set these switches to \code{True} in any MSSM-\footnote{In \fs a model
+  is MSSM-like if (i) its superpotential is approximately given by the
+  MSSM one, (ii) it implements $R$-parity conservation or something
+  equivalent, and (iii) contains two CP-even and CP-odd Higgs bosons,
+  where one CP-odd Higgs boson may be a Goldstone boson, all with an
+  MSSM-like coupling to $t$, $b$ and $\tau$.} and
+NMSSM-like\footnote{In \fs a model is NMSSM-like if (i) its
+  superpotential is approximately given by the NMSSM one, (ii) it
+  implements $R$-parity conservation or something equivalent, and
+  (iii) contains three CP-even and CP-odd Higgs bosons, where one or
+  two CP-odd Higgs bosons may be Goldstone bosons, all with an
+  NMSSM-like coupling to $t$, $b$ and $\tau$\@.  Examples for
+  NMSSM-like models are the USSM and the \ESSM.} model.  For models
+with a more extended Higgs sector we recommend that the leading log
+two-loop corrections are estimated by generalizing those of the MSSM
+or NMSSM.
 

### 2014-06-23 meta/SelfEnergies.m
correcting naming of 3rd gen. sfermion mass calculation functions

@@ -525,13 +525,13 @@ double msb_1, msb_2, theta_b;
 double mstau_1, mstau_2, theta_tau;
 double msnu_1, msnu_2, theta_nu;
 
-" <> TreeMasses`CallThirdGenerationHelperFunctionName[SARAH`TopQuark, "mst_1", "mst_2", "theta_t"] <>
+" <> TreeMasses`CallThirdGenerationHelperFunctionName[SARAH`TopSquark, "mst_1", "mst_2", "theta_t"] <>
 ";
-" <> TreeMasses`CallThirdGenerationHelperFunctionName[SARAH`BottomQuark, "msb_1", "msb_2", "theta_b"] <>
+" <> TreeMasses`CallThirdGenerationHelperFunctionName[SARAH`BottomSquark, "msb_1", "msb_2", "theta_b"] <>
 ";
-" <> TreeMasses`CallThirdGenerationHelperFunctionName[SARAH`Electron, "mstau_1", "mstau_2", "theta_tau"] <>
+" <> TreeMasses`CallThirdGenerationHelperFunctionName[SARAH`Selectron, "mstau_1", "mstau_2", "theta_tau"] <>
 ";
-" <> TreeMasses`CallThirdGenerationHelperFunctionName[SARAH`Neutrino, "msnu_1", "msnu_2", "theta_nu"] <>
+" <> TreeMasses`CallThirdGenerationHelperFunctionName[SARAH`Sneutrino, "msnu_1", "msnu_2", "theta_nu"] <>
 ";
 
 double mst1sq = Sqr(mst_1), mst2sq = Sqr(mst_2);
@@ -608,13 +608,13 @@ double msb_1, msb_2, theta_b;
 double mstau_1, mstau_2, theta_tau;
 double msnu_1, msnu_2, theta_nu;
 
-" <> TreeMasses`CallThirdGenerationHelperFunctionName[SARAH`TopQuark, "mst_1", "mst_2", "theta_t"] <>
+" <> TreeMasses`CallThirdGenerationHelperFunctionName[SARAH`TopSquark, "mst_1", "mst_2", "theta_t"] <>
 ";
-" <> TreeMasses`CallThirdGenerationHelperFunctionName[SARAH`BottomQuark, "msb_1", "msb_2", "theta_b"] <>
[…]

### 2014-06-23 meta/TreeMasses.m

@@ -988,7 +988,7 @@ CreateThirdGenerationHelperPrototype[fermion_] :=
     CConversion`ToValidCSymbolString[FlexibleSUSY`M[fermion]] <>
     "_3rd_generation(double&, double&, double&) const;\n";
 
-CreateThirdGenerationHelperFunction[fermion_ /; fermion === SARAH`TopQuark] :=
+CreateThirdGenerationHelperFunction[fermion_ /; fermion === SARAH`TopSquark] :=
     "void CLASSNAME::calculate_" <>
     CConversion`ToValidCSymbolString[FlexibleSUSY`M[fermion]] <>
     "_3rd_generation(double& msf1, double& msf2, double& theta) const {
@@ -1015,7 +1015,7 @@ CreateThirdGenerationHelperFunction[fermion_ /; fermion === SARAH`TopQuark] :=
 }
 ";
 
-CreateThirdGenerationHelperFunction[fermion_ /; fermion === SARAH`BottomQuark] :=
+CreateThirdGenerationHelperFunction[fermion_ /; fermion === SARAH`BottomSquark] :=
     "void CLASSNAME::calculate_" <>
     CConversion`ToValidCSymbolString[FlexibleSUSY`M[fermion]] <>
     "_3rd_generation(double& msf1, double& msf2, double& theta) const {
@@ -1042,7 +1042,7 @@ CreateThirdGenerationHelperFunction[fermion_ /; fermion === SARAH`BottomQuark] :
 }
 ";
 
-CreateThirdGenerationHelperFunction[fermion_ /; fermion === SARAH`Electron] :=
+CreateThirdGenerationHelperFunction[fermion_ /; fermion === SARAH`Selectron] :=
     "void CLASSNAME::calculate_" <>
     CConversion`ToValidCSymbolString[FlexibleSUSY`M[fermion]] <>
     "_3rd_generation(double& msf1, double& msf2, double& theta) const {
@@ -1069,7 +1069,7 @@ CreateThirdGenerationHelperFunction[fermion_ /; fermion === SARAH`Electron] :=
 }
 ";
 
-CreateThirdGenerationHelperFunction[fermion_ /; fermion === SARAH`Neutrino] :=
+CreateThirdGenerationHelperFunction[fermion_ /; fermion === SARAH`Sneutrino] :=
     "void CLASSNAME::calculate_" <>
     CConversion`ToValidCSymbolString[FlexibleSUSY`M[fermion]] <>
[…]

### 2014-10-07 doc/flexiblesusy-paper.tex
rename DRbar spectrum calculation function
as suggested by Peter.  The function calculates the DRbar mass
spectrum, and therefore should not be called
calculate_DRbar_parameters (which is the Softsusy name).

@@
 \author[dresden]{Dominik St\"ockinger}
-\author[dresden]{Alexander Voigt}
+\author[dresden]{Alexander Voigt\corref{cor1}}
+\ead{Alexander.Voigt@physik.tu-dresden.de}
+\cortext[cor1]{Corresponding author}
 \address[adelaide]{ARC Centre of Excellence for Particle Physics at 
@@
 \DRbar\ and pole mass spectra.  These two calculations are performed
-in the \code{calculate_DRbar_parameters()} and
+in the \code{calculate_DRbar_masses()} and
 \code{calculate_pole_masses()} functions, which make use of the mass
@@
    if (ewsb_loop_order > 0) {
-      model->calculate_DRbar_parameters();
+      model->calculate_DRbar_masses();
       tadpole[0] -= Re(model->tadpole_hh(0));
@@
   model_1->solve_ewsb_tree_level();
-  model_1->calculate_DRbar_parameters();
+  model_1->calculate_DRbar_masses();
   model_1->set_thresholds(3); model_1->set_loops(2);

### 2014-10-07 examples/customized-betas/MSSMcbs_two_scale_initial_guesser.cpp

@@ -175,7 +175,7 @@ void MSSMcbs_initial_guesser<Two_scale>::guess_soft_parameters()
    model->solve_ewsb_tree_level();
 
    // calculate tree-level spectrum
-   model->calculate_DRbar_parameters();
+   model->calculate_DRbar_masses();
 }
 
 } // namespace flexiblesusy

### 2014-10-07 examples/customized-betas/MSSMcbs_two_scale_low_scale_constraint.cpp

@@ -77,7 +77,7 @@ void MSSMcbs_low_scale_constraint<Two_scale>::apply()
    assert(model && "Error: MSSMcbs_low_scale_constraint:"
           " model pointer must not be zero");
 
-   model->calculate_DRbar_parameters();
+   model->calculate_DRbar_masses();
    update_scale();
    calculate_DRbar_gauge_couplings();
 

### 2014-10-07 examples/standalone-model/standalone.cpp

@@ -126,7 +126,7 @@ void self_energy_example()
    MSSM<Two_scale> mssm;
 
    setup(mssm);
-   mssm.calculate_DRbar_parameters();
+   mssm.calculate_DRbar_masses();
    // INFO(mssm);
 
    double p = Electroweak_constants::MZ;

### 2014-10-07 examples/tower/MSSMD5O_MSSMRHN_two_scale_initial_guesser.cpp

@@ -195,7 +195,7 @@ void MSSMD5O_MSSMRHN_initial_guesser<Two_scale>::guess_soft_parameters()
    model_1->solve_ewsb_tree_level();
 
    // calculate tree-level spectrum
-   model_1->calculate_DRbar_parameters();
+   model_1->calculate_DRbar_masses();
    model_1->set_thresholds(3);
    model_1->set_loops(2);
 }

### 2014-10-07 meta/Constraint.m

@@ -107,7 +107,7 @@ CreateMinimizationFunctionWrapper[className_String, functionName_String, dim_Str
 
       MODELCLASSNAME* model = static_cast<MODELCLASSNAME*>(parameters);
 " <> TextFormatting`IndentText[SetModelParametersFromGSLVector["model","x",parameters],6] <> "
-      model->calculate_DRbar_parameters();
+      model->calculate_DRbar_masses();
 " <> TextFormatting`IndentText[Parameters`CreateLocalConstRefs[function],6] <> "
       return " <> CConversion`RValueToCFormString[function] <> ";
    }
@@ -136,7 +136,7 @@ CreateRootFinderFunctionWrapper[className_String, functionName_String, dim_Strin
 
       MODELCLASSNAME* model = static_cast<MODELCLASSNAME*>(parameters);
 " <> TextFormatting`IndentText[SetModelParametersFromGSLVector["model","x",parameters],6] <> "
-      model->calculate_DRbar_parameters();
+      model->calculate_DRbar_masses();
 " <> TextFormatting`IndentText[Parameters`CreateLocalConstRefs[function],6] <> "
 " <> TextFormatting`IndentText[SetGSLVectorFromExpressions["f",function],6] <> "
       return GSL_SUCCESS;

### 2014-10-07 test/test_MSSM.hpp

@@ -186,7 +186,7 @@ void setup_MSSM(MSSM<Two_scale>& m, MssmSoftsusy& s, MSSM_input_parameters& inpu
    s.setMw(s.displayMwRun());
 
    ensure_tree_level_ewsb(m);
-   m.calculate_DRbar_parameters();
+   m.calculate_DRbar_masses();
 
    ensure_tree_level_ewsb(s);
    s.calcDrBarPars();

### 2014-10-07 test/test_MSSMNoFV.hpp

@@ -127,10 +127,10 @@ void setup_MSSM_models(T1& m1, T2& m2, const TInput& input)
    m2.set_vd(vd);
 
    ensure_tree_level_ewsb(m1);
-   m1.calculate_DRbar_parameters();
+   m1.calculate_DRbar_masses();
 
    ensure_tree_level_ewsb(m2);
-   m2.calculate_DRbar_parameters();
+   m2.calculate_DRbar_masses();
 }
 
 #endif

### 2014-10-07 test/test_MSSMNoFV_low_scale_constraint.cpp

@@ -28,8 +28,8 @@ BOOST_AUTO_TEST_CASE( test_delta_alpha )
    QedQcd oneset;
    setup_MSSM_models(mssm, mssmnofv, input_mssmnofv);
 
-   mssm.calculate_DRbar_parameters();
-   mssmnofv.calculate_DRbar_parameters();
+   mssm.calculate_DRbar_masses();
+   mssmnofv.calculate_DRbar_masses();
 
    MSSM_low_scale_constraint<Two_scale> constraint_mssm(input_mssm, oneset);
    constraint_mssm.set_model(&mssm);

### 2014-10-07 test/test_MSSM_higgs_iteration.cpp

@@ -114,7 +114,7 @@ BOOST_AUTO_TEST_CASE( test_MSSM_higgs_iteration )
          model->set_vd(vd);
          model->set_vu(vu);
 
-         model->calculate_DRbar_parameters();
+         model->calculate_DRbar_masses();
          model->calculate_Mhh_pole();
          model->calculate_MVZ_pole();
 
@@ -136,7 +136,7 @@ BOOST_AUTO_TEST_CASE( test_MSSM_higgs_iteration )
    BOOST_MESSAGE("Predicted tan(beta) = " << model.get_vu() / model.get_vd());
 
    // check how close we got
-   model.calculate_DRbar_parameters();
+   model.calculate_DRbar_masses();
    model.calculate_Mhh_pole();
    model.calculate_MVZ_pole();
 

### 2014-10-07 test/test_MSSM_model.cpp

@@ -125,7 +125,7 @@ void test_beta_function_equality(const SoftParsMssm& a, const MSSM_soft_paramete
 void compare_tree_level_masses(MssmSoftsusy s, MSSM<Two_scale> m)
 {
    ensure_tree_level_ewsb(m);
-   m.calculate_DRbar_parameters();
+   m.calculate_DRbar_masses();
    s.calcDrBarPars();
 
    const double beta = atan(s.displayTanb());
@@ -894,7 +894,7 @@ void compare_top_self_energy(MssmSoftsusy s, MSSM<Two_scale> m)
    s.setMu(MZ);
    s.calcDrBarPars();
    m.set_scale(MZ);
-   m.calculate_DRbar_parameters();
+   m.calculate_DRbar_masses();
 
    const double mtpole = s.displayDataSet().displayPoleMt();
    const double softsusy_mtop = s.calcRunningMt();
@@ -911,7 +911,7 @@ void compare_bot_self_energy(MssmSoftsusy s, MSSM<Two_scale> m)
    s.setMu(MZ);
    s.calcDrBarPars();
    m.set_scale(MZ);
-   m.calculate_DRbar_parameters();
+   m.calculate_DRbar_masses();
 
    const double mb_ms_bar = s.displayDataSet().displayMass(mBottom);
    const double softsusy_mbot = s.calcRunningMb();
@@ -925,7 +925,7 @@ void compare_tau_self_energy(MssmSoftsusy s, MSSM<Two_scale> m)
    s.setMu(MZ);
    s.calcDrBarPars();
    m.set_scale(MZ);
-   m.calculate_DRbar_parameters();
+   m.calculate_DRbar_masses();
 
    const double mtaupole = s.displayDataSet().displayPoleMtau();
    const double softsusy_mtau = s.calcRunningMtau();
@@ -943,7 +943,7 @@ void compare_self_energies(MssmSoftsusy s, MSSM<Two_scale> m)
 {
    ensure_tree_level_ewsb(m);
    s.calcDrBarPars();
-   m.calculate_DRbar_parameters();
+   m.calculate_DRbar_masses();
 
    TEST_EQUALITY(s.displayMu(), m.get_scale());
 
@@ -968,7 +968,7 @@ void compare_tadpoles(MssmSoftsusy s, MSSM<Two_scale> m)
 {
    ensure_tree_level_ewsb(m);
    s.calcDrBarPars();
-   m.calculate_DRbar_parameters();
[…]

### 2014-10-07 test/test_MSSM_spectrum.cpp

@@ -53,7 +53,7 @@ void MSSM_precise_gauge_couplings_low_scale_constraint::apply()
 
    // run MSSM_low_scale_constraint::apply(), without the gauge
    // couplings
-   model->calculate_DRbar_parameters();
+   model->calculate_DRbar_masses();
    update_scale();
    calculate_DRbar_gauge_couplings();
 
@@ -144,7 +144,7 @@ void MSSM_softsusy_ewsb_susy_scale_constraint::apply()
           " model pointer must not be zero");
 
    // save old model parmeters
-   model->calculate_DRbar_parameters();
+   model->calculate_DRbar_masses();
    const MSSM<Two_scale> mssm(*model);
 
    MSSM_susy_scale_constraint<Two_scale>::apply();
@@ -690,7 +690,7 @@ void MSSM_iterative_low_scale_constraint::apply()
    assert(model && "Error: MSSM_low_scale_constraint:"
           " model pointer must not be zero");
 
-   model->calculate_DRbar_parameters();
+   model->calculate_DRbar_masses();
    update_scale();
    calculate_DRbar_gauge_couplings();
    calculate_DRbar_yukawa_couplings();
@@ -712,7 +712,7 @@ void MSSM_iterative_low_scale_constraint::apply()
          model->set_vd(vd);
          model->set_vu(vu);
 
-         model->calculate_DRbar_parameters();
+         model->calculate_DRbar_masses();
          model->calculate_Mhh_pole();
          model->calculate_MVZ_pole();
 

### 2014-10-07 test/test_NMSSM_ewsb.cpp

@@ -81,7 +81,7 @@ BOOST_AUTO_TEST_CASE( test_NMSSM_one_loop_tadpoles )
    setup_NMSSM_const(m, s, input);
 
    s.calcDrBarPars();
-   m.calculate_DRbar_parameters();
+   m.calculate_DRbar_masses();
 
    const double mt = s.displayDrBarPars().mt;
    const double sinthDRbar = s.calcSinthdrbar();
@@ -134,7 +134,7 @@ BOOST_AUTO_TEST_CASE( test_NMSSM_one_loop_ewsb )
    s.setMh2Squared(m.get_mHu2());
 
    s.calcDrBarPars();
-   m.calculate_DRbar_parameters();
+   m.calculate_DRbar_masses();
 
    const double mt = s.displayDrBarPars().mt;
    const int signMu = 1;
@@ -184,7 +184,7 @@ BOOST_AUTO_TEST_CASE( test_NMSSM_two_loop_tadpoles )
    setup_NMSSM_const(m, s, input);
 
    s.calcDrBarPars();
-   m.calculate_DRbar_parameters();
+   m.calculate_DRbar_masses();
 
    const double mt = s.displayDrBarPars().mt;
    const double sinthDRbar = s.calcSinthdrbar();
@@ -249,7 +249,7 @@ BOOST_AUTO_TEST_CASE( test_NMSSM_two_loop_ewsb )
    s.setMh2Squared(m.get_mHu2());
 
    s.calcDrBarPars();
-   m.calculate_DRbar_parameters();
+   m.calculate_DRbar_masses();
 
    const double mt = s.displayDrBarPars().mt;
    const int signMu = 1;

### 2014-10-07 test/test_NMSSM_low_scale_constraint.cpp

@@ -28,7 +28,7 @@ BOOST_AUTO_TEST_CASE( test_delta_alpha )
    setup_NMSSM_const(m, s, input);
    s.setData(oneset);
 
-   m.calculate_DRbar_parameters();
+   m.calculate_DRbar_masses();
    s.calcDrBarPars();
 
    NMSSM_low_scale_constraint<Two_scale> constraint(input, oneset);
@@ -66,7 +66,7 @@ BOOST_AUTO_TEST_CASE( test_low_energy_constraint )
    setup_NMSSM_const(m, s, input);
    s.setData(oneset);
 
-   m.calculate_DRbar_parameters();
+   m.calculate_DRbar_masses();
    s.calcDrBarPars();
 
    NMSSM_low_scale_constraint<Two_scale> constraint(input, oneset);

### 2014-10-07 test/test_NMSSM_one_loop_spectrum.cpp

@@ -56,7 +56,7 @@ void ensure_one_loop_ewsb(NMSSM<Two_scale>& m, NmssmSoftsusy& s)
    const double precision = 1.0e-5;
 
    s.calcDrBarPars();
-   m.calculate_DRbar_parameters();
+   m.calculate_DRbar_masses();
 
    const double mt = s.displayDrBarPars().mt;
    const int signMu = 1;
@@ -98,7 +98,7 @@ BOOST_AUTO_TEST_CASE( test_NMSSM_pole_masses )
    softsusy::numHiggsMassLoops = 1;
    s.physical(1);
    m.set_pole_mass_loop_order(1);
-   m.calculate_DRbar_parameters();
+   m.calculate_DRbar_masses();
    m.calculate_pole_masses();
 
    if (m.get_problems().have_problem()) {

### 2014-10-07 test/test_NMSSM_self_energies.cpp

@@ -41,7 +41,7 @@ BOOST_AUTO_TEST_CASE( test_NMSSM_self_energy_neutral_higgs )
    s.setMh2Squared(m.get_mHu2());
 
    s.calcDrBarPars();
-   m.calculate_DRbar_parameters();
+   m.calculate_DRbar_masses();
 
    // check tree-level
    DoubleVector hh_ss(s.displayDrBarPars().mh0);

### 2014-10-07 test/test_NMSSM_tree_level_spectrum.cpp

@@ -26,7 +26,7 @@ BOOST_AUTO_TEST_CASE( test_NMSSM_tree_level_masses )
    NmssmSoftsusy s;
    setup_NMSSM_const(m, s, input);
 
-   m.calculate_DRbar_parameters();
+   m.calculate_DRbar_masses();
    s.calcDrBarPars();
 
    // neutral CP even Higgs

### 2014-10-07 test/test_SMSSM_ewsb.cpp

@@ -110,7 +110,7 @@ BOOST_AUTO_TEST_CASE( test_SMSSM_one_loop_tadpoles )
    softsusy::Z3 = false;
    softsusy::numRewsbLoops = 1;
    s.calcDrBarPars();
-   m.calculate_DRbar_parameters();
+   m.calculate_DRbar_masses();
 
    // chech that there this is a valid point
    if (m.get_problems().have_problem()) {
@@ -181,7 +181,7 @@ BOOST_AUTO_TEST_CASE( test_SMSSM_one_loop_ewsb )
    s.setMh2Squared(m.get_mHu2());
 
    s.calcDrBarPars();
-   m.calculate_DRbar_parameters();
+   m.calculate_DRbar_masses();
 
    // chech that there this is a valid point
    if (m.get_problems().have_problem()) {

### 2014-10-07 test/test_SMSSM_one_loop_spectrum.cpp

@@ -56,7 +56,7 @@ void ensure_one_loop_ewsb(SMSSM<Two_scale>& m, NmssmSoftsusy& s)
    const double precision = 1.0e-5;
 
    s.calcDrBarPars();
-   m.calculate_DRbar_parameters();
+   m.calculate_DRbar_masses();
 
    const double mt = s.displayDrBarPars().mt;
    const int signMu = 1;
@@ -109,7 +109,7 @@ BOOST_AUTO_TEST_CASE( test_SMSSM_pole_masses )
    softsusy::numHiggsMassLoops = 1;
    s.physical(1);
    m.set_pole_mass_loop_order(1);
-   m.calculate_DRbar_parameters();
+   m.calculate_DRbar_masses();
    m.calculate_pole_masses();
 
    if (m.get_problems().have_problem()) {
@@ -258,7 +258,7 @@ BOOST_AUTO_TEST_CASE( test_self_energies )
    softsusy::MIXING = 1;
    ensure_tree_level_ewsb(m, s, input);
 
-   m.calculate_DRbar_parameters();
+   m.calculate_DRbar_masses();
    s.calcDrBarPars();
 
    if (m.get_problems().have_problem()) {

### 2014-10-07 test/test_SMSSM_tree_level_spectrum.cpp

@@ -21,7 +21,7 @@ BOOST_AUTO_TEST_CASE( test_SMSSM_tree_level_masses )
    NmssmSoftsusy s;
    setup_SMSSM(m, s, input);
 
-   m.calculate_DRbar_parameters();
+   m.calculate_DRbar_masses();
    m.reorder_DRbar_masses();
    s.calcDrBarPars();
 

### 2014-10-07 test/test_sfermions.cpp

@@ -110,7 +110,7 @@ void setup_mssm_models(MSSM<Two_scale>& m, MssmSoftsusy& softSusy)
    set_mssm_parameters(m, softSusy);
 
    ensure_tree_level_ewsb(m);
-   m.calculate_DRbar_parameters();
+   m.calculate_DRbar_masses();
    softSusy.calcDrBarPars();
 }
 
@@ -203,7 +203,7 @@ BOOST_AUTO_TEST_CASE( test_stop_different_sign )
    m.set_Mu(1200.0);
    s.setSusyMu(1200.0);
    ensure_tree_level_ewsb(m);
-   m.calculate_DRbar_parameters();
+   m.calculate_DRbar_masses();
    s.calcDrBarPars();
 
    DoubleVector mf(2);

### 2014-10-13 doc/flexiblesusy-paper.tex
updated paper to discuss issues with twoloop Higgs corrections a little more and to describe a change I *propose* we make.

@@
 
-\fs allows to add leading two-loop contributions to the CP-even Higgs
+\fs allows the user to add leading two-loop contributions to the CP-even Higgs
 tadpoles and self-energies.  For MSSM-like models (with two CP-even
@@
 corrections of the order $O(y_t^2 g_3^2)$, $O(y_b^2 g_3^2)$ from
-\cite{Degrassi:2009yq}, plus MSSM-like contributions of the order
-$O(y_\tau^4)$, $O(y_t^4 + y_t^2 y_b^2 + y_b^4)$
-\cite{Brignole:2001jy,Dedes:2003km}.
+\cite{Degrassi:2009yq}.  If one also wishes to include the MSSM contributions of the order $O(y_\tau^4)$, $O(y_t^4 + y_t^2 y_b^2 + y_b^4)$ as well
+\cite{Brignole:2001jy,Dedes:2003km} then one can also set both\footnote{Please note that with both of thse options set true it is the NMSSM corrections for $O(y_t^2 g_3^2)$, $O(y_b^2 g_3^2)$ which will be applied rather the the MSSM corrections for that order.} \code{UseHiggs2LoopMSSM = True} and \code{UseHiggs2LoopNMSSM = True}.
+
+ In the case of the NMSSM the default option is to have both flags set
+ to true as we feel that it is valuable to have these MSSM corrections
+ for scenarios where there singlet mixing is very small and in
+ particular for cross checks against the MSSM when close to the MSSM
+ limit of the model.  However in cases where the singlet mixing is
+ large the result at $O(y_t^4 + y_t^2 y_b^2 + y_b^4)$ and
+ $O(y_\tau^4)$ will not be complete and including these corrections
+ could in principle even make the numerical result further away from
+ the correct two loop result at that order if there is a cancellation
+ with the missing contributions.  So while including such partial
+ two-loop corrections is does not change the formal accuracy the user
+ should choose whether or not to include these corrections based on
+ the physics they study.  Similarly in models that go beyond the NMSSM
+ the user must decide based the physics whether on not these
+ corrections will give the leading two-loop corrections in their
+ model.
 
@@
 in the MSSM or \code{UseHiggs2LoopNMSSM = True} in the NMSSM in the
-model file.  This enables two-loop Higgs FORTRAN routines supplied by
-P.~Slavich from
+model file.  The former enables two-loop Higgs FORTRAN routines
+supplied by P.~Slavich from
 \cite{Degrassi:2001yf,Brignole:2001jy,Dedes:2002dy,Brignole:2002bz,Dedes:2003km}
 which add two-loop corrections of $\oatas$, $\oabas$, $\oatq$,
-$\oabq$, $\oatauq$ and $\oatab$.  Something similar is done for the
-NMSSM, but in this case the NMSSM $\oatas$, $\oabas$ pieces come from
-\cite{Degrassi:2009yq}, while for $\oatq$, $\oabq$, $\oatauq$ and
-$\oatab$ the MSSM pieces are used though it should be understood that
-these are not complete in the NMSSM.  Since the Higgs mass is a very
-important measurement and the two-loop corrections can be larger than
-the current experimental error \cite{Degrassi:2009yq} we recommend to
-set these switches to \code{True} in any MSSM-\footnote{In \fs a model
-  is MSSM-like if (i) its superpotential is approximately given by the
+$\oabq$, $\oatauq$ and $\oatab$.  The latter adds corrections
+calculated in the NMSSM at $\oatas$, $\oabas$ from
+\cite{Degrassi:2009yq}. If wants to include the $\oatq$, $\oabq$,
+$\oatauq$ and $\oatab$ the MSSM pieces are used though it NMSSM
+corrections at $\oatas$, $\oabas$ along with the MSSM corrections at
+$\oatq$, $\oabq$, $\oatauq$ and $\oatab$ then both flags
+\code{UseHiggs2LoopMSSM} and \code{UseHiggs2LoopNMSSM} should be set
+to \code{True}.  Since the Higgs mass is a very important measurement
+and the two-loop corrections can be larger than the current
+experimental error \cite{Degrassi:2009yq} we recommend to set these
+switches to \code{True} in any MSSM-\footnote{In \fs a model is
+  MSSM-like if (i) its superpotential is approximately given by the
   MSSM one, (ii) it implements $R$-parity conservation or something
@@
   NMSSM-like coupling to $t$, $b$ and $\tau$\@.  Examples for
-  NMSSM-like models are the USSM and the \ESSM.} model.  For models
+  NMSSM-like models are the USSM and the \ESSM.} model.  However in
+such models the user should still consider whether these corrections
+are really the leading corrections in the model or if there are other
+potentially large two-loop corrections.  For models
 with a more extended Higgs sector we recommend that the leading log

### 2014-10-13 doc/flexiblesusy-paper.tex
small edit to first sentence of introduction to address point 1 from referee 2.

@@
 space-time symmetries of the Poincar\'e
-group \cite{Coleman:1967ad,Haag:1974qh}, leading many to suspect that
+group, which still has scattering in the resulting quantum theory \cite{Coleman:1967ad,Haag:1974qh}. This leads many to suspect that
 SUSY may be realized in nature in some form. In particular

### 2014-10-14 doc/flexiblesusy-paper.tex
address referee 1 comment 4a: diagonalization precision

@@
 of diagonalization, \code{Eigen::SelfAdjointEigenSolver} from Eigen is
-used regardless of the matrix size.
+used regardless of the matrix size.  Note, that \fs uses double
+precision floating point data types with $15$ significant digits to
+store the mass matrices and the mass eigenvalues.  In case a particle
+multiplet contains a very split mass hierarchy, where the mass
+difference between the smallest and the largest mass in the multiplet
+is of the order or greater than 10 orders of magnitude, double
+precision data types are no longer sufficient.  In this case we
+recommend to either split the multiplet into sub-multiplets with
+smaller mass hierarchies, or integrate out the heavy states.
 

### 2014-10-14 doc/flexiblesusy-paper.tex
address referee 1 comment 4b: SUSY model restrictions
Added comments about the supported gauge groups and Yukawa coupling
structure at low energies.

@@
 
+The low-scale constraint matches the SUSY model to the Standard Model.
+Currently \fs allows only SUSY models with semisimple gauge groups,
+which contain the Standard Model gauge group $\SUc\times\SUL\times\UY$
+as factor.  This enables \fs to directly identify the strong,
+left-handed and hypercharge gauge couplings $g_i$ ($i=1,2,3$).
+
 The low-scale constraint automatically calculates the \DRbar\ gauge
@@
 %
-Afterwards, SARAH's expression for the Weinberg angle $\theta_W$ in
-terms of $M_{W,\text{susy}}^{\text{\DRbar}}(M_Z)$,
-$M_{Z,\text{susy}}^{\text{\DRbar}}(M_Z)$ is used to calculate
-$\theta_W$ in the SUSY model in the \DRbar\ scheme.  In the MSSM, for
-example, it yields
+Afterwards, the user-defined expression for the Weinberg angle
+$\theta_W$ in terms of $M_{W,\text{susy}}^{\text{\DRbar}}(M_Z)$ and
+$M_{Z,\text{susy}}^{\text{\DRbar}}(M_Z)$ (defined by the user in the
+SARAH model file) is used to calculate $\theta_W$ in the SUSY model in
+the \DRbar\ scheme.  In the MSSM, for example, it yields
 %
@@
 where $v_T$ is the vacuum expectation value of the scalar Higgs
-triplet field.
+triplet field.  In models with an additional $U(1)'$ gauge group, the
+additional $Z'$ gauge boson can mix with the $Z$ boson.  In such
+models the Weinberg angle can be defined by parametrizing the
+$Z$--$Z'$ mixing matrix as
+%
+\begin{align}
+  \begin{pmatrix}
+    \cos\theta_W & -\sin\theta_W \cos\theta_W' &  \sin\theta_W \sin\theta_W' \\
+    \sin\theta_W &  \cos\theta_W \cos\theta_W' & -\cos\theta_W \sin\theta_W' \\
+    0 & \sin\theta_W' & \cos\theta_W'
+  \end{pmatrix},
+\end{align}
+%
+where $\theta_W'$ is the $Z$--$Z'$ mixing angle.
 The running \DRbar\ $W$ and $Z$ boson masses are calculated in each
@@
 
-At the low-scale the \DRbar\ Yukawa coupling matrices
-$y_f^{\text{\DRbar}}(M_Z)$ ($f=u,d,e$) in the SUSY model are
-calculated automatically if the user has set $y_f$ to the value
-\code{Automatic} in the \fs model file.  This is done for example in
+The considered SUSY model is required to contain the three generations
+of Standard Model quarks and leptons.  If these particles acquire
+their masses due to Yukawa interactions with Higgs doublets, then the
+$3\times 3$ Yukawa matrices $y_f$ ($f=u,d,e$) can be calculated
+automatically in the \DRbar\ scheme at the scale $M_Z$ from the known
+Standard Model fermion masses by setting $y_f$ to the value
+\code{Automatic} in the \fs model file.
+%
+This is done for example in
 the CMSSM model file in \secref{sec:modfile}.  In this case \fs

### 2014-10-14 doc/flexiblesusy-paper.tex
expand undefined TeX macros

@@
 Currently \fs allows only SUSY models with semisimple gauge groups,
-which contain the Standard Model gauge group $\SUc\times\SUL\times\UY$
-as factor.  This enables \fs to directly identify the strong,
+which contain the Standard Model gauge group $SU(3)\times SU(2)\times
+U(1)$ as factor.  This enables \fs to directly identify the strong,
 left-handed and hypercharge gauge couplings $g_i$ ($i=1,2,3$).

### 2014-10-17 doc/flexiblesusy-paper.tex
Changed description of the two loop Higgs corrections in section 5 to match new plan which we agreed upon and Alex has implemented.  still need to fix later section.

@@
 
-\fs allows the user to add leading two-loop contributions to the CP-even Higgs
-tadpoles and self-energies.  For MSSM-like models (with two CP-even
-Higgs bosons, one CP-odd Higgs boson, one neutral Goldstone boson)
-these corrections can be enabled by setting \code{UseHiggs2LoopMSSM =
-  True} in the model file and by defining the effective $\mu$-term
-\code{EffectiveMu = \\[Mu]}.  This will add the zero-momentum
-corrections of the order $O(y_t^4 + y_b^2 y_t^2 + y_b^4)$, $O(y_t^2
-g_3^2)$, $O(y_b^2 g_3^2)$, $O(y_\tau^4)$ from
+\fs allows the user to add leading two-loop contributions to the
+CP-even Higgs tadpoles and self-energies.  For MSSM-like models (with
+two CP-even Higgs bosons, one CP-odd Higgs boson, one neutral
+Goldstone boson) routines dor caluclating these corrections will be
+generated by setting \code{UseHiggs2LoopMSSM = True} in the model file
+and by defining the effective $\mu$-term \code{EffectiveMu = \\[Mu]}.
+This will add the zero-momentum corrections of the order $O(y_t^4 +
+y_b^2 y_t^2 + y_b^4)$, $O(y_t^2 g_3^2)$, $O(y_b^2 g_3^2)$,
+$O(y_\tau^4)$ from
 \cite{Degrassi:2001yf,Brignole:2001jy,Dedes:2002dy,Brignole:2002bz,Dedes:2003km}.
@@
 Higgs bosons, one neutral Goldstone boson) the two-loop contributions
-are enabled by setting \code{UseHiggs2LoopNMSSM = True} and by
+are generated by setting \code{UseHiggs2LoopNMSSM = True} and by
 defining the effective $\mu$-term like \code{EffectiveMu = \\[Lambda]
@@
 corrections of the order $O(y_t^2 g_3^2)$, $O(y_b^2 g_3^2)$ from
-\cite{Degrassi:2009yq}.  If one also wishes to include the MSSM contributions of the order $O(y_\tau^4)$, $O(y_t^4 + y_t^2 y_b^2 + y_b^4)$ as well
-\cite{Brignole:2001jy,Dedes:2003km} then one can also set both\footnote{Please note that with both of thse options set true it is the NMSSM corrections for $O(y_t^2 g_3^2)$, $O(y_b^2 g_3^2)$ which will be applied rather the the MSSM corrections for that order.} \code{UseHiggs2LoopMSSM = True} and \code{UseHiggs2LoopNMSSM = True}.
-
- In the case of the NMSSM the default option is to have both flags set
- to true as we feel that it is valuable to have these MSSM corrections
- for scenarios where there singlet mixing is very small and in
- particular for cross checks against the MSSM when close to the MSSM
- limit of the model.  However in cases where the singlet mixing is
- large the result at $O(y_t^4 + y_t^2 y_b^2 + y_b^4)$ and
- $O(y_\tau^4)$ will not be complete and including these corrections
- could in principle even make the numerical result further away from
- the correct two loop result at that order if there is a cancellation
- with the missing contributions.  So while including such partial
- two-loop corrections is does not change the formal accuracy the user
- should choose whether or not to include these corrections based on
- the physics they study.  Similarly in models that go beyond the NMSSM
- the user must decide based the physics whether on not these
+\cite{Degrassi:2009yq} and also the MSSM contributions of the order
+$O(y_\tau^4)$, $O(y_t^4 + y_t^2 y_b^2 + y_b^4)$ as well
+\cite{Brignole:2001jy,Dedes:2003km} which only represent a partial
+correction for that order in the NMSSM, but can be a good approximation
+when singlet mixing is very small.
+
+ The corrections can then be used in the calculation of the Higgs
+ masses, when appropriate settings are selected in the SLHA file. Note
+ that even in the NMSSM the user miust make an important decision as
+ to whether or not to enable the generated MSSM corrections which are
+ incomplete in the NMSSM.  We feel that it is valuable to have
+ these MSSM corrections for scenarios where there singlet mixing is
+ very small and in particular for cross checks against the MSSM when
+ close to the MSSM limit of the model.  However in cases where the
+ singlet mixing is large the result at $O(y_t^4 + y_t^2 y_b^2 +
+ y_b^4)$ and $O(y_\tau^4)$ will not be complete and including these
+ corrections could in principle even make the numerical result further
+ away from the correct two loop result at that order if there is a
+ cancellation with the missing contributions.  So while including such
+ partial two-loop corrections does not change the formal accuracy
+ the user should choose whether or not to include these corrections
+ based on the physics they study.  Similarly in models that go beyond
+ the NMSSM the user must decide based the physics whether on not these
  corrections will give the leading two-loop corrections in their

### 2014-10-17 doc/flexiblesusy-paper.tex
Changed description of the two loop Higgs corrections in section 6.5 to match new plan which we agreed upon and Alex has implemented.

@@
 in the MSSM or \code{UseHiggs2LoopNMSSM = True} in the NMSSM in the
-model file.  The former enables two-loop Higgs FORTRAN routines
-supplied by P.~Slavich from
+model file.  The former provides routines that call the two-loop Higgs
+FORTRAN routines supplied by P.~Slavich from
 \cite{Degrassi:2001yf,Brignole:2001jy,Dedes:2002dy,Brignole:2002bz,Dedes:2003km}
-which add two-loop corrections of $\oatas$, $\oabas$, $\oatq$,
-$\oabq$, $\oatauq$ and $\oatab$.  The latter adds corrections
-calculated in the NMSSM at $\oatas$, $\oabas$ from
-\cite{Degrassi:2009yq}. If wants to include the $\oatq$, $\oabq$,
-$\oatauq$ and $\oatab$ the MSSM pieces are used though it NMSSM
-corrections at $\oatas$, $\oabas$ along with the MSSM corrections at
-$\oatq$, $\oabq$, $\oatauq$ and $\oatab$ then both flags
-\code{UseHiggs2LoopMSSM} and \code{UseHiggs2LoopNMSSM} should be set
-to \code{True}.  Since the Higgs mass is a very important measurement
-and the two-loop corrections can be larger than the current
-experimental error \cite{Degrassi:2009yq} we recommend to set these
-switches to \code{True} in any MSSM-\footnote{In \fs a model is
-  MSSM-like if (i) its superpotential is approximately given by the
-  MSSM one, (ii) it implements $R$-parity conservation or something
-  equivalent, and (iii) contains two CP-even and CP-odd Higgs bosons,
-  where one CP-odd Higgs boson may be a Goldstone boson, all with an
-  MSSM-like coupling to $t$, $b$ and $\tau$.} and
-NMSSM-like\footnote{In \fs a model is NMSSM-like if (i) its
-  superpotential is approximately given by the NMSSM one, (ii) it
+for calculating corrections of $\oatas$, $\oabas$, $\oatq$, $\oabq$,
+$\oatauq$ and $\oatab$.  The latter adds corrections calculated in the
+NMSSM at $\oatas$, $\oabas$ from \cite{Degrassi:2009yq} and partial
+corrections for the order $\oatq$, $\oabq$, $\oatauq$ and $\oatab$ from the
+MSSM.  These corrections can then be included in the
+calculation of the Higgs pole masses when running the spectrum
+generator by setting the appropriate SLHA flags.
+
+{\color{red} To include all two loop corrections which have been
+  generated the user may simply set entry 4 of the FlexibleSUSY block
+  in the SLHA file to 2.  The user may also want to include only some of
+  the corrections generated, for example in the NMSSM the user may decide not to use the partial corrections at order $\oatq$, $\oabq$, $\oatauq$ and $\oatab$ which have been computed in the MSSM.
+In that case the entries 9,10,11 and 12 may be used to switch on and off corrections at specific orders, 1 meaning the correction is included and 2 means the correction is not.
+Alex: can you check this part matches what you did and maybe fox details a little and also we should decide if it belongs here or if describe SLHA stuff later.}
+
+Since the Higgs mass is a very important measurement and
+the two-loop corrections can be larger than the current experimental
+error \cite{Degrassi:2009yq} we recommend to set these switches to
+\code{True} in any MSSM-\footnote{In \fs a model is MSSM-like if (i)
+  its superpotential is approximately given by the MSSM one, (ii) it
   implements $R$-parity conservation or something equivalent, and
-  (iii) contains three CP-even and CP-odd Higgs bosons, where one or
-  two CP-odd Higgs bosons may be Goldstone bosons, all with an
-  NMSSM-like coupling to $t$, $b$ and $\tau$\@.  Examples for
-  NMSSM-like models are the USSM and the \ESSM.} model.  However in
-such models the user should still consider whether these corrections
-are really the leading corrections in the model or if there are other
-potentially large two-loop corrections.  For models
+  (iii) contains two CP-even and CP-odd Higgs bosons, where one CP-odd
+  Higgs boson may be a Goldstone boson, all with an MSSM-like coupling
+  to $t$, $b$ and $\tau$.} and NMSSM-like\footnote{In \fs a model is
+  NMSSM-like if (i) its superpotential is approximately given by the
+  NMSSM one, (ii) it implements $R$-parity conservation or something
+  equivalent, and (iii) contains three CP-even and CP-odd Higgs
+  bosons, where one or two CP-odd Higgs bosons may be Goldstone
+  bosons, all with an NMSSM-like coupling to $t$, $b$ and $\tau$\@.
+  Examples for NMSSM-like models are the USSM and the \ESSM.} model.
+However in such models the user should still consider whether these
+corrections are really the leading corrections in the model or if
+there are other potentially large two-loop corrections.  For models
 with a more extended Higgs sector we recommend that the leading log

### 2014-10-17 doc/flexiblesusy-paper.tex
add new default FlexibleSUSY SLHA input options

@@
     6   2                    # beta-functions loop order
+    7   1                    # threshold corrections loop order
+    8   1                    # Higgs 2-loop corrections O(alpha_t alpha_s)
+    9   1                    # Higgs 2-loop corrections O(alpha_b alpha_s)
+   10   1                    # Higgs 2-loop corrections O(alpha_t^2 + alpha_t alpha_b + alpha_b^2)
+   11   1                    # Higgs 2-loop corrections O(alpha_tau^2)
 Block SPhenoInput            # SPheno specific input

### 2014-10-17 test/test_MSSM_low_scale_constraint_flavour.cpp
put MSSM fermion flavour mixing test into separate file

@@ -12,97 +12,40 @@
 #include "MSSM_two_scale_model.hpp"
 #include "MSSM_two_scale_low_scale_constraint.hpp"
 #include "softsusy.h"
+#include "flavoursoft.h"
 #include "wrappers.hpp"
 #include "ew_input.hpp"
+#include "ckm.hpp"
 
-DoubleVector calculate_gauge_couplings(MSSM<Two_scale> model,
-                                       MSSM_low_scale_constraint<Two_scale> constraint,
-                                       double scale)
-{
-   model.set_scale(scale);
-   constraint.set_model(&model);
-   constraint.apply();
-
-   DoubleVector g(3);
-   g(1) = model.get_g1();
-   g(2) = model.get_g2();
-   g(3) = model.get_g3();
-
-   return g;
-}
-
-BOOST_AUTO_TEST_CASE( test_threshold_corrections )
-{
-   MSSM<Two_scale> m; MssmSoftsusy s;
-   MSSM_input_parameters input;
-   QedQcd oneset;
-   setup_MSSM(m, s, input);
-
-   MSSM_low_scale_constraint<Two_scale> constraint(input, oneset);
-
-   const double Q1 = constraint.get_scale();
-   const double Q2 = 2. * Q1;
-   const double gut_normalization = 3./5.;
-   DoubleVector g_old(3);
-   g_old(1) = Electroweak_constants::g1;
-   g_old(2) = Electroweak_constants::g2;
-   g_old(3) = Electroweak_constants::g3;
-   DoubleVector prefactor(3);
-   for (int i = 1; i <= 3; i++)
-      prefactor(i) = 1. / (oneOver16PiSqr * Power(g_old(i),3));
-
-   const DoubleVector g_Q1(calculate_gauge_couplings(m, constraint, Q1));
-   const DoubleVector g_Q2(calculate_gauge_couplings(m, constraint, Q2));
-
-   const DoubleVector beta_numeric((g_Q1 - g_Q2) * prefactor * (1. / log(Q1/Q2)));
-   DoubleVector beta_SM(3);
-   beta_SM(1) = 41./6. * gut_normalization;
-   beta_SM(2) = -19./6.;
-   beta_SM(3) = -7. - 2./3;
[…]

### 2014-10-23 doc/flexiblesusy-paper.tex
modified the two loop corrections desscription a little.

--- 2014-10-17 doc/flexiblesusy-paper.tex
@@
 \def\oabatau{O(\ab \atau)}
+\def\oatplusabsq{O((\at+\ab)^2)}
 \def\oas{O(\as)}
@@
 
-{\color{red} To include all two loop corrections which have been
-  generated the user may simply set entry 4 of the FlexibleSUSY block
-  in the SLHA file to 2.  The user may also want to include only some of
-  the corrections generated, for example in the NMSSM the user may decide not to use the partial corrections at order $\oatq$, $\oabq$, $\oatauq$ and $\oatab$ which have been computed in the MSSM.
-In that case the entries 9,10,11 and 12 may be used to switch on and off corrections at specific orders, 1 meaning the correction is included and 2 means the correction is not.
-Alex: can you check this part matches what you did and maybe fox details a little and also we should decide if it belongs here or if describe SLHA stuff later.}
+{\color{red} When two loop corrections have been enabled in the SLHA
+  file by setting entry 4 of the FlexibleSUSY block to 2 the user may
+  also select individual corrections. Entries 9,10,11 and 12,
+  corrspond to two loop corrections of order $\oatas$, $\oabas$,
+  $\oatplusabsq$ and $\oatauq$ respectively and will be disabled when
+  the corresponding entry is set to zero. In this way, for example, in
+  the NMSSM the user may decide not to use the partial corrections at
+  order $\oatplusabsq$ and $\oatauq$, all of which have
+  only been computed in the MSSM. }
 

### 2014-10-23 doc/flexiblesusy-paper.tex
modified a little more and fixed typo.

@@
 
-{\color{red} When two loop corrections have been enabled in the SLHA
+  When two loop corrections have been enabled in the SLHA
   file by setting entry 4 of the FlexibleSUSY block to 2 the user may
   also select individual corrections. Entries 9,10,11 and 12,
-  corrspond to two loop corrections of order $\oatas$, $\oabas$,
+  correspond to two loop corrections of order $\oatas$, $\oabas$,
   $\oatplusabsq$ and $\oatauq$ respectively and will be disabled when
@@
   order $\oatplusabsq$ and $\oatauq$, all of which have
-  only been computed in the MSSM. }
-
-Since the Higgs mass is a very important measurement and
-the two-loop corrections can be larger than the current experimental
-error \cite{Degrassi:2009yq} we recommend to set these switches to
+  only been computed in the MSSM.
+
+Since the Higgs mass is a very important measurement and the two-loop
+corrections can be larger than the current experimental error
+\cite{Degrassi:2009yq} we recommend to set these switches to
 \code{True} in any MSSM-\footnote{In \fs a model is MSSM-like if (i)
@@
 corrections are really the leading corrections in the model or if
-there are other potentially large two-loop corrections.  For models
-with a more extended Higgs sector we recommend that the leading log
-two-loop corrections are estimated by generalizing those of the MSSM
-or NMSSM.
+there are other potentially large two-loop corrections which are
+missing.  For models with a more extended Higgs sector we recommend
+that the leading log two-loop corrections are estimated by
+generalizing those of the MSSM or NMSSM.
 

### 2014-10-23 doc/flexiblesusy-paper.tex
added a footnote to section 5 where the NMSSM/MSSM corrections issue is dicussed.

@@
 correction for that order in the NMSSM, but can be a good approximation
-when singlet mixing is very small.
+when singlet mixing is very small\footnote{These corrections may be disabled in the SLHA file, as described in section \secref{sec:PoleMasses}.}.
 
@@
   When two loop corrections have been enabled in the SLHA
-  file by setting entry 4 of the FlexibleSUSY block to 2 the user may
-  also select individual corrections. Entries 9,10,11 and 12,
+  file by setting entry $4$ of the FlexibleSUSY block to $2$ the user may
+  also select individual corrections. Entries $9$,$10$,$11$ and $12$,
   correspond to two loop corrections of order $\oatas$, $\oabas$,

### 2014-10-23 doc/flexiblesusy-paper.tex
Fix a typo in the paper

@@
 \section{Introduction}
+
 Supersymmetry (SUSY) provides the only non-trivial way to extend the
@@
 Detailed phenomenological studies have been carried out for scenarios
-within the minimal supersymmetric standard model (MSSM).  Such work
-has been greatly aided by public spectrum generators for the MSSM
+within the minimal supersymmetric standard model (MSSM), for a review
+see \cite{Chung:2003fi}.  Such work has been greatly aided by public
+spectrum generators for the MSSM
 \cite{Allanach:2001kg,Porod:2003um,Djouadi:2002ze,Baer:1993ae,Chowdhury:2011zr},
@@
 parameters. Beyond the MSSM there are also two public spectrum
-generators \cite{Ellwanger:2006rn,Allanach:2013kza} for the next to
-minimal supersymmetric standard model (NMSSM) \cite{NMSSM} (or for recent
-reviews see \cite{Ellwanger:2009dp,Maniatis:2009re}).
+generators \cite{Ellwanger:2004xm, Ellwanger:2005dv,
+  Ellwanger:2006rn,Ellwanger:2008py, Allanach:2013kza} for the next to
+minimal supersymmetric standard model (NMSSM)  (for
+recent reviews see \cite{Ellwanger:2009dp, Maniatis:2009re}).
 
@@
 These developments motivate the exploration of non-minimal SUSY models
-which ameliorate the naturalness problems, as can happen in the \ESSM
-\cite{King:2005jy,King:2005my,King:2007uj,Athron:2010zz,Nevzorov:2012hs,Nevzorov:2013tta,Nevzorov:2013ixa}, USSM
-\cite{Fayet:1977yc,Suematsu:1994qm,Cvetic:1995rj,deCarlos:1997yv,Cvetic:1997ky,Demir:1998dk,Langacker:1998tc,Erler:2002pr,Choi:2006fz,Ham:2007wc,Langacker:2008yv,Ham:2008xf,Kalinowski:2008iq} or from other gauge extensions \cite{Batra:2003nj, Medina:2009ey, Bharucha:2013ela},
-by raising the tree level Higgs mass. At the same time they can also
+which ameliorate the naturalness problems by raising the tree level
+Higgs mass, as can happen in the USSM
+\cite{Fayet:1977yc,Cvetic:1997ky,Langacker:2008yv}, \ESSM
+\cite{King:2005jy,Athron:2010zz} and similar models
+\cite{Nevzorov:2012hs}, or from other gauge extensions
+\cite{Batra:2003nj, Bharucha:2013ela}. At the same time they can also
 motivate models that are developed with a fresh perspective, based on
@@
 \fs already comes with plenty of predefined models: the CMSSM (simply
-called \code{MSSM}), the NMSSM \cite{NMSSM} in it's $Z_3$-symmetric form (called \code{NMSSM}), $Z_3$-violating NMSSM (\code{SMSSM}), the USSM (\code{UMSSM})
-\cite{Fayet:1977yc,Suematsu:1994qm,Cvetic:1995rj,deCarlos:1997yv,Cvetic:1997ky,Demir:1998dk,Langacker:1998tc,Erler:2002pr,Choi:2006fz,Ham:2007wc,Langacker:2008yv,Ham:2008xf,Kalinowski:2008iq},
+called \code{MSSM}), the NMSSM in its $Z_3$-symmetric form (called \code{NMSSM}), $Z_3$-violating NMSSM (\code{SMSSM}), the USSM (\code{UMSSM}),
 the NUHM \ESSM (\code{E6SSM}) \cite{Athron:2007en}, the
@@
 
-SUSYScale = Sqrt[M[Su[1]]*M[Su[6]]];
+SUSYScale = Sqrt[Product[M[Su[i]]^(Abs[ZU[i,3]]^2 + Abs[ZU[i,6]]^2), {i,6}]];
 
@@
   EWSB conditions and calculates the pole mass spectrum.  The
-  SUSY-scale, $M_S$, is usually defined as $M_S =
-  \sqrt{m_{\tilde{t}_1}m_{\tilde{t}_2}}$.  However, in the example
-  above, where sfermion flavour violation is enabled, it has the value
-  $M_S = \sqrt{m_{\tilde{u}_1}m_{\tilde{u}_6}}$, where
-  $m_{\tilde{u}_1}$ and $m_{\tilde{u}_6}$ are the \DRbar\ masses of
-  the lightest and heaviest up-type squark, respectively.
+  SUSY-scale, $M_S$, is defined as
+  \begin{align}
+    M_S = \sqrt{\prod_{i=1}^6 m_{\tilde{u}_i}^{|(Z_u)_{i3}|^2 + |(Z_u)_{i6}|^2}} ,
+    \label{eq:definition_of_MS}
+  \end{align}
+  where $m_{\tilde{u}_i}$ is the \DRbar\ mass of the $i$th up-type
+  squark and $Z_u$ is the up-type squark mixing matrix.  The
+  definition \eqref{eq:definition_of_MS} is equivalent to the usual
+  choice $M_S = \sqrt{m_{\tilde{t}_1}m_{\tilde{t}_2}}$ without squark
+  flavour mixing.
 %
@@
 \fs allows the user to add leading two-loop contributions to the
-CP-even Higgs tadpoles and self-energies.  For MSSM-like models (with
-two CP-even Higgs bosons, one CP-odd Higgs boson, one neutral
-Goldstone boson) routines dor caluclating these corrections will be
-generated by setting \code{UseHiggs2LoopMSSM = True} in the model file
-and by defining the effective $\mu$-term \code{EffectiveMu = \\[Mu]}.
-This will add the zero-momentum corrections of the order $O(y_t^4 +
-y_b^2 y_t^2 + y_b^4)$, $O(y_t^2 g_3^2)$, $O(y_b^2 g_3^2)$,
-$O(y_\tau^4)$ from
+CP-even and CP-odd Higgs self-energies as well as to the CP-even Higgs
+tadpoles.  For MSSM-like models (with two CP-even Higgs bosons, one
+CP-odd Higgs boson, one neutral Goldstone boson) routines dor
+caluclating these corrections will be generated by setting
+\code{UseHiggs2LoopMSSM = True} in the model file and by defining the
+effective $\mu$-term \code{EffectiveMu = \\[Mu]}.  This will add the
+zero-momentum corrections of the order $O(y_t^4 + y_b^2 y_t^2 +
+y_b^4)$, $O(y_t^2 g_3^2)$, $O(y_b^2 g_3^2)$, $O(y_\tau^4)$ from
 \cite{Degrassi:2001yf,Brignole:2001jy,Dedes:2002dy,Brignole:2002bz,Dedes:2003km}.
@@
 \cite{Brignole:2001jy,Dedes:2003km} which only represent a partial
-correction for that order in the NMSSM, but can be a good approximation
-when singlet mixing is very small\footnote{These corrections may be disabled in the SLHA file, as described in section \secref{sec:PoleMasses}.}.
+correction for that order in the NMSSM, but can be a good
+approximation when singlet mixing is very small\footnote{These
+  corrections may be disabled in the SLHA file, as described in
+  section \secref{sec:PoleMasses}.}.
 
@@
   conditions at the mass scale $M_S$ of the SUSY particles.  The value
-  of $M_S$ is defined by the model file variable \code{SUSYScale}.  In
-  the example model file in \secref{sec:modfile} it is defined to be
-  $\sqrt{m_{\tilde{u}_1} m_{\tilde{u}_6}}$, where $m_{\tilde{u}_1}$ is
-  the mass of the lightest, and $m_{\tilde{u}_6}$ is the mass of the
-  heaviest up-type squark.  The \code{apply()} function for this
-  constraint sets the model parameters to the values defined in
-  \code{SUSYScaleInput}.  Afterwards, \code{apply()} solves the EWSB
-  equations at the loop level by adjusting the parameters given in
+  of $M_S$ is defined in the model file variable \code{SUSYScale}.  In
+  the example model file in \secref{sec:modfile} it is given by the
+  expression written in Eq.~\eqref{eq:definition_of_MS}.  The
+  \code{apply()} function for this constraint sets the model
+  parameters to the values defined in \code{SUSYScaleInput}.
+  Afterwards, \code{apply()} solves the EWSB equations at the loop
+  level by adjusting the parameters given in
   \code{EWSBOutputParameters} such that the effective Higgs potential
@@
     \begin{align}
-      M_S = \sqrt{m_{\tilde{u}_1}m_{\tilde{u}_6}},
+      M_S = \sqrt{\prod_{i=1}^6 m_{\tilde{u}_i}^{|(Z_u)_{i3}|^2 + |(Z_u)_{i6}|^2}} ,
     \end{align}
-    where $m_{\tilde{u}_1}$ and $m_{\tilde{u}_6}$ are the lightest and
-    heaviest up-type squarks, respectively.
+    where $m_{\tilde{u}_i}$ is the \DRbar\ mass of the $i$th up-type
+    squark.
   \item Impose the SUSY-scale constraint (\code{SUSYScaleInput}).  In
@@
 spectrum generator for a constrained version of the \ESSM
-\cite{Athron:2009ue, Athron:2009bs, Athron:2011wu, Athron:2012sq,
-  Athron:2012pw, Athron:2013ipa}.  The RGEs could be compared directly
-in unit tests and were found to match.  The handwritten code doesn't
-include full one-loop self-energies or tadpoles so tests on these were
-not carried out.  Although the generators assume different constraints
-they could be compared by using the output of the CE$_6$SSM generator
-as an input to \fs and the spectra were found to be in reasonable
-agreement, given the different levels of precision.
+\cite{Athron:2009ue, Athron:2009bs, Athron:2012pw}.  The RGEs could be
+compared directly in unit tests and were found to match.  The
+handwritten code doesn't include full one-loop self-energies or
+tadpoles so tests on these were not carried out.  Although the
+generators assume different constraints they could be compared by
+using the output of the CE$_6$SSM generator as an input to \fs and the
+spectra were found to be in reasonable agreement, given the different
+levels of precision.
 
@@
   %717 citations counted in INSPIRE as of 21 Feb 2014
+\bibitem{Chung:2003fi} D.~J.~H.~Chung, L.~L.~Everett, G.~L.~Kane,
+  S.~F.~King, J.~D.~Lykken and L.~T.~Wang,
+  %``The Soft supersymmetry breaking Lagrangian: Theory and applications,''
+  Phys.\ Rept.\  {\bf 407}, 1 (2005)
+  [hep-ph/0312378].
+  %%CITATION = HEP-PH/0312378;%%
+  %389 citations counted in INSPIRE as of 23 Oct 2014
 
@@
 
+%\cite{Ellwanger:2004xm}
+\bibitem{Ellwanger:2004xm} 
+  U.~Ellwanger, J.~F.~Gunion and C.~Hugonie,
+  %``NMHDECAY: A Fortran code for the Higgs masses, couplings and decay widths in the NMSSM,''
+  JHEP {\bf 0502}, 066 (2005)
+  [hep-ph/0406215].
+  %%CITATION = HEP-PH/0406215;%%
+  %256 citations counted in INSPIRE as of 23 Oct 2014
+
+%\cite{Ellwanger:2005dv}
+\bibitem{Ellwanger:2005dv} 
+  U.~Ellwanger and C.~Hugonie,
+  %``NMHDECAY 2.0: An Updated program for sparticle masses, Higgs masses, couplings and decay widths in the NMSSM,''
+  Comput.\ Phys.\ Commun.\  {\bf 175}, 290 (2006)
+  [hep-ph/0508022].
+  %%CITATION = HEP-PH/0508022;%%
+  %227 citations counted in INSPIRE as of 23 Oct 2014
+
 %\cite{Ellwanger:2006rn}
@@
   %81 citations counted in INSPIRE as of 03 Nov 2013
+
+  %\cite{Ellwanger:2008py}
+\bibitem{Ellwanger:2008py} 
+  U.~Ellwanger, C.-C.~Jean-Louis and A.~M.~Teixeira,
+  %``Phenomenology of the General NMSSM with Gauge Mediated Supersymmetry Breaking,''
+  JHEP {\bf 0805}, 044 (2008)
+  [arXiv:0803.2962 [hep-ph]].
+  %%CITATION = ARXIV:0803.2962;%%
+  %34 citations counted in INSPIRE as of 23 Oct 2014
+
 %\cite{Allanach:2013kza}
@@
   %2 citations counted in INSPIRE as of 15 Jun 2014
-\bibitem{NMSSM} P. Fayet, Nucl. Phys. B \textbf{90} (1975) 104; Phys. Lett.
-B \textbf{64} (1976) 159; Phys. Lett. B \textbf{69} (1977) 489 and Phys. Lett. B
-\textbf{84} (1979) 416; H.P. Nilles, M. Srednicki and D. Wyler, Phys. Lett. B
-\textbf{120} (1983) 346; J.M. Frere, D.R. Jones and S. Raby, Nucl. Phys. B
-\textbf{222} (1983) 11; J.P. Derendinger and C.A. Savoy, Nucl. Phys. B
-\textbf{237} (1984) 307;  A.I. Veselov, M.I. Vysotsky and K.A. Ter-Martirosian,
-Sov. Phys. JETP \textbf{63} (1986) 489; J.R. Ellis, J.F. Gunion, H.E. Haber, L.
-Roszkowski and F. Zwirner, Phys. Rev. D \textbf{39}  (1989) 844; M. Drees, Int.
-J. Mod. Phys. A \textbf{4}  (1989) 3635; U. Ellwanger, M. Rausch de
-Traubenberg and C.A. Savoy, Phys. 
-Lett. B \textbf{315} (1993) 331, Z. Phys. C {\bf 67} (1995) 665 and Nucl. Phys.
-B \textbf{492} (1997) 307; U.~Ellwanger, Phys.\ Lett.\  B {\bf 303} (1993) 271; P.
-Pandita, Z. Phys. C \textbf{59} (1993) 575; T. Elliott, S.F. King and P.L.
-White, Phys. Rev. D {\bf 49} (1994) 2435; S.F. King and P.L. White, Phys. Rev. D
-\textbf{52} (1995) 4183;  F.~Franke and H.~Fraas, Int.\ J.\ Mod.\ Phys.\  A {\bf
-12} (1997) 479.   D.~J.~Miller, R.~Nevzorov and P.~M.~Zerwas,  Nucl.\ Phys.\ B {\bf 681}, 3 (2004) [hep-ph/0304049].
 %\cite{Ellwanger:2009dp}
@@
 
-%\cite{King:2005jy}
-\bibitem{King:2005jy} 
-  S.~F.~King, S.~Moretti and R.~Nevzorov,
-  %``Theory and phenomenology of an exceptional supersymmetric standard model,''
-  Phys.\ Rev.\ D {\bf 73}, 035009 (2006)
-  [hep-ph/0510419].
-  %%CITATION = HEP-PH/0510419;%%
-  %130 citations counted in INSPIRE as of 03 Nov 2013
-
-%\cite{King:2005my}
-\bibitem{King:2005my} 
-  S.~F.~King, S.~Moretti and R.~Nevzorov,
-  %``Exceptional supersymmetric standard model,''
-  Phys.\ Lett.\ B {\bf 634}, 278 (2006)
-  [hep-ph/0511256].
-  %%CITATION = HEP-PH/0511256;%%
-  %90 citations counted in INSPIRE as of 09 Jun 2014
-
-%\cite{King:2007uj}
-\bibitem{King:2007uj} 
-  S.~F.~King, S.~Moretti and R.~Nevzorov,
-  %``Gauge coupling unification in the exceptional supersymmetric standard model,''
-  Phys.\ Lett.\ B {\bf 650}, 57 (2007)
-  [hep-ph/0701064].
-  %%CITATION = HEP-PH/0701064;%%
-  %46 citations counted in INSPIRE as of 09 Jun 2014
-
-%\cite{Athron:2010zz}
-\bibitem{Athron:2010zz} 
-  P.~Athron, J.~P.~Hall, R.~Howl, S.~F.~King, D.~J.~Miller, S.~Moretti and R.~Nevzorov,
-  %``Aspects of the exceptional supersymmetric standard model,''
-  Nucl.\ Phys.\ Proc.\ Suppl.\  {\bf 200-202}, 120 (2010).
-  %%CITATION = NUPHZ,200-202,120;%%
-  %5 citations counted in INSPIRE as of 09 Jun 2014
-
-%\cite{Nevzorov:2012hs}
-\bibitem{Nevzorov:2012hs} 
-  R.~Nevzorov,
-  %``E6 Inspired SUSY Models with Exact Custodial Symmetry,''
-  Phys.\ Rev.\ D {\bf 87}, 015029 (2013)
-  [arXiv:1205.5967 [hep-ph]].
-  %%CITATION = ARXIV:1205.5967;%%
-  %4 citations counted in INSPIRE as of 13 Jun 2014
-
-%\cite{Nevzorov:2013tta}
-\bibitem{Nevzorov:2013tta} 
-  R.~Nevzorov and S.~Pakvasa,
-  %``Exotic Higgs decays in the $E_6$ inspired SUSY models,''
-  Phys.\ Lett.\ B {\bf 728}, 210 (2014)
-  [arXiv:1308.1021 [hep-ph]].
-  %%CITATION = ARXIV:1308.1021;%%
-  %2 citations counted in INSPIRE as of 13 Jun 2014
-
-%\cite{Nevzorov:2013ixa}
-\bibitem{Nevzorov:2013ixa} 
-  R.~Nevzorov,
-  %``Quasi-fixed point scenarios and the Higgs mass in the E6 inspired SUSY models,''
-  Phys.\ Rev.\ D {\bf 89}, 055010 (2014)
-  [arXiv:1309.4738 [hep-ph]].
-  %%CITATION = ARXIV:1309.4738;%%
-  %1 citations counted in INSPIRE as of 13 Jun 2014
-
 %\cite{Fayet:1977yc}
@@
 %\cite{Suematsu:1994qm}
-\bibitem{Suematsu:1994qm} 
-  D.~Suematsu and Y.~Yamagishi,
-  %``Radiative symmetry breaking in a supersymmetric model with an extra U(1),''
-  Int.\ J.\ Mod.\ Phys.\ A {\bf 10}, 4521 (1995)
-  [hep-ph/9411239].
-  %%CITATION = HEP-PH/9411239;%%
-  %116 citations counted in INSPIRE as of 09 Jun 2014
-
-  %\cite{Cvetic:1995rj}
-\bibitem{Cvetic:1995rj} 
-  M.~Cvetic and P.~Langacker,
-  %``Implications of Abelian extended gauge structures from string models,''
-  Phys.\ Rev.\ D {\bf 54}, 3570 (1996)
-  [hep-ph/9511378].
-  %%CITATION = HEP-PH/9511378;%%
-  %207 citations counted in INSPIRE as of 09 Jun 2014
-
-\bibitem{deCarlos:1997yv} 
-  B.~de Carlos and J.~R.~Espinosa,
-  %``Cold dark matter candidate in a class of supersymmetric models with an extra U(1),''
-  Phys.\ Lett.\ B {\bf 407}, 12 (1997)
-  [hep-ph/9705315].
-  %%CITATION = HEP-PH/9705315;%%
-  %51 citations counted in INSPIRE as of 09 Jun 2014
-%\cite{Cvetic:1997ky}
+
 \bibitem{Cvetic:1997ky}
@@
 
-%\cite{Demir:1998dk}
-\bibitem{Demir:1998dk} 
-  D.~A.~Demir and N.~K.~Pak,
-  %``One loop effects in supergravity models with an additional U(1),''
-  Phys.\ Rev.\ D {\bf 57}, 6609 (1998)
-  [hep-ph/9809357].
-  %%CITATION = HEP-PH/9809357;%%
-  %19 citations counted in INSPIRE as of 09 Jun 2014
-
-%\cite{Langacker:1998tc}
-\bibitem{Langacker:1998tc} 
-  P.~Langacker and J.~Wang,
-  %``U(1)-prime symmetry breaking in supersymmetric E(6) models,''
-  Phys.\ Rev.\ D {\bf 58}, 115010 (1998)
-  [hep-ph/9804428].
-  %%CITATION = HEP-PH/9804428;%%
-  %108 citations counted in INSPIRE as of 09 Jun 2014
-
-%\cite{Erler:2002pr}
-\bibitem{Erler:2002pr}
-  J.~Erler, P.~Langacker and T.~-j.~Li,
-  %``The $Z$ - $Z^\prime$ mass hierarchy in a supersymmetric model with a secluded U(1) -prime breaking sector,''
-  Phys.\ Rev.\ D {\bf 66} (2002) 015002
-  [hep-ph/0205001].
-  %%CITATION = HEP-PH/0205001;%%
-  %84 citations counted in INSPIRE as of 04 Jun 2014
-
-%\cite{Choi:2006fz}
-\bibitem{Choi:2006fz} 
-  S.~Y.~Choi, H.~E.~Haber, J.~Kalinowski and P.~M.~Zerwas,
-  %``The Neutralino sector in the U(1)-extended supersymmetric standard model,''
-  Nucl.\ Phys.\ B {\bf 778}, 85 (2007)
-  [hep-ph/0612218].
-  %%CITATION = HEP-PH/0612218;%%
-  %43 citations counted in INSPIRE as of 09 Jun 2014
-
-%\cite{Ham:2007wc}
-\bibitem{Ham:2007wc} 
-  S.~W.~Ham and S.~K.~OH,
-  %``Electroweak phase transition in MSSM with U(1)-prime in explicit CP violation scenario,''
-  Phys.\ Rev.\ D {\bf 76}, 095018 (2007)
-  [arXiv:0708.1785 [hep-ph]].
-  %%CITATION = ARXIV:0708.1785;%%
-  %20 citations counted in INSPIRE as of 09 Jun 2014
-
 %\cite{Langacker:2008yv}
@@
 
-%\cite{Ham:2008xf}
-\bibitem{Ham:2008xf} 
-  S.~W.~Ham, T.~Hur, P.~Ko and S.~K.~Oh,
-  %``Neutral scalar Higgs bosons in the USSM at the LHC,''
-  J.\ Phys.\ G {\bf 35}, 095007 (2008)
-  [arXiv:0801.2361 [hep-ph]].
-  %%CITATION = ARXIV:0801.2361;%%
-  %10 citations counted in INSPIRE as of 09 Jun 2014
-
-%\cite{Kalinowski:2008iq}
-\bibitem{Kalinowski:2008iq} 
-  J.~Kalinowski, S.~F.~King and J.~P.~Roberts,
-  %``Neutralino Dark Matter in the USSM,''
-  JHEP {\bf 0901}, 066 (2009)
-  [arXiv:0811.2204 [hep-ph]].
-  %%CITATION = ARXIV:0811.2204;%%
-  %30 citations counted in INSPIRE as of 09 Jun 2014
-%\cite{Bharucha:2013ela}
+%\cite{King:2005jy}
+\bibitem{King:2005jy} 
+  S.~F.~King, S.~Moretti and R.~Nevzorov,
+  %``Theory and phenomenology of an exceptional supersymmetric standard model,''
+  Phys.\ Rev.\ D {\bf 73}, 035009 (2006)
+  [hep-ph/0510419].
+  %%CITATION = HEP-PH/0510419;%%
+  %130 citations counted in INSPIRE as of 03 Nov 2013
+
+%\cite{Athron:2010zz}
+\bibitem{Athron:2010zz} 
+  P.~Athron, J.~P.~Hall, R.~Howl, S.~F.~King, D.~J.~Miller, S.~Moretti and R.~Nevzorov,
+  %``Aspects of the exceptional supersymmetric standard model,''
+  Nucl.\ Phys.\ Proc.\ Suppl.\  {\bf 200-202}, 120 (2010).
+  %%CITATION = NUPHZ,200-202,120;%%
+  %5 citations counted in INSPIRE as of 09 Jun 2014
+
+%\cite{Nevzorov:2012hs}
+\bibitem{Nevzorov:2012hs} 
+  R.~Nevzorov,
+  %``E6 Inspired SUSY Models with Exact Custodial Symmetry,''
+  Phys.\ Rev.\ D {\bf 87}, 015029 (2013)
+  [arXiv:1205.5967 [hep-ph]].
+  %%CITATION = ARXIV:1205.5967;%%
+  %4 citations counted in INSPIRE as of 13 Jun 2014
 
@@
   %174 citations counted in INSPIRE as of 12 Jun 2014
-%\cite{Medina:2009ey}
-\bibitem{Medina:2009ey} 
-  A.~D.~Medina, N.~R.~Shah and C.~E.~M.~Wagner,
-  %``A Heavy Higgs and a Light Sneutrino NLSP in the MSSM with Enhanced SU(2) D-terms,''
-  Phys.\ Rev.\ D {\bf 80}, 015001 (2009)
-  [arXiv:0904.1625 [hep-ph]].
-  %%CITATION = ARXIV:0904.1625;%%
-  %9 citations counted in INSPIRE as of 12 Jun 2014
 
@@
 
-%\cite{Athron:2011wu}
-\bibitem{Athron:2011wu} 
-  P.~Athron, S.~F.~King, D.~J.~Miller, S.~Moretti and R.~Nevzorov,
-  %``LHC Signatures of the Constrained Exceptional Supersymmetric Standard Model,''
-  Phys.\ Rev.\ D {\bf 84}, 055006 (2011)
-  [arXiv:1102.4363 [hep-ph]].
-  %%CITATION = ARXIV:1102.4363;%%
-  %25 citations counted in INSPIRE as of 09 Jun 2014
-
-%\cite{Athron:2012sq}
-\bibitem{Athron:2012sq} 
-  P.~Athron, S.~F.~King, D.~J.~Miller, S.~Moretti and R.~Nevzorov,
-  %``Constrained Exceptional Supersymmetric Standard Model with a Higgs Near 125 GeV,''
-  Phys.\ Rev.\ D {\bf 86}, 095003 (2012)
-  [arXiv:1206.5028 [hep-ph]].
-  %%CITATION = ARXIV:1206.5028;%%
-  %21 citations counted in INSPIRE as of 09 Jun 2014
-
 %\cite{Athron:2012pw}
@@
 
-  %\cite{Athron:2013ipa}
-\bibitem{Athron:2013ipa} 
-  P.~Athron, M.~Binjonaid and S.~F.~King,
-  %``Fine Tuning in the Constrained Exceptional Supersymmetric Standard Model,''
-  Phys.\ Rev.\ D {\bf 87}, no. 11, 115023 (2013)
-  [arXiv:1302.5291 [hep-ph]].
-  %%CITATION = ARXIV:1302.5291;%%
-  %10 citations counted in INSPIRE as of 09 Jun 2014
-
 \end{thebibliography}

### 2014-10-24 doc/flexiblesusy-paper.tex
added new references for NMSSMTools.

--- 2014-10-23 doc/flexiblesusy-paper.tex
@@
 parameters. Beyond the MSSM there are also two public spectrum
-generators \cite{Ellwanger:2006rn,Allanach:2013kza} for the next to
+generators \cite{Ellwanger:2004xm, Ellwanger:2005dv, Ellwanger:2006rn,Ellwanger:2008py, Allanach:2013kza} for the next to
 minimal supersymmetric standard model (NMSSM) \cite{NMSSM} (or for recent
-reviews see \cite{Ellwanger:2009dp,Maniatis:2009re}).
+reviews see \cite{Ellwanger:2009dp, Maniatis:2009re}).
 

### 2014-10-24 doc/flexiblesusy-paper.tex
Added MSSM review and removed NMSSM references other than the reviews in line wioth referee request.

@@
 parameters. Beyond the MSSM there are also two public spectrum
-generators \cite{Ellwanger:2004xm, Ellwanger:2005dv, Ellwanger:2006rn,Ellwanger:2008py, Allanach:2013kza} for the next to
-minimal supersymmetric standard model (NMSSM) \cite{NMSSM} (or for recent
-reviews see \cite{Ellwanger:2009dp, Maniatis:2009re}).
+generators \cite{Ellwanger:2004xm, Ellwanger:2005dv,
+  Ellwanger:2006rn,Ellwanger:2008py, Allanach:2013kza} for the next to
+minimal supersymmetric standard model (NMSSM)  (for
+recent reviews see \cite{Ellwanger:2009dp, Maniatis:2009re}).
 

### 2014-10-24 doc/flexiblesusy-paper.tex
cut references to E6SSM, USSM, and other gauge extensions.

@@
 These developments motivate the exploration of non-minimal SUSY models
-which ameliorate the naturalness problems, as can happen in the \ESSM
-\cite{King:2005jy,King:2005my,King:2007uj,Athron:2010zz,Nevzorov:2012hs,Nevzorov:2013tta,Nevzorov:2013ixa}, USSM
-\cite{Fayet:1977yc,Suematsu:1994qm,Cvetic:1995rj,deCarlos:1997yv,Cvetic:1997ky,Demir:1998dk,Langacker:1998tc,Erler:2002pr,Choi:2006fz,Ham:2007wc,Langacker:2008yv,Ham:2008xf,Kalinowski:2008iq} or from other gauge extensions \cite{Batra:2003nj, Medina:2009ey, Bharucha:2013ela},
+which ameliorate the naturalness problems, as can happen in the USSM
+\cite{Fayet:1977yc,Cvetic:1997ky,Langacker:2008yv},  \ESSM
+\cite{King:2005jy,Athron:2010zz} and similar models \cite{Nevzorov:2012hs}, or from other gauge extensions \cite{Batra:2003nj, Bharucha:2013ela},
 by raising the tree level Higgs mass. At the same time they can also
@@
 \fs already comes with plenty of predefined models: the CMSSM (simply
-called \code{MSSM}), the NMSSM \cite{NMSSM} in it's $Z_3$-symmetric form (called \code{NMSSM}), $Z_3$-violating NMSSM (\code{SMSSM}), the USSM (\code{UMSSM})
-\cite{Fayet:1977yc,Suematsu:1994qm,Cvetic:1995rj,deCarlos:1997yv,Cvetic:1997ky,Demir:1998dk,Langacker:1998tc,Erler:2002pr,Choi:2006fz,Ham:2007wc,Langacker:2008yv,Ham:2008xf,Kalinowski:2008iq},
+called \code{MSSM}), the NMSSM \cite{NMSSM} in it's $Z_3$-symmetric form (called \code{NMSSM}), $Z_3$-violating NMSSM (\code{SMSSM}), the USSM (\code{UMSSM}),
 the NUHM \ESSM (\code{E6SSM}) \cite{Athron:2007en}, the

### 2014-10-24 doc/flexiblesusy-paper.tex
removed a now defunct \cite{NMSSM} that I had missed.

@@
 These developments motivate the exploration of non-minimal SUSY models
-which ameliorate the naturalness problems, as can happen in the USSM
+which ameliorate the naturalness problems by raising the tree level
+Higgs mass, as can happen in the USSM
 \cite{Fayet:1977yc,Cvetic:1997ky,Langacker:2008yv}, \ESSM
-\cite{King:2005jy,Athron:2010zz} and similar models \cite{Nevzorov:2012hs}, or from other gauge extensions \cite{Batra:2003nj, Bharucha:2013ela},
-by raising the tree level Higgs mass. At the same time they can also
+\cite{King:2005jy,Athron:2010zz} and similar models
+\cite{Nevzorov:2012hs}, or from other gauge extensions
+\cite{Batra:2003nj, Bharucha:2013ela}. At the same time they can also
 motivate models that are developed with a fresh perspective, based on
@@
 \fs already comes with plenty of predefined models: the CMSSM (simply
-called \code{MSSM}), the NMSSM \cite{NMSSM} in it's $Z_3$-symmetric form (called \code{NMSSM}), $Z_3$-violating NMSSM (\code{SMSSM}), the USSM (\code{UMSSM}),
+called \code{MSSM}), the NMSSM in it's $Z_3$-symmetric form (called \code{NMSSM}), $Z_3$-violating NMSSM (\code{SMSSM}), the USSM (\code{UMSSM}),
 the NUHM \ESSM (\code{E6SSM}) \cite{Athron:2007en}, the

### 2014-10-25 doc/flexiblesusy-paper.tex
strted making edits for point 2.  Not finished.

--- 2014-10-24 doc/flexiblesusy-paper.tex
@@
 contributions for the MSSM and NMSSM).
+
+The standard input and output of the generated code is given in SLHA
+format and an extension of this to incorporate new blocks for models
+which go beyond the MSSM and NMSSM. We use the well known SLHAea
+\cite{slhaea} to create these SLHA files and the extra blocks may be chosen
+by the user. Since the SLHA standard is intended precisely for the
+purpose, MSSM and NMSSM codes should be compliant with this, and we
+have tested that HDECAY\cite{hdecay} can understand the SLHA output of
+FlexibleSUSY.
+
+For models which go beyond this there is no fixed standard for all
+names.  Nonetheless extending the SLHA framework and allowing user
+control of the new block names should greatly simplify the process of
+linking the output to a users own calculator or any new public tools
+that are developed for that model.
 
@@
 
+\bibitem{slhaea} F.S.~Thomas http://fthomas.github.io/slhaea/
+
+\bibitem{hdecay}  A. Djouadi, M. Spira and P.M. Zerwas, Phys. Lett.  B {\bf 264}
+(1991) 440 and  Z. Phys. C {\bf 70} (1996) 427;  M. Spira {\it et al.}, Nucl.\
+Phys.\ B {\bf 453} (1995) 17;  A.~Djouadi, J.~Kalinowski and M.~Spira, Comput.
+Phys. Commun. {\bf 108} (1998) 56; J.~M.~Butterworth, A.~Arbey, L.~Basso, S.~Belov, A.~Bharucha, F.~Braam, A.~Buckley and M.~Campanelli {\it et al.},
+  %``The Tools and Monte Carlo working group Summary Report,''
+  arXiv:1003.1643 [hep-ph].
+
 %\cite{Degrassi:2009yq}

### 2014-10-25 doc/flexiblesusy-paper.tex

@@
 \fs already comes with plenty of predefined models: the CMSSM (simply
-called \code{MSSM}), the NMSSM in it's $Z_3$-symmetric form (called \code{NMSSM}), $Z_3$-violating NMSSM (\code{SMSSM}), the USSM (\code{UMSSM}),
+called \code{MSSM}), the NMSSM in its $Z_3$-symmetric form (called \code{NMSSM}), $Z_3$-violating NMSSM (\code{SMSSM}), the USSM (\code{UMSSM}),
 the NUHM \ESSM (\code{E6SSM}) \cite{Athron:2007en}, the

### 2014-10-27 doc/flexiblesusy-paper.tex
give expressions for the neglected higher order terms exlicitly
This addresses Reviewer 1 issue A1.

@@
 multiplets, \fs provides two approximative procedures with a shorter
-run-time.  The used procedure can be set in the model file for each
-field.  These approximations work as follows:
+run-time in addition to the iterative procedure.  In the \fs model
+file the two approximative procedures are called \code{LowPrecision}
+and \code{MediumPrecision}.  The iterative procedure is called
+\code{HighPrecision}.  The procedure to be used can be set in the
+model file for each field.  The default setting is
+%
+\begin{lstlisting}
+DefaultPoleMassPrecision = MediumPrecision;
+HighPoleMassPrecision    = {hh, Ah, Hpm};
+MediumPoleMassPrecision  = {};
+LowPoleMassPrecision     = {};
+\end{lstlisting}
+%
+In the variable \code{DefaultPoleMassPrecision} the default pole mass
+calculation precision to be used for all particles is selected.
+Possible values are \code{LowPrecision}, \code{MediumPrecision} and
+\code{HighPrecision}.  The values \code{LowPrecision} and
+\code{MediumPrecision} correspond to the two approximations described
+below and \code{HighPrecision} corresponds to the iterative
+determination of the pole masses.  In the variables
+\code{HighPoleMassPrecision}, \code{MediumPoleMassPrecision} and
+\code{LowPoleMassPrecision} the pole mass calculation precision can be
+changed for individual particles.  The settings in these variables
+overwrite the setting in \code{DefaultPoleMassPrecision} for these
+particles.  In the above example the pole masses of all particles are
+calculated with \code{MediumPrecision}, except for the Higgs boson
+pole masses, which are calculated with the iterative procedure
+(\code{HighPrecision}).
+
+The three different pole mass calculation procedures work as follows:
 %
 \begin{itemize}
-\item \code{LowPoleMassPrecision}: This option provides the
-  lowest precision but is also the fastest one.  Here the one-loop
-  mass matrix $m_{f,1L}^\text{low}$ is calculated exactly once as
+\item \code{LowPrecision}: This option provides the lowest precision
+  but is also the fastest one.  Here the one-loop mass matrix
+  $m_{f,1L}^\text{low}$ is calculated exactly once as
 %
@@
   the eigenvalues are interpreted as pole masses $m_{f_i}^\pole$.
-  This method neglects what are formally of two-loop order.  This
-  method is imprecise if the self-energy corrections to the mass
-  matrix are large or the tree-level mass spectrum of the multiplet is
-  very split.
-
-\item \code{MediumPoleMassPrecision} (default): This option
+  This method neglects terms of the form
+  %
+  \begin{align}
+    \left[(m_{f_k}^\pole)^2 - m_{f_i}m_{f_j}\right]
+    \left.\frac{\partial m_{f,1L}(p^2)}{\partial
+        p^2}\right|_{p^2=m_{f_i}m_{f_j}} ,
+  \end{align}
+  %
+  which are formally of two-loop order.  The method is imprecise if
+  the self-energy corrections to the mass matrix are large or the
+  tree-level mass spectrum of the multiplet is very split.
+
+\item \code{MediumPrecision} (default): This option
   provides calculation with medium precision with a medium execution
@@
   $(m_{f,1L}^\text{medium})^{(k)}$ is diagonalized and the $k$th
-  eigenvalue is interpreted as pole mass $m_{f_k}^\pole$.  This method
-  is imprecise if the self-energy corrections to the mass matrix are
-  large.
-
-\item \code{HighPoleMassPrecision}: This option solves
+  eigenvalue is interpreted as pole mass $m_{f_k}^\pole$.  Thereby for
+  the $k$th eigenvalue two-loop terms of the form
+  %
+  \begin{align}
+    \left[(m_{f_k}^\pole)^2 - m_{f_k}^2\right] \left.\frac{\partial
+        m_{f,1L}(p^2)}{\partial p^2}\right|_{p^2=m_{f_k}^2}
+  \end{align}
+  %
+  are neglected.  This method is imprecise if the self-energy
+  corrections to the mass matrix are large.  Note, that this method is
+  used in Softsusy to calculate the pole masses of the non-Higgs
+  fields.
+
+\item \code{HighPrecision}: This option solves
   Eq.~\eqref{eq:pole-mass-def} exactly by iterating over the momentum
@@
   Here the one-loop mass matrix $m_{f,1L}^\text{high}$ is diagonalized
-  $n$ times, as in the case of \code{MediumPoleMassPrecision},
+  $n$ times, as in the case of \code{MediumPrecision},
   resulting in $n$ pole masses $m_{f_k}^\pole$ ($k = 1,\ldots,n$).

### 2014-10-27 doc/flexiblesusy-paper.tex
add table with pole mass and time comparison
round times
describe pole mass calculation comparison table

@@
 %
+A numerical comparison of the three different methods for a specific
+CMSSM parameter point can be found in
+\tabref{tab:pole-mass-precision-comparison}.  One finds that (i) the
+calculated lightest CP-even Higgs pole mass is sensitive to the chosen
+pole mass calculation method due to the large loop corrections.
+Especially, if two-loop contributions to the Higgs tadpoles and
+self-energies are added, the two approximations \code{LowPrecision}
+and \code{MediumPrecision} must not be used, because they neglect
+terms of two-loop order.  (ii) Since the gluino does not mix with
+other particles, there is no difference between \code{LowPrecision}
+and \code{MediumPrecision}.  Not neglecting the two-loop terms by
+using \code{HighPrecision} changes the gluino mass about $1\unit{\%}$.
+(iii) Since the up-type squarks form a $6$-plet, the calculation of
+their pole masses is dramatically increased from $2.57\unit{ms}$ to
+$60.9\unit{ms}$ between \code{LowPrecision} and \code{HighPrecision}.
+However, since the change in the lightest up-type squark mass between
+the three different methods is only of the order $\unit{0.4\%}$, one
+can consider calculating then with \code{MediumPrecision} only.
+%
+\begin{table}[tbh]
+  \centering
+  \begin{tabular}{llll}
+    \toprule
+    & \code{LowPrecision}
+    & \code{MediumPrecision}
+    & \code{HighPrecision}\\
+    \midrule
+    $m_h$
+    & $1441\unit{GeV}$ ($0.96\unit{ms}$)
+    & $128\unit{GeV}$ ($2.53\unit{ms}$)
+    & $126\unit{GeV}$ ($9.79\unit{ms}$) \\
+    $m_{\tilde{g}}$
+    & $10167\unit{GeV}$ ($0.10\unit{ms}$)
+    & $10167\unit{GeV}$ ($0.10\unit{ms}$)
+    & $10251\unit{GeV}$ ($0.41\unit{ms}$) \\
+    $m_{\tilde{u}_1}$
+    & $8702\unit{GeV}$ ($2.57\unit{ms}$)
+    & $8714\unit{GeV}$ ($15.7\unit{ms}$)
+    & $8734\unit{GeV}$ ($60.9\unit{ms}$) \\
+    \bottomrule
+  \end{tabular}
+  \caption{Comparison of pole masses with different calculation
+    methods for the CMSSM with $m_0=8\unit{TeV}$, $M_{1/2}=A_0=5\unit{TeV}$,
+    $\tan\beta=20$, $\sign\mu=+1$.  In brackets the time used to calculate
+    the pole masses of the whole multiplet is given.}
+  \label{tab:pole-mass-precision-comparison}
+\end{table}
+%
+
 If higher accuracy is required, two-loop corrections to the

### 2014-10-27 doc/flexiblesusy-paper.tex
explain finding of analytic solutions to the EWSB equations
This addresses Reviewer 1, comment B3.

@@
 VEV corresponding to the Higgs fields $H_i$ and $t_i$ is the one-loop
-tadpole diagram of $H_i$.  The electroweak symmetry breaking
-conditions \eqref{eq:one-loop-ewsb-eq} are solved simultaneously using
-the iterative multi-dimensional root finder algorithm
-\code{gsl_multiroot_fsolver_hybrid} from the GNU Scientific Library
-(GSL).  If no root can be found, the
+tadpole diagram of $H_i$.  Already at the tree-level ($t_i = 0$) Eqs.\
+\eqref{eq:one-loop-ewsb-eq} can have multiple solutions, depending on
+which paramters are chosen to be fixed by theses equations.  For
+example in the real MSSM Eqs.\ \eqref{eq:one-loop-ewsb-eq} can only
+determine $|\mu|$, while the sign of $\mu$ is not fixed.  Currently
+\fs handles only real model parameters and renormalizable theories,
+which restricts the number of possible multiple solutions of Eqs.\
+\eqref{eq:one-loop-ewsb-eq} to be less or equal to four times the
+number of equations.  In some cases, as for example in the real MSSM,
+% where $\mu$ is chosen to be fixed by the EWSB equations
+these different solutions might be related by one or more free global
+signs in the form
+%
+\begin{align}
+  \mu = \sign\mu \cdot \sqrt{\ldots} .
+\end{align}
+%
+\fs recognizes such cases and introduces these global signs as
+additional free parameters.  The case of the real MSSM, where only
+$|\mu|$ is fixed by the EWSB equations, is therefore handled
+automatically.  If, however, the different solutions are not related
+by global signs, then \fs writes all solutions to the file
+\code{models/<model>/<model>_tree_level_EWSB_solution.m}.  The user
+has then the option to pick a particular solution by setting it in the
+\code{TreeLevelEWSBSolution} variable in the model file.  This is for
+example done in the $Z_3$-violating NMSSM (\code{SMSSM}), where $\mu$
+is chosen to be fixed by the EWSB equations: There the tree-level
+solution for $\mu$ has the form
+%
+\begin{align}
+  \mu = -\frac{v_s \lambda}{\sqrt{2}} + \sign X \cdot \sqrt{\ldots}
+\end{align}
+%
+with a non-global $\sign X$.  In this model the user can parametrize
+the tree-level solution in terms of $\sign X$ in the
+\code{TreeLevelEWSBSolution} variable and add $\sign X$ by hand to the
+list of free input parameters, see
+\code{model_files/SMSSM/FlexibleSUSY.m.in} for an example.  If the
+user decides to not pick a particular solution by leaving the variable
+\code{TreeLevelEWSBSolution} empty, \fs tries to find a solution to
+the tree-level EWSB equations numerically via an iteration.  In this
+case, however, the user does not have the option to chose between the
+different solutions.
+
+%
+If a solution of the tree-level EWSB equations has been found, the
+one-loop equations \eqref{eq:one-loop-ewsb-eq} are solved
+simultaneously using the iterative multi-dimensional root finder
+algorithm \code{gsl_multiroot_fsolver_hybrid} from the GNU Scientific
+Library (GSL).  If no root can be found, the
 \code{gsl_multiroot_fsolver_hybrids} algorithm is tried as

### 2014-10-29 doc/flexiblesusy-paper.tex
fix typo theses --> these.
added more details on testing with SUSY-HITS.

@@
 
-The standard input and output of the generated code is given in SLHA
-format and an extension of this to incorporate new blocks for models
-which go beyond the MSSM and NMSSM. We use the well known SLHAea
-\cite{slhaea} to create these SLHA files and the extra blocks may be chosen
-by the user. Since the SLHA standard is intended precisely for the
-purpose, MSSM and NMSSM codes should be compliant with this, and we
-have tested that HDECAY\cite{hdecay} can understand the SLHA output of
-FlexibleSUSY.
+The standard input and output of the generated code is given in
+SLHA-style format where the user has control of the Block names by
+editing the SARAH files and may also add extra blocks in the
+FlexibleSUSY.m model file which is decribed in section 5. We use the
+well known SLHAea \cite{slhaea} to create these SLHA files and with a
+simple modifications these can be passed directly to another code
+which uses them as an alternative to writing out the SLHA file. Since the
+SLHA standard is intended precisely for the purpose, MSSM and NMSSM
+codes should be compliant with this.  We have tested that the versions
+of HDECAY\cite{hdecay} and SDECAY\cite{sdecay} inside SUSY-HIT
+\cite{susyhit} can understand the SLHA output of FlexibleSUSY and give
+matching output to that from SOFTSUSY.
 
@@
 \eqref{eq:one-loop-ewsb-eq} can have multiple solutions, depending on
-which paramters are chosen to be fixed by theses equations.  For
+which paramters are chosen to be fixed by these equations.  For
 example in the real MSSM Eqs.\ \eqref{eq:one-loop-ewsb-eq} can only
@@
   arXiv:1003.1643 [hep-ph].
+%\cite{Muhlleitner:2003vg}
+\bibitem{sdecay} 
+  M.~Muhlleitner, A.~Djouadi and Y.~Mambrini,
+  %``SDECAY: A Fortran code for the decays of the supersymmetric particles in the MSSM,''
+  Comput.\ Phys.\ Commun.\  {\bf 168}, 46 (2005)
+  [hep-ph/0311167].
+  %%CITATION = HEP-PH/0311167;%%
+  %235 citations counted in INSPIRE as of 29 Oct 2014
+
+%\cite{Djouadi:2006bz}
+\bibitem{susyhit} 
+  A.~Djouadi, M.~M.~Muhlleitner and M.~Spira,
+  %``Decays of supersymmetric particles: The Program SUSY-HIT (SUspect-SdecaY-Hdecay-InTerface),''
+  Acta Phys.\ Polon.\ B {\bf 38}, 635 (2007)
+  [hep-ph/0609292].
+  %%CITATION = HEP-PH/0609292;%%
+  %238 citations counted in INSPIRE as of 29 Oct 2014
 

### 2014-10-30 doc/flexiblesusy-paper.tex
simplify statement about slhaea.

@@
 editing the SARAH files and may also add extra blocks in the
-FlexibleSUSY.m model file which is decribed in section 5. We use the
-well known SLHAea \cite{slhaea} to create these SLHA files and with a
-simple modifications these can be passed directly to another code
-which uses them as an alternative to writing out the SLHA file. Since the
-SLHA standard is intended precisely for the purpose, MSSM and NMSSM
-codes should be compliant with this.  We have tested that the versions
-of HDECAY\cite{hdecay} and SDECAY\cite{sdecay} inside SUSY-HIT
+FlexibleSUSY.m model file which is decribed in section 5. We use
+SLHAea \cite{slhaea} to create these SLHA files and with a simple
+modifications these can be passed directly to another code which uses
+them as an alternative to writing out the SLHA file. Since the SLHA
+standard is intended precisely for the purpose, MSSM and NMSSM codes
+should be compliant with this.  We have tested that the versions of
+HDECAY\cite{hdecay} and SDECAY\cite{sdecay} inside SUSY-HIT
 \cite{susyhit} can understand the SLHA output of FlexibleSUSY and give

### 2014-11-03 doc/flexiblesusy-paper.tex
swap sections 8.1 (speed test) and 8.2 (numeric tests)
explain difference to Softsusy (w/o flavour violation)
The differences are:

 * calculation of the weak mixing angle (FS uses MW and MZ, while
   Softsusy uses the muon decay)

 * some two-loop threshold corrections in Softsusy, missing in FS

 * some three-loop beta-functions in Softsusy

@@
 
+\subsection{Numeric tests}
+
+To check the correctness of \fs's generated spectrum generators
+extensive unit testing against Softsusy's MSSM and NMSSM
+implementations (both $Z_3$-invariant and $Z_3$-violating variants)
+was carried out.  These unit tests systematically compare all
+tree-level mass matrices, EWSB equations, one- and two-loop
+$\beta$-functions, one- and two-loop self-energies and one- and
+two-loop tadpoles numerically for certain MSSM and NMSSM parameter
+points.  All tested expressions were found to agree within double
+machine precision.\footnote{Due to the systematic and detailed tests
+  several bugs in Softsusy, SARAH and \fs could be identified and
+  corrected.}  Furthermore, the overall pole mass spectrum after the
+full fixed-point iteration has finished is compared, and was found to
+agree at the sub-permille level.  The origin of the sub-permille level
+difference between \fs and Softsusy is the different determination of
+the \DRbar\ weak mixing angle
+$\theta_{W,\text{susy}}^{\text{\DRbar}}(M_Z)$: \fs calculates
+$\theta_W$ from $M_W^2$ and $M_Z^2$ as described in
+\secref{sec:calculation-of-gauge-couplings}, while Softsusy determines
+$\theta_W$ from the muon decay constant $G_\mu$, which is more
+precise.  Furthermore, in Softsusy $3.5.0$ some three-loop
+$\beta$-functions and two-loop threshold corrections for the gauge and
+Yukawa couplings can be enabled in the MSSM to increase the accuracy
+of the RG running and the \DRbar\ gauge and Yukawa couplings at $M_Z$
+\cite{}.  These corrections are not implemented in \fs so far.
+
+The complete set of unit tests is shipped with \fs and can be found in
+the \code{test/} directory.  The tests can be run with the command
+\code{make execute-tests}.  All unit tests are carried out nightly in
+order to continuously check the correctness of the meta code and the
+generated spectrum generators.  The test results can be found at
+\url{http://iktp.tu-dresden.de/~avoigt/FlexibleSUSY/FlexibleSUSY/test.html}
+
+The \fs generated NUHM \ESSM\ spectrum generator has also been
+compared against a handwritten one for a constrained version of the
+\ESSM \cite{Athron:2009ue, Athron:2009bs, Athron:2012pw}.  The
+$\beta$-functions were systematically compared in unit tests and were
+found to match at the ??? level.  The handwritten code doesn't include
+full one-loop self-energies or tadpoles, so tests on these were not
+carried out.  Although the generators assume different constraints
+they could be compared by using the output of the CE$_6$SSM generator
+as an input to \fs and the spectra were found to be in reasonable
+agreement, given the different levels of precision.
+
+In addition \fs has already undergone some user testing. This includes
+analytic tests of the $R$-symmetric low-energy model (MRSSM) and
+alternative $E_6$-inspired SUSY scenarios.  The users who have helped us
+with this are thanked in the acknowledgements.
+
+We also compared the run-time of \fs against SPheno, Softsusy and the
+SARAH generated MSSM spectrum generator SPhenoMSSM.  The test results
+can be found in \secref{sec:run-time comparison}.
+
 \subsection{Run-time comparison}
@@
 separate thread, and therefore benefits from multi-core CPUs.
-
-\subsection{Numeric tests}
-
-To check the correctness of \fs's generated spectrum generators
-extensive unit testing against Softsusy's MSSM and NMSSM
-implementations (both $Z_3$-invariant and $Z_3$-violating variants)
-was carried out. We checked mass matrices, EWSB equations, one- and two-loop
-$\beta$-functions, one- and two-loop self-energies and one- and
-two-loop tadpoles and found all to agree within double machine
-precision.  We also compared the overall pole mass spectrum after the
-full fixed-point iteration has finished, and found the spectra to
-agree at the sub-permille level.
-
-\fs generated NUHM \ESSM has also been compared against a handwritten
-spectrum generator for a constrained version of the \ESSM
-\cite{Athron:2009ue, Athron:2009bs, Athron:2012pw}.  The RGEs could be
-compared directly in unit tests and were found to match.  The
-handwritten code doesn't include full one-loop self-energies or
-tadpoles so tests on these were not carried out.  Although the
-generators assume different constraints they could be compared by
-using the output of the CE$_6$SSM generator as an input to \fs and the
-spectra were found to be in reasonable agreement, given the different
-levels of precision.
-
-In addition \fs has already undergone some user testing. This includes
-analytic tests of the $R$-symmetric low-energy model (MRSSM) and
-alternative $E_6$-inspired SUSY scenarios.  The users who have helped us
-with this are thanked in the acknowledgements.
-
-We also compared the run-time of \fs against SPheno, Softsusy and the
-SARAH generated MSSM spectrum generator SPhenoMSSM.  The test results
-can be found in \secref{sec:run-time comparison}.
 

### 2014-11-03 doc/flexiblesusy-paper.tex
explicitly state the kind of flavour violation used in the speed test

@@
 \begin{itemize}
-\item \emph{Without flavour violation:} Disallowing flavour violation
-  simplifies the calculation of the pole masses, because
-  flavour-off-diagonal sfermion self-energy matrix elements don't need
-  to be calculated.  Here we compare \fs's non-flavour violating CMSSM
-  spectrum generator FlexibleSUSY-NoFV (version 1.0.0) against SPheno
-  (version 3.2.4) and Softsusy (version 3.4.0).
-%
-\item \emph{With flavour violation:} Allowing for flavour violation in
-  general increases the run-time of spectrum generators, because the
-  full $6\times 6$ sfermion self-energy matrices have to be
-  calculated.  Here we compare FlexibleSUSY-FV (version 1.0.0) and
-  SPhenoMSSM (generated with SARAH 4.1.0 and linked against SPheno
-  3.2.4).  Both spectrum generators are based on SARAH's MSSM model
-  file, which allows for flavour violation.
+\item \emph{Without sfermion flavour violation:} Disallowing sfermion
+  flavour violation simplifies the calculation of the pole masses,
+  because flavour-off-diagonal sfermion self-energy matrix elements
+  don't need to be calculated.  Here we compare \fs's non-flavour
+  violating CMSSM spectrum generator FlexibleSUSY-NoFV (version 1.0.0)
+  against SPheno (version 3.2.4) and Softsusy (version 3.4.0).
+%
+\item \emph{With sfermion flavour violation:} Allowing for sfermion
+  flavour violation in general increases the run-time of spectrum
+  generators, because the full $6\times 6$ sfermion self-energy
+  matrices have to be calculated.  Here we compare FlexibleSUSY-FV
+  (version 1.0.0) and SPhenoMSSM (generated with SARAH 4.1.0 and
+  linked against SPheno 3.2.4).  Both spectrum generators are based on
+  SARAH's MSSM model file, which allows for sfermion flavour
+  violation.
 \end{itemize}
@@
     for random parameter points.  The first three rows show
-    spectrum generators which disallow flavour violation.  Rows
-    $4$--$5$ contain flavour violating spectrum generators, based
+    spectrum generators which disallow sfermion flavour violation.  Rows
+    $4$--$5$ contain sfermion flavour violating spectrum generators, based
     on SARAH's MSSM model file.}

### 2014-11-03 doc/flexiblesusy-paper.tex
correcting nightly test results url
add refeence for Softsusy 3.5.0

@@
 Yukawa couplings can be enabled in the MSSM to increase the accuracy
-of the RG running and the \DRbar\ gauge and Yukawa couplings at $M_Z$
-\cite{}.  These corrections are not implemented in \fs so far.
+of the RG running and the determination of the \DRbar\ gauge and
+Yukawa couplings at $M_Z$ \cite{Allanach:2014nba}.  These corrections
+are not implemented in \fs so far.
 
@@
 generated spectrum generators.  The test results can be found at
-\url{http://iktp.tu-dresden.de/~avoigt/FlexibleSUSY/FlexibleSUSY/test.html}
+\url{http://iktp.tu-dresden.de/~avoigt/FlexibleSUSY/test.xhtml}
 
@@
 
+\bibitem{Allanach:2014nba}
+  B.~C.~Allanach, A.~Bednyakov and R.~Ruiz de Austri,
+  %``Higher Order Corrections and Unification in the Minimal Supersymmetric Standard Model: SOFTSUSY3.5.0,''
+  arXiv:1407.6130 [hep-ph].
+  %%CITATION = ARXIV:1407.6130;%%
+  %2 citations counted in INSPIRE as of 03 Nov 2014
+
 %\cite{Athron:2009ue}

### 2014-11-04 doc/flexiblesusy-paper.tex
describe used SG settings for the speed-test
The CKM matrix is set to unity, CP-violating phases are set to zero,
R-parity violation is disabled.

@@
 implementations (both $Z_3$-invariant and $Z_3$-violating variants)
-was carried out.  These unit tests systematically compare all
+has been carried out.  These unit tests systematically compare all
 tree-level mass matrices, EWSB equations, one- and two-loop
 $\beta$-functions, one- and two-loop self-energies and one- and
-two-loop tadpoles numerically for certain MSSM and NMSSM parameter
+two-loop tadpoles numerically for various MSSM and NMSSM parameter
 points.  All tested expressions were found to agree within double
@@
   corrected.}  Furthermore, the overall pole mass spectrum after the
-full fixed-point iteration has finished is compared, and was found to
-agree at the sub-permille level.  The origin of the sub-permille level
-difference between \fs and Softsusy is the different determination of
-the \DRbar\ weak mixing angle
-$\theta_{W,\text{susy}}^{\text{\DRbar}}(M_Z)$: \fs calculates
-$\theta_W$ from $M_W^2$ and $M_Z^2$ as described in
-\secref{sec:calculation-of-gauge-couplings}, while Softsusy determines
-$\theta_W$ from the muon decay constant $G_\mu$, which is more
-precise.  Furthermore, in Softsusy $3.5.0$ some three-loop
-$\beta$-functions and two-loop threshold corrections for the gauge and
-Yukawa couplings can be enabled in the MSSM to increase the accuracy
-of the RG running and the determination of the \DRbar\ gauge and
-Yukawa couplings at $M_Z$ \cite{Allanach:2014nba}.  These corrections
-are not implemented in \fs so far.
-
-The complete set of unit tests is shipped with \fs and can be found in
-the \code{test/} directory.  The tests can be run with the command
-\code{make execute-tests}.  All unit tests are carried out nightly in
-order to continuously check the correctness of the meta code and the
-generated spectrum generators.  The test results can be found at
+full fixed-point iteration has finished has been compared, and was
+found to agree at the sub-permille level.  The origin of the
+sub-permille level difference between \fs and Softsusy is the
+different determination of the weak mixing angle
+$\theta_{W,\text{susy}}^{\text{\DRbar}}(M_Z)$ in the SUSY model in the
+\DRbar\ scheme: \fs calculates
+$\theta_{W,\text{susy}}^{\text{\DRbar}}(M_Z)$ from $M_W$ and $M_Z$, as
+described in \secref{sec:calculation-of-gauge-couplings}, while
+Softsusy determines $\theta_{W,\text{susy}}^{\text{\DRbar}}(M_Z)$ from
+the muon decay constant $G_\mu$.  The approach used in Softsusy
+results in more precise \DRbar\ gauge couplings at the $M_Z$ scale,
+because the muon decay constant is known with a higher accuracy than
+the $W$-boson mass.  Furthermore, in Softsusy $3.5.0$ some three-loop
+$\beta$-functions and two-loop threshold corrections can be enabled in
+the MSSM to increase the accuracy of the RG running and the
+determination of the \DRbar\ gauge and Yukawa couplings at $M_Z$
+\cite{Allanach:2014nba}.  These corrections are not implemented in \fs
+so far.  The complete set of unit tests is shipped with \fs and can be
+found in the \code{test/} directory.  The tests can be run with the
+command \code{make execute-tests}.  All unit tests are carried out
+nightly in order to continuously check the correctness of the meta
+code and the generated spectrum generators for the shipped models.
+The nightly test results can be found at
 \url{http://iktp.tu-dresden.de/~avoigt/FlexibleSUSY/test.xhtml}
@@
 We also compared the run-time of \fs against SPheno, Softsusy and the
-SARAH generated MSSM spectrum generator SPhenoMSSM.  The test results
-can be found in \secref{sec:run-time comparison}.
+SARAH generated MSSM spectrum generator SPhenoMSSM.  The results of
+the comparison can be found in \secref{sec:run-time comparison}.
 
@@
 %
-\fs and Softsusy are compiled with g++ 4.8.0 and Intel ifort
-13.1.3 20130607.  SPheno and SPhenoMSSM are compiled with Intel ifort
-13.1.3 20130607.\footnote{Intel's ifort compiler decreases the
-  run-time of SPheno and SPhenoMSSM by approximately a factor $1.5$,
-  compared to gfortran.}
-
-For the run-time comparison we're generating $2\cdot 10^{4}$ random
-CMSSM parameter points with $m_0\in [50,1000]\unit{GeV}$, $m_{1/2}\in
+For the run-time comparison the CKM matrix is set to unity, all CP
+violating phases are set to zero and R-parity violation is disabled.
+\fs and Softsusy are compiled with g++ 4.8.0 and Intel ifort 13.1.3
+20130607.  SPheno and SPhenoMSSM are compiled with Intel ifort 13.1.3
+20130607.\footnote{Intel's ifort compiler decreases the run-time of
+  SPheno and SPhenoMSSM by approximately a factor $1.5$, compared to
+  gfortran.}  We're generating $2\cdot 10^{4}$ random CMSSM parameter
+points with $m_0\in [50,1000]\unit{GeV}$, $m_{1/2}\in
 [50,1000]\unit{GeV}$, $\tan\beta\in [1,100]$, $\sign\mu\in \{-1,+1\}$
@@
 $\tan\beta$, $\sign\mu$, $A_0$ in form of a \code{MINPAR} block to the
-SLHA template file given in
-\ref{sec:speed-test-slha-template-file}.  The resulting SLHA input
-file is passed to each spectrum generator and the (wall-clock) time is
-measured until the program has finished.  The average run-times for
-three different CPU types can be found in
+SLHA template file given in \ref{sec:speed-test-slha-template-file}.
+The resulting SLHA input file is passed to each spectrum generator and
+the (wall-clock) time is measured until the program has finished.  The
+average run-times for three different CPU types can be found in
 \tabref{tab:run-time-comparison}.  The first column shows the run-time

### 2014-11-05 doc/flexiblesusy-paper.tex
note that CP-violating phases are set to zero
because only real model parameters are currently allowed.
reformulate caption in Table 2
This addresses referee 2, comment 2

@@
 the \code{SMINPUTS} block of the SLHA input file \cite{Skands:2003cj}.
-The CKM mixing matrix is currently assumed to be diagonal.  The third
-generation quark masses are calculated in the \DRbar scheme from the
-SLHA user input quantities $m_t^\text{pole}$,
-$m_{b,\text{SM}}^{\text{\MSbar}}(M_Z)$ and
+The CKM mixing matrix is currently set to unity and CP-violating
+phases are set to zero.  The third generation quark masses are
+calculated in the \DRbar scheme from the SLHA user input quantities
+$m_t^\text{pole}$, $m_{b,\text{SM}}^{\text{\MSbar}}(M_Z)$ and
 $m_{\tau,\text{SM}}^{\text{\MSbar}}(M_Z)$ \cite{Skands:2003cj}.  In
@@
   \caption{Average run-time of CMSSM spectrum generators
-    for random parameter points.  The first three rows show
-    spectrum generators which disallow sfermion flavour violation.  Rows
-    $4$--$5$ contain sfermion flavour violating spectrum generators, based
+    for random parameter points.  The first three rows show three
+    spectrum generators with disabled sfermion flavour violation.  Rows
+    $4$--$5$ show two spectrum generators with enabled sfermion flavour violation, both based
     on SARAH's MSSM model file.}

### 2014-11-05 doc/flexiblesusy-paper.tex
added new subsection on current limitations to deal with referee 2's point 3 and also further help with referee 1's poiunt A4 at the same time.

@@
 
-\subsection*{Design goals}
+\subsection{Design goals}
 
@@
 scale.
+
+\subsection{Current limitations and future extensions}
+Although we try to handle as many models as possible there are still
+some limitations to what can be done. In it's current form \fs assumes
+all couplings are are real, therefore it is limited to CP conserving
+versions of SUSY models.   Although it is implicit in the
+title we would like to stress that currently we cannot provide a
+spectrum generator for non-SUSY models.  This also means that like
+other spectrum generators it is difficult to reliably predict the mass
+spectrum in extreme cases such as Split-SUSY
+\cite{Wells:2003tf,ArkaniHamed:2004fb,Giudice:2004tc} where the mass
+scales of the SUSY model are drastically split leading to very large
+logarithms which are not resummed.  In such cases the tower of
+effective theories offers the best possibility for a solution.
+However for this to work first \fs must be extended to include
+non-SUSY models.  Finally the gauge group of the model should be
+semi-simple and contain the standard model gauge group as factor so
+that we can match to the SM at low energies.
+
+Future releases which extend \fs beyond each of these limitations are
+already planned.
 
@@
   \caption{Average run-time of CMSSM spectrum generators
-    for random parameter points.  The first three rows show three
-    spectrum generators with disabled sfermion flavour violation.  Rows
-    $4$--$5$ show two spectrum generators with enabled sfermion flavour violation, both based
+    for random parameter points.  The first three rows show
+    spectrum generators which disallow sfermion flavour violation.  Rows
+    $4$--$5$ contain sfermion flavour violating spectrum generators, based
     on SARAH's MSSM model file.}
@@
 
+%\cite{Wells:2003tf}
+\bibitem{Wells:2003tf} 
+  J.~D.~Wells,
+  %``Implications of supersymmetry breaking with a little hierarchy between gauginos and scalars,''
+  hep-ph/0306127.
+  %%CITATION = HEP-PH/0306127;%%
+  %81 citations counted in INSPIRE as of 05 Nov 2014
+
+%\cite{ArkaniHamed:2004fb}
+\bibitem{ArkaniHamed:2004fb} 
+  N.~Arkani-Hamed and S.~Dimopoulos,
+  %``Supersymmetric unification without low energy supersymmetry and signatures for fine-tuning at the LHC,''
+  JHEP {\bf 0506}, 073 (2005)
+  [hep-th/0405159].
+  %%CITATION = HEP-TH/0405159;%%
+  %772 citations counted in INSPIRE as of 05 Nov 2014
+%\cite{Giudice:2004tc}
+\bibitem{Giudice:2004tc}
+  G.~F.~Giudice and A.~Romanino,
+  %``Split supersymmetry,''
+  Nucl.\ Phys.\ B {\bf 699} (2004) 65
+   [Erratum-ibid.\ B {\bf 706} (2005) 65]
+  [hep-ph/0406088].
+  %%CITATION = HEP-PH/0406088;%%
+  %576 citations counted in INSPIRE as of 05 Nov 2014
+
 %\cite{Degrassi:2009yq}

### 2014-11-06 doc/flexiblesusy-paper.tex
adding sdown, selectron, neutralino to pole mass table
align columns to the right
discuss Table 1 in more detail

@@
     + m_{\chi} \Big( \Sigma_\chi^L(p^2) + \Sigma_\chi^{R,T}(p^2) \Big)
-  \Big], \\
+  \Big], \label{eq:majorana-loop-corrections} \\
   &\text{Dirac fermions } \psi: &
@@
   the self-energy corrections to the mass matrix are large or the
-  tree-level mass spectrum of the multiplet is very split.
+  tree-level mass spectrum of the multiplet is very split.  Note: We
+  strongly discourage the use of this method for the determination the
+  Higgs pole masses, as the result will very imprecise due to the
+  large loop corrections.  \fs will print a warning if this method is
+  used for any Higgs boson.
 
@@
 CMSSM parameter point can be found in
-\tabref{tab:pole-mass-precision-comparison}.  One finds that (i) the
-calculated lightest CP-even Higgs pole mass is sensitive to the chosen
-pole mass calculation method due to the large loop corrections.
-Especially, if two-loop contributions to the Higgs tadpoles and
-self-energies are added, the two approximations \code{LowPrecision}
-and \code{MediumPrecision} must not be used, because they neglect
-terms of two-loop order.  (ii) Since the gluino does not mix with
-other particles, there is no difference between \code{LowPrecision}
-and \code{MediumPrecision}.  Not neglecting the two-loop terms by
-using \code{HighPrecision} changes the gluino mass about $1\unit{\%}$.
-(iii) Since the up-type squarks form a $6$-plet, the calculation of
-their pole masses is dramatically increased from $2.57\unit{ms}$ to
-$60.9\unit{ms}$ between \code{LowPrecision} and \code{HighPrecision}.
-However, since the change in the lightest up-type squark mass between
-the three different methods is only of the order $\unit{0.4\%}$, one
-can consider calculating then with \code{MediumPrecision} only.
+\tabref{tab:pole-mass-precision-comparison}.  One finds that
+%
+(i) the calculated lightest CP-even Higgs pole mass, $m_h$, differs
+about $2\unit{GeV}$ between \code{MediumPrecision} and
+\code{HighPrecision}, due to large loop corrections.  Since the
+experimental Higgs mass uncertainty is currently around
+$0.4\unit{GeV}$ \cite{Beringer:1900zz}, we strongly recommend the use
+of \code{HighPrecision} to calculate the Higgs pole mass (this is the
+default).  Especially, if two-loop contributions to the Higgs tadpoles
+and self-energies are added \code{MediumPrecision} must not be used,
+because it neglects terms of two-loop order.  The Higgs boson mass for
+the \code{LowPrecision} method is not given in the table, as it will
+lead to an imprecise result and is therefore strongly discouraged to
+be used, see \secref{sec:PoleMasses}.
+%
+(ii) The gluino pole mass, $m_{\tilde{g}}$, is given in the second
+row.  Since the gluino does not mix with other particles, there is no
+difference between \code{LowPrecision} and \code{MediumPrecision}.
+Not neglecting the two-loop terms by using \code{HighPrecision}
+increases the gluino mass about $1\unit{\%}$ for this parameter point.
+%
+(iii) The pole mass of the lightest neutralino,
+$m_{\tilde{\chi}_1^0}$, is given in the third row.  Since the
+momentum-dependent loop-corrections to the lightest neutralino mass
+are small for this parameter point, it's pole mass varies only in the
+sub-GeV range between the three methods.  However, the run-time of the
+\code{LowPrecision} method is more than a factor $10$ smaller than of
+the \code{HighPrecision}, due to the complicated structure of the loop
+corrections in Eq.~\eqref{eq:majorana-loop-corrections}.
+%
+(iv) The pole masses of the lightest sfermions are given in rows
+$4$--$6$.  Since these particles are contained in $6$-plets, the
+run-time for the calculation of their pole masses is dramatically
+increased by around a factor $20$ between \code{LowPrecision} and
+\code{HighPrecision}.  However, since the change in the lightest
+sfermion masses between the three different methods is less than
+$\unit{0.4\%}$, one can consider calculating then with
+\code{MediumPrecision} only.
 %
@@
   \centering
-  \begin{tabular}{llll}
+  \begin{tabular}{lrrr}
     \toprule
@@
     $m_h$
-    & $1441\unit{GeV}$ ($0.96\unit{ms}$)
-    & $128\unit{GeV}$ ($2.53\unit{ms}$)
-    & $126\unit{GeV}$ ($9.79\unit{ms}$) \\
+    & --
+    & $128.3\unit{GeV}$ ($2.45\unit{ms}$)
+    & $126.3\unit{GeV}$ ($9.53\unit{ms}$) \\
     $m_{\tilde{g}}$
@@
     & $10167\unit{GeV}$ ($0.10\unit{ms}$)
-    & $10251\unit{GeV}$ ($0.41\unit{ms}$) \\
+    & $10251\unit{GeV}$ ($0.38\unit{ms}$) \\
+    $m_{\tilde{\chi}_1^0}$
+    & $2295\unit{GeV}$ ($4.17\unit{ms}$)
+    & $2295\unit{GeV}$ ($16.1\unit{ms}$)
+    & $2295\unit{GeV}$ ($64.2\unit{ms}$) \\
     $m_{\tilde{u}_1}$
-    & $8702\unit{GeV}$ ($2.57\unit{ms}$)
-    & $8714\unit{GeV}$ ($15.7\unit{ms}$)
-    & $8734\unit{GeV}$ ($60.9\unit{ms}$) \\
+    & $8703\unit{GeV}$ ($2.45\unit{ms}$)
+    & $8714\unit{GeV}$ ($14.9\unit{ms}$)
+    & $8734\unit{GeV}$ ($58.5\unit{ms}$) \\
+    $m_{\tilde{d}_1}$
+    & $10442\unit{GeV}$ ($2.48\unit{ms}$)
+    & $10391\unit{GeV}$ ($15.0\unit{ms}$)
+    & $10418\unit{GeV}$ ($59.9\unit{ms}$) \\
+    $m_{\tilde{e}_1}$
+    & $7913\unit{GeV}$ ($1.94\unit{ms}$)
+    & $7909\unit{GeV}$ ($11.2\unit{ms}$)
+    & $7909\unit{GeV}$ ($34.1\unit{ms}$) \\
     \bottomrule

### 2014-11-06 doc/flexiblesusy-paper.tex
explain the manual setting of EWSB solutions in more detail

@@
 \begin{align}
-  0 = \frac{\partial V^\text{tree}}{\partial v_i} - t_i,
+  0 &= \frac{\partial V^\text{tree}}{\partial v_i} - t_i,
+  & & (i=1,\ldots,N)
   \label{eq:one-loop-ewsb-eq}
@@
 \eqref{eq:one-loop-ewsb-eq} can have multiple solutions, depending on
-which paramters are chosen to be fixed by these equations.  For
-example in the real MSSM Eqs.\ \eqref{eq:one-loop-ewsb-eq} can only
-determine $|\mu|$, while the sign of $\mu$ is not fixed.  Currently
-\fs handles only real model parameters and renormalizable theories,
-which restricts the number of possible multiple solutions of Eqs.\
-\eqref{eq:one-loop-ewsb-eq} to be less or equal to four times the
-number of equations.  In some cases, as for example in the real MSSM,
-% where $\mu$ is chosen to be fixed by the EWSB equations
-these different solutions might be related by one or more free global
-signs in the form
+which paramters are chosen to be fixed by these equations.  A
+well-known example is the real MSSM, where the $\mu$-parameter is
+chosen to be fixed by the EWSB equations.  In this model, the EWSB
+equations, Eqs.\ \eqref{eq:one-loop-ewsb-eq} can only determine
+$|\mu|$, while the sign of $\mu$ is not fixed.  This results in two
+solutions for $\mu$ of the form
 %
 \begin{align}
-  \mu = \sign\mu \cdot \sqrt{\ldots} .
+  \mu = \sign\mu \cdot |\mu| .
+  \label{eq:EWSB-solution-for-mu}
 \end{align}
 %
-\fs recognizes such cases and introduces these global signs as
-additional free parameters.  The case of the real MSSM, where only
-$|\mu|$ is fixed by the EWSB equations, is therefore handled
-automatically.  If, however, the different solutions are not related
-by global signs, then \fs writes all solutions to the file
+Currently, \fs handles only real model parameters and renormalizable
+theories, which restricts the number of possible multiple solutions of
+Eqs.\ \eqref{eq:one-loop-ewsb-eq} to be less or equal to $4N$.  In
+some cases, as for example in the real MSSM, the different solutions
+are related by one or more global signs as in
+Eq.~\eqref{eq:EWSB-solution-for-mu}.  \fs recognizes such cases and
+introduces these global signs as additional free parameters to allow
+the user to chose between the different solutions.  The case of the
+real MSSM Eq.~\eqref{eq:EWSB-solution-for-mu} is therefore handled
+automatically in \fs, because $\sign\mu$ is automatically introduced
+as an additional free parameter.
+
+If, however, the different solutions are not related by global signs,
+then \fs writes all solutions to the file
 \code{models/<model>/<model>_tree_level_EWSB_solution.m}.  The user
@@
 \code{TreeLevelEWSBSolution} variable in the model file.  This is for
-example done in the $Z_3$-violating NMSSM (\code{SMSSM}), where $\mu$
-is chosen to be fixed by the EWSB equations: There the tree-level
-solution for $\mu$ has the form
+example the case in the $Z_3$-violating NMSSM (\code{SMSSM}), where
+$\mu$ is chosen to be fixed by the EWSB equations: There the
+tree-level solution for $\mu$ has the form
 %
@@
 %
-with a non-global $\sign X$.  In this model the user can parametrize
-the tree-level solution in terms of $\sign X$ in the
-\code{TreeLevelEWSBSolution} variable and add $\sign X$ by hand to the
-list of free input parameters, see
-\code{model_files/SMSSM/FlexibleSUSY.m.in} for an example.  If the
-user decides to not pick a particular solution by leaving the variable
-\code{TreeLevelEWSBSolution} empty, \fs tries to find a solution to
-the tree-level EWSB equations numerically via an iteration.  In this
-case, however, the user does not have the option to chose between the
-different solutions.
+with a non-global $\sign X$.  When \fs solves the tree-level EWSB
+equations for $\mu$, $B\mu$ and $\xi_S$ it finds the two solutions and
+writes them to \code{models/SMSSM/SMSSM_tree_level_EWSB_solution.m} in
+the form
+%
+\begin{lstlisting}[language=Mathematica]
+{
+   {{B[\[Mu]] -> ...}},
+   {
+      {\[Mu] -> (-20*Sqrt[2]*vd*vS*(\[Lambda] + conj[\[Lambda]])
+                 - Sqrt[...])/(80*vd)},
+      {\[Mu] -> (-20*Sqrt[2]*vd*vS*(\[Lambda] + conj[\[Lambda]])
+                 + Sqrt[...])/(80*vd)}
+   },
+   {{L[L1] -> ...}}
+}
+\end{lstlisting}
+%
+In the above solutions the dots stand for the full expression, which
+is left out here for better readability.  Inspecting the two above
+solutions for $\mu$, one finds that both can be parametrized by an
+additional free sign in front of the \code{Sqrt[...]}.  The user can
+now introduce an additional free sign by hand in the \code{MINPAR}
+block
+%
+\begin{lstlisting}[language=Mathematica]
+MINPAR = {
+   {1, m0},
+   {2, m12},
+   {3, TanBeta},
+   {4, Sign[X]},  (* <-- additional free sign *)
+   {5, Azero}
+};
+\end{lstlisting}
+%
+and set the the tree-level solution, parametrized in terms of
+\code{Sign[X]}, in the \code{TreeLevelEWSBSolution} variable in
+the model file:
+%
+\begin{lstlisting}
+TreeLevelEWSBSolution = {
+   { B[\[Mu]], ... },
+   { \[Mu], (-20*Sqrt[2]*vd*vS*(\[Lambda] + conj[\[Lambda]])
+             + Sign[X] * Sqrt[...])/(80*vd) },
+   { L[L1], ... }
+};
+\end{lstlisting}
+%
+One can now chose between the two solutions by setting entry $4$ in
+the \code{MINPAR} block of the SLHA input file to either $+1$ or $-1$.
+See \code{model_files/SMSSM/FlexibleSUSY.m.in} for a complete example
+model file.  If the user decides to not pick a particular solution by
+leaving the variable \code{TreeLevelEWSBSolution} empty, \fs tries to
+find a solution to the tree-level EWSB equations numerically via an
+iteration.  In this case, however, the user does not have the option
+to chose between the different solutions.
 

### 2014-11-06 doc/flexiblesusy-paper.tex
Move to appendix sections on C++ code adaptation

@@
 \renewcommand{\ptitle}[1]{}
+\newcommand{\Zv}{\mathbf{\backslash}\mkern-11.0mu{Z}}
 \makeatletter
@@
 some limitations to what can be done. In it's current form \fs assumes
-all couplings are are real, therefore it is limited to CP conserving
+all couplings are are real, therefore it is limited to $CP$-conserving
 versions of SUSY models.   Although it is implicit in the
@@
 \fs allows the user to add leading two-loop contributions to the
-CP-even and CP-odd Higgs self-energies as well as to the CP-even Higgs
-tadpoles.  For MSSM-like models (with two CP-even Higgs bosons, one
-CP-odd Higgs boson, one neutral Goldstone boson) routines dor
+$CP$-even and $CP$-odd Higgs self-energies as well as to the $CP$-even Higgs
+tadpoles.  For MSSM-like models (with two $CP$-even Higgs bosons, one
+$CP$-odd Higgs boson, one neutral Goldstone boson) routines dor
 caluclating these corrections will be generated by setting
@@
 \cite{Degrassi:2001yf,Brignole:2001jy,Dedes:2002dy,Brignole:2002bz,Dedes:2003km}.
-For NMSSM-like models (with three CP-even Higgs bosons, two CP-odd
+For NMSSM-like models (with three $CP$-even Higgs bosons, two $CP$-odd
 Higgs bosons, one neutral Goldstone boson) the two-loop contributions
@@
 the \code{SMINPUTS} block of the SLHA input file \cite{Skands:2003cj}.
-The CKM mixing matrix is currently set to unity and CP-violating
+The CKM mixing matrix is currently set to unity and $CP$-violating
 phases are set to zero.  The third generation quark masses are
@@
 %
-(i) the calculated lightest CP-even Higgs pole mass, $m_h$, differs
+(i) the calculated lightest $CP$-even Higgs pole mass, $m_h$, differs
 about $2\unit{GeV}$ between \code{MediumPrecision} and
@@
   implements $R$-parity conservation or something equivalent, and
-  (iii) contains two CP-even and CP-odd Higgs bosons, where one CP-odd
+  (iii) contains two $CP$-even and $CP$-odd Higgs bosons, where one $CP$-odd
   Higgs boson may be a Goldstone boson, all with an MSSM-like coupling
@@
   NMSSM one, (ii) it implements $R$-parity conservation or something
-  equivalent, and (iii) contains three CP-even and CP-odd Higgs
-  bosons, where one or two CP-odd Higgs bosons may be Goldstone
+  equivalent, and (iii) contains three $CP$-even and $CP$-odd Higgs
+  bosons, where one or two $CP$-odd Higgs bosons may be Goldstone
   bosons, all with an NMSSM-like coupling to $t$, $b$ and $\tau$\@.
@@
 
-\subsubsection{Stacking models in a tower of effective theories}
+The procedure described above is essentially all that is needed
+to employ new spectrum generator components,
+which may be composed from scratch or through
+a linkage to external routines.
+In doing so,
+notice that there was an evident limit on the scope of
+modules which one had to deal with.
+For instance, it was clear from the outset
+that one does not have to go through the code of the central
+fixed-point iteration engine, \code{RGFlow}.
+This manifests the power of the clear separation among objects
+each with its well-defined distinct role.
+This is just like the fact that one does not need to access the internals
+of the \code{std::sort} function in the C++ Standard Library.
+It might be entertaining to complete the analogy by
+mapping the model objects in \code{RGFlow}
+to the elements that \code{std::sort} sorts
+and the boundary condition objects
+to the comparator function.
+
+\section{Tests and comparisons with other spectrum generators}
+\label{sec:comparison}
+
+\subsection{Numeric tests}
+
+To check the correctness of \fs's generated spectrum generators
+extensive unit testing against Softsusy's MSSM and NMSSM
+implementations (both $Z_3$-invariant and $Z_3$-violating variants)
+has been carried out.  These unit tests systematically compare all
+tree-level mass matrices, EWSB equations, one- and two-loop
+$\beta$-functions, one- and two-loop self-energies and one- and
+two-loop tadpoles numerically for the CMSSM and NMSSM parameter points
+given in \tabref{tab:unit-test-parameter-points}.
+%
+\begin{table}[tbh]
+  \centering
+  \begin{tabularx}{\textwidth}{lX}
+    \toprule
+    SUSY Model & Tested parameter points\\
+    \midrule CMSSM & $m_0 = 125\unit{GeV}$, $M_{1/2} = 500\unit{GeV}$,
+    $\tan\beta = 10$, $A_0 = 0$, $\sign\mu = \pm 1$
+    \\
+    $Z_3$-NMSSM & $m_0 = \{250,300\}\unit{GeV}$, $M_{1/2} =
+    200\unit{GeV}$, $\tan\beta = 10$, $A_0 = -500\unit{GeV}$,
+    $\sign\mu_\text{eff} = +1$, $\lambda = 0.1$
+    \\
+    $\Zv_3$-NMSSM & $m_0 = 540\unit{GeV}$, $M_{1/2} = 200\unit{GeV}$,
+    $\tan\beta = 10$, $A_0 = -350\unit{GeV}$, $\sign\mu_\text{eff} =
+    \pm 1$, $\lambda = \kappa = 0.1$, $s = 1\unit{TeV}$, $\mu' =
+    290\unit{GeV}$, $m_S^{\prime 2} = 400\unit{GeV}$, $\xi_F =
+    300\unit{GeV}$
+    \\
+    \bottomrule
+  \end{tabularx}
+  \caption{CMSSM and NMSSM parameter points used for the unit tests
+    against Softsusy.
+    We follow the notation of the NMSSM model parameters used in
+    \cite{Ellwanger:2009dp,Allanach:2013kza}.}
+  \label{tab:unit-test-parameter-points}
+\end{table}
+%
+All tested expressions were found to agree within double machine
+precision.\footnote{Due to the systematic and detailed tests several
+  bugs in Softsusy, SARAH and \fs could be identified and corrected.}
+Furthermore, the output of the iterative procedures which solve the
+one- and two-loop corrected tadpole equations
+\eqref{eq:one-loop-ewsb-eq} to find the minimum of the effective Higgs
+potential were compared numerically for these parameter points and
+found to agree within machine precision as well.  Finally, the overall
+pole mass spectrum and mixing after the full fixed-point iteration has
+finished has been compared, and was found to agree at the sub-permille
+level.  The origin of the sub-permille level difference between \fs
+and Softsusy is the different determination of the weak mixing angle
+$\theta_{W,\text{susy}}^{\text{\DRbar}}(M_Z)$ in the SUSY model in the
+\DRbar\ scheme: \fs calculates
+$\theta_{W,\text{susy}}^{\text{\DRbar}}(M_Z)$ from $M_W$ and $M_Z$, as
+described in \secref{sec:calculation-of-gauge-couplings}, while
+Softsusy determines $\theta_{W,\text{susy}}^{\text{\DRbar}}(M_Z)$ from
+the muon decay constant $G_\mu$.  The approach used in Softsusy
+results in more precise \DRbar\ gauge couplings at the $M_Z$ scale,
+because the muon decay constant is known with a higher accuracy than
+the $W$-boson mass.  Furthermore, in Softsusy $3.5.0$ some three-loop
+$\beta$-functions and two-loop threshold corrections can be enabled in
+the MSSM to increase the accuracy of the RG running and the
+determination of the \DRbar\ gauge and Yukawa couplings at $M_Z$
+\cite{Allanach:2014nba}.  These corrections are not implemented in \fs
+so far.  The complete set of unit tests is shipped with \fs and can be
+found in the \code{test/} directory.  The tests can be run with the
+command \code{make execute-tests}.  All unit tests are carried out
+nightly in order to continuously check the correctness of the meta
+code and the generated spectrum generators for the shipped models.
+The nightly test results can be found at
+\url{http://iktp.tu-dresden.de/~avoigt/FlexibleSUSY/test.xhtml}
+
+The \fs generated NUHM \ESSM\ spectrum generator has also been
+compared against a handwritten one for a constrained version of the
+\ESSM \cite{Athron:2009ue, Athron:2009bs, Athron:2012pw}.  The
+$\beta$-functions were systematically compared in unit tests and were
+found to match at the ??? level.  The handwritten code doesn't include
+full one-loop self-energies or tadpoles, so tests on these were not
+carried out.  Although the generators assume different constraints
+they could be compared by using the output of the CE$_6$SSM generator
+as an input to \fs and the spectra were found to be in reasonable
+agreement, given the different levels of precision.
+
+In addition \fs has already undergone some user testing. This includes
+analytic tests of the $R$-symmetric low-energy model (MRSSM) and
+alternative $E_6$-inspired SUSY scenarios.  The users who have helped us
+with this are thanked in the acknowledgements.
+
+We also compared the run-time of \fs against SPheno, Softsusy and the
+SARAH generated MSSM spectrum generator SPhenoMSSM.  The results of
+the comparison can be found in \secref{sec:run-time comparison}.
+
+\subsection{Run-time comparison}
+\label{sec:run-time comparison}
+
+One of \fs's design goals is a short run-time.  In this section we
+demonstrate that this goal was achieved by comparing the run-time of
+two different sets of CMSSM spectrum generators:
+%
+\begin{itemize}
+\item \emph{Without sfermion flavour violation:} Disallowing sfermion
+  flavour violation simplifies the calculation of the pole masses,
+  because flavour-off-diagonal sfermion self-energy matrix elements
+  don't need to be calculated.  Here we compare \fs's non-flavour
+  violating CMSSM spectrum generator FlexibleSUSY-NoFV (version 1.0.0)
+  against SPheno (version 3.2.4) and Softsusy (version 3.4.0).
+%
+\item \emph{With sfermion flavour violation:} Allowing for sfermion
+  flavour violation in general increases the run-time of spectrum
+  generators, because the full $6\times 6$ sfermion self-energy
+  matrices have to be calculated.  Here we compare FlexibleSUSY-FV
+  (version 1.0.0) and SPhenoMSSM (generated with SARAH 4.1.0 and
+  linked against SPheno 3.2.4).  Both spectrum generators are based on
+  SARAH's MSSM model file, which allows for sfermion flavour
+  violation.
+\end{itemize}
+%
+For the run-time comparison the CKM matrix is set to unity, all
+$CP$-violating phases are set to zero and $R$-parity violation is
+disabled.  \fs and Softsusy are compiled with g++ 4.8.0 and Intel
+ifort 13.1.3 20130607.  SPheno and SPhenoMSSM are compiled with Intel
+ifort 13.1.3 20130607.\footnote{Intel's ifort compiler decreases the
+  run-time of SPheno and SPhenoMSSM by approximately a factor $1.5$,
+  compared to gfortran.}  We're generating $2\cdot 10^{4}$ random
+CMSSM parameter points with $m_0\in [50,1000]\unit{GeV}$, $m_{1/2}\in
+[50,1000]\unit{GeV}$, $\tan\beta\in [1,100]$, $\sign\mu\in \{-1,+1\}$
+and $A_0\in [-1000,$ $1000]\unit{GeV}$.  For each point an SLHA input
+file is created by appending the values of $m_0$, $m_{1/2}$,
+$\tan\beta$, $\sign\mu$, $A_0$ in form of a \code{MINPAR} block to the
+SLHA template file given in \ref{sec:speed-test-slha-template-file}.
+The resulting SLHA input file is passed to each spectrum generator and
+the (wall-clock) time is measured until the program has finished.  The
+average run-times for three different CPU types can be found in
+\tabref{tab:run-time-comparison}.  The first column shows the run-time
+on a Intel Core2 Duo (P8600, $2.40\unit{GHz}$) where only one core was
+enabled.  The second column shows the run-time on the same processor
+where both cores were enabled.  In the third column a machine with two
+Intel Xeon CPUs (L5640, $2.27\unit{GHz}$, $6$ cores) was used.
+%
+\begin{table}[tbh]
+  \centering
+  \begin{tabular}{llll}
+    \toprule
+                            & Intel Core2 Duo    & Intel Core2 Duo   & $2$ $\times$ Intel Xeon\\
+                            & (P8600, $1$ core)  & (P8600, $2$ cores)& (L5640, $6$ cores)\\
+    \midrule
+    FlexibleSUSY-NoFV 1.0.0 & $0.086\unit{s}$    & $0.079\unit{s}$   & $0.060\unit{s}$\\
+    SPheno 3.2.4            & $0.119\unit{s}$    & $0.114\unit{s}$   & $0.101\unit{s}$\\
+    Softsusy 3.4.0          & $0.175\unit{s}$    & $0.171\unit{s}$   & $0.147\unit{s}$\\
+    \midrule
+    FlexibleSUSY-FV 1.0.0   & $0.150\unit{s}$    & $0.113\unit{s}$   & $0.074\unit{s}$\\
+    SPhenoMSSM 4.1.0        & $0.415\unit{s}$    & $0.401\unit{s}$   & $0.370\unit{s}$\\
+    \bottomrule
+  \end{tabular}
+  \caption{Average run-time of CMSSM spectrum generators
+    for random parameter points.  The first three rows show three
+    spectrum generators with disabled sfermion flavour violation.  Rows
+    $4$--$5$ show two spectrum generators with enabled sfermion flavour violation, both based
+    on SARAH's MSSM model file.}
+  \label{tab:run-time-comparison}
+\end{table}
+
+Under both the non-flavour violating spectrum generators (first three
+rows) as well as the flavour violating ones (4th and 5th row) we find
+that \fs is significantly fastest.  Compared to SPheno,
+FlexibleSUSY-NoFV is faster by a factor $1.4$--$1.7$, and compared to
+Softsusy around a factor $2$--$2.5$.  Under the flavour violating
+spectrum generators FlexibleSUSY-FV is faster than SPhenoMSSM by a
+factor $2.8$--$5$.  Reason for the long run-time of SPhenoMSSM is the
+long calculation duration of the two-loop $\beta$-functions.  Here \fs
+benefits a lot from Eigen's well-optimizable matrix expressions.  We
+also find that increasing the number of CPU cores reduces the run-time
+of \fs.  The reason is that \fs calculates each pole mass in a
+separate thread, and therefore benefits from multi-core CPUs.
+
+\section{Conclusions}
+
+We have presented \fs, a Mathematica and C++ package, which generates fast and
+modular spectrum generators for any user specified SUSY model.  \fs is
+distributed with a large number of predefined models for the CMSSM,
+NMSSM, USSM, \ESSM, MRSSM etc., which can be generated immediately
+without any editing.  In particular the CMSSM and NMSSM spectrum
+generators constitute a fast and reliable alternative to the existing
+publicly available spectrum generators, Softsusy, SPheno and
+NMSPEC.
+
+We have described how the generated source code can be influenced at
+two different levels: The \mathematica level where the user provides a
+model file, and the C++ level where the generated objects can be
+easily exchanged, extended, modified and reused.  This provides great
+flexibility for creating custom spectrum generators for both the most
+common and most extraordinary models.  We have demonstrated these
+features in detailed examples for the NUHMSSM, right handed neutrinos
+and on adding three-loop RGEs and two-loop matching for the strong
+gauge coupling.
+
+The generated code has been extensively tested against Softsusy, and
+additional tests have been carried out for non-minimal models, the
+\ESSM and MRSSM.  Speed tests have also been performed against
+Softsusy, SPheno and SPheno-like MSSM code generated by SARAH,
+demonstrating that \fs runs faster than all three.
+
+As a result \fs enables fast exploitation of new SUSY models with high
+precision and reliability.
+
+\section*{Acknowledgments}
+
+A.V.\ would like to thank Florian Staub for countless explanations of
+SARAH's internals, discussions, and exceptionally fast bug
+fixing. P.A.~would like to thank Roman Nevzorov for useful discussions
+about challenges in non-minimal SUSY models and the GAMBIT
+collaboration for helpful suggestions, particularly regarding
+lowering our compiler prerequisites. The authors would also like to
+thank Lewis Tunstall for helping with early tests against
+Next-to-Minimal Softsusy; Sophie Underwood for discovering problems
+when introducing couplings with a single family index; Gregor Hellwig
+for writing the first version of the E6SSM model file; Philip Diessner
+for further adaptions to the E6SSM model file and for supplying it to
+us and for for testing and identifying several bugs in \fs and in
+SARAH, in work on the MRSSM; Ulrik Günther for compilation tests on
+Mac OS X and Dylan Harries for spotting a bug in the configuration
+script.
+%
+J.P. acknowledges support from the MEC and FEDER (EC) Grants
+FPA2011--23596 and the Generalitat Valenciana under grant PROMETEOII/2013/017.
+This work has been supported by the German Research Foundation DFG
+through Grant No.~STO876/2-1.
+
+\appendix
+
+\section{Examples of C++ code adaptation}
+\label{sec:examples of c++ code adaptation}
+
+\subsection{Stacking models in a tower of effective theories}
 \label{sec:tower construction}
@@
 
-\subsubsection{Integrating custom-built C++ components}
+\subsection{Integrating custom-built C++ components}
 \label{sec:integrating-custom-built}
@@
 
-The procedure described above is essentially all that is needed
-to employ new spectrum generator components,
-which may be composed from scratch or through
-a linkage to external routines.
-In doing so,
-notice that there was an evident limit on the scope of
-modules which one had to deal with.
-For instance, it was clear from the outset
-that one does not have to go through the code of the central
-fixed-point iteration engine, \code{RGFlow}.
-This manifests the power of the clear separation among objects
-each with its well-defined distinct role.
-This is just like the fact that one does not need to access the internals
-of the \code{std::sort} function in the C++ Standard Library.
-It might be entertaining to complete the analogy by
-mapping the model objects in \code{RGFlow}
-to the elements that \code{std::sort} sorts
-and the boundary condition objects
-to the comparator function.
-
-\section{Tests and comparisons with other spectrum generators}
-\label{sec:comparison}
-
-\subsection{Numeric tests}
-
-To check the correctness of \fs's generated spectrum generators
-extensive unit testing against Softsusy's MSSM and NMSSM
-implementations (both $Z_3$-invariant and $Z_3$-violating variants)
-has been carried out.  These unit tests systematically compare all
-tree-level mass matrices, EWSB equations, one- and two-loop
-$\beta$-functions, one- and two-loop self-energies and one- and
-two-loop tadpoles numerically for various MSSM and NMSSM parameter
-points.  All tested expressions were found to agree within double
-machine precision.\footnote{Due to the systematic and detailed tests
-  several bugs in Softsusy, SARAH and \fs could be identified and
-  corrected.}  Furthermore, the overall pole mass spectrum after the
-full fixed-point iteration has finished has been compared, and was
-found to agree at the sub-permille level.  The origin of the
-sub-permille level difference between \fs and Softsusy is the
-different determination of the weak mixing angle
-$\theta_{W,\text{susy}}^{\text{\DRbar}}(M_Z)$ in the SUSY model in the
-\DRbar\ scheme: \fs calculates
-$\theta_{W,\text{susy}}^{\text{\DRbar}}(M_Z)$ from $M_W$ and $M_Z$, as
-described in \secref{sec:calculation-of-gauge-couplings}, while
-Softsusy determines $\theta_{W,\text{susy}}^{\text{\DRbar}}(M_Z)$ from
-the muon decay constant $G_\mu$.  The approach used in Softsusy
-results in more precise \DRbar\ gauge couplings at the $M_Z$ scale,
-because the muon decay constant is known with a higher accuracy than
-the $W$-boson mass.  Furthermore, in Softsusy $3.5.0$ some three-loop
-$\beta$-functions and two-loop threshold corrections can be enabled in
-the MSSM to increase the accuracy of the RG running and the
-determination of the \DRbar\ gauge and Yukawa couplings at $M_Z$
-\cite{Allanach:2014nba}.  These corrections are not implemented in \fs
-so far.  The complete set of unit tests is shipped with \fs and can be
-found in the \code{test/} directory.  The tests can be run with the
-command \code{make execute-tests}.  All unit tests are carried out
-nightly in order to continuously check the correctness of the meta
-code and the generated spectrum generators for the shipped models.
-The nightly test results can be found at
-\url{http://iktp.tu-dresden.de/~avoigt/FlexibleSUSY/test.xhtml}
-
-The \fs generated NUHM \ESSM\ spectrum generator has also been
-compared against a handwritten one for a constrained version of the
-\ESSM \cite{Athron:2009ue, Athron:2009bs, Athron:2012pw}.  The
-$\beta$-functions were systematically compared in unit tests and were
-found to match at the ??? level.  The handwritten code doesn't include
-full one-loop self-energies or tadpoles, so tests on these were not
-carried out.  Although the generators assume different constraints
-they could be compared by using the output of the CE$_6$SSM generator
-as an input to \fs and the spectra were found to be in reasonable
-agreement, given the different levels of precision.
-
-In addition \fs has already undergone some user testing. This includes
-analytic tests of the $R$-symmetric low-energy model (MRSSM) and
-alternative $E_6$-inspired SUSY scenarios.  The users who have helped us
-with this are thanked in the acknowledgements.
-
-We also compared the run-time of \fs against SPheno, Softsusy and the
-SARAH generated MSSM spectrum generator SPhenoMSSM.  The results of
-the comparison can be found in \secref{sec:run-time comparison}.
-
-\subsection{Run-time comparison}
-\label{sec:run-time comparison}
-
-One of \fs's design goals is a short run-time.  In this section we
-demonstrate that this goal was achieved by comparing the run-time of
-two different sets of CMSSM spectrum generators:
-%
-\begin{itemize}
-\item \emph{Without sfermion flavour violation:} Disallowing sfermion
-  flavour violation simplifies the calculation of the pole masses,
-  because flavour-off-diagonal sfermion self-energy matrix elements
-  don't need to be calculated.  Here we compare \fs's non-flavour
-  violating CMSSM spectrum generator FlexibleSUSY-NoFV (version 1.0.0)
-  against SPheno (version 3.2.4) and Softsusy (version 3.4.0).
-%
-\item \emph{With sfermion flavour violation:} Allowing for sfermion
-  flavour violation in general increases the run-time of spectrum
-  generators, because the full $6\times 6$ sfermion self-energy
-  matrices have to be calculated.  Here we compare FlexibleSUSY-FV
-  (version 1.0.0) and SPhenoMSSM (generated with SARAH 4.1.0 and
-  linked against SPheno 3.2.4).  Both spectrum generators are based on
-  SARAH's MSSM model file, which allows for sfermion flavour
-  violation.
-\end{itemize}
-%
-For the run-time comparison the CKM matrix is set to unity, all CP
-violating phases are set to zero and R-parity violation is disabled.
-\fs and Softsusy are compiled with g++ 4.8.0 and Intel ifort 13.1.3
-20130607.  SPheno and SPhenoMSSM are compiled with Intel ifort 13.1.3
-20130607.\footnote{Intel's ifort compiler decreases the run-time of
-  SPheno and SPhenoMSSM by approximately a factor $1.5$, compared to
-  gfortran.}  We're generating $2\cdot 10^{4}$ random CMSSM parameter
-points with $m_0\in [50,1000]\unit{GeV}$, $m_{1/2}\in
-[50,1000]\unit{GeV}$, $\tan\beta\in [1,100]$, $\sign\mu\in \{-1,+1\}$
-and $A_0\in [-1000,$ $1000]\unit{GeV}$.  For each point an SLHA input
-file is created by appending the values of $m_0$, $m_{1/2}$,
-$\tan\beta$, $\sign\mu$, $A_0$ in form of a \code{MINPAR} block to the
-SLHA template file given in \ref{sec:speed-test-slha-template-file}.
-The resulting SLHA input file is passed to each spectrum generator and
-the (wall-clock) time is measured until the program has finished.  The
-average run-times for three different CPU types can be found in
-\tabref{tab:run-time-comparison}.  The first column shows the run-time
-on a Intel Core2 Duo (P8600, $2.40\unit{GHz}$) where only one core was
-enabled.  The second column shows the run-time on the same processor
-where both cores were enabled.  In the third column a machine with two
-Intel Xeon CPUs (L5640, $2.27\unit{GHz}$, $6$ cores) was used.
-%
-\begin{table}[tbh]
-  \centering
-  \begin{tabular}{llll}
-    \toprule
-                            & Intel Core2 Duo    & Intel Core2 Duo   & $2$ $\times$ Intel Xeon\\
-                            & (P8600, $1$ core)  & (P8600, $2$ cores)& (L5640, $6$ cores)\\
-    \midrule
-    FlexibleSUSY-NoFV 1.0.0 & $0.086\unit{s}$    & $0.079\unit{s}$   & $0.060\unit{s}$\\
-    SPheno 3.2.4            & $0.119\unit{s}$    & $0.114\unit{s}$   & $0.101\unit{s}$\\
-    Softsusy 3.4.0          & $0.175\unit{s}$    & $0.171\unit{s}$   & $0.147\unit{s}$\\
-    \midrule
-    FlexibleSUSY-FV 1.0.0   & $0.150\unit{s}$    & $0.113\unit{s}$   & $0.074\unit{s}$\\
-    SPhenoMSSM 4.1.0        & $0.415\unit{s}$    & $0.401\unit{s}$   & $0.370\unit{s}$\\
-    \bottomrule
-  \end{tabular}
-  \caption{Average run-time of CMSSM spectrum generators
-    for random parameter points.  The first three rows show three
-    spectrum generators with disabled sfermion flavour violation.  Rows
-    $4$--$5$ show two spectrum generators with enabled sfermion flavour violation, both based
-    on SARAH's MSSM model file.}
-  \label{tab:run-time-comparison}
-\end{table}
-
-Under both the non-flavour violating spectrum generators (first three
-rows) as well as the flavour violating ones (4th and 5th row) we find
-that \fs is significantly fastest.  Compared to SPheno,
-FlexibleSUSY-NoFV is faster by a factor $1.4$--$1.7$, and compared to
-Softsusy around a factor $2$--$2.5$.  Under the flavour violating
-spectrum generators FlexibleSUSY-FV is faster than SPhenoMSSM by a
-factor $2.8$--$5$.  Reason for the long run-time of SPhenoMSSM is the
-long calculation duration of the two-loop $\beta$-functions.  Here \fs
-benefits a lot from Eigen's well-optimizable matrix expressions.  We
-also find that increasing the number of CPU cores reduces the run-time
-of \fs.  The reason is that \fs calculates each pole mass in a
-separate thread, and therefore benefits from multi-core CPUs.
-
-\section{Conclusions}
-
-We have presented \fs, a Mathematica and C++ package, which generates fast and
-modular spectrum generators for any user specified SUSY model.  \fs is
-distributed with a large number of predefined models for the CMSSM,
-NMSSM, USSM, \ESSM, MRSSM etc., which can be generated immediately
-without any editing.  In particular the CMSSM and NMSSM spectrum
-generators constitute a fast and reliable alternative to the existing
-publicly available spectrum generators, Softsusy, SPheno and
-NMSPEC.
-
-We have described how the generated source code can be influenced at
-two different levels: The \mathematica level where the user provides a
-model file, and the C++ level where the generated objects can be
-easily exchanged, extended, modified and reused.  This provides great
-flexibility for creating custom spectrum generators for both the most
-common and most extraordinary models.  We have demonstrated these
-features in detailed examples for the NUHMSSM, right handed neutrinos
-and on adding three-loop RGEs and two-loop matching for the strong
-gauge coupling.
-
-The generated code has been extensively tested against Softsusy, and
-additional tests have been carried out for non-minimal models, the
-\ESSM and MRSSM.  Speed tests have also been performed against
-Softsusy, SPheno and SPheno-like MSSM code generated by SARAH,
-demonstrating that \fs runs faster than all three.
-
-As a result \fs enables fast exploitation of new SUSY models with high
-precision and reliability.
-
-\section*{Acknowledgments}
-
-A.V.\ would like to thank Florian Staub for countless explanations of
-SARAH's internals, discussions, and exceptionally fast bug
-fixing. P.A.~would like to thank Roman Nevzorov for useful discussions
-about challenges in non-minimal SUSY models and the GAMBIT
-collaboration for helpful suggestions, particularly regarding
-lowering our compiler prerequisites. The authors would also like to
-thank Lewis Tunstall for helping with early tests against
-Next-to-Minimal Softsusy; Sophie Underwood for discovering problems
-when introducing couplings with a single family index; Gregor Hellwig
-for writing the first version of the E6SSM model file; Philip Diessner
-for further adaptions to the E6SSM model file and for supplying it to
-us and for for testing and identifying several bugs in \fs and in
-SARAH, in work on the MRSSM; Ulrik Günther for compilation tests on
-Mac OS X and Dylan Harries for spotting a bug in the configuration
-script.
-%
-J.P. acknowledges support from the MEC and FEDER (EC) Grants
-FPA2011--23596 and the Generalitat Valenciana under grant PROMETEOII/2013/017.
-This work has been supported by the German Research Foundation DFG
-through Grant No.~STO876/2-1.
-
-\appendix
-
 \section{Speed test SLHA input file}

### 2014-11-09 doc/flexiblesusy-paper.tex
fix typographical errors in the paper

@@
 \renewcommand{\ptitle}[1]{}
-\newcommand{\Zv}{\mathbf{\backslash}\mkern-11.0mu{Z}}
 \makeatletter
@@
 
-The procedure described above is essentially all that is needed
-to employ new spectrum generator components,
-which may be composed from scratch or through
-a linkage to external routines.
-In doing so,
-notice that there was an evident limit on the scope of
-modules which one had to deal with.
-For instance, it was clear from the outset
-that one does not have to go through the code of the central
-fixed-point iteration engine, \code{RGFlow}.
-This manifests the power of the clear separation among objects
-each with its well-defined distinct role.
-This is just like the fact that one does not need to access the internals
-of the \code{std::sort} function in the C++ Standard Library.
-It might be entertaining to complete the analogy by
-mapping the model objects in \code{RGFlow}
-to the elements that \code{std::sort} sorts
-and the boundary condition objects
-to the comparator function.
-
-\section{Tests and comparisons with other spectrum generators}
-\label{sec:comparison}
-
-\subsection{Numeric tests}
-
-To check the correctness of \fs's generated spectrum generators
-extensive unit testing against Softsusy's MSSM and NMSSM
-implementations (both $Z_3$-invariant and $Z_3$-violating variants)
-has been carried out.  These unit tests systematically compare all
-tree-level mass matrices, EWSB equations, one- and two-loop
-$\beta$-functions, one- and two-loop self-energies and one- and
-two-loop tadpoles numerically for the CMSSM and NMSSM parameter points
-given in \tabref{tab:unit-test-parameter-points}.
-%
-\begin{table}[tbh]
-  \centering
-  \begin{tabularx}{\textwidth}{lX}
-    \toprule
-    SUSY Model & Tested parameter points\\
-    \midrule CMSSM & $m_0 = 125\unit{GeV}$, $M_{1/2} = 500\unit{GeV}$,
-    $\tan\beta = 10$, $A_0 = 0$, $\sign\mu = \pm 1$
-    \\
-    $Z_3$-NMSSM & $m_0 = \{250,300\}\unit{GeV}$, $M_{1/2} =
-    200\unit{GeV}$, $\tan\beta = 10$, $A_0 = -500\unit{GeV}$,
-    $\sign\mu_\text{eff} = +1$, $\lambda = 0.1$
-    \\
-    $\Zv_3$-NMSSM & $m_0 = 540\unit{GeV}$, $M_{1/2} = 200\unit{GeV}$,
-    $\tan\beta = 10$, $A_0 = -350\unit{GeV}$, $\sign\mu_\text{eff} =
-    \pm 1$, $\lambda = \kappa = 0.1$, $s = 1\unit{TeV}$, $\mu' =
-    290\unit{GeV}$, $m_S^{\prime 2} = 400\unit{GeV}$, $\xi_F =
-    300\unit{GeV}$
-    \\
-    \bottomrule
-  \end{tabularx}
-  \caption{CMSSM and NMSSM parameter points used for the unit tests
-    against Softsusy.
-    We follow the notation of the NMSSM model parameters used in
-    \cite{Ellwanger:2009dp,Allanach:2013kza}.}
-  \label{tab:unit-test-parameter-points}
-\end{table}
-%
-All tested expressions were found to agree within double machine
-precision.\footnote{Due to the systematic and detailed tests several
-  bugs in Softsusy, SARAH and \fs could be identified and corrected.}
-Furthermore, the output of the iterative procedures which solve the
-one- and two-loop corrected tadpole equations
-\eqref{eq:one-loop-ewsb-eq} to find the minimum of the effective Higgs
-potential were compared numerically for these parameter points and
-found to agree within machine precision as well.  Finally, the overall
-pole mass spectrum and mixing after the full fixed-point iteration has
-finished has been compared, and was found to agree at the sub-permille
-level.  The origin of the sub-permille level difference between \fs
-and Softsusy is the different determination of the weak mixing angle
-$\theta_{W,\text{susy}}^{\text{\DRbar}}(M_Z)$ in the SUSY model in the
-\DRbar\ scheme: \fs calculates
-$\theta_{W,\text{susy}}^{\text{\DRbar}}(M_Z)$ from $M_W$ and $M_Z$, as
-described in \secref{sec:calculation-of-gauge-couplings}, while
-Softsusy determines $\theta_{W,\text{susy}}^{\text{\DRbar}}(M_Z)$ from
-the muon decay constant $G_\mu$.  The approach used in Softsusy
-results in more precise \DRbar\ gauge couplings at the $M_Z$ scale,
-because the muon decay constant is known with a higher accuracy than
-the $W$-boson mass.  Furthermore, in Softsusy $3.5.0$ some three-loop
-$\beta$-functions and two-loop threshold corrections can be enabled in
-the MSSM to increase the accuracy of the RG running and the
-determination of the \DRbar\ gauge and Yukawa couplings at $M_Z$
-\cite{Allanach:2014nba}.  These corrections are not implemented in \fs
-so far.  The complete set of unit tests is shipped with \fs and can be
-found in the \code{test/} directory.  The tests can be run with the
-command \code{make execute-tests}.  All unit tests are carried out
-nightly in order to continuously check the correctness of the meta
-code and the generated spectrum generators for the shipped models.
-The nightly test results can be found at
-\url{http://iktp.tu-dresden.de/~avoigt/FlexibleSUSY/test.xhtml}
-
-The \fs generated NUHM \ESSM\ spectrum generator has also been
-compared against a handwritten one for a constrained version of the
-\ESSM \cite{Athron:2009ue, Athron:2009bs, Athron:2012pw}.  The
-$\beta$-functions were systematically compared in unit tests and were
-found to match at the ??? level.  The handwritten code doesn't include
-full one-loop self-energies or tadpoles, so tests on these were not
-carried out.  Although the generators assume different constraints
-they could be compared by using the output of the CE$_6$SSM generator
-as an input to \fs and the spectra were found to be in reasonable
-agreement, given the different levels of precision.
-
-In addition \fs has already undergone some user testing. This includes
-analytic tests of the $R$-symmetric low-energy model (MRSSM) and
-alternative $E_6$-inspired SUSY scenarios.  The users who have helped us
-with this are thanked in the acknowledgements.
-
-We also compared the run-time of \fs against SPheno, Softsusy and the
-SARAH generated MSSM spectrum generator SPhenoMSSM.  The results of
-the comparison can be found in \secref{sec:run-time comparison}.
-
-\subsection{Run-time comparison}
-\label{sec:run-time comparison}
-
-One of \fs's design goals is a short run-time.  In this section we
-demonstrate that this goal was achieved by comparing the run-time of
-two different sets of CMSSM spectrum generators:
-%
-\begin{itemize}
-\item \emph{Without sfermion flavour violation:} Disallowing sfermion
-  flavour violation simplifies the calculation of the pole masses,
-  because flavour-off-diagonal sfermion self-energy matrix elements
-  don't need to be calculated.  Here we compare \fs's non-flavour
-  violating CMSSM spectrum generator FlexibleSUSY-NoFV (version 1.0.0)
-  against SPheno (version 3.2.4) and Softsusy (version 3.4.0).
-%
-\item \emph{With sfermion flavour violation:} Allowing for sfermion
-  flavour violation in general increases the run-time of spectrum
-  generators, because the full $6\times 6$ sfermion self-energy
-  matrices have to be calculated.  Here we compare FlexibleSUSY-FV
-  (version 1.0.0) and SPhenoMSSM (generated with SARAH 4.1.0 and
-  linked against SPheno 3.2.4).  Both spectrum generators are based on
-  SARAH's MSSM model file, which allows for sfermion flavour
-  violation.
-\end{itemize}
-%
-For the run-time comparison the CKM matrix is set to unity, all
-$CP$-violating phases are set to zero and $R$-parity violation is
-disabled.  \fs and Softsusy are compiled with g++ 4.8.0 and Intel
-ifort 13.1.3 20130607.  SPheno and SPhenoMSSM are compiled with Intel
-ifort 13.1.3 20130607.\footnote{Intel's ifort compiler decreases the
-  run-time of SPheno and SPhenoMSSM by approximately a factor $1.5$,
-  compared to gfortran.}  We're generating $2\cdot 10^{4}$ random
-CMSSM parameter points with $m_0\in [50,1000]\unit{GeV}$, $m_{1/2}\in
-[50,1000]\unit{GeV}$, $\tan\beta\in [1,100]$, $\sign\mu\in \{-1,+1\}$
-and $A_0\in [-1000,$ $1000]\unit{GeV}$.  For each point an SLHA input
-file is created by appending the values of $m_0$, $m_{1/2}$,
-$\tan\beta$, $\sign\mu$, $A_0$ in form of a \code{MINPAR} block to the
-SLHA template file given in \ref{sec:speed-test-slha-template-file}.
-The resulting SLHA input file is passed to each spectrum generator and
-the (wall-clock) time is measured until the program has finished.  The
-average run-times for three different CPU types can be found in
-\tabref{tab:run-time-comparison}.  The first column shows the run-time
-on a Intel Core2 Duo (P8600, $2.40\unit{GHz}$) where only one core was
-enabled.  The second column shows the run-time on the same processor
-where both cores were enabled.  In the third column a machine with two
-Intel Xeon CPUs (L5640, $2.27\unit{GHz}$, $6$ cores) was used.
-%
-\begin{table}[tbh]
-  \centering
-  \begin{tabular}{llll}
-    \toprule
-                            & Intel Core2 Duo    & Intel Core2 Duo   & $2$ $\times$ Intel Xeon\\
-                            & (P8600, $1$ core)  & (P8600, $2$ cores)& (L5640, $6$ cores)\\
-    \midrule
-    FlexibleSUSY-NoFV 1.0.0 & $0.086\unit{s}$    & $0.079\unit{s}$   & $0.060\unit{s}$\\
-    SPheno 3.2.4            & $0.119\unit{s}$    & $0.114\unit{s}$   & $0.101\unit{s}$\\
-    Softsusy 3.4.0          & $0.175\unit{s}$    & $0.171\unit{s}$   & $0.147\unit{s}$\\
-    \midrule
-    FlexibleSUSY-FV 1.0.0   & $0.150\unit{s}$    & $0.113\unit{s}$   & $0.074\unit{s}$\\
-    SPhenoMSSM 4.1.0        & $0.415\unit{s}$    & $0.401\unit{s}$   & $0.370\unit{s}$\\
-    \bottomrule
-  \end{tabular}
-  \caption{Average run-time of CMSSM spectrum generators
-    for random parameter points.  The first three rows show three
-    spectrum generators with disabled sfermion flavour violation.  Rows
-    $4$--$5$ show two spectrum generators with enabled sfermion flavour violation, both based
-    on SARAH's MSSM model file.}
-  \label{tab:run-time-comparison}
-\end{table}
-
-Under both the non-flavour violating spectrum generators (first three
-rows) as well as the flavour violating ones (4th and 5th row) we find
-that \fs is significantly fastest.  Compared to SPheno,
-FlexibleSUSY-NoFV is faster by a factor $1.4$--$1.7$, and compared to
-Softsusy around a factor $2$--$2.5$.  Under the flavour violating
-spectrum generators FlexibleSUSY-FV is faster than SPhenoMSSM by a
-factor $2.8$--$5$.  Reason for the long run-time of SPhenoMSSM is the
-long calculation duration of the two-loop $\beta$-functions.  Here \fs
-benefits a lot from Eigen's well-optimizable matrix expressions.  We
-also find that increasing the number of CPU cores reduces the run-time
-of \fs.  The reason is that \fs calculates each pole mass in a
-separate thread, and therefore benefits from multi-core CPUs.
-
-\section{Conclusions}
-
-We have presented \fs, a Mathematica and C++ package, which generates fast and
-modular spectrum generators for any user specified SUSY model.  \fs is
-distributed with a large number of predefined models for the CMSSM,
-NMSSM, USSM, \ESSM, MRSSM etc., which can be generated immediately
-without any editing.  In particular the CMSSM and NMSSM spectrum
-generators constitute a fast and reliable alternative to the existing
-publicly available spectrum generators, Softsusy, SPheno and
-NMSPEC.
-
-We have described how the generated source code can be influenced at
-two different levels: The \mathematica level where the user provides a
-model file, and the C++ level where the generated objects can be
-easily exchanged, extended, modified and reused.  This provides great
-flexibility for creating custom spectrum generators for both the most
-common and most extraordinary models.  We have demonstrated these
-features in detailed examples for the NUHMSSM, right handed neutrinos
-and on adding three-loop RGEs and two-loop matching for the strong
-gauge coupling.
-
-The generated code has been extensively tested against Softsusy, and
-additional tests have been carried out for non-minimal models, the
-\ESSM and MRSSM.  Speed tests have also been performed against
-Softsusy, SPheno and SPheno-like MSSM code generated by SARAH,
-demonstrating that \fs runs faster than all three.
-
-As a result \fs enables fast exploitation of new SUSY models with high
-precision and reliability.
-
-\section*{Acknowledgments}
-
-A.V.\ would like to thank Florian Staub for countless explanations of
-SARAH's internals, discussions, and exceptionally fast bug
-fixing. P.A.~would like to thank Roman Nevzorov for useful discussions
-about challenges in non-minimal SUSY models and the GAMBIT
-collaboration for helpful suggestions, particularly regarding
-lowering our compiler prerequisites. The authors would also like to
-thank Lewis Tunstall for helping with early tests against
-Next-to-Minimal Softsusy; Sophie Underwood for discovering problems
-when introducing couplings with a single family index; Gregor Hellwig
-for writing the first version of the E6SSM model file; Philip Diessner
-for further adaptions to the E6SSM model file and for supplying it to
-us and for for testing and identifying several bugs in \fs and in
-SARAH, in work on the MRSSM; Ulrik Günther for compilation tests on
-Mac OS X and Dylan Harries for spotting a bug in the configuration
-script.
-%
-J.P. acknowledges support from the MEC and FEDER (EC) Grants
-FPA2011--23596 and the Generalitat Valenciana under grant PROMETEOII/2013/017.
-This work has been supported by the German Research Foundation DFG
-through Grant No.~STO876/2-1.
-
-\appendix
-
-\section{Examples of C++ code adaptation}
-\label{sec:examples of c++ code adaptation}
-
-\subsection{Stacking models in a tower of effective theories}
+\subsubsection{Stacking models in a tower of effective theories}
 \label{sec:tower construction}
@@
 
-\subsection{Integrating custom-built C++ components}
+\subsubsection{Integrating custom-built C++ components}
 \label{sec:integrating-custom-built}
@@
 
+The procedure described above is essentially all that is needed
+to employ new spectrum generator components,
+which may be composed from scratch or through
+a linkage to external routines.
+In doing so,
+notice that there was an evident limit on the scope of
+modules which one had to deal with.
+For instance, it was clear from the outset
+that one does not have to go through the code of the central
+fixed-point iteration engine, \code{RGFlow}.
+This manifests the power of the clear separation among objects
+each with its well-defined distinct role.
+This is just like the fact that one does not need to access the internals
+of the \code{std::sort} function in the C++ Standard Library.
+It might be entertaining to complete the analogy by
+mapping the model objects in \code{RGFlow}
+to the elements that \code{std::sort} sorts
+and the boundary condition objects
+to the comparator function.
+
+\section{Tests and comparisons with other spectrum generators}
+\label{sec:comparison}
+
+\subsection{Numeric tests}
+
+To check the correctness of \fs's generated spectrum generators
+extensive unit testing against Softsusy's MSSM and NMSSM
+implementations (both $Z_3$-invariant and $Z_3$-violating variants)
+has been carried out.  These unit tests systematically compare all
+tree-level mass matrices, EWSB equations, one- and two-loop
+$\beta$-functions, one- and two-loop self-energies and one- and
+two-loop tadpoles numerically for various MSSM and NMSSM parameter
+points.  All tested expressions were found to agree within double
+machine precision.\footnote{Due to the systematic and detailed tests
+  several bugs in Softsusy, SARAH and \fs could be identified and
+  corrected.}  Furthermore, the overall pole mass spectrum after the
+full fixed-point iteration has finished has been compared, and was
+found to agree at the sub-permille level.  The origin of the
+sub-permille level difference between \fs and Softsusy is the
+different determination of the weak mixing angle
+$\theta_{W,\text{susy}}^{\text{\DRbar}}(M_Z)$ in the SUSY model in the
+\DRbar\ scheme: \fs calculates
+$\theta_{W,\text{susy}}^{\text{\DRbar}}(M_Z)$ from $M_W$ and $M_Z$, as
+described in \secref{sec:calculation-of-gauge-couplings}, while
+Softsusy determines $\theta_{W,\text{susy}}^{\text{\DRbar}}(M_Z)$ from
+the muon decay constant $G_\mu$.  The approach used in Softsusy
+results in more precise \DRbar\ gauge couplings at the $M_Z$ scale,
+because the muon decay constant is known with a higher accuracy than
+the $W$-boson mass.  Furthermore, in Softsusy $3.5.0$ some three-loop
+$\beta$-functions and two-loop threshold corrections can be enabled in
+the MSSM to increase the accuracy of the RG running and the
+determination of the \DRbar\ gauge and Yukawa couplings at $M_Z$
+\cite{Allanach:2014nba}.  These corrections are not implemented in \fs
+so far.  The complete set of unit tests is shipped with \fs and can be
+found in the \code{test/} directory.  The tests can be run with the
+command \code{make execute-tests}.  All unit tests are carried out
+nightly in order to continuously check the correctness of the meta
+code and the generated spectrum generators for the shipped models.
+The nightly test results can be found at
+\url{http://iktp.tu-dresden.de/~avoigt/FlexibleSUSY/test.xhtml}
+
+The \fs generated NUHM \ESSM\ spectrum generator has also been
+compared against a handwritten one for a constrained version of the
+\ESSM \cite{Athron:2009ue, Athron:2009bs, Athron:2012pw}.  The
+$\beta$-functions were systematically compared in unit tests and were
+found to match at the ??? level.  The handwritten code doesn't include
+full one-loop self-energies or tadpoles, so tests on these were not
+carried out.  Although the generators assume different constraints
+they could be compared by using the output of the CE$_6$SSM generator
+as an input to \fs and the spectra were found to be in reasonable
+agreement, given the different levels of precision.
+
+In addition \fs has already undergone some user testing. This includes
+analytic tests of the $R$-symmetric low-energy model (MRSSM) and
+alternative $E_6$-inspired SUSY scenarios.  The users who have helped us
+with this are thanked in the acknowledgements.
+
+We also compared the run-time of \fs against SPheno, Softsusy and the
+SARAH generated MSSM spectrum generator SPhenoMSSM.  The results of
+the comparison can be found in \secref{sec:run-time comparison}.
+
+\subsection{Run-time comparison}
+\label{sec:run-time comparison}
+
+One of \fs's design goals is a short run-time.  In this section we
+demonstrate that this goal was achieved by comparing the run-time of
+two different sets of CMSSM spectrum generators:
+%
+\begin{itemize}
+\item \emph{Without sfermion flavour violation:} Disallowing sfermion
+  flavour violation simplifies the calculation of the pole masses,
+  because flavour-off-diagonal sfermion self-energy matrix elements
+  don't need to be calculated.  Here we compare \fs's non-flavour
+  violating CMSSM spectrum generator FlexibleSUSY-NoFV (version 1.0.0)
+  against SPheno (version 3.2.4) and Softsusy (version 3.4.0).
+%
+\item \emph{With sfermion flavour violation:} Allowing for sfermion
+  flavour violation in general increases the run-time of spectrum
+  generators, because the full $6\times 6$ sfermion self-energy
+  matrices have to be calculated.  Here we compare FlexibleSUSY-FV
+  (version 1.0.0) and SPhenoMSSM (generated with SARAH 4.1.0 and
+  linked against SPheno 3.2.4).  Both spectrum generators are based on
+  SARAH's MSSM model file, which allows for sfermion flavour
+  violation.
+\end{itemize}
+%
+For the run-time comparison the CKM matrix is set to unity, all
+$CP$-violating phases are set to zero and $R$-parity violation is
+disabled.  \fs and Softsusy are compiled with g++ 4.8.0 and Intel
+ifort 13.1.3 20130607.  SPheno and SPhenoMSSM are compiled with Intel
+ifort 13.1.3 20130607.\footnote{Intel's ifort compiler decreases the
+  run-time of SPheno and SPhenoMSSM by approximately a factor $1.5$,
+  compared to gfortran.}  We're generating $2\cdot 10^{4}$ random
+CMSSM parameter points with $m_0\in [50,1000]\unit{GeV}$, $m_{1/2}\in
+[50,1000]\unit{GeV}$, $\tan\beta\in [1,100]$, $\sign\mu\in \{-1,+1\}$
+and $A_0\in [-1000,$ $1000]\unit{GeV}$.  For each point an SLHA input
+file is created by appending the values of $m_0$, $m_{1/2}$,
+$\tan\beta$, $\sign\mu$, $A_0$ in form of a \code{MINPAR} block to the
+SLHA template file given in \ref{sec:speed-test-slha-template-file}.
+The resulting SLHA input file is passed to each spectrum generator and
+the (wall-clock) time is measured until the program has finished.  The
+average run-times for three different CPU types can be found in
+\tabref{tab:run-time-comparison}.  The first column shows the run-time
+on a Intel Core2 Duo (P8600, $2.40\unit{GHz}$) where only one core was
+enabled.  The second column shows the run-time on the same processor
+where both cores were enabled.  In the third column a machine with two
+Intel Xeon CPUs (L5640, $2.27\unit{GHz}$, $6$ cores) was used.
+%
+\begin{table}[tbh]
+  \centering
+  \begin{tabular}{llll}
+    \toprule
+                            & Intel Core2 Duo    & Intel Core2 Duo   & $2$ $\times$ Intel Xeon\\
+                            & (P8600, $1$ core)  & (P8600, $2$ cores)& (L5640, $6$ cores)\\
+    \midrule
+    FlexibleSUSY-NoFV 1.0.0 & $0.086\unit{s}$    & $0.079\unit{s}$   & $0.060\unit{s}$\\
+    SPheno 3.2.4            & $0.119\unit{s}$    & $0.114\unit{s}$   & $0.101\unit{s}$\\
+    Softsusy 3.4.0          & $0.175\unit{s}$    & $0.171\unit{s}$   & $0.147\unit{s}$\\
+    \midrule
+    FlexibleSUSY-FV 1.0.0   & $0.150\unit{s}$    & $0.113\unit{s}$   & $0.074\unit{s}$\\
+    SPhenoMSSM 4.1.0        & $0.415\unit{s}$    & $0.401\unit{s}$   & $0.370\unit{s}$\\
+    \bottomrule
+  \end{tabular}
+  \caption{Average run-time of CMSSM spectrum generators
+    for random parameter points.  The first three rows show three
+    spectrum generators with disabled sfermion flavour violation.  Rows
+    $4$--$5$ show two spectrum generators with enabled sfermion flavour violation, both based
+    on SARAH's MSSM model file.}
+  \label{tab:run-time-comparison}
+\end{table}
+
+Under both the non-flavour violating spectrum generators (first three
+rows) as well as the flavour violating ones (4th and 5th row) we find
+that \fs is significantly fastest.  Compared to SPheno,
+FlexibleSUSY-NoFV is faster by a factor $1.4$--$1.7$, and compared to
+Softsusy around a factor $2$--$2.5$.  Under the flavour violating
+spectrum generators FlexibleSUSY-FV is faster than SPhenoMSSM by a
+factor $2.8$--$5$.  Reason for the long run-time of SPhenoMSSM is the
+long calculation duration of the two-loop $\beta$-functions.  Here \fs
+benefits a lot from Eigen's well-optimizable matrix expressions.  We
+also find that increasing the number of CPU cores reduces the run-time
+of \fs.  The reason is that \fs calculates each pole mass in a
+separate thread, and therefore benefits from multi-core CPUs.
+
+\section{Conclusions}
+
+We have presented \fs, a Mathematica and C++ package, which generates fast and
+modular spectrum generators for any user specified SUSY model.  \fs is
+distributed with a large number of predefined models for the CMSSM,
+NMSSM, USSM, \ESSM, MRSSM etc., which can be generated immediately
+without any editing.  In particular the CMSSM and NMSSM spectrum
+generators constitute a fast and reliable alternative to the existing
+publicly available spectrum generators, Softsusy, SPheno and
+NMSPEC.
+
+We have described how the generated source code can be influenced at
+two different levels: The \mathematica level where the user provides a
+model file, and the C++ level where the generated objects can be
+easily exchanged, extended, modified and reused.  This provides great
+flexibility for creating custom spectrum generators for both the most
+common and most extraordinary models.  We have demonstrated these
+features in detailed examples for the NUHMSSM, right handed neutrinos
+and on adding three-loop RGEs and two-loop matching for the strong
+gauge coupling.
+
+The generated code has been extensively tested against Softsusy, and
+additional tests have been carried out for non-minimal models, the
+\ESSM and MRSSM.  Speed tests have also been performed against
+Softsusy, SPheno and SPheno-like MSSM code generated by SARAH,
+demonstrating that \fs runs faster than all three.
+
+As a result \fs enables fast exploitation of new SUSY models with high
+precision and reliability.
+
+\section*{Acknowledgments}
+
+A.V.\ would like to thank Florian Staub for countless explanations of
+SARAH's internals, discussions, and exceptionally fast bug
+fixing. P.A.~would like to thank Roman Nevzorov for useful discussions
+about challenges in non-minimal SUSY models and the GAMBIT
+collaboration for helpful suggestions, particularly regarding
+lowering our compiler prerequisites. The authors would also like to
+thank Lewis Tunstall for helping with early tests against
+Next-to-Minimal Softsusy; Sophie Underwood for discovering problems
+when introducing couplings with a single family index; Gregor Hellwig
+for writing the first version of the E6SSM model file; Philip Diessner
+for further adaptions to the E6SSM model file and for supplying it to
+us and for for testing and identifying several bugs in \fs and in
+SARAH, in work on the MRSSM; Ulrik Günther for compilation tests on
+Mac OS X and Dylan Harries for spotting a bug in the configuration
+script.
+%
+J.P. acknowledges support from the MEC and FEDER (EC) Grants
+FPA2011--23596 and the Generalitat Valenciana under grant PROMETEOII/2013/017.
+This work has been supported by the German Research Foundation DFG
+through Grant No.~STO876/2-1.
+
+\appendix
+
 \section{Speed test SLHA input file}

### 2014-11-09 test/test_MSSM_model.cpp
increasing MSSM one- and two-loop EWSB test precision

@@ -1199,6 +1199,8 @@ void test_ewsb_tree(MSSM<Two_scale> model, MssmSoftsusy softSusy)
    softSusy.rewsbTreeLevel(signMu);
    TEST_CLOSE(softSusy.displayM3Squared(), model.get_BMu(), 5.0);
    TEST_CLOSE(softSusy.displaySusyMu(), model.get_Mu(), 0.1);
+   TEST_CLOSE_REL(softSusy.displayM3Squared(), model.get_BMu(), 0.0004);
+   TEST_CLOSE_REL(softSusy.displaySusyMu(), model.get_Mu(), 0.0002);
 }
 
 void test_ewsb_1loop(MSSM<Two_scale> model, MssmSoftsusy softSusy)
@@ -1230,8 +1232,9 @@ void test_ewsb_1loop(MSSM<Two_scale> model, MssmSoftsusy softSusy)
 
    softsusy::numRewsbLoops = 1;
    softSusy.rewsb(signMu, softSusy.displayDrBarPars().mt, pars);
-   TEST_CLOSE_REL(softSusy.displayM3Squared(), model.get_BMu(), 0.0004);
-   TEST_CLOSE(softSusy.displaySusyMu(), model.get_Mu(), 0.1);
+   TEST_CLOSE_REL(softSusy.displayM3Squared(), model.get_BMu(), 1.0e-10);
+   TEST_CLOSE_REL(softSusy.displaySusyMu(), model.get_Mu(), 1.0e-10);
+   TEST_CLOSE(softSusy.displaySusyMu(), model.get_Mu(), 1.0e-10);
 }
 
 void test_ewsb_2loop(MSSM<Two_scale> model, MssmSoftsusy softSusy)
@@ -1271,6 +1274,7 @@ void test_ewsb_2loop(MSSM<Two_scale> model, MssmSoftsusy softSusy)
    softsusy::numRewsbLoops = 2;
    softSusy.rewsb(signMu, softSusy.displayDrBarPars().mt, pars);
    TEST_CLOSE_REL(softSusy.displayM3Squared(), model.get_BMu(), 1.0e-9);
+   TEST_CLOSE_REL(softSusy.displaySusyMu(), model.get_Mu(), 1.0e-10);
    TEST_CLOSE(softSusy.displaySusyMu(), model.get_Mu(), 1.0e-10);
 }
 

### 2014-11-11 doc/flexiblesusy-paper.tex
adding table with unit tested (N)MSSM parameter points
I also added a note about the tests of the EWSB iterations.
Update description of MSSMD5O + MSSMRHN tower in paper

--- 2014-11-06 doc/flexiblesusy-paper.tex
@@
 \subsubsection{Extending existing models}
+\label{sec:extending existing models}
 
@@
 those with additional gauge-singlet fields.
-In what follows, a supersymmetric type-I see-saw model
+In what follows, a supersymmetric type-I see-saw scenario
 \cite{see-saw} 
-shall be considered in which
-three neutral (heavy) chiral superfields are introduced.
+shall be considered.
+For this, two extensions of the MSSM are introduced:
+MSSMRHN with three extra neutral (heavy) chiral superfields,
+and MSSMD5O with the dimension-5 neutrino mass operator
+added to the superpotential.
+Both models are included in the package.
 
 % To create a new model, one should name it.
-In the package, this model is named \code{MSSMRHN}, standing for
+The name MSSMRHN of the first model stands for
 the MSSM plus right-handed neutrinos.
@@
 HighScaleInput = {
-  {Mv, LHInput[Mv]}, {B[Mv], LHInput[B[Mv]]},
-  {Yv, LHInput[Yv]}, {T[Yv], Azero*Yv},
   {mv2, UNITMATRIX[3] m0^2},
+  {T[Yv], Azero*Yv},
+  {B[Mv], LHInput[B[Mv]]},
   ...
@@
 \end{numlstlisting}
-In this particular example,
-the new parameters shown in lines 2--3
-are constrained at the $M_X$ scale
-to the values given in the SLHA input file.
-In a data-driven approach,
-they might alternatively be fixed from a matching against
-a low-energy theory containing the dimension-5
-neutrino mass operator in conjunction with supplementary constraints.
+
+The second model is called MSSMD5O, standing for
+the MSSM including the dimension-5 operator.
+Obviously, one can compose it by adding
+the additional term to the superpotential in \code{MSSMD5O.m}:
+\begin{numlstlisting}
+SuperPotential = Yu u.q.Hu - Yd d.q.Hd - Ye e.l.Hd + \[Mu] Hu.Hd \
+               + WOp/2 l.Hu.l.Hu;
+\end{numlstlisting}
+where line 2 contains the dimension-5 operator
+multiplied by its coeffficient matrix \code{WOp}.
+The declarations of the neutrino Dirac spinors and mixing are
+very similar to those in the MSSMRHN,
+except that \code{FvR} is absent.
+
+One can specify the low-scale constraints on \code{WOp}
+in \code{model_files/MSSMRHN/FlexibleSUSY.m.in}:
+\begin{numlstlisting}
+EXTPAR = {
+  {1, mv1}, {2, mv2}, {3, mv3},
+  {4, ThetaV12}, {5, ThetaV13}, {6, ThetaV23},
+  {7, YvDiag1}, {8, YvDiag2}, {9, YvDiag3}
+};
+
+UPMNS = Module[{
+    s12 = Sin @ ThetaV12, c12 = Cos @ ThetaV12,
+    s13 = Sin @ ThetaV13, c13 = Cos @ ThetaV13,
+    s23 = Sin @ ThetaV23, c23 = Cos @ ThetaV23
+  },
+  {{  c12 c13              ,  s12 c13              , s13     },
+   { -s12 c23 - c12 s23 s13,  c12 c23 - s12 s23 s13, s23 c13 },
+   {  s12 s23 - c12 c23 s13, -c12 s23 - s12 c23 s13, c23 c13 }}
+];
+
+mv = conj[UPMNS].DiagonalMatrix[{mv1, mv2, mv3}].Transpose[conj @ UPMNS];
+
+LowScaleInput = Join[
+  { (* MSSM low-scale constraints *) },
+  Flatten[Table[{WOp[i,j], mv[[i,j]] / (vu/Sqrt[2])^2}, {i,3}, {j,3}], 1]
+];
+
+InitialGuessAtLowScale = Join[
+  { (* MSSM initial guesses at low scale *) },
+  Flatten[Table[{WOp[i,j], mv[[i,j]] / (vu/Sqrt[2])^2}, {i,3}, {j,3}], 1]
+];
+\end{numlstlisting}
+The \code{EXTPAR} list contains the input parameters
+to be read from the corresponding SLHA block.
+The low-energy neutrino mass eigenvalues and mixing angles are
+declared in lines 2--3.
+They are followed by the neutrino Yukawa eigenvalues,
+which shall be used as part of the matching condition
+described in \ref{sec:tower construction}.
+The constraint on and the initial guess of \code{WOp}
+in lines 21 and 26 should be self-explanatory.
 
 With the above set of input files,
-\fs can generate the C++ code and compile it
-to produce \code{libMSSMRHN}.
-These products shall be employed as the implementation of the theory that
-underlies the MSSM in the next subsection.
+\fs can generate the C++ class libraries,
+\code{libMSSMRHN} and \code{libMSSMD5O}\@.
+These products shall be employed as the two effective theories
+in the implementation of the see-saw mechanism.
+To this end, one further needs to code at the C++ level,
+as explained in the next subsection and \ref{sec:tower construction}.
 
@@
 of a spectrum generator, set out in \secref{sec:SpecGenStruct}.
-In what follows,
-it shall be demonstrated that
+In \ref{sec:examples of c++ code adaptation},
+two examples are presented for demonstrating that
 the clean class structure serves as firm guidance on the job.
 
-The procedure described above is essentially all that is needed
-to employ new spectrum generator components,
+The first project in \ref{sec:tower construction} is
+to build a spectrum generator that can handle
+a tower of multiple effective field theories.
+The aim is to take a first step towards
+a study of slepton-mediated lepton flavour violation
+due to radiative corrections in the type-I supersymmetric see-saw model
+\cite{Borzumati:1986qx}.
+To this end, \code{MSSMRHN} is stacked on top of \code{MSSMD5O}.
+The preparation of these two models has been covered in
+\secref{sec:extending existing models}.
+
+The spectrum generator shall accept
+the low-energy neutrino masses and mixing angles
+which determine \code{WOp}, the coefficients
+of the $L H_u L H_u$ operator.
+These $6 \times 6$ coefficients evolve to
+the right-handed neutrino mass scale at which
+they are matched to the neutrino Yukawa couplings \code{Yv} and
+the right-handed neutrino masses \code{Mv}.
+Since there are more degrees of freedom in the pair of
+\code{Yv} and \code{Mv} than in \code{WOp},
+one needs supplementary conditions in addition to the see-saw relation.
+In the presented matching code, it is assumed that all mixing in \code{WOp}
+stems from the left-handed rotation of \code{Yv} whose eigenvalues are
+fixed to those specified by the user.
+The non-trivial flavour structure of the neutrino Yukawa couplings then
+causes running slepton mass matrices to acquire flavour-violating elements.
+The output from the spectrum generator includes
+the slepton mass matrices as well as
+the resulting mass eigenvalues and mixing.
+One might pass this outcome on to another routine
+to calculate rates of lepton flavour violating processes.
+
+For brevity,
+threshold corrections are ignored
+in the specimen matching code connecting the two models
+as well as in the low-scale boundary condition on \code{WOp}
+from the neutrino oscillation data.
+Therefore, the result maintains only the accuracy of one-loop RGEs,
+even though two-loop $\beta$-functions are computed by each model class.
+For a full accuracy of two-loop RGEs, one can incorporate the omitted
+one-loop corrections into the constraint classes.
+The way to implement them should be self-evident from the code structure.
+
+The second project in \ref{sec:integrating-custom-built} shows
+% is essentially all that is needed
+how one can employ new spectrum generator components,
 which may be composed from scratch or through
 a linkage to external routines.
-In doing so,
-notice that there was an evident limit on the scope of
-modules which one had to deal with.
-For instance, it was clear from the outset
+In the procedure,
+it would be noticed that there is an evident limit on the scope of
+modules which one has to deal with.
+For instance, it is clear from the outset
 that one does not have to go through the code of the central
@@
 
+In what follows, technical details of \fs programming at the C++ level are
+set out which supplement the outline given
+in \secref{sec:adapting-cpp-code}.
+
 \subsection{Stacking models in a tower of effective theories}
@@
 using a well-known configuration in which
-the higher-energy theory is the MSSMRHN discussed above which
-gives rise to the MSSM as the lower-energy effective theory.
+the higher-energy theory is the MSSMRHN which
+gives rise to the MSSMD5O as the lower-energy effective theory.
 The relevant classes are sketched
@@
   \draw (MZ) +(-.1,0) -- +(+.1,0); \draw[<-] (MZ) ++(-.8,0) -- ++(-.5,0)
-     node[left,draw] {\code{MSSM_low_scale_constraint}};
+     node[left,draw] {\code{MSSMD5O_low_scale_constraint}};
   \draw (MS) +(-.1,0) -- +(+.1,0); \draw[<-] (MS) ++(-.8,0) -- ++(-.5,0)
-     node[left,draw] {\code{MSSM_susy_scale_constraint}};
+     node[left,draw] {\code{MSSMD5O_susy_scale_constraint}};
   \draw (Mv) +(-.1,0) -- +(+.1,0); \draw[<-] (Mv) ++(-.8,0) -- ++(-.5,0)
-     node[left,draw] {\code{MSSM_MSSMRHN_matching}};
+     node[left,draw] {\code{MSSMD5O_MSSMRHN_matching}};
   \draw (MX) +(-.1,0) -- +(+.1,0); \draw[<-] (MX) ++(-.8,0) -- ++(-.5,0)
@@
 
-  \draw (MvRR) +(0,-.03) rectangle (MZR) node[pos=.5] {\code{MSSM}};
+  \draw (MvRR) +(0,-.03) rectangle (MZR) node[pos=.5] {\code{MSSMD5O}};
   \draw (MvRR) +(0, .03) rectangle (MXR) node[pos=.5] {\code{MSSMRHN}};
@@
 the $M_\nu$ scale at which the right-handed neutrinos are decoupled.
-Below this scale, the \code{MSSM} object takes over.
+Below this scale, the \code{MSSMD5O} object takes over.
 On the left of the vertical axis,
@@
 \begin{numlstlisting}
-$ ./createmodel --name=MSSM
+$ ./createmodel --name=MSSMD5O
 $ ./createmodel --name=MSSMRHN
-$ ./configure --with-models=MSSM,MSSMRHN
+$ ./configure --with-models=MSSMD5O,MSSMRHN
 $ make
@@
 CPPFLAGS  := -I. $(INCCONFIG) $(INCFLEXI) $(INCLEGACY) $(INCSLHAEA) \
-             $(INCMSSM) $(INCMSSMRHN)
+             $(INCMSSMD5O) $(INCMSSMRHN)
 
 TOWER_SRC := run_tower.cpp \
-	     MSSM_MSSMRHN_two_scale_matching.cpp \
-	     MSSM_MSSMRHN_two_scale_initial_guesser.cpp
+	     MSSMD5O_MSSMRHN_two_scale_matching.cpp \
+	     MSSMD5O_MSSMRHN_two_scale_initial_guesser.cpp
 
@@
 
-run_tower.x: $(TOWER_OBJ) $(LIBMSSM) $(LIBMSSMRHN) $(LIBFLEXI) $(LIBLEGACY)
-  $(CXX) -o $@ $^ $(LOOPFUNCLIBS) $(GSLLIBS) $(BOOSTTHREADLIBS) $(THREADLIBS) $(LAPACKLIBS) $(FLIBS)
-\end{numlstlisting}
+run_tower.x: $(TOWER_OBJ) $(LIBMSSMD5O) $(LIBMSSMRHN) $(LIBFLEXI) $(LIBLEGACY)
+  $(CXX) -o $@ $^ $(LOOPFUNCLIBS) $(GSLLIBS) $(BOOSTTHREADLIBS) $(THREADLIBS) $(LAPACKLIBS) $(BLASLIBS) $(FLIBS)
+\end{numlstlisting}%% $
 The include directives in line 2 tell the compiler
-where to find the headers for either \code{MSSM} or \code{MSSMRHN}.
+where to find the headers for either \code{MSSMD5O} or \code{MSSMRHN}.
 The \code{.cpp} files in lines 4--6 and the \code{.hpp} files
@@
 Obviously, the executable \code{run_tower.x}, in line 10, depends on both
-\code{$(LIBMSSM)} and \code{$(LIBMSSMRHN)}
+\code{$(LIBMSSMD5O)} and \code{$(LIBMSSMRHN)}
 that implement the auto-generated
@@
 one can extend % make a straightforward extension to
-\code{run_MSSM.cpp} or \code{run_MSSMRHN.cpp}
+\code{run_MSSMD5O.cpp} or \code{run_MSSMRHN.cpp}
 produced in either model directory.
@@
 \begin{numlstlisting}
-#include "MSSM_MSSMRHN_spectrum_generator.hpp"
+#include "MSSMD5O_MSSMRHN_spectrum_generator.hpp"
 
@@
   QedQcd oneset;
-  MSSM_input_parameters    input_1;
+  MSSMD5O_input_parameters input_1;
   MSSMRHN_input_parameters input_2;
@@
   typedef Two_scale algorithm_type;
-  MSSM_MSSMRHN_spectrum_generator<algorithm_type> spectrum_generator;
+  MSSMD5O_MSSMRHN_spectrum_generator<algorithm_type> spectrum_generator;
   // set up spectrum_generator;
@@
 The crucial point is the definition of the
-\code{MSSM_MSSMRHN_spectrum_generator} object in line 12, which
+\code{MSSMD5O_MSSMRHN_spectrum_generator} object in line 12, which
 creates and drives the \code{RGFlow} object in \figref{fig:class
@@
 \code{run()} member function in line 14.  It is defined in
-\code{MSSM_MSSMRHN_spectrum_generator.hpp} and reads:
+\code{MSSMD5O_MSSMRHN_spectrum_generator.hpp} and reads:
 %
 \begin{numlstlisting}[name=SGrun,language=C++]
-template<class T> void MSSM_MSSMRHN_spectrum_generator<T>::run
+template<class T> void MSSMD5O_MSSMRHN_spectrum_generator<T>::run
 (const QedQcd& oneset,
- const MSSM_input_parameters& input_1, const MSSMRHN_input_parameters& input_2)
+ const MSSMD5O_input_parameters& input_1, const MSSMRHN_input_parameters& input_2)
 {
   high_scale_constraint_2.clear(); // of type MSSMRHN_high_scale_constraint<T>
-  susy_scale_constraint_1.clear(); // of type MSSM_susy_scale_constraint<T>
-  low_scale_constraint_1 .clear(); // of type MSSM_low_scale_constraint<T>
-  matching.reset();                // of type MSSM_MSSMRHN_matching<T>
+  susy_scale_constraint_1.clear(); // of type MSSMD5O_susy_scale_constraint<T>
+  low_scale_constraint_1 .clear(); // of type MSSMD5O_low_scale_constraint<T>
+  matching.reset();                // of type MSSMD5O_MSSMRHN_matching<T>
   high_scale_constraint_2.set_input_parameters(input_2);
@@
   low_scale_constraint_1 .set_input_parameters(input_1);
-  matching.set_upper_input_parameters(input_2);
+  matching.set_lower_input_parameters(input_1);
   high_scale_constraint_2.initialize();
@@
 This piece of code is nearly a verbatim copy of
-the corresponding part of \code{MSSM_spectrum_generator.hpp}.
+the corresponding part of \code{MSSMD5O_spectrum_generator.hpp}.
 The only differences are that the type of
@@
 One then constructs a list of
-the constraints on \code{MSSM}:
+the constraints on \code{MSSMD5O}:
 \begin{numlstlisting}[name=SGrun]
@@
 \end{numlstlisting}
-and initializes the \code{MSSM} object:
+and initializes the \code{MSSMD5O} object:
 \begin{numlstlisting}[name=SGrun,language=C++]
-  model_1.clear();                 // of type MSSM<T>
+  model_1.clear();                 // of type MSSMD5O<T>
   model_1.set_input_parameters(input_1);
@@
 out of auto-generated
-\code{MSSM_convergence_tester} and \code{MSSMRHN_convergence_tester}:
+\code{MSSMD5O_convergence_tester} and \code{MSSMRHN_convergence_tester}:
 \begin{numlstlisting}[name=SGrun]
-  MSSM_convergence_tester<T> convergence_tester_1(&model_1, precision_goal);
+  MSSMD5O_convergence_tester<T> convergence_tester_1(&model_1, precision_goal);
   MSSMRHN_convergence_tester<T> convergence_tester_2(&model_2, precision_goal);
@@
 \begin{numlstlisting}[name=SGrun]
-  MSSM_MSSMRHN_initial_guesser<T> initial_guesser
-    (&model_1, &model_2, input_2, oneset,
+  MSSMD5O_MSSMRHN_initial_guesser<T> initial_guesser
+    (&model_1, &model_2, input_1, oneset,
      low_scale_constraint_1, susy_scale_constraint_1, high_scale_constraint_2,
@@
   susy_scale_1 = susy_scale_constraint_1.get_scale();
-  model_1.run_to(susy_scale_1);    // of type MSSM<T>
+  model_1.run_to(susy_scale_1);    // of type MSSMD5O<T>
   model_1.calculate_spectrum();
@@
 In the present example, the class is declared in the header
-\code{MSSM_MSSMRHN_two_scale_matching.hpp}:
+\code{MSSMD5O_MSSMRHN_two_scale_matching.hpp}:
 \begin{numlstlisting}
-template<> class MSSM_MSSMRHN_matching<Two_scale> : public Matching<Two_scale> {
+template<> class MSSMD5O_MSSMRHN_matching<Two_scale> : public Matching<Two_scale> {
 public:
-  MSSM_MSSMRHN_matching();
-  MSSM_MSSMRHN_matching(const MSSMRHN_input_parameters&);
+  MSSMD5O_MSSMRHN_matching();
+  MSSMD5O_MSSMRHN_matching(const MSSMD5O_input_parameters&);
   void match_low_to_high_scale_model();
@@
   double get_initial_scale_guess() const;
-  void set_upper_input_parameters(const MSSMRHN_input_parameters&);
+  void set_lower_input_parameters(const MSSMD5O_input_parameters&);
   void set_scale(double);
@@
 private:
-  MSSM   <Two_scale> *lower;
+  MSSMD5O<Two_scale> *lower;
   MSSMRHN<Two_scale> *upper;
@@
 As lines 4 and 10 indicate,
-this class takes an \code{MSSMRHN_input_parameters} object as input.
-The \code{MvInput} field thereof is referenced
-for the initial guess of the matching scale
-in line 25 of \code{MSSM_MSSMRHN_two_scale_matching.cpp}:
+this class takes an \code{MSSMD5O_input_parameters} object as input.
+The low-energy neutrino data therein is referenced by
+\code{make_initial_scale_guess}
+starting from line 23 of \code{MSSMD5O_MSSMRHN_two_scale_matching.cpp}:
 \begin{numlstlisting}
-void MSSM_MSSMRHN_matching<Two_scale>::match_low_to_high_scale_model()
+void MSSMD5O_MSSMRHN_matching<Two_scale>::invert_seesaw_formula
+(const Eigen::Matrix3d& WOp, const Eigen::Vector3d& YvDiag,
+ Eigen::Matrix3d& Yv, Eigen::Matrix3d& Mv)
 {
+  Eigen::Matrix3cd uh;
+  Eigen::Array3d s;
+  fs_diagonalize_symmetric(WOp, s, uh);
+  Eigen::Matrix3d U = uh.adjoint().real();
+  Eigen::Vector3d YvDiagInv(1, 1, 1);
+  YvDiagInv.array() /= YvDiag.array();
+  Eigen::Matrix3d YvInv = U * YvDiagInv.asDiagonal();
+  Mv = (YvInv.transpose() * WOp * YvInv).inverse();
+  Yv = YvDiag.asDiagonal() * U.adjoint();
+}
+
+void MSSMD5O_MSSMRHN_matching<Two_scale>::set_lower_input_parameters
+(const MSSMD5O_input_parameters& inputPars_)
+{
+  inputPars = inputPars_;
+  make_initial_scale_guess();
+}
+
+void MSSMD5O_MSSMRHN_matching<Two_scale>::make_initial_scale_guess()
+{
+  Eigen::Matrix3d WOp;
+  // fill WOp with elements in terms of data from inputPars;
+  Eigen::Vector3d YvDiag;
+  YvDiag << inputPars.YvDiag1, inputPars.YvDiag2, inputPars.YvDiag3;
+  Eigen::Matrix3d Yv;
+  Eigen::Matrix3d Mv;
+  invert_seesaw_formula(WOp, YvDiag, Yv, Mv);
+  double RHN_scale = pow(abs(Mv.determinant()), 1.0/3);
+  scale = initial_scale_guess = RHN_scale;
+}
+\end{numlstlisting}
+To guess the matching scale, this function estimates
+the right-handed neutrino mass matrix \code{Mv}
+from \code{WOp}, the dimension-5 operator coefficients,
+and \code{YvDiag}, the neutrino Yukawa eigenvalues,
+by calling \code{invert_seesaw_formula}.
+Since the mapping,
+$\mbox{\code{WOp}} \rightarrow (\mbox{\code{Yv}}, \mbox{\code{Mv}})$,
+is not unique, \code{invert_seesaw_formula}
+opts to impose the following additional constraints:
+the eigenvalues of \code{Yv} are fixed to the user input (line 10),
+and all mixing in \code{WOp} originates from the rotation of the SU(2)
+doublets (line 11).
+
+The actual matching process takes place in the two functions
+\code{match_low_to_high_scale_model} and \code{match_high_to_low_scale_model}:
+\begin{numlstlisting}
+void MSSMD5O_MSSMRHN_matching<Two_scale>::match_low_to_high_scale_model()
+{
+  Eigen::Vector3d YvDiag;
+  YvDiag << inputPars.YvDiag1, inputPars.YvDiag2, inputPars.YvDiag3;
+  Eigen::Matrix3d Yv;
+  Eigen::Matrix3d Mv;
+  invert_seesaw_formula(lower->get_WOp(), YvDiag, Yv, Mv);
+  upper->set_Yv(Yv);
+  upper->set_Mv(Mv);
+
   upper->set_Yd(lower->get_Yd());
@@
 
-void MSSM_MSSMRHN_matching<Two_scale>::match_high_to_low_scale_model()
+void MSSMD5O_MSSMRHN_matching<Two_scale>::match_high_to_low_scale_model()
 {
   update_scale();
+
+  const auto& Yv = upper->get_Yv();
+  const auto& Mv = upper->get_Mv();
+  lower->set_WOp(Yv.transpose() * Mv.inverse() * Yv);
+
   lower->set_Yd(upper->get_Yd());
@@
 
-void MSSM_MSSMRHN_matching<Two_scale>::set_upper_input_parameters
-(const MSSMRHN_input_parameters& inputPars_)
-{
-  inputPars = inputPars_;
-  make_initial_scale_guess();
-}
-
-void MSSM_MSSMRHN_matching<Two_scale>::make_initial_scale_guess()
-{
-  double RHN_scale = pow(abs(inputPars.MvInput.determinant()), 1.0/3);
-  scale = initial_scale_guess = RHN_scale;
-}
-
-void MSSM_MSSMRHN_matching<Two_scale>::update_scale()
+void MSSMD5O_MSSMRHN_matching<Two_scale>::update_scale()
 {
@@
 \end{numlstlisting}
-Recall that \code{MvInput} is the high-scale boundary value of \code{Mv}.
-If one had opted for an alternative strategy to fix \code{Mv},
-then \code{MSSM_MSSMRHN_matching} might have required a different input.
-Subsequently, the matching scale is updated at each iteration
+For the low-to-high matching, \code{invert_seesaw_formula} is called in line 7,
+this time with \code{WOp} at the matching scale.
+The high-to-low matching function contains the well-known
+see-saw formula in line 22.
+The matching scale is updated at each iteration
 to be the geometric mean of the running \code{Mv} eigenvalues
 in lines 31--32.
-The actual matching process takes place in the two functions
-\code{match_low_to_high_scale_model} and \code{match_high_to_low_scale_model}.
-For brevity,
-their examples shown above simply copy each coupling
-upwards or downwards, neglecting threshold corrections.
-The way to incorporate these corrections should be self-evident
-from the code structure.
 
@@
 One can extend the already available
-\code{MSSM_initial_guesser} class. % in a straightforward manner.
+\code{MSSMD5O_initial_guesser} class. % in a straightforward manner.
 The essential task is done by the following member function:
-\begin{numlstlisting}[name=MSSM_MSSMRHN_two_scale_initial_guesser.cpp]
-void MSSM_MSSMRHN_initial_guesser<Two_scale>::guess()
+\begin{numlstlisting}[name=MSSMD5O_MSSMRHN_two_scale_initial_guesser.cpp]
+void MSSMD5O_MSSMRHN_initial_guesser<Two_scale>::guess()
 {
@@
 \end{numlstlisting}
-Compared to the MSSM case, the differences are that the type of
+Compared to the MSSMD5O case, the differences are that the type of
 \code{high_constraint_2} is \linebreak {\code{MSSMRHN_high_scale_constraint<Two_scale>} and that \code{matching_scale_guess} has been inserted.
 Due to this intermediate scale, the initial run-up is divided into two steps, with a matching procedure in-between:
-\begin{numlstlisting}[name=MSSM_MSSMRHN_two_scale_initial_guesser.cpp,language=C++]
-  model_1->run_to(matching_scale_guess); // of type MSSM<Two_scale>
+\begin{numlstlisting}[name=MSSMD5O_MSSMRHN_two_scale_initial_guesser.cpp,language=C++]
+  model_1->run_to(matching_scale_guess); // of type MSSMD5O<Two_scale>
   matching.set_models(model_1, model_2);
@@
 and the remaining undetermined parameters are guessed:
-\begin{numlstlisting}[name=MSSM_MSSMRHN_two_scale_initial_guesser.cpp]
+\begin{numlstlisting}[name=MSSMD5O_MSSMRHN_two_scale_initial_guesser.cpp]
   high_constraint_2.set_model(model_2);
@@
 The initial two-step run-down again involves a matching process:
-\begin{numlstlisting}[name=MSSM_MSSMRHN_two_scale_initial_guesser.cpp]
+\begin{numlstlisting}[name=MSSMD5O_MSSMRHN_two_scale_initial_guesser.cpp]
   model_2->run_to(matching_scale_guess);
@@
 \end{numlstlisting}
-At the low scale where \code{MSSM} is valid,
-the code is the same as in \code{MSSM_initial_guesser}:
-\begin{numlstlisting}[name=MSSM_MSSMRHN_two_scale_initial_guesser.cpp]
+At the low scale where \code{MSSMD5O} is valid,
+the code is the same as in \code{MSSMD5O_initial_guesser}:
+\begin{numlstlisting}[name=MSSMD5O_MSSMRHN_two_scale_initial_guesser.cpp]
   model_1->solve_ewsb_tree_level();
@@
 \begin{numlstlisting}
-Block YvIN                   # neutrino Yukawas at MX
-1 1  0.1568611               # Yv(1,1)
-1 2  0.6400513               # Yv(1,2)
-1 3  0.7521494               # Yv(1,3)
-2 1  0.4663838               # Yv(2,1)
-...                          # remaining 5 entries
-Block MvIN                   # heavy neutrino masses at MX
-1 1  4.150000E+14            # Mv(1,1)
-...                          # remaining 8 entries
+Block EXTPAR  # Input parameters
+    1	5.0E-11		     # mv1
+    2	5.07523E-11	     # mv2
+    3	6.96419E-11	     # mv3
+    4	0.586168	     # ThetaV12
+    5	0.157512	     # ThetaV13
+    6	0.705053	     # ThetaV23
+    7	0.6		     # YvDiag1
+    8	0.8		     # YvDiag2
+    9	1.0		     # YvDiag3
 Block BMvIN                  # right-handed sneutrino bilinear terms
@@
 \end{numlstlisting}
-The file in the package contains the values that approximately reproduce
-the observed neutrino masses and mixing angles
-\cite{Beringer:1900zz}
-through the see-saw mechanism.
+The file in the package contains the values % that approximately reproduce
+consistent with
+the observed neutrino mass-squared differences and mixing angles
+\cite{Beringer:1900zz}.
 
@@
 In the output,
-a part of significant physical interest might be:
+a part of the main interest is:
 \begin{numlstlisting}
-Block MSL2 Q=   8.97114431E+02
-  1  1     1.25652698E+05   # ml2(1,1)
-  1  2    -7.64196328E+01   # ml2(1,2)
-  1  3    -7.08429254E+01   # ml2(1,3)
-  2  1    -7.64196328E+01   # ml2(2,1)
-  2  2     1.25388732E+05   # ml2(2,2)
-  2  3    -3.15865242E+02   # ml2(2,3)
-  3  1    -7.08429254E+01   # ml2(3,1)
-  3  2    -3.15865242E+02   # ml2(3,2)
-  3  3     1.24556532E+05   # ml2(3,3)
+Block MSL2 Q= 8.82028104E+02
+  1  1     1.26231305E+05   # ml2(1,1)
+  1  2    -5.08365953E+02   # ml2(1,2)
+  1  3     4.67463327E+01   # ml2(1,3)
+  2  1    -5.08365953E+02   # ml2(2,1)
+  2  2     1.25328663E+05   # ml2(2,2)
+  2  3    -6.93488070E+02   # ml2(2,3)
+  3  1     4.67463327E+01   # ml2(3,1)
+  3  2    -6.93488070E+02   # ml2(3,2)
+  3  3     1.24229816E+05   # ml2(3,3)
 \end{numlstlisting}

### 2014-11-13 doc/flexiblesusy-paper.tex
Sort references in paper
clarify what is meant by "global sign" in the EWSB solutions

@@
 some cases, as for example in the real MSSM, the different solutions
-are related by one or more global signs as in
+are related by one or more global signs for the parameters, as for example in
 Eq.~\eqref{eq:EWSB-solution-for-mu}.  \fs recognizes such cases and
@@
 \code{TreeLevelEWSBSolution} variable in the model file.  This is for
-example the case in the $Z_3$-violating NMSSM (\code{SMSSM}), where
+example the case in the $Z_3$-violating NMSSM (\code{SMSSM}), which is shipped with \fs, where
 $\mu$ is chosen to be fixed by the EWSB equations: There the
@@
 %
-with a non-global $\sign X$.  When \fs solves the tree-level EWSB
+with $\sign X$ as free parameter.  When \fs solves the tree-level EWSB
 equations for $\mu$, $B\mu$ and $\xi_S$ it finds the two solutions and
@@
 
-The spectrum generator shall accept
+Since each model class has its own $\beta$-functions,
+the spectrum generator contains two different sets of RGEs
+that are connected by a matching object.
+The program shall accept
 the low-energy neutrino masses and mixing angles
@@
 
-\bibitem{3-loop MSSM betas}
-  I.~Jack and D.~R.~T.~Jones, \ptitle{
-    3-loop MSSM beta functions,}
-  \url{http://www.liv.ac.uk/~dij/betas}.
-
-\bibitem{Harlander:2005wm}
-  R.~Harlander, L.~Mihaila and M.~Steinhauser,
-  %``Two-loop matching coefficients for the strong coupling in the MSSM,''
-  Phys.\ Rev.\ D {\bf 72} (2005) 095009
-  [hep-ph/0509048].
-  %%CITATION = HEP-PH/0509048;%%
-  %27 citations counted in INSPIRE as of 05 Jun 2014
-
 \bibitem{Allanach:2014nba}
@@
 
+\bibitem{3-loop MSSM betas}
+  I.~Jack and D.~R.~T.~Jones, \ptitle{
+    3-loop MSSM beta functions,}
+  \url{http://www.liv.ac.uk/~dij/betas}.
+
+\bibitem{Harlander:2005wm}
+  R.~Harlander, L.~Mihaila and M.~Steinhauser,
+  %``Two-loop matching coefficients for the strong coupling in the MSSM,''
+  Phys.\ Rev.\ D {\bf 72} (2005) 095009
+  [hep-ph/0509048].
+  %%CITATION = HEP-PH/0509048;%%
+  %27 citations counted in INSPIRE as of 05 Jun 2014
+
 \end{thebibliography}

### 2014-11-14 doc/flexiblesusy-paper.tex
added more detail on E6SSM testing.

@@
 $\beta$-functions were systematically compared in unit tests and were
-found to match at the ??? level.  The handwritten code doesn't include
-full one-loop self-energies or tadpoles, so tests on these were not
-carried out.  Although the generators assume different constraints
-they could be compared by using the output of the CE$_6$SSM generator
-as an input to \fs and the spectra were found to be in reasonable
-agreement, given the different levels of precision.
+found to match within numerical precision.  The handwritten code
+doesn't include full one-loop self-energies or tadpoles, so tests on
+these were not carried out.  Although the generators assume different
+constraints and solve the boundary value problem with completely
+different algorithms they could be compared by using the output of
+the CE$_6$SSM generator as an input to \fs and the spectra were found
+to be in reasonable agreement ($\approx 10\%$), given the different
+levels of precision.
 

### 2014-11-14 doc/flexiblesusy-paper.tex
added goodsell dirac gaugino rges paper and added jack+jones paper and paper with them and martin etc explaining discrepancies.

@@
 $\beta$-functions of a general supersymmetric model
-\cite{Jones:1974pg,Jones:1983vk,West:1984dg,Martin:1993yx,Yamada:1993ga,MV94,Fonseca:2011vn,Yam94,Sperling:2013eva,Sperling:2013xqa}
+\cite{Jones:1974pg,Jones:1983vk,West:1984dg,Martin:1993yx,Yamada:1993ga,Jack:1994kd,Jack:1994rk,Yam94,MV94,Fonseca:2011vn,Goodsell:2012fm,Sperling:2013eva,Sperling:2013xqa}
 allows to split these parameters into two classes:
@@
 
+%\cite{Yam94}
+\bibitem{Yam94} 
+  Y.~Yamada,
+  %``Two loop renormalization group equations for soft SUSY breaking scalar interactions: Supergraph method,''
+  Phys.\ Rev.\ D {\bf 50}, 3537 (1994)
+  [hep-ph/9401241].
+  %%CITATION = HEP-PH/9401241;%%
+  %225 citations counted in INSPIRE as of 24 Sep 2013
+%\cite{}
+\bibitem{Jack:1994kd} 
+  I.~Jack and D.~R.~T.~Jones,
+  %``Soft supersymmetry breaking and finiteness,''
+  Phys.\ Lett.\ B {\bf 333}, 372 (1994)
+  [hep-ph/9405233].
+  %%CITATION = HEP-PH/9405233;%%
+  %195 citations counted in INSPIRE as of 14 Nov 2014
+
+%\cite{Jack:1994rk}
+\bibitem{Jack:1994rk} 
+  I.~Jack, D.~R.~T.~Jones, S.~P.~Martin, M.~T.~Vaughn and Y.~Yamada,
+  %``Decoupling of the epsilon scalar mass in softly broken supersymmetry,''
+  Phys.\ Rev.\ D {\bf 50}, 5481 (1994)
+  [hep-ph/9407291].
+  %%CITATION = HEP-PH/9407291;%%
+  %170 citations counted in INSPIRE as of 14 Nov 2014
+
 %\cite{Fonseca:2011vn}
@@
 
-%\cite{Yam94}
-\bibitem{Yam94} 
-  Y.~Yamada,
-  %``Two loop renormalization group equations for soft SUSY breaking scalar interactions: Supergraph method,''
-  Phys.\ Rev.\ D {\bf 50}, 3537 (1994)
-  [hep-ph/9401241].
-  %%CITATION = HEP-PH/9401241;%%
-  %225 citations counted in INSPIRE as of 24 Sep 2013
+%\cite{Goodsell:2012fm}
+\bibitem{Goodsell:2012fm} 
+  M.~D.~Goodsell,
+  %``Two-loop RGEs with Dirac gaugino masses,''
+  JHEP {\bf 1301}, 066 (2013)
+  [arXiv:1206.6697 [hep-ph]].
+  %%CITATION = ARXIV:1206.6697;%%
+  %23 citations counted in INSPIRE as of 14 Nov 2014
 

### 2014-11-14 doc/flexiblesusy-paper.tex
describe the definition of the tested NMSSM models

@@
 $\beta$-functions, one- and two-loop self-energies and one- and
-two-loop tadpoles numerically for the CMSSM and NMSSM parameter points
-given in \tabref{tab:unit-test-parameter-points}.
+two-loop tadpoles numerically for the CMSSM, the semi-constrained
+$Z_3$-invariant NMSSM ($Z_3$-NMSSM)\footnote{With the semi-constrained
+  $Z_3$-invariant NMSSM ($Z_3$-NMSSM) we denote a constrained variant
+  of the NMSSM with universal gaugino masses $M_{1/2}$, universal
+  trilinear couplings $A_0$ and universal MSSM-like soft-breaking
+  squared scalar masses $m_0^2$ at the GUT scale.  The soft-breaking
+  singlet mass $m_S^2$, the trilinear singlet superpotential coupling
+  $\kappa$ and the singlet VEV $s$ are fixed by the EWSB conditions at
+  the SUSY scale.  The $Z_3$-NMSSM has the $6$ free parameters
+  $(m_0^2, M_{1/2}, A_0, \tan\beta, \sign\mu_\text{eff}, \lambda)$.}
+and the constrained $Z_3$-violating NMSSM
+($\Zv_3$-NMSSM)\footnote{With the constrained $Z_3$-violating NMSSM
+  ($\Zv_3$-NMSSM) we denote a constrained variant of the
+  $Z_3$-violating NMSSM with universal gaugino masses $M_{1/2}$,
+  universal trilinear couplings $A_0$ and universal MSSM-like
+  soft-breaking squared scalar masses $m_0^2$ at the GUT scale.  The
+  $\mu$-parameter, its soft-breaking equivalent $B\mu$ and the
+  soft-breaking singlet tadpole coupling $\xi_S$ are fixed by the EWSB
+  conditions at the SUSY scale.  The $\Zv_3$-NMSSM has the $11$ free
+  parameters $(m_0^2, M_{1/2}, A_0, \tan\beta, \sign\mu_\text{eff},
+  \lambda, \kappa, s, \mu', m_S^{\prime 2}, \xi_F)$.}  parameter
+points given in \tabref{tab:unit-test-parameter-points}.
 %
@@
   \end{tabularx}
-  \caption{CMSSM and NMSSM parameter points used for the unit tests
-    against Softsusy.
+  \caption{CMSSM, semi-constrained
+    $Z_3$-invariant NMSSM ($Z_3$-NMSSM) and constrained
+    $Z_3$-violating NMSSM ($\Zv_3$-NMSSM) parameter points
+    used for the unit tests against Softsusy.
     We follow the notation of the NMSSM model parameters used in

### 2014-11-18 doc/flexiblesusy-paper.tex
update my address
slightly reformulate SLHA compatibility with decay calulators

@@
 \author[dresden]{Dominik St\"ockinger}
-\author[dresden]{Alexander Voigt\corref{cor1}}
-\ead{Alexander.Voigt@physik.tu-dresden.de}
+\author[desy]{Alexander Voigt\corref{cor1}}
+\ead{Alexander.Voigt@desy.de}
 \cortext[cor1]{Corresponding author}
@@
 TU Dresden, Zellescher Weg 19, 01069 Dresden, Germany}
+\address[desy]{Deutsches Elektronen-Synchrotron (DESY), 22607 Hamburg, Germany}
 
@@
 
-The standard input and output of the generated code is given in
-SLHA-style format where the user has control of the Block names by
-editing the SARAH files and may also add extra blocks in the
-FlexibleSUSY.m model file which is decribed in section 5. We use
-SLHAea \cite{slhaea} to create these SLHA files and with a simple
-modifications these can be passed directly to another code which uses
-them as an alternative to writing out the SLHA file. Since the SLHA
-standard is intended precisely for the purpose, MSSM and NMSSM codes
-should be compliant with this.  We have tested that the versions of
-HDECAY\cite{hdecay} and SDECAY\cite{sdecay} inside SUSY-HIT
-\cite{susyhit} can understand the SLHA output of FlexibleSUSY and give
-matching output to that from SOFTSUSY.
+The standard input and output of the generated spectrum generator is
+the SLHA format \cite{Skands:2003cj,Allanach:2008qq}, which is
+intended for the communication between MSSM and NMSSM spectrum
+generators and observable calculators.  The user has control over the
+SLHA block names by editing the SARAH model files and may also add
+extra blocks in the \code{FlexibleSUSY.m} model file, as described in
+\secref{sec:modfile}.
+%
+% We use SLHAea \cite{slhaea} to create these SLHA files and with
+% simple modifications these can be passed directly to another code
+% which uses them as an alternative to writing out the SLHA file.
+%
+We have tested that the HDECAY $3.4$ \cite{hdecay} and SDECAY $1.4$
+\cite{sdecay}, which are shipped with SUSY-HIT \cite{susyhit},
+understand the SLHA output of \fs and give matching output to that
+from Softsusy.
 

### 2014-11-18 doc/flexiblesusy-paper.tex
slightly reformulate SLHA block control for non-(N)MSSM models

@@
 from Softsusy.
-
-For models which go beyond this there is no fixed standard for all
-names.  Nonetheless extending the SLHA framework and allowing user
-control of the new block names should greatly simplify the process of
-linking the output to a users own calculator or any new public tools
-that are developed for that model.
+%
+For models which go beyond the MSSM and NMSSM no fixed standard exists
+which specifies the spectrum generator input and output format.  For
+this reason, \fs allows the user to control the SLHA input and output
+blocks in order to simplify the process of passing the output of a
+custom \fs-generated spectrum generator or any new public tool
+developed for that model.
 

### 2014-11-18 doc/flexiblesusy-paper.tex
describe new feature ExtraSLHAOutputBlocks

@@
 LowPoleMassPrecision     = {};
+
+(* optional *)
+ExtraSLHAOutputBlocks = {
+   {ALPHA, {{ArcCos[Pole[ZH[1,2]]]}}},
+   {HMIX , {{1, \[Mu]},
+            {2, vu / vd},
+            {3, Sqrt[vu^2 + vd^2]},
+            {4, M[Ah[2]]^2},
+            {101, B[\[Mu]]},
+            {102, vd},
+            {103, vu} } },
+   {Au,    {{1, 1, T[Yu][1,1] / Yu[1,1]},
+            {2, 2, T[Yu][2,2] / Yu[2,2]},
+            {3, 3, T[Yu][3,3] / Yu[3,3]} } },
+   {Ad,    {{1, 1, T[Yd][1,1] / Yd[1,1]},
+            {2, 2, T[Yd][2,2] / Yd[2,2]},
+            {3, 3, T[Yd][3,3] / Yd[3,3]} } },
+   {Ae,    {{1, 1, T[Ye][1,1] / Ye[1,1]},
+            {2, 2, T[Ye][2,2] / Ye[2,2]},
+            {3, 3, T[Ye][3,3] / Ye[3,3]} } },
+   {MSOFT, {{1, MassB},
+            {2, MassWB},
+            {3, MassG},
+            {21, mHd2},
+            {22, mHu2},
+            {31, Sqrt[ml2[1,1]]},
+            {32, Sqrt[ml2[2,2]]},
+            {33, Sqrt[ml2[3,3]]},
+            {34, Sqrt[me2[1,1]]},
+            {35, Sqrt[me2[2,2]]},
+            {36, Sqrt[me2[3,3]]},
+            {41, Sqrt[mq2[1,1]]},
+            {42, Sqrt[mq2[2,2]]},
+            {43, Sqrt[mq2[3,3]]},
+            {44, Sqrt[mu2[1,1]]},
+            {45, Sqrt[mu2[2,2]]},
+            {46, Sqrt[mu2[3,3]]},
+            {47, Sqrt[md2[1,1]]},
+            {48, Sqrt[md2[2,2]]},
+            {49, Sqrt[md2[3,3]]} } }
+};
 \end{lstlisting}
@@
 selectron) are considered to be LSP candidates.
+
+Finally, with the variable \code{ExtraSLHAOutputBlocks} the user can
+define extra SLHA output blocks.  The values of the block entries will
+be calculated at the output scale, which is specified in entry $12$ in
+the SLHA input block \code{MODSEL}.  In the example model file above
+the following six extra SLHA compliant output blocks are defined:
+\code{ALPHA}, \code{HMIX}, \code{Au}, \code{Ad}, \code{Ae} and
+\code{MSOFT}.  The \code{ALPHA} output block contains the CP-even
+Higgs pole mass mixing angle as the only entry.  The \code{HMIX}
+output block contains the $\mu$-parameter, the ratio $\tan\beta =
+v_u/v_d$, the combination $v=\sqrt{v_u^2 + v_d^2}$, the squared mass
+of the CP-odd Higgs $m_A^2$, the soft-breaking parameter $B\mu$ and
+the values of $v_u$ and $v_d$, all defined in the \DRbar\ scheme.  In
+an analogue way four more output blocks for the soft-breaking \DRbar\
+parameters are defined.  For a CMSSM example parameter point with $m_0
+= 125\unit{GeV}$, $M_{1/2} = 500\unit{GeV}$, $\tan\beta = 10$,
+$\sign\mu = +1$ and $A_0 = 0$ the \fs-generated CMSSM spectrum
+generator writes the so defined blocks to the output in the form
+%
+\begin{lstlisting}
+Block ALPHA Q= 8.76740936E+02
+           1.06784138E-01   # ArcCos(Pole(ZH(0,1)))
+Block HMIX Q= 8.76740936E+02
+     1     6.31218384E+02   # Mu
+     2     9.67312621E+00   # vu/vd
+     3     2.44053433E+02   # Sqrt(Sqr(vd) + Sqr(vu))
+     4     5.36777230E+05   # Sqr(MAh(1))
+   101     5.49048159E+04   # BMu
+   102     2.50962986E+01   # vd
+   103     2.42759663E+02   # vu
+Block Au Q= 8.76740936E+02
+  1  1    -1.14477419E+03   # TYu(0,0)/Yu(0,0)
+  2  2    -1.14476911E+03   # TYu(1,1)/Yu(1,1)
+  3  3    -8.83902977E+02   # TYu(2,2)/Yu(2,2)
+Block Ad Q= 8.76740936E+02
+  1  1    -1.40026447E+03   # TYd(0,0)/Yd(0,0)
+  2  2    -1.40025976E+03   # TYd(1,1)/Yd(1,1)
+  3  3    -1.30885006E+03   # TYd(2,2)/Yd(2,2)
+Block Ae Q= 8.76740936E+02
+  1  1    -3.00005426E+02   # TYe(0,0)/Ye(0,0)
+  2  2    -3.00000006E+02   # TYe(1,1)/Ye(1,1)
+  3  3    -2.98364373E+02   # TYe(2,2)/Ye(2,2)
+Block MSOFT Q= 8.76740936E+02
+     1     2.09018579E+02   # MassB
+     2     3.88257873E+02   # MassWB
+     3     1.11544211E+03   # MassG
+    21     1.09683411E+05   # mHd2
+    22    -3.85898988E+05   # mHu2
+    31     3.54416224E+02   # Sqrt(ml2(0,0))
+    32     3.54412891E+02   # Sqrt(ml2(1,1))
+    33     3.53407333E+02   # Sqrt(ml2(2,2))
+    34     2.22035720E+02   # Sqrt(me2(0,0))
+    35     2.22024873E+02   # Sqrt(me2(1,1))
+    36     2.18731935E+02   # Sqrt(me2(2,2))
+    41     1.02053367E+03   # Sqrt(mq2(0,0))
+    42     1.02053107E+03   # Sqrt(mq2(1,1))
+    43     9.40760849E+02   # Sqrt(mq2(2,2))
+    44     9.82565930E+02   # Sqrt(mu2(0,0))
+    45     9.82563167E+02   # Sqrt(mu2(1,1))
+    46     8.09126982E+02   # Sqrt(mu2(2,2))
+    47     9.77979392E+02   # Sqrt(md2(0,0))
+    48     9.77976666E+02   # Sqrt(md2(1,1))
+    49     9.73121951E+02   # Sqrt(md2(2,2))
+\end{lstlisting}
 

### 2014-11-18 doc/flexiblesusy-paper.tex
fixing small typos

@@
 
-  When two loop corrections have been enabled in the SLHA
-  file by setting entry $4$ of the FlexibleSUSY block to $2$ the user may
-  also select individual corrections. Entries $9$,$10$,$11$ and $12$,
-  correspond to two loop corrections of order $\oatas$, $\oabas$,
-  $\oatplusabsq$ and $\oatauq$ respectively and will be disabled when
-  the corresponding entry is set to zero. In this way, for example, in
-  the NMSSM the user may decide not to use the partial corrections at
-  order $\oatplusabsq$ and $\oatauq$, all of which have
-  only been computed in the MSSM.
+When two-loop corrections have been enabled in the SLHA file by
+setting entry $4$ of the \code{FlexibleSUSY} block to $2$ the user may
+also select individual corrections.  The \code{FlexibleSUSY} block
+entries $9$, $10$, $11$ and $12$, correspond to two-loop corrections
+of the order $\oatas$, $\oabas$, $\oatplusabsq$ and $\oatauq$
+respectively and will be disabled when the corresponding entry is set
+to zero.  In this way, for example, in the NMSSM the user may decide
+not to use the partial corrections at order $\oatplusabsq$ and
+$\oatauq$, all of which have only been computed in the MSSM.
 

### 2014-11-18 doc/flexiblesusy-paper.tex
slightly reformulate "Current limitations"

@@
 \subsection{Current limitations and future extensions}
-Although we try to handle as many models as possible there are still
+Although we try to handle as many models as possible, there are still
 some limitations to what can be done. In it's current form \fs assumes
-all couplings are are real, therefore it is limited to $CP$-conserving
-versions of SUSY models.   Although it is implicit in the
-title we would like to stress that currently we cannot provide a
-spectrum generator for non-SUSY models.  This also means that like
-other spectrum generators it is difficult to reliably predict the mass
+all couplings are real, therefore it is limited to $CP$-conserving
+versions of SUSY models.  Although it is implicit in the title, we
+would like to stress that currently we cannot provide a spectrum
+generator for non-SUSY models.  This also means that like other
+spectrum generators it is difficult to reliably predict the mass
 spectrum in extreme cases such as Split-SUSY
-\cite{Wells:2003tf,ArkaniHamed:2004fb,Giudice:2004tc} where the mass
-scales of the SUSY model are drastically split leading to very large
+\cite{Wells:2003tf,ArkaniHamed:2004fb,Giudice:2004tc}, where the mass
+scales of the SUSY model are drastically split, leading to very large
 logarithms which are not resummed.  In such cases the tower of
 effective theories offers the best possibility for a solution.
-However for this to work first \fs must be extended to include
-non-SUSY models.  Finally the gauge group of the model should be
-semi-simple and contain the standard model gauge group as factor so
-that we can match to the SM at low energies.
-
+However, for this to work \fs must be extended to include non-SUSY
+models.  Finally, the gauge group of the model is currently restricted
+to be semi-simple and to contain the Standard Model gauge group as
+factor, so that the SUSY model can be directly matched to the SM at
+low energies.
+%
 Future releases which extend \fs beyond each of these limitations are

### 2014-11-19 doc/flexiblesusy-paper.tex
changed sentence just before Design Goal to reinstate original meaning - we pass the outut *to* other codes both private and public.

@@
 blocks in order to simplify the process of passing the output of a
-custom \fs-generated spectrum generator or any new public tool
-developed for that model.
+custom \fs-generated spectrum generator to any private or new public
+tool developed for that model.
 

### 2014-11-19 doc/flexiblesusy-paper.tex
Fixed teh way we introduce the two loop higgs corrections, sicne editd had made the start of this sentence misleading.

@@
 
-If higher accuracy is required, two-loop corrections to the
-self-energies can be added by setting \code{UseHiggs2LoopMSSM = True}
-in the MSSM or \code{UseHiggs2LoopNMSSM = True} in the NMSSM in the
-model file.  The former provides routines that call the two-loop Higgs
+For the Higgs states two-loop corrections to the self-energies can
+optionally be added by setting \code{UseHiggs2LoopMSSM = True} in the
+MSSM or \code{UseHiggs2LoopNMSSM = True} in the NMSSM in the model
+file.  The former provides routines that call the two-loop Higgs
 FORTRAN routines supplied by P.~Slavich from
@@
 NMSSM at $\oatas$, $\oabas$ from \cite{Degrassi:2009yq} and partial
-corrections for the order $\oatq$, $\oabq$, $\oatauq$ and $\oatab$ from the
-MSSM.  These corrections can then be included in the
+corrections for the order $\oatq$, $\oabq$, $\oatauq$ and $\oatab$
+from the MSSM.  These corrections can then be included in the
 calculation of the Higgs pole masses when running the spectrum

### 2014-11-19 doc/flexiblesusy-paper.tex
altered wording describing E6SSM mass comparison.

@@
 the CE$_6$SSM generator as an input to \fs and the spectra were found
-to be in reasonable agreement ($\approx 10\%$), given the different
-levels of precision.
+to be in reasonable agreement  given the different levels of precision with deviatiuons in the mass spectra $\lesssim 10\%$.
 

### 2014-11-19 doc/flexiblesusy-paper.tex
capitalize Standard Model

@@
 SUSY may be realized in nature in some form. In particular
-supersymmetric extensions of the standard model (SM) where SUSY is broken
+supersymmetric extensions of the Standard Model (SM) where SUSY is broken
 at the TeV scale have been proposed to solve the hierarchy
@@
 Detailed phenomenological studies have been carried out for scenarios
-within the minimal supersymmetric standard model (MSSM), for a review
+within the minimal supersymmetric Standard Model (MSSM), for a review
 see \cite{Chung:2003fi}.  Such work has been greatly aided by public
@@
   Ellwanger:2006rn,Ellwanger:2008py, Allanach:2013kza} for the next to
-minimal supersymmetric standard model (NMSSM)  (for
+minimal supersymmetric Standard Model (NMSSM)  (for
 recent reviews see \cite{Ellwanger:2009dp, Maniatis:2009re}).
@@
 (see e.g.~\cite{Aad:2013wta,Chatrchyan:2014lfa}). No other signature
-of beyond the standard model (BSM) physics has been observed, leaving
+of beyond the Standard Model (BSM) physics has been observed, leaving
 the fundamental questions which motivated BSM physics

### 2014-11-20 doc/flexiblesusy-paper.tex
note about internally used SLHAea class and adding reference
add heaviest neutralino to pole mass comparison table to avoid confusion

@@
 \secref{sec:modfile}.
-%
-% We use SLHAea \cite{slhaea} to create these SLHA files and with
-% simple modifications these can be passed directly to another code
-% which uses them as an alternative to writing out the SLHA file.
-%
+Internally, \fs uses the SLHAea library \cite{slhaea} to read and
+write the SLHA files.  The internally stored SLHAea object can be
+passed to other programs at the C++ level for inter program
+communication.
 We have tested that the HDECAY $3.4$ \cite{hdecay} and SDECAY $1.4$
@@
 %
-(iii) The pole mass of the lightest neutralino,
-$m_{\tilde{\chi}_1^0}$, is given in the third row.  Since the
+(iii) The pole masses of the lightest and heaviest neutralinos,
+$m_{\tilde{\chi}_1^0}$ and $m_{\tilde{\chi}_4^0}$, are given in the rows
+$3$--$4$.  Since the
 momentum-dependent loop-corrections to the lightest neutralino mass
@@
 (iv) The pole masses of the lightest sfermions are given in rows
-$4$--$6$.  Since these particles are contained in $6$-plets, the
+$5$--$7$.  Since these particles are contained in $6$-plets, the
 run-time for the calculation of their pole masses is dramatically
@@
     & $2295\unit{GeV}$ ($64.2\unit{ms}$) \\
+    $m_{\tilde{\chi}_4^0}$
+    & $4181\unit{GeV}$ ($4.17\unit{ms}$)
+    & $4174\unit{GeV}$ ($16.1\unit{ms}$)
+    & $4174\unit{GeV}$ ($64.2\unit{ms}$) \\
     $m_{\tilde{u}_1}$

### 2014-11-21 doc/flexiblesusy-paper.tex
Minor corrections to paper

@@
 Although we try to handle as many models as possible, there are still
-some limitations to what can be done. In it's current form \fs assumes
+some limitations to what can be done. In its current form \fs assumes
 all couplings are real, therefore it is limited to $CP$-conserving
@@
 tadpoles.  For MSSM-like models (with two $CP$-even Higgs bosons, one
-$CP$-odd Higgs boson, one neutral Goldstone boson) routines dor
-caluclating these corrections will be generated by setting
+$CP$-odd Higgs boson, one neutral Goldstone boson) routines for
+calculating these corrections will be generated by setting
 \code{UseHiggs2LoopMSSM = True} in the model file and by defining the
@@
  masses, when appropriate settings are selected in the SLHA file. Note
- that even in the NMSSM the user miust make an important decision as
+ that even in the NMSSM the user must make an important decision as
  to whether or not to enable the generated MSSM corrections which are
  incomplete in the NMSSM.  We feel that it is valuable to have
- these MSSM corrections for scenarios where there singlet mixing is
+ these MSSM corrections for scenarios where singlet mixing is
  very small and in particular for cross checks against the MSSM when
@@
 the values of $v_u$ and $v_d$, all defined in the \DRbar\ scheme.  In
-an analogue way four more output blocks for the soft-breaking \DRbar\
+an analogous way four more output blocks for the soft-breaking \DRbar\
 parameters are defined.  For a CMSSM example parameter point with $m_0
@@
 %
-One can now chose between the two solutions by setting entry $4$ in
+One can now choose between the two solutions by setting entry $4$ in
 the \code{MINPAR} block of the SLHA input file to either $+1$ or $-1$.
@@
 iteration.  In this case, however, the user does not have the option
-to chose between the different solutions.
+to choose between the different solutions.
 
@@
   tree-level mass spectrum of the multiplet is very split.  Note: We
-  strongly discourage the use of this method for the determination the
-  Higgs pole masses, as the result will very imprecise due to the
+  strongly discourage the use of this method for the determination of the
+  Higgs pole masses, as the result will be very imprecise due to the
   large loop corrections.  \fs will print a warning if this method is
@@
 momentum-dependent loop-corrections to the lightest neutralino mass
-are small for this parameter point, it's pole mass varies only in the
+are small for this parameter point, its pole mass varies only in the
 sub-GeV range between the three methods.  However, the run-time of the
@@
 sfermion masses between the three different methods is less than
-$\unit{0.4\%}$, one can consider calculating then with
+$\unit{0.4\%}$, one can consider calculating them with
 \code{MediumPrecision} only.
@@
 However in such models the user should still consider whether these
-corrections are really the leading corrections in the model or if
+corrections are really the leading corrections in the model or %if
 there are other potentially large two-loop corrections which are
@@
     We follow the notation of the NMSSM model parameters used in
-    \cite{Ellwanger:2009dp,Allanach:2013kza}.}
+    \cite{Allanach:2013kza,Ellwanger:2009dp}.}
   \label{tab:unit-test-parameter-points}
@@
 the CE$_6$SSM generator as an input to \fs and the spectra were found
-to be in reasonable agreement  given the different levels of precision with deviatiuons in the mass spectra $\lesssim 10\%$.
+to be in reasonable agreement  given the different levels of precision with deviations in the mass spectra $\lesssim 10\%$.
 

### 2014-11-21 doc/flexiblesusy-paper.tex
Fix a typo in paper
Update package version number in paper

@@
 \url{http://flexiblesusy.hepforge.org}.  To download and install
-version 1.0.1 run:
+version 1.0.3 run:
 %
 \begin{lstlisting}[language=bash]
-$ wget https://www.hepforge.org/archive/flexiblesusy/FlexibleSUSY-1.0.1.tar.gz
-$ tar -xf FlexibleSUSY-1.0.1.tar.gz
-$ cd FlexibleSUSY-1.0.1
+$ wget https://www.hepforge.org/archive/flexiblesusy/FlexibleSUSY-1.0.3.tar.gz
+$ tar -xf FlexibleSUSY-1.0.3.tar.gz
+$ cd FlexibleSUSY-1.0.3
 \end{lstlisting}%% $
@@
 introduces these global signs as additional free parameters to allow
-the user to chose between the different solutions.  The case of the
+the user to choose between the different solutions.  The case of the
 real MSSM Eq.~\eqref{eq:EWSB-solution-for-mu} is therefore handled

### 2014-12-10 doc/flexiblesusy-paper.tex
update masses for Table 1 (polem mass precision comparison)
The new CMSSM parameter point was proposed by Peter and reads

Block MINPAR                 # Input parameters
    1   1000                 # m0
    2   1000                 # m12
    3   10                   # TanBeta
    4   1                    # SignMu
    5   -3300                # Azero

@@
 $\Delta m_t^{(1),\text{qcd}}$ and $\Delta m_t^{(2),\text{qcd}}$ are
-taken from \cite{Bednyakov:2002sf} and read
+taken from \cite{Avdeev:1997sz,Bednyakov:2002sf} and read
 %
@@
 The \DRbar mass of the bottom quark is calculated as
-\cite{Baer:2002ek,Skands:2003cj}
+\cite{Avdeev:1997sz,Baer:2002ek,Skands:2003cj}
 %
@@
 (i) the calculated lightest $CP$-even Higgs pole mass, $m_h$, differs
-about $2\unit{GeV}$ between \code{MediumPrecision} and
+about $1.2\unit{GeV}$ between \code{MediumPrecision} and
 \code{HighPrecision}, due to large loop corrections.  Since the
@@
 Not neglecting the two-loop terms by using \code{HighPrecision}
-increases the gluino mass about $1\unit{\%}$ for this parameter point.
+increases the gluino mass about $0.5\unit{\%}$ for this parameter point.
 %
@@
 sfermion masses between the three different methods is less than
-$\unit{0.4\%}$, one can consider calculating them with
+$\unit{0.3\%}$, one can consider calculating them with
 \code{MediumPrecision} only.
@@
     & --
-    & $128.3\unit{GeV}$ ($2.45\unit{ms}$)
-    & $126.3\unit{GeV}$ ($9.53\unit{ms}$) \\
+    & $125.3\unit{GeV}$ ($2.40\unit{ms}$)
+    & $124.1\unit{GeV}$ ($9.57\unit{ms}$) \\
     $m_{\tilde{g}}$
-    & $10167\unit{GeV}$ ($0.10\unit{ms}$)
-    & $10167\unit{GeV}$ ($0.10\unit{ms}$)
-    & $10251\unit{GeV}$ ($0.38\unit{ms}$) \\
+    & $2218\unit{GeV}$ ($0.12\unit{ms}$)
+    & $2218\unit{GeV}$ ($0.12\unit{ms}$)
+    & $2231\unit{GeV}$ ($0.40\unit{ms}$) \\
     $m_{\tilde{\chi}_1^0}$
-    & $2295\unit{GeV}$ ($4.17\unit{ms}$)
-    & $2295\unit{GeV}$ ($16.1\unit{ms}$)
-    & $2295\unit{GeV}$ ($64.2\unit{ms}$) \\
+    & $429\unit{GeV}$ ($4.02\unit{ms}$)
+    & $429\unit{GeV}$ ($16.4\unit{ms}$)
+    & $429\unit{GeV}$ ($48.3\unit{ms}$) \\
     $m_{\tilde{\chi}_4^0}$
-    & $4181\unit{GeV}$ ($4.17\unit{ms}$)
-    & $4174\unit{GeV}$ ($16.1\unit{ms}$)
-    & $4174\unit{GeV}$ ($64.2\unit{ms}$) \\
+    & $1943\unit{GeV}$ ($4.02\unit{ms}$)
+    & $1944\unit{GeV}$ ($16.4\unit{ms}$)
+    & $1944\unit{GeV}$ ($48.3\unit{ms}$) \\
     $m_{\tilde{u}_1}$
-    & $8703\unit{GeV}$ ($2.45\unit{ms}$)
-    & $8714\unit{GeV}$ ($14.9\unit{ms}$)
-    & $8734\unit{GeV}$ ($58.5\unit{ms}$) \\
+    & $1055\unit{GeV}$ ($2.50\unit{ms}$)
+    & $1081\unit{GeV}$ ($15.1\unit{ms}$)
+    & $1085\unit{GeV}$ ($59.1\unit{ms}$) \\
     $m_{\tilde{d}_1}$
-    & $10442\unit{GeV}$ ($2.48\unit{ms}$)
-    & $10391\unit{GeV}$ ($15.0\unit{ms}$)
-    & $10418\unit{GeV}$ ($59.9\unit{ms}$) \\
+    & $1801\unit{GeV}$ ($2.54\unit{ms}$)
+    & $1778\unit{GeV}$ ($15.3\unit{ms}$)
+    & $1783\unit{GeV}$ ($59.7\unit{ms}$) \\
     $m_{\tilde{e}_1}$
-    & $7913\unit{GeV}$ ($1.94\unit{ms}$)
-    & $7909\unit{GeV}$ ($11.2\unit{ms}$)
-    & $7909\unit{GeV}$ ($34.1\unit{ms}$) \\
+    & $1019\unit{GeV}$ ($1.90\unit{ms}$)
+    & $1018\unit{GeV}$ ($11.3\unit{ms}$)
+    & $1018\unit{GeV}$ ($22.6\unit{ms}$) \\
     \bottomrule
@@
   \caption{Comparison of pole masses with different calculation
-    methods for the CMSSM with $m_0=8\unit{TeV}$, $M_{1/2}=A_0=5\unit{TeV}$,
-    $\tan\beta=20$, $\sign\mu=+1$.  In brackets the time used to calculate
+    methods for the CMSSM with $m_0=M_{1/2}=1\unit{TeV}$, $A_0=-3.3\unit{TeV}$,
+    $\tan\beta=10$, $\sign\mu=+1$.  In brackets the time used to calculate
     the pole masses of the whole multiplet is given.}
@@
 
+%\cite{Avdeev:1997sz}
+\bibitem{Avdeev:1997sz}
+  L.~V.~Avdeev and M.~Y.~Kalmykov,
+  %``Pole masses of quarks in dimensional reduction,''
+  Nucl.\ Phys.\ B {\bf 502} (1997) 419
+  [hep-ph/9701308].
+  %%CITATION = HEP-PH/9701308;%%
+  %66 citations counted in INSPIRE as of 28 Nov 2014
+
 %\cite{Bednyakov:2002sf}

### 2015-01-28 doc/flexiblesusy-paper.tex
relax minimum required boost version to 1.41.0
adding table of contents
add little line spacings between Program Summary items

@@
 
+\clearpage
 \section{Program Summary}
-\noindent{\em Program title:} \fs\\ {\em Program obtainable from:}
-         {\tt http://flexiblesusy.hepforge.org/}\\ {\em Distribution
-           format:}\/ tar.gz\\ {\em Programming language:} {\tt
-           C++, Wolfram/\mathematica, FORTRAN, Bourne shell}\\ {\em Computer:}\/ Personal computer\\ {\em Operating
-           system:}\/ Tested on Linux 3.x, Mac OS X\\
+\noindent{\em Program title:} \fs\\[0.5em] {\em Program obtainable from:}
+         {\tt http://flexiblesusy.hepforge.org/}\\[0.5em] {\em Distribution
+           format:}\/ tar.gz\\[0.5em] {\em Programming language:} {\tt
+           C++, Wolfram/\mathematica, FORTRAN, Bourne shell}\\[0.5em]
+         {\em Computer:}\/ Personal computer\\[0.5em] {\em Operating
+           system:}\/ Tested on Linux 3.x, Mac OS X\\[0.5em]
          {\em External routines:}\/ SARAH 4.0.4, Boost library,
-         Eigen, LAPACK\\ {\em
-           Typical running time:}\/ 0.06-0.2 seconds per parameter
-         point.\\ {\em Nature of problem:}\/ Determining the mass
+         Eigen, LAPACK\\[0.5em] {\em
+           Typical running time:}\/ 0.06--0.2 seconds per parameter
+         point\\[0.5em] {\em Nature of problem:}\/ Determining the mass
          spectrum and mixings for any supersymmetric model. The
@@
          renormalization group equations forming a large set of
-         coupled first-order differential equations. \\ {\em Solution method:}\/
+         coupled first-order differential equations. \\[0.5em] {\em Solution method:}\/
          Nested iterative algorithm and numerical minimization of the
-         Higgs potential.  \\ {\em Restrictions:} The couplings must
+         Higgs potential.  \\[0.5em] {\em Restrictions:} The couplings must
          remain perturbative at all scales between the highest and
          lowest boundary condition.  \fs~ assumes that all couplings
-         of the model are real (i.e.\ $CP-$conserving). Due to the
+         of the model are real (i.e.\ $CP$-conserving). Due to the
          modular nature of the generated code adaption and extension
          to overcome restrictions in scope is quite straightforward.
+
+\clearpage
+\tableofcontents
 
@@
   \url{http://eigen.tuxfamily.org}
-\item Boost library, version 1.36.0 or higher
+\item Boost library, version 1.37.0 or higher
   \url{http://www.boost.org}

### 2015-01-31 doc/flexiblesusy-paper.tex
adding self-energy symbol to be absolutely clear
describe symbol m_t to be absolutely clear
Note: The meaning of m_t was not obvious to Pietro Slavich, so, let's
be absolutely clear here.
add acknowledgments to DH

@@
 contributions.  The labels $L,R,S$ denote the left-, right- and
-non-polarized part of the self-energy.  The separated QCD corrections
+non-polarized part of the self-energy, $\Sigma_{t}$.  The separated QCD corrections
 $\Delta m_t^{(1),\text{qcd}}$ and $\Delta m_t^{(2),\text{qcd}}$ are
@@
 \begin{align}
-  \Delta m_t^{(1),\text{qcd}} &= -\frac{g_3^2}{12 \pi^2} \left[5-3 \log\left(\frac{m_t^2}{\mu^2}\right)\right],\\
+  \Delta m_t^{(1),\text{qcd}} &= -\frac{g_3^2}{12 \pi^2} \left[5-3 \log\left(\frac{m_t^2}{\mu^2}\right)\right],
+  \label{eq:top-selfenergy-qcd-1L}\\
   \begin{split}
@@
   \end{split}
+  \label{eq:top-selfenergy-qcd-2L}
 \end{align}
+%
+In Eqs.\ \eqref{eq:top-selfenergy-qcd-1L} and
+\eqref{eq:top-selfenergy-qcd-2L} $m_t$ denotes the \DRbar mass of the
+top quark.
 %
@@
 Mac OS X and Dylan Harries for spotting a bug in the configuration
-script.
+script and creating an EWSB fixed-point iteration implementation for
+the CNE$_6$SSM, which served as a prototype for implementing a
+generalized EWSB fixed-point iteration.
 %

### 2015-02-03 doc/flexiblesusy-paper.tex
fixing typos and language style according to CPC

@@
 \eqref{eq:one-loop-ewsb-eq} can have multiple solutions, depending on
-which paramters are chosen to be fixed by these equations.  A
+which parameters are chosen to be fixed by these equations.  A
 well-known example is the real MSSM, where the $\mu$-parameter is
@@
 %
-where we're again choosing the phases of $U$ and $V$ such that all mass
+where we are again choosing the phases of $U$ and $V$ such that all mass
 eigenvalues are positive.
@@
 where line 2 contains the dimension-5 operator
-multiplied by its coeffficient matrix \code{WOp}.
+multiplied by its coefficient matrix \code{WOp}.
 The declarations of the neutrino Dirac spinors and mixing are
@@
 found to match within numerical precision.  The handwritten code
-doesn't include full one-loop self-energies or tadpoles, so tests on
+does not include full one-loop self-energies or tadpoles, so tests on
 these were not carried out.  Although the generators assume different
@@
   because flavour-off-diagonal sfermion self-energy matrix elements
-  don't need to be calculated.  Here we compare \fs's non-flavour
+  do not need to be calculated.  Here we compare \fs's non-flavour
   violating CMSSM spectrum generator FlexibleSUSY-NoFV (version 1.0.0)
@@
   run-time of SPheno and SPhenoMSSM by approximately a factor $1.5$,
-  compared to gfortran.}  We're generating $2\cdot 10^{4}$ random
+  compared to gfortran.}  We are generating $2\cdot 10^{4}$ random
 CMSSM parameter points with $m_0\in [50,1000]\unit{GeV}$, $m_{1/2}\in

### 2015-02-09 doc/flexiblesusy-paper.tex
fixing typos found by Jae-hyeon Park

@@
 
-None of the fundamental motivations of supersymmetry require
+None of the fundamental motivations of supersymmetry requires
 minimality, and specific alternatives to (or extensions of) the MSSM
@@
   algorithms usually need $O(10)$ iterations between the high and the
-  low scale to find a set of parameters consistent will all boundary
+  low scale to find a set of parameters consistent with all boundary
   conditions with a $0.01\%$ precision goal.  During each iteration
@@
 \code{model_files/<model-file-name>/FlexibleSUSY.m.in} file, where the
-directory \code{<model-file-name>} is the specified by the
+directory \code{<model-file-name>} is specified by the
 \code{--model-file=<model-file-name>} option.  If no such option is
@@
 %
-The variables \code{LowScaleInput}, \code{SUSYScaleInput}
+The variables \code{LowScaleInput}, \code{SUSYScaleInput} and
 \code{HighScaleInput}, which list the parameter settings for imposing
@@
  based on the physics they study.  Similarly in models that go beyond
- the NMSSM the user must decide based the physics whether on not these
+ the NMSSM the user must decide based on the physics whether on not these
  corrections will give the leading two-loop corrections in their
@@
 \code{ALPHA}, \code{HMIX}, \code{Au}, \code{Ad}, \code{Ae} and
-\code{MSOFT}.  The \code{ALPHA} output block contains the CP-even
+\code{MSOFT}.  The \code{ALPHA} output block contains the $CP$-even
 Higgs pole mass mixing angle as the only entry.  The \code{HMIX}
@@
 v_u/v_d$, the combination $v=\sqrt{v_u^2 + v_d^2}$, the squared mass
-of the CP-odd Higgs $m_A^2$, the soft-breaking parameter $B\mu$ and
+of the $CP$-odd Higgs $m_A^2$, the soft-breaking parameter $B\mu$ and
 the values of $v_u$ and $v_d$, all defined in the \DRbar\ scheme.  In
@@
   linear scalar terms, soft bilinear scalar interactions, soft
-  trilinear scalar interactions, soft gaugino mass terms and soft
+  trilinear scalar interactions, soft quadlinear scalar interactions,
+  soft gaugino mass terms and soft
   scalar squared masses.
@@
 where $V^\text{tree}$ is the tree-level Higgs potential, $v_i$ is the
-VEV corresponding to the Higgs fields $H_i$ and $t_i$ is the one-loop
+VEV corresponding to the Higgs field $H_i$ and $t_i$ is the one-loop
 tadpole diagram of $H_i$.  Already at the tree-level ($t_i = 0$) Eqs.\
@@
 theories, which restricts the number of possible multiple solutions of
-Eqs.\ \eqref{eq:one-loop-ewsb-eq} to be less or equal to $4N$.  In
+Eqs.\ \eqref{eq:one-loop-ewsb-eq} to be less than or equal to $4N$.  In
 some cases, as for example in the real MSSM, the different solutions
@@
 non-trivial to judge whether this is because there is no physical
-solution for the given parameter space point or whether a solution
+solution for the given parameter space point or a solution
 exists but the fixed point iteration is unable to find the solution.
@@
 new models a physical understanding of the model is still essential
-and this can help the user determine why the such problems arise.
+and this can help the user determine why such problems arise.
 
@@
 The nightly test results can be found at
-\url{http://iktp.tu-dresden.de/~avoigt/FlexibleSUSY/test.xhtml}
+\url{https://www.desy.de/~alvoigt/FlexibleSUSY/test.xhtml}
 
@@
 \tabref{tab:run-time-comparison}.  The first column shows the run-time
-on a Intel Core2 Duo (P8600, $2.40\unit{GHz}$) where only one core was
+on an Intel Core2 Duo (P8600, $2.40\unit{GHz}$) where only one core was
 enabled.  The second column shows the run-time on the same processor
@@
 
-Under both the non-flavour violating spectrum generators (first three
+Among both the non-flavour violating spectrum generators (first three
 rows) as well as the flavour violating ones (4th and 5th row) we find
@@
 FlexibleSUSY-NoFV is faster by a factor $1.4$--$1.7$, and compared to
-Softsusy around a factor $2$--$2.5$.  Under the flavour violating
+Softsusy around a factor $2$--$2.5$.  Between the flavour violating
 spectrum generators FlexibleSUSY-FV is faster than SPhenoMSSM by a

### 2015-02-09 doc/flexiblesusy-paper.tex
fixing typo found by Jae-hyeon Park

@@
  based on the physics they study.  Similarly in models that go beyond
- the NMSSM the user must decide based on the physics whether on not these
+ the NMSSM the user must decide based on the physics whether or not these
  corrections will give the leading two-loop corrections in their

### 2015-04-17 test/test_CMSSMCPV_ewsb.cpp
tadpole[2] is not exactly zero due to higher order effects.
tadpole[0,1,3] are fixed by the three EWSB eqs.  tadpole[2] is not
directly fixed by any EWSB eq.  However, the relation

   tadpole[2]/tadpole[3] = vu/vd           (1)

is valid, up to higher order effects.  After the iteration tadpole[2]
is not exactly zero, because during the iteration the relation (1) is
spoiled by higher order effects: During the iteration Mu and BMu pick
up tree-level and one-loop tadpole contributions.  The new values for
Mu and BMu in the next iteration step are then used to recalculate the
tree-level spectrum

@@ -79,10 +79,10 @@ BOOST_AUTO_TEST_CASE( test_CMSSMCPV_ewsb_one_loop )
 
    m.tadpole_equations(tadpole);
 
-   BOOST_CHECK(Abs(tadpole[0]) > 1.);
-   BOOST_CHECK(Abs(tadpole[1]) > 1.);
-   BOOST_CHECK(Abs(tadpole[2]) > 1.);
-   BOOST_CHECK(Abs(tadpole[3]) > 1.);
+   BOOST_CHECK(Abs(tadpole[0]) > 1000.);
+   BOOST_CHECK(Abs(tadpole[1]) > 1000.);
+   BOOST_CHECK(Abs(tadpole[2]) > 1000.);
+   BOOST_CHECK(Abs(tadpole[3]) > 100. );
 
    m.set_ewsb_iteration_precision(precision);
    const int error = m.solve_ewsb_one_loop();
@@ -91,10 +91,27 @@ BOOST_AUTO_TEST_CASE( test_CMSSMCPV_ewsb_one_loop )
 
    m.tadpole_equations(tadpole);
 
-   BOOST_CHECK_SMALL(Abs(tadpole[0]), 0.003);
-   BOOST_CHECK_SMALL(Abs(tadpole[1]), 0.0005);
-   BOOST_CHECK_SMALL(Abs(tadpole[2]), precision);
-   BOOST_CHECK_SMALL(Abs(tadpole[3]), precision);
+   // tadpole[0,1,3] are fixed by the three EWSB eqs.  tadpole[2] is
+   // not directly fixed by any EWSB eq.  However, the relation
+   //
+   //    tadpole[2]/tadpole[3] = vu/vd           (1)
+   //
+   // is valid, up to higher order effects.  After the iteration
+   // tadpole[2] is not exactly zero, because during the iteration the
+   // relation (1) is spoiled by higher order effects: During the
+   // iteration Mu and BMu pick up tree-level and one-loop tadpole
+   // contributions.  The new values for Mu and BMu in the next
+   // iteration step are then used to recalculate the tree-level
+   // spectrum, which is then used to calculate the one-loop tadpoles.
+   // I.e. during this iterative procedure higher order contributions
+   // are put into Mu and BMu, which leads to a violation of (1).  For
[…]

### 2017-03-15 meta/FlexibleSUSY.m
allowing for multiple parameters

@@ -291,7 +291,7 @@ HaveEWSBSolver[solver_] := MemberQ[FlexibleSUSY`FSEWSBSolvers, solver];
 
 HaveBVPSolver[solver_] := MemberQ[FlexibleSUSY`FSBVPSolvers, solver];
 
-PrintHeadline[text_] :=
+PrintHeadline[text__] :=
     Block[{},
           Print[""];
           Print[Style["---------------------------------", Bold]];

### 2017-03-15 meta/FlexibleSUSY.m
nicer start messages

@@ -2089,7 +2089,7 @@ ReadPoleMassPrecisions[defaultPrecision_Symbol, highPrecisionList_List,
 LoadModelFile[file_String] :=
     Module[{},
            If[FileExists[file],
-              Print["Loading model file ", file];
+              PrintHeadline["Loading FlexibleSUSY model file ", file];
               Get[file];
               CheckModelFileSettings[];
               ,
@@ -2217,6 +2217,7 @@ MakeFlexibleSUSY[OptionsPattern[]] :=
             extraSLHAOutputBlocks, effectiveCouplings ={}, extraVertices = {},
             vertexRules, vertexRuleFileName, effectiveCouplingsFileName,
             Lat$massMatrices, spectrumGeneratorFiles = {}, spectrumGeneratorInputFile},
+           PrintHeadline["Starting FlexibleSUSY"];
            (* check if SARAH`Start[] was called *)
            If[!ValueQ[Model`Name],
               Print["Error: Model`Name is not defined.  Did you call SARAH`Start[\"Model\"]?"];
@@ -2232,7 +2233,7 @@ MakeFlexibleSUSY[OptionsPattern[]] :=
            (* load model file *)
            LoadModelFile[OptionValue[InputFile]];
            Print["FlexibleSUSY model file loaded"];
-           Print["  Model: ", FlexibleSUSY`FSModelName];
+           Print["  Model: ", Style[FlexibleSUSY`FSModelName, Blue]];
            Print["  Model file: ", OptionValue[InputFile]];
            Print["  Model output directory: ", FSOutputDir];
 

### 2017-03-15 doc/flexiblesusy-paper.tex

@@
 \begin{lstlisting}[language=bash]
-$ ./createmodel --name=MSSM
-$ ./configure --with-models=MSSM
+$ ./createmodel --name=CMSSM
+$ ./configure --with-models=CMSSM
 $ make
@@
 %
-The first command creates the model directory \code{models/MSSM/}
+The first command creates the model directory \code{models/CMSSM/}
 together with a CMSSM model file accompanied by
@@
 \begin{lstlisting}[language=bash]
-$ cd models/MSSM 
-$ ./run_MSSM.x --slha-input-file=LesHouches.in.MSSM
+$ cd models/CMSSM
+$ ./run_CMSSM.x --slha-input-file=LesHouches.in.CMSSM
 \end{lstlisting}
@@
 parameter point given in the SLHA input file
-\code{LesHouches.in.MSSM}.  Afterwards, the pole mass
+\code{LesHouches.in.CMSSM}.  Afterwards, the pole mass
 spectrum and mixing matrices are calculated and written to the standard
@@
 
-LowScale = SM[MZ];
-
-LowScaleFirstGuess = SM[MZ];
+LowScale = LowEnergyConstant[MZ];
+
+LowScaleFirstGuess = LowEnergyConstant[MZ];
 
@@
 InitialGuessAtLowScale = {
-   {vd, SM[vev] Cos[ArcTan[TanBeta]]},
-   {vu, SM[vev] Sin[ArcTan[TanBeta]]},
+   {vd, LowEnergyConstant[vev] Cos[ArcTan[TanBeta]]},
+   {vu, LowEnergyConstant[vev] Sin[ArcTan[TanBeta]]},
    {Yu, Automatic},
@@
 FSMinimize[{vd,vu},
-           (SM[MZ] - Pole[M[VZ]])^2 / STANDARDDEVIATION[MZ]^2 +
-           (SM[MH] - Pole[M[hh[1]]])^2 / STANDARDDEVIATION[MH]^2]
+           (LowEnergyConstant[MZ] - Pole[M[VZ]])^2 / STANDARDDEVIATION[MZ]^2 +
+           (LowEnergyConstant[MH] - Pole[M[hh[1]]])^2 / STANDARDDEVIATION[MH]^2]
   \end{lstlisting}
@@
     \chi^2(v_d,v_u) =
-    \frac{(\texttt{SM[MZ]}-m_Z^\pole)^2}{\sigma_{m_Z}^2} +
-    \frac{(\texttt{SM[MH]}-m_{h_1}^\pole)^2}{\sigma_{m_h}^2}
+    \frac{(\texttt{LowEnergyConstant[MZ]}-m_Z^\pole)^2}{\sigma_{m_Z}^2} +
+    \frac{(\texttt{LowEnergyConstant[MH]}-m_{h_1}^\pole)^2}{\sigma_{m_h}^2}
   \end{align}
   %
-  is minimal.  The constants \code{SM[MZ]}, \code{SM[MH]},
+  is minimal.  The constants \code{LowEnergyConstant[MZ]}, \code{LowEnergyConstant[MH]},
   $\sigma_{m_Z}$ and $\sigma_{m_h}$ are defined in
@@
   \begin{align}
-    \texttt{SM[MZ]} &= 91.1876, &
-    \texttt{SM[MH]} &= 125.9, \\
+    \texttt{LowEnergyConstant[MZ]} &= 91.1876, &
+    \texttt{LowEnergyConstant[MH]} &= 125.9, \\
     \sigma_{m_Z} &= 0.0021, &
@@
 FSFindRoot[{vd,vu},
-           {SM[MZ] - Pole[M[VZ]], SM[MH] - Pole[M[hh[1]]]}]
+           {LowEnergyConstant[MZ] - Pole[M[VZ]], LowEnergyConstant[MH] - Pole[M[hh[1]]]}]
   \end{lstlisting}
@@
     \begin{pmatrix}
-      \texttt{SM[MZ]} - m_Z^\pole \\
-      \texttt{SM[MH]} - m_{h_1}^\pole
+      \texttt{LowEnergyConstant[MZ]} - m_Z^\pole \\
+      \texttt{LowEnergyConstant[MH]} - m_{h_1}^\pole
     \end{pmatrix}
@@
 \appendix
+\renewcommand{\thesection}{\Alph{section}}
 
@@
     6   2                    # beta-functions loop order
-    7   1                    # threshold corrections loop order
+    7   2                    # threshold corrections loop order
     8   1                    # Higgs 2-loop corrections O(alpha_t alpha_s)

### 2017-03-15 meta/FlexibleSUSY.m
Fixed merge issue

@@ -1466,25 +1466,15 @@ WriteObservables[extraSLHAOutputBlocks_, files_List] :=
 (* Write the GMM2 c++ files *)
 WriteGMuonMinus2Class[vertexRules_List, files_List] :=
     Module[{particles, muonFunctionPrototypes, diagrams, vertexFunctionData,
-<<<<<<< HEAD
-        definitions, calculationCode},
+        definitions, calculationCode, getMSUSY, getQED2L},
            particles = GMuonMinus2`GMuonMinus2CreateParticles[];
            muonFunctionPrototypes = GMuonMinus2`GMuonMinus2CreateMuonFunctions[vertexRules][[1]];
            diagrams = GMuonMinus2`GMuonMinus2CreateDiagrams[];
            vertexFunctionData = GMuonMinus2`GMuonMinus2CreateVertexFunctionData[vertexRules];
            definitions = GMuonMinus2`GMuonMinus2CreateDefinitions[vertexRules];
            calculationCode = GMuonMinus2`GMuonMinus2CreateCalculation[];
-=======
-        definitions, calculationCode, getMSUSY, getQED2L},
-           particles = GMuonMinus2`CreateParticles[];
-           muonFunctionPrototypes = GMuonMinus2`CreateMuonFunctions[vertexRules][[1]];
-           diagrams = GMuonMinus2`CreateDiagrams[];
-           vertexFunctionData = GMuonMinus2`CreateVertexFunctionData[vertexRules];
-           definitions = GMuonMinus2`CreateDefinitions[vertexRules];
-           calculationCode = GMuonMinus2`CreateCalculation[];
            getMSUSY = GMuonMinus2`GetMSUSY[];
            getQED2L = GMuonMinus2`GetQED2L[];
->>>>>>> feature-2.0
 
            WriteOut`ReplaceInFiles[files,
                                    { "@GMuonMinus2_Particles@"               -> particles,
@@ -2230,7 +2220,7 @@ Options[MakeFlexibleSUSY] :=
 
 MakeFlexibleSUSY[OptionsPattern[]] :=
     Module[{nPointFunctions, runInputFile, initialGuesserInputFile,
-            gmm2Vertices = {},
+            gmm2Vertices = {}, edmFields,
             susyBetaFunctions, susyBreakingBetaFunctions,
             numberOfSusyParameters, anomDim,
             inputParameters (* list of 3-component lists of the form {name, block, type} *),
[…]

### 2017-10-10 doc/flexiblesusy-2.0.tex
adding paper draft

\documentclass[final,3p,11pt,pdflatex]{elsarticle}
\usepackage[utf8x]{inputenc}      % input font encoding
\usepackage{amsmath,amssymb,mathtools}
\usepackage[T1]{fontenc}          % output font encoding
\usepackage{booktabs,tabularx}
\usepackage{graphicx,subfig}
\usepackage{xspace}
\usepackage[usenames]{xcolor}\definecolor{fscolor}{RGB}{44,118,255}
\usepackage{tikz,tikz-uml}
\usetikzlibrary{arrows,calc,matrix,positioning,shapes}
\usepackage{listings}
\usepackage[absolute]{textpos}
\usepackage[many]{tcolorbox}
\usepackage{xparse}
\usepackage[font=small,labelfont=bf,format=plain,margin=0.05\textwidth]{caption}
\usepackage{bbm}
\bibstyle{elsarticle-num}
% source code highlighting
\lstset{breaklines=true,
  breakatwhitespace=true,
  stepnumber=1,
  basicstyle=\ttfamily\footnotesize,
  commentstyle=\ttfamily\color{gray},
  prebreak={\textbackslash},
  breakindent=10pt,
  breakautoindent=false,
  showspaces=false,
  showstringspaces=false,
  frame=shadowbox,
  rulesepcolor=\color{gray},
  rulesep=0.1em,
  abovecaptionskip=0em,
  aboveskip=1.5em,
  belowcaptionskip=0.5em,
  belowskip=1em,
}

\geometry{top=2.2cm,left=3cm,right=3cm,bottom=4cm,footskip=3em}

\usepackage[pdftitle={FlexibleSUSY 2.0: Extensions to investigate the phenomenology of
  SUSY and non-SUSY models},
  pdfauthor={Peter Athron, Markus Bach, Dylan Harries, Thomas Kwasnitza, Jae-hyeon Park,
    Dominik Stockinger, Alexander Voigt, Jobst Ziebell},
  pdfkeywords={FlexibleSUSY,supersymmetry,spectrum,generator,MSSM,NMSSM,E6SSM},
  bookmarks=true, linktocpage, colorlinks=true, allbordercolors=white,
  allcolors=fscolor]{hyperref}
\biboptions{sort&compress}
\allowdisplaybreaks

%macros
\newcommand{\modelname}[1]{\texttt{#1}\@\xspace}
\newcommand{\sarah}{\texttt{SARAH}\@\xspace}
\newcommand{\spheno}{\texttt{SPheno}\@\xspace}
\newcommand{\suspect}{\texttt{SuSpect}\@\xspace}
\newcommand{\isasusy}{\texttt{ISASUSY}\@\xspace}
\newcommand{\suseflav}{\texttt{SuSeFLAV}\@\xspace}
\newcommand{\nmspec}{\texttt{NMSPEC}\@\xspace}
\newcommand{\fs}{\texttt{FlexibleSUSY}\@\xspace}
\newcommand{\fsbreak}{\texttt{Flex\-ib\-le\-SUSY}\@\xspace}
\newcommand{\HSSUSY}{\modelname{HSSUSY}}
\newcommand{\susyhd}{\texttt{SusyHD}\@\xspace}
\newcommand{\MhEFT}{\texttt{MhEFT}\@\xspace}
\newcommand{\softsusy}{\texttt{SOFTSUSY}\@\xspace}
\newcommand{\micromegas}{\texttt{micrOMEGAS}\@\xspace}
\newcommand{\darksusy}{\texttt{DarkSUSY}\@\xspace}
\newcommand{\calchep}{\texttt{CALCHEP}\@\xspace}
\newcommand{\lanhep}{\texttt{LANHEP}\@\xspace}
\newcommand{\madgraph}{\texttt{MadGraph}\@\xspace}
\newcommand{\whizard}{\texttt{WHIZARD}\@\xspace}
\newcommand{\sherpa}{\texttt{SHERPA}\@\xspace}
\newcommand{\FAFC}{\texttt{FeynArts}\@\xspace}
\newcommand{\helac}{\texttt{HELAC}\@\xspace}
\newcommand{\herwig}{\texttt{HERWIG}\@\xspace}
\newcommand{\pythia}{\texttt{PYTHIA}\@\xspace}
\newcommand{\checkmate}{\texttt{CheckMate}\@\xspace}
\newcommand{\madanalysis}{\texttt{MadAnalysis}\@\xspace}
\newcommand{\smodels}{\texttt{SModelS}\@\xspace}
\newcommand{\fastlim}{\texttt{Fastlim}\@\xspace}
\newcommand{\colliderbit}{\texttt{ColliderBit}\@\xspace}
\newcommand{\higgsbounds}{\texttt{HiggsBounds}\@\xspace}
\newcommand{\higgssignals}{\texttt{HiggsSignals}\@\xspace}
\newcommand{\htm}{\texttt{H3m}\@\xspace}
\newcommand{\Himalaya}{\texttt{Himalaya}\@\xspace}
\newcommand{\FeynHiggs}{\texttt{FeynHiggs}\@\xspace}
\newcommand{\FeynRules}{\texttt{FeynRules}\@\xspace}
\newcommand{\NMSSMCalc}{\texttt{NMSSMCalc}\@\xspace}
\newcommand{\multinest}{\texttt{MultiNest}\@\xspace}
\newcommand{\CPsuperH}{\texttt{CPsuperH}\@\xspace}
\newcommand{\GAMBIT}{\texttt{GAMBIT}\@\xspace}
\newcommand{\GMTCalc}{\texttt{GM2Calc}\@\xspace}
\newcommand{\fsone}{\fs 1.0\@\xspace}
\newcommand{\fstwo}{\fs 2.0\@\xspace}
\newcommand{\fsh}{\texttt{FlexibleSUSY+}\Himalaya\xspace}
\newcommand{\fbsm}{\texttt{FlexibleBSM}\@\xspace}
\newcommand{\fcpv}{\texttt{FlexibleCPV}\@\xspace}
\newcommand{\fmw}{\texttt{FlexibleMW}\@\xspace}
\newcommand{\famu}{\texttt{FlexibleAMU}\@\xspace}
\newcommand{\fsas}{\texttt{FlexibleSAS}\@\xspace}
\newcommand{\feft}{\texttt{Flex\-ib\-le\-EFT\-Higgs}\@\xspace}
\newcommand{\mathematica}{\texttt{Ma\-the\-ma\-ti\-ca}\xspace}
\newcommand{\LibraryLink}{\texttt{LibraryLink}\xspace}
\newcommand{\ESSM}{E$_6$SSM\@\xspace}
\newcommand{\code}[1]{\lstinline|#1|}  % inline source code
\newcommand{\ol}[1]{\overline{#1}}
\newcommand{\MSbar}{\ensuremath{\ol{\text{MS}}}\xspace}
\newcommand{\DRbar}{\ensuremath{\ol{\text{DR}}}\xspace}
\newcommand{\MDRbar}{\ensuremath{\ol{\text{MDR}}}\xspace}
\newcommand{\unit}[1]{\,\text{#1}}      % units
\newcommand{\userinput}{\text{input}}
\newcommand{\pole}{\text{pole}}
\newcommand{\Lagr}{\mathcal{L}}
\newcommand{\SM}{\ensuremath{\text{SM}}\xspace}
\newcommand{\BSM}{\ensuremath{\text{BSM}}\xspace}
\newcommand{\THDM}{\ensuremath{\text{THDM}}\xspace}
\newcommand{\THDMII}{\ensuremath{\text{THDM-II}}\xspace}
\newcommand{\MSSM}{\ensuremath{\text{MSSM}}\xspace}
\newcommand{\MS}{\ensuremath{M_S}\xspace}
\newcommand{\MEWSB}{\ensuremath{M_\text{EWSB}}\xspace}
\newcommand{\Qmatch}{\ensuremath{Q_\text{match}}}
\newcommand{\QpoleBSM}{\ensuremath{Q_\text{pole,BSM}}}
\newcommand{\QpoleSM}{\ensuremath{Q_\text{pole,SM}}}
\newcommand{\QED}{\ensuremath{\text{QED}}}
\newcommand{\unity}{\mathbf{1}}
\newcommand{\amu}{\ensuremath{a_\mu}\xspace}
\newcommand{\amuBSM}{\ensuremath{\amu^{\BSM}}\xspace}
\newcommand{\amuMSSM}{\ensuremath{\amu^{\MSSM}}\xspace}
\newcommand{\edm}[1]{\ensuremath{d_{#1}}\xspace}
\newcommand{\edmBSM}[1]{\ensuremath{\edm{#1}^{\BSM}}\xspace}
\newcommand{\deltaVB}{\ensuremath{\delta_{\text{VB}}}\xspace}
\newcommand{\figref}[1]{\figurename~\ref{#1}}
\newcommand{\secref}[1]{Section~\ref{#1}}
\newcommand{\appref}[1]{Appendix~\ref{#1}}
\newcommand{\tabref}[1]{\tablename~\ref{#1}}
\newcommand{\exref}[1]{Example~\ref{#1}}
\newcommand{\ptitle}[1]{\emph{#1}}
\renewcommand{\ptitle}[1]{}
\newcommand{\Zv}{\mathbf{\backslash}\mkern-11.0mu{Z}}
\newcommand{\scoeff}[2]{[ c^{#1}_{#2}(Q) ]}
\newcommand{\barlog}{\overline{\log}}
\newcommand{\GSM}{\ensuremath{G_{\SM}}\xspace}
\newcommand{\CP}{\ensuremath{CP}\xspace}
\newcommand{\azero}{\ensuremath{A_0}\xspace}
\newcommand{\mhalf}{\ensuremath{M_{1/2}}\xspace}
\newcommand{\mzero}{\ensuremath{m_0}\xspace}
\newcommand{\SQCD}{SUSY-QCD\xspace}

\newcommand{\multilinecell}[2][c]{%
  \begin{tabular}[#1]{@{}c@{}}#2\end{tabular}}

\newtcolorbox[
  auto counter
]{example}[1][]{%
  breakable,
  before skip=1em,
  after skip=1em,
  enhanced,
  colback=white,
  colbacktitle=white,
  arc=0pt,
  leftrule=1.5pt,
  rightrule=0.5pt,
  toprule=0.5pt,
  bottomrule=0.5pt,
  titlerule=0.5pt,
  colframe=fscolor,
  fonttitle=\bfseries\normalcolor,
  overlay={},
  parbox=false,
  title=Example~\thetcbcounter,
  #1
}

% comments
\newcommand{\avnote}[1]{\textcolor{blue}{AV: #1}}
\newcommand{\dhnote}[1]{\textcolor{green}{DH: #1}}
\newcommand{\panote}[1]{\textcolor{red}{PA: #1}}
\newcommand{\jpnote}[1]{\textcolor{magenta}{JP: #1}}
\newcommand{\mbnote}[1]{\textcolor{orange}{MB: #1}}

\NewDocumentEnvironment{OptionTable}{m m O{llXX}}{%
\table[tbh!]
  \tabularx{\textwidth}{#3}%
    \toprule
    Symbol & Default value & Allowed values & Description \\
    \midrule
}{\endtabularx\caption{#1}\label{#2}\endtable}

\makeatletter
\lstnewenvironment{numlstlisting}[1][]{%
  \lstset{%
    #1,
    numbers=left,
    firstnumber=auto,
    numberstyle=\tiny\sffamily}%
  \csname\@lst @SetFirstNumber\endcsname
}{%
  \csname \@lst @SaveFirstNumber\endcsname
}
\makeatother
\DeclareMathOperator{\diag}{diag}
\DeclareMathOperator{\sign}{sign}
\DeclareMathOperator{\re}{Re}
\DeclareMathOperator{\im}{Im}
\def\at{\alpha_t}
\def\ab{\alpha_b}
\def\as{\alpha_s}
\def\atau{\alpha_{\tau}}
\def\aem{\alpha_{\text{em}}}
\def\oat{O(\at)}
\def\oab{O(\ab)}
\def\oatau{O(\atau)}
\def\oatab{O(\at\ab)}
\def\oatas{O(\at\as)}
\def\oabas{O(\ab\as)}
\def\oatababq{O(\at\ab + \ab^2)}
\def\oatqatababq{O(\at^2 + \at\ab + \ab^2)}
\def\oatasatq{O(\at\as + \at^2)}
\def\oatasabas{O(\at\as +\ab\as)}
\def\oatasabasatq{O(\at\as + \at^2 +\ab\as)}
\def\oatq{O(\at^2)}
\def\oabq{O(\ab^2)}
\def\oatauq{O(\atau^2)}
\def\oabatau{O(\ab \atau)}
\def\oatplusabsq{O((\at+\ab)^2)}
\def\oas{O(\as)}
\def\oatauqatab{O(\atau^2 +\ab \atau )}

\journal{Computer Physics Communications}
\begin{document}
\begin{frontmatter}
 \vspace*{0.5cm}
 \title{\Large\bf FlexibleSUSY 2.0: Extensions to investigate the phenomenology of SUSY and non-SUSY models}

\author[Monash]{Peter Athron}
\author[dresden]{Markus Bach}
\author[adelaide,prague]{Dylan Harries}
\author[dresden]{Thomas Kwasnitza}
\author[kiasquc]{Jae-hyeon Park}
\author[dresden]{Dominik St\"ockinger}
\author[rwth]{Alexander Voigt\corref{cor1}}
\ead{Alexander.Voigt@physik.rwth-aachen.de}
\cortext[cor1]{Corresponding author}
\author[dresden]{Jobst Ziebell}
\address[Monash]{ARC Centre of Excellence for Particle Physics at
  the Terascale, School of Physics, Monash University, Melbourne,
  Victoria 3800, Australia}
\address[dresden]{Institut f\"ur Kern- und Teilchenphysik,
TU Dresden, Zellescher Weg 19, 01069 Dresden, Germany}
\address[adelaide]{ARC Centre of Excellence for Particle Physics at
the Terascale, Department of Physics, The University of Adelaide,
Adelaide, South Australia 5005, Australia}
\address[prague]{Institute of Particle and Nuclear Physics, Faculty of
  Mathematics and Physics, Charles University in Prague, V
  Hole\v{s}ovi\v{c}k\'{a}ch 2, 180 00 Praha 8, Czech Republic}
\address[kiasquc]{Quantum Universe Center,
Korea Institute for Advanced Study,
85 Hoegiro Dongdaemungu,
Seoul 02455, Republic of Korea}
\address[rwth]{Institute for Theoretical Particle Physics and Cosmology, RWTH Aachen University, 52074 Aachen, Germany}

  \begin{abstract}
    We document major new features and improvements
    of \fs, a \mathematica and
    C++ package that generates fast and precise spectrum
    generators. These extensions significantly increase the generality
    and capabilities of the \fs package, while maintaining the elegant
    structure and easy to use interfaces. The \fbsm extension makes it
    possible to also create spectrum generators for non-supersymmetric
    extensions of the Standard Model. The \fcpv extension adds the
    option of complex parameters to the spectrum generators, allowing
    the study of many interesting models with new sources of \CP
    violation. \fmw computes the decay of the muon for the generated
    model and thereby allows \fs to predict the mass of the $W$ boson
    from the input parameters by using the more precise electroweak
    input of $\{ G_F, M_Z, \aem \}$ instead of $\{ M_W, M_Z,
    \aem \}$. The \famu extension provides a calculator of the
    anomalous magnetic moment of the muon in any model \fs can
    generate a spectrum for.  \fsas introduces a new solver for the
    boundary value problem which makes use of semi-analytic expressions
    for dimensionful parameters to find solutions in models
    where the classic two-scale solver will not work such as the constrained
    E$_6$SSM\@. \feft is a hybrid calculation of the Higgs mass which
    combines the virtues of both effective field theory calculations
    and fixed-order calculations.  All of these extensions are
    included in \fstwo, which is released simultaneously with this
    manual.
  \end{abstract}

\begin{keyword}
sparticle,
supersymmetry,
Higgs,
renormalization group equations
\PACS 12.60.Jv
\PACS 14.80.Ly
\end{keyword}
\end{frontmatter}

% report numbers
\begin{textblock*}{10em}(\textwidth,1.5cm)
\raggedleft\noindent\footnotesize
KIAS--Q17043 \\
TTK--17--31 \\
CoEPP--MN--17--16
\end{textblock*}

\clearpage
\newgeometry{top=2.5cm,left=3cm,right=3cm,bottom=4cm,footskip=3em}
\section*{Program Summary}
\noindent
{\em Program title:} \fs\\[0.5em]
{\em Program obtainable from:} \url{http://flexiblesusy.hepforge.org/}\\[0.5em]
{\em Distribution format:} tar.gz\\[0.5em]
{\em Programming language:} C++, Wolfram/Mathematica, FORTRAN, Bourne shell\\[0.5em]
{\em Computer:} Personal computer\\[0.5em]
{\em Operating system:} Tested on FreeBSD, Linux, Mac OS X, Windows/Cygwin\\[0.5em]
{\em External routines:} \sarah, Boost library, Eigen library, GNU Scientific Library\\[0.5em]
{\em Optional external routines:} BLAS, Himalaya, LAPACK, LoopTools, SQLite, TSIL\\[0.5em]
{\em Typical running time:} 0.02--0.2 seconds per parameter point\\[0.5em]
{\em Nature of problem:}
%
Determining the mass spectrum, mixings and further observables for an arbitrary extension
of the Standard Model, input by the user. The generated code must find
simultaneous solutions to constraints that are specified at two or
more different renormalization scales, which are connected by
renormalization group equations forming a large set of coupled
first-order differential
equations. \\[0.5em]
%
{\em Solution method:}
%
Nested iterative algorithm and numerical
minimization of the Higgs potential.\\[0.5em]
%
{\em Restrictions:}
%
The couplings must remain perturbative at all scales between the
highest and lowest boundary condition.  Tensor-like Lagrangian
parameters of rank 3 are currently not supported.  The automatic
determination of the Standard Model-like gauge and Yukawa couplings is
only supported for models that have the Standard Model gauge group
$SU(3)_C\times SU(2)_L\times U(1)_Y$ as a gauge symmetry group factor.
However, due to the modular nature of the generated code, adapting and
extending it to overcome restrictions in scope is quite straightforward.

\clearpage
\tableofcontents

\newpage
\section{Introduction}

Popular and well studied new physics extensions of the Standard Model
(SM) are coming under increasing pressure from the searches
at the Large Hadron Collider (LHC) as well as other current experiments.
At the same time, there remain many outstanding physics problems that are not
solved by the SM and require new physics.  For example, the origin of dark
matter is still unexplained and it requires some new particle
to fit the relic density.  In addition, the gauge structure and
fractional charges of the SM have no explanation, but some
observations hint at the possibility of a grand unified theory (GUT)
from which these can be derived after the breakdown of the GUT gauge
group.  Finally, the stability of the weak scale, which is $17$ orders
of magnitude smaller than the Planck scale, has no explanation within
the SM and has been the driving motivation for the construction and study
of concepts that go beyond the Standard Model (BSM), most notably supersymmetry
(SUSY) \cite{Weinberg:1975gm, Weinberg:1979bn, Gildener:1976ai,Susskind:1978ms,
  tHooft:1980xss}.

These issues strongly motivate the development of new ideas and new
models.  However, in principle there are a huge number of models that
can solve some, or all, of these problems and many more that may be
motivated by other principles that have not yet gained widespread
interest in high energy physics (HEP)\@. Previously, most work has
been done on the simplest variants, on scenarios that are easiest to
test and on those that individual researchers consider to be the very
best motivated models. Examples include the Minimal Supersymmetric
Standard Model\footnote{For a review see Ref.\ \cite{Chung:2003fi}.}
(MSSM), scalar singlet dark matter models (SSDM)
\cite{Silveira:1985rk,McDonald:1993ex,Burgess:2000yq}, type II two
Higgs doublet models (THDM-II) \cite{Lee:1973iz,Glashow:1976nt,
  Donoghue:1978cj}, and universal extra dimensions
\cite{Appelquist:2000nn}.  Now, it is becoming increasingly well
motivated to also examine more complicated model variants, explore
scenarios that are calculationally difficult or hard to observe, or
motivated from an entirely new perspective.  Such advanced models can
avoid phenomenological or conceptual difficulties of simpler models,
and might provide explanations for the lack of experimental evidence
for new physics at the LHC.

Faced with this challenge, it is important to reduce
the calculational hurdle required to explore a new model and
look at its phenomenology as much as possible. This makes it easier for
new ideas to get developed and for many more models to
be studied together for much more general conclusions.

\fs \cite{Athron:2014yba,Athron:2016fuq} already made a significant
push in this direction, allowing the automatic creation of a spectrum
generator for a very wide range of supersymmetric models.  \fs uses
\sarah \cite{Staub:2010ty,Staub:2009bi,Staub:2010jh,Staub:2012pb,
  Staub:2013tta} to obtain \mathematica expressions for model dependent
components, the 2-loop renormalization group equations (RGEs),
1-loop self energies, 1-loop tadpoles, mass matrices and
electroweak symmetry breaking (EWSB) conditions. \fs then translates these
expressions into C++ routines and embeds them inside a code structure
for solving the boundary value problem (BVP)\@.  It also uses some numerical
routines from \softsusy \cite{Allanach:2001kg,Allanach:2013kza} and is
heavily unit-tested against the MSSM and Next-to-MSSM versions of
\softsusy every night to avoid bugs.

Spectrum generators determine the pole masses and couplings of a
particular model from assumptions about the parameters at some
high scale, such as the grand unification scale, or directly from
on-shell parameters or running parameters\footnote{Usually defined in
  the \MSbar or \DRbar scheme.} given at the new physics scale.
This requires computing self energies, tadpole corrections to the EWSB
conditions and threshold corrections to SM-like gauge and Yukawa
couplings.  In the case of high scale assumptions a BVP must be
solved, which additionally requires integrating the RGEs for the model.
These codes are essential for testing hypotheses about the parameters of
the model, and determining if they lead to phenomenologically viable masses.

Spectrum generators are widely used in studies of supersymmetry, and for the
MSSM and the Next-to-MSSM\footnote{For a review of the model see
  Refs.\ \cite{Ellwanger:2009dp, Maniatis:2009re}.} (NMSSM) there are
a number of them that solve the BVP for particular choices of breaking
mechanism inspired boundary conditions (\softsusy
\cite{Allanach:2001kg,Allanach:2009bv, Allanach:2011de,
  Allanach:2013kza, Allanach:2014nba,
  Allanach:2016rxd,Allanach:2017hcf}, \suspect \cite{Djouadi:2002ze},
\spheno \cite{Porod:2003um,Porod:2011nf}, \isasusy \cite{Baer:1993ae},
\nmspec \cite{Ellwanger:2004xm,Ellwanger:2005dv, Ellwanger:2006rn} and
\suseflav \cite{Chowdhury:2011zr}) and several more that start from \DRbar
parameters at the SUSY scale or on-shell inputs (\FeynHiggs
\cite{Heinemeyer:1998yj, Heinemeyer:1998np, Degrassi:2002fi,
  Frank:2006yh, Hahn:2013ria, Drechsel:2016jdg, Bahl:2016brp,
  Bahl:2017aev}, \NMSSMCalc \cite{Baglio:2013iia, King:2015oxa} and
\CPsuperH \cite{Lee:2003nta, Lee:2007gn, Lee:2012wa}). This means that
the parameter space of the MSSM (and to a lesser extent the NMSSM) has
been extensively explored in a fast and reliable way.

However, for other models few, if any, public software packages exist.
Until recently this meant that one would have to spend a very long
time writing and testing code in order to explore the parameter space
of the model and do phenomenological investigations, and even then
bugs are more likely if one uses a private tool rather than a well
tested public one.  With the recent development of \fs and generated
modules for \spheno from \sarah \cite{Porod:2003um,
  Porod:2011nf,Staub:2017jnp,Staub:2009bi,Staub:2010jh,Staub:2012pb,
  Staub:2013tta,Athron:2014yba} it is now possible to obtain spectrum
generators in a much wider range of models.  \fs can generate spectrum
generators of various kinds, with boundary conditions at some high
scale or purely with low-energy input.  This push towards calculators
that are not model specific has also happened for other major types of
analyses.  For example, one may create new models in \micromegas
\cite{Belanger:2001fz,
  Belanger:2004yn,Belanger:2006is,Belanger:2008sj,Belanger:2010gh,
  Belanger:2013oya,Belanger:2014vza} using \calchep \cite{Boos:1994xb,
  Belyaev:2012qa} and \lanhep \cite{Semenov:1996es,Semenov:1998eb,
  Semenov:2002jw,Semenov:2008jy, Semenov:2010qt} to study the relic
density and direct and indirect detection of dark matter. \FeynRules
\cite{Christensen:2008py,Alloul:2013bka} can
also be used to generate the Feynman rules after inputting the
Lagrangian, in a similar manner to \sarah and \lanhep. The output of
these codes can be used in matrix element generators, such as \calchep,
\madgraph
\cite{Stelzer:1994ta,Maltoni:2002qb,Alwall:2007st,Alwall:2011uj,
  Alwall:2014hca}, \whizard \cite{Moretti:2001zz,Kilian:2007gr},
\sherpa \cite{Gleisberg:2003xi,Gleisberg:2008ta,Hoeche:2012yf}, \FAFC
\cite{Hahn:1998yk,Hahn:2000kx,Hahn:2006zy,Hahn:2010zi,Hahn:2016ebn}
and \helac \cite{Kanaki:2000ey,Cafarella:2007pc}, with showering and
hadronization handled by the event generators \herwig
\cite{Bahr:2008pv, Bellm:2015jjp}, \pythia
\cite{Bengtsson:1987kr,Sjostrand:2014zea, Sjostrand:2006za} and
\sherpa. Results from collider experiments can be applied to different
BSM models using re-interpretation tools, with some examples being \higgsbounds
\cite{Bechtle:2008jh,Bechtle:2011sb,Bechtle:2013wla, Bechtle:2015pma},
\higgssignals \cite{HiggsSignals}, \checkmate
\cite{Drees:2013wra,Dercks:2016npn}, \smodels \cite{Kraml:2013mwa,
  Kraml:2014sna,Ambrogi:2017neo}, \fastlim \cite{Papucci:2014rja},
\madanalysis\ \cite{Conte:2012fm,Conte:2014zja,Dumont:2014tja}, and
the native re-interpretation of \colliderbit \cite{Balazs:2017moi}.
Finally, very recently \GAMBIT \cite{Athron:2017ard} has been released
which uses auto-generated code from \fs \cite{Workgroup:2017bkh} and
\micromegas \cite{Workgroup:2017lvb}, along with many other packages,
to perform global fits of a user implemented BSM model.

Indeed \fs has been used extensively to study new physics models, on
its own or in combination with some of the codes mentioned above, in a
large number of physics studies, see, e.g.,
Refs.\ \cite{Diessner:2014ksa,
  Athron:2014pua,Athron:2015tsa,Diessner:2015yna,Staub:2015aea,Athron:2015vxg,
  Diessner:2015iln,Bagnaschi:2015pwa,Staub:2016dxq,Athron:2016fuq,
  deFlorian:2016spz,Athron:2016gor,Athron:2016qqb,Drechsel:2016htw,
  Bagnaschi:2017xid,Athron:2017yua,Athron:2017kgt,Athron:2017qdc,
  Workgroup:2017bkh,Chakravarty:2017hcy,Harlander:2017kuc,Ellis:2017hdw}.
Nonetheless, in the first release of \fs there were still a number of
limitations that restricted the models and phenomenology that could be studied
and the precision of the calculations.

In this paper, we document extensions to \fs, all now available in
version 2.0, that make a substantial push further in expanding the
number of models that can be explored and the observables that can be
calculated within them.
Now the list of models that can be investigated with
\fstwo also includes non-supersymmetric models (\fbsm), those with
complex parameters (\fcpv) and, with the new BVP solver
\fsas, constrained versions of certain non-minimal SUSY models like the NMSSM
\cite{Ellwanger:2009dp, Maniatis:2009re,Ellwanger:1993xa,Elliott:1994ht,
  King:1995vk,Ellwanger:1995ru,Ellwanger:1996gw} and the Exceptional
Supersymmetric Standard Model (\ESSM)
\cite{King:2005jy,King:2005my,Athron:2010zz, Athron:2009ue,Athron:2009bs} that
cannot be solved using the two-scale BVP solver
\cite{Barger:1993gh} approach implemented in public spectrum
generators.
%
In addition, \fstwo extends its repertoire of calculations.
The user may now calculate the mass of the $W$ boson as a
prediction of the model (\fmw) with the Fermi
constant treated as an input, as well as the anomalous magnetic
moment of the muon (\famu) and some electric dipole moments as
an application of \fcpv. Furthermore, we document and update our \feft
calculation of the Higgs pole mass, the physics of which has been discussed
in Ref.~\cite{Athron:2016fuq}.  Besides these new physics features, we
also document the new \mathematica interface.

In \secref{sec:download} we give a quick start guide explaining how to
download and compile the code using basic options.  In
\secref{sec:Mathematica_interface} we present the \mathematica
interface of \fs's spectrum generators and in
\secref{sec:model_file_extensions} we describe the new model file
extensions.  The
following Sections \ref{sec:fbsm}--\ref{sec:feft} document
each of the major extensions to \fs in a
modular fashion so that readers
can easily skip to a particular section.  These extensions are
as follows.  \fbsm, documented in \secref{sec:fbsm}, allows \fs to work in
non-SUSY models as well as SUSY models; the calculation of the
anomalous magnetic moment of the muon is
given in \secref{sec:famu}; \fcpv, described in
\secref{sec:fcpv}, introduces complex parameters to \fs so that \CP
violating effects may be considered; \fmw calculates the
muon decay, allowing \fs to predict $M_W$, and
is documented in \secref{sec:fmw}; a new BVP solver \fsas is
introduced in \secref{sec:fsas} and finally the hybrid Higgs mass
calculation, \feft, which combines the benefits of both effective field theory
and fixed-order calculations is documented in \secref{sec:feft}.
After this, we briefly summarize remaining limitations of the code in
\secref{sec:limitations} before adding concluding remarks in
\secref{sec:conclusions}.
%
In the appendices we provide a reference of all input parameters and
configuration options of \fstwo as well as new models.

\section{Quick start}
\label{sec:download}

\subsection{Requirements}

The build process of a custom spectrum generator using \fs requires
the following:
%
\begin{itemize}
\item \mathematica, version 7 or higher
\item \sarah, version 4.11.0 or higher [\url{http://sarah.hepforge.org}]
\item C++11 compatible compiler (\texttt{g++} 4.8.5 or higher,
  \texttt{clang++} 3.8 or higher, \texttt{icpc} 15.0 or higher)
\item FORTRAN compiler (\texttt{gfortran}, \texttt{ifort} etc.)
\item Eigen library, version 3.1 or higher
  [\url{http://eigen.tuxfamily.org}]
\item Boost library, version 1.37.0 or higher
  [\url{http://www.boost.org}]
\item GNU scientific library [\url{http://www.gnu.org/software/gsl}]
\end{itemize}
%
Optional:
%
\begin{itemize}
\item an implementation of LAPACK [\url{http://www.netlib.org/lapack}]
  such as ATLAS [\url{http://math-atlas.sourceforge.net}] or
  Intel Math Kernel Library\\\ [\url{http://software.intel.com/intel-mkl}]
\item LoopTools, version 2.8 or higher
  [\url{http://www.feynarts.de/looptools}]
\end{itemize}

\subsection{Downloading \fs and generating a first spectrum generator}
\label{sec:quick-start-cmssm}
\fs is available as gzipped tarball on
[\url{http://flexiblesusy.hepforge.org}] or
[\url{https://github.com/FlexibleSUSY/FlexibleSUSY}] under the version
control system \texttt{git}.
%
To download and install \fstwo as a gzipped tarball run at the command
line:
%
\begin{lstlisting}[language=bash]
$ wget https://www.hepforge.org/archive/flexiblesusy/FlexibleSUSY-2.0.0.tar.gz
$ tar -xf FlexibleSUSY-2.0.0.tar.gz
$ cd FlexibleSUSY-2.0.0
\end{lstlisting}%% $
%
\fstwo is distributed with a huge selection of predefined models
including: the CMSSM (called \modelname{CMSSM}), the semi-constrained and
fully constrained NMSSM (\modelname{NMSSM}, \modelname{CNMSSM}), the
$R$-symmetric MSSM (\modelname{MRSSM}) \cite{Kribs:2007ac}, the NUHM \ESSM
(\modelname{E6SSM}) \cite{Athron:2007en}, high-scale SUSY (\HSSUSY)
\cite{Bagnaschi:2015pwa}, split-SUSY (\modelname{SplitMSSM})
\cite{Bagnaschi:2015pwa} and the THDM-II (\modelname{THDMII}).  See the
contents of \code{model_files/} for all predefined model files.

A spectrum generator for any of these models can be built with just
three commands.  For example, the CMSSM spectrum generator can be
created by running the following shell commands:
%
\begin{lstlisting}[language=bash]
$ ./createmodel --name=CMSSM
$ ./configure --with-models=CMSSM
$ make
\end{lstlisting}%% $
%
The \code{createmodel} command creates the model directory
\code{models/CMSSM/} where the code will be generated and adds the
CMSSM model file and an example SLHA input file to it.  The \code{configure}
script sets up the \code{Makefile}, checking the system requirements
and dependencies.  For more options see \code{./configure
  --help}.  The last command creates the code for the spectrum
generator of this model and compiles it.
%
This generated spectrum generator can then be run from the command
line as:
%
\begin{lstlisting}[language=bash]
$ cd models/CMSSM
$ ./run_CMSSM.x --slha-input-file=LesHouches.in.CMSSM \
                --slha-output-file=LesHouches.out.CMSSM
\end{lstlisting}
%
The spectrum generator reads the CMSSM input parameters from the SLHA input file
\code{LesHouches.in.CMSSM} and first solves a BVP to find a
set of running parameters at the \code{SUSYScale} that are consistent with all
boundary conditions specified in the model file, and then calculates the pole
masses, mixing matrices and potentially further observables.
The mass spectrum etc.\ obtained in this way is output
in SLHA format \cite{Skands:2003cj,Allanach:2008qq} as
\code{LesHouches.out.CMSSM}.  See \code{./run_CMSSM.x --help} for more
options.

\fs also provides a \mathematica interface, introduced in version 1.7.0,
to call the generated spectrum generators.  For each spectrum
generator, an example \mathematica script named
\code{models/<model>/run_<model>.m} is created for illustration.  For
example, the \modelname{CMSSM} spectrum generator can be called from within
\mathematica like this:
%
\begin{lstlisting}
Get["models/CMSSM/CMSSM_librarylink.m"];

handle = FSCMSSMOpenHandle[
    fsModelParameters -> {
        m0 -> 125,
        m12 -> 500,
        TanBeta -> 10,
        SignMu -> 1,
        Azero -> 0
    }
];

spectrum = FSCMSSMCalculateSpectrum[handle];
FSCMSSMCloseHandle[handle];

Print[spectrum];
\end{lstlisting}
%
Execute \code{?FSCMSSMOpenHandle} for a list of all allowed options and
input parameters.  In the example above, the \code{spectrum} variable
contains the mass spectrum and the running parameters in the form of a
list of replacement rules:
%
\begin{lstlisting}
{CMSSM -> {
   Pole[M[Glu]] -> 1147.35,
   Pole[M[Sd]] -> {957.993, 997.56, 1000.49, 1000.5, 1045.93, 1045.94},
   Pole[M[Sv]] -> {350.753, 351.913, 351.917},
   Pole[M[Su]] -> {796.653, 1002.67, 1003.96, 1005.06, 1043.07, 1043.07},
   Pole[M[Se]] -> {222.916, 229.983, 230.008, 360.842, 360.846, 361.978},
   Pole[M[hh]] -> {114.836, 713.119},
   Pole[M[Ah]] -> {88.5864, 712.848},
   Pole[M[Hpm]] -> {77.2642, 717.628},
   Pole[M[Chi]] -> {204.054, 385.012, 629.649, 643.612},
   Pole[M[Cha]] -> {385.017, 643.924},
   Pole[M[VWm]] -> 80.3935, ...}
}
\end{lstlisting}
%
More details about the \mathematica interface as well as a neat
example of running \HSSUSY through it can be found in Sections
\ref{sec:Mathematica_interface} and \ref{sec:application_HSSUSY},
respectively.

\subsection{Spectrum generators for alternative models}
\label{sec:quick-start-alternative-models}
If the user instead wants to create a spectrum generator for a model
for which there is no pre-existing model file distributed in
\fs, then the model file can be written. Before a \fs
model file can be written for the spectrum generator, there must exist
\sarah model files, which \fs uses to obtain model dependent
information. \sarah also comes with many pre-defined models, but if an
appropriate model is not available, the users may create their own
\sarah model files and add them to the directory
\code{sarah/<modelname>/} in \fs.  For the writing of a \sarah model
file we refer the reader to the extensive \sarah documentation, for
example Refs.\
\cite{Staub:2017jnp,Staub:2009bi,Staub:2010jh,Staub:2012pb,
  Staub:2013tta,Athron:2014yba}.

Creating a new \fs model file is straightforward. Full details are
given in the original \fs manual Ref.~\cite{Athron:2014yba}.
Here we just repeat a basic example
in the context of the NMSSM that illustrates the main points:
%
The semi-constrained NMSSM (\modelname{NMSSM}) distributed in \fs has
all the soft-breaking trilinear scalar couplings set to a unified
$\azero$ at the GUT scale.  However, often $A_\lambda$ and $A_\kappa$ are
taken to be non-universal in semi-constrained variants of the NMSSM
since the non-universality of the soft singlet mass already violates
the standard universality assumptions of constrained
models.\footnote{See \secref{sec:fsas} for a new approach that
  allows the fully constrained NMSSM to be solved.} To allow separate
values for $A_\lambda$ and $A_\kappa$ at the GUT scale the \fs model
file \code{model_files/NMSSM/FlexibleSUSY.m.in} should be changed from
%
\begin{lstlisting}[language=Mathematica]
EXTPAR = { {61, LambdaInput} };

HighScaleInput = {
  ...
  {T[\[Kappa]], Azero \[Kappa]},
  {T[\[Lambda]], Azero LambdaInput}
   ...
};
\end{lstlisting}
%
into
%
\begin{lstlisting}[language=Mathematica]
EXTPAR = { {61, LambdaInput},
           {63, ALambdaInput},
           {64, AKappaInput} };

HighScaleInput = {
  ...
  {T[\[Kappa]], AKappaInput \[Kappa]},
  {T[\[Lambda]], ALambdaInput LambdaInput},
  ...
};
\end{lstlisting}
%
The GUT scale values of $A_\lambda$ and $A_\kappa$ can then be specified in the SLHA input file in the \code{EXTPAR} block by entries $63$ and $64$,
%
\begin{lstlisting}
Block EXTPAR
   61   0.1                  # LambdaInput
   63   -100                 # ALambdaInput
   64   -300                 # AKappaInput

\end{lstlisting}

\section{\mathematica interface}
\label{sec:Mathematica_interface}

The spectrum generators created with \fs can be called from within
\mathematica.  To do that, first the spectrum generator must be built,
as described in \secref{sec:quick-start-cmssm}.  Afterwards, the
provided \mathematica interface functions for the model must be
loaded.  For a given model \code{<model>} this is done by including
the following file in the \mathematica session:
%
\begin{lstlisting}
Get["models/<model>/<model>_librarylink.m"];
\end{lstlisting}
%
This script loads the library
\code{models/<model>/<model>_librarylink.so} into the \mathematica
session (assuming the user is in the \code{FlexibleSUSY/} directory).
Afterwards, the \mathematica interface functions listed in
\tabref{tab:mma_ifce_functions} are available.
%
\begin{table}[tbh]
  \centering
  \begin{tabularx}{\textwidth}{lX}
    \toprule
    Function & Description \\

    \midrule \code{FS<model>OpenHandle[...]} & Takes all model input
    parameters as argument and returns a ``handle'' (a reference) to
    the given parameter point, the associated mass spectrum and observables.\\

    \code{FS<model>CloseHandle[handle]} & Releases the resources
    associated to a given handle.\\

    \code{FS<model>CalculateSpectrum[handle]} & Calculates the mass
    spectrum for a given handle.\\

    \code{FS<model>CalculateObservables[handle]} & Calculates the
    observables for a given handle.\\

    \code{FS<model>ToSLHA[handle]} & Returns a string containing the
    mass spectrum and observables associated to a given handle in SLHA format.\\

    \code{FS<model>Set[handle, ...]} & Changes the input parameters
    associated to a given handle.\\

    \code{FS<model>GetSettings[handle]} & Returns the spectrum generator
    settings (precision goal, loop orders, etc.) associated to
    a given handle.\\

    \code{FS<model>GetSMInputParameters[handle]} & Returns the SM
    input parameters associated to a given handle.\\

    \code{FS<model>GetInputParameters[handle]} & Returns the
    model-specific input parameters associated to a given handle.\\

    \code{FS<model>GetProblems[handle]} & Returns a list of problems
    that occurred when calculating the spectrum.\\

    \code{FS<model>GetWarnings[handle]} & Returns a list of warnings
    that occurred when calculating the spectrum.\\
    \bottomrule
  \end{tabularx}
  \caption{\mathematica interface functions provided for a \fs model with the name \code{<model>}.}
  \label{tab:mma_ifce_functions}
\end{table}
%

To run the spectrum generator for a given parameter point, a handle to that point must be
created first, using the \code{FS<model>OpenHandle[...]} function.
%
The returned handle represents a reference to the given parameter
point, the associated mass spectrum and observables.  The concept of
handles allows the user to calculate mass spectra for different
parameter points in parallel using multiple \mathematica kernels: Each
kernel can open a handle to a separate parameter point, calculate the
mass spectrum and finally close the handle.  In this way there
is no ambiguity in the parameter point used by each kernel.
\exref{ex:HSSUSY_parallel} in \secref{sec:application_HSSUSY} illustrates the usage of handles by
performing a parallel scan over the MSSM parameter space with \HSSUSY.
%
In the most general form, the \code{FS<model>OpenHandle[...]} function
can take the following three arguments:
%
\begin{lstlisting}
handle = FS<model>OpenHandle[
   fsSettings -> {...},
   fsSMParameters -> {...},
   fsModelParameters -> {...}
]
\end{lstlisting}
%
The \code{fsSettings} symbol can be used to set the spectrum generator
options.  All possible options are listed in
\tabref{tab:sg_mma_options} in \appref{app:fs_config_options}.
%
The \code{fsSMParameters} symbol can be used to set the SM
input parameters.  The possible SM input parameters are
listed in \tabref{tab:sm_mma_options} in \appref{app:sm_input_parameters}.
%
The \code{fsModelParameters} symbol can be used to set the \BSM
model-specific input parameters.  Unspecified model input parameters
are set to zero by default.  The names of the model input parameters
are identical to the ones specified in the \code{MINPAR},
\code{IMMINPAR}, \code{EXTPAR}, \code{IMEXTPAR} and
\code{FSAuxiliaryParameterInfo} variables in the \fs model file.  The
settings, the SM and the \BSM input parameters
associated to a given handle can be obtained using the
\code{FS<model>GetSettings[]}, \code{FS<model>GetSMInputParameters[]}
and \code{FS<model>GetInputParameters[]} functions, respectively.  The
opened handle can then be used to calculate the mass spectrum and the
observables:
%
\begin{lstlisting}
spectrum    = FS<model>CalculateSpectrum[handle];
observables = FS<model>CalculateObservables[handle];
FS<model>CloseHandle[handle];
\end{lstlisting}
%
Finally, the handle should be closed to release the associated
resources by calling the function \code{FS<model>CloseHandle[handle]}.
%
\begin{example}
  In the CMSSM, the \BSM model-specific input parameters are named as
  \code{m0}, \code{m12}, \code{TanBeta}, \code{SignMu} and
  \code{Azero}, see the \modelname{CMSSM} model file provided with \fstwo.
  Thus, an example \mathematica session for the CMSSM could look like:
%
\begin{lstlisting}
Get["models/CMSSM/CMSSM_librarylink.m"];

handle = FSCMSSMOpenHandle[
    fsSettings -> {
        poleMassLoopOrder -> 2,
        ewsbLoopOrder -> 2,
        thresholdCorrectionsLoopOrder -> 2,
        betaFunctionLoopOrder -> 3
    },
    fsSMParameters -> {
        Mt -> 173.34,
        alphaSMZ -> 0.1184
    },
    fsModelParameters -> {
        m0 -> 125,
        m12 -> 500,
        TanBeta -> 10,
        SignMu -> 1,
        Azero -> 0
    }
];

spectrum    = FSCMSSMCalculateSpectrum[handle];
observables = FSCMSSMCalculateObservables[handle];

FSCMSSMCloseHandle[handle];
\end{lstlisting}
\end{example}
%
The output of \code{FS<model>CalculateSpectrum[handle]} is a list
that contains the pole mass spectrum as well as the running masses
and parameters at the chosen output scale.
%
The running parameters are named as defined in the \sarah model.  For
example, \code{g1}, \code{g2}, \code{g3} usually denote the running
gauge couplings and \code{Yu}, \code{Yd}, \code{Ye} the running Yukawa
couplings.  The running masses are denoted as \code{M[<p>]}, where
\code{<p>} is the name of the particle as defined in the \sarah model
file.  All running parameters and masses are given at the parameter
output scale, \code{SCALE}.  The pole masses and mixing matrices carry
the additional \code{Pole[]} head.  For example, \code{Pole[M[hh]]}
usually denotes the pole mass(es) of the Higgs boson(s).

In the \mathematica output, the running parameters, the masses and
mixing matrices are defined in the \sarah convention, \emph{not} in the
SLHA convention.  This means in particular that the Yukawa matrices,
the soft-breaking squark mass matrices and the soft-breaking trilinear
couplings are \emph{not} defined in the (super)-CKM basis.  In the
\mathematica output, the particle masses are always non-negative and mixing
matrices are in general complex.
%
\begin{example}
  In the CMSSM, the output of \code{FSCMSSMCalculateSpectrum[handle]}
  may look like (skipping some entries for brevity):
%
\begin{lstlisting}
{
  CMSSM -> {
    M[Glu] -> 1116.4857717819132,
    M[Sd] -> {929.5770939936384, 963.6803089181217,
              965.7750791635142, 965.7786645820956,
              1010.5301444258299, 1010.5317308607175},
    M[Su] -> {770.2929836944288, 969.5389940603632,
              969.5446720936367, 975.4073015489867,
              1007.5435846398124, 1007.5443071514123},
    M[Se] -> {219.54939719808144, 226.45230058860656,
              226.4768407746941, 356.2452631868304,
              356.2501661663376, 357.5576772510361},
    M[hh] -> {88.16467333922309, 726.2603417238729},
    M[Ah] -> {90.09835220027803, 726.0229889725828},
    M[Hpm] -> {78.48914789145176, 730.2533306006959},
    M[Chi] -> {207.1963755793879, 375.7416364302936,
               627.5178023483583, 641.6676783271736},
    M[Cha] -> {375.56991892585705, 641.3578484531205},
    ...
    ZH -> {{0.10592570722508611, 0.9943740465985952},
           {0.9943740465985952, -0.10592570722508611}},
    ...
    Pole[M[Glu]] -> 1147.3536227374905,
    Pole[M[Sd]] -> {957.9934299811302, 997.5603867095314,
                    1000.4932601265115, 1000.4969819618583,
                    1045.9354429433467, 1045.9372472457565},
    Pole[M[Su]] -> {796.653619369782, 1002.6690741336473,
                    1003.9614916607435, 1005.0642702137084,
                    1043.0672831732345, 1043.067920812505},
    Pole[M[Se]] -> {222.90126096766593, 229.9832415178622,
                    230.00840279144913, 360.84198174065307,
                    360.8462569384804, 361.9798562942742},
    Pole[M[hh]] -> {114.83583179574276, 713.1187313487922},
    Pole[M[Ah]] -> {88.58641341930426, 712.8473602456997},
    Pole[M[Hpm]] -> {77.26414997655887, 717.6270868215212},
    Pole[M[Chi]] ->  {204.05370940499517, 385.0116889026496,
                      629.6500252267041, 643.6127224060953},
    Pole[M[Cha]] -> {385.0164604772902, 643.924798526633},
    Pole[ZH] -> {{0.1066307364997843, 0.9942986905520461},
                 {0.9942986905520461, -0.1066307364997843}},
    ...
    Yd -> {{0.00013999141660755535, 0., 0.},
           {0., 0.0030650771019348657, 0.},
           {0., 0., 0.1317656023934078}},
    Ye -> {{0.00002895462310608183, 0., 0.},
           {0., 0.005986898727723582, 0.},
           {0., 0., 0.1006931798596906}},
    Yu -> {{7.267255094462218*^-6, 0., 0.},
           {0., 0.0033082824973678956, 0.},
           {0., 0., 0.8606532901364391}},
    \[Mu] -> 624.160899893032,
    g1 -> 0.4679063156949638,
    g2 -> 0.6430285180350706,
    g3 -> 1.0655340318624051,
    vd -> 25.099612589273388,
    vu -> 242.8296409176676,
    T[Yd] -> {{-0.19442921534444055, 0., 0.},
              {0., -4.256965123994014, 0.},
              {0., 0., -171.13078241755716}},
    T[Ye] -> {{-0.008660431147847696, 0., 0.},
              {0., -1.790668166947902, 0.},
              {0., 0., -29.953008563262035}},
    T[Yu] -> {{-0.008251796354121698, 0., 0.},
              {0., -3.7564603713866354, 0.},
              {0., 0., -755.7309107649228}},
    B[\[Mu]] -> 53907.68839928095,
    mq2 -> {{1.0178399270038805*^6, 0., 0.},
            {0., 1.0178348758707164*^6, 0.},
            {0., 0., 865711.3590482193}},
    ml2 -> {{124853.43144557778, 0., 0.},
            {0., 124850.91737364854, 0.},
            {0., 0., 124143.31667980803}},
    mHd2 -> 109509.1756551005,
    mHu2 -> -377534.5544643501,
    md2 -> {{932089.8366766714, 0., 0.},
            {0., 932084.7501918572, 0.},
            {0., 0., 923097.970074961}},
    mu2 -> {{941294.0634724408, 0., 0.},
            {0., 941288.914981822, 0.},
            {0., 0., 639354.6597906639}},
    me2 -> {{49375.97115879859, 0., 0.},
            {0., 49370.8410356076, 0.},
            {0., 0., 47926.68179837005}},
    MassB -> 209.15138268684439,
    MassWB -> 387.9365053016521,
    MassG -> 1116.4857717819132,
    SCALE -> 866.8060753250803
  }
}
\end{lstlisting}
\end{example}
%
The function \code{FS<model>CalculateObservables[handle]} returns the
observables for a given handle in the form of a list of replacement rules.
%
\begin{example}
  In the CMSSM, the output of \code{FSCMSSMCalculateObservables[handle]}
  might look like
%
\begin{lstlisting}
{
  CMSSM -> {
    FlexibleSUSYObservable`aMuon -> 8.46316731749956*^-10,
    FlexibleSUSYObservable`CpHiggsPhotonPhoton ->
      {0.000029645107712580034 - 2.1094207517495443*^-7*I,
       7.984875731996049*^-7 + 9.125974009524432*^-7*I},
     FlexibleSUSYObservable`CpHiggsGluonGluon ->
       {-0.00006704582692080136 - 2.68821584064419*^-6*I,
        2.8335074218510527*^-6 + 4.966199588265877*^-6*I},
     FlexibleSUSYObservable`CpPseudoScalarPhotonPhoton ->
       1.066762880186648*^-6 - 8.27198259619208*^-7*I,
     FlexibleSUSYObservable`CpPseudoScalarGluonGluon ->
       6.825816917497379*^-6 + 8.151893730825134*^-7*I
  }
}
\end{lstlisting}
\end{example}
%
The symbol \code{aMuon} represents the anomalous magnetic moment of
the muon, $\amu$, calculated as described in \secref{sec:famu}.  The
symbols \code{CpHiggsPhotonPhoton}, \code{CpHiggsGluonGluon},
\code{CpPseudoScalarPhotonPhoton} and \code{CpPseudoScalarGluonGluon}
denote the effective couplings of the physical \CP-even and \CP-odd Higgs
boson(s) to two photons and gluons, respectively, as described in
Ref.~\cite{Staub:2016dxq}.  Note that if the \CP-even or \CP-odd Higgs states
are multiplets, as is the case for the \CP-even Higgs in the MSSM, for
example, the relevant couplings are calculated for all members of the
multiplet and the result is returned as a list.

The calculated spectrum can be printed in an SLHA-compatible format
using the \code{FS<model>ToSLHA[]} function.
%
\begin{example}
  For the CMSSM, an example output of \code{FSCMSSMToSLHA[handle]}
  could look like:
\begin{lstlisting}
Block SPINFO
     1   FlexibleSUSY
     2   2.0.0
     5   CMSSM
     9   4.11.0
Block MASS
   1000021     1.14735362E+03   # Glu
        24     8.03935152E+01   # VWm
   1000024     3.85016460E+02   # Cha(1)
   1000037     6.43924799E+02   # Cha(2)
        25     1.14835832E+02   # hh(1)
        35     7.13118731E+02   # hh(2)
...
\end{lstlisting}
\end{example}
%
In \secref{sec:application_HSSUSY} several examples can be found that
illustrate how to perform parameter scans and uncertainty
estimates using the \mathematica interface of \fs's spectrum
generators.

\section{\fs model file extensions}
\label{sec:model_file_extensions}

\subsection{Model-specific higher-order contributions}
\label{sec:model_specific_contributions}
To improve the accuracy in some specific models, \fs provides a few
model file switches to enable further higher-order contributions in
the calculation of the running parameters, the $\beta$ functions or
the Higgs pole mass.  \tabref{tab:fbsm_options} lists all switches
available in \fstwo, which are explained in the following subsections.
%
These switches can usually be enabled in many models, providing that
the corresponding requirements are fulfilled.  However, the user
should be aware that contributions may be missing if the switches are
enabled in models beyond their scope of application.

\begin{OptionTable}{\fs model file switches to enable/disable
    model-specific higher-order contributions in the SM,
    MSSM, NMSSM and split-MSSM.}{tab:fbsm_options}[lllX]
  %
  \multicolumn{4}{c}{\textit{SM}}\\
  \midrule
  \code{UseHiggs2LoopSM} & \code{False} & \code{True} or \code{False}
  & 2-loop contributions $O(\at^2 + \at \as)$ to $M_h$
  in
  the SM \\
  %
  \code{UseHiggs3LoopSM} & \code{False} & \code{True} or \code{False} &
  3-loop contributions $O(\at^3 + \at^2 \as \allowbreak + \at \as^2)$ to $M_h$ in
  the SM \\
  %
  \code{UseSM3LoopRGEs} & \code{False} & \code{True} or \code{False} &
  3-loop RGEs in the SM \\
  %
  \code{UseYukawa3LoopQCD} & \code{Automatic} & \multilinecell[t]{\code{True} or \code{False}\\ or \code{Automatic}} &
  2-loop and 3-loop QCD contributions $O(\as^2 + \as^3)$ to the \MSbar $y_t$ in the SM \\
  %
  \code{UseSMAlphaS3Loop} & \code{False} & \code{True} or \code{False}
  & 2-loop and 3-loop QCD threshold corrections
  $O(\as^2 + \as^3)$
  to $\as$ in the SM \\
  %
  \midrule
  \multicolumn{4}{c}{\textit{MSSM}}\\
  \midrule
  %
  \code{UseHiggs2LoopMSSM} & \code{False} & \code{True} or \code{False} &
  2-loop contributions $O((\at + \ab)\as \allowbreak + (\at +
  \ab)^2 \allowbreak + \atau^2)$ to $M_h$, $M_H$ and $M_A$ in
  the MSSM \\
  %
  \code{UseHiggs3LoopMSSM} & \code{False} & \code{True} or
  \code{False} & 3-loop contributions
  $O(\at\as^2 \allowbreak + \ab\as^2)$ to $M_h$ in
  the MSSM (requires \Himalaya) \\
  %
  \code{UseMSSM3LoopRGEs} & \code{False} & \code{True} or \code{False} &
  3-loop RGEs in the MSSM \\
  %
  \code{UseMSSMYukawa2Loop} & \code{False} & \code{True} or \code{False} &
  2-loop \SQCD contribution $O(\as^2)$ to the \DRbar $y_b$ and $y_t$ in the MSSM \\
  %
  \code{UseMSSMAlphaS2Loop} & \code{False} & \code{True} or
  \code{False} & 2-loop \SQCD contribution
  $O(\as^2 + \as\at + \as\ab)$
  to $\as$ in the MSSM \\
  %
  \midrule
  \multicolumn{4}{c}{\textit{NMSSM}}\\
  \midrule
  %
  \code{UseHiggs2LoopNMSSM} & \code{False} & \code{True} or \code{False} &
  2-loop contributions $O((\at + \ab)\as \allowbreak + (\at +
  \ab)^2 \allowbreak + \atau^2)$ to $M_{h_i}$ and $M_{A_i}$ in
  the NMSSM \\
  %
  \midrule
  \multicolumn{4}{c}{\textit{split-MSSM}}\\
  \midrule
  %
  \code{UseHiggs3LoopSplit} & \code{False} & \code{True} or \code{False} &
  3-loop contributions $O(\at\as^2)$ to $M_h$, $M_H$ and
  $M_A$ in the split-MSSM \\
  %
  \bottomrule
\end{OptionTable}

%%%%%%%%%%%%%%%%%%%%%% SM corrections %%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{SM-specific higher-order contributions}

\paragraph{2-loop and 3-loop contributions to the SM Higgs mass}

In the SM, 2-loop and leading 3-loop contributions to the Higgs pole
mass are known in the \MSbar scheme
\cite{Degrassi:2012ry,Martin:2014cxa,Vega:2015fna}.  In \fs the 2-loop
contributions of $O(\at^2 + \at \as)$
\cite{Degrassi:2012ry,Martin:2014cxa} and the 3-loop
$O(\at^3 + \at^2 \as + \at \as^2)$ contributions \cite{Martin:2014cxa}
can be taken into account to calculate the SM-like Higgs pole mass for
non-SUSY models with only one Higgs.  In order to enable these loop
contributions, the following switches must be set in the model file:
%
\begin{lstlisting}
UseHiggs2LoopSM = True;
UseHiggs3LoopSM = True;
\end{lstlisting}
%
The enabled 2-loop SM contributions of
$O(\at^2 + \at \as)$ to the Higgs mass read
%
\begin{align}
  (\Delta m_h^2)_{2L} &= (\Delta m_h^2)_{2L}^{(\at^2)} + (\Delta m_h^2)_{2L}^{(\at\as)} \,, \\
  (\Delta m_h^2)_{2L}^{(\at^2)} &=
  \frac{2 t}{(4\pi)^4} \left[
    -3 y_t^4 \left(3 \;\barlog^2(t) - 7 \;\barlog(t) + 2 + \frac{\pi^2}{3}\right)
  \right] , \label{eq:2L_Mh_atat}\\
  (\Delta m_h^2)_{2L}^{(\at\as)} &=
      \frac{g_3^2 y_t^2}{(4\pi)^4} \;\bigg[
        \frac{37}{3} p^2 - \frac{122 p^4}{135 t} - 4 (5 p^2 - 8 t) \;\barlog(t) -
        12 (p^2 - 8 t) \;\barlog^2(t) \bigg] \label{eq:2L_Mh_atas} \\
  &\stackrel{\mathmakebox[\widthof{=}]{p^2=0}}{=} \
  \frac{2 t}{(4\pi)^4} \left[
    16 g_3^2 y_t^2 \left(3 \;\barlog^2(t) + \barlog(t)\right)
  \right] \,.
\end{align}
%
In Eqs.~\eqref{eq:2L_Mh_atat} ff.\ the abbreviations
$t \equiv m_t^2$ and $\barlog(t) \equiv \log(t/Q^2)$ have been used
for brevity, analogous to the notation of Ref.~\cite{Martin:2014cxa}.  Note
that the momentum dependence is included in the 2-loop expression of
Eq.~\eqref{eq:2L_Mh_atas} in order to generate implicit 3-loop
$O(\at^2\as)$ terms.
%
The explicit 3-loop effective potential contributions of
$O(\at^3 + \at^2 \as + \at \as^2)$ to the
Higgs mass included by \fs read
%
\begin{align}
\begin{split}
  (\Delta m_h^2)_{3L} &=
  \frac{m_t^2}{(4\pi)^6} \Bigg[
     g_3^4 y_t^2 \left(248.122 + 839.197 \;\barlog(t) + 160 \;\barlog^2(t) - 736 \;\barlog^3(t)\right) \\
     &\phantom{=\frac{m_t^2}{(4\pi)^6} \Bigg[}
     + g_3^2 y_t^4 \left(2764.365 + 1283.716 \;\barlog(t) - 360 \;\barlog^2(t) + 240 \;\barlog^3(t)\right) \\
     &\phantom{=\frac{m_t^2}{(4\pi)^6} \Bigg[}
     + y_t^6 \Big(-3199.017 + 36 \;\barlog(h) - 2653.511 \;\barlog(t) + 756 \;\barlog(h) \;\barlog(t)\\
     &\phantom{=\frac{m_t^2}{(4\pi)^6} \Bigg[}\qquad~~
     + \frac{27}{2} \;\barlog^2(t) + 324 \;\barlog(h) \;\barlog^2(t) - 225 \;\barlog^3(t)\Big)
  \Bigg]
\end{split}
\end{align}
%
with $h\equiv m_h^2$.  Note that the 3-loop Higgs mass calculation in
the \MSbar scheme in \fs is only complete at $O(\at\as^2)$.  The
3-loop contributions of $O(\at^3 + \at^2 \as)$ are currently incomplete
because they would require the \MSbar top Yukawa coupling to be determined
from the top pole mass at the 2-loop $O(\at^2 + \at\as)$.
However, these 2-loop contributions to the top Yukawa coupling are
currently not available in \fs. Furthermore, the proper inclusion of
corrections to the Higgs pole mass at $O(\at^3)$
would require the extension of Eq.~\eqref{eq:2L_Mh_atat} by momentum
dependent $O(\at^2)$ terms. Alternatively, the evaluation of the Higgs mass at
the renormalization scale $Q^2 = t$ implies that the neglected
3-loop contributions at $O(\at^3 t)$ vanish
\cite{Martin:2014cxa}.\footnote{Note that terms of $O(\at^3 h)$ are
  neglected here and in Ref.\ \cite{Martin:2014cxa}.}
Likewise, Eq.~\eqref{eq:2L_Mh_atat} neglects contributions of $O(\at^2h)$ which are
subdominant in comparison to the implemented $O(\at^2t)$ corrections.
In contrast, the QCD corrections to the pole mass at 2-loop level, Eq.~\eqref{eq:2L_Mh_atas}, involve
terms proportional to the quartic Higgs coupling $\lambda$ up to $O(\as\at h^2/t)$, which are not
neglected here.

The 2-loop and 3-loop Higgs mass contributions from above can be
enabled at runtime by setting the following flags in the SLHA input
file:
%
\begin{lstlisting}
Block FlexibleSUSY
    4   3         # pole mass loop order
    5   3         # EWSB loop order
    7   2         # threshold corrections loop order
    8   1         # Higgs 2-loop corrections O(alpha_t alpha_s)
   10   1         # Higgs 2-loop corrections O(alpha_t^2)
   24   122111221 # individual threshold correction loop orders
   26   1         # Higgs 3-loop corrections O(alpha_t alpha_s^2)
   28   1         # Higgs 3-loop corrections O(alpha_t^2 alpha_s)
   29   1         # Higgs 3-loop corrections O(alpha_t^3)
\end{lstlisting}
%
In \fs's \mathematica interface, the above SLHA configuration options
correspond to
%
\begin{lstlisting}[language=Mathematica]
handle = FS<model>OpenHandle[
    fsSettings -> {poleMassLoopOrder -> 3,
                   ewsbLoopOrder -> 3,
                   thresholdCorrectionsLoopOrder -> 2,
                   higgs2loopCorrectionAtAs -> 1,
                   higgs2loopCorrectionAtAt -> 1,
                   thresholdCorrections -> 122111221,
                   higgs3loopCorrectionAtAsAs -> 1,
                   higgs3loopCorrectionAtAtAs -> 1,
                   higgs3loopCorrectionAtAtAt -> 1}
    ...];
\end{lstlisting}
%
See \appref{app:fs_config_options} for a list and description of
all of \fs's configuration options.
%
In \figref{fig:SM_thresholds} the impact of the leading 3-loop
Standard Model corrections of $O(\at^3 + \at^2 \as + \at \as^2)$ on
the light \CP-even Higgs pole mass in the MSSM in the pure effective
field theory (EFT) calculation of \HSSUSY is shown with the red
dashed-dotted line.  The line shows the predicted 3-loop Higgs mass
relative to the one calculated at the 2-loop level as a function of
the SUSY scale \MS, by taking into account the 1-loop threshold
correction for $\as^{\SM}(M_Z)$ and the 2-loop QCD correction to
$y_t^{\SM}(M_Z)$.  We find that the explicit 3-loop Standard Model
contributions to the Higgs mass lead to a small positive shift by
around $30\unit{MeV}$.
%
\begin{figure}[tbh]
  \centering
  \includegraphics[width=0.5\textwidth]{doc/plots/HSSUSY/HSSUSY_thresholds}
  \caption{Effect of different SM corrections on the Higgs
    pole mass in the pure EFT calculation of \HSSUSY for
    $\tan\beta = 5$ and $X_t = \sqrt{6}\MS$ as a function of the SUSY
    scale \MS.  The orange band marks the region where \HSSUSY
    predicts a Higgs mass compatible with the experimental value of
    $(125.09 \pm 0.32)\unit{GeV}$ \cite{Aad:2015zhl}.}
  \label{fig:SM_thresholds}
\end{figure}

\paragraph{3-loop renormalization group running in the SM}

In the SM, the 3-loop $\beta$ functions are known
\cite{Mihaila:2012fm,Bednyakov:2012rb,Bednyakov:2012en,Chetyrkin:2012rz,Bednyakov:2013eba}.
\fs allows the user to take these 3-loop $\beta$ functions into
account in the running of the \MSbar SM parameters by
setting the following switch in the model file:
%
\begin{lstlisting}
UseSM3LoopRGEs = True;
\end{lstlisting}
%
The expression for the 3-loop $\beta$ function of the $\mu^2$ parameter in
the SM has been extracted from the file \code{smh3l.m} available on
the arXiv page of Ref.~\cite{Bednyakov:2013eba}.  The remaining 3-loop
$\beta$ functions have been kindly provided by the authors of \susyhd
1.0.1 \cite{Vega:2015fna}.  To use 3-loop running at runtime, the
following flag should also be set in the SLHA input file:
%
\begin{lstlisting}
Block FlexibleSUSY
    6   3         # beta-functions loop order
\end{lstlisting}
%
In \fs's \mathematica interface, the above SLHA configuration
corresponds to
%
\begin{lstlisting}[language=Mathematica]
handle = FS<model>OpenHandle[
    fsSettings -> {betaFunctionLoopOrder -> 3}
    ...];
\end{lstlisting}

\paragraph{3-loop QCD corrections to the \MSbar top Yukawa coupling in the SM}

The pure SM QCD contribution to the \MSbar top Yukawa
coupling is known up to the 3-loop level of $O(\as^3)$
\cite{Chetyrkin:1999qi,Melnikov:2000qh}.  This 3-loop QCD expression
can be taken into account in \fs to extract the \MSbar top Yukawa
coupling in the SM from the top pole mass $M_t$ by setting
the following switch in the model file:
%
\begin{lstlisting}
UseYukawa3LoopQCD = True;
\end{lstlisting}
%
By default, \code{UseYukawa3LoopQCD} is set to \code{Automatic}, which
means that the 3-loop QCD contribution to the top Yukawa coupling is
taken into account only if the model is \MSbar renormalized.  To take
this 3-loop correction into account at runtime, the following flags
should be set in the SLHA input file:
%
\begin{lstlisting}
Block FlexibleSUSY
    7   3         # threshold corrections loop order
   24   123111221 # individual threshold correction loop orders
\end{lstlisting}
%
In \fs's \mathematica interface, the above SLHA configuration options
correspond to
%
\begin{lstlisting}[language=Mathematica]
handle = FS<model>OpenHandle[
    fsSettings -> {thresholdCorrectionsLoopOrder -> 3,
                   thresholdCorrections -> 123111221}
    ...];
\end{lstlisting}
%
The 3-loop QCD contribution is combined with the full 1-loop and QCD
2-loop contribution as
%
\begin{align}
\begin{split}
  m_t(Q) &= M_t + \re\Sigma_{t}^S(p^2=M_t^2,Q) \\
  &\phantom{={}} + M_t \Big[ \re\Sigma_{t}^L(p^2=M_t^2,Q) +
    \re\Sigma_{t}^R(p^2=M_t^2,Q) \\
  &\phantom{={} + M_t \Big[}
    + \Delta m_t^{(1),\text{qcd}}(Q) + \Delta m_t^{(2),\text{qcd}}(Q) + \Delta m_t^{(3),\text{qcd}}(Q) \Big]
  \,,
\end{split}
\end{align}
%
where $\Sigma_t^{S,L,R}(p^2,Q)$ denote the scalar, left- and
right-handed parts of the \MSbar-renormalized 1-loop top quark self
energy, $M_t$ is the top pole mass and
%
\begin{align}
  \Delta m_t^{(1),\text{qcd}}(Q) &=
  -\frac{g_3^2}{12 \pi^2} \left[4 - 3 \;\barlog(t)\right], \\
  \begin{split}
    \Delta m_t^{(2),\text{qcd}}(Q) &= \left(\Delta
      m_t^{(1),\text{qcd}}(Q)\right)^2 - \frac{g_3^4}{4608 \pi^4}
    \Bigg[396 \;\barlog^2(t) - 1452 \;\barlog(t) \\
    &\phantom{={}}
    - 48 \zeta(3) + 2053 + 16 \pi^2 (1+\log 4)\Bigg] ,
  \end{split} \\
\begin{split}
  \Delta m_t^{(3),\text{qcd}}(Q) &= \frac{g_3^6}{2430\, (4\pi)^6}
  \Bigg\{
  48600 \;\barlog^3(t) - 208980 \;\barlog^2(t) \\
  &\phantom{=\;} +540 \Big[-1560\,\zeta (3)+2993+40 \pi ^2 (1+\log 4)\Big] \barlog(t)\\
  &\phantom{=\;} +15 \left[69120\,\text{Li}_4\left(\frac{1}{2}\right)+113040\,\zeta(3)-
    94800 \,\zeta(5)-280853+2880 \,\log^4 2\right]\\
  &\phantom{=\;} +4 \pi^2 \Big[129510 \,\zeta(3)-388781 + 240 (733+24 \log 2) \,\log 2\Big] - 10500 \pi^4
  \Bigg\}
\end{split}\\
  &\approx
  \frac{g_3^6}{(4\pi)^6}\, 20 \left[\barlog^3(t)
    - \frac{43}{10} \;\barlog^2(t)
    + 22.8874 \;\barlog(t)
    -172.937 \right] \,.
\end{align}
%
In \figref{fig:SM_thresholds} the impact of the 3-loop correction to
the \MSbar top Yukawa coupling in the SM $y_t^{\SM}(M_Z)$
on the prediction of the light \CP-even Higgs pole mass in the MSSM in
the pure EFT calculation of \HSSUSY is shown as the green
dashed-double-dotted line.  As already discussed in
Ref.~\cite{Vega:2015fna}, we find that the inclusion of the 3-loop
correction to $y_t^{\SM}(M_Z)$ reduces the Higgs mass by up to
$500\unit{MeV}$.  Note that this is formally a (partial) 4-loop
effect on the Higgs mass, which is beyond the current accuracy of
\HSSUSY.

\paragraph{3-loop QCD corrections to the \MSbar strong coupling in the SM}

The pure SM QCD contribution to the \MSbar strong coupling is known up
to the 3-loop level of $O(\as^3)$
\cite{Rodrigo:1993hc,Rodrigo:1997zd,Chetyrkin:1997un,Chetyrkin:2000yt}.
This 3-loop QCD expression can be taken into account in \fs to extract
the \MSbar strong coupling in the SM from the input value
$\as^{\SM(5)}(M_Z)$ by setting the following switch in the
model file:
%
\begin{lstlisting}
UseSMAlphaS3Loop = True;
\end{lstlisting}
%
To take this 3-loop threshold correction into account at runtime, the
following flags should also be set in the SLHA input file:
%
\begin{lstlisting}
Block FlexibleSUSY
    7   3         # threshold corrections loop order
   24   123111321 # individual threshold correction loop orders
\end{lstlisting}
%
In \fs's \mathematica interface, the above SLHA configuration options
correspond to
%
\begin{lstlisting}[language=Mathematica]
handle = FS<model>OpenHandle[
    fsSettings -> {thresholdCorrectionsLoopOrder -> 3,
                   thresholdCorrections -> 123111321}
    ...];
\end{lstlisting}
%
The 3-loop QCD contributions are combined as
%
\begin{align}
  \as^{\SM}(Q) &=
  \frac{\as^{\SM(5)}(Q)}{1 - \Delta\as^{1L}(Q)
    - \Delta\as^{2L}(Q) - \Delta\as^{3L}(Q)} \,,
%  &= \as^{\SM(5)}(Q)\Big[ 1 + \Delta_1 + \Delta_2 + \Delta_3 + O(\hbar^4) \Big] \,,
\end{align}
%
where $\Delta\as^{nL}(Q)$ are the $n$-loop threshold corrections,
which read \cite{Chetyrkin:2000yt}
%
\begin{align}
  \Delta\as^{1L}(Q) &= \Delta_1 \,,\\
  \Delta\as^{2L}(Q) &= \Delta_2 - \Delta_1^2 \,,\\
  \Delta\as^{3L}(Q) &= \Delta_3 + \Delta_1^3 - 2\Delta_1\Delta_2 \,,\\
  \Delta_1 &= \left(\frac{\as^{\SM(5)}(Q)}{\pi}\right) \frac{L}{6} \,, \\
  \Delta_2 &= \left(\frac{\as^{\SM(5)}(Q)}{\pi}\right)^2 \Bigg[
  -\frac{11}{72} + \frac{11}{24} L + \frac{1}{36} L^2
  \Bigg] \,,\\
  \Delta_3 &=
  \left(\frac{\as^{\SM(5)}(Q)}{\pi}\right)^3 \Bigg[
  \frac{1}{216} L^3
  + \frac{167}{576} L^2
  + \frac{2645}{1728}L
  + n_f \left(\frac{1}{36} L^2
  - \frac{67}{576} L
  + \frac{2633}{31104}\right) \nonumber \\
  &\phantom{={}}
  + \frac{82043}{27648} \zeta_3
  - \frac{564731}{124416}
  \Bigg] \,, \\
  &\approx
  \left(\frac{\as^{\SM(5)}(Q)}{\pi}\right)^3 \Bigg[
  0.00462963 L^3
  + 0.428819 L^2
  + 0.94907 L
  - 0.54880 \Bigg] \,,
\end{align}
%
with $n_f = 5$, $L = \log(Q^2/(m_t^{\SM}(Q))^2)$ and $m_t^{\SM}(Q)$ being the \MSbar
top quark mass in the SM at the scale $Q$.

\figref{fig:SM_thresholds} shows the impact of the 2- and 3-loop QCD
threshold corrections for $\as^{\SM}(M_Z)$ on the light \CP-even Higgs
pole mass in the MSSM, as predicted by the pure EFT calculation of
\HSSUSY as a function of the SUSY scale \MS.  We find that the
inclusion of $\Delta\as^{2L}$ (blue dashed line) reduces the Higgs
mass by up to $40\unit{MeV}$, depending on the SUSY scale.  Taking
into account $\Delta\as^{3L}$ (blue dotted line) reduces the Higgs
mass further by around $3\unit{MeV}$.

%%%%%%%%%%%%%%%%%%%%%% MSSM corrections %%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{(N)MSSM-specific higher-order contributions}
\label{sec:MSSM-specific_ho_corrections}
\paragraph{2-loop contributions to the (N)MSSM Higgs masses}

As already described in Ref.~\cite{Athron:2014yba}, the known dominant 2-loop
effective potential contributions to the Higgs masses in the (N)MSSM
of $O((\at + \ab)\as + (\at +
\ab)^2 + \atau^2)$
\cite{Degrassi:2001yf,Brignole:2001jy,Dedes:2002dy,Brignole:2002bz,Dedes:2003km,Degrassi:2009yq}
can be taken into account by \fs by setting the corresponding switches
in the model files:
%
\begin{lstlisting}
UseHiggs2LoopMSSM = True;
UseHiggs2LoopNMSSM = True;
\end{lstlisting}
%
These flags can be enabled in all real (N)MSSM-like models with 2(3)
\CP-even Higgs bosons, 1(2) \CP-odd Higgs boson(s) and 1 electrically neutral
Goldstone boson.  In addition to these flags, the (effective) $\mu$
parameter in the convention of Ref.~\cite{Haber:1984rc} and the effective
squared \CP-odd Higgs mass $m_A^2$ of the model must be identified.
In the \modelname{NMSSM} model file of \sarah they would read for
example:
%
\begin{lstlisting}
EffectiveMu = \[Lambda] vS / Sqrt[2];
EffectiveMASqr = (T[\[Lambda]] vS / Sqrt[2] + 0.5 \[Lambda] \[Kappa] vS^2) (vu^2 + vd^2) / (vu vd);
\end{lstlisting}
%
These 2-loop contributions can then be enabled at runtime by setting
the following flags in the SLHA input file:
%
\begin{lstlisting}
Block FlexibleSUSY
    4   2         # pole mass loop order
    5   2         # EWSB loop order
    7   2         # threshold corrections loop order
    8   1         # Higgs 2-loop corrections O(alpha_t alpha_s)
    9   1         # Higgs 2-loop corrections O(alpha_b alpha_s)
   10   1         # Higgs 2-loop corrections O((alpha_t + alpha_b)^2)
   11   1         # Higgs 2-loop corrections O(alpha_tau^2)
   24   122111221 # individual threshold correction loop orders
\end{lstlisting}
%
In \fs's \mathematica interface, the above SLHA configuration options
correspond to
%
\begin{lstlisting}[language=Mathematica]
handle = FS<model>OpenHandle[
    fsSettings -> {poleMassLoopOrder -> 2,
                   ewsbLoopOrder -> 2,
                   thresholdCorrectionsLoopOrder -> 2,
                   higgs2loopCorrectionAtAs -> 1,
                   higgs2loopCorrectionAbAs -> 1,
                   higgs2loopCorrectionAtAt -> 1,
                   higgs2loopCorrectionAtauAtau -> 1,
                   thresholdCorrections -> 122111221}
    ...];
\end{lstlisting}

\paragraph{3-loop contributions to the light \CP-even MSSM Higgs mass}

The 3-loop contributions to the light \CP-even Higgs mass in the MSSM
have been calculated to $O(\at\as^2)$ in the \DRbar and
\MDRbar scheme \cite{Harlander:2008ju,Kant:2010tf,Kunz:2014gya}.  The
expressions are available in the public spectrum generator \htm
\cite{h3murl}, where they are added to the 2-loop on-shell result of
\FeynHiggs
\cite{Heinemeyer:1998yj,Heinemeyer:1998np,Degrassi:2002fi,Frank:2006yh,Hahn:2013ria,Drechsel:2016jdg,Bahl:2016brp,Bahl:2017aev}.
In Ref.~\cite{Harlander:2017kuc} the 3-loop contributions of
$O(\at\as^2 + \ab\as^2)$ have been studied for the first time in a
pure \DRbar MSSM spectrum generator and have been made available in
the public C++ library \Himalaya \cite{himalaya}.

The explicit 3-loop contributions of $O(\at\as^2 + \ab\as^2)$ to the
light \CP-even Higgs mass from \Himalaya can be taken into account in
\fs by setting the following flag in the \fs model file:
%
\begin{lstlisting}[language=Mathematica]
UseHiggs3LoopMSSM = True;
\end{lstlisting}
%
The model is required to be real MSSM-like with two \CP-even Higgs bosons and
one electrically neutral Goldstone boson.
%
However, in a pure \DRbar calculation, another source of such 3-loop
contributions originates from the 2-loop \SQCD contribution to the MSSM
\DRbar top Yukawa coupling, which must be included.  Then, however, in
order to be consistent with respect to the loop orders of the running and
decoupling, also 3-loop renormalization group running should be
performed and the 2-loop threshold correction $\Delta\as^{2L}$ of
the strong coupling should be included, see below.  Therefore, we
strongly recommend setting the following flags in addition in the \fs
model file:
%
\begin{lstlisting}[language=Mathematica]
UseHiggs2LoopMSSM = True;  (* 2-loop contribution to Higgs mass *)
EffectiveMu = \[Mu];       (* specify mu parameter, see above *)
UseMSSM3LoopRGEs = True;   (* 3-loop running *)
UseMSSMYukawa2Loop = True; (* 2-loop SUSY-QCD correction to yt *)
UseMSSMAlphaS2Loop = True; (* 2-loop threshold correction to alpha_s *)
\end{lstlisting}
%
When building \fs, the path to the \Himalaya headers and to the
\Himalaya library must be specified:
%
\begin{lstlisting}[language=bash]
$ ./configure --with-models=[...] \
   --enable-himalaya \
   --with-himalaya-incdir=$HIMALAYA_PATH/source/include \
   --with-himalaya-libdir=$HIMALAYA_PATH/build
$ make
\end{lstlisting}
%
where \code{$HIMALAYA_PATH} is the \Himalaya directory. %$
To calculate the light \CP-even Higgs mass in the MSSM at the 3-loop
level with \fs, the following flags must be set at runtime: In the
SLHA input file we recommend setting at least
%
\begin{lstlisting}
Block FlexibleSUSY
    4   3         # pole mass loop order
    5   3         # EWSB loop order
    6   3         # beta-functions loop order
    7   2         # threshold corrections loop order
   24   122111221 # individual threshold correction loop orders
   25   0         # ren. scheme for 3L corrections (0 = DR, 1 = MDR)
   26   1         # Higgs 3-loop corrections O(alpha_t alpha_s^2)
   27   1         # Higgs 3-loop corrections O(alpha_b alpha_s^2)
\end{lstlisting}
%
In \fs's \mathematica interface, the above SLHA configuration options
correspond to
%
\begin{lstlisting}[language=Mathematica]
handle = FS<model>OpenHandle[
    fsSettings -> {poleMassLoopOrder -> 3,
                   ewsbLoopOrder -> 3,
                   betaFunctionLoopOrder -> 3,
                   thresholdCorrectionsLoopOrder -> 2,
                   thresholdCorrections -> 122111221,
                   higgs3loopCorrectionRenScheme -> 0,
                   higgs3loopCorrectionAtAsAs -> 1,
                   higgs3loopCorrectionAbAsAs -> 1}
    ...];
\end{lstlisting}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.49\textwidth]{doc/plots/Himalaya/scan_Mh_MS_TB-5_Xt-0}\hfill
  \includegraphics[width=0.49\textwidth]{doc/plots/Himalaya/scan_Mh_MS_TB-5_Xt-0_delta_alpha}\\
  \includegraphics[width=0.49\textwidth]{doc/plots/Himalaya/scan_Mh_Xt_TB-5_MS-2000}\hfill
  \includegraphics[width=0.49\textwidth]{doc/plots/Himalaya/scan_Mh_Xt_TB-5_MS-2000_delta_alpha}
  \caption{Comparison of different 2-loop and 3-loop calculations of
    the lightest \CP-even Higgs mass in the MSSM with different \fs
    spectrum generators for $\tan\beta = 5$.  In the top row we set
    $X_t = 0$ and in the bottom row we use $\MS = 2\unit{TeV}$.}
  \label{fig:Mh-himalaya}
\end{figure}
%
\figref{fig:Mh-himalaya} shows a comparison of the different light
\CP-even Higgs mass calculations available in \fs.  The scenario is
chosen such that all soft-breaking \DRbar mass parameters, the
superpotential
$\mu$ parameter and the \DRbar \CP-odd Higgs mass are equal to the SUSY
scale $\MS$ at the scale $Q =
\MS$ and all sfermion mixing parameters
$X_f$ are set to zero, except for the stop mixing parameter
$X_t$.  Furthermore, we set $M_t =
173.34\unit{GeV}$, $\as^{\SM(5)}(M_Z) =
0.1184$ and $\aem^{\SM(5)}(M_Z) =
1/127.944$.  The top row of \figref{fig:Mh-himalaya} shows the Higgs
mass as a function of $\MS$ for $X_t = 0$ and $\tan\beta =
5$.  In the left panel of the top row the classic 2-loop fixed-order
calculation with \fs \cite{Athron:2014yba} is shown as the blue dashed
line and the pure EFT calculation with \HSSUSY is shown as the black
dotted line.  The reason why these two curves do not deviate from each
other logarithmically when
$\MS$ increases is an accidental cancellation of large
logarithmic contributions in the fixed-order calculation
\cite{Athron:2016fuq}.  The green dashed-dotted line shows the
improved version of the \feft calculation \cite{Athron:2016fuq}, which
interpolates
between the fixed-order calculation at low \MS and the pure EFT
calculation at large \MS.  See \secref{sec:feft} for a discussion of
the improvements of \feft.  The red solid line shows the 3-loop
calculation up to $O(\at\as^2 +
\ab\as^2)$ with \fsh \cite{Harlander:2017kuc}.  In this calculation, the
full 2-loop \SQCD contributions to the \DRbar top and bottom Yukawa
couplings $y_{t,b}(M_Z)$ \cite{Bednyakov:2007vm,Bednyakov:2002sf,Bednyakov:2005kt}
are taken into account.  The sum of these contributions
and the explicit 3-loop contribution to the \CP-even
Higgs mass matrix leads to a downward shift of the predicted Higgs
mass by $1$--$2\unit{GeV}$, depending on \MS, compared to the classic 2-loop
calculation with \fs (blue dashed line).  As also discussed in
Ref.~\cite{Harlander:2017kuc}, this brings the prediction closer to the pure
EFT calculation, which is expected to lead to a more precise result above
the TeV scale.  In the left panel of the bottom row of
\figref{fig:Mh-himalaya}, the calculated Higgs mass is shown as a
function of the stop mixing parameter $X_t$ for fixed
$\MS = 2\unit{TeV}$.  Also for non-zero
$X_t$, one finds a reduction of the Higgs mass by
$1$--$3\unit{GeV}$, which brings the prediction closer to the one of
the pure EFT calculation.

In the right panels of \figref{fig:Mh-himalaya}, the effect of the 2-loop
\SQCD threshold correction $\Delta\as^{2L}$
\cite{Harlander:2005wm,Bauer:2008bj,Bednyakov:2010ni} to the strong
coupling $\as$ (see below) on the 3-loop Higgs mass calculation is shown for
the same scenario as in the corresponding left panels.  The inclusion
of $\Delta\as^{2L}$ is formally a 4-loop effect on the light
\CP-even Higgs mass in the MSSM\@.  However, $\Delta\as^{2L}$ should
be taken into account for a consistent running and decoupling
procedure with 3-loop renormalization group running.  The red solid
lines in the right panels correspond to the 3-loop calculation of
Ref.~\cite{Harlander:2017kuc}, which uses only the 1-loop threshold correction
$\Delta\as^{1L}$.  These red solid lines are the same as in the
corresponding left panels.  The effect of including
$\Delta\as^{2L}$ is shown as the red dashed line.  We find that the
inclusion of this 2-loop threshold correction leads to an upwards
shift of the Higgs mass by up to $2\unit{GeV}$, depending on \MS and
$X_t$.

\paragraph{3-loop renormalization group running in the MSSM}

In the MSSM, the 3-loop $\beta$ functions are also known
\cite{Jack:2003sx,Jack:2004ch}.  \fs allows the user to take these 3-loop
$\beta$ functions into account in the running of the \DRbar MSSM parameters by
setting the following switch in the model file:
%
\begin{lstlisting}
UseMSSM3LoopRGEs = True;
\end{lstlisting}
%
To use 3-loop running at runtime, the following flag should also be set in
the SLHA input file:
%
\begin{lstlisting}
Block FlexibleSUSY
    6   3         # beta-functions loop order
\end{lstlisting}
%
In \fs's \mathematica interface, the above SLHA configuration
corresponds to
%
\begin{lstlisting}[language=Mathematica]
handle = FS<model>OpenHandle[
    fsSettings -> {betaFunctionLoopOrder -> 3}
    ...];
\end{lstlisting}
%
The expressions for the 3-loop $\beta$ functions have been extracted
from the official FORM file provided by the authors of
Refs.~\cite{Jack:2003sx,Jack:2004ch}.\footnote{\url{https://www.liverpool.ac.uk/~dij/betas/allgennb.log}}
We have numerically compared the expressions with the ones implemented
in \softsusy 3.7.0 \cite{Allanach:2014nba} and found exact agreement.
%
The effect of the 3-loop RGEs on the Higgs pole mass in the MSSM is of
the order of a few $100\unit{MeV}$ as discussed in
Ref.~\cite{Allanach:2014nba} and is shown in \figref{fig:MSSMMuBMu}.
%
Due to the complexity of the 3-loop $\beta$ functions, the runtime of
the MSSM spectrum generators is increased by a factor of $4$--$5$ if
3-loop running is enabled in the MSSM\@.

\paragraph{2-loop \SQCD corrections to the \DRbar top and bottom Yukawa couplings in the MSSM}
In the MSSM, the full 2-loop \SQCD corrections of $O(\as^2)$
to the \DRbar top Yukawa coupling
\cite{Bednyakov:2007vm,Bednyakov:2002sf,Bednyakov:2005kt} as well as
the 2-loop \SQCD $O(\as^2)$ and partial electroweak contributions
to the \DRbar bottom Yukawa coupling \cite{Bednyakov:2009wt} are
known.  These 2-loop corrections have already been made available in
\softsusy~3.7.0 \cite{Allanach:2014nba}.  The 2-loop \SQCD
corrections of $O(\as^2)$ are now also incorporated in \fstwo and can be
used by setting the following switch in the model file:\footnote{We
  kindly thank Alexander Bednyakov for providing the 2-loop \SQCD
  expressions.}
%
\begin{lstlisting}
UseMSSMYukawa2Loop = True;
\end{lstlisting}
%
To take these 2-loop contributions into account at runtime, the
following flags should also be set in the SLHA input file:
%
\begin{lstlisting}
Block FlexibleSUSY
    7   2         # threshold corrections loop order
   24   122111221 # individual threshold correction loop orders
\end{lstlisting}
%
In \fs's \mathematica interface, the above SLHA configuration options
correspond to
%
\begin{lstlisting}[language=Mathematica]
handle = FS<model>OpenHandle[
    fsSettings -> {thresholdCorrectionsLoopOrder -> 2,
                   thresholdCorrections -> 122111221}
    ...];
\end{lstlisting}
%
In \figref{fig:MSSMMuBMu} we show in blue the effect of the full
2-loop \SQCD corrections on the lightest \CP-even Higgs pole mass in
the MSSM\@.  The blue dashed line shows the effect in \fstwo and the
crosses in \softsusy~4.0.1.
As can be seen from the figure, the full 2-loop \SQCD corrections are
negative and can affect the Higgs mass by several GeV, as has been
observed in Ref.~\cite{Allanach:2014nba}.  In the left panel of
\figref{fig:MSSMMuBMu}, the shift in the Higgs mass is shown as a
function of the SUSY scale.  For scales above $\approx 2\unit{TeV}$ we
find a logarithmic shift in $M_h$, which is caused by new large
logarithms originating from the 2-loop contribution of the SUSY
particles to the \DRbar top Yukawa coupling.  We also find that these
new logarithms alone would spoil the accidental cancellation of large
logarithms described in Ref.~\cite{Athron:2016fuq}.
%
Note that in the MSSM the effect of the 2-loop \SQCD corrections to
the top and bottom Yukawa couplings is a partial 3-loop contribution
to the light \CP-even Higgs pole mass.  Thus, these 2-loop \SQCD corrections
must be taken into account if the explicit 3-loop Higgs mass
contributions of $O(\at \as^2 + \ab \as^2)$ from
\Himalaya are used, see above.
%
\begin{figure}[tbh]
  \centering
  \includegraphics[width=0.49\textwidth]{doc/plots/MSSMMuBMu/scan_MSSMMuBMu_Mh_MS_TB-5_Xt-0}\hfill
  \includegraphics[width=0.49\textwidth]{doc/plots/MSSMMuBMu/scan_MSSMMuBMu_Mh_Xt_TB-5_MS-2000}
  \caption{Effect of the 3-loop RGEs for all MSSM parameters (red),
    2-loop \SQCD contributions to $y_b$ and $y_t$ (blue) and 2-loop
    \SQCD contributions to $\as$ (green) on the lightest \CP-even
    Higgs pole mass in the MSSM\@.  The lines show the shift in the
    Higgs pole mass in \fstwo compared to the one obtained with 2-loop
    RGEs, full 1-loop \SQCD + 2-loop SM-QCD contributions to $y_t$ and
    1-loop \SQCD contributions to $\as$.  The crosses show the
    corresponding shift obtained with \softsusy~4.0.1.  In the left
    panel we use $\tan\beta = 5$ and $X_t = 0$ and in the right panel
    we fix $\tan\beta = 5$ and $\MS = 2\unit{TeV}$.}
  \label{fig:MSSMMuBMu}
\end{figure}

\paragraph{2-loop \SQCD corrections to the \DRbar strong gauge coupling in the MSSM}

In the MSSM, the full 2-loop \SQCD corrections of
$O(\as^2 + \as\at + \as\ab)$ to the \DRbar strong gauge coupling are known
\cite{Harlander:2005wm,Bauer:2008bj,Bednyakov:2010ni} and have been
made available in \softsusy~3.7.0 \cite{Allanach:2014nba}.  In \fstwo,
these corrections can be taken into account by setting in the model
file:\footnote{We kindly thank Ben Allanach and Alexander Bednyakov
  for providing the 2-loop \SQCD expressions.}
%
\begin{lstlisting}
UseMSSMAlphaS2Loop = True;
\end{lstlisting}
%
To take these 2-loop threshold corrections into account at runtime, the
following flags should also be set in the SLHA input file:
%
\begin{lstlisting}
Block FlexibleSUSY
    7   2         # threshold corrections loop order
   24   122111221 # individual threshold correction loop orders
\end{lstlisting}
%
In \fs's \mathematica interface, the above SLHA configuration options
correspond to
%
\begin{lstlisting}[language=Mathematica]
handle = FS<model>OpenHandle[
    fsSettings -> {thresholdCorrectionsLoopOrder -> 2,
                   thresholdCorrections -> 122111221}
    ...];
\end{lstlisting}
%
In \figref{fig:MSSMMuBMu} we show in green the effect of the 2-loop
\SQCD corrections to $\as$ on the lightest \CP-even Higgs pole
mass in the MSSM\@.  The green dashed-dotted line shows the effect in
\fstwo and the crosses in \softsusy~4.0.1.  Both implementations agree
exactly.  We furthermore find that the inclusion of the 2-loop
threshold corrections to $\as$ leads to a logarithmic enhancement
of the Higgs mass as a function of the SUSY scale.  The enhancement is
around $+1\unit{GeV}$ for $\MS\approx 2\unit{TeV}$ and maximal stop
mixing.
%
Note that in the MSSM the effect of the 2-loop \SQCD corrections to
the strong coupling is formally a partial 4-loop contribution to the
Higgs pole mass.

\subsubsection{Split-MSSM-specific higher-order contributions}
\paragraph{3-loop contribution to the Higgs mass in the split-MSSM}

In the split-MSSM \cite{Bagnaschi:2014rsa}, part of the 3-loop
contribution of $O(\at \as^2)$
to the SM-like Higgs pole mass is known in the \MSbar
scheme \cite{Benakli:2013msa}.  \fs allows the user to take this
3-loop contribution into account by setting the following switch in
the model file:
%
\begin{lstlisting}
UseHiggs3LoopSplit = True;
\end{lstlisting}
%
This is used in the distributed \modelname{SplitMSSM} model file. For use in
other models, it is a requirement that the model contains a single
Higgs boson and a gluino.  The enabled 3-loop Higgs pole mass
contribution reads
%
\begin{align}
  (\Delta m_h^2)_{3L} &= \frac{64 g_3^4 y_t^2}{(4\pi)^6} m_t^2 \log^3\frac{m_{\tilde{g}}^2}{Q^2} \,,
\end{align}
%
where $m_{\tilde{g}}$ is the \MSbar gluino mass.  To take this
contribution into account at runtime, the following flags should also be
set in the SLHA input file:
%
\begin{lstlisting}
Block FlexibleSUSY
    4   3         # pole mass loop order
    5   3         # EWSB loop order
    7   2         # threshold corrections loop order
   24   122111221 # individual threshold correction loop orders
   26   1         # Higgs 3-loop corrections O(alpha_t alpha_s^2)
\end{lstlisting}
%
In \fs's \mathematica interface, the above SLHA configuration options
correspond to
%
\begin{lstlisting}[language=Mathematica]
handle = FS<model>OpenHandle[
    fsSettings -> {poleMassLoopOrder -> 3,
                   ewsbLoopOrder -> 3,
                   thresholdCorrectionsLoopOrder -> 2,
                   thresholdCorrections -> 122111221,
                   higgs3loopCorrectionAtAsAs -> 1}
    ...];
\end{lstlisting}

\subsection{New features for definition of boundary conditions}
In \fstwo, the expressions to define boundary conditions are allowed
to be more complicated and to involve trigonometric functions,
dilogarithms, branches and more.  This is particularly useful for defining
high-scale boundary conditions that match a model to its UV-completion.  An
example is \fs's \HSSUSY model file, which implements the known
1- and 2-loop high-scale matching condition on the quartic Higgs
coupling of the SM against the MSSM at the SUSY scale
\cite{Bagnaschi:2014rsa,Vega:2015fna,Bagnaschi:2017xid}.  The list of
special functions and symbols to be used in boundary conditions can be
found in \tabref{tab:bcs_functions}.
%
\begin{table}[h!]
  \centering
 \resizebox{\textwidth}{!}{%
  \begin{tabularx}{1.05\textwidth}{lX}
    \toprule
    Function & Description \\
    \midrule

    \code{Abs[a_]} &
    Returns the magnitude of a real or complex number \code{a}, $|a|$.
    If \code{a} is a vector, a vector is returned with \code{Abs} applied to each element. \\

    \code{AbsSqr[a_]} &
    Returns the squared magnitude of a real or complex number \code{a}, $|a|^2$. \\

    \code{AbsSqrt[a_]} &
    Returns the square root of the magnitude of \code{a}, $\sqrt{|a|}$. \\

    \code{ArcSin[a_]}, \code{ArcCos[a_]} &
    Returns $\arcsin a$ and $\arccos a$, respectively. \\

    \code{ArcTan[a_]} & Returns $\arctan a$. \\

    \code{Arg[z_]} &
    Returns the phase angle of a complex number, $\arg z$. \\

    \code{Cbrt[a_]} &
    Returns the cubic root of \code{a}, $\sqrt[3]{a}$. \\

    \code{Conjugate[a_]} &
    Returns the complex conjugate of \code{a}. \\

    \code{Exp[a_]} & Returns $e^a$ for real or complex \code{a}. \\

    \code{FiniteLog[a_]} &
    Returns $\log a$ if $\log a$ is well-defined, otherwise returns $0$. \\

    \code{FSThrow[msg_]} & Throws an exception of type
    \code{PhysicalError} with the message \code{msg}. \\

    \code{I} & Imaginary unit. \\

    \code{If[cond_, a_, b_]} &
    If \code{cond} is true, \code{a} is returned, otherwise \code{b}. \\

    \code{Im[a_]} &
    Returns the imaginary part of \code{a}. \\

    \code{IsClose[a_, b_, eps_]} &
    Returns \code{True} if \code{Abs[a - b]} $<$ \code{eps}, otherwise \code{False}. \\

    \code{IsCloseRel[a_, b_, eps_]} &
    Returns \code{True} if \code{Abs[(a - b)/a]} $<$ \code{eps}, otherwise \code{False}. \\

    \code{IsFinite[a_]} &
    Returns \code{True} if \code{a} is neither \code{nan} nor \code{inf}. \\

    \code{KroneckerDelta[i_, j_]} &
    Returns the Kronecker $\delta_{ij}$. \\

    \code{Log[a_], ComplexLog[a_]} &
    Returns the natural logarithm for real and complex arguments, respectively. \\

    \code{Max[a_, ...]} &
    Returns the maximum of all given arguments. \\

    \code{Min[a_, ...]} &
    Returns the minimum of all given arguments. \\

    \code{Not[cond_]} &
    Returns the logical negation of \code{cond}. \\

    \code{PolyLog[2, z_]} &
    Returns the dilogarithm of the real or complex number \code{z}. \\

    \code{Re[a_]} &
    Returns the real part of \code{a}. \\

    \code{Round[a_]} &
    Returns \code{Floor[a + 0.5]} if \code{a} $\geq 0$, otherwise
    \code{Floor[a - 0.5]}. \\

    \code{Print<type>[msg_]} & Prints a debug, info, error, warning or
    fatal message, depending on the \code{<type>}, and returns zero.
    \code{<type>} can be \code{DEBUG}, \code{INFO}, \code{WARNING},
    \code{ERROR} or \code{FATAL}.  \code{PrintFATAL[msg]}
    throws an exception after \code{msg} has been printed.\\

    \code{SCALE} & Returns the renormalization scale at which
    the boundary condition is imposed. \\

    \code{Sign[a_]} &
    Returns $1$ if \code{a} $\geq 0$, otherwise $-1$. \\

    \code{SignedAbsSqrt[a_]} &
    Returns \code{Sign[a]*Sqrt[Abs[a]]}. \\

    \code{Sin[a_]}, \code{Cos[a_]}, \code{Tan[a_]} &
    Returns $\sin a$, $\cos a$ and $\tan a$, respectively. \\

    \code{Total[vec_]} &
    Returns the sum of all elements of \code{vec}, $\sum_i v_i$. \\

    \code{UnitStep[value_]} &
    Returns $0$ if \code{value} $<0$, $1$ otherwise. \\

    \multilinecell[t]{\code{Which[test1_, value1_,}\\\code{\ test2_, value2_, ...]}} &
    If \code{test1} is true, \code{value1} is returned, otherwise
    if \code{test2} is true, \code{value2} is returned, etc. \\

    \code{ZeroSqrt[a_]} &
    Returns $\sqrt{a}$ if $a>0$, otherwise returns $0$. \\

    \bottomrule
  \end{tabularx}}
  \caption{Available special functions and symbols in the boundary conditions.}
  \label{tab:bcs_functions}
\end{table}

\section{\fbsm extension}
\label{sec:fbsm}

Since version 1.1.0, \fs can generate spectrum generators not
only for SUSY models, but also for non-SUSY models.  We document this
feature here for the first time.  In Subsections
\ref{sec:application_THDM} and \ref{sec:application_HSSUSY}, we will
describe important applications of this feature to the two-Higgs
doublet model and to an effective low-energy theory of the MSSM
(\HSSUSY).

The generated non-SUSY spectrum generators have the same features as the SUSY
spectrum generators:
%
\begin{itemize}
\item The running gauge and Yukawa couplings of the non-SUSY model are
  calculated automatically at the 1-loop level from the known
  low-energy SM parameters
  $\aem^{\SM(5)}(M_Z)$, $\as^{\SM(5)}(M_Z)$
  and from the known quark and lepton masses as well as $M_Z$ and either $G_F$
  or $M_W$.  2-loop and 3-loop QCD corrections can be taken into
  account to determine the running top Yukawa coupling of the model,
  see \secref{sec:model_specific_contributions}.
\item Up to three boundary conditions can be specified to fix the
  running parameters of the model at different user-defined scales.
\item 2-loop renormalization group running is used between the scales
  at which the boundary conditions are imposed.\footnote{Note that
    the $\beta$ functions of scalar tadpole terms in
    non-supersymmetric models \cite{Goodsell:2012fm} are currently not
    generated by \sarah.  For this reason, such tadpole terms do not
    run in \sarah/\spheno or \fs.} In the SM and in the MSSM, also
  3-loop running is available, see
  \secref{sec:model_specific_contributions}.
\item The pole mass spectrum is calculated at the full 1-loop level,
  taking into account all BSM contributions.  Some 2-loop and 3-loop
  corrections can be added in specific non-SUSY models, see
  \secref{sec:model_specific_contributions}.
\end{itemize}

\subsection{Setting up a \fbsm model}

The \fs user interface for creating spectrum generators for non-SUSY
models is exactly the same as in the case of SUSY models, except that
all non-SUSY parameters are defined in the \MSbar scheme.  In
particular, at the low-energy scale \fs automatically determines
the gauge and Yukawa couplings of the non-SUSY model in the \MSbar
scheme.

\subsection{Determination of the \MSbar gauge and Yukawa couplings}
\label{sec:determination_of_g_Y}

If the considered BSM model has a gauge symmetry with the SM gauge group $SU(3)_C\times SU(2)_L\times U(1)_Y$ as a factor,
then \fs automatically fixes the three corresponding normalized gauge
couplings $g_1$, $g_2$ and $g_3$ at the low-energy boundary condition
from the given input parameters $M_Z$, $\aem^{\SM(5)}(M_Z)$,
$\as^{\SM(5)}(M_Z)$ and $G_F$ or $M_W$ as
%
\begin{align}
  g_1(Q) &= N_{g_Y} \, g_Y(Q) \,,
  & g_Y(Q) &= \frac{\sqrt{4\pi\aem(Q)}}{\cos\theta_W(Q)} \,,\\
  g_2(Q) &= N_{g_L} \, g_L(Q) \,,
  & g_L(Q) &= \frac{\sqrt{4\pi\aem(Q)}}{\sin\theta_W(Q)} \,,\\
  g_3(Q) &= N_{g_{\text{s}}} \, g_{\text{s}}(Q) \,,
  & g_{\text{s}}(Q) &= \sqrt{4\pi\as(Q)} \,.
\end{align}
%
Here, $\aem$ and $\as$
denote the \MSbar electromagnetic and strong coupling constants of the
non-SUSY model, respectively, and $\theta_W$ is the \MSbar\ weak
mixing angle.  The coefficients $N_{g_i}$ denote the potential
normalization factors defined in the \sarah model file.  The
renormalization scale $Q$, at which the gauge couplings are calculated,
can be specified using the \code{LowScale} variable in the model file.
The coupling constants of the model are related to the corresponding
ones of the SM with five active quark flavors,
$\aem^{\SM(5)}(Q)$ and $\as^{\SM(5)}(Q)$, which are input,
via the relations
%
\begin{align}
  \aem(Q) &=
  \frac{\aem^{\SM(5)}(Q)}{1 - \Delta\aem(Q)} \,,\\
  \as(Q) &=
  \frac{\as^{\SM(5)}(Q)}{1 - \Delta\as(Q)} \,.
\end{align}
%
The threshold corrections $\Delta\alpha_i(Q)$ have the form
%
\begin{align}
  \Delta\aem(Q) &=
  \frac{\aem}{2\pi} \sum_i C_i^{\text{em}} \log\frac{m_i}{Q}
  \,,\label{eq:alpha_em_general_threshold} \\
  \Delta\as(Q) &=
  \frac{\as}{2\pi} \sum_i C_i^{\text{s}} \log\frac{m_i}{Q} \,,
\end{align}
%
where the sum runs over all non-SM particles plus the top quark and
the constants $C_i^{\text{em}}$ and $C_i^{\text{s}}$ depend on the
representation of the particle $i$ with respect to the Lorentz and gauge group.
%
The \MSbar\ weak mixing angle $\theta_W$ in the non-SUSY model is
determined either
%
\begin{itemize}
\item from the Fermi constant $G_F$ and $M_Z$ using the iterative
  approach described in Ref.~\cite{Degrassi:1990tu} taking into account the
  full 1-loop corrections and leading 2-loop SM
  corrections to $\Delta\hat{\rho}$ and $\Delta\hat{r}$, see
  \secref{sec:fmw}.
\item or from the running $W$ and $Z$ masses, which
  are obtained from the corresponding pole masses via a 1-loop
  calculation.  See \secref{sec:fmw} for more details.
\end{itemize}
%
If the considered BSM model does not contain the SM gauge
group as a factor, it is of course still possible to fix the gauge
couplings at the low-energy boundary condition by defining them to be
input parameters.
%
\begin{example}
  In a left-right-symmetric model with the gauge group $SU(3)_C\times
  SU(2)_L\times SU(2)_R\times U(1)_L\times U(1)_R$, one could for
  example fix the BSM gauge couplings $g_3$, $g_L$, $g_R$, $g_{1L}$
  and $g_{1R}$ by SM-like gauge couplings $g_3^\SM$,
  $g_2^\SM$ and $g_1^\SM$ which are given as input via the
  \code{EXTPAR} block:
%
\begin{lstlisting}
EXTPAR = {
  {100, g1SMInput},
  {101, g2SMInput},
  {102, g3SMInput}
};

LowScaleInput = {
   {g3, g3SMInput},
   {gL, g2SMInput},
   {gR, g2SMInput},
   {g1L, g1SMInput g2SMInput / Sqrt[2 (-g1SMInput^2 + g2SMInput^2)]},
   {g1R, g1SMInput g2SMInput / Sqrt[2 (-g1SMInput^2 + g2SMInput^2)]},
   {g1L1R, 0},
   {g1R1L, 0}
};
\end{lstlisting}
Note that, in addition to fixing the BSM gauge couplings, in this example
the off-diagonal gauge couplings $g_{1L1R}$ and $g_{1R1L}$ that arise due to
$U(1)$ mixing are being set to zero at the low-energy scale.
\end{example}

The \MSbar\ Yukawa couplings $Y_f(Q)$ of the SM-like
fermions $f$ in the non-SUSY model are determined from the
corresponding \MSbar\ masses $m_f(Q)$ using the tree-level
relation. For example, in the SM this relation reads
%
\begin{align}
  y_f^{\SM}(Q) = \frac{\sqrt{2} m_f^{\SM}(Q)}{v^{\SM}(Q)} \,,
\end{align}
%
with $f = u,d,c,s,t,b,e,\mu,\tau$.
The running top quark \MSbar\ mass in the non-SUSY model, $m_t(Q)$, is
calculated from the top pole mass $M_t$
using the full 1-loop self energy plus 2-loop SM QCD
corrections,
%
\begin{align}
\begin{split}
  m_t(Q) &= M_t + \re\Sigma_{t}^S(p^2=M_t^2,Q) \\
  &\phantom{={}} + M_t \Big[ \re\Sigma_{t}^L(p^2=M_t^2,Q) +
    \re\Sigma_{t}^R(p^2=M_t^2,Q) \\
  &\phantom{={} + M_t \Big[}
    + \Delta m_t^{(1),\text{qcd}}(Q) + \Delta m_t^{(2),\text{qcd}}(Q) \Big]
  \,,
\end{split}
\end{align}
%
where $\Sigma_{t}^{S,L,R}$ denote the scalar, left- and right-handed
parts of the 1-loop top self energy in the \MSbar\ scheme in the
non-SUSY model without the gluon
contribution, and $\Delta m_t^{(1),\text{qcd}}$ and $\Delta
m_t^{(2),\text{qcd}}$ are 1- and 2-loop gluon corrections taken from
Ref.~\cite{Fleischer:1998dw},
%
\begin{align}
  \Delta m_t^{(1),\text{qcd}}(Q) &=
  -\frac{g_3^2}{12 \pi^2} \left[4 - 3 \log\frac{m_t^2}{Q^2}\right],
  \label{eq:top-selfenergy-qcd-1L}\\
  \begin{split}
    \Delta m_t^{(2),\text{qcd}}(Q) &= \left(\Delta
      m_t^{(1),\text{qcd}}(Q)\right)^2 - \frac{g_3^4}{4608 \pi^4}
    \Bigg[396 \log^2\frac{m_t^2}{Q^2} \\
    &\phantom{={}} - 1452 \log\frac{m_t^2}{Q^2}
    - 48 \zeta(3) + 2053 + 16 \pi^2 (1+\log 4)\Bigg] \,.
  \end{split}
  \label{eq:top-selfenergy-qcd-2L}
\end{align}
%
In the SM, 3-loop QCD contributions can be taken into
account as well, see \secref{sec:model_specific_contributions}.
%
The bottom quark \MSbar\ mass in the non-SUSY model, $m_b(Q)$, is
obtained from the \MSbar\ mass $m_b^{\SM(5)}(m_b)$ in the SM with $5$ active quark flavors by first evolving
$m_b^{\SM(5)}(m_b)$ to the scale $Q$ using the 1-loop QED and 3-loop
QCD RGEs.  Afterwards, $m_b^{\SM(5)}(Q)$ is converted to $m_b(Q)$ as
%
\begin{align}
  m_b(Q) &= \frac{m_b^{\SM(5)}(Q)}{1 - \Delta m_b} \,, \\
  \Delta m_b &= \re\Sigma_{b}^S(p^2=(m_b^{\SM(5)})^2,Q)/m_b \notag\\
  &\phantom{={}}+ \re\Sigma_{b}^L(p^2=(m_b^{\SM(5)})^2,Q)
  + \re\Sigma_{b}^R(p^2=(m_b^{\SM(5)})^2,Q) \,,
  \label{eq:bottom-conversion}
\end{align}
%
where $\Sigma_{b}^{S,L,R}$ are the scalar, left- and right-handed
parts of the 1-loop bottom quark self energy in the \MSbar\ scheme, in
which all loops that contain only SM(5) particles are omitted.  Finally, the
\MSbar\ mass of the $\tau$ lepton, $m_\tau(Q)$, is calculated by first
identifying the $\tau$ pole mass, $M_\tau$, with the \MSbar\ mass in
the SM with 5 active quark flavors at the scale $M_\tau$,
%
\begin{align}
  m_\tau^{\SM(5)}(M_\tau) &=  M_\tau \,.
\end{align}
%
In this identification, the 1-loop SM electroweak
corrections to $m_\tau^{\SM(5)}(M_\tau)$ are neglected.  Afterwards,
$m_\tau^{\SM(5)}(M_\tau)$ is evolved to the scale $Q$ using the 1-loop QED RGE
and $m_\tau^{\SM(5)}(Q)$ is converted to $m_\tau(Q)$ as
%
\begin{align}
  m_\tau(Q) &= \frac{m_\tau^{\SM(5)}(Q)}{1 - \Delta m_\tau} \,, \\
  \Delta m_\tau &=
  \re\Sigma_{\tau}^S(p^2=(m_\tau^{\SM(5)})^2,Q) / m_\tau^{\SM(5)}(Q) \notag \\
  &\phantom{={}}
  + \re\Sigma_{\tau}^L(p^2=(m_\tau^{\SM(5)})^2,Q)
    + \re\Sigma_{\tau}^R(p^2=(m_\tau^{\SM(5)})^2,Q) \,,
\end{align}
%
where $\Sigma_{\tau}^{S,L,R}$ are the scalar, left- and right-handed
parts of the 1-loop $\tau$ self energy in the \MSbar\ scheme,
from which all loops that contain only SM(5) particles are omitted.

In most models, it is necessary to also fix the running SM-like vacuum
expectation value (VEV), $v$, at the low-energy scale.  For this purpose \fs
provides the symbols \code{MZMSbar} and \code{MWMSbar} in the model
file to access the \MSbar $W$ and $Z$ masses $m_W$ and $m_Z$ in the
non-SUSY model at the low-energy scale.  These running masses can be
used to calculate the \MSbar vacuum expectation value $v$, as for
example in the SM,
%
\begin{align}
  v(Q) &= \frac{2 m_Z(Q)}{\sqrt{3 g_1^2(Q) / 5+ g_2^2(Q)}} \,.
\end{align}

\subsection{Structure of the generated code}

In analogy to SUSY models, the parameters of a non-supersymmetric
model are distributed among two classes in the model class
hierarchy, see \figref{fig:parameter-classes}:
%
\begin{figure}[tbh]
  \centering
  \tikzumlset{fill class=white}
  \begin{tikzpicture}
    \umlclass[x=0, y=8, type=interface]{Beta\_function}{
    }{
      + \umlvirt{get()}\\
      + \umlvirt{set()}\\
      + \umlvirt{beta()}\\
      + run\_to()
    }
    \umlclass[x=0, y=4]{<model>\_susy\_parameters}{
      -- dimensionless parameters
    }{
      + get()\\
      + set()\\
      + beta()\\
      + run\_to()
    }
    \umlclass[x=0, y=0]{<model>\_soft\_parameters}{
      -- parameters with mass dimension > 0
    }{
      + get()\\
      + set()\\
      + beta()\\
      + run\_to()
    }
    \umlinherit{<model>\_susy\_parameters}{Beta\_function}
    \umlinherit{<model>\_soft\_parameters}{<model>\_susy\_parameters}
  \end{tikzpicture}
  \caption{Model class hierarchy.}
  \label{fig:parameter-classes}
\end{figure}
%
At the top of the model class hierarchy stands the
\code{Beta_function} interface class, which defines the interface for
the RGE integrator of the model parameters.  It provides the interface
function \code{run_to()}, which integrates the RGEs up to a given
scale using an adaptive Runge-Kutta algorithm.  The Runge-Kutta
algorithm makes use of the virtual functions \code{get()},
\code{set()} and \code{beta()} to obtain the model parameters at the
current renormalization scale, set the model parameters to new values
or calculate the $\beta$ functions.  These virtual functions are
implemented by the derived classes, \code{<model>_susy_parameters} and
\code{<model>_soft_parameters}.  Dimensionless parameters, like
gauge, Yukawa or quartic scalar couplings, are contained in the
\code{<model>_susy_parameters} class.  Parameters with mass dimension
greater than zero are contained in the \code{<model>_soft_parameters}
class.  The distribution of the model parameters between these two
classes reflects the dependency of the
$\beta$ functions of the dimensionful parameters upon the
dimensionless parameters.  Furthermore, it allows
the RGEs of the dimensionless parameters to be integrated independently of the
dimensionful parameters.

\subsection{Application: High-scale MSSM with light Higgs sector}
\label{sec:application_THDM}

As an application of \fbsm we consider the Higgs pole mass prediction
in an MSSM scenario with very heavy sfermions, Higgsinos and gauginos
at the SUSY scale $\MS$, but a light Higgs sector.  If $\MS$ is larger
than a few TeV, an EFT approach should be considered and the heavy SUSY
particles should be integrated out at $\MS$.  The resulting EFT below $\MS$ is
the Two-Higgs-Doublet-Model (\THDM).  Our aim is to calculate the Higgs pole
masses in this effective \THDM, where the quartic Higgs couplings are fixed by the
MSSM at $\MS$.\footnote{This \fs setup was also used in
  Ref.~\cite{Bagnaschi:2015pwa} to study the vacuum stability at very
  high SUSY scales in different \THDM variants with the MSSM as a
  supersymmetric UV completion.}  We use a \THDM of type II here, for
which the full 1- and leading 2-loop threshold corrections at the SUSY
scale are known \cite{Haber:1993an,Gorbahn:2009pp,Lee:2015uza}.

In order to construct such an EFT setup we have to build the \THDM-II
with \sarah.  We start by specifying the gauge group, the field
content and the Lagrangian:
%
\begin{lstlisting}
(* gauge groups *)
Gauge[[1]]={B,   U[1], hypercharge, g1,False};
Gauge[[2]]={WB, SU[2], left,        g2,True};
Gauge[[3]]={G,  SU[3], color,       g3,False};

(* field content *)
FermionFields[[1]] = {q , 3, {uL,dL}   , 1/6, 2,  3};
FermionFields[[2]] = {l , 3, {vL,eL}   ,-1/2, 2,  1};
FermionFields[[3]] = {d , 3, conj[dR]  , 1/3, 1, -3};
FermionFields[[4]] = {u , 3, conj[uR]  ,-2/3, 1, -3};
FermionFields[[5]] = {e , 3, conj[eR]  ,   1, 1,  1};
ScalarFields[[1]]  = {H1, 1, {H1p, H10}, 1/2, 2,  1};
ScalarFields[[2]]  = {H2, 1, {H2p, H20}, 1/2, 2,  1};

DEFINITION[GaugeES][Additional] = {
    {LagHC  , { AddHC->True  }},
    {LagNoHC, { AddHC->False }}
};

LagNoHC = -(M112 conj[H1].H1 + M222 conj[H2].H2 \
            + Lambda1 conj[H1].H1.conj[H1].H1 \
            + Lambda2 conj[H2].H2.conj[H2].H2 \
            + Lambda3 conj[H2].H2.conj[H1].H1 \
            + Lambda4 conj[H2].H1.conj[H1].H2 );

LagHC = -(-M122 conj[H1].H2
          + Lambda5/2 conj[H2].H1.conj[H2].H1
          + Lambda6 conj[H1].H1.conj[H1].H2
          + Lambda7 conj[H2].H2.conj[H1].H2
          + Yd conj[H1].d.q + Ye conj[H1].e.l + Yu H2.u.q);
\end{lstlisting}
%
The neutral components of the two Higgs doublets acquire vacuum
expectation values $v_1$ and $v_2$:
%
\begin{lstlisting}
DEFINITION[EWSB][VEVs] = {
    {H10, {v1, 1/Sqrt[2]},
          {sigma1, \[ImaginaryI]/Sqrt[2]}, {phi1, 1/Sqrt[2]}},
    {H20, {v2, 1/Sqrt[2]},
          {sigma2, \[ImaginaryI]/Sqrt[2]}, {phi2, 1/Sqrt[2]}}
};
\end{lstlisting}
%
and the Higgs field components mix to \CP-even, \CP-odd and charged
Higgs mass eigenstates \code{hh}, \code{Ah} and \code{Hm},
respectively:
%
\begin{lstlisting}
DEFINITION[EWSB][MatterSector] = {
    {{phi1, phi2}          , {hh, ZH}},
    {{sigma1, sigma2}      , {Ah, ZA}},
    {{conj[H1p], conj[H2p]}, {Hm, ZP}}
};
\end{lstlisting}
%
Now we need a \fs model file in which we specify the boundary
conditions for all \THDM-II parameters.  As input we use the \MSbar
parameter
$\tan\beta^{\THDMII}(M_t)$, the MSSM \DRbar parameters
$\mu(\MS)$ and $A_f(\MS)$
($f=t,b,\tau$) at the SUSY scale, and the \MSbar \CP-odd Higgs mass
$m_A^{\THDMII}(\MEWSB)$, where \MEWSB is the scale of the electroweak
symmetry breaking, at which we calculate the light \CP-even Higgs pole mass in the
end:
%
\begin{lstlisting}
MINPAR = {
    {3, TanBeta}
};

EXTPAR = {
    {0, MSUSY},
    {1, MEWSB},
    {2, MuInput},
    {6, MAInput},
    {7, AtInput},
    {8, AbInput},
    {9, AtauInput}
};
\end{lstlisting}
%
At the low-energy scale
$M_t$, we let \fs calculate the gauge and Yukawa couplings of the
\THDMII and we fix the two Higgs VEVs using
$m_Z^{\THDMII}(M_t)$ and $\tan\beta^{\THDMII}(M_t)$:
%
\begin{lstlisting}
LowScale = LowEnergyConstant[MT];

LowScaleInput = {
   {Yu, Automatic},
   {Yd, Automatic},
   {Ye, Automatic},
   {v1, 2 MZMSbar / Sqrt[GUTNormalization[g1]^2 g1^2 + g2^2] \
        Cos[ArcTan[TanBeta]]},
   {v2, 2 MZMSbar / Sqrt[GUTNormalization[g1]^2 g1^2 + g2^2] \
        Sin[ArcTan[TanBeta]]}
};
\end{lstlisting}
%
At the scale $\MEWSB$, we fix the \code{M122} parameter using the input
value of $m_A^{\THDMII}(\MEWSB)$ and we impose the EWSB conditions by
fixing \code{M112} and \code{M222}:
%
\begin{lstlisting}
EWSBOutputParameters = { M112, M222 };

SUSYScale = MEWSB;

SUSYScaleInput = {
    {M122, MAInput^2 Sin[ArcTan[v2/v1]] Cos[ArcTan[v2/v1]]},
    FSSolveEWSBFor[EWSBOutputParameters]
};
\end{lstlisting}
%
Finally, we need to fix the quartic Higgs couplings of the \THDMII at
the scale $\MS$.  The necessary relations between the MSSM parameters
and the quartic Higgs couplings of the \THDMII are known at the full
1-loop and leading 2-loop level
\cite{Haber:1993an,Gorbahn:2009pp,Lee:2015uza}.
We can use expressions from these references to write the boundary conditions
on the quartic Higgs couplings at the SUSY scale,
shown in lines 23--57 of \appref{app:THDMIIMSSMBC},
in terms of the 1- and 2-loop threshold corrections,
shown in lines 110--398 of the same listing
which displays the complete
\fs model file.
\figref{fig:THDMIIMSSMBC_Mh_MS} shows the lightest \CP-even
Higgs pole mass calculated at the 1-loop level with \fs in this EFT
setup as a function of $\tan\beta^{\THDMII}(M_t)$ and $\MS$ for
$m_A^{\THDMII}(\MEWSB) = 200\unit{GeV}$, $\MEWSB = M_t$ and maximal
stop mixing.  The figure shows that using this setup, \fs can reproduce
the results presented in the left panels of Figure~2 of
Ref.~\cite{Lee:2015uza}.  This EFT model is
distributed with the \fs package under the name \modelname{THDMIIMSSMBC}.
%
\begin{figure}[tbh]
  \centering
  \includegraphics[width=0.49\textwidth]{doc/plots/THDMIIMSSMBC/THDMIIMSSMBC_Mh_MS_TB_MA-200_Xt-0}\hfill
  \includegraphics[width=0.49\textwidth]{doc/plots/THDMIIMSSMBC/THDMIIMSSMBC_Mh_MS_TB_MA-200_Xt-sqrt6}
  \caption{Lightest \CP-even Higgs pole mass calculated at the 1-loop
    level in the effective \THDM setup as a function of $\tan\beta$ and
    $\MS$ for $m_A(M_t) = 200\unit{GeV}$, $A_t = \mu/\tan\beta +
    X_t$, $A_b = A_\tau = A_t$, $\mu = \MS$, $M_t = 173.34\unit{GeV}$
    and $\as^{\SM(5)}(M_Z) = 0.1184$.  The left panel shows the results for $X_t
    = 0$ and the right panel for $X_t = \sqrt{6}\MS$.  The solid line
    corresponds to a Higgs pole mass of $125\unit{GeV}$ and the dashed
    lines to $124\unit{GeV}$ and $126\unit{GeV}$, respectively.}
  \label{fig:THDMIIMSSMBC_Mh_MS}
\end{figure}

\subsection{Application: High-scale MSSM (\HSSUSY)}
\label{sec:application_HSSUSY}

\fbsm has already been applied in
Refs.~\cite{Bagnaschi:2015pwa,Athron:2016fuq,Bagnaschi:2017xid} to perform a
pure EFT calculation of the lightest \CP-even Higgs mass in the MSSM,
assuming that all SUSY particles are integrated out at a heavy SUSY
scale $\MS$.  The \fs spectrum generator constructed for this purpose
is called \HSSUSY and is based on \sarah's SM model file
(\modelname{SM}).  In \fs 1.2.3, \HSSUSY implemented the 1-loop and leading
2-loop threshold corrections of $O(\at\as + \at^2)$ to the quartic Higgs
coupling of the SM from
Refs.~\cite{Bagnaschi:2014rsa,Vega:2015fna} at the SUSY scale.  In
\fstwo, the generalized 2-loop expressions of $O(\at^2)$ for general stop
masses as well as the new 2-loop contributions from Ref.~\cite{Bagnaschi:2017xid},
which involve the bottom and tau Yukawa couplings, are
included.  As a result, the version of \HSSUSY included in \fstwo uses
the 2-loop threshold corrections of
$O(\at\as + \ab\as + (\at+\ab)^2 + \atau^2)$ for general SUSY spectra.

In the \HSSUSY model file
(\code{model_files/HSSUSY/FlexibleSUSY.m.in}), these threshold
corrections are implemented as \mathematica expressions in the
high-scale boundary condition:
%
\begin{lstlisting}
HighScaleInput = {
    {\[Lambda], lambdaTree                     (* tree-level *)
                + lambda1LReg + lambda1LPhi    (* 1-loop *)
                + lambda1LChi1 + lambda1LChi2  (* 1-loop *)
                + lambda1Lbottom + lambda1Ltau (* 1-loop *)
                + ...                          (* 2-loop *)
    }
};

(* arXiv:1407.4081, Eq. (3) *)
lambdaTree = 1/4 (g2^2 + 3/5 g1^2) Cos[2 ArcTan[TanBeta]]^2;

(* arXiv:1407.4081, Eq. (9) *)
lambda1LReg = 1/(4 Pi)^2 (
    - 9/100 g1^4 - 3/10 g1^2 g2^2
    - (3/4 - Cos[2 ArcTan[TanBeta]]^2/6) * g2^4
    );

(* arXiv:1407.4081, Eq. (10) *)
lambda1LPhi = 1/(4 Pi)^2 (
    3 Yu[3,3]^2 (
        Yu[3,3]^2
        + 1/2 (g2^2-g1^2/5) Cos[2 ArcTan[TanBeta]]
      ) Log[msq2[3,3]/SCALE^2]
    + 3 Yu[3,3]^2 (
        Yu[3,3]^2
        + 2/5 g1^2 Cos[2 ArcTan[TanBeta]]
      ) Log[msu2[3,3]/SCALE^2]
    + Cos[2 ArcTan[TanBeta]]^2/300 (
        3 (g1^4 + 25 g2^4) (
            + Log[msq2[1,1]/SCALE^2]
            + Log[msq2[2,2]/SCALE^2]
            + Log[msq2[3,3]/SCALE^2]
        )
        + ...
      )
    + ...
    );

...
\end{lstlisting}
%
The full expressions for the threshold corrections can be found in the
\HSSUSY model file.  In addition, \HSSUSY makes use of the known
3-loop SM $\beta$ functions, 3-loop corrections to the
running top Yukawa coupling, 3-loop threshold corrections to the
strong coupling and up to 3-loop corrections to the Higgs pole mass:
%
\begin{lstlisting}
UseSM3LoopRGEs    = True; (* 3-loop RGEs *)
UseYukawa3LoopQCD = True; (* 3-loop thresholds for yt *)
UseSMAlphaS3Loop  = True; (* 3-loop thresholds for alpha_s *)
UseHiggs2LoopSM   = True; (* 2-loop contributions to Mh *)
UseHiggs3LoopSM   = True; (* 3-loop contributions to Mh *)
\end{lstlisting}
%
With all these corrections enabled, \HSSUSY can be regarded as an
improved variant of \susyhd \cite{Vega:2015fna}, the difference being
that \HSSUSY includes generalized 2-loop threshold corrections also
involving $\ab$ and $\atau$, which are not present in \susyhd.  In
\secref{sec:feft}, \HSSUSY is also compared to the fixed-order
calculation in the full MSSM as well as to \feft.  The following
Examples \ref{ex:HSSUSY_parallel}--\ref{ex:HSSUSY_EFT_uncertainty}
show how \HSSUSY can be run and how different sources of uncertainty
can be estimated.\footnote{The example scripts can be found in the
  \code{doc/examples-2.0/} sub-directory of the \fs package.}
%
\begin{example}[label=ex:HSSUSY_parallel]
  For illustration, we show in this example the Higgs mass prediction
  in the MSSM with \HSSUSY.  In the following example script, a scan
  over the relative stop mixing parameter $X_t/\MS$ is performed for
  $\tan\beta = 5$ and three different values of the SUSY scale $\MS$.
  This is done in parallel on all available CPU cores.
  %
\begin{lstlisting}
Get["models/HSSUSY/HSSUSY_librarylink.m"];

CalcMh[TB_, Xt_, MS_] := Module[{handle, spectrum},
    handle = FSHSSUSYOpenHandle[
        fsSettings -> {
            precisionGoal -> 1.*^-5,
            calculateStandardModelMasses -> 1,
            poleMassLoopOrder -> 2,
            ewsbLoopOrder -> 2,
            betaFunctionLoopOrder -> 3,
            thresholdCorrectionsLoopOrder -> 2,
            poleMassScale -> 173.34
        },
        fsModelParameters -> {
            TanBeta -> TB,
            MEWSB -> 173.34,
            MSUSY -> MS,
            M1Input -> MS,
            M2Input -> MS,
            M3Input -> MS,
            MuInput -> MS,
            mAInput -> MS,
            AtInput -> (Xt + 1/TB) MS,
            msq2 -> MS^2 IdentityMatrix[3],
            msu2 -> MS^2 IdentityMatrix[3],
            msd2 -> MS^2 IdentityMatrix[3],
            msl2 -> MS^2 IdentityMatrix[3],
            mse2 -> MS^2 IdentityMatrix[3],
            LambdaLoopOrder -> 2,
            TwoLoopAtAs -> 1,
            TwoLoopAbAs -> 1,
            TwoLoopAtAb -> 1,
            TwoLoopAtauAtau -> 1,
            TwoLoopAtAt -> 1
        }
    ];
    spec = FSHSSUSYCalculateSpectrum[handle];
    FSHSSUSYCloseHandle[handle];
    If[spec =!= $Failed, Pole[M[hh]] /. (HSSUSY /. spec), 0]
];

LaunchKernels[];
DistributeDefinitions[CalcMh];

data = {
    ParallelMap[{#, CalcMh[5, #, 1000 ]}&, Range[-3.5, 3.5, 0.1]],
    ParallelMap[{#, CalcMh[5, #, 2000 ]}&, Range[-3.5, 3.5, 0.1]],
    ParallelMap[{#, CalcMh[5, #, 10000]}&, Range[-3.5, 3.5, 0.1]]
};
\end{lstlisting}%$
%
When plotting this data, the following figure results:
%
\begin{center}
  \includegraphics[width=0.5\textwidth]{doc/plots/HSSUSY/Mh_Xt}
\end{center}
\end{example}

\begin{example}[label=ex:HSSUSY_uncertainty]
  In this example, we slightly extend \exref{ex:HSSUSY_parallel}
  to make a partial uncertainty estimate of the Higgs pole
  mass predicted by \HSSUSY at the 2-loop level.  We do this by varying the
  threshold correction loop orders which determine the running \MSbar
  top Yukawa coupling
  $y_t$ and the strong coupling in the SM from 2-loop to
  3-loop.  In addition, the renormalization scale,
  at which the Higgs pole mass is calculated,
  is varied by a factor $2$.  The uncertainty
  estimated in this way is referred to as ``Standard Model
  uncertainty'' in the literature \cite{Vega:2015fna,Athron:2016fuq}
  and is one part of the full uncertainty of the EFT calculation
  of \HSSUSY.
  %
  \begin{lstlisting}
Get["models/HSSUSY/HSSUSY_librarylink.m"];

(* generate logarithmically spaced range [start, stop] *)
LogRange[start_, stop_, steps_] :=
    Exp /@ Range[Log[start], Log[stop],
                 (Log[stop] - Log[start])/steps];

(* generate logarithmically spaced range [Q / 2, 2 Q] *)
GenerateScales[Q_] := LogRange[Q/2, 2 Q, 10];

CalcMh[MS_, TB_, Xt_, ytLoops_, asLoops_, Qpole_] :=
    Module[{handle, spec},
    handle = FSHSSUSYOpenHandle[
        fsSettings -> {
            precisionGoal -> 1.*^-5,
            calculateStandardModelMasses -> 1,
            poleMassLoopOrder -> 2,
            ewsbLoopOrder -> 2,
            betaFunctionLoopOrder -> 3,
            thresholdCorrectionsLoopOrder -> 3,
            poleMassScale -> Qpole,
            thresholdCorrections -> 120111021 +
                ytLoops * 10^6 + asLoops * 10^2
        },
        fsModelParameters -> {
            TanBeta -> TB,
            MEWSB -> 173.34,
            MSUSY -> MS,
            M1Input -> MS,
            M2Input -> MS,
            M3Input -> MS,
            MuInput -> MS,
            mAInput -> MS,
            AtInput -> (Xt + 1/TB) * MS,
            msq2 -> MS^2 IdentityMatrix[3],
            msu2 -> MS^2 IdentityMatrix[3],
            msd2 -> MS^2 IdentityMatrix[3],
            msl2 -> MS^2 IdentityMatrix[3],
            mse2 -> MS^2 IdentityMatrix[3],
            LambdaLoopOrder -> 2,
            TwoLoopAtAs -> 1,
            TwoLoopAbAs -> 1,
            TwoLoopAtAb -> 1,
            TwoLoopAtauAtau -> 1,
            TwoLoopAtAt -> 1
        }
    ];
    spec = FSHSSUSYCalculateSpectrum[handle];
    FSHSSUSYCloseHandle[handle];
    If[spec =!= $Failed, Pole[M[hh]] /. (HSSUSY /. spec), 0]
];

(* calculate Higgs mass with uncertainty estimate *)
CalcDMh[MS_, TB_, Xt_] :=
    Module[{Mh, MhYt3L, MhAs3L, varyQpole, DMh},
           Mh     = CalcMh[MS, TB, Xt, 2, 2, 0];
           MhYt3L = CalcMh[MS, TB, Xt, 3, 2, 0];
           MhAs3L = CalcMh[MS, TB, Xt, 2, 3, 0];
           varyQpole = CalcMh[MS, TB, Xt, 2, 2, #]& /@
                       GenerateScales[173.34];
           (* combine uncertainty estimates *)
           DMh = Max[Abs[Max[varyQpole] - Mh],
                     Abs[Min[varyQpole] - Mh]] +
                 Abs[Mh - MhYt3L] + Abs[Mh - MhAs3L];
           { Mh, DMh }
          ];

LaunchKernels[];
DistributeDefinitions[CalcDMh];

data = {
    ParallelMap[{#, CalcDMh[1000 , 5, #]}&, Range[-3.5, 3.5, 0.1]],
    ParallelMap[{#, CalcDMh[2000 , 5, #]}&, Range[-3.5, 3.5, 0.1]],
    ParallelMap[{#, CalcDMh[10000, 5, #]}&, Range[-3.5, 3.5, 0.1]]
};
\end{lstlisting}%$
%
In the function \code{CalcDMh[]}, the three sources of uncertainty are
combined linearly.  When drawing the uncertainty $\Delta
M_h^{\SM}$, estimated in this way, around the central value as $M_h
\pm \Delta M_h^{\SM}$, the following figure results:
\begin{center}
  \includegraphics[width=0.5\textwidth]{doc/plots/HSSUSY/Mh_Xt_uncertainty}
\end{center}
%
We find that the uncertainty $\Delta
M_h^{\SM}$ is around or below $500\unit{MeV}$.
\end{example}

\begin{example}[label=ex:HSSUSY_EFT_uncertainty]
  In this example, we make an estimate of the ``EFT uncertainty''
  for the Higgs pole mass predicted by the pure-EFT calculation of
  \HSSUSY in the MSSM\@.  The pure-EFT calculation performed by \HSSUSY
  neglects terms of
  $O(v^2/\MS^2)$.  We estimate these missing terms by multiplying the
  individual 1-loop contributions by the term $(1 +
  \texttt{DeltaEFT}\times
  v^2/\MS^2)$, where \texttt{DeltaEFT} is an input parameter of
  \HSSUSY.  By varying \texttt{DeltaEFT} between $0$ and
  $1$ we obtain an estimate of the effect of these missing terms.
  This method has also been used in
  Refs.~\cite{Vega:2015fna,Athron:2016fuq,Bagnaschi:2017xid}.  The
  following code snippet illustrates how to calculate the Higgs mass
  with \HSSUSY and estimate the ``EFT uncertainty'' as a function of
  the SUSY scale for maximal stop mixing, $X_t =
  \sqrt{6}\MS$ and $\tan\beta = 5$.
  %
  \begin{lstlisting}
Get["models/HSSUSY/HSSUSY_librarylink.m"];

(* generate logarithmically spaced range [start, stop] *)
LogRange[start_, stop_, steps_] :=
    Exp /@ Range[Log[start], Log[stop],
                 (Log[stop] - Log[start])/steps];

CalcMh[MS_, TB_, Xt_, deltaEFT_] :=
    Module[{handle, spec},
    handle = FSHSSUSYOpenHandle[
        fsSettings -> {
            precisionGoal -> 1.*^-5,
            calculateStandardModelMasses -> 1,
            poleMassLoopOrder -> 2,
            ewsbLoopOrder -> 2,
            betaFunctionLoopOrder -> 3,
            thresholdCorrectionsLoopOrder -> 3,
            thresholdCorrections -> 122111221
        },
        fsModelParameters -> {
            TanBeta -> TB,
            MEWSB -> 173.34,
            MSUSY -> MS,
            M1Input -> MS,
            M2Input -> MS,
            M3Input -> MS,
            MuInput -> MS,
            mAInput -> MS,
            AtInput -> (Xt + 1/TB) * MS,
            msq2 -> MS^2 IdentityMatrix[3],
            msu2 -> MS^2 IdentityMatrix[3],
            msd2 -> MS^2 IdentityMatrix[3],
            msl2 -> MS^2 IdentityMatrix[3],
            mse2 -> MS^2 IdentityMatrix[3],
            LambdaLoopOrder -> 2,
            TwoLoopAtAs -> 1,
            TwoLoopAbAs -> 1,
            TwoLoopAtAb -> 1,
            TwoLoopAtauAtau -> 1,
            TwoLoopAtAt -> 1,
            DeltaEFT -> deltaEFT
        }
    ];
    spec = FSHSSUSYCalculateSpectrum[handle];
    FSHSSUSYCloseHandle[handle];
    If[spec =!= $Failed, Pole[M[hh]] /. (HSSUSY /. spec), 0]
];

(* calculate Higgs mass with uncertainty estimate *)
CalcDMh[MS_, TB_, Xt_] :=
    Module[{Mh, MhEFT},
           Mh    = CalcMh[MS, TB, Xt, 0];
           MhEFT = CalcMh[MS, TB, Xt, 1];
           { Mh, Abs[Mh - MhEFT] }
          ];

LaunchKernels[];
DistributeDefinitions[CalcDMh];

data = ParallelMap[{#, Sequence @@ CalcDMh[#, 5, Sqrt[6]]}&,
                   LogRange[173.34, 10^5, 60]];
\end{lstlisting}%$
%
In the function \code{CalcDMh[]}, the difference between the Higgs masses
calculated with $\texttt{DeltaEFT} = 0$ and $\texttt{DeltaEFT} =
1$ is used as the uncertainty estimate ($\texttt{DeltaEFT} =
0$ corresponds to the standard \HSSUSY calculation).  When drawing the
uncertainty estimated in this way symmetrically around the central
value, the following figure results:
%
\begin{center}
  \includegraphics[width=0.5\textwidth]{doc/plots/HSSUSY/Mh_MS_EFT_uncertainty}
\end{center}
%
As expected, we find that the ``EFT uncertainty'' decreases as
$\MS$ increases and falls below $150\unit{MeV}$ for $\MS \gtrsim 2\unit{TeV}$.
\end{example}

\subsection{Tests and comparisons with other spectrum generators}

We have performed various direct and indirect tests of \fbsm and its
components to ensure the correctness of the code:
%
\begin{itemize}
\item We have performed an analytic comparison of the split-MSSM RGEs
  generated with \sarah 4.5.3 against the RGEs presented in
  Ref.~\cite{Benakli:2013msa} and we found exact agreement.
\item We have performed a detailed numerical comparison of \fs's
  \HSSUSY model against \susyhd 1.0.1 and found excellent agreement
  \cite{Athron:2016fuq}.  The small differences between the two
  programs are of $O(100\unit{MeV})$ and originate from a
  different determination of $y_t$ in the SM at the low-energy scale,
  a different procedure to calculate the Higgs pole mass in the SM and
  the inclusion of additional 2-loop corrections in \HSSUSY which
  involve $\ab$ and $\atau$.
\item The effective \THDM shown in \secref{sec:application_THDM} and
  \appref{app:THDMIIMSSMBC} reproduces the results of
  Ref.~\cite{Lee:2015uza} and \MhEFT 1.0 \cite{MhEFT} for scenarios with
  heavy Higgsinos and gauginos, see for example
  \figref{fig:THDMIIMSSMBC_Mh_MS}.
\item The \fs package contains various EFT scenarios of the MSSM with
  boundary conditions from the literature (\HSSUSY,
  \modelname{SplitMSSM}, \modelname{THDMIIMSSMBC}, \modelname{HTHDMIIMSSMBC},
  \modelname{HGTHDMIIMSSMBC}).  For all these models, we have performed
  various analytic tests of the implemented MSSM boundary conditions,
  checking for example the renormalization scale dependence, relations
  among the parameters and threshold corrections at the
  1-loop level.
\item We have checked the numeric equality of the 3-loop MSSM $\beta$
  functions implemented in \fs and \softsusy 3.7.0.
\item We have checked the correctness of the renormalization scale
  dependent part of the 2-loop QCD corrections to the \MSbar top
  Yukawa coupling in the SM by deriving the 2-loop
  threshold corrections for $y_t$ from the SM to the MSSM
  and checking that no large logarithms appear
  \cite{ThomasKwasnitza:2016yqj}.
\item We have also analytically checked the correctness of the
  renormalization scale dependent part of the 2-loop and 3-loop QCD
  corrections to the \MSbar top Yukawa coupling in the SM
  by proving the renormalization scale invariance of the top quark
  pole mass in the SM at the 3-loop QCD level.
\end{itemize}

\section{\famu extension}
\label{sec:famu}

\fstwo introduces a calculation of the BSM contributions to the
anomalous magnetic moment of the muon, \amuBSM, in the \MSbar/\DRbar
scheme at the 1-loop level in the model under consideration plus the
universal 2-loop QED contributions
\cite{Degrassi:1998es,vonWeitershausen:2010zr}.  The 1-loop
diagram types that are taken into account are shown in
\figref{fig:amu_diagram_types}.
%
\begin{figure}[tbh]
  \centering
  \subfloat[Type FFS]{
    \includegraphics{doc/Feynman/amuFFS}
    \label{fig:amu_diagram_I}
  }
  \qquad
  \subfloat[Type SSF]{
    \includegraphics{doc/Feynman/amuFSS}
    \label{fig:amu_diagram_II}
  }
  \caption{Generic diagram types implemented in \fs to calculate the
    1-loop contribution to $\amuBSM$.  The solid line in the loop
    represents any contributing non-SM fermion $F$ and
    the dashed line any contributing scalar particle $S$.}
  \label{fig:amu_diagram_types}
\end{figure}
%
The general result for these two Feynman diagram types reads, in a
notation based on
Refs.~\cite{Stockinger:2006zn,Martin:2001st,Fargnoli:2013zia},
%
\begin{align}
  \label{eq:amu BSM 1L FFS}
  \amu^{\BSM\text{,1L,FFS}} &=
  \frac{c}{(4\pi)^2} \frac{M_\mu m_\mu}{m_{S}^2}\left(
    \frac{1}{12} A_{SF} F_1^C(x) + \frac{m_F}{3m_\mu} B_{SF} F_2^C(x)
  \right) \,, \\
  \label{eq:amu BSM 1L SSF}
  \amu^{\BSM\text{,1L,SSF}} &=
  - \frac{c}{(4\pi)^2} \frac{M_\mu m_\mu}{m_{S}^2}\left(
    \frac{1}{12} A_{SF} F_1^N(x) + \frac{m_F}{6m_\mu} B_{SF} F_2^N(x)
  \right) \,,
\end{align}
%
where $x=m_F^2/m_{S}^2$, $M_\mu$ is the muon pole mass,
$m_\mu$ is the muon \MSbar/\DRbar mass and $c$ denotes the electric charge of
the particle coupling to the photon divided by the muon charge.  The
$A_{SF}$ and $B_{SF}$ constants are defined as
%
\begin{align}
  \label{eq:ASF}
  A_{SF} &= z_L z_L^* + z_R z_R^* \,,\\
  \label{eq:BSF}
  B_{SF} &= z_L z_R^* + z_R z_L^* \,,
\end{align}
%
where $z_L$ and $z_R$ are the left- and right-handed components of the
scalar--fermion--muon vertex, $\Gamma_{SF\mu} =
ie(z_L P_L + z_R P_R)$.
%
The loop functions $F_i^C(x)$ and $F_i^N(x)$ read
%
\begin{align}
  F_1^C(x) &= \frac{2}{(1-x)^4} \Big[ 2 + 3x - 6x^2 + x^3 + 6x\log x \Big], \\
  F_2^C(x) &= \frac{3}{2(1-x)^3} \Big[ -3 + 4x - x^2 - 2\log x \Big], \\
  F_1^N(x) &= \frac{2}{(1-x)^4} \Big[ 1 - 6x + 3x^2 + 2x^3 - 6x^2\log x \Big], \\
  F_2^N(x) &= \frac{3}{(1-x)^3} \Big[ 1 - x^2 + 2x\log x \Big].
\end{align}
%
To calculate $\amu^{\BSM\text{,1L}}$, \fs sums over all possible
instantiations of these two 1-loop diagram types with at least one
non-SM particle in the loop,
%
\begin{align}
  \label{eq:amu BSM 1L}
  \amu^{\BSM\text{,1L}} =
  \sum_{\substack{\text{FFS}\\\text{diagrams}}} \amu^{\BSM\text{,1L,FFS}}
  + \sum_{\substack{\text{SSF}\\\text{diagrams}}} \amu^{\BSM\text{,1L,SSF}} \,.
\end{align}
%
The calculation is performed at the renormalization scale $\MS$, which
is defined as the running mass of the lightest electrically charged
\BSM particle contributing to $\amu^{\BSM\text{,1L}}$.  In addition to
this 1-loop \BSM contribution, \famu adds the universal leading
logarithmic 2-loop QED contribution \cite{Degrassi:1998es},
%
\begin{align}
  \amu^{\QED,\text{2L}} = \amu^{\BSM\text{,1L}} \times \Bigg[
    16 \, \frac{\aem^\BSM(\MS)}{4\pi}
    \log\frac{M_\mu}{\MS}
  \Bigg] \,.
\end{align}
%
The overall result for \amuBSM is then given by the sum of 1-loop and
the 2-loop QED contribution,
%
\begin{align}
  \amuBSM = \amu^{\BSM\text{,1L}} + \amu^{\QED,\text{2L}} \,.
\end{align}

\subsection{Choosing \famu in the model file}

In order to calculate \amuBSM with \fs, the
\code{FlexibleSUSYObservable`aMuon} symbol has to be added to an
output block, see \tabref{tab:famu_options} and
\exref{ex:FlexibleAMU_in_model_file}.
%
\begin{OptionTable}{\famu model file options.}{tab:famu_options}[lllX]
    \multilinecell[t]{\code{FlexibleSUSYObservable`}\\\code{aMuon}}
    & -- & -- & Represents \amuBSM calculated with \famu \\
    \multilinecell[t]{\code{FlexibleSUSYObservable`}\\\code{aMuonGM2Calc}}
    & -- & -- & Represents \amuMSSM calculated with \GMTCalc at 2-loop level \\
    \multilinecell[t]{\code{FlexibleSUSYObservable`}\\\code{aMuonGM2CalcUncertainty}}
    & -- & -- & Represents the uncertainty $\Delta\amuMSSM$ calculated with \GMTCalc \\
    \bottomrule
\end{OptionTable}

\begin{example}[label=ex:FlexibleAMU_in_model_file]
In the following code snippet, the \code{FlexibleSUSYObservable`aMuon} symbol
is added to an output block named \code{FlexibleSUSYLowEnergy}.
The calculated \amuBSM will be written to the entry $21$ of this
block.
%
\begin{lstlisting}
ExtraSLHAOutputBlocks = {
   {FlexibleSUSYLowEnergy,
           {{21, FlexibleSUSYObservable`aMuon} } }
};
\end{lstlisting}
\end{example}

Note that \amuBSM is treated as an observable in \fs.  The
calculation of observables can be enabled/disabled using the flag $15$
of the \code{FlexibleSUSY} block.

As an alternative to \famu, \GMTCalc \cite{Athron:2015rva} can be used
to calculate \amuMSSM in \fs at the 2-loop level employing results from
Refs.~\cite{Fargnoli:2013zia,Fargnoli:2013zda,vonWeitershausen:2010zr}
in MSSM models without flavor violation. This MSSM-specific option was
first introduced in \fsbreak 1.3.0.  In order to let \GMTCalc calculate the
anomalous magnetic moment of the muon, the symbol
\code{FlexibleSUSYObservable`aMuonGM2Calc} must be added to an output
block.  In addition, the symbol
\code{FlexibleSUSYObservable`aMuonGM2CalcUncertainty} can be used to
calculate the estimated corresponding theory uncertainty for \amuMSSM.

\begin{example}
Calculating \amuMSSM with \GMTCalc in the \modelname{CMSSMNoFV} is
enabled by defining:
%
\begin{lstlisting}
ExtraSLHAOutputBlocks = {
   {FlexibleSUSYLowEnergy,
           {{1, FlexibleSUSYObservable`aMuonGM2Calc},
            {2, FlexibleSUSYObservable`aMuonGM2CalcUncertainty} } }
};
\end{lstlisting}
\end{example}

\GMTCalc is incorporated into \fs in the form of an addon.
In order to enable the \GMTCalc addon, the
\code{--with-addons=GM2Calc} argument can be passed to the
\code{configure} script during the \fs configuration step.

\subsection{Structure of \famu code}

The C++ interface for \famu is defined in the
\code{<model>_a_muon.hpp} file.  The interface consists of a single
function, which takes a model object as the argument and returns the value
of \amuBSM.

\begin{example}
In the \modelname{CMSSM}, the interface function reads:
%
\begin{lstlisting}
namespace CMSSM_a_muon {
   double calculate_a_muon(const CMSSM_mass_eigenstates& model);
}
\end{lstlisting}
\end{example}

\subsection{Tests and comparisons with other calculations}

We have tested \famu in the MSSM against \GMTCalc 1.3.3
\cite{Athron:2015rva}, which is currently the most precise code
available to calculate \amuMSSM.  We usually find a $1$--$10\%$
deviation depending on the parameter choice, see
\tabref{tab:test_famu}, except for specific parameter points, where
the deviation is much larger due to a large renormalization scale
dependence, see below.  The $1$--$10\%$ deviation is caused mainly by
the fact that \GMTCalc calculates \amuMSSM in an on-shell scheme and
includes all known 2-loop corrections, while \famu performs the
calculation in the \MSbar/\DRbar scheme at the SUSY scale and only
includes the leading logarithmic 2-loop QED correction.  A comparison
of \famu with \spheno 4.0.2 shows differences of up to $10\%$.  These
differences are caused by the fact that \famu includes the leading
logarithmic 2-loop QED contribution and by the different choice of the
renormalization scale: \spheno performs the calculation at the $Z$ pole
mass scale, while \famu performs it at $\MS$.

In specific parameter scenarios, where the \MSbar/\DRbar smuon masses
show a high sensitivity to the renormalization scale, the predictions
of $\amuBSM$ in \famu and \spheno are expected to have a large theory
uncertainty, see for example BM4$^\prime$ in \tabref{tab:test_famu}
and see the discussion in Ref.~\cite{Athron:2015rva}.  One
reason for this large uncertainty are the missing 2-loop corrections
in \famu and \spheno, which would (if included) cancel the
renormalization scale dependence at the 2-loop level.  To illustrate
the sensitivity on the renormalization scale, we show in the third
column of \tabref{tab:test_famu} the variation of $\amuBSM$ in \famu
when the scale is varied in the interval $[\MS/2,2\MS]$.  For BM4$^\prime$,
$\amuBSM$ varies by around $66\%$, which indicates a very imprecise
prediction for this point.  Such a large scale uncertainty is avoided
in \GMTCalc due to the renormalization in an on-shell scheme.  Another
reason for the larger uncertainty in \famu and \spheno is the
renormalization scheme used to renormalize the smuon masses: In the
\MSbar/\DRbar scheme, smuon self energy contributions which are
quadratic in the BSM particle masses can lead to large 1-loop
corrections to the smuon masses.  Such large corrections are avoided in
the on-shell scheme, where they are absorbed into the smuon mass
counter term.
%
\begin{table}[tbh]
  \centering
  \begin{tabular}{lrrrr}
    \toprule
    Point & \famu & \famu           & \GMTCalc & \spheno \\
          &       & scale variation &          &         \\
    \midrule
    SPS1a  & $29.77$ & $0.46$ & $29.31 \pm 2.36$ & $31.00$ \\
    SPS1b  & $32.46$ & $0.45$ & $32.38 \pm 2.40$ & $32.68$ \\
    SPS3   & $13.80$ & $0.12$ & $13.52 \pm 2.33$ & $14.99$ \\
    SPS4   & $50.02$ & $1.02$ & $52.45 \pm 2.64$ & $45.64$ \\
    BM1$^\prime$  & $42.08$ & $1.58$ & $42.34 \pm 2.33$ & $43.72$ \\
    BM2$^\prime$  & $25.79$ & $0.10$ & $25.67 \pm 2.32$ & $26.16$ \\
    BM3$^\prime$  & $27.81$ & $0.68$ & $27.95 \pm 2.34$ & $27.98$ \\
    BM4$^\prime$  &  $8.11$ & $5.41$ & $33.11 \pm 2.31$ &  $2.19$ \\
    \bottomrule
  \end{tabular}
  \caption{Comparison of $\amuMSSM\cdot 10^{10}$ calculated with \famu,
    \GMTCalc 1.3.3 and \spheno 4.0.2 for the CMSSM benchmark points
    presented in Ref.~\cite{Allanach:2002nj} and the parameter points shown
    in \tabref{tab:bm_amu}.  The third column shows the variation
    of $\amuBSM$ when the renormalization scale is varied by a factor $2$.
    For BM4$^\prime$ the value of \amu calculated by the \DRbar
    programs \famu and \spheno suffers from a high renormalization scale
    sensitivity due to the large values of $M_2$ and $(m_{\tilde{l}})_{ii}$,
    which leads to a very imprecise result and to a huge deviation
    compared to \GMTCalc.}
  \label{tab:test_famu}
\end{table}
%
\begin{table}[tbh]
  \centering
  \begin{tabular}{lrrrr}
    \toprule
                & BM1$^\prime$  & BM2$^\prime$   & BM3$^\prime$   & BM4$^\prime$ \\
    \midrule
    $\mu$ / GeV & $350$ & $1300$ & $2000$ & $-160$ \\
    $\tan\beta$ & $40$  & $40$   & $40$   & $50$ \\
    $M_1$ / GeV & $150$ & $150$  & $150$  & $140$ \\
    $M_2$ / GeV & $300$ & $300$  & $300$  & $2000$ \\
    $(m_{\tilde{e}})_{ii}$ / GeV & $400$ & $400$  & $400$  & $200$ \\
    $(m_{\tilde{l}})_{ii}$ / GeV & $400$ & $400$  & $400$  & $2000$ \\
    $(m_{\tilde{q}})_{ii}$, $(m_{\tilde{u}})_{ii}$, $(m_{\tilde{d}})_{ii}$ / GeV
    & $400$ & $600$  & $700$ & $2000$ \\
    \bottomrule
  \end{tabular}
  \caption{Definition of the MSSM benchmark points BM1$^\prime$--BM4$^\prime$, inspired
    by the points BM1--BM4 presented in Ref.~\cite{Fargnoli:2013zda}.
    All parameters are defined in the \DRbar scheme at the scale
    $Q = 454.7\unit{GeV}$, except for $\tan\beta$, which is defined at $M_Z$.
    The trilinear couplings and off-diagonal
    elements of the sfermion mass parameters are set to zero and we
    have fixed $m_A = 2\unit{TeV}$.}
  \label{tab:bm_amu}
\end{table}

\section{\fcpv extension}
\label{sec:fcpv}
\subsection{Setting up a \fcpv model}

Since \fs 1.1.0, the model parameters are no longer restricted to be real,
but can be complex.  Whether a parameter is real or complex is
specified in the corresponding \sarah model file.  Parameters can be
forced to be treated as real in \fs by adding them to the
\code{RealParameters} list in the \fs model file, see
\tabref{tab:fcpv_options} and the following examples.  For
compatibility with \fs 1.0, the \code{RealParameters} list is by default
set to \code{\{All\}} meaning that all parameters are assumed to be real.
%
\begin{OptionTable}{\fcpv model file options.}{tab:fcpv_options}
    \code{RealParameters} & \code{\{ All \}} &
    List of model parameters or \code{\{\}} or \code{\{ All \}} &
    List of parameters to be treated as real\\
    \bottomrule
\end{OptionTable}

\begin{example}
In the MSSM, the $\mu$ parameter, the Yukawa
couplings, the soft-breaking trilinear couplings, the soft-breaking
scalar mass parameters, the soft-breaking gaugino mass parameters and
the $B\mu$ parameter can be complex.  In order to choose all of these
parameters to be complex, except for $B\mu$, one can set
%
\begin{lstlisting}
RealParameters = { B[\[Mu]] };
\end{lstlisting}
\end{example}

\begin{example}
In order to treat all MSSM parameters defined in
the \sarah model file for the MSSM as complex, set \code{RealParameters} to the
empty list:
%
\begin{lstlisting}
RealParameters = {};
\end{lstlisting}
\end{example}

\subsection{Application: CMSSM with \CP-violation}

In \sarah's \modelname{MSSM} model, the phase factor between the two Higgs
doublets is set to zero.  Therefore, this model does not allow for
\CP-violation in the Higgs sector.  In order to enable \CP-violation in
the MSSM Higgs sector, \sarah's \modelname{MSSM/CPV} model file can be
used, which allows for a non-zero relative phase factor $e^{i\eta}$
between the Higgs doublets.  In the \modelname{MSSM/CPV}, there are three
linearly independent EWSB equations.  Therefore, in this model three
EWSB output parameters have to be chosen.  In \fs's \modelname{CMSSMCPV}
model file, these are chosen to be $\re B\mu$, $\im B\mu$ and $|\mu|$
by setting the \code{EWSBOutputParameters} variable to
%
\begin{lstlisting}
EWSBOutputParameters = { Re[B[\[Mu]]], Im[B[\[Mu]]], \[Mu] };
\end{lstlisting}
%
Since only the magnitude of the $\mu$ parameter is fixed by the EWSB
equations, \fs introduces the phase of $\mu$ as a free parameter,
\code{Phase[\\[Mu]]} = $e^{i\phi_\mu}$.  This phase should be specified in
an SLHA-2 compliant way by reading the real and imaginary parts of
$e^{i\phi_\mu} = \cos\phi_\mu + i \sin\phi_\mu$ from the \code{MINPAR}
and \code{IMMINPAR} block entries $4$ and fixing $e^{i\phi_\mu}$ at the
SUSY scale:
%
\begin{lstlisting}
MINPAR   = { {4, CosPhiMu} };
IMMINPAR = { {4, SinPhiMu} };

SUSYScaleInput = {
    {Phase[\[Mu]], CosPhiMu + I SinPhiMu}
};
\end{lstlisting}
%
In the \modelname{CMSSMCPV}, the phase $\eta$ is read from the \code{EXTPAR}
block entry $100$ and also chosen to be input at the SUSY scale:
%
\begin{lstlisting}
EXTPAR = {
    {100, etaInput}
};

SUSYScaleInput = {
    {eta, etaInput}
};
\end{lstlisting}
%
The complete \fs \modelname{CMSSMCPV} model file can be found in
\appref{app:CMSSMCPV}.

\subsection{Application: Electric dipole moments of fermions}

\fstwo can calculate $\edmBSM{f}$, new physics contributions to the
electric dipole moment (EDM) of a fermion $f$, in the given model
in the \DRbar scheme at the 1-loop level.
The procedure is very similar to the calculation of \amuBSM
described in \secref{sec:famu}.
This is expected from the following effective Lagrangian,
\begin{equation}
  \Delta \mathcal{L}_\text{eff} =
  - \frac{D_f}{2}\bar{f}_L \sigma_{\mu\nu} f_R F^{\mu\nu} + \mathrm{h.c.},
\end{equation}
where the real and the imaginary parts of the Wilson coefficient
$D_f$ are proportional to the magnetic
(see, e.g., Refs.~\cite{Cho:2001hx,Baek:2002cc})
and the electric (see, e.g., Ref.~\cite{Pokorski:1999hz}) dipole moments
of $f$, respectively.
More precisely,
\begin{equation}
  a_f = - \frac{2 M_f}{e} \re D_f, \qquad
  \edm{f} = \im D_f,
\end{equation}
where $M_f$ is the pole mass of $f$.
It is then obvious that
the EDM is given by the sum of all FFS-type and SSF-type diagrams,
\begin{align}
  \edm{f}^{\BSM\text{,1L}} =
  \sum_{\substack{\text{FFS}\\\text{diagrams}}}
  \edm{f}^{\BSM\text{,1L,FFS}}
  + \sum_{\substack{\text{SSF}\\\text{diagrams}}}
  \edm{f}^{\BSM\text{,1L,SSF}}
  \,,
\end{align}
as in Eq.\ \eqref{eq:amu BSM 1L}.  The type of diagram refers to those
shown in \figref{fig:amu_diagram_types}, resulting in the contributions,
\begin{align}
  \frac{1}{e}\,\edm{f}^{\BSM\text{,1L,FFS}} &=
  \frac{c}{(4\pi)^2}
    \frac{m_F}{6{m_{S}^2}} \widetilde{B}_{SF} F_2^C(x)
  \,, \\
  \frac{1}{e}\,\edm{f}^{\BSM\text{,1L,SSF}} &=
  - \frac{c}{(4\pi)^2}
    \frac{m_F}{12{m_{S}^2}} \widetilde{B}_{SF} F_2^N(x)
  \,,
\end{align}
which are essentially the second terms of Eq.\ \eqref{eq:amu BSM 1L FFS} and
Eq.\ \eqref{eq:amu BSM 1L SSF}, respectively, divided by $2 M_f$, except that
the coupling factor is instead
\begin{equation}
  \widetilde{B}_{SF} = 2 \im ( z_L z_R^* ) .
\end{equation}
There are no imaginary parts corresponding to
the first terms of Eq.\ \eqref{eq:amu BSM 1L FFS} and Eq.\ \eqref{eq:amu BSM 1L SSF}
as can be guessed from Eq.\ \eqref{eq:ASF}.

One can have the EDM of particle \code{f} calculated
by adding to an output block the form: \code{FlexibleSUSYObservable`EDM[f_]}.
Then $\edm{f}$ is reported in units of $\unit{GeV}^{-1}$.
\begin{example}
The output includes
the EDMs of the electron, muon, and tau if
the following code snippet is inserted into the \fs model file:
\begin{lstlisting}
ExtraSLHAOutputBlocks = {
   {FlexibleSUSYLowEnergy,
           {{23, FlexibleSUSYObservable`EDM[Fe[1]]},
            {24, FlexibleSUSYObservable`EDM[Fe[2]]},
            {25, FlexibleSUSYObservable`EDM[Fe[3]]} } }
};
\end{lstlisting}
\end{example}

\begin{figure}
  \centering
  \includegraphics[width=0.49\textwidth]{doc/plots/MSSMCPV/contour_de_PhiMuM12}\hfill
  \includegraphics[width=0.49\textwidth]{doc/plots/MSSMCPV/contour_de_PhiMuTe}
  \caption{Left panel: contours of the electron EDM in the
    \CP-violating MSSM at the 1-loop level as a function of
    the phases of $\mhalf = M_1 = M_2$ and $\mu$,
    with $\mu A_e$ kept positive real.
    Right panel: same plot except for the vertical axis showing
    the $A_e$ phase and $\mu \mhalf$ kept positive real.
    In both panels, the other relevant parameters are:
    $\tan\beta = 10$, $|\mu| = |\mhalf| = \mzero = |A_e| = 2\unit{TeV}$.
    The thick red lines show the experimental upper bound on $|\edm{e}|$
    \cite{Baron:2013eja}.}
  \label{fig:EDM plots}
\end{figure}
Using a model file thus configured, one can do a quick test of
the electron EDM evaluation in the \CP-violating MSSM for instance.
In this model, only $\arg(\mu M_{1,2,3})$ and $\arg (\mu A_f)$ are
physical among the flavor-conserving phases
apart from those already present in the SM
(see, e.g., Refs.~\cite{Pokorski:1999hz,Stockinger:2006zn}).
For simplicity,
the gaugino masses are assumed to have an equal (complex) value
$\mhalf$, and both soft masses of the left- and right-handed
selectron shall be $\mzero$.
The moduli of $\mu$, $\mhalf$, $\mzero$, and $A_e$
are specified at the SUSY scale,
all of which including the scale are set to $2\unit{TeV}$.
Finally, we set $\tan\beta = 10$ at the SUSY scale.

The resulting electron EDM is displayed as contours in
\figref{fig:EDM plots}.
In the left panel, $\phi_{\mhalf} \equiv \arg \mhalf$ and
$\phi_\mu \equiv \arg \mu$ are varied
while $\arg(\mu A_e)$ is constrained to be zero.
From the directions of the contours it is clear
that \fs correctly reproduces the behaviour of
$\edm{e}$ depending only on the rephasing invariant $\arg(\mu \mhalf)$.
In the right panel, the roles of $\phi_{A_e} \equiv \arg A_e$ and
$\phi_{\mhalf}$ are swapped so that $\arg(\mu \mhalf)$ stays at zero.
Again, the contours verify that
$\edm{e}$ from \fs is determined by the physical phase $\arg(\mu A_e)$.
For reference,
the experimental upper limit on $|\edm{e}|$ at the 90\% confidence level
is shown as the thick red lines \cite{Baron:2013eja}.

\subsection{Tests and comparisons with other spectrum generators}
%
In the left panel of \figref{fig:CMSSMCPV_plots}, we show the lightest
Higgs pole mass calculated in the \CP-violating CMSSM at the 1-loop
level with \fstwo and \spheno~4.0.2 as a function of the phase angle
of the complex $\mu$ parameter, $\phi_\mu \equiv \arg \mu$.
We use a low-energy scenario with $\mzero = \mhalf = 500\unit{GeV}$,
$\tan\beta = 10$ and $\azero = 0$.  Even though this scenario is
excluded, the figure illustrates that both \fs and \spheno show the
same behaviour of the Higgs pole mass as a function of $\phi_\mu$.
The $\phi_\mu$-independent shift of around $0.7\unit{GeV}$ between the
Higgs masses calculated by the two programs is mainly caused by the
different treatment of higher-order corrections to
the running \DRbar top Yukawa coupling, which has been discussed in
Refs.~\cite{Staub:2015aea,ThomasKwasnitza:2016yqj}.
%
\begin{figure}[tbh]
  \centering
  \includegraphics[width=0.49\textwidth]{doc/plots/CMSSMCPV/scan_Mh_PhiMu_m0-500_M12-500_TB-10_A0-0}\hfill
  \includegraphics[width=0.49\textwidth]{doc/plots/CMSSMCPV/scan_TYu33_Q_m0-500_M12-500_TB-10_A0-500_PhiMu-0_PhiA0-Pi4}
  \caption{Left panel: lightest Higgs pole mass calculated in the
    \CP-violating CMSSM at the 1-loop level with \fstwo and
    \spheno 4.0.2 as a function of the phase angle of the $\mu$
    parameter for $\mzero = \mhalf = 500\unit{GeV}$, $\tan\beta = 10$
    and $\azero = 0$.  Right panel: renormalization group running of the
    trilinear coupling $(T^u)_{33}$ for $\mzero = \mhalf =
    500\unit{GeV}$, $\tan\beta = 10$, $\phi_\mu = 0$ and $\azero = 500\,
    e^{i \pi/4} \unit{GeV}$.}
  \label{fig:CMSSMCPV_plots}
\end{figure}

The right panel of \figref{fig:CMSSMCPV_plots} shows the
renormalization group running of the $(3,3)$-component of the complex
trilinear coupling $T^u$ for the same CMSSM scenario, except that we
use $\phi_\mu = 0$ and $\azero = 500\, e^{i \pi/4} \unit{GeV}$.  The
lines show the running in \fstwo and the dots the
running in \spheno 4.0.2.  We find very good agreement of
the running of $(T^u)_{33}$ between the two programs over the shown
$14$ orders of magnitude. The maximum deviation between \fs and
\spheno is around $2\%$ for $Q \approx 10^{16} \unit{GeV}$ and is
mainly caused by a different determination of the
dimensionless MSSM parameters at the electroweak scale.

\section{\fmw extension}
\label{sec:fmw}
\fmw is a major new feature which is released in \fstwo.
It allows a more accurate determination of the electroweak gauge
couplings $g_1$ and $g_2$, the prediction of the $W$ pole mass and thus
enables direct comparisons with electroweak precision data.

The running electroweak gauge couplings are related to the running
electromagnetic coupling~$e$, the weak mixing angle~$\theta_W$ and the
normalization factors~$N$ by
%
\begin{align}
  g_1 &= N_{g_Y}\,\frac{e}{\cos\theta_W}\,,
  & g_2 &= N_{g_L}\,\frac{e}{\sin\theta_W} \,.
\end{align}
%
The electromagnetic gauge coupling can be obtained from the known
value of $\aem^{\SM(5)}(M_Z)$ as described in
\secref{sec:fbsm} using the known general form of the threshold
correction, Eq.\ \eqref{eq:alpha_em_general_threshold}.
The weak mixing angle, together with the $W$ and $Z$ pole masses and
the muon decay constant $G_F$, form a set of four electroweak
precision quantities. Within the theory, only two of them are
independent. Hence, the running weak
mixing angle can be calculated in different ways:
%
\begin{enumerate}
\item Using the $W$ and $Z$ pole masses $M_W$ and $M_Z$ as input and
  calculating the cosine of the running weak mixing angle for example
  as:
  \begin{align}
    \cos^2\theta_W = \frac{m_W^2}{m_Z^2} \,,
    \label{eq:cosW_from_MWMZ}
  \end{align}
  where $m_W$ and $m_Z$ are the running $W$ and $Z$ masses which are
  obtained from the corresponding pole masses by
  \begin{align}
    m_V^2(Q) = M_V^2 + \re \Sigma_{V,T}(p^2 = M_V^2, Q) \,, \qquad (V=W,Z)
  \end{align}
  with $\Sigma_{V,T}$ being the transverse part of the vector boson
  self energy. In models with Higgs triplets having the
  vacuum expectation value $v_T$ for example,
  Eq.~\eqref{eq:cosW_from_MWMZ} is adapted accordingly:
  \begin{align}
    \cos^2\theta_W = \frac{m_W^2 - g_2^2 v_T^2}{m_Z^2} \,.
  \end{align}
  This approach was implemented in the first release of \fs \cite{Athron:2014yba},
  as a generalization of the calculation in Ref.\ \cite{Athron:2012pw}.
\item Using the $Z$ pole mass and the measured muon decay constant
  $G_F$ as input and determining $\theta_W$ and the $W$ pole mass
  as described below.
\end{enumerate}
%
The first approach has the advantage that it can be easily applied to
any \BSM model with a $W$ and $Z$ boson, because the 1-loop calculation of
the running $W$ and $Z$ masses can be fully automatized since only
self energies are necessary.  However, this approach has the
disadvantage that the parametric uncertainty of the calculated
electroweak gauge couplings is then limited by that
of the measured $W$ pole mass of the order $0.02\%$ \cite{Olive:2016xmw}.
This prohibits a meaningful computation of other electroweak precision
observables for which more precise experimental data exist.

The second approach is more complicated to automatize for all \BSM
models, because also 1-loop vertex and box diagrams contributing to
the muon decay have to be taken into account.  However, the approach
has the advantage that the parametric uncertainty of the calculated
electroweak gauge couplings is related to those
of $G_F$ and $M_Z$, which are of the order $0.00005\%$
and $0.002\%$, respectively \cite{Olive:2016xmw}. As a result, the $W$
pole mass is now a meaningful prediction, which can be used to
constrain BSM models.

Before version 2.0, only the first approach could be used in \fs for
all \BSM models.  The second approach was only available in models
which are SM-like or (N)MSSM-like, an option introduced in \fs 1.1.0 (as described in Ref.\ \cite{Staub:2015aea}).
\fstwo is now able to apply the second approach to all \BSM models
which have a $W$ and a $Z$ boson and which have the SM gauge
group as a gauge group factor.

For the implementation of the second approach, \fs uses a
generalization of the procedure presented in Ref.~\cite{Degrassi:1990tu}
for the SM, which has been adapted to the MSSM in Ref.~\cite{Pierce:1996zz}.
The running weak mixing angle is extracted from the relation
\cite{Degrassi:1990tu}
%
\begin{align}
  \sin^2\theta_W \cos^2\theta_W =
  \frac{\pi\,\aem}
   {\sqrt{2}\,M_Z^2\,G_F\,\hat\rho_\text{tree} \left(1-\Delta\hat{r}\right)}\,,
  \label{eq:muon_decay_master}
\end{align}
%
where
%
\begin{align}
  \Delta\hat{r} &= \Delta\hat{r}_\text{1L} + \Delta\hat{r}_\text{2L}^\SM \\
  &\text{with}\;\;
  \Delta\hat{r}_\text{1L} = \frac{1}{1-\Delta\hat\rho}\,\frac{\re\Sigma_{W,T}(0)}{M_W^2}
  - \frac{\re\Sigma_{Z,T}(M_Z^2)}{M_Z^2} + \deltaVB\,,
  \label{eq:Delta_r_hat}\\
  \Delta\hat\rho &= \frac{1}{1 + \frac{\re\Sigma_{Z,T}(M_Z^2)}{M_Z^2}}
  \left[\frac{\re\Sigma_{Z,T}(M_Z^2)}{M_Z^2} - \frac{\re\Sigma_{W,T}(M_W^2)}{M_W^2}
   + \Delta\hat\rho_\text{2L}^\SM\right].
   \label{eq:Delta_rho_hat}
\end{align}
%
In the occurring self energies $\Sigma$, the top quark mass is chosen
to be the pole mass in order to include partial 2-loop corrections not
contained in $\Delta\hat{r}_\text{2L}^\SM$ and $\Delta\hat\rho_\text{2L}^\SM$
\cite{Fanchiotti:1992tu}.
Since $\Delta\hat{r}$ and $\Delta\hat\rho$ themselves depend
on $\theta_W$, an iteration including these equations has to be
performed to get a self-consistent solution.

The quantity
%
\begin{align}
  \hat\rho_\text{tree} &= \rho_0\,\frac{m_{Z,\SM}^2}{m_{Z,\text{mix}}^2}
  \label{eq:rho_hat_tree}
\end{align}
%
introduces two different generalizations. On the one hand, corrections
from higher dimensional Higgs multiplets are included via \cite{Langacker:1991an}
%
\begin{align}
  \rho_0 &= \frac{\sum_i \left(t_i^2 - t_{3i}^2 + t_i \right)|v_{\varphi_i}|^2}
   {\sum_i 2\,t_{3i}^2\,|v_{\varphi_i}|^2}\,,
\end{align}
%
where the sums run over all neutral Higgs fields~$\varphi_i$ with vacuum
expectation value~$v_{\varphi_i}$, weak isospin~$t_i$ and its third
component~$t_{3i}$. On the other hand, corrections from extra $U(1)$
gauge groups are included via the ratio of the SM-like tree-level $Z$
boson mass $m_{Z,\SM}$ and the tree-level $Z$ boson mass $m_{Z,\text{mix}}$
including mixing with additional $Z^\prime$ bosons
\cite{Degrassi:1989mu,Leike:1991if}.

The leading SM 2-loop contributions to $\Delta\hat{r}$ and
$\Delta\hat\rho$ \cite{Fanchiotti:1992tu,Pierce:1996zz} are given
by\footnote{The specific numerical values in these formulas depend
on the renormalization scale, which is assumed to be $M_Z$, and on
the choice of the top pole mass in the self energies from
Eqs.~\eqref{eq:Delta_r_hat} and \eqref{eq:Delta_rho_hat}.}
%
\begin{align}
 \begin{split}
  \Delta\hat{r}_\text{2L}^\SM &= \frac{\aem \,\as^{\SM(5)}}
   {4 \pi^2 \sin^2\theta_W \cos^2\theta_W}
  \left[2.145\,\frac{M_t^2}{M_Z^2} + 0.575 \log\mathopen{}\left(\frac{M_t}{M_Z}
  \right)\mathclose{} - 0.224 - 0.144\,\frac{M_Z^2}{M_t^2}\right] \\
  & \phantom{={}} - \delta_\text{Higgs} \,\frac{1 - \Delta\hat{r}_\text{1L}}
   {1 - \Delta\hat\rho}\,,
   \label{eq:Delta_r_hat_SM_2L}
 \end{split} \\[1em]
  \Delta\hat\rho_\text{2L}^\SM &= \frac{\aem \,\as^{\SM(5)}}
   {4 \pi^2 \sin^2\theta_W}
  \left[-2.145\,\frac{M_t^2}{M_W^2} + 1.262 \log\mathopen{}\left(\frac{M_t}{M_Z}
  \right)\mathclose{} - 2.24 - 0.85\,\frac{M_Z^2}{M_t^2}\right] + \delta_\text{Higgs}\,,
  \label{eq:Delta_rho_hat_SM_2L}
\end{align}
%
where we have generalized the Higgs dependent part as
%
\begin{align}
   \delta_\text{Higgs} &= 3 \left(\frac{G_F \,M_t \,v_\SM}{8 \pi^2 \sqrt{2}}\right)^2
   \,\sum_{i}\left(|a_{\varphi_i tt}|^2 - |b_{\varphi_i tt}|^2\right)
   \rho^{(2)}\mathopen{}\left(\frac{m_{\varphi_i}}{M_t}\right)\mathclose{}
\end{align}
%
to include corrections from all neutral Higgs fields $\varphi_i$
coupling to the top quark via the vertex
$(\mathbbm{1}\,a_{\varphi_i tt} + \gamma_5\,b_{\varphi_i tt})$.
For this generalization to work properly, the SM-like vacuum
expectation value $v_\SM$ has to be defined and normalized
to the value $\approx 246$ GeV in the corresponding \sarah model file.

The model-specific correction in Eq.~\eqref{eq:Delta_r_hat}
consists of an SM and a BSM part,
%
\begin{align}
   \deltaVB &= \deltaVB^\SM + \deltaVB^\BSM\,,
\end{align}
%
where the SM contribution $\deltaVB^\SM$
\cite{Degrassi:1990tu}, originating
from diagrams with additional internal gauge bosons, is given by
Eq.~(C.12) from Ref.~\cite{Pierce:1996zz} considering the replacement
$\hat\rho \rightarrow 1/(1-\Delta\hat\rho)$.
The BSM contribution $\deltaVB^\BSM$ contains corrections from
1-loop external wave-function renormalizations, vertex and
box diagrams, which are put together as in Eq.~(C.13) from the
same reference. For the different corrections, \fs considers the
diagram types shown in \figref{fig:delta_VB_diagram_types} and sums
over all possible instantiations of these by inserting the valid
combinations of particles into the loop.\footnote{ Note that
self energy, vertex and box diagrams with internal vector bosons
are not considered outside of $\deltaVB^\SM$.}
%
\begin{figure}[tbh]
  \centering
  \subfloat[]{
  \quad\includegraphics{doc/Feynman/deltaVBWave}\quad
  \label{fig:delta_VB_wave_diagrams}
  }
  \subfloat[]{
  \qquad\includegraphics{doc/Feynman/deltaVBVertex1}\quad
  \includegraphics{doc/Feynman/deltaVBVertex2}\qquad
  \label{fig:delta_VB_vertex_diagrams}
  }
  \subfloat[]{
  \quad\includegraphics{doc/Feynman/deltaVBBox1}\quad
  \label{fig:delta_VB_box_diagrams}
  }
  \caption{Generic diagram types implemented in \fs to calculate
  the BSM 1-loop contributions to the muon decay, $\deltaVB^\BSM$.
  The solid and dashed lines in the loops represent any valid
  combination of fermions and scalars.
  \protect\subref{fig:delta_VB_wave_diagrams} wave-function
  renormalization diagrams: $f$ stands for $e$, $\mu$, $\nu_e$
  or~$\nu_\mu$.
  \protect\subref{fig:delta_VB_vertex_diagrams} vertex diagrams:
  $e_i$ stands for $e$ or~$\mu$.
  \protect\subref{fig:delta_VB_box_diagrams} box diagrams:
  the triplet $(f_1, f_2, f_3)$ stands for any permutation of
  $e$, $\nu_e$ and $\nu_\mu$.}
  \label{fig:delta_VB_diagram_types}
\end{figure}
%

After the weak mixing angle~$\theta_W$ has been calculated via
Eq.~\eqref{eq:muon_decay_master}, the $W$ pole mass can be
computed as
%
\begin{align}
   M_W = \sqrt{M_Z^2\,\cos^2\theta_W\,\frac{\hat\rho_\text{tree}}{1-\Delta\hat\rho}}\,.
   \label{eq:MW_calculation}
\end{align}
%

\subsection{Choosing \fmw in the model file}
%
The method to determine the weak mixing angle can be selected by
setting the following variable in the \fs model file:
%
\begin{lstlisting}
(* possible values: Automatic, FSFermiConstant or FSMassW *)
FSWeakMixingAngleInput = Automatic;
\end{lstlisting}
%
By default, \code{FSWeakMixingAngleInput} is set to \code{Automatic},
in which case the weak mixing angle is determined from the muon decay,
if all conditions are fulfilled, otherwise the $W$ mass is used.
Further possible values are \code{FSFermiConstant} and
\code{FSMassW} to explicitly select the muon decay or the $W$ mass method,
respectively.

\subsection{Structure of \fmw code}
%
The C++ interface for the determination of the running weak mixing
angle~$\theta_W$ and the $W$ pole mass via muon decay is provided by the class
\code{<model>_weinberg_angle} defined in
\code{<model>_weinberg_angle.hpp}.  To construct an object from this
class, two arguments have to be provided: a model object, which
represents the set of running BSM parameters, and a struct of type
\code{Sm_parameters}. The latter is defined within the
\code{<model>_weinberg_angle} class and contains the required SM parameters,
namely the Fermi constant $G_F$ as well as the pole
masses $M_W$, $M_Z$, $M_t$ and $\as^{\SM(5)}(M_t)$.
The values of $\sin\theta_W$ and $M_W$ are calculated and returned by
the class member function \code{calculate()}, which includes all of the
Eqs.~\eqref{eq:muon_decay_master}--\eqref{eq:MW_calculation} and the
necessary iteration.\footnote{The input $W$ pole mass is used as an
initial value while the function \code{calculate()} returns a more fitting
one. By updating the utilized value of $M_W$ during the iteration of the
spectrum generator, a self-consistent solution is ensured.}
%
\begin{example}
  In the \modelname{MSSM}, the running weak mixing angle~$\theta_W$ and the
  $W$ pole mass for a given \code{model} object can be calculated with
  the following code:
%
\begin{lstlisting}
MSSM_weinberg_angle::Sm_parameters sm_pars;
sm_pars.fermi_constant = 1.1663787e-05;
sm_pars.mw_pole = 80.385;
sm_pars.mz_pole = 91.1876;
sm_pars.mt_pole = 173.34;
sm_pars.alpha_s = 0.1079;

MSSM_weinberg_angle weinberg(model, sm_pars);

const auto sw_mw     = weinberg.calculate();
const double theta_w = std::asin(sw_mw.first);
const double MW      = sw_mw.second;
\end{lstlisting}
\end{example}
%

\subsection{Tests and comparisons with other spectrum generators}
%
The implementation of the muon decay method for the determination
of $\theta_W$ and $M_W$ provided as \fmw has been tested in the SM
and CMSSM by comparing to the results obtained using the algorithm
from \softsusy, which has been added in \fs 1.1.0. We have found
excellent agreement between the two implementations and also added
unit tests performing these comparisons to the \fs test suite.

Furthermore, the automatically generated expression for the
generalized $\hat\rho_\text{tree}$ given in Eq.~\eqref{eq:rho_hat_tree}
has been analytically checked for many models, such as the UMSSM,
MRSSM and \ESSM.

Finally, we compared numerical results for the $W$ pole mass in the
CMSSM and MRSSM obtained with \fmw to the ones from \spheno code
generated by \sarah~4.12.2. \figref{fig:FlexibleMW_vs_SPheno} shows $M_W$
as a function of $\mhalf$ (CMSSM, see left panel) or the superpotential
parameter $\Lambda_u$ (MRSSM, see right panel) while all
the other parameters are specified as described in the caption.
In the case of the CMSSM, additionally the results calculated with
\spheno~4.0.3 are plotted. For both models, there is a large
discrepancy between the \fmw values illustrated by the blue solid
line and the results from \sarah/\spheno presented by the red dashed
line. A thorough comparison of the two implementations has revealed
two major differences. On the one hand, \sarah/\spheno partly uses the
\DRbar top mass in the self energies occurring in
Eqs.~\eqref{eq:Delta_r_hat} and \eqref{eq:Delta_rho_hat} as well as
the SM 2-loop corrections Eqs.~\eqref{eq:Delta_r_hat_SM_2L} and
\eqref{eq:Delta_rho_hat_SM_2L} while \fmw always uses the top pole mass,
as suggested by Ref.~\cite{Fanchiotti:1992tu}. Not utilizing the top
pole mass in all of these formulas spoils the correctness of the
included SM 2-loop corrections. On the other hand, the
\sarah/\spheno code contains an inconsistency in the final computation of
$M_W$, which is similar to Eq.~\eqref{eq:MW_calculation} but partly
neglects the SM 2-loop correction to $\Delta\hat\rho$.
This inconsistency is also existent in the original \spheno code that,
for the CMSSM, produces the results depicted by the red dashed-double-dotted
line in the left panel of \figref{fig:FlexibleMW_vs_SPheno}.
After fixing these issues within the \sarah/\spheno and \spheno code,
we get the modified results illustrated by the green dotted and
dashed-dotted line, respectively. These show good agreement with the
values from \fmw for both the CMSSM and MRSSM.
The remaining small discrepancies are well
understood and mainly caused by minor differences in the implemented
formulas and the various utilized \DRbar parameters. In addition,
the \sarah/\spheno and \spheno codes use $\as^{\SM(5)}(M_Z)$ while \fmw uses
$\as^{\SM(5)}(M_t)$ as preferred by Ref.~\cite{Fanchiotti:1992tu}.
%
\begin{figure}[tbh]
  \centering
  \includegraphics[width=0.49\textwidth]{doc/plots/FlexibleMW/scan_MW_M12}\hfill
  \includegraphics[width=0.49\textwidth]{doc/plots/FlexibleMW/scan_MW_LamTU}
  \caption{Left panel: $W$ pole mass in the CMSSM as a function of
  $\mhalf$ with $\mzero = 1\unit{TeV}$, $\tan\beta = 10$,
  $\sign\mu = +1$ and $\azero = 0$. Right panel: $W$ pole mass in
  the MRSSM as a function of $\Lambda_u$ with the other parameters
  fixed as for BMP1 in Table~2 from Ref.~\cite{Diessner:2014ksa}.
  For both panels, the differences between the lines are explained
  in the text.}
  \label{fig:FlexibleMW_vs_SPheno}
\end{figure}
%

\section{\fsas extension}
\label{sec:fsas}
In this section, we introduce a new BVP
solver, a major new feature released in \fstwo. The specification of the
running parameters of a BSM model at multiple scales connected by RGEs
constitutes a BVP that must be solved in order to compute the mass spectrum.
\fs was designed with the intention of allowing multiple solvers for this
problem. In \fsone only one supported solver was
distributed, the two-scale solver, which uses a fixed point iteration
with boundary conditions set at both the high scale and the low scale
and is the one used in MSSM spectrum generators. However, additional
solvers are important because convergence of valid points is not
guaranteed with the two-scale solver and viable regions of the
parameter space can be missed if the only option is the two-scale
solver.  This is already true in the CMSSM where it has also been
shown that there can even be multiple solutions to the BVP
\cite{Allanach:2013cda,Allanach:2013yua}, of which the two-scale solver will
only find at most one.  Furthermore, the two-scale
solver cannot find solutions at all for a large class of models, which
includes the fully constrained NMSSM (CNMSSM) \cite{Ellwanger:1993xa,
  Ellwanger:1995ru,Ellwanger:1996gw} and constrained \ESSM
(C\ESSM) \cite{Athron:2009bs}.

A basic problem in these models is that parameters exist that would
naturally be computed at the weak scale by EWSB conditions but
are at the same time constrained by GUT scale conditions.\footnote{In
  principle this could be avoided, e.g., in the CNMSSM and C\ESSM by including
  $\tan \beta$ or $\lambda$ in the set of parameters fixed by minimization
  conditions.  However, $\tan\beta$ affects the top, bottom and $\tau$ Yukawa
  couplings at tree level, and $\lambda$ appears in the 1-loop RGEs for these
  couplings and therefore can have a significant impact on the RG evolution.
  This makes it difficult to obtain a convergent fixed point iteration with
  such a setup.} In the constrained NMSSM, where the soft scalar mass,
$m_S$, is typically an EWSB output, solutions have previously been
found by varying one of the input parameters, $\tan \beta$, and tuning
this until a solution is found where $m_S$ is sufficiently close to
the universal scalar mass, $\mzero$ \cite{Djouadi:2008yj,Djouadi:2008uj}.
For the constrained \ESSM again the soft scalar masses need to be EWSB
outputs.  In that case a new BVP solver using semi-analytic solutions
was invented to find solutions \cite{Athron:2009ue,Athron:2009bs} and
extended and improved in further studies
\cite{Athron:2011wu,Athron:2012sq,Athron:2012pw}, though the precision
was limited due to analytical approximations used in the code.  In
general, the challenges in finding solutions to the mass spectrum in
BSM models can be quite different to those found in the MSSM and
therefore additional solvers can be of great benefit in increasing the
scope of \fs.

In \fstwo, an additional BVP solver has been added that solves the BVP
by using semi-analytic solutions to the RGEs for a subset of the
running parameters in the model.  The semi-analytic algorithm takes
advantage of the hierarchical structure of the RGEs in any model: the parameters
can be split into a sequence of sets such that the running of the parameters
in each set is independent of all of the parameters in the following
sets.  For example, in SUSY models the SUSY preserving parameters run
independently of the soft breaking parameters; in general renormalizable
models the dimensionless parameters run independently of the mass
parameters.\footnote{This partitioning is already reflected to some extent in
  the original C++ class structure for the parameters and
RGEs, which for SUSY (non-SUSY) models is split up so that
the RGEs for the SUSY preserving (dimensionless) parameters can be integrated
separately to those for the soft (dimensionful) parameters; see
\secref{sec:fbsm}.}
It is clear from the general form of the RGEs
\cite{Machacek:1983tz,Machacek:1983fi,Machacek:1984zw,Martin:1993zk,
  Yamada:1994id,Jack:1997eh,Luo:2002ti,Fonseca:2011vn,Goodsell:2012fm,
  Fonseca:2013bua,Sperling:2013eva,Sperling:2013xqa} that further
divisions are possible on the basis of the mass dimension of the
parameters.  The simplest case is the evolution of the mass dimension
one parameters $m^i(Q)$, which is described by a system of linear homogeneous
differential equations that can be solved in the form
\begin{align}
m^i(Q) &= \scoeff{}{}^i_j m^j(Q_0) ,
\end{align}
i.e., in terms of a linear combination of initial values and a set of
dimensionless coefficients that only depend on dimensionless parameters.
The RGEs for higher mass dimension parameters can then be cast as linear,
non-homogeneous systems by substituting in the solutions for the lower
mass dimension parameters, for which the general solutions can easily
be written down.\footnote{See, for example, the derivation given in
  Ref.~\cite{Athron:2009bs} for the case of a SUSY model with real
  soft parameters.}  In the CMSSM, doing so leads to well-known expressions
for the soft masses in terms of the universal parameters $\mzero$, $\mhalf$,
and $\azero$.  For instance, the solutions for the soft scalar masses read
(see also Eq.~\eqref{eq:cmssm-semi-analytic-solns} below)
\begin{equation*}
  m_i^2(Q) = a_i(Q) \mzero^2 + b_i(Q) \mhalf^2
  + c_i(Q) \mhalf \azero + d_i(Q) \azero^2 \, ,
\end{equation*}
where the dimensionless coefficients $a_i$, $b_i$, $c_i$, and $d_i$ are
determined by the running of the gauge and Yukawa couplings, and are computed
numerically.  The basic idea of the semi-analytic solver is to generalize this
approach to other models: first, the analytic forms of the solutions are
determined from the boundary conditions, and then the appearing coefficients
are determined numerically.  Thus the solver obtains a collection of
semi-analytic solutions which express parameters at the scale $Q$ directly in
terms of parameters at the boundary scale $Q_0$.

To make the discussion concrete, consider a general SUSY model with bilinear
and linear superpotential parameters $\mu^{ij}$ and $L^i$, together with a set
of soft breaking gaugino masses $M_a$, scalar trilinear, bilinear and linear
couplings $T^{ijk}$, $b^{ij}$ and $t^i$, and a set of soft scalar masses
$(m^2)_i^j$.  The semi-analytic solutions are obtained from the general 2-loop
RGEs given in Refs.~\cite{Martin:1993zk,Yamada:1994id,Fonseca:2011vn}.
Since the superpotential parameters are only multiplicatively
renormalized, the semi-analytic solutions for the parameters $\mu^{ij}$
and $L^i$ are particularly simple, taking the form
\begin{align}
  \mu^{ij}(Q) &= \scoeff{\mu}{\mu}^{ij}_{kp} \mu^{kp}(Q_0) \, ,
  \label{eq:general-susy-bilinear-solution} \\
  L^i(Q) &= \scoeff{L}{L}^i_p L^p(Q_0) \, ,
  \label{eq:general-susy-linear-solution}
\end{align}
where the coefficients $\scoeff{\mu}{\mu}$ and $\scoeff{L}{L}$ satisfy the
initial conditions $[c^\mu_\mu(Q_0)]^{ij}_{kp} = \delta^i_k
\delta^j_p$, $[c^L_L(Q_0)]^i_p = \delta^i_p$. Again, the point is that
the coefficients $c(Q)$ can be determined once the
running dimensionless SUSY parameters are known, and the equations then express
parameters at $Q$ in  terms of parameters at $Q_0$ with numerically known
coefficients.

The semi-analytic solutions for the dimension one soft breaking parameters can
be written as
\begin{align}
  T^{ijk}(Q) &= \scoeff{T}{T}^{ijk}_{lmn} T^{lmn}(Q_0)
  + \scoeff{T}{M}^{ijkb} M_b(Q_0) \, ,
  \label{eq:general-soft-trilinear-solution} \\
  M_a(Q) &= \scoeff{M}{T}_{almn} T^{lmn}(Q_0)
  + \scoeff{M}{M}_a^b M_b(Q_0) \, .
  \label{eq:general-gaugino-mass-solution}
\end{align}
Upon substituting these solutions into the RGEs for the soft breaking bilinears
$b^{ij}$ and the soft scalar masses $(m^2)_i^j$, one finds the semi-analytic
solutions
\begin{equation} \label{eq:general-soft-bilinear-solution}
  b^{ij}(Q) = \scoeff{b}{b}^{ij}_{kl} b^{kl}(Q_0)
  + \scoeff{b}{\mu T}^{ij}_{abmno} \mu^{ab}(Q_0) T^{mno}(Q_0)
  + \scoeff{b}{\mu M}^{ijc}_{ab} \mu^{ab}(Q_0) M_c(Q_0)
\end{equation}
and
\begin{align}
  (m^2)_i^j(Q) &= \scoeff{m^2}{m^2}^{jl}_{ik} (m^2)^k_l(Q_0)
  + \scoeff{m^2}{MM^*}_i^{jab} M_a(Q_0) M_b^*(Q_0) \nonumber \\
  & \quad {} + \scoeff{m^2}{T M^*}_{iklm}^{ja} T^{klm}(Q_0) M_a^*(Q_0)
  + \scoeff{m^2}{T^* M}_{iklm}^{ja} T^{klm*}(Q_0)
  M_a(Q_0) \nonumber \\
  & \quad {} + \scoeff{m^2}{T^* T}_{iklmnop}^j T^{klm*}(Q_0) T^{nop}(Q_0) \, .
  \label{eq:general-soft-scalar-mass-solution}
\end{align}
Finally, the semi-analytic solutions for the soft breaking linear parameters
$t^i$ are given by
\begin{align}
  t^i(Q) &= \scoeff{t}{t}^i_j t^j(Q_0) + \scoeff{t}{L T}^i_{jklm} L^j(Q_0)
  T^{klm}(Q_0) + \scoeff{t}{LM}^{ia}_j L^j(Q_0) M_a(Q_0) \nonumber \\
  & \quad {} + \scoeff{t}{\mu \mu T}^i_{jklmnop} \mu^{jk}(Q_0) \mu^{lm}(Q_0)
  T^{nop}(Q_0) + \scoeff{t}{\mu \mu M}^{ia}_{jklm} \mu^{jk}(Q_0) \mu^{lm}(Q_0)
  M_a(Q_0)
  \nonumber \\
  & \quad {} + \scoeff{t}{\mu b}^i_{jklm} \mu^{jk}(Q_0) b^{lm}(Q_0)
  + \scoeff{t}{\mu^* M M^*}^{iab}_{jk} \mu^{jk*}(Q_0) M_a(Q_0) M_b^*(Q_0)
  \nonumber \\
  & \quad {} + \scoeff{t}{\mu^* T T^*}^i_{jklmnopq} \mu^{jk*}(Q_0) T^{lmn}(Q_0)
  T^{opq*}(Q_0) \nonumber \\
  & \quad {} + \scoeff{t}{\mu^* T M^*}^{ia}_{jklmn} \mu^{jk*}(Q_0) T^{lmn}(Q_0)
  M_a^*(Q_0) \nonumber \\
  & \quad {} + \scoeff{t}{\mu^* T^* M}^{ia}_{jklmn} \mu^{jk*}(Q_0) T^{lmn*}(Q_0)
  M_a(Q_0) \nonumber \\
  & \quad {} + \scoeff{t}{b^* T}^i_{jklmn} b^{jk*}(Q_0) T^{lmn}(Q_0)
  + \scoeff{t}{b^* M}^{ia}_{jk} b^{jk*}(Q_0) M_a(Q_0) \nonumber \\
  & \quad {} + \scoeff{t}{\mu^* m^2}^{il}_{jkm} \mu^{jk*}(Q_0) (m^2)^m_l(Q_0)
  \, . \label{eq:general-soft-linear-solution}
\end{align}
For models in which Dirac gaugino masses $m_{Da}^i$ are also present, the
solutions for the parameters $t^i$ are modified, $t^i \to t^i + \Delta t^i$,
where $\Delta t^i$ is of the form
\begin{align}
  \Delta t^i(Q) &= \scoeff{t}{m_D m^2}^{iaj}_{kl} m_{Da}^k(Q_0) (m^2)_j^l(Q_0)
  + \scoeff{t}{m_D M M^*}^{iabc}_j m_{Da}^j(Q_0) M_b(Q_0) M_c^*(Q_0)
  \nonumber \\
  & \quad {} + \scoeff{t}{m_D T M^*}^{iab}_{jklm} m_{Da}^j(Q_0) T^{klm}(Q_0)
  M_b^*(Q_0) \nonumber \\
  & \quad {} + \scoeff{t}{m_D T^* M}^{iab}_{jklm} m_{Da}^j(Q_0) T^{klm*}(Q_0)
  M_b(Q_0) \nonumber \\
  & \quad {} + \scoeff{t}{m_D T^* T}^{ia}_{jklmnop} m_{Da}^j(Q_0) T^{klm*}(Q_0)
  T^{nop}(Q_0) \nonumber \\
  & \quad {} + \scoeff{t}{m_D^* m_D^* T}^{iab}_{jklmn} m_{Da}^{j*}(Q_0)
  m_{Db}^{k*}(Q_0) T^{lmn}(Q_0) \nonumber \\
  & \quad {} + \scoeff{t}{m_D^* m_D^* M}^{iabc}_{jk} m_{Da}^{j*}(Q_0)
  m_{Db}^{k*}(Q_0) M_c(Q_0) \nonumber \\
  & \quad {} + \scoeff{t}{m_D m_D \mu}^{iab}_{jklm} m_{Da}^j(Q_0) m_{Db}^k(Q_0)
  \mu^{lm}(Q_0) \, . \label{eq:extra-soft-trilinear-solution}
\end{align}
The semi-analytic solutions for the Dirac gaugino masses themselves follow
from the known general 2-loop RGEs \cite{Goodsell:2012fm}, and can be
written in the form
\begin{equation} \label{eq:general-dirac-gaugino-solutions}
  m_{Da}^i(Q) = \scoeff{m_{Da}}{m_{Da}}^i_j m^j_{Da}(Q_0) \, .
\end{equation}

In non-SUSY models, the semi-analytic solutions follow from the known
results for the 2-loop RGEs in a general gauge theory \cite{Machacek:1983tz,
  Machacek:1983fi,Machacek:1984zw,Luo:2002ti,Fonseca:2013bua}.  For a non-SUSY
model containing a set of real scalar trilinear couplings $h^{ijk}$ and squared
scalar masses $(m^2)^{ij}$, and a set of fermion masses $(M_f)^{ij}$, the
semi-analytic solutions for the mass dimension one parameters read
\begin{align}
  h^{ijk}(Q) &= \scoeff{h}{h}^{ijk}_{lmn} h^{lmn}(Q_0)
  + \scoeff{h}{M_f}^{ijk}_{lm} (M_f)^{lm}(Q_0)
  + \scoeff{h}{M_f^*}^{ijk}_{lm} (M_f)^{lm*}(Q_0) \, ,
  \label{eq:general-dimension-one-trilinear-solution} \\
  (M_f)^{ij}(Q) &= \scoeff{M_f}{h}^{ij}_{lmn} h^{lmn}(Q_0)
  + \scoeff{M_f}{M_f}^{ij}_{lm} (M_f)^{lm}(Q_0)
  + \scoeff{M_f}{M_f^*}^{ij}_{lm} (M_f)^{lm*}(Q_0) \, .
  \label{eq:general-dimension-one-fermion-mass-solution}
\end{align}
In general, all of the dimension one parameters must be considered together,
unlike in SUSY models where they can be separated into superpotential and soft
breaking masses.  After substituting the solutions
Eq.~\eqref{eq:general-dimension-one-trilinear-solution} and
Eq.~\eqref{eq:general-dimension-one-fermion-mass-solution} into the RGEs for
the squared scalar masses, the semi-analytic solutions for the scalar masses
are found to be
\begin{align}
  (m^2)^{ij}(Q) &= \scoeff{m^2}{m^2}^{ij}_{kl} (m^2)^{kl}(Q_0)
  + \scoeff{m^2}{h h}^{ij}_{klmnop} h^{klm}(Q_0) h^{nop}(Q_0) \nonumber \\
  & \quad {} + \scoeff{m^2}{h M_f}^{ij}_{klmpq} h^{klm}(Q_0) (M_f)^{pq}(Q_0)
  + \scoeff{m^2}{h M_f^*}^{ij}_{klmpq} h^{klm}(Q_0) (M_f)^{pq*}(Q_0)
  \nonumber \\
  & \quad {} + \scoeff{m^2}{M_f M_f}^{ij}_{klmn} (M_f)^{kl}(Q_0) (M_f)^{mn}(Q_0)
  \nonumber \\
  & \quad {} + \scoeff{m^2}{M_f M_f^*}^{ij}_{klmn} (M_f)^{kl}(Q_0)
  (M_f)^{mn*}(Q_0) \nonumber \\
  & \quad {} + \scoeff{m^2}{M_f^* M_f^*}^{ij}_{klmn} (M_f)^{kl*}(Q_0)
  (M_f)^{mn*}(Q_0) \, .
  \label{eq:general-scalar-mass-solution}
\end{align}

The solver algorithm implemented by \fsas automatically determines
the above semi-analytic solutions in the model,\footnote{In non-SUSY models,
  \sarah currently does not calculate RGEs for linear scalar couplings
  $L^i$, which have been given in, e.g., Ref.~\cite{Goodsell:2012fm}, and
  therefore the semi-analytic solutions for these parameters are also not used
  in \fs.} given the set of boundary conditions at some scale.  Since the
required coefficients may be determined knowing only the running of the
dimensionless or SUSY preserving parameters, each step of the main fixed point
iteration is split up into two parts.  Firstly, the BVP for the dimensionless
parameters is solved iteratively.  The semi-analytic coefficients at any scale
can then be calculated.  In this second stage, the soft breaking or
dimensionful parameters are expanded in terms of the semi-analytic solutions to
the RGEs.  The low-energy EWSB conditions and masses are thus expressed
explicitly in terms of the boundary values at $Q_0$, allowing, for example,
unknown quantities at one scale to be directly constrained at another.

\subsection{Choosing \fsas in the model file}
\label{sec:chosing_SAS}

The BVP solver algorithms that are applicable in a given model can be
specified in the \fs model file using the list
\code{FSBVPSolvers}.  The elements of this list correspond to
the desired BVP solvers that should be enabled for the model, identified
using the predefined symbols \code{TwoScaleSolver} for the two-scale
algorithm and \code{SemiAnalyticSolver} for the semi-analytic solver.  By
default, if a list of BVP solvers is not specified in the model file, only
the two-scale algorithm is enabled, as summarized in \tabref{tab:fsas_options}.
%
\begin{OptionTable}{\fsas model file options.}{tab:fsas_options}
  \code{FSBVPSolvers} & \code{\{ TwoScaleSolver \}} & Non-empty list
  containing \code{TwoScaleSolver} or \code{SemiAnalyticSolver} or
  both &
  List of BVP solvers to be used \\
  \code{TwoScaleSolver} & -- & -- & Represents the two-scale BVP solver \\
  \code{SemiAnalyticSolver} & -- & -- & Represents the semi-analytic BVP
  solver \\
  \bottomrule
\end{OptionTable}

In addition to specifying that the semi-analytic solver should be used,
the user should ensure that the boundary conditions for the model are
compatible with its use.  Currently, this requires that those parameters
that will be expanded using semi-analytic solutions of the RGEs
are fixed in the same boundary condition.  So in SUSY models, the
boundary values for all of the soft SUSY breaking parameters should be given
at a single scale, while in non-SUSY models the same should be done for all
of the dimensionful parameters.  The expressions for the boundary values
must also only be polynomials in any dimensionful  parameters, such as
universal scalar masses.

Boundary values for the running parameters of the model can be specified
in terms of input parameters such as those defined in the \code{MINPAR}
and \code{EXTPAR} variables.  In \fstwo, the user can also define extra
parameters by using the list \code{FSAuxiliaryParameterInfo}.   Each entry in
this list should contain the name of the extra parameter being defined and a
list of its properties.  The possible properties for the
new parameters are specified in \tabref{tab:parameter_info}.  In particular,
the mass dimensions of the new parameters may be specified to allow for
simplifying the forms of the semi-analytic solutions, in which dimensionless
parameters can be absorbed into the definitions of the semi-analytic
coefficients.  Input or auxiliary parameters that are not scalars may be
defined by setting the \code{ParameterDimensions} property to a list of the
form \code{\{M,N\}} for an $M \times N$ matrix or \code{\{N\}} for an
$N$-dimensional vector.  A value of \code{\{1\}} corresponds to a scalar.  Note
that in versions of \fs prior to version 2.0, this functionality was available
for input parameters using the variable \code{FSExtraInputParameters}.
However, please note that this variable has been removed in \fstwo; definitions
that were previously given in \code{FSExtraInputParameters} must now be given
in the new list \code{FSAuxiliaryParameterInfo}.
%
\begin{OptionTable}{Allowed properties for extra parameters}{tab:parameter_info}
  \code{InputParameter} & \code{False} & \code{True} or \code{False}
  & Indicates whether the new parameter is an input parameter \\
  \code{ParameterDimensions} & \code{\{1\}} & A list of the form \code{\{M,N\}}
  or \code{\{N\}} & Specifies the dimensions of the parameter \\
  \code{MassDimension} & -- & A non-negative integer & Specifies the mass
  dimension of the parameter \\
  \code{LesHouches} & -- & A symbol or string, or a list of the form
  \code{\{block, entry\}} & Specifies the SLHA block from which the parameter
  should be read if it is an input parameter \\
  \bottomrule
\end{OptionTable}

\begin{example}
In the MSSM, input parameters giving the values of the
soft SUSY breaking trilinears as $3 \times 3$ matrices can be defined:
%
\begin{lstlisting}
FSAuxiliaryParameterInfo = {
   {Ae  , { LesHouches -> AEIN,
            ParameterDimensions -> {3,3},
            InputParameter -> True
          } },
   {Ad  , { LesHouches -> ADIN,
            ParameterDimensions -> {3,3},
            InputParameter -> True
          } },
   {Au  , { LesHouches -> AUIN,
            ParameterDimensions -> {3,3},
            InputParameter -> True
          } }
};
\end{lstlisting}
\end{example}

Extra parameters defined in this way can be used in the specification of the
boundary conditions for running model parameters, or indeed themselves be
fixed in the boundary conditions.  A special case is if the new parameters
are to be fixed by the EWSB conditions.  To facilitate this usage, the new
variables \code{EWSBInitialGuess} and \code{EWSBSubstitutions} may be
given in the model file.  The former allows for explicit initial
guesses to be provided for the EWSB output parameters, in the same format
as used for specifying the boundary conditions.  The latter, a list of
two-component lists, can be used to define any substitutions that should
be made in the EWSB equations before attempting to solve them.

\begin{example}
In the so-called VCMSSM \cite{Ellis:2003pz,Ellis:2004qe}, the value of the soft
breaking bilinear $B\mu$ is fixed at the GUT scale, $M_X$, in terms of the
universal soft trilinear $\azero$ and scalar mass $\mzero$ according to
$B\mu(M_X) = \mu(M_X) (\mzero + \azero)$.  Consequently, $B\mu$ can no longer
be fixed to ensure proper EWSB, as is usually done in the CMSSM\@.  Instead,
$|\mu|^2$ and $\tan\beta$ are used as EWSB outputs.  This can be achieved in a
\modelname{VCMSSM} model file through the definitions:
%
\begin{lstlisting}
FSAuxiliaryParameterInfo = {
   {TanBeta, { ParameterDimensions -> {1},
               MassDimension -> 0 } },
   {MuSq,    { ParameterDimensions -> {1},
               MassDimension -> 2 } },
   {vMSSM,   { ParameterDimensions -> {1},
               MassDimension -> 1 } }
};

EWSBOutputParameters = { TanBeta, MuSq };

EWSBSubstitutions = {
   {vd, vMSSM Cos[ArcTan[TanBeta]]},
   {vu, vMSSM Sin[ArcTan[TanBeta]]},
   {\[Mu], Sign[\[Mu]] Sqrt[MuSq]}
};

EWSBInitialGuess = {
   {TanBeta, vu / vd},
   {MuSq, \[Mu]^2}
};

SUSYScaleInput = {
   {vMSSM, Sqrt[vd^2 + vu^2]},
   FSSolveEWSBFor[EWSBOutputParameters]
};
\end{lstlisting}
\end{example}

Note that \fs automatically substitutes the semi-analytic solutions into
the EWSB conditions for the soft or dimensionful parameters.  Therefore, it
is not necessary for the user to explicitly provide them.  For the purpose of
making use of the semi-analytic algorithm, the only required addition to the
model file is the inclusion of \code{SemiAnalyticSolver} in the list
\code{FSBVPSolvers}.

\begin{example}
The CNMSSM is characterized by universal soft scalar masses at the
high scale.  Typically, this constraint is relaxed somewhat to allow the
soft singlet mass $m_S^2$ to differ from the common scalar mass $\mzero^2$ at
the high scale.  This is done to allow fixing $m_S^2$ using the EWSB conditions
to ensure correct EWSB, and is the approach taken in the \modelname{NMSSM} model
included with \fs, as well as in other public spectrum generators such as
\softsusy, \spheno and \nmspec.  Universality of the soft scalar masses can
alternatively be maintained by using the semi-analytic BVP solver, which is
achieved in the \modelname{CNMSSM} model file by specifying the BVP solvers to
be used,
%
\begin{lstlisting}
FSBVPSolvers = { SemiAnalyticSolver };
\end{lstlisting}
In this set-up, $\mzero$ ceases to be an input parameter and is fixed by the
EWSB conditions.  This is achieved in the model file by removing $\mzero$ from
the list of input parameters and defining it as an additional parameter, as
follows:
%
\begin{lstlisting}
(* CNMSSM input parameters *)

MINPAR = {
   {2, m12},
   {3, TanBeta},
   {4, Sign[vS]},
   {5, Azero}
};

EXTPAR = {
   {61, LambdaInput}
};

FSAuxiliaryParameterInfo = {
   {m0Sq,        { ParameterDimensions -> {1},
                   MassDimension -> 2 } },
   {LambdaInput, { ParameterDimensions -> {1},
                   MassDimension -> 0 } }
};
\end{lstlisting}
Note that the definition of \code{LambdaInput} in
\code{FSAuxiliaryParameterInfo} is not compulsory, but allows the semi-analytic
solutions to be simplified by using the fact that it is a dimensionless
parameter.  The value of $\mzero^2$ is then determined from the EWSB conditions
by setting
%
\begin{lstlisting}
EWSBOutputParameters = { \[Kappa], vS, m0Sq };
\end{lstlisting}
To impose the universality constraint, the condition $m_S^2 = \mzero^2$ at the
GUT scale must be added to the high-scale boundary condition, by defining
%
\begin{lstlisting}
HighScaleInput={
   {T[Ye], Azero*Ye},
   {T[Yd], Azero*Yd},
   {T[Yu], Azero*Yu},
   {mq2, UNITMATRIX[3] m0Sq},
   {ml2, UNITMATRIX[3] m0Sq},
   {md2, UNITMATRIX[3] m0Sq},
   {mu2, UNITMATRIX[3] m0Sq},
   {me2, UNITMATRIX[3] m0Sq},
   {mHu2, m0Sq},
   {mHd2, m0Sq},
   {ms2, m0Sq},
   {\[Lambda], LambdaInput},
   {T[\[Kappa]], Azero \[Kappa]},
   {T[\[Lambda]], Azero LambdaInput},
   {MassB, m12},
   {MassWB,m12},
   {MassG,m12}
};
\end{lstlisting}
The remainder of the model implementation is otherwise rather similar to that
for the NMSSM solved using the two-scale solver.  The full \modelname{CNMSSM}
model file is given in \appref{app:CNMSSM}.
\end{example}

\subsection{Structure of the generated code}

In keeping with the design goal of \fs to produce generated code that is highly
modular in nature, the implementation of the BVP solvers is separated from
the details of specific physics models.  In \fs, a general BVP solver algorithm
is represented by the templated \code{RGFlow<T>} class.  Particular algorithms
are provided as specializations of this class, with the two-scale and
semi-analytic solvers corresponding to the classes \code{RGFlow<Two_scale>}
and \code{RGFlow<Semi_analytic>}, respectively.  Each realizes an abstract
implementation of the appropriate algorithm, with no dependence on the details
of any particular model.  The required model-dependent information is provided
by separate classes representing the model and boundary and matching conditions,
which are linked to the desired BVP solver class.  New algorithms can easily be
added simply by writing additional specializations of the \code{RGFlow} class.

The semi-analytic solver algorithm requires two nested iterations.  An
inner iteration, carried out at each step, determines consistent values for
the SUSY preserving (in SUSY models) or dimensionless parameters (in non-SUSY
models) at the low- and high-scale boundaries.  Updated estimates for these
scales are simultaneously calculated during the iteration if necessary, for
example if the high scale $M_X$ is defined in the model file by gauge
unification, $g_1(M_X) = g_2(M_X)$.  Once this has converged, the resulting
estimate for these parameters is used to compute the semi-analytic solutions
for the soft SUSY breaking or dimensionful parameters, at which point the
EWSB conditions may be solved and the \DRbar/\MSbar mass spectrum
calculated.  The new values of the soft or dimensionful parameters are then used
in the inner iteration for computing the required threshold corrections.
This sequence of steps is illustrated in \figref{fig:semi-analytic-algorithm}.
For a single high-scale model such as the CMSSM, the algorithm proceeds as
follows:
%
\subparagraph{Initial guess:} The initial guess involves a first run of the
inner iteration.  In all of the steps below, threshold corrections are ignored.
%
\begin{enumerate}
\item The known values of the SM gauge couplings at the scale $M_Z$ are
  used to estimate the values of $g_1$, $g_2$ and $g_3$ at the scale
  $M_t$, ignoring threshold corrections.
\item The user-defined initial guess at the low scale, as given in
  \code{InitialGuessAtLowScale}, is imposed at the scale $M_t$.
\item The SUSY preserving or dimensionless parameters are run to the initial
  guess for $M_X$, given by \code{HighScaleFirstGuess}, and the
  high-scale boundary condition for these parameters, defined in
  \code{HighScaleInput}, is imposed.  The initial guess at the high scale,
  defined in \code{InitialGuessAtHighScale}, is then applied.
\item \label{initial-guess-low-scale} The model is run to the guess for the
  low scale, initially set to the value defined in \code{LowScaleFirstGuess},
  and the low-scale boundary conditions for the SUSY preserving or
  dimensionless parameters defined in \code{LowScaleInput} are applied.
\item The model is run to the current guess for $M_X$.
  \begin{enumerate}
  \item If necessary, the guess for $M_X$ is updated.  For example, in
    the CMSSM with $M_X$ defined to be the scale at which $g_1(M_X) = g_2(M_X)$,
    a new estimate for $M_X$ is calculated according to
    \begin{equation*}
      M_X^\prime = M_X \exp\left( \frac{g_2(M_X)-g_1(M_X)}
      {\beta_{g_1} - \beta_{g_2}} \right ) \, .
    \end{equation*}
  \item The high-scale boundary conditions for the SUSY preserving or
    dimensionless parameters are applied.
  \end{enumerate}
\item If not converged, goto \ref{initial-guess-low-scale}.
\item The model is run to the guess for the low scale.  The semi-analytic
  solutions are calculated at this scale.
\item The EWSB equations are solved at tree level.
\item The \DRbar/\MSbar mass spectrum is calculated.
\end{enumerate}
%
At this stage, initial guesses for all of the model parameters, boundary
condition scales and the \DRbar/\MSbar mass spectrum are available.  The full
iteration now starts, in which the full set of threshold corrections are
applied.
%
\subparagraph{Thresholds iteration:}
%
\begin{enumerate}
\item \label{outer-iteration-step-one} The SUSY preserving or dimensionless
  parameters are determined in an inner iteration analogous to that in the
  initial guess, namely:
  \begin{enumerate}
  \item \label{inner-iteration-step-one} All model parameters are run to the
    low scale (\code{LowScale}) and the \DRbar/\MSbar mass spectrum is
    calculated.
  \item The low scale is recalculated if it is not fixed.
  \item The SM gauge couplings are calculated in the model, including the
    appropriate threshold corrections.
  \item The user-defined constraints for the SUSY preserving or dimensionless
    parameters are applied.
  \item All model parameters are run to the high scale (\code{HighScale}).
  \item The high scale is recalculated if necessary.
  \item The user-defined boundary conditions at this scale for the
    SUSY preserving or dimensionless parameters are applied.
  \item The model parameters are run to the SUSY scale (\code{SUSYScale})
    and the SUSY scale is updated if necessary.
  \item The boundary conditions for the SUSY preserving or dimensionless
    parameters are applied.
  \item If not converged, goto \ref{inner-iteration-step-one}.
  \end{enumerate}
\item All model parameters are run to the scale at which the EWSB equations
  are to be solved.
  \begin{enumerate}
  \item The coefficients in the semi-analytic solutions are determined
    at this scale, using the current estimate for the scale at which the
    relevant boundary conditions are imposed.  For example, in the CMSSM, the
    semi-analytic solutions for the soft gaugino masses, trilinears, scalar
    masses and bilinear take the form
    \begin{equation}
      \begin{aligned}
        M_i(Q) &= p_i(Q) \azero + q_i(Q) \mhalf \, , \\
        T_i(Q) &= e_i(Q) \azero + f_i(Q) \mhalf \, , \\
        m_i^2(Q) &= a_i(Q) \mzero^2 + b_i(Q) \mhalf^2
        + c_i(Q) \mhalf \azero + d_i(Q) \azero^2 \, , \\
        B\mu(Q) &= u(Q) B\mu(M_X) + v(Q) \mu(M_X) \mhalf
        + w(Q) \mu(M_X) \azero \, .
      \end{aligned} \label{eq:cmssm-semi-analytic-solns}
    \end{equation}
    The coefficients are determined numerically by varying the values of
    $\mhalf$, $\azero$, $\mzero$ and $B\mu(M_X)$ and integrating the RGEs from
    $M_X$ to $Q$.  For example, the coefficients $p_i(Q)$, $e_i(Q)$, $d_i(Q)$
    and $w(Q)$ are obtained by keeping only $\azero \neq 0$.  A similar approach
    is followed to successively obtain all of the remaining coefficients.
  \item The calculated semi-analytic solutions are used to set the values of
    the soft SUSY breaking or dimensionful parameters at this scale.
  \item The \DRbar/\MSbar mass spectrum is calculated and the scale at which
    EWSB occurs is updated.
  \item The EWSB conditions are solved at the loop level.
  \end{enumerate}
\item If not converged, goto \ref{outer-iteration-step-one}.  Otherwise
  the iteration finishes.
\end{enumerate}
%
If the iteration converges, all running parameters in the model are determined
between the low  and high scales.  The remainder of the calculation, that is,
the calculation of the pole mass spectrum and observables, then proceeds in
the same way as in the two-scale algorithm.  Alternatively, the iteration
may fail to converge or may encounter problems that render the parameter point
physically invalid.  As for the two-scale solver, the specific problems that
are encountered for a given parameter space point are stored in the
\code{Problems} class, and may be accessed using the \code{get_problems()}
function of the model class.\footnote{Note that in \fstwo a separate class,
\code{BVP_solver_problems}, is used to store those problems that are associated
only with failures of the BVP solver algorithm, such as a failure to converge,
and which do not necessarily mean the parameter point is ruled out.  A summary
combining all of the problems that arise during a run of the spectrum generator
can then be obtained by calling the \code{get_problems()} method of the spectrum
generator class.}

\begin{figure}[tbh]
  \centering
  \begin{tikzpicture}[node distance = 1cm, auto]
    \tikzstyle{block} = [rectangle, draw, text width=15em,
      text centered, minimum height=3em, minimum width=10em]
    \tikzstyle{arrow} = [draw, -latex, thick]
    \node[block] (MZ)
         {Calculate $g_i(M_Z)$, $y_f(M_Z)$, apply low-scale boundary
           conditions for SUSY parameters};
    \node[block, below = 2em of MZ] (susy) {Iterate SUSY parameters};
    \path[arrow] (MZ) --node {} (susy);
    \node[block, below = 2em of susy] (coeffs)
         {Determine semi-analytic solutions at scale of EWSB};
    \path[arrow] (susy) --node {} (coeffs);
    \node[block, below = 2em of coeffs] (MS) {Solve EWSB, apply boundary
      conditions at this scale};
    \path[arrow] (coeffs) --node {} (MS);
    \node[block, below = 2em of MS] (thresh) {Apply threshold corrections};
    \path[arrow] (MS) --node {} (thresh);
    \node[rectangle, draw, text width=10em,
          text centered, minimum height=3em, minimum width=10em,
          left = 2em of MZ] (guess) {Initial guess};
    \path[arrow] (guess) --node {} (MZ);
    \path[-latex, thick] (thresh.east) edge[bend right=50] (MZ.east);
    \node[rectangle, draw, text width=10em,
          text centered, minimum height=3em, minimum width=10em,
          left = 2em of thresh] (pole) {Calculate pole masses};
    \node[inner sep=0pt, above = 3.5em of pole] (convergence) {};
    \path[draw,thick,dashed,above] (MS) --node {Converged?} (convergence);
    \path[arrow,dashed] (convergence) --node {} (pole);
  \end{tikzpicture}
  \caption{Semi-analytic algorithm for calculating the mass spectrum in a
    SUSY model; in a non-SUSY model, the SUSY parameters are replaced by
    the dimensionless model parameters instead.}
  \label{fig:semi-analytic-algorithm}
\end{figure}

\subsection{Tests and comparisons with other spectrum generators}

The semi-analytic algorithm provided by \fsas has been tested in the
CMSSM, CNMSSM and the C\ESSM by comparing to the results
obtained using the existing two-scale solver.  We have
carried out consistency checks between the two by confirming that the same
solution can be found in both solvers, provided it is a stable fixed
point in both.  In each of the models, the existing model solved using
the two-scale algorithm has been compared with versions of the model
using alternative boundary conditions.  For example, in the CMSSM,
instead of the traditional approach of fixing $|\mu|^2$ and $B\mu$
using the EWSB conditions, the value of $\mu$ is provided as an input
and $\mzero^2$ and $B\mu(M_X)$ are determined from the EWSB
conditions.\footnote{This alternative approach is useful in scenarios where
  one wishes to have direct control over the Higgsino masses, and therefore
  the composition of the lightest neutralino in the CMSSM, as was done in
  Ref.\ \cite{Athron:2016gor}.}  The benchmark points used as inputs for the
semi-analytic solver in each model are displayed in
\tabref{tab:semi_analytic_bms}.  In all three models, the
running parameters and pole mass spectra are found to differ at or
below the level of $0.1$\%.  Unit tests that perform these
comparisons have also been added to the extensive \fs test suite that
runs on a nightly basis.
%
\begin{table}[tbh]
  \centering
  \begin{tabularx}{\textwidth}{lX}
    \toprule
    Model & Unit test benchmark points for the semi-analytic solver\\
    \midrule
    CMSSM & $\mhalf = 500$ GeV, $\tan\beta = 10$, $\azero = 0$ GeV,
    $\mu(\MS) = 623.36$ GeV\\
    CNMSSM & $\mhalf = 133.33$ GeV, $\tan\beta = 10$,
    $\sign \mu_{\text{eff}} = -1$, $\azero = -300$ GeV, $\lambda(M_X) =-0.05$\\
    C\ESSM & $\tan\beta = 10$, $\lambda_3(M_X) = 0.12$, $\kappa(M_X) = 0.2$,
    $\mu^\prime(M_X) = 10$ TeV, $B^\prime\mu^\prime(M_X) = 0$ GeV$^2$,
    $s(\MS) = 4$ TeV, $\lambda_{1,2}(M_X) = 0.1$\\
    \bottomrule
  \end{tabularx}
  \caption{Input parameter values used for the unit tests comparing the
    results of the two-scale and semi-analytic algorithms in the CMSSM,
    CNMSSM and C\ESSM.  The notation for the CNMSSM follows that in
    Refs.~\cite{Allanach:2013kza,Ellwanger:2009dp}, while for the
    C\ESSM we use the notation of Ref.~\cite{Athron:2009bs}.}
  \label{tab:semi_analytic_bms}
\end{table}

In addition to carrying out unit tests on individual benchmark points, we have
also performed extensive scans in the CMSSM to confirm that the semi-analytic
solver produces results in agreement with the two-scale solver.  For most
points, this is found to be the case; however, we have also observed important
exceptions where non-negligible differences are found between the two
solvers.  In these cases, one solver may fail to converge to a stable
solution, or multiple solutions are present \cite{Allanach:2013cda} with
different stability properties in the two solvers.  We have checked that such
points nevertheless satisfy the boundary conditions imposed at each
scale and are indeed valid solutions.  More generally we have checked that
solutions found in one solver also correspond to fixed point solutions of the
other solver algorithm; that is, they satisfy all of the boundary conditions
so that the parameter values remain unchanged after applying a single step of
the iteration.

From these tests, we have found that in some cases the agreement
between the two solvers can depend quite sensitively on small differences
between them.  To illustrate this, in \figref{fig:semi_analytic_differences}
we show the percentage changes in the \DRbar mass spectrum in the CMSSM after
running points obtained using the two-scale solver through a single step of
the semi-analytic solver; if the point is also a fixed point of the latter,
this change should be negligible.  In this
scan, the change after a single iteration can be on the level of several
percent for a small number of points, reaching between $20$\% and $30$\% for
some exceptional points.
%
\begin{figure}[tbh]
  \centering
  \includegraphics[width=0.6\textwidth]{%
    doc/plots/semi_analytic_rel_errs/single_step_errors}
  \caption{Percentage changes in the CMSSM \DRbar mass spectrum after
    applying a single step of the semi-analytic solver to points obtained
    using the two-scale solver in a linear scan over $\mhalf \in
    [0,300]$ GeV, $\mzero \in [0,2]$ TeV, $\azero = 0$ GeV, $\tan\beta = 40$ and
    $\mu > 0$.  When run using the default Runge-Kutta algorithm provided
    by \fs and allowing the iteration to stop as soon as the precision goal of
    $10^{-4}$ is reached (red circles), changes between $1$\% and $30$\% are
    found for a small number of points.  By using the $8^{\text{th}}$ order
    Runge-Kutta integrator and ensuring convergence is reached in the
    estimate for $M_X$ by forcing 40 iterations in the two-scale algorithm,
    these differences are reduced below the level of $0.001$\% (blue
    diamonds).}
  \label{fig:semi_analytic_differences}
\end{figure}
%
That these points initially appear not to be fixed points of the semi-analytic
solver arises primarily from the fact that the semi-analytic coefficients,
and hence the EWSB solution, are sensitive to the estimate for the
high scale $M_X$, as well as numerical errors in the integration of the RGEs.
In particular, for the default convergence criteria imposed by \fs, the
two-scale solver's estimate for $M_X$ is not close enough to convergence,
leading to significant differences in the calculated low-energy soft
parameters.  By requiring convergence in the estimate for $M_X$, together with
using a higher-order Runge-Kutta integration\footnote{By default, \fs makes
  use of an adaptive $5^{\text{th}}$ order algorithm; to perform this test we
  have also implemented into \fstwo an $8^{\text{th}}$ order solver that makes
  use of the Runge-Kutta-Fehlberg method provided by the Boost library
  \texttt{odeint}.  This higher-order solver is available to the user by
  choosing it at the C++ level.} and demanding a higher precision for the
obtained EWSB solution, the change after one iteration is reduced below the
permille level.  Thus, it is important to be aware that differences in the
convergence properties of the two solvers can have an impact on the solutions
found, even if a given point would be a fixed point of both solvers.

The typical runtimes for the two solvers in the CMSSM are compared in
\figref{fig:semi_analytic_benchmark}.  The distributions are obtained
by randomly sampling the CMSSM input parameters
$\mzero\in [0.2,1]\unit{TeV}$, $\mhalf\in [0.2,1]\unit{TeV}$,
$\tan\beta\in [2,30]$, $\sign\mu\in\{-1,+1\}$ and
$\azero\in [-1,1]\unit{TeV}$.
The runtime of the semi-analytic solver is increased on average by a factor of
$\approx 3$ compared to the two-scale solver.  This increase is
mostly due to the increased number of iterations performed by the semi-analytic
solver.  For each outer iteration of the semi-analytic solver, the inner
iteration typically runs through a similar number of steps as for a full run
of the two-scale solver, with this number decreasing as convergence is
approached on each outer iteration.  Consequently, the total number of
iterations for the semi-analytic solver tends to be larger than that for the
two-scale solver by a similar factor.  There is also an additional cost
associated with running between scales to compute the semi-analytic
coefficients.
%
\begin{figure}[tbh]
  \centering
  \includegraphics[width=0.5\textwidth]{%
    doc/plots/semi_analytic_benchmark/i7-4702MQ}
  \caption{Runtime distributions of the CMSSM spectrum generators created with
    \fstwo using the two-scale and semi-analytic BVP solvers, obtained on
    an Intel i7-4702MQ CPU\@.  The distributions are normalized to have integral
    of unity.}
\label{fig:semi_analytic_benchmark}
\end{figure}

While the semi-analytic solver suffers from an increased runtime compared to
the two-scale solver, it is also able to provide complementary coverage of the
parameter space to that of the two-scale solver.  This is demonstrated in the left panel of
\figref{fig:cmssm_solution_regions} in the CMSSM, where the solutions found
by each solver are plotted in the $\mzero - \mu$ plane.
%
\begin{figure}
  \centering
  \includegraphics[width=0.48\textwidth]{%
    doc/plots/CMSSM_solution_regions/CMSSM_solution_regions}
  \includegraphics[width=0.48\textwidth]{%
    doc/plots/CMSSM_solution_regions/CMSSM_higgs_mass}
  \caption{Left panel: CMSSM solutions found by the two-scale solver (blue diamonds)
    and the semi-analytic solver (red circles) in the $\mzero - \mu$ plane.
    For both solvers, $\mhalf \in [0,1]\unit{TeV}$ is randomly sampled,
    while $\tan\beta = 10$ and $\azero = 0$ GeV\@.  The solutions obtained
    using the two-scale solver are found by randomly sampling
    $\mzero \in [0, 6]\unit{TeV}$ with $\sign \mu = 1$, and those for the
    semi-analytic solver are found by randomly sampling $\mu \in [0,1150]
    \unit{GeV}$.  Right panel: Change in the calculated lightest \CP-even
    Higgs mass $M_{h_1}$ from a reference value of $121$ GeV as a
    function of $\mu$ at fixed $\mzero = 4$ TeV for the two solvers.  The
    solutions shown correspond to a vertical slice at fixed $\mzero$ in the
    left-hand plot; $\mhalf$ is again varied in $[0,1]\unit{TeV}$,
    $\tan\beta = 10$ and $\azero = 0$ GeV.}
  \label{fig:cmssm_solution_regions}
\end{figure}
%
In this case, the use of the semi-analytic solver allows for a large number
of solutions to be found in the focus point region \cite{Chan:1997bi,
  Feng:1999hg,Feng:1999mn} at small values of $\mu \ll \mzero$, where the
two-scale solver is unable to find convergent solutions.  Conversely,
the two-scale solver is more effective for finding solutions with small
$\mzero$.  In general terms, the regions of parameter space in which the two
solvers are effective need not overlap, and the use of both in tandem allows
for a more complete picture of the parameter space to be obtained.  Moreover,
in the regions in which both solvers do find solutions, there is excellent
agreement between the two algorithms.\footnote{Provided the same solution is
  found if multiple solutions exist.}  This is illustrated in the right
panel of \figref{fig:cmssm_solution_regions}, where the lightest \CP-even
Higgs mass, expressed as the difference from a reference value of $121$ GeV,
is plotted for fixed $\mzero = 4$ TeV\@.  If both solvers find a solution for
this value of $\mzero$, the two values of the Higgs mass agree very well.

It is also evident from \figref{fig:cmssm_solution_regions} that the use of
both solvers allows features in the parameter space to be picked up that would
be missed by either solver alone.  In this case, in 
\figref{fig:cmssm_solution_regions} the two-scale solver only finds a single 
solution for each value of $\mhalf$, for fixed $\sign\mu = 1$, while the 
semi-analytic solver in some cases finds multiple solutions.  These
solutions have different values of $|\mu|$ for the same value of $\mhalf$,
leading to the sharp feature at low values of $\mu$ evident in the right panel
of \figref{fig:cmssm_solution_regions}.  The existence of multiple solutions to
the CMSSM BVP, and the inability of the ordinary two-scale fixed point iteration
to find all such solutions, are well-known and have previously been studied in
Refs.~\cite{Allanach:2013cda,Allanach:2013yua}.  In
\figref{fig:cmssm_multiple_solutions} we compare the results obtained using
the two BVP solvers in \fstwo to those found using the modified version of
\softsusy employed in Ref.~\cite{Allanach:2013cda}.
%
\begin{figure}
  \centering
  \includegraphics[width=0.6\textwidth]{%
    doc/plots/CMSSM_multiple_solutions/CMSSM_multiple_solutions}
  \caption{Solutions obtained using the two-scale and semi-analytic solvers
    in the CMSSM for fixed $\mhalf = 660$ GeV, $\azero = 0$ GeV and
    $\tan\beta = 40$.  For comparison, we also reproduce the curve shown in
    Figure~5 of Ref.~\cite{Allanach:2013cda}, where a modified version of
    \softsusy 3.3.7 was used to allow multiple solutions to be found to the
    CMSSM boundary conditions.  The inset plot shows the region $2790\unit{GeV}
    \leq \mzero \leq 2820\unit{GeV}$, $-555\unit{GeV} \leq \mu
    \leq -520\unit{GeV}$ in more detail.}
  \label{fig:cmssm_multiple_solutions}
\end{figure}
%
As expected, in this region of the parameter space the two-scale solver
produced by \fs finds at most only two solutions, corresponding to the two
possible signs of $\mu$.  The automatically generated semi-analytic solver
enables additional solutions to be found, which are in good agreement with
those found using the modified version of \softsusy.  Slight differences in
the values of $\mu$ and $\mzero$, and where solutions are found, arise from
small differences in the incorporated corrections between the two codes.  In
particular, the semi-analytic solver does not find valid solutions for
$\mu < -545$ GeV due to the tree-level mass of the \CP-odd Higgs becoming
tachyonic.  In general, multiple solutions at a given parameter point
can have significantly different phenomenological properties
\cite{Allanach:2013yua}, so that finding them is important to completely
characterize a model.  The availability of multiple solvers improves the
ability of \fs to locate additional solutions, without requiring extensive
modifications to the generated code.

The longer runtime of the semi-analytic solver can also be an
acceptable tradeoff if the model of interest cannot easily be
handled using the two-scale solver.  Constrained models such as the
CNMSSM and C\ESSM are examples of this.  To demonstrate the applicability of
the semi-analytic solver to these models, we have performed
scans over the parameter spaces of these models.

First we have performed scans in the CNMSSM to demonstrate that with the
semi-analytic solver we can sample the parameter space effectively, making
\fstwo the first public spectrum generator that can do this ``out of the box''.
We performed a four dimensional scan of the CNMSSM, using the model file
provided in \appref{app:CNMSSM} to produce a CNMSSM spectrum generator that
was then linked to \multinest 3.10 for efficient sampling of the parameter
space.  The input parameters were varied over the ranges
\begin{gather}
  2 \leq \tan \beta \leq 50, \\
  0\unit{TeV} \leq \mhalf \leq 5\unit{TeV}, \\
  -5\unit{TeV} \leq \azero \leq 0\unit{TeV}, \\
  -0.3 \, \leq \lambda(M_X) \leq 0.3, \\
  \sign \mu_{\textrm{eff}}  = +1,
\end{gather}
and the log likelihood function provided to \multinest was defined to be
\begin{equation*}
  \log L = -\frac{1}{5}({M_h^{\SM}}/{\unit{GeV}} - 125.09)^2 ,
\end{equation*}
where $M_h^{\SM}$ is the SM-like Higgs mass.  In this way the scan is directed
towards solutions with the observed Higgs mass of $125.09$ GeV\@.

\begin{figure}[tbh]
  \centering
  \includegraphics[width=0.48\textwidth]{%
     doc/plots/CNMSSM/cnmssm_4D-m0sqpos-full-m0_m12_sort-A0.pdf}
   \includegraphics[width=0.48\textwidth]{%
     doc/plots/CNMSSM/cnmssm_4D-m0sqpos-full_mhZoom_lam_m12_sort-mh.pdf} \\
  \includegraphics[width=0.48\textwidth]{%
    doc/plots/CNMSSM/cnmssm_4D-full-squark_mass_m0sqgtr0_m0_m12.pdf}
   \includegraphics[width=0.48\textwidth]{%
    doc/plots/CNMSSM/cnmssm_4D-m0sqgtr0-full-msu6.pdf}
   \caption{CNMSSM solutions obtained using \fstwo and \multinest 3.10 shown in
     the $\mzero-\mhalf$ plane with $\azero$ as a color contour (top left
     panel); $\mhalf-\lambda$ plane with the Higgs masses satisfying
     $M_h^{\SM} > 100$ GeV shown as a color contour (top right
     panel); $\mzero-\mhalf$ plane with the mass of the lightest up-type squark, which is
     predominantly stop (bottom left panel), and the $\mhalf-\azero$ plane with
     the heaviest up-type squark mass (bottom right panel).}
   \label{fig:cnmssm_solutions}
\end{figure}

The results are shown in \figref{fig:cnmssm_solutions}, where all points are
required to have $\mzero^2 > 0$.  As can be seen in the top left panel, the range
of $\mzero$ values found is considerably smaller than the range allowed for the
universal soft masses, $\mhalf$ and $\azero$, which are inputs.  The reason for
this is that in the CNMSSM a non-zero singlet VEV must be generated by
developing the correct shape of the scalar potential.  This could be achieved
in the usual way through a negative quadratic term if the soft breaking singlet
mass squared, $m_S^2$, is negative.  In fact, since the NMSSM scalar potential
also contains the cubic singlet terms $\kappa A_\kappa S^3 / 3 + \mathrm{h.c.}$,
the requirement $m_S^2 < 0$ can be relaxed so that an approximate condition for
generating the correct shape of the potential reads $A_\kappa^2 \gtrsim
9 m_S^2$ \cite{Ellwanger:1996gw,Derendinger:1983bz}.  If $\mzero \gg
\azero / 3$, this condition is satisfied only if $m_S^2$ is driven to be
sufficiently small during the RG evolution from the GUT scale to the SUSY
scale.  This in turn can be achieved for large enough values of the
superpotential cubic singlet coupling $\kappa$ and the singlet-Higgs coupling
$\lambda$.

However, in this constrained model, where the soft trilinears are not free
input parameters at the SUSY scale, it is also the case that large values of $\lambda$
always generate substantial singlet mixing that can reduce the lightest
\CP-even Higgs mass.  As a result, a $125$ GeV Higgs mass is obtained with small
values of $\lambda$, to avoid this mixing, as well as large $\mhalf$, as can be
seen in the top right panel of \figref{fig:cnmssm_solutions}.  For such small
singlet Yukawa couplings, the RG flow between the GUT and SUSY scales leaves
$m_S^2(\MS)$ and $A_\kappa(\MS)$ close to their GUT scale values, namely
$\mzero^2$ and $\azero$.  As a result, $\mzero$ is heavily constrained if
the condition $A_\kappa^2 \gtrsim 9 m_S^2$ is to be satisfied in the absence of
significant RG evolution for $m_S$, as the top left panel of
\figref{fig:cnmssm_solutions} demonstrates.

The consequences of such small $\mzero$ values can be seen in the bottom left
and bottom right panels of \figref{fig:cnmssm_solutions}, where we plot the
lightest and heaviest up-type squark mass, respectively, to illustrate that the
squark masses are now predominantly set by the universal gaugino mass, with
little influence from $\mzero$ or $\azero$.  Furthermore, the squarks are always
lighter than the gluino, which has a mass $\approx 2\, \mhalf$.

\begin{figure}[tbh]
  \centering
  \includegraphics[width=0.48\textwidth]{%
     doc/plots/CE6SSM/ce6ssm_tb10-scan-l-k-vs_combined_m0_m12_ccA0.pdf}
   \includegraphics[width=0.48\textwidth]{%
     doc/plots/CE6SSM/ce6ssm_tb10-scan-l-k-vs_combined_m0_m12_ccmsu1.pdf} \\
  \includegraphics[width=0.48\textwidth]{%
    doc/plots/CE6SSM/ce6ssm_tb10-scan-l-k-vs_combined_m0_m12_cclam.pdf}
   \includegraphics[width=0.48\textwidth]{%
    doc/plots/CE6SSM/ce6ssm_tb10-scan-l-k-vs_combined_m0_m12_cckap.pdf}
   \caption{C\ESSM solutions obtained using \fstwo and \multinest 3.10 shown in
     the $\mzero-\mhalf$ plane with $\azero$, the lightest up-type squark mass
     $M_{\tilde{u}_1}$,  $|\lambda|$ and $|\kappa|$ as color contours.}
  \label{fig:ce6ssm_solutions}
\end{figure}

Compared to the CNMSSM, we find that the situation in the C\ESSM is rather
different.  The C\ESSM is an alternative to the CNMSSM in which an elementary
$\mu$ term is forbidden by a $U(1)_N$ gauge symmetry, and complete $E_6$ matter
supermultiplets are included to ensure anomaly cancellation.  As with the
CNMSSM, the two-scale solver is ineffective for finding solutions because EWSB
needs to have a soft mass as an output, but one can make a spectrum generator
with the semi-analytic solver where the universal GUT scale masses $\mzero$,
$\mhalf$ and $\azero$ are EWSB outputs.\footnote{The definitions of the
  new Lagrangian parameters and GUT scale constraints in the C\ESSM that we
  use may be found in Ref.~\cite{Athron:2009bs}.}  Using the \modelname{C\ESSM}
model file provided with \fstwo, we carried out a scan in which the
singlet-Higgs Yukawa coupling, $\lambda(M_X) \equiv \lambda_3(M_X)$, and the
exotic Yukawa coupling, $\kappa(M_X)$, were varied over $[-1,1]$ and the singlet
VEV, $s$, was varied over $[1,500] \unit{TeV}$.  The results of this scan are
shown in \figref{fig:ce6ssm_solutions}.

In contrast to the CNMSSM, in the C\ESSM it is very easy to have large
$\mzero$ values as there is an exotic Yukawa coupling between the singlet and
extra colored matter introduced to avoid gauge anomalies that drives the soft
singlet mass negative, providing a radiative symmetry breaking mechanism.
Additionally, whereas $\mzero \lesssim \mhalf$ in the CNMSSM, in this model
it is typically the case that $\mzero$ is larger than $\mhalf$.  This is
in qualitative agreement with the literature \cite{Athron:2009ue,
  Athron:2009bs,Athron:2012sq,Athron:2012pw}, and arises because the new
colored matter results in heavily modified RGEs in which the 1-loop
$\beta$ function of the strong coupling now vanishes.  The squark masses are
mostly set by $\mzero$ as a result, as is shown in the top right panel of
\figref{fig:ce6ssm_solutions}, while the bottom panels show that the range for
$\lambda$ and $\kappa$ at the GUT scale is very wide, despite strong
constraints on $\lambda$ at the electroweak scale coming from requirements for correct
EWSB\@.   We do not make a detailed quantitative comparison to previous work in
the literature here, but note that this is the first time that the C\ESSM
results have been presented with the same level of precision (full 2-loop RGEs,
1-loop pole masses) as is standard in the CMSSM and significant quantitative
differences are to be expected.

For a more precise comparison between calculations performed at the
same level of precision, we have also performed scans in a recently proposed
variant of the \ESSM, the so-called CS\ESSM \cite{Athron:2015vxg,
  Athron:2016gor}. Here the results obtained using \fstwo have been checked for
agreement with those obtained from a hand-written prototype of the
semi-analytic solver that was implemented for the studies in
Refs.~\cite{Athron:2015vxg,Athron:2016gor}.  The solutions found using the
generated CS\ESSM spectrum generator are compared with those found in
Ref.~\cite{Athron:2016gor} in \figref{fig:cse6ssm_m12_Azero_plane}.  The viable
solution regions and values of the model parameters are found to be in very
good agreement with the results obtained using the earlier code.
%
\begin{figure}[tbh]
  \centering
  \includegraphics[width=0.48\textwidth]{%
    doc/plots/CSE6SSM/CSE6SSM_m12_Azero_m0}
  \includegraphics[width=0.48\textwidth]{%
    doc/plots/CSE6SSM/CSE6SSM_m12_Azero_m0_handwritten}
  \caption{CS\ESSM solutions obtained using the spectrum generator
    automatically generated by \fstwo (left panel) and using the prototype
    spectrum generator used in the numerical analysis of
    Ref.~\cite{Athron:2016gor} (right panel), showing good agreement between
    the two codes.  Note that here we have not applied the limits on the Higgs
    mass or dark matter relic density that lead to additional restrictions on
    the parameter space, as discussed in Ref.~\cite{Athron:2016gor}.}
  \label{fig:cse6ssm_m12_Azero_plane}
\end{figure}

\section{\feft}
\label{sec:feft}

 \feft is a method to predict the lightest Higgs pole mass in any \BSM
 model accurately for both high and low new physics scales $\MS$ and
 was presented first in Ref.~\cite{Athron:2016fuq}.  An implementation
 of this method was first released in \fs 1.7.0.  Here we present an
 upgrade at the next-to-leading order and next-to-leading logarithmic
 (NLO+NLL) accuracy, which we release in \fstwo.

\feft combines an EFT approach with a diagrammatic
calculation, allowing for an all-order resummation of large logarithms
of the ratio $\MS/m_t$, together with the inclusion of all
non-logarithmic 1-loop contributions.  In particular, all
power-suppressed 1-loop contributions of $O(v^2/\MS^2)$ are
included in \feft, which would otherwise be neglected in a pure EFT
calculation.  Thanks to these properties, \feft maintains the accuracy
at all scales: For low scales the prediction agrees with a
fixed-order calculation; for large scales it agrees with a pure EFT
calculation. In the intermediate region, where the $O(v^2/\MS^2)$
terms are small, but still non-negligible, \feft gives the correct
fixed-order result plus higher-order logarithms, thus resolving the
ambiguity between the fixed-order and the pure EFT approach.

In Ref.~\cite{Athron:2016fuq} already several versions of the \feft
approach have been extensively discussed and compared with existing
calculations of the lightest Higgs boson mass in the MSSM and other
supersymmetric models. The approach has also been implemented recently
in \sarah/\spheno \cite{Staub:2017jnp}, including 2-loop corrections in
the matching. The version implemented in  \fstwo contains
additional improvements resulting in a higher accuracy. In the
following, we briefly summarize the main idea of \feft, then explain
the details of the implemented version and how to use it. For further
details of the approach and a detailed comparison of theoretical
uncertainties, we refer to Ref.~\cite{Athron:2016fuq}.

\subsection{Basic matching condition}

\feft performs a matching of the \BSM model to the SM,
thereby determining the quartic Higgs coupling $\lambda$ of the
SM\@.  The basic ingredient of \feft to fix $\lambda$ at
the matching scale is a Higgs pole mass matching condition
%
\begin{align}
  (M_h^\SM)^2 = (M_h^\BSM)^2 \,,
  \label{eq:Higgs_pole_mass_matching}
\end{align}
%
where $M_h^\SM$ is the Higgs pole mass calculated in the SM at the 1-loop level and $M_h^\BSM$ is the corresponding SM-like
Higgs pole mass in the \BSM model, also at the 1-loop
level.
Generally, the Higgs pole mass is computed in any \BSM model
by solving the following
equation (or a suitable matrix generalization):
%
\begin{align}
  (M_h^\BSM)^2 = (m_h^\BSM)^2 - \re\Sigma_h^{\BSM}(p^2) + \frac{t_h^\BSM}{v} \,,
  \label{eq:Mh_mom_iteration}
\end{align}
%
where $m_h^\BSM$ is the respective tree-level mass and
$\Sigma_h^{\BSM}$ and $t_h^\BSM$ are the Higgs self
energy and tadpole in the \MSbar/\DRbar scheme.
In principle, in an all-order calculation,  the
Higgs self energy has to be evaluated at the momentum $p^2 =
(M_h^\SM)^2=(M_h^\BSM)^2$.
  From this condition the quartic Higgs coupling of the SM can be extracted as
%
\begin{align}
  \lambda = \frac{1}{v^2} \left[
     (M_h^\BSM)^2 + \re\Sigma_h^{\SM}((M_h^\SM)^2) - \frac{t_h^\SM}{v}
  \right] \,.
  \label{eq:lambda_from_Mh_matching}
\end{align}
%

This matching is equivalent
to the one of pure EFT calculations \cite{Bagnaschi:2014rsa,Vega:2015fna} at the
1-loop level, up to power-suppressed terms. Correspondingly, the resulting
Higgs boson mass is exact at the 1-loop level and takes into account
all leading logarithms \cite{Athron:2016fuq}.

\subsection{New matching procedure in \fstwo}
\label{sec:FS-2_matching_procedure}

\fstwo has an improved implementation of the approach, which is
still exact at the 1-loop level but also
correctly resums next-to-leading logarithms.
This improvement originates from an amended
matching procedure. As mentioned before, the matching procedure
in Eq.~\eqref{eq:lambda_from_Mh_matching}
is equivalent to a pure EFT matching at the 1-loop level. However,
depending on implementation details, it can differ by terms of
2-loop or higher order. If these spurious 2-loop terms contain
 (next-to-leading) large logarithms, they spoil the correct
resummation of (next-to-leading) logarithms by RGE running.

By construction, all versions of \feft discussed in
Ref.~\cite{Athron:2016fuq} and Ref.~\cite{Staub:2017jnp} are correct at the
leading logarithmic level, however not all subleading logarithms are correctly
included.
%

In the following, we discuss the two potential origins of these subleading
logarithms and how they are avoided by the improved  implementation in
\fstwo.

\paragraph{Insertion of 1-loop parameters into 1-loop \BSM self energies or
  tadpoles}

The first potential source of large 2-loop logarithms in the matching
procedure is the insertion of parameters, which have been obtained from the SM
via a 1-loop matching, into the 1-loop self energies or tadpoles of the \BSM
model.  We illustrate this effect with the most important parameter, the top
Yukawa coupling:
The running top Yukawa coupling of the \BSM model  $y_t^\BSM$ is
determined by a matching as\footnote{We ignore potential tree-level
  factors here for brevity.}
%
\begin{align}
  y_t^\BSM = y_t^\SM + \Delta y_t \,, \label{eq:yt_SM_BSM}
\end{align}
%
where $\Delta y_t$ is of 1-loop order (but without large logarithms).
At the same time, the Higgs pole mass calculations on the left-hand side and
right-hand side of Eq.~\eqref{eq:Higgs_pole_mass_matching} are of the form
%
\begin{align}
  (M_h^\SM)^2 &= (m_h^\SM)^2 + \propto \frac{(v^\SM)^2(y_t^\SM)^4}{(4\pi)^2} \log\frac{m_t^\SM}{Q} + \cdots \,,
  \label{eq:Mh_top_SM}\\
  (M_h^\BSM)^2 &= (m_h^\BSM)^2 + \propto \frac{(v^\BSM)^2(y_t^\BSM)^4}{(4\pi)^2} \log\frac{m_t^\BSM}{Q} + \cdots \,,
  \label{eq:Mh_top_BSM}
\end{align}
%
where the matching scale $Q$ is of the order $\MS$.
If Eq.\ \eqref{eq:Mh_top_SM} and Eq.\ \eqref{eq:Mh_top_BSM} are set equal and
the relation Eq.\ \eqref{eq:yt_SM_BSM} is inserted, potentially
large 2-loop terms for example of the form
%
\begin{align}
  \frac{4(v^\SM)^2(y_t^\SM)^3 \Delta y_t}{(4\pi)^2} \log\frac{m_t^\SM}{Q} + \cdots
\label{eq:twolooplogsyt}
\end{align}
%
remain. Such terms effectively shift the quartic Higgs coupling of the
SM by  next-to-leading logarithmic 2-loop terms.%
\footnote{If the self energies are evaluated at the
  2-loop level, the problem repeats itself one order higher, i.e., the
  term in Eq.~\eqref{eq:twolooplogsyt} is cancelled but similar terms of
  next-to-leading logarithmic 3-loop order remain.}

In order to avoid large higher-order logarithms originating from the
insertion of 1-loop parameters into 1-loop \BSM self energies and
tadpoles, \fstwo maintains two different sets of running \BSM
parameters: One parameter set which has been obtained from the
SM using a tree-level matching, and another set
from the SM using  1-loop matching.
%
The tree-level parameter set is used to evaluate the 1-loop self
energies and tadpoles on the right-hand side of
Eq.~\eqref{eq:Mh_mom_iteration}.  In this way, no terms like the ones
in Eq.~\eqref{eq:twolooplogsyt} are generated.
%
The 1-loop parameter set is used to evaluate the tree-level Higgs
mass (matrix) of the \BSM model.  In this way, the desired 1-loop
corrections to the quartic Higgs coupling $\lambda$ are generated.

\paragraph{Momentum iteration}

The second source of large 2-loop logarithms in the matching has to
do with the momentum argument of the self energies entering
Eqs.~\eqref{eq:Higgs_pole_mass_matching} and \eqref{eq:Mh_mom_iteration}.
Writing $p^2=(m_h^\BSM)^2+\Delta p^2$, we see
that the momentum argument of Eq.~\eqref{eq:Mh_mom_iteration}
contains the 1-loop term $\Delta p^2$ (which
also involves large logarithms).
%
The difference between the left-hand side and the right-hand side of the
matching condition, Eq.~\eqref{eq:Higgs_pole_mass_matching},
then contains 2-loop terms, which can be expanded as
%
\begin{align}
\left.\left(
\frac{\partial}{\partial
        p^2}\re\Sigma_h^{\BSM}(p^2)
-
\frac{\partial}{\partial
        p^2}\re\Sigma_h^{\SM}(p^2)
\right)
\right|_{p^2 = (m_h^\BSM)^2}
\,\Delta p^2
\,.
\end{align}
%
If the self energies
are evaluated at the 1-loop level and $p^2$ is determined as
described above, these terms do not cancel against anything.
Like the terms discussed in Eq.~\eqref{eq:twolooplogsyt}, these terms
would then lead to large 2-loop next-to-leading logarithms in the
determination of $\lambda$.%

To avoid large higher-order logarithmic contributions coming from the
momentum argument, \fstwo does not  perform the usual momentum
iteration when the Higgs pole masses in the SM and in the
\BSM model are calculated at the matching scale for
Eq.~\eqref{eq:Higgs_pole_mass_matching}.  Instead, the SM Higgs pole mass at
the matching scale is now calculated as
%
\begin{align}
  (M_h^\SM)^2 = (m_h^\SM)^2 - \re\Sigma_h^{\SM}((m_h^\BSM)^2) + \frac{t_h^\SM}{v} \,,
  \label{eq:Higgs_pole_mass_matching_new}
\end{align}
%
where the self energy momentum is set to the tree-level
\MSbar/\DRbar Higgs mass $m_h^\BSM$ in the \BSM model at the matching
scale, which is calculated in terms of running \BSM parameters which
have been obtained by a tree-level matching.
%
A similar expression is used to calculate the Higgs pole mass in the
\BSM model $M_h^\BSM$, where we also insert $p^2 = (m_h^\BSM)^2$ as the
self energy momentum in order to enable cancellation of
momentum-dependent terms.  For example, if the \BSM Higgs is a
singlet, the \BSM Higgs pole mass is calculated as
%
\begin{align}
  (M_h^\BSM)^2 = (m_h^\BSM)^2 - \re\Sigma_h^{\BSM}((m_h^\BSM)^2) + \frac{t_h^\BSM}{v} \,.
  \label{eq:Higgs_pole_mass_matching_new_BSM_singlet}
\end{align}
%
On the other hand, if the \BSM Higgs is a multiplet and the $k$-th
element is the SM-like Higgs, then the SM-like \BSM Higgs pole mass is
the $k$-th eigenvalue of the loop-corrected mass matrix $\mathsf{M}_h$
in the interaction eigenstate basis,
%
\begin{align}
  (\mathsf{M}_h)_{ij} =
  (m_h^\BSM)^2_{ij} - \re\Sigma_{h,ij}^{\BSM}((m_{h_k}^\BSM)^2) + \frac{t_{h_i}^\BSM}{v_i}\delta_{ij}  \,.
  \label{eq:Higgs_pole_mass_matching_new_BSM_multiplet}
\end{align}
%

By employing this new matching procedure, \feft consistently avoids
large higher-order logarithms and thereby
resums the leading and next-to-leading logarithms and includes all
non-logarithmic 1-loop contributions.

\subsection{Comparison of old and improved \feft implementations}

In \figref{fig:feft_comparison} we show a comparison of the predicted
lightest \CP-even Higgs mass in the MSSM between the old \feft
implementation of \fs 1.7.4 (red dotted line) and the improved version
in \fs 2.0 (red solid line).  For small SUSY scales of $\MS <
300\unit{GeV}$, we find that the improved version still reproduces the
fixed-order calculation.
%
As can be seen in the left panel of \figref{fig:feft_comparison} for
vanishing stop mixing, $X_t = 0$, both the old and the improved
version closely reproduce the 2-loop pure EFT calculation with
\HSSUSY: For SUSY scales above $10\unit{TeV}$, the old version
deviates from \HSSUSY-2L by around $600\unit{MeV}$ while the
improved one deviates by around $10\unit{MeV}$.  This is due to the
fact that for $X_t = 0$, the 2-loop threshold correction to the
quartic Higgs coupling at the SUSY scale is negligible.
%
However, for maximal stop mixing, $X_t/\MS=\sqrt{6}$, which is the
region where the old implementation showed the largest theoretical
uncertainty, we find up to $3\unit{GeV}$ difference between the old
and the improved implementation, see the right panel of
\figref{fig:feft_comparison}.  This difference manifests the consequences of the
different treatment of higher-order terms in the two versions,
especially the inclusion of large 2-loop logarithms in the old
implementation.

The figure shows furthermore that the improved version (which
performs a 1-loop calculation) is now able to perfectly reproduce the
1-loop pure EFT calculation of \HSSUSY (blue crosses) for
\emph{arbitrary} stop mixing and SUSY scales above $\approx
1\unit{TeV}$.  This is in contrast to the old version, which shows a
stronger deviation of around $2\unit{GeV}$ from the 1-loop pure EFT
calculation for large stop mixing, see the right panel of
\figref{fig:feft_comparison}.  Compared to the 2-loop pure EFT
calculation of \HSSUSY (blue dashed line), both the old and the
improved version deviate by around $1$--$2\unit{GeV}$ for non-zero
stop mixing.  This deviation can be attributed to genuine 2-loop
contributions.  Note that \HSSUSY does not reproduce the Higgs mass
prediction of the fixed-order calculation for
$\MS \lesssim 1\unit{TeV}$, because it neglects terms of
$O(v^2/\MS^2)$.
%
\begin{figure}[tbh]
  \centering
  \includegraphics[width=0.49\textwidth]{{{doc/plots/FlexibleEFTHiggs/scan_Mh_MS_TB-5_Xt-0}}} \hfill
  \includegraphics[width=0.49\textwidth]{{{doc/plots/FlexibleEFTHiggs/scan_Mh_Xt_TB-5_MS-2000}}}
  \caption{Comparison of the predicted lightest \CP-even Higgs pole mass in the
    MSSM using the \feft implementations of \fs 1.7.4 and 2.0 for
    $\tan\beta = 5$.  In the left panel we use $X_t = 0$ and in the
    right panel $\MS = 2\unit{TeV}$.}
  \label{fig:feft_comparison}
\end{figure}

\subsection{Choosing \feft in the model file}

In order to build a \feft spectrum generator, the
\code{FlexibleEFTHiggs} flag can be set to \code{True} in the model
file, see \tabref{tab:feft_info}:
%
\begin{lstlisting}
FlexibleEFTHiggs = True;
\end{lstlisting}
%
In \feft spectrum generators, the low-energy boundary condition cannot
be modified, because it is fixed internally to perform a matching of
the $\SM(5)$ to the full SM\@.  The considered \BSM model is
matched to the SM at the \BSM matching scale, $\Qmatch$,
which is set to the \code{SUSYScale} by default.  In this matching,
the running normalized gauge couplings $g_i^\BSM(\Qmatch)$
($i=1,2,3$), the Yukawa coupling matrices $Y_f^\BSM(\Qmatch)$
($f=u,d,e$) and the SM-like vacuum expectation value $v^\BSM(\Qmatch)$
of the \BSM model are determined automatically from the matching
conditions
%
\begin{subequations}
\begin{align}
  (M_V^\BSM)^2 &= (M_V^\SM)^2, \qquad V = W, Z,
  \label{eq:WZ-matching} \\
  M_f^\BSM &= M_f^\SM, \qquad f = e, \mu, \tau, u, d, c, s, t, b,
  \label{eq:f-matching} \\
  \aem^\BSM(\Qmatch) &= \aem^\SM(\Qmatch)\times (1 + \Delta\aem),
  \label{eq:aem-matching} \\
  \as^\BSM(\Qmatch) &= \as^\SM(\Qmatch)\times (1 + \Delta\as),
  \label{eq:as-matching}
\end{align}%
\label{eq:FSEFTHiggs_matching}%
\end{subequations}%
%
where $\Delta\aem$ and $\Delta\as$ are the known 1-loop
threshold corrections \cite{Hall:1980kf}, including potential \MSbar
to \DRbar conversion terms \cite{Martin:1993yx}.
\fs imposes the individual matching conditions in
Eqs.~\eqref{eq:FSEFTHiggs_matching} at the appropriate loop orders
such that no large 2-loop logarithms are generated, as
described in \secref{sec:FS-2_matching_procedure}.  For example, to
obtain the correct Higgs mass in the MSSM at the 1-loop level,
Eqs.~\eqref{eq:WZ-matching} and \eqref{eq:aem-matching} are imposed at
the 1-loop level while Eqs.~\eqref{eq:f-matching} and
\eqref{eq:as-matching} are imposed at the tree level.  The SM-like
vacuum expectation value $v^\BSM(\Qmatch)$ of the \BSM model is
defined as
%
\begin{align}
  v^\BSM(\Qmatch) = \frac{2 m_Z^\BSM(\Qmatch)}{\sqrt{\big(g_Y^\BSM(\Qmatch)\big)^2 + \big(g_2^\BSM(\Qmatch)\big)^2}},
  \label{eq:feft_vev_definition}
\end{align}
%
where $m_Z^\BSM(\Qmatch)$ is the running $Z$ boson mass and
$g_Y^\BSM(\Qmatch)$ and $g_2^\BSM(\Qmatch)$ are the running
electroweak gauge couplings in the \BSM model at the matching scale.

The running \BSM model parameters can be given as input at either the
\code{SUSYScale} or at the \code{HighScale}.  The following example
demonstrates how to fix the \DRbar parameters of the MSSM at the SUSY
scale.
%
\begin{example}[label=ex:MSSMEFTHiggs]
In the \feft/MSSM model (\code{MSSMEFTHiggs})
the soft-breaking MSSM parameters, the $\mu$ parameter and $\tan\beta$
are input at the SUSY scale, \MS.  Thus, the boundary condition at the SUSY
scale has the form
%
\begin{lstlisting}
SUSYScaleInput = {
    {vu, Sqrt[vu^2 + vd^2] Sin[ArcTan[TanBeta]]},
    {vd, Sqrt[vu^2 + vd^2] Cos[ArcTan[TanBeta]]},
    {MassB, M1Input},
    {MassWB, M2Input},
    {MassG, M3Input},
    {mq2, mq2Input},
    {mu2, mu2Input},
    {md2, md2Input},
    {ml2, ml2Input},
    {me2, me2Input},
    {\[Mu], MuInput},
    {B[\[Mu]], mAInput^2/(TanBeta + 1/TanBeta)},
    {T[Yu], AuInput Yu},
    {T[Yd], AdInput Yd},
    {T[Ye], AeInput Ye}
};
\end{lstlisting}
%
The symbols \code{TanBeta}, \code{M1Input}, \code{M2Input},
\code{M3Input}, \code{mq2Input}, \code{mu2Input}, \code{md2Input},
\code{ml2Input}, \code{me2Input}, \code{MuInput}, \code{mAInput},
\code{AuInput}, \code{AdInput}, \code{AeInput} describe the MSSM input
parameters $\tan\beta^{\DRbar}(\MS)$, $M_i(\MS)$, $m_{\tilde{f}}^2(\MS)$
($f=q,u,d,l,e$), $\mu(\MS)$, $m_A(\MS)$ and $A_f(\MS)$
($f=u,d,e$) in the \DRbar scheme at the SUSY scale.

Note that no explicit SUSY scale boundary condition for the gauge
couplings, $g_1$, $g_2$ and $g_3$, and Yukawa couplings, $Y_u$, $Y_d$
and $Y_e$, of the MSSM has to be specified, because they are all fixed
automatically at \Qmatch\ using the \feft matching conditions, Eqs.\
\eqref{eq:FSEFTHiggs_matching}.

However, there is a subtlety with the vacuum expectation values: In the
above boundary condition, the input value $\tan\beta^{\DRbar}(\MS)$ is
used to fix the ratio of $v_u(\MS)$ and $v_d(\MS)$.  However, their
magnitude $\sqrt{v_u^2 + v_d^2}$ is unfixed so far.  To fix
it, we can use the value of $v^\MSSM(\Qmatch)$, which is
automatically determined by \fs at the matching scale, see
Eq.~\eqref{eq:feft_vev_definition}.  Therefore, we want to set
%
\begin{subequations}
\begin{align}
  v_u(\Qmatch) &= v^\MSSM(\Qmatch) \sin\beta^{\DRbar}(\Qmatch) \,, \\
  v_d(\Qmatch) &= v^\MSSM(\Qmatch) \cos\beta^{\DRbar}(\Qmatch) \,.
\end{align}%
\label{eq:VEV_matching}%
\end{subequations}%
%
Such a matching is not done automatically by \feft.  The user must
specify how the VEVs of any Higgs fields that have electroweak
interactions are related to the electroweak VEV, $v^{\BSM}(\Qmatch)$,
which is given above for the MSSM\@.  To do this, the model file has an
additional constraint list: \code{MatchingScaleInput}.  Conditions
to relate or fix model parameters at the matching scale
$\Qmatch$ can be expressed using the \code{MatchingScaleInput} list.  This can actually be used to set any BSM parameter or to override the automatic \feft matching conditions if the user wishes.  However, it is only required that the user specifies the matching for the VEVs here.
To express the relations of Eqs.~\eqref{eq:VEV_matching}, we set
%
\begin{lstlisting}
MatchingScaleInput = {
    {vu, VEV Sin[ArcTan[vu/vd]]},
    {vd, VEV Cos[ArcTan[vu/vd]]}
};
\end{lstlisting}
%
The symbol \code{VEV} is reserved by \fs and represents the running
SM-like vacuum expectation value $v^\MSSM(\Qmatch)$ at the matching
scale, as defined in Eq.~\eqref{eq:feft_vev_definition}.
\end{example}
%
\begin{OptionTable}{\feft model file options}{tab:feft_info}[lllX]
  \code{FlexibleEFTHiggs} & \code{False} & \code{True} or \code{False} &
  Flag to enable/ disable \feft \\
  \code{VEV} & -- & -- & SM-like VEV in the \BSM model, $v^\BSM(\Qmatch)$ \\
  \code{MatchingScaleInput} & \code{\{\}} & list of 2-tuples &
  boundary conditions for \BSM parameters at the matching scale $\Qmatch$\\
  \bottomrule
\end{OptionTable}

Once the model file is written, a spectrum generator can be created and
run in the usual way.  For example, to build the model described in
\exref{ex:MSSMEFTHiggs}, one may run:
%
\begin{lstlisting}[language=bash]
$ ./createmodel --name=MSSMEFTHiggs
$ ./configure --with-models=MSSMEFTHiggs
$ make
\end{lstlisting}%% $
%
These commands create the \fs spectrum generator for the
\modelname{MSSMEFTHiggs} model.  The generated spectrum generator can then
be run from the command line as
%
\begin{lstlisting}[language=bash]
$ cd models/MSSMEFTHiggs
$ ./run_MSSMEFTHiggs.x --slha-input-file=LesHouches.in.MSSMEFTHiggs
\end{lstlisting}%% $
%

The only difference with the SLHA interface is that there are new
\feft-specific options in the SLHA file. In \feft, the pole masses of
the BSM particles are calculated at the scale $\QpoleBSM$, which is
set to the \code{SUSYScale} by default.  The scale $\QpoleBSM$ can be
changed by setting the entry \code{FlexibleSUSY[17]} to a non-zero
value in the SLHA input file.  Similarly, in the \mathematica interface
$\QpoleBSM$ can be changed by setting \code{poleMassScale} to a
non-zero value.
%
The pole masses of the SM particles are calculated at the
scale $\QpoleSM$, which is set to $M_t$ by default.  The scale
$\QpoleSM$ can be changed by setting the entry \code{FlexibleSUSY[18]}
to a non-zero value in the SLHA input file.  In the \mathematica
interface, $\QpoleSM$ can be changed by setting \code{eftPoleMassScale}
to a non-zero value.
%
The matching scale $\Qmatch$ is set to the \code{SUSYScale} by
default.  It can be changed by setting the entry \code{FlexibleSUSY[19]} to a
non-zero value in the SLHA input file.  In the \mathematica interface,
$\Qmatch$ can be changed by setting \code{eftMatchingScale} to a
non-zero value.

\begin{example}[label=ex:MSSMEFTHiggs_uncertainty_estimate]
  This example demonstrates how a partial uncertainty estimate of the
  lightest Higgs pole mass can be made with \feft.  The
  uncertainty is estimated by varying the matching scale $\Qmatch$ and
  the scale $\QpoleSM$, at which the lightest Higgs pole mass is
  calculated, both by a factor 2.
  %
  \begin{lstlisting}
Get["models/MSSMEFTHiggs/MSSMEFTHiggs_librarylink.m"];

Mtpole = 173.34;

(* generate logarithmically spaced range [start, stop] *)
LogRange[start_, stop_, steps_] :=
    Exp /@ Range[Log[start], Log[stop],
                 (Log[stop] - Log[start])/steps];

(* generate logarithmically spaced range [Q / 2, 2 Q] *)
GenerateScales[Q_] := LogRange[Q/2, 2 Q, 10];

(* run MSSMEFTHiggs spectrum generator *)
RunMSSMEFTHiggs[MS_, TB_, Xt_, Qpole_, Qmatch_] :=
    Module[{handle, spectrum},
           handle = FSMSSMEFTHiggsOpenHandle[
               fsSettings -> {
                   precisionGoal -> 1.*^-5,
                   maxIterations -> 10000,
                   poleMassLoopOrder -> 2,
                   ewsbLoopOrder -> 2,
                   betaFunctionLoopOrder -> 3,
                   thresholdCorrectionsLoopOrder -> 2,
                   poleMassScale -> 0,
                   eftPoleMassScale -> Qpole,
                   eftMatchingScale -> Qmatch,
                   eftMatchingLoopOrderUp -> 1,
                   eftMatchingLoopOrderDown -> 1,
                   calculateBSMMasses -> 0
               },
               fsSMParameters -> {
                   Mt -> Mtpole
               },
               fsModelParameters -> {
                   MSUSY   -> MS,
                   M1Input -> MS,
                   M2Input -> MS,
                   M3Input -> MS,
                   MuInput -> MS,
                   mAInput -> MS,
                   TanBeta -> TB,
                   mq2Input -> MS^2 IdentityMatrix[3],
                   mu2Input -> MS^2 IdentityMatrix[3],
                   md2Input -> MS^2 IdentityMatrix[3],
                   ml2Input -> MS^2 IdentityMatrix[3],
                   me2Input -> MS^2 IdentityMatrix[3],
                   AuInput -> {{MS/TB, 0    , 0},
                               {0    , MS/TB, 0},
                               {0    , 0    , MS/TB + Xt MS}},
                   AdInput -> MS TB IdentityMatrix[3],
                   AeInput -> MS TB IdentityMatrix[3]
               }
           ];
           spectrum = FSMSSMEFTHiggsCalculateSpectrum[handle];
           FSMSSMEFTHiggsCloseHandle[handle];
           spectrum
          ];

(* extract lightest Higgs pole mass Pole[M[hh]] from spectrum *)
RunMSSMEFTHiggsMh[pars__] :=
    (Pole[M[hh]] /. (MSSMEFTHiggs /. RunMSSMEFTHiggs[pars]))[[1]];

(* calculate Higgs mass and perform scale variation *)
RunMSSMEFTHiggsUncertainty[MS_, TB_, Xt_] :=
    Module[{MhMean, DMh, varyQpole, varyQmatch},
           MhMean = RunMSSMEFTHiggsMh[MS, TB, Xt, 0, 0];
           varyQpole  = RunMSSMEFTHiggsMh[MS, TB, Xt, #, 0]& /@
                        GenerateScales[Mtpole];
           varyQmatch = RunMSSMEFTHiggsMh[MS, TB, Xt, 0, #]& /@
                        GenerateScales[MS];
           (* combine uncertainty estimates *)
           DMh = Max[Abs[Max[varyQpole] - MhMean],
                     Abs[Min[varyQpole] - MhMean]] +
                 Max[Abs[Max[varyQmatch] - MhMean],
                     Abs[Min[varyQmatch] - MhMean]];
           { MhMean, DMh }
          ];

{Mh, DMh} = RunMSSMEFTHiggsUncertainty[2500, 20, Sqrt[6]];

Print["Mh = (", Mh, " +- ", DMh, ") GeV"];
  \end{lstlisting}
  The output of the script could read
  \begin{lstlisting}
Mh = (125.047 \pm 1.52741) GeV
  \end{lstlisting}
\end{example}

\section{Current limitations and workarounds}
\label{sec:limitations}

Currently, the models and scenarios which can be constructed with
\fstwo are limited to the following cases:
%
\begin{itemize}
\item The couplings of the model(s) must remain perturbative at all
  scales between the highest and lowest boundary condition.
\item The considered models are required to have a gauge symmetry
  that has the SM gauge group
  $\GSM = SU(3)_C\times SU(2)_L\times U(1)_Y$ as a factor.  This
  restriction is currently necessary to perform an unambiguous
  matching of the model to the $\SM(5)$ at the low-energy scale.  If a
  model that does not have $\GSM$ as a gauge group
  factor is considered, then the gauge couplings of the model must be fixed
  by hand in one of the boundary conditions.  See
  \secref{sec:determination_of_g_Y} for an example of a model with a
  left-right symmetry.
\item Tensor-like Lagrangian parameters of rank 3, which would arise
  in $R$-parity violating SUSY models for example, are currently not
  supported.  As a workaround, the rank 3 tensor-like couplings could
  be decomposed into a sum of terms with rank 2 matrix-like couplings.
\item The extraction of the running Yukawa couplings in models where
  a 4th generation of fermions mixes with the SM fermions is currently
  not supported.  As a workaround, the running Yukawa couplings can
  manually be fixed by using the running SM fermion masses, which can
  be accessed in the model file via the \code{upQuarksDRbar} =
  $\diag(m_u, m_c, m_t)$, \code{downQuarksDRbar} =
  $\diag(m_d, m_s, m_b)$ and \code{downLeptonsDRbar} =
  $\diag(m_e, \allowbreak m_\mu, \allowbreak m_\tau)$ symbols.  In the
  $\mu\nu$SSM \cite{LopezFogliani:2005yw} (\sarah/\fs model name:
  \modelname{munuSSM}), for example, the Yukawa coupling matrix $Y_e$ of
  the down-type leptons can approximately be fixed at the low-energy
  scale as
  \begin{align}
    Y_e(Q) = \frac{\sqrt{2}}{v_d} \diag(m_e, m_\mu, m_\tau) \,,
  \end{align}
  which is expressed in the \fs model file as
  \begin{lstlisting}
LowScaleInput = {
   {Ye, Sqrt[2] downLeptonsDRbar / vd},
   ...
};
  \end{lstlisting}
\end{itemize}
%
Due to the modular nature of the generated code, adaptation and extension
to overcome restrictions in scope are quite straightforward.

\section{Conclusions}
\label{sec:conclusions}

In order to study the vast zoo of models beyond the SM,
tools for each model are necessary to calculate the mass spectrum and observables.
\fs is a meta-tool for automatized generation of such tools which reliably operate at high precision and speed
for a broad class of BSM models.

In this paper, we have presented all of the substantial updates to \fs
available in version 2.0.  These include many model-specific
higher-order corrections, as well as extensions to support non-SUSY
models (\fbsm), models with complex parameters (\fcpv) and a new
solver which allows the EWSB
outputs to be defined at the high scale (\fsas).  Furthermore, \fs can
now calculate in any given
model: the anomalous magnetic moment of the muon (\famu) as well as the muon decay and the $W$ mass (\fmw),
including partial 2-loop contributions.  \fstwo also comes with an
update of the hybrid EFT/fixed-order calculation of the Higgs mass
(\feft) with a higher-order log resummation.

Altogether, these represent a significant extension to the
calculations that can be performed in BSM models at high precision. To
illustrate the variety of potential applications of \fstwo, we have
presented many physics examples.  These include large-scale parameter
scans performed efficiently on multiple CPU cores (see
\secref{sec:application_HSSUSY}), and the construction of low-energy
effective field theories of SUSY models with complicated boundary
conditions at the matching scale (see Sections
\ref{sec:application_THDM} and \ref{sec:application_HSSUSY}).  Indeed,
\fs is already being used extensively for such cases, including global
fits by the \GAMBIT collaboration and major studies of precision Higgs
mass predictions.  Furthermore, the modularity of the generated
spectrum generators allows easy implementation of model-specific
higher-order corrections, which has been done in the past to include
3-loop Higgs mass contributions from the \Himalaya library (see
\secref{sec:MSSM-specific_ho_corrections}) and to add power-suppressed
terms of $O(v^2/\MS^2)$ to \HSSUSY.
Furthermore, we have illustrated how to calculate the anomalous
magnetic moment of the muon and electric dipole moments with \fstwo in
Sections \ref{sec:famu}--\ref{sec:fcpv}.  Various physics applications
for \fsas have been presented in \secref{sec:fsas}, which include the
study of multiple solutions to the boundary value problem of the CMSSM
as well as parameter scans in the CNMSSM and C\ESSM.

The \fs system has been extensively tested for correctness against
results from the literature and other spectrum generators.  In
addition, speed tests have been carried out, with results proving its
effectiveness in large-scale scans.  The auto-generated C++ code is
designed in such a way that users can easily read and reuse
its components to develop their own analysis tools.

\section*{Acknowledgments}

We especially thank Florian Staub and Mark Goodsell for many helpful
discussions and technical help with \sarah.  Furthermore, we kindly
thank Ben Allanach for providing components of \softsusy, which are
included in \fs.  We also thank Pietro Slavich and Emanuele Bagnaschi
for providing the 2-loop contributions to the Higgs mass(es) in the
full MSSM and in \HSSUSY.
%
Furthermore, we thank Ben
Allanach and Alexander Bednyakov for providing the 2-loop \SQCD
contributions to the \DRbar top, bottom and $\tau$ Yukawa couplings as
well as to the strong coupling in the MSSM and both Alexander
Bednyakov and the authors of \susyhd for providing the 3-loop SM $\beta$
functions.  We also thank the authors of the \Himalaya
package for providing 3-loop contributions to the light \CP-even Higgs mass
in the MSSM\@.
%
We kindly thank and acknowledge the role of Tom Steudtner for writing
the first version of \feft.
%
A.V.\ is very grateful to the KUTS working group, especially Emanuele
Bagnaschi, Henning Bahl, Johannes Braathen, Mark Goodsell, Robert
Harlander, Luminita Mihaila, Pietro Slavich, Florian Staub and Georg
Weiglein, for countless insights and helpful discussions about Higgs
mass precision predictions in supersymmetry over the last years.
%
P.A.\ is very grateful to his \GAMBIT collaborators, especially Ben
Farmer and James McKay, for helpful suggestions and feedback on
\fs. P.A., A.V.\ and D.H.\ thank their collaborators on the
``Precision tools and models to narrow in on the $750\unit{GeV}$
diphoton resonance'' project for helpful comments and advice that have
influenced the development of \fs.  Similarly, P.A.\ and A.V.\ also
thank their collaborators in the ``Higgs mass predictions of public
NMSSM spectrum generators'' project for the helpful discussions during
that project.  P.A.\ and D.H.\ warmly thank Roman Nevzorov for many
helpful discussions about the semi-analytic solver approach and
collaboration with him which had a lot of impact on the development of
this code.  The work of P.A.\ and D.H.\ was supported by the Australian
Research Council through the ARC Centre of Excellence for Particle Physics
at the Terascale (CoEPP) (grant CE110001104).  The work of P.A. is also
supported by the Australian Research Council Future Fellowship grant
FT160100274.  The work of D.H.\ was supported by the University of Adelaide,
and through an Australian Government Research Training Program Scholarship.
D.H.\ also acknowledges financial support from the Grant Agency of the Czech
Republic (GACR), contract 17-04902S.  This work has also been supported by the
German Research Foundation DFG through Grant No.~STO876/2.

\appendix
\renewcommand*{\thesection}{\Alph{section}}

\section{SM input parameters}
\label{app:sm_input_parameters}
%
The list of all SM input parameters is given
in \tabref{tab:sm_mma_options}.  In the SLHA interface of \fs, the
SM parameters can be given in four different input blocks:
%
\begin{itemize}
\item The block \texttt{SMINPUTS} contains the electromagnetic and
  strong coupling, the Fermi constant and the masses of SM
  particles, as defined in the SLHA-2 standard \cite{Allanach:2008qq}.
  The definition of the individual block entries are shown in the
  first part of \tabref{tab:sm_mma_options}.  If a parameter is
  omitted from the \texttt{SMINPUTS} block, then it is set to the
  default value defined in \tabref{tab:sm_mma_options}.
\item The CKM matrix is given as input in the block \texttt{VCKMIN} in
  the Wolfenstein parametrization, as defined in
  Ref.~\cite{Allanach:2008qq}.  If the \texttt{VCKMIN} block or an entry is
  missing, the corresponding parameter is set to zero.
\item The PMNS matrix is given as input in the block
  \texttt{UPMNSIN}, as defined in Ref.~\cite{Allanach:2008qq}.  If the
  \texttt{UPMNSIN} block or an entry is missing, the corresponding
  parameter is set to zero.
\item For special applications such as the calculation of
  $\amu$ with \GMTCalc, further input parameters are
  needed.  These can be given in the \texttt{FlexibleSUSYInput} block.
  The block entries are defined in \tabref{tab:sm_mma_options}.  If
  the block or a block entry is missing, the input parameters are set
  to their default values as defined in the table.
\end{itemize}
%
In SLHA format the blocks with their respective default values read:
%
\begin{lstlisting}
Block SMINPUTS               # Standard Model inputs
    1   1.279160000e+02      # alpha^(-1) SM MSbar(MZ)
    2   1.166378700e-05      # G_Fermi
    3   1.184000000e-01      # alpha_s(MZ) SM MSbar
    4   9.118760000e+01      # MZ(pole)
    5   4.180000000e+00      # mb(mb) SM MSbar
    6   1.733400000e+02      # mtop(pole)
    7   1.776990000e+00      # mtau(pole)
    8   0.000000000e+00      # mnu3(pole)
    9   8.038500000e+01      # MW pole
   11   5.109989020e-04      # melectron(pole)
   12   0.000000000e+00      # mnu1(pole)
   13   1.056583570e-01      # mmuon(pole)
   14   0.000000000e+00      # mnu2(pole)
   21   4.750000000e-03      # md(2 GeV) MS-bar
   22   2.400000000e-03      # mu(2 GeV) MS-bar
   23   1.040000000e-01      # ms(2 GeV) MS-bar
   24   1.270000000e+00      # mc(mc) MS-bar
Block VCKMIN                 # CKM matrix input (Wolfenstein parameters)
    1                 0      # lambda(MZ) SM DR-bar
    2                 0      # A(MZ) SM DR-bar
    3                 0      # rhobar(MZ) SM DR-bar
    4                 0      # etabar(MZ) SM DR-bar
Block UPMNSIN                # PMNS matrix input
    1                 0      # theta_12
    2                 0      # theta_23
    3                 0      # theta_13
    4                 0      # delta
    5                 0      # alpha_1
    6                 0      # alpha_2
Block FlexibleSUSYInput
    0   0.00729735           # alpha_em(0)
    1   125.09               # Mh pole
\end{lstlisting}

In \fs's \mathematica interface, the SM parameters must be
passed to the \code{FS<model>OpenHandle[fsSMParameters -> \{...\}]}
function in the form of replacement rules.  The symbols associated to the
SM input parameters are given in
\tabref{tab:sm_mma_options}.  Unset parameters are set to their
default values defined in the table.  Note that in the \mathematica
interface of \fs, the CKM matrix parameters are given in the
exact parametrization in terms of the angles $\theta_{12}$,
$\theta_{23}$, $\theta_{13}$ and $\delta$ \cite{Olive:2016xmw}.  A
call of \code{FS<model>OpenHandle[]} with all parameters set to their
respective default values would read:
%
\begin{lstlisting}{Mathematica}
handle = FS<model>OpenHandle[
    fsSMParameters -> {
        alphaEmMZ -> 1/127.916,      (* SMINPUTS[1] *)
        GF -> 1.166378700*^-5,       (* SMINPUTS[2] *)
        alphaSMZ -> 0.1184,          (* SMINPUTS[3] *)
        MZ -> 91.1876,               (* SMINPUTS[4] *)
        mbmb -> 4.18,                (* SMINPUTS[5] *)
        Mt -> 173.34,                (* SMINPUTS[6] *)
        Mtau -> 1.77699,             (* SMINPUTS[7] *)
        Mv3 -> 0,                    (* SMINPUTS[8] *)
        MW -> 80.385,                (* SMINPUTS[9] *)
        Me -> 0.000510998902,        (* SMINPUTS[11] *)
        Mv1 -> 0,                    (* SMINPUTS[12] *)
        Mm -> 0.1056583715,          (* SMINPUTS[13] *)
        Mv2 -> 0,                    (* SMINPUTS[14] *)
        md2GeV -> 0.00475,           (* SMINPUTS[21] *)
        mu2GeV -> 0.0024,            (* SMINPUTS[22] *)
        ms2GeV -> 0.104,             (* SMINPUTS[23] *)
        mcmc -> 1.27,                (* SMINPUTS[24] *)
        CKMTheta12 -> 0,
        CKMTheta23 -> 0,
        CKMTheta13 -> 0,
        CKMDelta -> 0,
        PMNSTheta12 -> 0,            (* UPMNSIN[1] *)
        PMNSTheta23 -> 0,            (* UPMNSIN[2] *)
        PMNSTheta13 -> 0,            (* UPMNSIN[3] *)
        PMNSDelta -> 0,              (* UPMNSIN[4] *)
        PMNSAlpha1 -> 0,             (* UPMNSIN[5] *)
        PMNSAlpha2 -> 0,             (* UPMNSIN[6] *)
        alphaEm0 -> 1/137.035999074, (* FlexibleSUSYInput[0] *)
        Mh -> 125.09                 (* FlexibleSUSYInput[1] *)
    }
]
\end{lstlisting}
%
\begin{table}[tbh!]
  \centering
  \resizebox{\textwidth}{!}{%
  \begin{tabular}{llll}
    \toprule
    Index & \mathematica symbol & Default & Description \\
    \midrule
    \multicolumn{4}{c}{\texttt{Block SMINPUTS}}\\
    \midrule
     1 & \code{alphaEmMZ} & $1/127.916$ & electromagnetic coupling, $\aem^{\SM(5)}(M_Z)$ \\
     2 & \code{GF} & $1.1663787\cdot 10^{-5}$ & Fermi coupling constant, $G_F\times\text{GeV}^2$ \\
     3 & \code{alphaSMZ} & $0.1184$ & strong coupling, $\as^{\SM(5)}(M_Z)$ \\
     4 & \code{MZ} & $91.1876$ & $Z$ pole mass, $M_Z/\text{GeV}$ \\
     5 & \code{mbmb} & $4.18$ & running bottom mass, $m_b^{\SM(5)}(m_b)/\text{GeV}$ \\
     6 & \code{Mt} & $173.34$ & top pole mass, $M_t/\text{GeV}$ \\
     7 & \code{Mtau} & $1.77699$ & $\tau$ pole mass, $M_\tau/\text{GeV}$ \\
     8 & \code{Mv3} & $0$ & heaviest neutrino pole mass, $M_{\nu_3}/\text{GeV}$ \\
     9 & \code{MW} & $80.385$ & $W$ pole mass, $M_W/\text{GeV}$ \\
    11 & \code{Me} & $0.000510998902$ & electron pole mass, $M_e/\text{GeV}$ \\
    12 & \code{Mv1} & $0$ & lightest neutrino pole mass, $M_{\nu_1}/\text{GeV}$ \\
    13 & \code{Mm} & $0.1056583715$ & muon pole mass, $M_\mu/\text{GeV}$ \\
    14 & \code{Mv2} & $0$ & 2nd lightest neutrino pole mass, $M_{\nu_2}/\text{GeV}$ \\
    21 & \code{md2GeV} & $0.00475$ & running down mass, $m_d(2\unit{GeV})/\text{GeV}$ \\
    22 & \code{mu2GeV} & $0.0024$ & running up mass, $m_u(2\unit{GeV})/\text{GeV}$ \\
    23 & \code{ms2GeV} & $0.104$ & running strange mass, $m_s(2\unit{GeV})/\text{GeV}$ \\
    24 & \code{mcmc} & $1.27$ & running charm mass, $m_c^{\SM(4)}(m_c)/\text{GeV}$ \\
    \midrule
    \multicolumn{4}{c}{\texttt{Block VCKMIN}}\\
    \midrule
     1 & & $0$ & CKM Wolfenstein parameter $\lambda$ \\
     2 & & $0$ & CKM Wolfenstein parameter $A$ \\
     3 & & $0$ & CKM Wolfenstein parameter $\bar{\rho}$ \\
     4 & & $0$ & CKM Wolfenstein parameter $\eta$ \\
    \midrule
       & \code{CKMTheta12} & $0$ & CKM matrix parameter $\theta_{12}$ \\
       & \code{CKMTheta23} & $0$ & CKM matrix parameter $\theta_{23}$ \\
       & \code{CKMTheta13} & $0$ & CKM matrix parameter $\theta_{13}$ \\
       & \code{CKMDelta} & $0$ & CKM matrix parameter $\delta$ \\
    \midrule
    \multicolumn{4}{c}{\texttt{Block UPMNSIN}}\\
    \midrule
     1 & \code{PMNSTheta12} & $0$ & PMNS solar angle $\theta_{12}$ \\
     2 & \code{PMNSTheta23} & $0$ & PMNS atmospheric angle $\theta_{23}$ \\
     3 & \code{PMNSTheta13} & $0$ & PMNS matrix parameter $\theta_{13}$ \\
     4 & \code{PMNSDelta}   & $0$ & PMNS Dirac phase $\delta$ \\
     5 & \code{PMNSAlpha1}  & $0$ & PMNS 1$^\text{st}$ Majorana phase $\alpha_1$ \\
     6 & \code{PMNSAlpha2}  & $0$ & PMNS 2$^\text{nd}$ Majorana phase $\alpha_2$ \\
    \midrule
    \multicolumn{4}{c}{\texttt{Block FlexibleSUSYInput}}\\
    \midrule
     0 & \code{alphaEm0} & $1/137.035999074$ & $\aem$ in the Thomson limit \\
     1 & \code{Mh} & $125.09$ & SM Higgs pole mass $M_h/\text{GeV}$ \\
    \bottomrule
  \end{tabular}}
  \caption{SLHA input block entries and \mathematica symbols to
    specify the SM input parameters.  The first column
    represents the index in the corresponding SLHA input block and the
    second column the symbol used in the \mathematica interface.}
  \label{tab:sm_mma_options}
\end{table}

\section{FlexibleSUSY configuration options}
\label{app:fs_config_options}

\fs provides many configuration options to switch on/off contributions
and choose/fine-tune the solver algorithm(s).  All runtime
configuration options are listed in \tabref{tab:sg_mma_options}.
%
In the SLHA interface of \fs, the configuration options are read from
the \fs block.  In addition, some information is also read
from the \texttt{MODSEL} block, see below.  In the SLHA format all \fs
configuration blocks with their respective default values read:
%
\begin{lstlisting}
Block MODSEL
   12   0            # output scale of running parameters (0 = SUSY scale)
Block FlexibleSUSY
    0   1e-04        # precision goal
    1   0            # max. iterations (0 = automatic)
    2   0            # solver (0 = all, 1 = two-scale, 2 = semi-analytic)
    3   0            # calculate SM pole masses
    4   2            # pole mass loop order
    5   2            # EWSB loop order
    6   3            # beta-functions loop order
    7   2            # threshold corrections loop order
    8   1            # Higgs 2L corrections O(alpha_t alpha_s)
    9   1            # Higgs 2L corrections O(alpha_b alpha_s)
   10   1            # Higgs 2L corrections O((alpha_t + alpha_b)^2)
   11   1            # Higgs 2L corrections O(alpha_tau^2)
   12   0            # force output
   13   1            # Top quark 2L corrections QCD
   14   1e-11        # beta-function zero threshold
   15   0            # calculate observables (a_muon, ...)
   16   0            # force positive majorana masses
   17   0            # pole mass renormalization scale (0 = SUSY scale)
   18   0            # pole mass renormalization scale in the EFT
                     # (0 = min(SUSY scale, Mt))
   19   0            # EFT matching scale (0 = SUSY scale)
   20   2            # EFT loop order for upwards matching
   21   1            # EFT loop order for downwards matching
   22   0            # EFT index of SM-like Higgs in the BSM model
   23   1            # calculate BSM pole masses
   24   123111321    # individual threshold correction loop orders
   25   0            # ren. scheme for Higgs 3L corrections
                     # (0 = DR, 1 = MDR)
   26   1            # Higgs 3L corrections O(alpha_t alpha_s^2)
   27   1            # Higgs 3L corrections O(alpha_b alpha_s^2)
   28   1            # Higgs 3L corrections O(alpha_t^2 alpha_s)
   29   1            # Higgs 3L corrections O(alpha_t^3)
   28   1            # Higgs 3L corrections O(alpha_t^2 alpha_s)
   29   1            # Higgs 3L corrections O(alpha_t^3)
\end{lstlisting}
%
In the \mathematica interface of \fs, the configuration options are
passed to the function \code{FS<model>OpenHandle[fsSettings ->
  \{...\}]} in form of replacement rules.  The symbols associated to
the configuration options are given in \tabref{tab:sg_mma_options}.
Unset parameters are set to their default values defined in the table.
A call of \code{FS<model>OpenHandle[]} with all configuration options
set to their default values would read:
%
\begin{lstlisting}
handle = FS<model>OpenHandle[
    fsSettings -> {
        precisionGoal -> 1.*^-4,           (* FlexibleSUSY[0] *)
        maxIterations -> 0,                (* FlexibleSUSY[1] *)
        solver -> 0,                       (* FlexibleSUSY[2] *)
        calculateStandardModelMasses -> 0, (* FlexibleSUSY[3] *)
        poleMassLoopOrder -> 2,            (* FlexibleSUSY[4] *)
        ewsbLoopOrder -> 2,                (* FlexibleSUSY[5] *)
        betaFunctionLoopOrder -> 3,        (* FlexibleSUSY[6] *)
        thresholdCorrectionsLoopOrder -> 2,(* FlexibleSUSY[7] *)
        higgs2loopCorrectionAtAs -> 1,     (* FlexibleSUSY[8] *)
        higgs2loopCorrectionAbAs -> 1,     (* FlexibleSUSY[9] *)
        higgs2loopCorrectionAtAt -> 1,     (* FlexibleSUSY[10] *)
        higgs2loopCorrectionAtauAtau -> 1, (* FlexibleSUSY[11] *)
        forceOutput -> 0,                  (* FlexibleSUSY[12] *)
        topPoleQCDCorrections -> 1,        (* FlexibleSUSY[13] *)
        betaZeroThreshold -> 1.*^-11,      (* FlexibleSUSY[14] *)
        forcePositiveMasses -> 0,          (* FlexibleSUSY[16] *)
        poleMassScale -> 0,                (* FlexibleSUSY[17] *)
        eftPoleMassScale -> 0,             (* FlexibleSUSY[18] *)
        eftMatchingScale -> 0,             (* FlexibleSUSY[19] *)
        eftMatchingLoopOrderUp -> 2,       (* FlexibleSUSY[20] *)
        eftMatchingLoopOrderDown -> 1,     (* FlexibleSUSY[21] *)
        eftHiggsIndex -> 0,                (* FlexibleSUSY[22] *)
        calculateBSMMasses -> 1,           (* FlexibleSUSY[23] *)
        thresholdCorrections -> 123111321, (* FlexibleSUSY[24] *)
        higgs3loopCorrectionRenScheme -> 0,(* FlexibleSUSY[25] *)
        higgs3loopCorrectionAtAsAs -> 1,   (* FlexibleSUSY[26] *)
        higgs3loopCorrectionAbAsAs -> 1,   (* FlexibleSUSY[27] *)
        higgs3loopCorrectionAtAtAs -> 1,   (* FlexibleSUSY[28] *)
        higgs3loopCorrectionAtAtAt -> 1,   (* FlexibleSUSY[29] *)
        parameterOutputScale -> 0          (* MODSEL[12] *)
    }
]
\end{lstlisting}
%
\begin{table}
  \centering
  \resizebox{\textwidth}{!}{%
  \begin{tabularx}{1.2\textwidth}{lllX}
    \toprule
    Index & \mathematica symbol & Default & Description \\
    \midrule
    \multicolumn{4}{c}{\texttt{Block FlexibleSUSY}}\\
    \midrule
     0 & \code{precisionGoal}                 & \code{1.*^-4} &
    precision goal of RG running and mass spectrum\\
     1 & \code{maxIterations}                 & \code{0} &
    maximum number of iterations for the running between the scales ($0$ = automatic)\\
     2 & \code{solver}                        & \code{0} &
    BVP solver ($0$ = all, $1$ = two-scale solver, $2$ = semi-analytic solver)\\
     3 & \code{calculateStandardModelMasses}  & \code{0} &
    switch to enable/disable calculation of pole masses of SM particles ($0$ = disabled)\\
     4 & \code{poleMassLoopOrder}             & \code{2} &
    pole mass loop order\\
     5 & \code{ewsbLoopOrder}                 & \code{2} &
    EWSB loop order (should be set equal to the pole mass loop order)\\
     6 & \code{betaFunctionLoopOrder}         & \code{3} &
    loop order for renormalization group running\\
     7 & \code{thresholdCorrectionsLoopOrder} & \code{2} &
    global switch for loop order of threshold corrections when
    converting the $\SM(5)$ parameters to the BSM parameters\\
     8 & \code{higgs2loopCorrectionAtAs}      & \code{1} &
    enable/disable 2-loop corrections $O(\at\as)$ to $M_{h,H,A}$\\
     9 & \code{higgs2loopCorrectionAbAs}      & \code{1} &
    enable/disable 2-loop corrections $O(\ab\as)$ to $M_{h,H,A}$\\
    10 & \code{higgs2loopCorrectionAtAt}      & \code{1} &
    enable/disable 2-loop corrections $O(\at^2)$ to $M_{h,H,A}$\\
    11 & \code{higgs2loopCorrectionAtauAtau}  & \code{1} &
    enable/disable 2-loop corrections $O(\atau\atau)$ to $M_{h,H,A}$\\
    12 & \code{forceOutput}                   & \code{0} &
    force output, even if problems occurred\\
    13 & \code{topPoleQCDCorrections}         & \code{1} &
    QCD corrections to calculate $M_t$ (0 = 1-loop, 1 = 2-loop, 2 = 3-loop)\\
    14 & \code{betaZeroThreshold}             & \code{1.*^-11} &
    below this threshold $\beta$ functions are treated as zero\\
    15 & & \code{0} & enable/disable calculation of observables\\
    16 & \code{forcePositiveMasses}           & \code{0} &
    make Majorana masses positive (violates SLHA)\\
    17 & \code{poleMassScale}                 & \code{0} &
    scale at which pole masses are calculated (0 = SUSY scale)\\
    18 & \code{eftPoleMassScale}              & \code{0} &
    scale at which SM pole masses are calculated in \feft (0 = $M_t$)\\
    19 & \code{eftMatchingScale}              & \code{0} &
    matching scale in \feft (0 = SUSY scale)\\
    20 & \code{eftMatchingLoopOrderUp}        & -- &
    ignored\\
    21 & \code{eftMatchingLoopOrderDown}      & \code{1} &
    loop order for $\Delta\lambda$ in \feft\\
    22 & \code{eftHiggsIndex}                 & \code{0} &
    index of SM-like Higgs in BSM Higgs multiplet in \feft\\
    23 & \code{calculateBSMMasses}            & \code{1} &
    enable/disable calculation of BSM pole masses\\
    24 & \code{thresholdCorrections}          & \code{123111321} &
    individual threshold correction loop orders, see \tabref{tab:sg_mma_tc}\\
    25 & \code{higgs3loopCorrectionRenScheme} & \code{0} &
    renormalization scheme for 3-loop MSSM Higgs corrections:
    0 = \DRbar, 1 = \MDRbar\\
    26 & \code{higgs3loopCorrectionAtAsAs}    & \code{1} &
    enable/disable 3-loop corrections $O(\at\as^2)$ to $M_{h,H,A}$\\
    27 & \code{higgs3loopCorrectionAbAsAs}    & \code{1} &
    enable/disable 3-loop corrections $O(\ab\as^2)$ to $M_{h,H,A}$\\
    28 & \code{higgs3loopCorrectionAtAtAs}    & \code{1} &
    enable/disable 3-loop corrections $O(\at^2\as)$ to $M_{h}$\\
    29 & \code{higgs3loopCorrectionAtAtAt}    & \code{1} &
    enable/disable 3-loop corrections $O(\at^3)$ to $M_{h}$\\
    \midrule
    \multicolumn{4}{c}{\texttt{Block MODSEL}}\\
    \midrule
    12 & \code{parameterOutputScale}          & \code{0} &
    output scale for running parameters (0 = SUSY scale)\\
    \bottomrule
  \end{tabularx}}
  \caption{SLHA input block entries and corresponding \mathematica
    symbols to specify the configuration options for \fs's spectrum
    generators.}
  \label{tab:sg_mma_options}
\end{table}
%
\begin{table}[tbh]
  \centering
  \begin{tabular}{lll}
    \toprule
    digit position $n$    & default value & parameter \\
    (from the right) & (prefactor of $10^n$) & \\
    \midrule
    $0$ & $1$ (1-loop) & $\alpha^\BSM_{\text{em}}$ \\
    $1$ & $2$ (2-loop) & $\sin^\BSM(\theta_W)$ \\
    $2$ & $3$ (3-loop) & $\as^\BSM$ \\
    $3$ & $1$ (1-loop) & $m^\BSM_Z$ \\
    $4$ & $1$ (1-loop) & $m^\BSM_W$ \\
    $5$ & $1$ (1-loop) & $m^\BSM_h$ \\
    $6$ & $3$ (3-loop) & $m^\BSM_t$ \\
    $7$ & $2$ (2-loop) & $m^\BSM_b$ \\
    $8$ & $1$ (1-loop) & $m^\BSM_\tau$ \\
    \bottomrule
  \end{tabular}
  \caption{Specification of individual loop orders of threshold
    corrections for the model parameters at the low-energy scale using
    the field \texttt{FlexibleSUSY[24]} in the SLHA interface or the
    symbol \texttt{thresholdCorrections} in the \mathematica
    interface, respectively.  The digit position is counted from the
    right, starting at $0$.  Setting all loop orders to their default
    values results in the integer $123111321$.}
  \label{tab:sg_mma_tc}
\end{table}
%
The individual configuration options have the following meaning:
%
\begin{itemize}
\item[\texttt{FlexibleSUSY[0]},] \texttt{precisionGoal}: This option
  describes the numeric precision of the renormalization group
  running, the mass spectrum calculation, the electroweak symmetry
  breaking and the calculation of the observables.  For most models a
  precision of $10^{-4}$ is sufficient.  For models with various
  3-loop corrections, like \HSSUSY or an MSSM spectrum generator, a
  precision of $10^{-5}$ might be better.

\item[\texttt{FlexibleSUSY[1]},] \texttt{maxIterations}: This option
  describes the maximum number of iterations for the renormalization
  group running between the various scales.  If it is set to $0$, then
  the maximum number of iterations $N_{\text{max,it}}$ is chosen
  according to the precision goal $p$ (see above) as
  \begin{align}
    N_{\text{max,it}} = - 10\log_{10} p \,.
  \end{align}

\item[\texttt{FlexibleSUSY[2]},] \texttt{solver}: This option chooses
  the BVP solver to be used.  If set to $0$, all
  solvers that have been enabled in the model file (see
  \secref{sec:chosing_SAS}) are used, starting with the two-scale
  solver.  If set to $1$, the two-scale solver is used, if it has been
  enabled in the model file.  If set to $2$, then the semi-analytic
  solver is used, if it has been enabled in the model file.  If a
  solver that has not been enabled in the model file is chosen, \fs
  stops with an error.

\item[\texttt{FlexibleSUSY[3]},]
  \texttt{calculateStandardModelMasses}: This option allows the user
  to enable/disable the calculation of the pole masses of the SM
  particles.  Note that this switch must be set to $1$ in
  \HSSUSY to calculate the Higgs pole mass, because in \HSSUSY the
  Higgs pole mass is calculated in the SM\@.

\item[\texttt{FlexibleSUSY[4]},] \texttt{poleMassLoopOrder}: This
  option allows the user to select the loop order at which the pole
  masses are calculated.  If set to $0$, the running tree-level masses
  are output.  If set to $1$, the pole masses are calculated at the
  1-loop level.  If set to $2$ or $3$, then model-specific 2-loop or
  3-loop corrections are taken into account, respectively, if they are
  enabled in the model file (see
  \secref{sec:model_specific_contributions}).  \textbf{Important
    note:} In order to obtain a consistent pole mass spectrum, the
  loop order of the electroweak symmetry breaking (see
  \texttt{FlexibleSUSY[5]}, \texttt{ewsbLoopOrder}) must be set to the
  \emph{same} value as the pole mass loop order!

\item[\texttt{FlexibleSUSY[5]},] \texttt{ewsbLoopOrder}: This option
  allows the user to select the loop order at which the electroweak
  symmetry breaking (EWSB) equations are solved.  If set to $0$, the
  EWSB equations are solved at the tree-level.  If set to $1$, the
  EWSB equations are solved at the 1-loop level.  If set to $2$ or
  $3$, then model-specific 2-loop or 3-loop corrections are taken into
  account, respectively, if they are enabled in the model file (see
  \secref{sec:model_specific_contributions}).  \textbf{Important
    note:} In order to obtain a consistent pole mass spectrum, the
  loop order of the electroweak symmetry breaking must be set to the
  \emph{same} value as the pole mass loop order (see
  \texttt{FlexibleSUSY[4]}, \texttt{poleMassLoopOrder})!

\item[\texttt{FlexibleSUSY[6]},] \texttt{betaFunctionLoopOrder}: With
  this option the user can select the loop level of the $\beta$
  functions used to integrate the RGEs.  If set to $1$, 1-loop $\beta$
  functions are used.  If set to $2$, 2-loop $\beta$ functions are used.  If
  set to $3$, then model-specific 3-loop $\beta$ functions are used (see
  \secref{sec:model_specific_contributions}).  Note that \sarah can
  generate 2-loop $\beta$ functions for all model parameters (except
  scalar tadpole terms), so 2-loop running can always be used.

\item[\texttt{FlexibleSUSY[7]},]
  \texttt{thresholdCorrectionsLoopOrder}: With this option the user
  can choose the \emph{maximum} loop level of the threshold corrections
  used to determine the running gauge couplings $g_1$, $g_2$, $g_3$
  and the running Yukawa coupling matrices $Y_u$, $Y_d$, $Y_e$ of the
  \BSM model at the low-energy scale (\code{LowScale}) from the given
  SM input parameters ($\aem^{\SM(5)}(M_Z)$,
  $\as^{\SM(5)}(M_Z)$, $G_F$, $M_Z$, $M_t$, $m_b^{\SM(5)}(m_b)$,
  $m_c^{\SM(4)}(m_c)$, \ldots).  See \secref{sec:determination_of_g_Y}
  and \ref{sec:model_specific_contributions} for a description on how
  the running gauge and Yukawa couplings are calculated and how
  model-specific higher-order corrections can be included.  If the
  threshold corrections loop order is set to $1$, then no 2-loop
  threshold corrections or higher are taken into account.  If set to
  $2$, then no 3-loop threshold corrections or higher are taken into
  account.  If set to $3$, then no 4-loop threshold corrections are
  taken into account.
  %
  Note that threshold corrections for individual parameters can be
  disabled by using \texttt{FlexibleSUSY[24]} or
  \texttt{thresholdCorrections}, respectively.

\item[\texttt{FlexibleSUSY[8]},] \texttt{higgs2loopCorrectionAtAs}:
  With this option the 2-loop contributions to the Higgs pole mass(es)
  of $O(\at\as)$ can be enabled/disabled.  Note that this
  option has an effect only if 2-loop contributions have been
  activated in the \fs model file.  See
  \secref{sec:model_specific_contributions} for details on how to
  activate 2-loop contributions to the Higgs pole mass(es) of
  $O(\at\as)$ in the SM, (N)MSSM or split-MSSM\@.

\item[\texttt{FlexibleSUSY[9]},] \texttt{higgs2loopCorrectionAbAs}:
  With this option the 2-loop contributions to the Higgs pole mass(es)
  of $O(\ab\as)$ can be enabled/disabled.  Note that this
  option has an effect only if 2-loop contributions have been
  activated in the \fs model file.  See
  \secref{sec:model_specific_contributions} for details on how to
  activate 2-loop contributions of $O(\ab\as)$ to the Higgs pole
  mass(es) in the (N)MSSM\@.

\item[\texttt{FlexibleSUSY[10]},] \texttt{higgs2loopCorrectionAtAt}:
  With this option the 2-loop contributions to the Higgs pole mass(es)
  of $O(\at^2)$ or $O((\at+\ab)^2)$ can be enabled/disabled.
  Note that this option has an effect only if 2-loop contributions
  have been activated in the \fs model file.  See
  \secref{sec:model_specific_contributions} for details on how to
  activate 2-loop contributions of these orders to the Higgs pole
  mass(es) in the SM or (N)MSSM\@.

\item[\texttt{FlexibleSUSY[11]},]
  \texttt{higgs2loopCorrectionAtauAtau}: With this option the 2-loop
  contributions to the Higgs pole mass(es) of $O(\atau^2)$
  can be enabled/disabled.  Note that this option has an effect only
  if 2-loop contributions have been activated in the \fs model file.
  See \secref{sec:model_specific_contributions} for details on how to
  activate 2-loop contributions of $O(\atau^2)$ to the Higgs pole
  mass(es) in the (N)MSSM\@.

\item[\texttt{FlexibleSUSY[12]},] \texttt{forceOutput}: This option
  allows the user to force an output of \fs, even if a physical
  problem has occurred (tachyon, non-perturbative parameter, no EWSB,
  \ldots).  If set to $0$, \fs does not give an output if a problem
  has occurred.  If set to $1$, an output is always given, even if a
  problem has occurred.  Please be very careful and check for
  potential warnings/problems when forcing the output!

\item[\texttt{FlexibleSUSY[13]},] \texttt{topPoleQCDCorrections}: With
  this option the user can enable additional loop contributions when
  the top quark \emph{pole mass} is re-calculated.  Note that the top pole mass
  is only re-calculated if \texttt{FlexibleSUSY[3]} or
  \texttt{calculateStandardModelMasses} is set to $1$.  If set to
  $0$, then 1-loop (SUSY-)QCD contributions are taken into account (but
  only if \texttt{FlexibleSUSY[4]} or \texttt{poleMassLoopOrder} is
  set to $1$).  If set to $1$, then 2-loop (SUSY-)QCD contributions are
  taken into account (but only if \texttt{FlexibleSUSY[4]} or
  \texttt{poleMassLoopOrder} is set to $2$).  If set to $2$, then
  3-loop (SUSY-)QCD contributions are taken into account (but only if
  \texttt{FlexibleSUSY[4]} or \texttt{poleMassLoopOrder} is set to
  $3$).

\item[\texttt{FlexibleSUSY[14]},] \texttt{betaZeroThreshold}: With
  this option a numerical threshold can be defined below which a
  $\beta$ function is treated as being exactly zero.  A small but
  non-zero threshold can avoid numerical problems when integrating the
  RGEs.

\item[\texttt{FlexibleSUSY[15]}:] With this option the calculation of
  the observables ($\amu$, EDMs, effective couplings of
  $h\rightarrow\gamma\gamma$ and $h\rightarrow gg$) can be
  enabled/disabled.  Note that in the \mathematica interface, the
  observables are calculated by the function
  \code{FS<model>CalculateObservables[]}, see
  \secref{sec:Mathematica_interface}.

\item[\texttt{FlexibleSUSY[16]},] \texttt{forcePositiveMasses}: With
  this option, the masses of Majorana fermions can be forced to be
  positive in the SLHA output of \fs.  If set to $1$, then Majorana
  masses are always positive, but the corresponding mixing matrices
  are in general complex (note that this violates the SLHA
  convention).  If set to $0$, then the Majorana masses can be
  positive or negative, but the corresponding mixing matrices are
  guaranteed to be real (SLHA convention).

\item[\texttt{FlexibleSUSY[17]},] \texttt{poleMassScale}: With this
  option the user can choose the scale (in GeV) at which the pole mass
  spectrum is calculated.  If set to $0$, then the value assigned to
  the \code{SUSYScale} variable in the model file is used.  In \feft,
  the pole mass scale is defined to be the scale at which the pole
  masses in the \emph{full} \BSM model are calculated.  To vary the
  scale at which the pole masses in the effective theory (the
  SM) are calculated, use \texttt{FlexibleSUSY[18]} or
  \texttt{eftPoleMassScale} in the SLHA or Mathematica interfaces,
  respectively.

\item[\texttt{FlexibleSUSY[18]},] \texttt{eftPoleMassScale}: This
  option applies only to \feft models.  With this option the user can
  choose the scale (in GeV) at which the pole masses in the effective
  field theory (i.e., in the SM) are calculated.  If the
  scale is set to $0$, then $Q = M_t$ is used.  This option can be
  used to estimate a partial uncertainty of the Higgs mass
  prediction in \feft by varying the pole mass scale around $Q = M_t$,
  see \exref{ex:MSSMEFTHiggs_uncertainty_estimate}.

\item[\texttt{FlexibleSUSY[19]},] \texttt{eftMatchingScale}: This
  option applies only to \feft models.  With this option the user can
  specify the scale at which the matching of the \BSM model to the
  effective theory (i.e., the SM) is performed.  If the
  scale is set to $0$, then the value assigned to the \code{SUSYScale}
  variable in the \fs model file is used.  This option can be used to
  estimate a partial uncertainty of the Higgs mass prediction
  in \feft by varying the matching scale around $Q = \MS$, see
  \exref{ex:MSSMEFTHiggs_uncertainty_estimate}.

\item[\texttt{FlexibleSUSY[20]},] \texttt{eftMatchingLoopOrderUp}:
  This option is ignored in \fstwo.

\item[\texttt{FlexibleSUSY[21]},] \texttt{eftMatchingLoopOrderDown}:
  This option applies only to \feft models.  With this option the user
  can select the loop order at which the quartic Higgs coupling
  $\lambda$ of the SM is fixed when matching the \BSM
  model to the SM in \feft.  If set to $0$, then $\lambda$
  is fixed using only tree-level matching.  If set to $1$
  (recommended), then $\lambda$ is fixed by a 1-loop matching
  condition.

\item[\texttt{FlexibleSUSY[22]},] \texttt{eftHiggsIndex}: This option
  applies only to \feft models.  With this option the user can choose
  which field in the Higgs multiplet of the \BSM model corresponds to
  the SM-like Higgs.  If set to $0$, the lightest field in the Higgs
  multiplet is interpreted as SM-like Higgs.  If set to $1$, the
  2$^{\text{nd}}$ lightest field is interpreted as SM-like Higgs, etc.
  The chosen field is then used in the matching condition
  $M_h^{\SM} = M_{h_i}^{\BSM}$, where $i$ is the index of the chosen
  field in the Higgs multiplet ($i=0,1,2,\ldots$).

\item[\texttt{FlexibleSUSY[23]},] \texttt{calculateBSMMasses}: This
  option allows the user to enable/disable the calculation of the pole
  masses of the \BSM particles.  If set to $0$, then the \BSM pole
  masses are not calculated.  If set to $1$, then the \BSM pole masses
  are calculated.  This option is useful in \feft for example: If one
  is only interested in the prediction of the SM-like Higgs pole mass,
  then this option can be set to $0$ to suppress the calculation of
  the heavy Higgs bosons, the charginos, neutralinos and sfermions.

\item[\texttt{FlexibleSUSY[24]},] \texttt{thresholdCorrections}: With
  this option the user has a finer control over the threshold
  corrections to the individual model parameters.  The value assigned
  to this option is an integer number, where each digit (with respect to base
  $10$) represents the threshold correction loop order for a
  particular running \BSM parameter.  The association between the
  digits and the parameters as well as the default loop orders is
  shown in \tabref{tab:sg_mma_tc}.
  %
  \begin{example}
    The following table shows example values for the integer number
    which specifies the individual threshold correction loop orders,
    together with the list of included loop corrections.
    \begin{center}
    \begin{tabular}{rl}
      \toprule
      integer & used threshold corrections \\
      \midrule
      0       & no threshold corrections, everything at tree-level \\
      1       & only $\Delta\aem^{1L}$, everything else at tree-level \\
      100     & only $\Delta\as^{1L}$, everything else at tree-level \\
      101     & only $\Delta\aem^{1L}$ and  $\Delta\as^{1L}$, everything else at tree-level \\
      3000101 & only $\Delta\aem^{1L}$ and  $\Delta\as^{1L}$ and  $\Delta y_t^{3L}$, everything else at tree-level \\
      \bottomrule
    \end{tabular}
  \end{center}
  \end{example}
  %
  If the field \texttt{FlexibleSUSY[24]} or the symbol
  \texttt{thresholdCorrections} is omitted, then the whole field is
  set to the default value given in \tabref{tab:sg_mma_tc}.  If the
  field is not omitted, then all loop orders must be given.  Note
  that setting the loop orders larger than the value set in
  \texttt{FlexibleSUSY[7]} or \texttt{thresholdCorrectionsLoopOrder}
  has no effect, see above.
  %
  \begin{example}[label=ex:HSSUSY_thresholds]
    In the model file of \HSSUSY 3-loop QCD corrections to the running
    top Yukawa coupling $y_t$ are enabled (\code{UseYukawa3LoopQCD =
      True}).  Switching between the 2-loop and 3-loop QCD corrections
    to $y_t$ can be used to estimate a partial uncertainty of
    the 2-loop Higgs pole mass.  In order to do this,
    \code{FSHSSUSYOpenHandle[]} must be called twice, setting
    \begin{align*}
      \texttt{thresholdCorrections} \rightarrow 123111321 \\
      \texttt{thresholdCorrections} \rightarrow 122111321
    \end{align*}
    respectively, and setting each time
    \code{thresholdCorrectionsLoopOrder -> 3} to enable the 3-loop
    corrections globally.  Note that the digit at the $6$th position
    from the right (the prefactor of $10^6$) has been changed from $3$
    to $2$ to change the threshold correction loop order of $y_t$ from
    3-loop to 2-loop.  \exref{ex:HSSUSY_uncertainty} makes use
    of this method to estimate a partial uncertainty of
    \HSSUSY based on changing the threshold correction loop orders for
    $y_t$ and $\as$ in the SM\@.
  \end{example}

\item[\texttt{FlexibleSUSY[25]},]
  \texttt{higgs3loopCorrectionRenScheme}: This option applies only to
  MSSM models in which the 3-loop Higgs pole mass contributions from
  \Himalaya are enabled (the flag \code{UseHiggs3LoopMSSM = True} is
  set in the model file), see
  \secref{sec:model_specific_contributions}.  With this option the
  user can choose between the \DRbar and \MDRbar renormalization
  scheme.  If this option is set to $0$, then the \DRbar scheme is
  used.  If set to $1$, the \MDRbar scheme is used.

\item[\texttt{FlexibleSUSY[26]},] \texttt{higgs3loopCorrectionAtAsAs}:
  With this option the user can enable/disable 3-loop contributions to
  the Higgs pole mass(es) of $O(\at\as^2)$.  Note that this option
  has an effect only if model-specific 3-loop contributions to the
  Higgs pole mass(es) are enabled in the \fs model file.  See
  \secref{sec:model_specific_contributions} on how to enable 3-loop
  contributions of this order in the SM and in the MSSM\@.

\item[\texttt{FlexibleSUSY[27]},] \texttt{higgs3loopCorrectionAbAsAs}:
  With this option the user can enable/disable 3-loop contributions to
  the Higgs pole mass(es) of $O(\ab\as^2)$.  Note that this option
  has an effect only if model-specific 3-loop contributions to the
  Higgs pole mass(es) of this order are enabled in the \fs model file.
  See \secref{sec:model_specific_contributions} on how to enable
  3-loop contributions of this order in the SM and in the
  MSSM\@.

\item[\texttt{FlexibleSUSY[28]},] \texttt{higgs3loopCorrectionAtAtAs}:
  With this option the user can enable/disable 3-loop contributions to
  the Higgs pole mass(es) of $O(\at^2\as)$.  Note that this option
  has an effect only if model-specific 3-loop contributions to the
  Higgs pole mass(es) of this order are enabled in the \fs model file.
  See \secref{sec:model_specific_contributions} on how to enable
  3-loop contributions of this order in the SM\@.

\item[\texttt{FlexibleSUSY[29]},] \texttt{higgs3loopCorrectionAtAtAt}:
  With this option the user can enable/disable 3-loop contributions to
  the Higgs pole mass(es) of $O(\at^3)$.  Note that this option has
  an effect only if model-specific 3-loop contributions to the Higgs
  pole mass(es) of this order are enabled in the \fs model file.  See
  \secref{sec:model_specific_contributions} on how to enable 3-loop
  contributions of this order in the SM\@.

\item[\texttt{MODSEL[12]},] \texttt{parameterOutputScale}: With this
  option the scale (in GeV) can be specified, at which the running
  \MSbar/\DRbar model parameters are output.  If set to $0$, then the
  running parameters are output at the scale assigned to the
  \code{SUSYScale} variable in the \fs model file.
\end{itemize}

\section{CMSSMCPV model file}
\label{app:CMSSMCPV}
\lstinputlisting[language=Mathematica]{model_files/CMSSMCPV/FlexibleSUSY.m.in}

\section{\THDM model file}
\label{app:THDMIIMSSMBC}
\lstinputlisting[language=Mathematica,numbers=left,numberstyle=\tiny\sffamily]{model_files/THDMIIMSSMBC/FlexibleSUSY.m.in}

\section{CNMSSM model file}
\label{app:CNMSSM}
\lstinputlisting[language=Mathematica]{model_files/CNMSSM/FlexibleSUSY.m.in}

\clearpage

\section*{References}
\addcontentsline{toc}{section}{References}

\bibliographystyle{doc/JHEP}
\bibliography{doc/flexiblesusy-2.0}

\end{document}

### 2017-12-20 doc/flexiblesusy-2.0.tex

@@
 renormalization group equations
-\PACS 12.60.Jv
-\PACS 14.80.Ly
+\PACS 12.60.Cn, 12.60.Fr, 12.60.Jv, 12.60.-i
+\PACS 14.80.Bn, 14.80.Da, 14.80.Ec, 14.80.Fd, 14.80.Ly, 14.80.Nb, 14.80.Sv
 \end{keyword}
@@
 \begin{lstlisting}[language=bash]
-$ wget https://www.hepforge.org/archive/flexiblesusy/FlexibleSUSY-2.0.0.tar.gz
-$ tar -xf FlexibleSUSY-2.0.0.tar.gz
-$ cd FlexibleSUSY-2.0.0
+$ wget https://www.hepforge.org/archive/flexiblesusy/FlexibleSUSY-2.0.1.tar.gz
+$ tar -xf FlexibleSUSY-2.0.1.tar.gz
+$ cd FlexibleSUSY-2.0.1
 \end{lstlisting}%% $
@@
      1   FlexibleSUSY
-     2   2.0.0
+     2   2.0.1
      5   CMSSM
@@
   \code{UseHiggs2LoopSM} & \code{False} & \code{True} or \code{False}
-  & 2-loop contributions $O(\at^2 + \at \as)$ to $M_h$
-  in
-  the SM \\
+  & 2-loop contributions $O(\at^2 + \at \as)$ to $M_h$ in the SM \\
   %
@@
   $O(\as^2 + \as^3)$
-  to $\as$ in the SM \\
+  to the \MSbar $\as$ in the SM \\
   %
@@
   $O(\as^2 + \as\at + \as\ab)$
-  to $\as$ in the MSSM \\
+  to the \DRbar $\as$ in the MSSM \\
   %
@@
 $t \equiv m_t^2$ and $\barlog(t) \equiv \log(t/Q^2)$ have been used
-for brevity, analogous to the notation of Ref.~\cite{Martin:2014cxa}.  Note
+for brevity, analogous to the notation of Ref.~\cite{Martin:2014cxa},
+where $m_t$ is the \MSbar top mass in the SM\@.  Note
 that the momentum dependence is included in the 2-loop expression of
@@
 \begin{align}
-\begin{split}
   (\Delta m_h^2)_{3L} &=
   \frac{m_t^2}{(4\pi)^6} \Bigg[
-     g_3^4 y_t^2 \left(248.122 + 839.197 \;\barlog(t) + 160 \;\barlog^2(t) - 736 \;\barlog^3(t)\right) \\
+     g_3^4 y_t^2 \left(248.122 + 839.197 \;\barlog(t) + 160 \;\barlog^2(t) - 736 \;\barlog^3(t)\right) \nonumber \\
      &\phantom{=\frac{m_t^2}{(4\pi)^6} \Bigg[}
-     + g_3^2 y_t^4 \left(2764.365 + 1283.716 \;\barlog(t) - 360 \;\barlog^2(t) + 240 \;\barlog^3(t)\right) \\
+     + g_3^2 y_t^4 \left(2764.365 + 1283.716 \;\barlog(t) - 360 \;\barlog^2(t) + 240 \;\barlog^3(t)\right) \nonumber \\
      &\phantom{=\frac{m_t^2}{(4\pi)^6} \Bigg[}
-     + y_t^6 \Big(-3199.017 + 36 \;\barlog(h) - 2653.511 \;\barlog(t) + 756 \;\barlog(h) \;\barlog(t)\\
+     + y_t^6 \Big(-3199.017 + 36 \;\barlog(h) - 2653.511 \;\barlog(t) + 756 \;\barlog(h) \;\barlog(t) \nonumber \\
      &\phantom{=\frac{m_t^2}{(4\pi)^6} \Bigg[}\qquad~~
@@
   \Bigg]
-\end{split}
 \end{align}
 %
-with $h\equiv m_h^2$.  Note that the 3-loop Higgs mass calculation in
-the \MSbar scheme in \fs is only complete at $O(\at\as^2)$.  The
+where $h\equiv m_h^2$ is the squared \MSbar Higgs mass in the SM\@.
+Note that the 3-loop Higgs mass calculation in
+the \MSbar scheme in \fsbreak is only complete at $O(\at\as^2)$.  The
 3-loop contributions of $O(\at^3 + \at^2 \as)$ are currently incomplete
@@
 %
+\\\begin{minipage}{\linewidth}
 \begin{lstlisting}
@@
 \end{lstlisting}
+\end{minipage}
 %
@@
   \,,
-\end{split}
+\end{split} \label{eq:mt_MSbar}
 \end{align}
@@
 parameter in the convention of Ref.~\cite{Haber:1984rc} and the effective
-squared \CP-odd Higgs mass $m_A^2$ of the model must be identified.
+squared \CP-odd Higgs tree-level mass $m_A^2$ of the model must be identified.
 In the \modelname{NMSSM} model file of \sarah they would read for
@@
 scale $\MS$ at the scale $Q =
-\MS$ and all sfermion mixing parameters
+\MS$ and all \DRbar sfermion mixing parameters
 $X_f$ are set to zero, except for the stop mixing parameter
@@
 \begin{align}
-  (\Delta m_h^2)_{3L} &= \frac{64 g_3^4 y_t^2}{(4\pi)^6} m_t^2 \log^3\frac{m_{\tilde{g}}^2}{Q^2} \,,
+  (\Delta m_h^2)_{3L} &= \frac{64 g_3^4 y_t^2}{(4\pi)^6} m_t^2\;\barlog^3(g) \,,
 \end{align}
 %
-where $m_{\tilde{g}}$ is the \MSbar gluino mass.  To take this
-contribution into account at runtime, the following flags should also be
-set in the SLHA input file:
+where $g = m_{\tilde{g}}^2$ and $m_{\tilde{g}}$ is the \MSbar gluino
+mass.  Furthermore, for consistency the 2-loop gluino contribution in
+the calculation of the \MSbar top mass from Ref.\
+\cite{Benakli:2013msa} is taken into account by adding the following
+term to Eq.\ \eqref{eq:mt_MSbar} or \eqref{eq:mt_MSbar_default},
+respectively:
+%
+\begin{align}
+  \Delta m_t^{(2),\text{split-qcd}}(Q) =
+  -\frac{g_3^4}{(4\pi)^4} M_t \left\{
+     \frac{89}{9}
+     + 4\;\barlog(g) \left[
+        \frac{13}{3}
+        + \barlog(g)
+        - 2\;\barlog(t)
+     \right]
+  \right\} \,.
+\end{align}
+%
+To take these contributions into account at runtime, the following flags
+should also be set in the SLHA input file:
 %
@@
 %
+\\\begin{minipage}{\linewidth}
 \begin{lstlisting}[language=Mathematica]
@@
 \end{lstlisting}
+\end{minipage}
 
@@
 If the considered BSM model has a gauge symmetry with the SM gauge group $SU(3)_C\times SU(2)_L\times U(1)_Y$ as a factor,
-then \fs automatically fixes the three corresponding normalized gauge
+then \fs automatically fixes the three corresponding normalized running gauge
 couplings $g_1$, $g_2$ and $g_3$ at the low-energy boundary condition
@@
 %
-where the sum runs over all non-SM particles plus the top quark and
-the constants $C_i^{\text{em}}$ and $C_i^{\text{s}}$ depend on the
+where the sum runs over all non-SM particles plus the top quark
+with running \MSbar masses $m_i(Q)$.
+The constants $C_i^{\text{em}}$ and $C_i^{\text{s}}$ depend on the
 representation of the particle $i$ with respect to the Lorentz and gauge group.
@@
   SU(2)_L\times SU(2)_R\times U(1)_L\times U(1)_R$, one could for
-  example fix the BSM gauge couplings $g_3$, $g_L$, $g_R$, $g_{1L}$
-  and $g_{1R}$ by SM-like gauge couplings $g_3^\SM$,
+  example fix the running BSM gauge couplings $g_3$, $g_L$, $g_R$, $g_{1L}$
+  and $g_{1R}$ by the running SM-like gauge couplings $g_3^\SM$,
   $g_2^\SM$ and $g_1^\SM$ which are given as input via the
@@
 %
-with $f = u,d,c,s,t,b,e,\mu,\tau$.
+with $f = u,d,c,s,t,b,e,\mu,\tau$ and the \MSbar vacuum expectation value $v^{\SM}(Q)$.
 The running top quark \MSbar\ mass in the non-SUSY model, $m_t(Q)$, is
@@
   \,,
-\end{split}
+\end{split} \label{eq:mt_MSbar_default}
 \end{align}
@@
   in the MSSM with \HSSUSY.  In the following example script, a scan
-  over the relative stop mixing parameter $X_t/\MS$ is performed for
-  $\tan\beta = 5$ and three different values of the SUSY scale $\MS$.
+  over the relative \DRbar stop mixing parameter $X_t/\MS$ is performed for
+  $\tan\beta(\MS) = 5$ and three different values of the SUSY scale $\MS$.
   This is done in parallel on all available CPU cores.
@@
 \begin{itemize}
-\item We have performed an analytic comparison of the split-MSSM RGEs
-  generated with \sarah 4.5.3 against the RGEs presented in
+\item We have performed an analytic comparison of the RGEs generated
+  with \sarah 4.5.3 for \fs's split-MSSM model file
+  (\modelname{SplitMSSM}) against the RGEs presented in
   Ref.~\cite{Benakli:2013msa} and we found exact agreement.
@@
 %
-where $x=m_F^2/m_{S}^2$, $M_\mu$ is the muon pole mass,
+where $x=m_F^2/m_{S}^2$ is the squared running mass ratio between $F$ and $S$, $M_\mu$ is the muon pole mass,
 $m_\mu$ is the muon \MSbar/\DRbar mass and $c$ denotes the electric charge of
@@
 scalar--fermion--muon vertex, $\Gamma_{SF\mu} =
-ie(z_L P_L + z_R P_R)$.
+ie(z_L P_L + z_R P_R)$ with $e$ being the \MSbar/\DRbar electromagnetic coupling constant.
 %
@@
 \end{equation}
-where $M_f$ is the pole mass of $f$.
+where $M_f$ is the pole mass of $f$
+and $e$ is the running electromagnetic coupling constant in the BSM model.
 It is then obvious that
@@
 as can be guessed from Eq.\ \eqref{eq:ASF}.
+The calculation is performed at the renormalization scale $\MS$ specified
+in the model file, which is typically set to the stop mass scale
+in supersymmetric models.
 
@@
 $\mhalf$, and both soft masses of the left- and right-handed
-selectron shall be $\mzero$.
+selectron as well as the approximate tree-level heavy Higgs masses%
+\footnote{$\sqrt{\re\!B\mu / (\cos\!\beta \sin\!\beta)}$.}
+shall be $\mzero$.
 The moduli of $\mu$, $\mhalf$, $\mzero$, and $A_e$
@@
 all of which including the scale are set to $2\unit{TeV}$.
-Finally, we set $\tan\beta = 10$ at the SUSY scale.
+Finally, we set $\tan\beta = 10$ at the $M_Z$ scale.
 
@@
   \frac{\pi\,\aem}
-   {\sqrt{2}\,M_Z^2\,G_F\,\hat\rho_\text{tree} \left(1-\Delta\hat{r}\right)}\,,
+   {\sqrt{2}\,M_Z^2\,G_F\,\hat\rho_\text{tree} \left(1-\Delta\hat{r}\right)}\,
   \label{eq:muon_decay_master}
@@
 %
-where
+at the scale $M_Z$,
+where $\aem$ is the electromagnetic coupling of the \BSM model in the
+\MSbar/\DRbar scheme and
 %
@@
 In the occurring self energies $\Sigma$, the top quark mass is chosen
-to be the pole mass in order to include partial 2-loop corrections not
+to be the pole mass $M_t$ in order to include partial 2-loop corrections not
 contained in $\Delta\hat{r}_\text{2L}^\SM$ and $\Delta\hat\rho_\text{2L}^\SM$
@@
 %
-where we have generalized the Higgs dependent part as
+where the SM strong coupling $\as^{\SM(5)}$ is taken at the scale $M_t$ and
+we have generalized the Higgs dependent part as
 %
@@
 dashed-dotted line, respectively. These show good agreement with the
-values from \fmw for both the CMSSM and MRSSM.
+values from \fmw for both the CMSSM and MRSSM\@.
 The remaining small discrepancies are well
@@
     \unit{GeV}$.  Right panel: Change in the calculated lightest \CP-even
-    Higgs mass $M_{h_1}$ from a reference value of $121$ GeV as a
+    Higgs pole mass $M_{h_1}$ from a reference value of $121$ GeV as a
     function of $\mu$ at fixed $\mzero = 4$ TeV for the two solvers.  The
@@
 \end{equation*}
-where $M_h^{\SM}$ is the SM-like Higgs mass.  In this way the scan is directed
-towards solutions with the observed Higgs mass of $125.09$ GeV\@.
+where $M_h^{\SM}$ is the SM-like Higgs pole mass.  In this way the
+scan is directed towards solutions with the observed Higgs mass of
+$125.09$ GeV\@.
 
@@
     doc/plots/CNMSSM/cnmssm_4D-m0sqgtr0-full-msu6.pdf}
-   \caption{CNMSSM solutions obtained using \fstwo and \multinest 3.10 shown in
-     the $\mzero-\mhalf$ plane with $\azero$ as a color contour (top left
-     panel); $\mhalf-\lambda$ plane with the Higgs masses satisfying
-     $M_h^{\SM} > 100$ GeV shown as a color contour (top right
-     panel); $\mzero-\mhalf$ plane with the mass of the lightest up-type squark, which is
-     predominantly stop (bottom left panel), and the $\mhalf-\azero$ plane with
-     the heaviest up-type squark mass (bottom right panel).}
+   \caption{CNMSSM solutions obtained using \fstwo and \multinest 3.10
+     shown in the $\mzero-\mhalf$ plane with $\azero$ as a color
+     contour (top left panel); $\mhalf-\lambda$ plane showing the
+     SM-like Higgs pole mass $M_h^{\SM} $, as a color contour, with
+     the range restricted to $M_h^{\SM} > 100$ GeV for clarity (top
+     right panel); $\mzero-\mhalf$ plane with the mass of the lightest
+     up-type squark, which is predominantly stop (bottom left panel),
+     and the $\mhalf-\azero$ plane with the heaviest up-type squark
+     mass (bottom right panel).}
    \label{fig:cnmssm_solutions}
@@
    \caption{C\ESSM solutions obtained using \fstwo and \multinest 3.10 shown in
-     the $\mzero-\mhalf$ plane with $\azero$, the lightest up-type squark mass
+     the $\mzero-\mhalf$ plane with $\azero$, the lightest up-type squark pole mass
      $M_{\tilde{u}_1}$,  $|\lambda|$ and $|\kappa|$ as color contours.}
@@
 %
-where the matching scale $Q$ is of the order $\MS$.
-If Eq.\ \eqref{eq:Mh_top_SM} and Eq.\ \eqref{eq:Mh_top_BSM} are set equal and
-the relation Eq.\ \eqref{eq:yt_SM_BSM} is inserted, potentially
+where the matching scale $Q$ is of the order $\MS$ and we have also
+introduced $m_h^\SM$, $m_t^\SM$ and $m_t^\BSM$ for the running SM
+Higgs mass, running SM top mass and running BSM top mass,
+keeping our convention of using an upper case `$M$' for pole masses
+and lower case `$m$' for running tree-level masses.  If
+Eq.\ \eqref{eq:Mh_top_SM} and Eq.\ \eqref{eq:Mh_top_BSM} are set equal
+and the relation Eq.\ \eqref{eq:yt_SM_BSM} is inserted, potentially
 large 2-loop terms for example of the form
@@
 scale, which is calculated in terms of running \BSM parameters which
-have been obtained by a tree-level matching.
+have been obtained by a tree level matching.
 %
@@
 ($f=u,d,e$) and the SM-like vacuum expectation value $v^\BSM(\Qmatch)$
-of the \BSM model are determined automatically from the matching
-conditions
+of the \BSM model are determined automatically from the following matching
+conditions on pole masses and running couplings
 %
@@
 %
-The pole masses of the SM particles are calculated at the
-scale $\QpoleSM$, which is set to $M_t$ by default.  The scale
-$\QpoleSM$ can be changed by setting the entry \code{FlexibleSUSY[18]}
-to a non-zero value in the SLHA input file.  In the \mathematica
-interface, $\QpoleSM$ can be changed by setting \code{eftPoleMassScale}
-to a non-zero value.
+The pole masses of the SM particles are calculated at the scale
+$\QpoleSM$, which is set to the top pole mass $M_t$ by default.  The
+scale $\QpoleSM$ can be changed by setting the entry
+\code{FlexibleSUSY[18]} to a non-zero value in the SLHA input file.
+In the \mathematica interface, $\QpoleSM$ can be changed by setting
+\code{eftPoleMassScale} to a non-zero value.
 %
@@
   \begin{lstlisting}
-Mh = (125.047 \pm 1.52741) GeV
+Mh = (125.047 +- 1.52741) GeV
   \end{lstlisting}
@@
 \item The extraction of the running Yukawa couplings in models where
-  a 4th generation of fermions mixes with the SM fermions is currently
+  a 4$^\text{th}$ generation of fermions mixes with the SM fermions is currently
   not supported.  As a workaround, the running Yukawa couplings can
@@
 D.H.\ also acknowledges financial support from the Grant Agency of the Czech
-Republic (GACR), contract 17-04902S.  This work has also been supported by the
+Republic (GACR), contract 17-04902S\@.  This work has also been supported by the
 German Research Foundation DFG through Grant No.~STO876/2.
@@
   particles, as defined in the SLHA-2 standard \cite{Allanach:2008qq}.
-  The definition of the individual block entries are shown in the
+  The definitions of the individual block entries are shown in the
   first part of \tabref{tab:sm_mma_options}.  If a parameter is
@@
    12   0.000000000e+00      # mnu1(pole)
-   13   1.056583570e-01      # mmuon(pole)
+   13   1.056583715e-01      # mmuon(pole)
    14   0.000000000e+00      # mnu2(pole)
@@
 SM input parameters are given in
-\tabref{tab:sm_mma_options}.  Unset parameters are set to their
+\tabref{tab:sm_mma_options}.\footnote{Note that in \fs's \mathematica
+  interface, the fine structure constant $\aem^{\SM(5)}(M_Z)$ is input,
+  not its inverse as in the SLHA standard.}
+Unset parameters are set to their
 default values defined in the table.  Note that in the \mathematica
@@
     13 & \code{Mm} & $0.1056583715$ & muon pole mass, $M_\mu/\text{GeV}$ \\
-    14 & \code{Mv2} & $0$ & 2nd lightest neutrino pole mass, $M_{\nu_2}/\text{GeV}$ \\
+    14 & \code{Mv2} & $0$ & 2$^\text{nd}$ lightest neutrino pole mass, $M_{\nu_2}/\text{GeV}$ \\
     21 & \code{md2GeV} & $0.00475$ & running down mass, $m_d(2\unit{GeV})/\text{GeV}$ \\
@@
 from the \texttt{MODSEL} block, see below.  In the SLHA format all \fs
-configuration blocks with their respective default values read:
+configuration entries with their respective default values read:
 %
@@
    29   1            # Higgs 3L corrections O(alpha_t^3)
-   28   1            # Higgs 3L corrections O(alpha_t^2 alpha_s)
-   29   1            # Higgs 3L corrections O(alpha_t^3)
 \end{lstlisting}
@@
 the configuration options are given in \tabref{tab:sg_mma_options}.
-Unset parameters are set to their default values defined in the table.
+Unset options are set to their default values defined in the table.
 A call of \code{FS<model>OpenHandle[]} with all configuration options
@@
     11 & \code{higgs2loopCorrectionAtauAtau}  & \code{1} &
-    enable/disable 2-loop corrections $O(\atau\atau)$ to $M_{h,H,A}$\\
+    enable/disable 2-loop corrections $O(\atau^2)$ to $M_{h,H,A}$\\
     12 & \code{forceOutput}                   & \code{0} &
@@
     25 & \code{higgs3loopCorrectionRenScheme} & \code{0} &
-    renormalization scheme for 3-loop MSSM Higgs corrections:
-    0 = \DRbar, 1 = \MDRbar\\
+    renormalization scheme for 3-loop MSSM Higgs corrections
+    (0 = \DRbar, 1 = \MDRbar)\\
     26 & \code{higgs3loopCorrectionAtAsAs}    & \code{1} &
@@
     symbols to specify the configuration options for \fs's spectrum
-    generators.}
+    generators.  The symbols $M_{h,H,A}$ and $M_t$ denote the Higgs
+    and top quark pole masses, respectively.}
   \label{tab:sg_mma_options}
@@
   \caption{Specification of individual loop orders of threshold
-    corrections for the model parameters at the low-energy scale using
-    the field \texttt{FlexibleSUSY[24]} in the SLHA interface or the
+    corrections for extracting the running masses, couplings and
+    Weinberg angle in the BSM model at the low-energy scale using the
+    field \texttt{FlexibleSUSY[24]} in the SLHA interface or the
     symbol \texttt{thresholdCorrections} in the \mathematica
@@
   precision of $10^{-4}$ is sufficient.  For models with various
-  3-loop corrections, like \HSSUSY or an MSSM spectrum generator, a
+  3-loop corrections, like \HSSUSY or MSSM-like models, a
   precision of $10^{-5}$ might be better.
@@
   \secref{sec:chosing_SAS}) are used, starting with the two-scale
-  solver.  If set to $1$, the two-scale solver is used, if it has been
+  solver.  If set to $1$, the two-scale solver is used if it has been
   enabled in the model file.  If set to $2$, then the semi-analytic
-  solver is used, if it has been enabled in the model file.  If a
+  solver is used if it has been enabled in the model file.  If a
   solver that has not been enabled in the model file is chosen, \fs
@@
   1-loop level.  If set to $2$ or $3$, then model-specific 2-loop or
-  3-loop corrections are taken into account, respectively, if they are
+  3-loop corrections are taken into account, respectively, if they have been
   enabled in the model file (see
@@
   symmetry breaking (EWSB) equations are solved.  If set to $0$, the
-  EWSB equations are solved at the tree-level.  If set to $1$, the
+  EWSB equations are solved at the tree level.  If set to $1$, the
   EWSB equations are solved at the 1-loop level.  If set to $2$ or
   $3$, then model-specific 2-loop or 3-loop corrections are taken into
-  account, respectively, if they are enabled in the model file (see
+  account, respectively, if they have been enabled in the model file (see
   \secref{sec:model_specific_contributions}).  \textbf{Important
@@
 \item[\texttt{FlexibleSUSY[16]},] \texttt{forcePositiveMasses}: With
-  this option, the masses of Majorana fermions can be forced to be
+  this option the masses of Majorana fermions can be forced to be
   positive in the SLHA output of \fs.  If set to $1$, then Majorana
@@
   SM) are calculated, use \texttt{FlexibleSUSY[18]} or
-  \texttt{eftPoleMassScale} in the SLHA or Mathematica interfaces,
+  \texttt{eftPoleMassScale} in the SLHA or \mathematica interface,
   respectively.
@@
   then this option can be set to $0$ to suppress the calculation of
-  the heavy Higgs bosons, the charginos, neutralinos and sfermions.
+  the masses of the heavy Higgs bosons, the charginos, neutralinos and sfermions.
 
@@
   particular running \BSM parameter.  The association between the
-  digits and the parameters as well as the default loop orders is
+  digits and the parameters as well as the default loop orders are
   shown in \tabref{tab:sg_mma_tc}.
@@
       \midrule
-      0       & no threshold corrections, everything at tree-level \\
-      1       & only $\Delta\aem^{1L}$, everything else at tree-level \\
-      100     & only $\Delta\as^{1L}$, everything else at tree-level \\
-      101     & only $\Delta\aem^{1L}$ and  $\Delta\as^{1L}$, everything else at tree-level \\
-      3000101 & only $\Delta\aem^{1L}$ and  $\Delta\as^{1L}$ and  $\Delta y_t^{3L}$, everything else at tree-level \\
+      0       & no threshold corrections, everything at tree level \\
+      1       & only $\Delta\aem^{1L}$, everything else at tree level \\
+      100     & only $\Delta\as^{1L}$, everything else at tree level \\
+      101     & only $\Delta\aem^{1L}$ and  $\Delta\as^{1L}$, everything else at tree level \\
+      3000101 & only $\Delta\aem^{1L}$ and  $\Delta\as^{1L}$ and  $\Delta y_t^{3L}$, everything else at tree level \\
       \bottomrule
@@
   If the field \texttt{FlexibleSUSY[24]} or the symbol
-  \texttt{thresholdCorrections} is omitted, then the whole field is
+  \texttt{thresholdCorrections} is omitted, then the whole option is
   set to the default value given in \tabref{tab:sg_mma_tc}.  If the
@@
   \begin{example}[label=ex:HSSUSY_thresholds]
-    In the model file of \HSSUSY 3-loop QCD corrections to the running
+    In the model file of \HSSUSY, 3-loop QCD corrections to the running
     top Yukawa coupling $y_t$ are enabled (\code{UseYukawa3LoopQCD =
@@
     \code{thresholdCorrectionsLoopOrder -> 3} to enable the 3-loop
-    corrections globally.  Note that the digit at the $6$th position
+    corrections globally.  Note that the digit at the 6$^\text{th}$ position
     from the right (the prefactor of $10^6$) has been changed from $3$
@@
   has an effect only if model-specific 3-loop contributions to the
-  Higgs pole mass(es) are enabled in the \fs model file.  See
+  Higgs pole mass(es) of this order have been enabled in the \fs model file.  See
   \secref{sec:model_specific_contributions} on how to enable 3-loop
@@
   has an effect only if model-specific 3-loop contributions to the
-  Higgs pole mass(es) of this order are enabled in the \fs model file.
+  Higgs pole mass(es) of this order have been enabled in the \fs model file.
   See \secref{sec:model_specific_contributions} on how to enable
@@
   an effect only if model-specific 3-loop contributions to the Higgs
-  pole mass(es) of this order are enabled in the \fs model file.  See
+  pole mass(es) of this order have been enabled in the \fs model file.  See
   \secref{sec:model_specific_contributions} on how to enable 3-loop

### 2018-02-22 doc/flexiblesusy-2.0.tex
update FS-2.0 paper (version re-submitted to CPC, commit 1d558aec475d)

@@
 \newcommand{\MS}{\ensuremath{M_S}\xspace}
+\newcommand{\MLCP}{\ensuremath{M_\text{LCP}}\xspace}
 \newcommand{\MEWSB}{\ensuremath{M_\text{EWSB}}\xspace}
@@
   \begin{abstract}
-    We document major new features and improvements
-    of \fs, a \mathematica and
-    C++ package that generates fast and precise spectrum
-    generators. These extensions significantly increase the generality
-    and capabilities of the \fs package, while maintaining the elegant
-    structure and easy to use interfaces. The \fbsm extension makes it
-    possible to also create spectrum generators for non-supersymmetric
-    extensions of the Standard Model. The \fcpv extension adds the
-    option of complex parameters to the spectrum generators, allowing
-    the study of many interesting models with new sources of \CP
-    violation. \fmw computes the decay of the muon for the generated
-    model and thereby allows \fs to predict the mass of the $W$ boson
-    from the input parameters by using the more precise electroweak
-    input of $\{ G_F, M_Z, \aem \}$ instead of $\{ M_W, M_Z,
-    \aem \}$. The \famu extension provides a calculator of the
-    anomalous magnetic moment of the muon in any model \fs can
-    generate a spectrum for.  \fsas introduces a new solver for the
-    boundary value problem which makes use of semi-analytic expressions
-    for dimensionful parameters to find solutions in models
-    where the classic two-scale solver will not work such as the constrained
-    E$_6$SSM\@. \feft is a hybrid calculation of the Higgs mass which
-    combines the virtues of both effective field theory calculations
-    and fixed-order calculations.  All of these extensions are
-    included in \fstwo, which is released simultaneously with this
-    manual.
+    We document major new features and improvements of \fs, a
+    \mathematica and C++ package with a dependency on the external
+    package \sarah, that generates fast and precise spectrum
+    generators. The extensions presented here significantly increase
+    the generality and capabilities of the \fs package, which already
+    works with a wide class of models, while
+    maintaining an elegant structure and easy to use interfaces. The
+    \fbsm extension makes it possible to also create spectrum
+    generators for non-supersymmetric extensions of the Standard
+    Model. The \fcpv extension adds the option of complex parameters
+    to the spectrum generators, allowing the study of many interesting
+    models with new sources of \CP violation. \fmw computes the decay
+    of the muon for the generated model and thereby allows \fs to
+    predict the mass of the $W$ boson from the input parameters by
+    using the more precise electroweak input of $\{ G_F, M_Z, \aem \}$
+    instead of $\{ M_W, M_Z, \aem \}$. The \famu extension provides a
+    calculator of the anomalous magnetic moment of the muon in any
+    model \fs can generate a spectrum for.  \fsas introduces a new
+    solver for the boundary value problem which makes use of
+    semi-analytic expressions for dimensionful parameters to find
+    solutions in models where the classic two-scale solver will not
+    work such as the constrained E$_6$SSM\@. \feft is a hybrid
+    calculation of the Higgs mass which combines the virtues of both
+    effective field theory calculations and fixed-order calculations.
+    All of these extensions are included in \fstwo, which is released
+    simultaneously with this manual.%% \sarah, an external dependency
+    %% providing model dependent information, can be automaically
+    %% installed using an installation script distributed with \fs.
   \end{abstract}
@@
 Popular and well studied new physics extensions of the Standard Model
-(SM) are coming under increasing pressure from the searches
-at the Large Hadron Collider (LHC) as well as other current experiments.
-At the same time, there remain many outstanding physics problems that are not
-solved by the SM and require new physics.  For example, the origin of dark
-matter is still unexplained and it requires some new particle
+(SM) are coming under increasing pressure from the searches at the
+Large Hadron Collider (LHC) as well as other current experiments.  At
+the same time, there remain many outstanding physics problems that are
+not solved by the SM and require new physics.  For example, the origin
+of dark matter is still unexplained and it requires some new particle
 to fit the relic density.  In addition, the gauge structure and
@@
 of magnitude smaller than the Planck scale, has no explanation within
-the SM and has been the driving motivation for the construction and study
-of concepts that go beyond the Standard Model (BSM), most notably supersymmetry
-(SUSY) \cite{Weinberg:1975gm, Weinberg:1979bn, Gildener:1976ai,Susskind:1978ms,
-  tHooft:1980xss}.
+the SM \cite{Weinberg:1975gm, Weinberg:1979bn,
+  Gildener:1976ai,Susskind:1978ms, tHooft:1980xss} and has been the
+driving motivation for the construction and study of concepts that go
+beyond the Standard Model (BSM), most notably supersymmetry (SUSY) \cite{Miyazawa:1966mfa,Miyazawa:1968zz,Ramond:1971gb,Golfand:1971iw,Gervais:1971ji,Neveu:1971rx,Neveu:1971iv,Volkov:1973ix,Wess:1973kz,Wess:1974tw,Salam:1974yz} which ensures that the quadratic corrections from fermions and bosons in the loop diagrams cancel at all orders in perturbation theory \cite{Dimopoulos:1981au,Witten:1981nf,Dine:1981za,Dimopoulos:1981zb,Kaul:1981hi}.
 
@@
 heavily unit-tested against the MSSM and Next-to-MSSM versions of
-\softsusy every night to avoid bugs.
+\softsusy every night to ensure bugs are avoided in updates.
 
@@
 %
-\fstwo is distributed with a huge selection of predefined models
-including: the CMSSM (called \modelname{CMSSM}), the semi-constrained and
-fully constrained NMSSM (\modelname{NMSSM}, \modelname{CNMSSM}), the
-$R$-symmetric MSSM (\modelname{MRSSM}) \cite{Kribs:2007ac}, the NUHM \ESSM
-(\modelname{E6SSM}) \cite{Athron:2007en}, high-scale SUSY (\HSSUSY)
-\cite{Bagnaschi:2015pwa}, split-SUSY (\modelname{SplitMSSM})
-\cite{Bagnaschi:2015pwa} and the THDM-II (\modelname{THDMII}).  See the
+\fstwo is distributed with a huge selection of predefined ``models'', 
+including several MSSM and NMSSM scenarios such as the CMSSM (called
+\modelname{CMSSM}), high-scale SUSY and split-SUSY (called \HSSUSY and
+\modelname{SplitMSSM}; these  models have been created for
+Ref.~\cite{Bagnaschi:2015pwa}), the semi-constrained and
+fully constrained NMSSM (\modelname{NMSSM}, \modelname{CNMSSM}). The
+distribution also contains BSM models such as the
+$R$-symmetric MSSM (\modelname{MRSSM}; for a definition of the model
+see Ref.~\cite{Kribs:2007ac}), the NUHM \ESSM
+(\modelname{E6SSM}, see \cite{King:2005jy,King:2005my,Athron:2010zz,
+  Athron:2009ue,Athron:2009bs,Athron:2007en}) and the two-Higgs 
+doublet model type II
+(\modelname{THDMII}).  See the 
 contents of \code{model_files/} for all predefined model files.
@@
 \code{LesHouches.in.CMSSM} and first solves a BVP to find a
-set of running parameters at the \code{SUSYScale} that are consistent with all
+set of running parameters at the SUSY scale that are consistent with all
 boundary conditions specified in the model file, and then calculates the pole
@@
   \code{UseHiggs3LoopSplit} & \code{False} & \code{True} or \code{False} &
-  3-loop contributions $O(\at\as^2)$ to $M_h$, $M_H$ and
-  $M_A$ in the split-MSSM \\
+  3-loop contributions $O(\at\as^2)$ to $M_h$ in the split-MSSM \\
   %
@@
 mass are known in the \MSbar scheme
-\cite{Degrassi:2012ry,Martin:2014cxa,Vega:2015fna}.  In \fs the 2-loop
+\cite{Degrassi:2012ry,Vega:2015fna,Martin:2014cxa}.  In \fs the 2-loop
 contributions of $O(\at^2 + \at \as)$
-\cite{Degrassi:2012ry,Martin:2014cxa} and the 3-loop
+\cite{Vega:2015fna,Martin:2014cxa} and the 3-loop
 $O(\at^3 + \at^2 \as + \at \as^2)$ contributions \cite{Martin:2014cxa}
@@
 %
+The 2-loop and 3-loop
+contributions enter the mass of the Higgs boson as
+\begin{align} 
+ \begin{split}
+   M_h^2 &= m_h^2 + (\Delta m_h^2)_{1L}(p^2)
+   + (\Delta m_h^2)_{2L}(p^2) +(\Delta m_h^2)_{3L} \,,
+ \end{split} \label{eq:Mh_polp^2}
+\end{align}
+where $m_h^2$ and $(\Delta m_h)_{1L}(p^2)$ correspond to the tree-level 
+and a 1-loop expression, respectively.
 The enabled 2-loop SM contributions of
@@
 \begin{align}
-  (\Delta m_h^2)_{2L} &= (\Delta m_h^2)_{2L}^{(\at^2)} + (\Delta m_h^2)_{2L}^{(\at\as)} \,, \\
+  (\Delta m_h^2)_{2L}(p^2) &= (\Delta m_h^2)_{2L}^{(\at^2)} + 
+     (\Delta m_h^2)_{2L}^{(\at\as)}(p^2) \,, \\
   (\Delta m_h^2)_{2L}^{(\at^2)} &=
@@
   \right] , \label{eq:2L_Mh_atat}\\
-  (\Delta m_h^2)_{2L}^{(\at\as)} &=
+  (\Delta m_h^2)_{2L}^{(\at\as)} (p^2)&=
       \frac{g_3^2 y_t^2}{(4\pi)^4} \;\bigg[
@@
   \right] \,.
+  \label{eq:2L_Mh_atas p2=0}
 \end{align}
 %
-In Eqs.~\eqref{eq:2L_Mh_atat} ff.\ the abbreviations
-$t \equiv m_t^2$ and $\barlog(t) \equiv \log(t/Q^2)$ have been used
-for brevity, analogous to the notation of Ref.~\cite{Martin:2014cxa},
-where $m_t$ is the \MSbar top mass in the SM\@.  Note
-that the momentum dependence is included in the 2-loop expression of
-Eq.~\eqref{eq:2L_Mh_atas} in order to generate implicit 3-loop
-$O(\at^2\as)$ terms.
+In Eqs.~\eqref{eq:2L_Mh_atat} ff.\ the abbreviations $t \equiv m_t^2$
+and $\barlog(t) \equiv \log(t/Q^2)$ have been used for brevity, 
+analogous to the notation of Ref.~\cite{Martin:2014cxa}, where $m_t$ 
+is the \MSbar top mass in the SM\@. 
+By performing a momentum iteration in the computation of the Higgs 
+pole mass, the momentum entering the 1-loop self energy in 
+Eq.~\eqref{eq:Mh_polp^2} consists of a tree-level part and a loop 
+correction.
+In combination with Eq.~\eqref{eq:2L_Mh_atat}, the latter yields the 
+complete Higgs mass contribution at $O(\at^2 t)$ which is identical 
+to the corresponding correction given in Ref.~\cite{Degrassi:2012ry}.
+The 2-loop self energy at $O(\at\as)$ \cite{Martin:2014cxa} together
+with the occurring integral functions from Ref.~\cite{Martin:2003qz}
+have been evaluated for a small external momentum argument. Neglecting
+higher orders, the expansion in powers of the momentum over the \MSbar
+top mass up to $O(p^4/t)$ results in the Eq.~\eqref{eq:2L_Mh_atas}. 
+Note that the momentum dependence is included in the 2-loop expression 
+in order to generate implicit 3-loop $O(\at^2\as)$ terms.
 %
@@
      + \frac{27}{2} \;\barlog^2(t) + 324 \;\barlog(h) \;\barlog^2(t) - 225 \;\barlog^3(t)\Big)
-  \Bigg]
+  \Bigg] \,,
 \end{align}
@@
 \begin{align}
+  \label{eq:top-selfenergy-qcd-1L}
   \Delta m_t^{(1),\text{qcd}}(Q) &=
   -\frac{g_3^2}{12 \pi^2} \left[4 - 3 \;\barlog(t)\right], \\
+  \label{eq:top-selfenergy-qcd-2L}
   \begin{split}
@@
 \cite{Degrassi:2001yf,Brignole:2001jy,Dedes:2002dy,Brignole:2002bz,Dedes:2003km,Degrassi:2009yq}
-can be taken into account by \fs by setting the corresponding switches
-in the model files:
+can be taken into account by \fs.
+%
+Note, however, that the implemented NMSSM 2-loop contributions of
+$O((\at + \ab)^2 + \atau^2)$ are currently available only in the
+MSSM-limit.  Furthermore, NMSSM-specific 2-loop corrections to the
+running vacuum expectation value beyond the MSSM limit, which would be
+required for a consistent treatment of the $O((\at + \ab)\as)$
+contributions, are currently not implemented.
+%
+To take the implemented 2-loop corrections to the Higgs masses in the
+(N)MSSM into account, the following switches must be set in the model
+file:
 %
@@
 coupling $\as$ (see below) on the 3-loop Higgs mass calculation is shown for
-the same scenario as in the corresponding left panels.  The inclusion
-of $\Delta\as^{2L}$ is formally a 4-loop effect on the light
+the same scenario as in the corresponding left panels.
+The threshold correction $\Delta\as^{2L}$ is included at the scale
+$Q = M_Z$, the same scale at which all dimensionless MSSM \DRbar
+parameters are determined from the Standard Model input parameters.
+The inclusion of $\Delta\as^{2L}$ is formally a 4-loop effect on the light
 \CP-even Higgs mass in the MSSM\@.  However, $\Delta\as^{2L}$ should
@@
 shift of the Higgs mass by up to $2\unit{GeV}$, depending on \MS and
-$X_t$.
+$X_t$.  Note, that large 4-loop contributions of multiple GeV hint at
+a large theoretical uncertainty of the fixed-order calculation of the
+light \CP-even Higgs pole mass in parameter regions with multi-TeV
+stop masses.
 
@@
 \cite{Benakli:2013msa} is taken into account by adding the following
-term to Eq.\ \eqref{eq:mt_MSbar} or \eqref{eq:mt_MSbar_default},
-respectively:
+term to Eq.\ \eqref{eq:mt_MSbar}:
 %
@@
     Returns the cubic root of \code{a}, $\sqrt[3]{a}$. \\
+
+    \code{CKM}, \code{PMNS} &
+    CKM and PMNS matrices, respectively, as defined in \cite{Allanach:2008qq}. \\
 
@@
 the gauge and Yukawa couplings of the non-SUSY model in the \MSbar
-scheme.
+scheme.  For gauge-dependent quantities like running masses and VEVs,
+\fs adopts the Feynman gauge, where all gauge fixing parameters
+$\xi_i$ are set to unity.
 
@@
 using the full 1-loop self energy plus 2-loop SM QCD
-corrections,
-%
-\begin{align}
-\begin{split}
-  m_t(Q) &= M_t + \re\Sigma_{t}^S(p^2=M_t^2,Q) \\
-  &\phantom{={}} + M_t \Big[ \re\Sigma_{t}^L(p^2=M_t^2,Q) +
-    \re\Sigma_{t}^R(p^2=M_t^2,Q) \\
-  &\phantom{={} + M_t \Big[}
-    + \Delta m_t^{(1),\text{qcd}}(Q) + \Delta m_t^{(2),\text{qcd}}(Q) \Big]
-  \,,
-\end{split} \label{eq:mt_MSbar_default}
-\end{align}
-%
-where $\Sigma_{t}^{S,L,R}$ denote the scalar, left- and right-handed
-parts of the 1-loop top self energy in the \MSbar\ scheme in the
-non-SUSY model without the gluon
-contribution, and $\Delta m_t^{(1),\text{qcd}}$ and $\Delta
-m_t^{(2),\text{qcd}}$ are 1- and 2-loop gluon corrections taken from
-Ref.~\cite{Fleischer:1998dw},
-%
-\begin{align}
-  \Delta m_t^{(1),\text{qcd}}(Q) &=
-  -\frac{g_3^2}{12 \pi^2} \left[4 - 3 \log\frac{m_t^2}{Q^2}\right],
-  \label{eq:top-selfenergy-qcd-1L}\\
-  \begin{split}
-    \Delta m_t^{(2),\text{qcd}}(Q) &= \left(\Delta
-      m_t^{(1),\text{qcd}}(Q)\right)^2 - \frac{g_3^4}{4608 \pi^4}
-    \Bigg[396 \log^2\frac{m_t^2}{Q^2} \\
-    &\phantom{={}} - 1452 \log\frac{m_t^2}{Q^2}
-    - 48 \zeta(3) + 2053 + 16 \pi^2 (1+\log 4)\Bigg] \,.
-  \end{split}
-  \label{eq:top-selfenergy-qcd-2L}
-\end{align}
-%
+corrections as shown in Eq.\ \eqref{eq:mt_MSbar}.
 In the SM, 3-loop QCD contributions can be taken into
@@
 the 2-loop threshold corrections of
-$O(\at\as + \ab\as + (\at+\ab)^2 + \atau^2)$ for general SUSY spectra.
+$O(\at\as + \ab\as + (\at+\ab)^2 + \ab\atau + \atau^2)$ for general
+SUSY spectra.
 
@@
 %
-The calculation is performed at the renormalization scale $\MS$, which
+The calculation is performed at the renormalization scale $\MLCP$, which
 is defined as the running mass of the lightest electrically charged
@@
   \amu^{\QED,\text{2L}} = \amu^{\BSM\text{,1L}} \times \Bigg[
-    16 \, \frac{\aem^\BSM(\MS)}{4\pi}
-    \log\frac{M_\mu}{\MS}
+    16 \, \frac{\aem^\BSM(\MLCP)}{4\pi}
+    \log\frac{M_\mu}{\MLCP}
   \Bigg] \,.
@@
 includes all known 2-loop corrections, while \famu performs the
-calculation in the \MSbar/\DRbar scheme at the SUSY scale and only
+calculation in the \MSbar/\DRbar scheme at the lightest charged BSM particle scale $\MLCP$ and only
 includes the leading logarithmic 2-loop QED correction.  A comparison
@@
 renormalization scale: \spheno performs the calculation at the $Z$ pole
-mass scale, while \famu performs it at $\MS$.
+mass scale, while \famu performs it at $\MLCP$.
 
@@
 column of \tabref{tab:test_famu} the variation of $\amuBSM$ in \famu
-when the scale is varied in the interval $[\MS/2,2\MS]$.  For BM4$^\prime$,
+when the scale is varied in the interval $[\MLCP/2,2\MLCP]$.  For BM4$^\prime$,
 $\amuBSM$ varies by around $66\%$, which indicates a very imprecise
@@
     in \tabref{tab:bm_amu}.  The third column shows the variation
-    of $\amuBSM$ when the renormalization scale is varied by a factor $2$.
+    of $\amuBSM$ when the renormalization scale is varied between
+    $\MLCP/2$ and $2\MLCP$.
     For BM4$^\prime$ the value of \amu calculated by the \DRbar
@@
 %
-at the scale $M_Z$,
+with the renormalization scale consistently being set to $M_Z$,
 where $\aem$ is the electromagnetic coupling of the \BSM model in the
@@
 to the value $\approx 246$ GeV in the corresponding \sarah model file.
+The utilized expansions of the function $\rho^{(2)}$ can be found in
+Ref.~\cite{Fleischer:1993ub}.
 
@@
 from diagrams with additional internal gauge bosons, is given by
-Eq.~(C.12) from Ref.~\cite{Pierce:1996zz} considering the replacement
-$\hat\rho \rightarrow 1/(1-\Delta\hat\rho)$.
+Eq.~(C.12) from Ref.~\cite{Pierce:1996zz} with the replacement
+$\hat\rho \rightarrow 1/(1-\Delta\hat\rho)$.\footnote{The expression
+  for $\deltaVB^\SM$ implemented in \fs has been extended by a term
+  expressing the scale dependence so that the calculation can be
+  performed consistently at the 1-loop level at scales different from
+  $M_Z$.  However, the 2-loop contributions
+  $\Delta\hat{r}_\text{2L}^\SM$ and $\Delta\hat\rho_\text{2L}^\SM$ are
+  omitted if $Q \neq M_Z$, because their explicit scale dependence is
+  currently not taken into account.}
 The BSM contribution $\deltaVB^\BSM$ contains corrections from
@@
 below the level of $0.1$\%.  Unit tests that perform these
-comparisons have also been added to the extensive \fs test suite that
-runs on a nightly basis.
+comparisons have also been added to the \fs test suite.
 %
@@
 solution, or multiple solutions are present \cite{Allanach:2013cda} with
-different stability properties in the two solvers.  We have checked that such
-points nevertheless satisfy the boundary conditions imposed at each
-scale and are indeed valid solutions.  More generally we have checked that
-solutions found in one solver also correspond to fixed point solutions of the
-other solver algorithm; that is, they satisfy all of the boundary conditions
-so that the parameter values remain unchanged after applying a single step of
-the iteration.
+different stability properties in the two solvers.  In this latter case, note
+that both solvers always return the first solution to
+which they converge,\footnote{When multiple convergent solutions exist,
+  the one first obtained will depend, for instance, on the initial guess
+  used for the iteration.} with the iteration
+stopping immediately once a convergent solution is obtained.  That is, neither
+method attempts to find all possible solutions for the given input parameters
+or automatically select between multiple fixed points.  Since a given solution
+might not be a stable fixed point of both iterations, the two solvers need not
+converge to the same solution, leading to the observed differences.  We have
+checked that such points nevertheless satisfy the boundary conditions imposed at
+each scale and are indeed valid solutions of the BVP.  More generally we have
+checked that solutions found in one solver also correspond to (not necessarily
+stable) fixed point solutions of the other solver algorithm; that is, they
+satisfy all of the boundary conditions so that the parameter values remain
+unchanged after applying a single step of the iteration.
 
@@
 the two-scale solver, it is also able to provide complementary coverage of the
-parameter space to that of the two-scale solver.  This is demonstrated in the left panel of
-\figref{fig:cmssm_solution_regions} in the CMSSM, where the solutions found
-by each solver are plotted in the $\mzero - \mu$ plane.
+parameter space to that of the two-scale solver.  This is demonstrated in the
+left panel of \figref{fig:cmssm_solution_regions} in the CMSSM, where the
+solutions found by each solver are plotted in the $\mzero - \mu$ plane.
 %
@@
     doc/plots/CMSSM_solution_regions/CMSSM_higgs_mass}
-  \caption{Left panel: CMSSM solutions found by the two-scale solver (blue diamonds)
-    and the semi-analytic solver (red circles) in the $\mzero - \mu$ plane.
-    For both solvers, $\mhalf \in [0,1]\unit{TeV}$ is randomly sampled,
-    while $\tan\beta = 10$ and $\azero = 0$ GeV\@.  The solutions obtained
-    using the two-scale solver are found by randomly sampling
-    $\mzero \in [0, 6]\unit{TeV}$ with $\sign \mu = 1$, and those for the
-    semi-analytic solver are found by randomly sampling $\mu \in [0,1150]
-    \unit{GeV}$.  Right panel: Change in the calculated lightest \CP-even
-    Higgs pole mass $M_{h_1}$ from a reference value of $121$ GeV as a
-    function of $\mu$ at fixed $\mzero = 4$ TeV for the two solvers.  The
-    solutions shown correspond to a vertical slice at fixed $\mzero$ in the
-    left-hand plot; $\mhalf$ is again varied in $[0,1]\unit{TeV}$,
-    $\tan\beta = 10$ and $\azero = 0$ GeV.}
+  \caption{Left panel: CMSSM solutions found by the two-scale solver
+    (blue diamonds) and the semi-analytic solver (red circles) in the
+    $\mzero - \mu$ plane.  For both solvers, $\mhalf \in [0,1]\unit{TeV}$
+    is randomly sampled, while $\tan\beta = 10$ and $\azero = 0$ GeV\@.
+    The solutions obtained using the two-scale solver are found by
+    randomly sampling $\mzero \in [0, 6]\unit{TeV}$ with $\sign \mu = 1$,
+    and those for the semi-analytic solver are found by randomly sampling
+    $\mu \in [0,1150] \unit{GeV}$.  Right panel: Change in the calculated
+    lightest \CP-even Higgs pole mass $M_{h_1}$ from a reference value of
+    $121$ GeV as a function of $\mu$ at fixed $\mzero = 4$ TeV for the two
+    solvers.  The solutions shown correspond to a vertical slice at fixed
+    $\mzero$ in the left-hand plot; $\mhalf$ is again varied in
+    $[0,1]\unit{TeV}$, $\tan\beta = 10$ and $\azero = 0$ GeV.}
   \label{fig:cmssm_solution_regions}
@@
 the two-scale solver is more effective for finding solutions with small
-$\mzero$.  In general terms, the regions of parameter space in which the two
-solvers are effective need not overlap, and the use of both in tandem allows
-for a more complete picture of the parameter space to be obtained.  Moreover,
-in the regions in which both solvers do find solutions, there is excellent
-agreement between the two algorithms.\footnote{Provided the same solution is
-  found if multiple solutions exist.}  This is illustrated in the right
-panel of \figref{fig:cmssm_solution_regions}, where the lightest \CP-even
-Higgs mass, expressed as the difference from a reference value of $121$ GeV,
-is plotted for fixed $\mzero = 4$ TeV\@.  If both solvers find a solution for
-this value of $\mzero$, the two values of the Higgs mass agree very well.
+$\mzero$.  This highlights the fact that, for some parameter points,
+the semi-analytic solver generated by \fs might not find a solution
+where the two-scale solver is able to, and vice versa.  As noted above, this
+can be due to the point in question having different stability properties
+under the two different algorithms.  For example, the CMSSM point with
+$m_0 = 125\unit{GeV}$, $M_{1/2} = 300\unit{GeV}$, $\tan\beta = 10$,
+$\sign \mu = 1$ and $A_0 = 0\unit{GeV}$  is successfully solved using the
+two-scale solver, with $\mu \approx 395\unit{GeV}$ required for correct EWSB.
+The same point, when run using the semi-analytic solver with the two-scale
+solution for $\mu$, fails to converge; for this choice of input parameters
+and initial guess, the iteration enters a periodic orbit in which the
+approximations to the solution are close to, but do not correspond to, the
+fixed point found by the two-scale solver.  To avoid the observed cyclic
+behaviour here, it is necessary to either fine-tune the provided input
+parameters or otherwise modify the initial guess made by \fs.  Due to the
+differing choice of input and output parameters between the two algorithms
+in general, certain regions of the parameter space can also be susceptible to
+high levels of numerical sensitivity in one approach and not the other,
+again necessitating significant fine-tuning.  In the CMSSM, for instance,
+large cancellations are typically required in
+order to produce a small value of $m_0$ when using the semi-analytic solver.
+In such cases, $\mu$ must be carefully fine-tuned to obtain a valid solution.
+On the other hand, when using the two-scale solver one has direct control over
+the value of $m_0$, and it is not necessary to fine-tune in order to obtain
+the desired small value of $m_0$.  The situation is reversed in regions of
+parameter space with small values of $\mu$ and large $m_0$, where one now
+has direct control over the fine-tuned value of $\mu$ in the semi-analytic
+solver but not in the two-scale solver.  In general terms, the regions of
+parameter space in which the two solvers are effective need not overlap, and
+the use of both in tandem allows for a more complete picture of the parameter
+space to be obtained.  Moreover, in the regions in which both solvers do find
+solutions, there is excellent agreement between the two algorithms.\footnote{
+  Provided the same solution is found if multiple solutions exist.}  This is
+illustrated in the right panel of \figref{fig:cmssm_solution_regions}, where
+the lightest \CP-even Higgs mass, expressed as the difference from a reference
+value of $121$ GeV, is plotted for fixed $\mzero = 4$ TeV\@.  If both solvers
+find a solution for this value of $\mzero$, the two values of the Higgs mass
+agree very well.
 
@@
 prediction of the fixed-order calculation for
-$\MS \lesssim 1\unit{TeV}$, because it neglects terms of
-$O(v^2/\MS^2)$.
+$\MS \lesssim 400\unit{GeV}$ in the shown scenario with $X_t = 0$,
+because of the neglected terms of $O(v^2/\MS^2)$.  In other scenarios
+the $O(v^2/\MS^2)$ terms may be important up to
+$\MS \approx 1\unit{TeV}$.
 %
@@
   missing, the corresponding parameter is set to zero.
+  The given CKM matrix elements can be accessed in the \fs model file
+  via the \texttt{CKM} symbol to set the Yukawa matrices at the
+  low-energy scale.  See \fs's \modelname{CMSSMCKM} model for an example.
 \item The PMNS matrix is given as input in the block
@@
   parameter is set to zero.
+  The given PMNS matrix elements can be accessed in the \fs model file
+  via the \texttt{PMNS} symbol to fix potential neutrino mass
+  parameters at the low-energy scale.
 \item For special applications such as the calculation of
@@
 \begin{itemize}
+\setlength{\itemindent}{1cm}
 \item[\texttt{FlexibleSUSY[0]},] \texttt{precisionGoal}: This option
@@
   solvers that have been enabled in the model file (see
-  \secref{sec:chosing_SAS}) are used, starting with the two-scale
-  solver.  If set to $1$, the two-scale solver is used if it has been
-  enabled in the model file.  If set to $2$, then the semi-analytic
-  solver is used if it has been enabled in the model file.  If a
-  solver that has not been enabled in the model file is chosen, \fs
-  stops with an error.
+  \secref{sec:chosing_SAS}) are used.  In this case, each of the
+  enabled solvers will be tried in turn, in the same order as given in
+  \code{FSBVPSolvers}, until a solution is found, at which point \fs
+  will return this solution and no further solvers are tried.  In the
+  event that no solver obtains a valid solution, \fs reports the status of the
+  last solver tried.  Non-zero values of \texttt{FlexibleSUSY[2]}
+  select a single solver to be used.  If set to $1$, the two-scale
+  solver is used if it has been enabled in the model file.  If set
+  to $2$, then the semi-analytic solver is used if it has been enabled
+  in the model file.  If a solver that has not been enabled in the model
+  file is chosen, \fs stops with an error.
 

### 2018-04-06 doc/flexiblesusy-2.0.tex
update FS-2.0 paper to version accepted by CPC
with some additional information in the program summary section

@@
 \newgeometry{top=2.5cm,left=3cm,right=3cm,bottom=4cm,footskip=3em}
-\section*{Program Summary}
+\section*{New version program summary}
 \noindent
 {\em Program title:} \fs\\[0.5em]
+{\em Licensing provisions:} GPLv3\\[0.5em]
 {\em Program obtainable from:} \url{http://flexiblesusy.hepforge.org/}\\[0.5em]
-{\em Distribution format:} tar.gz\\[0.5em]
 {\em Programming language:} C++, Wolfram/Mathematica, FORTRAN, Bourne shell\\[0.5em]
-{\em Computer:} Personal computer\\[0.5em]
 {\em Operating system:} Tested on FreeBSD, Linux, Mac OS X, Windows/Cygwin\\[0.5em]
@@
 {\em Optional external routines:} BLAS, Himalaya, LAPACK, LoopTools, SQLite, TSIL\\[0.5em]
-{\em Typical running time:} 0.02--0.2 seconds per parameter point\\[0.5em]
+{\em Journal reference of previous version:} Comput.Phys.Commun.\ 190 (2015) 139-172\\[0.5em]
+{\em Does the new version supersede the previous version?:} yes \\[0.5em]
+{\em Reasons for the new version:} Program extension including new models, observables and algorithms\\[0.5em]
+{\em Summary of revisions:} Extension to non-supersymmetric models
+(\fbsm), complex parameters (\fcpv), prediction of $W$ boson mass from
+muon decay (\fmw), calculation of anomalous magnetic moment of the
+muon (\famu), semi-analytic boundary value problem solver (\fsas),
+improved hybrid Higgs mass calculation (\feft).\\[0.5em]
 {\em Nature of problem:}

### 2018-04-28 doc/flexiblesusy-2.0.tex
changes made by CPC

@@
 From the directions of the contours it is clear
-that \fs correctly reproduces the behaviour of
+that \fs correctly reproduces the behavior of
 $\edm{e}$ depending only on the rephasing invariant $\arg(\mu \mhalf)$.
@@
 excluded, the figure illustrates that both \fs and \spheno show the
-same behaviour of the Higgs pole mass as a function of $\phi_\mu$.
+same behavior of the Higgs pole mass as a function of $\phi_\mu$.
 The $\phi_\mu$-independent shift of around $0.7\unit{GeV}$ between the
@@
 fixed point found by the two-scale solver.  To avoid the observed cyclic
-behaviour here, it is necessary to either fine-tune the provided input
+behavior here, it is necessary to either fine-tune the provided input
 parameters or otherwise modify the initial guess made by \fs.  Due to the
@@
   2-loop level, the problem repeats itself one order higher, i.e., the
-  term in Eq.~\eqref{eq:twolooplogsyt} is cancelled but similar terms of
+  term in Eq.~\eqref{eq:twolooplogsyt} is canceled but similar terms of
   next-to-leading logarithmic 3-loop order remain.}

### 2019-03-04 doc/flexiblesusy-new_features.tex

--- 2018-02-22 doc/flexiblesusy-2.0.tex
@@
 
-\usepackage[pdftitle={FlexibleSUSY 2.0: Extensions to investigate the phenomenology of
-  SUSY and non-SUSY models},
-  pdfauthor={Peter Athron, Markus Bach, Dylan Harries, Thomas Kwasnitza, Jae-hyeon Park,
-    Dominik Stockinger, Alexander Voigt, Jobst Ziebell},
+\usepackage[pdftitle={New features in FlexibleSUSY since version 2.0},
+pdfauthor={Peter Athron, Markus Bach, Dylan Harries, Wojciech
+  Kotlarski, Thomas Kwasnitza, Jae-hyeon Park, Dominik Stockinger,
+  Alexander Voigt, Jobst Ziebell},
 pdfkeywords={FlexibleSUSY,supersymmetry,spectrum,generator,MSSM,NMSSM,E6SSM},
 bookmarks=true, linktocpage, colorlinks=true, allbordercolors=white,
-  allcolors=fscolor]{hyperref}
-\biboptions{sort&compress}
+allcolors=fscolor]{hyperref} \biboptions{sort&compress}
 \allowdisplaybreaks
@@
 \newcommand{\sherpa}{\texttt{SHERPA}\@\xspace}
+\newcommand{\TSIL}{\texttt{TSIL}\@\xspace}
+\newcommand{\LoopTools}{\texttt{LoopTools}\@\xspace}
 \newcommand{\FAFC}{\texttt{FeynArts}\@\xspace}
@@
  \vspace*{0.5cm}
- \title{\Large\bf FlexibleSUSY 2.0: Extensions to investigate the phenomenology of SUSY and non-SUSY models}
+ \title{\Large\bf New features in FlexibleSUSY since version 2.0}
 
@@
 \author[adelaide,prague]{Dylan Harries}
+\author[dresden]{Wojciech Kotlarski}
 \author[dresden]{Thomas Kwasnitza}
@@
   \begin{abstract}
-    We document major new features and improvements of \fs, a
-    \mathematica and C++ package with a dependency on the external
-    package \sarah, that generates fast and precise spectrum
-    generators. The extensions presented here significantly increase
-    the generality and capabilities of the \fs package, which already
-    works with a wide class of models, while
-    maintaining an elegant structure and easy to use interfaces. The
-    \fbsm extension makes it possible to also create spectrum
-    generators for non-supersymmetric extensions of the Standard
-    Model. The \fcpv extension adds the option of complex parameters
-    to the spectrum generators, allowing the study of many interesting
-    models with new sources of \CP violation. \fmw computes the decay
-    of the muon for the generated model and thereby allows \fs to
-    predict the mass of the $W$ boson from the input parameters by
-    using the more precise electroweak input of $\{ G_F, M_Z, \aem \}$
-    instead of $\{ M_W, M_Z, \aem \}$. The \famu extension provides a
-    calculator of the anomalous magnetic moment of the muon in any
-    model \fs can generate a spectrum for.  \fsas introduces a new
-    solver for the boundary value problem which makes use of
-    semi-analytic expressions for dimensionful parameters to find
-    solutions in models where the classic two-scale solver will not
-    work such as the constrained E$_6$SSM\@. \feft is a hybrid
-    calculation of the Higgs mass which combines the virtues of both
-    effective field theory calculations and fixed-order calculations.
-    All of these extensions are included in \fstwo, which is released
-    simultaneously with this manual.%% \sarah, an external dependency
-    %% providing model dependent information, can be automaically
-    %% installed using an installation script distributed with \fs.
+    We document major new features and improvements of \fs
+    \cite{Athron:2014yba,Athron:2017fvs}, a \mathematica and C++
+    package with a dependency on the external package \sarah, that
+    generates fast and precise spectrum generators.
   \end{abstract}
@@
 {\em Optional external routines:} BLAS, Himalaya, LAPACK, LoopTools, SQLite, TSIL\\[0.5em]
-{\em Typical running time:} 0.02--0.2 seconds per parameter point\\[0.5em]
+{\em Journal reference of previous version:} Comput.Phys.Commun.\ 230 (2018) 145-217\\[0.5em]
+{\em Does the new version supersede the previous version?:} yes \\[0.5em]
+{\em Reasons for the new version:} Program extension including new models, observables and algorithms\\[0.5em]
+{\em Summary of revisions:} $\mu\to e\gamma$\\[0.5em]
 {\em Nature of problem:}
 %
-Determining the mass spectrum, mixings and further observables for an arbitrary extension
-of the Standard Model, input by the user. The generated code must find
-simultaneous solutions to constraints that are specified at two or
-more different renormalization scales, which are connected by
-renormalization group equations forming a large set of coupled
-first-order differential
+Determining the mass spectrum, mixings and further observables for an
+arbitrary extension of the Standard Model, input by the user. The
+generated code must find simultaneous solutions to constraints that
+are specified at two or more different renormalization scales, which
+are connected by renormalization group equations forming a large set
+of coupled first-order differential
 equations. \\[0.5em]
@@
 
-\newpage
-\section{Introduction}
-
-Popular and well studied new physics extensions of the Standard Model
-(SM) are coming under increasing pressure from the searches at the
-Large Hadron Collider (LHC) as well as other current experiments.  At
-the same time, there remain many outstanding physics problems that are
-not solved by the SM and require new physics.  For example, the origin
-of dark matter is still unexplained and it requires some new particle
-to fit the relic density.  In addition, the gauge structure and
-fractional charges of the SM have no explanation, but some
-observations hint at the possibility of a grand unified theory (GUT)
-from which these can be derived after the breakdown of the GUT gauge
-group.  Finally, the stability of the weak scale, which is $17$ orders
-of magnitude smaller than the Planck scale, has no explanation within
-the SM \cite{Weinberg:1975gm, Weinberg:1979bn,
-  Gildener:1976ai,Susskind:1978ms, tHooft:1980xss} and has been the
-driving motivation for the construction and study of concepts that go
-beyond the Standard Model (BSM), most notably supersymmetry (SUSY) \cite{Miyazawa:1966mfa,Miyazawa:1968zz,Ramond:1971gb,Golfand:1971iw,Gervais:1971ji,Neveu:1971rx,Neveu:1971iv,Volkov:1973ix,Wess:1973kz,Wess:1974tw,Salam:1974yz} which ensures that the quadratic corrections from fermions and bosons in the loop diagrams cancel at all orders in perturbation theory \cite{Dimopoulos:1981au,Witten:1981nf,Dine:1981za,Dimopoulos:1981zb,Kaul:1981hi}.
-
-These issues strongly motivate the development of new ideas and new
-models.  However, in principle there are a huge number of models that
-can solve some, or all, of these problems and many more that may be
-motivated by other principles that have not yet gained widespread
-interest in high energy physics (HEP)\@. Previously, most work has
-been done on the simplest variants, on scenarios that are easiest to
-test and on those that individual researchers consider to be the very
-best motivated models. Examples include the Minimal Supersymmetric
-Standard Model\footnote{For a review see Ref.\ \cite{Chung:2003fi}.}
-(MSSM), scalar singlet dark matter models (SSDM)
-\cite{Silveira:1985rk,McDonald:1993ex,Burgess:2000yq}, type II two
-Higgs doublet models (THDM-II) \cite{Lee:1973iz,Glashow:1976nt,
-  Donoghue:1978cj}, and universal extra dimensions
-\cite{Appelquist:2000nn}.  Now, it is becoming increasingly well
-motivated to also examine more complicated model variants, explore
-scenarios that are calculationally difficult or hard to observe, or
-motivated from an entirely new perspective.  Such advanced models can
-avoid phenomenological or conceptual difficulties of simpler models,
-and might provide explanations for the lack of experimental evidence
-for new physics at the LHC.
-
-Faced with this challenge, it is important to reduce
-the calculational hurdle required to explore a new model and
-look at its phenomenology as much as possible. This makes it easier for
-new ideas to get developed and for many more models to
-be studied together for much more general conclusions.
-
-\fs \cite{Athron:2014yba,Athron:2016fuq} already made a significant
-push in this direction, allowing the automatic creation of a spectrum
-generator for a very wide range of supersymmetric models.  \fs uses
-\sarah \cite{Staub:2010ty,Staub:2009bi,Staub:2010jh,Staub:2012pb,
-  Staub:2013tta} to obtain \mathematica expressions for model dependent
-components, the 2-loop renormalization group equations (RGEs),
-1-loop self energies, 1-loop tadpoles, mass matrices and
-electroweak symmetry breaking (EWSB) conditions. \fs then translates these
-expressions into C++ routines and embeds them inside a code structure
-for solving the boundary value problem (BVP)\@.  It also uses some numerical
-routines from \softsusy \cite{Allanach:2001kg,Allanach:2013kza} and is
-heavily unit-tested against the MSSM and Next-to-MSSM versions of
-\softsusy every night to ensure bugs are avoided in updates.
-
-Spectrum generators determine the pole masses and couplings of a
-particular model from assumptions about the parameters at some
-high scale, such as the grand unification scale, or directly from
-on-shell parameters or running parameters\footnote{Usually defined in
-  the \MSbar or \DRbar scheme.} given at the new physics scale.
-This requires computing self energies, tadpole corrections to the EWSB
-conditions and threshold corrections to SM-like gauge and Yukawa
-couplings.  In the case of high scale assumptions a BVP must be
-solved, which additionally requires integrating the RGEs for the model.
-These codes are essential for testing hypotheses about the parameters of
-the model, and determining if they lead to phenomenologically viable masses.
-
-Spectrum generators are widely used in studies of supersymmetry, and for the
-MSSM and the Next-to-MSSM\footnote{For a review of the model see
-  Refs.\ \cite{Ellwanger:2009dp, Maniatis:2009re}.} (NMSSM) there are
-a number of them that solve the BVP for particular choices of breaking
-mechanism inspired boundary conditions (\softsusy
-\cite{Allanach:2001kg,Allanach:2009bv, Allanach:2011de,
-  Allanach:2013kza, Allanach:2014nba,
-  Allanach:2016rxd,Allanach:2017hcf}, \suspect \cite{Djouadi:2002ze},
-\spheno \cite{Porod:2003um,Porod:2011nf}, \isasusy \cite{Baer:1993ae},
-\nmspec \cite{Ellwanger:2004xm,Ellwanger:2005dv, Ellwanger:2006rn} and
-\suseflav \cite{Chowdhury:2011zr}) and several more that start from \DRbar
-parameters at the SUSY scale or on-shell inputs (\FeynHiggs
-\cite{Heinemeyer:1998yj, Heinemeyer:1998np, Degrassi:2002fi,
-  Frank:2006yh, Hahn:2013ria, Drechsel:2016jdg, Bahl:2016brp,
-  Bahl:2017aev}, \NMSSMCalc \cite{Baglio:2013iia, King:2015oxa} and
-\CPsuperH \cite{Lee:2003nta, Lee:2007gn, Lee:2012wa}). This means that
-the parameter space of the MSSM (and to a lesser extent the NMSSM) has
-been extensively explored in a fast and reliable way.
-
-However, for other models few, if any, public software packages exist.
-Until recently this meant that one would have to spend a very long
-time writing and testing code in order to explore the parameter space
-of the model and do phenomenological investigations, and even then
-bugs are more likely if one uses a private tool rather than a well
-tested public one.  With the recent development of \fs and generated
-modules for \spheno from \sarah \cite{Porod:2003um,
-  Porod:2011nf,Staub:2017jnp,Staub:2009bi,Staub:2010jh,Staub:2012pb,
-  Staub:2013tta,Athron:2014yba} it is now possible to obtain spectrum
-generators in a much wider range of models.  \fs can generate spectrum
-generators of various kinds, with boundary conditions at some high
-scale or purely with low-energy input.  This push towards calculators
-that are not model specific has also happened for other major types of
-analyses.  For example, one may create new models in \micromegas
-\cite{Belanger:2001fz,
-  Belanger:2004yn,Belanger:2006is,Belanger:2008sj,Belanger:2010gh,
-  Belanger:2013oya,Belanger:2014vza} using \calchep \cite{Boos:1994xb,
-  Belyaev:2012qa} and \lanhep \cite{Semenov:1996es,Semenov:1998eb,
-  Semenov:2002jw,Semenov:2008jy, Semenov:2010qt} to study the relic
-density and direct and indirect detection of dark matter. \FeynRules
-\cite{Christensen:2008py,Alloul:2013bka} can
-also be used to generate the Feynman rules after inputting the
-Lagrangian, in a similar manner to \sarah and \lanhep. The output of
-these codes can be used in matrix element generators, such as \calchep,
-\madgraph
-\cite{Stelzer:1994ta,Maltoni:2002qb,Alwall:2007st,Alwall:2011uj,
-  Alwall:2014hca}, \whizard \cite{Moretti:2001zz,Kilian:2007gr},
-\sherpa \cite{Gleisberg:2003xi,Gleisberg:2008ta,Hoeche:2012yf}, \FAFC
-\cite{Hahn:1998yk,Hahn:2000kx,Hahn:2006zy,Hahn:2010zi,Hahn:2016ebn}
-and \helac \cite{Kanaki:2000ey,Cafarella:2007pc}, with showering and
-hadronization handled by the event generators \herwig
-\cite{Bahr:2008pv, Bellm:2015jjp}, \pythia
-\cite{Bengtsson:1987kr,Sjostrand:2014zea, Sjostrand:2006za} and
-\sherpa. Results from collider experiments can be applied to different
-BSM models using re-interpretation tools, with some examples being \higgsbounds
-\cite{Bechtle:2008jh,Bechtle:2011sb,Bechtle:2013wla, Bechtle:2015pma},
-\higgssignals \cite{HiggsSignals}, \checkmate
-\cite{Drees:2013wra,Dercks:2016npn}, \smodels \cite{Kraml:2013mwa,
-  Kraml:2014sna,Ambrogi:2017neo}, \fastlim \cite{Papucci:2014rja},
-\madanalysis\ \cite{Conte:2012fm,Conte:2014zja,Dumont:2014tja}, and
-the native re-interpretation of \colliderbit \cite{Balazs:2017moi}.
-Finally, very recently \GAMBIT \cite{Athron:2017ard} has been released
-which uses auto-generated code from \fs \cite{Workgroup:2017bkh} and
-\micromegas \cite{Workgroup:2017lvb}, along with many other packages,
-to perform global fits of a user implemented BSM model.
-
-Indeed \fs has been used extensively to study new physics models, on
-its own or in combination with some of the codes mentioned above, in a
-large number of physics studies, see, e.g.,
-Refs.\ \cite{Diessner:2014ksa,
-  Athron:2014pua,Athron:2015tsa,Diessner:2015yna,Staub:2015aea,Athron:2015vxg,
-  Diessner:2015iln,Bagnaschi:2015pwa,Staub:2016dxq,Athron:2016fuq,
-  deFlorian:2016spz,Athron:2016gor,Athron:2016qqb,Drechsel:2016htw,
-  Bagnaschi:2017xid,Athron:2017yua,Athron:2017kgt,Athron:2017qdc,
-  Workgroup:2017bkh,Chakravarty:2017hcy,Harlander:2017kuc,Ellis:2017hdw}.
-Nonetheless, in the first release of \fs there were still a number of
-limitations that restricted the models and phenomenology that could be studied
-and the precision of the calculations.
-
-In this paper, we document extensions to \fs, all now available in
-version 2.0, that make a substantial push further in expanding the
-number of models that can be explored and the observables that can be
-calculated within them.
-Now the list of models that can be investigated with
-\fstwo also includes non-supersymmetric models (\fbsm), those with
-complex parameters (\fcpv) and, with the new BVP solver
-\fsas, constrained versions of certain non-minimal SUSY models like the NMSSM
-\cite{Ellwanger:2009dp, Maniatis:2009re,Ellwanger:1993xa,Elliott:1994ht,
-  King:1995vk,Ellwanger:1995ru,Ellwanger:1996gw} and the Exceptional
-Supersymmetric Standard Model (\ESSM)
-\cite{King:2005jy,King:2005my,Athron:2010zz, Athron:2009ue,Athron:2009bs} that
-cannot be solved using the two-scale BVP solver
-\cite{Barger:1993gh} approach implemented in public spectrum
-generators.
-%
-In addition, \fstwo extends its repertoire of calculations.
-The user may now calculate the mass of the $W$ boson as a
-prediction of the model (\fmw) with the Fermi
-constant treated as an input, as well as the anomalous magnetic
-moment of the muon (\famu) and some electric dipole moments as
-an application of \fcpv. Furthermore, we document and update our \feft
-calculation of the Higgs pole mass, the physics of which has been discussed
-in Ref.~\cite{Athron:2016fuq}.  Besides these new physics features, we
-also document the new \mathematica interface.
-
-In \secref{sec:download} we give a quick start guide explaining how to
-download and compile the code using basic options.  In
-\secref{sec:Mathematica_interface} we present the \mathematica
-interface of \fs's spectrum generators and in
-\secref{sec:model_file_extensions} we describe the new model file
-extensions.  The
-following Sections \ref{sec:fbsm}--\ref{sec:feft} document
-each of the major extensions to \fs in a
-modular fashion so that readers
-can easily skip to a particular section.  These extensions are
-as follows.  \fbsm, documented in \secref{sec:fbsm}, allows \fs to work in
-non-SUSY models as well as SUSY models; the calculation of the
-anomalous magnetic moment of the muon is
-given in \secref{sec:famu}; \fcpv, described in
-\secref{sec:fcpv}, introduces complex parameters to \fs so that \CP
-violating effects may be considered; \fmw calculates the
-muon decay, allowing \fs to predict $M_W$, and
-is documented in \secref{sec:fmw}; a new BVP solver \fsas is
-introduced in \secref{sec:fsas} and finally the hybrid Higgs mass
-calculation, \feft, which combines the benefits of both effective field theory
-and fixed-order calculations is documented in \secref{sec:feft}.
-After this, we briefly summarize remaining limitations of the code in
-\secref{sec:limitations} before adding concluding remarks in
-\secref{sec:conclusions}.
-%
-In the appendices we provide a reference of all input parameters and
-configuration options of \fstwo as well as new models.
-
-\section{Quick start}
-\label{sec:download}
-
-\subsection{Requirements}
-
-The build process of a custom spectrum generator using \fs requires
-the following:
+\clearpage
+\section{External libraries}
+
+\subsection{New mandatory external libraries}
+
+Since version X.Y.Z \fs requires the following additional libraries to
+be installed:
 %
 \begin{itemize}
-\item \mathematica, version 7 or higher
-\item \sarah, version 4.11.0 or higher [\url{http://sarah.hepforge.org}]
-\item C++11 compatible compiler (\texttt{g++} 4.8.5 or higher,
-  \texttt{clang++} 3.8 or higher, \texttt{icpc} 15.0 or higher)
-\item FORTRAN compiler (\texttt{gfortran}, \texttt{ifort} etc.)
-\item Eigen library, version 3.1 or higher
-  [\url{http://eigen.tuxfamily.org}]
-\item Boost library, version 1.37.0 or higher
-  [\url{http://www.boost.org}]
-\item GNU scientific library [\url{http://www.gnu.org/software/gsl}]
+\item \LoopTools: required for calculating the 1-loop decays.
 \end{itemize}
-%
-Optional:
-%
+
+\subsection{New optional external libraries}
+
 \begin{itemize}
-\item an implementation of LAPACK [\url{http://www.netlib.org/lapack}]
-  such as ATLAS [\url{http://math-atlas.sourceforge.net}] or
-  Intel Math Kernel Library\\\ [\url{http://software.intel.com/intel-mkl}]
-\item LoopTools, version 2.8 or higher
-  [\url{http://www.feynarts.de/looptools}]
+\item \TSIL: Since version X.Y.Z \fs can calculate 2-loop corretions
+  to SM \MSbar top quark mass of the order $O(\at\as + \at^2)$ from
+  Ref.~\cite{Martin:2016xsp}.  The calculation of these corrections
+  requires linking the \TSIL library \cite{Martin:2005qm} to \fs.
 \end{itemize}
 
-\subsection{Downloading \fs and generating a first spectrum generator}
-\label{sec:quick-start-cmssm}
-\fs is available as gzipped tarball on
-[\url{http://flexiblesusy.hepforge.org}] or
-[\url{https://github.com/FlexibleSUSY/FlexibleSUSY}] under the version
-control system \texttt{git}.
-%
-To download and install \fstwo as a gzipped tarball run at the command
-line:
-%
-\begin{lstlisting}[language=bash]
-$ wget https://www.hepforge.org/archive/flexiblesusy/FlexibleSUSY-2.0.1.tar.gz
-$ tar -xf FlexibleSUSY-2.0.1.tar.gz
-$ cd FlexibleSUSY-2.0.1
-\end{lstlisting}%% $
-%
-\fstwo is distributed with a huge selection of predefined ``models'', 
-including several MSSM and NMSSM scenarios such as the CMSSM (called
-\modelname{CMSSM}), high-scale SUSY and split-SUSY (called \HSSUSY and
-\modelname{SplitMSSM}; these  models have been created for
-Ref.~\cite{Bagnaschi:2015pwa}), the semi-constrained and
-fully constrained NMSSM (\modelname{NMSSM}, \modelname{CNMSSM}). The
-distribution also contains BSM models such as the
-$R$-symmetric MSSM (\modelname{MRSSM}; for a definition of the model
-see Ref.~\cite{Kribs:2007ac}), the NUHM \ESSM
-(\modelname{E6SSM}, see \cite{King:2005jy,King:2005my,Athron:2010zz,
-  Athron:2009ue,Athron:2009bs,Athron:2007en}) and the two-Higgs 
-doublet model type II
-(\modelname{THDMII}).  See the 
-contents of \code{model_files/} for all predefined model files.
-
-A spectrum generator for any of these models can be built with just
-three commands.  For example, the CMSSM spectrum generator can be
-created by running the following shell commands:
-%
-\begin{lstlisting}[language=bash]
-$ ./createmodel --name=CMSSM
-$ ./configure --with-models=CMSSM
-$ make
-\end{lstlisting}%% $
-%
-The \code{createmodel} command creates the model directory
-\code{models/CMSSM/} where the code will be generated and adds the
-CMSSM model file and an example SLHA input file to it.  The \code{configure}
-script sets up the \code{Makefile}, checking the system requirements
-and dependencies.  For more options see \code{./configure
-  --help}.  The last command creates the code for the spectrum
-generator of this model and compiles it.
-%
-This generated spectrum generator can then be run from the command
-line as:
-%
-\begin{lstlisting}[language=bash]
-$ cd models/CMSSM
-$ ./run_CMSSM.x --slha-input-file=LesHouches.in.CMSSM \
-                --slha-output-file=LesHouches.out.CMSSM
-\end{lstlisting}
-%
-The spectrum generator reads the CMSSM input parameters from the SLHA input file
-\code{LesHouches.in.CMSSM} and first solves a BVP to find a
-set of running parameters at the SUSY scale that are consistent with all
-boundary conditions specified in the model file, and then calculates the pole
-masses, mixing matrices and potentially further observables.
-The mass spectrum etc.\ obtained in this way is output
-in SLHA format \cite{Skands:2003cj,Allanach:2008qq} as
-\code{LesHouches.out.CMSSM}.  See \code{./run_CMSSM.x --help} for more
-options.
-
-\fs also provides a \mathematica interface, introduced in version 1.7.0,
-to call the generated spectrum generators.  For each spectrum
-generator, an example \mathematica script named
-\code{models/<model>/run_<model>.m} is created for illustration.  For
-example, the \modelname{CMSSM} spectrum generator can be called from within
-\mathematica like this:
-%
-\begin{lstlisting}
-Get["models/CMSSM/CMSSM_librarylink.m"];
-
-handle = FSCMSSMOpenHandle[
-    fsModelParameters -> {
-        m0 -> 125,
-        m12 -> 500,
-        TanBeta -> 10,
-        SignMu -> 1,
-        Azero -> 0
-    }
-];
-
-spectrum = FSCMSSMCalculateSpectrum[handle];
-FSCMSSMCloseHandle[handle];
-
-Print[spectrum];
-\end{lstlisting}
-%
-Execute \code{?FSCMSSMOpenHandle} for a list of all allowed options and
-input parameters.  In the example above, the \code{spectrum} variable
-contains the mass spectrum and the running parameters in the form of a
-list of replacement rules:
-%
-\begin{lstlisting}
-{CMSSM -> {
-   Pole[M[Glu]] -> 1147.35,
-   Pole[M[Sd]] -> {957.993, 997.56, 1000.49, 1000.5, 1045.93, 1045.94},
-   Pole[M[Sv]] -> {350.753, 351.913, 351.917},
-   Pole[M[Su]] -> {796.653, 1002.67, 1003.96, 1005.06, 1043.07, 1043.07},
-   Pole[M[Se]] -> {222.916, 229.983, 230.008, 360.842, 360.846, 361.978},
-   Pole[M[hh]] -> {114.836, 713.119},
-   Pole[M[Ah]] -> {88.5864, 712.848},
-   Pole[M[Hpm]] -> {77.2642, 717.628},
-   Pole[M[Chi]] -> {204.054, 385.012, 629.649, 643.612},
-   Pole[M[Cha]] -> {385.017, 643.924},
-   Pole[M[VWm]] -> 80.3935, ...}
-}
-\end{lstlisting}
-%
-More details about the \mathematica interface as well as a neat
-example of running \HSSUSY through it can be found in Sections
-\ref{sec:Mathematica_interface} and \ref{sec:application_HSSUSY},
-respectively.
-
-\subsection{Spectrum generators for alternative models}
-\label{sec:quick-start-alternative-models}
-If the user instead wants to create a spectrum generator for a model
-for which there is no pre-existing model file distributed in
-\fs, then the model file can be written. Before a \fs
-model file can be written for the spectrum generator, there must exist
-\sarah model files, which \fs uses to obtain model dependent
-information. \sarah also comes with many pre-defined models, but if an
-appropriate model is not available, the users may create their own
-\sarah model files and add them to the directory
-\code{sarah/<modelname>/} in \fs.  For the writing of a \sarah model
-file we refer the reader to the extensive \sarah documentation, for
-example Refs.\
-\cite{Staub:2017jnp,Staub:2009bi,Staub:2010jh,Staub:2012pb,
-  Staub:2013tta,Athron:2014yba}.
-
-Creating a new \fs model file is straightforward. Full details are
-given in the original \fs manual Ref.~\cite{Athron:2014yba}.
-Here we just repeat a basic example
-in the context of the NMSSM that illustrates the main points:
-%
-The semi-constrained NMSSM (\modelname{NMSSM}) distributed in \fs has
-all the soft-breaking trilinear scalar couplings set to a unified
-$\azero$ at the GUT scale.  However, often $A_\lambda$ and $A_\kappa$ are
-taken to be non-universal in semi-constrained variants of the NMSSM
-since the non-universality of the soft singlet mass already violates
-the standard universality assumptions of constrained
-models.\footnote{See \secref{sec:fsas} for a new approach that
-  allows the fully constrained NMSSM to be solved.} To allow separate
-values for $A_\lambda$ and $A_\kappa$ at the GUT scale the \fs model
-file \code{model_files/NMSSM/FlexibleSUSY.m.in} should be changed from
-%
-\begin{lstlisting}[language=Mathematica]
-EXTPAR = { {61, LambdaInput} };
-
-HighScaleInput = {
-  ...
-  {T[\[Kappa]], Azero \[Kappa]},
-  {T[\[Lambda]], Azero LambdaInput}
-   ...
-};
-\end{lstlisting}
-%
-into
-%
-\begin{lstlisting}[language=Mathematica]
-EXTPAR = { {61, LambdaInput},
-           {63, ALambdaInput},
-           {64, AKappaInput} };
-
-HighScaleInput = {
-  ...
-  {T[\[Kappa]], AKappaInput \[Kappa]},
-  {T[\[Lambda]], ALambdaInput LambdaInput},
-  ...
-};
-\end{lstlisting}
-%
-The GUT scale values of $A_\lambda$ and $A_\kappa$ can then be specified in the SLHA input file in the \code{EXTPAR} block by entries $63$ and $64$,
-%
-\begin{lstlisting}
-Block EXTPAR
-   61   0.1                  # LambdaInput
-   63   -100                 # ALambdaInput
-   64   -300                 # AKappaInput
-
-\end{lstlisting}
-
-\section{\mathematica interface}
-\label{sec:Mathematica_interface}
-
-The spectrum generators created with \fs can be called from within
-\mathematica.  To do that, first the spectrum generator must be built,
-as described in \secref{sec:quick-start-cmssm}.  Afterwards, the
-provided \mathematica interface functions for the model must be
-loaded.  For a given model \code{<model>} this is done by including
-the following file in the \mathematica session:
-%
-\begin{lstlisting}
-Get["models/<model>/<model>_librarylink.m"];
-\end{lstlisting}
-%
-This script loads the library
-\code{models/<model>/<model>_librarylink.so} into the \mathematica
-session (assuming the user is in the \code{FlexibleSUSY/} directory).
-Afterwards, the \mathematica interface functions listed in
-\tabref{tab:mma_ifce_functions} are available.
-%
-\begin{table}[tbh]
-  \centering
-  \begin{tabularx}{\textwidth}{lX}
-    \toprule
-    Function & Description \\
-
-    \midrule \code{FS<model>OpenHandle[...]} & Takes all model input
-    parameters as argument and returns a ``handle'' (a reference) to
-    the given parameter point, the associated mass spectrum and observables.\\
-
-    \code{FS<model>CloseHandle[handle]} & Releases the resources
-    associated to a given handle.\\
-
-    \code{FS<model>CalculateSpectrum[handle]} & Calculates the mass
-    spectrum for a given handle.\\
-
-    \code{FS<model>CalculateObservables[handle]} & Calculates the
-    observables for a given handle.\\
-
-    \code{FS<model>ToSLHA[handle]} & Returns a string containing the
-    mass spectrum and observables associated to a given handle in SLHA format.\\
-
-    \code{FS<model>Set[handle, ...]} & Changes the input parameters
-    associated to a given handle.\\
-
-    \code{FS<model>GetSettings[handle]} & Returns the spectrum generator
-    settings (precision goal, loop orders, etc.) associated to
-    a given handle.\\
-
-    \code{FS<model>GetSMInputParameters[handle]} & Returns the SM
-    input parameters associated to a given handle.\\
-
-    \code{FS<model>GetInputParameters[handle]} & Returns the
-    model-specific input parameters associated to a given handle.\\
-
-    \code{FS<model>GetProblems[handle]} & Returns a list of problems
-    that occurred when calculating the spectrum.\\
-
-    \code{FS<model>GetWarnings[handle]} & Returns a list of warnings
-    that occurred when calculating the spectrum.\\
-    \bottomrule
-  \end{tabularx}
-  \caption{\mathematica interface functions provided for a \fs model with the name \code{<model>}.}
-  \label{tab:mma_ifce_functions}
-\end{table}
-%
-
-To run the spectrum generator for a given parameter point, a handle to that point must be
-created first, using the \code{FS<model>OpenHandle[...]} function.
-%
-The returned handle represents a reference to the given parameter
-point, the associated mass spectrum and observables.  The concept of
-handles allows the user to calculate mass spectra for different
-parameter points in parallel using multiple \mathematica kernels: Each
-kernel can open a handle to a separate parameter point, calculate the
-mass spectrum and finally close the handle.  In this way there
-is no ambiguity in the parameter point used by each kernel.
-\exref{ex:HSSUSY_parallel} in \secref{sec:application_HSSUSY} illustrates the usage of handles by
-performing a parallel scan over the MSSM parameter space with \HSSUSY.
-%
-In the most general form, the \code{FS<model>OpenHandle[...]} function
-can take the following three arguments:
-%
-\begin{lstlisting}
-handle = FS<model>OpenHandle[
-   fsSettings -> {...},
-   fsSMParameters -> {...},
-   fsModelParameters -> {...}
-]
-\end{lstlisting}
-%
-The \code{fsSettings} symbol can be used to set the spectrum generator
-options.  All possible options are listed in
-\tabref{tab:sg_mma_options} in \appref{app:fs_config_options}.
-%
-The \code{fsSMParameters} symbol can be used to set the SM
-input parameters.  The possible SM input parameters are
-listed in \tabref{tab:sm_mma_options} in \appref{app:sm_input_parameters}.
-%
-The \code{fsModelParameters} symbol can be used to set the \BSM
-model-specific input parameters.  Unspecified model input parameters
-are set to zero by default.  The names of the model input parameters
-are identical to the ones specified in the \code{MINPAR},
-\code{IMMINPAR}, \code{EXTPAR}, \code{IMEXTPAR} and
-\code{FSAuxiliaryParameterInfo} variables in the \fs model file.  The
-settings, the SM and the \BSM input parameters
-associated to a given handle can be obtained using the
-\code{FS<model>GetSettings[]}, \code{FS<model>GetSMInputParameters[]}
-and \code{FS<model>GetInputParameters[]} functions, respectively.  The
-opened handle can then be used to calculate the mass spectrum and the
-observables:
-%
-\begin{lstlisting}
-spectrum    = FS<model>CalculateSpectrum[handle];
-observables = FS<model>CalculateObservables[handle];
-FS<model>CloseHandle[handle];
-\end{lstlisting}
-%
-Finally, the handle should be closed to release the associated
-resources by calling the function \code{FS<model>CloseHandle[handle]}.
-%
-\begin{example}
-  In the CMSSM, the \BSM model-specific input parameters are named as
-  \code{m0}, \code{m12}, \code{TanBeta}, \code{SignMu} and
-  \code{Azero}, see the \modelname{CMSSM} model file provided with \fstwo.
-  Thus, an example \mathematica session for the CMSSM could look like:
-%
-\begin{lstlisting}
-Get["models/CMSSM/CMSSM_librarylink.m"];
-
-handle = FSCMSSMOpenHandle[
-    fsSettings -> {
-        poleMassLoopOrder -> 2,
-        ewsbLoopOrder -> 2,
-        thresholdCorrectionsLoopOrder -> 2,
-        betaFunctionLoopOrder -> 3
-    },
-    fsSMParameters -> {
-        Mt -> 173.34,
-        alphaSMZ -> 0.1184
-    },
-    fsModelParameters -> {
-        m0 -> 125,
-        m12 -> 500,
-        TanBeta -> 10,
-        SignMu -> 1,
-        Azero -> 0
-    }
-];
-
-spectrum    = FSCMSSMCalculateSpectrum[handle];
-observables = FSCMSSMCalculateObservables[handle];
-
-FSCMSSMCloseHandle[handle];
-\end{lstlisting}
-\end{example}
-%
-The output of \code{FS<model>CalculateSpectrum[handle]} is a list
-that contains the pole mass spectrum as well as the running masses
-and parameters at the chosen output scale.
-%
-The running parameters are named as defined in the \sarah model.  For
-example, \code{g1}, \code{g2}, \code{g3} usually denote the running
-gauge couplings and \code{Yu}, \code{Yd}, \code{Ye} the running Yukawa
-couplings.  The running masses are denoted as \code{M[<p>]}, where
-\code{<p>} is the name of the particle as defined in the \sarah model
-file.  All running parameters and masses are given at the parameter
-output scale, \code{SCALE}.  The pole masses and mixing matrices carry
-the additional \code{Pole[]} head.  For example, \code{Pole[M[hh]]}
-usually denotes the pole mass(es) of the Higgs boson(s).
-
-In the \mathematica output, the running parameters, the masses and
-mixing matrices are defined in the \sarah convention, \emph{not} in the
-SLHA convention.  This means in particular that the Yukawa matrices,
-the soft-breaking squark mass matrices and the soft-breaking trilinear
-couplings are \emph{not} defined in the (super)-CKM basis.  In the
-\mathematica output, the particle masses are always non-negative and mixing
-matrices are in general complex.
-%
-\begin{example}
-  In the CMSSM, the output of \code{FSCMSSMCalculateSpectrum[handle]}
-  may look like (skipping some entries for brevity):
-%
-\begin{lstlisting}
-{
-  CMSSM -> {
-    M[Glu] -> 1116.4857717819132,
-    M[Sd] -> {929.5770939936384, 963.6803089181217,
-              965.7750791635142, 965.7786645820956,
-              1010.5301444258299, 1010.5317308607175},
-    M[Su] -> {770.2929836944288, 969.5389940603632,
-              969.5446720936367, 975.4073015489867,
-              1007.5435846398124, 1007.5443071514123},
-    M[Se] -> {219.54939719808144, 226.45230058860656,
-              226.4768407746941, 356.2452631868304,
-              356.2501661663376, 357.5576772510361},
-    M[hh] -> {88.16467333922309, 726.2603417238729},
-    M[Ah] -> {90.09835220027803, 726.0229889725828},
-    M[Hpm] -> {78.48914789145176, 730.2533306006959},
-    M[Chi] -> {207.1963755793879, 375.7416364302936,
-               627.5178023483583, 641.6676783271736},
-    M[Cha] -> {375.56991892585705, 641.3578484531205},
-    ...
-    ZH -> {{0.10592570722508611, 0.9943740465985952},
-           {0.9943740465985952, -0.10592570722508611}},
-    ...
-    Pole[M[Glu]] -> 1147.3536227374905,
-    Pole[M[Sd]] -> {957.9934299811302, 997.5603867095314,
-                    1000.4932601265115, 1000.4969819618583,
-                    1045.9354429433467, 1045.9372472457565},
-    Pole[M[Su]] -> {796.653619369782, 1002.6690741336473,
-                    1003.9614916607435, 1005.0642702137084,
-                    1043.0672831732345, 1043.067920812505},
-    Pole[M[Se]] -> {222.90126096766593, 229.9832415178622,
-                    230.00840279144913, 360.84198174065307,
-                    360.8462569384804, 361.9798562942742},
-    Pole[M[hh]] -> {114.83583179574276, 713.1187313487922},
-    Pole[M[Ah]] -> {88.58641341930426, 712.8473602456997},
-    Pole[M[Hpm]] -> {77.26414997655887, 717.6270868215212},
-    Pole[M[Chi]] ->  {204.05370940499517, 385.0116889026496,
-                      629.6500252267041, 643.6127224060953},
-    Pole[M[Cha]] -> {385.0164604772902, 643.924798526633},
-    Pole[ZH] -> {{0.1066307364997843, 0.9942986905520461},
-                 {0.9942986905520461, -0.1066307364997843}},
-    ...
-    Yd -> {{0.00013999141660755535, 0., 0.},
-           {0., 0.0030650771019348657, 0.},
-           {0., 0., 0.1317656023934078}},
-    Ye -> {{0.00002895462310608183, 0., 0.},
-           {0., 0.005986898727723582, 0.},
-           {0., 0., 0.1006931798596906}},
-    Yu -> {{7.267255094462218*^-6, 0., 0.},
-           {0., 0.0033082824973678956, 0.},
-           {0., 0., 0.8606532901364391}},
-    \[Mu] -> 624.160899893032,
-    g1 -> 0.4679063156949638,
-    g2 -> 0.6430285180350706,
-    g3 -> 1.0655340318624051,
-    vd -> 25.099612589273388,
-    vu -> 242.8296409176676,
-    T[Yd] -> {{-0.19442921534444055, 0., 0.},
-              {0., -4.256965123994014, 0.},
-              {0., 0., -171.13078241755716}},
-    T[Ye] -> {{-0.008660431147847696, 0., 0.},
-              {0., -1.790668166947902, 0.},
-              {0., 0., -29.953008563262035}},
-    T[Yu] -> {{-0.008251796354121698, 0., 0.},
-              {0., -3.7564603713866354, 0.},
-              {0., 0., -755.7309107649228}},
-    B[\[Mu]] -> 53907.68839928095,
-    mq2 -> {{1.0178399270038805*^6, 0., 0.},
-            {0., 1.0178348758707164*^6, 0.},
-            {0., 0., 865711.3590482193}},
-    ml2 -> {{124853.43144557778, 0., 0.},
-            {0., 124850.91737364854, 0.},
-            {0., 0., 124143.31667980803}},
-    mHd2 -> 109509.1756551005,
-    mHu2 -> -377534.5544643501,
-    md2 -> {{932089.8366766714, 0., 0.},
-            {0., 932084.7501918572, 0.},
-            {0., 0., 923097.970074961}},
-    mu2 -> {{941294.0634724408, 0., 0.},
-            {0., 941288.914981822, 0.},
-            {0., 0., 639354.6597906639}},
-    me2 -> {{49375.97115879859, 0., 0.},
-            {0., 49370.8410356076, 0.},
-            {0., 0., 47926.68179837005}},
-    MassB -> 209.15138268684439,
-    MassWB -> 387.9365053016521,
-    MassG -> 1116.4857717819132,
-    SCALE -> 866.8060753250803
-  }
-}
-\end{lstlisting}
-\end{example}
-%
-The function \code{FS<model>CalculateObservables[handle]} returns the
-observables for a given handle in the form of a list of replacement rules.
-%
-\begin{example}
-  In the CMSSM, the output of \code{FSCMSSMCalculateObservables[handle]}
-  might look like
-%
-\begin{lstlisting}
-{
-  CMSSM -> {
-    FlexibleSUSYObservable`aMuon -> 8.46316731749956*^-10,
-    FlexibleSUSYObservable`CpHiggsPhotonPhoton ->
-      {0.000029645107712580034 - 2.1094207517495443*^-7*I,
-       7.984875731996049*^-7 + 9.125974009524432*^-7*I},
-     FlexibleSUSYObservable`CpHiggsGluonGluon ->
-       {-0.00006704582692080136 - 2.68821584064419*^-6*I,
-        2.8335074218510527*^-6 + 4.966199588265877*^-6*I},
-     FlexibleSUSYObservable`CpPseudoScalarPhotonPhoton ->
-       1.066762880186648*^-6 - 8.27198259619208*^-7*I,
-     FlexibleSUSYObservable`CpPseudoScalarGluonGluon ->
-       6.825816917497379*^-6 + 8.151893730825134*^-7*I
-  }
-}
-\end{lstlisting}
-\end{example}
-%
-The symbol \code{aMuon} represents the anomalous magnetic moment of
-the muon, $\amu$, calculated as described in \secref{sec:famu}.  The
-symbols \code{CpHiggsPhotonPhoton}, \code{CpHiggsGluonGluon},
-\code{CpPseudoScalarPhotonPhoton} and \code{CpPseudoScalarGluonGluon}
-denote the effective couplings of the physical \CP-even and \CP-odd Higgs
-boson(s) to two photons and gluons, respectively, as described in
-Ref.~\cite{Staub:2016dxq}.  Note that if the \CP-even or \CP-odd Higgs states
-are multiplets, as is the case for the \CP-even Higgs in the MSSM, for
-example, the relevant couplings are calculated for all members of the
-multiplet and the result is returned as a list.
-
-The calculated spectrum can be printed in an SLHA-compatible format
-using the \code{FS<model>ToSLHA[]} function.
-%
-\begin{example}
-  For the CMSSM, an example output of \code{FSCMSSMToSLHA[handle]}
-  could look like:
-\begin{lstlisting}
-Block SPINFO
-     1   FlexibleSUSY
-     2   2.0.1
-     5   CMSSM
-     9   4.11.0
-Block MASS
-   1000021     1.14735362E+03   # Glu
-        24     8.03935152E+01   # VWm
-   1000024     3.85016460E+02   # Cha(1)
-   1000037     6.43924799E+02   # Cha(2)
-        25     1.14835832E+02   # hh(1)
-        35     7.13118731E+02   # hh(2)
-...
-\end{lstlisting}
-\end{example}
-%
-In \secref{sec:application_HSSUSY} several examples can be found that
-illustrate how to perform parameter scans and uncertainty
-estimates using the \mathematica interface of \fs's spectrum
-generators.
-
-\section{\fs model file extensions}
-\label{sec:model_file_extensions}
-
-\subsection{Model-specific higher-order contributions}
-\label{sec:model_specific_contributions}
-To improve the accuracy in some specific models, \fs provides a few
-model file switches to enable further higher-order contributions in
-the calculation of the running parameters, the $\beta$ functions or
-the Higgs pole mass.  \tabref{tab:fbsm_options} lists all switches
-available in \fstwo, which are explained in the following subsections.
-%
-These switches can usually be enabled in many models, providing that
-the corresponding requirements are fulfilled.  However, the user
-should be aware that contributions may be missing if the switches are
-enabled in models beyond their scope of application.
-
-\begin{OptionTable}{\fs model file switches to enable/disable
-    model-specific higher-order contributions in the SM,
-    MSSM, NMSSM and split-MSSM.}{tab:fbsm_options}[lllX]
-  %
-  \multicolumn{4}{c}{\textit{SM}}\\
-  \midrule
-  \code{UseHiggs2LoopSM} & \code{False} & \code{True} or \code{False}
-  & 2-loop contributions $O(\at^2 + \at \as)$ to $M_h$ in the SM \\
-  %
-  \code{UseHiggs3LoopSM} & \code{False} & \code{True} or \code{False} &
-  3-loop contributions $O(\at^3 + \at^2 \as \allowbreak + \at \as^2)$ to $M_h$ in
-  the SM \\
-  %
-  \code{UseSM3LoopRGEs} & \code{False} & \code{True} or \code{False} &
-  3-loop RGEs in the SM \\
-  %
-  \code{UseYukawa3LoopQCD} & \code{Automatic} & \multilinecell[t]{\code{True} or \code{False}\\ or \code{Automatic}} &
-  2-loop and 3-loop QCD contributions $O(\as^2 + \as^3)$ to the \MSbar $y_t$ in the SM \\
-  %
-  \code{UseSMAlphaS3Loop} & \code{False} & \code{True} or \code{False}
-  & 2-loop and 3-loop QCD threshold corrections
-  $O(\as^2 + \as^3)$
-  to the \MSbar $\as$ in the SM \\
-  %
-  \midrule
-  \multicolumn{4}{c}{\textit{MSSM}}\\
-  \midrule
-  %
-  \code{UseHiggs2LoopMSSM} & \code{False} & \code{True} or \code{False} &
-  2-loop contributions $O((\at + \ab)\as \allowbreak + (\at +
-  \ab)^2 \allowbreak + \atau^2)$ to $M_h$, $M_H$ and $M_A$ in
-  the MSSM \\
-  %
-  \code{UseHiggs3LoopMSSM} & \code{False} & \code{True} or
-  \code{False} & 3-loop contributions
-  $O(\at\as^2 \allowbreak + \ab\as^2)$ to $M_h$ in
-  the MSSM (requires \Himalaya) \\
-  %
-  \code{UseMSSM3LoopRGEs} & \code{False} & \code{True} or \code{False} &
-  3-loop RGEs in the MSSM \\
-  %
-  \code{UseMSSMYukawa2Loop} & \code{False} & \code{True} or \code{False} &
-  2-loop \SQCD contribution $O(\as^2)$ to the \DRbar $y_b$ and $y_t$ in the MSSM \\
-  %
-  \code{UseMSSMAlphaS2Loop} & \code{False} & \code{True} or
-  \code{False} & 2-loop \SQCD contribution
-  $O(\as^2 + \as\at + \as\ab)$
-  to the \DRbar $\as$ in the MSSM \\
-  %
-  \midrule
-  \multicolumn{4}{c}{\textit{NMSSM}}\\
-  \midrule
-  %
-  \code{UseHiggs2LoopNMSSM} & \code{False} & \code{True} or \code{False} &
-  2-loop contributions $O((\at + \ab)\as \allowbreak + (\at +
-  \ab)^2 \allowbreak + \atau^2)$ to $M_{h_i}$ and $M_{A_i}$ in
-  the NMSSM \\
-  %
-  \midrule
-  \multicolumn{4}{c}{\textit{split-MSSM}}\\
-  \midrule
-  %
-  \code{UseHiggs3LoopSplit} & \code{False} & \code{True} or \code{False} &
-  3-loop contributions $O(\at\as^2)$ to $M_h$ in the split-MSSM \\
-  %
-  \bottomrule
-\end{OptionTable}
-
-%%%%%%%%%%%%%%%%%%%%%% SM corrections %%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{SM-specific higher-order contributions}
-
-\paragraph{2-loop and 3-loop contributions to the SM Higgs mass}
-
-In the SM, 2-loop and leading 3-loop contributions to the Higgs pole
-mass are known in the \MSbar scheme
-\cite{Degrassi:2012ry,Vega:2015fna,Martin:2014cxa}.  In \fs the 2-loop
-contributions of $O(\at^2 + \at \as)$
-\cite{Vega:2015fna,Martin:2014cxa} and the 3-loop
-$O(\at^3 + \at^2 \as + \at \as^2)$ contributions \cite{Martin:2014cxa}
-can be taken into account to calculate the SM-like Higgs pole mass for
-non-SUSY models with only one Higgs. In order to enable these loop
-contributions, the following switches must be set in the model file:
-%
-\begin{lstlisting}
-UseHiggs2LoopSM = True;
-UseHiggs3LoopSM = True;
-\end{lstlisting}
-%
-The 2-loop and 3-loop
-contributions enter the mass of the Higgs boson as
-\begin{align} 
- \begin{split}
-   M_h^2 &= m_h^2 + (\Delta m_h^2)_{1L}(p^2)
-   + (\Delta m_h^2)_{2L}(p^2) +(\Delta m_h^2)_{3L} \,,
- \end{split} \label{eq:Mh_polp^2}
-\end{align}
-where $m_h^2$ and $(\Delta m_h)_{1L}(p^2)$ correspond to the tree-level 
-and a 1-loop expression, respectively.
-The enabled 2-loop SM contributions of
-$O(\at^2 + \at \as)$ to the Higgs mass read
-%
-\begin{align}
-  (\Delta m_h^2)_{2L}(p^2) &= (\Delta m_h^2)_{2L}^{(\at^2)} + 
-     (\Delta m_h^2)_{2L}^{(\at\as)}(p^2) \,, \\
-  (\Delta m_h^2)_{2L}^{(\at^2)} &=
-  \frac{2 t}{(4\pi)^4} \left[
-    -3 y_t^4 \left(3 \;\barlog^2(t) - 7 \;\barlog(t) + 2 + \frac{\pi^2}{3}\right)
-  \right] , \label{eq:2L_Mh_atat}\\
-  (\Delta m_h^2)_{2L}^{(\at\as)} (p^2)&=
-      \frac{g_3^2 y_t^2}{(4\pi)^4} \;\bigg[
-        \frac{37}{3} p^2 - \frac{122 p^4}{135 t} - 4 (5 p^2 - 8 t) \;\barlog(t) -
-        12 (p^2 - 8 t) \;\barlog^2(t) \bigg] \label{eq:2L_Mh_atas} \\
-  &\stackrel{\mathmakebox[\widthof{=}]{p^2=0}}{=} \
-  \frac{2 t}{(4\pi)^4} \left[
-    16 g_3^2 y_t^2 \left(3 \;\barlog^2(t) + \barlog(t)\right)
-  \right] \,.
-  \label{eq:2L_Mh_atas p2=0}
-\end{align}
-%
-In Eqs.~\eqref{eq:2L_Mh_atat} ff.\ the abbreviations $t \equiv m_t^2$
-and $\barlog(t) \equiv \log(t/Q^2)$ have been used for brevity, 
-analogous to the notation of Ref.~\cite{Martin:2014cxa}, where $m_t$ 
-is the \MSbar top mass in the SM\@. 
-By performing a momentum iteration in the computation of the Higgs 
-pole mass, the momentum entering the 1-loop self energy in 
-Eq.~\eqref{eq:Mh_polp^2} consists of a tree-level part and a loop 
-correction.
-In combination with Eq.~\eqref{eq:2L_Mh_atat}, the latter yields the 
-complete Higgs mass contribution at $O(\at^2 t)$ which is identical 
-to the corresponding correction given in Ref.~\cite{Degrassi:2012ry}.
-The 2-loop self energy at $O(\at\as)$ \cite{Martin:2014cxa} together
-with the occurring integral functions from Ref.~\cite{Martin:2003qz}
-have been evaluated for a small external momentum argument. Neglecting
-higher orders, the expansion in powers of the momentum over the \MSbar
-top mass up to $O(p^4/t)$ results in the Eq.~\eqref{eq:2L_Mh_atas}. 
-Note that the momentum dependence is included in the 2-loop expression 
-in order to generate implicit 3-loop $O(\at^2\as)$ terms.
-%
-The explicit 3-loop effective potential contributions of
-$O(\at^3 + \at^2 \as + \at \as^2)$ to the
-Higgs mass included by \fs read
-%
-\begin{align}
-  (\Delta m_h^2)_{3L} &=
-  \frac{m_t^2}{(4\pi)^6} \Bigg[
-     g_3^4 y_t^2 \left(248.122 + 839.197 \;\barlog(t) + 160 \;\barlog^2(t) - 736 \;\barlog^3(t)\right) \nonumber \\
-     &\phantom{=\frac{m_t^2}{(4\pi)^6} \Bigg[}
-     + g_3^2 y_t^4 \left(2764.365 + 1283.716 \;\barlog(t) - 360 \;\barlog^2(t) + 240 \;\barlog^3(t)\right) \nonumber \\
-     &\phantom{=\frac{m_t^2}{(4\pi)^6} \Bigg[}
-     + y_t^6 \Big(-3199.017 + 36 \;\barlog(h) - 2653.511 \;\barlog(t) + 756 \;\barlog(h) \;\barlog(t) \nonumber \\
-     &\phantom{=\frac{m_t^2}{(4\pi)^6} \Bigg[}\qquad~~
-     + \frac{27}{2} \;\barlog^2(t) + 324 \;\barlog(h) \;\barlog^2(t) - 225 \;\barlog^3(t)\Big)
-  \Bigg] \,,
-\end{align}
-%
-where $h\equiv m_h^2$ is the squared \MSbar Higgs mass in the SM\@.
-Note that the 3-loop Higgs mass calculation in
-the \MSbar scheme in \fsbreak is only complete at $O(\at\as^2)$.  The
-3-loop contributions of $O(\at^3 + \at^2 \as)$ are currently incomplete
-because they would require the \MSbar top Yukawa coupling to be determined
-from the top pole mass at the 2-loop $O(\at^2 + \at\as)$.
-However, these 2-loop contributions to the top Yukawa coupling are
-currently not available in \fs. Furthermore, the proper inclusion of
-corrections to the Higgs pole mass at $O(\at^3)$
-would require the extension of Eq.~\eqref{eq:2L_Mh_atat} by momentum
-dependent $O(\at^2)$ terms. Alternatively, the evaluation of the Higgs mass at
-the renormalization scale $Q^2 = t$ implies that the neglected
-3-loop contributions at $O(\at^3 t)$ vanish
-\cite{Martin:2014cxa}.\footnote{Note that terms of $O(\at^3 h)$ are
-  neglected here and in Ref.\ \cite{Martin:2014cxa}.}
-Likewise, Eq.~\eqref{eq:2L_Mh_atat} neglects contributions of $O(\at^2h)$ which are
-subdominant in comparison to the implemented $O(\at^2t)$ corrections.
-In contrast, the QCD corrections to the pole mass at 2-loop level, Eq.~\eqref{eq:2L_Mh_atas}, involve
-terms proportional to the quartic Higgs coupling $\lambda$ up to $O(\as\at h^2/t)$, which are not
-neglected here.
-
-The 2-loop and 3-loop Higgs mass contributions from above can be
-enabled at runtime by setting the following flags in the SLHA input
-file:
-%
-\begin{lstlisting}
-Block FlexibleSUSY
-    4   3         # pole mass loop order
-    5   3         # EWSB loop order
-    7   2         # threshold corrections loop order
-    8   1         # Higgs 2-loop corrections O(alpha_t alpha_s)
-   10   1         # Higgs 2-loop corrections O(alpha_t^2)
-   24   122111221 # individual threshold correction loop orders
-   26   1         # Higgs 3-loop corrections O(alpha_t alpha_s^2)
-   28   1         # Higgs 3-loop corrections O(alpha_t^2 alpha_s)
-   29   1         # Higgs 3-loop corrections O(alpha_t^3)
-\end{lstlisting}
-%
-In \fs's \mathematica interface, the above SLHA configuration options
-correspond to
-%
-\begin{lstlisting}[language=Mathematica]
-handle = FS<model>OpenHandle[
-    fsSettings -> {poleMassLoopOrder -> 3,
-                   ewsbLoopOrder -> 3,
-                   thresholdCorrectionsLoopOrder -> 2,
-                   higgs2loopCorrectionAtAs -> 1,
-                   higgs2loopCorrectionAtAt -> 1,
-                   thresholdCorrections -> 122111221,
-                   higgs3loopCorrectionAtAsAs -> 1,
-                   higgs3loopCorrectionAtAtAs -> 1,
-                   higgs3loopCorrectionAtAtAt -> 1}
-    ...];
-\end{lstlisting}
-%
-See \appref{app:fs_config_options} for a list and description of
-all of \fs's configuration options.
-%
-In \figref{fig:SM_thresholds} the impact of the leading 3-loop
-Standard Model corrections of $O(\at^3 + \at^2 \as + \at \as^2)$ on
-the light \CP-even Higgs pole mass in the MSSM in the pure effective
-field theory (EFT) calculation of \HSSUSY is shown with the red
-dashed-dotted line.  The line shows the predicted 3-loop Higgs mass
-relative to the one calculated at the 2-loop level as a function of
-the SUSY scale \MS, by taking into account the 1-loop threshold
-correction for $\as^{\SM}(M_Z)$ and the 2-loop QCD correction to
-$y_t^{\SM}(M_Z)$.  We find that the explicit 3-loop Standard Model
-contributions to the Higgs mass lead to a small positive shift by
-around $30\unit{MeV}$.
-%
-\begin{figure}[tbh]
-  \centering
-  \includegraphics[width=0.5\textwidth]{doc/plots/HSSUSY/HSSUSY_thresholds}
-  \caption{Effect of different SM corrections on the Higgs
-    pole mass in the pure EFT calculation of \HSSUSY for
-    $\tan\beta = 5$ and $X_t = \sqrt{6}\MS$ as a function of the SUSY
-    scale \MS.  The orange band marks the region where \HSSUSY
-    predicts a Higgs mass compatible with the experimental value of
-    $(125.09 \pm 0.32)\unit{GeV}$ \cite{Aad:2015zhl}.}
-  \label{fig:SM_thresholds}
-\end{figure}
-
-\paragraph{3-loop renormalization group running in the SM}
-
-In the SM, the 3-loop $\beta$ functions are known
-\cite{Mihaila:2012fm,Bednyakov:2012rb,Bednyakov:2012en,Chetyrkin:2012rz,Bednyakov:2013eba}.
-\fs allows the user to take these 3-loop $\beta$ functions into
-account in the running of the \MSbar SM parameters by
-setting the following switch in the model file:
-%
-\begin{lstlisting}
-UseSM3LoopRGEs = True;
-\end{lstlisting}
-%
-The expression for the 3-loop $\beta$ function of the $\mu^2$ parameter in
-the SM has been extracted from the file \code{smh3l.m} available on
-the arXiv page of Ref.~\cite{Bednyakov:2013eba}.  The remaining 3-loop
-$\beta$ functions have been kindly provided by the authors of \susyhd
-1.0.1 \cite{Vega:2015fna}.  To use 3-loop running at runtime, the
-following flag should also be set in the SLHA input file:
-%
-\begin{lstlisting}
-Block FlexibleSUSY
-    6   3         # beta-functions loop order
-\end{lstlisting}
-%
-In \fs's \mathematica interface, the above SLHA configuration
-corresponds to
-%
-\begin{lstlisting}[language=Mathematica]
-handle = FS<model>OpenHandle[
-    fsSettings -> {betaFunctionLoopOrder -> 3}
-    ...];
-\end{lstlisting}
-
-\paragraph{3-loop QCD corrections to the \MSbar top Yukawa coupling in the SM}
-
-The pure SM QCD contribution to the \MSbar top Yukawa
-coupling is known up to the 3-loop level of $O(\as^3)$
-\cite{Chetyrkin:1999qi,Melnikov:2000qh}.  This 3-loop QCD expression
-can be taken into account in \fs to extract the \MSbar top Yukawa
-coupling in the SM from the top pole mass $M_t$ by setting
-the following switch in the model file:
-%
-\begin{lstlisting}
-UseYukawa3LoopQCD = True;
-\end{lstlisting}
-%
-By default, \code{UseYukawa3LoopQCD} is set to \code{Automatic}, which
-means that the 3-loop QCD contribution to the top Yukawa coupling is
-taken into account only if the model is \MSbar renormalized.  To take
-this 3-loop correction into account at runtime, the following flags
-should be set in the SLHA input file:
-%
-\\\begin{minipage}{\linewidth}
-\begin{lstlisting}
-Block FlexibleSUSY
-    7   3         # threshold corrections loop order
-   24   123111221 # individual threshold correction loop orders
-\end{lstlisting}
-\end{minipage}
-%
-In \fs's \mathematica interface, the above SLHA configuration options
-correspond to
-%
-\begin{lstlisting}[language=Mathematica]
-handle = FS<model>OpenHandle[
-    fsSettings -> {thresholdCorrectionsLoopOrder -> 3,
-                   thresholdCorrections -> 123111221}
-    ...];
-\end{lstlisting}
-%
-The 3-loop QCD contribution is combined with the full 1-loop and QCD
-2-loop contribution as
-%
-\begin{align}
-\begin{split}
-  m_t(Q) &= M_t + \re\Sigma_{t}^S(p^2=M_t^2,Q) \\
-  &\phantom{={}} + M_t \Big[ \re\Sigma_{t}^L(p^2=M_t^2,Q) +
-    \re\Sigma_{t}^R(p^2=M_t^2,Q) \\
-  &\phantom{={} + M_t \Big[}
-    + \Delta m_t^{(1),\text{qcd}}(Q) + \Delta m_t^{(2),\text{qcd}}(Q) + \Delta m_t^{(3),\text{qcd}}(Q) \Big]
-  \,,
-\end{split} \label{eq:mt_MSbar}
-\end{align}
-%
-where $\Sigma_t^{S,L,R}(p^2,Q)$ denote the scalar, left- and
-right-handed parts of the \MSbar-renormalized 1-loop top quark self
-energy, $M_t$ is the top pole mass and
-%
-\begin{align}
-  \label{eq:top-selfenergy-qcd-1L}
-  \Delta m_t^{(1),\text{qcd}}(Q) &=
-  -\frac{g_3^2}{12 \pi^2} \left[4 - 3 \;\barlog(t)\right], \\
-  \label{eq:top-selfenergy-qcd-2L}
-  \begin{split}
-    \Delta m_t^{(2),\text{qcd}}(Q) &= \left(\Delta
-      m_t^{(1),\text{qcd}}(Q)\right)^2 - \frac{g_3^4}{4608 \pi^4}
-    \Bigg[396 \;\barlog^2(t) - 1452 \;\barlog(t) \\
-    &\phantom{={}}
-    - 48 \zeta(3) + 2053 + 16 \pi^2 (1+\log 4)\Bigg] ,
-  \end{split} \\
-\begin{split}
-  \Delta m_t^{(3),\text{qcd}}(Q) &= \frac{g_3^6}{2430\, (4\pi)^6}
-  \Bigg\{
-  48600 \;\barlog^3(t) - 208980 \;\barlog^2(t) \\
-  &\phantom{=\;} +540 \Big[-1560\,\zeta (3)+2993+40 \pi ^2 (1+\log 4)\Big] \barlog(t)\\
-  &\phantom{=\;} +15 \left[69120\,\text{Li}_4\left(\frac{1}{2}\right)+113040\,\zeta(3)-
-    94800 \,\zeta(5)-280853+2880 \,\log^4 2\right]\\
-  &\phantom{=\;} +4 \pi^2 \Big[129510 \,\zeta(3)-388781 + 240 (733+24 \log 2) \,\log 2\Big] - 10500 \pi^4
-  \Bigg\}
-\end{split}\\
-  &\approx
-  \frac{g_3^6}{(4\pi)^6}\, 20 \left[\barlog^3(t)
-    - \frac{43}{10} \;\barlog^2(t)
-    + 22.8874 \;\barlog(t)
-    -172.937 \right] \,.
-\end{align}
-%
-In \figref{fig:SM_thresholds} the impact of the 3-loop correction to
-the \MSbar top Yukawa coupling in the SM $y_t^{\SM}(M_Z)$
-on the prediction of the light \CP-even Higgs pole mass in the MSSM in
-the pure EFT calculation of \HSSUSY is shown as the green
-dashed-double-dotted line.  As already discussed in
-Ref.~\cite{Vega:2015fna}, we find that the inclusion of the 3-loop
-correction to $y_t^{\SM}(M_Z)$ reduces the Higgs mass by up to
-$500\unit{MeV}$.  Note that this is formally a (partial) 4-loop
-effect on the Higgs mass, which is beyond the current accuracy of
-\HSSUSY.
-
-\paragraph{3-loop QCD corrections to the \MSbar strong coupling in the SM}
-
-The pure SM QCD contribution to the \MSbar strong coupling is known up
-to the 3-loop level of $O(\as^3)$
-\cite{Rodrigo:1993hc,Rodrigo:1997zd,Chetyrkin:1997un,Chetyrkin:2000yt}.
-This 3-loop QCD expression can be taken into account in \fs to extract
-the \MSbar strong coupling in the SM from the input value
-$\as^{\SM(5)}(M_Z)$ by setting the following switch in the
-model file:
-%
-\begin{lstlisting}
-UseSMAlphaS3Loop = True;
-\end{lstlisting}
-%
-To take this 3-loop threshold correction into account at runtime, the
-following flags should also be set in the SLHA input file:
-%
-\begin{lstlisting}
-Block FlexibleSUSY
-    7   3         # threshold corrections loop order
-   24   123111321 # individual threshold correction loop orders
-\end{lstlisting}
-%
-In \fs's \mathematica interface, the above SLHA configuration options
-correspond to
-%
-\begin{lstlisting}[language=Mathematica]
-handle = FS<model>OpenHandle[
-    fsSettings -> {thresholdCorrectionsLoopOrder -> 3,
-                   thresholdCorrections -> 123111321}
-    ...];
-\end{lstlisting}
-%
-The 3-loop QCD contributions are combined as
-%
-\begin{align}
-  \as^{\SM}(Q) &=
-  \frac{\as^{\SM(5)}(Q)}{1 - \Delta\as^{1L}(Q)
-    - \Delta\as^{2L}(Q) - \Delta\as^{3L}(Q)} \,,
-%  &= \as^{\SM(5)}(Q)\Big[ 1 + \Delta_1 + \Delta_2 + \Delta_3 + O(\hbar^4) \Big] \,,
-\end{align}
-%
-where $\Delta\as^{nL}(Q)$ are the $n$-loop threshold corrections,
-which read \cite{Chetyrkin:2000yt}
-%
-\begin{align}
-  \Delta\as^{1L}(Q) &= \Delta_1 \,,\\
-  \Delta\as^{2L}(Q) &= \Delta_2 - \Delta_1^2 \,,\\
-  \Delta\as^{3L}(Q) &= \Delta_3 + \Delta_1^3 - 2\Delta_1\Delta_2 \,,\\
-  \Delta_1 &= \left(\frac{\as^{\SM(5)}(Q)}{\pi}\right) \frac{L}{6} \,, \\
-  \Delta_2 &= \left(\frac{\as^{\SM(5)}(Q)}{\pi}\right)^2 \Bigg[
-  -\frac{11}{72} + \frac{11}{24} L + \frac{1}{36} L^2
-  \Bigg] \,,\\
-  \Delta_3 &=
-  \left(\frac{\as^{\SM(5)}(Q)}{\pi}\right)^3 \Bigg[
-  \frac{1}{216} L^3
-  + \frac{167}{576} L^2
-  + \frac{2645}{1728}L
-  + n_f \left(\frac{1}{36} L^2
-  - \frac{67}{576} L
-  + \frac{2633}{31104}\right) \nonumber \\
-  &\phantom{={}}
-  + \frac{82043}{27648} \zeta_3
-  - \frac{564731}{124416}
-  \Bigg] \,, \\
-  &\approx
-  \left(\frac{\as^{\SM(5)}(Q)}{\pi}\right)^3 \Bigg[
-  0.00462963 L^3
-  + 0.428819 L^2
-  + 0.94907 L
-  - 0.54880 \Bigg] \,,
-\end{align}
-%
-with $n_f = 5$, $L = \log(Q^2/(m_t^{\SM}(Q))^2)$ and $m_t^{\SM}(Q)$ being the \MSbar
-top quark mass in the SM at the scale $Q$.
-
-\figref{fig:SM_thresholds} shows the impact of the 2- and 3-loop QCD
-threshold corrections for $\as^{\SM}(M_Z)$ on the light \CP-even Higgs
-pole mass in the MSSM, as predicted by the pure EFT calculation of
-\HSSUSY as a function of the SUSY scale \MS.  We find that the
-inclusion of $\Delta\as^{2L}$ (blue dashed line) reduces the Higgs
-mass by up to $40\unit{MeV}$, depending on the SUSY scale.  Taking
-into account $\Delta\as^{3L}$ (blue dotted line) reduces the Higgs
-mass further by around $3\unit{MeV}$.
-
-%%%%%%%%%%%%%%%%%%%%%% MSSM corrections %%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{(N)MSSM-specific higher-order contributions}
-\label{sec:MSSM-specific_ho_corrections}
-\paragraph{2-loop contributions to the (N)MSSM Higgs masses}
-
-As already described in Ref.~\cite{Athron:2014yba}, the known dominant 2-loop
-effective potential contributions to the Higgs masses in the (N)MSSM
-of $O((\at + \ab)\as + (\at +
-\ab)^2 + \atau^2)$
-\cite{Degrassi:2001yf,Brignole:2001jy,Dedes:2002dy,Brignole:2002bz,Dedes:2003km,Degrassi:2009yq}
-can be taken into account by \fs.
-%
-Note, however, that the implemented NMSSM 2-loop contributions of
-$O((\at + \ab)^2 + \atau^2)$ are currently available only in the
-MSSM-limit.  Furthermore, NMSSM-specific 2-loop corrections to the
-running vacuum expectation value beyond the MSSM limit, which would be
-required for a consistent treatment of the $O((\at + \ab)\as)$
-contributions, are currently not implemented.
-%
-To take the implemented 2-loop corrections to the Higgs masses in the
-(N)MSSM into account, the following switches must be set in the model
-file:
-%
-\begin{lstlisting}
-UseHiggs2LoopMSSM = True;
-UseHiggs2LoopNMSSM = True;
-\end{lstlisting}
-%
-These flags can be enabled in all real (N)MSSM-like models with 2(3)
-\CP-even Higgs bosons, 1(2) \CP-odd Higgs boson(s) and 1 electrically neutral
-Goldstone boson.  In addition to these flags, the (effective) $\mu$
-parameter in the convention of Ref.~\cite{Haber:1984rc} and the effective
-squared \CP-odd Higgs tree-level mass $m_A^2$ of the model must be identified.
-In the \modelname{NMSSM} model file of \sarah they would read for
-example:
-%
-\begin{lstlisting}
-EffectiveMu = \[Lambda] vS / Sqrt[2];
-EffectiveMASqr = (T[\[Lambda]] vS / Sqrt[2] + 0.5 \[Lambda] \[Kappa] vS^2) (vu^2 + vd^2) / (vu vd);
-\end{lstlisting}
-%
-These 2-loop contributions can then be enabled at runtime by setting
-the following flags in the SLHA input file:
-%
-\begin{lstlisting}
-Block FlexibleSUSY
-    4   2         # pole mass loop order
-    5   2         # EWSB loop order
-    7   2         # threshold corrections loop order
-    8   1         # Higgs 2-loop corrections O(alpha_t alpha_s)
-    9   1         # Higgs 2-loop corrections O(alpha_b alpha_s)
-   10   1         # Higgs 2-loop corrections O((alpha_t + alpha_b)^2)
-   11   1         # Higgs 2-loop corrections O(alpha_tau^2)
-   24   122111221 # individual threshold correction loop orders
-\end{lstlisting}
-%
-In \fs's \mathematica interface, the above SLHA configuration options
-correspond to
-%
-\begin{lstlisting}[language=Mathematica]
-handle = FS<model>OpenHandle[
-    fsSettings -> {poleMassLoopOrder -> 2,
-                   ewsbLoopOrder -> 2,
-                   thresholdCorrectionsLoopOrder -> 2,
-                   higgs2loopCorrectionAtAs -> 1,
-                   higgs2loopCorrectionAbAs -> 1,
-                   higgs2loopCorrectionAtAt -> 1,
-                   higgs2loopCorrectionAtauAtau -> 1,
-                   thresholdCorrections -> 122111221}
-    ...];
-\end{lstlisting}
-
-\paragraph{3-loop contributions to the light \CP-even MSSM Higgs mass}
-
-The 3-loop contributions to the light \CP-even Higgs mass in the MSSM
-have been calculated to $O(\at\as^2)$ in the \DRbar and
-\MDRbar scheme \cite{Harlander:2008ju,Kant:2010tf,Kunz:2014gya}.  The
-expressions are available in the public spectrum generator \htm
-\cite{h3murl}, where they are added to the 2-loop on-shell result of
-\FeynHiggs
-\cite{Heinemeyer:1998yj,Heinemeyer:1998np,Degrassi:2002fi,Frank:2006yh,Hahn:2013ria,Drechsel:2016jdg,Bahl:2016brp,Bahl:2017aev}.
-In Ref.~\cite{Harlander:2017kuc} the 3-loop contributions of
-$O(\at\as^2 + \ab\as^2)$ have been studied for the first time in a
-pure \DRbar MSSM spectrum generator and have been made available in
-the public C++ library \Himalaya \cite{himalaya}.
-
-The explicit 3-loop contributions of $O(\at\as^2 + \ab\as^2)$ to the
-light \CP-even Higgs mass from \Himalaya can be taken into account in
-\fs by setting the following flag in the \fs model file:
-%
-\begin{lstlisting}[language=Mathematica]
-UseHiggs3LoopMSSM = True;
-\end{lstlisting}
-%
-The model is required to be real MSSM-like with two \CP-even Higgs bosons and
-one electrically neutral Goldstone boson.
-%
-However, in a pure \DRbar calculation, another source of such 3-loop
-contributions originates from the 2-loop \SQCD contribution to the MSSM
-\DRbar top Yukawa coupling, which must be included.  Then, however, in
-order to be consistent with respect to the loop orders of the running and
-decoupling, also 3-loop renormalization group running should be
-performed and the 2-loop threshold correction $\Delta\as^{2L}$ of
-the strong coupling should be included, see below.  Therefore, we
-strongly recommend setting the following flags in addition in the \fs
-model file:
-%
-\begin{lstlisting}[language=Mathematica]
-UseHiggs2LoopMSSM = True;  (* 2-loop contribution to Higgs mass *)
-EffectiveMu = \[Mu];       (* specify mu parameter, see above *)
-UseMSSM3LoopRGEs = True;   (* 3-loop running *)
-UseMSSMYukawa2Loop = True; (* 2-loop SUSY-QCD correction to yt *)
-UseMSSMAlphaS2Loop = True; (* 2-loop threshold correction to alpha_s *)
-\end{lstlisting}
-%
-When building \fs, the path to the \Himalaya headers and to the
-\Himalaya library must be specified:
-%
-\begin{lstlisting}[language=bash]
-$ ./configure --with-models=[...] \
-   --enable-himalaya \
-   --with-himalaya-incdir=$HIMALAYA_PATH/source/include \
-   --with-himalaya-libdir=$HIMALAYA_PATH/build
-$ make
-\end{lstlisting}
-%
-where \code{$HIMALAYA_PATH} is the \Himalaya directory. %$
-To calculate the light \CP-even Higgs mass in the MSSM at the 3-loop
-level with \fs, the following flags must be set at runtime: In the
-SLHA input file we recommend setting at least
-%
-\begin{lstlisting}
-Block FlexibleSUSY
-    4   3         # pole mass loop order
-    5   3         # EWSB loop order
-    6   3         # beta-functions loop order
-    7   2         # threshold corrections loop order
-   24   122111221 # individual threshold correction loop orders
-   25   0         # ren. scheme for 3L corrections (0 = DR, 1 = MDR)
-   26   1         # Higgs 3-loop corrections O(alpha_t alpha_s^2)
-   27   1         # Higgs 3-loop corrections O(alpha_b alpha_s^2)
-\end{lstlisting}
-%
-In \fs's \mathematica interface, the above SLHA configuration options
-correspond to
-%
-\begin{lstlisting}[language=Mathematica]
-handle = FS<model>OpenHandle[
-    fsSettings -> {poleMassLoopOrder -> 3,
-                   ewsbLoopOrder -> 3,
-                   betaFunctionLoopOrder -> 3,
-                   thresholdCorrectionsLoopOrder -> 2,
-                   thresholdCorrections -> 122111221,
-                   higgs3loopCorrectionRenScheme -> 0,
-                   higgs3loopCorrectionAtAsAs -> 1,
-                   higgs3loopCorrectionAbAsAs -> 1}
-    ...];
-\end{lstlisting}
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width=0.49\textwidth]{doc/plots/Himalaya/scan_Mh_MS_TB-5_Xt-0}\hfill
-  \includegraphics[width=0.49\textwidth]{doc/plots/Himalaya/scan_Mh_MS_TB-5_Xt-0_delta_alpha}\\
-  \includegraphics[width=0.49\textwidth]{doc/plots/Himalaya/scan_Mh_Xt_TB-5_MS-2000}\hfill
-  \includegraphics[width=0.49\textwidth]{doc/plots/Himalaya/scan_Mh_Xt_TB-5_MS-2000_delta_alpha}
-  \caption{Comparison of different 2-loop and 3-loop calculations of
-    the lightest \CP-even Higgs mass in the MSSM with different \fs
-    spectrum generators for $\tan\beta = 5$.  In the top row we set
-    $X_t = 0$ and in the bottom row we use $\MS = 2\unit{TeV}$.}
-  \label{fig:Mh-himalaya}
-\end{figure}
-%
-\figref{fig:Mh-himalaya} shows a comparison of the different light
-\CP-even Higgs mass calculations available in \fs.  The scenario is
-chosen such that all soft-breaking \DRbar mass parameters, the
-superpotential
-$\mu$ parameter and the \DRbar \CP-odd Higgs mass are equal to the SUSY
-scale $\MS$ at the scale $Q =
-\MS$ and all \DRbar sfermion mixing parameters
-$X_f$ are set to zero, except for the stop mixing parameter
-$X_t$.  Furthermore, we set $M_t =
-173.34\unit{GeV}$, $\as^{\SM(5)}(M_Z) =
-0.1184$ and $\aem^{\SM(5)}(M_Z) =
-1/127.944$.  The top row of \figref{fig:Mh-himalaya} shows the Higgs
-mass as a function of $\MS$ for $X_t = 0$ and $\tan\beta =
-5$.  In the left panel of the top row the classic 2-loop fixed-order
-calculation with \fs \cite{Athron:2014yba} is shown as the blue dashed
-line and the pure EFT calculation with \HSSUSY is shown as the black
-dotted line.  The reason why these two curves do not deviate from each
-other logarithmically when
-$\MS$ increases is an accidental cancellation of large
-logarithmic contributions in the fixed-order calculation
-\cite{Athron:2016fuq}.  The green dashed-dotted line shows the
-improved version of the \feft calculation \cite{Athron:2016fuq}, which
-interpolates
-between the fixed-order calculation at low \MS and the pure EFT
-calculation at large \MS.  See \secref{sec:feft} for a discussion of
-the improvements of \feft.  The red solid line shows the 3-loop
-calculation up to $O(\at\as^2 +
-\ab\as^2)$ with \fsh \cite{Harlander:2017kuc}.  In this calculation, the
-full 2-loop \SQCD contributions to the \DRbar top and bottom Yukawa
-couplings $y_{t,b}(M_Z)$ \cite{Bednyakov:2007vm,Bednyakov:2002sf,Bednyakov:2005kt}
-are taken into account.  The sum of these contributions
-and the explicit 3-loop contribution to the \CP-even
-Higgs mass matrix leads to a downward shift of the predicted Higgs
-mass by $1$--$2\unit{GeV}$, depending on \MS, compared to the classic 2-loop
-calculation with \fs (blue dashed line).  As also discussed in
-Ref.~\cite{Harlander:2017kuc}, this brings the prediction closer to the pure
-EFT calculation, which is expected to lead to a more precise result above
-the TeV scale.  In the left panel of the bottom row of
-\figref{fig:Mh-himalaya}, the calculated Higgs mass is shown as a
-function of the stop mixing parameter $X_t$ for fixed
-$\MS = 2\unit{TeV}$.  Also for non-zero
-$X_t$, one finds a reduction of the Higgs mass by
-$1$--$3\unit{GeV}$, which brings the prediction closer to the one of
-the pure EFT calculation.
-
-In the right panels of \figref{fig:Mh-himalaya}, the effect of the 2-loop
-\SQCD threshold correction $\Delta\as^{2L}$
-\cite{Harlander:2005wm,Bauer:2008bj,Bednyakov:2010ni} to the strong
-coupling $\as$ (see below) on the 3-loop Higgs mass calculation is shown for
-the same scenario as in the corresponding left panels.
-The threshold correction $\Delta\as^{2L}$ is included at the scale
-$Q = M_Z$, the same scale at which all dimensionless MSSM \DRbar
-parameters are determined from the Standard Model input parameters.
-The inclusion of $\Delta\as^{2L}$ is formally a 4-loop effect on the light
-\CP-even Higgs mass in the MSSM\@.  However, $\Delta\as^{2L}$ should
-be taken into account for a consistent running and decoupling
-procedure with 3-loop renormalization group running.  The red solid
-lines in the right panels correspond to the 3-loop calculation of
-Ref.~\cite{Harlander:2017kuc}, which uses only the 1-loop threshold correction
-$\Delta\as^{1L}$.  These red solid lines are the same as in the
-corresponding left panels.  The effect of including
-$\Delta\as^{2L}$ is shown as the red dashed line.  We find that the
-inclusion of this 2-loop threshold correction leads to an upwards
-shift of the Higgs mass by up to $2\unit{GeV}$, depending on \MS and
-$X_t$.  Note, that large 4-loop contributions of multiple GeV hint at
-a large theoretical uncertainty of the fixed-order calculation of the
-light \CP-even Higgs pole mass in parameter regions with multi-TeV
-stop masses.
-
-\paragraph{3-loop renormalization group running in the MSSM}
-
-In the MSSM, the 3-loop $\beta$ functions are also known
-\cite{Jack:2003sx,Jack:2004ch}.  \fs allows the user to take these 3-loop
-$\beta$ functions into account in the running of the \DRbar MSSM parameters by
-setting the following switch in the model file:
-%
-\begin{lstlisting}
-UseMSSM3LoopRGEs = True;
-\end{lstlisting}
-%
-To use 3-loop running at runtime, the following flag should also be set in
-the SLHA input file:
-%
-\\\begin{minipage}{\linewidth}
-\begin{lstlisting}
-Block FlexibleSUSY
-    6   3         # beta-functions loop order
-\end{lstlisting}
-\end{minipage}
-%
-In \fs's \mathematica interface, the above SLHA configuration
-corresponds to
-%
-\begin{lstlisting}[language=Mathematica]
-handle = FS<model>OpenHandle[
-    fsSettings -> {betaFunctionLoopOrder -> 3}
-    ...];
-\end{lstlisting}
-%
-The expressions for the 3-loop $\beta$ functions have been extracted
-from the official FORM file provided by the authors of
-Refs.~\cite{Jack:2003sx,Jack:2004ch}.\footnote{\url{https://www.liverpool.ac.uk/~dij/betas/allgennb.log}}
-We have numerically compared the expressions with the ones implemented
-in \softsusy 3.7.0 \cite{Allanach:2014nba} and found exact agreement.
-%
-The effect of the 3-loop RGEs on the Higgs pole mass in the MSSM is of
-the order of a few $100\unit{MeV}$ as discussed in
-Ref.~\cite{Allanach:2014nba} and is shown in \figref{fig:MSSMMuBMu}.
-%
-Due to the complexity of the 3-loop $\beta$ functions, the runtime of
-the MSSM spectrum generators is increased by a factor of $4$--$5$ if
-3-loop running is enabled in the MSSM\@.
-
-\paragraph{2-loop \SQCD corrections to the \DRbar top and bottom Yukawa couplings in the MSSM}
-In the MSSM, the full 2-loop \SQCD corrections of $O(\as^2)$
-to the \DRbar top Yukawa coupling
-\cite{Bednyakov:2007vm,Bednyakov:2002sf,Bednyakov:2005kt} as well as
-the 2-loop \SQCD $O(\as^2)$ and partial electroweak contributions
-to the \DRbar bottom Yukawa coupling \cite{Bednyakov:2009wt} are
-known.  These 2-loop corrections have already been made available in
-\softsusy~3.7.0 \cite{Allanach:2014nba}.  The 2-loop \SQCD
-corrections of $O(\as^2)$ are now also incorporated in \fstwo and can be
-used by setting the following switch in the model file:\footnote{We
-  kindly thank Alexander Bednyakov for providing the 2-loop \SQCD
-  expressions.}
-%
-\begin{lstlisting}
-UseMSSMYukawa2Loop = True;
-\end{lstlisting}
-%
-To take these 2-loop contributions into account at runtime, the
-following flags should also be set in the SLHA input file:
-%
-\begin{lstlisting}
-Block FlexibleSUSY
-    7   2         # threshold corrections loop order
-   24   122111221 # individual threshold correction loop orders
-\end{lstlisting}
-%
-In \fs's \mathematica interface, the above SLHA configuration options
-correspond to
-%
-\begin{lstlisting}[language=Mathematica]
-handle = FS<model>OpenHandle[
-    fsSettings -> {thresholdCorrectionsLoopOrder -> 2,
-                   thresholdCorrections -> 122111221}
-    ...];
-\end{lstlisting}
-%
-In \figref{fig:MSSMMuBMu} we show in blue the effect of the full
-2-loop \SQCD corrections on the lightest \CP-even Higgs pole mass in
-the MSSM\@.  The blue dashed line shows the effect in \fstwo and the
-crosses in \softsusy~4.0.1.
-As can be seen from the figure, the full 2-loop \SQCD corrections are
-negative and can affect the Higgs mass by several GeV, as has been
-observed in Ref.~\cite{Allanach:2014nba}.  In the left panel of
-\figref{fig:MSSMMuBMu}, the shift in the Higgs mass is shown as a
-function of the SUSY scale.  For scales above $\approx 2\unit{TeV}$ we
-find a logarithmic shift in $M_h$, which is caused by new large
-logarithms originating from the 2-loop contribution of the SUSY
-particles to the \DRbar top Yukawa coupling.  We also find that these
-new logarithms alone would spoil the accidental cancellation of large
-logarithms described in Ref.~\cite{Athron:2016fuq}.
-%
-Note that in the MSSM the effect of the 2-loop \SQCD corrections to
-the top and bottom Yukawa couplings is a partial 3-loop contribution
-to the light \CP-even Higgs pole mass.  Thus, these 2-loop \SQCD corrections
-must be taken into account if the explicit 3-loop Higgs mass
-contributions of $O(\at \as^2 + \ab \as^2)$ from
-\Himalaya are used, see above.
-%
-\begin{figure}[tbh]
-  \centering
-  \includegraphics[width=0.49\textwidth]{doc/plots/MSSMMuBMu/scan_MSSMMuBMu_Mh_MS_TB-5_Xt-0}\hfill
-  \includegraphics[width=0.49\textwidth]{doc/plots/MSSMMuBMu/scan_MSSMMuBMu_Mh_Xt_TB-5_MS-2000}
-  \caption{Effect of the 3-loop RGEs for all MSSM parameters (red),
-    2-loop \SQCD contributions to $y_b$ and $y_t$ (blue) and 2-loop
-    \SQCD contributions to $\as$ (green) on the lightest \CP-even
-    Higgs pole mass in the MSSM\@.  The lines show the shift in the
-    Higgs pole mass in \fstwo compared to the one obtained with 2-loop
-    RGEs, full 1-loop \SQCD + 2-loop SM-QCD contributions to $y_t$ and
-    1-loop \SQCD contributions to $\as$.  The crosses show the
-    corresponding shift obtained with \softsusy~4.0.1.  In the left
-    panel we use $\tan\beta = 5$ and $X_t = 0$ and in the right panel
-    we fix $\tan\beta = 5$ and $\MS = 2\unit{TeV}$.}
-  \label{fig:MSSMMuBMu}
-\end{figure}
-
-\paragraph{2-loop \SQCD corrections to the \DRbar strong gauge coupling in the MSSM}
-
-In the MSSM, the full 2-loop \SQCD corrections of
-$O(\as^2 + \as\at + \as\ab)$ to the \DRbar strong gauge coupling are known
-\cite{Harlander:2005wm,Bauer:2008bj,Bednyakov:2010ni} and have been
-made available in \softsusy~3.7.0 \cite{Allanach:2014nba}.  In \fstwo,
-these corrections can be taken into account by setting in the model
-file:\footnote{We kindly thank Ben Allanach and Alexander Bednyakov
-  for providing the 2-loop \SQCD expressions.}
-%
-\begin{lstlisting}
-UseMSSMAlphaS2Loop = True;
-\end{lstlisting}
-%
-To take these 2-loop threshold corrections into account at runtime, the
-following flags should also be set in the SLHA input file:
-%
-\begin{lstlisting}
-Block FlexibleSUSY
-    7   2         # threshold corrections loop order
-   24   122111221 # individual threshold correction loop orders
-\end{lstlisting}
-%
-In \fs's \mathematica interface, the above SLHA configuration options
-correspond to
-%
-\begin{lstlisting}[language=Mathematica]
-handle = FS<model>OpenHandle[
-    fsSettings -> {thresholdCorrectionsLoopOrder -> 2,
-                   thresholdCorrections -> 122111221}
-    ...];
-\end{lstlisting}
-%
-In \figref{fig:MSSMMuBMu} we show in green the effect of the 2-loop
-\SQCD corrections to $\as$ on the lightest \CP-even Higgs pole
-mass in the MSSM\@.  The green dashed-dotted line shows the effect in
-\fstwo and the crosses in \softsusy~4.0.1.  Both implementations agree
-exactly.  We furthermore find that the inclusion of the 2-loop
-threshold corrections to $\as$ leads to a logarithmic enhancement
-of the Higgs mass as a function of the SUSY scale.  The enhancement is
-around $+1\unit{GeV}$ for $\MS\approx 2\unit{TeV}$ and maximal stop
-mixing.
-%
-Note that in the MSSM the effect of the 2-loop \SQCD corrections to
-the strong coupling is formally a partial 4-loop contribution to the
-Higgs pole mass.
-
-\subsubsection{Split-MSSM-specific higher-order contributions}
-\paragraph{3-loop contribution to the Higgs mass in the split-MSSM}
-
-In the split-MSSM \cite{Bagnaschi:2014rsa}, part of the 3-loop
-contribution of $O(\at \as^2)$
-to the SM-like Higgs pole mass is known in the \MSbar
-scheme \cite{Benakli:2013msa}.  \fs allows the user to take this
-3-loop contribution into account by setting the following switch in
-the model file:
-%
-\begin{lstlisting}
-UseHiggs3LoopSplit = True;
-\end{lstlisting}
-%
-This is used in the distributed \modelname{SplitMSSM} model file. For use in
-other models, it is a requirement that the model contains a single
-Higgs boson and a gluino.  The enabled 3-loop Higgs pole mass
-contribution reads
-%
-\begin{align}
-  (\Delta m_h^2)_{3L} &= \frac{64 g_3^4 y_t^2}{(4\pi)^6} m_t^2\;\barlog^3(g) \,,
-\end{align}
-%
-where $g = m_{\tilde{g}}^2$ and $m_{\tilde{g}}$ is the \MSbar gluino
-mass.  Furthermore, for consistency the 2-loop gluino contribution in
-the calculation of the \MSbar top mass from Ref.\
-\cite{Benakli:2013msa} is taken into account by adding the following
-term to Eq.\ \eqref{eq:mt_MSbar}:
-%
-\begin{align}
-  \Delta m_t^{(2),\text{split-qcd}}(Q) =
-  -\frac{g_3^4}{(4\pi)^4} M_t \left\{
-     \frac{89}{9}
-     + 4\;\barlog(g) \left[
-        \frac{13}{3}
-        + \barlog(g)
-        - 2\;\barlog(t)
-     \right]
-  \right\} \,.
-\end{align}
-%
-To take these contributions into account at runtime, the following flags
-should also be set in the SLHA input file:
-%
-\begin{lstlisting}
-Block FlexibleSUSY
-    4   3         # pole mass loop order
-    5   3         # EWSB loop order
-    7   2         # threshold corrections loop order
-   24   122111221 # individual threshold correction loop orders
-   26   1         # Higgs 3-loop corrections O(alpha_t alpha_s^2)
-\end{lstlisting}
-%
-In \fs's \mathematica interface, the above SLHA configuration options
-correspond to
-%
-\\\begin{minipage}{\linewidth}
-\begin{lstlisting}[language=Mathematica]
-handle = FS<model>OpenHandle[
-    fsSettings -> {poleMassLoopOrder -> 3,
-                   ewsbLoopOrder -> 3,
-                   thresholdCorrectionsLoopOrder -> 2,
-                   thresholdCorrections -> 122111221,
-                   higgs3loopCorrectionAtAsAs -> 1}
-    ...];
-\end{lstlisting}
-\end{minipage}
-
-\subsection{New features for definition of boundary conditions}
-In \fstwo, the expressions to define boundary conditions are allowed
-to be more complicated and to involve trigonometric functions,
-dilogarithms, branches and more.  This is particularly useful for defining
-high-scale boundary conditions that match a model to its UV-completion.  An
-example is \fs's \HSSUSY model file, which implements the known
-1- and 2-loop high-scale matching condition on the quartic Higgs
-coupling of the SM against the MSSM at the SUSY scale
-\cite{Bagnaschi:2014rsa,Vega:2015fna,Bagnaschi:2017xid}.  The list of
-special functions and symbols to be used in boundary conditions can be
-found in \tabref{tab:bcs_functions}.
-%
-\begin{table}[h!]
-  \centering
- \resizebox{\textwidth}{!}{%
-  \begin{tabularx}{1.05\textwidth}{lX}
-    \toprule
-    Function & Description \\
-    \midrule
-
-    \code{Abs[a_]} &
-    Returns the magnitude of a real or complex number \code{a}, $|a|$.
-    If \code{a} is a vector, a vector is returned with \code{Abs} applied to each element. \\
-
-    \code{AbsSqr[a_]} &
-    Returns the squared magnitude of a real or complex number \code{a}, $|a|^2$. \\
-
-    \code{AbsSqrt[a_]} &
-    Returns the square root of the magnitude of \code{a}, $\sqrt{|a|}$. \\
-
-    \code{ArcSin[a_]}, \code{ArcCos[a_]} &
-    Returns $\arcsin a$ and $\arccos a$, respectively. \\
-
-    \code{ArcTan[a_]} & Returns $\arctan a$. \\
-
-    \code{Arg[z_]} &
-    Returns the phase angle of a complex number, $\arg z$. \\
-
-    \code{Cbrt[a_]} &
-    Returns the cubic root of \code{a}, $\sqrt[3]{a}$. \\
-
-    \code{CKM}, \code{PMNS} &
-    CKM and PMNS matrices, respectively, as defined in \cite{Allanach:2008qq}. \\
-
-    \code{Conjugate[a_]} &
-    Returns the complex conjugate of \code{a}. \\
-
-    \code{Exp[a_]} & Returns $e^a$ for real or complex \code{a}. \\
-
-    \code{FiniteLog[a_]} &
-    Returns $\log a$ if $\log a$ is well-defined, otherwise returns $0$. \\
-
-    \code{FSThrow[msg_]} & Throws an exception of type
-    \code{PhysicalError} with the message \code{msg}. \\
-
-    \code{I} & Imaginary unit. \\
-
-    \code{If[cond_, a_, b_]} &
-    If \code{cond} is true, \code{a} is returned, otherwise \code{b}. \\
-
-    \code{Im[a_]} &
-    Returns the imaginary part of \code{a}. \\
-
-    \code{IsClose[a_, b_, eps_]} &
-    Returns \code{True} if \code{Abs[a - b]} $<$ \code{eps}, otherwise \code{False}. \\
-
-    \code{IsCloseRel[a_, b_, eps_]} &
-    Returns \code{True} if \code{Abs[(a - b)/a]} $<$ \code{eps}, otherwise \code{False}. \\
-
-    \code{IsFinite[a_]} &
-    Returns \code{True} if \code{a} is neither \code{nan} nor \code{inf}. \\
-
-    \code{KroneckerDelta[i_, j_]} &
-    Returns the Kronecker $\delta_{ij}$. \\
-
-    \code{Log[a_], ComplexLog[a_]} &
-    Returns the natural logarithm for real and complex arguments, respectively. \\
-
-    \code{Max[a_, ...]} &
-    Returns the maximum of all given arguments. \\
-
-    \code{Min[a_, ...]} &
-    Returns the minimum of all given arguments. \\
-
-    \code{Not[cond_]} &
-    Returns the logical negation of \code{cond}. \\
-
-    \code{PolyLog[2, z_]} &
-    Returns the dilogarithm of the real or complex number \code{z}. \\
-
-    \code{Re[a_]} &
-    Returns the real part of \code{a}. \\
-
-    \code{Round[a_]} &
-    Returns \code{Floor[a + 0.5]} if \code{a} $\geq 0$, otherwise
-    \code{Floor[a - 0.5]}. \\
-
-    \code{Print<type>[msg_]} & Prints a debug, info, error, warning or
-    fatal message, depending on the \code{<type>}, and returns zero.
-    \code{<type>} can be \code{DEBUG}, \code{INFO}, \code{WARNING},
-    \code{ERROR} or \code{FATAL}.  \code{PrintFATAL[msg]}
-    throws an exception after \code{msg} has been printed.\\
-
-    \code{SCALE} & Returns the renormalization scale at which
-    the boundary condition is imposed. \\
-
-    \code{Sign[a_]} &
-    Returns $1$ if \code{a} $\geq 0$, otherwise $-1$. \\
-
-    \code{SignedAbsSqrt[a_]} &
-    Returns \code{Sign[a]*Sqrt[Abs[a]]}. \\
-
-    \code{Sin[a_]}, \code{Cos[a_]}, \code{Tan[a_]} &
-    Returns $\sin a$, $\cos a$ and $\tan a$, respectively. \\
-
-    \code{Total[vec_]} &
-    Returns the sum of all elements of \code{vec}, $\sum_i v_i$. \\
-
-    \code{UnitStep[value_]} &
-    Returns $0$ if \code{value} $<0$, $1$ otherwise. \\
-
-    \multilinecell[t]{\code{Which[test1_, value1_,}\\\code{\ test2_, value2_, ...]}} &
-    If \code{test1} is true, \code{value1} is returned, otherwise
-    if \code{test2} is true, \code{value2} is returned, etc. \\
-
-    \code{ZeroSqrt[a_]} &
-    Returns $\sqrt{a}$ if $a>0$, otherwise returns $0$. \\
-
-    \bottomrule
-  \end{tabularx}}
-  \caption{Available special functions and symbols in the boundary conditions.}
-  \label{tab:bcs_functions}
-\end{table}
-
-\section{\fbsm extension}
-\label{sec:fbsm}
-
-Since version 1.1.0, \fs can generate spectrum generators not
-only for SUSY models, but also for non-SUSY models.  We document this
-feature here for the first time.  In Subsections
-\ref{sec:application_THDM} and \ref{sec:application_HSSUSY}, we will
-describe important applications of this feature to the two-Higgs
-doublet model and to an effective low-energy theory of the MSSM
-(\HSSUSY).
-
-The generated non-SUSY spectrum generators have the same features as the SUSY
-spectrum generators:
-%
-\begin{itemize}
-\item The running gauge and Yukawa couplings of the non-SUSY model are
-  calculated automatically at the 1-loop level from the known
-  low-energy SM parameters
-  $\aem^{\SM(5)}(M_Z)$, $\as^{\SM(5)}(M_Z)$
-  and from the known quark and lepton masses as well as $M_Z$ and either $G_F$
-  or $M_W$.  2-loop and 3-loop QCD corrections can be taken into
-  account to determine the running top Yukawa coupling of the model,
-  see \secref{sec:model_specific_contributions}.
-\item Up to three boundary conditions can be specified to fix the
-  running parameters of the model at different user-defined scales.
-\item 2-loop renormalization group running is used between the scales
-  at which the boundary conditions are imposed.\footnote{Note that
-    the $\beta$ functions of scalar tadpole terms in
-    non-supersymmetric models \cite{Goodsell:2012fm} are currently not
-    generated by \sarah.  For this reason, such tadpole terms do not
-    run in \sarah/\spheno or \fs.} In the SM and in the MSSM, also
-  3-loop running is available, see
-  \secref{sec:model_specific_contributions}.
-\item The pole mass spectrum is calculated at the full 1-loop level,
-  taking into account all BSM contributions.  Some 2-loop and 3-loop
-  corrections can be added in specific non-SUSY models, see
-  \secref{sec:model_specific_contributions}.
-\end{itemize}
-
-\subsection{Setting up a \fbsm model}
-
-The \fs user interface for creating spectrum generators for non-SUSY
-models is exactly the same as in the case of SUSY models, except that
-all non-SUSY parameters are defined in the \MSbar scheme.  In
-particular, at the low-energy scale \fs automatically determines
-the gauge and Yukawa couplings of the non-SUSY model in the \MSbar
-scheme.  For gauge-dependent quantities like running masses and VEVs,
-\fs adopts the Feynman gauge, where all gauge fixing parameters
-$\xi_i$ are set to unity.
-
-\subsection{Determination of the \MSbar gauge and Yukawa couplings}
-\label{sec:determination_of_g_Y}
-
-If the considered BSM model has a gauge symmetry with the SM gauge group $SU(3)_C\times SU(2)_L\times U(1)_Y$ as a factor,
-then \fs automatically fixes the three corresponding normalized running gauge
-couplings $g_1$, $g_2$ and $g_3$ at the low-energy boundary condition
-from the given input parameters $M_Z$, $\aem^{\SM(5)}(M_Z)$,
-$\as^{\SM(5)}(M_Z)$ and $G_F$ or $M_W$ as
-%
-\begin{align}
-  g_1(Q) &= N_{g_Y} \, g_Y(Q) \,,
-  & g_Y(Q) &= \frac{\sqrt{4\pi\aem(Q)}}{\cos\theta_W(Q)} \,,\\
-  g_2(Q) &= N_{g_L} \, g_L(Q) \,,
-  & g_L(Q) &= \frac{\sqrt{4\pi\aem(Q)}}{\sin\theta_W(Q)} \,,\\
-  g_3(Q) &= N_{g_{\text{s}}} \, g_{\text{s}}(Q) \,,
-  & g_{\text{s}}(Q) &= \sqrt{4\pi\as(Q)} \,.
-\end{align}
-%
-Here, $\aem$ and $\as$
-denote the \MSbar electromagnetic and strong coupling constants of the
-non-SUSY model, respectively, and $\theta_W$ is the \MSbar\ weak
-mixing angle.  The coefficients $N_{g_i}$ denote the potential
-normalization factors defined in the \sarah model file.  The
-renormalization scale $Q$, at which the gauge couplings are calculated,
-can be specified using the \code{LowScale} variable in the model file.
-The coupling constants of the model are related to the corresponding
-ones of the SM with five active quark flavors,
-$\aem^{\SM(5)}(Q)$ and $\as^{\SM(5)}(Q)$, which are input,
-via the relations
-%
-\begin{align}
-  \aem(Q) &=
-  \frac{\aem^{\SM(5)}(Q)}{1 - \Delta\aem(Q)} \,,\\
-  \as(Q) &=
-  \frac{\as^{\SM(5)}(Q)}{1 - \Delta\as(Q)} \,.
-\end{align}
-%
-The threshold corrections $\Delta\alpha_i(Q)$ have the form
-%
-\begin{align}
-  \Delta\aem(Q) &=
-  \frac{\aem}{2\pi} \sum_i C_i^{\text{em}} \log\frac{m_i}{Q}
-  \,,\label{eq:alpha_em_general_threshold} \\
-  \Delta\as(Q) &=
-  \frac{\as}{2\pi} \sum_i C_i^{\text{s}} \log\frac{m_i}{Q} \,,
-\end{align}
-%
-where the sum runs over all non-SM particles plus the top quark
-with running \MSbar masses $m_i(Q)$.
-The constants $C_i^{\text{em}}$ and $C_i^{\text{s}}$ depend on the
-representation of the particle $i$ with respect to the Lorentz and gauge group.
-%
-The \MSbar\ weak mixing angle $\theta_W$ in the non-SUSY model is
-determined either
-%
-\begin{itemize}
-\item from the Fermi constant $G_F$ and $M_Z$ using the iterative
-  approach described in Ref.~\cite{Degrassi:1990tu} taking into account the
-  full 1-loop corrections and leading 2-loop SM
-  corrections to $\Delta\hat{\rho}$ and $\Delta\hat{r}$, see
-  \secref{sec:fmw}.
-\item or from the running $W$ and $Z$ masses, which
-  are obtained from the corresponding pole masses via a 1-loop
-  calculation.  See \secref{sec:fmw} for more details.
-\end{itemize}
-%
-If the considered BSM model does not contain the SM gauge
-group as a factor, it is of course still possible to fix the gauge
-couplings at the low-energy boundary condition by defining them to be
-input parameters.
-%
-\begin{example}
-  In a left-right-symmetric model with the gauge group $SU(3)_C\times
-  SU(2)_L\times SU(2)_R\times U(1)_L\times U(1)_R$, one could for
-  example fix the running BSM gauge couplings $g_3$, $g_L$, $g_R$, $g_{1L}$
-  and $g_{1R}$ by the running SM-like gauge couplings $g_3^\SM$,
-  $g_2^\SM$ and $g_1^\SM$ which are given as input via the
-  \code{EXTPAR} block:
-%
-\begin{lstlisting}
-EXTPAR = {
-  {100, g1SMInput},
-  {101, g2SMInput},
-  {102, g3SMInput}
-};
-
-LowScaleInput = {
-   {g3, g3SMInput},
-   {gL, g2SMInput},
-   {gR, g2SMInput},
-   {g1L, g1SMInput g2SMInput / Sqrt[2 (-g1SMInput^2 + g2SMInput^2)]},
-   {g1R, g1SMInput g2SMInput / Sqrt[2 (-g1SMInput^2 + g2SMInput^2)]},
-   {g1L1R, 0},
-   {g1R1L, 0}
-};
-\end{lstlisting}
-Note that, in addition to fixing the BSM gauge couplings, in this example
-the off-diagonal gauge couplings $g_{1L1R}$ and $g_{1R1L}$ that arise due to
-$U(1)$ mixing are being set to zero at the low-energy scale.
-\end{example}
-
-The \MSbar\ Yukawa couplings $Y_f(Q)$ of the SM-like
-fermions $f$ in the non-SUSY model are determined from the
-corresponding \MSbar\ masses $m_f(Q)$ using the tree-level
-relation. For example, in the SM this relation reads
-%
-\begin{align}
-  y_f^{\SM}(Q) = \frac{\sqrt{2} m_f^{\SM}(Q)}{v^{\SM}(Q)} \,,
-\end{align}
-%
-with $f = u,d,c,s,t,b,e,\mu,\tau$ and the \MSbar vacuum expectation value $v^{\SM}(Q)$.
-The running top quark \MSbar\ mass in the non-SUSY model, $m_t(Q)$, is
-calculated from the top pole mass $M_t$
-using the full 1-loop self energy plus 2-loop SM QCD
-corrections as shown in Eq.\ \eqref{eq:mt_MSbar}.
-In the SM, 3-loop QCD contributions can be taken into
-account as well, see \secref{sec:model_specific_contributions}.
-%
-The bottom quark \MSbar\ mass in the non-SUSY model, $m_b(Q)$, is
-obtained from the \MSbar\ mass $m_b^{\SM(5)}(m_b)$ in the SM with $5$ active quark flavors by first evolving
-$m_b^{\SM(5)}(m_b)$ to the scale $Q$ using the 1-loop QED and 3-loop
-QCD RGEs.  Afterwards, $m_b^{\SM(5)}(Q)$ is converted to $m_b(Q)$ as
-%
-\begin{align}
-  m_b(Q) &= \frac{m_b^{\SM(5)}(Q)}{1 - \Delta m_b} \,, \\
-  \Delta m_b &= \re\Sigma_{b}^S(p^2=(m_b^{\SM(5)})^2,Q)/m_b \notag\\
-  &\phantom{={}}+ \re\Sigma_{b}^L(p^2=(m_b^{\SM(5)})^2,Q)
-  + \re\Sigma_{b}^R(p^2=(m_b^{\SM(5)})^2,Q) \,,
-  \label{eq:bottom-conversion}
-\end{align}
-%
-where $\Sigma_{b}^{S,L,R}$ are the scalar, left- and right-handed
-parts of the 1-loop bottom quark self energy in the \MSbar\ scheme, in
-which all loops that contain only SM(5) particles are omitted.  Finally, the
-\MSbar\ mass of the $\tau$ lepton, $m_\tau(Q)$, is calculated by first
-identifying the $\tau$ pole mass, $M_\tau$, with the \MSbar\ mass in
-the SM with 5 active quark flavors at the scale $M_\tau$,
-%
-\begin{align}
-  m_\tau^{\SM(5)}(M_\tau) &=  M_\tau \,.
-\end{align}
-%
-In this identification, the 1-loop SM electroweak
-corrections to $m_\tau^{\SM(5)}(M_\tau)$ are neglected.  Afterwards,
-$m_\tau^{\SM(5)}(M_\tau)$ is evolved to the scale $Q$ using the 1-loop QED RGE
-and $m_\tau^{\SM(5)}(Q)$ is converted to $m_\tau(Q)$ as
-%
-\begin{align}
-  m_\tau(Q) &= \frac{m_\tau^{\SM(5)}(Q)}{1 - \Delta m_\tau} \,, \\
-  \Delta m_\tau &=
-  \re\Sigma_{\tau}^S(p^2=(m_\tau^{\SM(5)})^2,Q) / m_\tau^{\SM(5)}(Q) \notag \\
-  &\phantom{={}}
-  + \re\Sigma_{\tau}^L(p^2=(m_\tau^{\SM(5)})^2,Q)
-    + \re\Sigma_{\tau}^R(p^2=(m_\tau^{\SM(5)})^2,Q) \,,
-\end{align}
-%
-where $\Sigma_{\tau}^{S,L,R}$ are the scalar, left- and right-handed
-parts of the 1-loop $\tau$ self energy in the \MSbar\ scheme,
-from which all loops that contain only SM(5) particles are omitted.
-
-In most models, it is necessary to also fix the running SM-like vacuum
-expectation value (VEV), $v$, at the low-energy scale.  For this purpose \fs
-provides the symbols \code{MZMSbar} and \code{MWMSbar} in the model
-file to access the \MSbar $W$ and $Z$ masses $m_W$ and $m_Z$ in the
-non-SUSY model at the low-energy scale.  These running masses can be
-used to calculate the \MSbar vacuum expectation value $v$, as for
-example in the SM,
-%
-\begin{align}
-  v(Q) &= \frac{2 m_Z(Q)}{\sqrt{3 g_1^2(Q) / 5+ g_2^2(Q)}} \,.
-\end{align}
-
-\subsection{Structure of the generated code}
-
-In analogy to SUSY models, the parameters of a non-supersymmetric
-model are distributed among two classes in the model class
-hierarchy, see \figref{fig:parameter-classes}:
-%
-\begin{figure}[tbh]
-  \centering
-  \tikzumlset{fill class=white}
-  \begin{tikzpicture}
-    \umlclass[x=0, y=8, type=interface]{Beta\_function}{
-    }{
-      + \umlvirt{get()}\\
-      + \umlvirt{set()}\\
-      + \umlvirt{beta()}\\
-      + run\_to()
-    }
-    \umlclass[x=0, y=4]{<model>\_susy\_parameters}{
-      -- dimensionless parameters
-    }{
-      + get()\\
-      + set()\\
-      + beta()\\
-      + run\_to()
-    }
-    \umlclass[x=0, y=0]{<model>\_soft\_parameters}{
-      -- parameters with mass dimension > 0
-    }{
-      + get()\\
-      + set()\\
-      + beta()\\
-      + run\_to()
-    }
-    \umlinherit{<model>\_susy\_parameters}{Beta\_function}
-    \umlinherit{<model>\_soft\_parameters}{<model>\_susy\_parameters}
-  \end{tikzpicture}
-  \caption{Model class hierarchy.}
-  \label{fig:parameter-classes}
-\end{figure}
-%
-At the top of the model class hierarchy stands the
-\code{Beta_function} interface class, which defines the interface for
-the RGE integrator of the model parameters.  It provides the interface
-function \code{run_to()}, which integrates the RGEs up to a given
-scale using an adaptive Runge-Kutta algorithm.  The Runge-Kutta
-algorithm makes use of the virtual functions \code{get()},
-\code{set()} and \code{beta()} to obtain the model parameters at the
-current renormalization scale, set the model parameters to new values
-or calculate the $\beta$ functions.  These virtual functions are
-implemented by the derived classes, \code{<model>_susy_parameters} and
-\code{<model>_soft_parameters}.  Dimensionless parameters, like
-gauge, Yukawa or quartic scalar couplings, are contained in the
-\code{<model>_susy_parameters} class.  Parameters with mass dimension
-greater than zero are contained in the \code{<model>_soft_parameters}
-class.  The distribution of the model parameters between these two
-classes reflects the dependency of the
-$\beta$ functions of the dimensionful parameters upon the
-dimensionless parameters.  Furthermore, it allows
-the RGEs of the dimensionless parameters to be integrated independently of the
-dimensionful parameters.
-
-\subsection{Application: High-scale MSSM with light Higgs sector}
-\label{sec:application_THDM}
-
-As an application of \fbsm we consider the Higgs pole mass prediction
-in an MSSM scenario with very heavy sfermions, Higgsinos and gauginos
-at the SUSY scale $\MS$, but a light Higgs sector.  If $\MS$ is larger
-than a few TeV, an EFT approach should be considered and the heavy SUSY
-particles should be integrated out at $\MS$.  The resulting EFT below $\MS$ is
-the Two-Higgs-Doublet-Model (\THDM).  Our aim is to calculate the Higgs pole
-masses in this effective \THDM, where the quartic Higgs couplings are fixed by the
-MSSM at $\MS$.\footnote{This \fs setup was also used in
-  Ref.~\cite{Bagnaschi:2015pwa} to study the vacuum stability at very
-  high SUSY scales in different \THDM variants with the MSSM as a
-  supersymmetric UV completion.}  We use a \THDM of type II here, for
-which the full 1- and leading 2-loop threshold corrections at the SUSY
-scale are known \cite{Haber:1993an,Gorbahn:2009pp,Lee:2015uza}.
-
-In order to construct such an EFT setup we have to build the \THDM-II
-with \sarah.  We start by specifying the gauge group, the field
-content and the Lagrangian:
-%
-\begin{lstlisting}
-(* gauge groups *)
-Gauge[[1]]={B,   U[1], hypercharge, g1,False};
-Gauge[[2]]={WB, SU[2], left,        g2,True};
-Gauge[[3]]={G,  SU[3], color,       g3,False};
-
-(* field content *)
-FermionFields[[1]] = {q , 3, {uL,dL}   , 1/6, 2,  3};
-FermionFields[[2]] = {l , 3, {vL,eL}   ,-1/2, 2,  1};
-FermionFields[[3]] = {d , 3, conj[dR]  , 1/3, 1, -3};
-FermionFields[[4]] = {u , 3, conj[uR]  ,-2/3, 1, -3};
-FermionFields[[5]] = {e , 3, conj[eR]  ,   1, 1,  1};
-ScalarFields[[1]]  = {H1, 1, {H1p, H10}, 1/2, 2,  1};
-ScalarFields[[2]]  = {H2, 1, {H2p, H20}, 1/2, 2,  1};
-
-DEFINITION[GaugeES][Additional] = {
-    {LagHC  , { AddHC->True  }},
-    {LagNoHC, { AddHC->False }}
-};
-
-LagNoHC = -(M112 conj[H1].H1 + M222 conj[H2].H2 \
-            + Lambda1 conj[H1].H1.conj[H1].H1 \
-            + Lambda2 conj[H2].H2.conj[H2].H2 \
-            + Lambda3 conj[H2].H2.conj[H1].H1 \
-            + Lambda4 conj[H2].H1.conj[H1].H2 );
-
-LagHC = -(-M122 conj[H1].H2
-          + Lambda5/2 conj[H2].H1.conj[H2].H1
-          + Lambda6 conj[H1].H1.conj[H1].H2
-          + Lambda7 conj[H2].H2.conj[H1].H2
-          + Yd conj[H1].d.q + Ye conj[H1].e.l + Yu H2.u.q);
-\end{lstlisting}
-%
-The neutral components of the two Higgs doublets acquire vacuum
-expectation values $v_1$ and $v_2$:
-%
-\begin{lstlisting}
-DEFINITION[EWSB][VEVs] = {
-    {H10, {v1, 1/Sqrt[2]},
-          {sigma1, \[ImaginaryI]/Sqrt[2]}, {phi1, 1/Sqrt[2]}},
-    {H20, {v2, 1/Sqrt[2]},
-          {sigma2, \[ImaginaryI]/Sqrt[2]}, {phi2, 1/Sqrt[2]}}
-};
-\end{lstlisting}
-%
-and the Higgs field components mix to \CP-even, \CP-odd and charged
-Higgs mass eigenstates \code{hh}, \code{Ah} and \code{Hm},
-respectively:
-%
-\begin{lstlisting}
-DEFINITION[EWSB][MatterSector] = {
-    {{phi1, phi2}          , {hh, ZH}},
-    {{sigma1, sigma2}      , {Ah, ZA}},
-    {{conj[H1p], conj[H2p]}, {Hm, ZP}}
-};
-\end{lstlisting}
-%
-Now we need a \fs model file in which we specify the boundary
-conditions for all \THDM-II parameters.  As input we use the \MSbar
-parameter
-$\tan\beta^{\THDMII}(M_t)$, the MSSM \DRbar parameters
-$\mu(\MS)$ and $A_f(\MS)$
-($f=t,b,\tau$) at the SUSY scale, and the \MSbar \CP-odd Higgs mass
-$m_A^{\THDMII}(\MEWSB)$, where \MEWSB is the scale of the electroweak
-symmetry breaking, at which we calculate the light \CP-even Higgs pole mass in the
-end:
-%
-\begin{lstlisting}
-MINPAR = {
-    {3, TanBeta}
-};
-
-EXTPAR = {
-    {0, MSUSY},
-    {1, MEWSB},
-    {2, MuInput},
-    {6, MAInput},
-    {7, AtInput},
-    {8, AbInput},
-    {9, AtauInput}
-};
-\end{lstlisting}
-%
-At the low-energy scale
-$M_t$, we let \fs calculate the gauge and Yukawa couplings of the
-\THDMII and we fix the two Higgs VEVs using
-$m_Z^{\THDMII}(M_t)$ and $\tan\beta^{\THDMII}(M_t)$:
-%
-\begin{lstlisting}
-LowScale = LowEnergyConstant[MT];
-
-LowScaleInput = {
-   {Yu, Automatic},
-   {Yd, Automatic},
-   {Ye, Automatic},
-   {v1, 2 MZMSbar / Sqrt[GUTNormalization[g1]^2 g1^2 + g2^2] \
-        Cos[ArcTan[TanBeta]]},
-   {v2, 2 MZMSbar / Sqrt[GUTNormalization[g1]^2 g1^2 + g2^2] \
-        Sin[ArcTan[TanBeta]]}
-};
-\end{lstlisting}
-%
-At the scale $\MEWSB$, we fix the \code{M122} parameter using the input
-value of $m_A^{\THDMII}(\MEWSB)$ and we impose the EWSB conditions by
-fixing \code{M112} and \code{M222}:
-%
-\begin{lstlisting}
-EWSBOutputParameters = { M112, M222 };
-
-SUSYScale = MEWSB;
-
-SUSYScaleInput = {
-    {M122, MAInput^2 Sin[ArcTan[v2/v1]] Cos[ArcTan[v2/v1]]},
-    FSSolveEWSBFor[EWSBOutputParameters]
-};
-\end{lstlisting}
-%
-Finally, we need to fix the quartic Higgs couplings of the \THDMII at
-the scale $\MS$.  The necessary relations between the MSSM parameters
-and the quartic Higgs couplings of the \THDMII are known at the full
-1-loop and leading 2-loop level
-\cite{Haber:1993an,Gorbahn:2009pp,Lee:2015uza}.
-We can use expressions from these references to write the boundary conditions
-on the quartic Higgs couplings at the SUSY scale,
-shown in lines 23--57 of \appref{app:THDMIIMSSMBC},
-in terms of the 1- and 2-loop threshold corrections,
-shown in lines 110--398 of the same listing
-which displays the complete
-\fs model file.
-\figref{fig:THDMIIMSSMBC_Mh_MS} shows the lightest \CP-even
-Higgs pole mass calculated at the 1-loop level with \fs in this EFT
-setup as a function of $\tan\beta^{\THDMII}(M_t)$ and $\MS$ for
-$m_A^{\THDMII}(\MEWSB) = 200\unit{GeV}$, $\MEWSB = M_t$ and maximal
-stop mixing.  The figure shows that using this setup, \fs can reproduce
-the results presented in the left panels of Figure~2 of
-Ref.~\cite{Lee:2015uza}.  This EFT model is
-distributed with the \fs package under the name \modelname{THDMIIMSSMBC}.
-%
-\begin{figure}[tbh]
-  \centering
-  \includegraphics[width=0.49\textwidth]{doc/plots/THDMIIMSSMBC/THDMIIMSSMBC_Mh_MS_TB_MA-200_Xt-0}\hfill
-  \includegraphics[width=0.49\textwidth]{doc/plots/THDMIIMSSMBC/THDMIIMSSMBC_Mh_MS_TB_MA-200_Xt-sqrt6}
-  \caption{Lightest \CP-even Higgs pole mass calculated at the 1-loop
-    level in the effective \THDM setup as a function of $\tan\beta$ and
-    $\MS$ for $m_A(M_t) = 200\unit{GeV}$, $A_t = \mu/\tan\beta +
-    X_t$, $A_b = A_\tau = A_t$, $\mu = \MS$, $M_t = 173.34\unit{GeV}$
-    and $\as^{\SM(5)}(M_Z) = 0.1184$.  The left panel shows the results for $X_t
-    = 0$ and the right panel for $X_t = \sqrt{6}\MS$.  The solid line
-    corresponds to a Higgs pole mass of $125\unit{GeV}$ and the dashed
-    lines to $124\unit{GeV}$ and $126\unit{GeV}$, respectively.}
-  \label{fig:THDMIIMSSMBC_Mh_MS}
-\end{figure}
-
-\subsection{Application: High-scale MSSM (\HSSUSY)}
-\label{sec:application_HSSUSY}
-
-\fbsm has already been applied in
-Refs.~\cite{Bagnaschi:2015pwa,Athron:2016fuq,Bagnaschi:2017xid} to perform a
-pure EFT calculation of the lightest \CP-even Higgs mass in the MSSM,
-assuming that all SUSY particles are integrated out at a heavy SUSY
-scale $\MS$.  The \fs spectrum generator constructed for this purpose
-is called \HSSUSY and is based on \sarah's SM model file
-(\modelname{SM}).  In \fs 1.2.3, \HSSUSY implemented the 1-loop and leading
-2-loop threshold corrections of $O(\at\as + \at^2)$ to the quartic Higgs
-coupling of the SM from
-Refs.~\cite{Bagnaschi:2014rsa,Vega:2015fna} at the SUSY scale.  In
-\fstwo, the generalized 2-loop expressions of $O(\at^2)$ for general stop
-masses as well as the new 2-loop contributions from Ref.~\cite{Bagnaschi:2017xid},
-which involve the bottom and tau Yukawa couplings, are
-included.  As a result, the version of \HSSUSY included in \fstwo uses
-the 2-loop threshold corrections of
-$O(\at\as + \ab\as + (\at+\ab)^2 + \ab\atau + \atau^2)$ for general
-SUSY spectra.
-
-In the \HSSUSY model file
-(\code{model_files/HSSUSY/FlexibleSUSY.m.in}), these threshold
-corrections are implemented as \mathematica expressions in the
-high-scale boundary condition:
-%
-\begin{lstlisting}
-HighScaleInput = {
-    {\[Lambda], lambdaTree                     (* tree-level *)
-                + lambda1LReg + lambda1LPhi    (* 1-loop *)
-                + lambda1LChi1 + lambda1LChi2  (* 1-loop *)
-                + lambda1Lbottom + lambda1Ltau (* 1-loop *)
-                + ...                          (* 2-loop *)
-    }
-};
-
-(* arXiv:1407.4081, Eq. (3) *)
-lambdaTree = 1/4 (g2^2 + 3/5 g1^2) Cos[2 ArcTan[TanBeta]]^2;
-
-(* arXiv:1407.4081, Eq. (9) *)
-lambda1LReg = 1/(4 Pi)^2 (
-    - 9/100 g1^4 - 3/10 g1^2 g2^2
-    - (3/4 - Cos[2 ArcTan[TanBeta]]^2/6) * g2^4
-    );
-
-(* arXiv:1407.4081, Eq. (10) *)
-lambda1LPhi = 1/(4 Pi)^2 (
-    3 Yu[3,3]^2 (
-        Yu[3,3]^2
-        + 1/2 (g2^2-g1^2/5) Cos[2 ArcTan[TanBeta]]
-      ) Log[msq2[3,3]/SCALE^2]
-    + 3 Yu[3,3]^2 (
-        Yu[3,3]^2
-        + 2/5 g1^2 Cos[2 ArcTan[TanBeta]]
-      ) Log[msu2[3,3]/SCALE^2]
-    + Cos[2 ArcTan[TanBeta]]^2/300 (
-        3 (g1^4 + 25 g2^4) (
-            + Log[msq2[1,1]/SCALE^2]
-            + Log[msq2[2,2]/SCALE^2]
-            + Log[msq2[3,3]/SCALE^2]
-        )
-        + ...
-      )
-    + ...
-    );
-
-...
-\end{lstlisting}
-%
-The full expressions for the threshold corrections can be found in the
-\HSSUSY model file.  In addition, \HSSUSY makes use of the known
-3-loop SM $\beta$ functions, 3-loop corrections to the
-running top Yukawa coupling, 3-loop threshold corrections to the
-strong coupling and up to 3-loop corrections to the Higgs pole mass:
-%
-\begin{lstlisting}
-UseSM3LoopRGEs    = True; (* 3-loop RGEs *)
-UseYukawa3LoopQCD = True; (* 3-loop thresholds for yt *)
-UseSMAlphaS3Loop  = True; (* 3-loop thresholds for alpha_s *)
-UseHiggs2LoopSM   = True; (* 2-loop contributions to Mh *)
-UseHiggs3LoopSM   = True; (* 3-loop contributions to Mh *)
-\end{lstlisting}
-%
-With all these corrections enabled, \HSSUSY can be regarded as an
-improved variant of \susyhd \cite{Vega:2015fna}, the difference being
-that \HSSUSY includes generalized 2-loop threshold corrections also
-involving $\ab$ and $\atau$, which are not present in \susyhd.  In
-\secref{sec:feft}, \HSSUSY is also compared to the fixed-order
-calculation in the full MSSM as well as to \feft.  The following
-Examples \ref{ex:HSSUSY_parallel}--\ref{ex:HSSUSY_EFT_uncertainty}
-show how \HSSUSY can be run and how different sources of uncertainty
-can be estimated.\footnote{The example scripts can be found in the
-  \code{doc/examples-2.0/} sub-directory of the \fs package.}
-%
-\begin{example}[label=ex:HSSUSY_parallel]
-  For illustration, we show in this example the Higgs mass prediction
-  in the MSSM with \HSSUSY.  In the following example script, a scan
-  over the relative \DRbar stop mixing parameter $X_t/\MS$ is performed for
-  $\tan\beta(\MS) = 5$ and three different values of the SUSY scale $\MS$.
-  This is done in parallel on all available CPU cores.
-  %
-\begin{lstlisting}
-Get["models/HSSUSY/HSSUSY_librarylink.m"];
-
-CalcMh[TB_, Xt_, MS_] := Module[{handle, spectrum},
-    handle = FSHSSUSYOpenHandle[
-        fsSettings -> {
-            precisionGoal -> 1.*^-5,
-            calculateStandardModelMasses -> 1,
-            poleMassLoopOrder -> 2,
-            ewsbLoopOrder -> 2,
-            betaFunctionLoopOrder -> 3,
-            thresholdCorrectionsLoopOrder -> 2,
-            poleMassScale -> 173.34
-        },
-        fsModelParameters -> {
-            TanBeta -> TB,
-            MEWSB -> 173.34,
-            MSUSY -> MS,
-            M1Input -> MS,
-            M2Input -> MS,
-            M3Input -> MS,
-            MuInput -> MS,
-            mAInput -> MS,
-            AtInput -> (Xt + 1/TB) MS,
-            msq2 -> MS^2 IdentityMatrix[3],
-            msu2 -> MS^2 IdentityMatrix[3],
-            msd2 -> MS^2 IdentityMatrix[3],
-            msl2 -> MS^2 IdentityMatrix[3],
-            mse2 -> MS^2 IdentityMatrix[3],
-            LambdaLoopOrder -> 2,
-            TwoLoopAtAs -> 1,
-            TwoLoopAbAs -> 1,
-            TwoLoopAtAb -> 1,
-            TwoLoopAtauAtau -> 1,
-            TwoLoopAtAt -> 1
-        }
-    ];
-    spec = FSHSSUSYCalculateSpectrum[handle];
-    FSHSSUSYCloseHandle[handle];
-    If[spec =!= $Failed, Pole[M[hh]] /. (HSSUSY /. spec), 0]
-];
-
-LaunchKernels[];
-DistributeDefinitions[CalcMh];
-
-data = {
-    ParallelMap[{#, CalcMh[5, #, 1000 ]}&, Range[-3.5, 3.5, 0.1]],
-    ParallelMap[{#, CalcMh[5, #, 2000 ]}&, Range[-3.5, 3.5, 0.1]],
-    ParallelMap[{#, CalcMh[5, #, 10000]}&, Range[-3.5, 3.5, 0.1]]
-};
-\end{lstlisting}%$
-%
-When plotting this data, the following figure results:
-%
-\begin{center}
-  \includegraphics[width=0.5\textwidth]{doc/plots/HSSUSY/Mh_Xt}
-\end{center}
-\end{example}
-
-\begin{example}[label=ex:HSSUSY_uncertainty]
-  In this example, we slightly extend \exref{ex:HSSUSY_parallel}
-  to make a partial uncertainty estimate of the Higgs pole
-  mass predicted by \HSSUSY at the 2-loop level.  We do this by varying the
-  threshold correction loop orders which determine the running \MSbar
-  top Yukawa coupling
-  $y_t$ and the strong coupling in the SM from 2-loop to
-  3-loop.  In addition, the renormalization scale,
-  at which the Higgs pole mass is calculated,
-  is varied by a factor $2$.  The uncertainty
-  estimated in this way is referred to as ``Standard Model
-  uncertainty'' in the literature \cite{Vega:2015fna,Athron:2016fuq}
-  and is one part of the full uncertainty of the EFT calculation
-  of \HSSUSY.
-  %
-  \begin{lstlisting}
-Get["models/HSSUSY/HSSUSY_librarylink.m"];
-
-(* generate logarithmically spaced range [start, stop] *)
-LogRange[start_, stop_, steps_] :=
-    Exp /@ Range[Log[start], Log[stop],
-                 (Log[stop] - Log[start])/steps];
-
-(* generate logarithmically spaced range [Q / 2, 2 Q] *)
-GenerateScales[Q_] := LogRange[Q/2, 2 Q, 10];
-
-CalcMh[MS_, TB_, Xt_, ytLoops_, asLoops_, Qpole_] :=
-    Module[{handle, spec},
-    handle = FSHSSUSYOpenHandle[
-        fsSettings -> {
-            precisionGoal -> 1.*^-5,
-            calculateStandardModelMasses -> 1,
-            poleMassLoopOrder -> 2,
-            ewsbLoopOrder -> 2,
-            betaFunctionLoopOrder -> 3,
-            thresholdCorrectionsLoopOrder -> 3,
-            poleMassScale -> Qpole,
-            thresholdCorrections -> 120111021 +
-                ytLoops * 10^6 + asLoops * 10^2
-        },
-        fsModelParameters -> {
-            TanBeta -> TB,
-            MEWSB -> 173.34,
-            MSUSY -> MS,
-            M1Input -> MS,
-            M2Input -> MS,
-            M3Input -> MS,
-            MuInput -> MS,
-            mAInput -> MS,
-            AtInput -> (Xt + 1/TB) * MS,
-            msq2 -> MS^2 IdentityMatrix[3],
-            msu2 -> MS^2 IdentityMatrix[3],
-            msd2 -> MS^2 IdentityMatrix[3],
-            msl2 -> MS^2 IdentityMatrix[3],
-            mse2 -> MS^2 IdentityMatrix[3],
-            LambdaLoopOrder -> 2,
-            TwoLoopAtAs -> 1,
-            TwoLoopAbAs -> 1,
-            TwoLoopAtAb -> 1,
-            TwoLoopAtauAtau -> 1,
-            TwoLoopAtAt -> 1
-        }
-    ];
-    spec = FSHSSUSYCalculateSpectrum[handle];
-    FSHSSUSYCloseHandle[handle];
-    If[spec =!= $Failed, Pole[M[hh]] /. (HSSUSY /. spec), 0]
-];
-
-(* calculate Higgs mass with uncertainty estimate *)
-CalcDMh[MS_, TB_, Xt_] :=
-    Module[{Mh, MhYt3L, MhAs3L, varyQpole, DMh},
-           Mh     = CalcMh[MS, TB, Xt, 2, 2, 0];
-           MhYt3L = CalcMh[MS, TB, Xt, 3, 2, 0];
-           MhAs3L = CalcMh[MS, TB, Xt, 2, 3, 0];
-           varyQpole = CalcMh[MS, TB, Xt, 2, 2, #]& /@
-                       GenerateScales[173.34];
-           (* combine uncertainty estimates *)
-           DMh = Max[Abs[Max[varyQpole] - Mh],
-                     Abs[Min[varyQpole] - Mh]] +
-                 Abs[Mh - MhYt3L] + Abs[Mh - MhAs3L];
-           { Mh, DMh }
-          ];
-
-LaunchKernels[];
-DistributeDefinitions[CalcDMh];
-
-data = {
-    ParallelMap[{#, CalcDMh[1000 , 5, #]}&, Range[-3.5, 3.5, 0.1]],
-    ParallelMap[{#, CalcDMh[2000 , 5, #]}&, Range[-3.5, 3.5, 0.1]],
-    ParallelMap[{#, CalcDMh[10000, 5, #]}&, Range[-3.5, 3.5, 0.1]]
-};
-\end{lstlisting}%$
-%
-In the function \code{CalcDMh[]}, the three sources of uncertainty are
-combined linearly.  When drawing the uncertainty $\Delta
-M_h^{\SM}$, estimated in this way, around the central value as $M_h
-\pm \Delta M_h^{\SM}$, the following figure results:
-\begin{center}
-  \includegraphics[width=0.5\textwidth]{doc/plots/HSSUSY/Mh_Xt_uncertainty}
-\end{center}
-%
-We find that the uncertainty $\Delta
-M_h^{\SM}$ is around or below $500\unit{MeV}$.
-\end{example}
-
-\begin{example}[label=ex:HSSUSY_EFT_uncertainty]
-  In this example, we make an estimate of the ``EFT uncertainty''
-  for the Higgs pole mass predicted by the pure-EFT calculation of
-  \HSSUSY in the MSSM\@.  The pure-EFT calculation performed by \HSSUSY
-  neglects terms of
-  $O(v^2/\MS^2)$.  We estimate these missing terms by multiplying the
-  individual 1-loop contributions by the term $(1 +
-  \texttt{DeltaEFT}\times
-  v^2/\MS^2)$, where \texttt{DeltaEFT} is an input parameter of
-  \HSSUSY.  By varying \texttt{DeltaEFT} between $0$ and
-  $1$ we obtain an estimate of the effect of these missing terms.
-  This method has also been used in
-  Refs.~\cite{Vega:2015fna,Athron:2016fuq,Bagnaschi:2017xid}.  The
-  following code snippet illustrates how to calculate the Higgs mass
-  with \HSSUSY and estimate the ``EFT uncertainty'' as a function of
-  the SUSY scale for maximal stop mixing, $X_t =
-  \sqrt{6}\MS$ and $\tan\beta = 5$.
-  %
-  \begin{lstlisting}
-Get["models/HSSUSY/HSSUSY_librarylink.m"];
-
-(* generate logarithmically spaced range [start, stop] *)
-LogRange[start_, stop_, steps_] :=
-    Exp /@ Range[Log[start], Log[stop],
-                 (Log[stop] - Log[start])/steps];
-
-CalcMh[MS_, TB_, Xt_, deltaEFT_] :=
-    Module[{handle, spec},
-    handle = FSHSSUSYOpenHandle[
-        fsSettings -> {
-            precisionGoal -> 1.*^-5,
-            calculateStandardModelMasses -> 1,
-            poleMassLoopOrder -> 2,
-            ewsbLoopOrder -> 2,
-            betaFunctionLoopOrder -> 3,
-            thresholdCorrectionsLoopOrder -> 3,
-            thresholdCorrections -> 122111221
-        },
-        fsModelParameters -> {
-            TanBeta -> TB,
-            MEWSB -> 173.34,
-            MSUSY -> MS,
-            M1Input -> MS,
-            M2Input -> MS,
-            M3Input -> MS,
-            MuInput -> MS,
-            mAInput -> MS,
-            AtInput -> (Xt + 1/TB) * MS,
-            msq2 -> MS^2 IdentityMatrix[3],
-            msu2 -> MS^2 IdentityMatrix[3],
-            msd2 -> MS^2 IdentityMatrix[3],
-            msl2 -> MS^2 IdentityMatrix[3],
-            mse2 -> MS^2 IdentityMatrix[3],
-            LambdaLoopOrder -> 2,
-            TwoLoopAtAs -> 1,
-            TwoLoopAbAs -> 1,
-            TwoLoopAtAb -> 1,
-            TwoLoopAtauAtau -> 1,
-            TwoLoopAtAt -> 1,
-            DeltaEFT -> deltaEFT
-        }
-    ];
-    spec = FSHSSUSYCalculateSpectrum[handle];
-    FSHSSUSYCloseHandle[handle];
-    If[spec =!= $Failed, Pole[M[hh]] /. (HSSUSY /. spec), 0]
-];
-
-(* calculate Higgs mass with uncertainty estimate *)
-CalcDMh[MS_, TB_, Xt_] :=
-    Module[{Mh, MhEFT},
-           Mh    = CalcMh[MS, TB, Xt, 0];
-           MhEFT = CalcMh[MS, TB, Xt, 1];
-           { Mh, Abs[Mh - MhEFT] }
-          ];
-
-LaunchKernels[];
-DistributeDefinitions[CalcDMh];
-
-data = ParallelMap[{#, Sequence @@ CalcDMh[#, 5, Sqrt[6]]}&,
-                   LogRange[173.34, 10^5, 60]];
-\end{lstlisting}%$
-%
-In the function \code{CalcDMh[]}, the difference between the Higgs masses
-calculated with $\texttt{DeltaEFT} = 0$ and $\texttt{DeltaEFT} =
-1$ is used as the uncertainty estimate ($\texttt{DeltaEFT} =
-0$ corresponds to the standard \HSSUSY calculation).  When drawing the
-uncertainty estimated in this way symmetrically around the central
-value, the following figure results:
-%
-\begin{center}
-  \includegraphics[width=0.5\textwidth]{doc/plots/HSSUSY/Mh_MS_EFT_uncertainty}
-\end{center}
-%
-As expected, we find that the ``EFT uncertainty'' decreases as
-$\MS$ increases and falls below $150\unit{MeV}$ for $\MS \gtrsim 2\unit{TeV}$.
-\end{example}
-
-\subsection{Tests and comparisons with other spectrum generators}
-
-We have performed various direct and indirect tests of \fbsm and its
-components to ensure the correctness of the code:
-%
-\begin{itemize}
-\item We have performed an analytic comparison of the RGEs generated
-  with \sarah 4.5.3 for \fs's split-MSSM model file
-  (\modelname{SplitMSSM}) against the RGEs presented in
-  Ref.~\cite{Benakli:2013msa} and we found exact agreement.
-\item We have performed a detailed numerical comparison of \fs's
-  \HSSUSY model against \susyhd 1.0.1 and found excellent agreement
-  \cite{Athron:2016fuq}.  The small differences between the two
-  programs are of $O(100\unit{MeV})$ and originate from a
-  different determination of $y_t$ in the SM at the low-energy scale,
-  a different procedure to calculate the Higgs pole mass in the SM and
-  the inclusion of additional 2-loop corrections in \HSSUSY which
-  involve $\ab$ and $\atau$.
-\item The effective \THDM shown in \secref{sec:application_THDM} and
-  \appref{app:THDMIIMSSMBC} reproduces the results of
-  Ref.~\cite{Lee:2015uza} and \MhEFT 1.0 \cite{MhEFT} for scenarios with
-  heavy Higgsinos and gauginos, see for example
-  \figref{fig:THDMIIMSSMBC_Mh_MS}.
-\item The \fs package contains various EFT scenarios of the MSSM with
-  boundary conditions from the literature (\HSSUSY,
-  \modelname{SplitMSSM}, \modelname{THDMIIMSSMBC}, \modelname{HTHDMIIMSSMBC},
-  \modelname{HGTHDMIIMSSMBC}).  For all these models, we have performed
-  various analytic tests of the implemented MSSM boundary conditions,
-  checking for example the renormalization scale dependence, relations
-  among the parameters and threshold corrections at the
-  1-loop level.
-\item We have checked the numeric equality of the 3-loop MSSM $\beta$
-  functions implemented in \fs and \softsusy 3.7.0.
-\item We have checked the correctness of the renormalization scale
-  dependent part of the 2-loop QCD corrections to the \MSbar top
-  Yukawa coupling in the SM by deriving the 2-loop
-  threshold corrections for $y_t$ from the SM to the MSSM
-  and checking that no large logarithms appear
-  \cite{ThomasKwasnitza:2016yqj}.
-\item We have also analytically checked the correctness of the
-  renormalization scale dependent part of the 2-loop and 3-loop QCD
-  corrections to the \MSbar top Yukawa coupling in the SM
-  by proving the renormalization scale invariance of the top quark
-  pole mass in the SM at the 3-loop QCD level.
-\end{itemize}
-
-\section{\famu extension}
-\label{sec:famu}
-
-\fstwo introduces a calculation of the BSM contributions to the
-anomalous magnetic moment of the muon, \amuBSM, in the \MSbar/\DRbar
-scheme at the 1-loop level in the model under consideration plus the
-universal 2-loop QED contributions
-\cite{Degrassi:1998es,vonWeitershausen:2010zr}.  The 1-loop
-diagram types that are taken into account are shown in
-\figref{fig:amu_diagram_types}.
-%
-\begin{figure}[tbh]
-  \centering
-  \subfloat[Type FFS]{
-    \includegraphics{doc/Feynman/amuFFS}
-    \label{fig:amu_diagram_I}
-  }
-  \qquad
-  \subfloat[Type SSF]{
-    \includegraphics{doc/Feynman/amuFSS}
-    \label{fig:amu_diagram_II}
-  }
-  \caption{Generic diagram types implemented in \fs to calculate the
-    1-loop contribution to $\amuBSM$.  The solid line in the loop
-    represents any contributing non-SM fermion $F$ and
-    the dashed line any contributing scalar particle $S$.}
-  \label{fig:amu_diagram_types}
-\end{figure}
-%
-The general result for these two Feynman diagram types reads, in a
-notation based on
-Refs.~\cite{Stockinger:2006zn,Martin:2001st,Fargnoli:2013zia},
-%
-\begin{align}
-  \label{eq:amu BSM 1L FFS}
-  \amu^{\BSM\text{,1L,FFS}} &=
-  \frac{c}{(4\pi)^2} \frac{M_\mu m_\mu}{m_{S}^2}\left(
-    \frac{1}{12} A_{SF} F_1^C(x) + \frac{m_F}{3m_\mu} B_{SF} F_2^C(x)
-  \right) \,, \\
-  \label{eq:amu BSM 1L SSF}
-  \amu^{\BSM\text{,1L,SSF}} &=
-  - \frac{c}{(4\pi)^2} \frac{M_\mu m_\mu}{m_{S}^2}\left(
-    \frac{1}{12} A_{SF} F_1^N(x) + \frac{m_F}{6m_\mu} B_{SF} F_2^N(x)
-  \right) \,,
-\end{align}
-%
-where $x=m_F^2/m_{S}^2$ is the squared running mass ratio between $F$ and $S$, $M_\mu$ is the muon pole mass,
-$m_\mu$ is the muon \MSbar/\DRbar mass and $c$ denotes the electric charge of
-the particle coupling to the photon divided by the muon charge.  The
-$A_{SF}$ and $B_{SF}$ constants are defined as
-%
-\begin{align}
-  \label{eq:ASF}
-  A_{SF} &= z_L z_L^* + z_R z_R^* \,,\\
-  \label{eq:BSF}
-  B_{SF} &= z_L z_R^* + z_R z_L^* \,,
-\end{align}
-%
-where $z_L$ and $z_R$ are the left- and right-handed components of the
-scalar--fermion--muon vertex, $\Gamma_{SF\mu} =
-ie(z_L P_L + z_R P_R)$ with $e$ being the \MSbar/\DRbar electromagnetic coupling constant.
-%
-The loop functions $F_i^C(x)$ and $F_i^N(x)$ read
-%
-\begin{align}
-  F_1^C(x) &= \frac{2}{(1-x)^4} \Big[ 2 + 3x - 6x^2 + x^3 + 6x\log x \Big], \\
-  F_2^C(x) &= \frac{3}{2(1-x)^3} \Big[ -3 + 4x - x^2 - 2\log x \Big], \\
-  F_1^N(x) &= \frac{2}{(1-x)^4} \Big[ 1 - 6x + 3x^2 + 2x^3 - 6x^2\log x \Big], \\
-  F_2^N(x) &= \frac{3}{(1-x)^3} \Big[ 1 - x^2 + 2x\log x \Big].
-\end{align}
-%
-To calculate $\amu^{\BSM\text{,1L}}$, \fs sums over all possible
-instantiations of these two 1-loop diagram types with at least one
-non-SM particle in the loop,
-%
-\begin{align}
-  \label{eq:amu BSM 1L}
-  \amu^{\BSM\text{,1L}} =
-  \sum_{\substack{\text{FFS}\\\text{diagrams}}} \amu^{\BSM\text{,1L,FFS}}
-  + \sum_{\substack{\text{SSF}\\\text{diagrams}}} \amu^{\BSM\text{,1L,SSF}} \,.
-\end{align}
-%
-The calculation is performed at the renormalization scale $\MLCP$, which
-is defined as the running mass of the lightest electrically charged
-\BSM particle contributing to $\amu^{\BSM\text{,1L}}$.  In addition to
-this 1-loop \BSM contribution, \famu adds the universal leading
-logarithmic 2-loop QED contribution \cite{Degrassi:1998es},
-%
-\begin{align}
-  \amu^{\QED,\text{2L}} = \amu^{\BSM\text{,1L}} \times \Bigg[
-    16 \, \frac{\aem^\BSM(\MLCP)}{4\pi}
-    \log\frac{M_\mu}{\MLCP}
-  \Bigg] \,.
-\end{align}
-%
-The overall result for \amuBSM is then given by the sum of 1-loop and
-the 2-loop QED contribution,
-%
-\begin{align}
-  \amuBSM = \amu^{\BSM\text{,1L}} + \amu^{\QED,\text{2L}} \,.
-\end{align}
-
-\subsection{Choosing \famu in the model file}
-
-In order to calculate \amuBSM with \fs, the
-\code{FlexibleSUSYObservable`aMuon} symbol has to be added to an
-output block, see \tabref{tab:famu_options} and
-\exref{ex:FlexibleAMU_in_model_file}.
-%
-\begin{OptionTable}{\famu model file options.}{tab:famu_options}[lllX]
-    \multilinecell[t]{\code{FlexibleSUSYObservable`}\\\code{aMuon}}
-    & -- & -- & Represents \amuBSM calculated with \famu \\
-    \multilinecell[t]{\code{FlexibleSUSYObservable`}\\\code{aMuonGM2Calc}}
-    & -- & -- & Represents \amuMSSM calculated with \GMTCalc at 2-loop level \\
-    \multilinecell[t]{\code{FlexibleSUSYObservable`}\\\code{aMuonGM2CalcUncertainty}}
-    & -- & -- & Represents the uncertainty $\Delta\amuMSSM$ calculated with \GMTCalc \\
-    \bottomrule
-\end{OptionTable}
-
-\begin{example}[label=ex:FlexibleAMU_in_model_file]
-In the following code snippet, the \code{FlexibleSUSYObservable`aMuon} symbol
-is added to an output block named \code{FlexibleSUSYLowEnergy}.
-The calculated \amuBSM will be written to the entry $21$ of this
-block.
-%
-\begin{lstlisting}
-ExtraSLHAOutputBlocks = {
-   {FlexibleSUSYLowEnergy,
-           {{21, FlexibleSUSYObservable`aMuon} } }
-};
-\end{lstlisting}
-\end{example}
-
-Note that \amuBSM is treated as an observable in \fs.  The
-calculation of observables can be enabled/disabled using the flag $15$
-of the \code{FlexibleSUSY} block.
-
-As an alternative to \famu, \GMTCalc \cite{Athron:2015rva} can be used
-to calculate \amuMSSM in \fs at the 2-loop level employing results from
-Refs.~\cite{Fargnoli:2013zia,Fargnoli:2013zda,vonWeitershausen:2010zr}
-in MSSM models without flavor violation. This MSSM-specific option was
-first introduced in \fsbreak 1.3.0.  In order to let \GMTCalc calculate the
-anomalous magnetic moment of the muon, the symbol
-\code{FlexibleSUSYObservable`aMuonGM2Calc} must be added to an output
-block.  In addition, the symbol
-\code{FlexibleSUSYObservable`aMuonGM2CalcUncertainty} can be used to
-calculate the estimated corresponding theory uncertainty for \amuMSSM.
-
-\begin{example}
-Calculating \amuMSSM with \GMTCalc in the \modelname{CMSSMNoFV} is
-enabled by defining:
-%
-\begin{lstlisting}
-ExtraSLHAOutputBlocks = {
-   {FlexibleSUSYLowEnergy,
-           {{1, FlexibleSUSYObservable`aMuonGM2Calc},
-            {2, FlexibleSUSYObservable`aMuonGM2CalcUncertainty} } }
-};
-\end{lstlisting}
-\end{example}
-
-\GMTCalc is incorporated into \fs in the form of an addon.
-In order to enable the \GMTCalc addon, the
-\code{--with-addons=GM2Calc} argument can be passed to the
-\code{configure} script during the \fs configuration step.
-
-\subsection{Structure of \famu code}
-
-The C++ interface for \famu is defined in the
-\code{<model>_a_muon.hpp} file.  The interface consists of a single
-function, which takes a model object as the argument and returns the value
-of \amuBSM.
-
-\begin{example}
-In the \modelname{CMSSM}, the interface function reads:
-%
-\begin{lstlisting}
-namespace CMSSM_a_muon {
-   double calculate_a_muon(const CMSSM_mass_eigenstates& model);
-}
-\end{lstlisting}
-\end{example}
-
-\subsection{Tests and comparisons with other calculations}
-
-We have tested \famu in the MSSM against \GMTCalc 1.3.3
-\cite{Athron:2015rva}, which is currently the most precise code
-available to calculate \amuMSSM.  We usually find a $1$--$10\%$
-deviation depending on the parameter choice, see
-\tabref{tab:test_famu}, except for specific parameter points, where
-the deviation is much larger due to a large renormalization scale
-dependence, see below.  The $1$--$10\%$ deviation is caused mainly by
-the fact that \GMTCalc calculates \amuMSSM in an on-shell scheme and
-includes all known 2-loop corrections, while \famu performs the
-calculation in the \MSbar/\DRbar scheme at the lightest charged BSM particle scale $\MLCP$ and only
-includes the leading logarithmic 2-loop QED correction.  A comparison
-of \famu with \spheno 4.0.2 shows differences of up to $10\%$.  These
-differences are caused by the fact that \famu includes the leading
-logarithmic 2-loop QED contribution and by the different choice of the
-renormalization scale: \spheno performs the calculation at the $Z$ pole
-mass scale, while \famu performs it at $\MLCP$.
-
-In specific parameter scenarios, where the \MSbar/\DRbar smuon masses
-show a high sensitivity to the renormalization scale, the predictions
-of $\amuBSM$ in \famu and \spheno are expected to have a large theory
-uncertainty, see for example BM4$^\prime$ in \tabref{tab:test_famu}
-and see the discussion in Ref.~\cite{Athron:2015rva}.  One
-reason for this large uncertainty are the missing 2-loop corrections
-in \famu and \spheno, which would (if included) cancel the
-renormalization scale dependence at the 2-loop level.  To illustrate
-the sensitivity on the renormalization scale, we show in the third
-column of \tabref{tab:test_famu} the variation of $\amuBSM$ in \famu
-when the scale is varied in the interval $[\MLCP/2,2\MLCP]$.  For BM4$^\prime$,
-$\amuBSM$ varies by around $66\%$, which indicates a very imprecise
-prediction for this point.  Such a large scale uncertainty is avoided
-in \GMTCalc due to the renormalization in an on-shell scheme.  Another
-reason for the larger uncertainty in \famu and \spheno is the
-renormalization scheme used to renormalize the smuon masses: In the
-\MSbar/\DRbar scheme, smuon self energy contributions which are
-quadratic in the BSM particle masses can lead to large 1-loop
-corrections to the smuon masses.  Such large corrections are avoided in
-the on-shell scheme, where they are absorbed into the smuon mass
-counter term.
-%
-\begin{table}[tbh]
-  \centering
-  \begin{tabular}{lrrrr}
-    \toprule
-    Point & \famu & \famu           & \GMTCalc & \spheno \\
-          &       & scale variation &          &         \\
-    \midrule
-    SPS1a  & $29.77$ & $0.46$ & $29.31 \pm 2.36$ & $31.00$ \\
-    SPS1b  & $32.46$ & $0.45$ & $32.38 \pm 2.40$ & $32.68$ \\
-    SPS3   & $13.80$ & $0.12$ & $13.52 \pm 2.33$ & $14.99$ \\
-    SPS4   & $50.02$ & $1.02$ & $52.45 \pm 2.64$ & $45.64$ \\
-    BM1$^\prime$  & $42.08$ & $1.58$ & $42.34 \pm 2.33$ & $43.72$ \\
-    BM2$^\prime$  & $25.79$ & $0.10$ & $25.67 \pm 2.32$ & $26.16$ \\
-    BM3$^\prime$  & $27.81$ & $0.68$ & $27.95 \pm 2.34$ & $27.98$ \\
-    BM4$^\prime$  &  $8.11$ & $5.41$ & $33.11 \pm 2.31$ &  $2.19$ \\
-    \bottomrule
-  \end{tabular}
-  \caption{Comparison of $\amuMSSM\cdot 10^{10}$ calculated with \famu,
-    \GMTCalc 1.3.3 and \spheno 4.0.2 for the CMSSM benchmark points
-    presented in Ref.~\cite{Allanach:2002nj} and the parameter points shown
-    in \tabref{tab:bm_amu}.  The third column shows the variation
-    of $\amuBSM$ when the renormalization scale is varied between
-    $\MLCP/2$ and $2\MLCP$.
-    For BM4$^\prime$ the value of \amu calculated by the \DRbar
-    programs \famu and \spheno suffers from a high renormalization scale
-    sensitivity due to the large values of $M_2$ and $(m_{\tilde{l}})_{ii}$,
-    which leads to a very imprecise result and to a huge deviation
-    compared to \GMTCalc.}
-  \label{tab:test_famu}
-\end{table}
-%
-\begin{table}[tbh]
-  \centering
-  \begin{tabular}{lrrrr}
-    \toprule
-                & BM1$^\prime$  & BM2$^\prime$   & BM3$^\prime$   & BM4$^\prime$ \\
-    \midrule
-    $\mu$ / GeV & $350$ & $1300$ & $2000$ & $-160$ \\
-    $\tan\beta$ & $40$  & $40$   & $40$   & $50$ \\
-    $M_1$ / GeV & $150$ & $150$  & $150$  & $140$ \\
-    $M_2$ / GeV & $300$ & $300$  & $300$  & $2000$ \\
-    $(m_{\tilde{e}})_{ii}$ / GeV & $400$ & $400$  & $400$  & $200$ \\
-    $(m_{\tilde{l}})_{ii}$ / GeV & $400$ & $400$  & $400$  & $2000$ \\
-    $(m_{\tilde{q}})_{ii}$, $(m_{\tilde{u}})_{ii}$, $(m_{\tilde{d}})_{ii}$ / GeV
-    & $400$ & $600$  & $700$ & $2000$ \\
-    \bottomrule
-  \end{tabular}
-  \caption{Definition of the MSSM benchmark points BM1$^\prime$--BM4$^\prime$, inspired
-    by the points BM1--BM4 presented in Ref.~\cite{Fargnoli:2013zda}.
-    All parameters are defined in the \DRbar scheme at the scale
-    $Q = 454.7\unit{GeV}$, except for $\tan\beta$, which is defined at $M_Z$.
-    The trilinear couplings and off-diagonal
-    elements of the sfermion mass parameters are set to zero and we
-    have fixed $m_A = 2\unit{TeV}$.}
-  \label{tab:bm_amu}
-\end{table}
-
-\section{\fcpv extension}
-\label{sec:fcpv}
-\subsection{Setting up a \fcpv model}
-
-Since \fs 1.1.0, the model parameters are no longer restricted to be real,
-but can be complex.  Whether a parameter is real or complex is
-specified in the corresponding \sarah model file.  Parameters can be
-forced to be treated as real in \fs by adding them to the
-\code{RealParameters} list in the \fs model file, see
-\tabref{tab:fcpv_options} and the following examples.  For
-compatibility with \fs 1.0, the \code{RealParameters} list is by default
-set to \code{\{All\}} meaning that all parameters are assumed to be real.
-%
-\begin{OptionTable}{\fcpv model file options.}{tab:fcpv_options}
-    \code{RealParameters} & \code{\{ All \}} &
-    List of model parameters or \code{\{\}} or \code{\{ All \}} &
-    List of parameters to be treated as real\\
-    \bottomrule
-\end{OptionTable}
-
-\begin{example}
-In the MSSM, the $\mu$ parameter, the Yukawa
-couplings, the soft-breaking trilinear couplings, the soft-breaking
-scalar mass parameters, the soft-breaking gaugino mass parameters and
-the $B\mu$ parameter can be complex.  In order to choose all of these
-parameters to be complex, except for $B\mu$, one can set
-%
-\begin{lstlisting}
-RealParameters = { B[\[Mu]] };
-\end{lstlisting}
-\end{example}
-
-\begin{example}
-In order to treat all MSSM parameters defined in
-the \sarah model file for the MSSM as complex, set \code{RealParameters} to the
-empty list:
-%
-\begin{lstlisting}
-RealParameters = {};
-\end{lstlisting}
-\end{example}
-
-\subsection{Application: CMSSM with \CP-violation}
-
-In \sarah's \modelname{MSSM} model, the phase factor between the two Higgs
-doublets is set to zero.  Therefore, this model does not allow for
-\CP-violation in the Higgs sector.  In order to enable \CP-violation in
-the MSSM Higgs sector, \sarah's \modelname{MSSM/CPV} model file can be
-used, which allows for a non-zero relative phase factor $e^{i\eta}$
-between the Higgs doublets.  In the \modelname{MSSM/CPV}, there are three
-linearly independent EWSB equations.  Therefore, in this model three
-EWSB output parameters have to be chosen.  In \fs's \modelname{CMSSMCPV}
-model file, these are chosen to be $\re B\mu$, $\im B\mu$ and $|\mu|$
-by setting the \code{EWSBOutputParameters} variable to
-%
-\begin{lstlisting}
-EWSBOutputParameters = { Re[B[\[Mu]]], Im[B[\[Mu]]], \[Mu] };
-\end{lstlisting}
-%
-Since only the magnitude of the $\mu$ parameter is fixed by the EWSB
-equations, \fs introduces the phase of $\mu$ as a free parameter,
-\code{Phase[\\[Mu]]} = $e^{i\phi_\mu}$.  This phase should be specified in
-an SLHA-2 compliant way by reading the real and imaginary parts of
-$e^{i\phi_\mu} = \cos\phi_\mu + i \sin\phi_\mu$ from the \code{MINPAR}
-and \code{IMMINPAR} block entries $4$ and fixing $e^{i\phi_\mu}$ at the
-SUSY scale:
-%
-\begin{lstlisting}
-MINPAR   = { {4, CosPhiMu} };
-IMMINPAR = { {4, SinPhiMu} };
-
-SUSYScaleInput = {
-    {Phase[\[Mu]], CosPhiMu + I SinPhiMu}
-};
-\end{lstlisting}
-%
-In the \modelname{CMSSMCPV}, the phase $\eta$ is read from the \code{EXTPAR}
-block entry $100$ and also chosen to be input at the SUSY scale:
-%
-\begin{lstlisting}
-EXTPAR = {
-    {100, etaInput}
-};
-
-SUSYScaleInput = {
-    {eta, etaInput}
-};
-\end{lstlisting}
-%
-The complete \fs \modelname{CMSSMCPV} model file can be found in
-\appref{app:CMSSMCPV}.
-
-\subsection{Application: Electric dipole moments of fermions}
-
-\fstwo can calculate $\edmBSM{f}$, new physics contributions to the
-electric dipole moment (EDM) of a fermion $f$, in the given model
-in the \DRbar scheme at the 1-loop level.
-The procedure is very similar to the calculation of \amuBSM
-described in \secref{sec:famu}.
-This is expected from the following effective Lagrangian,
-\begin{equation}
-  \Delta \mathcal{L}_\text{eff} =
-  - \frac{D_f}{2}\bar{f}_L \sigma_{\mu\nu} f_R F^{\mu\nu} + \mathrm{h.c.},
-\end{equation}
-where the real and the imaginary parts of the Wilson coefficient
-$D_f$ are proportional to the magnetic
-(see, e.g., Refs.~\cite{Cho:2001hx,Baek:2002cc})
-and the electric (see, e.g., Ref.~\cite{Pokorski:1999hz}) dipole moments
-of $f$, respectively.
-More precisely,
-\begin{equation}
-  a_f = - \frac{2 M_f}{e} \re D_f, \qquad
-  \edm{f} = \im D_f,
-\end{equation}
-where $M_f$ is the pole mass of $f$
-and $e$ is the running electromagnetic coupling constant in the BSM model.
-It is then obvious that
-the EDM is given by the sum of all FFS-type and SSF-type diagrams,
-\begin{align}
-  \edm{f}^{\BSM\text{,1L}} =
-  \sum_{\substack{\text{FFS}\\\text{diagrams}}}
-  \edm{f}^{\BSM\text{,1L,FFS}}
-  + \sum_{\substack{\text{SSF}\\\text{diagrams}}}
-  \edm{f}^{\BSM\text{,1L,SSF}}
-  \,,
-\end{align}
-as in Eq.\ \eqref{eq:amu BSM 1L}.  The type of diagram refers to those
-shown in \figref{fig:amu_diagram_types}, resulting in the contributions,
-\begin{align}
-  \frac{1}{e}\,\edm{f}^{\BSM\text{,1L,FFS}} &=
-  \frac{c}{(4\pi)^2}
-    \frac{m_F}{6{m_{S}^2}} \widetilde{B}_{SF} F_2^C(x)
-  \,, \\
-  \frac{1}{e}\,\edm{f}^{\BSM\text{,1L,SSF}} &=
-  - \frac{c}{(4\pi)^2}
-    \frac{m_F}{12{m_{S}^2}} \widetilde{B}_{SF} F_2^N(x)
-  \,,
-\end{align}
-which are essentially the second terms of Eq.\ \eqref{eq:amu BSM 1L FFS} and
-Eq.\ \eqref{eq:amu BSM 1L SSF}, respectively, divided by $2 M_f$, except that
-the coupling factor is instead
-\begin{equation}
-  \widetilde{B}_{SF} = 2 \im ( z_L z_R^* ) .
-\end{equation}
-There are no imaginary parts corresponding to
-the first terms of Eq.\ \eqref{eq:amu BSM 1L FFS} and Eq.\ \eqref{eq:amu BSM 1L SSF}
-as can be guessed from Eq.\ \eqref{eq:ASF}.
-The calculation is performed at the renormalization scale $\MS$ specified
-in the model file, which is typically set to the stop mass scale
-in supersymmetric models.
-
-One can have the EDM of particle \code{f} calculated
-by adding to an output block the form: \code{FlexibleSUSYObservable`EDM[f_]}.
-Then $\edm{f}$ is reported in units of $\unit{GeV}^{-1}$.
-\begin{example}
-The output includes
-the EDMs of the electron, muon, and tau if
-the following code snippet is inserted into the \fs model file:
-\begin{lstlisting}
-ExtraSLHAOutputBlocks = {
-   {FlexibleSUSYLowEnergy,
-           {{23, FlexibleSUSYObservable`EDM[Fe[1]]},
-            {24, FlexibleSUSYObservable`EDM[Fe[2]]},
-            {25, FlexibleSUSYObservable`EDM[Fe[3]]} } }
-};
-\end{lstlisting}
-\end{example}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=0.49\textwidth]{doc/plots/MSSMCPV/contour_de_PhiMuM12}\hfill
-  \includegraphics[width=0.49\textwidth]{doc/plots/MSSMCPV/contour_de_PhiMuTe}
-  \caption{Left panel: contours of the electron EDM in the
-    \CP-violating MSSM at the 1-loop level as a function of
-    the phases of $\mhalf = M_1 = M_2$ and $\mu$,
-    with $\mu A_e$ kept positive real.
-    Right panel: same plot except for the vertical axis showing
-    the $A_e$ phase and $\mu \mhalf$ kept positive real.
-    In both panels, the other relevant parameters are:
-    $\tan\beta = 10$, $|\mu| = |\mhalf| = \mzero = |A_e| = 2\unit{TeV}$.
-    The thick red lines show the experimental upper bound on $|\edm{e}|$
-    \cite{Baron:2013eja}.}
-  \label{fig:EDM plots}
-\end{figure}
-Using a model file thus configured, one can do a quick test of
-the electron EDM evaluation in the \CP-violating MSSM for instance.
-In this model, only $\arg(\mu M_{1,2,3})$ and $\arg (\mu A_f)$ are
-physical among the flavor-conserving phases
-apart from those already present in the SM
-(see, e.g., Refs.~\cite{Pokorski:1999hz,Stockinger:2006zn}).
-For simplicity,
-the gaugino masses are assumed to have an equal (complex) value
-$\mhalf$, and both soft masses of the left- and right-handed
-selectron as well as the approximate tree-level heavy Higgs masses%
-\footnote{$\sqrt{\re\!B\mu / (\cos\!\beta \sin\!\beta)}$.}
-shall be $\mzero$.
-The moduli of $\mu$, $\mhalf$, $\mzero$, and $A_e$
-are specified at the SUSY scale,
-all of which including the scale are set to $2\unit{TeV}$.
-Finally, we set $\tan\beta = 10$ at the $M_Z$ scale.
-
-The resulting electron EDM is displayed as contours in
-\figref{fig:EDM plots}.
-In the left panel, $\phi_{\mhalf} \equiv \arg \mhalf$ and
-$\phi_\mu \equiv \arg \mu$ are varied
-while $\arg(\mu A_e)$ is constrained to be zero.
-From the directions of the contours it is clear
-that \fs correctly reproduces the behaviour of
-$\edm{e}$ depending only on the rephasing invariant $\arg(\mu \mhalf)$.
-In the right panel, the roles of $\phi_{A_e} \equiv \arg A_e$ and
-$\phi_{\mhalf}$ are swapped so that $\arg(\mu \mhalf)$ stays at zero.
-Again, the contours verify that
-$\edm{e}$ from \fs is determined by the physical phase $\arg(\mu A_e)$.
-For reference,
-the experimental upper limit on $|\edm{e}|$ at the 90\% confidence level
-is shown as the thick red lines \cite{Baron:2013eja}.
-
-\subsection{Tests and comparisons with other spectrum generators}
-%
-In the left panel of \figref{fig:CMSSMCPV_plots}, we show the lightest
-Higgs pole mass calculated in the \CP-violating CMSSM at the 1-loop
-level with \fstwo and \spheno~4.0.2 as a function of the phase angle
-of the complex $\mu$ parameter, $\phi_\mu \equiv \arg \mu$.
-We use a low-energy scenario with $\mzero = \mhalf = 500\unit{GeV}$,
-$\tan\beta = 10$ and $\azero = 0$.  Even though this scenario is
-excluded, the figure illustrates that both \fs and \spheno show the
-same behaviour of the Higgs pole mass as a function of $\phi_\mu$.
-The $\phi_\mu$-independent shift of around $0.7\unit{GeV}$ between the
-Higgs masses calculated by the two programs is mainly caused by the
-different treatment of higher-order corrections to
-the running \DRbar top Yukawa coupling, which has been discussed in
-Refs.~\cite{Staub:2015aea,ThomasKwasnitza:2016yqj}.
-%
-\begin{figure}[tbh]
-  \centering
-  \includegraphics[width=0.49\textwidth]{doc/plots/CMSSMCPV/scan_Mh_PhiMu_m0-500_M12-500_TB-10_A0-0}\hfill
-  \includegraphics[width=0.49\textwidth]{doc/plots/CMSSMCPV/scan_TYu33_Q_m0-500_M12-500_TB-10_A0-500_PhiMu-0_PhiA0-Pi4}
-  \caption{Left panel: lightest Higgs pole mass calculated in the
-    \CP-violating CMSSM at the 1-loop level with \fstwo and
-    \spheno 4.0.2 as a function of the phase angle of the $\mu$
-    parameter for $\mzero = \mhalf = 500\unit{GeV}$, $\tan\beta = 10$
-    and $\azero = 0$.  Right panel: renormalization group running of the
-    trilinear coupling $(T^u)_{33}$ for $\mzero = \mhalf =
-    500\unit{GeV}$, $\tan\beta = 10$, $\phi_\mu = 0$ and $\azero = 500\,
-    e^{i \pi/4} \unit{GeV}$.}
-  \label{fig:CMSSMCPV_plots}
-\end{figure}
-
-The right panel of \figref{fig:CMSSMCPV_plots} shows the
-renormalization group running of the $(3,3)$-component of the complex
-trilinear coupling $T^u$ for the same CMSSM scenario, except that we
-use $\phi_\mu = 0$ and $\azero = 500\, e^{i \pi/4} \unit{GeV}$.  The
-lines show the running in \fstwo and the dots the
-running in \spheno 4.0.2.  We find very good agreement of
-the running of $(T^u)_{33}$ between the two programs over the shown
-$14$ orders of magnitude. The maximum deviation between \fs and
-\spheno is around $2\%$ for $Q \approx 10^{16} \unit{GeV}$ and is
-mainly caused by a different determination of the
-dimensionless MSSM parameters at the electroweak scale.
-
-\section{\fmw extension}
-\label{sec:fmw}
-\fmw is a major new feature which is released in \fstwo.
-It allows a more accurate determination of the electroweak gauge
-couplings $g_1$ and $g_2$, the prediction of the $W$ pole mass and thus
-enables direct comparisons with electroweak precision data.
-
-The running electroweak gauge couplings are related to the running
-electromagnetic coupling~$e$, the weak mixing angle~$\theta_W$ and the
-normalization factors~$N$ by
-%
-\begin{align}
-  g_1 &= N_{g_Y}\,\frac{e}{\cos\theta_W}\,,
-  & g_2 &= N_{g_L}\,\frac{e}{\sin\theta_W} \,.
-\end{align}
-%
-The electromagnetic gauge coupling can be obtained from the known
-value of $\aem^{\SM(5)}(M_Z)$ as described in
-\secref{sec:fbsm} using the known general form of the threshold
-correction, Eq.\ \eqref{eq:alpha_em_general_threshold}.
-The weak mixing angle, together with the $W$ and $Z$ pole masses and
-the muon decay constant $G_F$, form a set of four electroweak
-precision quantities. Within the theory, only two of them are
-independent. Hence, the running weak
-mixing angle can be calculated in different ways:
-%
-\begin{enumerate}
-\item Using the $W$ and $Z$ pole masses $M_W$ and $M_Z$ as input and
-  calculating the cosine of the running weak mixing angle for example
-  as:
-  \begin{align}
-    \cos^2\theta_W = \frac{m_W^2}{m_Z^2} \,,
-    \label{eq:cosW_from_MWMZ}
-  \end{align}
-  where $m_W$ and $m_Z$ are the running $W$ and $Z$ masses which are
-  obtained from the corresponding pole masses by
-  \begin{align}
-    m_V^2(Q) = M_V^2 + \re \Sigma_{V,T}(p^2 = M_V^2, Q) \,, \qquad (V=W,Z)
-  \end{align}
-  with $\Sigma_{V,T}$ being the transverse part of the vector boson
-  self energy. In models with Higgs triplets having the
-  vacuum expectation value $v_T$ for example,
-  Eq.~\eqref{eq:cosW_from_MWMZ} is adapted accordingly:
-  \begin{align}
-    \cos^2\theta_W = \frac{m_W^2 - g_2^2 v_T^2}{m_Z^2} \,.
-  \end{align}
-  This approach was implemented in the first release of \fs \cite{Athron:2014yba},
-  as a generalization of the calculation in Ref.\ \cite{Athron:2012pw}.
-\item Using the $Z$ pole mass and the measured muon decay constant
-  $G_F$ as input and determining $\theta_W$ and the $W$ pole mass
-  as described below.
-\end{enumerate}
-%
-The first approach has the advantage that it can be easily applied to
-any \BSM model with a $W$ and $Z$ boson, because the 1-loop calculation of
-the running $W$ and $Z$ masses can be fully automatized since only
-self energies are necessary.  However, this approach has the
-disadvantage that the parametric uncertainty of the calculated
-electroweak gauge couplings is then limited by that
-of the measured $W$ pole mass of the order $0.02\%$ \cite{Olive:2016xmw}.
-This prohibits a meaningful computation of other electroweak precision
-observables for which more precise experimental data exist.
-
-The second approach is more complicated to automatize for all \BSM
-models, because also 1-loop vertex and box diagrams contributing to
-the muon decay have to be taken into account.  However, the approach
-has the advantage that the parametric uncertainty of the calculated
-electroweak gauge couplings is related to those
-of $G_F$ and $M_Z$, which are of the order $0.00005\%$
-and $0.002\%$, respectively \cite{Olive:2016xmw}. As a result, the $W$
-pole mass is now a meaningful prediction, which can be used to
-constrain BSM models.
-
-Before version 2.0, only the first approach could be used in \fs for
-all \BSM models.  The second approach was only available in models
-which are SM-like or (N)MSSM-like, an option introduced in \fs 1.1.0 (as described in Ref.\ \cite{Staub:2015aea}).
-\fstwo is now able to apply the second approach to all \BSM models
-which have a $W$ and a $Z$ boson and which have the SM gauge
-group as a gauge group factor.
-
-For the implementation of the second approach, \fs uses a
-generalization of the procedure presented in Ref.~\cite{Degrassi:1990tu}
-for the SM, which has been adapted to the MSSM in Ref.~\cite{Pierce:1996zz}.
-The running weak mixing angle is extracted from the relation
-\cite{Degrassi:1990tu}
-%
-\begin{align}
-  \sin^2\theta_W \cos^2\theta_W =
-  \frac{\pi\,\aem}
-   {\sqrt{2}\,M_Z^2\,G_F\,\hat\rho_\text{tree} \left(1-\Delta\hat{r}\right)}\,
-  \label{eq:muon_decay_master}
-\end{align}
-%
-with the renormalization scale consistently being set to $M_Z$,
-where $\aem$ is the electromagnetic coupling of the \BSM model in the
-\MSbar/\DRbar scheme and
-%
-\begin{align}
-  \Delta\hat{r} &= \Delta\hat{r}_\text{1L} + \Delta\hat{r}_\text{2L}^\SM \\
-  &\text{with}\;\;
-  \Delta\hat{r}_\text{1L} = \frac{1}{1-\Delta\hat\rho}\,\frac{\re\Sigma_{W,T}(0)}{M_W^2}
-  - \frac{\re\Sigma_{Z,T}(M_Z^2)}{M_Z^2} + \deltaVB\,,
-  \label{eq:Delta_r_hat}\\
-  \Delta\hat\rho &= \frac{1}{1 + \frac{\re\Sigma_{Z,T}(M_Z^2)}{M_Z^2}}
-  \left[\frac{\re\Sigma_{Z,T}(M_Z^2)}{M_Z^2} - \frac{\re\Sigma_{W,T}(M_W^2)}{M_W^2}
-   + \Delta\hat\rho_\text{2L}^\SM\right].
-   \label{eq:Delta_rho_hat}
-\end{align}
-%
-In the occurring self energies $\Sigma$, the top quark mass is chosen
-to be the pole mass $M_t$ in order to include partial 2-loop corrections not
-contained in $\Delta\hat{r}_\text{2L}^\SM$ and $\Delta\hat\rho_\text{2L}^\SM$
-\cite{Fanchiotti:1992tu}.
-Since $\Delta\hat{r}$ and $\Delta\hat\rho$ themselves depend
-on $\theta_W$, an iteration including these equations has to be
-performed to get a self-consistent solution.
-
-The quantity
-%
-\begin{align}
-  \hat\rho_\text{tree} &= \rho_0\,\frac{m_{Z,\SM}^2}{m_{Z,\text{mix}}^2}
-  \label{eq:rho_hat_tree}
-\end{align}
-%
-introduces two different generalizations. On the one hand, corrections
-from higher dimensional Higgs multiplets are included via \cite{Langacker:1991an}
-%
-\begin{align}
-  \rho_0 &= \frac{\sum_i \left(t_i^2 - t_{3i}^2 + t_i \right)|v_{\varphi_i}|^2}
-   {\sum_i 2\,t_{3i}^2\,|v_{\varphi_i}|^2}\,,
-\end{align}
-%
-where the sums run over all neutral Higgs fields~$\varphi_i$ with vacuum
-expectation value~$v_{\varphi_i}$, weak isospin~$t_i$ and its third
-component~$t_{3i}$. On the other hand, corrections from extra $U(1)$
-gauge groups are included via the ratio of the SM-like tree-level $Z$
-boson mass $m_{Z,\SM}$ and the tree-level $Z$ boson mass $m_{Z,\text{mix}}$
-including mixing with additional $Z^\prime$ bosons
-\cite{Degrassi:1989mu,Leike:1991if}.
-
-The leading SM 2-loop contributions to $\Delta\hat{r}$ and
-$\Delta\hat\rho$ \cite{Fanchiotti:1992tu,Pierce:1996zz} are given
-by\footnote{The specific numerical values in these formulas depend
-on the renormalization scale, which is assumed to be $M_Z$, and on
-the choice of the top pole mass in the self energies from
-Eqs.~\eqref{eq:Delta_r_hat} and \eqref{eq:Delta_rho_hat}.}
-%
-\begin{align}
- \begin{split}
-  \Delta\hat{r}_\text{2L}^\SM &= \frac{\aem \,\as^{\SM(5)}}
-   {4 \pi^2 \sin^2\theta_W \cos^2\theta_W}
-  \left[2.145\,\frac{M_t^2}{M_Z^2} + 0.575 \log\mathopen{}\left(\frac{M_t}{M_Z}
-  \right)\mathclose{} - 0.224 - 0.144\,\frac{M_Z^2}{M_t^2}\right] \\
-  & \phantom{={}} - \delta_\text{Higgs} \,\frac{1 - \Delta\hat{r}_\text{1L}}
-   {1 - \Delta\hat\rho}\,,
-   \label{eq:Delta_r_hat_SM_2L}
- \end{split} \\[1em]
-  \Delta\hat\rho_\text{2L}^\SM &= \frac{\aem \,\as^{\SM(5)}}
-   {4 \pi^2 \sin^2\theta_W}
-  \left[-2.145\,\frac{M_t^2}{M_W^2} + 1.262 \log\mathopen{}\left(\frac{M_t}{M_Z}
-  \right)\mathclose{} - 2.24 - 0.85\,\frac{M_Z^2}{M_t^2}\right] + \delta_\text{Higgs}\,,
-  \label{eq:Delta_rho_hat_SM_2L}
-\end{align}
-%
-where the SM strong coupling $\as^{\SM(5)}$ is taken at the scale $M_t$ and
-we have generalized the Higgs dependent part as
-%
-\begin{align}
-   \delta_\text{Higgs} &= 3 \left(\frac{G_F \,M_t \,v_\SM}{8 \pi^2 \sqrt{2}}\right)^2
-   \,\sum_{i}\left(|a_{\varphi_i tt}|^2 - |b_{\varphi_i tt}|^2\right)
-   \rho^{(2)}\mathopen{}\left(\frac{m_{\varphi_i}}{M_t}\right)\mathclose{}
-\end{align}
-%
-to include corrections from all neutral Higgs fields $\varphi_i$
-coupling to the top quark via the vertex
-$(\mathbbm{1}\,a_{\varphi_i tt} + \gamma_5\,b_{\varphi_i tt})$.
-For this generalization to work properly, the SM-like vacuum
-expectation value $v_\SM$ has to be defined and normalized
-to the value $\approx 246$ GeV in the corresponding \sarah model file.
-The utilized expansions of the function $\rho^{(2)}$ can be found in
-Ref.~\cite{Fleischer:1993ub}.
-
-The model-specific correction in Eq.~\eqref{eq:Delta_r_hat}
-consists of an SM and a BSM part,
-%
-\begin{align}
-   \deltaVB &= \deltaVB^\SM + \deltaVB^\BSM\,,
-\end{align}
-%
-where the SM contribution $\deltaVB^\SM$
-\cite{Degrassi:1990tu}, originating
-from diagrams with additional internal gauge bosons, is given by
-Eq.~(C.12) from Ref.~\cite{Pierce:1996zz} with the replacement
-$\hat\rho \rightarrow 1/(1-\Delta\hat\rho)$.\footnote{The expression
-  for $\deltaVB^\SM$ implemented in \fs has been extended by a term
-  expressing the scale dependence so that the calculation can be
-  performed consistently at the 1-loop level at scales different from
-  $M_Z$.  However, the 2-loop contributions
-  $\Delta\hat{r}_\text{2L}^\SM$ and $\Delta\hat\rho_\text{2L}^\SM$ are
-  omitted if $Q \neq M_Z$, because their explicit scale dependence is
-  currently not taken into account.}
-The BSM contribution $\deltaVB^\BSM$ contains corrections from
-1-loop external wave-function renormalizations, vertex and
-box diagrams, which are put together as in Eq.~(C.13) from the
-same reference. For the different corrections, \fs considers the
-diagram types shown in \figref{fig:delta_VB_diagram_types} and sums
-over all possible instantiations of these by inserting the valid
-combinations of particles into the loop.\footnote{ Note that
-self energy, vertex and box diagrams with internal vector bosons
-are not considered outside of $\deltaVB^\SM$.}
-%
-\begin{figure}[tbh]
-  \centering
-  \subfloat[]{
-  \quad\includegraphics{doc/Feynman/deltaVBWave}\quad
-  \label{fig:delta_VB_wave_diagrams}
-  }
-  \subfloat[]{
-  \qquad\includegraphics{doc/Feynman/deltaVBVertex1}\quad
-  \includegraphics{doc/Feynman/deltaVBVertex2}\qquad
-  \label{fig:delta_VB_vertex_diagrams}
-  }
-  \subfloat[]{
-  \quad\includegraphics{doc/Feynman/deltaVBBox1}\quad
-  \label{fig:delta_VB_box_diagrams}
-  }
-  \caption{Generic diagram types implemented in \fs to calculate
-  the BSM 1-loop contributions to the muon decay, $\deltaVB^\BSM$.
-  The solid and dashed lines in the loops represent any valid
-  combination of fermions and scalars.
-  \protect\subref{fig:delta_VB_wave_diagrams} wave-function
-  renormalization diagrams: $f$ stands for $e$, $\mu$, $\nu_e$
-  or~$\nu_\mu$.
-  \protect\subref{fig:delta_VB_vertex_diagrams} vertex diagrams:
-  $e_i$ stands for $e$ or~$\mu$.
-  \protect\subref{fig:delta_VB_box_diagrams} box diagrams:
-  the triplet $(f_1, f_2, f_3)$ stands for any permutation of
-  $e$, $\nu_e$ and $\nu_\mu$.}
-  \label{fig:delta_VB_diagram_types}
-\end{figure}
-%
-
-After the weak mixing angle~$\theta_W$ has been calculated via
-Eq.~\eqref{eq:muon_decay_master}, the $W$ pole mass can be
-computed as
-%
-\begin{align}
-   M_W = \sqrt{M_Z^2\,\cos^2\theta_W\,\frac{\hat\rho_\text{tree}}{1-\Delta\hat\rho}}\,.
-   \label{eq:MW_calculation}
-\end{align}
-%
-
-\subsection{Choosing \fmw in the model file}
-%
-The method to determine the weak mixing angle can be selected by
-setting the following variable in the \fs model file:
-%
-\begin{lstlisting}
-(* possible values: Automatic, FSFermiConstant or FSMassW *)
-FSWeakMixingAngleInput = Automatic;
-\end{lstlisting}
-%
-By default, \code{FSWeakMixingAngleInput} is set to \code{Automatic},
-in which case the weak mixing angle is determined from the muon decay,
-if all conditions are fulfilled, otherwise the $W$ mass is used.
-Further possible values are \code{FSFermiConstant} and
-\code{FSMassW} to explicitly select the muon decay or the $W$ mass method,
-respectively.
-
-\subsection{Structure of \fmw code}
-%
-The C++ interface for the determination of the running weak mixing
-angle~$\theta_W$ and the $W$ pole mass via muon decay is provided by the class
-\code{<model>_weinberg_angle} defined in
-\code{<model>_weinberg_angle.hpp}.  To construct an object from this
-class, two arguments have to be provided: a model object, which
-represents the set of running BSM parameters, and a struct of type
-\code{Sm_parameters}. The latter is defined within the
-\code{<model>_weinberg_angle} class and contains the required SM parameters,
-namely the Fermi constant $G_F$ as well as the pole
-masses $M_W$, $M_Z$, $M_t$ and $\as^{\SM(5)}(M_t)$.
-The values of $\sin\theta_W$ and $M_W$ are calculated and returned by
-the class member function \code{calculate()}, which includes all of the
-Eqs.~\eqref{eq:muon_decay_master}--\eqref{eq:MW_calculation} and the
-necessary iteration.\footnote{The input $W$ pole mass is used as an
-initial value while the function \code{calculate()} returns a more fitting
-one. By updating the utilized value of $M_W$ during the iteration of the
-spectrum generator, a self-consistent solution is ensured.}
-%
-\begin{example}
-  In the \modelname{MSSM}, the running weak mixing angle~$\theta_W$ and the
-  $W$ pole mass for a given \code{model} object can be calculated with
-  the following code:
-%
-\begin{lstlisting}
-MSSM_weinberg_angle::Sm_parameters sm_pars;
-sm_pars.fermi_constant = 1.1663787e-05;
-sm_pars.mw_pole = 80.385;
-sm_pars.mz_pole = 91.1876;
-sm_pars.mt_pole = 173.34;
-sm_pars.alpha_s = 0.1079;
-
-MSSM_weinberg_angle weinberg(model, sm_pars);
-
-const auto sw_mw     = weinberg.calculate();
-const double theta_w = std::asin(sw_mw.first);
-const double MW      = sw_mw.second;
-\end{lstlisting}
-\end{example}
-%
-
-\subsection{Tests and comparisons with other spectrum generators}
-%
-The implementation of the muon decay method for the determination
-of $\theta_W$ and $M_W$ provided as \fmw has been tested in the SM
-and CMSSM by comparing to the results obtained using the algorithm
-from \softsusy, which has been added in \fs 1.1.0. We have found
-excellent agreement between the two implementations and also added
-unit tests performing these comparisons to the \fs test suite.
-
-Furthermore, the automatically generated expression for the
-generalized $\hat\rho_\text{tree}$ given in Eq.~\eqref{eq:rho_hat_tree}
-has been analytically checked for many models, such as the UMSSM,
-MRSSM and \ESSM.
-
-Finally, we compared numerical results for the $W$ pole mass in the
-CMSSM and MRSSM obtained with \fmw to the ones from \spheno code
-generated by \sarah~4.12.2. \figref{fig:FlexibleMW_vs_SPheno} shows $M_W$
-as a function of $\mhalf$ (CMSSM, see left panel) or the superpotential
-parameter $\Lambda_u$ (MRSSM, see right panel) while all
-the other parameters are specified as described in the caption.
-In the case of the CMSSM, additionally the results calculated with
-\spheno~4.0.3 are plotted. For both models, there is a large
-discrepancy between the \fmw values illustrated by the blue solid
-line and the results from \sarah/\spheno presented by the red dashed
-line. A thorough comparison of the two implementations has revealed
-two major differences. On the one hand, \sarah/\spheno partly uses the
-\DRbar top mass in the self energies occurring in
-Eqs.~\eqref{eq:Delta_r_hat} and \eqref{eq:Delta_rho_hat} as well as
-the SM 2-loop corrections Eqs.~\eqref{eq:Delta_r_hat_SM_2L} and
-\eqref{eq:Delta_rho_hat_SM_2L} while \fmw always uses the top pole mass,
-as suggested by Ref.~\cite{Fanchiotti:1992tu}. Not utilizing the top
-pole mass in all of these formulas spoils the correctness of the
-included SM 2-loop corrections. On the other hand, the
-\sarah/\spheno code contains an inconsistency in the final computation of
-$M_W$, which is similar to Eq.~\eqref{eq:MW_calculation} but partly
-neglects the SM 2-loop correction to $\Delta\hat\rho$.
-This inconsistency is also existent in the original \spheno code that,
-for the CMSSM, produces the results depicted by the red dashed-double-dotted
-line in the left panel of \figref{fig:FlexibleMW_vs_SPheno}.
-After fixing these issues within the \sarah/\spheno and \spheno code,
-we get the modified results illustrated by the green dotted and
-dashed-dotted line, respectively. These show good agreement with the
-values from \fmw for both the CMSSM and MRSSM\@.
-The remaining small discrepancies are well
-understood and mainly caused by minor differences in the implemented
-formulas and the various utilized \DRbar parameters. In addition,
-the \sarah/\spheno and \spheno codes use $\as^{\SM(5)}(M_Z)$ while \fmw uses
-$\as^{\SM(5)}(M_t)$ as preferred by Ref.~\cite{Fanchiotti:1992tu}.
-%
-\begin{figure}[tbh]
-  \centering
-  \includegraphics[width=0.49\textwidth]{doc/plots/FlexibleMW/scan_MW_M12}\hfill
-  \includegraphics[width=0.49\textwidth]{doc/plots/FlexibleMW/scan_MW_LamTU}
-  \caption{Left panel: $W$ pole mass in the CMSSM as a function of
-  $\mhalf$ with $\mzero = 1\unit{TeV}$, $\tan\beta = 10$,
-  $\sign\mu = +1$ and $\azero = 0$. Right panel: $W$ pole mass in
-  the MRSSM as a function of $\Lambda_u$ with the other parameters
-  fixed as for BMP1 in Table~2 from Ref.~\cite{Diessner:2014ksa}.
-  For both panels, the differences between the lines are explained
-  in the text.}
-  \label{fig:FlexibleMW_vs_SPheno}
-\end{figure}
-%
-
-\section{\fsas extension}
-\label{sec:fsas}
-In this section, we introduce a new BVP
-solver, a major new feature released in \fstwo. The specification of the
-running parameters of a BSM model at multiple scales connected by RGEs
-constitutes a BVP that must be solved in order to compute the mass spectrum.
-\fs was designed with the intention of allowing multiple solvers for this
-problem. In \fsone only one supported solver was
-distributed, the two-scale solver, which uses a fixed point iteration
-with boundary conditions set at both the high scale and the low scale
-and is the one used in MSSM spectrum generators. However, additional
-solvers are important because convergence of valid points is not
-guaranteed with the two-scale solver and viable regions of the
-parameter space can be missed if the only option is the two-scale
-solver.  This is already true in the CMSSM where it has also been
-shown that there can even be multiple solutions to the BVP
-\cite{Allanach:2013cda,Allanach:2013yua}, of which the two-scale solver will
-only find at most one.  Furthermore, the two-scale
-solver cannot find solutions at all for a large class of models, which
-includes the fully constrained NMSSM (CNMSSM) \cite{Ellwanger:1993xa,
-  Ellwanger:1995ru,Ellwanger:1996gw} and constrained \ESSM
-(C\ESSM) \cite{Athron:2009bs}.
-
-A basic problem in these models is that parameters exist that would
-naturally be computed at the weak scale by EWSB conditions but
-are at the same time constrained by GUT scale conditions.\footnote{In
-  principle this could be avoided, e.g., in the CNMSSM and C\ESSM by including
-  $\tan \beta$ or $\lambda$ in the set of parameters fixed by minimization
-  conditions.  However, $\tan\beta$ affects the top, bottom and $\tau$ Yukawa
-  couplings at tree level, and $\lambda$ appears in the 1-loop RGEs for these
-  couplings and therefore can have a significant impact on the RG evolution.
-  This makes it difficult to obtain a convergent fixed point iteration with
-  such a setup.} In the constrained NMSSM, where the soft scalar mass,
-$m_S$, is typically an EWSB output, solutions have previously been
-found by varying one of the input parameters, $\tan \beta$, and tuning
-this until a solution is found where $m_S$ is sufficiently close to
-the universal scalar mass, $\mzero$ \cite{Djouadi:2008yj,Djouadi:2008uj}.
-For the constrained \ESSM again the soft scalar masses need to be EWSB
-outputs.  In that case a new BVP solver using semi-analytic solutions
-was invented to find solutions \cite{Athron:2009ue,Athron:2009bs} and
-extended and improved in further studies
-\cite{Athron:2011wu,Athron:2012sq,Athron:2012pw}, though the precision
-was limited due to analytical approximations used in the code.  In
-general, the challenges in finding solutions to the mass spectrum in
-BSM models can be quite different to those found in the MSSM and
-therefore additional solvers can be of great benefit in increasing the
-scope of \fs.
-
-In \fstwo, an additional BVP solver has been added that solves the BVP
-by using semi-analytic solutions to the RGEs for a subset of the
-running parameters in the model.  The semi-analytic algorithm takes
-advantage of the hierarchical structure of the RGEs in any model: the parameters
-can be split into a sequence of sets such that the running of the parameters
-in each set is independent of all of the parameters in the following
-sets.  For example, in SUSY models the SUSY preserving parameters run
-independently of the soft breaking parameters; in general renormalizable
-models the dimensionless parameters run independently of the mass
-parameters.\footnote{This partitioning is already reflected to some extent in
-  the original C++ class structure for the parameters and
-RGEs, which for SUSY (non-SUSY) models is split up so that
-the RGEs for the SUSY preserving (dimensionless) parameters can be integrated
-separately to those for the soft (dimensionful) parameters; see
-\secref{sec:fbsm}.}
-It is clear from the general form of the RGEs
-\cite{Machacek:1983tz,Machacek:1983fi,Machacek:1984zw,Martin:1993zk,
-  Yamada:1994id,Jack:1997eh,Luo:2002ti,Fonseca:2011vn,Goodsell:2012fm,
-  Fonseca:2013bua,Sperling:2013eva,Sperling:2013xqa} that further
-divisions are possible on the basis of the mass dimension of the
-parameters.  The simplest case is the evolution of the mass dimension
-one parameters $m^i(Q)$, which is described by a system of linear homogeneous
-differential equations that can be solved in the form
-\begin{align}
-m^i(Q) &= \scoeff{}{}^i_j m^j(Q_0) ,
-\end{align}
-i.e., in terms of a linear combination of initial values and a set of
-dimensionless coefficients that only depend on dimensionless parameters.
-The RGEs for higher mass dimension parameters can then be cast as linear,
-non-homogeneous systems by substituting in the solutions for the lower
-mass dimension parameters, for which the general solutions can easily
-be written down.\footnote{See, for example, the derivation given in
-  Ref.~\cite{Athron:2009bs} for the case of a SUSY model with real
-  soft parameters.}  In the CMSSM, doing so leads to well-known expressions
-for the soft masses in terms of the universal parameters $\mzero$, $\mhalf$,
-and $\azero$.  For instance, the solutions for the soft scalar masses read
-(see also Eq.~\eqref{eq:cmssm-semi-analytic-solns} below)
-\begin{equation*}
-  m_i^2(Q) = a_i(Q) \mzero^2 + b_i(Q) \mhalf^2
-  + c_i(Q) \mhalf \azero + d_i(Q) \azero^2 \, ,
-\end{equation*}
-where the dimensionless coefficients $a_i$, $b_i$, $c_i$, and $d_i$ are
-determined by the running of the gauge and Yukawa couplings, and are computed
-numerically.  The basic idea of the semi-analytic solver is to generalize this
-approach to other models: first, the analytic forms of the solutions are
-determined from the boundary conditions, and then the appearing coefficients
-are determined numerically.  Thus the solver obtains a collection of
-semi-analytic solutions which express parameters at the scale $Q$ directly in
-terms of parameters at the boundary scale $Q_0$.
-
-To make the discussion concrete, consider a general SUSY model with bilinear
-and linear superpotential parameters $\mu^{ij}$ and $L^i$, together with a set
-of soft breaking gaugino masses $M_a$, scalar trilinear, bilinear and linear
-couplings $T^{ijk}$, $b^{ij}$ and $t^i$, and a set of soft scalar masses
-$(m^2)_i^j$.  The semi-analytic solutions are obtained from the general 2-loop
-RGEs given in Refs.~\cite{Martin:1993zk,Yamada:1994id,Fonseca:2011vn}.
-Since the superpotential parameters are only multiplicatively
-renormalized, the semi-analytic solutions for the parameters $\mu^{ij}$
-and $L^i$ are particularly simple, taking the form
-\begin{align}
-  \mu^{ij}(Q) &= \scoeff{\mu}{\mu}^{ij}_{kp} \mu^{kp}(Q_0) \, ,
-  \label{eq:general-susy-bilinear-solution} \\
-  L^i(Q) &= \scoeff{L}{L}^i_p L^p(Q_0) \, ,
-  \label{eq:general-susy-linear-solution}
-\end{align}
-where the coefficients $\scoeff{\mu}{\mu}$ and $\scoeff{L}{L}$ satisfy the
-initial conditions $[c^\mu_\mu(Q_0)]^{ij}_{kp} = \delta^i_k
-\delta^j_p$, $[c^L_L(Q_0)]^i_p = \delta^i_p$. Again, the point is that
-the coefficients $c(Q)$ can be determined once the
-running dimensionless SUSY parameters are known, and the equations then express
-parameters at $Q$ in  terms of parameters at $Q_0$ with numerically known
-coefficients.
-
-The semi-analytic solutions for the dimension one soft breaking parameters can
-be written as
-\begin{align}
-  T^{ijk}(Q) &= \scoeff{T}{T}^{ijk}_{lmn} T^{lmn}(Q_0)
-  + \scoeff{T}{M}^{ijkb} M_b(Q_0) \, ,
-  \label{eq:general-soft-trilinear-solution} \\
-  M_a(Q) &= \scoeff{M}{T}_{almn} T^{lmn}(Q_0)
-  + \scoeff{M}{M}_a^b M_b(Q_0) \, .
-  \label{eq:general-gaugino-mass-solution}
-\end{align}
-Upon substituting these solutions into the RGEs for the soft breaking bilinears
-$b^{ij}$ and the soft scalar masses $(m^2)_i^j$, one finds the semi-analytic
-solutions
-\begin{equation} \label{eq:general-soft-bilinear-solution}
-  b^{ij}(Q) = \scoeff{b}{b}^{ij}_{kl} b^{kl}(Q_0)
-  + \scoeff{b}{\mu T}^{ij}_{abmno} \mu^{ab}(Q_0) T^{mno}(Q_0)
-  + \scoeff{b}{\mu M}^{ijc}_{ab} \mu^{ab}(Q_0) M_c(Q_0)
-\end{equation}
-and
-\begin{align}
-  (m^2)_i^j(Q) &= \scoeff{m^2}{m^2}^{jl}_{ik} (m^2)^k_l(Q_0)
-  + \scoeff{m^2}{MM^*}_i^{jab} M_a(Q_0) M_b^*(Q_0) \nonumber \\
-  & \quad {} + \scoeff{m^2}{T M^*}_{iklm}^{ja} T^{klm}(Q_0) M_a^*(Q_0)
-  + \scoeff{m^2}{T^* M}_{iklm}^{ja} T^{klm*}(Q_0)
-  M_a(Q_0) \nonumber \\
-  & \quad {} + \scoeff{m^2}{T^* T}_{iklmnop}^j T^{klm*}(Q_0) T^{nop}(Q_0) \, .
-  \label{eq:general-soft-scalar-mass-solution}
-\end{align}
-Finally, the semi-analytic solutions for the soft breaking linear parameters
-$t^i$ are given by
-\begin{align}
-  t^i(Q) &= \scoeff{t}{t}^i_j t^j(Q_0) + \scoeff{t}{L T}^i_{jklm} L^j(Q_0)
-  T^{klm}(Q_0) + \scoeff{t}{LM}^{ia}_j L^j(Q_0) M_a(Q_0) \nonumber \\
-  & \quad {} + \scoeff{t}{\mu \mu T}^i_{jklmnop} \mu^{jk}(Q_0) \mu^{lm}(Q_0)
-  T^{nop}(Q_0) + \scoeff{t}{\mu \mu M}^{ia}_{jklm} \mu^{jk}(Q_0) \mu^{lm}(Q_0)
-  M_a(Q_0)
-  \nonumber \\
-  & \quad {} + \scoeff{t}{\mu b}^i_{jklm} \mu^{jk}(Q_0) b^{lm}(Q_0)
-  + \scoeff{t}{\mu^* M M^*}^{iab}_{jk} \mu^{jk*}(Q_0) M_a(Q_0) M_b^*(Q_0)
-  \nonumber \\
-  & \quad {} + \scoeff{t}{\mu^* T T^*}^i_{jklmnopq} \mu^{jk*}(Q_0) T^{lmn}(Q_0)
-  T^{opq*}(Q_0) \nonumber \\
-  & \quad {} + \scoeff{t}{\mu^* T M^*}^{ia}_{jklmn} \mu^{jk*}(Q_0) T^{lmn}(Q_0)
-  M_a^*(Q_0) \nonumber \\
-  & \quad {} + \scoeff{t}{\mu^* T^* M}^{ia}_{jklmn} \mu^{jk*}(Q_0) T^{lmn*}(Q_0)
-  M_a(Q_0) \nonumber \\
-  & \quad {} + \scoeff{t}{b^* T}^i_{jklmn} b^{jk*}(Q_0) T^{lmn}(Q_0)
-  + \scoeff{t}{b^* M}^{ia}_{jk} b^{jk*}(Q_0) M_a(Q_0) \nonumber \\
-  & \quad {} + \scoeff{t}{\mu^* m^2}^{il}_{jkm} \mu^{jk*}(Q_0) (m^2)^m_l(Q_0)
-  \, . \label{eq:general-soft-linear-solution}
-\end{align}
-For models in which Dirac gaugino masses $m_{Da}^i$ are also present, the
-solutions for the parameters $t^i$ are modified, $t^i \to t^i + \Delta t^i$,
-where $\Delta t^i$ is of the form
-\begin{align}
-  \Delta t^i(Q) &= \scoeff{t}{m_D m^2}^{iaj}_{kl} m_{Da}^k(Q_0) (m^2)_j^l(Q_0)
-  + \scoeff{t}{m_D M M^*}^{iabc}_j m_{Da}^j(Q_0) M_b(Q_0) M_c^*(Q_0)
-  \nonumber \\
-  & \quad {} + \scoeff{t}{m_D T M^*}^{iab}_{jklm} m_{Da}^j(Q_0) T^{klm}(Q_0)
-  M_b^*(Q_0) \nonumber \\
-  & \quad {} + \scoeff{t}{m_D T^* M}^{iab}_{jklm} m_{Da}^j(Q_0) T^{klm*}(Q_0)
-  M_b(Q_0) \nonumber \\
-  & \quad {} + \scoeff{t}{m_D T^* T}^{ia}_{jklmnop} m_{Da}^j(Q_0) T^{klm*}(Q_0)
-  T^{nop}(Q_0) \nonumber \\
-  & \quad {} + \scoeff{t}{m_D^* m_D^* T}^{iab}_{jklmn} m_{Da}^{j*}(Q_0)
-  m_{Db}^{k*}(Q_0) T^{lmn}(Q_0) \nonumber \\
-  & \quad {} + \scoeff{t}{m_D^* m_D^* M}^{iabc}_{jk} m_{Da}^{j*}(Q_0)
-  m_{Db}^{k*}(Q_0) M_c(Q_0) \nonumber \\
-  & \quad {} + \scoeff{t}{m_D m_D \mu}^{iab}_{jklm} m_{Da}^j(Q_0) m_{Db}^k(Q_0)
-  \mu^{lm}(Q_0) \, . \label{eq:extra-soft-trilinear-solution}
-\end{align}
-The semi-analytic solutions for the Dirac gaugino masses themselves follow
-from the known general 2-loop RGEs \cite{Goodsell:2012fm}, and can be
-written in the form
-\begin{equation} \label{eq:general-dirac-gaugino-solutions}
-  m_{Da}^i(Q) = \scoeff{m_{Da}}{m_{Da}}^i_j m^j_{Da}(Q_0) \, .
-\end{equation}
-
-In non-SUSY models, the semi-analytic solutions follow from the known
-results for the 2-loop RGEs in a general gauge theory \cite{Machacek:1983tz,
-  Machacek:1983fi,Machacek:1984zw,Luo:2002ti,Fonseca:2013bua}.  For a non-SUSY
-model containing a set of real scalar trilinear couplings $h^{ijk}$ and squared
-scalar masses $(m^2)^{ij}$, and a set of fermion masses $(M_f)^{ij}$, the
-semi-analytic solutions for the mass dimension one parameters read
-\begin{align}
-  h^{ijk}(Q) &= \scoeff{h}{h}^{ijk}_{lmn} h^{lmn}(Q_0)
-  + \scoeff{h}{M_f}^{ijk}_{lm} (M_f)^{lm}(Q_0)
-  + \scoeff{h}{M_f^*}^{ijk}_{lm} (M_f)^{lm*}(Q_0) \, ,
-  \label{eq:general-dimension-one-trilinear-solution} \\
-  (M_f)^{ij}(Q) &= \scoeff{M_f}{h}^{ij}_{lmn} h^{lmn}(Q_0)
-  + \scoeff{M_f}{M_f}^{ij}_{lm} (M_f)^{lm}(Q_0)
-  + \scoeff{M_f}{M_f^*}^{ij}_{lm} (M_f)^{lm*}(Q_0) \, .
-  \label{eq:general-dimension-one-fermion-mass-solution}
-\end{align}
-In general, all of the dimension one parameters must be considered together,
-unlike in SUSY models where they can be separated into superpotential and soft
-breaking masses.  After substituting the solutions
-Eq.~\eqref{eq:general-dimension-one-trilinear-solution} and
-Eq.~\eqref{eq:general-dimension-one-fermion-mass-solution} into the RGEs for
-the squared scalar masses, the semi-analytic solutions for the scalar masses
-are found to be
-\begin{align}
-  (m^2)^{ij}(Q) &= \scoeff{m^2}{m^2}^{ij}_{kl} (m^2)^{kl}(Q_0)
-  + \scoeff{m^2}{h h}^{ij}_{klmnop} h^{klm}(Q_0) h^{nop}(Q_0) \nonumber \\
-  & \quad {} + \scoeff{m^2}{h M_f}^{ij}_{klmpq} h^{klm}(Q_0) (M_f)^{pq}(Q_0)
-  + \scoeff{m^2}{h M_f^*}^{ij}_{klmpq} h^{klm}(Q_0) (M_f)^{pq*}(Q_0)
-  \nonumber \\
-  & \quad {} + \scoeff{m^2}{M_f M_f}^{ij}_{klmn} (M_f)^{kl}(Q_0) (M_f)^{mn}(Q_0)
-  \nonumber \\
-  & \quad {} + \scoeff{m^2}{M_f M_f^*}^{ij}_{klmn} (M_f)^{kl}(Q_0)
-  (M_f)^{mn*}(Q_0) \nonumber \\
-  & \quad {} + \scoeff{m^2}{M_f^* M_f^*}^{ij}_{klmn} (M_f)^{kl*}(Q_0)
-  (M_f)^{mn*}(Q_0) \, .
-  \label{eq:general-scalar-mass-solution}
-\end{align}
-
-The solver algorithm implemented by \fsas automatically determines
-the above semi-analytic solutions in the model,\footnote{In non-SUSY models,
-  \sarah currently does not calculate RGEs for linear scalar couplings
-  $L^i$, which have been given in, e.g., Ref.~\cite{Goodsell:2012fm}, and
-  therefore the semi-analytic solutions for these parameters are also not used
-  in \fs.} given the set of boundary conditions at some scale.  Since the
-required coefficients may be determined knowing only the running of the
-dimensionless or SUSY preserving parameters, each step of the main fixed point
-iteration is split up into two parts.  Firstly, the BVP for the dimensionless
-parameters is solved iteratively.  The semi-analytic coefficients at any scale
-can then be calculated.  In this second stage, the soft breaking or
-dimensionful parameters are expanded in terms of the semi-analytic solutions to
-the RGEs.  The low-energy EWSB conditions and masses are thus expressed
-explicitly in terms of the boundary values at $Q_0$, allowing, for example,
-unknown quantities at one scale to be directly constrained at another.
-
-\subsection{Choosing \fsas in the model file}
-\label{sec:chosing_SAS}
-
-The BVP solver algorithms that are applicable in a given model can be
-specified in the \fs model file using the list
-\code{FSBVPSolvers}.  The elements of this list correspond to
-the desired BVP solvers that should be enabled for the model, identified
-using the predefined symbols \code{TwoScaleSolver} for the two-scale
-algorithm and \code{SemiAnalyticSolver} for the semi-analytic solver.  By
-default, if a list of BVP solvers is not specified in the model file, only
-the two-scale algorithm is enabled, as summarized in \tabref{tab:fsas_options}.
-%
-\begin{OptionTable}{\fsas model file options.}{tab:fsas_options}
-  \code{FSBVPSolvers} & \code{\{ TwoScaleSolver \}} & Non-empty list
-  containing \code{TwoScaleSolver} or \code{SemiAnalyticSolver} or
-  both &
-  List of BVP solvers to be used \\
-  \code{TwoScaleSolver} & -- & -- & Represents the two-scale BVP solver \\
-  \code{SemiAnalyticSolver} & -- & -- & Represents the semi-analytic BVP
-  solver \\
-  \bottomrule
-\end{OptionTable}
-
-In addition to specifying that the semi-analytic solver should be used,
-the user should ensure that the boundary conditions for the model are
-compatible with its use.  Currently, this requires that those parameters
-that will be expanded using semi-analytic solutions of the RGEs
-are fixed in the same boundary condition.  So in SUSY models, the
-boundary values for all of the soft SUSY breaking parameters should be given
-at a single scale, while in non-SUSY models the same should be done for all
-of the dimensionful parameters.  The expressions for the boundary values
-must also only be polynomials in any dimensionful  parameters, such as
-universal scalar masses.
-
-Boundary values for the running parameters of the model can be specified
-in terms of input parameters such as those defined in the \code{MINPAR}
-and \code{EXTPAR} variables.  In \fstwo, the user can also define extra
-parameters by using the list \code{FSAuxiliaryParameterInfo}.   Each entry in
-this list should contain the name of the extra parameter being defined and a
-list of its properties.  The possible properties for the
-new parameters are specified in \tabref{tab:parameter_info}.  In particular,
-the mass dimensions of the new parameters may be specified to allow for
-simplifying the forms of the semi-analytic solutions, in which dimensionless
-parameters can be absorbed into the definitions of the semi-analytic
-coefficients.  Input or auxiliary parameters that are not scalars may be
-defined by setting the \code{ParameterDimensions} property to a list of the
-form \code{\{M,N\}} for an $M \times N$ matrix or \code{\{N\}} for an
-$N$-dimensional vector.  A value of \code{\{1\}} corresponds to a scalar.  Note
-that in versions of \fs prior to version 2.0, this functionality was available
-for input parameters using the variable \code{FSExtraInputParameters}.
-However, please note that this variable has been removed in \fstwo; definitions
-that were previously given in \code{FSExtraInputParameters} must now be given
-in the new list \code{FSAuxiliaryParameterInfo}.
-%
-\begin{OptionTable}{Allowed properties for extra parameters}{tab:parameter_info}
-  \code{InputParameter} & \code{False} & \code{True} or \code{False}
-  & Indicates whether the new parameter is an input parameter \\
-  \code{ParameterDimensions} & \code{\{1\}} & A list of the form \code{\{M,N\}}
-  or \code{\{N\}} & Specifies the dimensions of the parameter \\
-  \code{MassDimension} & -- & A non-negative integer & Specifies the mass
-  dimension of the parameter \\
-  \code{LesHouches} & -- & A symbol or string, or a list of the form
-  \code{\{block, entry\}} & Specifies the SLHA block from which the parameter
-  should be read if it is an input parameter \\
-  \bottomrule
-\end{OptionTable}
-
-\begin{example}
-In the MSSM, input parameters giving the values of the
-soft SUSY breaking trilinears as $3 \times 3$ matrices can be defined:
-%
-\begin{lstlisting}
-FSAuxiliaryParameterInfo = {
-   {Ae  , { LesHouches -> AEIN,
-            ParameterDimensions -> {3,3},
-            InputParameter -> True
-          } },
-   {Ad  , { LesHouches -> ADIN,
-            ParameterDimensions -> {3,3},
-            InputParameter -> True
-          } },
-   {Au  , { LesHouches -> AUIN,
-            ParameterDimensions -> {3,3},
-            InputParameter -> True
-          } }
-};
-\end{lstlisting}
-\end{example}
-
-Extra parameters defined in this way can be used in the specification of the
-boundary conditions for running model parameters, or indeed themselves be
-fixed in the boundary conditions.  A special case is if the new parameters
-are to be fixed by the EWSB conditions.  To facilitate this usage, the new
-variables \code{EWSBInitialGuess} and \code{EWSBSubstitutions} may be
-given in the model file.  The former allows for explicit initial
-guesses to be provided for the EWSB output parameters, in the same format
-as used for specifying the boundary conditions.  The latter, a list of
-two-component lists, can be used to define any substitutions that should
-be made in the EWSB equations before attempting to solve them.
-
-\begin{example}
-In the so-called VCMSSM \cite{Ellis:2003pz,Ellis:2004qe}, the value of the soft
-breaking bilinear $B\mu$ is fixed at the GUT scale, $M_X$, in terms of the
-universal soft trilinear $\azero$ and scalar mass $\mzero$ according to
-$B\mu(M_X) = \mu(M_X) (\mzero + \azero)$.  Consequently, $B\mu$ can no longer
-be fixed to ensure proper EWSB, as is usually done in the CMSSM\@.  Instead,
-$|\mu|^2$ and $\tan\beta$ are used as EWSB outputs.  This can be achieved in a
-\modelname{VCMSSM} model file through the definitions:
-%
-\begin{lstlisting}
-FSAuxiliaryParameterInfo = {
-   {TanBeta, { ParameterDimensions -> {1},
-               MassDimension -> 0 } },
-   {MuSq,    { ParameterDimensions -> {1},
-               MassDimension -> 2 } },
-   {vMSSM,   { ParameterDimensions -> {1},
-               MassDimension -> 1 } }
-};
-
-EWSBOutputParameters = { TanBeta, MuSq };
-
-EWSBSubstitutions = {
-   {vd, vMSSM Cos[ArcTan[TanBeta]]},
-   {vu, vMSSM Sin[ArcTan[TanBeta]]},
-   {\[Mu], Sign[\[Mu]] Sqrt[MuSq]}
-};
-
-EWSBInitialGuess = {
-   {TanBeta, vu / vd},
-   {MuSq, \[Mu]^2}
-};
-
-SUSYScaleInput = {
-   {vMSSM, Sqrt[vd^2 + vu^2]},
-   FSSolveEWSBFor[EWSBOutputParameters]
-};
-\end{lstlisting}
-\end{example}
-
-Note that \fs automatically substitutes the semi-analytic solutions into
-the EWSB conditions for the soft or dimensionful parameters.  Therefore, it
-is not necessary for the user to explicitly provide them.  For the purpose of
-making use of the semi-analytic algorithm, the only required addition to the
-model file is the inclusion of \code{SemiAnalyticSolver} in the list
-\code{FSBVPSolvers}.
-
-\begin{example}
-The CNMSSM is characterized by universal soft scalar masses at the
-high scale.  Typically, this constraint is relaxed somewhat to allow the
-soft singlet mass $m_S^2$ to differ from the common scalar mass $\mzero^2$ at
-the high scale.  This is done to allow fixing $m_S^2$ using the EWSB conditions
-to ensure correct EWSB, and is the approach taken in the \modelname{NMSSM} model
-included with \fs, as well as in other public spectrum generators such as
-\softsusy, \spheno and \nmspec.  Universality of the soft scalar masses can
-alternatively be maintained by using the semi-analytic BVP solver, which is
-achieved in the \modelname{CNMSSM} model file by specifying the BVP solvers to
-be used,
-%
-\begin{lstlisting}
-FSBVPSolvers = { SemiAnalyticSolver };
-\end{lstlisting}
-In this set-up, $\mzero$ ceases to be an input parameter and is fixed by the
-EWSB conditions.  This is achieved in the model file by removing $\mzero$ from
-the list of input parameters and defining it as an additional parameter, as
-follows:
-%
-\begin{lstlisting}
-(* CNMSSM input parameters *)
-
-MINPAR = {
-   {2, m12},
-   {3, TanBeta},
-   {4, Sign[vS]},
-   {5, Azero}
-};
-
-EXTPAR = {
-   {61, LambdaInput}
-};
-
-FSAuxiliaryParameterInfo = {
-   {m0Sq,        { ParameterDimensions -> {1},
-                   MassDimension -> 2 } },
-   {LambdaInput, { ParameterDimensions -> {1},
-                   MassDimension -> 0 } }
-};
-\end{lstlisting}
-Note that the definition of \code{LambdaInput} in
-\code{FSAuxiliaryParameterInfo} is not compulsory, but allows the semi-analytic
-solutions to be simplified by using the fact that it is a dimensionless
-parameter.  The value of $\mzero^2$ is then determined from the EWSB conditions
-by setting
-%
-\begin{lstlisting}
-EWSBOutputParameters = { \[Kappa], vS, m0Sq };
-\end{lstlisting}
-To impose the universality constraint, the condition $m_S^2 = \mzero^2$ at the
-GUT scale must be added to the high-scale boundary condition, by defining
-%
-\begin{lstlisting}
-HighScaleInput={
-   {T[Ye], Azero*Ye},
-   {T[Yd], Azero*Yd},
-   {T[Yu], Azero*Yu},
-   {mq2, UNITMATRIX[3] m0Sq},
-   {ml2, UNITMATRIX[3] m0Sq},
-   {md2, UNITMATRIX[3] m0Sq},
-   {mu2, UNITMATRIX[3] m0Sq},
-   {me2, UNITMATRIX[3] m0Sq},
-   {mHu2, m0Sq},
-   {mHd2, m0Sq},
-   {ms2, m0Sq},
-   {\[Lambda], LambdaInput},
-   {T[\[Kappa]], Azero \[Kappa]},
-   {T[\[Lambda]], Azero LambdaInput},
-   {MassB, m12},
-   {MassWB,m12},
-   {MassG,m12}
-};
-\end{lstlisting}
-The remainder of the model implementation is otherwise rather similar to that
-for the NMSSM solved using the two-scale solver.  The full \modelname{CNMSSM}
-model file is given in \appref{app:CNMSSM}.
-\end{example}
-
-\subsection{Structure of the generated code}
-
-In keeping with the design goal of \fs to produce generated code that is highly
-modular in nature, the implementation of the BVP solvers is separated from
-the details of specific physics models.  In \fs, a general BVP solver algorithm
-is represented by the templated \code{RGFlow<T>} class.  Particular algorithms
-are provided as specializations of this class, with the two-scale and
-semi-analytic solvers corresponding to the classes \code{RGFlow<Two_scale>}
-and \code{RGFlow<Semi_analytic>}, respectively.  Each realizes an abstract
-implementation of the appropriate algorithm, with no dependence on the details
-of any particular model.  The required model-dependent information is provided
-by separate classes representing the model and boundary and matching conditions,
-which are linked to the desired BVP solver class.  New algorithms can easily be
-added simply by writing additional specializations of the \code{RGFlow} class.
-
-The semi-analytic solver algorithm requires two nested iterations.  An
-inner iteration, carried out at each step, determines consistent values for
-the SUSY preserving (in SUSY models) or dimensionless parameters (in non-SUSY
-models) at the low- and high-scale boundaries.  Updated estimates for these
-scales are simultaneously calculated during the iteration if necessary, for
-example if the high scale $M_X$ is defined in the model file by gauge
-unification, $g_1(M_X) = g_2(M_X)$.  Once this has converged, the resulting
-estimate for these parameters is used to compute the semi-analytic solutions
-for the soft SUSY breaking or dimensionful parameters, at which point the
-EWSB conditions may be solved and the \DRbar/\MSbar mass spectrum
-calculated.  The new values of the soft or dimensionful parameters are then used
-in the inner iteration for computing the required threshold corrections.
-This sequence of steps is illustrated in \figref{fig:semi-analytic-algorithm}.
-For a single high-scale model such as the CMSSM, the algorithm proceeds as
-follows:
-%
-\subparagraph{Initial guess:} The initial guess involves a first run of the
-inner iteration.  In all of the steps below, threshold corrections are ignored.
-%
-\begin{enumerate}
-\item The known values of the SM gauge couplings at the scale $M_Z$ are
-  used to estimate the values of $g_1$, $g_2$ and $g_3$ at the scale
-  $M_t$, ignoring threshold corrections.
-\item The user-defined initial guess at the low scale, as given in
-  \code{InitialGuessAtLowScale}, is imposed at the scale $M_t$.
-\item The SUSY preserving or dimensionless parameters are run to the initial
-  guess for $M_X$, given by \code{HighScaleFirstGuess}, and the
-  high-scale boundary condition for these parameters, defined in
-  \code{HighScaleInput}, is imposed.  The initial guess at the high scale,
-  defined in \code{InitialGuessAtHighScale}, is then applied.
-\item \label{initial-guess-low-scale} The model is run to the guess for the
-  low scale, initially set to the value defined in \code{LowScaleFirstGuess},
-  and the low-scale boundary conditions for the SUSY preserving or
-  dimensionless parameters defined in \code{LowScaleInput} are applied.
-\item The model is run to the current guess for $M_X$.
-  \begin{enumerate}
-  \item If necessary, the guess for $M_X$ is updated.  For example, in
-    the CMSSM with $M_X$ defined to be the scale at which $g_1(M_X) = g_2(M_X)$,
-    a new estimate for $M_X$ is calculated according to
-    \begin{equation*}
-      M_X^\prime = M_X \exp\left( \frac{g_2(M_X)-g_1(M_X)}
-      {\beta_{g_1} - \beta_{g_2}} \right ) \, .
-    \end{equation*}
-  \item The high-scale boundary conditions for the SUSY preserving or
-    dimensionless parameters are applied.
-  \end{enumerate}
-\item If not converged, goto \ref{initial-guess-low-scale}.
-\item The model is run to the guess for the low scale.  The semi-analytic
-  solutions are calculated at this scale.
-\item The EWSB equations are solved at tree level.
-\item The \DRbar/\MSbar mass spectrum is calculated.
-\end{enumerate}
-%
-At this stage, initial guesses for all of the model parameters, boundary
-condition scales and the \DRbar/\MSbar mass spectrum are available.  The full
-iteration now starts, in which the full set of threshold corrections are
-applied.
-%
-\subparagraph{Thresholds iteration:}
-%
-\begin{enumerate}
-\item \label{outer-iteration-step-one} The SUSY preserving or dimensionless
-  parameters are determined in an inner iteration analogous to that in the
-  initial guess, namely:
-  \begin{enumerate}
-  \item \label{inner-iteration-step-one} All model parameters are run to the
-    low scale (\code{LowScale}) and the \DRbar/\MSbar mass spectrum is
-    calculated.
-  \item The low scale is recalculated if it is not fixed.
-  \item The SM gauge couplings are calculated in the model, including the
-    appropriate threshold corrections.
-  \item The user-defined constraints for the SUSY preserving or dimensionless
-    parameters are applied.
-  \item All model parameters are run to the high scale (\code{HighScale}).
-  \item The high scale is recalculated if necessary.
-  \item The user-defined boundary conditions at this scale for the
-    SUSY preserving or dimensionless parameters are applied.
-  \item The model parameters are run to the SUSY scale (\code{SUSYScale})
-    and the SUSY scale is updated if necessary.
-  \item The boundary conditions for the SUSY preserving or dimensionless
-    parameters are applied.
-  \item If not converged, goto \ref{inner-iteration-step-one}.
-  \end{enumerate}
-\item All model parameters are run to the scale at which the EWSB equations
-  are to be solved.
-  \begin{enumerate}
-  \item The coefficients in the semi-analytic solutions are determined
-    at this scale, using the current estimate for the scale at which the
-    relevant boundary conditions are imposed.  For example, in the CMSSM, the
-    semi-analytic solutions for the soft gaugino masses, trilinears, scalar
-    masses and bilinear take the form
-    \begin{equation}
-      \begin{aligned}
-        M_i(Q) &= p_i(Q) \azero + q_i(Q) \mhalf \, , \\
-        T_i(Q) &= e_i(Q) \azero + f_i(Q) \mhalf \, , \\
-        m_i^2(Q) &= a_i(Q) \mzero^2 + b_i(Q) \mhalf^2
-        + c_i(Q) \mhalf \azero + d_i(Q) \azero^2 \, , \\
-        B\mu(Q) &= u(Q) B\mu(M_X) + v(Q) \mu(M_X) \mhalf
-        + w(Q) \mu(M_X) \azero \, .
-      \end{aligned} \label{eq:cmssm-semi-analytic-solns}
-    \end{equation}
-    The coefficients are determined numerically by varying the values of
-    $\mhalf$, $\azero$, $\mzero$ and $B\mu(M_X)$ and integrating the RGEs from
-    $M_X$ to $Q$.  For example, the coefficients $p_i(Q)$, $e_i(Q)$, $d_i(Q)$
-    and $w(Q)$ are obtained by keeping only $\azero \neq 0$.  A similar approach
-    is followed to successively obtain all of the remaining coefficients.
-  \item The calculated semi-analytic solutions are used to set the values of
-    the soft SUSY breaking or dimensionful parameters at this scale.
-  \item The \DRbar/\MSbar mass spectrum is calculated and the scale at which
-    EWSB occurs is updated.
-  \item The EWSB conditions are solved at the loop level.
-  \end{enumerate}
-\item If not converged, goto \ref{outer-iteration-step-one}.  Otherwise
-  the iteration finishes.
-\end{enumerate}
-%
-If the iteration converges, all running parameters in the model are determined
-between the low  and high scales.  The remainder of the calculation, that is,
-the calculation of the pole mass spectrum and observables, then proceeds in
-the same way as in the two-scale algorithm.  Alternatively, the iteration
-may fail to converge or may encounter problems that render the parameter point
-physically invalid.  As for the two-scale solver, the specific problems that
-are encountered for a given parameter space point are stored in the
-\code{Problems} class, and may be accessed using the \code{get_problems()}
-function of the model class.\footnote{Note that in \fstwo a separate class,
-\code{BVP_solver_problems}, is used to store those problems that are associated
-only with failures of the BVP solver algorithm, such as a failure to converge,
-and which do not necessarily mean the parameter point is ruled out.  A summary
-combining all of the problems that arise during a run of the spectrum generator
-can then be obtained by calling the \code{get_problems()} method of the spectrum
-generator class.}
-
-\begin{figure}[tbh]
-  \centering
-  \begin{tikzpicture}[node distance = 1cm, auto]
-    \tikzstyle{block} = [rectangle, draw, text width=15em,
-      text centered, minimum height=3em, minimum width=10em]
-    \tikzstyle{arrow} = [draw, -latex, thick]
-    \node[block] (MZ)
-         {Calculate $g_i(M_Z)$, $y_f(M_Z)$, apply low-scale boundary
-           conditions for SUSY parameters};
-    \node[block, below = 2em of MZ] (susy) {Iterate SUSY parameters};
-    \path[arrow] (MZ) --node {} (susy);
-    \node[block, below = 2em of susy] (coeffs)
-         {Determine semi-analytic solutions at scale of EWSB};
-    \path[arrow] (susy) --node {} (coeffs);
-    \node[block, below = 2em of coeffs] (MS) {Solve EWSB, apply boundary
-      conditions at this scale};
-    \path[arrow] (coeffs) --node {} (MS);
-    \node[block, below = 2em of MS] (thresh) {Apply threshold corrections};
-    \path[arrow] (MS) --node {} (thresh);
-    \node[rectangle, draw, text width=10em,
-          text centered, minimum height=3em, minimum width=10em,
-          left = 2em of MZ] (guess) {Initial guess};
-    \path[arrow] (guess) --node {} (MZ);
-    \path[-latex, thick] (thresh.east) edge[bend right=50] (MZ.east);
-    \node[rectangle, draw, text width=10em,
-          text centered, minimum height=3em, minimum width=10em,
-          left = 2em of thresh] (pole) {Calculate pole masses};
-    \node[inner sep=0pt, above = 3.5em of pole] (convergence) {};
-    \path[draw,thick,dashed,above] (MS) --node {Converged?} (convergence);
-    \path[arrow,dashed] (convergence) --node {} (pole);
-  \end{tikzpicture}
-  \caption{Semi-analytic algorithm for calculating the mass spectrum in a
-    SUSY model; in a non-SUSY model, the SUSY parameters are replaced by
-    the dimensionless model parameters instead.}
-  \label{fig:semi-analytic-algorithm}
-\end{figure}
-
-\subsection{Tests and comparisons with other spectrum generators}
-
-The semi-analytic algorithm provided by \fsas has been tested in the
-CMSSM, CNMSSM and the C\ESSM by comparing to the results
-obtained using the existing two-scale solver.  We have
-carried out consistency checks between the two by confirming that the same
-solution can be found in both solvers, provided it is a stable fixed
-point in both.  In each of the models, the existing model solved using
-the two-scale algorithm has been compared with versions of the model
-using alternative boundary conditions.  For example, in the CMSSM,
-instead of the traditional approach of fixing $|\mu|^2$ and $B\mu$
-using the EWSB conditions, the value of $\mu$ is provided as an input
-and $\mzero^2$ and $B\mu(M_X)$ are determined from the EWSB
-conditions.\footnote{This alternative approach is useful in scenarios where
-  one wishes to have direct control over the Higgsino masses, and therefore
-  the composition of the lightest neutralino in the CMSSM, as was done in
-  Ref.\ \cite{Athron:2016gor}.}  The benchmark points used as inputs for the
-semi-analytic solver in each model are displayed in
-\tabref{tab:semi_analytic_bms}.  In all three models, the
-running parameters and pole mass spectra are found to differ at or
-below the level of $0.1$\%.  Unit tests that perform these
-comparisons have also been added to the \fs test suite.
-%
-\begin{table}[tbh]
-  \centering
-  \begin{tabularx}{\textwidth}{lX}
-    \toprule
-    Model & Unit test benchmark points for the semi-analytic solver\\
-    \midrule
-    CMSSM & $\mhalf = 500$ GeV, $\tan\beta = 10$, $\azero = 0$ GeV,
-    $\mu(\MS) = 623.36$ GeV\\
-    CNMSSM & $\mhalf = 133.33$ GeV, $\tan\beta = 10$,
-    $\sign \mu_{\text{eff}} = -1$, $\azero = -300$ GeV, $\lambda(M_X) =-0.05$\\
-    C\ESSM & $\tan\beta = 10$, $\lambda_3(M_X) = 0.12$, $\kappa(M_X) = 0.2$,
-    $\mu^\prime(M_X) = 10$ TeV, $B^\prime\mu^\prime(M_X) = 0$ GeV$^2$,
-    $s(\MS) = 4$ TeV, $\lambda_{1,2}(M_X) = 0.1$\\
-    \bottomrule
-  \end{tabularx}
-  \caption{Input parameter values used for the unit tests comparing the
-    results of the two-scale and semi-analytic algorithms in the CMSSM,
-    CNMSSM and C\ESSM.  The notation for the CNMSSM follows that in
-    Refs.~\cite{Allanach:2013kza,Ellwanger:2009dp}, while for the
-    C\ESSM we use the notation of Ref.~\cite{Athron:2009bs}.}
-  \label{tab:semi_analytic_bms}
-\end{table}
-
-In addition to carrying out unit tests on individual benchmark points, we have
-also performed extensive scans in the CMSSM to confirm that the semi-analytic
-solver produces results in agreement with the two-scale solver.  For most
-points, this is found to be the case; however, we have also observed important
-exceptions where non-negligible differences are found between the two
-solvers.  In these cases, one solver may fail to converge to a stable
-solution, or multiple solutions are present \cite{Allanach:2013cda} with
-different stability properties in the two solvers.  In this latter case, note
-that both solvers always return the first solution to
-which they converge,\footnote{When multiple convergent solutions exist,
-  the one first obtained will depend, for instance, on the initial guess
-  used for the iteration.} with the iteration
-stopping immediately once a convergent solution is obtained.  That is, neither
-method attempts to find all possible solutions for the given input parameters
-or automatically select between multiple fixed points.  Since a given solution
-might not be a stable fixed point of both iterations, the two solvers need not
-converge to the same solution, leading to the observed differences.  We have
-checked that such points nevertheless satisfy the boundary conditions imposed at
-each scale and are indeed valid solutions of the BVP.  More generally we have
-checked that solutions found in one solver also correspond to (not necessarily
-stable) fixed point solutions of the other solver algorithm; that is, they
-satisfy all of the boundary conditions so that the parameter values remain
-unchanged after applying a single step of the iteration.
-
-From these tests, we have found that in some cases the agreement
-between the two solvers can depend quite sensitively on small differences
-between them.  To illustrate this, in \figref{fig:semi_analytic_differences}
-we show the percentage changes in the \DRbar mass spectrum in the CMSSM after
-running points obtained using the two-scale solver through a single step of
-the semi-analytic solver; if the point is also a fixed point of the latter,
-this change should be negligible.  In this
-scan, the change after a single iteration can be on the level of several
-percent for a small number of points, reaching between $20$\% and $30$\% for
-some exceptional points.
-%
-\begin{figure}[tbh]
-  \centering
-  \includegraphics[width=0.6\textwidth]{%
-    doc/plots/semi_analytic_rel_errs/single_step_errors}
-  \caption{Percentage changes in the CMSSM \DRbar mass spectrum after
-    applying a single step of the semi-analytic solver to points obtained
-    using the two-scale solver in a linear scan over $\mhalf \in
-    [0,300]$ GeV, $\mzero \in [0,2]$ TeV, $\azero = 0$ GeV, $\tan\beta = 40$ and
-    $\mu > 0$.  When run using the default Runge-Kutta algorithm provided
-    by \fs and allowing the iteration to stop as soon as the precision goal of
-    $10^{-4}$ is reached (red circles), changes between $1$\% and $30$\% are
-    found for a small number of points.  By using the $8^{\text{th}}$ order
-    Runge-Kutta integrator and ensuring convergence is reached in the
-    estimate for $M_X$ by forcing 40 iterations in the two-scale algorithm,
-    these differences are reduced below the level of $0.001$\% (blue
-    diamonds).}
-  \label{fig:semi_analytic_differences}
-\end{figure}
-%
-That these points initially appear not to be fixed points of the semi-analytic
-solver arises primarily from the fact that the semi-analytic coefficients,
-and hence the EWSB solution, are sensitive to the estimate for the
-high scale $M_X$, as well as numerical errors in the integration of the RGEs.
-In particular, for the default convergence criteria imposed by \fs, the
-two-scale solver's estimate for $M_X$ is not close enough to convergence,
-leading to significant differences in the calculated low-energy soft
-parameters.  By requiring convergence in the estimate for $M_X$, together with
-using a higher-order Runge-Kutta integration\footnote{By default, \fs makes
-  use of an adaptive $5^{\text{th}}$ order algorithm; to perform this test we
-  have also implemented into \fstwo an $8^{\text{th}}$ order solver that makes
-  use of the Runge-Kutta-Fehlberg method provided by the Boost library
-  \texttt{odeint}.  This higher-order solver is available to the user by
-  choosing it at the C++ level.} and demanding a higher precision for the
-obtained EWSB solution, the change after one iteration is reduced below the
-permille level.  Thus, it is important to be aware that differences in the
-convergence properties of the two solvers can have an impact on the solutions
-found, even if a given point would be a fixed point of both solvers.
-
-The typical runtimes for the two solvers in the CMSSM are compared in
-\figref{fig:semi_analytic_benchmark}.  The distributions are obtained
-by randomly sampling the CMSSM input parameters
-$\mzero\in [0.2,1]\unit{TeV}$, $\mhalf\in [0.2,1]\unit{TeV}$,
-$\tan\beta\in [2,30]$, $\sign\mu\in\{-1,+1\}$ and
-$\azero\in [-1,1]\unit{TeV}$.
-The runtime of the semi-analytic solver is increased on average by a factor of
-$\approx 3$ compared to the two-scale solver.  This increase is
-mostly due to the increased number of iterations performed by the semi-analytic
-solver.  For each outer iteration of the semi-analytic solver, the inner
-iteration typically runs through a similar number of steps as for a full run
-of the two-scale solver, with this number decreasing as convergence is
-approached on each outer iteration.  Consequently, the total number of
-iterations for the semi-analytic solver tends to be larger than that for the
-two-scale solver by a similar factor.  There is also an additional cost
-associated with running between scales to compute the semi-analytic
-coefficients.
-%
-\begin{figure}[tbh]
-  \centering
-  \includegraphics[width=0.5\textwidth]{%
-    doc/plots/semi_analytic_benchmark/i7-4702MQ}
-  \caption{Runtime distributions of the CMSSM spectrum generators created with
-    \fstwo using the two-scale and semi-analytic BVP solvers, obtained on
-    an Intel i7-4702MQ CPU\@.  The distributions are normalized to have integral
-    of unity.}
-\label{fig:semi_analytic_benchmark}
-\end{figure}
-
-While the semi-analytic solver suffers from an increased runtime compared to
-the two-scale solver, it is also able to provide complementary coverage of the
-parameter space to that of the two-scale solver.  This is demonstrated in the
-left panel of \figref{fig:cmssm_solution_regions} in the CMSSM, where the
-solutions found by each solver are plotted in the $\mzero - \mu$ plane.
-%
-\begin{figure}
-  \centering
-  \includegraphics[width=0.48\textwidth]{%
-    doc/plots/CMSSM_solution_regions/CMSSM_solution_regions}
-  \includegraphics[width=0.48\textwidth]{%
-    doc/plots/CMSSM_solution_regions/CMSSM_higgs_mass}
-  \caption{Left panel: CMSSM solutions found by the two-scale solver
-    (blue diamonds) and the semi-analytic solver (red circles) in the
-    $\mzero - \mu$ plane.  For both solvers, $\mhalf \in [0,1]\unit{TeV}$
-    is randomly sampled, while $\tan\beta = 10$ and $\azero = 0$ GeV\@.
-    The solutions obtained using the two-scale solver are found by
-    randomly sampling $\mzero \in [0, 6]\unit{TeV}$ with $\sign \mu = 1$,
-    and those for the semi-analytic solver are found by randomly sampling
-    $\mu \in [0,1150] \unit{GeV}$.  Right panel: Change in the calculated
-    lightest \CP-even Higgs pole mass $M_{h_1}$ from a reference value of
-    $121$ GeV as a function of $\mu$ at fixed $\mzero = 4$ TeV for the two
-    solvers.  The solutions shown correspond to a vertical slice at fixed
-    $\mzero$ in the left-hand plot; $\mhalf$ is again varied in
-    $[0,1]\unit{TeV}$, $\tan\beta = 10$ and $\azero = 0$ GeV.}
-  \label{fig:cmssm_solution_regions}
-\end{figure}
-%
-In this case, the use of the semi-analytic solver allows for a large number
-of solutions to be found in the focus point region \cite{Chan:1997bi,
-  Feng:1999hg,Feng:1999mn} at small values of $\mu \ll \mzero$, where the
-two-scale solver is unable to find convergent solutions.  Conversely,
-the two-scale solver is more effective for finding solutions with small
-$\mzero$.  This highlights the fact that, for some parameter points,
-the semi-analytic solver generated by \fs might not find a solution
-where the two-scale solver is able to, and vice versa.  As noted above, this
-can be due to the point in question having different stability properties
-under the two different algorithms.  For example, the CMSSM point with
-$m_0 = 125\unit{GeV}$, $M_{1/2} = 300\unit{GeV}$, $\tan\beta = 10$,
-$\sign \mu = 1$ and $A_0 = 0\unit{GeV}$  is successfully solved using the
-two-scale solver, with $\mu \approx 395\unit{GeV}$ required for correct EWSB.
-The same point, when run using the semi-analytic solver with the two-scale
-solution for $\mu$, fails to converge; for this choice of input parameters
-and initial guess, the iteration enters a periodic orbit in which the
-approximations to the solution are close to, but do not correspond to, the
-fixed point found by the two-scale solver.  To avoid the observed cyclic
-behaviour here, it is necessary to either fine-tune the provided input
-parameters or otherwise modify the initial guess made by \fs.  Due to the
-differing choice of input and output parameters between the two algorithms
-in general, certain regions of the parameter space can also be susceptible to
-high levels of numerical sensitivity in one approach and not the other,
-again necessitating significant fine-tuning.  In the CMSSM, for instance,
-large cancellations are typically required in
-order to produce a small value of $m_0$ when using the semi-analytic solver.
-In such cases, $\mu$ must be carefully fine-tuned to obtain a valid solution.
-On the other hand, when using the two-scale solver one has direct control over
-the value of $m_0$, and it is not necessary to fine-tune in order to obtain
-the desired small value of $m_0$.  The situation is reversed in regions of
-parameter space with small values of $\mu$ and large $m_0$, where one now
-has direct control over the fine-tuned value of $\mu$ in the semi-analytic
-solver but not in the two-scale solver.  In general terms, the regions of
-parameter space in which the two solvers are effective need not overlap, and
-the use of both in tandem allows for a more complete picture of the parameter
-space to be obtained.  Moreover, in the regions in which both solvers do find
-solutions, there is excellent agreement between the two algorithms.\footnote{
-  Provided the same solution is found if multiple solutions exist.}  This is
-illustrated in the right panel of \figref{fig:cmssm_solution_regions}, where
-the lightest \CP-even Higgs mass, expressed as the difference from a reference
-value of $121$ GeV, is plotted for fixed $\mzero = 4$ TeV\@.  If both solvers
-find a solution for this value of $\mzero$, the two values of the Higgs mass
-agree very well.
-
-It is also evident from \figref{fig:cmssm_solution_regions} that the use of
-both solvers allows features in the parameter space to be picked up that would
-be missed by either solver alone.  In this case, in
-\figref{fig:cmssm_solution_regions} the two-scale solver only finds a single
-solution for each value of $\mhalf$, for fixed $\sign\mu = 1$, while the
-semi-analytic solver in some cases finds multiple solutions.  These
-solutions have different values of $|\mu|$ for the same value of $\mhalf$,
-leading to the sharp feature at low values of $\mu$ evident in the right panel
-of \figref{fig:cmssm_solution_regions}.  The existence of multiple solutions to
-the CMSSM BVP, and the inability of the ordinary two-scale fixed point iteration
-to find all such solutions, are well-known and have previously been studied in
-Refs.~\cite{Allanach:2013cda,Allanach:2013yua}.  In
-\figref{fig:cmssm_multiple_solutions} we compare the results obtained using
-the two BVP solvers in \fstwo to those found using the modified version of
-\softsusy employed in Ref.~\cite{Allanach:2013cda}.
-%
-\begin{figure}
-  \centering
-  \includegraphics[width=0.6\textwidth]{%
-    doc/plots/CMSSM_multiple_solutions/CMSSM_multiple_solutions}
-  \caption{Solutions obtained using the two-scale and semi-analytic solvers
-    in the CMSSM for fixed $\mhalf = 660$ GeV, $\azero = 0$ GeV and
-    $\tan\beta = 40$.  For comparison, we also reproduce the curve shown in
-    Figure~5 of Ref.~\cite{Allanach:2013cda}, where a modified version of
-    \softsusy 3.3.7 was used to allow multiple solutions to be found to the
-    CMSSM boundary conditions.  The inset plot shows the region $2790\unit{GeV}
-    \leq \mzero \leq 2820\unit{GeV}$, $-555\unit{GeV} \leq \mu
-    \leq -520\unit{GeV}$ in more detail.}
-  \label{fig:cmssm_multiple_solutions}
-\end{figure}
-%
-As expected, in this region of the parameter space the two-scale solver
-produced by \fs finds at most only two solutions, corresponding to the two
-possible signs of $\mu$.  The automatically generated semi-analytic solver
-enables additional solutions to be found, which are in good agreement with
-those found using the modified version of \softsusy.  Slight differences in
-the values of $\mu$ and $\mzero$, and where solutions are found, arise from
-small differences in the incorporated corrections between the two codes.  In
-particular, the semi-analytic solver does not find valid solutions for
-$\mu < -545$ GeV due to the tree-level mass of the \CP-odd Higgs becoming
-tachyonic.  In general, multiple solutions at a given parameter point
-can have significantly different phenomenological properties
-\cite{Allanach:2013yua}, so that finding them is important to completely
-characterize a model.  The availability of multiple solvers improves the
-ability of \fs to locate additional solutions, without requiring extensive
-modifications to the generated code.
-
-The longer runtime of the semi-analytic solver can also be an
-acceptable tradeoff if the model of interest cannot easily be
-handled using the two-scale solver.  Constrained models such as the
-CNMSSM and C\ESSM are examples of this.  To demonstrate the applicability of
-the semi-analytic solver to these models, we have performed
-scans over the parameter spaces of these models.
-
-First we have performed scans in the CNMSSM to demonstrate that with the
-semi-analytic solver we can sample the parameter space effectively, making
-\fstwo the first public spectrum generator that can do this ``out of the box''.
-We performed a four dimensional scan of the CNMSSM, using the model file
-provided in \appref{app:CNMSSM} to produce a CNMSSM spectrum generator that
-was then linked to \multinest 3.10 for efficient sampling of the parameter
-space.  The input parameters were varied over the ranges
-\begin{gather}
-  2 \leq \tan \beta \leq 50, \\
-  0\unit{TeV} \leq \mhalf \leq 5\unit{TeV}, \\
-  -5\unit{TeV} \leq \azero \leq 0\unit{TeV}, \\
-  -0.3 \, \leq \lambda(M_X) \leq 0.3, \\
-  \sign \mu_{\textrm{eff}}  = +1,
-\end{gather}
-and the log likelihood function provided to \multinest was defined to be
-\begin{equation*}
-  \log L = -\frac{1}{5}({M_h^{\SM}}/{\unit{GeV}} - 125.09)^2 ,
-\end{equation*}
-where $M_h^{\SM}$ is the SM-like Higgs pole mass.  In this way the
-scan is directed towards solutions with the observed Higgs mass of
-$125.09$ GeV\@.
-
-\begin{figure}[tbh]
-  \centering
-  \includegraphics[width=0.48\textwidth]{%
-     doc/plots/CNMSSM/cnmssm_4D-m0sqpos-full-m0_m12_sort-A0.pdf}
-   \includegraphics[width=0.48\textwidth]{%
-     doc/plots/CNMSSM/cnmssm_4D-m0sqpos-full_mhZoom_lam_m12_sort-mh.pdf} \\
-  \includegraphics[width=0.48\textwidth]{%
-    doc/plots/CNMSSM/cnmssm_4D-full-squark_mass_m0sqgtr0_m0_m12.pdf}
-   \includegraphics[width=0.48\textwidth]{%
-    doc/plots/CNMSSM/cnmssm_4D-m0sqgtr0-full-msu6.pdf}
-   \caption{CNMSSM solutions obtained using \fstwo and \multinest 3.10
-     shown in the $\mzero-\mhalf$ plane with $\azero$ as a color
-     contour (top left panel); $\mhalf-\lambda$ plane showing the
-     SM-like Higgs pole mass $M_h^{\SM} $, as a color contour, with
-     the range restricted to $M_h^{\SM} > 100$ GeV for clarity (top
-     right panel); $\mzero-\mhalf$ plane with the mass of the lightest
-     up-type squark, which is predominantly stop (bottom left panel),
-     and the $\mhalf-\azero$ plane with the heaviest up-type squark
-     mass (bottom right panel).}
-   \label{fig:cnmssm_solutions}
-\end{figure}
-
-The results are shown in \figref{fig:cnmssm_solutions}, where all points are
-required to have $\mzero^2 > 0$.  As can be seen in the top left panel, the range
-of $\mzero$ values found is considerably smaller than the range allowed for the
-universal soft masses, $\mhalf$ and $\azero$, which are inputs.  The reason for
-this is that in the CNMSSM a non-zero singlet VEV must be generated by
-developing the correct shape of the scalar potential.  This could be achieved
-in the usual way through a negative quadratic term if the soft breaking singlet
-mass squared, $m_S^2$, is negative.  In fact, since the NMSSM scalar potential
-also contains the cubic singlet terms $\kappa A_\kappa S^3 / 3 + \mathrm{h.c.}$,
-the requirement $m_S^2 < 0$ can be relaxed so that an approximate condition for
-generating the correct shape of the potential reads $A_\kappa^2 \gtrsim
-9 m_S^2$ \cite{Ellwanger:1996gw,Derendinger:1983bz}.  If $\mzero \gg
-\azero / 3$, this condition is satisfied only if $m_S^2$ is driven to be
-sufficiently small during the RG evolution from the GUT scale to the SUSY
-scale.  This in turn can be achieved for large enough values of the
-superpotential cubic singlet coupling $\kappa$ and the singlet-Higgs coupling
-$\lambda$.
-
-However, in this constrained model, where the soft trilinears are not free
-input parameters at the SUSY scale, it is also the case that large values of $\lambda$
-always generate substantial singlet mixing that can reduce the lightest
-\CP-even Higgs mass.  As a result, a $125$ GeV Higgs mass is obtained with small
-values of $\lambda$, to avoid this mixing, as well as large $\mhalf$, as can be
-seen in the top right panel of \figref{fig:cnmssm_solutions}.  For such small
-singlet Yukawa couplings, the RG flow between the GUT and SUSY scales leaves
-$m_S^2(\MS)$ and $A_\kappa(\MS)$ close to their GUT scale values, namely
-$\mzero^2$ and $\azero$.  As a result, $\mzero$ is heavily constrained if
-the condition $A_\kappa^2 \gtrsim 9 m_S^2$ is to be satisfied in the absence of
-significant RG evolution for $m_S$, as the top left panel of
-\figref{fig:cnmssm_solutions} demonstrates.
-
-The consequences of such small $\mzero$ values can be seen in the bottom left
-and bottom right panels of \figref{fig:cnmssm_solutions}, where we plot the
-lightest and heaviest up-type squark mass, respectively, to illustrate that the
-squark masses are now predominantly set by the universal gaugino mass, with
-little influence from $\mzero$ or $\azero$.  Furthermore, the squarks are always
-lighter than the gluino, which has a mass $\approx 2\, \mhalf$.
-
-\begin{figure}[tbh]
-  \centering
-  \includegraphics[width=0.48\textwidth]{%
-     doc/plots/CE6SSM/ce6ssm_tb10-scan-l-k-vs_combined_m0_m12_ccA0.pdf}
-   \includegraphics[width=0.48\textwidth]{%
-     doc/plots/CE6SSM/ce6ssm_tb10-scan-l-k-vs_combined_m0_m12_ccmsu1.pdf} \\
-  \includegraphics[width=0.48\textwidth]{%
-    doc/plots/CE6SSM/ce6ssm_tb10-scan-l-k-vs_combined_m0_m12_cclam.pdf}
-   \includegraphics[width=0.48\textwidth]{%
-    doc/plots/CE6SSM/ce6ssm_tb10-scan-l-k-vs_combined_m0_m12_cckap.pdf}
-   \caption{C\ESSM solutions obtained using \fstwo and \multinest 3.10 shown in
-     the $\mzero-\mhalf$ plane with $\azero$, the lightest up-type squark pole mass
-     $M_{\tilde{u}_1}$,  $|\lambda|$ and $|\kappa|$ as color contours.}
-  \label{fig:ce6ssm_solutions}
-\end{figure}
-
-Compared to the CNMSSM, we find that the situation in the C\ESSM is rather
-different.  The C\ESSM is an alternative to the CNMSSM in which an elementary
-$\mu$ term is forbidden by a $U(1)_N$ gauge symmetry, and complete $E_6$ matter
-supermultiplets are included to ensure anomaly cancellation.  As with the
-CNMSSM, the two-scale solver is ineffective for finding solutions because EWSB
-needs to have a soft mass as an output, but one can make a spectrum generator
-with the semi-analytic solver where the universal GUT scale masses $\mzero$,
-$\mhalf$ and $\azero$ are EWSB outputs.\footnote{The definitions of the
-  new Lagrangian parameters and GUT scale constraints in the C\ESSM that we
-  use may be found in Ref.~\cite{Athron:2009bs}.}  Using the \modelname{C\ESSM}
-model file provided with \fstwo, we carried out a scan in which the
-singlet-Higgs Yukawa coupling, $\lambda(M_X) \equiv \lambda_3(M_X)$, and the
-exotic Yukawa coupling, $\kappa(M_X)$, were varied over $[-1,1]$ and the singlet
-VEV, $s$, was varied over $[1,500] \unit{TeV}$.  The results of this scan are
-shown in \figref{fig:ce6ssm_solutions}.
-
-In contrast to the CNMSSM, in the C\ESSM it is very easy to have large
-$\mzero$ values as there is an exotic Yukawa coupling between the singlet and
-extra colored matter introduced to avoid gauge anomalies that drives the soft
-singlet mass negative, providing a radiative symmetry breaking mechanism.
-Additionally, whereas $\mzero \lesssim \mhalf$ in the CNMSSM, in this model
-it is typically the case that $\mzero$ is larger than $\mhalf$.  This is
-in qualitative agreement with the literature \cite{Athron:2009ue,
-  Athron:2009bs,Athron:2012sq,Athron:2012pw}, and arises because the new
-colored matter results in heavily modified RGEs in which the 1-loop
-$\beta$ function of the strong coupling now vanishes.  The squark masses are
-mostly set by $\mzero$ as a result, as is shown in the top right panel of
-\figref{fig:ce6ssm_solutions}, while the bottom panels show that the range for
-$\lambda$ and $\kappa$ at the GUT scale is very wide, despite strong
-constraints on $\lambda$ at the electroweak scale coming from requirements for correct
-EWSB\@.   We do not make a detailed quantitative comparison to previous work in
-the literature here, but note that this is the first time that the C\ESSM
-results have been presented with the same level of precision (full 2-loop RGEs,
-1-loop pole masses) as is standard in the CMSSM and significant quantitative
-differences are to be expected.
-
-For a more precise comparison between calculations performed at the
-same level of precision, we have also performed scans in a recently proposed
-variant of the \ESSM, the so-called CS\ESSM \cite{Athron:2015vxg,
-  Athron:2016gor}. Here the results obtained using \fstwo have been checked for
-agreement with those obtained from a hand-written prototype of the
-semi-analytic solver that was implemented for the studies in
-Refs.~\cite{Athron:2015vxg,Athron:2016gor}.  The solutions found using the
-generated CS\ESSM spectrum generator are compared with those found in
-Ref.~\cite{Athron:2016gor} in \figref{fig:cse6ssm_m12_Azero_plane}.  The viable
-solution regions and values of the model parameters are found to be in very
-good agreement with the results obtained using the earlier code.
-%
-\begin{figure}[tbh]
-  \centering
-  \includegraphics[width=0.48\textwidth]{%
-    doc/plots/CSE6SSM/CSE6SSM_m12_Azero_m0}
-  \includegraphics[width=0.48\textwidth]{%
-    doc/plots/CSE6SSM/CSE6SSM_m12_Azero_m0_handwritten}
-  \caption{CS\ESSM solutions obtained using the spectrum generator
-    automatically generated by \fstwo (left panel) and using the prototype
-    spectrum generator used in the numerical analysis of
-    Ref.~\cite{Athron:2016gor} (right panel), showing good agreement between
-    the two codes.  Note that here we have not applied the limits on the Higgs
-    mass or dark matter relic density that lead to additional restrictions on
-    the parameter space, as discussed in Ref.~\cite{Athron:2016gor}.}
-  \label{fig:cse6ssm_m12_Azero_plane}
-\end{figure}
-
-\section{\feft}
-\label{sec:feft}
-
- \feft is a method to predict the lightest Higgs pole mass in any \BSM
- model accurately for both high and low new physics scales $\MS$ and
- was presented first in Ref.~\cite{Athron:2016fuq}.  An implementation
- of this method was first released in \fs 1.7.0.  Here we present an
- upgrade at the next-to-leading order and next-to-leading logarithmic
- (NLO+NLL) accuracy, which we release in \fstwo.
-
-\feft combines an EFT approach with a diagrammatic
-calculation, allowing for an all-order resummation of large logarithms
-of the ratio $\MS/m_t$, together with the inclusion of all
-non-logarithmic 1-loop contributions.  In particular, all
-power-suppressed 1-loop contributions of $O(v^2/\MS^2)$ are
-included in \feft, which would otherwise be neglected in a pure EFT
-calculation.  Thanks to these properties, \feft maintains the accuracy
-at all scales: For low scales the prediction agrees with a
-fixed-order calculation; for large scales it agrees with a pure EFT
-calculation. In the intermediate region, where the $O(v^2/\MS^2)$
-terms are small, but still non-negligible, \feft gives the correct
-fixed-order result plus higher-order logarithms, thus resolving the
-ambiguity between the fixed-order and the pure EFT approach.
-
-In Ref.~\cite{Athron:2016fuq} already several versions of the \feft
-approach have been extensively discussed and compared with existing
-calculations of the lightest Higgs boson mass in the MSSM and other
-supersymmetric models. The approach has also been implemented recently
-in \sarah/\spheno \cite{Staub:2017jnp}, including 2-loop corrections in
-the matching. The version implemented in  \fstwo contains
-additional improvements resulting in a higher accuracy. In the
-following, we briefly summarize the main idea of \feft, then explain
-the details of the implemented version and how to use it. For further
-details of the approach and a detailed comparison of theoretical
-uncertainties, we refer to Ref.~\cite{Athron:2016fuq}.
-
-\subsection{Basic matching condition}
-
-\feft performs a matching of the \BSM model to the SM,
-thereby determining the quartic Higgs coupling $\lambda$ of the
-SM\@.  The basic ingredient of \feft to fix $\lambda$ at
-the matching scale is a Higgs pole mass matching condition
-%
-\begin{align}
-  (M_h^\SM)^2 = (M_h^\BSM)^2 \,,
-  \label{eq:Higgs_pole_mass_matching}
-\end{align}
-%
-where $M_h^\SM$ is the Higgs pole mass calculated in the SM at the 1-loop level and $M_h^\BSM$ is the corresponding SM-like
-Higgs pole mass in the \BSM model, also at the 1-loop
-level.
-Generally, the Higgs pole mass is computed in any \BSM model
-by solving the following
-equation (or a suitable matrix generalization):
-%
-\begin{align}
-  (M_h^\BSM)^2 = (m_h^\BSM)^2 - \re\Sigma_h^{\BSM}(p^2) + \frac{t_h^\BSM}{v} \,,
-  \label{eq:Mh_mom_iteration}
-\end{align}
-%
-where $m_h^\BSM$ is the respective tree-level mass and
-$\Sigma_h^{\BSM}$ and $t_h^\BSM$ are the Higgs self
-energy and tadpole in the \MSbar/\DRbar scheme.
-In principle, in an all-order calculation,  the
-Higgs self energy has to be evaluated at the momentum $p^2 =
-(M_h^\SM)^2=(M_h^\BSM)^2$.
-  From this condition the quartic Higgs coupling of the SM can be extracted as
-%
-\begin{align}
-  \lambda = \frac{1}{v^2} \left[
-     (M_h^\BSM)^2 + \re\Sigma_h^{\SM}((M_h^\SM)^2) - \frac{t_h^\SM}{v}
-  \right] \,.
-  \label{eq:lambda_from_Mh_matching}
-\end{align}
-%
-
-This matching is equivalent
-to the one of pure EFT calculations \cite{Bagnaschi:2014rsa,Vega:2015fna} at the
-1-loop level, up to power-suppressed terms. Correspondingly, the resulting
-Higgs boson mass is exact at the 1-loop level and takes into account
-all leading logarithms \cite{Athron:2016fuq}.
-
-\subsection{New matching procedure in \fstwo}
-\label{sec:FS-2_matching_procedure}
-
-\fstwo has an improved implementation of the approach, which is
-still exact at the 1-loop level but also
-correctly resums next-to-leading logarithms.
-This improvement originates from an amended
-matching procedure. As mentioned before, the matching procedure
-in Eq.~\eqref{eq:lambda_from_Mh_matching}
-is equivalent to a pure EFT matching at the 1-loop level. However,
-depending on implementation details, it can differ by terms of
-2-loop or higher order. If these spurious 2-loop terms contain
- (next-to-leading) large logarithms, they spoil the correct
-resummation of (next-to-leading) logarithms by RGE running.
-
-By construction, all versions of \feft discussed in
-Ref.~\cite{Athron:2016fuq} and Ref.~\cite{Staub:2017jnp} are correct at the
-leading logarithmic level, however not all subleading logarithms are correctly
-included.
-%
-
-In the following, we discuss the two potential origins of these subleading
-logarithms and how they are avoided by the improved  implementation in
-\fstwo.
-
-\paragraph{Insertion of 1-loop parameters into 1-loop \BSM self energies or
-  tadpoles}
-
-The first potential source of large 2-loop logarithms in the matching
-procedure is the insertion of parameters, which have been obtained from the SM
-via a 1-loop matching, into the 1-loop self energies or tadpoles of the \BSM
-model.  We illustrate this effect with the most important parameter, the top
-Yukawa coupling:
-The running top Yukawa coupling of the \BSM model  $y_t^\BSM$ is
-determined by a matching as\footnote{We ignore potential tree-level
-  factors here for brevity.}
-%
-\begin{align}
-  y_t^\BSM = y_t^\SM + \Delta y_t \,, \label{eq:yt_SM_BSM}
-\end{align}
-%
-where $\Delta y_t$ is of 1-loop order (but without large logarithms).
-At the same time, the Higgs pole mass calculations on the left-hand side and
-right-hand side of Eq.~\eqref{eq:Higgs_pole_mass_matching} are of the form
-%
-\begin{align}
-  (M_h^\SM)^2 &= (m_h^\SM)^2 + \propto \frac{(v^\SM)^2(y_t^\SM)^4}{(4\pi)^2} \log\frac{m_t^\SM}{Q} + \cdots \,,
-  \label{eq:Mh_top_SM}\\
-  (M_h^\BSM)^2 &= (m_h^\BSM)^2 + \propto \frac{(v^\BSM)^2(y_t^\BSM)^4}{(4\pi)^2} \log\frac{m_t^\BSM}{Q} + \cdots \,,
-  \label{eq:Mh_top_BSM}
-\end{align}
-%
-where the matching scale $Q$ is of the order $\MS$ and we have also
-introduced $m_h^\SM$, $m_t^\SM$ and $m_t^\BSM$ for the running SM
-Higgs mass, running SM top mass and running BSM top mass,
-keeping our convention of using an upper case `$M$' for pole masses
-and lower case `$m$' for running tree-level masses.  If
-Eq.\ \eqref{eq:Mh_top_SM} and Eq.\ \eqref{eq:Mh_top_BSM} are set equal
-and the relation Eq.\ \eqref{eq:yt_SM_BSM} is inserted, potentially
-large 2-loop terms for example of the form
-%
-\begin{align}
-  \frac{4(v^\SM)^2(y_t^\SM)^3 \Delta y_t}{(4\pi)^2} \log\frac{m_t^\SM}{Q} + \cdots
-\label{eq:twolooplogsyt}
-\end{align}
-%
-remain. Such terms effectively shift the quartic Higgs coupling of the
-SM by  next-to-leading logarithmic 2-loop terms.%
-\footnote{If the self energies are evaluated at the
-  2-loop level, the problem repeats itself one order higher, i.e., the
-  term in Eq.~\eqref{eq:twolooplogsyt} is cancelled but similar terms of
-  next-to-leading logarithmic 3-loop order remain.}
-
-In order to avoid large higher-order logarithms originating from the
-insertion of 1-loop parameters into 1-loop \BSM self energies and
-tadpoles, \fstwo maintains two different sets of running \BSM
-parameters: One parameter set which has been obtained from the
-SM using a tree-level matching, and another set
-from the SM using  1-loop matching.
-%
-The tree-level parameter set is used to evaluate the 1-loop self
-energies and tadpoles on the right-hand side of
-Eq.~\eqref{eq:Mh_mom_iteration}.  In this way, no terms like the ones
-in Eq.~\eqref{eq:twolooplogsyt} are generated.
-%
-The 1-loop parameter set is used to evaluate the tree-level Higgs
-mass (matrix) of the \BSM model.  In this way, the desired 1-loop
-corrections to the quartic Higgs coupling $\lambda$ are generated.
-
-\paragraph{Momentum iteration}
-
-The second source of large 2-loop logarithms in the matching has to
-do with the momentum argument of the self energies entering
-Eqs.~\eqref{eq:Higgs_pole_mass_matching} and \eqref{eq:Mh_mom_iteration}.
-Writing $p^2=(m_h^\BSM)^2+\Delta p^2$, we see
-that the momentum argument of Eq.~\eqref{eq:Mh_mom_iteration}
-contains the 1-loop term $\Delta p^2$ (which
-also involves large logarithms).
-%
-The difference between the left-hand side and the right-hand side of the
-matching condition, Eq.~\eqref{eq:Higgs_pole_mass_matching},
-then contains 2-loop terms, which can be expanded as
-%
-\begin{align}
-\left.\left(
-\frac{\partial}{\partial
-        p^2}\re\Sigma_h^{\BSM}(p^2)
--
-\frac{\partial}{\partial
-        p^2}\re\Sigma_h^{\SM}(p^2)
-\right)
-\right|_{p^2 = (m_h^\BSM)^2}
-\,\Delta p^2
-\,.
-\end{align}
-%
-If the self energies
-are evaluated at the 1-loop level and $p^2$ is determined as
-described above, these terms do not cancel against anything.
-Like the terms discussed in Eq.~\eqref{eq:twolooplogsyt}, these terms
-would then lead to large 2-loop next-to-leading logarithms in the
-determination of $\lambda$.%
-
-To avoid large higher-order logarithmic contributions coming from the
-momentum argument, \fstwo does not  perform the usual momentum
-iteration when the Higgs pole masses in the SM and in the
-\BSM model are calculated at the matching scale for
-Eq.~\eqref{eq:Higgs_pole_mass_matching}.  Instead, the SM Higgs pole mass at
-the matching scale is now calculated as
-%
-\begin{align}
-  (M_h^\SM)^2 = (m_h^\SM)^2 - \re\Sigma_h^{\SM}((m_h^\BSM)^2) + \frac{t_h^\SM}{v} \,,
-  \label{eq:Higgs_pole_mass_matching_new}
-\end{align}
-%
-where the self energy momentum is set to the tree-level
-\MSbar/\DRbar Higgs mass $m_h^\BSM$ in the \BSM model at the matching
-scale, which is calculated in terms of running \BSM parameters which
-have been obtained by a tree level matching.
-%
-A similar expression is used to calculate the Higgs pole mass in the
-\BSM model $M_h^\BSM$, where we also insert $p^2 = (m_h^\BSM)^2$ as the
-self energy momentum in order to enable cancellation of
-momentum-dependent terms.  For example, if the \BSM Higgs is a
-singlet, the \BSM Higgs pole mass is calculated as
-%
-\begin{align}
-  (M_h^\BSM)^2 = (m_h^\BSM)^2 - \re\Sigma_h^{\BSM}((m_h^\BSM)^2) + \frac{t_h^\BSM}{v} \,.
-  \label{eq:Higgs_pole_mass_matching_new_BSM_singlet}
-\end{align}
-%
-On the other hand, if the \BSM Higgs is a multiplet and the $k$-th
-element is the SM-like Higgs, then the SM-like \BSM Higgs pole mass is
-the $k$-th eigenvalue of the loop-corrected mass matrix $\mathsf{M}_h$
-in the interaction eigenstate basis,
-%
-\begin{align}
-  (\mathsf{M}_h)_{ij} =
-  (m_h^\BSM)^2_{ij} - \re\Sigma_{h,ij}^{\BSM}((m_{h_k}^\BSM)^2) + \frac{t_{h_i}^\BSM}{v_i}\delta_{ij}  \,.
-  \label{eq:Higgs_pole_mass_matching_new_BSM_multiplet}
-\end{align}
-%
-
-By employing this new matching procedure, \feft consistently avoids
-large higher-order logarithms and thereby
-resums the leading and next-to-leading logarithms and includes all
-non-logarithmic 1-loop contributions.
-
-\subsection{Comparison of old and improved \feft implementations}
-
-In \figref{fig:feft_comparison} we show a comparison of the predicted
-lightest \CP-even Higgs mass in the MSSM between the old \feft
-implementation of \fs 1.7.4 (red dotted line) and the improved version
-in \fs 2.0 (red solid line).  For small SUSY scales of $\MS <
-300\unit{GeV}$, we find that the improved version still reproduces the
-fixed-order calculation.
-%
-As can be seen in the left panel of \figref{fig:feft_comparison} for
-vanishing stop mixing, $X_t = 0$, both the old and the improved
-version closely reproduce the 2-loop pure EFT calculation with
-\HSSUSY: For SUSY scales above $10\unit{TeV}$, the old version
-deviates from \HSSUSY-2L by around $600\unit{MeV}$ while the
-improved one deviates by around $10\unit{MeV}$.  This is due to the
-fact that for $X_t = 0$, the 2-loop threshold correction to the
-quartic Higgs coupling at the SUSY scale is negligible.
-%
-However, for maximal stop mixing, $X_t/\MS=\sqrt{6}$, which is the
-region where the old implementation showed the largest theoretical
-uncertainty, we find up to $3\unit{GeV}$ difference between the old
-and the improved implementation, see the right panel of
-\figref{fig:feft_comparison}.  This difference manifests the consequences of the
-different treatment of higher-order terms in the two versions,
-especially the inclusion of large 2-loop logarithms in the old
-implementation.
-
-The figure shows furthermore that the improved version (which
-performs a 1-loop calculation) is now able to perfectly reproduce the
-1-loop pure EFT calculation of \HSSUSY (blue crosses) for
-\emph{arbitrary} stop mixing and SUSY scales above $\approx
-1\unit{TeV}$.  This is in contrast to the old version, which shows a
-stronger deviation of around $2\unit{GeV}$ from the 1-loop pure EFT
-calculation for large stop mixing, see the right panel of
-\figref{fig:feft_comparison}.  Compared to the 2-loop pure EFT
-calculation of \HSSUSY (blue dashed line), both the old and the
-improved version deviate by around $1$--$2\unit{GeV}$ for non-zero
-stop mixing.  This deviation can be attributed to genuine 2-loop
-contributions.  Note that \HSSUSY does not reproduce the Higgs mass
-prediction of the fixed-order calculation for
-$\MS \lesssim 400\unit{GeV}$ in the shown scenario with $X_t = 0$,
-because of the neglected terms of $O(v^2/\MS^2)$.  In other scenarios
-the $O(v^2/\MS^2)$ terms may be important up to
-$\MS \approx 1\unit{TeV}$.
-%
-\begin{figure}[tbh]
-  \centering
-  \includegraphics[width=0.49\textwidth]{{{doc/plots/FlexibleEFTHiggs/scan_Mh_MS_TB-5_Xt-0}}} \hfill
-  \includegraphics[width=0.49\textwidth]{{{doc/plots/FlexibleEFTHiggs/scan_Mh_Xt_TB-5_MS-2000}}}
-  \caption{Comparison of the predicted lightest \CP-even Higgs pole mass in the
-    MSSM using the \feft implementations of \fs 1.7.4 and 2.0 for
-    $\tan\beta = 5$.  In the left panel we use $X_t = 0$ and in the
-    right panel $\MS = 2\unit{TeV}$.}
-  \label{fig:feft_comparison}
-\end{figure}
-
-\subsection{Choosing \feft in the model file}
-
-In order to build a \feft spectrum generator, the
-\code{FlexibleEFTHiggs} flag can be set to \code{True} in the model
-file, see \tabref{tab:feft_info}:
-%
-\begin{lstlisting}
-FlexibleEFTHiggs = True;
-\end{lstlisting}
-%
-In \feft spectrum generators, the low-energy boundary condition cannot
-be modified, because it is fixed internally to perform a matching of
-the $\SM(5)$ to the full SM\@.  The considered \BSM model is
-matched to the SM at the \BSM matching scale, $\Qmatch$,
-which is set to the \code{SUSYScale} by default.  In this matching,
-the running normalized gauge couplings $g_i^\BSM(\Qmatch)$
-($i=1,2,3$), the Yukawa coupling matrices $Y_f^\BSM(\Qmatch)$
-($f=u,d,e$) and the SM-like vacuum expectation value $v^\BSM(\Qmatch)$
-of the \BSM model are determined automatically from the following matching
-conditions on pole masses and running couplings
-%
-\begin{subequations}
-\begin{align}
-  (M_V^\BSM)^2 &= (M_V^\SM)^2, \qquad V = W, Z,
-  \label{eq:WZ-matching} \\
-  M_f^\BSM &= M_f^\SM, \qquad f = e, \mu, \tau, u, d, c, s, t, b,
-  \label{eq:f-matching} \\
-  \aem^\BSM(\Qmatch) &= \aem^\SM(\Qmatch)\times (1 + \Delta\aem),
-  \label{eq:aem-matching} \\
-  \as^\BSM(\Qmatch) &= \as^\SM(\Qmatch)\times (1 + \Delta\as),
-  \label{eq:as-matching}
-\end{align}%
-\label{eq:FSEFTHiggs_matching}%
-\end{subequations}%
-%
-where $\Delta\aem$ and $\Delta\as$ are the known 1-loop
-threshold corrections \cite{Hall:1980kf}, including potential \MSbar
-to \DRbar conversion terms \cite{Martin:1993yx}.
-\fs imposes the individual matching conditions in
-Eqs.~\eqref{eq:FSEFTHiggs_matching} at the appropriate loop orders
-such that no large 2-loop logarithms are generated, as
-described in \secref{sec:FS-2_matching_procedure}.  For example, to
-obtain the correct Higgs mass in the MSSM at the 1-loop level,
-Eqs.~\eqref{eq:WZ-matching} and \eqref{eq:aem-matching} are imposed at
-the 1-loop level while Eqs.~\eqref{eq:f-matching} and
-\eqref{eq:as-matching} are imposed at the tree level.  The SM-like
-vacuum expectation value $v^\BSM(\Qmatch)$ of the \BSM model is
-defined as
-%
-\begin{align}
-  v^\BSM(\Qmatch) = \frac{2 m_Z^\BSM(\Qmatch)}{\sqrt{\big(g_Y^\BSM(\Qmatch)\big)^2 + \big(g_2^\BSM(\Qmatch)\big)^2}},
-  \label{eq:feft_vev_definition}
-\end{align}
-%
-where $m_Z^\BSM(\Qmatch)$ is the running $Z$ boson mass and
-$g_Y^\BSM(\Qmatch)$ and $g_2^\BSM(\Qmatch)$ are the running
-electroweak gauge couplings in the \BSM model at the matching scale.
-
-The running \BSM model parameters can be given as input at either the
-\code{SUSYScale} or at the \code{HighScale}.  The following example
-demonstrates how to fix the \DRbar parameters of the MSSM at the SUSY
-scale.
-%
-\begin{example}[label=ex:MSSMEFTHiggs]
-In the \feft/MSSM model (\code{MSSMEFTHiggs})
-the soft-breaking MSSM parameters, the $\mu$ parameter and $\tan\beta$
-are input at the SUSY scale, \MS.  Thus, the boundary condition at the SUSY
-scale has the form
-%
-\begin{lstlisting}
-SUSYScaleInput = {
-    {vu, Sqrt[vu^2 + vd^2] Sin[ArcTan[TanBeta]]},
-    {vd, Sqrt[vu^2 + vd^2] Cos[ArcTan[TanBeta]]},
-    {MassB, M1Input},
-    {MassWB, M2Input},
-    {MassG, M3Input},
-    {mq2, mq2Input},
-    {mu2, mu2Input},
-    {md2, md2Input},
-    {ml2, ml2Input},
-    {me2, me2Input},
-    {\[Mu], MuInput},
-    {B[\[Mu]], mAInput^2/(TanBeta + 1/TanBeta)},
-    {T[Yu], AuInput Yu},
-    {T[Yd], AdInput Yd},
-    {T[Ye], AeInput Ye}
-};
-\end{lstlisting}
-%
-The symbols \code{TanBeta}, \code{M1Input}, \code{M2Input},
-\code{M3Input}, \code{mq2Input}, \code{mu2Input}, \code{md2Input},
-\code{ml2Input}, \code{me2Input}, \code{MuInput}, \code{mAInput},
-\code{AuInput}, \code{AdInput}, \code{AeInput} describe the MSSM input
-parameters $\tan\beta^{\DRbar}(\MS)$, $M_i(\MS)$, $m_{\tilde{f}}^2(\MS)$
-($f=q,u,d,l,e$), $\mu(\MS)$, $m_A(\MS)$ and $A_f(\MS)$
-($f=u,d,e$) in the \DRbar scheme at the SUSY scale.
-
-Note that no explicit SUSY scale boundary condition for the gauge
-couplings, $g_1$, $g_2$ and $g_3$, and Yukawa couplings, $Y_u$, $Y_d$
-and $Y_e$, of the MSSM has to be specified, because they are all fixed
-automatically at \Qmatch\ using the \feft matching conditions, Eqs.\
-\eqref{eq:FSEFTHiggs_matching}.
-
-However, there is a subtlety with the vacuum expectation values: In the
-above boundary condition, the input value $\tan\beta^{\DRbar}(\MS)$ is
-used to fix the ratio of $v_u(\MS)$ and $v_d(\MS)$.  However, their
-magnitude $\sqrt{v_u^2 + v_d^2}$ is unfixed so far.  To fix
-it, we can use the value of $v^\MSSM(\Qmatch)$, which is
-automatically determined by \fs at the matching scale, see
-Eq.~\eqref{eq:feft_vev_definition}.  Therefore, we want to set
-%
-\begin{subequations}
-\begin{align}
-  v_u(\Qmatch) &= v^\MSSM(\Qmatch) \sin\beta^{\DRbar}(\Qmatch) \,, \\
-  v_d(\Qmatch) &= v^\MSSM(\Qmatch) \cos\beta^{\DRbar}(\Qmatch) \,.
-\end{align}%
-\label{eq:VEV_matching}%
-\end{subequations}%
-%
-Such a matching is not done automatically by \feft.  The user must
-specify how the VEVs of any Higgs fields that have electroweak
-interactions are related to the electroweak VEV, $v^{\BSM}(\Qmatch)$,
-which is given above for the MSSM\@.  To do this, the model file has an
-additional constraint list: \code{MatchingScaleInput}.  Conditions
-to relate or fix model parameters at the matching scale
-$\Qmatch$ can be expressed using the \code{MatchingScaleInput} list.
-This can actually be used to set any BSM parameter or to override the automatic \feft matching conditions if the user wishes.
-However, it is only required that the user specifies the matching for the VEVs here.
-To express the relations of Eqs.~\eqref{eq:VEV_matching}, we set
-%
-\begin{lstlisting}
-MatchingScaleInput = {
-    {vu, VEV Sin[ArcTan[vu/vd]]},
-    {vd, VEV Cos[ArcTan[vu/vd]]}
-};
-\end{lstlisting}
-%
-The symbol \code{VEV} is reserved by \fs and represents the running
-SM-like vacuum expectation value $v^\MSSM(\Qmatch)$ at the matching
-scale, as defined in Eq.~\eqref{eq:feft_vev_definition}.
-\end{example}
-%
-\begin{OptionTable}{\feft model file options}{tab:feft_info}[lllX]
-  \code{FlexibleEFTHiggs} & \code{False} & \code{True} or \code{False} &
-  Flag to enable/ disable \feft \\
-  \code{VEV} & -- & -- & SM-like VEV in the \BSM model, $v^\BSM(\Qmatch)$ \\
-  \code{MatchingScaleInput} & \code{\{\}} & list of 2-tuples &
-  boundary conditions for \BSM parameters at the matching scale $\Qmatch$\\
-  \bottomrule
-\end{OptionTable}
-
-Once the model file is written, a spectrum generator can be created and
-run in the usual way.  For example, to build the model described in
-\exref{ex:MSSMEFTHiggs}, one may run:
-%
-\begin{lstlisting}[language=bash]
-$ ./createmodel --name=MSSMEFTHiggs
-$ ./configure --with-models=MSSMEFTHiggs
-$ make
-\end{lstlisting}%% $
-%
-These commands create the \fs spectrum generator for the
-\modelname{MSSMEFTHiggs} model.  The generated spectrum generator can then
-be run from the command line as
-%
-\begin{lstlisting}[language=bash]
-$ cd models/MSSMEFTHiggs
-$ ./run_MSSMEFTHiggs.x --slha-input-file=LesHouches.in.MSSMEFTHiggs
-\end{lstlisting}%% $
-%
-
-The only difference with the SLHA interface is that there are new
-\feft-specific options in the SLHA file. In \feft, the pole masses of
-the BSM particles are calculated at the scale $\QpoleBSM$, which is
-set to the \code{SUSYScale} by default.  The scale $\QpoleBSM$ can be
-changed by setting the entry \code{FlexibleSUSY[17]} to a non-zero
-value in the SLHA input file.  Similarly, in the \mathematica interface
-$\QpoleBSM$ can be changed by setting \code{poleMassScale} to a
-non-zero value.
-%
-The pole masses of the SM particles are calculated at the scale
-$\QpoleSM$, which is set to the top pole mass $M_t$ by default.  The
-scale $\QpoleSM$ can be changed by setting the entry
-\code{FlexibleSUSY[18]} to a non-zero value in the SLHA input file.
-In the \mathematica interface, $\QpoleSM$ can be changed by setting
-\code{eftPoleMassScale} to a non-zero value.
-%
-The matching scale $\Qmatch$ is set to the \code{SUSYScale} by
-default.  It can be changed by setting the entry \code{FlexibleSUSY[19]} to a
-non-zero value in the SLHA input file.  In the \mathematica interface,
-$\Qmatch$ can be changed by setting \code{eftMatchingScale} to a
-non-zero value.
-
-\begin{example}[label=ex:MSSMEFTHiggs_uncertainty_estimate]
-  This example demonstrates how a partial uncertainty estimate of the
-  lightest Higgs pole mass can be made with \feft.  The
-  uncertainty is estimated by varying the matching scale $\Qmatch$ and
-  the scale $\QpoleSM$, at which the lightest Higgs pole mass is
-  calculated, both by a factor 2.
-  %
-  \begin{lstlisting}
-Get["models/MSSMEFTHiggs/MSSMEFTHiggs_librarylink.m"];
-
-Mtpole = 173.34;
-
-(* generate logarithmically spaced range [start, stop] *)
-LogRange[start_, stop_, steps_] :=
-    Exp /@ Range[Log[start], Log[stop],
-                 (Log[stop] - Log[start])/steps];
-
-(* generate logarithmically spaced range [Q / 2, 2 Q] *)
-GenerateScales[Q_] := LogRange[Q/2, 2 Q, 10];
-
-(* run MSSMEFTHiggs spectrum generator *)
-RunMSSMEFTHiggs[MS_, TB_, Xt_, Qpole_, Qmatch_] :=
-    Module[{handle, spectrum},
-           handle = FSMSSMEFTHiggsOpenHandle[
-               fsSettings -> {
-                   precisionGoal -> 1.*^-5,
-                   maxIterations -> 10000,
-                   poleMassLoopOrder -> 2,
-                   ewsbLoopOrder -> 2,
-                   betaFunctionLoopOrder -> 3,
-                   thresholdCorrectionsLoopOrder -> 2,
-                   poleMassScale -> 0,
-                   eftPoleMassScale -> Qpole,
-                   eftMatchingScale -> Qmatch,
-                   eftMatchingLoopOrderUp -> 1,
-                   eftMatchingLoopOrderDown -> 1,
-                   calculateBSMMasses -> 0
-               },
-               fsSMParameters -> {
-                   Mt -> Mtpole
-               },
-               fsModelParameters -> {
-                   MSUSY   -> MS,
-                   M1Input -> MS,
-                   M2Input -> MS,
-                   M3Input -> MS,
-                   MuInput -> MS,
-                   mAInput -> MS,
-                   TanBeta -> TB,
-                   mq2Input -> MS^2 IdentityMatrix[3],
-                   mu2Input -> MS^2 IdentityMatrix[3],
-                   md2Input -> MS^2 IdentityMatrix[3],
-                   ml2Input -> MS^2 IdentityMatrix[3],
-                   me2Input -> MS^2 IdentityMatrix[3],
-                   AuInput -> {{MS/TB, 0    , 0},
-                               {0    , MS/TB, 0},
-                               {0    , 0    , MS/TB + Xt MS}},
-                   AdInput -> MS TB IdentityMatrix[3],
-                   AeInput -> MS TB IdentityMatrix[3]
-               }
-           ];
-           spectrum = FSMSSMEFTHiggsCalculateSpectrum[handle];
-           FSMSSMEFTHiggsCloseHandle[handle];
-           spectrum
-          ];
-
-(* extract lightest Higgs pole mass Pole[M[hh]] from spectrum *)
-RunMSSMEFTHiggsMh[pars__] :=
-    (Pole[M[hh]] /. (MSSMEFTHiggs /. RunMSSMEFTHiggs[pars]))[[1]];
-
-(* calculate Higgs mass and perform scale variation *)
-RunMSSMEFTHiggsUncertainty[MS_, TB_, Xt_] :=
-    Module[{MhMean, DMh, varyQpole, varyQmatch},
-           MhMean = RunMSSMEFTHiggsMh[MS, TB, Xt, 0, 0];
-           varyQpole  = RunMSSMEFTHiggsMh[MS, TB, Xt, #, 0]& /@
-                        GenerateScales[Mtpole];
-           varyQmatch = RunMSSMEFTHiggsMh[MS, TB, Xt, 0, #]& /@
-                        GenerateScales[MS];
-           (* combine uncertainty estimates *)
-           DMh = Max[Abs[Max[varyQpole] - MhMean],
-                     Abs[Min[varyQpole] - MhMean]] +
-                 Max[Abs[Max[varyQmatch] - MhMean],
-                     Abs[Min[varyQmatch] - MhMean]];
-           { MhMean, DMh }
-          ];
-
-{Mh, DMh} = RunMSSMEFTHiggsUncertainty[2500, 20, Sqrt[6]];
-
-Print["Mh = (", Mh, " +- ", DMh, ") GeV"];
-  \end{lstlisting}
-  The output of the script could read
-  \begin{lstlisting}
-Mh = (125.047 +- 1.52741) GeV
-  \end{lstlisting}
-\end{example}
-
-\section{Current limitations and workarounds}
-\label{sec:limitations}
-
-Currently, the models and scenarios which can be constructed with
-\fstwo are limited to the following cases:
-%
-\begin{itemize}
-\item The couplings of the model(s) must remain perturbative at all
-  scales between the highest and lowest boundary condition.
-\item The considered models are required to have a gauge symmetry
-  that has the SM gauge group
-  $\GSM = SU(3)_C\times SU(2)_L\times U(1)_Y$ as a factor.  This
-  restriction is currently necessary to perform an unambiguous
-  matching of the model to the $\SM(5)$ at the low-energy scale.  If a
-  model that does not have $\GSM$ as a gauge group
-  factor is considered, then the gauge couplings of the model must be fixed
-  by hand in one of the boundary conditions.  See
-  \secref{sec:determination_of_g_Y} for an example of a model with a
-  left-right symmetry.
-\item Tensor-like Lagrangian parameters of rank 3, which would arise
-  in $R$-parity violating SUSY models for example, are currently not
-  supported.  As a workaround, the rank 3 tensor-like couplings could
-  be decomposed into a sum of terms with rank 2 matrix-like couplings.
-\item The extraction of the running Yukawa couplings in models where
-  a 4$^\text{th}$ generation of fermions mixes with the SM fermions is currently
-  not supported.  As a workaround, the running Yukawa couplings can
-  manually be fixed by using the running SM fermion masses, which can
-  be accessed in the model file via the \code{upQuarksDRbar} =
-  $\diag(m_u, m_c, m_t)$, \code{downQuarksDRbar} =
-  $\diag(m_d, m_s, m_b)$ and \code{downLeptonsDRbar} =
-  $\diag(m_e, \allowbreak m_\mu, \allowbreak m_\tau)$ symbols.  In the
-  $\mu\nu$SSM \cite{LopezFogliani:2005yw} (\sarah/\fs model name:
-  \modelname{munuSSM}), for example, the Yukawa coupling matrix $Y_e$ of
-  the down-type leptons can approximately be fixed at the low-energy
-  scale as
-  \begin{align}
-    Y_e(Q) = \frac{\sqrt{2}}{v_d} \diag(m_e, m_\mu, m_\tau) \,,
-  \end{align}
-  which is expressed in the \fs model file as
-  \begin{lstlisting}
-LowScaleInput = {
-   {Ye, Sqrt[2] downLeptonsDRbar / vd},
-   ...
-};
-  \end{lstlisting}
-\end{itemize}
-%
-Due to the modular nature of the generated code, adaptation and extension
-to overcome restrictions in scope are quite straightforward.
-
-\section{Conclusions}
-\label{sec:conclusions}
-
-In order to study the vast zoo of models beyond the SM,
-tools for each model are necessary to calculate the mass spectrum and observables.
-\fs is a meta-tool for automatized generation of such tools which reliably operate at high precision and speed
-for a broad class of BSM models.
-
-In this paper, we have presented all of the substantial updates to \fs
-available in version 2.0.  These include many model-specific
-higher-order corrections, as well as extensions to support non-SUSY
-models (\fbsm), models with complex parameters (\fcpv) and a new
-solver which allows the EWSB
-outputs to be defined at the high scale (\fsas).  Furthermore, \fs can
-now calculate in any given
-model: the anomalous magnetic moment of the muon (\famu) as well as the muon decay and the $W$ mass (\fmw),
-including partial 2-loop contributions.  \fstwo also comes with an
-update of the hybrid EFT/fixed-order calculation of the Higgs mass
-(\feft) with a higher-order log resummation.
-
-Altogether, these represent a significant extension to the
-calculations that can be performed in BSM models at high precision. To
-illustrate the variety of potential applications of \fstwo, we have
-presented many physics examples.  These include large-scale parameter
-scans performed efficiently on multiple CPU cores (see
-\secref{sec:application_HSSUSY}), and the construction of low-energy
-effective field theories of SUSY models with complicated boundary
-conditions at the matching scale (see Sections
-\ref{sec:application_THDM} and \ref{sec:application_HSSUSY}).  Indeed,
-\fs is already being used extensively for such cases, including global
-fits by the \GAMBIT collaboration and major studies of precision Higgs
-mass predictions.  Furthermore, the modularity of the generated
-spectrum generators allows easy implementation of model-specific
-higher-order corrections, which has been done in the past to include
-3-loop Higgs mass contributions from the \Himalaya library (see
-\secref{sec:MSSM-specific_ho_corrections}) and to add power-suppressed
-terms of $O(v^2/\MS^2)$ to \HSSUSY.
-Furthermore, we have illustrated how to calculate the anomalous
-magnetic moment of the muon and electric dipole moments with \fstwo in
-Sections \ref{sec:famu}--\ref{sec:fcpv}.  Various physics applications
-for \fsas have been presented in \secref{sec:fsas}, which include the
-study of multiple solutions to the boundary value problem of the CMSSM
-as well as parameter scans in the CNMSSM and C\ESSM.
-
-The \fs system has been extensively tested for correctness against
-results from the literature and other spectrum generators.  In
-addition, speed tests have been carried out, with results proving its
-effectiveness in large-scale scans.  The auto-generated C++ code is
-designed in such a way that users can easily read and reuse
-its components to develop their own analysis tools.
-
-\section*{Acknowledgments}
-
-We especially thank Florian Staub and Mark Goodsell for many helpful
-discussions and technical help with \sarah.  Furthermore, we kindly
-thank Ben Allanach for providing components of \softsusy, which are
-included in \fs.  We also thank Pietro Slavich and Emanuele Bagnaschi
-for providing the 2-loop contributions to the Higgs mass(es) in the
-full MSSM and in \HSSUSY.
-%
-Furthermore, we thank Ben
-Allanach and Alexander Bednyakov for providing the 2-loop \SQCD
-contributions to the \DRbar top, bottom and $\tau$ Yukawa couplings as
-well as to the strong coupling in the MSSM and both Alexander
-Bednyakov and the authors of \susyhd for providing the 3-loop SM $\beta$
-functions.  We also thank the authors of the \Himalaya
-package for providing 3-loop contributions to the light \CP-even Higgs mass
-in the MSSM\@.
-%
-We kindly thank and acknowledge the role of Tom Steudtner for writing
-the first version of \feft.
-%
-A.V.\ is very grateful to the KUTS working group, especially Emanuele
-Bagnaschi, Henning Bahl, Johannes Braathen, Mark Goodsell, Robert
-Harlander, Luminita Mihaila, Pietro Slavich, Florian Staub and Georg
-Weiglein, for countless insights and helpful discussions about Higgs
-mass precision predictions in supersymmetry over the last years.
-%
-P.A.\ is very grateful to his \GAMBIT collaborators, especially Ben
-Farmer and James McKay, for helpful suggestions and feedback on
-\fs. P.A., A.V.\ and D.H.\ thank their collaborators on the
-``Precision tools and models to narrow in on the $750\unit{GeV}$
-diphoton resonance'' project for helpful comments and advice that have
-influenced the development of \fs.  Similarly, P.A.\ and A.V.\ also
-thank their collaborators in the ``Higgs mass predictions of public
-NMSSM spectrum generators'' project for the helpful discussions during
-that project.  P.A.\ and D.H.\ warmly thank Roman Nevzorov for many
-helpful discussions about the semi-analytic solver approach and
-collaboration with him which had a lot of impact on the development of
-this code.  The work of P.A.\ and D.H.\ was supported by the Australian
-Research Council through the ARC Centre of Excellence for Particle Physics
-at the Terascale (CoEPP) (grant CE110001104).  The work of P.A. is also
-supported by the Australian Research Council Future Fellowship grant
-FT160100274.  The work of D.H.\ was supported by the University of Adelaide,
-and through an Australian Government Research Training Program Scholarship.
-D.H.\ also acknowledges financial support from the Grant Agency of the Czech
-Republic (GACR), contract 17-04902S\@.  This work has also been supported by the
-German Research Foundation DFG through Grant No.~STO876/2.
-
-\appendix
-\renewcommand*{\thesection}{\Alph{section}}
-
-\section{SM input parameters}
-\label{app:sm_input_parameters}
-%
-The list of all SM input parameters is given
-in \tabref{tab:sm_mma_options}.  In the SLHA interface of \fs, the
-SM parameters can be given in four different input blocks:
-%
-\begin{itemize}
-\item The block \texttt{SMINPUTS} contains the electromagnetic and
-  strong coupling, the Fermi constant and the masses of SM
-  particles, as defined in the SLHA-2 standard \cite{Allanach:2008qq}.
-  The definitions of the individual block entries are shown in the
-  first part of \tabref{tab:sm_mma_options}.  If a parameter is
-  omitted from the \texttt{SMINPUTS} block, then it is set to the
-  default value defined in \tabref{tab:sm_mma_options}.
-\item The CKM matrix is given as input in the block \texttt{VCKMIN} in
-  the Wolfenstein parametrization, as defined in
-  Ref.~\cite{Allanach:2008qq}.  If the \texttt{VCKMIN} block or an entry is
-  missing, the corresponding parameter is set to zero.
-  The given CKM matrix elements can be accessed in the \fs model file
-  via the \texttt{CKM} symbol to set the Yukawa matrices at the
-  low-energy scale.  See \fs's \modelname{CMSSMCKM} model for an example.
-\item The PMNS matrix is given as input in the block
-  \texttt{UPMNSIN}, as defined in Ref.~\cite{Allanach:2008qq}.  If the
-  \texttt{UPMNSIN} block or an entry is missing, the corresponding
-  parameter is set to zero.
-  The given PMNS matrix elements can be accessed in the \fs model file
-  via the \texttt{PMNS} symbol to fix potential neutrino mass
-  parameters at the low-energy scale.
-\item For special applications such as the calculation of
-  $\amu$ with \GMTCalc, further input parameters are
-  needed.  These can be given in the \texttt{FlexibleSUSYInput} block.
-  The block entries are defined in \tabref{tab:sm_mma_options}.  If
-  the block or a block entry is missing, the input parameters are set
-  to their default values as defined in the table.
-\end{itemize}
-%
-In SLHA format the blocks with their respective default values read:
-%
-\begin{lstlisting}
-Block SMINPUTS               # Standard Model inputs
-    1   1.279160000e+02      # alpha^(-1) SM MSbar(MZ)
-    2   1.166378700e-05      # G_Fermi
-    3   1.184000000e-01      # alpha_s(MZ) SM MSbar
-    4   9.118760000e+01      # MZ(pole)
-    5   4.180000000e+00      # mb(mb) SM MSbar
-    6   1.733400000e+02      # mtop(pole)
-    7   1.776990000e+00      # mtau(pole)
-    8   0.000000000e+00      # mnu3(pole)
-    9   8.038500000e+01      # MW pole
-   11   5.109989020e-04      # melectron(pole)
-   12   0.000000000e+00      # mnu1(pole)
-   13   1.056583715e-01      # mmuon(pole)
-   14   0.000000000e+00      # mnu2(pole)
-   21   4.750000000e-03      # md(2 GeV) MS-bar
-   22   2.400000000e-03      # mu(2 GeV) MS-bar
-   23   1.040000000e-01      # ms(2 GeV) MS-bar
-   24   1.270000000e+00      # mc(mc) MS-bar
-Block VCKMIN                 # CKM matrix input (Wolfenstein parameters)
-    1                 0      # lambda(MZ) SM DR-bar
-    2                 0      # A(MZ) SM DR-bar
-    3                 0      # rhobar(MZ) SM DR-bar
-    4                 0      # etabar(MZ) SM DR-bar
-Block UPMNSIN                # PMNS matrix input
-    1                 0      # theta_12
-    2                 0      # theta_23
-    3                 0      # theta_13
-    4                 0      # delta
-    5                 0      # alpha_1
-    6                 0      # alpha_2
-Block FlexibleSUSYInput
-    0   0.00729735           # alpha_em(0)
-    1   125.09               # Mh pole
-\end{lstlisting}
-
-In \fs's \mathematica interface, the SM parameters must be
-passed to the \code{FS<model>OpenHandle[fsSMParameters -> \{...\}]}
-function in the form of replacement rules.  The symbols associated to the
-SM input parameters are given in
-\tabref{tab:sm_mma_options}.\footnote{Note that in \fs's \mathematica
-  interface, the fine structure constant $\aem^{\SM(5)}(M_Z)$ is input,
-  not its inverse as in the SLHA standard.}
-Unset parameters are set to their
-default values defined in the table.  Note that in the \mathematica
-interface of \fs, the CKM matrix parameters are given in the
-exact parametrization in terms of the angles $\theta_{12}$,
-$\theta_{23}$, $\theta_{13}$ and $\delta$ \cite{Olive:2016xmw}.  A
-call of \code{FS<model>OpenHandle[]} with all parameters set to their
-respective default values would read:
-%
-\begin{lstlisting}{Mathematica}
-handle = FS<model>OpenHandle[
-    fsSMParameters -> {
-        alphaEmMZ -> 1/127.916,      (* SMINPUTS[1] *)
-        GF -> 1.166378700*^-5,       (* SMINPUTS[2] *)
-        alphaSMZ -> 0.1184,          (* SMINPUTS[3] *)
-        MZ -> 91.1876,               (* SMINPUTS[4] *)
-        mbmb -> 4.18,                (* SMINPUTS[5] *)
-        Mt -> 173.34,                (* SMINPUTS[6] *)
-        Mtau -> 1.77699,             (* SMINPUTS[7] *)
-        Mv3 -> 0,                    (* SMINPUTS[8] *)
-        MW -> 80.385,                (* SMINPUTS[9] *)
-        Me -> 0.000510998902,        (* SMINPUTS[11] *)
-        Mv1 -> 0,                    (* SMINPUTS[12] *)
-        Mm -> 0.1056583715,          (* SMINPUTS[13] *)
-        Mv2 -> 0,                    (* SMINPUTS[14] *)
-        md2GeV -> 0.00475,           (* SMINPUTS[21] *)
-        mu2GeV -> 0.0024,            (* SMINPUTS[22] *)
-        ms2GeV -> 0.104,             (* SMINPUTS[23] *)
-        mcmc -> 1.27,                (* SMINPUTS[24] *)
-        CKMTheta12 -> 0,
-        CKMTheta23 -> 0,
-        CKMTheta13 -> 0,
-        CKMDelta -> 0,
-        PMNSTheta12 -> 0,            (* UPMNSIN[1] *)
-        PMNSTheta23 -> 0,            (* UPMNSIN[2] *)
-        PMNSTheta13 -> 0,            (* UPMNSIN[3] *)
-        PMNSDelta -> 0,              (* UPMNSIN[4] *)
-        PMNSAlpha1 -> 0,             (* UPMNSIN[5] *)
-        PMNSAlpha2 -> 0,             (* UPMNSIN[6] *)
-        alphaEm0 -> 1/137.035999074, (* FlexibleSUSYInput[0] *)
-        Mh -> 125.09                 (* FlexibleSUSYInput[1] *)
-    }
-]
-\end{lstlisting}
-%
-\begin{table}[tbh!]
-  \centering
-  \resizebox{\textwidth}{!}{%
-  \begin{tabular}{llll}
-    \toprule
-    Index & \mathematica symbol & Default & Description \\
-    \midrule
-    \multicolumn{4}{c}{\texttt{Block SMINPUTS}}\\
-    \midrule
-     1 & \code{alphaEmMZ} & $1/127.916$ & electromagnetic coupling, $\aem^{\SM(5)}(M_Z)$ \\
-     2 & \code{GF} & $1.1663787\cdot 10^{-5}$ & Fermi coupling constant, $G_F\times\text{GeV}^2$ \\
-     3 & \code{alphaSMZ} & $0.1184$ & strong coupling, $\as^{\SM(5)}(M_Z)$ \\
-     4 & \code{MZ} & $91.1876$ & $Z$ pole mass, $M_Z/\text{GeV}$ \\
-     5 & \code{mbmb} & $4.18$ & running bottom mass, $m_b^{\SM(5)}(m_b)/\text{GeV}$ \\
-     6 & \code{Mt} & $173.34$ & top pole mass, $M_t/\text{GeV}$ \\
-     7 & \code{Mtau} & $1.77699$ & $\tau$ pole mass, $M_\tau/\text{GeV}$ \\
-     8 & \code{Mv3} & $0$ & heaviest neutrino pole mass, $M_{\nu_3}/\text{GeV}$ \\
-     9 & \code{MW} & $80.385$ & $W$ pole mass, $M_W/\text{GeV}$ \\
-    11 & \code{Me} & $0.000510998902$ & electron pole mass, $M_e/\text{GeV}$ \\
-    12 & \code{Mv1} & $0$ & lightest neutrino pole mass, $M_{\nu_1}/\text{GeV}$ \\
-    13 & \code{Mm} & $0.1056583715$ & muon pole mass, $M_\mu/\text{GeV}$ \\
-    14 & \code{Mv2} & $0$ & 2$^\text{nd}$ lightest neutrino pole mass, $M_{\nu_2}/\text{GeV}$ \\
-    21 & \code{md2GeV} & $0.00475$ & running down mass, $m_d(2\unit{GeV})/\text{GeV}$ \\
-    22 & \code{mu2GeV} & $0.0024$ & running up mass, $m_u(2\unit{GeV})/\text{GeV}$ \\
-    23 & \code{ms2GeV} & $0.104$ & running strange mass, $m_s(2\unit{GeV})/\text{GeV}$ \\
-    24 & \code{mcmc} & $1.27$ & running charm mass, $m_c^{\SM(4)}(m_c)/\text{GeV}$ \\
-    \midrule
-    \multicolumn{4}{c}{\texttt{Block VCKMIN}}\\
-    \midrule
-     1 & & $0$ & CKM Wolfenstein parameter $\lambda$ \\
-     2 & & $0$ & CKM Wolfenstein parameter $A$ \\
-     3 & & $0$ & CKM Wolfenstein parameter $\bar{\rho}$ \\
-     4 & & $0$ & CKM Wolfenstein parameter $\eta$ \\
-    \midrule
-       & \code{CKMTheta12} & $0$ & CKM matrix parameter $\theta_{12}$ \\
-       & \code{CKMTheta23} & $0$ & CKM matrix parameter $\theta_{23}$ \\
-       & \code{CKMTheta13} & $0$ & CKM matrix parameter $\theta_{13}$ \\
-       & \code{CKMDelta} & $0$ & CKM matrix parameter $\delta$ \\
-    \midrule
-    \multicolumn{4}{c}{\texttt{Block UPMNSIN}}\\
-    \midrule
-     1 & \code{PMNSTheta12} & $0$ & PMNS solar angle $\theta_{12}$ \\
-     2 & \code{PMNSTheta23} & $0$ & PMNS atmospheric angle $\theta_{23}$ \\
-     3 & \code{PMNSTheta13} & $0$ & PMNS matrix parameter $\theta_{13}$ \\
-     4 & \code{PMNSDelta}   & $0$ & PMNS Dirac phase $\delta$ \\
-     5 & \code{PMNSAlpha1}  & $0$ & PMNS 1$^\text{st}$ Majorana phase $\alpha_1$ \\
-     6 & \code{PMNSAlpha2}  & $0$ & PMNS 2$^\text{nd}$ Majorana phase $\alpha_2$ \\
-    \midrule
-    \multicolumn{4}{c}{\texttt{Block FlexibleSUSYInput}}\\
-    \midrule
-     0 & \code{alphaEm0} & $1/137.035999074$ & $\aem$ in the Thomson limit \\
-     1 & \code{Mh} & $125.09$ & SM Higgs pole mass $M_h/\text{GeV}$ \\
-    \bottomrule
-  \end{tabular}}
-  \caption{SLHA input block entries and \mathematica symbols to
-    specify the SM input parameters.  The first column
-    represents the index in the corresponding SLHA input block and the
-    second column the symbol used in the \mathematica interface.}
-  \label{tab:sm_mma_options}
-\end{table}
-
-\section{FlexibleSUSY configuration options}
-\label{app:fs_config_options}
-
-\fs provides many configuration options to switch on/off contributions
-and choose/fine-tune the solver algorithm(s).  All runtime
-configuration options are listed in \tabref{tab:sg_mma_options}.
-%
-In the SLHA interface of \fs, the configuration options are read from
-the \fs block.  In addition, some information is also read
-from the \texttt{MODSEL} block, see below.  In the SLHA format all \fs
-configuration entries with their respective default values read:
-%
-\begin{lstlisting}
-Block MODSEL
-   12   0            # output scale of running parameters (0 = SUSY scale)
-Block FlexibleSUSY
-    0   1e-04        # precision goal
-    1   0            # max. iterations (0 = automatic)
-    2   0            # solver (0 = all, 1 = two-scale, 2 = semi-analytic)
-    3   0            # calculate SM pole masses
-    4   2            # pole mass loop order
-    5   2            # EWSB loop order
-    6   3            # beta-functions loop order
-    7   2            # threshold corrections loop order
-    8   1            # Higgs 2L corrections O(alpha_t alpha_s)
-    9   1            # Higgs 2L corrections O(alpha_b alpha_s)
-   10   1            # Higgs 2L corrections O((alpha_t + alpha_b)^2)
-   11   1            # Higgs 2L corrections O(alpha_tau^2)
-   12   0            # force output
-   13   1            # Top quark 2L corrections QCD
-   14   1e-11        # beta-function zero threshold
-   15   0            # calculate observables (a_muon, ...)
-   16   0            # force positive majorana masses
-   17   0            # pole mass renormalization scale (0 = SUSY scale)
-   18   0            # pole mass renormalization scale in the EFT
-                     # (0 = min(SUSY scale, Mt))
-   19   0            # EFT matching scale (0 = SUSY scale)
-   20   2            # EFT loop order for upwards matching
-   21   1            # EFT loop order for downwards matching
-   22   0            # EFT index of SM-like Higgs in the BSM model
-   23   1            # calculate BSM pole masses
-   24   123111321    # individual threshold correction loop orders
-   25   0            # ren. scheme for Higgs 3L corrections
-                     # (0 = DR, 1 = MDR)
-   26   1            # Higgs 3L corrections O(alpha_t alpha_s^2)
-   27   1            # Higgs 3L corrections O(alpha_b alpha_s^2)
-   28   1            # Higgs 3L corrections O(alpha_t^2 alpha_s)
-   29   1            # Higgs 3L corrections O(alpha_t^3)
-\end{lstlisting}
-%
-In the \mathematica interface of \fs, the configuration options are
-passed to the function \code{FS<model>OpenHandle[fsSettings ->
-  \{...\}]} in form of replacement rules.  The symbols associated to
-the configuration options are given in \tabref{tab:sg_mma_options}.
-Unset options are set to their default values defined in the table.
-A call of \code{FS<model>OpenHandle[]} with all configuration options
-set to their default values would read:
-%
-\begin{lstlisting}
-handle = FS<model>OpenHandle[
-    fsSettings -> {
-        precisionGoal -> 1.*^-4,           (* FlexibleSUSY[0] *)
-        maxIterations -> 0,                (* FlexibleSUSY[1] *)
-        solver -> 0,                       (* FlexibleSUSY[2] *)
-        calculateStandardModelMasses -> 0, (* FlexibleSUSY[3] *)
-        poleMassLoopOrder -> 2,            (* FlexibleSUSY[4] *)
-        ewsbLoopOrder -> 2,                (* FlexibleSUSY[5] *)
-        betaFunctionLoopOrder -> 3,        (* FlexibleSUSY[6] *)
-        thresholdCorrectionsLoopOrder -> 2,(* FlexibleSUSY[7] *)
-        higgs2loopCorrectionAtAs -> 1,     (* FlexibleSUSY[8] *)
-        higgs2loopCorrectionAbAs -> 1,     (* FlexibleSUSY[9] *)
-        higgs2loopCorrectionAtAt -> 1,     (* FlexibleSUSY[10] *)
-        higgs2loopCorrectionAtauAtau -> 1, (* FlexibleSUSY[11] *)
-        forceOutput -> 0,                  (* FlexibleSUSY[12] *)
-        topPoleQCDCorrections -> 1,        (* FlexibleSUSY[13] *)
-        betaZeroThreshold -> 1.*^-11,      (* FlexibleSUSY[14] *)
-        forcePositiveMasses -> 0,          (* FlexibleSUSY[16] *)
-        poleMassScale -> 0,                (* FlexibleSUSY[17] *)
-        eftPoleMassScale -> 0,             (* FlexibleSUSY[18] *)
-        eftMatchingScale -> 0,             (* FlexibleSUSY[19] *)
-        eftMatchingLoopOrderUp -> 2,       (* FlexibleSUSY[20] *)
-        eftMatchingLoopOrderDown -> 1,     (* FlexibleSUSY[21] *)
-        eftHiggsIndex -> 0,                (* FlexibleSUSY[22] *)
-        calculateBSMMasses -> 1,           (* FlexibleSUSY[23] *)
-        thresholdCorrections -> 123111321, (* FlexibleSUSY[24] *)
-        higgs3loopCorrectionRenScheme -> 0,(* FlexibleSUSY[25] *)
-        higgs3loopCorrectionAtAsAs -> 1,   (* FlexibleSUSY[26] *)
-        higgs3loopCorrectionAbAsAs -> 1,   (* FlexibleSUSY[27] *)
-        higgs3loopCorrectionAtAtAs -> 1,   (* FlexibleSUSY[28] *)
-        higgs3loopCorrectionAtAtAt -> 1,   (* FlexibleSUSY[29] *)
-        parameterOutputScale -> 0          (* MODSEL[12] *)
-    }
-]
-\end{lstlisting}
-%
-\begin{table}
-  \centering
-  \resizebox{\textwidth}{!}{%
-  \begin{tabularx}{1.2\textwidth}{lllX}
-    \toprule
-    Index & \mathematica symbol & Default & Description \\
-    \midrule
-    \multicolumn{4}{c}{\texttt{Block FlexibleSUSY}}\\
-    \midrule
-     0 & \code{precisionGoal}                 & \code{1.*^-4} &
-    precision goal of RG running and mass spectrum\\
-     1 & \code{maxIterations}                 & \code{0} &
-    maximum number of iterations for the running between the scales ($0$ = automatic)\\
-     2 & \code{solver}                        & \code{0} &
-    BVP solver ($0$ = all, $1$ = two-scale solver, $2$ = semi-analytic solver)\\
-     3 & \code{calculateStandardModelMasses}  & \code{0} &
-    switch to enable/disable calculation of pole masses of SM particles ($0$ = disabled)\\
-     4 & \code{poleMassLoopOrder}             & \code{2} &
-    pole mass loop order\\
-     5 & \code{ewsbLoopOrder}                 & \code{2} &
-    EWSB loop order (should be set equal to the pole mass loop order)\\
-     6 & \code{betaFunctionLoopOrder}         & \code{3} &
-    loop order for renormalization group running\\
-     7 & \code{thresholdCorrectionsLoopOrder} & \code{2} &
-    global switch for loop order of threshold corrections when
-    converting the $\SM(5)$ parameters to the BSM parameters\\
-     8 & \code{higgs2loopCorrectionAtAs}      & \code{1} &
-    enable/disable 2-loop corrections $O(\at\as)$ to $M_{h,H,A}$\\
-     9 & \code{higgs2loopCorrectionAbAs}      & \code{1} &
-    enable/disable 2-loop corrections $O(\ab\as)$ to $M_{h,H,A}$\\
-    10 & \code{higgs2loopCorrectionAtAt}      & \code{1} &
-    enable/disable 2-loop corrections $O(\at^2)$ to $M_{h,H,A}$\\
-    11 & \code{higgs2loopCorrectionAtauAtau}  & \code{1} &
-    enable/disable 2-loop corrections $O(\atau^2)$ to $M_{h,H,A}$\\
-    12 & \code{forceOutput}                   & \code{0} &
-    force output, even if problems occurred\\
-    13 & \code{topPoleQCDCorrections}         & \code{1} &
-    QCD corrections to calculate $M_t$ (0 = 1-loop, 1 = 2-loop, 2 = 3-loop)\\
-    14 & \code{betaZeroThreshold}             & \code{1.*^-11} &
-    below this threshold $\beta$ functions are treated as zero\\
-    15 & & \code{0} & enable/disable calculation of observables\\
-    16 & \code{forcePositiveMasses}           & \code{0} &
-    make Majorana masses positive (violates SLHA)\\
-    17 & \code{poleMassScale}                 & \code{0} &
-    scale at which pole masses are calculated (0 = SUSY scale)\\
-    18 & \code{eftPoleMassScale}              & \code{0} &
-    scale at which SM pole masses are calculated in \feft (0 = $M_t$)\\
-    19 & \code{eftMatchingScale}              & \code{0} &
-    matching scale in \feft (0 = SUSY scale)\\
-    20 & \code{eftMatchingLoopOrderUp}        & -- &
-    ignored\\
-    21 & \code{eftMatchingLoopOrderDown}      & \code{1} &
-    loop order for $\Delta\lambda$ in \feft\\
-    22 & \code{eftHiggsIndex}                 & \code{0} &
-    index of SM-like Higgs in BSM Higgs multiplet in \feft\\
-    23 & \code{calculateBSMMasses}            & \code{1} &
-    enable/disable calculation of BSM pole masses\\
-    24 & \code{thresholdCorrections}          & \code{123111321} &
-    individual threshold correction loop orders, see \tabref{tab:sg_mma_tc}\\
-    25 & \code{higgs3loopCorrectionRenScheme} & \code{0} &
-    renormalization scheme for 3-loop MSSM Higgs corrections
-    (0 = \DRbar, 1 = \MDRbar)\\
-    26 & \code{higgs3loopCorrectionAtAsAs}    & \code{1} &
-    enable/disable 3-loop corrections $O(\at\as^2)$ to $M_{h,H,A}$\\
-    27 & \code{higgs3loopCorrectionAbAsAs}    & \code{1} &
-    enable/disable 3-loop corrections $O(\ab\as^2)$ to $M_{h,H,A}$\\
-    28 & \code{higgs3loopCorrectionAtAtAs}    & \code{1} &
-    enable/disable 3-loop corrections $O(\at^2\as)$ to $M_{h}$\\
-    29 & \code{higgs3loopCorrectionAtAtAt}    & \code{1} &
-    enable/disable 3-loop corrections $O(\at^3)$ to $M_{h}$\\
-    \midrule
-    \multicolumn{4}{c}{\texttt{Block MODSEL}}\\
-    \midrule
-    12 & \code{parameterOutputScale}          & \code{0} &
-    output scale for running parameters (0 = SUSY scale)\\
-    \bottomrule
-  \end{tabularx}}
-  \caption{SLHA input block entries and corresponding \mathematica
-    symbols to specify the configuration options for \fs's spectrum
-    generators.  The symbols $M_{h,H,A}$ and $M_t$ denote the Higgs
-    and top quark pole masses, respectively.}
-  \label{tab:sg_mma_options}
-\end{table}
-%
-\begin{table}[tbh]
-  \centering
-  \begin{tabular}{lll}
-    \toprule
-    digit position $n$    & default value & parameter \\
-    (from the right) & (prefactor of $10^n$) & \\
-    \midrule
-    $0$ & $1$ (1-loop) & $\alpha^\BSM_{\text{em}}$ \\
-    $1$ & $2$ (2-loop) & $\sin^\BSM(\theta_W)$ \\
-    $2$ & $3$ (3-loop) & $\as^\BSM$ \\
-    $3$ & $1$ (1-loop) & $m^\BSM_Z$ \\
-    $4$ & $1$ (1-loop) & $m^\BSM_W$ \\
-    $5$ & $1$ (1-loop) & $m^\BSM_h$ \\
-    $6$ & $3$ (3-loop) & $m^\BSM_t$ \\
-    $7$ & $2$ (2-loop) & $m^\BSM_b$ \\
-    $8$ & $1$ (1-loop) & $m^\BSM_\tau$ \\
-    \bottomrule
-  \end{tabular}
-  \caption{Specification of individual loop orders of threshold
-    corrections for extracting the running masses, couplings and
-    Weinberg angle in the BSM model at the low-energy scale using the
-    field \texttt{FlexibleSUSY[24]} in the SLHA interface or the
-    symbol \texttt{thresholdCorrections} in the \mathematica
-    interface, respectively.  The digit position is counted from the
-    right, starting at $0$.  Setting all loop orders to their default
-    values results in the integer $123111321$.}
-  \label{tab:sg_mma_tc}
-\end{table}
-%
-The individual configuration options have the following meaning:
-%
-\begin{itemize}
-\setlength{\itemindent}{1cm}
-\item[\texttt{FlexibleSUSY[0]},] \texttt{precisionGoal}: This option
-  describes the numeric precision of the renormalization group
-  running, the mass spectrum calculation, the electroweak symmetry
-  breaking and the calculation of the observables.  For most models a
-  precision of $10^{-4}$ is sufficient.  For models with various
-  3-loop corrections, like \HSSUSY or MSSM-like models, a
-  precision of $10^{-5}$ might be better.
-
-\item[\texttt{FlexibleSUSY[1]},] \texttt{maxIterations}: This option
-  describes the maximum number of iterations for the renormalization
-  group running between the various scales.  If it is set to $0$, then
-  the maximum number of iterations $N_{\text{max,it}}$ is chosen
-  according to the precision goal $p$ (see above) as
-  \begin{align}
-    N_{\text{max,it}} = - 10\log_{10} p \,.
-  \end{align}
-
-\item[\texttt{FlexibleSUSY[2]},] \texttt{solver}: This option chooses
-  the BVP solver to be used.  If set to $0$, all
-  solvers that have been enabled in the model file (see
-  \secref{sec:chosing_SAS}) are used.  In this case, each of the
-  enabled solvers will be tried in turn, in the same order as given in
-  \code{FSBVPSolvers}, until a solution is found, at which point \fs
-  will return this solution and no further solvers are tried.  In the
-  event that no solver obtains a valid solution, \fs reports the status of the
-  last solver tried.  Non-zero values of \texttt{FlexibleSUSY[2]}
-  select a single solver to be used.  If set to $1$, the two-scale
-  solver is used if it has been enabled in the model file.  If set
-  to $2$, then the semi-analytic solver is used if it has been enabled
-  in the model file.  If a solver that has not been enabled in the model
-  file is chosen, \fs stops with an error.
-
-\item[\texttt{FlexibleSUSY[3]},]
-  \texttt{calculateStandardModelMasses}: This option allows the user
-  to enable/disable the calculation of the pole masses of the SM
-  particles.  Note that this switch must be set to $1$ in
-  \HSSUSY to calculate the Higgs pole mass, because in \HSSUSY the
-  Higgs pole mass is calculated in the SM\@.
-
-\item[\texttt{FlexibleSUSY[4]},] \texttt{poleMassLoopOrder}: This
-  option allows the user to select the loop order at which the pole
-  masses are calculated.  If set to $0$, the running tree-level masses
-  are output.  If set to $1$, the pole masses are calculated at the
-  1-loop level.  If set to $2$ or $3$, then model-specific 2-loop or
-  3-loop corrections are taken into account, respectively, if they have been
-  enabled in the model file (see
-  \secref{sec:model_specific_contributions}).  \textbf{Important
-    note:} In order to obtain a consistent pole mass spectrum, the
-  loop order of the electroweak symmetry breaking (see
-  \texttt{FlexibleSUSY[5]}, \texttt{ewsbLoopOrder}) must be set to the
-  \emph{same} value as the pole mass loop order!
-
-\item[\texttt{FlexibleSUSY[5]},] \texttt{ewsbLoopOrder}: This option
-  allows the user to select the loop order at which the electroweak
-  symmetry breaking (EWSB) equations are solved.  If set to $0$, the
-  EWSB equations are solved at the tree level.  If set to $1$, the
-  EWSB equations are solved at the 1-loop level.  If set to $2$ or
-  $3$, then model-specific 2-loop or 3-loop corrections are taken into
-  account, respectively, if they have been enabled in the model file (see
-  \secref{sec:model_specific_contributions}).  \textbf{Important
-    note:} In order to obtain a consistent pole mass spectrum, the
-  loop order of the electroweak symmetry breaking must be set to the
-  \emph{same} value as the pole mass loop order (see
-  \texttt{FlexibleSUSY[4]}, \texttt{poleMassLoopOrder})!
-
-\item[\texttt{FlexibleSUSY[6]},] \texttt{betaFunctionLoopOrder}: With
-  this option the user can select the loop level of the $\beta$
-  functions used to integrate the RGEs.  If set to $1$, 1-loop $\beta$
-  functions are used.  If set to $2$, 2-loop $\beta$ functions are used.  If
-  set to $3$, then model-specific 3-loop $\beta$ functions are used (see
-  \secref{sec:model_specific_contributions}).  Note that \sarah can
-  generate 2-loop $\beta$ functions for all model parameters (except
-  scalar tadpole terms), so 2-loop running can always be used.
-
-\item[\texttt{FlexibleSUSY[7]},]
-  \texttt{thresholdCorrectionsLoopOrder}: With this option the user
-  can choose the \emph{maximum} loop level of the threshold corrections
-  used to determine the running gauge couplings $g_1$, $g_2$, $g_3$
-  and the running Yukawa coupling matrices $Y_u$, $Y_d$, $Y_e$ of the
-  \BSM model at the low-energy scale (\code{LowScale}) from the given
-  SM input parameters ($\aem^{\SM(5)}(M_Z)$,
-  $\as^{\SM(5)}(M_Z)$, $G_F$, $M_Z$, $M_t$, $m_b^{\SM(5)}(m_b)$,
-  $m_c^{\SM(4)}(m_c)$, \ldots).  See \secref{sec:determination_of_g_Y}
-  and \ref{sec:model_specific_contributions} for a description on how
-  the running gauge and Yukawa couplings are calculated and how
-  model-specific higher-order corrections can be included.  If the
-  threshold corrections loop order is set to $1$, then no 2-loop
-  threshold corrections or higher are taken into account.  If set to
-  $2$, then no 3-loop threshold corrections or higher are taken into
-  account.  If set to $3$, then no 4-loop threshold corrections are
-  taken into account.
-  %
-  Note that threshold corrections for individual parameters can be
-  disabled by using \texttt{FlexibleSUSY[24]} or
-  \texttt{thresholdCorrections}, respectively.
-
-\item[\texttt{FlexibleSUSY[8]},] \texttt{higgs2loopCorrectionAtAs}:
-  With this option the 2-loop contributions to the Higgs pole mass(es)
-  of $O(\at\as)$ can be enabled/disabled.  Note that this
-  option has an effect only if 2-loop contributions have been
-  activated in the \fs model file.  See
-  \secref{sec:model_specific_contributions} for details on how to
-  activate 2-loop contributions to the Higgs pole mass(es) of
-  $O(\at\as)$ in the SM, (N)MSSM or split-MSSM\@.
-
-\item[\texttt{FlexibleSUSY[9]},] \texttt{higgs2loopCorrectionAbAs}:
-  With this option the 2-loop contributions to the Higgs pole mass(es)
-  of $O(\ab\as)$ can be enabled/disabled.  Note that this
-  option has an effect only if 2-loop contributions have been
-  activated in the \fs model file.  See
-  \secref{sec:model_specific_contributions} for details on how to
-  activate 2-loop contributions of $O(\ab\as)$ to the Higgs pole
-  mass(es) in the (N)MSSM\@.
-
-\item[\texttt{FlexibleSUSY[10]},] \texttt{higgs2loopCorrectionAtAt}:
-  With this option the 2-loop contributions to the Higgs pole mass(es)
-  of $O(\at^2)$ or $O((\at+\ab)^2)$ can be enabled/disabled.
-  Note that this option has an effect only if 2-loop contributions
-  have been activated in the \fs model file.  See
-  \secref{sec:model_specific_contributions} for details on how to
-  activate 2-loop contributions of these orders to the Higgs pole
-  mass(es) in the SM or (N)MSSM\@.
-
-\item[\texttt{FlexibleSUSY[11]},]
-  \texttt{higgs2loopCorrectionAtauAtau}: With this option the 2-loop
-  contributions to the Higgs pole mass(es) of $O(\atau^2)$
-  can be enabled/disabled.  Note that this option has an effect only
-  if 2-loop contributions have been activated in the \fs model file.
-  See \secref{sec:model_specific_contributions} for details on how to
-  activate 2-loop contributions of $O(\atau^2)$ to the Higgs pole
-  mass(es) in the (N)MSSM\@.
-
-\item[\texttt{FlexibleSUSY[12]},] \texttt{forceOutput}: This option
-  allows the user to force an output of \fs, even if a physical
-  problem has occurred (tachyon, non-perturbative parameter, no EWSB,
-  \ldots).  If set to $0$, \fs does not give an output if a problem
-  has occurred.  If set to $1$, an output is always given, even if a
-  problem has occurred.  Please be very careful and check for
-  potential warnings/problems when forcing the output!
-
-\item[\texttt{FlexibleSUSY[13]},] \texttt{topPoleQCDCorrections}: With
-  this option the user can enable additional loop contributions when
-  the top quark \emph{pole mass} is re-calculated.  Note that the top pole mass
-  is only re-calculated if \texttt{FlexibleSUSY[3]} or
-  \texttt{calculateStandardModelMasses} is set to $1$.  If set to
-  $0$, then 1-loop (SUSY-)QCD contributions are taken into account (but
-  only if \texttt{FlexibleSUSY[4]} or \texttt{poleMassLoopOrder} is
-  set to $1$).  If set to $1$, then 2-loop (SUSY-)QCD contributions are
-  taken into account (but only if \texttt{FlexibleSUSY[4]} or
-  \texttt{poleMassLoopOrder} is set to $2$).  If set to $2$, then
-  3-loop (SUSY-)QCD contributions are taken into account (but only if
-  \texttt{FlexibleSUSY[4]} or \texttt{poleMassLoopOrder} is set to
-  $3$).
-
-\item[\texttt{FlexibleSUSY[14]},] \texttt{betaZeroThreshold}: With
-  this option a numerical threshold can be defined below which a
-  $\beta$ function is treated as being exactly zero.  A small but
-  non-zero threshold can avoid numerical problems when integrating the
-  RGEs.
-
-\item[\texttt{FlexibleSUSY[15]}:] With this option the calculation of
-  the observables ($\amu$, EDMs, effective couplings of
-  $h\rightarrow\gamma\gamma$ and $h\rightarrow gg$) can be
-  enabled/disabled.  Note that in the \mathematica interface, the
-  observables are calculated by the function
-  \code{FS<model>CalculateObservables[]}, see
-  \secref{sec:Mathematica_interface}.
-
-\item[\texttt{FlexibleSUSY[16]},] \texttt{forcePositiveMasses}: With
-  this option the masses of Majorana fermions can be forced to be
-  positive in the SLHA output of \fs.  If set to $1$, then Majorana
-  masses are always positive, but the corresponding mixing matrices
-  are in general complex (note that this violates the SLHA
-  convention).  If set to $0$, then the Majorana masses can be
-  positive or negative, but the corresponding mixing matrices are
-  guaranteed to be real (SLHA convention).
-
-\item[\texttt{FlexibleSUSY[17]},] \texttt{poleMassScale}: With this
-  option the user can choose the scale (in GeV) at which the pole mass
-  spectrum is calculated.  If set to $0$, then the value assigned to
-  the \code{SUSYScale} variable in the model file is used.  In \feft,
-  the pole mass scale is defined to be the scale at which the pole
-  masses in the \emph{full} \BSM model are calculated.  To vary the
-  scale at which the pole masses in the effective theory (the
-  SM) are calculated, use \texttt{FlexibleSUSY[18]} or
-  \texttt{eftPoleMassScale} in the SLHA or \mathematica interface,
-  respectively.
-
-\item[\texttt{FlexibleSUSY[18]},] \texttt{eftPoleMassScale}: This
-  option applies only to \feft models.  With this option the user can
-  choose the scale (in GeV) at which the pole masses in the effective
-  field theory (i.e., in the SM) are calculated.  If the
-  scale is set to $0$, then $Q = M_t$ is used.  This option can be
-  used to estimate a partial uncertainty of the Higgs mass
-  prediction in \feft by varying the pole mass scale around $Q = M_t$,
-  see \exref{ex:MSSMEFTHiggs_uncertainty_estimate}.
-
-\item[\texttt{FlexibleSUSY[19]},] \texttt{eftMatchingScale}: This
-  option applies only to \feft models.  With this option the user can
-  specify the scale at which the matching of the \BSM model to the
-  effective theory (i.e., the SM) is performed.  If the
-  scale is set to $0$, then the value assigned to the \code{SUSYScale}
-  variable in the \fs model file is used.  This option can be used to
-  estimate a partial uncertainty of the Higgs mass prediction
-  in \feft by varying the matching scale around $Q = \MS$, see
-  \exref{ex:MSSMEFTHiggs_uncertainty_estimate}.
-
-\item[\texttt{FlexibleSUSY[20]},] \texttt{eftMatchingLoopOrderUp}:
-  This option is ignored in \fstwo.
-
-\item[\texttt{FlexibleSUSY[21]},] \texttt{eftMatchingLoopOrderDown}:
-  This option applies only to \feft models.  With this option the user
-  can select the loop order at which the quartic Higgs coupling
-  $\lambda$ of the SM is fixed when matching the \BSM
-  model to the SM in \feft.  If set to $0$, then $\lambda$
-  is fixed using only tree-level matching.  If set to $1$
-  (recommended), then $\lambda$ is fixed by a 1-loop matching
-  condition.
-
-\item[\texttt{FlexibleSUSY[22]},] \texttt{eftHiggsIndex}: This option
-  applies only to \feft models.  With this option the user can choose
-  which field in the Higgs multiplet of the \BSM model corresponds to
-  the SM-like Higgs.  If set to $0$, the lightest field in the Higgs
-  multiplet is interpreted as SM-like Higgs.  If set to $1$, the
-  2$^{\text{nd}}$ lightest field is interpreted as SM-like Higgs, etc.
-  The chosen field is then used in the matching condition
-  $M_h^{\SM} = M_{h_i}^{\BSM}$, where $i$ is the index of the chosen
-  field in the Higgs multiplet ($i=0,1,2,\ldots$).
-
-\item[\texttt{FlexibleSUSY[23]},] \texttt{calculateBSMMasses}: This
-  option allows the user to enable/disable the calculation of the pole
-  masses of the \BSM particles.  If set to $0$, then the \BSM pole
-  masses are not calculated.  If set to $1$, then the \BSM pole masses
-  are calculated.  This option is useful in \feft for example: If one
-  is only interested in the prediction of the SM-like Higgs pole mass,
-  then this option can be set to $0$ to suppress the calculation of
-  the masses of the heavy Higgs bosons, the charginos, neutralinos and sfermions.
-
-\item[\texttt{FlexibleSUSY[24]},] \texttt{thresholdCorrections}: With
-  this option the user has a finer control over the threshold
-  corrections to the individual model parameters.  The value assigned
-  to this option is an integer number, where each digit (with respect to base
-  $10$) represents the threshold correction loop order for a
-  particular running \BSM parameter.  The association between the
-  digits and the parameters as well as the default loop orders are
-  shown in \tabref{tab:sg_mma_tc}.
-  %
-  \begin{example}
-    The following table shows example values for the integer number
-    which specifies the individual threshold correction loop orders,
-    together with the list of included loop corrections.
-    \begin{center}
-    \begin{tabular}{rl}
-      \toprule
-      integer & used threshold corrections \\
-      \midrule
-      0       & no threshold corrections, everything at tree level \\
-      1       & only $\Delta\aem^{1L}$, everything else at tree level \\
-      100     & only $\Delta\as^{1L}$, everything else at tree level \\
-      101     & only $\Delta\aem^{1L}$ and  $\Delta\as^{1L}$, everything else at tree level \\
-      3000101 & only $\Delta\aem^{1L}$ and  $\Delta\as^{1L}$ and  $\Delta y_t^{3L}$, everything else at tree level \\
-      \bottomrule
-    \end{tabular}
-  \end{center}
-  \end{example}
-  %
-  If the field \texttt{FlexibleSUSY[24]} or the symbol
-  \texttt{thresholdCorrections} is omitted, then the whole option is
-  set to the default value given in \tabref{tab:sg_mma_tc}.  If the
-  field is not omitted, then all loop orders must be given.  Note
-  that setting the loop orders larger than the value set in
-  \texttt{FlexibleSUSY[7]} or \texttt{thresholdCorrectionsLoopOrder}
-  has no effect, see above.
-  %
-  \begin{example}[label=ex:HSSUSY_thresholds]
-    In the model file of \HSSUSY, 3-loop QCD corrections to the running
-    top Yukawa coupling $y_t$ are enabled (\code{UseYukawa3LoopQCD =
-      True}).  Switching between the 2-loop and 3-loop QCD corrections
-    to $y_t$ can be used to estimate a partial uncertainty of
-    the 2-loop Higgs pole mass.  In order to do this,
-    \code{FSHSSUSYOpenHandle[]} must be called twice, setting
-    \begin{align*}
-      \texttt{thresholdCorrections} \rightarrow 123111321 \\
-      \texttt{thresholdCorrections} \rightarrow 122111321
-    \end{align*}
-    respectively, and setting each time
-    \code{thresholdCorrectionsLoopOrder -> 3} to enable the 3-loop
-    corrections globally.  Note that the digit at the 6$^\text{th}$ position
-    from the right (the prefactor of $10^6$) has been changed from $3$
-    to $2$ to change the threshold correction loop order of $y_t$ from
-    3-loop to 2-loop.  \exref{ex:HSSUSY_uncertainty} makes use
-    of this method to estimate a partial uncertainty of
-    \HSSUSY based on changing the threshold correction loop orders for
-    $y_t$ and $\as$ in the SM\@.
-  \end{example}
-
-\item[\texttt{FlexibleSUSY[25]},]
-  \texttt{higgs3loopCorrectionRenScheme}: This option applies only to
-  MSSM models in which the 3-loop Higgs pole mass contributions from
-  \Himalaya are enabled (the flag \code{UseHiggs3LoopMSSM = True} is
-  set in the model file), see
-  \secref{sec:model_specific_contributions}.  With this option the
-  user can choose between the \DRbar and \MDRbar renormalization
-  scheme.  If this option is set to $0$, then the \DRbar scheme is
-  used.  If set to $1$, the \MDRbar scheme is used.
-
-\item[\texttt{FlexibleSUSY[26]},] \texttt{higgs3loopCorrectionAtAsAs}:
-  With this option the user can enable/disable 3-loop contributions to
-  the Higgs pole mass(es) of $O(\at\as^2)$.  Note that this option
-  has an effect only if model-specific 3-loop contributions to the
-  Higgs pole mass(es) of this order have been enabled in the \fs model file.  See
-  \secref{sec:model_specific_contributions} on how to enable 3-loop
-  contributions of this order in the SM and in the MSSM\@.
-
-\item[\texttt{FlexibleSUSY[27]},] \texttt{higgs3loopCorrectionAbAsAs}:
-  With this option the user can enable/disable 3-loop contributions to
-  the Higgs pole mass(es) of $O(\ab\as^2)$.  Note that this option
-  has an effect only if model-specific 3-loop contributions to the
-  Higgs pole mass(es) of this order have been enabled in the \fs model file.
-  See \secref{sec:model_specific_contributions} on how to enable
-  3-loop contributions of this order in the SM and in the
-  MSSM\@.
-
-\item[\texttt{FlexibleSUSY[28]},] \texttt{higgs3loopCorrectionAtAtAs}:
-  With this option the user can enable/disable 3-loop contributions to
-  the Higgs pole mass(es) of $O(\at^2\as)$.  Note that this option
-  has an effect only if model-specific 3-loop contributions to the
-  Higgs pole mass(es) of this order have been enabled in the \fs model file.
-  See \secref{sec:model_specific_contributions} on how to enable
-  3-loop contributions of this order in the SM\@.
-
-\item[\texttt{FlexibleSUSY[29]},] \texttt{higgs3loopCorrectionAtAtAt}:
-  With this option the user can enable/disable 3-loop contributions to
-  the Higgs pole mass(es) of $O(\at^3)$.  Note that this option has
-  an effect only if model-specific 3-loop contributions to the Higgs
-  pole mass(es) of this order have been enabled in the \fs model file.  See
-  \secref{sec:model_specific_contributions} on how to enable 3-loop
-  contributions of this order in the SM\@.
-
-\item[\texttt{MODSEL[12]},] \texttt{parameterOutputScale}: With this
-  option the scale (in GeV) can be specified, at which the running
-  \MSbar/\DRbar model parameters are output.  If set to $0$, then the
-  running parameters are output at the scale assigned to the
-  \code{SUSYScale} variable in the \fs model file.
-\end{itemize}
-
-\section{CMSSMCPV model file}
-\label{app:CMSSMCPV}
-\lstinputlisting[language=Mathematica]{model_files/CMSSMCPV/FlexibleSUSY.m.in}
-
-\section{\THDM model file}
-\label{app:THDMIIMSSMBC}
-\lstinputlisting[language=Mathematica,numbers=left,numberstyle=\tiny\sffamily]{model_files/THDMIIMSSMBC/FlexibleSUSY.m.in}
-
-\section{CNMSSM model file}
-\label{app:CNMSSM}
-\lstinputlisting[language=Mathematica]{model_files/CNMSSM/FlexibleSUSY.m.in}
-
 \clearpage
+\section{$\mu\to e\gamma$}
 
@@
 \bibliographystyle{doc/JHEP}
-\bibliography{doc/flexiblesusy-2.0}
+\bibliography{doc/flexiblesusy-new_features}
 

### 2019-03-18 doc/Manuals/flexiblesusy-2.0.tex
start tidying up doc directory

--- 2018-04-28 doc/flexiblesusy-2.0.tex
@@
   \centering
-  \includegraphics[width=0.5\textwidth]{doc/plots/HSSUSY/HSSUSY_thresholds}
+  \includegraphics[width=0.5\textwidth]{doc/Manuals/plots/HSSUSY/HSSUSY_thresholds}
   \caption{Effect of different SM corrections on the Higgs
@@
   \centering
-  \includegraphics[width=0.49\textwidth]{doc/plots/Himalaya/scan_Mh_MS_TB-5_Xt-0}\hfill
-  \includegraphics[width=0.49\textwidth]{doc/plots/Himalaya/scan_Mh_MS_TB-5_Xt-0_delta_alpha}\\
-  \includegraphics[width=0.49\textwidth]{doc/plots/Himalaya/scan_Mh_Xt_TB-5_MS-2000}\hfill
-  \includegraphics[width=0.49\textwidth]{doc/plots/Himalaya/scan_Mh_Xt_TB-5_MS-2000_delta_alpha}
+  \includegraphics[width=0.49\textwidth]{doc/Manuals/plots/Himalaya/scan_Mh_MS_TB-5_Xt-0}\hfill
+  \includegraphics[width=0.49\textwidth]{doc/Manuals/plots/Himalaya/scan_Mh_MS_TB-5_Xt-0_delta_alpha}\\
+  \includegraphics[width=0.49\textwidth]{doc/Manuals/plots/Himalaya/scan_Mh_Xt_TB-5_MS-2000}\hfill
+  \includegraphics[width=0.49\textwidth]{doc/Manuals/plots/Himalaya/scan_Mh_Xt_TB-5_MS-2000_delta_alpha}
   \caption{Comparison of different 2-loop and 3-loop calculations of
@@
   \centering
-  \includegraphics[width=0.49\textwidth]{doc/plots/MSSMMuBMu/scan_MSSMMuBMu_Mh_MS_TB-5_Xt-0}\hfill
-  \includegraphics[width=0.49\textwidth]{doc/plots/MSSMMuBMu/scan_MSSMMuBMu_Mh_Xt_TB-5_MS-2000}
+  \includegraphics[width=0.49\textwidth]{doc/Manuals/plots/MSSMMuBMu/scan_MSSMMuBMu_Mh_MS_TB-5_Xt-0}\hfill
+  \includegraphics[width=0.49\textwidth]{doc/Manuals/plots/MSSMMuBMu/scan_MSSMMuBMu_Mh_Xt_TB-5_MS-2000}
   \caption{Effect of the 3-loop RGEs for all MSSM parameters (red),
@@
   \centering
-  \includegraphics[width=0.49\textwidth]{doc/plots/THDMIIMSSMBC/THDMIIMSSMBC_Mh_MS_TB_MA-200_Xt-0}\hfill
-  \includegraphics[width=0.49\textwidth]{doc/plots/THDMIIMSSMBC/THDMIIMSSMBC_Mh_MS_TB_MA-200_Xt-sqrt6}
+  \includegraphics[width=0.49\textwidth]{doc/Manuals/plots/THDMIIMSSMBC/THDMIIMSSMBC_Mh_MS_TB_MA-200_Xt-0}\hfill
+  \includegraphics[width=0.49\textwidth]{doc/Manuals/plots/THDMIIMSSMBC/THDMIIMSSMBC_Mh_MS_TB_MA-200_Xt-sqrt6}
   \caption{Lightest \CP-even Higgs pole mass calculated at the 1-loop
@@
 can be estimated.\footnote{The example scripts can be found in the
-  \code{doc/examples-2.0/} sub-directory of the \fs package.}
+  \code{examples-2.0/} sub-directory of the \fs package.}
 %
@@
 \begin{center}
-  \includegraphics[width=0.5\textwidth]{doc/plots/HSSUSY/Mh_Xt}
+  \includegraphics[width=0.5\textwidth]{doc/Manuals/plots/HSSUSY/Mh_Xt}
 \end{center}
@@
 \begin{center}
-  \includegraphics[width=0.5\textwidth]{doc/plots/HSSUSY/Mh_Xt_uncertainty}
+  \includegraphics[width=0.5\textwidth]{doc/Manuals/plots/HSSUSY/Mh_Xt_uncertainty}
 \end{center}
@@
 \begin{center}
-  \includegraphics[width=0.5\textwidth]{doc/plots/HSSUSY/Mh_MS_EFT_uncertainty}
+  \includegraphics[width=0.5\textwidth]{doc/Manuals/plots/HSSUSY/Mh_MS_EFT_uncertainty}
 \end{center}
@@
   \subfloat[Type FFS]{
-    \includegraphics{doc/Feynman/amuFFS}
+    \includegraphics{doc/Manuals/Feynman/amuFFS}
     \label{fig:amu_diagram_I}
@@
   \subfloat[Type SSF]{
-    \includegraphics{doc/Feynman/amuFSS}
+    \includegraphics{doc/Manuals/Feynman/amuFSS}
     \label{fig:amu_diagram_II}
@@
   \centering
-  \includegraphics[width=0.49\textwidth]{doc/plots/MSSMCPV/contour_de_PhiMuM12}\hfill
-  \includegraphics[width=0.49\textwidth]{doc/plots/MSSMCPV/contour_de_PhiMuTe}
+  \includegraphics[width=0.49\textwidth]{doc/Manuals/plots/MSSMCPV/contour_de_PhiMuM12}\hfill
+  \includegraphics[width=0.49\textwidth]{doc/Manuals/plots/MSSMCPV/contour_de_PhiMuTe}
   \caption{Left panel: contours of the electron EDM in the
@@
   \centering
-  \includegraphics[width=0.49\textwidth]{doc/plots/CMSSMCPV/scan_Mh_PhiMu_m0-500_M12-500_TB-10_A0-0}\hfill
-  \includegraphics[width=0.49\textwidth]{doc/plots/CMSSMCPV/scan_TYu33_Q_m0-500_M12-500_TB-10_A0-500_PhiMu-0_PhiA0-Pi4}
+  \includegraphics[width=0.49\textwidth]{doc/Manuals/plots/CMSSMCPV/scan_Mh_PhiMu_m0-500_M12-500_TB-10_A0-0}\hfill
+  \includegraphics[width=0.49\textwidth]{doc/Manuals/plots/CMSSMCPV/scan_TYu33_Q_m0-500_M12-500_TB-10_A0-500_PhiMu-0_PhiA0-Pi4}
   \caption{Left panel: lightest Higgs pole mass calculated in the
@@
   \subfloat[]{
-  \quad\includegraphics{doc/Feynman/deltaVBWave}\quad
+  \quad\includegraphics{doc/Manuals/Feynman/deltaVBWave}\quad
   \label{fig:delta_VB_wave_diagrams}
@@
   \subfloat[]{
-  \qquad\includegraphics{doc/Feynman/deltaVBVertex1}\quad
-  \includegraphics{doc/Feynman/deltaVBVertex2}\qquad
+  \qquad\includegraphics{doc/Manuals/Feynman/deltaVBVertex1}\quad
+  \includegraphics{doc/Manuals/Feynman/deltaVBVertex2}\qquad
   \label{fig:delta_VB_vertex_diagrams}
@@
   \subfloat[]{
-  \quad\includegraphics{doc/Feynman/deltaVBBox1}\quad
+  \quad\includegraphics{doc/Manuals/Feynman/deltaVBBox1}\quad
   \label{fig:delta_VB_box_diagrams}
@@
   \centering
-  \includegraphics[width=0.49\textwidth]{doc/plots/FlexibleMW/scan_MW_M12}\hfill
-  \includegraphics[width=0.49\textwidth]{doc/plots/FlexibleMW/scan_MW_LamTU}
+  \includegraphics[width=0.49\textwidth]{doc/Manuals/plots/FlexibleMW/scan_MW_M12}\hfill
+  \includegraphics[width=0.49\textwidth]{doc/Manuals/plots/FlexibleMW/scan_MW_LamTU}
   \caption{Left panel: $W$ pole mass in the CMSSM as a function of
@@
   \includegraphics[width=0.6\textwidth]{%
-    doc/plots/semi_analytic_rel_errs/single_step_errors}
+    doc/Manuals/plots/semi_analytic_rel_errs/single_step_errors}
   \caption{Percentage changes in the CMSSM \DRbar mass spectrum after
@@
   \includegraphics[width=0.5\textwidth]{%
-    doc/plots/semi_analytic_benchmark/i7-4702MQ}
+    doc/Manuals/plots/semi_analytic_benchmark/i7-4702MQ}
   \caption{Runtime distributions of the CMSSM spectrum generators created with
@@
   \includegraphics[width=0.48\textwidth]{%
-    doc/plots/CMSSM_solution_regions/CMSSM_solution_regions}
+    doc/Manuals/plots/CMSSM_solution_regions/CMSSM_solution_regions}
   \includegraphics[width=0.48\textwidth]{%
-    doc/plots/CMSSM_solution_regions/CMSSM_higgs_mass}
+    doc/Manuals/plots/CMSSM_solution_regions/CMSSM_higgs_mass}
   \caption{Left panel: CMSSM solutions found by the two-scale solver
@@
   \includegraphics[width=0.6\textwidth]{%
-    doc/plots/CMSSM_multiple_solutions/CMSSM_multiple_solutions}
+    doc/Manuals/plots/CMSSM_multiple_solutions/CMSSM_multiple_solutions}
   \caption{Solutions obtained using the two-scale and semi-analytic solvers
@@
   \includegraphics[width=0.48\textwidth]{%
-     doc/plots/CNMSSM/cnmssm_4D-m0sqpos-full-m0_m12_sort-A0.pdf}
+    doc/Manuals/plots/CNMSSM/cnmssm_4D-m0sqpos-full-m0_m12_sort-A0.pdf}
    \includegraphics[width=0.48\textwidth]{%
-     doc/plots/CNMSSM/cnmssm_4D-m0sqpos-full_mhZoom_lam_m12_sort-mh.pdf} \\
+    doc/Manuals/plots/CNMSSM/cnmssm_4D-m0sqpos-full_mhZoom_lam_m12_sort-mh.pdf} \\
   \includegraphics[width=0.48\textwidth]{%
-    doc/plots/CNMSSM/cnmssm_4D-full-squark_mass_m0sqgtr0_m0_m12.pdf}
+    doc/Manuals/plots/CNMSSM/cnmssm_4D-full-squark_mass_m0sqgtr0_m0_m12.pdf}
    \includegraphics[width=0.48\textwidth]{%
-    doc/plots/CNMSSM/cnmssm_4D-m0sqgtr0-full-msu6.pdf}
+    doc/Manuals/plots/CNMSSM/cnmssm_4D-m0sqgtr0-full-msu6.pdf}
    \caption{CNMSSM solutions obtained using \fstwo and \multinest 3.10
@@
   \includegraphics[width=0.48\textwidth]{%
-     doc/plots/CE6SSM/ce6ssm_tb10-scan-l-k-vs_combined_m0_m12_ccA0.pdf}
+    doc/Manuals/plots/CE6SSM/ce6ssm_tb10-scan-l-k-vs_combined_m0_m12_ccA0.pdf}
    \includegraphics[width=0.48\textwidth]{%
-     doc/plots/CE6SSM/ce6ssm_tb10-scan-l-k-vs_combined_m0_m12_ccmsu1.pdf} \\
+    doc/Manuals/plots/CE6SSM/ce6ssm_tb10-scan-l-k-vs_combined_m0_m12_ccmsu1.pdf} \\
   \includegraphics[width=0.48\textwidth]{%
-    doc/plots/CE6SSM/ce6ssm_tb10-scan-l-k-vs_combined_m0_m12_cclam.pdf}
+    doc/Manuals/plots/CE6SSM/ce6ssm_tb10-scan-l-k-vs_combined_m0_m12_cclam.pdf}
    \includegraphics[width=0.48\textwidth]{%
-    doc/plots/CE6SSM/ce6ssm_tb10-scan-l-k-vs_combined_m0_m12_cckap.pdf}
+    doc/Manuals/plots/CE6SSM/ce6ssm_tb10-scan-l-k-vs_combined_m0_m12_cckap.pdf}
    \caption{C\ESSM solutions obtained using \fstwo and \multinest 3.10 shown in
@@
   \includegraphics[width=0.48\textwidth]{%
-    doc/plots/CSE6SSM/CSE6SSM_m12_Azero_m0}
+    doc/Manuals/plots/CSE6SSM/CSE6SSM_m12_Azero_m0}
   \includegraphics[width=0.48\textwidth]{%
-    doc/plots/CSE6SSM/CSE6SSM_m12_Azero_m0_handwritten}
+    doc/Manuals/plots/CSE6SSM/CSE6SSM_m12_Azero_m0_handwritten}
   \caption{CS\ESSM solutions obtained using the spectrum generator
@@
   \centering
-  \includegraphics[width=0.49\textwidth]{{{doc/plots/FlexibleEFTHiggs/scan_Mh_MS_TB-5_Xt-0}}} \hfill
-  \includegraphics[width=0.49\textwidth]{{{doc/plots/FlexibleEFTHiggs/scan_Mh_Xt_TB-5_MS-2000}}}
+  \includegraphics[width=0.49\textwidth]{{{doc/Manuals/plots/FlexibleEFTHiggs/scan_Mh_MS_TB-5_Xt-0}}} \hfill
+  \includegraphics[width=0.49\textwidth]{{{doc/Manuals/plots/FlexibleEFTHiggs/scan_Mh_Xt_TB-5_MS-2000}}}
   \caption{Comparison of the predicted lightest \CP-even Higgs pole mass in the
@@
 
-\bibliographystyle{doc/JHEP}
-\bibliography{doc/flexiblesusy-2.0}
+\bibliographystyle{doc/Manuals/JHEP}
+\bibliography{doc/Manuals/flexiblesusy-2.0}
 

### 2019-03-18 doc/manuals/flexiblesusy-2.0.tex
rename doc/Models -> doc/models and doc/Manual -> doc/manual

--- 2018-04-28 doc/flexiblesusy-2.0.tex
@@
   \centering
-  \includegraphics[width=0.5\textwidth]{doc/plots/HSSUSY/HSSUSY_thresholds}
+  \includegraphics[width=0.5\textwidth]{doc/manuals/plots/HSSUSY/HSSUSY_thresholds}
   \caption{Effect of different SM corrections on the Higgs
@@
   \centering
-  \includegraphics[width=0.49\textwidth]{doc/plots/Himalaya/scan_Mh_MS_TB-5_Xt-0}\hfill
-  \includegraphics[width=0.49\textwidth]{doc/plots/Himalaya/scan_Mh_MS_TB-5_Xt-0_delta_alpha}\\
-  \includegraphics[width=0.49\textwidth]{doc/plots/Himalaya/scan_Mh_Xt_TB-5_MS-2000}\hfill
-  \includegraphics[width=0.49\textwidth]{doc/plots/Himalaya/scan_Mh_Xt_TB-5_MS-2000_delta_alpha}
+  \includegraphics[width=0.49\textwidth]{doc/manuals/plots/Himalaya/scan_Mh_MS_TB-5_Xt-0}\hfill
+  \includegraphics[width=0.49\textwidth]{doc/manuals/plots/Himalaya/scan_Mh_MS_TB-5_Xt-0_delta_alpha}\\
+  \includegraphics[width=0.49\textwidth]{doc/manuals/plots/Himalaya/scan_Mh_Xt_TB-5_MS-2000}\hfill
+  \includegraphics[width=0.49\textwidth]{doc/manuals/plots/Himalaya/scan_Mh_Xt_TB-5_MS-2000_delta_alpha}
   \caption{Comparison of different 2-loop and 3-loop calculations of
@@
   \centering
-  \includegraphics[width=0.49\textwidth]{doc/plots/MSSMMuBMu/scan_MSSMMuBMu_Mh_MS_TB-5_Xt-0}\hfill
-  \includegraphics[width=0.49\textwidth]{doc/plots/MSSMMuBMu/scan_MSSMMuBMu_Mh_Xt_TB-5_MS-2000}
+  \includegraphics[width=0.49\textwidth]{doc/manuals/plots/MSSMMuBMu/scan_MSSMMuBMu_Mh_MS_TB-5_Xt-0}\hfill
+  \includegraphics[width=0.49\textwidth]{doc/manuals/plots/MSSMMuBMu/scan_MSSMMuBMu_Mh_Xt_TB-5_MS-2000}
   \caption{Effect of the 3-loop RGEs for all MSSM parameters (red),
@@
   \centering
-  \includegraphics[width=0.49\textwidth]{doc/plots/THDMIIMSSMBC/THDMIIMSSMBC_Mh_MS_TB_MA-200_Xt-0}\hfill
-  \includegraphics[width=0.49\textwidth]{doc/plots/THDMIIMSSMBC/THDMIIMSSMBC_Mh_MS_TB_MA-200_Xt-sqrt6}
+  \includegraphics[width=0.49\textwidth]{doc/manuals/plots/THDMIIMSSMBC/THDMIIMSSMBC_Mh_MS_TB_MA-200_Xt-0}\hfill
+  \includegraphics[width=0.49\textwidth]{doc/manuals/plots/THDMIIMSSMBC/THDMIIMSSMBC_Mh_MS_TB_MA-200_Xt-sqrt6}
   \caption{Lightest \CP-even Higgs pole mass calculated at the 1-loop
@@
 \begin{center}
-  \includegraphics[width=0.5\textwidth]{doc/plots/HSSUSY/Mh_Xt}
+  \includegraphics[width=0.5\textwidth]{doc/manuals/plots/HSSUSY/Mh_Xt}
 \end{center}
@@
 \begin{center}
-  \includegraphics[width=0.5\textwidth]{doc/plots/HSSUSY/Mh_Xt_uncertainty}
+  \includegraphics[width=0.5\textwidth]{doc/manuals/plots/HSSUSY/Mh_Xt_uncertainty}
 \end{center}
@@
 \begin{center}
-  \includegraphics[width=0.5\textwidth]{doc/plots/HSSUSY/Mh_MS_EFT_uncertainty}
+  \includegraphics[width=0.5\textwidth]{doc/manuals/plots/HSSUSY/Mh_MS_EFT_uncertainty}
 \end{center}
@@
   \subfloat[Type FFS]{
-    \includegraphics{doc/Feynman/amuFFS}
+    \includegraphics{doc/manuals/Feynman/amuFFS}
     \label{fig:amu_diagram_I}
@@
   \subfloat[Type SSF]{
-    \includegraphics{doc/Feynman/amuFSS}
+    \includegraphics{doc/manuals/Feynman/amuFSS}
     \label{fig:amu_diagram_II}
@@
   \centering
-  \includegraphics[width=0.49\textwidth]{doc/plots/MSSMCPV/contour_de_PhiMuM12}\hfill
-  \includegraphics[width=0.49\textwidth]{doc/plots/MSSMCPV/contour_de_PhiMuTe}
+  \includegraphics[width=0.49\textwidth]{doc/manuals/plots/MSSMCPV/contour_de_PhiMuM12}\hfill
+  \includegraphics[width=0.49\textwidth]{doc/manuals/plots/MSSMCPV/contour_de_PhiMuTe}
   \caption{Left panel: contours of the electron EDM in the
@@
   \centering
-  \includegraphics[width=0.49\textwidth]{doc/plots/CMSSMCPV/scan_Mh_PhiMu_m0-500_M12-500_TB-10_A0-0}\hfill
-  \includegraphics[width=0.49\textwidth]{doc/plots/CMSSMCPV/scan_TYu33_Q_m0-500_M12-500_TB-10_A0-500_PhiMu-0_PhiA0-Pi4}
+  \includegraphics[width=0.49\textwidth]{doc/manuals/plots/CMSSMCPV/scan_Mh_PhiMu_m0-500_M12-500_TB-10_A0-0}\hfill
+  \includegraphics[width=0.49\textwidth]{doc/manuals/plots/CMSSMCPV/scan_TYu33_Q_m0-500_M12-500_TB-10_A0-500_PhiMu-0_PhiA0-Pi4}
   \caption{Left panel: lightest Higgs pole mass calculated in the
@@
   \subfloat[]{
-  \quad\includegraphics{doc/Feynman/deltaVBWave}\quad
+  \quad\includegraphics{doc/manuals/Feynman/deltaVBWave}\quad
   \label{fig:delta_VB_wave_diagrams}
@@
   \subfloat[]{
-  \qquad\includegraphics{doc/Feynman/deltaVBVertex1}\quad
-  \includegraphics{doc/Feynman/deltaVBVertex2}\qquad
+  \qquad\includegraphics{doc/manuals/Feynman/deltaVBVertex1}\quad
+  \includegraphics{doc/manuals/Feynman/deltaVBVertex2}\qquad
   \label{fig:delta_VB_vertex_diagrams}
@@
   \subfloat[]{
-  \quad\includegraphics{doc/Feynman/deltaVBBox1}\quad
+  \quad\includegraphics{doc/manuals/Feynman/deltaVBBox1}\quad
   \label{fig:delta_VB_box_diagrams}
@@
   \centering
-  \includegraphics[width=0.49\textwidth]{doc/plots/FlexibleMW/scan_MW_M12}\hfill
-  \includegraphics[width=0.49\textwidth]{doc/plots/FlexibleMW/scan_MW_LamTU}
+  \includegraphics[width=0.49\textwidth]{doc/manuals/plots/FlexibleMW/scan_MW_M12}\hfill
+  \includegraphics[width=0.49\textwidth]{doc/manuals/plots/FlexibleMW/scan_MW_LamTU}
   \caption{Left panel: $W$ pole mass in the CMSSM as a function of
@@
   \includegraphics[width=0.6\textwidth]{%
-    doc/plots/semi_analytic_rel_errs/single_step_errors}
+    doc/manuals/plots/semi_analytic_rel_errs/single_step_errors}
   \caption{Percentage changes in the CMSSM \DRbar mass spectrum after
@@
   \includegraphics[width=0.5\textwidth]{%
-    doc/plots/semi_analytic_benchmark/i7-4702MQ}
+    doc/manuals/plots/semi_analytic_benchmark/i7-4702MQ}
   \caption{Runtime distributions of the CMSSM spectrum generators created with
@@
   \includegraphics[width=0.48\textwidth]{%
-    doc/plots/CMSSM_solution_regions/CMSSM_solution_regions}
+    doc/manuals/plots/CMSSM_solution_regions/CMSSM_solution_regions}
   \includegraphics[width=0.48\textwidth]{%
-    doc/plots/CMSSM_solution_regions/CMSSM_higgs_mass}
+    doc/manuals/plots/CMSSM_solution_regions/CMSSM_higgs_mass}
   \caption{Left panel: CMSSM solutions found by the two-scale solver
@@
   \includegraphics[width=0.6\textwidth]{%
-    doc/plots/CMSSM_multiple_solutions/CMSSM_multiple_solutions}
+    doc/manuals/plots/CMSSM_multiple_solutions/CMSSM_multiple_solutions}
   \caption{Solutions obtained using the two-scale and semi-analytic solvers
@@
   \includegraphics[width=0.48\textwidth]{%
-     doc/plots/CNMSSM/cnmssm_4D-m0sqpos-full-m0_m12_sort-A0.pdf}
+    doc/manuals/plots/CNMSSM/cnmssm_4D-m0sqpos-full-m0_m12_sort-A0.pdf}
    \includegraphics[width=0.48\textwidth]{%
-     doc/plots/CNMSSM/cnmssm_4D-m0sqpos-full_mhZoom_lam_m12_sort-mh.pdf} \\
+    doc/manuals/plots/CNMSSM/cnmssm_4D-m0sqpos-full_mhZoom_lam_m12_sort-mh.pdf} \\
   \includegraphics[width=0.48\textwidth]{%
-    doc/plots/CNMSSM/cnmssm_4D-full-squark_mass_m0sqgtr0_m0_m12.pdf}
+    doc/manuals/plots/CNMSSM/cnmssm_4D-full-squark_mass_m0sqgtr0_m0_m12.pdf}
    \includegraphics[width=0.48\textwidth]{%
-    doc/plots/CNMSSM/cnmssm_4D-m0sqgtr0-full-msu6.pdf}
+    doc/manuals/plots/CNMSSM/cnmssm_4D-m0sqgtr0-full-msu6.pdf}
    \caption{CNMSSM solutions obtained using \fstwo and \multinest 3.10
@@
   \includegraphics[width=0.48\textwidth]{%
-     doc/plots/CE6SSM/ce6ssm_tb10-scan-l-k-vs_combined_m0_m12_ccA0.pdf}
+    doc/manuals/plots/CE6SSM/ce6ssm_tb10-scan-l-k-vs_combined_m0_m12_ccA0.pdf}
    \includegraphics[width=0.48\textwidth]{%
-     doc/plots/CE6SSM/ce6ssm_tb10-scan-l-k-vs_combined_m0_m12_ccmsu1.pdf} \\
+    doc/manuals/plots/CE6SSM/ce6ssm_tb10-scan-l-k-vs_combined_m0_m12_ccmsu1.pdf} \\
   \includegraphics[width=0.48\textwidth]{%
-    doc/plots/CE6SSM/ce6ssm_tb10-scan-l-k-vs_combined_m0_m12_cclam.pdf}
+    doc/manuals/plots/CE6SSM/ce6ssm_tb10-scan-l-k-vs_combined_m0_m12_cclam.pdf}
    \includegraphics[width=0.48\textwidth]{%
-    doc/plots/CE6SSM/ce6ssm_tb10-scan-l-k-vs_combined_m0_m12_cckap.pdf}
+    doc/manuals/plots/CE6SSM/ce6ssm_tb10-scan-l-k-vs_combined_m0_m12_cckap.pdf}
    \caption{C\ESSM solutions obtained using \fstwo and \multinest 3.10 shown in
@@
   \includegraphics[width=0.48\textwidth]{%
-    doc/plots/CSE6SSM/CSE6SSM_m12_Azero_m0}
+    doc/manuals/plots/CSE6SSM/CSE6SSM_m12_Azero_m0}
   \includegraphics[width=0.48\textwidth]{%
-    doc/plots/CSE6SSM/CSE6SSM_m12_Azero_m0_handwritten}
+    doc/manuals/plots/CSE6SSM/CSE6SSM_m12_Azero_m0_handwritten}
   \caption{CS\ESSM solutions obtained using the spectrum generator
@@
   \centering
-  \includegraphics[width=0.49\textwidth]{{{doc/plots/FlexibleEFTHiggs/scan_Mh_MS_TB-5_Xt-0}}} \hfill
-  \includegraphics[width=0.49\textwidth]{{{doc/plots/FlexibleEFTHiggs/scan_Mh_Xt_TB-5_MS-2000}}}
+  \includegraphics[width=0.49\textwidth]{{{doc/manuals/plots/FlexibleEFTHiggs/scan_Mh_MS_TB-5_Xt-0}}} \hfill
+  \includegraphics[width=0.49\textwidth]{{{doc/manuals/plots/FlexibleEFTHiggs/scan_Mh_Xt_TB-5_MS-2000}}}
   \caption{Comparison of the predicted lightest \CP-even Higgs pole mass in the
@@
 
-\bibliographystyle{doc/JHEP}
-\bibliography{doc/flexiblesusy-2.0}
+\bibliographystyle{doc/manuals/JHEP}
+\bibliography{doc/manuals/flexiblesusy-2.0}
 
